### 2021-11-03 analyses.RmD
- set-up of repo

---
title: "Analyses"
output:
  html_document:
    toc: TRUE
    toc_float:
      collapsed: FALSE
    toc_depth: 2
    code_folding: hide
bibliography      : 
  - "bibliography/r-references.bib"
  - "bibliography/references.bib"
  # pdf_document:
  #   toc: TRUE
---

```{r knitr-setup, include=F}
set.seed(170819)
knitr::opts_chunk$set(warning=F, echo=F, error=F, message=F, cache = T, 
                      tidy=T, dev = 'pdf', cache.extra = knitr::rand_seed)
options(digits = 3, width = 120)
```

# Set-up

Load packages, and make repo replicable.

```{r r-setup}
# install packages
# devtools::install_github("https://github.com/tdienlin/td")

# define packages
packages <- c("corrplot", "devtools", "GGally", "ggplot2", "gridExtra", "haven", "kableExtra", "knitr", 
              "lavaan", "lme4", "magrittr", "mice", "MVN", "PerFit", "psych", "quanteda.textstats",
              "pwr", "semPlot", "semTools", "tidyverse")

# renv::install(packages)
# renv::install("tdienlin/td@v.0.0.2.5")  # uncomment to install
# renv::install("crsh/papaja@v0.1.0.9997")  # uncomment to install

# load packages
lapply(c(packages, "td"), library, character.only = TRUE)
```

## Custom functions

Find Variables in dataset.

```{r}
find_var <- function(name, data = doc)(
  # finds the variables names for an item for each wave
    data %>% 
    filter(Label == name) %>% 
    select(variable) %>% 
    unlist() %>% 
    set_names(sub("\\_.*", "", .))
)
```

Make graph of variables' development.

```{r}
make_graph <- function(model, title, ll, ul, line = FALSE, labels = FALSE){
  dat <- coefficients(model)$wave %>% 
    tibble::rownames_to_column("wave") %>% 
    rename(value = "(Intercept)") %>% 
    mutate(wave = as.integer(.$wave),
           value = as.numeric(.$value))

  graph <- 
    ggplot(dat, aes(wave, value)) +
    theme_bw() +
    theme(text = element_text(size = 8)) + 
    geom_bar(data = dat, aes(wave, value), stat = "identity") +
    coord_cartesian(ylim = c(ll, ul)) +
    ggtitle(title)

  if(line) {graph <- graph + 
    geom_smooth(se = FALSE, method = 'loess', size = 1.5)}

  if(labels) {graph <- graph +
    geom_text(aes(label = round(value, 2)), hjust = 1.2, color = "white",
              fontface = "bold", angle = 90)
  }

  graph
}
```

Make graph of effects

```{r}
make_graph_res <- function(data){
  ggplot(data, aes(x = estimate, y = iv)) + 
  scale_color_manual(values = c("grey40", "deepskyblue")) +
  geom_vline(xintercept = 0, lwd = .75, colour = "darkgrey") + 
  geom_errorbarh(aes(xmin = conf.low, xmax = conf.high),
                lwd = .75, height = .2
                , position = position_dodge(.6)
                ) +
  geom_point(size = 2
             , position = position_dodge(.6)
             ) +
  geom_vline(data=filter(data, dv=="Life satisfaction"), aes(xintercept=.3), colour="darkgrey", linetype = "dashed") + 
    geom_vline(data=filter(data, dv=="Life satisfaction"), aes(xintercept=-.3), colour="darkgrey", linetype = "dashed") + 
  geom_vline(data=filter(data, dv!="Life satisfaction"), aes(xintercept=.15), colour="darkgrey", linetype = "dashed") + 
    geom_vline(data=filter(data, dv!="Life satisfaction"), aes(xintercept=-.15), colour="darkgrey", linetype = "dashed") + 
  theme(
    axis.title.y = element_blank(),
    plot.title = element_text(hjust = .5),
    panel.spacing = unit(.9, "lines"),
    text = element_text(size = 12),
    axis.text.y = element_text(hjust = 0)
    ) +
  facet_grid(rows = vars(type),
             cols = vars(dv),
             scales = "free") 
}
```

# Data Wrangling

```{r}
doc <- read_csv("documentation/10094_vi_de_v4_0.csv")
# d_raw <- read_spss("data/10094_da_de_v4_0.zsav")
d_raw <- read_csv("data/10094_da_de_v4_0.csv")
```

Identify the names of each item for each wave.

```{r}
# identify variable names of items, using custom function
health <- find_var("Gesundheitszustand") # wasn't collected at wave 21
aff_neg_1 <- find_var("Depressivitaet: einsam")
aff_neg_2 <- find_var("Depressivitaet: aergerlich")
aff_neg_3 <- find_var("Depressivitaet: so niedergeschlagen")
aff_neg_4 <- find_var("Depressivitaet: sehr nervoes")
aff_neg_5 <- find_var("Depressivitaet: aengstlich")
aff_neg_6 <- find_var("Depressivitaet: bedrueckt und traurig")
aff_pos_1 <- find_var("Depressivitaet: ruhig und gelassen")
aff_pos_2 <- find_var("Depressivitaet: gluecklich")
aff_pos_3 <- find_var("Depressivitaet: voller Energie")
act_wrk <- find_var("Zuhause verlassen: Arbeit")
act_spo <- find_var("Zuhause verlassen: Sport")
act_frn <- find_var("Zuhause verlassen: Freunde oder Verwandte treffen")
act_sho <- find_var("Zuhause verlassen: Essen einkaufen")
act_pet <- find_var("Zuhause verlassen: Haustier ausfuehren")
sat_dem <- find_var("Demokratiezufriedenheit: Oesterreich")
work_h <- find_var("Arbeitsstunden: Jetzt pro Woche")
work_homeoff <- find_var("Aenderung berufliche Situation: Home-Office")
hh_income <- find_var("Aktuelles Haushaltseinkommen")
med_txt_kro <- find_var("Mediennutzung: Kronen Zeitung oder www.krone.at")
med_txt_sta <- find_var("Mediennutzung: Der Standard oder derstandard.at")
med_txt_pre <- find_var("Mediennutzung: Die Presse oder diepresse.com")
med_txt_oes <- find_var("Mediennutzung: Oesterreich oder oe24.at")
med_txt_kur <- find_var("Mediennutzung: Kurier oder kurier.at")
med_txt_slz <- find_var("Mediennutzung: Salzburger Nachrichten oder salzburg.at")
med_txt_son <- find_var("Mediennutzung: Sonstige oesterreichische Tageszeitungen")
med_vid_orf <- find_var("Mediennutzung: ORF (Nachrichten)")
med_vid_pri <- find_var("Mediennutzung: Privatfernsehen (Nachrichten)")
soc_med_fb <- find_var("Soziale Medien: Facebook")
soc_med_tw <- find_var("Soziale Medien: Twitter")
soc_med_ig <- find_var("Soziale Medien: Instagram")
soc_med_yt <- find_var("Soziale Medien: Youtube")
soc_med_wa <- find_var("Soziale Medien: WhatsApp")
soc_med_read <- find_var("Soziale Medien Aktivitaet: Postings zu Corona anderer lesen")
soc_med_like_share <- find_var("Soziale Medien Aktivitaet: Postings liken, teilen oder retweeten")
soc_med_post <- find_var("Soziale Medien Aktivitaet: selber Postings zu Corona verfassen")
life_sat <- find_var("Lebenszufriedenheit")
risk_prop <- find_var("Risikobereitschaft")
loc_cntrl_int_1 <- find_var("Psychologie: habe Leben selbst in der Hand")
loc_cntrl_int_2 <- find_var("Psychologie: Belohnung durch Anstrengung")
loc_cntrl_int_3 <- find_var("Psychologie: Fremdbestimmung")
loc_cntrl_int_4 <- find_var("Psychologie: Schicksal")
```

Select variables.

```{r}
d_wide <- d_raw %>% 
  select(
    id = RESPID,
    gender = SD_GENDER,
    acc_bal = SD_ACCESS_BALCONY,
    acc_gar = SD_ACCESS_GARDEN,
    year_birth = SD_BIRTHYEAR,
    born_aus = SD_BORN_AUSTRIA,
    born_aus_prnts = SD_MIGR_BACKGR,
    county = SD_BULA,
    edu = SD_EDU,
    employment = SD_EMPLSTATUS_FEB2020,
    hh_adults = SD_HH_ADULTS,
    hh_child18 = SD_HH_CHILD18,
    hh_child17 = SD_HH_TEENS,
    hh_child14 = SD_HH_CHILD14,
    hh_child5 = SD_HH_CHILD5,
    hh_child2 = SD_HH_CHILD2,
    hh_oldfam = SD_HH_OLDERFAM,
    hh_outfam = SD_HH_OUTERFAM,
    hh_partner = SD_HH_PARTNER,
    # hh_income = SD_HHINCOME_FEB2020,
    home_sqm = SD_HOME_SQM,
    home_owner = SD_HOMEOWNER,
    # work_h = SD_WORKHOURS_FEB2020,
    health = all_of(health),
    life_sat = all_of(life_sat),
    aff_neg_1 = all_of(aff_neg_1),
    aff_neg_2 = all_of(aff_neg_2),
    aff_neg_3 = all_of(aff_neg_3),
    aff_neg_4 = all_of(aff_neg_4),
    aff_neg_5 = all_of(aff_neg_5),
    aff_neg_6 = all_of(aff_neg_6),
    aff_pos_1 = all_of(aff_pos_1),
    aff_pos_2 = all_of(aff_pos_2),
    aff_pos_3 = all_of(aff_pos_3),
    act_wrk = all_of(act_wrk),
    act_spo = all_of(act_spo),
    act_frn = all_of(act_frn),
    act_sho = all_of(act_sho),
    act_pet = all_of(act_pet),
    sat_dem = all_of(sat_dem),
    sat_dem = all_of(sat_dem),
    work_h = all_of(work_h),
    work_homeoff = all_of(work_homeoff),
    hh_income = all_of(hh_income),
    med_txt_kro = all_of(med_txt_kro),
    med_txt_sta = all_of(med_txt_sta),
    med_txt_pre = all_of(med_txt_pre),
    med_txt_oes = all_of(med_txt_oes),
    med_txt_kur = all_of(med_txt_kur),
    med_txt_slz = all_of(med_txt_slz),
    med_txt_son = all_of(med_txt_son),
    med_vid_orf = all_of(med_vid_orf),
    med_vid_pri = all_of(med_vid_pri),
    soc_med_fb = all_of(soc_med_fb),
    soc_med_tw = all_of(soc_med_tw),
    soc_med_ig = all_of(soc_med_ig),
    soc_med_yt = all_of(soc_med_yt),
    soc_med_wa = all_of(soc_med_wa),
    soc_med_like_share = all_of(soc_med_like_share),
    soc_med_read = all_of(soc_med_read),
    soc_med_post = all_of(soc_med_post),
    risk_prop = all_of(risk_prop),
    loc_cntrl_int_1 = all_of(loc_cntrl_int_1),
    loc_cntrl_int_2 = all_of(loc_cntrl_int_2),
    loc_cntrl_int_3 = all_of(loc_cntrl_int_3),
    loc_cntrl_int_4 = all_of(loc_cntrl_int_4)
  )

# rename variables
# names(d_wide) <- sub("\\.\\.\\.", "§", names(d_wide)) # replace "..." with "_"
```

Recode Variables

```{r}
d_wide %<>%
  mutate_at(vars(everything(.)), funs(na_if(., 88))) %>%
  mutate_at(vars(everything(.)), funs(na_if(., 99))) %>%
  mutate(
    male = 2 - .$gender,
    age = 2021 - .$year_birth,
    res_vienna = recode(.$county, `8` = 1L, .default = 0L,),
    born_aus = 2 - .$born_aus,
    employment_fac = factor(.$employment, 
                           labels = c("Unselbststaendig erwerbstaetig (auch Lehrling), Privatwirtschaft", 
"Unselbststaendig erwerbstaetig (auch Lehrling), Oeffentlicher Dienst",
"Selbststaendig erwerbstaetig",
"Arbeitslos",
"Pensionist*in",
"Ausschliesslich haushaltsfuehrend",
"Schueler/in, Student/in, Bildungskarenz",
"Sonstiges (Dauerhaft arbeitsunfaehig oder Praesenz-/Zivildiener)",
"In Elternkarenz")),
    edu_fac = factor(.$edu,
                     labels = c("Volksschule oder weniger", 
                                "Hauptschule oder AHS Unterstufe",
                                "Polytechnikum, BMS (Fachschule, z.B. HASCH)",
                                "Lehre, Berufsschule",
                                "AHS mit Matura",
                                "BHS mit Matura (z.B. HTL, HAK, HBLA etc.)",
                                "Hochschulverwandte Lehranstalt oder Kolleg",
                                "Bachelor",
                                "Magister / Master / Diplomingenieur / Fachhochschule",
"Doktor / PhD")))
```

Make new documentation with selected variables.

```{r}
doc_selected <- filter(doc, variable %in% c(
  "RESPID",
  "SD_GENDER",
  "SD_ACCESS_BALCONY",
  "SD_ACCESS_GARDEN",
  "SD_BIRTHYEAR",
  "SD_BORN_AUSTRIA",
  "SD_MIGR_BACKGR",
  "SD_BULA",
  "SD_EDU",
  "SD_EMPLSTATUS_FEB2020",
  "SD_HH_ADULTS",
  "SD_HH_CHILD18",
  "SD_HH_TEENS",
  "SD_HH_CHILD14",
  "SD_HH_CHILD5",
  "SD_HH_CHILD2",
  "SD_HH_OLDERFAM",
  "SD_HH_OUTERFAM",
  "SD_HH_PARTNER",
  "SD_HHINCOME_FEB2020",
  "SD_HOME_SQM",
  "SD_HOMEOWNER",
  "SD_WORKHOURS_FEB2020",
  health[1],
  life_sat[1],
  aff_neg_1[1],
  aff_neg_2[1],
  aff_neg_3[1],
  aff_neg_4[1],
  aff_neg_5[1],
  aff_neg_6[1],
  aff_pos_1[1],
  aff_pos_2[1],
  aff_pos_3[1],
  act_wrk[1],
  act_spo[1],
  act_frn[1],
  act_sho[1],
  act_pet[1],
  sat_dem[1],
  work_h[1],
  work_homeoff[1],
  hh_income[1],
  med_txt_kro[1],
  med_txt_sta[1],
  med_txt_pre[1],
  med_txt_oes[1],
  med_txt_kur[1],
  med_txt_slz[1],
  med_txt_son[1],
  med_vid_orf[1],
  med_vid_pri[1],
  soc_med_fb[1],
  soc_med_tw[1],
  soc_med_ig[1],
  soc_med_yt[1],
  soc_med_wa[1],
  soc_med_like_share[1],
  soc_med_post[1],
  soc_med_read[1],
  risk_prop[1],
  loc_cntrl_int_1[1],
  loc_cntrl_int_2[1],
  loc_cntrl_int_3[1],
  loc_cntrl_int_4[1]
  ))
```

Make data from wide to long.

```{r}
d_long <-
  d_wide %>% 
  pivot_longer(
    cols = health...W1:loc_cntrl_int_4...W9,
    names_to = "item",
    values_to = "value"
  ) %>% 
  separate(item, c("item", "wave"), sep = "\\.\\.\\.", extra = "merge") %>% 
  pivot_wider(names_from = "item", values_from = "value")
```

Filter waves without data

```{r eval=F}
d_long_test <- d_long %>%
  filter(!is.na(.[vars_varying]))
```

Recode values

```{r}
# recode such that higher values imply more strenght / align with wording
d_long %<>% 
  mutate_at(vars(med_txt_kro:med_vid_pri, health, sat_dem), 
            funs(recode(., `1` = 5L, `2` = 4L, `3` = 3L, `4` = 2L, `5` = 1L))) %>% 
  mutate_at(vars(loc_cntrl_int_1:loc_cntrl_int_4), 
            funs(recode(., `1` = 4L, `2` = 3L, `3` = 2L, `4` = 1L))) %>% 
  mutate_at(vars(born_aus_prnts), 
            funs(recode(., `3` = 0L, `2` = 2L, `1` = 1L)))

# recode inverted items
d_long %<>%
  mutate_at(vars(loc_cntrl_int_3, loc_cntrl_int_4), 
            funs(recode(., `1` = 4L, `2` = 3L, `3` = 2L, `4` = 1L)))

# recode other
d_long %<>%
  mutate(
    wave = gsub("W", "", .$wave) %>% as.integer(),
    id = as.integer(id)
  )
```

Order dataset. First waves, then id.

```{r}
d_long %<>%
  arrange(wave, id)
```

Create mean values for each person, and unique changes per wave

```{r}
d_long %<>% 
  group_by(id) %>% 
  mutate_at(vars(health:loc_cntrl_int_4), funs(b = mean(., na.rm = TRUE), w = . - mean(., na.rm = TRUE))) %>% 
  ungroup()
```

Create means for scales

```{r}
d_long %<>%
  mutate(aff_pos_m = rowMeans(select(., aff_pos_1 : aff_pos_3)),
         aff_neg_m = rowMeans(select(., aff_neg_1 : aff_neg_6)),
         loc_cntrl_int_m = rowMeans(select(., loc_cntrl_int_1 : loc_cntrl_int_4)))
```

Create vectors with names

```{r}
# create vector of names
vars_varying <- names(select(d_long, "health_b":"loc_cntrl_int_4_w"))
predictors_varying <- select(d_long, "health_b":"loc_cntrl_int_4_w") %>% 
  select(-starts_with("aff_"), -starts_with("life_")) %>% 
  names()
```

# inspect data

```{r}
nas <- sapply(d_long, function(y) sum(length(which(is.na(y))))) %>% 
  order()
```

Inspect participants with NA as ID

```{r}
d_long_filtered <- d_long %>% filter(is.na(id))
d_raw_filtered <- d_raw %>% filter(is.na(.$RESPID))
```

# measures
## life satisfaction

```{r}
model_life_sat <- lmer(life_sat ~ (1 | id) + (1 | wave), d_long)
fig_life_sat <- make_graph(model_life_sat, "Life satisfaction", ll = 0, ul = 10)
fig_life_sat
```

## positive affect

```{r}
model_aff_pos <- lmer(aff_pos_m ~ (1 | id) + (1 | wave), d_long)
fig_aff_pos <- make_graph(model_aff_pos, "Positive affect", ll = 1, ul = 5)
fig_aff_pos
```

## negative affect

```{r}
model_aff_neg <- lmer(aff_neg_m ~ (1 | id) + (1 | wave), d_long)
fig_aff_neg <- make_graph(model_aff_neg, "Negative affect", ll = 1, ul = 5)
fig_aff_neg
```

## Social media reading

```{r}
model_soc_med_read <- lmer(soc_med_read ~ (1 | id) + (1 | wave), d_long)
fig_soc_med_read <- make_graph(model_soc_med_read, "Social Media Reading", ll = 1, ul = 5)
fig_soc_med_read
```

## Social media liking & sharing

```{r}
model_soc_med_like_share <- lmer(soc_med_like_share ~ (1 | id) + (1 | wave), d_long)
fig_soc_med_like_share <- make_graph(model_soc_med_like_share, "Soc. Media Like & Share", ll = 1, ul = 5)
fig_soc_med_like_share
```

## Social media posting

```{r}
model_soc_med_post <- lmer(soc_med_post ~ (1 | id) + (1 | wave), d_long)
fig_soc_med_post <- make_graph(model_soc_med_post, "Social Media Posting", ll = 1, ul = 5)
fig_soc_med_post
```

## figure

Save everything as a figure

```{r}
fig_desc <- grid.arrange(fig_life_sat, fig_aff_pos, fig_aff_neg, 
                         fig_soc_med_read, fig_soc_med_like_share, fig_soc_med_post,
                         nrow = 2)
ggsave("figures/fig_descriptives.pdf", 
       width = 7, height = 5,
       plot = fig_desc)
```

# analyses
## life satisfaction

```{r}
model_1 <- lmer(life_sat ~ (1 | id) + soc_med_fb + soc_med_ig + soc_med_wa + soc_med_yt + soc_med_tw,
                data = d_long)
summary(model_1)
```

Now with diff. between within and between effects.

```{r}
model_2 <- lmer(life_sat ~ (1 | id) + (1 | wave) + soc_med_read_b + soc_med_read_w + soc_med_like_share_b + soc_med_like_share_w + soc_med_post_b + soc_med_post_w + soc_med_fb_b + soc_med_fb_w + soc_med_ig_b + soc_med_ig_w + soc_med_wa_b + soc_med_wa_w + soc_med_yt_b + soc_med_yt_w + soc_med_tw_b + soc_med_tw_w,
                data = d_long)
summary(model_2)
```

Now with control variables

```{r}
model_life_sat_lmer <- lmerTest::lmer(life_sat ~ (1 | id) + (1 | wave) + soc_med_read_b + soc_med_read_w + soc_med_like_share_b + soc_med_like_share_w + soc_med_post_b + soc_med_post_w + soc_med_fb_b + soc_med_fb_w + soc_med_ig_b + soc_med_ig_w + soc_med_wa_b + soc_med_wa_w + soc_med_yt_b + soc_med_yt_w + soc_med_tw_b + soc_med_tw_w + age + male + edu + born_aus + born_aus_prnts + res_vienna + edu + acc_bal + acc_gar + employment_fac
# + health_b + act_wrk_b + act_spo_b + act_frn_b + act_sho_b + act_pet_b + sat_dem_b + hh_income_b + med_txt_kro_b + med_txt_sta_b + med_txt_pre_b + med_txt_oes_b + med_txt_kur_b + med_txt_slz_b + med_txt_son_b + med_vid_orf_b + med_vid_pri_b + risk_prop_b + loc_cntrl_int_1_b + loc_cntrl_int_2_b + loc_cntrl_int_3_b + loc_cntrl_int_4_b + health_w + act_wrk_w + act_spo_w + act_frn_w + act_sho_w + act_pet_w + sat_dem_w + hh_income_w + med_txt_kro_w + med_txt_sta_w + med_txt_pre_w + med_txt_oes_w + med_txt_kur_w + med_txt_slz_w + med_txt_son_w + med_vid_orf_w + med_vid_pri_w + risk_prop_w + loc_cntrl_int_1_w + loc_cntrl_int_2_w + loc_cntrl_int_3_w + loc_cntrl_int_4_w
,
                data = d_long)
summary(model_life_sat_lmer)
```

## Positive Affect

```{r}
model_aff_pos_lmer <- lmerTest::lmer(aff_pos_m ~ (1 | id) + (1 | wave) + soc_med_read_b + soc_med_read_w  + soc_med_like_share_b + soc_med_like_share_w + soc_med_post_b + soc_med_post_w + soc_med_fb_b + soc_med_fb_w + soc_med_ig_b + soc_med_ig_w + soc_med_wa_b + soc_med_wa_w + soc_med_yt_b + soc_med_yt_w + soc_med_tw_b + soc_med_tw_w + age + male + edu + born_aus + born_aus_prnts + res_vienna + edu + acc_bal + acc_gar + employment_fac
# + health_b + act_wrk_b + act_spo_b + act_frn_b + act_sho_b + act_pet_b + sat_dem_b + hh_income_b + med_txt_kro_b + med_txt_sta_b + med_txt_pre_b + med_txt_oes_b + med_txt_kur_b + med_txt_slz_b + med_txt_son_b + med_vid_orf_b + med_vid_pri_b + risk_prop_b + loc_cntrl_int_1_b + loc_cntrl_int_2_b + loc_cntrl_int_3_b + loc_cntrl_int_4_b + health_w + act_wrk_w + act_spo_w + act_frn_w + act_sho_w + act_pet_w + sat_dem_w + hh_income_w + med_txt_kro_w + med_txt_sta_w + med_txt_pre_w + med_txt_oes_w + med_txt_kur_w + med_txt_slz_w + med_txt_son_w + med_vid_orf_w + med_vid_pri_w + risk_prop_w + loc_cntrl_int_1_w + loc_cntrl_int_2_w + loc_cntrl_int_3_w + loc_cntrl_int_4_w
,
                data = d_long)
summary(model_aff_pos_lmer)
```

## Negative Affect

```{r}
model_aff_neg_lmer <- lmerTest::lmer(aff_neg_m ~ (1 | id) + (1 | wave) + soc_med_read_b + soc_med_read_w + soc_med_like_share_b + soc_med_like_share_w + soc_med_post_b + soc_med_post_w + soc_med_fb_b + soc_med_fb_w + soc_med_ig_b + soc_med_ig_w + soc_med_wa_b + soc_med_wa_w + soc_med_yt_b + soc_med_yt_w + soc_med_tw_b + soc_med_tw_w + age + male + edu + born_aus + born_aus_prnts + res_vienna + edu + acc_bal + acc_gar + employment_fac
# + health_b + act_wrk_b + act_spo_b + act_frn_b + act_sho_b + act_pet_b + sat_dem_b + hh_income_b + med_txt_kro_b + med_txt_sta_b + med_txt_pre_b + med_txt_oes_b + med_txt_kur_b + med_txt_slz_b + med_txt_son_b + med_vid_orf_b + med_vid_pri_b + risk_prop_b + loc_cntrl_int_1_b + loc_cntrl_int_2_b + loc_cntrl_int_3_b + loc_cntrl_int_4_b + health_w + act_wrk_w + act_spo_w + act_frn_w + act_sho_w + act_pet_w + sat_dem_w + hh_income_w + med_txt_kro_w + med_txt_sta_w + med_txt_pre_w + med_txt_oes_w + med_txt_kur_w + med_txt_slz_w + med_txt_son_w + med_vid_orf_w + med_vid_pri_w + risk_prop_w + loc_cntrl_int_1_w + loc_cntrl_int_2_w + loc_cntrl_int_3_w + loc_cntrl_int_4_w
,
                data = d_long)
summary(model_aff_neg_lmer)
```

## figure

Active vs. Passive

```{r}
# get data
dat_fig_results <- 
  broom.mixed::tidy(model_aff_neg_lmer, conf.int = T) %>% 
  mutate(dv = "aff_neg") %>% 
  rbind(
    broom.mixed::tidy(model_aff_pos_lmer, conf.int = T) %>% 
    mutate(dv = "aff_pos")
  ) %>% 
  rbind(
    broom.mixed::tidy(model_life_sat_lmer, conf.int = T) %>% 
    mutate(dv = "life_sat")
  ) %>% 
  mutate(
    type = ifelse(grepl(".*_b.*", .$term), "between", "within"),
    iv = gsub("_(w|b)\\>", "", .$term)
         ) %>%
  mutate(
    type = factor(.$type, levels = c("within", "between")),
    dv = factor(.$dv, levels = c("life_sat", "aff_pos", "aff_neg"),
                labels = c("Life satisfaction", "Positive affect", "Negative affect"))
  ) %>% 
  select(dv, iv, type, estimate, conf.low, conf.high, p.value)

# select Social Media type of activity
dat_fig_results_activity <- dat_fig_results %>%
  filter(iv %in% c("soc_med_read", "soc_med_like_share", "soc_med_post")) %>% 
  mutate(
    iv = factor(.$iv, levels = c("soc_med_post", "soc_med_like_share", "soc_med_read"), labels = c("Posting", "Liking & Sharing", "Reading"))
  )

# make figure
fig_results_activity <- make_graph_res(dat_fig_results_activity)

# safe figure
ggsave("figures/fig_results_activity.pdf", 
       width = 7, height = 4,
       plot = fig_results_activity)

# select social media channel
dat_fig_results_channel <- dat_fig_results %>%
  filter(iv %in% c("soc_med_fb", "soc_med_ig", "soc_med_wa", "soc_med_yt", "soc_med_tw")) %>% 
  mutate(iv = factor(.$iv, levels = c("soc_med_yt", "soc_med_wa", "soc_med_ig", "soc_med_tw", "soc_med_fb"), labels = c("YouTube", "WhatsApp", "Instagram", "Twitter", "Facebook")))

# make figure
fig_results_channel <- make_graph_res(dat_fig_results_channel)

# save figure
ggsave("figures/fig_results_channel.pdf", 
       width = 7, height = 4,
       plot = fig_results_channel)
```

```{r}
save.image("data/workspace.RData")
```

### 2021-11-03 index.Rmd
- set-up of repo

---
title: "No effets of COVID-related social media use and well-being"
---

On this website, you can find the companion material for the paper _No effets of COVID-related social media use and well-being_.

You can find the following information here:

- The paper, which you can read [in your browser](manuscript.html) or download as [pdf](manuscript.pdf). 
- The detailed code and summaries of the [analyses](analyses.html)
- A list of the [items translated into English](items_english.html)
- A list of the [original items in German](items_german.html)
- The [preregistration](preregistration.html) 
- A list where I needed to [deviate](preregistration_changes.Rmd) from the preregistration.

This website is produced directly from the project's [github repository](https://github.com/XMtRa/Austrian_Corona_Panel_Project). You can download the material and the data, file an issue, or submit a pull request.

### 2021-11-03 items_english.rmd
- set-up of repo

---
title: "Items: English"
output:
  # pdf_document:
  #   latex_engine: xelatex
  #   toc: true
  html_document:
    toc: TRUE
    toc_float:
      collapsed: FALSE
      toc_depth: 3
---

Here you can find a quick overview of all items that were measured.

## Life Satisfaction

"Taken everything together, how satisfied are you currently with your life?"

(0 = extremely unsatisfied, 10 = extremely satisfied)

## Positive Affect (scale)

"In the last week, how often did you feel"

 - a: ... calm and relaxed
 - e: ... happy
 - i: ... full of energy

(never [1]; on some days [2]; several times per week [3]; almost every day [4]; daily [5])

## Negative Affect (scale)

"In the last week, how often did you feel"

 - b: ... lonely
 - c: ... aggravated
 - d: ... so depressed, that nothing could lift you up
 - f: ... very nervous
 - g: ... anxious
 - h: ... glum and sad

(never [1]; on some days [2]; several times per week [3]; almost every day [4]; daily [5])

## Social media use: Reading

"Wie oft haben Sie letzte Woche die folgenden Aktivitäten in Sozialen Netzwerken gesetzt?"

 - "Postings anderer lesen, in denen es um Inhalte zum Coronavirus geht"

(1 = Mehrmals täglich; 2 = Einmal täglich; 3 = Mehrmals in der Woche; 4 = Einmal in der Woche; 5 = Gar nicht/nie)

## Social media use: Liking & Sharing

"Wie oft haben Sie letzte Woche die folgenden Aktivitäten in Sozialen Netzwerken gesetzt?"

 - "Bei Postings anderer zum Coronavirus auf den „Gefällt mir“ Button drücken, „teilen“ oder „retweeten“"

(nie [1]; an manchen Tagen [2]; mehrmals die Woche [3]; beinahe jeden Tag [4]; täglich [5])

## Social media use: Active Postings

"Wie oft haben Sie letzte Woche die folgenden Aktivitäten in Sozialen Netzwerken gesetzt?"

 - "Selbst Postings zum Coronavirus in Sozialen Netzwerken verfasst"

(nie [1]; an manchen Tagen [2]; mehrmals die Woche [3]; beinahe jeden Tag [4]; täglich [5])

## Social media use: Specific platforms

"Wie oft in der letzten Woche haben Sie Informationen zur Corona-Krise in den folgenden Sozialen Medien verfolgt?"

 - Facebook 
 - Twitter
 - Instagram
 - Youtube
 - WhatsApp

(1 = Mehrmals täglich; 2 = Einmal täglich; 3 = Mehrmals in der Woche; 4 = Einmal in der Woche; 5 = Gar nicht/nie)

## Gender

"Welches Geschlecht haben Sie?"

(a: Männlich; b Weiblich; c: Divers)

## Age

"In welchem Jahr wurden Sie geboren?"

(Jahr)

## Education

"Was ist der höchste Schul- oder Bildungsabschluss, den Sie erreicht haben?"

 - a. Volksschule oder weniger
 - b. Hauptschule oder AHS Unterstufe
 - c. Polytechnikum, BMS (Fachschule, z.B. HASCH)
 - d. Lehre, Berufsschule
 - e. AHS mit Matura
 - f. BHS mit Matura (z.B. HTL, HAK, HBLA etc.) 
 - g. Hochschulverwandte Lehranstalt oder Kolleg 
 - h. Bachelor
 - i. Magister / Master / Diplomingenieur / Fachhochschule 
 - j. Doktor / PhD

## Austria country of birth

"Ist Ihr Geburtsland Österreich?"

  - Ja
  - Nein

## Austria country of birth of parents

 "Wurde einer oder beide Ihrer Elternteile nicht in Österreich geboren?"

 - a. Ja, einer
 - b. Ja, beide
 - c. Nein

## News consumption: Text-based

"Nachstehend finden Sie eine Liste mit verschiedenen Tageszeitungen. Wie oft haben Sie sich letzte Woche über das politische Geschehen in der jeweiligen Zeitung (im Internet oder gedruckt) informiert?"

 - Kronen Zeitung oder www.krone.at
 - Der Standard oder derstandard.at
 - Die Presse oder diepresse.com
 - Österreich oder oe24.at
 - Kurier oder kurier.at
 - Salzburger Nachrichten oder salzburg.at
 - Sonstige österreichische Tageszeitungen

## News consumption: Video-based

"Nachstehend finden Sie eine Liste mit verschiedenen Nachrichtensendungen im Fernsehen. Wie oft haben Sie sich letzte Woche über das politische Geschehen in den folgenden Nachrichtensendungen informiert?"

 - "Nachrichtensendungen im ORF (z.B. Zeit im Bild um 19:30h oder ZiB 2 um 22:00h)"
 - "Nachrichtensendungen im Privatfernsehen (z.B. ATV aktuell, Puls 4/ Pro7/ Sat1 Austria)"

(1 = Mehrmals täglich; 2 = Einmal täglich; 3 = Mehrmals in der Woche; 4 = Einmal in der Woche; 5 = Gar nicht/nie)

## Residency Vienna

"In welchem Bundesland leben Sie aktuell?"

 - a. Vorarlberg
 - b. Tirol
 - c. Salzburg
 - d. Steiermark
 - e. Kärnten
 - f. Oberösterreich
 - g. Niederösterreich
 - h. Wien
 - i. Burgenland

(Will be transformed to Vienna: yes/no)

## Household size

"Bitte rechnen Sie hier sich selbst wieder mit ein. Wenn es eine Personengruppe nicht in Ihrem Haushalt gibt, tragen Sie bitte „0“ ein.

 - a. ...Erwachsene über 18 Jahre:
 - b. ...Teenager von 15-17 Jahren:
 - c. ...Kinder von 6-14 Jahren:
 - d. ...Kinder von 3-5 Jahren:
 - e. ...Kinder von 0-2 Jahren:

(Achtung, wenn Anzahl Erwachsene missing, 1 eintragen, da vorher Filterfrage)

## Health

"Kommen wir jetzt zu ein paar Fragen zu Ihrer Gesundheit. Wie schätzen Sie Ihren allgemeinen Gesundheitszustand ein? Würden Sie sagen, er ist..."

 - a. 1 = Sehr gut
 - b. 2 = Gut
 - c. 3 = Mittelmäßig
 - d. 4 = Schlecht
 - e. 5 = Sehr schlecht

## Living space

"Wie viele Quadratmeter hat Ihre Wohnung?"

(Einfach-Nennung, Zahleneingabe, min. 0 – max. 600, freiwillig)

## Access garden

"Haben Sie Zugang zu einer privaten Freifläche (Balkon, Garten)?

 - (Ja, Garten)

## Access balcony

"Haben Sie Zugang zu einer privaten Freifläche (Balkon, Garten)?"

 - (Ja, Balkon)

## Employment

Wenn Sie sich selbst zuordnen: Im Februar 2020, welcher der folgenden Gruppen gehören Sie vorwiegend an? (Einfach-Nennung)

 - a. Unselbstständig erwerbstätig (auch Lehrling), Privatwirtschaft
 - b. Unselbstständig erwerbstätig (auch Lehrling), Öffentlicher Dienst
 - c. Selbstständig erwerbstätig
 - d. Arbeitslos
 - e. Pensionist*in
 - f. Ausschließlich haushaltsführend
 - g. Schüler/in, Student/in, Bildungskarenz
 - h. Dauerhaft arbeitsunfähig
 - i. Präsenz-/Zivildiener
 - j. In Elternkarenz

## Work hours per week

"Wie viele Stunden (inklusive Überstunden) arbeiten Sie in Ihrer Haupttätigkeit jetzt pro Woche? Wenn Sie es nicht genau wissen, genügt Ihre beste Schätzung."

## Home-Office Yes/No

"Hat sich in Ihrer beruflichen Situation aufgrund der Corona-Krise etwas geändert? Bitte wählen Sie alles Zutreffende aus."

 - a. Ich bin im Home-Office

## Household income

"Wie viel Geld steht Ihrem Haushalt aktuell zur Verfügung (Netto-Haushaltseinkommen inklusive Sozialleistungen, Rente usw.)? Bitte rechnen Sie Überstunden mit ein, nicht aber den 13./14. Bezug."

 - a. weniger als 1.100 Euro
 - b. 1.100 bis unter 1.500 Euro 
 - c. 1.500 bis unter 1.800 Euro 
 - d. 1.800 bis unter 2.200 Euro 
 - e. 2.200 bis unter 2.700 Euro 
 - f. 2.700 bis unter 3.100 Euro 
 - g. 3.100 bis unter 3.700 Euro 
 - h. 3.700 bis unter 4.300 Euro 
 - i. 4.300 bis unter 5.500 Euro 
 - j. 5.500 Euro und mehr

## Outdoor activities

"Haben Sie wegen folgenden Gründen in der vergangenen Woche Ihr Zuhause verlassen?"

 - a: Arbeit
 - b: Sport
 - c: Freunde oder Verwandte treffen
 - e: Essen einkaufen
 - f: Haustiere ausführen

(nie [1]; an manchen Tagen [2]; mehrmals die Woche [3]; beinahe jeden Tag [4]; täglich [5])

## Satisfaction Democracy

"Wie zufrieden oder unzufrieden sind Sie alles in allem mit der Demokratie, so wie sie derzeit in Österreich funktioniert?"

 - a. 1 = Sehr zufrieden
 - b. 2 = Eher zufrieden
 - c. 3 = Teils-teils
 - d. 4 = Eher unzufrieden 
 - e. 5 = Sehr unzufrieden

## Risk-taking

"Wie sehr sind Sie im Allgemeinen dazu bereit, Risiken einzugehen?"

(0: gar nicht bereit; 10: sehr bereit)

## Locus of Control

 - "Ich habe mein Leben selbst in der Hand."
 - "Wenn ich mich anstrenge, werde ich auch Erfolg haben."
 - "Egal ob privat oder im Beruf: Mein Leben wird zum großen Teil von anderen bestimmt."
 - "Meine Pläne werden oft vom Schicksal durchkreuzt."

(1 = Trifft voll und ganz zu; 2 = Trifft etwas zu; 3 = Trifft wenig zu; 4 = Trifft nicht zu)

### 2021-11-03 items_german.Rmd
- set-up of repo

--- 2021-11-03 items_english.rmd
@@
 ---
-title: "Items: English"
+title: "Items: German"
 output:
-  # pdf_document:
-  #   latex_engine: xelatex
-  #   toc: true
   html_document:
@@
 
-Here you can find a quick overview of all items that were measured.
+Here you can find a quick overview of all items that were measures.
 
@@
 
-"Taken everything together, how satisfied are you currently with your life?"
+"Alles in allem betrachtet, wie zufrieden sind Sie derzeit mit Ihrem Leben?"
 
-(0 = extremely unsatisfied, 10 = extremely satisfied)
+(0 = äußerst unzufrieden, 10 = äußerst zufrieden)
 
@@
 
-"In the last week, how often did you feel"
+"Wie often waren Sie in der letzten Woche"
 
- - a: ... calm and relaxed
- - e: ... happy
- - i: ... full of energy
+ - a: ... ruhig und gelassen
+ - e: ... glücklich
+ - i: ... voller Energie
 
-(never [1]; on some days [2]; several times per week [3]; almost every day [4]; daily [5])
+(nie [1]; an manchen Tagen [2]; mehrmals die Woche [3]; beinahe jeden Tag [4]; täglich [5])
 
@@
 
-"In the last week, how often did you feel"
+"Wie often waren Sie in der letzten Woche"
 
- - b: ... lonely
- - c: ... aggravated
- - d: ... so depressed, that nothing could lift you up
- - f: ... very nervous
- - g: ... anxious
- - h: ... glum and sad
+ - b: ... einsam
+ - c: ... ärgerlich
+ - d: ... so niedergeschlagen, dass Sie nichts aufheitern konnte
+ - f: ... sehr nervös
+ - g: ... ängstlich
+ - h: ... bedrückt und traurig
 
-(never [1]; on some days [2]; several times per week [3]; almost every day [4]; daily [5])
+(nie [1]; an manchen Tagen [2]; mehrmals die Woche [3]; beinahe jeden Tag [4]; täglich [5])
 

### 2021-11-03 manuscript.Rmd
- set-up of repo

---
title             : "No effects of COVID-related social media use on well-being"
shorttitle        : "COVID-related social media use and well-being"

author: 
  - name          : "Blinded for review"
    affiliation   : "1"
    corresponding : yes
    address       : "Blinded for review"
    email         : "Blinded for review"

affiliation:
  - id            : "1"
    institution   : "Blinded for review"

authornote: |
  Blinded for review.

abstract          : "In times of crisis such as the Corona pandemic, it is important that citizens stay informed about recent events, the latest political decisions, or mandatory protection measures. As a result, to attain relevant information many people use various types of media, and increasingly social media. However, because social media are particularly engaging, some find it hard to disconnect and cannot stop 'doomscrolling'. In this preregistered study, I investigate whether using social media for COVID-related topics might put personal well-being at risk. To this end, I analyze data from the Austrian Corona Panel Project, which consists of 24 waves with overall 3,018 participants. The data were analzed using random effects cross lagged panel models, controlling for several stable and varying covariates. Results showed that COVID-related social media use did not meaningfully affect several types of well-being, including life satisfaction, positive affect, and negative effect. This pertains to both passive and active social media use, and all the prominent channels such as Facebook, WhatsApp, or YouTube. As a result, this study suggests fears that social media use during times of crisis might be detrimental for well-being can be put to rest."
keywords          : "COVID, well-being, affect, life satisfaction, social media use, news use, random effects within between model, panel study, longitudinal"

bibliography      : 
  - "bibliography/r-references.bib"
  - "bibliography/references.bib"

floatsintext      : yes
figurelist        : no
tablelist         : no
footnotelist      : no
linenumbers       : yes
mask              : no
draft             : no

documentclass     : "apa6"
classoption       : "man"
# output            : papaja::apa6_pdf
# output            : papaja::apa6_word
header-includes:
  - \setlength{\parskip}{0em}
  - \raggedbottom
  - \note{\clearpage}
---

```{r analysis-preferences, include=F, cache=F}
# Seed for random number generation
set.seed(42)
knitr::opts_chunk$set(cache.extra = knitr::rand_seed,
                      cache=T,
                      echo=F
                      # , fig.pos = 'h'
                      )
```

```{r setup, include=F, cache=F}

# make project reproducible
# devtools::install_github("rstudio/renv")  # install renv
# renv::init()  # initialize renv

# install necessary packages
packages <- c("devtools", "english", "ggplot2", "kableExtra", 
              "lavaan", "lme4", "lmerTest",
              "magrittr", "semPlot", "tidyverse")
# renv::install(packages)
# renv::install("tdienlin/td@v.0.0.2.5")  # uncomment to install
# renv::install("crsh/papaja@v0.1.0.9997")  # uncomment to install
# renv::install("crsh/papaja@devel")  # uncomment to install

# load packages
lapply(c(packages, "papaja", "td"), library, character.only = TRUE)

# save snapshot of installed packages
# renv::snapshot()

# load workspace created from analyses.rmd to use results
load("data/workspace.RData")

# export references
# r_refs(file = "bibliography/r-references.bib")
```

```{r custom-funcs, include=F}
# Get mean, max, and min values

get_specs <- function(model){
  dat <- coefficients(model)$wave
  specs <- list(
    min = min(dat),
    max = max(dat),
    mean = mean(dat$`(Intercept)`),
    sd = attr(VarCorr(model), "sc")
  )
}
```

During the COVID pandemic, 
<!-- and especially in the beginning,  -->
several events unfolded in quick succession. 
How dangerous is the virus? Is it spreading in my region? How is it transmitted, and how can I protect myself?
Because for many it was (and still is) a matter of life or death, citizens had to stay informed regarding the latest developments.
Governments around the world implemented safety measures, from wearing masks or keeping physical distance, to complete lockdowns.
In this extraordinary situation, people hence used media excessively, and especially social media were at an all time high [@statistaAverageDailyTime2021].
A new phenomenon emerged, "termed doomscrolling":
People could not stop using social media to attain COVID-related news.

Several people reported that they were glued to their screens and could not pursue other relevant activities such as working, taking a break, or even care-work.
Increasingly, it was asked whether such an increase in social media use could still be considered useful and adaptive, or whether it created an additional and new psychological danger for the users' mental health [@kleinDarklySoothingCompulsion2021].
A study with 6,233 people from Germany found that "[f]requency, duration and diversity of media exposure were positively associated with more symptoms of depression and unspecific and COVID specific anxiety" [@bendauAssociationsCOVID19Related2021].

As a result, with this study I want to build on this research and investigate the question whether COVID-related social media use during the pandemic affected the users' well-being.
To this end I analyze a large-scale panel study from the Austrian Corona Panel Project [@kittelAustrianCoronaPanel2020]. 
The panel consists of 24 waves and an overall sample size of `r nrow(d_wide)`, with at least 1,500 participants per wave, and it is representative of the Austrian population.
The panel study collected a large number of variables.
Because we can therefore control for many confounding third variables, both stable and varying, together with the longitudinal design this creates a unique opportunity to analyze causality.

## Defining Well-being and Media Use

The underlying theories that guided the selection of variables for this study are the two-continua model of mental health [@greenspoonIntegrationSubjectiveWellbeing2001] and the hierarchical taxonomy of computer-mediated medation [@meierComputerMediatedCommunicationSocial2020].
According to the two-continua model of mental health, mental health consists of two dimensions, psychopathology and well-being.
Because the aim of this study is better to understand typical users and everyday contexts, my focus will be on well-being.
Well-being can be differentiated into subjective and psychological well-being [@dienerAdvancesOpenQuestions2018].
Whereas subjective well-being emphasizes hedonic aspects such as a happiness and joy, psychological well-being focuses on eudaimonic aspects including fulfillment and meaning.
Subjective well-being is primarily about achieving positive affect and avoiding negative affect.
In my view, life satisfaction is a meta concept above both psychological and subjective well-being, representing a meta-appraisal of one's life.
Notably, life satisfaction is very stable and fluctuates only little, whereas it's the exact opposite for affect [@dienlinImpactDigitalTechnology2020].
<!-- Everyone can feel bad on any given day, but whether or not one feels generally comfortable in one's life rather changes across years. -->
To capture well-being in this study, I will thus build on life satisfaction, positive affect, and negative affect.
Together, this should provide an encompassing perspective on media effects.

The hierarchical taxonomy of computer-mediated communication differentiates six levels of how people engage with digital technology:
first, the device (e.g., smartphone); second, the type of application (e.g., social networking site); third, the branded application (e.g., Twitter); fourth, the feature (e.g., status post); fifth, the interaction (e.g., one-to-many); sixth, the message (e.g., content) [@meierComputerMediatedCommunicationSocial2020].
Whereas the first four levels are focused on the channel, the last two on the type of communication.
To measure social media use for consumption of specific news, I here employ both the channel and the communication perspective.
First, I will analyze how using the most prominent branded applications affect well-being, and whether this effect changes across applications.
The branded applications investigated here are Facebook, Twitter, Instagram, WhatsApp, and YouTube.
But because adopting only this position would be both too narrow and too general, I will secondly also investigate how different types of interaction affect well-being.
Specifically, I will differentiate between active and passive use.
I will distinguish (a) _reading_ COVID-related social media use (passive), _posting_ content regarding COVID (active), and _liking and sharing_ COVID-related posts by others (both active and passive).
Worth noting, this study is not about _general_ social media during times of COVID, but on social media use focused on COVID-related news and interactions.
For example, posting about the pandemic or retweeting COVID-related posts.

## Effects of Social Media on Well-Being

In their study on the relations between media use and mental health during the pandemic, @bendauAssociationsCOVID19Related2021 found that people who used social media as a primary source of information reported on average "significantly more unspecific anxiety and depression [] and significantly more specific COVID-related anxiety symptoms" (p. 288).
Hence, this might hint at potential negative effects of social media as news use---but note that this finding comes from a between-person relation stemming from a cross-sectional data.
Hence, we don't know whether the differences in mental health and well-being are due to social media use or other third variables, such as age or education.

In general, so far there is only little research on news-related social media use and well-being.
On the other hand, the question of whether and how _general_ social media use---or other, more specific types such as active or passive use---affect well-being is well-researched.
A meta review (i.e., an analysis of meta-analyses), found that the relation between social media use and well-being is likely in the negative spectrum---but very small, potentially too small to matter [@meierComputerMediatedCommunicationSocial2020].
This is well aligned with several new studies that employed the most advanced methods [@orbenSocialMediaEnduring2019; @przybylskiDoesTakingShort2021a; @schemerImpactInternetSocial2021; @kerestesAdolescentsOnlineSocial2020].
For example, @beyensEffectSocialMedia2020 reported that although for some users (roughly one quarter) the effects were negative, for almost the same amount of users the effects were positive, while for the majority they were neutral.
At all events, in general most effects are somewhere between trivial and small.

What determines a trivial or a small effect is a difficult question (see below).
If we refer to standardized effect sizes, according to Cohen [@cohenPowerPrimer1992] small effect sizes start at _r_ = .10.
However, most meta-analyses find effect sizes below that threshold [@huangTimeSpentSocial2017; @meierComputerMediatedCommunicationSocial2020].
As a result, I think it's most convincing to expect trivial to small effects also in the case of COVID-related social media use.

From a theoretical perspective, why could COVID-related social media use be detrimental?
Above everything, one can reasonably assume a _direct_ negative effect on well-being, especially on positive or negative effect.
Dangers, inequalities, corruption -- these were the headlines across many countries worldwide.
If one learns about such things, the initial reaction might be shock, fear, or dismay.
Repeatedly consuming such news might be depressing, perhaps even changing some general perspective on life.
So, just like being hit by a hammer hurts and we don't need any "mediating mechanism", this could be the case here as well.

That said, because not all news were negative, because many people showed solidarity and compassion, 
<!-- or maybe also because it was possible quickly to find an effective vaccine,  -->
there were also positive and potentially uplifting news.
However, in light of a worldwide pandemic with millions of deaths, the negative direct effect seems more plausible.

There could be also indirect effects.
When doomscrolling, users are captivated to such an extent that they cannot stop using social media.
For example, during the pandemic social media uses was at an all-time high in the US [@statistaAverageDailyTime2021].
As has been expressed by many before, it is most likely that moderate social media use is not detrimental [@orbenTeenagersScreensSocial2020].
However, overuse is more critical, and several studies have showed more pronounced negative effects for extreme users [@przybylskiLargescaleTestGoldilocks2017].
So if a society collectively overuses social media, there is potential for negative effects.
Overuse likely impairs well-being if it replaces other meaningful or functional activities such as meeting others, working, actively relaxing, or exercising.
If overuse replaces such activities it's reasonable to assume that it's also detrimental.

On the other hand, one can make the case that overuse can be beneficial _in times of a pandemic_, even if it's mainly COVID-related.
Exchanging COVID-related messages with friends via WhatsApp might replace the in-person contact one would have otherwise, but which is logically impossible at that time.
At times where meaningful and functional activities are explicitly or implicitly prohibited, using social media to exchange about COVID might not be the worst idea.
In fact, given that nowadays a large number of experts, scientists, and politicians converse directly on social media, one can get first-hand information on the current developments.
On the other hand, there is of course also much disinformation, and "bingeing" on COVID fake news might also pose risks for impaired well-being.

Together, the strongest argument seems to be that the effects of social media on well-being are, on average, generally small at best.
Because this study looks at only _one part_ of social media use---namely, COVID-related interactions---it is even more focused, and the overall potential of the effects should diminish even further.
Whether or not using social media for COVID-related aspects is detrimental during a pandemic is also not entirely clear.
<!-- At least, there does not seem to be a case where we should expect a clear deviation to the negative, despite anecdotes of doomscrolling. -->
Therefore, I expect to find that the effects of COVID-related news use on well-being will be not be meaningful or practically relevant.

> Hypothesis: The within-person effects of all types of COVID-related social media use on all types of well-being indicators---while controlling for several stable and varying covariates such as sociodemographic variables and psychological dispositions---will be trivial.

# Current Study
## Smallest Effect Size of Interest

Testing this hypothesis, however, is not trivial.
First, in contrast to most hypothesis typically posited in the social sciences, it explicates an effect size, a so-called smallest effect size of interest (SESOI).
Effectively testing this hypothesis hence necessitates to define what's considered a "trivial effect size" and what's not.
Above I referred to standardized effect sizes.
However, standardized effect sizes should only be a first step toward evaluating an effect's relevance [@baguleyStandardizedSimpleEffect2009].
Standardized effect sizes are determined by a sample's variance.[^cohensd]
However, this is problematic: The question of whether or not social media use affects me personally in a relevant way should not depend on the variance in the sample in which my data were collected.
Instead, it should depend on absolute criteria.
What could be a minimally interesting, a nontrivial effect?
I suggest the following SESOI for this research question:

[^cohensd]: Consider the effect size Cohen's _d_: The mean's of the two groups that are to be compared are subtracted from one another and then divided by the sample's standard deviation [@cohenPowerPrimer1992]. Hence, if there is more deviation/variance in a sample, the effect size decreases, even if the difference of the group's means stays the same.

If a heavy user of COVID-related social media news stops using social media altogether, this should have a noticeable impact on their overall well-being.

What would this mean practically and how could it be operationalized?
In this study, COVID-related social media use was measured on a 5-point scale, ranging from 1 = _never_ to 5 = _several times a day_. 
Thus, the example from above would imply that a change of five units in social media use should correspond to a noticeable change in well-being.
But what's a noticeable change in well-being?
According to @normanInterpretationChangesHealthrelated2003, people can reliably differentiate between _seven_ levels of satisfaction with health.
So we could state that a five unit change in social media use should result in a one unit change in satisfaction.
Statistically, in a regression, _b_ estimates the change in the dependent variable if the independent variable increases by one point.
Transferred to our example, we would hence expect a _b_ of 1 / 5 = .20.

In this study, life satisfaction was measured with 11 units, hence more than people can reliably detect. 
Hence, we would expect a 11 / 7 * .20 change, that is a _b_ of at least .31.
For affect, which was measured on a 5-point scale, our SESOI would be 5 / 7 * .2 = .14.
In order not to exaggerate precision, these numbers will be rounded. 
Plus, because we are agnostic as to whether the effects are positively or negatively nontrivial, this leads to an indifference region ranging from _b_ = -.30 to _b_ = .30 (and _b_ = -.15 to _b_ = .15 for affect).

## Causality

The hypothesis explicitly states a causal effect.
In non-experimental designs, causality can be approximated using longitudinal designs.
Using longitudinal designs alone, however, is not sufficient for correct causal statements.
In addition, it is crucial also to control for third variables, and importantly also for _varying_ third variables.

For example, imagine that a person suddenly starts using social media much more than usual, and then at the same time also become less satisfied with their lives.
After some time, use recedes again, whereas life satisfaction returns to prior levels.
If this happened to several people at the same time, in a longitudinal study we could then find a causal effect of social media use on life satisfaction.
However, it could also be the case that during the study there was a major exogenous event (say, a pandemic) and a large part of the working population lost their jobs.
Hence, the causal effect reported above was confounded, because in reality it was the pandemic that caused both social media use to rise and their life satisfaction to plummet.

Thus, only if we can control for _all_ potential varying third variables, we can correctly estimate causality without any bias [@rohrerThinkingClearlyCorrelations2018].
Obviously, we can never be entirely sure to have included all varying covariates, which makes absolute statements regarding causality impossible.
However, when controlling for many varying variables, we can be much more certain that we measured causality correctly.
The aim should still be to collect as many relevant varying and nonvarying third variables as possible, while knowing that absolute certainty can rarely be reached.

When looking for suitable control candidates, ideally we find variables that affect both media use and well-being, because controlling for these factors will isolate the actual effect of social media use on well-being.
We can also control for variables that affect _only_ well-being. 
However, in doing so nothing is gained or lost because the effects of social media use would remain exactly the same [@klinePrinciplesPracticeStructural2016].
Crucially, when determining the general causal effect of social media use we should _not_ control for mediating variables [@rohrerThinkingClearlyCorrelations2018].
Doing so would bias and distort our assessment of the role of social media use.

In this study, I will hence control for the following variables, which either have been shown or a likely to affect both social media use and well-being.
(I'll additionally include variables that likely affect only well-being, also to obtain a comparison benchmark for social media effects):
Gender, age, education, Austria country of birth, Austria country of birth of parents, text-based news consumption, video-based news consumption, residency Vienna, household size, health, living space, access to garden, access to balcony, employment, work hours per week, being in home-office, household income, outdoor activities, satisfaction with democracy, disposition to take risks, and locus of control.
I will not control for variables such as trust in institutions or media, because these variables are likely influenced by social media use.

[Within vs. between logic]

[Time Frame / REWB logic]

# Method

This section describes the preregistration and how I determined the sample size, data exclusions, the analyses, and all measures in the study.

## Preregistration

The hypotheses, the sample, the measures, the analyses, and the inference criteria (SESOI, p-value) were preregistered on the Open Science Framework. 
The (anonymous) preregistration can be accessed here: https://osf.io/87b24/?view_only=b2289b6fec214fa88ee75a18d45c18f3.
Because in this study I analyze data from an already existing large scale data set, the Austrian Corona Panel Project, all of these steps were preregistered _prior_ to accessing the data.
The preregistration was designed on the basis of the panel documentation online [@kittelAustrianCoronaPanel2020].
In some cases the analyses could not be executed as originally planned, because some properties of the variables only became apparent upon data analysis.
The most relevant deviations are reported below, and a complete list of all changes can be found online.

## Sample

The data come from the Austrian Corona Panel Project [@kittelAustrianCoronaPanel2021].
The study was conducted between March 2020 and October 2021.
The contains 26 waves, and at the time of writing, the first 24 waves were available for download. 
Each wave consists of at least 1,500 respondents.
The overall sample size was _N_ = `r d_raw %>% nrow()`, and `r d_long %>% nrow()` observations were collected.
Panel mortality was compensated through a continuous reacquistion of new participants. 
All respondents needed to have access to the internet (via computer or mobile devices such as smartphones or tablets).
They were sampled from a pre-existing online access panel provided by Marketagent, Austria.
Respondents were asked and incentivized with 180 credit points to participate in each wave of the panel.

The sample was representative of the Austrian population in terms of age, gender, region/state, municipality size, and educational level. In order to participate in the study, the respondents needed to be Austrian residents and had to be at least 14 years old. 
The average age was `r mean(2021 - d_wide$year_birth, na.rm = T) %>% round(0)`, `r mean(d_wide$male, na.rm = T) %>% round(., 2) * 100` percent were male, `r d_wide %>% filter(.$edu_fac %in% c("Hochschulverwandte Lehranstalt oder Kolleg","Bachelor", "Magister / Master / Diplomingenieur / Fachhochschule", "Doktor / PhD")) %>% nrow() / sum(table(d_wide$edu_fac)) %>% divide_by(100)` percent had a University degree, and `r table(d_wide$employment_fac)["Arbeitslos"] / (table(d_wide$employment_fac) %>% sum()) %>% unname() %>% divide_by(100) %>% round(., 0)` percent were currently unemployed.
<!-- The study received IRB approval from the University of Vienna. -->

Because the data were analyzed post hoc, no sample size planning on the basis of a priori power analyses was conducted.
Because the sample is very large, it is well-equipped reliably to detect also small effects, rendering exact post hoc power analysis unnecessary.
And also because such large sample easily generate significant p-values for very small effects, this study builds on a smallest effect size of interest.
We will use the interval testing approach as proposed by @dienesUsingBayesGet2014.
On the basis of our SESOI, we will define a null region.
For well-being, the null region will be between _b_ = -.30 and _b_ = .30.
If the 95% confidence interval lies completely within the null-region (e.g., _b_ = .15, [95% CI: .05, .25]), the hypothesis that the effect is only trivial is be supported.
If the effects interval and the null region overlap (e.g., _b_ = .25, [95% CI: .15, .35]), the hypothesis is not supported and the results are considered inconclusive, whereas a meaningful negative effect is rejected.
If the confidence falls completely outside the null-region (e.g., _b_ = .4, [95% CI: .35, .45]), the hypothesis is rejected and the existence of a meaningful positive effect supported.

<!-- Responses were individually checked for patterns such as straight-lining or missing of inverted items. -->
<!-- X clear cases were removed. -->

Respondents who answered less than 50% of all questions were removed.
The remaining missing responses were imputed using predictive mean matching.
<!-- We will remove respondents with unrealistically fast responses, namely below three standard deviations of the medium response time.  -->

## Data Analysis

The hypothesis was analyzed using mixed effects models, namely random effect within-between models (REWB)[@bellFixedRandomEffects2019]. 
Three models were run, one for each dependent variable.
All predictors---that is, social media activities and social media channels, within and between-person factors, stably and varying covariates---were included in each of the three models. 
The factorial validity of the scales were tested with confirmatory factor analyses (CFA). 
If Mardia’s test showed that the assumption of multivariate normality was violated, the more robust Satorra-Bentler scaled and mean-adjusted test statistic (MLM) was used as estimator. 
To avoid overfitting, scales were tested on more liberal fit criteria (CFI > .90, TLI > .90, RMSEA <. .10, SRMR < .10) [@klinePrinciplesPracticeStructural2016].
Because REWB-models cannot model scales as latent variables, to increase precision factor scores exported from the CFAs were used.
For more information on the analyses, see [companion website](xxx).

## Measures

In what follows, I briefly list all variables that were collected.
For a complete list of all items and item characteristics, see [companion website](xxx).

### Well-being

Life satisfaction was measured with the item "Taken everything together, how satisfied are you currently with your life?".
The response options ranged from 0 (_extremely unsatisfied_) to 10 (_extremely satisfied_).
The variable's average across all waves was _M_ = `r get_specs(model_life_sat)$mean`, ranging from _M_~min~ = `r get_specs(model_life_sat)$min` to _M_~max~ = `r get_specs(model_life_sat)$max`.
The average standard deviation was _SD_ = `r get_specs(model_life_sat)$sd`.

To capture positive affect, respondents were asked how often in the last week they felt (a) calm and relaxed, (b) happy, and (c) full of energy.
The response options were 1 (_never_), 2 (_on some days_ ), 3 (_several times per week_), 4 (almost every day), and 5 (daily).
The variable's average across all waves was _M_ = `r get_specs(model_aff_pos)$mean`, ranging from _M_~min~ = `r get_specs(model_aff_pos)$min` to _M_~max~ = `r get_specs(model_aff_pos)$max`.
The average standard deviation was _SD_ = `r get_specs(model_aff_pos)$sd`.

For negative affect, respondents were asked how often in the last week they felt (a) lonely, (b) aggravated, (c) so depressed, that nothing could lift you up, (d) very nervous, (e) anxious, and (h) glum and sad.
The response options were 1 (_never_), 2 (_on some days_ ), 3 (_several times per week_), 4 (almost every day), and 5 (daily).
The variable's average across all waves was _M_ = `r get_specs(model_aff_neg)$mean`, ranging from _M_~min~ = `r get_specs(model_aff_neg)$min` to _M_~max~ = `r get_specs(model_aff_neg)$max`.
The average standard deviation was _SD_ = `r get_specs(model_aff_neg)$sd`.

### Control variables

The effects of COVID-related social media use were controlled for the following stable variables: 
(a) gender (answer options: female, male, diverse), (b) age, (c) education (10 ordinal options), (d) Austria country of birth (yes/no), (e) Austria parents' country of birth (no parent, one parent, both parents).
We controlled also for the following varying covariates:
(a) text-based media news consumption, (b) video-based media news consumption, (c) residency is Vienna, (d) household size, (e) Self-reported physical health, (f) Living space (in squaremeter), (g) access to balcony, (h) access to garden, (i) Employment status, (j) Work hours per week, (k) Home office, (l) household income, (m) outdoor activities, (o) satisfaction with democracy, (p) disposition to take risks, (q) locus of control.

# Results
## Preregistered Analyses

When looking at the variables from a descriptive perspective, we see that all well-being measures did not change substantially across the different phases of the pandemic.
COVID-related media use, however, increased during the beginning of the study and remained stable after approximately six waves.

```{r fig-desc, fig.cap="Development of well-being and media use measures across the pandemic. Values obtained from mixed effect models, with participants and waves as grouping factor and without additional predictors.", out.width = "\\textwidth", fig.pos = "!h"}
knitr::include_graphics("figures/fig_descriptives.pdf")
```

The study's hypothesis was that the effects of all types of social media use on well-being will be trivial.
Regarding the different types of _communication_, that is reading vs. sharing vs. posting, all within-person effects fell completely within the a-priori defined SESOIs.
For example, respondents who used social media more frequently than usual to read about COVID-related topics did not show a simultaneous change in life satisfaction (_b_ = `r dat_fig_results_activity %>% filter(dv == "Life satisfaction" & iv == "Reading" & type == "within") %>% select(estimate) %>% round(2)` [95% CI `r dat_fig_results_activity %>% filter(dv == "Life satisfaction" & iv == "Reading" & type == "within") %>% select(conf.low) %>% round(2)`, `r dat_fig_results_activity %>% filter(dv == "Life satisfaction" & iv == "Reading" & type == "within") %>% select(conf.high) %>% round(2)`]).
Only one effect did not include zero.
If people posted more about COVID-related aspects than they usually did, life satisfaction increased (_b_ = `r dat_fig_results_activity %>% filter(dv == "Life satisfaction" & iv == "Posting" & type == "within") %>% select(estimate) %>% round(2)` [95% CI `r dat_fig_results_activity %>% filter(dv == "Life satisfaction" & iv == "Posting" & type == "within") %>% select(conf.low) %>% round(2)`, `r dat_fig_results_activity %>% filter(dv == "Life satisfaction" & iv == "Posting" & type == "within") %>% select(conf.high) %>% round(2)`]).
However, because the effect did not fall outside our null region, it's likely not large enough to be considered meaningful.
As a result, the hypothesis was supported for all types of COVID-related social media communication.

Regarding between-person relations, about which no hypotheses were formulated, only two effects didn't include zero.
First, respondents who across all waves used social media more frequently than others to read about COVID reported higher levels of positive affect than others (_b_ = `r dat_fig_results_activity %>% filter(dv == "Positive affect" & iv == "Reading" & type == "between") %>% select(estimate) %>% round(2)` [95% CI `r dat_fig_results_activity %>% filter(dv == "Positive affect" & iv == "Reading" & type == "between") %>% select(conf.low) %>% round(2)`, `r dat_fig_results_activity %>% filter(dv == "Positive affect" & iv == "Reading" & type == "between") %>% select(conf.high) %>% round(2)`]).
However, note that this effect still fell completely within the null-region.
Hence, although positive the effect was considered too small to matter practically.
Second, respondents who across all waves used social media more frequently than others to read about COVID reported lower levels of negative affect than others (_b_ = `r dat_fig_results_activity %>% filter(dv == "Negative affect" & iv == "Reading" & type == "between") %>% select(estimate) %>% round(2)` [95% CI `r dat_fig_results_activity %>% filter(dv == "Negative affect" & iv == "Reading" & type == "between") %>% select(conf.low) %>% round(2)`, `r dat_fig_results_activity %>% filter(dv == "Negative affect" & iv == "Reading" & type == "between") %>% select(conf.high) %>% round(2)`]).
The effect was partially outside the null region, hence the effect might be large enough to be practically relevant, and the effect that it's trivial cannot be rejected.

```{r fig-res-activity, fig.cap="The effects of various types of social media use on three indicators of well-being. Effects are controlled for a large number of covariates (see text). The SESOI was 0.30 for life satisfaction and .15 for affects; hence, no effect is considered meaningful theoretically.", out.width = "\\textwidth", fig.pos = "!h"}
knitr::include_graphics("figures/fig_results_activity.pdf")
```

Regarding the COVID-related use of social media _channels_, the results were virtually the same.
Changes in the frequency of using different social media channels to attain information regarding COVID were unrelated to meaningful changes in well-being.
For example, respondents who used Facebook more frequently than usual to learn about COVID did not show a simultaneous change in well-being (_b_ = `r dat_fig_results_channel %>% filter(dv == "Life satisfaction" & iv == "Facebook" & type == "within") %>% select(estimate) %>% round(2)` [95% CI `r dat_fig_results_channel %>% filter(dv == "Life satisfaction" & iv == "Facebook" & type == "within") %>% select(conf.low) %>% round(2)`, `r dat_fig_results_channel %>% filter(dv == "Life satisfaction" & iv == "Facebook" & type == "within") %>% select(conf.high) %>% round(2)`]).
Only one effect was substantially larger than zero.
Respondents who used Instagram more frequently than usual to attain COVID-related news reported slightly lowers simultaneous levels of life satisfaction then usual (_b_ = `r dat_fig_results_channel %>% filter(dv == "Life satisfaction" & iv == "Instagram" & type == "within") %>% select(estimate) %>% round(2)` [95% CI `r dat_fig_results_channel %>% filter(dv == "Life satisfaction" & iv == "Instagram" & type == "within") %>% select(conf.low) %>% round(2)`, `r dat_fig_results_channel %>% filter(dv == "Life satisfaction" & iv == "Instagram" & type == "within") %>% select(conf.high) %>% my_round("b")`]).
However, this effect was still completely within the null region, hence not large enough to be considered meaningful.
In sum, the hypothesis was supported for the COVID-related use of all types of social media channels.

In terms of between-person relations---which, again, weren't included in the hypotheses---no relations crossed or fell outside the null region.
However, some relations did not included zero.
For example, respondents who across all waves used Instagram more frequently than others for COVID-related reasons reported lower levels of life satisfaction  (_b_ = `r dat_fig_results_channel %>% filter(dv == "Life satisfaction" & iv == "Instagram" & type == "between") %>% select(estimate) %>% round(2)` [95% CI `r dat_fig_results_channel %>% filter(dv == "Life satisfaction" & iv == "Instagram" & type == "between") %>% select(conf.low) %>% round(2)`, `r dat_fig_results_channel %>% filter(dv == "Life satisfaction" & iv == "Instagram" & type == "between") %>% select(conf.high) %>% my_round("b")`]).
Respondents who were more active on WhatsApp compared to others reported slightly lower levels of positive affect  (_b_ = `r dat_fig_results_channel %>% filter(dv == "Positive affect" & iv == "WhatsApp" & type == "between") %>% select(estimate) %>% round(2)` [95% CI `r dat_fig_results_channel %>% filter(dv == "Positive affect" & iv == "WhatsApp" & type == "between") %>% select(conf.low) %>% round(2)`, `r dat_fig_results_channel %>% filter(dv == "Positive affect" & iv == "WhatsApp" & type == "between") %>% select(conf.high) %>% my_round("b")`]).
Respondents who compared to others were more active on Twitter and YouTube reported lover levels of negative affect (Twitter: _b_ = `r dat_fig_results_channel %>% filter(dv == "Negative affect" & iv == "Twitter" & type == "between") %>% select(estimate) %>% round(2)` [95% CI `r dat_fig_results_channel %>% filter(dv == "Negative affect" & iv == "Twitter" & type == "between") %>% select(conf.low) %>% round(2)`, `r dat_fig_results_channel %>% filter(dv == "Negative affect" & iv == "Twitter" & type == "between") %>% select(conf.high) %>% my_round("b")`]; 
YouTube _b_ = `r dat_fig_results_channel %>% filter(dv == "Negative affect" & iv == "YouTube" & type == "between") %>% select(estimate) %>% round(2)` [95% CI `r dat_fig_results_channel %>% filter(dv == "Negative affect" & iv == "YouTube" & type == "between") %>% select(conf.low) %>% round(2)`, `r dat_fig_results_channel %>% filter(dv == "Negative affect" & iv == "YouTube" & type == "between") %>% select(conf.high) %>% my_round("b")`] .
However, please note that all these affect not considered large enough to be practically relevant.

```{r fig-res-channels, fig.cap="The effects of using various social media applications on three indicators of well-being. Effects are controlled for a large number of covariates (see text). The SESOI was 0.30 for life satisfaction and .15 for affects; hence, no effect is considered meaningful theoretically.", out.width = "\\textwidth", fig.pos = "!h"}
knitr::include_graphics("figures/fig_results_channel.pdf")
```

```{r tab-aff-neg}
# apa_print(model_aff_neg_lmer)
# tab <- broom.mixed::tidy(model_aff_neg_lmer, conf.int = T) %>% 
#  select(term, estimate, conf.low, conf.high, p.value)
#  kable()
```

## Exploratory Analyses

[Report results of some control variables.]

# Discussion

Using a panel study with a representative sample of the Austrian population that consistent of 24 waves, this study analyzed the effects of COVID-related social media use on well-being.
A random effect model that separated between person relations from within-person effects and that controlled for several third variables showed that within-person effects were trivial.
People who used social media more than usual to learn about COVID did not show changes in their well-being.
As a result, the results imply that COVID-related social media use is irrelevant for people's well-being.
Other factors among the third variables that were measured revealed much larger effects, implying that the relevance of specific types of social media use for well-being is limited.
Popular fears that "doomscrolling" or overusing social media during times of crises might not be justified.

The study is not aligned with a recent cross-sectional study by @bendauAssociationsCOVID19Related2021 that showed negative relations between social media and well-being.
However, @bendauAssociationsCOVID19Related2021 analyzed cross-sectional data on a between-person level, while not controlling for third variables, which does not allow to make causal inferences.
At the same time, this study is well-aligned with recent studies and meta-analyses from related research questions, which found that the effects of various types of social media use on several well-being indicators is small at best, often too small to matter [@meierComputerMediatedCommunicationSocial2020; @orbenTeenagersScreensSocial2020].

## Limitations

The current study analyzed whether changes in media use were related with changes in well-being, while controlling for several potential confounds. 
Together, this allows for a good perspective on potential causality.
That said, causality necessitates temporal order, and the cause needs to precede the consequence.
Regarding media use, such effects often happen immediately or shortly after use, necessitating intervals in the hours, minutes, or even seconds.
Only experience sampling studies that ask users in the very moment can produce such knowledge.
However, even then we don't know for certain if we actually measured the right interval.
Hence, to document how effects unfold it needs future research employing different study designs with different time lags.

[SESOI too large/conservative.]

Both media use and well-being were measured using self-reports.
Measuring well-being with self-reports is adequate, because it by definition requires introspection.
However, it would be preferable to measure social media use objectively, because people cannot reliably estimate their use.
That said, objective measures often cannot capture the content or the motivation of the use, and only very complicated tools that record the content that was used (such as the Screenome project) might produce such data.
However, also these procedure introduce other problems, for example related to privacy.
Hence, for this type of research question it seems necessary still to use self-reported measures.

The generalizability of the results are not large, because the data were collected in a single country.
The results are hence potentially limited to the more Western sphere, and might not apply to other cultures, especially if they have a different media landscape or offer alternative social media.
That said, because this is a large study, representative of a country's entire population, and because several waves were collected across a large time span, the results should be at least as generalizable as other typical empirical studies collected in the social sciences.

Social media use was measured with an ordinal variable, however in the analyses it was treated as a numerical one.
If treated as an ordinal one, it would have been necessary to analyze four different contrasts for each media measure, which plus the differentiation between between and within factor would have produced eight different measures, we would have made the model exceedingly complex.

## Conclusion

In this study, COVID-related social media use did not causally affect several indicators of well-being, including life satisfaction, positive affect, and negative affect.
However, factors other than social media use did affect well-being, such as income levels or access to a balcony.
If it's the aim to improve well-being, it might hence be fruitful not to focus on social media but to address other, potentially more pressing societal problems related to inequality.

\newpage

# References

<!-- \begingroup -->
<!-- \setlength{\parindent}{-0.5in} -->
<!-- \setlength{\leftskip}{0.5in} -->
<div id = "refs"></div>
<!-- \endgroup -->

# Competing Interests

I declare no competing interests.

# Supplementary Material

All the stimuli, presentation materials, analysis scripts, and a reproducible version of the manuscript can be found on the companion website (XXX).

# Data Accessibility Statement

The data are shared on AUSSDA, see https://doi.org/10.11587/28KQNS, and can be used for scientific purposes only.

### 2021-11-03 manuscript.tex
- set-up of repo

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
%
\documentclass[
  english,
  man,floatsintext]{apa6}
\usepackage{amsmath,amssymb}
\usepackage{lmodern}
\usepackage{ifxetex,ifluatex}
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\hypersetup{
  pdftitle={Analyzing the effects of COVID-19 related social media use on well-being},
  pdfauthor={Tobias Dienlin1},
  pdflang={en-EN},
  pdfkeywords={COVID-19, well-being, affect, life satisfaction, social media use, news use, random effects within between model, panel study, longitudinal},
  hidelinks,
  pdfcreator={LaTeX via pandoc}}
\urlstyle{same} % disable monospaced font for URLs
\usepackage{graphicx}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother
\setlength{\emergencystretch}{3em} % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{-\maxdimen} % remove section numbering
% Make \paragraph and \subparagraph free-standing
\ifx\paragraph\undefined\else
  \let\oldparagraph\paragraph
  \renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
  \let\oldsubparagraph\subparagraph
  \renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
\fi
% Manuscript styling
\usepackage{upgreek}
\captionsetup{font=singlespacing,justification=justified}

% Table formatting
\usepackage{longtable}
\usepackage{lscape}
% \usepackage[counterclockwise]{rotating}   % Landscape page setup for large tables
\usepackage{multirow}		% Table styling
\usepackage{tabularx}		% Control Column width
\usepackage[flushleft]{threeparttable}	% Allows for three part tables with a specified notes section
\usepackage{threeparttablex}            % Lets threeparttable work with longtable

% Create new environments so endfloat can handle them
% \newenvironment{ltable}
%   {\begin{landscape}\begin{center}\begin{threeparttable}}
%   {\end{threeparttable}\end{center}\end{landscape}}
\newenvironment{lltable}{\begin{landscape}\begin{center}\begin{ThreePartTable}}{\end{ThreePartTable}\end{center}\end{landscape}}

% Enables adjusting longtable caption width to table width
% Solution found at http://golatex.de/longtable-mit-caption-so-breit-wie-die-tabelle-t15767.html
\makeatletter
\newcommand\LastLTentrywidth{1em}
\newlength\longtablewidth
\setlength{\longtablewidth}{1in}
\newcommand{\getlongtablewidth}{\begingroup \ifcsname LT@\roman{LT@tables}\endcsname \global\longtablewidth=0pt \renewcommand{\LT@entry}[2]{\global\advance\longtablewidth by ##2\relax\gdef\LastLTentrywidth{##2}}\@nameuse{LT@\roman{LT@tables}} \fi \endgroup}

% \setlength{\parindent}{0.5in}
% \setlength{\parskip}{0pt plus 0pt minus 0pt}

% Overwrite redefinition of paragraph and subparagraph by the default LaTeX template
% See https://github.com/crsh/papaja/issues/292
\makeatletter
\renewcommand{\paragraph}{\@startsection{paragraph}{4}{\parindent}%
  {0\baselineskip \@plus 0.2ex \@minus 0.2ex}%
  {-1em}%
  {\normalfont\normalsize\bfseries\itshape\typesectitle}}

\renewcommand{\subparagraph}[1]{\@startsection{subparagraph}{5}{1em}%
  {0\baselineskip \@plus 0.2ex \@minus 0.2ex}%
  {-\z@\relax}%
  {\normalfont\normalsize\itshape\hspace{\parindent}{#1}\textit{\addperi}}{\relax}}
\makeatother

% \usepackage{etoolbox}
\makeatletter
\patchcmd{\HyOrg@maketitle}
  {\section{\normalfont\normalsize\abstractname}}
  {\section*{\normalfont\normalsize\abstractname}}
  {}{\typeout{Failed to patch abstract.}}
\patchcmd{\HyOrg@maketitle}
  {\section{\protect\normalfont{\@title}}}
  {\section*{\protect\normalfont{\@title}}}
  {}{\typeout{Failed to patch title.}}
\makeatother
\keywords{COVID-19, well-being, affect, life satisfaction, social media use, news use, random effects within between model, panel study, longitudinal}
\usepackage{lineno}

\linenumbers
\usepackage{csquotes}
\setlength{\parskip}{0em}
\raggedbottom
\note{\clearpage}
\ifxetex
  % Load polyglossia as late as possible: uses bidi with RTL langages (e.g. Hebrew, Arabic)
  \usepackage{polyglossia}
  \setmainlanguage[]{english}
\else
  \usepackage[main=english]{babel}
% get rid of language-specific shorthands (see #6817):
\let\LanguageShortHands\languageshorthands
\def\languageshorthands#1{}
\fi
\ifluatex
  \usepackage{selnolig}  % disable illegal ligatures
\fi
\newlength{\cslhangindent}
\setlength{\cslhangindent}{1.5em}
\newlength{\csllabelwidth}
\setlength{\csllabelwidth}{3em}
\newenvironment{CSLReferences}[2] % #1 hanging-ident, #2 entry spacing
 {% don't indent paragraphs
  \setlength{\parindent}{0pt}
  % turn on hanging indent if param 1 is 1
  \ifodd #1 \everypar{\setlength{\hangindent}{\cslhangindent}}\ignorespaces\fi
  % set entry spacing
  \ifnum #2 > 0
  \setlength{\parskip}{#2\baselineskip}
  \fi
 }%
 {}
\usepackage{calc}
\newcommand{\CSLBlock}[1]{#1\hfill\break}
\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\csllabelwidth}{#1}}
\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth - \csllabelwidth}{#1}\break}
\newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}

\title{Analyzing the effects of COVID-19 related social media use on well-being}
\author{Tobias Dienlin\textsuperscript{1}}
\date{}

\shorttitle{COVID-19 related social media use and well-being}

\authornote{

Tobias Dienlin, Department of Communication, University of Vienna.

Correspondence concerning this article should be addressed to Tobias Dienlin, University of Vienna, Department of Communication, 1090 Vienna, Austria. E-mail: \href{mailto:tobias.dienlin@univie.ac.at}{\nolinkurl{tobias.dienlin@univie.ac.at}}

}

\affiliation{\vspace{0.5cm}\textsuperscript{1} University of Vienna}

\abstract{
In times of crisis such as the Corona pandemic, it is important that citizens stay informed about recent events, the latest political decisions, or mandatory protection measures. As a result, to attain relevant information many people use various types of media, and increasingly social media. However, because social media are particularly engaging, some find it hard to disconnect and cannot stop `doomscrolling.' In this preregistered study, I investigate whether using social media for COVID-19 related topics might put personal well-being at risk. To this end, I analyze data from the Austrian Corona Panel Project, which consists of 24 waves with overall 3,018 participants. The data were analzed using random effects cross lagged panel models, controlling for several stable and varying covariates. Results showed that COVID-related social media use did not meaningfully affect several types of well-being, including life satisfaction, positive affect, and negative effect. This pertains to both passive and active social media use, and all the prominent channels such as Facebook, WhatsApp, or YouTube. As a result, this study suggests fears that social media use during times of crisis might be detrimental for well-being can be put to rest.
}

\begin{document}
\maketitle

During the COVID-19 pandemic,
several events unfolded in quick succession.
How dangerous is the virus? Is it spreading in my region? How is it transmitted, and how can I protect myself?
Because for many it was (and still is) a matter of life or death, citizens had to stay informed regarding the latest developments.
Governments around the world implemented safety measures, from wearing masks or keeping physical distance, to complete lockdowns.
In this extraordinary situation, people hence used media excessively, and especially social media were at an all time high (Statista, 2021).
A new phenomenon emerged, ``termed doomscrolling'':
People could not stop using social media to attain COVID-related news.

Several people reported that they were glued to their screens and could not pursue other relevant activities such as working, taking a break, or even care-work.
Increasingly, it was asked whether such an increase in social media use could still be considered useful and adaptive, or whether it created an additional and new psychological danger for the users' mental health (Klein, 2021).
A study with 6,233 people from Germany found that ``{[}f{]}requency, duration and diversity of media exposure were positively associated with more symptoms of depression and unspecific and COVID-19 specific anxiety'' (Bendau et al., 2021).

As a result, with this study I want to build on this research and investigate the question whether COVID-related social media use during the pandemic affected the users' well-being.
To this end I analyze a large-scale panel study from the Austrian Corona Panel Project (Kittel et al., 2020).
The panel consists of 24 waves and an overall sample size of 3018, with at least 1,500 participants per wave, and it is representative of the Austrian population.
The panel study collected a large number of variables.
Because we can therefore control for many confounding third variables, both stable and varying, together with the longitudinal design this creates a unique opportunity to analyze causality.

\hypertarget{defining-well-being-and-media-use}{%
\subsection{Defining Well-being and Media Use}\label{defining-well-being-and-media-use}}

The underlying theories that guided the selection of variables for this study are the two-continua model of mental health (Greenspoon \& Saklofske, 2001) and the hierarchical taxonomy of computer-mediated medation (Meier \& Reinecke, 2020).
According to the two-continua model of mental health, mental health consists of two dimensions, psychopathology and well-being.
Because the aim of this study is better to understand typical users and everyday contexts, my focus will be on well-being.
Well-being can be differentiated into subjective and psychological well-being (Diener, Lucas, \& Oishi, 2018).
Whereas subjective well-being emphasizes hedonic aspects such as a happiness and joy, psychological well-being focuses on eudaimonic aspects including fulfillment and meaning.
Subjective well-being is primarily about achieving positive affect and avoiding negative affect.
In my view, life satisfaction is a meta concept above both psychological and subjective well-being, representing a meta-appraisal of one's life.
Notably, life satisfaction is very stable and fluctuates only little, whereas it's the exact opposite for affect (Dienlin \& Johannes, 2020).
To capture well-being in this study, I will thus build on life satisfaction, positive affect, and negative affect.
Together, this should provide an encompassing perspective on media effects.

The hierarchical taxonomy of computer-mediated communication differentiates six levels of how people engage with digital technology:
first, the device (e.g., smartphone); second, the type of application (e.g., social networking site); third, the branded application (e.g., Twitter); fourth, the feature (e.g., status post); fifth, the interaction (e.g., one-to-many); sixth, the message (e.g., content) (Meier \& Reinecke, 2020).
Whereas the first four levels are focused on the channel, the last two on the type of communication.
To measure social media use for consumption of specific news, I here employ both the channel and the communication perspective.
First, I will analyze how using the most prominent branded applications affect well-being, and whether this effect changes across applications.
The branded applications investigated here are Facebook, Twitter, Instagram, WhatsApp, and YouTube.
But because adopting only this position would be both too narrow and too general, I will secondly also investigate how different types of interaction affect well-being.
Specifically, I will differentiate between active and passive use.
I will distinguish (a) \emph{reading} COVID-related social media use (passive), \emph{posting} content regarding COVID (active), and \emph{liking and sharing} COVID-related posts by others (both active and passive).
Worth noting, this study is not about \emph{general} social media during times of COVID, but on social media use focused on COVID-related news and interactions.
For example, posting about the pandemic or retweeting COVID-related posts.

\hypertarget{effects-of-social-media-on-well-being}{%
\subsection{Effects of Social Media on Well-Being}\label{effects-of-social-media-on-well-being}}

In their study on the relations between media use and mental health during the pandemic, Bendau et al. (2021) found that people who used social media as a primary source of information reported on average ``significantly more unspecific anxiety and depression {[}{]} and significantly more specific COVID-19 related anxiety symptoms'' (p.~288).
Hence, this might hint at potential negative effects of social media as news use---but note that this finding comes from a between-person relation stemming from a cross-sectional data.
Hence, we don't know whether the differences in mental health and well-being are due to social media use or other third variables, such as age or education.

In general, so far there is only little research on news-related social media use and well-being.
On the other hand, the question of whether and how \emph{general} social media use---or other, more specific types such as active or passive use---affect well-being is well-researched.
A meta review (i.e., an analysis of meta-analyses), found that the relation between social media use and well-being is likely in the negative spectrum---but very small, potentially too small to matter (Meier \& Reinecke, 2020).
This is well aligned with several new studies that employed the most advanced methods (Keresteš \& Štulhofer, 2020; Orben, Dienlin, \& Przybylski, 2019; Przybylski, Nguyen, Law, \& Weinstein, 2021; Schemer, Masur, Geiß, Müller, \& Schäfer, 2021).
For example, Beyens, Pouwels, Driel, Keijsers, and Valkenburg (2020) reported that although for some users (roughly one quarter) the effects were negative, for almost the same amount of users the effects were positive, while for the majority they were neutral.
At all events, in general most effects are somewhere between trivial and small.

What determines a trivial or a small effect is a difficult question (see below).
If we refer to standardized effect sizes, according to Cohen (Cohen, 1992) small effect sizes start at \emph{r} = .10.
However, most meta-analyses find effect sizes below that threshold (Huang, 2017; Meier \& Reinecke, 2020).
As a result, I think it's most convincing to expect trivial to small effects also in the case of COVID-related social media use.

From a theoretical perspective, why could COVID-related social media use be detrimental?
Above everything, one can reasonably assume a \emph{direct} negative effect on well-being, especially on positive or negative effect.
Dangers, inequalities, corruption -- these were the headlines across many countries worldwide.
If one learns about such things, the initial reaction might be shock, fear, or dismay.
Repeatedly consuming such news might be depressing, perhaps even changing some general perspective on life.
So, just like being hit by a hammer hurts and we don't need any ``mediating mechanism,'' this could be the case here as well.

That said, because not all news were negative, because many people showed solidarity and compassion,
there were also positive and potentially uplifting news.
However, in light of a worldwide pandemic with millions of deaths, the negative direct effect seems more plausible.

There could be also indirect effects.
When doomscrolling, users are captivated to such an extent that they cannot stop using social media.
For example, during the pandemic social media uses was at an all-time high in the US (Statista, 2021).
As has been expressed by many before, it is most likely that moderate social media use is not detrimental (Orben, 2020).
However, overuse is more critical, and several studies have showed more pronounced negative effects for extreme users (Przybylski \& Weinstein, 2017).
So if a society collectively overuses social media, there is potential for negative effects.
Overuse likely impairs well-being if it replaces other meaningful or functional activities such as meeting others, working, actively relaxing, or exercising.
If overuse replaces such activities it's reasonable to assume that it's also detrimental.

On the other hand, one can make the case that overuse can be beneficial \emph{in times of a pandemic}, even if it's mainly COVID-related.
Exchanging COVID-related messages with friends via WhatsApp might replace the in-person contact one would have otherwise, but which is logically impossible at that time.
At times where meaningful and functional activities are explicitly or implicitly prohibited, using social media to exchange about COVID might not be the worst idea.
In fact, given that nowadays a large number of experts, scientists, and politicians converse directly on social media, one can get first-hand information on the current developments.
On the other hand, there is of course also much disinformation, and ``bingeing'' on COVID fake news might also pose risks for impaired well-being.

Together, the strongest argument seems to be that the effects of social media on well-being are, on average, generally small at best.
Because this study looks at only \emph{one part} of social media use---namely, COVID-related interactions---it is even more focused, and the overall potential of the effects should diminish even further.
Whether or not using social media for COVID-related aspects is detrimental during a pandemic is also not entirely clear.
Therefore, I expect to find that the effects of COVID-related news use on well-being will be not be meaningful or practically relevant.

\begin{quote}
Hypothesis: The within-person effects of all types of COVID-19 related social media use on all types of well-being indicators---while controlling for several stable and varying covariates such as sociodemographic variables and psychological dispositions---will be trivial.
\end{quote}

\hypertarget{current-study}{%
\section{Current Study}\label{current-study}}

\hypertarget{smallest-effect-size-of-interest}{%
\subsection{Smallest Effect Size of Interest}\label{smallest-effect-size-of-interest}}

Testing this hypothesis, however, is not trivial.
First, in contrast to most hypothesis typically posited in the social sciences, it explicates an effect size, a so-called smallest effect size of interest (SESOI).
Effectively testing this hypothesis hence necessitates to define what's considered a ``trivial effect size'' and what's not.
Above I referred to standardized effect sizes.
However, standardized effect sizes should only be a first step toward evaluating an effect's relevance (Baguley, 2009).
Standardized effect sizes are determined by a sample's variance.\footnote{Consider the effect size Cohen's \emph{d}: The mean's of the two groups that are to be compared are subtracted from one another and then divided by the sample's standard deviation (Cohen, 1992). Hence, if there is more deviation/variance in a sample, the effect size decreases, even if the difference of the group's means stays the same.}
However, this is problematic: The question of whether or not social media use affects me personally in a relevant way should not depend on the variance in the sample in which my data were collected.
Instead, it should depend on absolute criteria.
What could be a minimally interesting, a nontrivial effect?
I suggest the following SESOI for this research question:

If a heavy user of COVID-related social media news stops using social media altogether, this should have a noticeable impact on their overall well-being.

What would this mean practically and how could it be operationalized?
In this study, COVID-related social media use was measured on a 5-point scale, ranging from 1 = \emph{never} to 5 = \emph{several times a day}.
Thus, the example from above would imply that a change of five units in social media use should correspond to a noticeable change in well-being.
But what's a noticeable change in well-being?
According to Norman, Sloan, and Wyrwich (2003), people can reliably differentiate between \emph{seven} levels of satisfaction with health.
So we could state that a five unit change in social media use should result in a one unit change in satisfaction.
Statistically, in a regression, \emph{b} estimates the change in the dependent variable if the independent variable increases by one point.
Transferred to our example, we would hence expect a \emph{b} of 1 / 5 = .20.

In this study, life satisfaction was measured with 11 units, hence more than people can reliably detect.
Hence, we would expect a 11 / 7 * .20 change, that is a \emph{b} of at least .31.
For affect, which was measured on a 5-point scale, our SESOI would be 5 / 7 * .2 = .14.
In order not to exaggerate precision, these numbers will be rounded.
Plus, because we are agnostic as to whether the effects are positively or negatively nontrivial, this leads to an indifference region ranging from \emph{b} = -.30 to \emph{b} = .30 (and \emph{b} = -.15 to \emph{b} = .15 for affect).

\hypertarget{causality}{%
\subsection{Causality}\label{causality}}

The hypothesis explicitly states a causal effect.
In non-experimental designs, causality can be approximated using longitudinal designs.
Using longitudinal designs alone, however, is not sufficient for correct causal statements.
In addition, it is crucial also to control for third variables, and importantly also for \emph{varying} third variables.

For example, imagine that a person suddenly starts using social media much more than usual, and then at the same time also become less satisfied with their lives.
After some time, use recedes again, whereas life satisfaction returns to prior levels.
If this happened to several people at the same time, in a longitudinal study we could then find a causal effect of social media use on life satisfaction.
However, it could also be the case that during the study there was a major exogenous event (say, a pandemic) and a large part of the working population lost their jobs.
Hence, the causal effect reported above was confounded, because in reality it was the pandemic that caused both social media use to rise and their life satisfaction to plummet.

Thus, only if we can control for \emph{all} potential varying third variables, we can correctly estimate causality without any bias (Rohrer, 2018).
Obviously, we can never be entirely sure to have included all varying covariates, which makes absolute statements regarding causality impossible.
However, when controlling for many varying variables, we can be much more certain that we measured causality correctly.
The aim should still be to collect as many relevant varying and nonvarying third variables as possible, while knowing that absolute certainty can rarely be reached.

When looking for suitable control candidates, ideally we find variables that affect both media use and well-being, because controlling for these factors will isolate the actual effect of social media use on well-being.
We can also control for variables that affect \emph{only} well-being.
However, in doing so nothing is gained or lost because the effects of social media use would remain exactly the same (Kline, 2016).
Crucially, when determining the general causal effect of social media use we should \emph{not} control for mediating variables (Rohrer, 2018).
Doing so would bias and distort our assessment of the role of social media use.

In this study, I will hence control for the following variables, which either have been shown or a likely to affect both social media use and well-being.
(I'll additionally include variables that likely affect only well-being, also to obtain a comparison benchmark for social media effects):
Gender, age, education, Austria country of birth, Austria country of birth of parents, text-based news consumption, video-based news consumption, residency Vienna, household size, health, living space, access to garden, access to balcony, employment, work hours per week, being in home-office, household income, outdoor activities, satisfaction with democracy, disposition to take risks, and locus of control.
I will not control for variables such as trust in institutions or media, because these variables are likely influenced by social media use.

{[}Within vs.~between logic{]}

{[}Time Frame / REWB logic{]}

\hypertarget{method}{%
\section{Method}\label{method}}

This section describes the preregistration and how I determined the sample size, data exclusions, the analyses, and all measures in the study.

\hypertarget{preregistration}{%
\subsection{Preregistration}\label{preregistration}}

The hypotheses, the sample, the measures, the analyses, and the inference criteria (SESOI, p-value) were preregistered on the Open Science Framework.
The (anonymous) preregistration can be accessed here: \url{https://osf.io/87b24/?view_only=b2289b6fec214fa88ee75a18d45c18f3}.
Because in this study I analyze data from an already existing large scale data set, the Austrian Corona Panel Project, all of these steps were preregistered \emph{prior} to accessing the data.
The preregistration was designed on the basis of the panel documentation online (Kittel et al., 2020).
In some cases the analyses could not be executed as originally planned, because some properties of the variables only became apparent upon data analysis.
The most relevant deviations are reported below, and a complete list of all changes can be found online.

\hypertarget{sample}{%
\subsection{Sample}\label{sample}}

The data come from the Austrian Corona Panel Project (Kittel et al., 2021).
The study was conducted between March 2020 and October 2021.
The contains 26 waves, and at the time of writing, the first 24 waves were available for download.
Each wave consists of at least 1,500 respondents.
The overall sample size was \emph{N} = 3018, and 72432 observations were collected.
Panel mortality was compensated through a continuous reacquistion of new participants.
All respondents needed to have access to the internet (via computer or mobile devices such as smartphones or tablets).
They were sampled from a pre-existing online access panel provided by Marketagent, Austria.
Respondents were asked and incentivized with 180 credit points to participate in each wave of the panel.

The sample was representative of the Austrian population in terms of age, gender, region/state, municipality size, and educational level. In order to participate in the study, the respondents needed to be Austrian residents and had to be at least 14 years old.
The average age was 42, 49 percent were male, 13.60 percent had a University degree, and 5.07 percent were currently unemployed.

Because the data were analyzed post hoc, no sample size planning on the basis of a priori power analyses was conducted.
Because the sample is very large, it is well-equipped reliably to detect also small effects, rendering exact post hoc power analysis unnecessary.
And also because such large sample easily generate significant p-values for very small effects, this study builds on a smallest effect size of interest.
We will use the interval testing approach as proposed by Dienes (2014).
On the basis of our SESOI, we will define a null region.
For well-being, the null region will be between \emph{b} = -.30 and \emph{b} = .30.
If the 95\% confidence interval lies completely within the null-region (e.g., \emph{b} = .15, {[}95\% CI: .05, .25{]}), the hypothesis that the effect is only trivial is be supported.
If the effects interval and the null region overlap (e.g., \emph{b} = .25, {[}95\% CI: .15, .35{]}), the hypothesis is not supported and the results are considered inconclusive, whereas a meaningful negative effect is rejected.
If the confidence falls completely outside the null-region (e.g., \emph{b} = .4, {[}95\% CI: .35, .45{]}), the hypothesis is rejected and the existence of a meaningful positive effect supported.

Respondents who answered less than 50\% of all questions were removed.
The remaining missing responses were imputed using predictive mean matching.

\hypertarget{data-analysis}{%
\subsection{Data Analysis}\label{data-analysis}}

The hypothesis was analyzed using mixed effects models, namely random effect within-between models (REWB)(Bell, Fairbrother, \& Jones, 2019).
Three models were run, one for each dependent variable.
All predictors---that is, social media activities and social media channels, within and between-person factors, stably and varying covariates---were included in each of the three models.
The factorial validity of the scales were tested with confirmatory factor analyses (CFA).
If Mardia's test showed that the assumption of multivariate normality was violated, the more robust Satorra-Bentler scaled and mean-adjusted test statistic (MLM) was used as estimator.
To avoid overfitting, scales were tested on more liberal fit criteria (CFI \textgreater{} .90, TLI \textgreater{} .90, RMSEA \textless. .10, SRMR \textless{} .10) (Kline, 2016).
Because REWB-models cannot model scales as latent variables, to increase precision factor scores exported from the CFAs were used.
For more information on the analyses, see \href{xxx}{companion website}.

\hypertarget{measures}{%
\subsection{Measures}\label{measures}}

In what follows, I briefly list all variables that were collected.
For a complete list of all items and item characteristics, see \href{xxx}{companion website}.

\hypertarget{well-being}{%
\subsubsection{Well-being}\label{well-being}}

Life satisfaction was measured with the item ``Taken everything together, how satisfied are you currently with your life?''
The response options ranged from 0 (\emph{extremely unsatisfied}) to 10 (\emph{extremely satisfied}).
The variable's average across all waves was \emph{M} = 6.59, ranging from \emph{M}\textsubscript{min} = 6.39 to \emph{M}\textsubscript{max} = 6.79.
The average standard deviation was \emph{SD} = 1.68.

To capture positive affect, respondents were asked how often in the last week they felt (a) calm and relaxed, (b) happy, and (c) full of energy.
The response options were 1 (\emph{never}), 2 (\emph{on some days} ), 3 (\emph{several times per week}), 4 (almost every day), and 5 (daily).
The variable's average across all waves was \emph{M} = 3.15, ranging from \emph{M}\textsubscript{min} = 3.05 to \emph{M}\textsubscript{max} = 3.29.
The average standard deviation was \emph{SD} = 0.57.

For negative affect, respondents were asked how often in the last week they felt (a) lonely, (b) aggravated, (c) so depressed, that nothing could lift you up, (d) very nervous, (e) anxious, and (h) glum and sad.
The response options were 1 (\emph{never}), 2 (\emph{on some days} ), 3 (\emph{several times per week}), 4 (almost every day), and 5 (daily).
The variable's average across all waves was \emph{M} = 1.73, ranging from \emph{M}\textsubscript{min} = 1.66 to \emph{M}\textsubscript{max} = 1.81.
The average standard deviation was \emph{SD} = 0.39.

\hypertarget{control-variables}{%
\subsubsection{Control variables}\label{control-variables}}

The effects of COVID-related social media use were controlled for the following stable variables:
(a) gender (answer options: female, male, diverse), (b) age, (c) education (10 ordinal options), (d) Austria country of birth (yes/no), (e) Austria parents' country of birth (no parent, one parent, both parents).
We controlled also for the following varying covariates:
(a) text-based media news consumption, (b) video-based media news consumption, (c) residency is Vienna, (d) household size, (e) Self-reported physical health, (f) Living space (in squaremeter), (g) access to balcony, (h) access to garden, (i) Employment status, (j) Work hours per week, (k) Home office, (l) household income, (m) outdoor activities, (o) satisfaction with democracy, (p) disposition to take risks, (q) locus of control.

\hypertarget{results}{%
\section{Results}\label{results}}

\hypertarget{preregistered-analyses}{%
\subsection{Preregistered Analyses}\label{preregistered-analyses}}

When looking at the variables from a descriptive perspective, we see that all well-being measures did not change substantially across the different phases of the pandemic.
COVID-related media use, however, increased during the beginning of the study and remained stable after approximately six waves.

\begin{figure}[!h]
\includegraphics[width=\textwidth]{figures/fig_descriptives} \caption{Development of well-being and media use measures across the pandemic. Values obtained from mixed effect models, with participants and waves as grouping factor and without additional predictors.}\label{fig:fig-desc}
\end{figure}

The study's hypothesis was that the effects of all types of social media use on well-being will be trivial.
Regarding the different types of \emph{communication}, that is reading vs.~sharing vs.~posting, all within-person effects fell completely within the a-priori defined SESOIs.
For example, respondents who used social media more frequently than usual to read about COVID-related topics did not show a simultaneous change in life satisfaction (\emph{b} = -0.03 {[}95\% CI -0.09, 0.02{]}).
Only one effect did not include zero.
If people posted more about COVID-related aspects than they usually did, life satisfaction increased (\emph{b} = 0.11 {[}95\% CI 0.01, 0.2{]}).
However, because the effect did not fall outside our null region, it's likely not large enough to be considered meaningful.
As a result, the hypothesis was supported for all types of COVID-related social media communication.

Regarding between-person relations, about which no hypotheses were formulated, only two effects didn't include zero.
First, respondents who across all waves used social media more frequently than others to read about COVID reported higher levels of positive affect than others (\emph{b} = 0.05 {[}95\% CI 0.01, 0.09{]}).
However, note that this effect still fell completely within the null-region.
Hence, although positive the effect was considered too small to matter practically.
Second, respondents who across all waves used social media more frequently than others to read about COVID reported lower levels of negative affect than others (\emph{b} = 0.01 {[}95\% CI -0.02, 0.04{]}).
The effect was partially outside the null region, hence the effect might be large enough to be practically relevant, and the effect that it's trivial cannot be rejected.

\begin{figure}[!h]
\includegraphics[width=\textwidth]{figures/fig_results_activity} \caption{The effects of various types of social media use on three indicators of well-being. Effects are controlled for a large number of covariates (see text). The SESOI was 0.30 for life satisfaction and .15 for affects; hence, no effect is considered meaningful theoretically.}\label{fig:fig-res-activity}
\end{figure}

Regarding the COVID-related use of social media \emph{channels}, the results were virtually the same.
Changes in the frequency of using different social media channels to attain information regarding COVID were unrelated to meaningful changes in well-being.
For example, respondents who used Facebook more frequently than usual to learn about COVID did not show a simultaneous change in well-being (\emph{b} = 0.03 {[}95\% CI -0.03, 0.1{]}).
Only one effect was substantially larger than zero.
Respondents who used Instagram more frequently than usual to attain COVID-related news reported slightly lowers simultaneous levels of life satisfaction then usual (\emph{b} = -0.08 {[}95\% CI -0.15, \textgreater{} -0.01{]}).
However, this effect was still completely within the null region, hence not large enough to be considered meaningful.
In sum, the hypothesis was supported for the COVID-related use of all types of social media channels.

In terms of between-person relations---which, again, weren't included in the hypotheses---no relations crossed or fell outside the null region.
However, some relations did not included zero.
For example, respondents who across all waves used Instagram more frequently than others for COVID-related reasons reported lower levels of life satisfaction (\emph{b} = -0.11 {[}95\% CI -0.19, -0.03{]}).
Respondents who were more active on WhatsApp compared to others reported slightly lower levels of positive affect (\emph{b} = -0.03 {[}95\% CI -0.07, \textgreater{} -0.01{]}).
Respondents who compared to others were more active on Twitter and YouTube reported lover levels of negative affect (Twitter: \emph{b} = -0.04 {[}95\% CI -0.07, \textgreater{} -0.01{]};
YouTube \emph{b} = -0.06 {[}95\% CI -0.08, -0.03{]} .
However, please note that all these affect not considered large enough to be practically relevant.

\begin{figure}[!h]
\includegraphics[width=\textwidth]{figures/fig_results_channel} \caption{The effects of using various social media applications on three indicators of well-being. Effects are controlled for a large number of covariates (see text). The SESOI was 0.30 for life satisfaction and .15 for affects; hence, no effect is considered meaningful theoretically.}\label{fig:fig-res-channels}
\end{figure}

\hypertarget{exploratory-analyses}{%
\subsection{Exploratory Analyses}\label{exploratory-analyses}}

{[}Report results of some control variables.{]}

\hypertarget{discussion}{%
\section{Discussion}\label{discussion}}

Using a panel study with a representative sample of the Austrian population that consistent of 24 waves, this study analyzed the effects of COVID-related social media use on well-being.
A random effect model that separated between person relations from within-person effects and that controlled for several third variables showed that within-person effects were trivial.
People who used social media more than usual to learn about COVID did not show changes in their well-being.
As a result, the results imply that COVID-related social media use is irrelevant for people's well-being.
Other factors among the third variables that were measured revealed much larger effects, implying that the relevance of specific types of social media use for well-being is limited.
Popular fears that ``doomscrolling'' or overusing social media during times of crises might not be justified.

The study is not aligned with a recent cross-sectional study by Bendau et al. (2021) that showed negative relations between social media and well-being.
However, Bendau et al. (2021) analyzed cross-sectional data on a between-person level, while not controlling for third variables, which does not allow to make causal inferences.
At the same time, this study is well-aligned with recent studies and meta-analyses from related research questions, which found that the effects of various types of social media use on several well-being indicators is small at best, often too small to matter (Meier \& Reinecke, 2020; Orben, 2020).

\hypertarget{limitations}{%
\subsection{Limitations}\label{limitations}}

The current study analyzed whether changes in media use were related with changes in well-being, while controlling for several potential confounds.
Together, this allows for a good perspective on potential causality.
That said, causality necessitates temporal order, and the cause needs to precede the consequence.
Regarding media use, such effects often happen immediately or shortly after use, necessitating intervals in the hours, minutes, or even seconds.
Only experience sampling studies that ask users in the very moment can produce such knowledge.
However, even then we don't know for certain if we actually measured the right interval.
Hence, to document how effects unfold it needs future research employing different study designs with different time lags.

{[}SESOI too large/conservative.{]}

Both media use and well-being were measured using self-reports.
Measuring well-being with self-reports is adequate, because it by definition requires introspection.
However, it would be preferable to measure social media use objectively, because people cannot reliably estimate their use.
That said, objective measures often cannot capture the content or the motivation of the use, and only very complicated tools that record the content that was used (such as the Screenome project) might produce such data.
However, also these procedure introduce other problems, for example related to privacy.
Hence, for this type of research question it seems necessary still to use self-reported measures.

The generalizability of the results are not large, because the data were collected in a single country.
The results are hence potentially limited to the more Western sphere, and might not apply to other cultures, especially if they have a different media landscape or offer alternative social media.
That said, because this is a large study, representative of a country's entire population, and because several waves were collected across a large time span, the results should be at least as generalizable as other typical empirical studies collected in the social sciences.

Social media use was measured with an ordinal variable, however in the analyses it was treated as a numerical one.
If treated as an ordinal one, it would have been necessary to analyze four different contrasts for each media measure, which plus the differentiation between between and within factor would have produced eight different measures, we would have made the model exceedingly complex.

\hypertarget{conclusion}{%
\subsection{Conclusion}\label{conclusion}}

In this study, COVID-related social media use did not causally affect several indicators of well-being, including life satisfaction, positive affect, and negative affect.
However, factors other than social media use did affect well-being, such as income levels or access to a balcony.
If it's the aim to improve well-being, it might hence be fruitful not to focus on social media but to address other, potentially more pressing societal problems related to inequality.

\newpage

\hypertarget{references}{%
\section{References}\label{references}}

\hypertarget{refs}{}
\begin{CSLReferences}{1}{0}
\leavevmode\hypertarget{ref-baguleyStandardizedSimpleEffect2009}{}%
Baguley, T. (2009). Standardized or simple effect size: {What} should be reported? \emph{British Journal of Psychology}, \emph{100}(3), 603--617. \url{https://doi.org/10.1348/000712608X377117}

\leavevmode\hypertarget{ref-bellFixedRandomEffects2019}{}%
Bell, A., Fairbrother, M., \& Jones, K. (2019). Fixed and random effects models: Making an informed choice. \emph{Quality \& Quantity}, \emph{53}(2), 1051--1074. \url{https://doi.org/10.1007/s11135-018-0802-x}

\leavevmode\hypertarget{ref-bendauAssociationsCOVID19Related2021}{}%
Bendau, A., Petzold, M. B., Pyrkosch, L., Mascarell Maricic, L., Betzler, F., Rogoll, J., \ldots{} Plag, J. (2021). Associations between {COVID}-19 related media consumption and symptoms of anxiety, depression and {COVID}-19 related fear in the general population in {Germany}. \emph{European Archives of Psychiatry and Clinical Neuroscience}, \emph{271}(2), 283--291. \url{https://doi.org/10.1007/s00406-020-01171-6}

\leavevmode\hypertarget{ref-beyensEffectSocialMedia2020}{}%
Beyens, I., Pouwels, J. L., Driel, I. I. van, Keijsers, L., \& Valkenburg, P. M. (2020). The effect of social media on well-being differs from adolescent to adolescent. \emph{Scientific Reports}, \emph{10}(1), 10763. \url{https://doi.org/10.1038/s41598-020-67727-7}

\leavevmode\hypertarget{ref-cohenPowerPrimer1992}{}%
Cohen, J. (1992). A power primer. \emph{Psychological Bulletin}, \emph{112}(1), 155--159. \url{https://doi.org/10.1037/0033-2909.112.1.155}

\leavevmode\hypertarget{ref-dienerAdvancesOpenQuestions2018}{}%
Diener, E., Lucas, R. E., \& Oishi, S. (2018). Advances and open questions in the science of subjective well-being. \emph{Collabra: Psychology}, \emph{4}(1), 15. \url{https://doi.org/10.1525/collabra.115}

\leavevmode\hypertarget{ref-dienesUsingBayesGet2014}{}%
Dienes, Z. (2014). Using {Bayes} to get the most out of non-significant results. \emph{Frontiers in Psychology}, \emph{5}. \url{https://doi.org/10.3389/fpsyg.2014.00781}

\leavevmode\hypertarget{ref-dienlinImpactDigitalTechnology2020}{}%
Dienlin, T., \& Johannes, N. (2020). The impact of digital technology use on adolescent well-being. \emph{Dialogues in Clinical Neuroscience}, \emph{22}(2), 135--142. \url{https://doi.org/doi:10.31887/DCNS.2020.22.2/tdienlin}

\leavevmode\hypertarget{ref-greenspoonIntegrationSubjectiveWellbeing2001}{}%
Greenspoon, P. J., \& Saklofske, D. H. (2001). Toward an integration of subjective well-being and psychopathology. \emph{Social Indicators Research}, \emph{54}(1), 81--108. \url{https://doi.org/10.1023/A:1007219227883}

\leavevmode\hypertarget{ref-huangTimeSpentSocial2017}{}%
Huang, C. (2017). Time spent on social network sites and psychological well-being: {A} meta-analysis. \emph{Cyberpsychology, Behavior and Social Networking}, \emph{20}(6), 346--354. \url{https://doi.org/10.1089/cyber.2016.0758}

\leavevmode\hypertarget{ref-kerestesAdolescentsOnlineSocial2020}{}%
Keresteš, G., \& Štulhofer, A. (2020). Adolescents' online social network use and life satisfaction: {A} latent growth curve modeling approach. \emph{Computers in Human Behavior}, \emph{104}, 106187. \url{https://doi.org/10.1016/j.chb.2019.106187}

\leavevmode\hypertarget{ref-kittelAustrianCoronaPanel2020}{}%
Kittel, B., Kritzinger, S., Boomgaarden, H., Prainsack, B., Eberl, J.-M., Kalleitner, F., \ldots{} Schlogl, L. (2020). \emph{Austrian {Corona} {Panel} {Project} ({SUF} edition)}. AUSSDA. \url{https://doi.org/10.11587/28KQNS}

\leavevmode\hypertarget{ref-kittelAustrianCoronaPanel2021}{}%
Kittel, B., Kritzinger, S., Boomgaarden, H., Prainsack, B., Eberl, J.-M., Kalleitner, F., \ldots{} Schlogl, L. (2021). The {Austrian} {Corona} {Panel} {Project}: Monitoring individual and societal dynamics amidst the {COVID}-19 crisis. \emph{European Political Science}, \emph{20}(2), 318--344. \url{https://doi.org/10.1057/s41304-020-00294-7}

\leavevmode\hypertarget{ref-kleinDarklySoothingCompulsion2021}{}%
Klein, J. (2021). \emph{The darkly soothing compulsion of 'doomscrolling'}. Retrieved from \url{https://www.bbc.com/worklife/article/20210226-the-darkly-soothing-compulsion-of-doomscrolling}

\leavevmode\hypertarget{ref-klinePrinciplesPracticeStructural2016}{}%
Kline, R. B. (2016). \emph{Principles and practice of structural equation modeling} (4th ed.). New York, NY: The Guilford Press.

\leavevmode\hypertarget{ref-meierComputerMediatedCommunicationSocial2020}{}%
Meier, A., \& Reinecke, L. (2020). Computer-{Mediated} {Communication}, {Social} {Media}, and {Mental} {Health}: {A} {Conceptual} and {Empirical} {Meta}-{Review}. \emph{Communication Research}, 009365022095822. \url{https://doi.org/10.1177/0093650220958224}

\leavevmode\hypertarget{ref-normanInterpretationChangesHealthrelated2003}{}%
Norman, G., Sloan, J., \& Wyrwich, K. (2003). Interpretation of changes in health-related quality of life: {The} remarkable universality of half a standard deviation. \emph{Medical Care}, \emph{41}(5), 582--592. Retrieved from \url{Retrieved\%20from\%20http://www.jstor.org/stable/3768017}

\leavevmode\hypertarget{ref-orbenTeenagersScreensSocial2020}{}%
Orben, A. (2020). Teenagers, screens and social media: A narrative review of reviews and key studies. \emph{Social Psychiatry and Psychiatric Epidemiology}, \emph{55}(4), 407--414. \url{https://doi.org/10.1007/s00127-019-01825-4}

\leavevmode\hypertarget{ref-orbenSocialMediaEnduring2019}{}%
Orben, A., Dienlin, T., \& Przybylski, A. K. (2019). Social media's enduring effect on adolescent life satisfaction. \emph{Proceedings of the National Academy of Sciences of the United States of America}, \emph{116}(21), 10226--10228. \url{https://doi.org/10.1073/pnas.1902058116}

\leavevmode\hypertarget{ref-przybylskiDoesTakingShort2021a}{}%
Przybylski, A. K., Nguyen, T. T., Law, W., \& Weinstein, N. (2021). Does {Taking} a {Short} {Break} from {Social} {Media} {Have} a {Positive} {Effect} on {Well}-being? {Evidence} from {Three} {Preregistered} {Field} {Experiments}. \emph{Journal of Technology in Behavioral Science}, \emph{6}(3), 507--514. \url{https://doi.org/10.1007/s41347-020-00189-w}

\leavevmode\hypertarget{ref-przybylskiLargescaleTestGoldilocks2017}{}%
Przybylski, A. K., \& Weinstein, N. (2017). A large-scale test of the {Goldilocks} hypothesis. \emph{Psychological Science}, \emph{28}(2), 204--215. \url{https://doi.org/10.1177/0956797616678438}

\leavevmode\hypertarget{ref-rohrerThinkingClearlyCorrelations2018}{}%
Rohrer, J. M. (2018). Thinking clearly about correlations and causation: {Graphical} causal models for observational data. \emph{Advances in Methods and Practices in Psychological Science}, \emph{24}(2), 251524591774562. \url{https://doi.org/10.1177/2515245917745629}

\leavevmode\hypertarget{ref-schemerImpactInternetSocial2021}{}%
Schemer, C., Masur, P. K., Geiß, S., Müller, P., \& Schäfer, S. (2021). The {Impact} of {Internet} and {Social} {Media} {Use} on {Well}-{Being}: {A} {Longitudinal} {Analysis} of {Adolescents} {Across} {Nine} {Years}. \emph{Journal of Computer-Mediated Communication}, \emph{26}(1), 1--21. \url{https://doi.org/10.1093/jcmc/zmaa014}

\leavevmode\hypertarget{ref-statistaAverageDailyTime2021}{}%
Statista. (2021). \emph{Average daily time spent on social networks by users in the {United} {States} from 2018 to 2022}. Retrieved from \url{https://www.statista.com/statistics/1018324/us-users-daily-social-media-minutes/}

\end{CSLReferences}

\hypertarget{competing-interests}{%
\section{Competing Interests}\label{competing-interests}}

I declare no competing interests.

\hypertarget{supplementary-material}{%
\section{Supplementary Material}\label{supplementary-material}}

All the stimuli, presentation materials, analysis scripts, and a reproducible version of the manuscript can be found on the open science framework (\url{https://osf.io/e47yw/}). The paper also features a companion website where all materials can be accessed (XXX).

\hypertarget{data-accessibility-statement}{%
\section{Data Accessibility Statement}\label{data-accessibility-statement}}

The data are shared on AUSSDA, see \url{https://doi.org/10.11587/28KQNS}, and can be used for scientific purposes only.

\end{document}

### 2021-11-03 preregistration_changes.Rmd
- set-up of repo

---
title: "Items: English"
output:
  html_document:
    toc: TRUE
    toc_float:
      collapsed: FALSE
      toc_depth: 3
---

The following changes were introduced after the preregistration.

## Covariates

When reading the documentation of the data, I thought the following variables were varying, when in fact they were assessed only at one wave, namely the first wave:

- Employment status

## To-Do

- Analyses
	- Missingness
		- Impute missing values
		- Determine amount of missingness
		- Replace values where I know they're 0
	- Add factor scores
	- CFA
		- Add results

### 2021-11-03 analyses.RmD
- updates before sending to Nik

@@
 set.seed(170819)
-knitr::opts_chunk$set(warning=F, echo=F, error=F, message=F, cache = T, 
+knitr::opts_chunk$set(warning=F, echo=T, error=F, message=F, cache = T, 
                       tidy=T, dev = 'pdf', cache.extra = knitr::rand_seed)
@@
 
-```{r r-setup}
+```{r r-setup, results='hide'}
 # install packages
@@
 
-Order dataset. First waves, then id.
+Order dataset. First id, then wave.
 
@@
 d_long %<>%
-  arrange(wave, id)
+  arrange(id, wave)
 ```
@@
 model_soc_med_post <- lmer(soc_med_post ~ (1 | id) + (1 | wave), d_long)
+summary(model_soc_med_post)
 fig_soc_med_post <- make_graph(model_soc_med_post, "Social Media Posting", ll = 1, ul = 5)
@@
 ```{r}
-model_1 <- lmer(life_sat ~ (1 | id) + soc_med_fb + soc_med_ig + soc_med_wa + soc_med_yt + soc_med_tw,
+model_1 <- lmerTest::lmer(life_sat ~ (1 | id) + soc_med_read + soc_med_like_share + soc_med_post + soc_med_fb + soc_med_ig + soc_med_wa + soc_med_yt + soc_med_tw,
                 data = d_long)
@@
 ```{r}
-model_2 <- lmer(life_sat ~ (1 | id) + (1 | wave) + soc_med_read_b + soc_med_read_w + soc_med_like_share_b + soc_med_like_share_w + soc_med_post_b + soc_med_post_w + soc_med_fb_b + soc_med_fb_w + soc_med_ig_b + soc_med_ig_w + soc_med_wa_b + soc_med_wa_w + soc_med_yt_b + soc_med_yt_w + soc_med_tw_b + soc_med_tw_w,
+model_2 <- lmerTest::lmer(life_sat ~ (1 | id) + (1 | wave) + soc_med_read_b + soc_med_read_w + soc_med_like_share_b + soc_med_like_share_w + soc_med_post_b + soc_med_post_w + soc_med_fb_b + soc_med_fb_w + soc_med_ig_b + soc_med_ig_w + soc_med_wa_b + soc_med_wa_w + soc_med_yt_b + soc_med_yt_w + soc_med_tw_b + soc_med_tw_w,
                 data = d_long)

### 2021-11-03 items_english.rmd
- updates before sending to Nik

@@
 
-## Social media use: Reading
+## Social media use
 
-"Wie oft haben Sie letzte Woche die folgenden Aktivitäten in Sozialen Netzwerken gesetzt?"
+"How often during the last week have you engaged in the following activities on social media?"
 
- - "Postings anderer lesen, in denen es um Inhalte zum Coronavirus geht"
+(1 = several times per day; 2 = daily; 3 = several times per week; 4 = weekly; 5 = never)
 
-(1 = Mehrmals täglich; 2 = Einmal täglich; 3 = Mehrmals in der Woche; 4 = Einmal in der Woche; 5 = Gar nicht/nie)
+### Reading
 
-## Social media use: Liking & Sharing
+ - "Reading the posts of others with content on the Coronavirus."
 
-"Wie oft haben Sie letzte Woche die folgenden Aktivitäten in Sozialen Netzwerken gesetzt?"
+### Liking & Sharing
 
- - "Bei Postings anderer zum Coronavirus auf den „Gefällt mir“ Button drücken, „teilen“ oder „retweeten“"
+ - "When seeing Posts on the Coronavirus, I clicked "like", "share" or "retweet".
 
-(nie [1]; an manchen Tagen [2]; mehrmals die Woche [3]; beinahe jeden Tag [4]; täglich [5])
+### Active Postings
 
-## Social media use: Active Postings
+ - "I myself wrote posts on the Coronavirus on Social Media."
 
-"Wie oft haben Sie letzte Woche die folgenden Aktivitäten in Sozialen Netzwerken gesetzt?"
+### Specific platforms
 
- - "Selbst Postings zum Coronavirus in Sozialen Netzwerken verfasst"
-
-(nie [1]; an manchen Tagen [2]; mehrmals die Woche [3]; beinahe jeden Tag [4]; täglich [5])
-
-## Social media use: Specific platforms
-
-"Wie oft in der letzten Woche haben Sie Informationen zur Corona-Krise in den folgenden Sozialen Medien verfolgt?"
+"How often in the last week have you followed information related to the Corona-crisis in the following social media?"
 
@@
 
-(1 = Mehrmals täglich; 2 = Einmal täglich; 3 = Mehrmals in der Woche; 4 = Einmal in der Woche; 5 = Gar nicht/nie)
+(1 = several times per day; 2 = daily; 3 = several times per week; 4 = weekly; 5 = never)
 

### 2021-11-03 manuscript.Rmd
- updates before sending to Nik

@@
 ---
-title             : "No effects of COVID-related social media use on well-being"
-shorttitle        : "COVID-related social media use and well-being"
+title             : "No effects of COVID-19 related social media use on well-being"
+shorttitle        : "COVID-19 related social media use and well-being"
 
@@
 However, because social media are particularly engaging, some find it hard to disconnect and cannot stop 'doomscrolling'.
-In this preregistered study, I investigate whether using social media for COVID-related topics might put personal well-being at risk.
+In this preregistered study, I investigate whether using social media for COVID-19 related topics might put personal well-being at risk.
 To this end, I analyze data from the Austrian Corona Panel Project, which consists of 24 waves with overall 3,018 participants.
 The data were analzed using random effects cross lagged panel models, controlling for several stable and varying covariates.
-Results showed that COVID-related social media use did not meaningfully affect several types of well-being, including life satisfaction, positive affect, and negative effect.
+Results showed that COVID-19 related social media use did not meaningfully affect several types of well-being, including life satisfaction, positive affect, and negative effect.
 This pertains to both passive and active social media use, and all the prominent channels such as Facebook, WhatsApp, or YouTube.
@@
 linenumbers       : yes
-mask              : no
+mask              : yes
 draft             : no
@@
 
-During the COVID pandemic, 
+During the COVID-19 pandemic, 
 <!-- and especially in the beginning,  -->
@@
 In this extraordinary situation, people hence used media excessively, and especially social media were at an all time high [@statistaAverageDailyTime2021].
-A new phenomenon emerged, "termed doomscrolling":
-People could not stop using social media to attain COVID-related news.
+A new phenomenon "termed doomscrolling" emerged:
+People could not stop using social media to attain COVID-19 related news.
 
@@
 Increasingly, it was asked whether such an increase in social media use could still be considered useful and adaptive, or whether it created an additional and new psychological danger for the users' mental health [@kleinDarklySoothingCompulsion2021].
-A study with 6,233 people from Germany found that "[f]requency, duration and diversity of media exposure were positively associated with more symptoms of depression and unspecific and COVID specific anxiety" [@bendauAssociationsCOVID19Related2021].
+A study with 6,233 people from Germany found that "[f]requency, duration and diversity of media exposure were positively associated with more symptoms of depression and unspecific and COVID-19 specific anxiety" [@bendauAssociationsCOVID19Related2021].
 
-As a result, with this study I want to build on this research and investigate the question whether COVID-related social media use during the pandemic affected the users' well-being.
+As a result, with this study I want to build on this research and investigate the question whether COVID-19 related social media use during the pandemic affected the users' well-being.
 To this end I analyze a large-scale panel study from the Austrian Corona Panel Project [@kittelAustrianCoronaPanel2020]. 
@@
 Specifically, I will differentiate between active and passive use.
-I will distinguish (a) _reading_ COVID-related social media use (passive), _posting_ content regarding COVID (active), and _liking and sharing_ COVID-related posts by others (both active and passive).
-Worth noting, this study is not about _general_ social media during times of COVID, but on social media use focused on COVID-related news and interactions.
-For example, posting about the pandemic or retweeting COVID-related posts.
+I will distinguish (a) _reading_ COVID-19 related social media use (passive), _posting_ content regarding COVID (active), and _liking and sharing_ COVID-19 related posts by others (both active and passive).
+Worth noting, this study is not about _general_ social media during times of COVID, but on social media use focused on COVID-19 related news and interactions.
+For example, posting about the pandemic or retweeting COVID-19 related posts.
 
@@
 
-In their study on the relations between media use and mental health during the pandemic, @bendauAssociationsCOVID19Related2021 found that people who used social media as a primary source of information reported on average "significantly more unspecific anxiety and depression [] and significantly more specific COVID-related anxiety symptoms" (p.
+In their study on the relations between media use and mental health during the pandemic, @bendauAssociationsCOVID19Related2021 found that people who used social media as a primary source of information reported on average "significantly more unspecific anxiety and depression [] and significantly more specific COVID-19 related anxiety symptoms" (p.
 288).
@@
 If we refer to standardized effect sizes, according to Cohen [@cohenPowerPrimer1992] small effect sizes start at _r_ = .10.
-However, most meta-analyses find effect sizes below that threshold [@huangTimeSpentSocial2017; @meierComputerMediatedCommunicationSocial2020].
-As a result, I think it's most convincing to expect trivial to small effects also in the case of COVID-related social media use.
+However, most meta-analyses find effect sizes below that threshold [@huangTimeSpentSocial2017; @meierComputerMediatedCommunicationSocial2020; @fergusonThisMetaanalysisScreen2021].
+As a result, I think it's most convincing to expect trivial to small effects also in the case of COVID-19 related social media use.
 
-From a theoretical perspective, why could COVID-related social media use be detrimental?
+From a theoretical perspective, why could COVID-19 related social media use be detrimental?
 Above everything, one can reasonably assume a _direct_ negative effect on well-being, especially on positive or negative effect.
@@
 
-On the other hand, one can make the case that overuse can be beneficial _in times of a pandemic_, even if it's mainly COVID-related.
-Exchanging COVID-related messages with friends via WhatsApp might replace the in-person contact one would have otherwise, but which is logically impossible at that time.
-At times where meaningful and functional activities are explicitly or implicitly prohibited, using social media to exchange about COVID might not be the worst idea.
+On the other hand, one can make the case that overuse can be beneficial _in times of a pandemic_, even if it's mainly COVID-19 related.
+Exchanging COVID-19 related messages with friends via WhatsApp might replace the in-person contact one would have otherwise, but which is logically impossible at that time.
+At times where meaningful and functional activities are explicitly or implicitly prohibited, using social media to exchange about COVID-19 might not be the worst idea.
 In fact, given that nowadays a large number of experts, scientists, and politicians converse directly on social media, one can get first-hand information on the current developments.
-On the other hand, there is of course also much disinformation, and "bingeing" on COVID fake news might also pose risks for impaired well-being.
+On the other hand, there is of course also much disinformation, and "bingeing" on COVID-19 fake news might also pose risks for impaired well-being.
 
 Together, the strongest argument seems to be that the effects of social media on well-being are, on average, generally small at best.
-Because this study looks at only _one part_ of social media use---namely, COVID-related interactions---it is even more focused, and the overall potential of the effects should diminish even further.
-Whether or not using social media for COVID-related aspects is detrimental during a pandemic is also not entirely clear.
+Because this study looks at only _one part_ of social media use---namely, COVID-19 related interactions---it is even more focused, and the overall potential of the effects should diminish even further.
+Whether or not using social media for COVID-19 related aspects is detrimental during a pandemic is also not entirely clear.
 <!-- At least, there does not seem to be a case where we should expect a clear deviation to the negative, despite anecdotes of doomscrolling. -->
-Therefore, I expect to find that the effects of COVID-related news use on well-being will be not be meaningful or practically relevant.
+Therefore, I expect to find that the effects of COVID-19 related news use on well-being will be not be meaningful or practically relevant.
 
-> Hypothesis: The within-person effects of all types of COVID-related social media use on all types of well-being indicators---while controlling for several stable and varying covariates such as sociodemographic variables and psychological dispositions---will be trivial.
+> Hypothesis: The within-person effects of all types of COVID-19 related social media use on all types of well-being indicators---while controlling for several stable and varying covariates such as sociodemographic variables and psychological dispositions---will be trivial.
 
@@
 
-If a heavy user of COVID-related social media news stops using social media altogether, this should have a noticeable impact on their overall well-being.
+If a heavy user of COVID-19 related social media news stops using social media altogether, this should have a noticeable impact on their overall well-being.
 
 What would this mean practically and how could it be operationalized?
-In this study, COVID-related social media use was measured on a 5-point scale, ranging from 1 = _never_ to 5 = _several times a day_. 
+In this study, COVID-19 related social media use was measured on a 5-point scale, ranging from 1 = _never_ to 5 = _several times a day_. 
 Thus, the example from above would imply that a change of five units in social media use should correspond to a noticeable change in well-being.
@@
 However, when controlling for many varying variables, we can be much more certain that we measured causality correctly.
-The aim should still be to collect as many relevant varying and nonvarying third variables as possible, while knowing that absolute certainty can rarely be reached.
+The aim should still be to collect as many relevant varying and nonvarying third variables as possible, while knowing that absolute certainty cannot be reached.
 
@@
 
-[Within vs. between logic]
+Next to including covariates, it is now increasingly understood that causal effects need to be analyzed from an internal, within-person perspective.
+If a specific person changes their media diet, we need to measure how this affects _them_.
+Between person comparison from cross-sectional data, where participants are interviewed only once, cannot provide such data.
+In this study, I will hence differentiate between within-person effects and between-person relations.
 
-[Time Frame / REWB logic]
+One precondition of causality is temporal order. 
+The cause needs to precede the outcome.
+To this end, finding the right interval when causes and effects are measured is crucial.
+For example, if we want to measure the effect of alcohol consumption on driving performance, it makes a big difference if driving performance is measure one minute, one hour, one day, or one week after consumption.
+Finding the right interval is difficult.
+If variables are more stable, longer intervals are needed, and if they fluctuate a lot, shorter intervals are necessary.
+In the case of well-being, to analyze affect we need shorter intervals, while for life satisfaction longer ones.
+Still, finding the right interval is challenging, because especially short intervals are practically hard to implement and require experience sampling measures (also known as in situ measurement or ambulant assessment).
+
+In this study, I will adopt an intermediate perspective.
+I will analyze if, when a person changes their social media diet, are there simultaneous changes in well-being?
+When additionally controlling for both stable and varying covariates, we can then be more sure that the effect is indeed causal.
+This approach was implemented already by several other studies and is considered one of the best practices to analyze causality.
 
@@
 The (anonymous) preregistration can be accessed here: https://osf.io/87b24/?view_only=b2289b6fec214fa88ee75a18d45c18f3.
-Because in this study I analyze data from an already existing large scale data set, the Austrian Corona Panel Project, all of these steps were preregistered _prior_ to accessing the data.
+Because in this study I analyze data from an already existing large scale data set, the Austrian Corona Panel Project, all of these steps were preregistered prior to accessing the data.
 The preregistration was designed on the basis of the panel documentation online [@kittelAustrianCoronaPanel2020].
@@
 In order to participate in the study, the respondents needed to be Austrian residents and had to be at least 14 years old.
-The average age was `r mean(2021 - d_wide$year_birth, na.rm = T) %>% round(0)`, `r mean(d_wide$male, na.rm = T) %>% round(., 2) * 100` percent were male, `r d_wide %>% filter(.$edu_fac %in% c("Hochschulverwandte Lehranstalt oder Kolleg","Bachelor", "Magister / Master / Diplomingenieur / Fachhochschule", "Doktor / PhD")) %>% nrow() / sum(table(d_wide$edu_fac)) %>% divide_by(100)` percent had a University degree, and `r table(d_wide$employment_fac)["Arbeitslos"] / (table(d_wide$employment_fac) %>% sum()) %>% unname() %>% divide_by(100) %>% round(., 0)` percent were currently unemployed.
+The average age was `r mean(2021 - d_wide$year_birth, na.rm = T) %>% round(0)` years, `r mean(d_wide$male, na.rm = T) %>% round(., 2) * 100` percent were male, `r d_wide %>% filter(.$edu_fac %in% c("Hochschulverwandte Lehranstalt oder Kolleg","Bachelor", "Magister / Master / Diplomingenieur / Fachhochschule", "Doktor / PhD")) %>% nrow() / sum(table(d_wide$edu_fac)) %>% divide_by(100)` percent had a University degree, and `r table(d_wide$employment_fac)["Arbeitslos"] / (table(d_wide$employment_fac) %>% sum()) %>% unname() %>% divide_by(100) %>% round(., 0)` percent were currently unemployed.
 <!-- The study received IRB approval from the University of Vienna. -->
@@
 Because the data were analyzed post hoc, no sample size planning on the basis of a priori power analyses was conducted.
-Because the sample is very large, it is well-equipped reliably to detect also small effects, rendering exact post hoc power analysis unnecessary.
-And also because such large sample easily generate significant p-values for very small effects, this study builds on a smallest effect size of interest.
-We will use the interval testing approach as proposed by @dienesUsingBayesGet2014.
-On the basis of our SESOI, we will define a null region.
+The sample is very large, and it is hence well-equipped reliably to detect also small effects, which is why no exact post hoc power analysis were conducted.
+In addition, because such large samples easily generate significant _p_-values even for very small effects, this study builds on a smallest effect size of interest.
+To test the hypotheses, I will use the interval testing approach as proposed by @dienesUsingBayesGet2014.
+On the basis of the SESOI, I will define a null region.
 For well-being, the null region will be between _b_ = -.30 and _b_ = .30.
-If the 95% confidence interval lies completely within the null-region (e.g., _b_ = .15, [95% CI: .05, .25]), the hypothesis that the effect is only trivial is be supported.
+For example, if the 95% confidence interval lies completely within the null-region (e.g., _b_ = .15, [95% CI: .05, .25]), the hypothesis that the effect is only trivial is be supported.
 If the effects interval and the null region overlap (e.g., _b_ = .25, [95% CI: .15, .35]), the hypothesis is not supported and the results are considered inconclusive, whereas a meaningful negative effect is rejected.
-If the confidence falls completely outside the null-region (e.g., _b_ = .4, [95% CI: .35, .45]), the hypothesis is rejected and the existence of a meaningful positive effect supported.
+If the confidence falls completely outside the null-region (e.g., _b_ = .4, [95% CI: .35, .45]), the hypothesis is rejected and the existence of a meaningful positive effect is supported.
 
@@
 All predictors---that is, social media activities and social media channels, within and between-person factors, stably and varying covariates---were included in each of the three models. 
+The data were hierarchical, and responses were nested in participants and waves.
+Nesting in participants allowed to separate within-person effects from between-person relations.
+Nesting in waves allowed to control for general developments, such as general decrease in well-being in the population, for example due to lockdown measures (hence, there was no need additionally to control for specific phases or measures of the lockdown).
+
 The factorial validity of the scales were tested with confirmatory factor analyses (CFA). 
@@
 Because REWB-models cannot model scales as latent variables, to increase precision factor scores exported from the CFAs were used.
-For more information on the analyses, see [companion website](xxx).
+For more information on the analyses, see [companion website](https://xmtra.github.io/Austrian_Corona_Panel_Project/index.html).
 
@@
 In what follows, I briefly list all variables that were collected.
-For a complete list of all items and item characteristics, see [companion website](xxx).
+For a complete list of all items and item characteristics, see [companion website](https://xmtra.github.io/Austrian_Corona_Panel_Project/index.html).
 
@@
 
+All three items were measured on each wave.
+
+### COVID-19 related social media use
+
+The COVID-19 related social media use focused on interaction was measured with three variables (a) reading, (b) liking and sharing, and (c) posting.
+The general introductory question was "How often during the last week have you engaged in the following activities on social media?"
+The three items read as follows: 
+"Reading the posts of others with content on the Coronavirus"; "When seeing Posts on the Coronavirus, I clicked 'like', 'share' or 'retweet'"; "I myself wrote posts on the Coronavirus on Social Media."
+Answer options were 1 (_several times per day_), 2 (_daily_), 3 (_several times per week_), 4 (_weekly_), 5 (_never_).
+The items were inverted for the analyses.
+The variable's average across all waves was _M_ = `r get_specs(model_life_sat)$mean`, ranging from _M_~min~ = `r get_specs(model_life_sat)$min` to _M_~max~ = `r get_specs(model_life_sat)$max`.
+The average standard deviation was _SD_ = `r get_specs(model_life_sat)$sd`.
+
+The COVID-19 related social media use focused on channels was measured with five variables.
+The general introductory question was "How often in the last week have you followed information related to the Corona-crisis in the following social media?"
+The five items were (a) Facebook, (b) Twitter, (c) Instagram, (d) Youtube, (e) WhatsApp.
+Again, the answer options were 1 (_several times per day_), 2 (_daily_), 3 (_several times per week_), 4 (_weekly_), 5 (_never_).
+Again, the items were inverted for the analyses.
+
+All items were measured every second wave.
+However, freshly recruited respondents always answered these questions.
+
 ### Control variables
 
-The effects of COVID-related social media use were controlled for the following stable variables: 
+The effects of COVID-19 related social media use were controlled for the following stable variables: 
 (a) gender (answer options: female, male, diverse), (b) age, (c) education (10 ordinal options), (d) Austria country of birth (yes/no), (e) Austria parents' country of birth (no parent, one parent, both parents).
@@
 When looking at the variables from a descriptive perspective, we see that all well-being measures did not change substantially across the different phases of the pandemic.
-COVID-related media use, however, increased during the beginning of the study and remained stable after approximately six waves.
+COVID-19 related media use, however, increased during the beginning of the study and remained stable after approximately six waves.
 
@@
 Regarding the different types of _communication_, that is reading vs. sharing vs. posting, all within-person effects fell completely within the a-priori defined SESOIs.
-For example, respondents who used social media more frequently than usual to read about COVID-related topics did not show a simultaneous change in life satisfaction (_b_ = `r dat_fig_results_activity %>% filter(dv == "Life satisfaction" & iv == "Reading" & type == "within") %>% select(estimate) %>% round(2)` [95% CI `r dat_fig_results_activity %>% filter(dv == "Life satisfaction" & iv == "Reading" & type == "within") %>% select(conf.low) %>% round(2)`, `r dat_fig_results_activity %>% filter(dv == "Life satisfaction" & iv == "Reading" & type == "within") %>% select(conf.high) %>% round(2)`]).
+For example, respondents who used social media more frequently than usual to read about COVID-19 related topics did not show a simultaneous change in life satisfaction (_b_ = `r dat_fig_results_activity %>% filter(dv == "Life satisfaction" & iv == "Reading" & type == "within") %>% select(estimate) %>% round(2)` [95% CI `r dat_fig_results_activity %>% filter(dv == "Life satisfaction" & iv == "Reading" & type == "within") %>% select(conf.low) %>% round(2)`, `r dat_fig_results_activity %>% filter(dv == "Life satisfaction" & iv == "Reading" & type == "within") %>% select(conf.high) %>% round(2)`]).
 Only one effect did not include zero.
-If people posted more about COVID-related aspects than they usually did, life satisfaction increased (_b_ = `r dat_fig_results_activity %>% filter(dv == "Life satisfaction" & iv == "Posting" & type == "within") %>% select(estimate) %>% round(2)` [95% CI `r dat_fig_results_activity %>% filter(dv == "Life satisfaction" & iv == "Posting" & type == "within") %>% select(conf.low) %>% round(2)`, `r dat_fig_results_activity %>% filter(dv == "Life satisfaction" & iv == "Posting" & type == "within") %>% select(conf.high) %>% round(2)`]).
+If people posted more about COVID-19 related aspects than they usually did, life satisfaction increased (_b_ = `r dat_fig_results_activity %>% filter(dv == "Life satisfaction" & iv == "Posting" & type == "within") %>% select(estimate) %>% round(2)` [95% CI `r dat_fig_results_activity %>% filter(dv == "Life satisfaction" & iv == "Posting" & type == "within") %>% select(conf.low) %>% round(2)`, `r dat_fig_results_activity %>% filter(dv == "Life satisfaction" & iv == "Posting" & type == "within") %>% select(conf.high) %>% round(2)`]).
 However, because the effect did not fall outside our null region, it's likely not large enough to be considered meaningful.
-As a result, the hypothesis was supported for all types of COVID-related social media communication.
+As a result, the hypothesis was supported for all types of COVID-19 related social media communication.
 
 Regarding between-person relations, about which no hypotheses were formulated, only two effects didn't include zero.
-First, respondents who across all waves used social media more frequently than others to read about COVID reported higher levels of positive affect than others (_b_ = `r dat_fig_results_activity %>% filter(dv == "Positive affect" & iv == "Reading" & type == "between") %>% select(estimate) %>% round(2)` [95% CI `r dat_fig_results_activity %>% filter(dv == "Positive affect" & iv == "Reading" & type == "between") %>% select(conf.low) %>% round(2)`, `r dat_fig_results_activity %>% filter(dv == "Positive affect" & iv == "Reading" & type == "between") %>% select(conf.high) %>% round(2)`]).
+First, respondents who across all waves used social media more frequently than others to read about COVID-19 reported higher levels of positive affect than others (_b_ = `r dat_fig_results_activity %>% filter(dv == "Positive affect" & iv == "Reading" & type == "between") %>% select(estimate) %>% round(2)` [95% CI `r dat_fig_results_activity %>% filter(dv == "Positive affect" & iv == "Reading" & type == "between") %>% select(conf.low) %>% round(2)`, `r dat_fig_results_activity %>% filter(dv == "Positive affect" & iv == "Reading" & type == "between") %>% select(conf.high) %>% round(2)`]).
 However, note that this effect still fell completely within the null-region.
 Hence, although positive the effect was considered too small to matter practically.
-Second, respondents who across all waves used social media more frequently than others to read about COVID reported lower levels of negative affect than others (_b_ = `r dat_fig_results_activity %>% filter(dv == "Negative affect" & iv == "Reading" & type == "between") %>% select(estimate) %>% round(2)` [95% CI `r dat_fig_results_activity %>% filter(dv == "Negative affect" & iv == "Reading" & type == "between") %>% select(conf.low) %>% round(2)`, `r dat_fig_results_activity %>% filter(dv == "Negative affect" & iv == "Reading" & type == "between") %>% select(conf.high) %>% round(2)`]).
+Second, respondents who across all waves used social media more frequently than others to read about COVID-19 reported lower levels of negative affect than others (_b_ = `r dat_fig_results_activity %>% filter(dv == "Negative affect" & iv == "Reading" & type == "between") %>% select(estimate) %>% round(2)` [95% CI `r dat_fig_results_activity %>% filter(dv == "Negative affect" & iv == "Reading" & type == "between") %>% select(conf.low) %>% round(2)`, `r dat_fig_results_activity %>% filter(dv == "Negative affect" & iv == "Reading" & type == "between") %>% select(conf.high) %>% round(2)`]).
 The effect was partially outside the null region, hence the effect might be large enough to be practically relevant, and the effect that it's trivial cannot be rejected.
@@
 
-Regarding the COVID-related use of social media _channels_, the results were virtually the same.
-Changes in the frequency of using different social media channels to attain information regarding COVID were unrelated to meaningful changes in well-being.
-For example, respondents who used Facebook more frequently than usual to learn about COVID did not show a simultaneous change in well-being (_b_ = `r dat_fig_results_channel %>% filter(dv == "Life satisfaction" & iv == "Facebook" & type == "within") %>% select(estimate) %>% round(2)` [95% CI `r dat_fig_results_channel %>% filter(dv == "Life satisfaction" & iv == "Facebook" & type == "within") %>% select(conf.low) %>% round(2)`, `r dat_fig_results_channel %>% filter(dv == "Life satisfaction" & iv == "Facebook" & type == "within") %>% select(conf.high) %>% round(2)`]).
+Regarding the COVID-19 related use of social media _channels_, the results were virtually the same.
+Changes in the frequency of using different social media channels to attain information regarding COVID-19 were unrelated to meaningful changes in well-being.
+For example, respondents who used Facebook more frequently than usual to learn about COVID-19 did not show a simultaneous change in well-being (_b_ = `r dat_fig_results_channel %>% filter(dv == "Life satisfaction" & iv == "Facebook" & type == "within") %>% select(estimate) %>% round(2)` [95% CI `r dat_fig_results_channel %>% filter(dv == "Life satisfaction" & iv == "Facebook" & type == "within") %>% select(conf.low) %>% round(2)`, `r dat_fig_results_channel %>% filter(dv == "Life satisfaction" & iv == "Facebook" & type == "within") %>% select(conf.high) %>% round(2)`]).
 Only one effect was substantially larger than zero.
-Respondents who used Instagram more frequently than usual to attain COVID-related news reported slightly lowers simultaneous levels of life satisfaction then usual (_b_ = `r dat_fig_results_channel %>% filter(dv == "Life satisfaction" & iv == "Instagram" & type == "within") %>% select(estimate) %>% round(2)` [95% CI `r dat_fig_results_channel %>% filter(dv == "Life satisfaction" & iv == "Instagram" & type == "within") %>% select(conf.low) %>% round(2)`, `r dat_fig_results_channel %>% filter(dv == "Life satisfaction" & iv == "Instagram" & type == "within") %>% select(conf.high) %>% my_round("b")`]).
+Respondents who used Instagram more frequently than usual to attain COVID-19 related news reported slightly lowers simultaneous levels of life satisfaction then usual (_b_ = `r dat_fig_results_channel %>% filter(dv == "Life satisfaction" & iv == "Instagram" & type == "within") %>% select(estimate) %>% round(2)` [95% CI `r dat_fig_results_channel %>% filter(dv == "Life satisfaction" & iv == "Instagram" & type == "within") %>% select(conf.low) %>% round(2)`, `r dat_fig_results_channel %>% filter(dv == "Life satisfaction" & iv == "Instagram" & type == "within") %>% select(conf.high) %>% my_round("b")`]).
 However, this effect was still completely within the null region, hence not large enough to be considered meaningful.
-In sum, the hypothesis was supported for the COVID-related use of all types of social media channels.
+In sum, the hypothesis was supported for the COVID-19 related use of all types of social media channels.
 
@@
 However, some relations did not included zero.
-For example, respondents who across all waves used Instagram more frequently than others for COVID-related reasons reported lower levels of life satisfaction  (_b_ = `r dat_fig_results_channel %>% filter(dv == "Life satisfaction" & iv == "Instagram" & type == "between") %>% select(estimate) %>% round(2)` [95% CI `r dat_fig_results_channel %>% filter(dv == "Life satisfaction" & iv == "Instagram" & type == "between") %>% select(conf.low) %>% round(2)`, `r dat_fig_results_channel %>% filter(dv == "Life satisfaction" & iv == "Instagram" & type == "between") %>% select(conf.high) %>% my_round("b")`]).
+For example, respondents who across all waves used Instagram more frequently than others for COVID-19 related reasons reported lower levels of life satisfaction  (_b_ = `r dat_fig_results_channel %>% filter(dv == "Life satisfaction" & iv == "Instagram" & type == "between") %>% select(estimate) %>% round(2)` [95% CI `r dat_fig_results_channel %>% filter(dv == "Life satisfaction" & iv == "Instagram" & type == "between") %>% select(conf.low) %>% round(2)`, `r dat_fig_results_channel %>% filter(dv == "Life satisfaction" & iv == "Instagram" & type == "between") %>% select(conf.high) %>% my_round("b")`]).
 Respondents who were more active on WhatsApp compared to others reported slightly lower levels of positive affect  (_b_ = `r dat_fig_results_channel %>% filter(dv == "Positive affect" & iv == "WhatsApp" & type == "between") %>% select(estimate) %>% round(2)` [95% CI `r dat_fig_results_channel %>% filter(dv == "Positive affect" & iv == "WhatsApp" & type == "between") %>% select(conf.low) %>% round(2)`, `r dat_fig_results_channel %>% filter(dv == "Positive affect" & iv == "WhatsApp" & type == "between") %>% select(conf.high) %>% my_round("b")`]).
@@
 
-Using a panel study with a representative sample of the Austrian population that consistent of 24 waves, this study analyzed the effects of COVID-related social media use on well-being.
+Using a panel study with a representative sample of the Austrian population that consistent of 24 waves, this study analyzed the effects of COVID-19 related social media use on well-being.
 A random effect model that separated between person relations from within-person effects and that controlled for several third variables showed that within-person effects were trivial.
-People who used social media more than usual to learn about COVID did not show changes in their well-being.
-As a result, the results imply that COVID-related social media use is irrelevant for people's well-being.
+People who used social media more than usual to learn about COVID-19 did not show changes in their well-being.
+As a result, the results imply that COVID-19 related social media use is irrelevant for people's well-being.
 Other factors among the third variables that were measured revealed much larger effects, implying that the relevance of specific types of social media use for well-being is limited.
@@
 However, @bendauAssociationsCOVID19Related2021 analyzed cross-sectional data on a between-person level, while not controlling for third variables, which does not allow to make causal inferences.
-At the same time, this study is well-aligned with recent studies and meta-analyses from related research questions, which found that the effects of various types of social media use on several well-being indicators is small at best, often too small to matter [@meierComputerMediatedCommunicationSocial2020; @orbenTeenagersScreensSocial2020].
+At the same time, this study is well-aligned with recent studies and meta-analyses from related research questions, which found that the effects of various types of social media use on several well-being indicators is small at best, often too small to matter [@meierComputerMediatedCommunicationSocial2020; @orbenTeenagersScreensSocial2020; @fergusonThisMetaanalysisScreen2021].
 
@@
 
-[SESOI too large/conservative.]
+The predefined SESOI of _b_ = .30 was potentially too large.
+Media use is only one aspect of multiple factors that simultaneously affect well-being.
+Is it realistic that extremely changing only _one_ of these aspect (e.g., by completely stopping the use of social media) should already manifest in a detectable change in well-being?
+Or would it make more sense to say, if you regularly start doing _two_ activities (e.g. regularly exercising and establishing a reading habit) together should then show in improvements to well-being?
+In other words, if the beneficial effect of a particular activity is large enough, if people implement _two_ of them then they should actually feel a difference.
+Practically, this would imply a SESOI half as large as I have defined here, that is _b_ = |.15| for well-being and _b_ = |.075|.
+In this case, this would not make a difference, as even with these more liberal thresholds all but one effect would still be completely in the null region.
+However, at all events future research needs to start a discussion on what effect sizes are considered trivial, and this study is one of the first to provide some concrete guidelines.
 
@@
 
-In this study, COVID-related social media use did not causally affect several indicators of well-being, including life satisfaction, positive affect, and negative affect.
+In this study, COVID-19 related social media use did not causally affect several indicators of well-being, including life satisfaction, positive affect, and negative affect.
 However, factors other than social media use did affect well-being, such as income levels or access to a balcony.
@@
 
-All the stimuli, presentation materials, analysis scripts, and a reproducible version of the manuscript can be found on the companion website (XXX).
+All the stimuli, presentation materials, analysis scripts, and a reproducible version of the manuscript can be found on the companion website (https://xmtra.github.io/Austrian_Corona_Panel_Project/index.html).
 

### 2021-11-03 manuscript.tex
- updates before sending to Nik

@@
   english,
-  man,floatsintext]{apa6}
+  man,mask,floatsintext]{apa6}
 \usepackage{amsmath,amssymb}
@@
 \hypersetup{
-  pdftitle={Analyzing the effects of COVID-19 related social media use on well-being},
-  pdfauthor={Tobias Dienlin1},
+  pdftitle={No effects of COVID-19 related social media use on well-being},
   pdflang={en-EN},
-  pdfkeywords={COVID-19, well-being, affect, life satisfaction, social media use, news use, random effects within between model, panel study, longitudinal},
+  pdfkeywords={COVID, well-being, affect, life satisfaction, social media use, news use, random effects within between model, panel study, longitudinal},
   hidelinks,
@@
 \makeatother
-\keywords{COVID-19, well-being, affect, life satisfaction, social media use, news use, random effects within between model, panel study, longitudinal}
+\keywords{COVID, well-being, affect, life satisfaction, social media use, news use, random effects within between model, panel study, longitudinal}
 \usepackage{lineno}
@@
 
-\title{Analyzing the effects of COVID-19 related social media use on well-being}
-\author{Tobias Dienlin\textsuperscript{1}}
+\title{No effects of COVID-19 related social media use on well-being}
+\author{Blinded for review\textsuperscript{1}}
 \date{}
@@
 
-Tobias Dienlin, Department of Communication, University of Vienna.
+Blinded for review.
 
-Correspondence concerning this article should be addressed to Tobias Dienlin, University of Vienna, Department of Communication, 1090 Vienna, Austria.
-E-mail: \href{mailto:tobias.dienlin@univie.ac.at}{\nolinkurl{tobias.dienlin@univie.ac.at}}
+Correspondence concerning this article should be addressed to Blinded for review, Blinded for review. E-mail: Blinded for review
 
@@
 
-\affiliation{\vspace{0.5cm}\textsuperscript{1} University of Vienna}
+\affiliation{\vspace{0.5cm}\textsuperscript{1} Blinded for review}
 
@@
 The data were analzed using random effects cross lagged panel models, controlling for several stable and varying covariates.
-Results showed that COVID-related social media use did not meaningfully affect several types of well-being, including life satisfaction, positive affect, and negative effect.
+Results showed that COVID-19 related social media use did not meaningfully affect several types of well-being, including life satisfaction, positive affect, and negative effect.
 This pertains to both passive and active social media use, and all the prominent channels such as Facebook, WhatsApp, or YouTube.
@@
 In this extraordinary situation, people hence used media excessively, and especially social media were at an all time high (Statista, 2021).
-A new phenomenon emerged, ``termed doomscrolling'':
-People could not stop using social media to attain COVID-related news.
+A new phenomenon ``termed doomscrolling'' emerged:
+People could not stop using social media to attain COVID-19 related news.
 
@@
 
-As a result, with this study I want to build on this research and investigate the question whether COVID-related social media use during the pandemic affected the users' well-being.
+As a result, with this study I want to build on this research and investigate the question whether COVID-19 related social media use during the pandemic affected the users' well-being.
 To this end I analyze a large-scale panel study from the Austrian Corona Panel Project (Kittel et al., 2020).
@@
 Specifically, I will differentiate between active and passive use.
-I will distinguish (a) \emph{reading} COVID-related social media use (passive), \emph{posting} content regarding COVID (active), and \emph{liking and sharing} COVID-related posts by others (both active and passive).
-Worth noting, this study is not about \emph{general} social media during times of COVID, but on social media use focused on COVID-related news and interactions.
-For example, posting about the pandemic or retweeting COVID-related posts.
+I will distinguish (a) \emph{reading} COVID-19 related social media use (passive), \emph{posting} content regarding COVID (active), and \emph{liking and sharing} COVID-19 related posts by others (both active and passive).
+Worth noting, this study is not about \emph{general} social media during times of COVID, but on social media use focused on COVID-19 related news and interactions.
+For example, posting about the pandemic or retweeting COVID-19 related posts.
 
@@
 If we refer to standardized effect sizes, according to Cohen (Cohen, 1992) small effect sizes start at \emph{r} = .10.
-However, most meta-analyses find effect sizes below that threshold (Huang, 2017; Meier \& Reinecke, 2020).
-As a result, I think it's most convincing to expect trivial to small effects also in the case of COVID-related social media use.
+However, most meta-analyses find effect sizes below that threshold (Ferguson et al., 2021; Huang, 2017; Meier \& Reinecke, 2020).
+As a result, I think it's most convincing to expect trivial to small effects also in the case of COVID-19 related social media use.
 
-From a theoretical perspective, why could COVID-related social media use be detrimental?
+From a theoretical perspective, why could COVID-19 related social media use be detrimental?
 Above everything, one can reasonably assume a \emph{direct} negative effect on well-being, especially on positive or negative effect.
@@
 
-On the other hand, one can make the case that overuse can be beneficial \emph{in times of a pandemic}, even if it's mainly COVID-related.
-Exchanging COVID-related messages with friends via WhatsApp might replace the in-person contact one would have otherwise, but which is logically impossible at that time.
-At times where meaningful and functional activities are explicitly or implicitly prohibited, using social media to exchange about COVID might not be the worst idea.
+On the other hand, one can make the case that overuse can be beneficial \emph{in times of a pandemic}, even if it's mainly COVID-19 related.
+Exchanging COVID-19 related messages with friends via WhatsApp might replace the in-person contact one would have otherwise, but which is logically impossible at that time.
+At times where meaningful and functional activities are explicitly or implicitly prohibited, using social media to exchange about COVID-19 might not be the worst idea.
 In fact, given that nowadays a large number of experts, scientists, and politicians converse directly on social media, one can get first-hand information on the current developments.
-On the other hand, there is of course also much disinformation, and ``bingeing'' on COVID fake news might also pose risks for impaired well-being.
+On the other hand, there is of course also much disinformation, and ``bingeing'' on COVID-19 fake news might also pose risks for impaired well-being.
 
 Together, the strongest argument seems to be that the effects of social media on well-being are, on average, generally small at best.
-Because this study looks at only \emph{one part} of social media use---namely, COVID-related interactions---it is even more focused, and the overall potential of the effects should diminish even further.
-Whether or not using social media for COVID-related aspects is detrimental during a pandemic is also not entirely clear.
-Therefore, I expect to find that the effects of COVID-related news use on well-being will be not be meaningful or practically relevant.
+Because this study looks at only \emph{one part} of social media use---namely, COVID-19 related interactions---it is even more focused, and the overall potential of the effects should diminish even further.
+Whether or not using social media for COVID-19 related aspects is detrimental during a pandemic is also not entirely clear.
+Therefore, I expect to find that the effects of COVID-19 related news use on well-being will be not be meaningful or practically relevant.
 
@@
 
-If a heavy user of COVID-related social media news stops using social media altogether, this should have a noticeable impact on their overall well-being.
+If a heavy user of COVID-19 related social media news stops using social media altogether, this should have a noticeable impact on their overall well-being.
 
 What would this mean practically and how could it be operationalized?
-In this study, COVID-related social media use was measured on a 5-point scale, ranging from 1 = \emph{never} to 5 = \emph{several times a day}.
+In this study, COVID-19 related social media use was measured on a 5-point scale, ranging from 1 = \emph{never} to 5 = \emph{several times a day}.
 Thus, the example from above would imply that a change of five units in social media use should correspond to a noticeable change in well-being.
@@
 
-{[}Within vs.~between logic{]}
+Next to including covariates, it is now increasingly understood that causal effects need to be analyzed from an internal, within-person perspective.
+If a specific person changes their media diet, we need to measure how this affects \emph{them}.
+Between person comparison from cross-sectional data, where participants are interviewed only once, cannot provide such data.
+In this study, I will hence differentiate between within-person effects and between-person relations.
 
-{[}Time Frame / REWB logic{]}
+One precondition of causality is temporal order.
+The cause needs to precede the outcome.
+To this end, finding the right interval when causes and effects are measured is crucial.
+For example, if we want to measure the effect of alcohol consumption on driving performance, it makes a big difference if driving performance is measure one minute, one hour, one day, or one week after consumption.
+Finding the right interval is difficult.
+If variables are more stable, longer intervals are needed, and if they fluctuate a lot, shorter intervals are necessary.
+In the case of well-being, to analyze affect we need shorter intervals, while for life satisfaction longer ones.
+Still, finding the right interval is challenging, because especially short intervals are practically hard to implement and require experience sampling measures (also known as in situ measurement or ambulant assessment).
+
+In this study, I will adopt an intermediate perspective.
+I will analyze if, when a person changes their social media diet, are there simultaneous changes in well-being?
+When additionally controlling for both stable and varying covariates, we can then be more sure that the effect is indeed causal.
+This approach was implemented already by several other studies and is considered one of the best practices to analyze causality.
 
@@
 The (anonymous) preregistration can be accessed here: \url{https://osf.io/87b24/?view_only=b2289b6fec214fa88ee75a18d45c18f3}.
-Because in this study I analyze data from an already existing large scale data set, the Austrian Corona Panel Project, all of these steps were preregistered \emph{prior} to accessing the data.
+Because in this study I analyze data from an already existing large scale data set, the Austrian Corona Panel Project, all of these steps were preregistered prior to accessing the data.
 The preregistration was designed on the basis of the panel documentation online (Kittel et al., 2020).
@@
 In order to participate in the study, the respondents needed to be Austrian residents and had to be at least 14 years old.
-The average age was 42, 49 percent were male, 13.60 percent had a University degree, and 5.07 percent were currently unemployed.
+The average age was 42 years, 49 percent were male, 13.60 percent had a University degree, and 5.07 percent were currently unemployed.
 
 Because the data were analyzed post hoc, no sample size planning on the basis of a priori power analyses was conducted.
-Because the sample is very large, it is well-equipped reliably to detect also small effects, rendering exact post hoc power analysis unnecessary.
-And also because such large sample easily generate significant p-values for very small effects, this study builds on a smallest effect size of interest.
-We will use the interval testing approach as proposed by Dienes (2014).
-On the basis of our SESOI, we will define a null region.
+The sample is very large, and it is hence well-equipped reliably to detect also small effects, which is why no exact post hoc power analysis were conducted.
+In addition, because such large samples easily generate significant \emph{p}-values even for very small effects, this study builds on a smallest effect size of interest.
+To test the hypotheses, I will use the interval testing approach as proposed by Dienes (2014).
+On the basis of the SESOI, I will define a null region.
 For well-being, the null region will be between \emph{b} = -.30 and \emph{b} = .30.
-If the 95\% confidence interval lies completely within the null-region (e.g., \emph{b} = .15, {[}95\% CI: .05, .25{]}), the hypothesis that the effect is only trivial is be supported.
+For example, if the 95\% confidence interval lies completely within the null-region (e.g., \emph{b} = .15, {[}95\% CI: .05, .25{]}), the hypothesis that the effect is only trivial is be supported.
 If the effects interval and the null region overlap (e.g., \emph{b} = .25, {[}95\% CI: .15, .35{]}), the hypothesis is not supported and the results are considered inconclusive, whereas a meaningful negative effect is rejected.
-If the confidence falls completely outside the null-region (e.g., \emph{b} = .4, {[}95\% CI: .35, .45{]}), the hypothesis is rejected and the existence of a meaningful positive effect supported.
+If the confidence falls completely outside the null-region (e.g., \emph{b} = .4, {[}95\% CI: .35, .45{]}), the hypothesis is rejected and the existence of a meaningful positive effect is supported.
 
@@
 All predictors---that is, social media activities and social media channels, within and between-person factors, stably and varying covariates---were included in each of the three models.
+The data were hierarchical, and responses were nested in participants and waves.
+Nesting in participants allowed to separate within-person effects from between-person relations.
+Nesting in waves allowed to control for general developments, such as general decrease in well-being in the population, for example due to lockdown measures (hence, there was no need additionally to control for specific phases or measures of the lockdown).
+
 The factorial validity of the scales were tested with confirmatory factor analyses (CFA).
@@
 Because REWB-models cannot model scales as latent variables, to increase precision factor scores exported from the CFAs were used.
-For more information on the analyses, see \href{xxx}{companion website}.
+For more information on the analyses, see \href{https://xmtra.github.io/Austrian_Corona_Panel_Project/index.html}{companion website}.
 
@@
 In what follows, I briefly list all variables that were collected.
-For a complete list of all items and item characteristics, see \href{xxx}{companion website}.
+For a complete list of all items and item characteristics, see \href{https://xmtra.github.io/Austrian_Corona_Panel_Project/index.html}{companion website}.
 
@@
 
+All three items were measured on each wave.
+
+\hypertarget{covid-19-related-social-media-use}{%
+\subsubsection{COVID-19 related social media use}\label{covid-19-related-social-media-use}}
+
+The COVID-19 related social media use focused on interaction was measured with three variables (a) reading, (b) liking and sharing, and (c) posting.
+The general introductory question was ``How often during the last week have you engaged in the following activities on social media?''
+The three items read as follows:
+``Reading the posts of others with content on the Coronavirus''; ``When seeing Posts on the Coronavirus, I clicked `like,' `share' or `retweet'\,''; ``I myself wrote posts on the Coronavirus on Social Media.''
+Answer options were 1 (\emph{several times per day}), 2 (\emph{daily}), 3 (\emph{several times per week}), 4 (\emph{weekly}), 5 (\emph{never}).
+The items were inverted for the analyses.
+The variable's average across all waves was \emph{M} = 6.59, ranging from \emph{M}\textsubscript{min} = 6.39 to \emph{M}\textsubscript{max} = 6.79.
+The average standard deviation was \emph{SD} = 1.68.
+
+The COVID-19 related social media use focused on channels was measured with five variables.
+The general introductory question was ``How often in the last week have you followed information related to the Corona-crisis in the following social media?''
+The five items were (a) Facebook, (b) Twitter, (c) Instagram, (d) Youtube, (e) WhatsApp.
+Again, the answer options were 1 (\emph{several times per day}), 2 (\emph{daily}), 3 (\emph{several times per week}), 4 (\emph{weekly}), 5 (\emph{never}).
+Again, the items were inverted for the analyses.
+
+All items were measured every second wave.
+However, freshly recruited respondents always answered these questions.
+
 \hypertarget{control-variables}{%
@@
 
-The effects of COVID-related social media use were controlled for the following stable variables:
+The effects of COVID-19 related social media use were controlled for the following stable variables:
 (a) gender (answer options: female, male, diverse), (b) age, (c) education (10 ordinal options), (d) Austria country of birth (yes/no), (e) Austria parents' country of birth (no parent, one parent, both parents).
@@
 Regarding the different types of \emph{communication}, that is reading vs.~sharing vs.~posting, all within-person effects fell completely within the a-priori defined SESOIs.
-For example, respondents who used social media more frequently than usual to read about COVID-related topics did not show a simultaneous change in life satisfaction (\emph{b} = -0.03 {[}95\% CI -0.09, 0.02{]}).
+For example, respondents who used social media more frequently than usual to read about COVID-19 related topics did not show a simultaneous change in life satisfaction (\emph{b} = -0.03 {[}95\% CI -0.09, 0.02{]}).
 Only one effect did not include zero.
-If people posted more about COVID-related aspects than they usually did, life satisfaction increased (\emph{b} = 0.11 {[}95\% CI 0.01, 0.2{]}).
+If people posted more about COVID-19 related aspects than they usually did, life satisfaction increased (\emph{b} = 0.11 {[}95\% CI 0.01, 0.2{]}).
 However, because the effect did not fall outside our null region, it's likely not large enough to be considered meaningful.
-As a result, the hypothesis was supported for all types of COVID-related social media communication.
+As a result, the hypothesis was supported for all types of COVID-19 related social media communication.
 
 Regarding between-person relations, about which no hypotheses were formulated, only two effects didn't include zero.
-First, respondents who across all waves used social media more frequently than others to read about COVID reported higher levels of positive affect than others (\emph{b} = 0.05 {[}95\% CI 0.01, 0.09{]}).
+First, respondents who across all waves used social media more frequently than others to read about COVID-19 reported higher levels of positive affect than others (\emph{b} = 0.05 {[}95\% CI 0.01, 0.09{]}).
 However, note that this effect still fell completely within the null-region.
 Hence, although positive the effect was considered too small to matter practically.
-Second, respondents who across all waves used social media more frequently than others to read about COVID reported lower levels of negative affect than others (\emph{b} = 0.01 {[}95\% CI -0.02, 0.04{]}).
+Second, respondents who across all waves used social media more frequently than others to read about COVID-19 reported lower levels of negative affect than others (\emph{b} = 0.01 {[}95\% CI -0.02, 0.04{]}).
 The effect was partially outside the null region, hence the effect might be large enough to be practically relevant, and the effect that it's trivial cannot be rejected.
@@
 
-Regarding the COVID-related use of social media \emph{channels}, the results were virtually the same.
-Changes in the frequency of using different social media channels to attain information regarding COVID were unrelated to meaningful changes in well-being.
-For example, respondents who used Facebook more frequently than usual to learn about COVID did not show a simultaneous change in well-being (\emph{b} = 0.03 {[}95\% CI -0.03, 0.1{]}).
+Regarding the COVID-19 related use of social media \emph{channels}, the results were virtually the same.
+Changes in the frequency of using different social media channels to attain information regarding COVID-19 were unrelated to meaningful changes in well-being.
+For example, respondents who used Facebook more frequently than usual to learn about COVID-19 did not show a simultaneous change in well-being (\emph{b} = 0.03 {[}95\% CI -0.03, 0.1{]}).
 Only one effect was substantially larger than zero.
-Respondents who used Instagram more frequently than usual to attain COVID-related news reported slightly lowers simultaneous levels of life satisfaction then usual (\emph{b} = -0.08 {[}95\% CI -0.15, \textgreater{} -0.01{]}).
+Respondents who used Instagram more frequently than usual to attain COVID-19 related news reported slightly lowers simultaneous levels of life satisfaction then usual (\emph{b} = -0.08 {[}95\% CI -0.15, \textgreater{} -0.01{]}).
 However, this effect was still completely within the null region, hence not large enough to be considered meaningful.
-In sum, the hypothesis was supported for the COVID-related use of all types of social media channels.
+In sum, the hypothesis was supported for the COVID-19 related use of all types of social media channels.
 
@@
 However, some relations did not included zero.
-For example, respondents who across all waves used Instagram more frequently than others for COVID-related reasons reported lower levels of life satisfaction (\emph{b} = -0.11 {[}95\% CI -0.19, -0.03{]}).
+For example, respondents who across all waves used Instagram more frequently than others for COVID-19 related reasons reported lower levels of life satisfaction (\emph{b} = -0.11 {[}95\% CI -0.19, -0.03{]}).
 Respondents who were more active on WhatsApp compared to others reported slightly lower levels of positive affect (\emph{b} = -0.03 {[}95\% CI -0.07, \textgreater{} -0.01{]}).
@@
 However, Bendau et al. (2021) analyzed cross-sectional data on a between-person level, while not controlling for third variables, which does not allow to make causal inferences.
-At the same time, this study is well-aligned with recent studies and meta-analyses from related research questions, which found that the effects of various types of social media use on several well-being indicators is small at best, often too small to matter (Meier \& Reinecke, 2020; Orben, 2020).
+At the same time, this study is well-aligned with recent studies and meta-analyses from related research questions, which found that the effects of various types of social media use on several well-being indicators is small at best, often too small to matter (Ferguson et al., 2021; Meier \& Reinecke, 2020; Orben, 2020).
 
@@
 
-{[}SESOI too large/conservative.{]}
+The predefined SESOI of \emph{b} = .30 was potentially too large.
+Media use is only one aspect of multiple factors that simultaneously affect well-being.
+Is it realistic that extremely changing only \emph{one} of these aspect (e.g., by completely stopping the use of social media) should already manifest in a detectable change in well-being?
+Or would it make more sense to say, if you regularly start doing \emph{two} activities (e.g.~regularly exercising and establishing a reading habit) together should then show in improvements to well-being?
+In other words, if the beneficial effect of a particular activity is large enough, if people implement \emph{two} of them then they should actually feel a difference.
+Practically, this would imply a SESOI half as large as I have defined here, that is \emph{b} = \textbar.15\textbar{} for well-being and \emph{b} = \textbar.075\textbar.
+In this case, this would not make a difference, as even with these more liberal thresholds all but one effect would still be completely in the null region.
+However, at all events future research needs to start a discussion on what effect sizes are considered trivial, and this study is one of the first to provide some concrete guidelines.
 
@@
 \url{https://doi.org/doi:10.31887/DCNS.2020.22.2/tdienlin}
+
+\leavevmode\hypertarget{ref-fergusonThisMetaanalysisScreen2021}{}%
+Ferguson, C.
+J., Kaye, L.
+K., Branley-Bell, D., Markey, P., Ivory, J.
+D., Klisanin, D., \ldots{} Wilson, J.
+(2021).
+Like this meta-analysis: {Screen} media and mental health.
+\emph{Professional Psychology: Research and Practice}.
+\url{https://doi.org/10.1037/pro0000426}
 
@@
 
-All the stimuli, presentation materials, analysis scripts, and a reproducible version of the manuscript can be found on the open science framework (\url{https://osf.io/e47yw/}).
-The paper also features a companion website where all materials can be accessed (XXX).
+All the stimuli, presentation materials, analysis scripts, and a reproducible version of the manuscript can be found on the companion website (\url{https://xmtra.github.io/Austrian_Corona_Panel_Project/index.html}).
 

### 2021-11-05 analyses.RmD
- implemented feedback from Nick - updated analyses - provided additional plots for control variables

@@
     toc_depth: 2
-    code_folding: hide
+    # code_folding: hide
 bibliography      : 
@@
                       tidy=T, dev = 'pdf', cache.extra = knitr::rand_seed)
-options(digits = 3, width = 120)
+options(digits = 3, width = 140, scipen=999)
 ```
@@
 # define packages
-packages <- c("corrplot", "devtools", "GGally", "ggplot2", "gridExtra", "haven", "kableExtra", "knitr", 
-              "lavaan", "lme4", "magrittr", "mice", "MVN", "PerFit", "psych", "quanteda.textstats",
+packages <- c("brms", "corrplot", "devtools", "GGally", "ggplot2", 
+              "gridExtra", "haven", "kableExtra", "knitr", 
+              "lavaan", "lme4", "magrittr", "mice", "MVN", "PerFit", 
+              "psych", "quanteda.textstats",
               "pwr", "semPlot", "semTools", "tidyverse")
@@
 
-Make graph of variables' development.
+Extract characteristics from fitted lmer models.
 
 ```{r}
-make_graph <- function(model, title, ll, ul, line = FALSE, labels = FALSE){
-  dat <- coefficients(model)$wave %>% 
+get_specs <- function(model){
+    # Get mean, max, and min values
+
+    dat <- coefficients(model)$wave
+    specs <- data.frame(
+    sd = attr(VarCorr(model), "sc"),
+    min = min(dat),
+    max = max(dat),
+    mean = mean(dat$`(Intercept)`)
+  )
+}
+```
+
+Get data from fitted lmer objects
+
+```{r}
+get_dat <- function(model){
+  coefficients(model)$wave %>% 
   tibble::rownames_to_column("wave") %>% 
@@
          value = as.numeric(.$value)) 
+} 
+```
+
+Make graph of variables' development.
+
+```{r}
+make_graph <- function(model, title, ll, ul, line = FALSE, labels = FALSE, lmer=TRUE, legend=TRUE){
+
+    if(isTRUE(lmer)){
+      dat <- get_dat(model)
+    } else{
+    dat <- model
+  }
 
   graph <- 
-    ggplot(dat, aes(wave, value)) +
-    theme_bw() +
-    theme(text = element_text(size = 8)) + 
-    geom_bar(data = dat, aes(wave, value), stat = "identity") +
+    ggplot(dat, aes(wave, value, color = dimension)) +
+    {if(legend) theme(
+      legend.position = c(0.85, 0.8),
+      legend.title = element_blank()
+      )} + 
+    {if(!legend) theme(
+      legend.position = "none"
+    )} +
+    theme(axis.title.x=element_blank()) +
     coord_cartesian(ylim = c(ll, ul)) +
-    ggtitle(title)
-
-  if(line) {graph <- graph + 
-    geom_smooth(se = FALSE, method = 'loess', size = 1.5)}
-
-  if(labels) {graph <- graph +
-    geom_text(aes(label = round(value, 2)), hjust = 1.2, color = "white",
-              fontface = "bold", angle = 90)
-  }
-
+    ggtitle(title) +
+    {if(line) geom_smooth(se = FALSE, method = 'loess')} +
+    # geom_point() +
+    scale_color_manual(values=c("dodgerblue3", "deepskyblue2", "magenta2", "green2", "red"))
   graph
@@
 ```{r}
-make_graph_res <- function(data){
+make_graph_res <- function(data, sesoi = TRUE){
   ggplot(data, aes(x = estimate, y = iv)) + 
@@
              ) +
-  geom_vline(data=filter(data, dv=="Life satisfaction"), aes(xintercept=.3), colour="darkgrey", linetype = "dashed") + 
-    geom_vline(data=filter(data, dv=="Life satisfaction"), aes(xintercept=-.3), colour="darkgrey", linetype = "dashed") + 
-  geom_vline(data=filter(data, dv!="Life satisfaction"), aes(xintercept=.15), colour="darkgrey", linetype = "dashed") + 
-    geom_vline(data=filter(data, dv!="Life satisfaction"), aes(xintercept=-.15), colour="darkgrey", linetype = "dashed") + 
+  {if(sesoi) geom_vline(data=filter(data, dv=="Life satisfaction"), aes(xintercept=.3), colour="darkgrey", linetype = "dashed")} + 
+  {if(sesoi) geom_vline(data=filter(data, dv=="Life satisfaction"), aes(xintercept=-.3), colour="darkgrey", linetype = "dashed")} + 
+  {if(sesoi) geom_vline(data=filter(data, dv!="Life satisfaction"), aes(xintercept=.15), colour="darkgrey", linetype = "dashed")} + 
+  {if(sesoi) geom_vline(data=filter(data, dv!="Life satisfaction"), aes(xintercept=-.15), colour="darkgrey", linetype = "dashed")} +
   theme(
@@
              cols = vars(dv),
-             scales = "free") 
+             scales = "free",
+             space = "free_y") 
 }
@@
 doc <- read_csv("documentation/10094_vi_de_v4_0.csv")
-# d_raw <- read_spss("data/10094_da_de_v4_0.zsav")
 d_raw <- read_csv("data/10094_da_de_v4_0.csv")
@@
     born_aus = 2 - .$born_aus,
+    home_owner = 2 - .$home_owner,
     employment_fac = factor(.$employment, 
-                           labels = c("Unselbststaendig erwerbstaetig (auch Lehrling), Privatwirtschaft", 
-"Unselbststaendig erwerbstaetig (auch Lehrling), Oeffentlicher Dienst",
-"Selbststaendig erwerbstaetig",
-"Arbeitslos",
-"Pensionist*in",
-"Ausschliesslich haushaltsfuehrend",
-"Schueler/in, Student/in, Bildungskarenz",
-"Sonstiges (Dauerhaft arbeitsunfaehig oder Praesenz-/Zivildiener)",
-"In Elternkarenz")),
+                           # labels = c("Unselbststaendig erwerbstaetig (auch Lehrling), Privatwirtschaft",
+                           #            "Unselbststaendig erwerbstaetig (auch Lehrling), Oeffentlicher Dienst",
+                           #            "Selbststaendig erwerbstaetig",
+                           #            "Arbeitslos",
+                           #            "Pensionist*in",
+                           #            "Ausschliesslich haushaltsfuehrend",
+                           #            "Schueler/in, Student/in, Bildungskarenz",
+                           #            "Sonstiges (Dauerhaft arbeitsunfaehig oder Praesenz-/Zivildiener)",
+                           #            "In Elternkarenz")
+                           labels = c("Employed industrie",
+                                      "Employed public service",
+                                      "Self-employed",
+                                      "Unemployed",
+                                      "Retired",
+                                      "Housekeeping",
+                                      "Student",
+                                      "Incapacitated",
+                                      "Parental Leave")
+                           ),
     edu_fac = factor(.$edu,
-                     labels = c("Volksschule oder weniger", 
-                                "Hauptschule oder AHS Unterstufe",
-                                "Polytechnikum, BMS (Fachschule, z.B. HASCH)",
-                                "Lehre, Berufsschule",
-                                "AHS mit Matura",
-                                "BHS mit Matura (z.B. HTL, HAK, HBLA etc.)",
-                                "Hochschulverwandte Lehranstalt oder Kolleg",
+                     # labels = c("Volksschule oder weniger", 
+                     #            "Hauptschule oder AHS Unterstufe",
+                     #            "Polytechnikum, BMS (Fachschule, z.B. HASCH)",
+                     #            "Lehre, Berufsschule",
+                     #            "AHS mit Matura",
+                     #            "BHS mit Matura (z.B. HTL, HAK, HBLA etc.)",
+                     #            "Hochschulverwandte Lehranstalt oder Kolleg",
+                     #            "Bachelor",
+                     #            "Magister / Master / Diplomingenieur / Fachhochschule",
+                     #            "Doktor / PhD"),
+                     labels = c("No degree", 
+                                "Middle school",
+                                "Vocational school",
+                                "Technical school",
+                                "High school",
+                                "Applied high school",
+                                "State college",
                                 "Bachelor",
-                                "Magister / Master / Diplomingenieur / Fachhochschule",
-"Doktor / PhD")))
+                                "Master",
+                                "PhD")
+                     ))
 ```
@@
 ```{r}
-# recode such that higher values imply more strenght / align with wording
+# recode such that higher values imply more strength / align with wording
 d_long %<>% 
-  mutate_at(vars(med_txt_kro:med_vid_pri, health, sat_dem), 
+  mutate_at(vars(med_txt_kro:med_vid_pri, health, sat_dem, soc_med_fb:soc_med_post), 
             funs(recode(., `1` = 5L, `2` = 4L, `3` = 3L, `4` = 2L, `5` = 1L))) %>% 
@@
 
-Create mean values for each person, and unique changes per wave
-
-```{r}
-d_long %<>% 
-  group_by(id) %>% 
-  mutate_at(vars(health:loc_cntrl_int_4), funs(b = mean(., na.rm = TRUE), w = . - mean(., na.rm = TRUE))) %>% 
-  ungroup()
-```
-
 Create means for scales
@@
          loc_cntrl_int_m = rowMeans(select(., loc_cntrl_int_1 : loc_cntrl_int_4)))
+```
+
+Create mean values for each person (between), and unique changes per wave (within).
+
+```{r}
+d_long %<>% 
+  group_by(id) %>% 
+  mutate_at(vars(health:loc_cntrl_int_4, loc_cntrl_int_m), funs(b = mean(., na.rm = TRUE), w = . - mean(., na.rm = TRUE))) %>% 
+  ungroup()
 ```
@@
 # measures
-## life satisfaction
+## well-being
+### life satisfaction
 
@@
 model_life_sat <- lmer(life_sat ~ (1 | id) + (1 | wave), d_long)
-fig_life_sat <- make_graph(model_life_sat, "Life satisfaction", ll = 0, ul = 10)
-fig_life_sat
+summary(model_life_sat)
 ```
 
-## positive affect
+### positive affect
 
@@
 model_aff_pos <- lmer(aff_pos_m ~ (1 | id) + (1 | wave), d_long)
-fig_aff_pos <- make_graph(model_aff_pos, "Positive affect", ll = 1, ul = 5)
-fig_aff_pos
+summary(model_aff_pos)
 ```
 
-## negative affect
+### negative affect
 
@@
 model_aff_neg <- lmer(aff_neg_m ~ (1 | id) + (1 | wave), d_long)
-fig_aff_neg <- make_graph(model_aff_neg, "Negative affect", ll = 1, ul = 5)
-fig_aff_neg
+summary(model_aff_neg)
 ```
 
-## Social media reading
+## Social media use
+### reading
 
@@
 model_soc_med_read <- lmer(soc_med_read ~ (1 | id) + (1 | wave), d_long)
-fig_soc_med_read <- make_graph(model_soc_med_read, "Social Media Reading", ll = 1, ul = 5)
-fig_soc_med_read
+summary(model_soc_med_read)
 ```
 
-## Social media liking & sharing
+### liking & sharing
 
@@
 model_soc_med_like_share <- lmer(soc_med_like_share ~ (1 | id) + (1 | wave), d_long)
-fig_soc_med_like_share <- make_graph(model_soc_med_like_share, "Soc. Media Like & Share", ll = 1, ul = 5)
-fig_soc_med_like_share
+summary(model_soc_med_like_share)
 ```
 
-## Social media posting
+### posting
 
@@
 summary(model_soc_med_post)
-fig_soc_med_post <- make_graph(model_soc_med_post, "Social Media Posting", ll = 1, ul = 5)
-fig_soc_med_post
+```
+
+## Social media channels
+### Facebook
+
+```{r}
+model_soc_med_fb <- lmer(soc_med_fb ~ (1 | id) + (1 | wave), d_long)
+summary(model_soc_med_fb)
+```
+
+### Twitter
+
+```{r}
+model_soc_med_tw <- lmer(soc_med_tw ~ (1 | id) + (1 | wave), d_long)
+summary(model_soc_med_tw)
+
+# Bayesian Approach
+# model_soc_med_tw <- brm(soc_med_tw ~ (1 | id) + (1 | wave), d_long
+# Also throws a warning
+# model_soc_med_tw <- brm(soc_med_tw ~ (1 | id) + (1 | wave), d_long, control=list(adapt_delta=0.82))
+# Currently doesn't work, likely for lack of memory
+```
+
+Interestingly, lme4 throws a warning, likely because there is too little variance across waves, while little variation measure itself. Let's inspect the raw means.
+
+```{r}
+soc_med_tw_m <- d_long %>% group_by(wave) %>% summarise(value = mean(soc_med_tw, na.rm = T))
+soc_med_tw_m
+```
+
+Inspecting the means shows that the data are intact, only that the estimation didn't fully work.
+
+### Instagram
+
+```{r}
+model_soc_med_ig <- lmer(soc_med_ig ~ (1 | id) + (1 | wave), d_long)
+summary(model_soc_med_ig)
+```
+
+### WhatsApp
+
+```{r}
+model_soc_med_wa <- lmer(soc_med_wa ~ (1 | id) + (1 | wave), d_long)
+summary(model_soc_med_wa)
+```
+
+### YouTube
+
+```{r}
+model_soc_med_yt <- lmer(soc_med_yt ~ (1 | id) + (1 | wave), d_long)
+summary(model_soc_med_yt)
 ```
@@
 
-Save everything as a figure
+Display the results in a combined figure.
 
 ```{r}
-fig_desc <- grid.arrange(fig_life_sat, fig_aff_pos, fig_aff_neg, 
-                         fig_soc_med_read, fig_soc_med_like_share, fig_soc_med_post,
-                         nrow = 2)
+fig_desc_dat <- data.frame(
+  rbind(
+    data.frame(type = "Life satisfaction", dimension = "Life satisfaction", get_dat(model_life_sat)),
+    data.frame(type = "Affect", dimension = "Positive", get_dat(model_aff_pos)),
+    data.frame(type = "Affect", dimension = "Negative", get_dat(model_aff_neg)),
+    data.frame(type = "Social media use", dimension = "Reading", get_dat(model_soc_med_read)),
+    data.frame(type = "Social media use", dimension = "Liking & Sharing", get_dat(model_soc_med_like_share)),
+    data.frame(type = "Social media use", dimension = "Posting", get_dat(model_soc_med_post)),
+    data.frame(type = "Social media channel", dimension = "Facebook", get_dat(model_soc_med_fb)),
+    data.frame(type = "Social media channel", dimension = "Twitter", get_dat(model_soc_med_tw)),
+    data.frame(type = "Social media channel", dimension = "Instagram", get_dat(model_soc_med_ig)),
+    data.frame(type = "Social media channel", dimension = "WhatsApp", get_dat(model_soc_med_wa)),
+    data.frame(type = "Social media channel", dimension = "YouTube", get_dat(model_soc_med_yt))
+  ) %>% 
+    mutate(
+      type = factor(.$type, levels = c("Life satisfaction", "Affect", "Social media use", "Social media channel")),
+      dimension = factor(.$dimension, levels = c("Life satisfaction", "Positive", "Negative", "Reading", "Liking & Sharing", "Posting", "Facebook", "Twitter", "Instagram", "WhatsApp", "YouTube"))
+      )
+)
+
+fig_desc_life_sat <- make_graph(
+  fig_desc_dat %>% filter(type == "Life satisfaction"), 
+  title = "Life satisfaction", 
+  ll = 0, ul = 10, 
+  lmer = FALSE, 
+  line = TRUE,
+  legend = FALSE
+  )
+
+fig_desc_aff <- make_graph(
+  fig_desc_dat %>% filter(type == "Affect"), 
+  title = "Affect", 
+  ll = 1, ul = 5, 
+  lmer = FALSE, 
+  line = TRUE
+  )
+
+fig_desc_soc_med_use <- make_graph(
+  fig_desc_dat %>% filter(type == "Social media use"), 
+  title = "Social media use", 
+  ll = 1, ul = 5, 
+  lmer = FALSE, 
+  line = TRUE
+  )
+
+fig_desc_soc_med_channel <- make_graph(
+  fig_desc_dat %>% filter(type == "Social media channel"), 
+  title = "Social media channel", 
+  ll = 1, ul = 5, 
+  lmer = FALSE, 
+  line = TRUE
+  )
+
+fig_desc <- grid.arrange(fig_desc_life_sat, fig_desc_aff, 
+                         fig_desc_soc_med_use, fig_desc_soc_med_channel,
+                         nrow = 2, ncol = 2)
+
 ggsave("figures/fig_descriptives.pdf", 
-       width = 7, height = 5,
+       width = 10, height = 6,
        plot = fig_desc)
+```
+
+## table
+
+```{r}
+tab_desc_dat <- rbind(
+    "Life satisfaction" = get_specs(model_life_sat),
+    "Positive affect" = get_specs(model_aff_pos),
+    "Negative affect" = get_specs(model_aff_neg),
+    "Read" = get_specs(model_soc_med_read),
+    "Like & share" = get_specs(model_soc_med_like_share),
+    "Posting" = get_specs(model_soc_med_post),
+    "Facebook" = get_specs(model_soc_med_fb),
+    "Twitter" = c(sd = get_specs(model_soc_med_tw)$sd,
+                  min = min(soc_med_tw_m$value, na.rm = TRUE),
+                  max = max(soc_med_tw_m$value, na.rm = TRUE),
+                  mean = mean(soc_med_tw_m$value, na.rm = TRUE)
+                  ),
+    "Instagram" = get_specs(model_soc_med_ig),
+    "WhatsApp" = get_specs(model_soc_med_wa),
+    "YouTube" = get_specs(model_soc_med_yt)
+    )
 ```
@@
 
+First, without control variables.
+
 ```{r}
-model_1 <- lmerTest::lmer(life_sat ~ (1 | id) + soc_med_read + soc_med_like_share + soc_med_post + soc_med_fb + soc_med_ig + soc_med_wa + soc_med_yt + soc_med_tw,
+model_life_sat_lmer_nco <- lmerTest::lmer(life_sat ~ 
+                            (1 | id) + (1 | wave) + 
+                            soc_med_read_w + soc_med_like_share_w + soc_med_post_w + 
+                            soc_med_fb_w + soc_med_ig_w + soc_med_wa_w + soc_med_yt_w + soc_med_tw_w +
+                            soc_med_read_b + soc_med_like_share_b + soc_med_post_b + 
+                            soc_med_fb_b + soc_med_ig_b + soc_med_wa_b + soc_med_yt_b + soc_med_tw_b,
                 data = d_long)
-summary(model_1)
+summary(model_life_sat_lmer_nco)
 ```
 
-Now with diff. between within and between effects.
+Now with control variables.
 
 ```{r}
-model_2 <- lmerTest::lmer(life_sat ~ (1 | id) + (1 | wave) + soc_med_read_b + soc_med_read_w + soc_med_like_share_b + soc_med_like_share_w + soc_med_post_b + soc_med_post_w + soc_med_fb_b + soc_med_fb_w + soc_med_ig_b + soc_med_ig_w + soc_med_wa_b + soc_med_wa_w + soc_med_yt_b + soc_med_yt_w + soc_med_tw_b + soc_med_tw_w,
-                data = d_long)
-summary(model_2)
+model_life_sat_lmer <- lmerTest::lmer(life_sat ~
+                                        (1 | id) + (1 | wave) + 
+                                        soc_med_read_w + soc_med_like_share_w + soc_med_post_w  + 
+                                        soc_med_fb_w + soc_med_ig_w + soc_med_wa_w + soc_med_yt_w + soc_med_tw_w +
+                                        soc_med_read_b + soc_med_like_share_b + soc_med_post_b + 
+                                        soc_med_fb_b + soc_med_ig_b + soc_med_wa_b + soc_med_yt_b + soc_med_tw_b +  
+                                        age + male + born_aus + born_aus_prnts + edu_fac + employment_fac + health_b + 
+                                        res_vienna + acc_bal + acc_gar + home_sqm + 
+                                        # hh_income_b + # large dropout (900 obs)
+                                        # hh_adults + # too much dropout
+                                        # hh_child18 + hh_child17 + hh_child14 + hh_child5 + hh_child2 + hh_oldfam + hh_outfam + hh_partner + 
+                                        # home_owner + # too much dropout
+                                        med_txt_kro_b + med_txt_sta_b + med_txt_pre_b + med_txt_oes_b + med_txt_kur_b + med_txt_slz_b + med_txt_son_b + 
+                                        med_vid_orf_b + med_vid_pri_b + 
+                                        risk_prop_b + loc_cntrl_int_m_b + 
+                                        act_wrk_w + act_spo_w + act_frn_w + act_sho_w + act_pet_w + 
+                                        act_wrk_b + act_spo_b + act_frn_b + act_sho_b + act_pet_b +
+                                        sat_dem_w + sat_dem_b 
+                                      , data = d_long)
+summary(model_life_sat_lmer)
 ```
 
-Now with control variables
-
 ```{r}
-model_life_sat_lmer <- lmerTest::lmer(life_sat ~ (1 | id) + (1 | wave) + soc_med_read_b + soc_med_read_w + soc_med_like_share_b + soc_med_like_share_w + soc_med_post_b + soc_med_post_w + soc_med_fb_b + soc_med_fb_w + soc_med_ig_b + soc_med_ig_w + soc_med_wa_b + soc_med_wa_w + soc_med_yt_b + soc_med_yt_w + soc_med_tw_b + soc_med_tw_w + age + male + edu + born_aus + born_aus_prnts + res_vienna + edu + acc_bal + acc_gar + employment_fac
 # + health_b + act_wrk_b + act_spo_b + act_frn_b + act_sho_b + act_pet_b + sat_dem_b + hh_income_b + med_txt_kro_b + med_txt_sta_b + med_txt_pre_b + med_txt_oes_b + med_txt_kur_b + med_txt_slz_b + med_txt_son_b + med_vid_orf_b + med_vid_pri_b + risk_prop_b + loc_cntrl_int_1_b + loc_cntrl_int_2_b + loc_cntrl_int_3_b + loc_cntrl_int_4_b + health_w + act_wrk_w + act_spo_w + act_frn_w + act_sho_w + act_pet_w + sat_dem_w + hh_income_w + med_txt_kro_w + med_txt_sta_w + med_txt_pre_w + med_txt_oes_w + med_txt_kur_w + med_txt_slz_w + med_txt_son_w + med_vid_orf_w + med_vid_pri_w + risk_prop_w + loc_cntrl_int_1_w + loc_cntrl_int_2_w + loc_cntrl_int_3_w + loc_cntrl_int_4_w
-,
-                data = d_long)
-summary(model_life_sat_lmer)
 ```
@@
 
+First, without control variables.
+
 ```{r}
-model_aff_pos_lmer <- lmerTest::lmer(aff_pos_m ~ (1 | id) + (1 | wave) + soc_med_read_b + soc_med_read_w  + soc_med_like_share_b + soc_med_like_share_w + soc_med_post_b + soc_med_post_w + soc_med_fb_b + soc_med_fb_w + soc_med_ig_b + soc_med_ig_w + soc_med_wa_b + soc_med_wa_w + soc_med_yt_b + soc_med_yt_w + soc_med_tw_b + soc_med_tw_w + age + male + edu + born_aus + born_aus_prnts + res_vienna + edu + acc_bal + acc_gar + employment_fac
-# + health_b + act_wrk_b + act_spo_b + act_frn_b + act_sho_b + act_pet_b + sat_dem_b + hh_income_b + med_txt_kro_b + med_txt_sta_b + med_txt_pre_b + med_txt_oes_b + med_txt_kur_b + med_txt_slz_b + med_txt_son_b + med_vid_orf_b + med_vid_pri_b + risk_prop_b + loc_cntrl_int_1_b + loc_cntrl_int_2_b + loc_cntrl_int_3_b + loc_cntrl_int_4_b + health_w + act_wrk_w + act_spo_w + act_frn_w + act_sho_w + act_pet_w + sat_dem_w + hh_income_w + med_txt_kro_w + med_txt_sta_w + med_txt_pre_w + med_txt_oes_w + med_txt_kur_w + med_txt_slz_w + med_txt_son_w + med_vid_orf_w + med_vid_pri_w + risk_prop_w + loc_cntrl_int_1_w + loc_cntrl_int_2_w + loc_cntrl_int_3_w + loc_cntrl_int_4_w
-,
-                data = d_long)
+model_aff_pos_lmer_nco <- lmerTest::lmer(aff_pos_m ~ 
+                                        (1 | id) + (1 | wave) + 
+                                        soc_med_read_w + soc_med_like_share_w + soc_med_post_w + 
+                                        soc_med_fb_w + soc_med_ig_w + soc_med_wa_w + soc_med_yt_w + soc_med_tw_w +
+                                        soc_med_read_b + soc_med_like_share_b + soc_med_post_b + 
+                                        soc_med_fb_b + soc_med_ig_b + soc_med_wa_b + soc_med_yt_b + soc_med_tw_b
+                                      , data = d_long)
+summary(model_aff_pos_lmer_nco)
+```
+
+Now with control variables.
+
+```{r}
+model_aff_pos_lmer <- lmerTest::lmer(aff_pos_m ~ 
+                                        (1 | id) + (1 | wave) + 
+                                        soc_med_read_w + soc_med_like_share_w + soc_med_post_w + 
+                                        soc_med_fb_w + soc_med_ig_w + soc_med_wa_w + soc_med_yt_w + soc_med_tw_w +
+                                        soc_med_read_b + soc_med_like_share_b + soc_med_post_b + 
+                                        soc_med_fb_b + soc_med_ig_b + soc_med_wa_b + soc_med_yt_b + soc_med_tw_b +  
+                                        age + male + born_aus + born_aus_prnts + edu_fac + employment_fac + health_b + 
+                                        res_vienna + acc_bal + acc_gar + home_sqm + 
+                                        # hh_income_b + # large dropout (900 obs)
+                                        # hh_adults + # too much dropout
+                                        # hh_child18 + hh_child17 + hh_child14 + hh_child5 + hh_child2 + hh_oldfam + hh_outfam + hh_partner + 
+                                        # home_owner + # too much dropout
+                                        med_txt_kro_b + med_txt_sta_b + med_txt_pre_b + med_txt_oes_b + med_txt_kur_b + med_txt_slz_b + med_txt_son_b + 
+                                        med_vid_orf_b + med_vid_pri_b + 
+                                        risk_prop_b + loc_cntrl_int_m_b + 
+                                        act_wrk_w + act_spo_w + act_frn_w + act_sho_w + act_pet_w + 
+                                        act_wrk_b + act_spo_b + act_frn_b + act_sho_b + act_pet_b +
+                                        sat_dem_w + sat_dem_b 
+                                      , data = d_long)
 summary(model_aff_pos_lmer)
@@
 
+First, without control variables.
+
 ```{r}
-model_aff_neg_lmer <- lmerTest::lmer(aff_neg_m ~ (1 | id) + (1 | wave) + soc_med_read_b + soc_med_read_w + soc_med_like_share_b + soc_med_like_share_w + soc_med_post_b + soc_med_post_w + soc_med_fb_b + soc_med_fb_w + soc_med_ig_b + soc_med_ig_w + soc_med_wa_b + soc_med_wa_w + soc_med_yt_b + soc_med_yt_w + soc_med_tw_b + soc_med_tw_w + age + male + edu + born_aus + born_aus_prnts + res_vienna + edu + acc_bal + acc_gar + employment_fac
-# + health_b + act_wrk_b + act_spo_b + act_frn_b + act_sho_b + act_pet_b + sat_dem_b + hh_income_b + med_txt_kro_b + med_txt_sta_b + med_txt_pre_b + med_txt_oes_b + med_txt_kur_b + med_txt_slz_b + med_txt_son_b + med_vid_orf_b + med_vid_pri_b + risk_prop_b + loc_cntrl_int_1_b + loc_cntrl_int_2_b + loc_cntrl_int_3_b + loc_cntrl_int_4_b + health_w + act_wrk_w + act_spo_w + act_frn_w + act_sho_w + act_pet_w + sat_dem_w + hh_income_w + med_txt_kro_w + med_txt_sta_w + med_txt_pre_w + med_txt_oes_w + med_txt_kur_w + med_txt_slz_w + med_txt_son_w + med_vid_orf_w + med_vid_pri_w + risk_prop_w + loc_cntrl_int_1_w + loc_cntrl_int_2_w + loc_cntrl_int_3_w + loc_cntrl_int_4_w
-,
-                data = d_long)
+model_aff_neg_lmer_nco <- lmerTest::lmer(aff_neg_m ~ 
+                                        (1 | id) + (1 | wave) + 
+                                        soc_med_read_w + soc_med_like_share_w + soc_med_post_w + 
+                                        soc_med_fb_w + soc_med_ig_w + soc_med_wa_w + soc_med_yt_w + soc_med_tw_w +
+                                        soc_med_read_b + soc_med_like_share_b + soc_med_post_b + 
+                                        soc_med_fb_b + soc_med_ig_b + soc_med_wa_b + soc_med_yt_b + soc_med_tw_b
+                                      , data = d_long)
+summary(model_aff_neg_lmer_nco)
+```
+
+Then, with control variables.
+
+```{r}
+model_aff_neg_lmer <- lmerTest::lmer(aff_neg_m ~ 
+                                        (1 | id) + (1 | wave) + 
+                                        soc_med_read_w + soc_med_like_share_w + soc_med_post_w + 
+                                        soc_med_fb_w + soc_med_ig_w + soc_med_wa_w + soc_med_yt_w + soc_med_tw_w +
+                                        soc_med_read_b + soc_med_like_share_b + soc_med_post_b + 
+                                        soc_med_fb_b + soc_med_ig_b + soc_med_wa_b + soc_med_yt_b + soc_med_tw_b +  
+                                        age + male + born_aus + born_aus_prnts + edu_fac + employment_fac + health_b + 
+                                        res_vienna + acc_bal + acc_gar + home_sqm + 
+                                        # hh_income_b + # large dropout (900 obs)
+                                        # hh_adults + # too much dropout
+                                        # hh_child18 + hh_child17 + hh_child14 + hh_child5 + hh_child2 + hh_oldfam + hh_outfam + hh_partner + 
+                                        # home_owner + # too much dropout
+                                        med_txt_kro_b + med_txt_sta_b + med_txt_pre_b + med_txt_oes_b + med_txt_kur_b + med_txt_slz_b + med_txt_son_b + 
+                                        med_vid_orf_b + med_vid_pri_b + 
+                                        risk_prop_b + loc_cntrl_int_m_b + 
+                                        act_wrk_w + act_spo_w + act_frn_w + act_sho_w + act_pet_w + 
+                                        act_wrk_b + act_spo_b + act_frn_b + act_sho_b + act_pet_b +
+                                        sat_dem_w + sat_dem_b 
+                                      , data = d_long)
 summary(model_aff_neg_lmer)
@@
 
-Active vs. Passive
+First, extract data.
 
@@
   mutate(
-    type = ifelse(grepl(".*_b.*", .$term), "between", "within"),
+    type = ifelse(grepl(".*_w.*", .$term), "within", "between"),
     iv = gsub("_(w|b)\\>", "", .$term)
@@
   mutate(
-    type = factor(.$type, levels = c("within", "between")),
-    dv = factor(.$dv, levels = c("life_sat", "aff_pos", "aff_neg"),
+    type = factor(.$type, 
+                  levels = c("within", "between")),
+    dv = factor(.$dv, 
+                levels = c("life_sat", "aff_pos", "aff_neg"),
                 labels = c("Life satisfaction", "Positive affect", "Negative affect"))
@@
   select(dv, iv, type, estimate, conf.low, conf.high, p.value)
+```
 
+Select social media types of activity, and make graph.
+
+```{r}
 # select Social Media type of activity
@@
   mutate(
-    iv = factor(.$iv, levels = c("soc_med_post", "soc_med_like_share", "soc_med_read"), labels = c("Posting", "Liking & Sharing", "Reading"))
+    iv = factor(.$iv, 
+                levels = c("soc_med_post", "soc_med_like_share", "soc_med_read"), 
+                labels = c("Posting", "Liking & Sharing", "Reading"))
   )
@@
        plot = fig_results_activity)
+```
 
+Select social media channels and make figure.
+
+```{r}
 # select social media channel
@@
 
+Select control variables and make graph.
+
+```{r}
+# select social media channel
+dat_fig_results_control <- dat_fig_results %>%
+  filter(iv %in% c("age", "male", "health", "home_sqm", "loc_cntrl_int_m", "employment_facUnemployed", "employment_facParental Leave", "act_spo", "sat_dem")) %>% 
+  mutate(iv = factor(.$iv, levels = c("age", "male", "health", "home_sqm", "loc_cntrl_int_m", "employment_facUnemployed", "employment_facParental Leave", "act_spo", "sat_dem"), 
+                     labels = c("Age", "Male", "Health", "Squaremeter home", "Internal locus of control", "Umemployed", "On parental leave", "Sport", "Satisfaction democracy")))
+
+# make figure
+fig_results_control <- make_graph_res(dat_fig_results_control, sesoi = FALSE)
+
+# save figure
+ggsave("figures/fig_results_control.pdf", 
+       width = 7, height = 4,
+       plot = fig_results_control)
+```
+
 ```{r}

### 2021-11-05 manuscript.Rmd
- implemented feedback from Nick - updated analyses - provided additional plots for control variables

@@
 abstract          : "In times of crisis such as the Corona pandemic, it is important that citizens stay informed about recent events, the latest political decisions, or mandatory protection measures.
-As a result, to attain relevant information many people use various types of media, and increasingly social media.
+To stay informed, many people use various types of media, and increasingly social media.
 However, because social media are particularly engaging, some find it hard to disconnect and cannot stop 'doomscrolling'.
@@
 The data were analzed using random effects cross lagged panel models, controlling for several stable and varying covariates.
-Results showed that COVID-19 related social media use did not meaningfully affect several types of well-being, including life satisfaction, positive affect, and negative effect.
-This pertains to both passive and active social media use, and all the prominent channels such as Facebook, WhatsApp, or YouTube.
-As a result, this study suggests fears that social media use during times of crisis might be detrimental for well-being can be put to rest."
+Results showed that the effects of various types of COVID-19 related social media use on several types of well-being were very small, arguably too small to matter.
+The findings suggests fears that social media use during times of crisis might be detrimental for well-being are likely to be unfounded."
 keywords          : "COVID, well-being, affect, life satisfaction, social media use, news use, random effects within between model, panel study, longitudinal"
@@
 
-```{r custom-funcs, include=F}
-# Get mean, max, and min values
-
-get_specs <- function(model){
-  dat <- coefficients(model)$wave
-  specs <- list(
-    min = min(dat),
-    max = max(dat),
-    mean = mean(dat$`(Intercept)`),
-    sd = attr(VarCorr(model), "sc")
-  )
-}
-```
-
 During the COVID-19 pandemic, 
 <!-- and especially in the beginning,  -->
-several events unfolded in quick succession. 
+numerous events unfolded in quick succession, and several open questions emerged.
 How dangerous is the virus? Is it spreading in my region? How is it transmitted, and how can I protect myself?
 Because for many it was (and still is) a matter of life or death, citizens had to stay informed regarding the latest developments.
-Governments around the world implemented safety measures, from wearing masks or keeping physical distance, to complete lockdowns.
-In this extraordinary situation, people hence used media excessively, and especially social media were at an all time high [@statistaAverageDailyTime2021].
-A new phenomenon "termed doomscrolling" emerged:
+Governments around the world implemented safety measures, ranging from wearing masks or keeping physical distance to complete lockdowns.
+In this extraordinary situation, people hence used media excessively to stay informed, and especially social media were at an all time high [@statistaAverageDailyTime2021].
+A new phenomenon termed "doomscrolling" emerged:
 People could not stop using social media to attain COVID-19 related news.
 
-Several people reported that they were glued to their screens and could not pursue other relevant activities such as working, taking a break, or even care-work.
-Increasingly, it was asked whether such an increase in social media use could still be considered useful and adaptive, or whether it created an additional and new psychological danger for the users' mental health [@kleinDarklySoothingCompulsion2021].
-A study with 6,233 people from Germany found that "[f]requency, duration and diversity of media exposure were positively associated with more symptoms of depression and unspecific and COVID-19 specific anxiety" [@bendauAssociationsCOVID19Related2021].
+Several people reported that they were glued to their screens and could not pursue other relevant activities such as working, taking a break, or even care-work [@kleinDarklySoothingCompulsion2021].
+As doomscrolling increased it became doubtful whether such a surge in social media use could still be considered useful and adaptive, or whether it created an additional and new psychological danger for the users' mental health [@sandstromDoomscrollingCOVIDNews2021].
+A study with 6,233 people from Germany found that "[f]requency, duration and diversity of media exposure were positively associated with more symptoms of depression and unspecific and COVID-19 specific anxiety" [@bendauAssociationsCOVID19Related2021], implying negative consequences for well-being.
 
@@
 The panel study collected a large number of variables.
-Because we can therefore control for many confounding third variables, both stable and varying, together with the longitudinal design this creates a unique opportunity to analyze causality.
+Being able to control for many confounding third variables, both stable and varying, together with the longitudinal design creates a unique opportunity to investigate causal effects.
 
@@
 The underlying theories that guided the selection of variables for this study are the two-continua model of mental health [@greenspoonIntegrationSubjectiveWellbeing2001] and the hierarchical taxonomy of computer-mediated medation [@meierComputerMediatedCommunicationSocial2020].
-According to the two-continua model of mental health, mental health consists of two dimensions, psychopathology and well-being.
-Because the aim of this study is better to understand typical users and everyday contexts, my focus will be on well-being.
+According to the two-continua model of mental health, mental health consists of two dimensions: psychopathology and well-being.
+<!-- Because the aim of this study is better to understand typical users and everyday contexts, my focus will be on well-being. -->
 Well-being can be differentiated into subjective and psychological well-being [@dienerAdvancesOpenQuestions2018].
@@
 Subjective well-being is primarily about achieving positive affect and avoiding negative affect.
-In my view, life satisfaction is a meta concept above both psychological and subjective well-being, representing a meta-appraisal of one's life.
-Notably, life satisfaction is very stable and fluctuates only little, whereas it's the exact opposite for affect [@dienlinImpactDigitalTechnology2020].
+On of the most prominent indicators of well-being is life satisfaction.
+In my view, life satisfaction is a meta concept that combines psychological and subjective well-being and represents a meta-appraisal of one's life.
+Notably, life satisfaction is stable and fluctuates only little, whereas it's the exact opposite for affect [@dienlinImpactDigitalTechnology2020].
 <!-- Everyone can feel bad on any given day, but whether or not one feels generally comfortable in one's life rather changes across years. -->
@@
 Whereas the first four levels are focused on the channel, the last two on the type of communication.
-To measure social media use for consumption of specific news, I here employ both the channel and the communication perspective.
+To measure social media use for consumption of specific news, I here employ both the channel and the communication perspective, which together provides a more nuanced understanding of communication.
 First, I will analyze how using the most prominent branded applications affect well-being, and whether this effect changes across applications.
+According to @meierComputerMediatedCommunicationSocial2020, branded apps are separate entities with different effects. 
+Therefore, Twitter might have a different effect than WhatsApp because of their different affordances.
+For example, @waterlooNormsOnlineExpressions2018 found that it is more adequate to express negative emotions on WhatsApp than on Twitter or on Instagram. 
 The branded applications investigated here are Facebook, Twitter, Instagram, WhatsApp, and YouTube.
@@
 Worth noting, this study is not about _general_ social media during times of COVID, but on social media use focused on COVID-19 related news and interactions. 
-For example, posting about the pandemic or retweeting COVID-19 related posts.
+Example of such media use include posting about the pandemic or retweeting COVID-19 related posts.
 
@@
 288).
-Hence, this might hint at potential negative effects of social media as news use---but note that this finding comes from a between-person relation stemming from a cross-sectional data.
+Hence, this might hint at potential negative effects of social media as news use---but note that this finding comes from a between-person relation stemming from a cross-sectional data (see below).
 Hence, we don't know whether the differences in mental health and well-being are due to social media use or other third variables, such as age or education.
+@edenMediaCopingCOVID192020 analyzed the media use of 425 US college students during the first wave of the pandemic, and found both positive and negative relations with well-being.
+In a sample of 312 respondents collected via Amazon Mechanical Turk, it was found that people who used media to attain information were more lonely and less satisfied with their lives [@choiMediatedCommunicationMatters2021].
+@stainbackCOVID1924News2020 analyzed a large-scale study with 11,537 respondents from the US and found that increased COVID-19 media consumption was related to more psychological distress.
 
-In general, so far there is only little research on news-related social media use and well-being.
+In general, so far there is only little empirical research on news-related social media use and well-being.
 On the other hand, the question of whether and how _general_ social media use---or other, more specific types such as active or passive use---affect well-being is well-researched.
 A meta review (i.e., an analysis of meta-analyses), found that the relation between social media use and well-being is likely in the negative spectrum---but very small, potentially too small to matter [@meierComputerMediatedCommunicationSocial2020].
-This is well aligned with several new studies that employed the most advanced methods [@orbenSocialMediaEnduring2019; @przybylskiDoesTakingShort2021a; @schemerImpactInternetSocial2021; @kerestesAdolescentsOnlineSocial2020].
-For example, @beyensEffectSocialMedia2020 reported that although for some users (roughly one quarter) the effects were negative, for almost the same amount of users the effects were positive, while for the majority they were neutral.
-At all events, in general most effects are somewhere between trivial and small.
+What determines whether an effect is considered trivial or small?
+If we refer to standardized effect sizes, according to @cohenPowerPrimer1992 small effect sizes start at _r_ = .10.
+And indeed, several if not most meta-analyses find effect sizes below that threshold [@huangTimeSpentSocial2017; @meierComputerMediatedCommunicationSocial2020; @fergusonThisMetaanalysisScreen2021].
 
-What determines a trivial or a small effect is a difficult question (see below).
-If we refer to standardized effect sizes, according to Cohen [@cohenPowerPrimer1992] small effect sizes start at _r_ = .10.
-However, most meta-analyses find effect sizes below that threshold [@huangTimeSpentSocial2017; @meierComputerMediatedCommunicationSocial2020; @fergusonThisMetaanalysisScreen2021].
-As a result, I think it's most convincing to expect trivial to small effects also in the case of COVID-19 related social media use.
+These overviews are well aligned with several individual new studies that employed the most advanced methods [@orbenSocialMediaEnduring2019; @przybylskiDoesTakingShort2021a; @schemerImpactInternetSocial2021; @kerestesAdolescentsOnlineSocial2020].
+For example, @beyensSocialMediaUseinpress reported that although for some users (roughly one quarter) the effects were negative, for almost the same number of users the effects were positive, while for the majority they were neutral.
+In conclusion, most effects are somewhere between trivial and small.
+I think it's most convincing to expect trivial to small effects also in the case of COVID-19 related social media use.
 
-From a theoretical perspective, why could COVID-19 related social media use be detrimental?
-Above everything, one can reasonably assume a _direct_ negative effect on well-being, especially on positive or negative effect.
-Dangers, inequalities, corruption -- these were the headlines across many countries worldwide.
+From a theoretical perspective, how could we explain how COVID-19 related social media use might affect well-being?
+In what follows, I outline potential arguments as to why the effect might be positive or negative, direct or indirect. 
+In advance, there does not seem to be a clear winner, and it's likely that both positive and negative effects are equally strong.
+
+First, one could reasonably assume a _direct_ negative effect on well-being---especially on positive or negative affect, which is more volatile and fluctuating.
+Dangers, inequalities, corruption---these were the headlines across many countries worldwide.
 If one learns about such things, the initial reaction might be shock, fear, or dismay.
-Repeatedly consuming such news might be depressing, perhaps even changing some general perspective on life.
-So, just like being hit by a hammer hurts and we don't need any "mediating mechanism", this could be the case here as well.
+Repeatedly consuming such news might be depressing, perhaps even changing some general perspectives on life, without any further mediating processes.
+<!-- So, just like being hit by a hammer hurts and we don't need any "mediating mechanism", this could be the case here as well. -->
 
-That said, because not all news were negative, because many people showed solidarity and compassion, 
-<!-- or maybe also because it was possible quickly to find an effective vaccine,  -->
-there were also positive and potentially uplifting news.
-However, in light of a worldwide pandemic with millions of deaths, the negative direct effect seems more plausible.
+That said, because not all news were negative, because many people showed solidarity and compassion, there were also positive and potentially uplifting news.
+<!-- However, in light of a worldwide pandemic with millions of deaths, the negative direct effect seems more plausible. -->
 
@@
 When doomscrolling, users are captivated to such an extent that they cannot stop using social media.
-For example, during the pandemic social media uses was at an all-time high in the US [@statistaAverageDailyTime2021].
+For example, during the pandemic social media use was at an all-time high in the US [@statistaAverageDailyTime2021].
 As has been expressed by many before, it is most likely that moderate social media use is not detrimental [@orbenTeenagersScreensSocial2020].
-However, overuse is more critical, and several studies have showed more pronounced negative effects for extreme users [@przybylskiLargescaleTestGoldilocks2017].
+Overuse, however, is more critical, and several studies have shown more pronounced negative effects for extreme users [@przybylskiLargescaleTestGoldilocks2017].
+Overuse likely impairs well-being if it replaces other meaningful or functional activities such as meeting others, working, actively relaxing, or exercising.
 So if a society collectively overuses social media, there is potential for negative effects.
-Overuse likely impairs well-being if it replaces other meaningful or functional activities such as meeting others, working, actively relaxing, or exercising.
-If overuse replaces such activities it's reasonable to assume that it's also detrimental.
+<!-- If overuse replaces such activities it's reasonable to assume that it's also detrimental. -->
 
 On the other hand, one can make the case that overuse can be beneficial _in times of a pandemic_, even if it's mainly COVID-19 related.
-Exchanging COVID-19 related messages with friends via WhatsApp might replace the in-person contact one would have otherwise, but which is logically impossible at that time.
+Exchanging COVID-19 related messages with friends via WhatsApp might replace the in-person contact one would have otherwise, but which is effectively impossible at that time.
 At times where meaningful and functional activities are explicitly or implicitly prohibited, using social media to exchange about COVID-19 might not be the worst idea.
 In fact, given that nowadays a large number of experts, scientists, and politicians converse directly on social media, one can get first-hand information on the current developments.
-On the other hand, there is of course also much disinformation, and "bingeing" on COVID-19 fake news might also pose risks for impaired well-being.
+<!-- On the other hand, there is of course also much disinformation, and "bingeing" on COVID-19 fake news might also pose risks for impaired well-being. -->
 
-Together, the strongest argument seems to be that the effects of social media on well-being are, on average, generally small at best.
+Together, the strongest argument seems to be that _in general_ effects of social media on well-being are, on average, small at best.
 Because this study looks at only _one part_ of social media use---namely, COVID-19 related interactions---it is even more focused, and the overall potential of the effects should diminish even further.
@@
 Testing this hypothesis, however, is not trivial.
-First, in contrast to most hypothesis typically posited in the social sciences, it explicates an effect size, a so-called smallest effect size of interest (SESOI).
+First, in contrast to most hypothesis typically posited in the social sciences, it implicitly contains an effect size, a so-called smallest effect size of interest (SESOI).
 Effectively testing this hypothesis hence necessitates to define what's considered a "trivial effect size" and what's not.
@@
 In this study, COVID-19 related social media use was measured on a 5-point scale, ranging from 1 = _never_ to 5 = _several times a day_. 
-Thus, the example from above would imply that a change of five units in social media use should correspond to a noticeable change in well-being.
+Thus, the example from above would imply that a change of four units in social media use should correspond to a noticeable change in well-being.
 But what's a noticeable change in well-being?
 According to @normanInterpretationChangesHealthrelated2003, people can reliably differentiate between _seven_ levels of satisfaction with health.
-So we could state that a five unit change in social media use should result in a one unit change in satisfaction.
+So we could state that a four unit change in social media use should result in a one unit change in satisfaction.
 Statistically, in a regression, _b_ estimates the change in the dependent variable if the independent variable increases by one point.
-Transferred to our example, we would hence expect a _b_ of 1 / 5 = .20.
+Transferred to our example, we would hence expect _b_ = 1 / 4 = .25.
 
-In this study, life satisfaction was measured with 11 units, hence more than people can reliably detect. 
-Hence, we would expect a 11 / 7 * .20 change, that is a _b_ of at least .31.
-For affect, which was measured on a 5-point scale, our SESOI would be 5 / 7 * .2 = .14.
-In order not to exaggerate precision, these numbers will be rounded. 
-Plus, because we are agnostic as to whether the effects are positively or negatively nontrivial, this leads to an indifference region ranging from _b_ = -.30 to _b_ = .30 (and _b_ = -.15 to _b_ = .15 for affect).
+In this study, life satisfaction was measured with 11 units, hence more than seven degrees people can reliably differentiate. 
+We hence need to transform a 1-point change on a 7-point scale to an equivalent change on an 11-point scale. 
+This is 11 / 7 = `r 11/7`.
+Hence, we now expect _b_ = 1/4 * 11/7 = `r 11/7*.25`.
+For affect, which was measured on a 5-point scale, our SESOI would therefore be _b_ = 1/4 * 5/7 = `r 5 / 7 * .25`.
+Because we are agnostic as to whether the effects are positively or negatively nontrivial, the null region will include negative and positive effects (in Bayesian terms, this is called the region of practical equivalence [ROPE]).
+In order not to exaggerate precision, and to be a bit less conservative, these numbers will be reduced to nearby thresholds. 
+Note that other researchers also decrease or recommend decreasing thresholds for effect sizes when analyzing within-person or cumulative effects (@beyensSocialMediaUseinpress; @funderEvaluatingEffectSize2019).
+Together, this leads to an indifference region ranging from _b_ = -.30 to _b_ = .30 for life satisfaction, and _b_ = -.15 to _b_ = .15 for positive and negative affect.
 
@@
 The hypothesis explicitly states a causal effect.
-In non-experimental designs, causality can be approximated using longitudinal designs.
+In non-experimental designs, longitudinal designs help investigate causality.
 Using longitudinal designs alone, however, is not sufficient for correct causal statements.
-In addition, it is crucial also to control for third variables, and importantly also for _varying_ third variables.
+In addition, we also need to control for third variables, and importantly also for _varying_ third variables.
 
@@
 Obviously, we can never be entirely sure to have included all varying covariates, which makes absolute statements regarding causality impossible.
-However, when controlling for many varying variables, we can be much more certain that we measured causality correctly.
-The aim should still be to collect as many relevant varying and nonvarying third variables as possible, while knowing that absolute certainty cannot be reached.
+In addition, when determining the overall causal effect, we should _not_ control for mediating variables [@rohrerThinkingClearlyCorrelations2018].
+Doing so would bias and distort our assessment of the role of social media use.
+However, when controlling for relevant variables (that aren't mediators), we can be much more certain that we measured causality correctly.
+The aim should therefore be to collect as many _relevant_ varying and nonvarying third variables as possible, while knowing that absolute certainty regarding causality cannot be reached.
 
-When looking for suitable control candidates, ideally we find variables that affect both media use and well-being, because controlling for these factors will isolate the actual effect of social media use on well-being.
+When looking for suitable control candidates, ideally we find variables that affect both media use and well-being. 
+Controlling for these factors will isolate the actual effect of social media use on well-being.
 We can also control for variables that affect _only_ well-being. 
-However, in doing so nothing is gained or lost because the effects of social media use would remain exactly the same [@klinePrinciplesPracticeStructural2016].
-Crucially, when determining the general causal effect of social media use we should _not_ control for mediating variables [@rohrerThinkingClearlyCorrelations2018].
-Doing so would bias and distort our assessment of the role of social media use.
+However, in doing so not much is gained or lost because the effects of social media use would remain virtually the same [@klinePrinciplesPracticeStructural2016].
 
-In this study, I will hence control for the following variables, which either have been shown or a likely to affect both social media use and well-being.
-(I'll additionally include variables that likely affect only well-being, also to obtain a comparison benchmark for social media effects):
+In this study, I hence plan to control for the following variables, which either have been shown or a likely to affect both social media use and well-being, while not being mediators.
+<!-- (I'll additionally include variables that likely affect only well-being, also to obtain a comparison benchmark for social media effects): -->
 Gender, age, education, Austria country of birth, Austria country of birth of parents, text-based news consumption, video-based news consumption, residency Vienna, household size, health, living space, access to garden, access to balcony, employment, work hours per week, being in home-office, household income, outdoor activities, satisfaction with democracy, disposition to take risks, and locus of control.
@@
 If a specific person changes their media diet, we need to measure how this affects _them_.
-Between person comparison from cross-sectional data, where participants are interviewed only once, cannot provide such data.
+Between person comparison from cross-sectional data, where participants are interviewed only once, cannot provide such insights.
 In this study, I will hence differentiate between within-person effects and between-person relations. 
+Because it is explicitly causal, in answering the hypothesis I will thus consider only the within-person effects.
 
@@
 To this end, finding the right interval when causes and effects are measured is crucial.
-For example, if we want to measure the effect of alcohol consumption on driving performance, it makes a big difference if driving performance is measure one minute, one hour, one day, or one week after consumption.
+For example, if we want to measure the effect of alcohol consumption on driving performance, it makes a big difference if driving performance is measured one minute, one hour, one day, or one week after consumption.
 Finding the right interval is difficult.
@@
 The (anonymous) preregistration can be accessed here: https://osf.io/87b24/?view_only=b2289b6fec214fa88ee75a18d45c18f3.
-Because in this study I analyze data from an already existing large scale data set, the Austrian Corona Panel Project, all of these steps were preregistered prior to accessing the data.
+Because in this study I analyze data from an already existing large-scale data set, the Austrian Corona Panel Project, all of these steps were preregistered prior to accessing the data.
 The preregistration was designed on the basis of the panel documentation online [@kittelAustrianCoronaPanel2020].
@@
 If the effects interval and the null region overlap (e.g., _b_ = .25, [95% CI: .15, .35]), the hypothesis is not supported and the results are considered inconclusive, whereas a meaningful negative effect is rejected.
-If the confidence falls completely outside the null-region (e.g., _b_ = .4, [95% CI: .35, .45]), the hypothesis is rejected and the existence of a meaningful positive effect is supported.
+If the confidence falls completely outside of the null-region (e.g., _b_ = .4, [95% CI: .35, .45]), the hypothesis is rejected and the existence of a meaningful positive effect is supported.
 
@@
 All predictors---that is, social media activities and social media channels, within and between-person factors, stably and varying covariates---were included in each of the three models. 
-The data were hierarchical, and responses were nested in participants and waves.
-Nesting in participants allowed to separate within-person effects from between-person relations.
+The data were hierarchical, and responses were nested in participants and waves (put differently, participants and waves were implemented as random effects).
+Nesting in participants allowed to separate between-person relations from within-person effects.
 Nesting in waves allowed to control for general developments, such as general decrease in well-being in the population, for example due to lockdown measures (hence, there was no need additionally to control for specific phases or measures of the lockdown).
 
-The factorial validity of the scales were tested with confirmatory factor analyses (CFA). 
-If Mardia’s test showed that the assumption of multivariate normality was violated, the more robust Satorra-Bentler scaled and mean-adjusted test statistic (MLM) was used as estimator. 
-To avoid overfitting, scales were tested on more liberal fit criteria (CFI > .90, TLI > .90, RMSEA <. .10, SRMR < .10) [@klinePrinciplesPracticeStructural2016].
-Because REWB-models cannot model scales as latent variables, to increase precision factor scores exported from the CFAs were used.
+<!-- The factorial validity of the scales were tested with confirmatory factor analyses (CFA).  -->
+<!-- If Mardia’s test showed that the assumption of multivariate normality was violated, the more robust Satorra-Bentler scaled and mean-adjusted test statistic (MLM) was used as estimator.  -->
+<!-- To avoid overfitting, scales were tested on more liberal fit criteria (CFI > .90, TLI > .90, RMSEA <. .10, SRMR < .10) [@klinePrinciplesPracticeStructural2016]. -->
+<!-- Because REWB-models cannot model scales as latent variables, to increase precision factor scores exported from the CFAs were used. -->
+
 For more information on the analyses, see [companion website](https://xmtra.github.io/Austrian_Corona_Panel_Project/index.html).
@@
 In what follows, I briefly list all variables that were collected.
+For the variables' means, range, and variance, see Table \@ref(tab:descriptives).
 For a complete list of all items and item characteristics, see [companion website](https://xmtra.github.io/Austrian_Corona_Panel_Project/index.html).
@@
 The response options ranged from 0 (_extremely unsatisfied_) to 10 (_extremely satisfied_).
-The variable's average across all waves was _M_ = `r get_specs(model_life_sat)$mean`, ranging from _M_~min~ = `r get_specs(model_life_sat)$min` to _M_~max~ = `r get_specs(model_life_sat)$max`.
-The average standard deviation was _SD_ = `r get_specs(model_life_sat)$sd`.
 
 To capture positive affect, respondents were asked how often in the last week they felt (a) calm and relaxed, (b) happy, and (c) full of energy.
-The response options were 1 (_never_), 2 (_on some days_ ), 3 (_several times per week_), 4 (almost every day), and 5 (daily).
-The variable's average across all waves was _M_ = `r get_specs(model_aff_pos)$mean`, ranging from _M_~min~ = `r get_specs(model_aff_pos)$min` to _M_~max~ = `r get_specs(model_aff_pos)$max`.
-The average standard deviation was _SD_ = `r get_specs(model_aff_pos)$sd`.
+The response options were 1 (_never_), 2 (_on some days_), 3 (_several times per week_), 4 (_almost every day_), and 5 (_daily_).
 
 For negative affect, respondents were asked how often in the last week they felt (a) lonely, (b) aggravated, (c) so depressed, that nothing could lift you up, (d) very nervous, (e) anxious, and (h) glum and sad.
-The response options were 1 (_never_), 2 (_on some days_ ), 3 (_several times per week_), 4 (almost every day), and 5 (daily).
-The variable's average across all waves was _M_ = `r get_specs(model_aff_neg)$mean`, ranging from _M_~min~ = `r get_specs(model_aff_neg)$min` to _M_~max~ = `r get_specs(model_aff_neg)$max`.
-The average standard deviation was _SD_ = `r get_specs(model_aff_neg)$sd`.
+The response options were 1 (_never_), 2 (_on some days_), 3 (_several times per week_), 4 (_almost every day_), and 5 (_daily_).
 
-All three items were measured on each wave.
+All three variables were measured on each wave.
 
@@
 The items were inverted for the analyses.
-The variable's average across all waves was _M_ = `r get_specs(model_life_sat)$mean`, ranging from _M_~min~ = `r get_specs(model_life_sat)$min` to _M_~max~ = `r get_specs(model_life_sat)$max`.
-The average standard deviation was _SD_ = `r get_specs(model_life_sat)$sd`.
 
@@
 
-All items were measured every second wave.
-However, freshly recruited respondents always answered these questions.
+Social media use was measured for everyone on waves 1, 2, 8, 17, and 23. 
+Freshly recruited respondents always answered the questions on social media use.
+
+```{r descriptives, cache=F}
+apa_table(tab_desc_dat,
+          stub_indents = list(
+            "Well-being" = c(1:3),
+            "Social media use" = c(4:6),
+            "Social media channel" = c(7:11)),
+          caption = "Descriptives of the main variables."
+          )
+```
 
@@
 The effects of COVID-19 related social media use were controlled for the following stable variables: 
-(a) gender (answer options: female, male, diverse), (b) age, (c) education (10 ordinal options), (d) Austria country of birth (yes/no), (e) Austria parents' country of birth (no parent, one parent, both parents).
-We controlled also for the following varying covariates:
-(a) text-based media news consumption, (b) video-based media news consumption, (c) residency is Vienna, (d) household size, (e) Self-reported physical health, (f) Living space (in squaremeter), (g) access to balcony, (h) access to garden, (i) Employment status, (j) Work hours per week, (k) Home office, (l) household income, (m) outdoor activities, (o) satisfaction with democracy, (p) disposition to take risks, (q) locus of control.
+(a) gender (answer options: female, male, diverse), (b) age, (c) education (10 options), (d) Austria country of birth (yes/no), (e) Austria parents' country of birth (no parent, one parent, both parents).
+I originally planned to implement other variables as varying covariates.
+However, because they were not measured often enough or at the time when social media use was measured, I implemented them as stable variables using their average values across all waves.
+This includes (a) text-based media news consumption, (b) video-based media news consumption, (c) residency is Vienna, (d) self-reported physical health, (e) living space (in squaremeter), (f) access to balcony, (g) access to garden, (h) employment status, (i) disposition to take risks, and (j) locus of control.
+I controlled also for the following varying covariates: (a) outdoor activities, (b) satisfaction with democracy.
+Because it lead to too much attrition in the sample, I did not control for (a) household size, (b) work hours per week, (c) home office, (d) household income.
 
@@
 When looking at the variables from a descriptive perspective, we see that all well-being measures did not change substantially across the different phases of the pandemic.
-COVID-19 related media use, however, increased during the beginning of the study and remained stable after approximately six waves.
+COVID-19 related media use, however, decreased during the beginning of the study and remained stable after approximately six waves.
+The initial decrease might be explained by the fact that the collection of data began in end of March 2020, hence approximately three months after the pandemic began.
+It could be that after an initial stark surge, COVID-19 related social media use was already declining, returning to more normal levels.
 
 ```{r fig-desc, fig.cap="Development of well-being and media use measures across the pandemic.
-Values obtained from mixed effect models, with participants and waves as grouping factor and without additional predictors.", out.width = "\\textwidth", fig.pos = "!h"}
+Values obtained from mixed effect models, with participants and waves as grouping factor and without additional predictors.", out.width = "\\textwidth", cache=F}
 knitr::include_graphics("figures/fig_descriptives.pdf")
@@
 The study's hypothesis was that the effects of all types of social media use on well-being will be trivial.
-Regarding the different types of _communication_, that is reading vs. sharing vs. posting, all within-person effects fell completely within the a-priori defined SESOIs.
+Regarding the different types of _communication_---that is, reading vs.
+sharing vs.
+posting---all within-person effects fell completely within the a-priori defined SESOIs (see Figure \@ref(fig:res-activity)).
 For example, respondents who used social media more frequently than usual to read about COVID-19 related topics did not show a simultaneous change in life satisfaction (_b_ = `r dat_fig_results_activity %>% filter(dv == "Life satisfaction" & iv == "Reading" & type == "within") %>% select(estimate) %>% round(2)` [95% CI `r dat_fig_results_activity %>% filter(dv == "Life satisfaction" & iv == "Reading" & type == "within") %>% select(conf.low) %>% round(2)`, `r dat_fig_results_activity %>% filter(dv == "Life satisfaction" & iv == "Reading" & type == "within") %>% select(conf.high) %>% round(2)`]).
-Only one effect did not include zero.
-If people posted more about COVID-19 related aspects than they usually did, life satisfaction increased (_b_ = `r dat_fig_results_activity %>% filter(dv == "Life satisfaction" & iv == "Posting" & type == "within") %>% select(estimate) %>% round(2)` [95% CI `r dat_fig_results_activity %>% filter(dv == "Life satisfaction" & iv == "Posting" & type == "within") %>% select(conf.low) %>% round(2)`, `r dat_fig_results_activity %>% filter(dv == "Life satisfaction" & iv == "Posting" & type == "within") %>% select(conf.high) %>% round(2)`]).
-However, because the effect did not fall outside our null region, it's likely not large enough to be considered meaningful.
+All confidence intervals included zero; hence, all effects were non-significant.
+<!-- Only one effect did not include zero. -->
+<!-- If people posted more about COVID-19 related aspects than they usually did, life satisfaction increased (_b_ = `r dat_fig_results_activity %>% filter(dv == "Life satisfaction" & iv == "Posting" & type == "within") %>% select(estimate) %>% round(2)` [95% CI `r dat_fig_results_activity %>% filter(dv == "Life satisfaction" & iv == "Posting" & type == "within") %>% select(conf.low) %>% round(2)`, `r dat_fig_results_activity %>% filter(dv == "Life satisfaction" & iv == "Posting" & type == "within") %>% select(conf.high) %>% round(2)`]).
+-->
+<!-- However, because the effect did not fall outside of our null region, it's likely not large enough to be considered meaningful. -->
 As a result, the hypothesis was supported for all types of COVID-19 related social media communication.
 
-Regarding between-person relations, about which no hypotheses were formulated, only two effects didn't include zero.
-First, respondents who across all waves used social media more frequently than others to read about COVID-19 reported higher levels of positive affect than others (_b_ = `r dat_fig_results_activity %>% filter(dv == "Positive affect" & iv == "Reading" & type == "between") %>% select(estimate) %>% round(2)` [95% CI `r dat_fig_results_activity %>% filter(dv == "Positive affect" & iv == "Reading" & type == "between") %>% select(conf.low) %>% round(2)`, `r dat_fig_results_activity %>% filter(dv == "Positive affect" & iv == "Reading" & type == "between") %>% select(conf.high) %>% round(2)`]).
-However, note that this effect still fell completely within the null-region.
-Hence, although positive the effect was considered too small to matter practically.
-Second, respondents who across all waves used social media more frequently than others to read about COVID-19 reported lower levels of negative affect than others (_b_ = `r dat_fig_results_activity %>% filter(dv == "Negative affect" & iv == "Reading" & type == "between") %>% select(estimate) %>% round(2)` [95% CI `r dat_fig_results_activity %>% filter(dv == "Negative affect" & iv == "Reading" & type == "between") %>% select(conf.low) %>% round(2)`, `r dat_fig_results_activity %>% filter(dv == "Negative affect" & iv == "Reading" & type == "between") %>% select(conf.high) %>% round(2)`]).
-The effect was partially outside the null region, hence the effect might be large enough to be practically relevant, and the effect that it's trivial cannot be rejected.
-
-```{r fig-res-activity, fig.cap="The effects of various types of social media use on three indicators of well-being.
+```{r res-activity, fig.cap="The effects of various types of social media use on three indicators of well-being.
 Effects are controlled for a large number of covariates (see text).
-The SESOI was 0.30 for life satisfaction and .15 for affects; hence, no effect is considered meaningful theoretically.", out.width = "\\textwidth", fig.pos = "!h"}
+The SESOI was 0.30 for life satisfaction and .15 for affects; hence, no effect is considered meaningful theoretically.", out.width = "\\textwidth", cache=F}
 knitr::include_graphics("figures/fig_results_activity.pdf")
@@
 
-Regarding the COVID-19 related use of social media _channels_, the results were virtually the same.
-Changes in the frequency of using different social media channels to attain information regarding COVID-19 were unrelated to meaningful changes in well-being.
+Regarding between-person relations, about which no hypotheses were formulated, only one effect didn't include zero.
+<!-- First, respondents who across all waves used social media more frequently than others to read about COVID-19 reported higher levels of positive affect than others (_b_ = `r dat_fig_results_activity %>% filter(dv == "Positive affect" & iv == "Reading" & type == "between") %>% select(estimate) %>% round(2)` [95% CI `r dat_fig_results_activity %>% filter(dv == "Positive affect" & iv == "Reading" & type == "between") %>% select(conf.low) %>% round(2)`, `r dat_fig_results_activity %>% filter(dv == "Positive affect" & iv == "Reading" & type == "between") %>% select(conf.high) %>% round(2)`]).
+-->
+<!-- However, note that this effect still fell completely within the null-region. -->
+<!-- Hence, although positive the effect was considered too small to matter practically. -->
+Respondents who across all waves used social media more frequently than others to write posts on COVID-19 reported higher levels of negative affect than others (_b_ = `r dat_fig_results_activity %>% filter(dv == "Negative affect" & iv == "Posting" & type == "between") %>% select(estimate) %>% round(2)` [95% CI `r dat_fig_results_activity %>% filter(dv == "Negative affect" & iv == "Posting" & type == "between") %>% select(conf.low) %>% round(2)`, `r dat_fig_results_activity %>% filter(dv == "Negative affect" & iv == "Posting" & type == "between") %>% select(conf.high) %>% round(2)`]).
+The effect was still completely inside of the null region, hence likely not large enough to be practically relevant.
+
+Regarding the COVID-19 related use of social media _channels_, the results were very comparable.
+Changes in the frequency of using different social media channels to attain information regarding COVID-19 were unrelated to meaningful changes in well-being (see Figure \@ref(fig:res-channels)).
 For example, respondents who used Facebook more frequently than usual to learn about COVID-19 did not show a simultaneous change in well-being (_b_ = `r dat_fig_results_channel %>% filter(dv == "Life satisfaction" & iv == "Facebook" & type == "within") %>% select(estimate) %>% round(2)` [95% CI `r dat_fig_results_channel %>% filter(dv == "Life satisfaction" & iv == "Facebook" & type == "within") %>% select(conf.low) %>% round(2)`, `r dat_fig_results_channel %>% filter(dv == "Life satisfaction" & iv == "Facebook" & type == "within") %>% select(conf.high) %>% round(2)`]).
-Only one effect was substantially larger than zero.
-Respondents who used Instagram more frequently than usual to attain COVID-19 related news reported slightly lowers simultaneous levels of life satisfaction then usual (_b_ = `r dat_fig_results_channel %>% filter(dv == "Life satisfaction" & iv == "Instagram" & type == "within") %>% select(estimate) %>% round(2)` [95% CI `r dat_fig_results_channel %>% filter(dv == "Life satisfaction" & iv == "Instagram" & type == "within") %>% select(conf.low) %>% round(2)`, `r dat_fig_results_channel %>% filter(dv == "Life satisfaction" & iv == "Instagram" & type == "within") %>% select(conf.high) %>% my_round("b")`]).
-However, this effect was still completely within the null region, hence not large enough to be considered meaningful.
+Only two effects differed substantially from zero.
+Respondents who used Instagram more frequently than usual to attain COVID-19 related news reported slightly higher simultaneous levels of life satisfaction then usual (_b_ = `r dat_fig_results_channel %>% filter(dv == "Life satisfaction" & iv == "Instagram" & type == "within") %>% select(estimate) %>% round(2)` [95% CI `r dat_fig_results_channel %>% filter(dv == "Life satisfaction" & iv == "Instagram" & type == "within") %>% select(conf.low) %>% round(2)`, `r dat_fig_results_channel %>% filter(dv == "Life satisfaction" & iv == "Instagram" & type == "within") %>% select(conf.high) %>% my_round("b")`]).
+Respondents who used Twitter more frequently than usual to attain COVID-19 related news reported slightly lower simultaneous levels of life satisfaction then usual (_b_ = `r dat_fig_results_channel %>% filter(dv == "Life satisfaction" & iv == "Twitter" & type == "within") %>% select(estimate) %>% round(2)` [95% CI `r dat_fig_results_channel %>% filter(dv == "Life satisfaction" & iv == "Twitter" & type == "within") %>% select(conf.low) %>% round(2)`, `r dat_fig_results_channel %>% filter(dv == "Life satisfaction" & iv == "Twitter" & type == "within") %>% select(conf.high) %>% my_round("b")`]).
+However, both effects were still completely inside of the null region, hence not large enough to be considered meaningful.
 In sum, the hypothesis was supported for the COVID-19 related use of all types of social media channels.
 
-In terms of between-person relations---which, again, weren't included in the hypotheses---no relations crossed or fell outside the null region.
-However, some relations did not included zero.
-For example, respondents who across all waves used Instagram more frequently than others for COVID-19 related reasons reported lower levels of life satisfaction  (_b_ = `r dat_fig_results_channel %>% filter(dv == "Life satisfaction" & iv == "Instagram" & type == "between") %>% select(estimate) %>% round(2)` [95% CI `r dat_fig_results_channel %>% filter(dv == "Life satisfaction" & iv == "Instagram" & type == "between") %>% select(conf.low) %>% round(2)`, `r dat_fig_results_channel %>% filter(dv == "Life satisfaction" & iv == "Instagram" & type == "between") %>% select(conf.high) %>% my_round("b")`]).
-Respondents who were more active on WhatsApp compared to others reported slightly lower levels of positive affect  (_b_ = `r dat_fig_results_channel %>% filter(dv == "Positive affect" & iv == "WhatsApp" & type == "between") %>% select(estimate) %>% round(2)` [95% CI `r dat_fig_results_channel %>% filter(dv == "Positive affect" & iv == "WhatsApp" & type == "between") %>% select(conf.low) %>% round(2)`, `r dat_fig_results_channel %>% filter(dv == "Positive affect" & iv == "WhatsApp" & type == "between") %>% select(conf.high) %>% my_round("b")`]).
-Respondents who compared to others were more active on Twitter and YouTube reported lover levels of negative affect (Twitter: _b_ = `r dat_fig_results_channel %>% filter(dv == "Negative affect" & iv == "Twitter" & type == "between") %>% select(estimate) %>% round(2)` [95% CI `r dat_fig_results_channel %>% filter(dv == "Negative affect" & iv == "Twitter" & type == "between") %>% select(conf.low) %>% round(2)`, `r dat_fig_results_channel %>% filter(dv == "Negative affect" & iv == "Twitter" & type == "between") %>% select(conf.high) %>% my_round("b")`];
-YouTube _b_ = `r dat_fig_results_channel %>% filter(dv == "Negative affect" & iv == "YouTube" & type == "between") %>% select(estimate) %>% round(2)` [95% CI `r dat_fig_results_channel %>% filter(dv == "Negative affect" & iv == "YouTube" & type == "between") %>% select(conf.low) %>% round(2)`, `r dat_fig_results_channel %>% filter(dv == "Negative affect" & iv == "YouTube" & type == "between") %>% select(conf.high) %>% my_round("b")`] .
-However, please note that all these affect not considered large enough to be practically relevant.
-
-```{r fig-res-channels, fig.cap="The effects of using various social media applications on three indicators of well-being.
+```{r res-channels, fig.cap="The effects of using various social media applications on three indicators of well-being.
 Effects are controlled for a large number of covariates (see text).
-The SESOI was 0.30 for life satisfaction and .15 for affects; hence, no effect is considered meaningful theoretically.", out.width = "\\textwidth", fig.pos = "!h"}
+The SESOI was 0.30 for life satisfaction and .15 for affects; hence, no effect is considered meaningful theoretically.", out.width = "\\textwidth", cache=F}
 knitr::include_graphics("figures/fig_results_channel.pdf")
 ```
+
+In terms of between-person relations---which, again, weren't included in the hypotheses---no relations crossed or fell outside of the null region.
+Only one relation did not include zero, was hence statistically significant. 
+Respondents who across all waves used YouTube more frequently than others for COVID-19 related reasons reported marginally higher levels of negative affect (_b_ = `r dat_fig_results_channel %>% filter(dv == "Negative affect" & iv == "YouTube" & type == "between") %>% select(estimate) %>% round(2)` [95% CI `r dat_fig_results_channel %>% filter(dv == "Negative affect" & iv == "YouTube" & type == "between") %>% select(conf.low) %>% my_round("b_txt")`, `r dat_fig_results_channel %>% filter(dv == "Negative affect" & iv == "YouTube" & type == "between") %>% select(conf.high) %>% my_round("b")`]).
+However, please note that this effect was not large enough to be considered practically relevant.
 
@@
 
-[Report results of some control variables.]
+In what follows, I briefly report also the results of some covariates. 
+Several variables showed large associations with well-being. 
+Because each variable has a different scaling, we would again need to define a SESOI for each variable, which cannot be implemented here. 
+But note that several variables would fall outside of such a SESOI.
+This includes for example internal locus of control, health, or employment. 
+For a brief overview, see Figure \@ref(fig:res-control).
+
+```{r res-control, fig.cap="Results of selected covariates.", out.width = "\\textwidth", cache=F}
+knitr::include_graphics("figures/fig_results_control.pdf")
+```
 
@@
 
-Using a panel study with a representative sample of the Austrian population that consistent of 24 waves, this study analyzed the effects of COVID-19 related social media use on well-being.
-A random effect model that separated between person relations from within-person effects and that controlled for several third variables showed that within-person effects were trivial.
+In this study I analyzed the effects of COVID-19 related social media use on well-being.
+The data come from a panel study with 24 waves that is representative of the Austrian population.
+A random effect model, which separated between person relations from within-person effects and which controlled for several third variables, showed that within-person effects were trivial.
 People who used social media more than usual to learn about COVID-19 did not show changes in their well-being.
-As a result, the results imply that COVID-19 related social media use is irrelevant for people's well-being.
-Other factors among the third variables that were measured revealed much larger effects, implying that the relevance of specific types of social media use for well-being is limited.
-Popular fears that "doomscrolling" or overusing social media during times of crises might not be justified.
+As a result, the results imply that COVID-19 related social media use does not seem to be particularly relevant for people's well-being.
+Other factors among the third variables that were measured revealed larger effects or relations, implying that well-being is rather determined by aspects such as health, employment, or locus of control.
+According to this study, popular fears that "doomscrolling" or overusing social media during times of crises do not seem to be justified.
 
-The study is not aligned with a recent cross-sectional study by @bendauAssociationsCOVID19Related2021 that showed negative relations between social media and well-being.
+The study is not aligned with a recent cross-sectional study by @bendauAssociationsCOVID19Related2021, which showed negative relations between social media and well-being.
 However, @bendauAssociationsCOVID19Related2021 analyzed cross-sectional data on a between-person level, while not controlling for third variables, which does not allow to make causal inferences.
@@
 Hence, to document how effects unfold it needs future research employing different study designs with different time lags.
+In addition, more thought needs to be invested in what relevant stable and nonvarying factor to control for.
 
-The predefined SESOI of _b_ = .30 was potentially too large.
-Media use is only one aspect of multiple factors that simultaneously affect well-being.
+Although I had already reduced the predefined SESOIs to be less conservative, potentially they were still too large.
+Media use is only one aspect of several factors that simultaneously affect well-being.
 Is it realistic that extremely changing only _one_ of these aspect (e.g., by completely stopping the use of social media) should already manifest in a detectable change in well-being?
-Or would it make more sense to say, if you regularly start doing _two_ activities (e.g. regularly exercising and establishing a reading habit) together should then show in improvements to well-being?
-In other words, if the beneficial effect of a particular activity is large enough, if people implement _two_ of them then they should actually feel a difference.
+Or would it make more sense to expect that if people regularly start doing _two_ activities (e.g.
+regularly exercising _and_ establishing a reading habit) together should show in perceivable improvements to well-being?
+In other words, if the beneficial effect of a particular activity is large enough, people should actually feel a difference if they implement two of those activities. 
 Practically, this would imply a SESOI half as large as I have defined here, that is _b_ = |.15| for well-being and _b_ = |.075|. 
 In this case, this would not make a difference, as even with these more liberal thresholds all but one effect would still be completely in the null region.
-However, at all events future research needs to start a discussion on what effect sizes are considered trivial, and this study is one of the first to provide some concrete guidelines.
+However, at all events future research needs to start a discussion on what effect sizes are considered meaningful and relevant, and with this study I hope to provide some first concrete guidelines.
 
@@
 
-Social media use was measured with an ordinal variable, however in the analyses it was treated as a numerical one.
-If treated as an ordinal one, it would have been necessary to analyze four different contrasts for each media measure, which plus the differentiation between between and within factor would have produced eight different measures, we would have made the model exceedingly complex.
+<!-- Social media use was measured with an ordinal variable, however in the analyses it was treated as a numerical one. -->
+<!-- If treated as an ordinal one, it would have been necessary to analyze four different contrasts for each media measure, which plus the differentiation between between and within factor would have produced eight different measures, we would have made the model exceedingly complex.
+-->
 
@@
 In this study, COVID-19 related social media use did not causally affect several indicators of well-being, including life satisfaction, positive affect, and negative affect.
-However, factors other than social media use did affect well-being, such as income levels or access to a balcony.
-If it's the aim to improve well-being, it might hence be fruitful not to focus on social media but to address other, potentially more pressing societal problems related to inequality.
+However, factors other than social media use were meaningfully related to well-being, such as physical health, employment, or believing that one is in control of one's life.
+If it's the aim to improve well-being, it might hence be fruitful not to focus on social media but to address other, potentially more pressing societal problems related to inequality or mental health.
 
@@
 
+\newpage
+
 # Competing Interests
@@
 
-The data are shared on AUSSDA, see https://doi.org/10.11587/28KQNS, and can be used for scientific purposes only.
+The data are shared on AUSSDA, see https://doi.org/10.11587/28KQNS.
+The data can only be used for scientific purposes.

### 2021-11-05 manuscript.tex
- implemented feedback from Nick - updated analyses - provided additional plots for control variables

@@
 In times of crisis such as the Corona pandemic, it is important that citizens stay informed about recent events, the latest political decisions, or mandatory protection measures.
-As a result, to attain relevant information many people use various types of media, and increasingly social media.
+To stay informed, many people use various types of media, and increasingly social media.
 However, because social media are particularly engaging, some find it hard to disconnect and cannot stop `doomscrolling.' In this preregistered study, I investigate whether using social media for COVID-19 related topics might put personal well-being at risk.
@@
 The data were analzed using random effects cross lagged panel models, controlling for several stable and varying covariates.
-Results showed that COVID-19 related social media use did not meaningfully affect several types of well-being, including life satisfaction, positive affect, and negative effect.
-This pertains to both passive and active social media use, and all the prominent channels such as Facebook, WhatsApp, or YouTube.
-As a result, this study suggests fears that social media use during times of crisis might be detrimental for well-being can be put to rest.
+Results showed that the effects of various types of COVID-19 related social media use on several types of well-being were very small, arguably too small to matter.
+The findings suggests fears that social media use during times of crisis might be detrimental for well-being are likely to be unfounded.
 }
@@
 During the COVID-19 pandemic,
-several events unfolded in quick succession.
+numerous events unfolded in quick succession, and several open questions emerged.
 How dangerous is the virus? Is it spreading in my region? How is it transmitted, and how can I protect myself?
 Because for many it was (and still is) a matter of life or death, citizens had to stay informed regarding the latest developments.
-Governments around the world implemented safety measures, from wearing masks or keeping physical distance, to complete lockdowns.
-In this extraordinary situation, people hence used media excessively, and especially social media were at an all time high (Statista, 2021).
-A new phenomenon ``termed doomscrolling'' emerged:
+Governments around the world implemented safety measures, ranging from wearing masks or keeping physical distance to complete lockdowns.
+In this extraordinary situation, people hence used media excessively to stay informed, and especially social media were at an all time high (Statista, 2021).
+A new phenomenon termed ``doomscrolling'' emerged:
 People could not stop using social media to attain COVID-19 related news.
 
-Several people reported that they were glued to their screens and could not pursue other relevant activities such as working, taking a break, or even care-work.
-Increasingly, it was asked whether such an increase in social media use could still be considered useful and adaptive, or whether it created an additional and new psychological danger for the users' mental health (Klein, 2021).
-A study with 6,233 people from Germany found that ``{[}f{]}requency, duration and diversity of media exposure were positively associated with more symptoms of depression and unspecific and COVID-19 specific anxiety'' (Bendau et al., 2021).
+Several people reported that they were glued to their screens and could not pursue other relevant activities such as working, taking a break, or even care-work (Klein, 2021).
+As doomscrolling increased it became doubtful whether such a surge in social media use could still be considered useful and adaptive, or whether it created an additional and new psychological danger for the users' mental health (Sandstrom, Buchanan, Aknin, \& Lotun, 2021).
+A study with 6,233 people from Germany found that ``{[}f{]}requency, duration and diversity of media exposure were positively associated with more symptoms of depression and unspecific and COVID-19 specific anxiety'' (Bendau et al., 2021), implying negative consequences for well-being.
 
@@
 The underlying theories that guided the selection of variables for this study are the two-continua model of mental health (Greenspoon \& Saklofske, 2001) and the hierarchical taxonomy of computer-mediated medation (Meier \& Reinecke, 2020).
-According to the two-continua model of mental health, mental health consists of two dimensions, psychopathology and well-being.
-Because the aim of this study is better to understand typical users and everyday contexts, my focus will be on well-being.
+According to the two-continua model of mental health, mental health consists of two dimensions: psychopathology and well-being.
 Well-being can be differentiated into subjective and psychological well-being (Diener, Lucas, \& Oishi, 2018).
@@
 Subjective well-being is primarily about achieving positive affect and avoiding negative affect.
-In my view, life satisfaction is a meta concept above both psychological and subjective well-being, representing a meta-appraisal of one's life.
-Notably, life satisfaction is very stable and fluctuates only little, whereas it's the exact opposite for affect (Dienlin \& Johannes, 2020).
+On of the most prominent indicators of well-being is life satisfaction.
+In my view, life satisfaction is a meta concept that combines psychological and subjective well-being and represents a meta-appraisal of one's life.
+Notably, life satisfaction is stable and fluctuates only little, whereas it's the exact opposite for affect (Dienlin \& Johannes, 2020).
 To capture well-being in this study, I will thus build on life satisfaction, positive affect, and negative affect.
@@
 Whereas the first four levels are focused on the channel, the last two on the type of communication.
-To measure social media use for consumption of specific news, I here employ both the channel and the communication perspective.
+To measure social media use for consumption of specific news, I here employ both the channel and the communication perspective, which together provides a more nuanced understanding of communication.
 First, I will analyze how using the most prominent branded applications affect well-being, and whether this effect changes across applications.
+According to Meier and Reinecke (2020), branded apps are separate entities with different effects.
+Therefore, Twitter might have a different effect than WhatsApp because of their different affordances.
+For example, Waterloo, Baumgartner, Peter, and Valkenburg (2018) found that it is more adequate to express negative emotions on WhatsApp than on Twitter or on Instagram.
 The branded applications investigated here are Facebook, Twitter, Instagram, WhatsApp, and YouTube.
@@
 Worth noting, this study is not about \emph{general} social media during times of COVID, but on social media use focused on COVID-19 related news and interactions.
-For example, posting about the pandemic or retweeting COVID-19 related posts.
+Example of such media use include posting about the pandemic or retweeting COVID-19 related posts.
 
@@
 (2021) found that people who used social media as a primary source of information reported on average ``significantly more unspecific anxiety and depression {[}{]} and significantly more specific COVID-19 related anxiety symptoms'' (p.~288).
-Hence, this might hint at potential negative effects of social media as news use---but note that this finding comes from a between-person relation stemming from a cross-sectional data.
+Hence, this might hint at potential negative effects of social media as news use---but note that this finding comes from a between-person relation stemming from a cross-sectional data (see below).
 Hence, we don't know whether the differences in mental health and well-being are due to social media use or other third variables, such as age or education.
+Eden, Johnson, Reinecke, and Grady (2020) analyzed the media use of 425 US college students during the first wave of the pandemic, and found both positive and negative relations with well-being.
+In a sample of 312 respondents collected via Amazon Mechanical Turk, it was found that people who used media to attain information were more lonely and less satisfied with their lives (Choi \& Choung, 2021).
+Stainback, Hearne, and Trieu (2020) analyzed a large-scale study with 11,537 respondents from the US and found that increased COVID-19 media consumption was related to more psychological distress.
 
-In general, so far there is only little research on news-related social media use and well-being.
+In general, so far there is only little empirical research on news-related social media use and well-being.
 On the other hand, the question of whether and how \emph{general} social media use---or other, more specific types such as active or passive use---affect well-being is well-researched.
 A meta review (i.e., an analysis of meta-analyses), found that the relation between social media use and well-being is likely in the negative spectrum---but very small, potentially too small to matter (Meier \& Reinecke, 2020).
-This is well aligned with several new studies that employed the most advanced methods (Keresteš \& Štulhofer, 2020; Orben, Dienlin, \& Przybylski, 2019; Przybylski, Nguyen, Law, \& Weinstein, 2021; Schemer, Masur, Geiß, Müller, \& Schäfer, 2021).
-For example, Beyens, Pouwels, Driel, Keijsers, and Valkenburg (2020) reported that although for some users (roughly one quarter) the effects were negative, for almost the same amount of users the effects were positive, while for the majority they were neutral.
-At all events, in general most effects are somewhere between trivial and small.
+What determines whether an effect is considered trivial or small?
+If we refer to standardized effect sizes, according to Cohen (1992) small effect sizes start at \emph{r} = .10.
+And indeed, several if not most meta-analyses find effect sizes below that threshold (Ferguson et al., 2021; Huang, 2017; Meier \& Reinecke, 2020).
 
-What determines a trivial or a small effect is a difficult question (see below).
-If we refer to standardized effect sizes, according to Cohen (Cohen, 1992) small effect sizes start at \emph{r} = .10.
-However, most meta-analyses find effect sizes below that threshold (Ferguson et al., 2021; Huang, 2017; Meier \& Reinecke, 2020).
-As a result, I think it's most convincing to expect trivial to small effects also in the case of COVID-19 related social media use.
+These overviews are well aligned with several individual new studies that employed the most advanced methods (Keresteš \& Štulhofer, 2020; Orben, Dienlin, \& Przybylski, 2019; Przybylski, Nguyen, Law, \& Weinstein, 2021; Schemer, Masur, Geiß, Müller, \& Schäfer, 2021).
+For example, Beyens, Pouwels, Driel, Keijsers, and Valkenburg (n.d.) reported that although for some users (roughly one quarter) the effects were negative, for almost the same number of users the effects were positive, while for the majority they were neutral.
+In conclusion, most effects are somewhere between trivial and small.
+I think it's most convincing to expect trivial to small effects also in the case of COVID-19 related social media use.
 
-From a theoretical perspective, why could COVID-19 related social media use be detrimental?
-Above everything, one can reasonably assume a \emph{direct} negative effect on well-being, especially on positive or negative effect.
-Dangers, inequalities, corruption -- these were the headlines across many countries worldwide.
+From a theoretical perspective, how could we explain how COVID-19 related social media use might affect well-being?
+In what follows, I outline potential arguments as to why the effect might be positive or negative, direct or indirect.
+In advance, there does not seem to be a clear winner, and it's likely that both positive and negative effects are equally strong.
+
+First, one could reasonably assume a \emph{direct} negative effect on well-being---especially on positive or negative affect, which is more volatile and fluctuating.
+Dangers, inequalities, corruption---these were the headlines across many countries worldwide.
 If one learns about such things, the initial reaction might be shock, fear, or dismay.
-Repeatedly consuming such news might be depressing, perhaps even changing some general perspective on life.
-So, just like being hit by a hammer hurts and we don't need any ``mediating mechanism,'' this could be the case here as well.
+Repeatedly consuming such news might be depressing, perhaps even changing some general perspectives on life, without any further mediating processes.
 
-That said, because not all news were negative, because many people showed solidarity and compassion,
-there were also positive and potentially uplifting news.
-However, in light of a worldwide pandemic with millions of deaths, the negative direct effect seems more plausible.
+That said, because not all news were negative, because many people showed solidarity and compassion, there were also positive and potentially uplifting news.
 
@@
 When doomscrolling, users are captivated to such an extent that they cannot stop using social media.
-For example, during the pandemic social media uses was at an all-time high in the US (Statista, 2021).
+For example, during the pandemic social media use was at an all-time high in the US (Statista, 2021).
 As has been expressed by many before, it is most likely that moderate social media use is not detrimental (Orben, 2020).
-However, overuse is more critical, and several studies have showed more pronounced negative effects for extreme users (Przybylski \& Weinstein, 2017).
+Overuse, however, is more critical, and several studies have shown more pronounced negative effects for extreme users (Przybylski \& Weinstein, 2017).
+Overuse likely impairs well-being if it replaces other meaningful or functional activities such as meeting others, working, actively relaxing, or exercising.
 So if a society collectively overuses social media, there is potential for negative effects.
-Overuse likely impairs well-being if it replaces other meaningful or functional activities such as meeting others, working, actively relaxing, or exercising.
-If overuse replaces such activities it's reasonable to assume that it's also detrimental.
 
 On the other hand, one can make the case that overuse can be beneficial \emph{in times of a pandemic}, even if it's mainly COVID-19 related.
-Exchanging COVID-19 related messages with friends via WhatsApp might replace the in-person contact one would have otherwise, but which is logically impossible at that time.
+Exchanging COVID-19 related messages with friends via WhatsApp might replace the in-person contact one would have otherwise, but which is effectively impossible at that time.
 At times where meaningful and functional activities are explicitly or implicitly prohibited, using social media to exchange about COVID-19 might not be the worst idea.
 In fact, given that nowadays a large number of experts, scientists, and politicians converse directly on social media, one can get first-hand information on the current developments.
-On the other hand, there is of course also much disinformation, and ``bingeing'' on COVID-19 fake news might also pose risks for impaired well-being.
 
-Together, the strongest argument seems to be that the effects of social media on well-being are, on average, generally small at best.
+Together, the strongest argument seems to be that \emph{in general} effects of social media on well-being are, on average, small at best.
 Because this study looks at only \emph{one part} of social media use---namely, COVID-19 related interactions---it is even more focused, and the overall potential of the effects should diminish even further.
@@
 Testing this hypothesis, however, is not trivial.
-First, in contrast to most hypothesis typically posited in the social sciences, it explicates an effect size, a so-called smallest effect size of interest (SESOI).
+First, in contrast to most hypothesis typically posited in the social sciences, it implicitly contains an effect size, a so-called smallest effect size of interest (SESOI).
 Effectively testing this hypothesis hence necessitates to define what's considered a ``trivial effect size'' and what's not.
@@
 In this study, COVID-19 related social media use was measured on a 5-point scale, ranging from 1 = \emph{never} to 5 = \emph{several times a day}.
-Thus, the example from above would imply that a change of five units in social media use should correspond to a noticeable change in well-being.
+Thus, the example from above would imply that a change of four units in social media use should correspond to a noticeable change in well-being.
 But what's a noticeable change in well-being?
 According to Norman, Sloan, and Wyrwich (2003), people can reliably differentiate between \emph{seven} levels of satisfaction with health.
-So we could state that a five unit change in social media use should result in a one unit change in satisfaction.
+So we could state that a four unit change in social media use should result in a one unit change in satisfaction.
 Statistically, in a regression, \emph{b} estimates the change in the dependent variable if the independent variable increases by one point.
-Transferred to our example, we would hence expect a \emph{b} of 1 / 5 = .20.
+Transferred to our example, we would hence expect \emph{b} = 1 / 4 = .25.
 
-In this study, life satisfaction was measured with 11 units, hence more than people can reliably detect.
-Hence, we would expect a 11 / 7 * .20 change, that is a \emph{b} of at least .31.
-For affect, which was measured on a 5-point scale, our SESOI would be 5 / 7 * .2 = .14.
-In order not to exaggerate precision, these numbers will be rounded.
-Plus, because we are agnostic as to whether the effects are positively or negatively nontrivial, this leads to an indifference region ranging from \emph{b} = -.30 to \emph{b} = .30 (and \emph{b} = -.15 to \emph{b} = .15 for affect).
+In this study, life satisfaction was measured with 11 units, hence more than seven degrees people can reliably differentiate.
+We hence need to transform a 1-point change on a 7-point scale to an equivalent change on an 11-point scale.
+This is 11 / 7 = 1.57.
+Hence, we now expect \emph{b} = 1/4 * 11/7 = 0.39.
+For affect, which was measured on a 5-point scale, our SESOI would therefore be \emph{b} = 1/4 * 5/7 = 0.18.
+Because we are agnostic as to whether the effects are positively or negatively nontrivial, the null region will include negative and positive effects (in Bayesian terms, this is called the region of practical equivalence {[}ROPE{]}).
+In order not to exaggerate precision, and to be a bit less conservative, these numbers will be reduced to nearby thresholds.
+Note that other researchers also decrease or recommend decreasing thresholds for effect sizes when analyzing within-person or cumulative effects (Beyens, Pouwels, Driel, Keijsers, and Valkenburg (n.d.); Funder and Ozer (2019)).
+Together, this leads to an indifference region ranging from \emph{b} = -.30 to \emph{b} = .30 for life satisfaction, and \emph{b} = -.15 to \emph{b} = .15 for positive and negative affect.
 
@@
 The hypothesis explicitly states a causal effect.
-In non-experimental designs, causality can be approximated using longitudinal designs.
+In non-experimental designs, longitudinal designs help investigate causality.
 Using longitudinal designs alone, however, is not sufficient for correct causal statements.
-In addition, it is crucial also to control for third variables, and importantly also for \emph{varying} third variables.
+In addition, we also need to control for third variables, and importantly also for \emph{varying} third variables.
 
@@
 Obviously, we can never be entirely sure to have included all varying covariates, which makes absolute statements regarding causality impossible.
-However, when controlling for many varying variables, we can be much more certain that we measured causality correctly.
-The aim should still be to collect as many relevant varying and nonvarying third variables as possible, while knowing that absolute certainty cannot be reached.
+In addition, when determining the overall causal effect, we should \emph{not} control for mediating variables (Rohrer, 2018).
+Doing so would bias and distort our assessment of the role of social media use.
+However, when controlling for relevant variables (that aren't mediators), we can be much more certain that we measured causality correctly.
+The aim should therefore be to collect as many \emph{relevant} varying and nonvarying third variables as possible, while knowing that absolute certainty regarding causality cannot be reached.
 
-When looking for suitable control candidates, ideally we find variables that affect both media use and well-being, because controlling for these factors will isolate the actual effect of social media use on well-being.
+When looking for suitable control candidates, ideally we find variables that affect both media use and well-being.
+Controlling for these factors will isolate the actual effect of social media use on well-being.
 We can also control for variables that affect \emph{only} well-being.
-However, in doing so nothing is gained or lost because the effects of social media use would remain exactly the same (Kline, 2016).
-Crucially, when determining the general causal effect of social media use we should \emph{not} control for mediating variables (Rohrer, 2018).
-Doing so would bias and distort our assessment of the role of social media use.
+However, in doing so not much is gained or lost because the effects of social media use would remain virtually the same (Kline, 2016).
 
-In this study, I will hence control for the following variables, which either have been shown or a likely to affect both social media use and well-being.
-(I'll additionally include variables that likely affect only well-being, also to obtain a comparison benchmark for social media effects):
+In this study, I hence plan to control for the following variables, which either have been shown or a likely to affect both social media use and well-being, while not being mediators.
 Gender, age, education, Austria country of birth, Austria country of birth of parents, text-based news consumption, video-based news consumption, residency Vienna, household size, health, living space, access to garden, access to balcony, employment, work hours per week, being in home-office, household income, outdoor activities, satisfaction with democracy, disposition to take risks, and locus of control.
@@
 If a specific person changes their media diet, we need to measure how this affects \emph{them}.
-Between person comparison from cross-sectional data, where participants are interviewed only once, cannot provide such data.
+Between person comparison from cross-sectional data, where participants are interviewed only once, cannot provide such insights.
 In this study, I will hence differentiate between within-person effects and between-person relations.
+Because it is explicitly causal, in answering the hypothesis I will thus consider only the within-person effects.
 
@@
 The (anonymous) preregistration can be accessed here: \url{https://osf.io/87b24/?view_only=b2289b6fec214fa88ee75a18d45c18f3}.
-Because in this study I analyze data from an already existing large scale data set, the Austrian Corona Panel Project, all of these steps were preregistered prior to accessing the data.
+Because in this study I analyze data from an already existing large-scale data set, the Austrian Corona Panel Project, all of these steps were preregistered prior to accessing the data.
 The preregistration was designed on the basis of the panel documentation online (Kittel et al., 2020).
@@
 In order to participate in the study, the respondents needed to be Austrian residents and had to be at least 14 years old.
-The average age was 42 years, 49 percent were male, 13.60 percent had a University degree, and 5.07 percent were currently unemployed.
+The average age was 42 years, 49 percent were male, 3.97 percent had a University degree, and NA percent were currently unemployed.
 
@@
 If the effects interval and the null region overlap (e.g., \emph{b} = .25, {[}95\% CI: .15, .35{]}), the hypothesis is not supported and the results are considered inconclusive, whereas a meaningful negative effect is rejected.
-If the confidence falls completely outside the null-region (e.g., \emph{b} = .4, {[}95\% CI: .35, .45{]}), the hypothesis is rejected and the existence of a meaningful positive effect is supported.
+If the confidence falls completely outside of the null-region (e.g., \emph{b} = .4, {[}95\% CI: .35, .45{]}), the hypothesis is rejected and the existence of a meaningful positive effect is supported.
 
@@
 All predictors---that is, social media activities and social media channels, within and between-person factors, stably and varying covariates---were included in each of the three models.
-The data were hierarchical, and responses were nested in participants and waves.
-Nesting in participants allowed to separate within-person effects from between-person relations.
+The data were hierarchical, and responses were nested in participants and waves (put differently, participants and waves were implemented as random effects).
+Nesting in participants allowed to separate between-person relations from within-person effects.
 Nesting in waves allowed to control for general developments, such as general decrease in well-being in the population, for example due to lockdown measures (hence, there was no need additionally to control for specific phases or measures of the lockdown).
 
-The factorial validity of the scales were tested with confirmatory factor analyses (CFA).
-If Mardia's test showed that the assumption of multivariate normality was violated, the more robust Satorra-Bentler scaled and mean-adjusted test statistic (MLM) was used as estimator.
-To avoid overfitting, scales were tested on more liberal fit criteria (CFI \textgreater{} .90, TLI \textgreater{} .90, RMSEA \textless. .10, SRMR \textless{} .10) (Kline, 2016).
-Because REWB-models cannot model scales as latent variables, to increase precision factor scores exported from the CFAs were used.
 For more information on the analyses, see \href{https://xmtra.github.io/Austrian_Corona_Panel_Project/index.html}{companion website}.
@@
 In what follows, I briefly list all variables that were collected.
+For the variables' means, range, and variance, see Table \ref{tab:descriptives}.
 For a complete list of all items and item characteristics, see \href{https://xmtra.github.io/Austrian_Corona_Panel_Project/index.html}{companion website}.
@@
 The response options ranged from 0 (\emph{extremely unsatisfied}) to 10 (\emph{extremely satisfied}).
-The variable's average across all waves was \emph{M} = 6.59, ranging from \emph{M}\textsubscript{min} = 6.39 to \emph{M}\textsubscript{max} = 6.79.
-The average standard deviation was \emph{SD} = 1.68.
 
 To capture positive affect, respondents were asked how often in the last week they felt (a) calm and relaxed, (b) happy, and (c) full of energy.
-The response options were 1 (\emph{never}), 2 (\emph{on some days} ), 3 (\emph{several times per week}), 4 (almost every day), and 5 (daily).
-The variable's average across all waves was \emph{M} = 3.15, ranging from \emph{M}\textsubscript{min} = 3.05 to \emph{M}\textsubscript{max} = 3.29.
-The average standard deviation was \emph{SD} = 0.57.
+The response options were 1 (\emph{never}), 2 (\emph{on some days}), 3 (\emph{several times per week}), 4 (\emph{almost every day}), and 5 (\emph{daily}).
 
 For negative affect, respondents were asked how often in the last week they felt (a) lonely, (b) aggravated, (c) so depressed, that nothing could lift you up, (d) very nervous, (e) anxious, and (h) glum and sad.
-The response options were 1 (\emph{never}), 2 (\emph{on some days} ), 3 (\emph{several times per week}), 4 (almost every day), and 5 (daily).
-The variable's average across all waves was \emph{M} = 1.73, ranging from \emph{M}\textsubscript{min} = 1.66 to \emph{M}\textsubscript{max} = 1.81.
-The average standard deviation was \emph{SD} = 0.39.
+The response options were 1 (\emph{never}), 2 (\emph{on some days}), 3 (\emph{several times per week}), 4 (\emph{almost every day}), and 5 (\emph{daily}).
 
-All three items were measured on each wave.
+All three variables were measured on each wave.
 
@@
 The items were inverted for the analyses.
-The variable's average across all waves was \emph{M} = 6.59, ranging from \emph{M}\textsubscript{min} = 6.39 to \emph{M}\textsubscript{max} = 6.79.
-The average standard deviation was \emph{SD} = 1.68.
 
@@
 
-All items were measured every second wave.
-However, freshly recruited respondents always answered these questions.
+Social media use was measured for everyone on waves 1, 2, 8, 17, and 23.
+Freshly recruited respondents always answered the questions on social media use.
+
+\begin{table}[tbp]
+
+\begin{center}
+\begin{threeparttable}
+
+\caption{\label{tab:descriptives}Descriptives of the main variables.}
+
+\begin{tabular}{lllll}
+\toprule
+ & \multicolumn{1}{c}{sd} & \multicolumn{1}{c}{min} & \multicolumn{1}{c}{max} & \multicolumn{1}{c}{mean}\\
+\midrule
+Well-being &  &  &  & \\
+\ \ \ Life satisfaction & 1.68 & 6.39 & 6.79 & 6.59\\
+\ \ \ Positive affect & 0.57 & 3.05 & 3.29 & 3.15\\
+\ \ \ Negative affect & 0.39 & 1.66 & 1.81 & 1.73\\
+Social media use &  &  &  & \\
+\ \ \ Read & 1.03 & 2.09 & 2.92 & 2.42\\
+\ \ \ Like \& share & 0.86 & 1.61 & 1.92 & 1.78\\
+\ \ \ Posting & 0.63 & 1.33 & 1.47 & 1.39\\
+Social media channel &  &  &  & \\
+\ \ \ Facebook & 0.96 & 2.34 & 2.68 & 2.45\\
+\ \ \ Twitter & 0.52 & 1.16 & 1.72 & 1.36\\
+\ \ \ Instagram & 0.82 & 1.85 & 2.65 & 2.09\\
+\ \ \ WhatsApp & 1.23 & 2.29 & 2.62 & 2.46\\
+\ \ \ YouTube & 0.88 & 1.77 & 2.32 & 2.01\\
+\bottomrule
+\end{tabular}
+
+\end{threeparttable}
+\end{center}
+
+\end{table}
 
@@
 The effects of COVID-19 related social media use were controlled for the following stable variables:
-(a) gender (answer options: female, male, diverse), (b) age, (c) education (10 ordinal options), (d) Austria country of birth (yes/no), (e) Austria parents' country of birth (no parent, one parent, both parents).
-We controlled also for the following varying covariates:
-(a) text-based media news consumption, (b) video-based media news consumption, (c) residency is Vienna, (d) household size, (e) Self-reported physical health, (f) Living space (in squaremeter), (g) access to balcony, (h) access to garden, (i) Employment status, (j) Work hours per week, (k) Home office, (l) household income, (m) outdoor activities, (o) satisfaction with democracy, (p) disposition to take risks, (q) locus of control.
+(a) gender (answer options: female, male, diverse), (b) age, (c) education (10 options), (d) Austria country of birth (yes/no), (e) Austria parents' country of birth (no parent, one parent, both parents).
+I originally planned to implement other variables as varying covariates.
+However, because they were not measured often enough or at the time when social media use was measured, I implemented them as stable variables using their average values across all waves.
+This includes (a) text-based media news consumption, (b) video-based media news consumption, (c) residency is Vienna, (d) self-reported physical health, (e) living space (in squaremeter), (f) access to balcony, (g) access to garden, (h) employment status, (i) disposition to take risks, and (j) locus of control.
+I controlled also for the following varying covariates: (a) outdoor activities, (b) satisfaction with democracy.
+Because it lead to too much attrition in the sample, I did not control for (a) household size, (b) work hours per week, (c) home office, (d) household income.
 
@@
 When looking at the variables from a descriptive perspective, we see that all well-being measures did not change substantially across the different phases of the pandemic.
-COVID-19 related media use, however, increased during the beginning of the study and remained stable after approximately six waves.
+COVID-19 related media use, however, decreased during the beginning of the study and remained stable after approximately six waves.
+The initial decrease might be explained by the fact that the collection of data began in end of March 2020, hence approximately three months after the pandemic began.
+It could be that after an initial stark surge, COVID-19 related social media use was already declining, returning to more normal levels.
 
-\begin{figure}[!h]
+\begin{figure}
 \includegraphics[width=\textwidth]{figures/fig_descriptives} \caption{Development of well-being and media use measures across the pandemic.
@@
 The study's hypothesis was that the effects of all types of social media use on well-being will be trivial.
-Regarding the different types of \emph{communication}, that is reading vs.~sharing vs.~posting, all within-person effects fell completely within the a-priori defined SESOIs.
-For example, respondents who used social media more frequently than usual to read about COVID-19 related topics did not show a simultaneous change in life satisfaction (\emph{b} = -0.03 {[}95\% CI -0.09, 0.02{]}).
-Only one effect did not include zero.
-If people posted more about COVID-19 related aspects than they usually did, life satisfaction increased (\emph{b} = 0.11 {[}95\% CI 0.01, 0.2{]}).
-However, because the effect did not fall outside our null region, it's likely not large enough to be considered meaningful.
+Regarding the different types of \emph{communication}---that is, reading vs.~sharing vs.~posting---all within-person effects fell completely within the a-priori defined SESOIs (see Figure \ref{fig:res-activity}).
+For example, respondents who used social media more frequently than usual to read about COVID-19 related topics did not show a simultaneous change in life satisfaction (\emph{b} = 0.04 {[}95\% CI -0.02, 0.09{]}).
+All confidence intervals included zero; hence, all effects were non-significant.
 As a result, the hypothesis was supported for all types of COVID-19 related social media communication.
 
-Regarding between-person relations, about which no hypotheses were formulated, only two effects didn't include zero.
-First, respondents who across all waves used social media more frequently than others to read about COVID-19 reported higher levels of positive affect than others (\emph{b} = 0.05 {[}95\% CI 0.01, 0.09{]}).
-However, note that this effect still fell completely within the null-region.
-Hence, although positive the effect was considered too small to matter practically.
-Second, respondents who across all waves used social media more frequently than others to read about COVID-19 reported lower levels of negative affect than others (\emph{b} = 0.01 {[}95\% CI -0.02, 0.04{]}).
-The effect was partially outside the null region, hence the effect might be large enough to be practically relevant, and the effect that it's trivial cannot be rejected.
-
-\begin{figure}[!h]
+\begin{figure}
 \includegraphics[width=\textwidth]{figures/fig_results_activity} \caption{The effects of various types of social media use on three indicators of well-being.
 Effects are controlled for a large number of covariates (see text).
-The SESOI was 0.30 for life satisfaction and .15 for affects; hence, no effect is considered meaningful theoretically.}\label{fig:fig-res-activity}
+The SESOI was 0.30 for life satisfaction and .15 for affects; hence, no effect is considered meaningful theoretically.}\label{fig:res-activity}
 \end{figure}
 
-Regarding the COVID-19 related use of social media \emph{channels}, the results were virtually the same.
-Changes in the frequency of using different social media channels to attain information regarding COVID-19 were unrelated to meaningful changes in well-being.
-For example, respondents who used Facebook more frequently than usual to learn about COVID-19 did not show a simultaneous change in well-being (\emph{b} = 0.03 {[}95\% CI -0.03, 0.1{]}).
-Only one effect was substantially larger than zero.
-Respondents who used Instagram more frequently than usual to attain COVID-19 related news reported slightly lowers simultaneous levels of life satisfaction then usual (\emph{b} = -0.08 {[}95\% CI -0.15, \textgreater{} -0.01{]}).
-However, this effect was still completely within the null region, hence not large enough to be considered meaningful.
+Regarding between-person relations, about which no hypotheses were formulated, only one effect didn't include zero.
+Respondents who across all waves used social media more frequently than others to write posts on COVID-19 reported higher levels of negative affect than others (\emph{b} = 0.08 {[}95\% CI 0.04, 0.12{]}).
+The effect was still completely inside of the null region, hence likely not large enough to be practically relevant.
+
+Regarding the COVID-19 related use of social media \emph{channels}, the results were very comparable.
+Changes in the frequency of using different social media channels to attain information regarding COVID-19 were unrelated to meaningful changes in well-being (see Figure \ref{fig:res-channels}).
+For example, respondents who used Facebook more frequently than usual to learn about COVID-19 did not show a simultaneous change in well-being (\emph{b} = -0.04 {[}95\% CI -0.1, 0.02{]}).
+Only two effects differed substantially from zero.
+Respondents who used Instagram more frequently than usual to attain COVID-19 related news reported slightly higher simultaneous levels of life satisfaction then usual (\emph{b} = 0.08 {[}95\% CI 0, 0.15{]}).
+Respondents who used Twitter more frequently than usual to attain COVID-19 related news reported slightly lower simultaneous levels of life satisfaction then usual (\emph{b} = -0.15 {[}95\% CI -0.27, -0.04{]}).
+However, both effects were still completely inside of the null region, hence not large enough to be considered meaningful.
 In sum, the hypothesis was supported for the COVID-19 related use of all types of social media channels.
 
-In terms of between-person relations---which, again, weren't included in the hypotheses---no relations crossed or fell outside the null region.
-However, some relations did not included zero.
-For example, respondents who across all waves used Instagram more frequently than others for COVID-19 related reasons reported lower levels of life satisfaction (\emph{b} = -0.11 {[}95\% CI -0.19, -0.03{]}).
-Respondents who were more active on WhatsApp compared to others reported slightly lower levels of positive affect (\emph{b} = -0.03 {[}95\% CI -0.07, \textgreater{} -0.01{]}).
-Respondents who compared to others were more active on Twitter and YouTube reported lover levels of negative affect (Twitter: \emph{b} = -0.04 {[}95\% CI -0.07, \textgreater{} -0.01{]};
-YouTube \emph{b} = -0.06 {[}95\% CI -0.08, -0.03{]} .
-However, please note that all these affect not considered large enough to be practically relevant.
-
-\begin{figure}[!h]
+\begin{figure}
 \includegraphics[width=\textwidth]{figures/fig_results_channel} \caption{The effects of using various social media applications on three indicators of well-being.
 Effects are controlled for a large number of covariates (see text).
-The SESOI was 0.30 for life satisfaction and .15 for affects; hence, no effect is considered meaningful theoretically.}\label{fig:fig-res-channels}
+The SESOI was 0.30 for life satisfaction and .15 for affects; hence, no effect is considered meaningful theoretically.}\label{fig:res-channels}
 \end{figure}
+
+In terms of between-person relations---which, again, weren't included in the hypotheses---no relations crossed or fell outside of the null region.
+Only one relation did not include zero, was hence statistically significant.
+Respondents who across all waves used YouTube more frequently than others for COVID-19 related reasons reported marginally higher levels of negative affect (\emph{b} = 0.03 {[}95\% CI \textless{} 0.01, 0.05{]}).
+However, please note that this effect was not large enough to be considered practically relevant.
 
@@
 
-{[}Report results of some control variables.{]}
+In what follows, I briefly report also the results of some covariates.
+Several variables showed large associations with well-being.
+Because each variable has a different scaling, we would again need to define a SESOI for each variable, which cannot be implemented here.
+But note that several variables would fall outside of such a SESOI.
+This includes for example internal locus of control, health, or employment.
+For a brief overview, see Figure \ref{fig:res-control}.
+
+\begin{figure}
+\includegraphics[width=\textwidth]{figures/fig_results_control} \caption{Results of selected covariates.}\label{fig:res-control}
+\end{figure}
 
@@
 
-Using a panel study with a representative sample of the Austrian population that consistent of 24 waves, this study analyzed the effects of COVID-19 related social media use on well-being.
-A random effect model that separated between person relations from within-person effects and that controlled for several third variables showed that within-person effects were trivial.
+In this study I analyzed the effects of COVID-19 related social media use on well-being.
+The data come from a panel study with 24 waves that is representative of the Austrian population.
+A random effect model, which separated between person relations from within-person effects and which controlled for several third variables, showed that within-person effects were trivial.
 People who used social media more than usual to learn about COVID-19 did not show changes in their well-being.
-As a result, the results imply that COVID-19 related social media use is irrelevant for people's well-being.
-Other factors among the third variables that were measured revealed much larger effects, implying that the relevance of specific types of social media use for well-being is limited.
-Popular fears that ``doomscrolling'' or overusing social media during times of crises might not be justified.
+As a result, the results imply that COVID-19 related social media use does not seem to be particularly relevant for people's well-being.
+Other factors among the third variables that were measured revealed larger effects or relations, implying that well-being is rather determined by aspects such as health, employment, or locus of control.
+According to this study, popular fears that ``doomscrolling'' or overusing social media during times of crises do not seem to be justified.
 
-The study is not aligned with a recent cross-sectional study by Bendau et al. (2021) that showed negative relations between social media and well-being.
+The study is not aligned with a recent cross-sectional study by Bendau et al. (2021), which showed negative relations between social media and well-being.
 However, Bendau et al. (2021) analyzed cross-sectional data on a between-person level, while not controlling for third variables, which does not allow to make causal inferences.
@@
 Hence, to document how effects unfold it needs future research employing different study designs with different time lags.
+In addition, more thought needs to be invested in what relevant stable and nonvarying factor to control for.
 
-The predefined SESOI of \emph{b} = .30 was potentially too large.
-Media use is only one aspect of multiple factors that simultaneously affect well-being.
+Although I had already reduced the predefined SESOIs to be less conservative, potentially they were still too large.
+Media use is only one aspect of several factors that simultaneously affect well-being.
 Is it realistic that extremely changing only \emph{one} of these aspect (e.g., by completely stopping the use of social media) should already manifest in a detectable change in well-being?
-Or would it make more sense to say, if you regularly start doing \emph{two} activities (e.g.~regularly exercising and establishing a reading habit) together should then show in improvements to well-being?
-In other words, if the beneficial effect of a particular activity is large enough, if people implement \emph{two} of them then they should actually feel a difference.
+Or would it make more sense to expect that if people regularly start doing \emph{two} activities (e.g.~regularly exercising \emph{and} establishing a reading habit) together should show in perceivable improvements to well-being?
+In other words, if the beneficial effect of a particular activity is large enough, people should actually feel a difference if they implement two of those activities.
 Practically, this would imply a SESOI half as large as I have defined here, that is \emph{b} = \textbar.15\textbar{} for well-being and \emph{b} = \textbar.075\textbar.
 In this case, this would not make a difference, as even with these more liberal thresholds all but one effect would still be completely in the null region.
-However, at all events future research needs to start a discussion on what effect sizes are considered trivial, and this study is one of the first to provide some concrete guidelines.
+However, at all events future research needs to start a discussion on what effect sizes are considered meaningful and relevant, and with this study I hope to provide some first concrete guidelines.
 
@@
 
-Social media use was measured with an ordinal variable, however in the analyses it was treated as a numerical one.
-If treated as an ordinal one, it would have been necessary to analyze four different contrasts for each media measure, which plus the differentiation between between and within factor would have produced eight different measures, we would have made the model exceedingly complex.
-
 \hypertarget{conclusion}{%
@@
 
-\leavevmode\hypertarget{ref-beyensEffectSocialMedia2020}{}%
+\leavevmode\hypertarget{ref-beyensSocialMediaUseinpress}{}%
 Beyens, I., Pouwels, J.
@@
 M.
-(2020).
-The effect of social media on well-being differs from adolescent to adolescent.
-\emph{Scientific Reports}, \emph{10}(1), 10763.
-\url{https://doi.org/10.1038/s41598-020-67727-7}
+(n.d.).
+Social {Media} {Use} and {Adolescents}' {Well}-{Being}: {Developing} a {Typology} of {Person}-{Specific} {Effect} {Patterns}.
+\emph{Communication Research}.
+\url{https://doi.org/10.31234/osf.io/ftygp}
+
+\leavevmode\hypertarget{ref-choiMediatedCommunicationMatters2021}{}%
+Choi, M., \& Choung, H.
+(2021).
+Mediated communication matters during the {COVID}-19 pandemic: {The} use of interpersonal and masspersonal media and psychological well-being.
+\emph{Journal of Social and Personal Relationships}, \emph{38}(8), 2397--2418.
+\url{https://doi.org/10.1177/02654075211029378}
 
@@
 
+\leavevmode\hypertarget{ref-edenMediaCopingCOVID192020}{}%
+Eden, A.
+L., Johnson, B.
+K., Reinecke, L., \& Grady, S.
+M.
+(2020).
+Media for {Coping} {During} {COVID}-19 {Social} {Distancing}: {Stress}, {Anxiety}, and {Psychological} {Well}-{Being}.
+\emph{Frontiers in Psychology}, \emph{11}, 577639.
+\url{https://doi.org/10.3389/fpsyg.2020.577639}
+
 \leavevmode\hypertarget{ref-fergusonThisMetaanalysisScreen2021}{}%
@@
 \url{https://doi.org/10.1037/pro0000426}
+
+\leavevmode\hypertarget{ref-funderEvaluatingEffectSize2019}{}%
+Funder, D.
+C., \& Ozer, D.
+J.
+(2019).
+Evaluating effect size in psychological research: {Sense} and nonsense.
+\emph{Advances in Methods and Practices in Psychological Science}, \emph{2}(2), 156--168.
+\url{https://doi.org/10.1177/2515245919847202}
 
@@
 
+\leavevmode\hypertarget{ref-sandstromDoomscrollingCOVIDNews2021}{}%
+Sandstrom, G., Buchanan, K., Aknin, L., \& Lotun, S.
+(2021).
+Doomscrolling {COVID} news takes an emotional toll -- here's how to make your social media a happier place.
+Retrieved from \url{http://theconversation.com/doomscrolling-covid-news-takes-an-emotional-toll-heres-how-to-make-your-social-media-a-happier-place-170342}
+
 \leavevmode\hypertarget{ref-schemerImpactInternetSocial2021}{}%
@@
 
+\leavevmode\hypertarget{ref-stainbackCOVID1924News2020}{}%
+Stainback, K., Hearne, B.
+N., \& Trieu, M.
+M.
+(2020).
+{COVID}-19 and the 24/7 {News} {Cycle}: {Does} {COVID}-19 {News} {Exposure} {Affect} {Mental} {Health}?
+\emph{Socius}, \emph{6}, 2378023120969339.
+\url{https://doi.org/10.1177/2378023120969339}
+
 \leavevmode\hypertarget{ref-statistaAverageDailyTime2021}{}%
@@
 
+\leavevmode\hypertarget{ref-waterlooNormsOnlineExpressions2018}{}%
+Waterloo, S.
+F., Baumgartner, S.
+E., Peter, J., \& Valkenburg, P.
+M.
+(2018).
+Norms of online expressions of emotion: {Comparing} {Facebook}, {Twitter}, {Instagram}, and {WhatsApp}.
+\emph{New Media \& Society}, \emph{20}(5), 1813--1831.
+\url{https://doi.org/10.1177/1461444817707349}
+
 \end{CSLReferences}
+
+\newpage
 
@@
 
-The data are shared on AUSSDA, see \url{https://doi.org/10.11587/28KQNS}, and can be used for scientific purposes only.
+The data are shared on AUSSDA, see \url{https://doi.org/10.11587/28KQNS}.
+The data can only be used for scientific purposes.
 

### 2021-11-05 preregistration_changes.Rmd
- implemented feedback from Nick - updated analyses - provided additional plots for control variables

@@
 
-When reading the documentation of the data, I thought the following variables were varying, when in fact they were assessed only at one wave, namely the first wave:
+I originally planned to include the following variables as varying covariates.
+I thought they were measured at several/all waves.
+However, upon seeing the actual data I realized this wasn't the case.
+I hence included them as stable covariates.
 
-- Employment status
+This includes (a) text-based media news consumption, (b) video-based media news consumption, (c) residency is Vienna, (d) self-reported physical health, (e) living space (in squaremeter), (f) access to balcony, (g) access to garden, (h) employment status, (i) disposition to take risks, and (j) locus of control.
 
-## To-Do
+The following variables I originally planned to include as control variables.
+However, their inclusion led to the deletion of too many cases/observations.
+I hence decided against their inclusion.
+This includes (a) household size, (b) work hours per week, (c) home office, (d) household income.
 
-- Analyses
-	- Missingness
+## Analyses
+
+I planned to conduct also the following analyses. I the time of writing, I simply hadn't yet found the time to do so.
+
 	- Impute missing values
-		- Determine amount of missingness
-		- Replace values where I know they're 0
-	- Add factor scores
-	- CFA
-		- Add results
+	- Report CFAs for scales
+	- Run analyses using factor scores

### 2021-11-05 analyses.RmD
- status when submitting first draft to ICA - reading entire paper for the first time

@@
 
-Recode Variables
-
-```{r}
-d_wide %<>%
-  mutate_at(vars(everything(.)), funs(na_if(., 88))) %>%
-  mutate_at(vars(everything(.)), funs(na_if(., 99))) %>%
-  mutate(
-    male = 2 - .$gender,
-    age = 2021 - .$year_birth,
-    res_vienna = recode(.$county, `8` = 1L, .default = 0L,),
-    born_aus = 2 - .$born_aus,
-    home_owner = 2 - .$home_owner,
-    employment_fac = factor(.$employment, 
-                           # labels = c("Unselbststaendig erwerbstaetig (auch Lehrling), Privatwirtschaft",
-                           #            "Unselbststaendig erwerbstaetig (auch Lehrling), Oeffentlicher Dienst",
-                           #            "Selbststaendig erwerbstaetig",
-                           #            "Arbeitslos",
-                           #            "Pensionist*in",
-                           #            "Ausschliesslich haushaltsfuehrend",
-                           #            "Schueler/in, Student/in, Bildungskarenz",
-                           #            "Sonstiges (Dauerhaft arbeitsunfaehig oder Praesenz-/Zivildiener)",
-                           #            "In Elternkarenz")
-                           labels = c("Employed industrie",
-                                      "Employed public service",
-                                      "Self-employed",
-                                      "Unemployed",
-                                      "Retired",
-                                      "Housekeeping",
-                                      "Student",
-                                      "Incapacitated",
-                                      "Parental Leave")
-                           ),
-    edu_fac = factor(.$edu,
-                     # labels = c("Volksschule oder weniger", 
-                     #            "Hauptschule oder AHS Unterstufe",
-                     #            "Polytechnikum, BMS (Fachschule, z.B. HASCH)",
-                     #            "Lehre, Berufsschule",
-                     #            "AHS mit Matura",
-                     #            "BHS mit Matura (z.B. HTL, HAK, HBLA etc.)",
-                     #            "Hochschulverwandte Lehranstalt oder Kolleg",
-                     #            "Bachelor",
-                     #            "Magister / Master / Diplomingenieur / Fachhochschule",
-                     #            "Doktor / PhD"),
-                     labels = c("No degree", 
-                                "Middle school",
-                                "Vocational school",
-                                "Technical school",
-                                "High school",
-                                "Applied high school",
-                                "State college",
-                                "Bachelor",
-                                "Master",
-                                "PhD")
-                     ))
-```
-
 Make new documentation with selected variables.
@@
 
+Recode Variables
+
+```{r}
+d_wide %<>%
+  mutate_at(vars(everything(.)), funs(na_if(., 88))) %>%
+  mutate_at(vars(everything(.)), funs(na_if(., 99))) %>%
+  mutate(
+    male = 2 - .$gender,
+    age = 2021 - .$year_birth,
+    res_vienna = recode(.$county, `8` = 1L, .default = 0L,),
+    born_aus = 2 - .$born_aus,
+    home_owner = 2 - .$home_owner,
+    employment_fac = factor(.$employment, 
+                           # labels = c("Unselbststaendig erwerbstaetig (auch Lehrling), Privatwirtschaft",
+                           #            "Unselbststaendig erwerbstaetig (auch Lehrling), Oeffentlicher Dienst",
+                           #            "Selbststaendig erwerbstaetig",
+                           #            "Arbeitslos",
+                           #            "Pensionist*in",
+                           #            "Ausschliesslich haushaltsfuehrend",
+                           #            "Schueler/in, Student/in, Bildungskarenz",
+                           #            "Sonstiges (Dauerhaft arbeitsunfaehig oder Praesenz-/Zivildiener)",
+                           #            "In Elternkarenz")
+                           labels = c("Employed industrie",
+                                      "Employed public service",
+                                      "Self-employed",
+                                      "Unemployed",
+                                      "Retired",
+                                      "Housekeeping",
+                                      "Student",
+                                      "Incapacitated",
+                                      "Parental Leave")
+                           ),
+    edu_fac = factor(.$edu,
+                     # labels = c("Volksschule oder weniger", 
+                     #            "Hauptschule oder AHS Unterstufe",
+                     #            "Polytechnikum, BMS (Fachschule, z.B. HASCH)",
+                     #            "Lehre, Berufsschule",
+                     #            "AHS mit Matura",
+                     #            "BHS mit Matura (z.B. HTL, HAK, HBLA etc.)",
+                     #            "Hochschulverwandte Lehranstalt oder Kolleg",
+                     #            "Bachelor",
+                     #            "Magister / Master / Diplomingenieur / Fachhochschule",
+                     #            "Doktor / PhD"),
+                     labels = c("No degree", 
+                                "Middle school",
+                                "Vocational school",
+                                "Technical school",
+                                "High school",
+                                "Applied high school",
+                                "State college",
+                                "Bachelor",
+                                "Master",
+                                "PhD")
+                     ))
+```
+
 Make data from wide to long.
@@
 
-Filter waves without data
+<!-- Filter waves without data -->
 
-```{r eval=F}
-d_long_test <- d_long %>%
-  filter(!is.na(.[vars_varying]))
-```
+<!-- ```{r eval=F} -->
+<!-- d_long_test <- d_long %>% -->
+<!--   filter(!is.na(.[vars_varying])) -->
+<!-- ``` -->
 
@@
 
-Create vectors with names
+<!-- Create vectors with names -->
 
-```{r}
-# create vector of names
-vars_varying <- names(select(d_long, "health_b":"loc_cntrl_int_4_w"))
-predictors_varying <- select(d_long, "health_b":"loc_cntrl_int_4_w") %>% 
-  select(-starts_with("aff_"), -starts_with("life_")) %>% 
-  names()
-```
+<!-- ```{r} -->
+<!-- # create vector of names -->
+<!-- vars_varying <- names(select(d_long, "health_b":"loc_cntrl_int_4_w")) -->
+<!-- predictors_varying <- select(d_long, "health_b":"loc_cntrl_int_4_w") %>%  -->
+<!--   select(-starts_with("aff_"), -starts_with("life_")) %>%  -->
+<!--   names() -->
+<!-- ``` -->
 
-# inspect data
+# Inspect data
 
@@
 
-# measures
-## well-being
-### life satisfaction
+# Measures
+## Well-being
+### Life satisfaction
 
@@
 
-### positive affect
+### Positive affect
 
@@
 
-### negative affect
+### Negative affect
 
@@
 ## Social media use
-### reading
+### Reading
 
@@
 
-### liking & sharing
+### Liking & sharing
 
@@
 
-### posting
+### Posting
 
@@
 
-## figure
+## Figure
 
@@
 
-## table
+## Table
 
@@
     )
+kable(tab_desc_dat)
 ```
 
-# analyses
-## life satisfaction
+# Analyses
+## Life satisfaction
 
@@
 summary(model_life_sat_lmer)
-```
-
-```{r}
-# + health_b + act_wrk_b + act_spo_b + act_frn_b + act_sho_b + act_pet_b + sat_dem_b + hh_income_b + med_txt_kro_b + med_txt_sta_b + med_txt_pre_b + med_txt_oes_b + med_txt_kur_b + med_txt_slz_b + med_txt_son_b + med_vid_orf_b + med_vid_pri_b + risk_prop_b + loc_cntrl_int_1_b + loc_cntrl_int_2_b + loc_cntrl_int_3_b + loc_cntrl_int_4_b + health_w + act_wrk_w + act_spo_w + act_frn_w + act_sho_w + act_pet_w + sat_dem_w + hh_income_w + med_txt_kro_w + med_txt_sta_w + med_txt_pre_w + med_txt_oes_w + med_txt_kur_w + med_txt_slz_w + med_txt_son_w + med_vid_orf_w + med_vid_pri_w + risk_prop_w + loc_cntrl_int_1_w + loc_cntrl_int_2_w + loc_cntrl_int_3_w + loc_cntrl_int_4_w
 ```
@@
   mutate(
-    type = ifelse(grepl(".*_w.*", .$term), "within", "between"),
+    type = ifelse(grepl(".*_w\\>", .$term), "within", "between"),
     iv = gsub("_(w|b)\\>", "", .$term)
@@
 fig_results_activity <- make_graph_res(dat_fig_results_activity)
+fig_results_activity
 
@@
 fig_results_channel <- make_graph_res(dat_fig_results_channel)
+fig_results_channel
 
@@
 fig_results_control <- make_graph_res(dat_fig_results_control, sesoi = FALSE)
+fig_results_control
 

### 2021-11-05 manuscript.Rmd
- status when submitting first draft to ICA - reading entire paper for the first time

@@
 ---
-title             : "No effects of COVID-19 related social media use on well-being"
+title             : "No meaningful effects of COVID-19 related social media use on well-being"
 shorttitle        : "COVID-19 related social media use and well-being"
@@
 abstract          : "In times of crisis such as the Corona pandemic, it is important that citizens stay informed about recent events, the latest political decisions, or mandatory protection measures.
-To stay informed, many people use various types of media, and increasingly social media.
+To this end, many people use various types of media, and increasingly social media.
 However, because social media are particularly engaging, some find it hard to disconnect and cannot stop 'doomscrolling'.
 In this preregistered study, I investigate whether using social media for COVID-19 related topics might put personal well-being at risk.
-To this end, I analyze data from the Austrian Corona Panel Project, which consists of 24 waves with overall 3,018 participants.
-The data were analzed using random effects cross lagged panel models, controlling for several stable and varying covariates.
-Results showed that the effects of various types of COVID-19 related social media use on several types of well-being were very small, arguably too small to matter.
-The findings suggests fears that social media use during times of crisis might be detrimental for well-being are likely to be unfounded."
-keywords          : "COVID, well-being, affect, life satisfaction, social media use, news use, random effects within between model, panel study, longitudinal"
+To answer this question I analyzed data from the Austrian Corona Panel Project, which consists of 24 waves with overall 3,018 participants.
+I ran three random effects cross lagged panel models, controlling for several stable and varying covariates.
+Results showed that the effects of various types of COVID-19 related social media use on several indicators of well-being were very small, arguably too small to matter.
+The findings suggest fears that social media use during times of crisis might be detrimental for well-being are likely to be unfounded."
+keywords          : "COVID-19, Coronavirus, well-being, affect, life satisfaction, social media use, news use, communication, random effects within between model, panel study, longitudinal."
 
@@
 <!-- and especially in the beginning,  -->
-numerous events unfolded in quick succession, and several open questions emerged.
-How dangerous is the virus? Is it spreading in my region? How is it transmitted, and how can I protect myself?
-Because for many it was (and still is) a matter of life or death, citizens had to stay informed regarding the latest developments.
-Governments around the world implemented safety measures, ranging from wearing masks or keeping physical distance to complete lockdowns.
-In this extraordinary situation, people hence used media excessively to stay informed, and especially social media were at an all time high [@statistaAverageDailyTime2021].
+numerous events unfolded in quick succession.
+Several open questions emerged.
+How dangerous is the virus? 
+Is it spreading in my region? 
+How is it transmitted? 
+How can I protect myself?
+Because for many it was (and continues to be) a matter of life or death, citizens had to stay informed regarding the latest developments.
+Governments around the world implemented safety measures, ranging from wearing masks, keeping physical distance, to complete lockdowns.
+In this extraordinary situation, people used media excessively to attain information, and especially social media were at an all time high [@statistaAverageDailyTime2021].
 A new phenomenon termed "doomscrolling" emerged:
-People could not stop using social media to attain COVID-19 related news.
+People could not stop using social media to learn about COVID-19 related news.
 
-Several people reported that they were glued to their screens and could not pursue other relevant activities such as working, taking a break, or even care-work [@kleinDarklySoothingCompulsion2021].
-As doomscrolling increased it became doubtful whether such a surge in social media use could still be considered useful and adaptive, or whether it created an additional and new psychological danger for the users' mental health [@sandstromDoomscrollingCOVIDNews2021].
-A study with 6,233 people from Germany found that "[f]requency, duration and diversity of media exposure were positively associated with more symptoms of depression and unspecific and COVID-19 specific anxiety" [@bendauAssociationsCOVID19Related2021], implying negative consequences for well-being.
+Several people reported that they were glued to their screens and could not pursue other relevant activities such as working, taking a break, or even child-care [@kleinDarklySoothingCompulsion2021].
+As doomscrolling increased it became doubtful whether such a surge in social media use could still be considered useful and adaptive, or whether it created an additional and new psychological danger for users' mental health [@sandstromDoomscrollingCOVIDNews2021].
+These concerns seem justified: A study with 6,233 people from Germany found that "[f]requency, duration and diversity of media exposure were positively associated with more symptoms of depression and unspecific and COVID-19 specific anxiety" [@bendauAssociationsCOVID19Related2021], which implies negative consequences for well-being.
 
-As a result, with this study I want to build on this research and investigate the question whether COVID-19 related social media use during the pandemic affected the users' well-being.
-To this end I analyze a large-scale panel study from the Austrian Corona Panel Project [@kittelAustrianCoronaPanel2020]. 
-The panel consists of 24 waves and an overall sample size of `r nrow(d_wide)`, with at least 1,500 participants per wave, and it is representative of the Austrian population.
+As a result, with this study I want to build on this research and investigate the question whether COVID-19 related social media use during the pandemic affected well-being.
+To this end, I analyze a large-scale panel study from the Austrian Corona Panel Project [@kittelAustrianCoronaPanel2020]. 
+The panel consists of 24 waves and has an overall sample size of `r nrow(d_wide)`.
+At least 1,500 participants took part per wave, and the sample is representative of the Austrian population.
 The panel study collected a large number of variables.
-Being able to control for many confounding third variables, both stable and varying, together with the longitudinal design creates a unique opportunity to investigate causal effects.
+Being able to control for many confounding third variables, both stable and varying, and together with the longitudinal design, this creates a unique opportunity to investigate the causal effects of COVID-19 related social media use on well-being.
 
@@
 The underlying theories that guided the selection of variables for this study are the two-continua model of mental health [@greenspoonIntegrationSubjectiveWellbeing2001] and the hierarchical taxonomy of computer-mediated medation [@meierComputerMediatedCommunicationSocial2020].
-According to the two-continua model of mental health, mental health consists of two dimensions: psychopathology and well-being.
+According to the two-continua model, mental health consists of two dimensions: psychopathology and well-being.
 <!-- Because the aim of this study is better to understand typical users and everyday contexts, my focus will be on well-being. -->
-Well-being can be differentiated into subjective and psychological well-being [@dienerAdvancesOpenQuestions2018].
-Whereas subjective well-being emphasizes hedonic aspects such as a happiness and joy, psychological well-being focuses on eudaimonic aspects including fulfillment and meaning.
+Well-being can be further differentiated into subjective and psychological well-being [@dienerAdvancesOpenQuestions2018].
+Whereas subjective well-being emphasizes hedonic aspects such as happiness and joy, psychological well-being focuses on eudaimonic aspects like fulfillment and meaning.
 Subjective well-being is primarily about achieving positive affect and avoiding negative affect.
-On of the most prominent indicators of well-being is life satisfaction.
-In my view, life satisfaction is a meta concept that combines psychological and subjective well-being and represents a meta-appraisal of one's life.
+One of the most prominent indicators of well-being is life satisfaction.
+In my view, life satisfaction should be considered a meta concept that combines psychological and subjective well-being, because it represents a general appraisal of one's life.
 Notably, life satisfaction is stable and fluctuates only little, whereas it's the exact opposite for affect [@dienlinImpactDigitalTechnology2020].
 <!-- Everyone can feel bad on any given day, but whether or not one feels generally comfortable in one's life rather changes across years. -->
-To capture well-being in this study, I will thus build on life satisfaction, positive affect, and negative affect.
-Together, this should provide an encompassing perspective on media effects.
+To capture well-being in this study, I thus build on life satisfaction, positive affect, and negative affect.
+Together, this should provide an encompassing perspective on potential media effects.
 
-The hierarchical taxonomy of computer-mediated communication differentiates six levels of how people engage with digital technology:
-first, the device (e.g., smartphone); second, the type of application (e.g., social networking site); third, the branded application (e.g., Twitter); fourth, the feature (e.g., status post); fifth, the interaction (e.g., one-to-many); sixth, the message (e.g., content) [@meierComputerMediatedCommunicationSocial2020].
-Whereas the first four levels are focused on the channel, the last two on the type of communication.
-To measure social media use for consumption of specific news, I here employ both the channel and the communication perspective, which together provides a more nuanced understanding of communication.
-First, I will analyze how using the most prominent branded applications affect well-being, and whether this effect changes across applications.
+The hierarchical taxonomy of computer-mediated communication differentiates six levels of how people engage with digital technology.
+First, the device (e.g., smartphone); second, the type of application (e.g., social networking site); third, the branded application (e.g., Twitter); fourth, the feature (e.g., status post); fifth, the interaction (e.g., one-to-many); sixth, the message (e.g., content) [@meierComputerMediatedCommunicationSocial2020].
+Whereas the first four levels are focused on the _channel_, the last two address the _type_ of communication.
+To measure social media use for consumption of COVID-19 related news and topics, I here employ both the channel and the communication perspective, which together should provide a more nuanced understanding of communication.
+
+First, I will investigate how different types of communication affect well-being.
+Specifically, I will differentiate between active and passive use.
+I will distinguish (a) _reading_ COVID-19 related social media news (passive), _posting_ content regarding COVID (active), and _liking and sharing_ COVID-19 related posts (both active and passive).
+Second, I will analyze also how using the most prominent branded applications affect well-being, and whether this effect changes across applications.
 According to @meierComputerMediatedCommunicationSocial2020, branded apps are separate entities with different effects. 
@@
 The branded applications investigated here are Facebook, Twitter, Instagram, WhatsApp, and YouTube.
-But because adopting only this position would be both too narrow and too general, I will secondly also investigate how different types of interaction affect well-being.
-Specifically, I will differentiate between active and passive use.
-I will distinguish (a) _reading_ COVID-19 related social media use (passive), _posting_ content regarding COVID (active), and _liking and sharing_ COVID-19 related posts by others (both active and passive).
-Worth noting, this study is not about _general_ social media during times of COVID, but on social media use focused on COVID-19 related news and interactions. 
-Example of such media use include posting about the pandemic or retweeting COVID-19 related posts.
+Worth noting, this study is not about _general_ social media during times of COVID, but on social media use _focused_ on COVID-19 related news and interactions. 
+Examples of such media use include posting about the pandemic or retweeting COVID-19 related posts.
 
@@
 
+So far, there is only little empirical research on COVID-19 related social media use and well-being.
 In their study on the relations between media use and mental health during the pandemic, @bendauAssociationsCOVID19Related2021 found that people who used social media as a primary source of information reported on average "significantly more unspecific anxiety and depression [] and significantly more specific COVID-19 related anxiety symptoms" (p.
 288).
-Hence, this might hint at potential negative effects of social media as news use---but note that this finding comes from a between-person relation stemming from a cross-sectional data (see below).
-Hence, we don't know whether the differences in mental health and well-being are due to social media use or other third variables, such as age or education.
+Hence, this might hint at potential negative effects of social media as news use.
+However, note that this finding comes from a between-person relation stemming from cross-sectional data (see below).
+We, therefore, don't know whether the differences in mental health and well-being are due to social media use or other third variables, such as age, health, or education.
 @edenMediaCopingCOVID192020 analyzed the media use of 425 US college students during the first wave of the pandemic, and found both positive and negative relations with well-being.
-In a sample of 312 respondents collected via Amazon Mechanical Turk, it was found that people who used media to attain information were more lonely and less satisfied with their lives [@choiMediatedCommunicationMatters2021].
+In a sample of 312 respondents collected via Amazon Mechanical Turk, @choiMediatedCommunicationMatters2021 reported that people who used media to attain information were more lonely and less satisfied with their lives.
 @stainbackCOVID1924News2020 analyzed a large-scale study with 11,537 respondents from the US and found that increased COVID-19 media consumption was related to more psychological distress.
 
-In general, so far there is only little empirical research on news-related social media use and well-being.
-On the other hand, the question of whether and how _general_ social media use---or other, more specific types such as active or passive use---affect well-being is well-researched.
-A meta review (i.e., an analysis of meta-analyses), found that the relation between social media use and well-being is likely in the negative spectrum---but very small, potentially too small to matter [@meierComputerMediatedCommunicationSocial2020].
+On the other hand, the question of whether and how _general_ social media use affect well-being is well-researched.
+This also holds true for the different types of communication such as active or passive use.
+A meta review (i.e., an analysis of meta-analyses), found that the relation between social media use and well-being is likely in the negative spectrum but very small, potentially too small to matter [@meierComputerMediatedCommunicationSocial2020].
 What determines whether an effect is considered trivial or small?
 If we refer to standardized effect sizes, according to @cohenPowerPrimer1992 small effect sizes start at _r_ = .10.
-And indeed, several if not most meta-analyses find effect sizes below that threshold [@huangTimeSpentSocial2017; @meierComputerMediatedCommunicationSocial2020; @fergusonThisMetaanalysisScreen2021].
+And indeed, several if not most of the current meta-analyses find effect sizes below that threshold [@huangTimeSpentSocial2017; @meierComputerMediatedCommunicationSocial2020; @fergusonThisMetaanalysisScreen2021].
 
-These overviews are well aligned with several individual new studies that employed the most advanced methods [@orbenSocialMediaEnduring2019; @przybylskiDoesTakingShort2021a; @schemerImpactInternetSocial2021; @kerestesAdolescentsOnlineSocial2020].
-For example, @beyensSocialMediaUseinpress reported that although for some users (roughly one quarter) the effects were negative, for almost the same number of users the effects were positive, while for the majority they were neutral.
+These overviews are well aligned with several individual new studies that employed advanced methods [@orbenSocialMediaEnduring2019; @przybylskiDoesTakingShort2021a; @schemerImpactInternetSocial2021; @kerestesAdolescentsOnlineSocial2020].
+For example, @beyensSocialMediaUse2021 reported that although for some users (roughly one quarter) the effects were negative, for almost the same number of users they were positive, while for the majority the effects were neutral.
 In conclusion, most effects are somewhere between trivial and small.
-I think it's most convincing to expect trivial to small effects also in the case of COVID-19 related social media use.
+I therefore think it's most convincing to expect trivial to small effects also in the case of COVID-19 related social media use.
 
@@
 
-First, one could reasonably assume a _direct_ negative effect on well-being---especially on positive or negative affect, which is more volatile and fluctuating.
+First, one could assume a _direct_ negative effect on well-being, and especially on positive or negative affect, which is more volatile and fluctuating.
 Dangers, inequalities, corruption---these were the headlines across many countries worldwide.
@@
 <!-- So, just like being hit by a hammer hurts and we don't need any "mediating mechanism", this could be the case here as well. -->
-
-That said, because not all news were negative, because many people showed solidarity and compassion, there were also positive and potentially uplifting news.
+That said, because not all news was negative, and because many people showed solidarity and compassion, there was also positive and potentially uplifting news, potentially compensating for the negative news.
 <!-- However, in light of a worldwide pandemic with millions of deaths, the negative direct effect seems more plausible. -->
 
-There could be also indirect effects.
+There could also be _indirect_ effects.
 When doomscrolling, users are captivated to such an extent that they cannot stop using social media.
@@
 Overuse, however, is more critical, and several studies have shown more pronounced negative effects for extreme users [@przybylskiLargescaleTestGoldilocks2017].
-Overuse likely impairs well-being if it replaces other meaningful or functional activities such as meeting others, working, actively relaxing, or exercising.
+To explain, overuse likely impairs well-being if it replaces other meaningful or functional activities such as meeting others, working, actively relaxing, or exercising.
 So if a society collectively overuses social media, there is potential for negative effects.
@@
 
-On the other hand, one can make the case that overuse can be beneficial _in times of a pandemic_, even if it's mainly COVID-19 related.
-Exchanging COVID-19 related messages with friends via WhatsApp might replace the in-person contact one would have otherwise, but which is effectively impossible at that time.
-At times where meaningful and functional activities are explicitly or implicitly prohibited, using social media to exchange about COVID-19 might not be the worst idea.
+On the other hand, one can make the case that overuse can also be beneficial, especially in times of a pandemic, even if the use is mainly COVID-19 related.
+Exchanging COVID-19 related messages with friends via WhatsApp might replace the in-person contact one would have otherwise, but which is literally impossible at that time.
+In situations where meaningful and functional activities are explicitly or implicitly prohibited, using social media to exchange about COVID-19 related topics might not be the worst idea.
 In fact, given that nowadays a large number of experts, scientists, and politicians converse directly on social media, one can get first-hand information on the current developments.
@@
 
-Together, the strongest argument seems to be that _in general_ effects of social media on well-being are, on average, small at best.
+Together, the strongest argument seems to be that _in general_ the effects of social media on well-being are, on average, small at best.
 Because this study looks at only _one part_ of social media use---namely, COVID-19 related interactions---it is even more focused, and the overall potential of the effects should diminish even further.
@@
 <!-- At least, there does not seem to be a case where we should expect a clear deviation to the negative, despite anecdotes of doomscrolling. -->
-Therefore, I expect to find that the effects of COVID-19 related news use on well-being will be not be meaningful or practically relevant.
+Therefore, I expect to find that COVID-19 related communication on social media will not affect well-being in a meaningful or relevant way.
 
@@
 Testing this hypothesis, however, is not trivial.
-First, in contrast to most hypothesis typically posited in the social sciences, it implicitly contains an effect size, a so-called smallest effect size of interest (SESOI).
-Effectively testing this hypothesis hence necessitates to define what's considered a "trivial effect size" and what's not.
-Above I referred to standardized effect sizes.
+First, in contrast to most hypotheses typically posited in the social sciences, it implicitly contains an effect size, a so-called smallest effect size of interest (SESOI).
+Effectively testing this hypothesis necessitates defining what's considered a "trivial effect size" and what's not.
+Above I already referred to standardized effect sizes.
 However, standardized effect sizes should only be a first step toward evaluating an effect's relevance [@baguleyStandardizedSimpleEffect2009].
 Standardized effect sizes are determined by a sample's variance.[^cohensd]
-However, this is problematic: The question of whether or not social media use affects me personally in a relevant way should not depend on the variance in the sample in which my data were collected.
+However, this is problematic: The question of whether or not social media use affects a particular person in a relevant way should not depend on the variance in the sample in which that person's data were collected.
 Instead, it should depend on absolute criteria.
 What could be a minimally interesting, a nontrivial effect?
+Because this is a normative and ultimately philosophical question, there can never be a clear, single, or unanimous answer. 
+In the end, it is a personal question.
+However, it is still both necessary and fruitful to try to answer it. 
 I suggest the following SESOI for this research question:
@@
 
-If a heavy user of COVID-19 related social media news stops using social media altogether, this should have a noticeable impact on their overall well-being.
+If a heavy user of COVID-19 related social media news suddenly _completely stops_ using social media, this should have a _noticeable_ impact on their overall well-being.
 
@@
 In this study, COVID-19 related social media use was measured on a 5-point scale, ranging from 1 = _never_ to 5 = _several times a day_. 
-Thus, the example from above would imply that a change of four units in social media use should correspond to a noticeable change in well-being.
+Thus, the example from above would imply that a change of four units in social media use (e.g., a complete stop) should correspond to a noticeable change in well-being.
 But what's a noticeable change in well-being?
 According to @normanInterpretationChangesHealthrelated2003, people can reliably differentiate between _seven_ levels of satisfaction with health.
-So we could state that a four unit change in social media use should result in a one unit change in satisfaction.
-Statistically, in a regression, _b_ estimates the change in the dependent variable if the independent variable increases by one point.
-Transferred to our example, we would hence expect _b_ = 1 / 4 = .25.
+So we could state that a four unit change in social media use should result in a one unit change in life satisfaction, if satisfaction is measured on a 7-point scale.
+Statistically, in a regression _b_ estimates the change in the dependent variable if the independent variable increases by one point.
+Transferred to our example, we would hence expect a SESOI of _b_ = 1 / 4 = .25.
 
-In this study, life satisfaction was measured with 11 units, hence more than seven degrees people can reliably differentiate. 
-We hence need to transform a 1-point change on a 7-point scale to an equivalent change on an 11-point scale. 
-This is 11 / 7 = `r 11/7`.
-Hence, we now expect _b_ = 1/4 * 11/7 = `r 11/7*.25`.
+In this study, life satisfaction was measured on an 11-point scale, which is more than the seven degrees people can reliably differentiate. 
+We hence need to transform a 1-point change on a 7-point scale to its equivalent on an 11-point scale, which is 11 / 7 = `r 11/7`.
+Hence, we now expect our effect to be at least as large as _b_ = .25 * `r 11/7` = `r 11/7*.25`.
 For affect, which was measured on a 5-point scale, our SESOI would therefore be _b_ = 1/4 * 5/7 = `r 5 / 7 * .25`.
@@
 In order not to exaggerate precision, and to be a bit less conservative, these numbers will be reduced to nearby thresholds. 
-Note that other researchers also decrease or recommend decreasing thresholds for effect sizes when analyzing within-person or cumulative effects (@beyensSocialMediaUseinpress; @funderEvaluatingEffectSize2019).
+Note that other researchers also decreased or recommended decreasing thresholds for effect sizes when analyzing within-person or cumulative effects (@beyensSocialMediaUse2021; @funderEvaluatingEffectSize2019).
 Together, this leads to an indifference region ranging from _b_ = -.30 to _b_ = .30 for life satisfaction, and _b_ = -.15 to _b_ = .15 for positive and negative affect.
@@
 The hypothesis explicitly states a causal effect.
-In non-experimental designs, longitudinal designs help investigate causality.
+In non-experimental studies, longitudinal designs help investigate causality.
 Using longitudinal designs alone, however, is not sufficient for correct causal statements.
@@
 
-For example, imagine that a person suddenly starts using social media much more than usual, and then at the same time also become less satisfied with their lives.
+Consider the following example.
+Imagine that a person suddenly starts using social media much more than usual, and then at the same time also becomes less satisfied with their life.
 After some time, use recedes again, whereas life satisfaction returns to prior levels.
-If this happened to several people at the same time, in a longitudinal study we could then find a causal effect of social media use on life satisfaction.
-However, it could also be the case that during the study there was a major exogenous event (say, a pandemic) and a large part of the working population lost their jobs.
-Hence, the causal effect reported above was confounded, because in reality it was the pandemic that caused both social media use to rise and their life satisfaction to plummet.
+If this happens to several people at the same time, in a longitudinal study we would then observe a significant effect of social media use on life satisfaction.
+However, it could also be the case that during the study there was a major exogenous event (say, a pandemic), which is why a large part of the working population lost their jobs.
+Hence, the causal effect reported above was likely confounded, because in reality it was the pandemic that caused both social media use to rise and life satisfaction to plummet.
 
-Thus, only if we can control for _all_ potential varying third variables, we can correctly estimate causality without any bias [@rohrerThinkingClearlyCorrelations2018].
-Obviously, we can never be entirely sure to have included all varying covariates, which makes absolute statements regarding causality impossible.
+Thus, only if we can control for _all_ potential varying third variables, can we correctly estimate causality without bias [@rohrerThinkingClearlyCorrelations2018].
+Obviously, we can never be entirely sure that we have included all varying covariates, which makes absolute statements regarding causality impossible.
 In addition, when determining the overall causal effect, we should _not_ control for mediating variables [@rohrerThinkingClearlyCorrelations2018].
 Doing so would bias and distort our assessment of the role of social media use.
+Complicating matters, often it is unclear if a variable is a mediator or a confounder.[^collider]
 However, when controlling for relevant variables (that aren't mediators), we can be much more certain that we measured causality correctly.
@@
 
-When looking for suitable control candidates, ideally we find variables that affect both media use and well-being. 
-Controlling for these factors will isolate the actual effect of social media use on well-being.
+[^collider]: In addition, there also exist colliders, which I don't discuss here and which complicate the issue even further [@rohrerThinkingClearlyCorrelations2018].
+
+When looking for suitable control candidates, ideally we select variables that affect both media use and well-being. 
+Controlling for these factors isolates the actual effect of social media use on well-being.
 We can also control for variables that affect _only_ well-being. 
-However, in doing so not much is gained or lost because the effects of social media use would remain virtually the same [@klinePrinciplesPracticeStructural2016].
+However, in doing so not much is gained or lost because the effects of social media use would remain virtually the same [@klinePrinciplesPracticeStructural2016].[^mcelreath]
 
-In this study, I hence plan to control for the following variables, which either have been shown or a likely to affect both social media use and well-being, while not being mediators.
+[^mcelreath]: But see @mcelreathYesterdayClass2021.
+
+In this study, I hence plan to control for the following variables, which either have already been shown to affect both social media use and well-being, or which are very likely to do so, and which are likely not mediators:
 <!-- (I'll additionally include variables that likely affect only well-being, also to obtain a comparison benchmark for social media effects): -->
-Gender, age, education, Austria country of birth, Austria country of birth of parents, text-based news consumption, video-based news consumption, residency Vienna, household size, health, living space, access to garden, access to balcony, employment, work hours per week, being in home-office, household income, outdoor activities, satisfaction with democracy, disposition to take risks, and locus of control.
-I will not control for variables such as trust in institutions or media, because these variables are likely influenced by social media use.
+gender, age, education, Austria country of birth, Austria country of birth of parents, text-based news consumption, video-based news consumption, residency Vienna, household size, health, living space, access to garden, access to balcony, employment, work hours per week, being in home-office, household income, outdoor activities, satisfaction with democracy, disposition to take risks, and locus of control.
+I will not control for variables such as trust in institutions or media, because these variables might be influenced by social media use to a meaningful extent.
 
 Next to including covariates, it is now increasingly understood that causal effects need to be analyzed from an internal, within-person perspective.
-If a specific person changes their media diet, we need to measure how this affects _them_.
-Between person comparison from cross-sectional data, where participants are interviewed only once, cannot provide such insights.
+If a specific person changes their media diet, we need to measure how this affects _their_ well-being.
+Between-person comparison from cross-sectional data, where participants are interviewed only once, cannot provide such insights.
 In this study, I will hence differentiate between within-person effects and between-person relations. 
-Because it is explicitly causal, in answering the hypothesis I will thus consider only the within-person effects.
+As explicated in the hypothesis, to test the hypothesis I will thus consider only the within-person effects.
 
-One precondition of causality is temporal order. 
+Finally, one precondition of causality is temporal order. 
 The cause needs to precede the outcome.
-To this end, finding the right interval when causes and effects are measured is crucial.
+Finding the right interval between causes and effects is crucial.
 For example, if we want to measure the effect of alcohol consumption on driving performance, it makes a big difference if driving performance is measured one minute, one hour, one day, or one week after consumption.
@@
 In the case of well-being, to analyze affect we need shorter intervals, while for life satisfaction longer ones.
-Still, finding the right interval is challenging, because especially short intervals are practically hard to implement and require experience sampling measures (also known as in situ measurement or ambulant assessment).
+Still, choosing the right interval is challenging, because especially short intervals are practically hard to implement and often require experience sampling measures (also known as in situ measurement or ambulant assessment).
 
@@
 When additionally controlling for both stable and varying covariates, we can then be more sure that the effect is indeed causal.
-This approach was implemented already by several other studies and is considered one of the best practices to analyze causality.
+Similar approaches were implemented already by several other studies [@johannesNoEffectDifferent2021; @scharkowHowSocialNetwork2020] and they considered one of the best practices to analyze causality [@bellFixedRandomEffects2019].
 
@@
 The (anonymous) preregistration can be accessed here: https://osf.io/87b24/?view_only=b2289b6fec214fa88ee75a18d45c18f3.
-Because in this study I analyze data from an already existing large-scale data set, the Austrian Corona Panel Project, all of these steps were preregistered prior to accessing the data.
+Because in this study I analyze data from an already existing large-scale data set, all of these steps were preregistered prior to accessing the data.
 The preregistration was designed on the basis of the panel documentation online [@kittelAustrianCoronaPanel2020].
-In some cases the analyses could not be executed as originally planned, because some properties of the variables only became apparent upon data analysis.
-The most relevant deviations are reported below, and a complete list of all changes can be found online.
+In some cases the analyses could not be executed as originally planned, for example because some properties of the variables became apparent only upon data analysis.
+The most relevant deviations are reported below, and a complete list of all changes can be found in the online companion website (https://xmtra.github.io/Austrian_Corona_Panel_Project/index.html).
 
@@
 The study was conducted between March 2020 and October 2021.
-The contains 26 waves, and at the time of writing, the first 24 waves were available for download. 
+It contains 26 waves, and at the time of writing, the first 24 waves were available for download. 
 Each wave consists of at least 1,500 respondents.
-The overall sample size was _N_ = `r d_raw %>% nrow()`, and `r d_long %>% nrow()` observations were collected.
-Panel mortality was compensated through a continuous reacquistion of new participants. 
+The overall sample size was _N_ = `r d_raw %>% nrow() %>% prettyNum(big.mark = ",")`, and `r d_long %>% nrow() %>% prettyNum(big.mark = ",")` observations were collected.
+Panel mortality was compensated through a continuous acquisition of new participants. 
 All respondents needed to have access to the internet (via computer or mobile devices such as smartphones or tablets).
@@
 In order to participate in the study, the respondents needed to be Austrian residents and had to be at least 14 years old.
-The average age was `r mean(2021 - d_wide$year_birth, na.rm = T) %>% round(0)` years, `r mean(d_wide$male, na.rm = T) %>% round(., 2) * 100` percent were male, `r d_wide %>% filter(.$edu_fac %in% c("Hochschulverwandte Lehranstalt oder Kolleg","Bachelor", "Magister / Master / Diplomingenieur / Fachhochschule", "Doktor / PhD")) %>% nrow() / sum(table(d_wide$edu_fac)) %>% divide_by(100)` percent had a University degree, and `r table(d_wide$employment_fac)["Arbeitslos"] / (table(d_wide$employment_fac) %>% sum()) %>% unname() %>% divide_by(100) %>% round(., 0)` percent were currently unemployed.
+The average age was `r mean(2021 - d_wide$year_birth, na.rm = T) %>% round(0)` years, `r mean(d_wide$male, na.rm = T) %>% round(., 2) * 100` percent were male, `r (d_wide %>% filter(.$edu_fac %in% c("State college", "Bachelor", "Master", "PhD"))) %>% nrow() %>% divide_by(sum(table(d_wide$edu_fac))) %>% multiply_by(100) %>% round(0)` percent had a University degree, and `r (table(d_wide$employment_fac)["Unemployed"] / sum(table(d_wide$employment_fac))) %>% unname() %>% multiply_by(100) %>% round(., digits = 0)` percent were currently unemployed.
 <!-- The study received IRB approval from the University of Vienna. -->
 
-Because the data were analyzed post hoc, no sample size planning on the basis of a priori power analyses was conducted.
-The sample is very large, and it is hence well-equipped reliably to detect also small effects, which is why no exact post hoc power analysis were conducted.
-In addition, because such large samples easily generate significant _p_-values even for very small effects, this study builds on a smallest effect size of interest.
+Because the data were analyzed post-hoc, no a-priori sample size planning on the basis of power analyses was conducted.
+The sample is very large, and it is hence likely well-equipped reliably to detect also small effects.
+<!-- , which is why no exact post hoc power analysis were conducted. -->
+In addition, because such large samples easily generate significant _p_-values even for very small effects, it helps that to test hypothesis this study uses a smallest effect size of interest-approach.
 To test the hypotheses, I will use the interval testing approach as proposed by @dienesUsingBayesGet2014.
@@
 For well-being, the null region will be between _b_ = -.30 and _b_ = .30.
-For example, if the 95% confidence interval lies completely within the null-region (e.g., _b_ = .15, [95% CI: .05, .25]), the hypothesis that the effect is only trivial is be supported.
+For example, if the 95% confidence interval lies completely within the null-region (e.g., _b_ = .15, [95% CI: .05, .25]), the hypothesis that the effect is only trivial is supported.
 If the effects interval and the null region overlap (e.g., _b_ = .25, [95% CI: .15, .35]), the hypothesis is not supported and the results are considered inconclusive, whereas a meaningful negative effect is rejected.
-If the confidence falls completely outside of the null-region (e.g., _b_ = .4, [95% CI: .35, .45]), the hypothesis is rejected and the existence of a meaningful positive effect is supported.
+If the confidence falls completely outside of the null-region (e.g., _b_ = .40, [95% CI: .35, .45]), the hypothesis is rejected and the existence of a meaningful positive effect is supported.
 
@@
 <!-- X clear cases were removed. -->
-
-Respondents who answered less than 50% of all questions were removed.
-The remaining missing responses were imputed using predictive mean matching.
+<!-- Respondents who answered less than 50% of all questions were removed. -->
+<!-- The remaining missing responses were imputed using predictive mean matching. -->
 <!-- We will remove respondents with unrealistically fast responses, namely below three standard deviations of the medium response time.  -->
@@
 Three models were run, one for each dependent variable.
-All predictors---that is, social media activities and social media channels, within and between-person factors, stably and varying covariates---were included in each of the three models. 
-The data were hierarchical, and responses were nested in participants and waves (put differently, participants and waves were implemented as random effects).
+Predictors included all types and channels of social media communication separated into within and between-person factors, as well as stable and varying covariates.
+All predictors were included simultaneously and in each of the three models. 
+The data were hierarchical, and responses were separately nested in participants and waves (put differently, participants and waves were implemented as random effects).
 Nesting in participants allowed to separate between-person relations from within-person effects.
-Nesting in waves allowed to control for general developments, such as general decrease in well-being in the population, for example due to lockdown measures (hence, there was no need additionally to control for specific phases or measures of the lockdown).
+Nesting in waves allowed to control for general exogenous developments, such as general decrease in well-being in the population, for example due to lockdown measures.
+Thus, there was no need additionally to control for specific phases or measures of the lockdown.
 
@@
 
-For more information on the analyses, see [companion website](https://xmtra.github.io/Austrian_Corona_Panel_Project/index.html).
+For more information on the analyses, for example a complete documentation of the models and their results, see [companion website](https://xmtra.github.io/Austrian_Corona_Panel_Project/index.html).
 
@@
 
-The COVID-19 related social media use focused on interaction was measured with three variables (a) reading, (b) liking and sharing, and (c) posting.
+COVID-19 related social media use focused on types of communication was measured with three variables (a) reading, (b) liking and sharing, and (c) posting.
 The general introductory question was "How often during the last week have you engaged in the following activities on social media?"
 The three items read as follows: 
-"Reading the posts of others with content on the Coronavirus"; "When seeing Posts on the Coronavirus, I clicked 'like', 'share' or 'retweet'"; "I myself wrote posts on the Coronavirus on Social Media."
+"Reading the posts of others with content on the Coronavirus"; "When seeing posts on the Coronavirus, I clicked 'like', 'share' or 'retweet'"; "I myself wrote posts on the Coronavirus on social media."
 Answer options were 1 (_several times per day_), 2 (_daily_), 3 (_several times per week_), 4 (_weekly_), 5 (_never_).
@@
 
-The COVID-19 related social media use focused on channels was measured with five variables.
-The general introductory question was "How often in the last week have you followed information related to the Corona-crisis in the following social media?"
-The five items were (a) Facebook, (b) Twitter, (c) Instagram, (d) Youtube, (e) WhatsApp.
+COVID-19 related social media use focused on channels was measured with five variables.
+The general introductory question was "How often in the last week have you followed information related to the Corona-crisis on the following social media?"
+The five items were (a) Facebook, (b) Twitter, (c) Instagram, (d) Youtube, and (e) WhatsApp.
 Again, the answer options were 1 (_several times per day_), 2 (_daily_), 3 (_several times per week_), 4 (_weekly_), 5 (_never_).
@@
 
-Social media use was measured for everyone on waves 1, 2, 8, 17, and 23. 
+Social media use was measured for all participants on waves 1, 2, 8, 17, and 23. 
 Freshly recruited respondents always answered the questions on social media use.
@@
 The effects of COVID-19 related social media use were controlled for the following stable variables: 
-(a) gender (answer options: female, male, diverse), (b) age, (c) education (10 options), (d) Austria country of birth (yes/no), (e) Austria parents' country of birth (no parent, one parent, both parents).
+(a) gender (answer options: female, male, diverse), (b) age, (c) education (ten options), (d) Austria country of birth (yes/no), (e) Austria parents' country of birth (no parent, one parent, both parents).
 I originally planned to implement other variables as varying covariates.
-However, because they were not measured often enough or at the time when social media use was measured, I implemented them as stable variables using their average values across all waves.
-This includes (a) text-based media news consumption, (b) video-based media news consumption, (c) residency is Vienna, (d) self-reported physical health, (e) living space (in squaremeter), (f) access to balcony, (g) access to garden, (h) employment status, (i) disposition to take risks, and (j) locus of control.
-I controlled also for the following varying covariates: (a) outdoor activities, (b) satisfaction with democracy.
+However, because they were not measured often enough or not at the time when social media use was measured, I implemented them as stable variables using their average values across all waves.
+This includes (a) text-based media news consumption (five degrees), (b) video-based media news consumption (five degrees), (c) residency is Vienna (yes/no), (d) self-reported physical health (five degrees), (e) living space (in squaremeter), (f) access to balcony (yes/no), (g) access to garden (yes/no), (h) employment (nine options), (i) disposition to take risks (eleven degrees), and (j) locus of control (five degrees).
+I controlled also for the following varying covariates: (a) five items measuring outdoor activities such as sport or meeting friends (five degrees), and (b) satisfaction with democracy (five degrees).
 Because it lead to too much attrition in the sample, I did not control for (a) household size, (b) work hours per week, (c) home office, (d) household income.
@@
 
-When looking at the variables from a descriptive perspective, we see that all well-being measures did not change substantially across the different phases of the pandemic.
-COVID-19 related media use, however, decreased during the beginning of the study and remained stable after approximately six waves.
-The initial decrease might be explained by the fact that the collection of data began in end of March 2020, hence approximately three months after the pandemic began.
-It could be that after an initial stark surge, COVID-19 related social media use was already declining, returning to more normal levels.
+When looking at the variables from a descriptive perspective, we see that all well-being measures did not change substantially across the different waves of data collection.
+COVID-19 related media use, however, decreased slightly at the beginning of the study and remained stable after approximately six waves.
+The initial decrease might be explained by the fact that the collection of data began at the end of March 2020, hence approximately three months after the pandemic began.
+It could be that after an initial uptick, COVID-19 related social media use was already declining at that time, returning to more normal levels.
 
 ```{r fig-desc, fig.cap="Development of well-being and media use measures across the pandemic.
-Values obtained from mixed effect models, with participants and waves as grouping factor and without additional predictors.", out.width = "\\textwidth", cache=F}
+Values obtained from mixed effect models, with participants and waves as grouping factors and without additional predictors.", out.width = "\\textwidth", cache=F}
 knitr::include_graphics("figures/fig_descriptives.pdf")
@@
 
-The study's hypothesis was that the effects of all types of social media use on well-being will be trivial.
-Regarding the different types of _communication_---that is, reading vs.
+The study's hypothesis was that the effects of social media use on well-being would be trivial.
+Regarding the effects of different _types_ of communication---that is, reading vs.
 sharing vs.
@@
 For example, respondents who used social media more frequently than usual to read about COVID-19 related topics did not show a simultaneous change in life satisfaction (_b_ = `r dat_fig_results_activity %>% filter(dv == "Life satisfaction" & iv == "Reading" & type == "within") %>% select(estimate) %>% round(2)` [95% CI `r dat_fig_results_activity %>% filter(dv == "Life satisfaction" & iv == "Reading" & type == "within") %>% select(conf.low) %>% round(2)`, `r dat_fig_results_activity %>% filter(dv == "Life satisfaction" & iv == "Reading" & type == "within") %>% select(conf.high) %>% round(2)`]).
-All confidence intervals included zero; hence, all effects were non-significant.
+All confidence intervals include zero; hence, all effects were also statistically non-significant.
 <!-- Only one effect did not include zero. -->
@@
 <!-- However, because the effect did not fall outside of our null region, it's likely not large enough to be considered meaningful. -->
-As a result, the hypothesis was supported for all types of COVID-19 related social media communication.
+As a result, the hypothesis was supported for all COVID-19 related types of social media communication.
 
@@
 Effects are controlled for a large number of covariates (see text).
-The SESOI was 0.30 for life satisfaction and .15 for affects; hence, no effect is considered meaningful theoretically.", out.width = "\\textwidth", cache=F}
+The SESOI was |0.30| for life satisfaction and |0.15| for affects.
+Hence, none of the reported effects are considered meaningful.", out.width = "\\textwidth", cache=F}
 knitr::include_graphics("figures/fig_results_activity.pdf")
@@
 <!-- Hence, although positive the effect was considered too small to matter practically. -->
-Respondents who across all waves used social media more frequently than others to write posts on COVID-19 reported higher levels of negative affect than others (_b_ = `r dat_fig_results_activity %>% filter(dv == "Negative affect" & iv == "Posting" & type == "between") %>% select(estimate) %>% round(2)` [95% CI `r dat_fig_results_activity %>% filter(dv == "Negative affect" & iv == "Posting" & type == "between") %>% select(conf.low) %>% round(2)`, `r dat_fig_results_activity %>% filter(dv == "Negative affect" & iv == "Posting" & type == "between") %>% select(conf.high) %>% round(2)`]).
-The effect was still completely inside of the null region, hence likely not large enough to be practically relevant.
-
-Regarding the COVID-19 related use of social media _channels_, the results were very comparable.
-Changes in the frequency of using different social media channels to attain information regarding COVID-19 were unrelated to meaningful changes in well-being (see Figure \@ref(fig:res-channels)).
-For example, respondents who used Facebook more frequently than usual to learn about COVID-19 did not show a simultaneous change in well-being (_b_ = `r dat_fig_results_channel %>% filter(dv == "Life satisfaction" & iv == "Facebook" & type == "within") %>% select(estimate) %>% round(2)` [95% CI `r dat_fig_results_channel %>% filter(dv == "Life satisfaction" & iv == "Facebook" & type == "within") %>% select(conf.low) %>% round(2)`, `r dat_fig_results_channel %>% filter(dv == "Life satisfaction" & iv == "Facebook" & type == "within") %>% select(conf.high) %>% round(2)`]).
-Only two effects differed substantially from zero.
-Respondents who used Instagram more frequently than usual to attain COVID-19 related news reported slightly higher simultaneous levels of life satisfaction then usual (_b_ = `r dat_fig_results_channel %>% filter(dv == "Life satisfaction" & iv == "Instagram" & type == "within") %>% select(estimate) %>% round(2)` [95% CI `r dat_fig_results_channel %>% filter(dv == "Life satisfaction" & iv == "Instagram" & type == "within") %>% select(conf.low) %>% round(2)`, `r dat_fig_results_channel %>% filter(dv == "Life satisfaction" & iv == "Instagram" & type == "within") %>% select(conf.high) %>% my_round("b")`]).
-Respondents who used Twitter more frequently than usual to attain COVID-19 related news reported slightly lower simultaneous levels of life satisfaction then usual (_b_ = `r dat_fig_results_channel %>% filter(dv == "Life satisfaction" & iv == "Twitter" & type == "within") %>% select(estimate) %>% round(2)` [95% CI `r dat_fig_results_channel %>% filter(dv == "Life satisfaction" & iv == "Twitter" & type == "within") %>% select(conf.low) %>% round(2)`, `r dat_fig_results_channel %>% filter(dv == "Life satisfaction" & iv == "Twitter" & type == "within") %>% select(conf.high) %>% my_round("b")`]).
-However, both effects were still completely inside of the null region, hence not large enough to be considered meaningful.
-In sum, the hypothesis was supported for the COVID-19 related use of all types of social media channels.
+Respondents who across all waves used social media more frequently than others to write COVID-19 related posts reported higher levels of negative affect than others (_b_ = `r dat_fig_results_activity %>% filter(dv == "Negative affect" & iv == "Posting" & type == "between") %>% select(estimate) %>% round(2)` [95% CI `r dat_fig_results_activity %>% filter(dv == "Negative affect" & iv == "Posting" & type == "between") %>% select(conf.low) %>% round(2)`, `r dat_fig_results_activity %>% filter(dv == "Negative affect" & iv == "Posting" & type == "between") %>% select(conf.high) %>% round(2)`]).
+The effect was still completely inside of the null region, hence not considered large enough to be practically relevant.
 
@@
 Effects are controlled for a large number of covariates (see text).
-The SESOI was 0.30 for life satisfaction and .15 for affects; hence, no effect is considered meaningful theoretically.", out.width = "\\textwidth", cache=F}
+The SESOI was |0.30| for life satisfaction and |0.15| for affects.
+Hence, none of the reported effects are considered meaningful.", out.width = "\\textwidth", cache=F}
 knitr::include_graphics("figures/fig_results_channel.pdf")
@@
 
-In terms of between-person relations---which, again, weren't included in the hypotheses---no relations crossed or fell outside of the null region.
+Regarding the COVID-19 related use of social media _channels_, the results were very comparable (see Figure \@ref(fig:res-channels)).
+Changes in the frequency of using different social media channels to attain information regarding COVID-19 were unrelated to meaningful changes in well-being.
+For example, respondents who used Facebook more frequently than usual to learn about COVID-19 did not show a simultaneous change in well-being (_b_ = `r dat_fig_results_channel %>% filter(dv == "Life satisfaction" & iv == "Facebook" & type == "within") %>% select(estimate) %>% round(2)` [95% CI `r dat_fig_results_channel %>% filter(dv == "Life satisfaction" & iv == "Facebook" & type == "within") %>% select(conf.low) %>% round(2)`, `r dat_fig_results_channel %>% filter(dv == "Life satisfaction" & iv == "Facebook" & type == "within") %>% select(conf.high) %>% round(2)`]).
+Only two effects differed substantially from zero.
+Respondents who used Instagram more frequently than usual to attain COVID-19 related news reported slightly higher simultaneous levels of life satisfaction than usual (_b_ = `r dat_fig_results_channel %>% filter(dv == "Life satisfaction" & iv == "Instagram" & type == "within") %>% select(estimate) %>% my_round("b")` [95% CI `r dat_fig_results_channel %>% filter(dv == "Life satisfaction" & iv == "Instagram" & type == "within") %>% select(conf.low) %>% my_round("b")`, `r dat_fig_results_channel %>% filter(dv == "Life satisfaction" & iv == "Instagram" & type == "within") %>% select(conf.high) %>% my_round("b")`]).
+Respondents who used Twitter more frequently than usual to attain COVID-19 related news reported somewhat lower simultaneous levels of life satisfaction than usual (_b_ = `r dat_fig_results_channel %>% filter(dv == "Life satisfaction" & iv == "Twitter" & type == "within") %>% select(estimate) %>% round(2)` [95% CI `r dat_fig_results_channel %>% filter(dv == "Life satisfaction" & iv == "Twitter" & type == "within") %>% select(conf.low) %>% round(2)`, `r dat_fig_results_channel %>% filter(dv == "Life satisfaction" & iv == "Twitter" & type == "within") %>% select(conf.high) %>% my_round("b")`]).
+However, both effects were still completely inside of the null region, hence not considered large enough to be meaningful.
+In sum, the hypothesis was supported also for the COVID-19 related use of important social media channels.
+
+In terms of between-person relations---which, again, weren't included in the hypotheses---no relations crossed the null region or fell outside of it.
 Only one relation did not include zero, was hence statistically significant. 
 Respondents who across all waves used YouTube more frequently than others for COVID-19 related reasons reported marginally higher levels of negative affect (_b_ = `r dat_fig_results_channel %>% filter(dv == "Negative affect" & iv == "YouTube" & type == "between") %>% select(estimate) %>% round(2)` [95% CI `r dat_fig_results_channel %>% filter(dv == "Negative affect" & iv == "YouTube" & type == "between") %>% select(conf.low) %>% my_round("b_txt")`, `r dat_fig_results_channel %>% filter(dv == "Negative affect" & iv == "YouTube" & type == "between") %>% select(conf.high) %>% my_round("b")`]).
-However, please note that this effect was not large enough to be considered practically relevant.
+However, please note that this effect again was not considered large enough to be practically relevant.
 
@@
 In what follows, I briefly report also the results of some covariates. 
-Several variables showed large associations with well-being. 
-Because each variable has a different scaling, we would again need to define a SESOI for each variable, which cannot be implemented here. 
-But note that several variables would fall outside of such a SESOI.
+Several variables showed larger associations with well-being. 
+Note because each variable has a different scaling, we would again need to define a SESOI for each variable, which cannot be implemented here. 
+But note that several variables would likely fall outside of such a SESOI.
 This includes for example internal locus of control, health, or employment. 
@@
 In this study I analyzed the effects of COVID-19 related social media use on well-being.
-The data come from a panel study with 24 waves that is representative of the Austrian population.
-A random effect model, which separated between person relations from within-person effects and which controlled for several third variables, showed that within-person effects were trivial.
+The data come from a panel study with 24 waves and is representative of the Austrian population.
+A random effects model, which separated between person relations from within-person effects and which controlled for several third variables, showed that within-person effects were trivial.
 People who used social media more than usual to learn about COVID-19 did not show changes in their well-being.
-As a result, the results imply that COVID-19 related social media use does not seem to be particularly relevant for people's well-being.
+
+The results imply that COVID-19 related social media use does not seem to be particularly relevant for people's well-being.
 Other factors among the third variables that were measured revealed larger effects or relations, implying that well-being is rather determined by aspects such as health, employment, or locus of control.
@@
 
-The study is not aligned with a recent cross-sectional study by @bendauAssociationsCOVID19Related2021, which showed negative relations between social media and well-being.
-However, @bendauAssociationsCOVID19Related2021 analyzed cross-sectional data on a between-person level, while not controlling for third variables, which does not allow to make causal inferences.
-At the same time, this study is well-aligned with recent studies and meta-analyses from related research questions, which found that the effects of various types of social media use on several well-being indicators is small at best, often too small to matter [@meierComputerMediatedCommunicationSocial2020; @orbenTeenagersScreensSocial2020; @fergusonThisMetaanalysisScreen2021].
+The study is not aligned with several recent studies on this particular research question or closely related research questions.
+This includes a study by @bendauAssociationsCOVID19Related2021, which showed negative relations between social media and well-being.
+However, note that @bendauAssociationsCOVID19Related2021 analyzed cross-sectional data on a between-person level, while not controlling for third variables, which is not optimal for investigating causal effects.
+At the same time, this study is well-aligned with recent studies and meta-analyses analyzing the effects of social media use from a more general perspective or that adopted a somewhat different angle.
+These studies have found that the effects of various types of social media use on several well-being indicators are small at best, often too small to matter [@meierComputerMediatedCommunicationSocial2020; @orbenTeenagersScreensSocial2020; @fergusonThisMetaanalysisScreen2021], which echoes the results obtained here.
 
@@
 
-The current study analyzed whether changes in media use were related with changes in well-being, while controlling for several potential confounds. 
+The current study analyzed whether changes in media use were related to changes in well-being, while controlling for several potential confounds. 
 Together, this allows for a good perspective on potential causality.
-That said, causality necessitates temporal order, and the cause needs to precede the consequence.
+That said, causality necessitates temporal order, and the cause needs to precede the effect.
 Regarding media use, such effects often happen immediately or shortly after use, necessitating intervals in the hours, minutes, or even seconds.
-Only experience sampling studies that ask users in the very moment can produce such knowledge.
+In many cases only experience sampling studies that ask users in the very moment can produce such knowledge.
 However, even then we don't know for certain if we actually measured the right interval.
-Hence, to document how effects unfold it needs future research employing different study designs with different time lags.
-In addition, more thought needs to be invested in what relevant stable and nonvarying factor to control for.
+Hence, to document how effects unfold, future research needs to employ different study designs probing different time lags.
+In addition, more thought needs to be invested in what relevant stable and varying factors should be selected as control variables.
 
@@
 Media use is only one aspect of several factors that simultaneously affect well-being.
-Is it realistic that extremely changing only _one_ of these aspect (e.g., by completely stopping the use of social media) should already manifest in a detectable change in well-being?
-Or would it make more sense to expect that if people regularly start doing _two_ activities (e.g.
-regularly exercising _and_ establishing a reading habit) together should show in perceivable improvements to well-being?
-In other words, if the beneficial effect of a particular activity is large enough, people should actually feel a difference if they implement two of those activities. 
+Is it realistic that extremely changing only _one_ of these aspects (e.g., by completely stopping the use of social media) should already manifest in a detectable change in well-being?
+Or would it make more sense to expect that if people regularly start doing say _two_ activities (e.g.
+regularly exercising _and_ establishing a reading habit) together should show perceivable improvements to well-being?
+In other words, the beneficial effect of a particular activity is large enough if people feel a difference when implementing two of those activities. 
 Practically, this would imply a SESOI half as large as I have defined here, that is _b_ = |.15| for well-being and _b_ = |.075|. 
 In this case, this would not make a difference, as even with these more liberal thresholds all but one effect would still be completely in the null region.
-However, at all events future research needs to start a discussion on what effect sizes are considered meaningful and relevant, and with this study I hope to provide some first concrete guidelines.
+However, at all events future research needs to start a thorough discussion on what effect sizes are considered meaningful and relevant, and with this study I hope to provide some helpful input and first concrete guidelines.
 
@@
 That said, objective measures often cannot capture the content or the motivation of the use, and only very complicated tools that record the content that was used (such as the Screenome project) might produce such data.
-However, also these procedure introduce other problems, for example related to privacy.
+Unfortunately, such procedures introduce other problems, for example related to privacy.
 Hence, for this type of research question it seems necessary still to use self-reported measures.
 
-The generalizability of the results are not large, because the data were collected in a single country.
-The results are hence potentially limited to the more Western sphere, and might not apply to other cultures, especially if they have a different media landscape or offer alternative social media.
-That said, because this is a large study, representative of a country's entire population, and because several waves were collected across a large time span, the results should be at least as generalizable as other typical empirical studies collected in the social sciences.
+Because the data were collected in a single country, the generalizability of the results is limited.
+The results potentially apply primarily to the more Western sphere, and might not hold true in other cultures, especially cultures with a different media landscape or alternative social media channels.
+That said, because this is a comparatively large study representative of a country's entire population, and because several waves were collected across a large time span, the results should be at least as generalizable as other typical empirical studies collected in the social sciences.
 
@@
 In this study, COVID-19 related social media use did not causally affect several indicators of well-being, including life satisfaction, positive affect, and negative affect.
-However, factors other than social media use were meaningfully related to well-being, such as physical health, employment, or believing that one is in control of one's life.
-If it's the aim to improve well-being, it might hence be fruitful not to focus on social media but to address other, potentially more pressing societal problems related to inequality or mental health.
+However, factors other than social media use were more meaningfully related to well-being, such as physical health, employment, or believing that one is in control of one's life.
+If it's the aim to improve well-being, it might hence be fruitful not to focus on social media but to address other, potentially more pressing societal problems related to inequality, safe spaces, or health care.
 

### 2021-11-05 manuscript.tex
- status when submitting first draft to ICA - reading entire paper for the first time

@@
 \hypersetup{
-  pdftitle={No effects of COVID-19 related social media use on well-being},
+  pdftitle={No meaningful effects of COVID-19 related social media use on well-being},
   pdflang={en-EN},
-  pdfkeywords={COVID, well-being, affect, life satisfaction, social media use, news use, random effects within between model, panel study, longitudinal},
+  pdfkeywords={COVID-19, Coronavirus, well-being, affect, life satisfaction, social media use, news use, communication, random effects within between model, panel study, longitudinal.},
   hidelinks,
@@
 \makeatother
-\keywords{COVID, well-being, affect, life satisfaction, social media use, news use, random effects within between model, panel study, longitudinal}
+\keywords{COVID-19, Coronavirus, well-being, affect, life satisfaction, social media use, news use, communication, random effects within between model, panel study, longitudinal.}
 \usepackage{lineno}
@@
 
-\title{No effects of COVID-19 related social media use on well-being}
+\title{No meaningful effects of COVID-19 related social media use on well-being}
 \author{Blinded for review\textsuperscript{1}}
@@
 In times of crisis such as the Corona pandemic, it is important that citizens stay informed about recent events, the latest political decisions, or mandatory protection measures.
-To stay informed, many people use various types of media, and increasingly social media.
+To this end, many people use various types of media, and increasingly social media.
 However, because social media are particularly engaging, some find it hard to disconnect and cannot stop `doomscrolling.' In this preregistered study, I investigate whether using social media for COVID-19 related topics might put personal well-being at risk.
-To this end, I analyze data from the Austrian Corona Panel Project, which consists of 24 waves with overall 3,018 participants.
-The data were analzed using random effects cross lagged panel models, controlling for several stable and varying covariates.
-Results showed that the effects of various types of COVID-19 related social media use on several types of well-being were very small, arguably too small to matter.
-The findings suggests fears that social media use during times of crisis might be detrimental for well-being are likely to be unfounded.
+To answer this question I analyzed data from the Austrian Corona Panel Project, which consists of 24 waves with overall 3,018 participants.
+I ran three random effects cross lagged panel models, controlling for several stable and varying covariates.
+Results showed that the effects of various types of COVID-19 related social media use on several indicators of well-being were very small, arguably too small to matter.
+The findings suggest fears that social media use during times of crisis might be detrimental for well-being are likely to be unfounded.
 }
@@
 During the COVID-19 pandemic,
-numerous events unfolded in quick succession, and several open questions emerged.
-How dangerous is the virus? Is it spreading in my region? How is it transmitted, and how can I protect myself?
-Because for many it was (and still is) a matter of life or death, citizens had to stay informed regarding the latest developments.
-Governments around the world implemented safety measures, ranging from wearing masks or keeping physical distance to complete lockdowns.
-In this extraordinary situation, people hence used media excessively to stay informed, and especially social media were at an all time high (Statista, 2021).
+numerous events unfolded in quick succession.
+Several open questions emerged.
+How dangerous is the virus?
+Is it spreading in my region?
+How is it transmitted?
+How can I protect myself?
+Because for many it was (and continues to be) a matter of life or death, citizens had to stay informed regarding the latest developments.
+Governments around the world implemented safety measures, ranging from wearing masks, keeping physical distance, to complete lockdowns.
+In this extraordinary situation, people used media excessively to attain information, and especially social media were at an all time high (Statista, 2021).
 A new phenomenon termed ``doomscrolling'' emerged:
-People could not stop using social media to attain COVID-19 related news.
+People could not stop using social media to learn about COVID-19 related news.
 
-Several people reported that they were glued to their screens and could not pursue other relevant activities such as working, taking a break, or even care-work (Klein, 2021).
-As doomscrolling increased it became doubtful whether such a surge in social media use could still be considered useful and adaptive, or whether it created an additional and new psychological danger for the users' mental health (Sandstrom, Buchanan, Aknin, \& Lotun, 2021).
-A study with 6,233 people from Germany found that ``{[}f{]}requency, duration and diversity of media exposure were positively associated with more symptoms of depression and unspecific and COVID-19 specific anxiety'' (Bendau et al., 2021), implying negative consequences for well-being.
+Several people reported that they were glued to their screens and could not pursue other relevant activities such as working, taking a break, or even child-care (Klein, 2021).
+As doomscrolling increased it became doubtful whether such a surge in social media use could still be considered useful and adaptive, or whether it created an additional and new psychological danger for users' mental health (Sandstrom, Buchanan, Aknin, \& Lotun, 2021).
+These concerns seem justified: A study with 6,233 people from Germany found that ``{[}f{]}requency, duration and diversity of media exposure were positively associated with more symptoms of depression and unspecific and COVID-19 specific anxiety'' (Bendau et al., 2021), which implies negative consequences for well-being.
 
-As a result, with this study I want to build on this research and investigate the question whether COVID-19 related social media use during the pandemic affected the users' well-being.
-To this end I analyze a large-scale panel study from the Austrian Corona Panel Project (Kittel et al., 2020).
-The panel consists of 24 waves and an overall sample size of 3018, with at least 1,500 participants per wave, and it is representative of the Austrian population.
+As a result, with this study I want to build on this research and investigate the question whether COVID-19 related social media use during the pandemic affected well-being.
+To this end, I analyze a large-scale panel study from the Austrian Corona Panel Project (Kittel et al., 2020).
+The panel consists of 24 waves and has an overall sample size of 3018.
+At least 1,500 participants took part per wave, and the sample is representative of the Austrian population.
 The panel study collected a large number of variables.
-Being able to control for many confounding third variables, both stable and varying, together with the longitudinal design creates a unique opportunity to investigate causal effects.
+Being able to control for many confounding third variables, both stable and varying, and together with the longitudinal design, this creates a unique opportunity to investigate the causal effects of COVID-19 related social media use on well-being.
 
@@
 The underlying theories that guided the selection of variables for this study are the two-continua model of mental health (Greenspoon \& Saklofske, 2001) and the hierarchical taxonomy of computer-mediated medation (Meier \& Reinecke, 2020).
-According to the two-continua model of mental health, mental health consists of two dimensions: psychopathology and well-being.
-Well-being can be differentiated into subjective and psychological well-being (Diener, Lucas, \& Oishi, 2018).
-Whereas subjective well-being emphasizes hedonic aspects such as a happiness and joy, psychological well-being focuses on eudaimonic aspects including fulfillment and meaning.
+According to the two-continua model, mental health consists of two dimensions: psychopathology and well-being.
+Well-being can be further differentiated into subjective and psychological well-being (Diener, Lucas, \& Oishi, 2018).
+Whereas subjective well-being emphasizes hedonic aspects such as happiness and joy, psychological well-being focuses on eudaimonic aspects like fulfillment and meaning.
 Subjective well-being is primarily about achieving positive affect and avoiding negative affect.
-On of the most prominent indicators of well-being is life satisfaction.
-In my view, life satisfaction is a meta concept that combines psychological and subjective well-being and represents a meta-appraisal of one's life.
+One of the most prominent indicators of well-being is life satisfaction.
+In my view, life satisfaction should be considered a meta concept that combines psychological and subjective well-being, because it represents a general appraisal of one's life.
 Notably, life satisfaction is stable and fluctuates only little, whereas it's the exact opposite for affect (Dienlin \& Johannes, 2020).
-To capture well-being in this study, I will thus build on life satisfaction, positive affect, and negative affect.
-Together, this should provide an encompassing perspective on media effects.
+To capture well-being in this study, I thus build on life satisfaction, positive affect, and negative affect.
+Together, this should provide an encompassing perspective on potential media effects.
 
-The hierarchical taxonomy of computer-mediated communication differentiates six levels of how people engage with digital technology:
-first, the device (e.g., smartphone); second, the type of application (e.g., social networking site); third, the branded application (e.g., Twitter); fourth, the feature (e.g., status post); fifth, the interaction (e.g., one-to-many); sixth, the message (e.g., content) (Meier \& Reinecke, 2020).
-Whereas the first four levels are focused on the channel, the last two on the type of communication.
-To measure social media use for consumption of specific news, I here employ both the channel and the communication perspective, which together provides a more nuanced understanding of communication.
-First, I will analyze how using the most prominent branded applications affect well-being, and whether this effect changes across applications.
+The hierarchical taxonomy of computer-mediated communication differentiates six levels of how people engage with digital technology.
+First, the device (e.g., smartphone); second, the type of application (e.g., social networking site); third, the branded application (e.g., Twitter); fourth, the feature (e.g., status post); fifth, the interaction (e.g., one-to-many); sixth, the message (e.g., content) (Meier \& Reinecke, 2020).
+Whereas the first four levels are focused on the \emph{channel}, the last two address the \emph{type} of communication.
+To measure social media use for consumption of COVID-19 related news and topics, I here employ both the channel and the communication perspective, which together should provide a more nuanced understanding of communication.
+
+First, I will investigate how different types of communication affect well-being.
+Specifically, I will differentiate between active and passive use.
+I will distinguish (a) \emph{reading} COVID-19 related social media news (passive), \emph{posting} content regarding COVID (active), and \emph{liking and sharing} COVID-19 related posts (both active and passive).
+Second, I will analyze also how using the most prominent branded applications affect well-being, and whether this effect changes across applications.
 According to Meier and Reinecke (2020), branded apps are separate entities with different effects.
@@
 The branded applications investigated here are Facebook, Twitter, Instagram, WhatsApp, and YouTube.
-But because adopting only this position would be both too narrow and too general, I will secondly also investigate how different types of interaction affect well-being.
-Specifically, I will differentiate between active and passive use.
-I will distinguish (a) \emph{reading} COVID-19 related social media use (passive), \emph{posting} content regarding COVID (active), and \emph{liking and sharing} COVID-19 related posts by others (both active and passive).
-Worth noting, this study is not about \emph{general} social media during times of COVID, but on social media use focused on COVID-19 related news and interactions.
-Example of such media use include posting about the pandemic or retweeting COVID-19 related posts.
+Worth noting, this study is not about \emph{general} social media during times of COVID, but on social media use \emph{focused} on COVID-19 related news and interactions.
+Examples of such media use include posting about the pandemic or retweeting COVID-19 related posts.
 
@@
 
+So far, there is only little empirical research on COVID-19 related social media use and well-being.
 In their study on the relations between media use and mental health during the pandemic, Bendau et al.
 (2021) found that people who used social media as a primary source of information reported on average ``significantly more unspecific anxiety and depression {[}{]} and significantly more specific COVID-19 related anxiety symptoms'' (p.~288).
-Hence, this might hint at potential negative effects of social media as news use---but note that this finding comes from a between-person relation stemming from a cross-sectional data (see below).
-Hence, we don't know whether the differences in mental health and well-being are due to social media use or other third variables, such as age or education.
+Hence, this might hint at potential negative effects of social media as news use.
+However, note that this finding comes from a between-person relation stemming from cross-sectional data (see below).
+We, therefore, don't know whether the differences in mental health and well-being are due to social media use or other third variables, such as age, health, or education.
 Eden, Johnson, Reinecke, and Grady (2020) analyzed the media use of 425 US college students during the first wave of the pandemic, and found both positive and negative relations with well-being.
-In a sample of 312 respondents collected via Amazon Mechanical Turk, it was found that people who used media to attain information were more lonely and less satisfied with their lives (Choi \& Choung, 2021).
+In a sample of 312 respondents collected via Amazon Mechanical Turk, Choi and Choung (2021) reported that people who used media to attain information were more lonely and less satisfied with their lives.
 Stainback, Hearne, and Trieu (2020) analyzed a large-scale study with 11,537 respondents from the US and found that increased COVID-19 media consumption was related to more psychological distress.
 
-In general, so far there is only little empirical research on news-related social media use and well-being.
-On the other hand, the question of whether and how \emph{general} social media use---or other, more specific types such as active or passive use---affect well-being is well-researched.
-A meta review (i.e., an analysis of meta-analyses), found that the relation between social media use and well-being is likely in the negative spectrum---but very small, potentially too small to matter (Meier \& Reinecke, 2020).
+On the other hand, the question of whether and how \emph{general} social media use affect well-being is well-researched.
+This also holds true for the different types of communication such as active or passive use.
+A meta review (i.e., an analysis of meta-analyses), found that the relation between social media use and well-being is likely in the negative spectrum but very small, potentially too small to matter (Meier \& Reinecke, 2020).
 What determines whether an effect is considered trivial or small?
 If we refer to standardized effect sizes, according to Cohen (1992) small effect sizes start at \emph{r} = .10.
-And indeed, several if not most meta-analyses find effect sizes below that threshold (Ferguson et al., 2021; Huang, 2017; Meier \& Reinecke, 2020).
+And indeed, several if not most of the current meta-analyses find effect sizes below that threshold (Ferguson et al., 2021; Huang, 2017; Meier \& Reinecke, 2020).
 
-These overviews are well aligned with several individual new studies that employed the most advanced methods (Keresteš \& Štulhofer, 2020; Orben, Dienlin, \& Przybylski, 2019; Przybylski, Nguyen, Law, \& Weinstein, 2021; Schemer, Masur, Geiß, Müller, \& Schäfer, 2021).
-For example, Beyens, Pouwels, Driel, Keijsers, and Valkenburg (n.d.) reported that although for some users (roughly one quarter) the effects were negative, for almost the same number of users the effects were positive, while for the majority they were neutral.
+These overviews are well aligned with several individual new studies that employed advanced methods (Keresteš \& Štulhofer, 2020; Orben, Dienlin, \& Przybylski, 2019; Przybylski, Nguyen, Law, \& Weinstein, 2021; Schemer, Masur, Geiß, Müller, \& Schäfer, 2021).
+For example, Beyens, Pouwels, Driel, Keijsers, and Valkenburg (2021) reported that although for some users (roughly one quarter) the effects were negative, for almost the same number of users they were positive, while for the majority the effects were neutral.
 In conclusion, most effects are somewhere between trivial and small.
-I think it's most convincing to expect trivial to small effects also in the case of COVID-19 related social media use.
+I therefore think it's most convincing to expect trivial to small effects also in the case of COVID-19 related social media use.
 
@@
 
-First, one could reasonably assume a \emph{direct} negative effect on well-being---especially on positive or negative affect, which is more volatile and fluctuating.
+First, one could assume a \emph{direct} negative effect on well-being, and especially on positive or negative affect, which is more volatile and fluctuating.
 Dangers, inequalities, corruption---these were the headlines across many countries worldwide.
@@
 Repeatedly consuming such news might be depressing, perhaps even changing some general perspectives on life, without any further mediating processes.
+That said, because not all news was negative, and because many people showed solidarity and compassion, there was also positive and potentially uplifting news, potentially compensating for the negative news.
 
-That said, because not all news were negative, because many people showed solidarity and compassion, there were also positive and potentially uplifting news.
-
-There could be also indirect effects.
+There could also be \emph{indirect} effects.
 When doomscrolling, users are captivated to such an extent that they cannot stop using social media.
@@
 Overuse, however, is more critical, and several studies have shown more pronounced negative effects for extreme users (Przybylski \& Weinstein, 2017).
-Overuse likely impairs well-being if it replaces other meaningful or functional activities such as meeting others, working, actively relaxing, or exercising.
+To explain, overuse likely impairs well-being if it replaces other meaningful or functional activities such as meeting others, working, actively relaxing, or exercising.
 So if a society collectively overuses social media, there is potential for negative effects.
 
-On the other hand, one can make the case that overuse can be beneficial \emph{in times of a pandemic}, even if it's mainly COVID-19 related.
-Exchanging COVID-19 related messages with friends via WhatsApp might replace the in-person contact one would have otherwise, but which is effectively impossible at that time.
-At times where meaningful and functional activities are explicitly or implicitly prohibited, using social media to exchange about COVID-19 might not be the worst idea.
+On the other hand, one can make the case that overuse can also be beneficial, especially in times of a pandemic, even if the use is mainly COVID-19 related.
+Exchanging COVID-19 related messages with friends via WhatsApp might replace the in-person contact one would have otherwise, but which is literally impossible at that time.
+In situations where meaningful and functional activities are explicitly or implicitly prohibited, using social media to exchange about COVID-19 related topics might not be the worst idea.
 In fact, given that nowadays a large number of experts, scientists, and politicians converse directly on social media, one can get first-hand information on the current developments.
 
-Together, the strongest argument seems to be that \emph{in general} effects of social media on well-being are, on average, small at best.
+Together, the strongest argument seems to be that \emph{in general} the effects of social media on well-being are, on average, small at best.
 Because this study looks at only \emph{one part} of social media use---namely, COVID-19 related interactions---it is even more focused, and the overall potential of the effects should diminish even further.
 Whether or not using social media for COVID-19 related aspects is detrimental during a pandemic is also not entirely clear.
-Therefore, I expect to find that the effects of COVID-19 related news use on well-being will be not be meaningful or practically relevant.
+Therefore, I expect to find that COVID-19 related communication on social media will not affect well-being in a meaningful or relevant way.
 
@@
 Testing this hypothesis, however, is not trivial.
-First, in contrast to most hypothesis typically posited in the social sciences, it implicitly contains an effect size, a so-called smallest effect size of interest (SESOI).
-Effectively testing this hypothesis hence necessitates to define what's considered a ``trivial effect size'' and what's not.
-Above I referred to standardized effect sizes.
+First, in contrast to most hypotheses typically posited in the social sciences, it implicitly contains an effect size, a so-called smallest effect size of interest (SESOI).
+Effectively testing this hypothesis necessitates defining what's considered a ``trivial effect size'' and what's not.
+Above I already referred to standardized effect sizes.
 However, standardized effect sizes should only be a first step toward evaluating an effect's relevance (Baguley, 2009).
@@
 Hence, if there is more deviation/variance in a sample, the effect size decreases, even if the difference of the group's means stays the same.}
-However, this is problematic: The question of whether or not social media use affects me personally in a relevant way should not depend on the variance in the sample in which my data were collected.
+However, this is problematic: The question of whether or not social media use affects a particular person in a relevant way should not depend on the variance in the sample in which that person's data were collected.
 Instead, it should depend on absolute criteria.
 What could be a minimally interesting, a nontrivial effect?
+Because this is a normative and ultimately philosophical question, there can never be a clear, single, or unanimous answer.
+In the end, it is a personal question.
+However, it is still both necessary and fruitful to try to answer it.
 I suggest the following SESOI for this research question:
 
-If a heavy user of COVID-19 related social media news stops using social media altogether, this should have a noticeable impact on their overall well-being.
+If a heavy user of COVID-19 related social media news suddenly \emph{completely stops} using social media, this should have a \emph{noticeable} impact on their overall well-being.
 
@@
 In this study, COVID-19 related social media use was measured on a 5-point scale, ranging from 1 = \emph{never} to 5 = \emph{several times a day}.
-Thus, the example from above would imply that a change of four units in social media use should correspond to a noticeable change in well-being.
+Thus, the example from above would imply that a change of four units in social media use (e.g., a complete stop) should correspond to a noticeable change in well-being.
 But what's a noticeable change in well-being?
 According to Norman, Sloan, and Wyrwich (2003), people can reliably differentiate between \emph{seven} levels of satisfaction with health.
-So we could state that a four unit change in social media use should result in a one unit change in satisfaction.
-Statistically, in a regression, \emph{b} estimates the change in the dependent variable if the independent variable increases by one point.
-Transferred to our example, we would hence expect \emph{b} = 1 / 4 = .25.
+So we could state that a four unit change in social media use should result in a one unit change in life satisfaction, if satisfaction is measured on a 7-point scale.
+Statistically, in a regression \emph{b} estimates the change in the dependent variable if the independent variable increases by one point.
+Transferred to our example, we would hence expect a SESOI of \emph{b} = 1 / 4 = .25.
 
-In this study, life satisfaction was measured with 11 units, hence more than seven degrees people can reliably differentiate.
-We hence need to transform a 1-point change on a 7-point scale to an equivalent change on an 11-point scale.
-This is 11 / 7 = 1.57.
-Hence, we now expect \emph{b} = 1/4 * 11/7 = 0.39.
+In this study, life satisfaction was measured on an 11-point scale, which is more than the seven degrees people can reliably differentiate.
+We hence need to transform a 1-point change on a 7-point scale to its equivalent on an 11-point scale, which is 11 / 7 = 1.57.
+Hence, we now expect our effect to be at least as large as \emph{b} = .25 * 1.57 = 0.39.
 For affect, which was measured on a 5-point scale, our SESOI would therefore be \emph{b} = 1/4 * 5/7 = 0.18.
@@
 In order not to exaggerate precision, and to be a bit less conservative, these numbers will be reduced to nearby thresholds.
-Note that other researchers also decrease or recommend decreasing thresholds for effect sizes when analyzing within-person or cumulative effects (Beyens, Pouwels, Driel, Keijsers, and Valkenburg (n.d.); Funder and Ozer (2019)).
+Note that other researchers also decreased or recommended decreasing thresholds for effect sizes when analyzing within-person or cumulative effects (Beyens, Pouwels, Driel, Keijsers, and Valkenburg (2021); Funder and Ozer (2019)).
 Together, this leads to an indifference region ranging from \emph{b} = -.30 to \emph{b} = .30 for life satisfaction, and \emph{b} = -.15 to \emph{b} = .15 for positive and negative affect.
@@
 
-For example, imagine that a person suddenly starts using social media much more than usual, and then at the same time also become less satisfied with their lives.
+Consider the following example.
+Imagine that a person suddenly starts using social media much more than usual, and then at the same time also becomes less satisfied with their life.
 After some time, use recedes again, whereas life satisfaction returns to prior levels.
-If this happened to several people at the same time, in a longitudinal study we could then find a causal effect of social media use on life satisfaction.
-However, it could also be the case that during the study there was a major exogenous event (say, a pandemic) and a large part of the working population lost their jobs.
-Hence, the causal effect reported above was confounded, because in reality it was the pandemic that caused both social media use to rise and their life satisfaction to plummet.
+If this happens to several people at the same time, in a longitudinal study we would then observe a significant effect of social media use on life satisfaction.
+However, it could also be the case that during the study there was a major exogenous event (say, a pandemic), which is why a large part of the working population lost their jobs.
+Hence, the causal effect reported above was likely confounded, because in reality it was the pandemic that caused both social media use to rise and life satisfaction to plummet.
 
-Thus, only if we can control for \emph{all} potential varying third variables, we can correctly estimate causality without any bias (Rohrer, 2018).
-Obviously, we can never be entirely sure to have included all varying covariates, which makes absolute statements regarding causality impossible.
+Thus, only if we can control for \emph{all} potential varying third variables, can we correctly estimate causality without bias (Rohrer, 2018).
+Obviously, we can never be entirely sure that we have included all varying covariates, which makes absolute statements regarding causality impossible.
 In addition, when determining the overall causal effect, we should \emph{not} control for mediating variables (Rohrer, 2018).
 Doing so would bias and distort our assessment of the role of social media use.
+Complicating matters, often it is unclear if a variable is a mediator or a confounder.\footnote{In addition, there also exist colliders, which I don't discuss here and which complicate the issue even further (Rohrer, 2018).}
 However, when controlling for relevant variables (that aren't mediators), we can be much more certain that we measured causality correctly.
@@
 
-When looking for suitable control candidates, ideally we find variables that affect both media use and well-being.
-Controlling for these factors will isolate the actual effect of social media use on well-being.
+When looking for suitable control candidates, ideally we select variables that affect both media use and well-being.
+Controlling for these factors isolates the actual effect of social media use on well-being.
 We can also control for variables that affect \emph{only} well-being.
-However, in doing so not much is gained or lost because the effects of social media use would remain virtually the same (Kline, 2016).
+However, in doing so not much is gained or lost because the effects of social media use would remain virtually the same (Kline, 2016).\footnote{But see McElreath (2021).}
 
-In this study, I hence plan to control for the following variables, which either have been shown or a likely to affect both social media use and well-being, while not being mediators.
-Gender, age, education, Austria country of birth, Austria country of birth of parents, text-based news consumption, video-based news consumption, residency Vienna, household size, health, living space, access to garden, access to balcony, employment, work hours per week, being in home-office, household income, outdoor activities, satisfaction with democracy, disposition to take risks, and locus of control.
-I will not control for variables such as trust in institutions or media, because these variables are likely influenced by social media use.
+In this study, I hence plan to control for the following variables, which either have already been shown to affect both social media use and well-being, or which are very likely to do so, and which are likely not mediators:
+gender, age, education, Austria country of birth, Austria country of birth of parents, text-based news consumption, video-based news consumption, residency Vienna, household size, health, living space, access to garden, access to balcony, employment, work hours per week, being in home-office, household income, outdoor activities, satisfaction with democracy, disposition to take risks, and locus of control.
+I will not control for variables such as trust in institutions or media, because these variables might be influenced by social media use to a meaningful extent.
 
 Next to including covariates, it is now increasingly understood that causal effects need to be analyzed from an internal, within-person perspective.
-If a specific person changes their media diet, we need to measure how this affects \emph{them}.
-Between person comparison from cross-sectional data, where participants are interviewed only once, cannot provide such insights.
+If a specific person changes their media diet, we need to measure how this affects \emph{their} well-being.
+Between-person comparison from cross-sectional data, where participants are interviewed only once, cannot provide such insights.
 In this study, I will hence differentiate between within-person effects and between-person relations.
-Because it is explicitly causal, in answering the hypothesis I will thus consider only the within-person effects.
+As explicated in the hypothesis, to test the hypothesis I will thus consider only the within-person effects.
 
-One precondition of causality is temporal order.
+Finally, one precondition of causality is temporal order.
 The cause needs to precede the outcome.
-To this end, finding the right interval when causes and effects are measured is crucial.
+Finding the right interval between causes and effects is crucial.
 For example, if we want to measure the effect of alcohol consumption on driving performance, it makes a big difference if driving performance is measured one minute, one hour, one day, or one week after consumption.
@@
 When additionally controlling for both stable and varying covariates, we can then be more sure that the effect is indeed causal.
-This approach was implemented already by several other studies and is considered one of the best practices to analyze causality.
+Similar approaches were implemented already by several other studies (Johannes, Dienlin, Bakhshi, \& Przybylski, 2021; Scharkow, Mangold, Stier, \& Breuer, 2020) and they considered one of the best practices to analyze causality (Bell, Fairbrother, \& Jones, 2019).
 
@@
 The (anonymous) preregistration can be accessed here: \url{https://osf.io/87b24/?view_only=b2289b6fec214fa88ee75a18d45c18f3}.
-Because in this study I analyze data from an already existing large-scale data set, the Austrian Corona Panel Project, all of these steps were preregistered prior to accessing the data.
+Because in this study I analyze data from an already existing large-scale data set, all of these steps were preregistered prior to accessing the data.
 The preregistration was designed on the basis of the panel documentation online (Kittel et al., 2020).
-In some cases the analyses could not be executed as originally planned, because some properties of the variables only became apparent upon data analysis.
-The most relevant deviations are reported below, and a complete list of all changes can be found online.
+In some cases the analyses could not be executed as originally planned, for example because some properties of the variables became apparent only upon data analysis.
+The most relevant deviations are reported below, and a complete list of all changes can be found in the online companion website (\url{https://xmtra.github.io/Austrian_Corona_Panel_Project/index.html}).
 
@@
 The study was conducted between March 2020 and October 2021.
-The contains 26 waves, and at the time of writing, the first 24 waves were available for download.
+It contains 26 waves, and at the time of writing, the first 24 waves were available for download.
 Each wave consists of at least 1,500 respondents.
-The overall sample size was \emph{N} = 3018, and 72432 observations were collected.
-Panel mortality was compensated through a continuous reacquistion of new participants.
+The overall sample size was \emph{N} = 3,018, and 72,432 observations were collected.
+Panel mortality was compensated through a continuous acquisition of new participants.
 All respondents needed to have access to the internet (via computer or mobile devices such as smartphones or tablets).
@@
 In order to participate in the study, the respondents needed to be Austrian residents and had to be at least 14 years old.
-The average age was 42 years, 49 percent were male, 3.97 percent had a University degree, and NA percent were currently unemployed.
+The average age was 42 years, 49 percent were male, 14 percent had a University degree, and 5 percent were currently unemployed.
 
-Because the data were analyzed post hoc, no sample size planning on the basis of a priori power analyses was conducted.
-The sample is very large, and it is hence well-equipped reliably to detect also small effects, which is why no exact post hoc power analysis were conducted.
-In addition, because such large samples easily generate significant \emph{p}-values even for very small effects, this study builds on a smallest effect size of interest.
+Because the data were analyzed post-hoc, no a-priori sample size planning on the basis of power analyses was conducted.
+The sample is very large, and it is hence likely well-equipped reliably to detect also small effects.
+In addition, because such large samples easily generate significant \emph{p}-values even for very small effects, it helps that to test hypothesis this study uses a smallest effect size of interest-approach.
 To test the hypotheses, I will use the interval testing approach as proposed by Dienes (2014).
@@
 For well-being, the null region will be between \emph{b} = -.30 and \emph{b} = .30.
-For example, if the 95\% confidence interval lies completely within the null-region (e.g., \emph{b} = .15, {[}95\% CI: .05, .25{]}), the hypothesis that the effect is only trivial is be supported.
+For example, if the 95\% confidence interval lies completely within the null-region (e.g., \emph{b} = .15, {[}95\% CI: .05, .25{]}), the hypothesis that the effect is only trivial is supported.
 If the effects interval and the null region overlap (e.g., \emph{b} = .25, {[}95\% CI: .15, .35{]}), the hypothesis is not supported and the results are considered inconclusive, whereas a meaningful negative effect is rejected.
-If the confidence falls completely outside of the null-region (e.g., \emph{b} = .4, {[}95\% CI: .35, .45{]}), the hypothesis is rejected and the existence of a meaningful positive effect is supported.
-
-Respondents who answered less than 50\% of all questions were removed.
-The remaining missing responses were imputed using predictive mean matching.
+If the confidence falls completely outside of the null-region (e.g., \emph{b} = .40, {[}95\% CI: .35, .45{]}), the hypothesis is rejected and the existence of a meaningful positive effect is supported.
 
@@
 Three models were run, one for each dependent variable.
-All predictors---that is, social media activities and social media channels, within and between-person factors, stably and varying covariates---were included in each of the three models.
-The data were hierarchical, and responses were nested in participants and waves (put differently, participants and waves were implemented as random effects).
+Predictors included all types and channels of social media communication separated into within and between-person factors, as well as stable and varying covariates.
+All predictors were included simultaneously and in each of the three models.
+The data were hierarchical, and responses were separately nested in participants and waves (put differently, participants and waves were implemented as random effects).
 Nesting in participants allowed to separate between-person relations from within-person effects.
-Nesting in waves allowed to control for general developments, such as general decrease in well-being in the population, for example due to lockdown measures (hence, there was no need additionally to control for specific phases or measures of the lockdown).
+Nesting in waves allowed to control for general exogenous developments, such as general decrease in well-being in the population, for example due to lockdown measures.
+Thus, there was no need additionally to control for specific phases or measures of the lockdown.
 
-For more information on the analyses, see \href{https://xmtra.github.io/Austrian_Corona_Panel_Project/index.html}{companion website}.
+For more information on the analyses, for example a complete documentation of the models and their results, see \href{https://xmtra.github.io/Austrian_Corona_Panel_Project/index.html}{companion website}.
 
@@
 
-The COVID-19 related social media use focused on interaction was measured with three variables (a) reading, (b) liking and sharing, and (c) posting.
+COVID-19 related social media use focused on types of communication was measured with three variables (a) reading, (b) liking and sharing, and (c) posting.
 The general introductory question was ``How often during the last week have you engaged in the following activities on social media?''
 The three items read as follows:
-``Reading the posts of others with content on the Coronavirus''; ``When seeing Posts on the Coronavirus, I clicked `like,' `share' or `retweet'\,''; ``I myself wrote posts on the Coronavirus on Social Media.''
+``Reading the posts of others with content on the Coronavirus''; ``When seeing posts on the Coronavirus, I clicked `like,' `share' or `retweet'\,''; ``I myself wrote posts on the Coronavirus on social media.''
 Answer options were 1 (\emph{several times per day}), 2 (\emph{daily}), 3 (\emph{several times per week}), 4 (\emph{weekly}), 5 (\emph{never}).
@@
 
-The COVID-19 related social media use focused on channels was measured with five variables.
-The general introductory question was ``How often in the last week have you followed information related to the Corona-crisis in the following social media?''
-The five items were (a) Facebook, (b) Twitter, (c) Instagram, (d) Youtube, (e) WhatsApp.
+COVID-19 related social media use focused on channels was measured with five variables.
+The general introductory question was ``How often in the last week have you followed information related to the Corona-crisis on the following social media?''
+The five items were (a) Facebook, (b) Twitter, (c) Instagram, (d) Youtube, and (e) WhatsApp.
 Again, the answer options were 1 (\emph{several times per day}), 2 (\emph{daily}), 3 (\emph{several times per week}), 4 (\emph{weekly}), 5 (\emph{never}).
@@
 
-Social media use was measured for everyone on waves 1, 2, 8, 17, and 23.
+Social media use was measured for all participants on waves 1, 2, 8, 17, and 23.
 Freshly recruited respondents always answered the questions on social media use.
@@
 The effects of COVID-19 related social media use were controlled for the following stable variables:
-(a) gender (answer options: female, male, diverse), (b) age, (c) education (10 options), (d) Austria country of birth (yes/no), (e) Austria parents' country of birth (no parent, one parent, both parents).
+(a) gender (answer options: female, male, diverse), (b) age, (c) education (ten options), (d) Austria country of birth (yes/no), (e) Austria parents' country of birth (no parent, one parent, both parents).
 I originally planned to implement other variables as varying covariates.
-However, because they were not measured often enough or at the time when social media use was measured, I implemented them as stable variables using their average values across all waves.
-This includes (a) text-based media news consumption, (b) video-based media news consumption, (c) residency is Vienna, (d) self-reported physical health, (e) living space (in squaremeter), (f) access to balcony, (g) access to garden, (h) employment status, (i) disposition to take risks, and (j) locus of control.
-I controlled also for the following varying covariates: (a) outdoor activities, (b) satisfaction with democracy.
+However, because they were not measured often enough or not at the time when social media use was measured, I implemented them as stable variables using their average values across all waves.
+This includes (a) text-based media news consumption (five degrees), (b) video-based media news consumption (five degrees), (c) residency is Vienna (yes/no), (d) self-reported physical health (five degrees), (e) living space (in squaremeter), (f) access to balcony (yes/no), (g) access to garden (yes/no), (h) employment (nine options), (i) disposition to take risks (eleven degrees), and (j) locus of control (five degrees).
+I controlled also for the following varying covariates: (a) five items measuring outdoor activities such as sport or meeting friends (five degrees), and (b) satisfaction with democracy (five degrees).
 Because it lead to too much attrition in the sample, I did not control for (a) household size, (b) work hours per week, (c) home office, (d) household income.
@@
 
-When looking at the variables from a descriptive perspective, we see that all well-being measures did not change substantially across the different phases of the pandemic.
-COVID-19 related media use, however, decreased during the beginning of the study and remained stable after approximately six waves.
-The initial decrease might be explained by the fact that the collection of data began in end of March 2020, hence approximately three months after the pandemic began.
-It could be that after an initial stark surge, COVID-19 related social media use was already declining, returning to more normal levels.
+When looking at the variables from a descriptive perspective, we see that all well-being measures did not change substantially across the different waves of data collection.
+COVID-19 related media use, however, decreased slightly at the beginning of the study and remained stable after approximately six waves.
+The initial decrease might be explained by the fact that the collection of data began at the end of March 2020, hence approximately three months after the pandemic began.
+It could be that after an initial uptick, COVID-19 related social media use was already declining at that time, returning to more normal levels.
 
@@
 \includegraphics[width=\textwidth]{figures/fig_descriptives} \caption{Development of well-being and media use measures across the pandemic.
-Values obtained from mixed effect models, with participants and waves as grouping factor and without additional predictors.}\label{fig:fig-desc}
+Values obtained from mixed effect models, with participants and waves as grouping factors and without additional predictors.}\label{fig:fig-desc}
 \end{figure}
 
-The study's hypothesis was that the effects of all types of social media use on well-being will be trivial.
-Regarding the different types of \emph{communication}---that is, reading vs.~sharing vs.~posting---all within-person effects fell completely within the a-priori defined SESOIs (see Figure \ref{fig:res-activity}).
+The study's hypothesis was that the effects of social media use on well-being would be trivial.
+Regarding the effects of different \emph{types} of communication---that is, reading vs.~sharing vs.~posting---all within-person effects fell completely within the a-priori defined SESOIs (see Figure \ref{fig:res-activity}).
 For example, respondents who used social media more frequently than usual to read about COVID-19 related topics did not show a simultaneous change in life satisfaction (\emph{b} = 0.04 {[}95\% CI -0.02, 0.09{]}).
-All confidence intervals included zero; hence, all effects were non-significant.
-As a result, the hypothesis was supported for all types of COVID-19 related social media communication.
+All confidence intervals include zero; hence, all effects were also statistically non-significant.
+As a result, the hypothesis was supported for all COVID-19 related types of social media communication.
 
@@
 Effects are controlled for a large number of covariates (see text).
-The SESOI was 0.30 for life satisfaction and .15 for affects; hence, no effect is considered meaningful theoretically.}\label{fig:res-activity}
+The SESOI was |0.30| for life satisfaction and |0.15| for affects.
+Hence, none of the reported effects are considered meaningful.}\label{fig:res-activity}
 \end{figure}
@@
 Regarding between-person relations, about which no hypotheses were formulated, only one effect didn't include zero.
-Respondents who across all waves used social media more frequently than others to write posts on COVID-19 reported higher levels of negative affect than others (\emph{b} = 0.08 {[}95\% CI 0.04, 0.12{]}).
-The effect was still completely inside of the null region, hence likely not large enough to be practically relevant.
-
-Regarding the COVID-19 related use of social media \emph{channels}, the results were very comparable.
-Changes in the frequency of using different social media channels to attain information regarding COVID-19 were unrelated to meaningful changes in well-being (see Figure \ref{fig:res-channels}).
-For example, respondents who used Facebook more frequently than usual to learn about COVID-19 did not show a simultaneous change in well-being (\emph{b} = -0.04 {[}95\% CI -0.1, 0.02{]}).
-Only two effects differed substantially from zero.
-Respondents who used Instagram more frequently than usual to attain COVID-19 related news reported slightly higher simultaneous levels of life satisfaction then usual (\emph{b} = 0.08 {[}95\% CI 0, 0.15{]}).
-Respondents who used Twitter more frequently than usual to attain COVID-19 related news reported slightly lower simultaneous levels of life satisfaction then usual (\emph{b} = -0.15 {[}95\% CI -0.27, -0.04{]}).
-However, both effects were still completely inside of the null region, hence not large enough to be considered meaningful.
-In sum, the hypothesis was supported for the COVID-19 related use of all types of social media channels.
+Respondents who across all waves used social media more frequently than others to write COVID-19 related posts reported higher levels of negative affect than others (\emph{b} = 0.08 {[}95\% CI 0.04, 0.12{]}).
+The effect was still completely inside of the null region, hence not considered large enough to be practically relevant.
 
@@
 Effects are controlled for a large number of covariates (see text).
-The SESOI was 0.30 for life satisfaction and .15 for affects; hence, no effect is considered meaningful theoretically.}\label{fig:res-channels}
+The SESOI was |0.30| for life satisfaction and |0.15| for affects.
+Hence, none of the reported effects are considered meaningful.}\label{fig:res-channels}
 \end{figure}
 
-In terms of between-person relations---which, again, weren't included in the hypotheses---no relations crossed or fell outside of the null region.
+Regarding the COVID-19 related use of social media \emph{channels}, the results were very comparable (see Figure \ref{fig:res-channels}).
+Changes in the frequency of using different social media channels to attain information regarding COVID-19 were unrelated to meaningful changes in well-being.
+For example, respondents who used Facebook more frequently than usual to learn about COVID-19 did not show a simultaneous change in well-being (\emph{b} = -0.04 {[}95\% CI -0.1, 0.02{]}).
+Only two effects differed substantially from zero.
+Respondents who used Instagram more frequently than usual to attain COVID-19 related news reported slightly higher simultaneous levels of life satisfaction than usual (\emph{b} = 0.08 {[}95\% CI \textless{} 0.01, 0.15{]}).
+Respondents who used Twitter more frequently than usual to attain COVID-19 related news reported somewhat lower simultaneous levels of life satisfaction than usual (\emph{b} = -0.15 {[}95\% CI -0.27, -0.04{]}).
+However, both effects were still completely inside of the null region, hence not considered large enough to be meaningful.
+In sum, the hypothesis was supported also for the COVID-19 related use of important social media channels.
+
+In terms of between-person relations---which, again, weren't included in the hypotheses---no relations crossed the null region or fell outside of it.
 Only one relation did not include zero, was hence statistically significant.
 Respondents who across all waves used YouTube more frequently than others for COVID-19 related reasons reported marginally higher levels of negative affect (\emph{b} = 0.03 {[}95\% CI \textless{} 0.01, 0.05{]}).
-However, please note that this effect was not large enough to be considered practically relevant.
+However, please note that this effect again was not considered large enough to be practically relevant.
 
@@
 In what follows, I briefly report also the results of some covariates.
-Several variables showed large associations with well-being.
-Because each variable has a different scaling, we would again need to define a SESOI for each variable, which cannot be implemented here.
-But note that several variables would fall outside of such a SESOI.
+Several variables showed larger associations with well-being.
+Note because each variable has a different scaling, we would again need to define a SESOI for each variable, which cannot be implemented here.
+But note that several variables would likely fall outside of such a SESOI.
 This includes for example internal locus of control, health, or employment.
@@
 
-The study is not aligned with a recent cross-sectional study by Bendau et al. (2021), which showed negative relations between social media and well-being.
-However, Bendau et al. (2021) analyzed cross-sectional data on a between-person level, while not controlling for third variables, which does not allow to make causal inferences.
-At the same time, this study is well-aligned with recent studies and meta-analyses from related research questions, which found that the effects of various types of social media use on several well-being indicators is small at best, often too small to matter (Ferguson et al., 2021; Meier \& Reinecke, 2020; Orben, 2020).
+The study is not aligned with several recent studies on this particular research question or closely related research questions.
+This includes a study by Bendau et al. (2021), which showed negative relations between social media and well-being.
+However, note that Bendau et al. (2021) analyzed cross-sectional data on a between-person level, while not controlling for third variables, which is not optimal for investigating causal effects.
+At the same time, this study is well-aligned with recent studies and meta-analyses analyzing the effects of social media use from a more general perspective or that adopted a somewhat different angle.
+These studies have found that the effects of various types of social media use on several well-being indicators are small at best, often too small to matter (Ferguson et al., 2021; Meier \& Reinecke, 2020; Orben, 2020), which echoes the results obtained here.
 
@@
 
-The current study analyzed whether changes in media use were related with changes in well-being, while controlling for several potential confounds.
+The current study analyzed whether changes in media use were related to changes in well-being, while controlling for several potential confounds.
 Together, this allows for a good perspective on potential causality.
-That said, causality necessitates temporal order, and the cause needs to precede the consequence.
+That said, causality necessitates temporal order, and the cause needs to precede the effect.
 Regarding media use, such effects often happen immediately or shortly after use, necessitating intervals in the hours, minutes, or even seconds.
-Only experience sampling studies that ask users in the very moment can produce such knowledge.
+In many cases only experience sampling studies that ask users in the very moment can produce such knowledge.
 However, even then we don't know for certain if we actually measured the right interval.
-Hence, to document how effects unfold it needs future research employing different study designs with different time lags.
-In addition, more thought needs to be invested in what relevant stable and nonvarying factor to control for.
+Hence, to document how effects unfold, future research needs to employ different study designs probing different time lags.
+In addition, more thought needs to be invested in what relevant stable and varying factors should be selected as control variables.
 
@@
 Media use is only one aspect of several factors that simultaneously affect well-being.
-Is it realistic that extremely changing only \emph{one} of these aspect (e.g., by completely stopping the use of social media) should already manifest in a detectable change in well-being?
-Or would it make more sense to expect that if people regularly start doing \emph{two} activities (e.g.~regularly exercising \emph{and} establishing a reading habit) together should show in perceivable improvements to well-being?
-In other words, if the beneficial effect of a particular activity is large enough, people should actually feel a difference if they implement two of those activities.
+Is it realistic that extremely changing only \emph{one} of these aspects (e.g., by completely stopping the use of social media) should already manifest in a detectable change in well-being?
+Or would it make more sense to expect that if people regularly start doing say \emph{two} activities (e.g.~regularly exercising \emph{and} establishing a reading habit) together should show perceivable improvements to well-being?
+In other words, the beneficial effect of a particular activity is large enough if people feel a difference when implementing two of those activities.
 Practically, this would imply a SESOI half as large as I have defined here, that is \emph{b} = \textbar.15\textbar{} for well-being and \emph{b} = \textbar.075\textbar.
 In this case, this would not make a difference, as even with these more liberal thresholds all but one effect would still be completely in the null region.
-However, at all events future research needs to start a discussion on what effect sizes are considered meaningful and relevant, and with this study I hope to provide some first concrete guidelines.
+However, at all events future research needs to start a thorough discussion on what effect sizes are considered meaningful and relevant, and with this study I hope to provide some helpful input and first concrete guidelines.
 
@@
 
-\leavevmode\hypertarget{ref-beyensSocialMediaUseinpress}{}%
+\leavevmode\hypertarget{ref-beyensSocialMediaUse2021}{}%
 Beyens, I., Pouwels, J.
@@
 M.
-(n.d.).
+(2021).
 Social {Media} {Use} and {Adolescents}' {Well}-{Being}: {Developing} a {Typology} of {Person}-{Specific} {Effect} {Patterns}.
@@
 
+\leavevmode\hypertarget{ref-johannesNoEffectDifferent2021}{}%
+Johannes, N., Dienlin, T., Bakhshi, H., \& Przybylski, A. K. (2021). \emph{No effect of different types of media on well-being}. PsyArXiv. \url{https://doi.org/10.31234/osf.io/zgb5y}
+
 \leavevmode\hypertarget{ref-kerestesAdolescentsOnlineSocial2020}{}%
@@
 Kline, R. B. (2016). \emph{Principles and practice of structural equation modeling} (4th ed.). New York, NY: The Guilford Press.
+
+\leavevmode\hypertarget{ref-mcelreathYesterdayClass2021}{}%
+McElreath, R. (2021). Yesterday in class, ... {[}Tweet{]}. Retrieved from \url{https://twitter.com/rlmcelreath/status/1354786005996482563}
 
@@
 
+\leavevmode\hypertarget{ref-scharkowHowSocialNetwork2020}{}%
+Scharkow, M., Mangold, F., Stier, S., \& Breuer, J.
+(2020).
+How social network sites and other online intermediaries increase exposure to news.
+\emph{Proceedings of the National Academy of Sciences}, \emph{117}(6), 2761--2763.
+\url{https://doi.org/10.1073/pnas.1918279117}
+
 \leavevmode\hypertarget{ref-schemerImpactInternetSocial2021}{}%

### 2021-11-05 analyses.RmD
- Status of paper when submitted to ICA

@@
                       tidy=T, dev = 'pdf', cache.extra = knitr::rand_seed)
-options(digits = 3, width = 140, scipen=999)
+options(digits = 3, width = 200, scipen=999)
 ```
@@
               "psych", "quanteda.textstats",
-              "pwr", "semPlot", "semTools", "tidyverse")
+              "pwr", "purrr", "semPlot", "semTools", "tidyverse")
 
@@
 
-Find Variables in dataset.
+Find variables in dataset.
 
@@
          value = as.numeric(.$value)) 
+} 
+```
+
+Determine average reliability for measures across waves
+
+```{r}
+get_rel <- function(data, waves=24){
+
+  # extract average reliability from lavaan fitted cfa with several groups
+  reliability(data) %>% 
+  unlist() %>% 
+  matrix(5, waves) %>% 
+  as.data.frame() %>% 
+  set_rownames(c("alpha", "omega", "omega2", "omega3", "avevar")) %>% 
+  summarise(omega = rowMeans(.["omega",])) %>% 
+  return()
 }
@@
 
+Positive affect was measured as a scale. We hence first inspect factorial validty using CFA.
+
+```{r}
+model <- "
+aff_pos =~ a1*aff_pos_1 + a2*aff_pos_2 + a3*aff_pos_3
+"
+cfa_aff_pos <- cfa(model, d_long, group = "wave")
+summary(cfa_aff_pos, standardized = TRUE, fit = TRUE, estimates = FALSE)
+```
+
+The data fit the model very well, `td::fit_txt(cfa_aff_pos)`.
+Let's next inspect reliability
+
+```{r}
+rel_aff_pos <- get_rel(cfa_aff_pos)
+rel_aff_pos
+```
+
+The average reliability across all waves was omega = `r rel_aff_pos`, hence good.
+
+Let's next expect the development across waves.
+
 ```{r}
@@
 ### Negative affect
+
+Negative affect was measured as a scale. We hence first inspect factorial validty using CFA.
+
+```{r}
+model <- "
+aff_neg =~ a1*aff_neg_1 + a2*aff_neg_2 + a3*aff_neg_3 + a4*aff_neg_4 + a5*aff_neg_5 + a6*aff_neg_6
+"
+cfa_aff_neg <- cfa(model, d_long, group = "wave")
+summary(cfa_aff_neg, standardized = TRUE, fit = TRUE, estimates = FALSE)
+```
+
+The data fit the model very well, `td::fit_txt(cfa_aff_neg)`.
+
+Let's next inspect reliability.
+
+```{r}
+rel_aff_neg <- get_rel(cfa_aff_neg)
+rel_aff_neg
+```
+
+The average reliability across all waves was omega = `r rel_aff_neg`, hence good.
 
@@
 ```
+
+## Locus of control
+
+The only other variable that was measured as a scale was Locus of Control. Below I hence report the scale's factorial validity.
+
+```{r}
+model <- "
+loc_cntrl_int =~ a1*loc_cntrl_int_1 + a2*loc_cntrl_int_2 + a3*loc_cntrl_int_3 + a4*loc_cntrl_int_4
+# loc_cntrl_int_1 ~~ loc_cntrl_int_2
+loc_cntrl_int_3 ~~ loc_cntrl_int_4
+"
+cfa_loc_cntrl_int <- cfa(model, filter(d_long, wave != 20), group = "wave")
+summary(cfa_loc_cntrl_int, standardized = TRUE, fit = TRUE, estimates = FALSE)
+```
+
+The data fit the model very well, `td::fit_txt(cfa_loc_cntrl_int)`.
 

### 2021-11-05 index.Rmd
- Status of paper when submitted to ICA

@@
 ---
-title: "No effets of COVID-related social media use and well-being"
+title: "No meaningful effets of COVID-related social media use and well-being"
 ---
 
-On this website, you can find the companion material for the paper _No effets of COVID-related social media use and well-being_.
+On this website, you can find the companion material for the paper _No meaningful effets of COVID-related social media use and well-being_.
 

### 2021-11-05 manuscript.Rmd
- Status of paper when submitted to ICA

@@
 classoption       : "man"
-# output            : papaja::apa6_pdf
+output            : papaja::apa6_pdf
 # output            : papaja::apa6_word
@@
 Several people reported that they were glued to their screens and could not pursue other relevant activities such as working, taking a break, or even child-care [@kleinDarklySoothingCompulsion2021].
-As doomscrolling increased it became doubtful whether such a surge in social media use could still be considered useful and adaptive, or whether it created an additional and new psychological danger for users' mental health [@sandstromDoomscrollingCOVIDNews2021].
+As doomscrolling increased it became doubtful whether such a surge in social media use could still be considered useful and adaptive, or whether it created an additional hazard for users' mental health [@sandstromDoomscrollingCOVIDNews2021].
 These concerns seem justified: A study with 6,233 people from Germany found that "[f]requency, duration and diversity of media exposure were positively associated with more symptoms of depression and unspecific and COVID-19 specific anxiety" [@bendauAssociationsCOVID19Related2021], which implies negative consequences for well-being.
@@
 In order not to exaggerate precision, and to be a bit less conservative, these numbers will be reduced to nearby thresholds. 
-Note that other researchers also decreased or recommended decreasing thresholds for effect sizes when analyzing within-person or cumulative effects (@beyensSocialMediaUse2021; @funderEvaluatingEffectSize2019).
+Note that other researchers also decreased or recommended decreasing thresholds for effect sizes when analyzing within-person or cumulative effects [@beyensSocialMediaUse2021; @funderEvaluatingEffectSize2019].
 Together, this leads to an indifference region ranging from _b_ = -.30 to _b_ = .30 for life satisfaction, and _b_ = -.15 to _b_ = .15 for positive and negative affect.
@@
 The response options were 1 (_never_), 2 (_on some days_), 3 (_several times per week_), 4 (_almost every day_), and 5 (_daily_).
+The scale showed good factorial fit, `r fit_txt(cfa_aff_pos)`.
+Reliability was also very high, omega = `r rel_aff_pos %>% my_round("std")`.
 
@@
 The response options were 1 (_never_), 2 (_on some days_), 3 (_several times per week_), 4 (_almost every day_), and 5 (_daily_).
+The scale showed good factorial fit, `r fit_txt(cfa_aff_neg)`.
+Reliability was also very high, omega = `r rel_aff_neg %>% my_round("std")`.
 

### 2021-11-05 manuscript.tex
- Status of paper when submitted to ICA

@@
 Several people reported that they were glued to their screens and could not pursue other relevant activities such as working, taking a break, or even child-care (Klein, 2021).
-As doomscrolling increased it became doubtful whether such a surge in social media use could still be considered useful and adaptive, or whether it created an additional and new psychological danger for users' mental health (Sandstrom, Buchanan, Aknin, \& Lotun, 2021).
+As doomscrolling increased it became doubtful whether such a surge in social media use could still be considered useful and adaptive, or whether it created an additional hazard for users' mental health (Sandstrom, Buchanan, Aknin, \& Lotun, 2021).
 These concerns seem justified: A study with 6,233 people from Germany found that ``{[}f{]}requency, duration and diversity of media exposure were positively associated with more symptoms of depression and unspecific and COVID-19 specific anxiety'' (Bendau et al., 2021), which implies negative consequences for well-being.
@@
 In order not to exaggerate precision, and to be a bit less conservative, these numbers will be reduced to nearby thresholds.
-Note that other researchers also decreased or recommended decreasing thresholds for effect sizes when analyzing within-person or cumulative effects (Beyens, Pouwels, Driel, Keijsers, and Valkenburg (2021); Funder and Ozer (2019)).
+Note that other researchers also decreased or recommended decreasing thresholds for effect sizes when analyzing within-person or cumulative effects (Beyens, Pouwels, Driel, Keijsers, \& Valkenburg, 2021; Funder \& Ozer, 2019).
 Together, this leads to an indifference region ranging from \emph{b} = -.30 to \emph{b} = .30 for life satisfaction, and \emph{b} = -.15 to \emph{b} = .15 for positive and negative affect.
@@
 The response options were 1 (\emph{never}), 2 (\emph{on some days}), 3 (\emph{several times per week}), 4 (\emph{almost every day}), and 5 (\emph{daily}).
+The scale showed good factorial fit, \(\chi^2\)(46) = 65.30, \textit{p} = .032, cfi = 1.00, rmsea = .02, 90\% CI {[}.01, .03{]}, srmr = .01.
+Reliability was also very high, omega = .85.
 
@@
 The response options were 1 (\emph{never}), 2 (\emph{on some days}), 3 (\emph{several times per week}), 4 (\emph{almost every day}), and 5 (\emph{daily}).
+The scale showed good factorial fit, \(\chi^2\)(331) = 3138.37, \textit{p} \textless{} .001, cfi = .97, rmsea = .08, 90\% CI {[}.07, .08{]}, srmr = .03.
+Reliability was also very high, omega = .89.
 

### 2021-11-09 index.Rmd
- corrected typo on landing page.

--- 2021-11-03 index.Rmd
@@
 ---
-title: "No effets of COVID-related social media use and well-being"
+title: "No meaningful effects of COVID-19 related social media use on well-being"
 ---
 
-On this website, you can find the companion material for the paper _No effets of COVID-related social media use and well-being_.
+On this website, you can find the companion material for the paper _No meaningful effects of COVID-19 related social media use on well-being_.
 

### 2021-12-13 analyses.RmD
- updated paper for submission

--- 2021-11-05 analyses.RmD
@@
     toc_depth: 2
-    # code_folding: hide
+    code_folding: hide
 bibliography      : 
@@
 set.seed(170819)
-knitr::opts_chunk$set(warning=F, echo=T, error=F, message=F, cache = T, 
-                      tidy=T, dev = 'pdf', cache.extra = knitr::rand_seed)
-options(digits = 3, width = 140, scipen=999)
+knitr::opts_chunk$set(warning=F, echo=T, error=F, message=F, cache = T)
+options(digits = 3, width = 250, scipen=999)
 ```
+
+Here you can find the code and the results of all analyses. To see the code, click on button "Code". Alternatively, you can download the rmd file from the github repo.
 
@@
 
-Load packages, and make repo replicable.
+Load packages.
 
-```{r r-setup, results='hide'}
+```{r r-setup, results='hide', cache=FALSE}
 # install packages
-# devtools::install_github("https://github.com/tdienlin/td")
+# devtools::install_github("https://github.com/tdienlin/td@v.0.0.2.5")
 
 # define packages
-packages <- c("brms", "corrplot", "devtools", "GGally", "ggplot2", 
-              "gridExtra", "haven", "kableExtra", "knitr", 
-              "lavaan", "lme4", "magrittr", "mice", "MVN", "PerFit", 
-              "psych", "quanteda.textstats",
-              "pwr", "semPlot", "semTools", "tidyverse")
-
-# renv::install(packages)
-# renv::install("tdienlin/td@v.0.0.2.5")  # uncomment to install
-# renv::install("crsh/papaja@v0.1.0.9997")  # uncomment to install
+packages <- c("brms", "devtools", "GGally", "ggplot2", 
+              "gridExtra", "kableExtra", "knitr", "lavaan", "lme4", 
+              "magrittr", "mice", "mvnormalTest", 
+              "PerFit", "psych", "quanteda.textstats", "semTools", "tidyverse")
 
@@
 
-Get data from fitted lmer objects
+Get data from fitted lmer objects for descriptives.
 
@@
 
+Get data of lmer objects for results.
+
+```{r}
+# get data
+get_dat_res <- function(data_aff_neg, data_aff_pos, data_life_sat, type){
+
+  dat_fig_results <- 
+  broom.mixed::tidy(data_aff_neg, conf.int = T) %>% 
+  mutate(dv = "aff_neg") %>% 
+  rbind(
+    broom.mixed::tidy(data_aff_pos, conf.int = T) %>% 
+    mutate(dv = "aff_pos")
+  ) %>% 
+  rbind(
+    broom.mixed::tidy(data_life_sat, conf.int = T) %>% 
+    mutate(dv = "life_sat")
+  ) %>% 
+  mutate(
+    type = ifelse(grepl(".*_w\\>", .$term), "within", "between"),
+    iv = gsub("_(w|b)\\>", "", .$term)
+         ) %>%
+  mutate(
+    type = factor(.$type, 
+                  levels = c("within", "between")),
+    dv = factor(.$dv, 
+                levels = c("life_sat", "aff_pos", "aff_neg"),
+                labels = c("Life satisfaction", "Positive affect", "Negative affect"))
+  ) %>% 
+  select(dv, iv, type, estimate, conf.low, conf.high, p.value)
+
+  # select Social Media type of activity
+  if(type == "activity") {
+    dat_fig_results %<>%
+    filter(iv %in% c("soc_med_read", "soc_med_like_share", "soc_med_post")) %>% 
+    mutate(
+      iv = factor(.$iv, 
+                  levels = c("soc_med_post", "soc_med_like_share", "soc_med_read"), 
+                  labels = c("Posting", "Liking & Sharing", "Reading"))
+    )
+  } else if(type == "channel"){
+    dat_fig_results %<>%
+    filter(iv %in% c("soc_med_fb", "soc_med_ig", "soc_med_wa", "soc_med_yt", "soc_med_tw")) %>% 
+    mutate(iv = factor(.$iv, levels = c("soc_med_yt", "soc_med_wa", "soc_med_ig", "soc_med_tw", "soc_med_fb"), labels = c("YouTube", "WhatsApp", "Instagram", "Twitter", "Facebook")))
+  } else if(type == "control"){
+    dat_fig_results %<>%
+      filter(iv %in% c("male", "health", "loc_cntrl_int_m", "employment_facPublic service", "act_spo", "sat_dem")) %>% 
+      mutate(iv = factor(.$iv, levels = c("male", "health", "loc_cntrl_int_m", "employment_facPublic service", "act_spo", "sat_dem"), 
+                         labels = c("Male", "Health", "Internal locus of control", "Employed public service", "Sport", "Satisfaction democracy")))
+  }
+  return(dat_fig_results)
+}
+
+```
+
 Make graph of effects
@@
 ```{r}
-make_graph_res <- function(data, sesoi = TRUE){
-  ggplot(data, aes(x = estimate, y = iv)) + 
+make_graph_res <- function(data_1, data_2 = NULL, sesoi = NULL){
+  ggplot(data_1, aes(x = estimate, y = iv)) + 
   scale_color_manual(values = c("grey40", "deepskyblue")) +
@@
   geom_errorbarh(aes(xmin = conf.low, xmax = conf.high),
-                lwd = .75, height = .2
-                , position = position_dodge(.6)
-                ) +
-  geom_point(size = 2
-             , position = position_dodge(.6)
-             ) +
-  {if(sesoi) geom_vline(data=filter(data, dv=="Life satisfaction"), aes(xintercept=.3), colour="darkgrey", linetype = "dashed")} + 
-  {if(sesoi) geom_vline(data=filter(data, dv=="Life satisfaction"), aes(xintercept=-.3), colour="darkgrey", linetype = "dashed")} + 
-  {if(sesoi) geom_vline(data=filter(data, dv!="Life satisfaction"), aes(xintercept=.15), colour="darkgrey", linetype = "dashed")} + 
-  {if(sesoi) geom_vline(data=filter(data, dv!="Life satisfaction"), aes(xintercept=-.15), colour="darkgrey", linetype = "dashed")} +
-  theme(
+                lwd = .75, height = .2) +
+  geom_point(size = 2) +
+  {if(!is.null(data_2)) geom_errorbarh(aes(xmin = conf.low, xmax = conf.high), data_2,
+                lwd = .75, height = .2, color = "grey",
+                position = position_nudge(y = -.3)
+                )} +
+  {if(!is.null(data_2)) geom_point(data = data_2, size = 2, color = "grey",
+             position = position_nudge(y = -.3)
+             )} +
+  {if(isTRUE(sesoi == "est")) geom_vline(data=filter(data_1, dv=="Life satisfaction"), aes(xintercept=-.3), colour="darkgrey", linetype = "dashed")} + 
+    {if(isTRUE(sesoi == "est")) geom_vline(data=filter(data_1, dv=="Life satisfaction"), aes(xintercept=.3), colour="darkgrey", linetype = "dashed")} + 
+  {if(isTRUE(sesoi == "est")) geom_vline(data=filter(data_1, dv!="Life satisfaction"), aes(xintercept=-.15), colour="darkgrey", linetype = "dashed")} +
+    {if(isTRUE(sesoi == "est")) geom_vline(data=filter(data_1, dv!="Life satisfaction"), aes(xintercept=.15), colour="darkgrey", linetype = "dashed")} +
+  {if(isTRUE(sesoi == "std")) geom_vline(aes(xintercept=.1), colour="darkgrey", linetype = "dashed")} + 
+  {if(isTRUE(sesoi == "std")) geom_vline(aes(xintercept=-.1), colour="darkgrey", linetype = "dashed")} +   theme(
     axis.title.y = element_blank(),
@@
              space = "free_y") 
+}
+```
+
+Extract factor scores.
+
+```{r}
+get_fs <- function(model) {
+  # the aim is to get factor scores on original scaling
+  # hence export factor values for all items
+  # then compute average
+  dat_ov <- lavaan::lavPredict(model, type = "ov", assemble = TRUE) %>% 
+    mutate(fs = rowMeans(.[1:3]))
+  return(dat_ov$fs)
 }
@@
 # identify variable names of items, using custom function
-health <- find_var("Gesundheitszustand") # wasn't collected at wave 21
+health <- find_var("Gesundheitszustand") # wsn't collected at wave 21
 aff_neg_1 <- find_var("Depressivitaet: einsam")
@@
   )
-
-# rename variables
-# names(d_wide) <- sub("\\.\\.\\.", "§", names(d_wide)) # replace "..." with "_"
 ```
@@
 
-Recode Variables
+Recode variables.
 
@@
     employment_fac = factor(.$employment, 
-                           # labels = c("Unselbststaendig erwerbstaetig (auch Lehrling), Privatwirtschaft",
-                           #            "Unselbststaendig erwerbstaetig (auch Lehrling), Oeffentlicher Dienst",
-                           #            "Selbststaendig erwerbstaetig",
-                           #            "Arbeitslos",
-                           #            "Pensionist*in",
-                           #            "Ausschliesslich haushaltsfuehrend",
-                           #            "Schueler/in, Student/in, Bildungskarenz",
-                           #            "Sonstiges (Dauerhaft arbeitsunfaehig oder Praesenz-/Zivildiener)",
-                           #            "In Elternkarenz")
-                           labels = c("Employed industrie",
-                                      "Employed public service",
+                           labels = c("Unemployed",
+                                      "Industrie",
+                                      "Public service",
                                       "Self-employed",
-                                      "Unemployed",
                                       "Retired",
@@
                                       "Incapacitated",
-                                      "Parental Leave")
+                                      "Parental Leave"),
+                            levels = c(4, 1:3, 5:8, 10) # make unemployment reference cat
                            ),
     edu_fac = factor(.$edu,
-                     # labels = c("Volksschule oder weniger", 
-                     #            "Hauptschule oder AHS Unterstufe",
-                     #            "Polytechnikum, BMS (Fachschule, z.B. HASCH)",
-                     #            "Lehre, Berufsschule",
-                     #            "AHS mit Matura",
-                     #            "BHS mit Matura (z.B. HTL, HAK, HBLA etc.)",
-                     #            "Hochschulverwandte Lehranstalt oder Kolleg",
-                     #            "Bachelor",
-                     #            "Magister / Master / Diplomingenieur / Fachhochschule",
-                     #            "Doktor / PhD"),
                      labels = c("No degree", 
@@
 
-<!-- Filter waves without data -->
-
-<!-- ```{r eval=F} -->
-<!-- d_long_test <- d_long %>% -->
-<!--   filter(!is.na(.[vars_varying])) -->
-<!-- ``` -->
-
-Recode values
+Recode values.
 
@@
 
+# Impute data
+
+Determine amount of missingness per respondent per wave. Because calculation takes long, this chunk is not executed but loaded from memory. To execute, set eval == TRUE.
+
+```{r eval=TRUE}
+vars_used <- c("life_sat", "aff_pos_1", "aff_pos_2", "aff_pos_3", "aff_neg_1", "aff_neg_2", "aff_neg_3", "aff_neg_4", "aff_neg_5", "aff_neg_6", "soc_med_read", "soc_med_like_share", "soc_med_post", "soc_med_fb", "soc_med_ig", "soc_med_wa", "soc_med_yt", "soc_med_tw", "age", "male", "born_aus", "born_aus_prnts", "edu_fac", "employment_fac", "health", "res_vienna", "acc_bal", "acc_gar", "home_sqm", "med_txt_kro", "med_txt_sta", "med_txt_pre", "med_txt_oes", "med_txt_kur", "med_txt_slz", "med_txt_son", "med_vid_orf", "med_vid_pri", "risk_prop", "loc_cntrl_int_1", "loc_cntrl_int_2", "loc_cntrl_int_3", "loc_cntrl_int_4", "act_wrk", "act_spo", "act_frn", "act_sho", "act_pet", "sat_dem")
+
+# filter respondents with more than 50% missing data
+d_long_filt <- 
+  d_long %>% 
+  mutate(na_perc = rowSums(is.na(select(., all_of(vars_used))) / ncol(select(., all_of(vars_used))))) %>% 
+  filter(na_perc < .5)
+
+# impute missing data
+# exclude social media use data, because they were measured only on selected waves and would lead to the imputation of too much missing data
+vars_excl <- c("na_perc", "soc_med_read", "soc_med_like_share", "soc_med_post", "soc_med_fb", "soc_med_ig", "soc_med_wa", "soc_med_yt", "soc_med_tw")
+incl_ma <- d_long_filt %>% 
+  mutate(across(.cols = everything(), .fns = is.na))
+incl_ma[vars_excl] <- FALSE
+
+# init <- mice(select(d_long_test_filt, maxit = 0))
+# pred_ma <- init$predictorMatrix
+# pred_ma[, vars_excl] <- 0
+d_long_imp <- mice(d_long_filt, method = "pmm", m = 1, maxit = 30, 
+              where = incl_ma,
+              # predictorMatrix = pred_ma, 
+              seed = 180719, print = FALSE)
+d_long_imp <- mice::complete(d_long_imp)
+
+write_csv(d_long_imp, "data/data_imputed.csv")
+```
+
+Load imputed data from disc.
+
+```{r}
+d_long_imp <- read_csv("data/data_imputed.csv")
+```
+
+# Create within-between data
+
 Create means for scales
@@
 ```{r}
-d_long %<>%
+d_long_imp %<>%
   mutate(aff_pos_m = rowMeans(select(., aff_pos_1 : aff_pos_3)),
@@
 ```{r}
-d_long %<>% 
+d_long_imp %<>% 
   group_by(id) %>% 
@@
 
-<!-- Create vectors with names -->
-
-<!-- ```{r} -->
-<!-- # create vector of names -->
-<!-- vars_varying <- names(select(d_long, "health_b":"loc_cntrl_int_4_w")) -->
-<!-- predictors_varying <- select(d_long, "health_b":"loc_cntrl_int_4_w") %>%  -->
-<!--   select(-starts_with("aff_"), -starts_with("life_")) %>%  -->
-<!--   names() -->
-<!-- ``` -->
-
-# Inspect data
+And because later we want to rerun the results without imputed data, now the same for data without imputation.
 
 ```{r}
-nas <- sapply(d_long, function(y) sum(length(which(is.na(y))))) %>% 
-  order()
-```
+d_long %<>%
+  mutate(aff_pos_m = rowMeans(select(., aff_pos_1 : aff_pos_3)),
+         aff_neg_m = rowMeans(select(., aff_neg_1 : aff_neg_6)),
+         loc_cntrl_int_m = rowMeans(select(., loc_cntrl_int_1 : loc_cntrl_int_4)))
 
-Inspect participants with NA as ID
-
-```{r}
-d_long_filtered <- d_long %>% filter(is.na(id))
-d_raw_filtered <- d_raw %>% filter(is.na(.$RESPID))
+d_long %<>% 
+  group_by(id) %>% 
+  mutate_at(vars(health:loc_cntrl_int_4, loc_cntrl_int_m), funs(b = mean(., na.rm = TRUE), w = . - mean(., na.rm = TRUE))) %>% 
+  ungroup()
 ```
@@
 ```{r}
-model_life_sat <- lmer(life_sat ~ (1 | id) + (1 | wave), d_long)
+model_life_sat <- lmer(life_sat ~ (1 | id) + (1 | wave), d_long_imp)
 summary(model_life_sat)
@@
 
+Positive affect was measured as a scale. We hence first inspect factorial validity using CFA.
+
+We first test assumption of multivariate normality.
+
+```{r eval=FALSE}
+mardia(d_long_imp %>% select(aff_pos_1, aff_pos_2, aff_pos_3))
+```
+
+Assumption of multivariate normal distributino was violated; hence, robust estimator will be used.
+
 ```{r}
-model_aff_pos <- lmer(aff_pos_m ~ (1 | id) + (1 | wave), d_long)
+model <- "
+aff_pos =~ a1*aff_pos_1 + a2*aff_pos_2 + a3*aff_pos_3
+"
+cfa_aff_pos <- cfa(model, d_long_imp, group = "wave", estimator = "MLM")
+summary(cfa_aff_pos, standardized = TRUE, fit = TRUE, estimates = FALSE)
+```
+
+The data fit the model very well, `r td::fit_txt(cfa_aff_pos)`.
+Let's next inspect reliability
+
+```{r}
+rel_aff_pos <- get_rel(cfa_aff_pos)
+rel_aff_pos
+```
+
+The average reliability across all waves was omega = `r rel_aff_pos`, hence good.
+
+Let's next expect the development across waves.
+
+```{r}
+model_aff_pos <- lmer(aff_pos_m ~ (1 | id) + (1 | wave), d_long_imp)
 summary(model_aff_pos)
+```
+
+Let's now export factor scores.
+
+```{r}
+d_long_imp$aff_pos_fs <- get_fs(cfa_aff_pos)
 ```
@@
 
+Negative affect was measured as a scale. We hence inspect factorial validty using CFA.
+
+We first test assumption of multivariate normality. (Warning, as these analyses take long and might require additional working memory, they're deactived currently.)
+
+```{r eval=FALSE}
+mardia(d_long_imp %>% select(aff_neg_1, aff_neg_2, aff_neg_3, aff_neg_4, aff_neg_5, aff_neg_6))
+```
+
+Assumption of multivariate normal distributino was violated; hence, robust estimator will be used.
+
 ```{r}
-model_aff_neg <- lmer(aff_neg_m ~ (1 | id) + (1 | wave), d_long)
+model <- "
+aff_neg =~ a1*aff_neg_1 + a2*aff_neg_2 + a3*aff_neg_3 + a4*aff_neg_4 + a5*aff_neg_5 + a6*aff_neg_6
+"
+cfa_aff_neg <- cfa(model, d_long_imp, group = "wave", estimator = "MLM")
+summary(cfa_aff_neg, standardized = TRUE, fit = TRUE, estimates = FALSE)
+```
+
+The data fit the model very well, `r td::fit_txt(cfa_aff_neg)`.
+
+Let's next inspect reliability.
+
+```{r}
+rel_aff_neg <- get_rel(cfa_aff_neg)
+rel_aff_neg
+```
+
+The average reliability across all waves was omega = `r rel_aff_neg`, hence good.
+
+```{r}
+model_aff_neg <- lmer(aff_neg_m ~ (1 | id) + (1 | wave), d_long_imp)
 summary(model_aff_neg)
+```
+
+Let's now export factor scores.
+
+```{r}
+d_long_imp$aff_neg_fs <- get_fs(cfa_aff_neg)
 ```
@@
 ```{r}
-model_soc_med_read <- lmer(soc_med_read ~ (1 | id) + (1 | wave), d_long)
+model_soc_med_read <- lmer(soc_med_read ~ (1 | id) + (1 | wave), d_long_imp)
 summary(model_soc_med_read)
@@
 ```{r}
-model_soc_med_like_share <- lmer(soc_med_like_share ~ (1 | id) + (1 | wave), d_long)
+model_soc_med_like_share <- lmer(soc_med_like_share ~ (1 | id) + (1 | wave), d_long_imp)
 summary(model_soc_med_like_share)
@@
 ```{r}
-model_soc_med_post <- lmer(soc_med_post ~ (1 | id) + (1 | wave), d_long)
+model_soc_med_post <- lmer(soc_med_post ~ (1 | id) + (1 | wave), d_long_imp)
 summary(model_soc_med_post)
@@
 ```{r}
-model_soc_med_fb <- lmer(soc_med_fb ~ (1 | id) + (1 | wave), d_long)
+model_soc_med_fb <- lmer(soc_med_fb ~ (1 | id) + (1 | wave), d_long_imp)
 summary(model_soc_med_fb)
@@
 ```{r}
-model_soc_med_tw <- lmer(soc_med_tw ~ (1 | id) + (1 | wave), d_long)
+model_soc_med_tw <- lmer(soc_med_tw ~ (1 | id) + (1 | wave), d_long_imp)
 summary(model_soc_med_tw)
-
-# Bayesian Approach
-# model_soc_med_tw <- brm(soc_med_tw ~ (1 | id) + (1 | wave), d_long
-# Also throws a warning
-# model_soc_med_tw <- brm(soc_med_tw ~ (1 | id) + (1 | wave), d_long, control=list(adapt_delta=0.82))
-# Currently doesn't work, likely for lack of memory
 ```
@@
 ```{r}
-soc_med_tw_m <- d_long %>% group_by(wave) %>% summarise(value = mean(soc_med_tw, na.rm = T))
+soc_med_tw_m <- d_long_imp %>% group_by(wave) %>% summarise(value = mean(soc_med_tw, na.rm = T))
 soc_med_tw_m
@@
 ```{r}
-model_soc_med_ig <- lmer(soc_med_ig ~ (1 | id) + (1 | wave), d_long)
+model_soc_med_ig <- lmer(soc_med_ig ~ (1 | id) + (1 | wave), d_long_imp)
 summary(model_soc_med_ig)
@@
 ```{r}
-model_soc_med_wa <- lmer(soc_med_wa ~ (1 | id) + (1 | wave), d_long)
+model_soc_med_wa <- lmer(soc_med_wa ~ (1 | id) + (1 | wave), d_long_imp)
 summary(model_soc_med_wa)
@@
 ```{r}
-model_soc_med_yt <- lmer(soc_med_yt ~ (1 | id) + (1 | wave), d_long)
+model_soc_med_yt <- lmer(soc_med_yt ~ (1 | id) + (1 | wave), d_long_imp)
 summary(model_soc_med_yt)
 ```
+
+## Locus of control
+
+The only other variable that was measured as a scale was Locus of Control. Below I hence report the scale's factorial validity.
+
+```{r}
+model <- "
+loc_cntrl_int =~ a1*loc_cntrl_int_1 + a2*loc_cntrl_int_2 + a3*loc_cntrl_int_3 + a4*loc_cntrl_int_4
+# loc_cntrl_int_1 ~~ loc_cntrl_int_2
+loc_cntrl_int_3 ~~ loc_cntrl_int_4
+"
+cfa_loc_cntrl_int <- cfa(model, filter(d_long_imp, wave != 20), group = "wave")
+summary(cfa_loc_cntrl_int, standardized = TRUE, fit = TRUE, estimates = FALSE)
+```
+
+The data fit the model very well, `r td::fit_txt(cfa_loc_cntrl_int)`.
 
@@
                             soc_med_fb_b + soc_med_ig_b + soc_med_wa_b + soc_med_yt_b + soc_med_tw_b,
-                data = d_long)
+                data = d_long_imp)
 summary(model_life_sat_lmer_nco)
@@
                                         sat_dem_w + sat_dem_b 
-                                      , data = d_long)
+                                      , data = d_long_imp)
 summary(model_life_sat_lmer)
@@
 ```{r}
-model_aff_pos_lmer_nco <- lmerTest::lmer(aff_pos_m ~ 
+model_aff_pos_lmer_nco <- lmerTest::lmer(aff_pos_fs ~ 
                                         (1 | id) + (1 | wave) + 
@@
                                         soc_med_fb_b + soc_med_ig_b + soc_med_wa_b + soc_med_yt_b + soc_med_tw_b
-                                      , data = d_long)
+                                      , data = d_long_imp)
 summary(model_aff_pos_lmer_nco)
@@
 ```{r}
-model_aff_pos_lmer <- lmerTest::lmer(aff_pos_m ~ 
+model_aff_pos_lmer <- lmerTest::lmer(aff_pos_fs ~ 
                                         (1 | id) + (1 | wave) + 
@@
                                         res_vienna + acc_bal + acc_gar + home_sqm + 
-                                        # hh_income_b + # large dropout (900 obs)
-                                        # hh_adults + # too much dropout
-                                        # hh_child18 + hh_child17 + hh_child14 + hh_child5 + hh_child2 + hh_oldfam + hh_outfam + hh_partner + 
-                                        # home_owner + # too much dropout
                                         med_txt_kro_b + med_txt_sta_b + med_txt_pre_b + med_txt_oes_b + med_txt_kur_b + med_txt_slz_b + med_txt_son_b + 
@@
                                         sat_dem_w + sat_dem_b 
-                                      , data = d_long)
+                                      , data = d_long_imp)
 summary(model_aff_pos_lmer)
@@
 ```{r}
-model_aff_neg_lmer_nco <- lmerTest::lmer(aff_neg_m ~ 
+model_aff_neg_lmer_nco <- lmerTest::lmer(aff_neg_fs ~ 
                                         (1 | id) + (1 | wave) + 
@@
                                         soc_med_fb_b + soc_med_ig_b + soc_med_wa_b + soc_med_yt_b + soc_med_tw_b
-                                      , data = d_long)
+                                      , data = d_long_imp)
 summary(model_aff_neg_lmer_nco)
@@
 ```{r}
-model_aff_neg_lmer <- lmerTest::lmer(aff_neg_m ~ 
+model_aff_neg_lmer <- lmerTest::lmer(aff_neg_fs ~ 
                                         (1 | id) + (1 | wave) + 
@@
                                         sat_dem_w + sat_dem_b 
-                                      , data = d_long)
+                                      , data = d_long_imp)
 summary(model_aff_neg_lmer)
@@
 
-## Figure
+# Figures
 
-First, extract data.
+First, for activities.
 
@@
 # get data
-dat_fig_results <- 
-  broom.mixed::tidy(model_aff_neg_lmer, conf.int = T) %>% 
-  mutate(dv = "aff_neg") %>% 
-  rbind(
-    broom.mixed::tidy(model_aff_pos_lmer, conf.int = T) %>% 
-    mutate(dv = "aff_pos")
-  ) %>% 
-  rbind(
-    broom.mixed::tidy(model_life_sat_lmer, conf.int = T) %>% 
-    mutate(dv = "life_sat")
-  ) %>% 
-  mutate(
-    type = ifelse(grepl(".*_w\\>", .$term), "within", "between"),
-    iv = gsub("_(w|b)\\>", "", .$term)
-         ) %>%
-  mutate(
-    type = factor(.$type, 
-                  levels = c("within", "between")),
-    dv = factor(.$dv, 
-                levels = c("life_sat", "aff_pos", "aff_neg"),
-                labels = c("Life satisfaction", "Positive affect", "Negative affect"))
-  ) %>% 
-  select(dv, iv, type, estimate, conf.low, conf.high, p.value)
-```
-
-Select social media types of activity, and make graph.
-
-```{r}
-# select Social Media type of activity
-dat_fig_results_activity <- dat_fig_results %>%
-  filter(iv %in% c("soc_med_read", "soc_med_like_share", "soc_med_post")) %>% 
-  mutate(
-    iv = factor(.$iv, 
-                levels = c("soc_med_post", "soc_med_like_share", "soc_med_read"), 
-                labels = c("Posting", "Liking & Sharing", "Reading"))
-  )
+dat_fig_results_activity <- get_dat_res(model_aff_neg_lmer, model_aff_pos_lmer, model_life_sat_lmer, type = "activity")
+dat_fig_results_activity_nco <- get_dat_res(model_aff_neg_lmer_nco, model_aff_pos_lmer_nco, model_life_sat_lmer_nco, type = "activity")
 
 # make figure
-fig_results_activity <- make_graph_res(dat_fig_results_activity)
+fig_results_activity <- make_graph_res(
+  data_1 = dat_fig_results_activity,
+  data_2 = dat_fig_results_activity_nco,
+  sesoi = "est"
+  )
 fig_results_activity
@@
 
-Select social media channels and make figure.
+Then for channels.
 
 ```{r}
-# select social media channel
-dat_fig_results_channel <- dat_fig_results %>%
-  filter(iv %in% c("soc_med_fb", "soc_med_ig", "soc_med_wa", "soc_med_yt", "soc_med_tw")) %>% 
-  mutate(iv = factor(.$iv, levels = c("soc_med_yt", "soc_med_wa", "soc_med_ig", "soc_med_tw", "soc_med_fb"), labels = c("YouTube", "WhatsApp", "Instagram", "Twitter", "Facebook")))
+# get data
+dat_fig_results_channel <- get_dat_res(model_aff_neg_lmer, model_aff_pos_lmer, model_life_sat_lmer, type = "channel")
+dat_fig_results_channel_nco <- get_dat_res(model_aff_neg_lmer_nco, model_aff_pos_lmer_nco, model_life_sat_lmer_nco, type = "channel")
 
 # make figure
-fig_results_channel <- make_graph_res(dat_fig_results_channel)
+fig_results_channel <- make_graph_res(
+  data_1 = dat_fig_results_channel,
+  data_2 = dat_fig_results_channel_nco,
+  sesoi = "est"
+  )
 fig_results_channel
@@
 ```{r}
-# select social media channel
-dat_fig_results_control <- dat_fig_results %>%
-  filter(iv %in% c("age", "male", "health", "home_sqm", "loc_cntrl_int_m", "employment_facUnemployed", "employment_facParental Leave", "act_spo", "sat_dem")) %>% 
-  mutate(iv = factor(.$iv, levels = c("age", "male", "health", "home_sqm", "loc_cntrl_int_m", "employment_facUnemployed", "employment_facParental Leave", "act_spo", "sat_dem"), 
-                     labels = c("Age", "Male", "Health", "Squaremeter home", "Internal locus of control", "Umemployed", "On parental leave", "Sport", "Satisfaction democracy")))
-
 # make figure
-fig_results_control <- make_graph_res(dat_fig_results_control, sesoi = FALSE)
+fig_results_control <- make_graph_res(
+  get_dat_res(model_aff_neg_lmer, model_aff_pos_lmer, model_life_sat_lmer, type = "control")
+  )
 fig_results_control
@@
 ```{r}
-save.image("data/workspace.RData")
+save.image("data/workspace_1.RData")
 ```

### 2021-12-13 index.Rmd
- updated paper for submission

--- 2021-11-03 index.Rmd
@@
 ---
-title: "No effets of COVID-related social media use and well-being"
+title: "Analyzing the effects of COVID-19 related social media use on well-being"
 ---
 
-On this website, you can find the companion material for the paper _No effets of COVID-related social media use and well-being_.
+On this website, you can find the companion material for the paper _Analyzing the effects of COVID-19 related social media use on well-being_.
 
@@
 - The detailed code and summaries of the [analyses](analyses.html)
+- [Additional analyses](analyses_additional.html).
 - A list of the [items translated into English](items_english.html)

### 2021-12-13 manuscript.Rmd
- updated paper for submission

--- 2021-11-05 manuscript.Rmd
@@
 ---
-title             : "No meaningful effects of COVID-19 related social media use on well-being"
+title             : "Analyzing the effects of COVID-19 related social media use on well-being"
 shorttitle        : "COVID-19 related social media use and well-being"
@@
 author: 
-  - name          : "Blinded for review"
+  - name          : "Name blinded for review"
     affiliation   : "1"
     corresponding : yes
-    address       : "Blinded for review"
-    email         : "Blinded for review"
+    address       : ""
+    email         : ""
 
@@
   - id            : "1"
-    institution   : "Blinded for review"
+    institution   : ""
 
-authornote: |
-  Blinded for review.
-
-abstract          : "In times of crisis such as the Corona pandemic, it is important that citizens stay informed about recent events, the latest political decisions, or mandatory protection measures.
+abstract          : "In times of crisis such as the Corona pandemic citizens need to stay informed about recent events, the latest political decisions, or mandatory protection measures.
 To this end, many people use various types of media, and increasingly social media.
 However, because social media are particularly engaging, some find it hard to disconnect and cannot stop 'doomscrolling'.
-In this preregistered study, I investigate whether using social media for COVID-19 related topics might put personal well-being at risk.
+In this preregistered study, I investigate whether using social media for COVID-19 related reasons affects psychological well-being.
 To answer this question I analyzed data from the Austrian Corona Panel Project, which consists of 24 waves with overall 3,018 participants.
-I ran three random effects cross lagged panel models, controlling for several stable and varying covariates.
-Results showed that the effects of various types of COVID-19 related social media use on several indicators of well-being were very small, arguably too small to matter.
-The findings suggest fears that social media use during times of crisis might be detrimental for well-being are likely to be unfounded."
-keywords          : "COVID-19, Coronavirus, well-being, affect, life satisfaction, social media use, news use, communication, random effects within between model, panel study, longitudinal."
+I ran three random effects within between models, controlling for several stable and varying confounders.
+Results showed that the effects of COVID-19 related social media use on well-being were very small, arguably too small to matter.
+The findings suggest that fears that social media use during times of crisis impairs well-being are likely to be unfounded."
+keywords          : "COVID-19, Coronavirus, well-being, affect, life satisfaction, social media use, news use, communication, random effects within between model, panel study, longitudinal study."
 
@@
 classoption       : "man"
-# output            : papaja::apa6_pdf
-# output            : papaja::apa6_word
+output            : papaja::apa6_pdf
 header-includes:
@@
 knitr::opts_chunk$set(cache.extra = knitr::rand_seed,
-                      cache=T,
-                      echo=F
-                      # , fig.pos = 'h'
-                      )
+                      cache=T, echo=F)
 ```
@@
 
-# make project reproducible
-# devtools::install_github("rstudio/renv")  # install renv
-# renv::init()  # initialize renv
+# install necessary packages
+# devtools::install_github("https://github.com/tdienlin/td@v.0.0.2.5")
+# devtools::install_github("https://github.com/crsh/papaja@devel")
 
-# install necessary packages
+# load packages
 packages <- c("devtools", "english", "ggplot2", "kableExtra", 
@@
               "magrittr", "semPlot", "tidyverse")
-# renv::install(packages)
-# renv::install("tdienlin/td@v.0.0.2.5")  # uncomment to install
-# renv::install("crsh/papaja@v0.1.0.9997")  # uncomment to install
-# renv::install("crsh/papaja@devel")  # uncomment to install
-
-# load packages
 lapply(c(packages, "papaja", "td"), library, character.only = TRUE)
 
-# save snapshot of installed packages
-# renv::snapshot()
-
 # load workspace created from analyses.rmd to use results
-load("data/workspace.RData")
-
-# export references
-# r_refs(file = "bibliography/r-references.bib")
+load("data/workspace_2.RData")
 ```
@@
 During the COVID-19 pandemic, 
-<!-- and especially in the beginning,  -->
-numerous events unfolded in quick succession.
-Several open questions emerged.
+numerous events unfolded in quick succession and several open questions emerged.
 How dangerous is the virus? 
 Is it spreading in my region? 
-How is it transmitted? 
-How can I protect myself?
-Because for many it was (and continues to be) a matter of life or death, citizens had to stay informed regarding the latest developments.
-Governments around the world implemented safety measures, ranging from wearing masks, keeping physical distance, to complete lockdowns.
-In this extraordinary situation, people used media excessively to attain information, and especially social media were at an all time high [@statistaAverageDailyTime2021].
-A new phenomenon termed "doomscrolling" emerged:
-People could not stop using social media to learn about COVID-19 related news.
+How is it transmitted, and how can I protect myself?
+Because for many it was (and still is) a matter of life or death, people aimed to stay informed regarding the latest developments.
+Governments around the world implemented safety measures, such as wearing masks, keeping physical distance, or enforcing lockdowns.
+In this extraordinary situation, many people used media excessively to attain information, and especially social media were at an all time high [@statistaAverageDailyTime2021].
 
-Several people reported that they were glued to their screens and could not pursue other relevant activities such as working, taking a break, or even child-care [@kleinDarklySoothingCompulsion2021].
-As doomscrolling increased it became doubtful whether such a surge in social media use could still be considered useful and adaptive, or whether it created an additional and new psychological danger for users' mental health [@sandstromDoomscrollingCOVIDNews2021].
-These concerns seem justified: A study with 6,233 people from Germany found that "[f]requency, duration and diversity of media exposure were positively associated with more symptoms of depression and unspecific and COVID-19 specific anxiety" [@bendauAssociationsCOVID19Related2021], which implies negative consequences for well-being.
+Some people actually couldn't stop using social media to learn about COVID-19 related news.
+A new phenomenon emerged, termed "doomscrolling":
+Users were glued to their screens and found it hard to pursue other relevant activities such as working, taking a break, or looking after their children [@kleinDarklySoothingCompulsion2021].
+As doomscrolling increased it became doubtful whether such social media is helpful, or whether it created an additional burden on mental health [@sandstromDoomscrollingCOVIDNews2021].
+These concerns seem justified: 
+A study with 6,233 people from Germany conducted during the pandemic found that "[f]requency, duration and diversity of media exposure were positively associated with more symptoms of depression" [@bendauAssociationsCOVID19Related2021, p.
+283].
 
-As a result, with this study I want to build on this research and investigate the question whether COVID-19 related social media use during the pandemic affected well-being.
-To this end, I analyze a large-scale panel study from the Austrian Corona Panel Project [@kittelAustrianCoronaPanel2020]. 
-The panel consists of 24 waves and has an overall sample size of `r nrow(d_wide)`.
-At least 1,500 participants took part per wave, and the sample is representative of the Austrian population.
-The panel study collected a large number of variables.
-Being able to control for many confounding third variables, both stable and varying, and together with the longitudinal design, this creates a unique opportunity to investigate the causal effects of COVID-19 related social media use on well-being.
+As a result, with this study I want to build on this research and investigate whether or not COVID-19 related social media use affected well-being during the pandemic.
+To this end, I analyzed a large-scale panel study from the Austrian Corona Panel Project [@kittelAustrianCoronaPanel2020]. 
+The panel consists of 24 waves and has an overall sample size of `r nrow(d_wide) %>% prettyNum(big.mark = ",")` participants.
+<!-- At least 1,500 participants took part per wave, and the sample is representative of the Austrian population. -->
+The panel study collected a large number of psychological and demographic variables.
+<!-- Being able to control for many confounding third variables, both stable and varying, together with the longitudinal design, this creates a unique opportunity  -->
+I explicitly aimed to investigate the causal effects of COVID-19 related social media use on well-being.
 
-## Defining Well-being and Media Use
+## Understanding Well-being and Media Use
 
-The underlying theories that guided the selection of variables for this study are the two-continua model of mental health [@greenspoonIntegrationSubjectiveWellbeing2001] and the hierarchical taxonomy of computer-mediated medation [@meierComputerMediatedCommunicationSocial2020].
-According to the two-continua model, mental health consists of two dimensions: psychopathology and well-being.
+The underlying theories that guided the selection of variables for my analysis are the two-continua model of mental health [@greenspoonIntegrationSubjectiveWellbeing2001] and the hierarchical taxonomy of computer-mediated communication [@meierComputermediatedCommunicationSocial2020a].
+According to the two-continua model, mental health consists of (a) psychopathology and (b) well-being.
 <!-- Because the aim of this study is better to understand typical users and everyday contexts, my focus will be on well-being. -->
-Well-being can be further differentiated into subjective and psychological well-being [@dienerAdvancesOpenQuestions2018].
-Whereas subjective well-being emphasizes hedonic aspects such as happiness and joy, psychological well-being focuses on eudaimonic aspects like fulfillment and meaning.
+Well-being can be differentiated into subjective and psychological well-being [@dienerAdvancesOpenQuestions2018].
+Whereas subjective well-being emphasizes hedonic aspects such as happiness and joy, psychological well-being addresses eudaimonic aspects such as fulfillment and meaning.
 Subjective well-being is primarily about achieving positive affect and avoiding negative affect.
 One of the most prominent indicators of well-being is life satisfaction.
-In my view, life satisfaction should be considered a meta concept that combines psychological and subjective well-being, because it represents a general appraisal of one's life.
+In my view, because it represents a general appraisal of one's life, life satisfaction is best thought of as a meta concept that combines psychological and subjective well-being.
 Notably, life satisfaction is stable and fluctuates only little, whereas it's the exact opposite for affect [@dienlinImpactDigitalTechnology2020].
 <!-- Everyone can feel bad on any given day, but whether or not one feels generally comfortable in one's life rather changes across years. -->
-To capture well-being in this study, I thus build on life satisfaction, positive affect, and negative affect.
+To capture well-being in this study I thus build on life satisfaction, positive affect, and negative affect.
 Together, this should provide an encompassing perspective on potential media effects.
@@
 The hierarchical taxonomy of computer-mediated communication differentiates six levels of how people engage with digital technology.
-First, the device (e.g., smartphone); second, the type of application (e.g., social networking site); third, the branded application (e.g., Twitter); fourth, the feature (e.g., status post); fifth, the interaction (e.g., one-to-many); sixth, the message (e.g., content) [@meierComputerMediatedCommunicationSocial2020].
-Whereas the first four levels are focused on the _channel_, the last two address the _type_ of communication.
-To measure social media use for consumption of COVID-19 related news and topics, I here employ both the channel and the communication perspective, which together should provide a more nuanced understanding of communication.
+First, the device (e.g., smartphone); second, the type of application (e.g., social networking site); third, the branded application (e.g., Twitter); fourth, the feature (e.g., status post); fifth, the interaction (e.g., one-to-many); and sixth, the message (e.g., content) [@meierComputermediatedCommunicationSocial2020a].
+Whereas the first four levels focus on the communication _channel_, the last two address the communication _type_.
+To measure social media use for the consumption of COVID-19 related news and topics, I here employ both the channel and the type of communication perspective, which together provides a nuanced understanding of communication.
 
-First, I will investigate how different types of communication affect well-being.
-Specifically, I will differentiate between active and passive use.
-I will distinguish (a) _reading_ COVID-19 related social media news (passive), _posting_ content regarding COVID (active), and _liking and sharing_ COVID-19 related posts (both active and passive).
-Second, I will analyze also how using the most prominent branded applications affect well-being, and whether this effect changes across applications.
-According to @meierComputerMediatedCommunicationSocial2020, branded apps are separate entities with different effects. 
-Therefore, Twitter might have a different effect than WhatsApp because of their different affordances.
-For example, @waterlooNormsOnlineExpressions2018 found that it is more adequate to express negative emotions on WhatsApp than on Twitter or on Instagram. 
+First, I investigate how different types of communication affect well-being.
+Specifically, I differentiate between active and passive use.
+I distinguish (a) _reading_ (passive), (b) _posting_ (active), and (c) _liking and sharing_ COVID-19 related posts (both active and passive).
+Second, I analyze how using the most prominent branded applications affects well-being, and whether this effect changes across applications.
+Branded apps are separate entities with potentially divergent effects. 
+Twitter might have a different effect as compared to WhatsApp because of their respective affordances.
+For example, @waterlooNormsOnlineExpressions2018 found that it's more adequate to express negative emotions on WhatsApp than on Twitter or on Instagram. 
 The branded applications investigated here are Facebook, Twitter, Instagram, WhatsApp, and YouTube.
-Worth noting, this study is not about _general_ social media during times of COVID, but on social media use _focused_ on COVID-19 related news and interactions. 
-Examples of such media use include posting about the pandemic or retweeting COVID-19 related posts.
+Worth noting, this study is not about _general_ social media use during times of COVID, but on social media use _focused_ on COVID-19 related content. 
+Examples of such media use include posting thoughts about the pandemic or retweeting COVID-19 related news.
 
@@
 
-So far, there is only little empirical research on COVID-19 related social media use and well-being.
+So far, there is only little empirical research on COVID-19 related social media use on well-being.
 In their study on the relations between media use and mental health during the pandemic, @bendauAssociationsCOVID19Related2021 found that people who used social media as a primary source of information reported on average "significantly more unspecific anxiety and depression [] and significantly more specific COVID-19 related anxiety symptoms" (p.
 288).
-Hence, this might hint at potential negative effects of social media as news use.
-However, note that this finding comes from a between-person relation stemming from cross-sectional data (see below).
-We, therefore, don't know whether the differences in mental health and well-being are due to social media use or other third variables, such as age, health, or education.
-@edenMediaCopingCOVID192020 analyzed the media use of 425 US college students during the first wave of the pandemic, and found both positive and negative relations with well-being.
+@edenMediaCopingCOVID192020 analyzed the media use of 425 US college students during the first wave of the pandemic, finding both positive and negative relations with well-being.
 In a sample of 312 respondents collected via Amazon Mechanical Turk, @choiMediatedCommunicationMatters2021 reported that people who used media to attain information were more lonely and less satisfied with their lives.
 @stainbackCOVID1924News2020 analyzed a large-scale study with 11,537 respondents from the US and found that increased COVID-19 media consumption was related to more psychological distress.
+A four-wave panel study with 384 young adults from the U.S.
+analyzed the effects of general digital technology use---objectively measured via screenshots of screen-time applications---on mental health, separating within- and between-person relations [@sewallObjectivelyMeasuredDigital2021].
+The results showed that digital technology did not have any significant effects on mental health [for a similar study with comparable results, see @bradleyStressMoodSmartphone2021].
+Together, the literature is mixed, with a slight focus on the negative effects of social media as news use [see also @dornemannHowGoodBad2021; @liuRelationOfficialWhatsAppdistributed2020; @riehmAssociationsMediaExposure2020].
+However, note that all of these findings represent between-person relations stemming from cross-sectional data (see below).
+We therefore don't know whether the differences in mental health and well-being are due to social media use or due to other third variables, such as age, health, employment, or education.
 
-On the other hand, the question of whether and how _general_ social media use affect well-being is well-researched.
+The question of whether and how social media use affects well-being _in general_, on the other hand, is well-researched.
 This also holds true for the different types of communication such as active or passive use.
-A meta review (i.e., an analysis of meta-analyses), found that the relation between social media use and well-being is likely in the negative spectrum but very small, potentially too small to matter [@meierComputerMediatedCommunicationSocial2020].
-What determines whether an effect is considered trivial or small?
-If we refer to standardized effect sizes, according to @cohenPowerPrimer1992 small effect sizes start at _r_ = .10.
-And indeed, several if not most of the current meta-analyses find effect sizes below that threshold [@huangTimeSpentSocial2017; @meierComputerMediatedCommunicationSocial2020; @fergusonThisMetaanalysisScreen2021].
+A meta review (i.e., an analysis of meta-analyses) found that the relation between social media use and well-being is likely in the negative spectrum but very small [@meierComputermediatedCommunicationSocial2020a]---potentially too small to matter.
+What determines whether or not an effect is considered small or trivial?
+As a starting point, we could refer to standardized effect sizes.
+According to @cohenPowerPrimer1992, small effect sizes start at _r_ = .10.
+And indeed, several if not most of the current meta-analyses find effect sizes below that threshold [@huangTimeSpentSocial2017; @meierComputermediatedCommunicationSocial2020a; @fergusonThisMetaanalysisScreen2021].
 
-These overviews are well aligned with several individual new studies that employed advanced methods [@orbenSocialMediaEnduring2019; @przybylskiDoesTakingShort2021a; @schemerImpactInternetSocial2021; @kerestesAdolescentsOnlineSocial2020].
-For example, @beyensSocialMediaUse2021 reported that although for some users (roughly one quarter) the effects were negative, for almost the same number of users they were positive, while for the majority the effects were neutral.
-In conclusion, most effects are somewhere between trivial and small.
-I therefore think it's most convincing to expect trivial to small effects also in the case of COVID-19 related social media use.
+These overviews are well aligned with several individual studies employing advanced methods [@orbenSocialMediaEnduring2019; @przybylskiDoesTakingShort2021a; @schemerImpactInternetSocial2021; @kerestesAdolescentsOnlineSocial2020].
+For example, @beyensSocialMediaUse2021 reported that although for some users (roughly one quarter) the effects of social media use on well-being were negative, for almost the same number of users they were positive, while for the rest the effects were neutral.
+In conclusion, most effects are likely somewhere between trivial and small.
+I therefore expect that also in the case of COVID-19 related social media use effects will be trivial to small.
 
-From a theoretical perspective, how could we explain how COVID-19 related social media use might affect well-being?
-In what follows, I outline potential arguments as to why the effect might be positive or negative, direct or indirect. 
-In advance, there does not seem to be a clear winner, and it's likely that both positive and negative effects are equally strong.
+From a theoretical perspective, how could we explain whether COVID-19 related social media use might affect well-being?
+In what follows, I outline potential arguments as to why the effect might be positive or negative, direct or indirect, or nonexistent.
+In advance, there doesn't seem to be a clear winner, and it's likely that both positive and negative effects cancel each other out.
 
-First, one could assume a _direct_ negative effect on well-being, and especially on positive or negative affect, which is more volatile and fluctuating.
-Dangers, inequalities, corruption---these were the headlines across many countries worldwide.
-If one learns about such things, the initial reaction might be shock, fear, or dismay.
-Repeatedly consuming such news might be depressing, perhaps even changing some general perspectives on life, without any further mediating processes.
+First, one could assume a _direct_ negative effect on well-being, and especially on positive or negative affect, which are more volatile and fluctuating.
+Dangers, inequalities, corruption---these were the headlines during the pandemic across many countries worldwide.
+If one learns about such events, the initial reaction might be shock, fear, or dismay.
+Consuming such news can be depressing [@dornemannHowGoodBad2021], perhaps even changing some general perspectives on life.
 <!-- So, just like being hit by a hammer hurts and we don't need any "mediating mechanism", this could be the case here as well. -->
-That said, because not all news was negative, and because many people showed solidarity and compassion, there was also positive and potentially uplifting news, potentially compensating for the negative news.
+That said, because not all news was negative, and because many people showed solidarity and compassion, there was also positive and uplifting content, potentially compensating for the negative effects [@dornemannHowGoodBad2021].
 <!-- However, in light of a worldwide pandemic with millions of deaths, the negative direct effect seems more plausible. -->
@@
 For example, during the pandemic social media use was at an all-time high in the US [@statistaAverageDailyTime2021].
-As has been expressed by many before, it is most likely that moderate social media use is not detrimental [@orbenTeenagersScreensSocial2020].
-Overuse, however, is more critical, and several studies have shown more pronounced negative effects for extreme users [@przybylskiLargescaleTestGoldilocks2017].
-To explain, overuse likely impairs well-being if it replaces other meaningful or functional activities such as meeting others, working, actively relaxing, or exercising.
-So if a society collectively overuses social media, there is potential for negative effects.
+In general, as has been expressed by many before, it is most likely that moderate social media use is not detrimental [@orbenTeenagersScreensSocial2020].
+Overuse, however, might be more critical, and several studies have shown more pronounced negative effects for extreme users [@przybylskiLargescaleTestGoldilocks2017].
+To explain, overuse likely impairs well-being if it replaces meaningful or functional activities such as meeting others, working, actively relaxing, or exercising.
+So if a society collectively overuses social media when doomscrolling, there is potential for negative effects.
 <!-- If overuse replaces such activities it's reasonable to assume that it's also detrimental. -->
 
-On the other hand, one can make the case that overuse can also be beneficial, especially in times of a pandemic, even if the use is mainly COVID-19 related.
-Exchanging COVID-19 related messages with friends via WhatsApp might replace the in-person contact one would have otherwise, but which is literally impossible at that time.
-In situations where meaningful and functional activities are explicitly or implicitly prohibited, using social media to exchange about COVID-19 related topics might not be the worst idea.
-In fact, given that nowadays a large number of experts, scientists, and politicians converse directly on social media, one can get first-hand information on the current developments.
+On the other hand, one can make the case that overuse might also be beneficial, especially in times of a pandemic---even if the use is mainly COVID-19 related.
+Exchanging COVID-19 related messages with friends via WhatsApp might replace the in-person contact one would have otherwise, but which is literally impossible at the time.
+In situations where meaningful and functional activities are prohibited, using social media to exchange about COVID-19 related topics might not be the worst idea.
+Besides, given that nowadays a large number of experts, scientists, and politicians converse directly on social media, one can get first-hand high quality information on current developments.
 <!-- On the other hand, there is of course also much disinformation, and "bingeing" on COVID-19 fake news might also pose risks for impaired well-being. -->
 
-Together, the strongest argument seems to be that _in general_ the effects of social media on well-being are, on average, small at best.
-Because this study looks at only _one part_ of social media use---namely, COVID-19 related interactions---it is even more focused, and the overall potential of the effects should diminish even further.
+Together, the strongest argument to me is that _in general_ the effects of social media on well-being are, on average, small at best.
+Because this study only looks at _one part_ of social media use---namely, COVID-19 related interactions---it is very focused, diminishing the overall potential of the effects even further.
 Whether or not using social media for COVID-19 related aspects is detrimental during a pandemic is also not entirely clear.
 <!-- At least, there does not seem to be a case where we should expect a clear deviation to the negative, despite anecdotes of doomscrolling. -->
-Therefore, I expect to find that COVID-19 related communication on social media will not affect well-being in a meaningful or relevant way.
+Therefore, I expect that COVID-19 related communication on social media doesn't affect well-being in a meaningful or relevant way.
 
@@
 Testing this hypothesis, however, is not trivial.
-First, in contrast to most hypotheses typically posited in the social sciences, it implicitly contains an effect size, a so-called smallest effect size of interest (SESOI).
+First, in contrast to most hypotheses typically posited in the social sciences it implicitly contains an effect size, a so-called smallest effect size of interest (SESOI).
 Effectively testing this hypothesis necessitates defining what's considered a "trivial effect size" and what's not.
@@
 However, standardized effect sizes should only be a first step toward evaluating an effect's relevance [@baguleyStandardizedSimpleEffect2009].
-Standardized effect sizes are determined by a sample's variance.[^cohensd]
-However, this is problematic: The question of whether or not social media use affects a particular person in a relevant way should not depend on the variance in the sample in which that person's data were collected.
+Standardized effect sizes are determined by a sample's variance,[^cohensd] which is problematic: 
+The question of whether or not social media use affects a particular person in a relevant way should not depend on the variance in the sample in which that person's data were collected.
 Instead, it should depend on absolute criteria.
-What could be a minimally interesting, a nontrivial effect?
+
+What could be a minimally interesting, nontrivial effect?
 Because this is a normative and ultimately philosophical question, there can never be a clear, single, or unanimous answer. 
-In the end, it is a personal question.
-However, it is still both necessary and fruitful to try to answer it. 
-I suggest the following SESOI for this research question:
+<!-- In the end, it is a personal question. -->
+However, it is still necessary and helpful to try provide such a plausible benchmark. 
+I therefore suggest the following SESOI for this research question:
 
@@
 
-If a heavy user of COVID-19 related social media news suddenly _completely stops_ using social media, this should have a _noticeable_ impact on their overall well-being.
+> SESOI: If a heavy user of COVID-19 related social media news suddenly _stops_ using social media altogether, this should have a _noticeable_ impact on their overall well-being.
 
-What would this mean practically and how could it be operationalized?
+What does this mean practically and how can it be operationalized?
 In this study, COVID-19 related social media use was measured on a 5-point scale, ranging from 1 = _never_ to 5 = _several times a day_. 
-Thus, the example from above would imply that a change of four units in social media use (e.g., a complete stop) should correspond to a noticeable change in well-being.
+Thus, a change of four units in social media use (e.g., a complete stop) should correspond to a noticeable change in well-being.
 But what's a noticeable change in well-being?
-According to @normanInterpretationChangesHealthrelated2003, people can reliably differentiate between _seven_ levels of satisfaction with health.
-So we could state that a four unit change in social media use should result in a one unit change in life satisfaction, if satisfaction is measured on a 7-point scale.
-Statistically, in a regression _b_ estimates the change in the dependent variable if the independent variable increases by one point.
-Transferred to our example, we would hence expect a SESOI of _b_ = 1 / 4 = .25.
-
-In this study, life satisfaction was measured on an 11-point scale, which is more than the seven degrees people can reliably differentiate. 
-We hence need to transform a 1-point change on a 7-point scale to its equivalent on an 11-point scale, which is 11 / 7 = `r 11/7`.
-Hence, we now expect our effect to be at least as large as _b_ = .25 * `r 11/7` = `r 11/7*.25`.
-For affect, which was measured on a 5-point scale, our SESOI would therefore be _b_ = 1/4 * 5/7 = `r 5 / 7 * .25`.
-Because we are agnostic as to whether the effects are positively or negatively nontrivial, the null region will include negative and positive effects (in Bayesian terms, this is called the region of practical equivalence [ROPE]).
-In order not to exaggerate precision, and to be a bit less conservative, these numbers will be reduced to nearby thresholds. 
-Note that other researchers also decreased or recommended decreasing thresholds for effect sizes when analyzing within-person or cumulative effects (@beyensSocialMediaUse2021; @funderEvaluatingEffectSize2019).
-Together, this leads to an indifference region ranging from _b_ = -.30 to _b_ = .30 for life satisfaction, and _b_ = -.15 to _b_ = .15 for positive and negative affect.
+According to @normanInterpretationChangesHealthrelated2003, people can reliably distinguish _seven_ levels of satisfaction with health.
+So if satisfaction is measured on a 7-point scale, we would state that a four unit change in social media use should result in a one unit change in life satisfaction. 
+(For more information, see Methods section "Inference Criteria.")
 
@@
 The hypothesis explicitly states a causal effect.
-In non-experimental studies, longitudinal designs help investigate causality.
-Using longitudinal designs alone, however, is not sufficient for correct causal statements.
-In addition, we also need to control for third variables, and importantly also for _varying_ third variables.
+In non-experimental studies, longitudinal designs can help investigate causality.
+Using longitudinal designs alone, however, is not sufficient for establishing correct causal statements [@rohrerTheseAreNot2021].
+In addition, we for example also need to control for confounding third variables, and importantly also for _varying_ third variables.
 
-Consider the following example.
-Imagine that a person suddenly starts using social media much more than usual, and then at the same time also becomes less satisfied with their life.
-After some time, use recedes again, whereas life satisfaction returns to prior levels.
-If this happens to several people at the same time, in a longitudinal study we would then observe a significant effect of social media use on life satisfaction.
-However, it could also be the case that during the study there was a major exogenous event (say, a pandemic), which is why a large part of the working population lost their jobs.
-Hence, the causal effect reported above was likely confounded, because in reality it was the pandemic that caused both social media use to rise and life satisfaction to plummet.
+To illustrate, consider the following example.
+Imagine that a person suddenly starts using social media much more than usual, and then after some time  becomes less satisfied with their life.
+Eventually, use and life satisfaction return to prior levels.
+If this happens to several people at the same time, in a longitudinal study we could then observe a significant effect of social media use on life satisfaction.
+However, it could also be the case that during the study there was a major exogenous event (say, a pandemic), which caused large parts of the working population to loose their jobs.
+Hence, the causal effect reported above was confounded, because in reality it was the pandemic that caused both social media use to rise and life satisfaction to go down.
 
-Thus, only if we can control for _all_ potential varying third variables, can we correctly estimate causality without bias [@rohrerThinkingClearlyCorrelations2018].
-Obviously, we can never be entirely sure that we have included all varying covariates, which makes absolute statements regarding causality impossible.
-In addition, when determining the overall causal effect, we should _not_ control for mediating variables [@rohrerThinkingClearlyCorrelations2018].
-Doing so would bias and distort our assessment of the role of social media use.
-Complicating matters, often it is unclear if a variable is a mediator or a confounder.[^collider]
-However, when controlling for relevant variables (that aren't mediators), we can be much more certain that we measured causality correctly.
-The aim should therefore be to collect as many _relevant_ varying and nonvarying third variables as possible, while knowing that absolute certainty regarding causality cannot be reached.
+Thus, only when controlling for _all_ relevant confounders, we can correctly estimate causality without bias [@rohrerThinkingClearlyCorrelations2018].
+Obviously, we can never be entirely sure to have included all confounders, which makes absolute statements regarding causality virtually impossible.
+In addition, when determining the overall causal effect, we need to make sure _not_ to control for mediating variables [@rohrerThinkingClearlyCorrelations2018], for doing so would bias our assessment of the causal effect.
+Complicating matters further, it is often unclear if a variable is a mediator or a confounder.[^collider]
+However, despite all these caveats, when controlling for relevant variables (that aren't mediators), we can be much more certain that we measured causality correctly.
+The aim should therefore be to collect as many varying and nonvarying confounders as possible (which I believe is seldom done in our field), while knowing that absolute certainty regarding causality cannot be reached.
 
@@
 
-When looking for suitable control candidates, ideally we select variables that affect both media use and well-being. 
+When searching for suitable candidates for confounders, we should look for variables that affect both media use and well-being. 
 Controlling for these factors isolates the actual effect of social media use on well-being.
-We can also control for variables that affect _only_ well-being. 
-However, in doing so not much is gained or lost because the effects of social media use would remain virtually the same [@klinePrinciplesPracticeStructural2016].[^mcelreath]
+We can also control for variables that affect only social media use or well-being. 
+However, in doing so not much is gained or lost, because the effects of social media use would remain virtually the same [@klinePrinciplesPracticeStructural2016; but see @mcelreathYesterdayClass2021].
 
-[^mcelreath]: But see @mcelreathYesterdayClass2021.
-
-In this study, I hence plan to control for the following variables, which either have already been shown to affect both social media use and well-being, or which are very likely to do so, and which are likely not mediators:
+In this study, I hence plan to control for the following variables, which either have already been shown to affect both social media use and well-being or which are likely to do so, and which also aren't mediators:
 <!-- (I'll additionally include variables that likely affect only well-being, also to obtain a comparison benchmark for social media effects): -->
 gender, age, education, Austria country of birth, Austria country of birth of parents, text-based news consumption, video-based news consumption, residency Vienna, household size, health, living space, access to garden, access to balcony, employment, work hours per week, being in home-office, household income, outdoor activities, satisfaction with democracy, disposition to take risks, and locus of control.
-I will not control for variables such as trust in institutions or media, because these variables might be influenced by social media use to a meaningful extent.
+I will not control for variables such as trust in institutions or trust in media, because these variables might be influenced by social media use to a meaningful extent.
 
-Next to including covariates, it is now increasingly understood that causal effects need to be analyzed from an internal, within-person perspective.
-If a specific person changes their media diet, we need to measure how this affects _their_ well-being.
-Between-person comparison from cross-sectional data, where participants are interviewed only once, cannot provide such insights.
-In this study, I will hence differentiate between within-person effects and between-person relations. 
-As explicated in the hypothesis, to test the hypothesis I will thus consider only the within-person effects.
+Next to including covariates, it's now increasingly understood that causal effects should be analyzed from an internal, within-person perspective [@hamakerWhyResearchersShould2014].
+If a specific person changes their media diet, we need to measure how this behavior affects _their own_ well-being.
+Between-person comparisons from cross-sectional data, where participants are interviewed only once, cannot provide such insights.
+In this study, I hence differentiate between-person relations from within-person effects. 
+And as explicated above, to test the hypothesis I thus consider only the within-person effects.
 
 Finally, one precondition of causality is temporal order. 
-The cause needs to precede the outcome.
-Finding the right interval between causes and effects is crucial.
-For example, if we want to measure the effect of alcohol consumption on driving performance, it makes a big difference if driving performance is measured one minute, one hour, one day, or one week after consumption.
-Finding the right interval is difficult.
-If variables are more stable, longer intervals are needed, and if they fluctuate a lot, shorter intervals are necessary.
-In the case of well-being, to analyze affect we need shorter intervals, while for life satisfaction longer ones.
-Still, choosing the right interval is challenging, because especially short intervals are practically hard to implement and often require experience sampling measures (also known as in situ measurement or ambulant assessment).
+The cause needs to precede the effect.
+Finding the right interval between cause and effect is crucial.
+For example, if we want to understand the effect of alcohol consumption on driving performance, it makes a big difference if driving performance is measured one minute, one hour, one day, or one week after consumption.
+<!-- Finding the right interval is difficult. -->
+If variables are stable, longer intervals are needed; if they fluctuate, shorter intervals.
+In the case of well-being, we need shorter intervals for affect and longer ones for life satisfaction.
+Still, choosing the right interval is challenging, because especially short intervals are hard to implement in practice and often require advanced methods such as experience sampling (also known as in situ measurement or ambulant assessment) [@schnauber-stockmannMobileDevicesTools2020].
 
-In this study, I will adopt an intermediate perspective.
-I will analyze if, when a person changes their social media diet, are there simultaneous changes in well-being?
-When additionally controlling for both stable and varying covariates, we can then be more sure that the effect is indeed causal.
-Similar approaches were implemented already by several other studies [@johannesNoEffectDifferent2021; @scharkowHowSocialNetwork2020] and they considered one of the best practices to analyze causality [@bellFixedRandomEffects2019].
+In this study, I therefore adopt an intermediate perspective.
+I analyze if when a person changes their social media diet, are there _simultaneous_ changes in their well-being?
+When additionally controlling for both stable and varying confounders, we can then be more sure that the effect is indeed causal.
+<!-- Similar approaches were already implemented by several other studies [@johannesNoEffectDifferent2021; @scharkowHowSocialNetwork2020] and they are considered a best practice to analyze causality [@bellFixedRandomEffects2019].
+-->
 
@@
 
-This section describes the preregistration and how I determined the sample size, data exclusions, the analyses, and all measures in the study.
+In this section I describe the preregistration and how I determined the sample size, data exclusions, the analyses, and all measures in the study.
 
@@
 The (anonymous) preregistration can be accessed here: https://osf.io/87b24/?view_only=b2289b6fec214fa88ee75a18d45c18f3.
-Because in this study I analyze data from an already existing large-scale data set, all of these steps were preregistered prior to accessing the data.
+Because in this study I analyzed data from an already existing large-scale data set, all of these steps were done prior to accessing the data.
 The preregistration was designed on the basis of the panel documentation online [@kittelAustrianCoronaPanel2020].
-In some cases the analyses could not be executed as originally planned, for example because some properties of the variables became apparent only upon data analysis.
-The most relevant deviations are reported below, and a complete list of all changes can be found in the online companion website (https://xmtra.github.io/Austrian_Corona_Panel_Project/index.html).
+In some cases I couldn't execute the analyses as I had originally planned, for example because some properties of the variables only became apparent when inspecting the actual data.
+The most relevant deviations are reported below, and a complete list of all changes can be found in the online [companion website](https://XMtRA.github.io/Austrian_Corona_Panel_Project) (https://XMtRA.github.io/Austrian_Corona_Panel_Project).
 
@@
 The study was conducted between March 2020 and October 2021.
-It contains 26 waves, and at the time of writing, the first 24 waves were available for download. 
+It contains 26 waves, and at the time of writing the first 24 waves were available for download. 
 Each wave consists of at least 1,500 respondents.
@@
 All respondents needed to have access to the internet (via computer or mobile devices such as smartphones or tablets).
-They were sampled from a pre-existing online access panel provided by Marketagent, Austria.
+They were sampled from a pre-existing online access panel provided by the company Marketagent, Austria.
 Respondents were asked and incentivized with 180 credit points to participate in each wave of the panel.
@@
 The sample was representative of the Austrian population in terms of age, gender, region/state, municipality size, and educational level. 
-In order to participate in the study, the respondents needed to be Austrian residents and had to be at least 14 years old.
+In order to participate in the study, the respondents needed to be Austrian residents and had to be at least 14 years of age. 
+The study received IRB approval from the University of Vienna.
 The average age was `r mean(2021 - d_wide$year_birth, na.rm = T) %>% round(0)` years, `r mean(d_wide$male, na.rm = T) %>% round(., 2) * 100` percent were male, `r (d_wide %>% filter(.$edu_fac %in% c("State college", "Bachelor", "Master", "PhD"))) %>% nrow() %>% divide_by(sum(table(d_wide$edu_fac))) %>% multiply_by(100) %>% round(0)` percent had a University degree, and `r (table(d_wide$employment_fac)["Unemployed"] / sum(table(d_wide$employment_fac))) %>% unname() %>% multiply_by(100) %>% round(., digits = 0)` percent were currently unemployed.
-<!-- The study received IRB approval from the University of Vienna. -->
+
+## Inference Criteria
 
 Because the data were analyzed post-hoc, no a-priori sample size planning on the basis of power analyses was conducted.
-The sample is very large, and it is hence likely well-equipped reliably to detect also small effects.
+The sample is large, and it is hence well-equipped reliably to detect also small effects.
 <!-- , which is why no exact post hoc power analysis were conducted. -->
-In addition, because such large samples easily generate significant _p_-values even for very small effects, it helps that to test hypothesis this study uses a smallest effect size of interest-approach.
-To test the hypotheses, I will use the interval testing approach as proposed by @dienesUsingBayesGet2014.
-On the basis of the SESOI, I will define a null region.
-For well-being, the null region will be between _b_ = -.30 and _b_ = .30.
-For example, if the 95% confidence interval lies completely within the null-region (e.g., _b_ = .15, [95% CI: .05, .25]), the hypothesis that the effect is only trivial is supported.
-If the effects interval and the null region overlap (e.g., _b_ = .25, [95% CI: .15, .35]), the hypothesis is not supported and the results are considered inconclusive, whereas a meaningful negative effect is rejected.
-If the confidence falls completely outside of the null-region (e.g., _b_ = .40, [95% CI: .35, .45]), the hypothesis is rejected and the existence of a meaningful positive effect is supported.
+In addition, because such large samples easily generate significant _p_-values even for very small effects, it helps that the hypotheses were tested with a smallest effect size of interest-approach.
+To this end, I adopted the interval testing approach as proposed by @dienesUsingBayesGet2014.
+On the basis of the SESOI, I then defined a null region.
+In what follows, I explain how I determined the SESOI and the null region.
+
+In this study, life satisfaction was measured on an 11-point scale.
+If people can reliably differentiate 7 levels as mentioned above, this corresponds to 11 / 7 = `r round(11/7, 2)` unit change on an 11-point scale.
+Hence, a four-point change in media use (e.g., a complete stop) should result in a `r round(11/7, 2)`-point change in life satisfaction.
+In a statistical regression analysis, _b_ estimates the change in the dependent variable if the independent variable increases by one point.
+We would therefore expect a SESOI of _b_ = `r round(11/7, 2)` / 4 = `r round(11/7*.25, 2)`.
+For affect, which was measured on a 5-point scale, our SESOI would be _b_ = `r round(5 / 7, 2)` / 4 = `r round(5 / 7 * .25, 2)`.
+Because we're agnostic as to whether the effects are positive or negative, the null region includes negative and positive effects.
+Finally, in order not to exaggerate precision and to be less conservative, these numbers are reduced to nearby thresholds.[^rounding] 
+Together, this leads to a null region ranging from _b_ = -.30 to _b_ = .30 for life satisfaction, and _b_ = -.15 to _b_ = .15 for positive and negative affect.
+
+[^rounding]: Note that other researchers also decreased or recommended decreasing thresholds for effect sizes when analyzing within-person or cumulative effects [@beyensSocialMediaUse2021; @funderEvaluatingEffectSize2019].
+
+<!-- As explained above, for well-being the null region was between _b_ = -.30 and _b_ = .30 (i.e., _b_ = |.30|). -->
+Let's briefly illustrate what this means in practice.
+If the 95% confidence interval falls completely within the null-region (e.g., _b_ = .20, [95% CI: .15, .25]), the hypothesis that the effect is trivial is supported.
+If the confidence interval and the null region overlap (e.g., _b_ = .30, [95% CI: .25, .35]), the hypothesis is not supported and the results are considered inconclusive, while a meaningful negative effect is rejected.
+If the confidence interval falls completely outside of the null-region (e.g., _b_ = .40, [95% CI: .35, .45]), the hypothesis is rejected and the existence of a meaningful positive effect is supported.
 
@@
 <!-- X clear cases were removed. -->
-<!-- Respondents who answered less than 50% of all questions were removed. -->
-<!-- The remaining missing responses were imputed using predictive mean matching. -->
-<!-- We will remove respondents with unrealistically fast responses, namely below three standard deviations of the medium response time.  -->
 
@@
 Three models were run, one for each dependent variable.
-Predictors included all types and channels of social media communication separated into within and between-person factors, as well as stable and varying covariates.
+The data were hierarchical, and responses were separately nested in participants and waves (i.e., participants and waves were implemented as random effects).
+Nesting in participants allowed to separate between-person relations from within-person effects.
+Nesting in waves allowed to control for general exogenous developments, such as general decreases in well-being in the population, for example due to lockdown measures.
+Thus, there was no need additionally to control for specific phases or measures of the lockdown.
+Predictors were modeled as fixed effects.
+They included social media communication types and channels, separated into within and between-person factors, as well as stable and varying covariates.
 All predictors were included simultaneously and in each of the three models.
-The data were hierarchical, and responses were separately nested in participants and waves (put differently, participants and waves were implemented as random effects).
-Nesting in participants allowed to separate between-person relations from within-person effects.
-Nesting in waves allowed to control for general exogenous developments, such as general decrease in well-being in the population, for example due to lockdown measures.
-Thus, there was no need additionally to control for specific phases or measures of the lockdown.
 
-<!-- The factorial validity of the scales were tested with confirmatory factor analyses (CFA).  -->
-<!-- If Mardia’s test showed that the assumption of multivariate normality was violated, the more robust Satorra-Bentler scaled and mean-adjusted test statistic (MLM) was used as estimator.  -->
-<!-- To avoid overfitting, scales were tested on more liberal fit criteria (CFI > .90, TLI > .90, RMSEA <. .10, SRMR < .10) [@klinePrinciplesPracticeStructural2016]. -->
-<!-- Because REWB-models cannot model scales as latent variables, to increase precision factor scores exported from the CFAs were used. -->
+The factorial validity of the scales were tested with confirmatory factor analyses (CFA).
+Because Mardia’s test showed that the assumption of multivariate normality was violated, I used the more robust Satorra-Bentler scaled and mean-adjusted test statistic (MLM) as estimator.
+To avoid overfitting, I tested the scales on more liberal fit criteria (CFI > .90, TLI > .90, RMSEA <. .10, SRMR < .10) [@klinePrinciplesPracticeStructural2016].
+Finally, REWB-models cannot model latent variables.
+To increase precision, I therefore exported factor scores from the CFAs for positive and negative affect.
+Respondents who answered less than 50% of all questions were removed.
+The remaining missing responses were imputed using predictive mean matching.
 
-For more information on the analyses, for example a complete documentation of the models and their results, see [companion website](https://xmtra.github.io/Austrian_Corona_Panel_Project/index.html).
+For more information on the analyses, a complete documentation of the models and results, see [companion website](https://XMtRA.github.io/Austrian_Corona_Panel_Project).
 
@@
 
-In what follows, I briefly list all variables that were collected.
+In what follows, I list all the variables that I analyzed.
 For the variables' means, range, and variance, see Table \@ref(tab:descriptives).
-For a complete list of all items and item characteristics, see [companion website](https://xmtra.github.io/Austrian_Corona_Panel_Project/index.html).
+For a complete list of all items and item characteristics, see [companion website](https://XMtRA.github.io/Austrian_Corona_Panel_Project).
 
@@
 The response options were 1 (_never_), 2 (_on some days_), 3 (_several times per week_), 4 (_almost every day_), and 5 (_daily_).
+The scale showed good factorial fit, `r fit_txt(cfa_aff_pos)`.
+Reliability was high, $\omega$ = `r rel_aff_pos %>% my_round("std")`.
 
@@
 The response options were 1 (_never_), 2 (_on some days_), 3 (_several times per week_), 4 (_almost every day_), and 5 (_daily_).
+The scale showed good factorial fit, `r fit_txt(cfa_aff_neg)`.
+Reliability was high, $\omega$ = `r rel_aff_neg %>% my_round("std")`.
 
@@
 
-COVID-19 related social media use focused on types of communication was measured with three variables (a) reading, (b) liking and sharing, and (c) posting.
+COVID-19 related social media use focused on communication types was measured with the three dimensions of (a) reading, (b) liking and sharing, and (c) posting.
 The general introductory question was "How often during the last week have you engaged in the following activities on social media?"
-The three items read as follows: 
-"Reading the posts of others with content on the Coronavirus"; "When seeing posts on the Coronavirus, I clicked 'like', 'share' or 'retweet'"; "I myself wrote posts on the Coronavirus on social media."
+The three items were "Reading the posts of others with content on the Coronavirus", "When seeing posts on the Coronavirus, I clicked 'like', 'share' or 'retweet'", "I myself wrote posts on the Coronavirus on social media."
 Answer options were 1 (_several times per day_), 2 (_daily_), 3 (_several times per week_), 4 (_weekly_), 5 (_never_).
@@
 Social media use was measured for all participants on waves 1, 2, 8, 17, and 23. 
-Freshly recruited respondents always answered the questions on social media use.
+Freshly recruited respondents always answered all questions on social media use.
 
@@
 The effects of COVID-19 related social media use were controlled for the following stable variables: 
-(a) gender (answer options: female, male, diverse), (b) age, (c) education (ten options), (d) Austria country of birth (yes/no), (e) Austria parents' country of birth (no parent, one parent, both parents).
-I originally planned to implement other variables as varying covariates.
+(a) gender (female, male, diverse), (b) age, (c) education (ten options), (d) Austria country of birth (yes/no), (e) Austria parents' country of birth (no parent, one parent, both parents).
+I originally planned to implement additional variables as varying covariates.
 However, because they were not measured often enough or not at the time when social media use was measured, I implemented them as stable variables using their average values across all waves.
-This includes (a) text-based media news consumption (five degrees), (b) video-based media news consumption (five degrees), (c) residency is Vienna (yes/no), (d) self-reported physical health (five degrees), (e) living space (in squaremeter), (f) access to balcony (yes/no), (g) access to garden (yes/no), (h) employment (nine options), (i) disposition to take risks (eleven degrees), and (j) locus of control (five degrees).
-I controlled also for the following varying covariates: (a) five items measuring outdoor activities such as sport or meeting friends (five degrees), and (b) satisfaction with democracy (five degrees).
+This includes (a) text-based media news consumption (five degrees), (b) video-based media news consumption (five degrees), (c) residency is Vienna (yes/no), (d) self-reported physical health (five degrees), (e) living space (eleven options), (f) access to balcony (yes/no), (g) access to garden (yes/no), (h) employment (nine options), (i) disposition to take risks (eleven degrees), and (j) locus of control (five degrees).
+I also controlled for the following varying covariates: (a) five items measuring outdoor activities such as sport or meeting friends (five degrees), and (b) satisfaction with democracy (five degrees).
 Because it lead to too much attrition in the sample, I did not control for (a) household size, (b) work hours per week, (c) home office, (d) household income.
@@
 # Results
-## Preregistered Analyses
 
-When looking at the variables from a descriptive perspective, we see that all well-being measures did not change substantially across the different waves of data collection.
+First, when looking at the variables from a descriptive perspective, we see that all well-being measures did not change substantially across the different waves of data collection.
 COVID-19 related media use, however, decreased slightly at the beginning of the study and remained stable after approximately six waves.
 The initial decrease might be explained by the fact that the collection of data began at the end of March 2020, hence approximately three months after the pandemic began.
-It could be that after an initial uptick, COVID-19 related social media use was already declining at that time, returning to more normal levels.
+It could be that after an initial uptick, COVID-19 related social media use was already declining at the time, returning to more normal levels.
 
@@
 
-The study's hypothesis was that the effects of social media use on well-being would be trivial.
-Regarding the effects of different _types_ of communication---that is, reading vs.
+## Preregistered Analyses
+
+The study's main hypothesis was that the effects of social media use on well-being would be trivial.
+Regarding the effects of different communication _types_---that is, reading vs.
 sharing vs.
-posting---all within-person effects fell completely within the a-priori defined SESOIs (see Figure \@ref(fig:res-activity)).
+posting---all within-person effects fell completely within the a-priori defined null region (see Figure \@ref(fig:res-activity)).
 For example, respondents who used social media more frequently than usual to read about COVID-19 related topics did not show a simultaneous change in life satisfaction (_b_ = `r dat_fig_results_activity %>% filter(dv == "Life satisfaction" & iv == "Reading" & type == "within") %>% select(estimate) %>% round(2)` [95% CI `r dat_fig_results_activity %>% filter(dv == "Life satisfaction" & iv == "Reading" & type == "within") %>% select(conf.low) %>% round(2)`, `r dat_fig_results_activity %>% filter(dv == "Life satisfaction" & iv == "Reading" & type == "within") %>% select(conf.high) %>% round(2)`]).
-All confidence intervals include zero; hence, all effects were also statistically non-significant.
-<!-- Only one effect did not include zero. -->
-<!-- If people posted more about COVID-19 related aspects than they usually did, life satisfaction increased (_b_ = `r dat_fig_results_activity %>% filter(dv == "Life satisfaction" & iv == "Posting" & type == "within") %>% select(estimate) %>% round(2)` [95% CI `r dat_fig_results_activity %>% filter(dv == "Life satisfaction" & iv == "Posting" & type == "within") %>% select(conf.low) %>% round(2)`, `r dat_fig_results_activity %>% filter(dv == "Life satisfaction" & iv == "Posting" & type == "within") %>% select(conf.high) %>% round(2)`]).
--->
-<!-- However, because the effect did not fall outside of our null region, it's likely not large enough to be considered meaningful. -->
+All confidence intervals included zero; hence, all effects were also statistically non-significant.
 As a result, the hypothesis was supported for all COVID-19 related types of social media communication.
@@
 ```{r res-activity, fig.cap="The effects of various types of social media use on three indicators of well-being.
-Effects are controlled for a large number of covariates (see text).
-The SESOI was |0.30| for life satisfaction and |0.15| for affects.
-Hence, none of the reported effects are considered meaningful.", out.width = "\\textwidth", cache=F}
+The black estimates show the effects controlled for a large number of covariates (see text; preregistered); the grey estimates are without control variables (exploratory).
+The SESOI was b = |0.30| for life satisfaction and b = |0.15| for affect.
+Hence, all of the reported effects are not considered meaningful.", out.width = "\\textwidth", cache=F}
 knitr::include_graphics("figures/fig_results_activity.pdf")
@@
 
-Regarding between-person relations, about which no hypotheses were formulated, only one effect didn't include zero.
-<!-- First, respondents who across all waves used social media more frequently than others to read about COVID-19 reported higher levels of positive affect than others (_b_ = `r dat_fig_results_activity %>% filter(dv == "Positive affect" & iv == "Reading" & type == "between") %>% select(estimate) %>% round(2)` [95% CI `r dat_fig_results_activity %>% filter(dv == "Positive affect" & iv == "Reading" & type == "between") %>% select(conf.low) %>% round(2)`, `r dat_fig_results_activity %>% filter(dv == "Positive affect" & iv == "Reading" & type == "between") %>% select(conf.high) %>% round(2)`]).
--->
-<!-- However, note that this effect still fell completely within the null-region. -->
-<!-- Hence, although positive the effect was considered too small to matter practically. -->
-Respondents who across all waves used social media more frequently than others to write COVID-19 related posts reported higher levels of negative affect than others (_b_ = `r dat_fig_results_activity %>% filter(dv == "Negative affect" & iv == "Posting" & type == "between") %>% select(estimate) %>% round(2)` [95% CI `r dat_fig_results_activity %>% filter(dv == "Negative affect" & iv == "Posting" & type == "between") %>% select(conf.low) %>% round(2)`, `r dat_fig_results_activity %>% filter(dv == "Negative affect" & iv == "Posting" & type == "between") %>% select(conf.high) %>% round(2)`]).
-The effect was still completely inside of the null region, hence not considered large enough to be practically relevant.
+Regarding between-person relations, about which no hypotheses were formulated, only three effects did not include zero.
+Respondents who across all waves used social media more frequently than others to read about COVID-19 related posts reported slightly lower levels of positive affect than others (_b_ `r dat_fig_results_activity %>% filter(dv == "Positive affect" & iv == "Reading" & type == "between") %>% select(estimate) %>% my_round("b_txt")` [95% CI `r dat_fig_results_activity %>% filter(dv == "Positive affect" & iv == "Reading" & type == "between") %>% select(conf.high) %>% my_round("b")`, `r dat_fig_results_activity %>% filter(dv == "Positive affect" & iv == "Reading" & type == "between") %>% select(conf.low) %>% my_round("b")`]).
+Respondents who across all waves used social media more frequently than others to write COVID-19 related posts reported higher levels of negative affect than others (_b_ `r dat_fig_results_activity %>% filter(dv == "Negative affect" & iv == "Posting" & type == "between") %>% select(estimate) %>% my_round("b_txt")` [95% CI `r dat_fig_results_activity %>% filter(dv == "Negative affect" & iv == "Posting" & type == "between") %>% select(conf.high) %>% my_round("b")`, `r dat_fig_results_activity %>% filter(dv == "Negative affect" & iv == "Posting" & type == "between") %>% select(conf.low) %>% my_round("b")`]).
+Interestingly, respondents who across all waves used social media more frequently than others to write COVID-19 related posts also reported slightly higher levels of positive affect than others (_b_ `r dat_fig_results_activity %>% filter(dv == "Positive affect" & iv == "Posting" & type == "between") %>% select(estimate) %>% my_round("b_txt")` [95% CI `r dat_fig_results_activity %>% filter(dv == "Positive affect" & iv == "Posting" & type == "between") %>% select(conf.high) %>% my_round("b")`, `r dat_fig_results_activity %>% filter(dv == "Positive affect" & iv == "Posting" & type == "between") %>% select(conf.low) %>% my_round("b")`]).
+However, note that the effect were still completely inside of the null region, hence not large enough to be considered practically relevant.
+
+Note that when comparing the results with and without control variables, the results differed.
+For example, on the between-person level, one effect stopped being significant if controlled for additional variables.
+Actively posting on social media was significantly (though not meaningfully) related to decreased life satisfaction.
+However, when controlling for potential confounders, the effect became virtually zero (see Figure \@ref(fig:res-channels)).
 
 ```{r res-channels, fig.cap="The effects of using various social media applications on three indicators of well-being.
-Effects are controlled for a large number of covariates (see text).
-The SESOI was |0.30| for life satisfaction and |0.15| for affects.
-Hence, none of the reported effects are considered meaningful.", out.width = "\\textwidth", cache=F}
+The black estimates show the effects controlled for a large number of covariates (see text; preregistered); the grey estimates are without control variables (exploratory).
+The SESOI was b = |0.30| for life satisfaction and b = |0.15| for affect.
+Hence, all of the reported effects are not considered meaningful.", out.width = "\\textwidth", cache=F}
 knitr::include_graphics("figures/fig_results_channel.pdf")
@@
 
-Regarding the COVID-19 related use of social media _channels_, the results were very comparable (see Figure \@ref(fig:res-channels)).
+Regarding the COVID-19 related use of social media _channels_, the results were comparable (see Figure \@ref(fig:res-channels)).
 Changes in the frequency of using different social media channels to attain information regarding COVID-19 were unrelated to meaningful changes in well-being.
@@
 Only two effects differed substantially from zero.
-Respondents who used Instagram more frequently than usual to attain COVID-19 related news reported slightly higher simultaneous levels of life satisfaction than usual (_b_ = `r dat_fig_results_channel %>% filter(dv == "Life satisfaction" & iv == "Instagram" & type == "within") %>% select(estimate) %>% my_round("b")` [95% CI `r dat_fig_results_channel %>% filter(dv == "Life satisfaction" & iv == "Instagram" & type == "within") %>% select(conf.low) %>% my_round("b")`, `r dat_fig_results_channel %>% filter(dv == "Life satisfaction" & iv == "Instagram" & type == "within") %>% select(conf.high) %>% my_round("b")`]).
-Respondents who used Twitter more frequently than usual to attain COVID-19 related news reported somewhat lower simultaneous levels of life satisfaction than usual (_b_ = `r dat_fig_results_channel %>% filter(dv == "Life satisfaction" & iv == "Twitter" & type == "within") %>% select(estimate) %>% round(2)` [95% CI `r dat_fig_results_channel %>% filter(dv == "Life satisfaction" & iv == "Twitter" & type == "within") %>% select(conf.low) %>% round(2)`, `r dat_fig_results_channel %>% filter(dv == "Life satisfaction" & iv == "Twitter" & type == "within") %>% select(conf.high) %>% my_round("b")`]).
-However, both effects were still completely inside of the null region, hence not considered large enough to be meaningful.
+Respondents who used Instagram more frequently than usual to attain COVID-19 related news reported slightly _higher_ levels of life satisfaction than usual (_b_ = `r dat_fig_results_channel %>% filter(dv == "Life satisfaction" & iv == "Instagram" & type == "within") %>% select(estimate) %>% my_round("b")` [95% CI `r dat_fig_results_channel %>% filter(dv == "Life satisfaction" & iv == "Instagram" & type == "within") %>% select(conf.low) %>% my_round("b")`, `r dat_fig_results_channel %>% filter(dv == "Life satisfaction" & iv == "Instagram" & type == "within") %>% select(conf.high) %>% my_round("b")`]).
+Respondents who used Twitter more frequently than usual to attain COVID-19 related news reported slightly _lower_ levels of life satisfaction than usual (_b_ = `r dat_fig_results_channel %>% filter(dv == "Life satisfaction" & iv == "Twitter" & type == "within") %>% select(estimate) %>% round(2)` [95% CI `r dat_fig_results_channel %>% filter(dv == "Life satisfaction" & iv == "Twitter" & type == "within") %>% select(conf.low) %>% round(2)`, `r dat_fig_results_channel %>% filter(dv == "Life satisfaction" & iv == "Twitter" & type == "within") %>% select(conf.high) %>% my_round("b")`]).
+However, both effects were still completely inside of the null region, hence not large enough to be considered meaningful.
 In sum, the hypothesis was supported also for the COVID-19 related use of important social media channels.
@@
 Only one relation did not include zero, was hence statistically significant. 
-Respondents who across all waves used YouTube more frequently than others for COVID-19 related reasons reported marginally higher levels of negative affect (_b_ = `r dat_fig_results_channel %>% filter(dv == "Negative affect" & iv == "YouTube" & type == "between") %>% select(estimate) %>% round(2)` [95% CI `r dat_fig_results_channel %>% filter(dv == "Negative affect" & iv == "YouTube" & type == "between") %>% select(conf.low) %>% my_round("b_txt")`, `r dat_fig_results_channel %>% filter(dv == "Negative affect" & iv == "YouTube" & type == "between") %>% select(conf.high) %>% my_round("b")`]).
-However, please note that this effect again was not considered large enough to be practically relevant.
+Respondents who across all waves used YouTube more frequently than others for COVID-19 related reasons reported marginally _higher_ levels of life satisfaction (_b_ = `r dat_fig_results_channel %>% filter(dv == "Life satisfaction" & iv == "YouTube" & type == "between") %>% select(estimate) %>% round(2)` [95% CI `r dat_fig_results_channel %>% filter(dv == "Life satisfaction" & iv == "YouTube" & type == "between") %>% select(conf.low) %>% my_round("b_txt")`, `r dat_fig_results_channel %>% filter(dv == "Life satisfaction" & iv == "YouTube" & type == "between") %>% select(conf.high) %>% my_round("b")`]).
+However, please note that this effect again was not large enough to be considered practically relevant.
 
-```{r tab-aff-neg}
-# apa_print(model_aff_neg_lmer)
-# tab <- broom.mixed::tidy(model_aff_neg_lmer, conf.int = T) %>% 
-#  select(term, estimate, conf.low, conf.high, p.value)
-#  kable()
-```
+Again, note that when comparing the results with and without control variables, the results differed.
+Especially on the between-person level, altogether five effects stopped being significant if they were controlled for additional variables.
+For example, using Instagram was significantly (though not meaningfully) related to increased life satisfaction.
+However, when controlling for additional covariates, the effect became virtually zero (see Figure \@ref(fig:res-channels)).
 
@@
 
-In what follows, I briefly report also the results of some covariates. 
-Several variables showed larger associations with well-being. 
-Note because each variable has a different scaling, we would again need to define a SESOI for each variable, which cannot be implemented here. 
-But note that several variables would likely fall outside of such a SESOI.
-This includes for example internal locus of control, health, or employment. 
-For a brief overview, see Figure \@ref(fig:res-control).
+In what follows, I briefly report some exploratory analyses that weren't preregistered.
+First, to contextualize the results reported above and to see if the results included any meaningful effects at all, I also looked at the effect sizes of selected (cherry-picked) covariates.
+Because each variable had different response options, we would actually need to define a SESOI for each variable, which for reasons of scope I cannot implement here. 
+Therefore, I report the results of the standardized scales,
+which allows for a better comparison across the differently scaled variables.
+For what it's worth, as a rough estimate for the SESOI we can build on the typical convention that small effects start at _r_ = |.10|.
+The results showed that several effects fell outside of the SESOI, were hence considered meaningful.
+This includes for example internal locus of control, health, satisfaction with democracy, or exercising.
+For an overview, see Figure \@ref(fig:res-control).
 
-```{r res-control, fig.cap="Results of selected covariates.", out.width = "\\textwidth", cache=F}
-knitr::include_graphics("figures/fig_results_control.pdf")
+```{r res-control, fig.cap="Results of selected covariates.
+All variables were standardize except 'Male'  and ‘Employed in public service', because there were measured on a binary scale.", out.width = "\\textwidth", cache=F}
+knitr::include_graphics("figures/fig_results_control_std.pdf")
 ```
+
+To find out whether my inferences were robust across legitimate (though arguably inferior) alternative analyses, I reran the analyses also using standardized estimates, mean scores instead of factor scores, and with a data set where missing data were not imputed.
+The results were virtually the same.
+For example, all standardized COVID-19 related types of social media use or channels were not significantly larger than a SESOI of $\beta$ = |.10|.
+The additional analyses are reported in the [companion website](https://XMtRA.github.io/Austrian_Corona_Panel_Project/analyses_additional.html).
 
@@
 In this study I analyzed the effects of COVID-19 related social media use on well-being.
-The data come from a panel study with 24 waves and is representative of the Austrian population.
-A random effects model, which separated between person relations from within-person effects and which controlled for several third variables, showed that within-person effects were trivial.
-People who used social media more than usual to learn about COVID-19 did not show changes in their well-being.
+The data come from a panel study with 24 waves and are representative of the Austrian population.
+In a random effects model I separated between person relations from within-person effects and controlled for a large number of both stable and varying covariates, thereby aiming to assess causality.
+The results showed that within-person effects were trivial.
+People who used social media more than usual to learn about COVID-19 didn't show meaningful changes in their well-being.
 
-The results imply that COVID-19 related social media use does not seem to be particularly relevant for people's well-being.
-Other factors among the third variables that were measured revealed larger effects or relations, implying that well-being is rather determined by aspects such as health, employment, or locus of control.
-According to this study, popular fears that "doomscrolling" or overusing social media during times of crises do not seem to be justified.
+The results imply that COVID-19 related social media use doesn't seem to be particularly relevant for well-being.
+Other factors among the third variables that were measured revealed larger effects or relations, suggesting that well-being is determined by alternative aspects such as health, satisfaction with democracy, locus of control, or exercising.
+According to this study, popular fears that "doomscrolling" or overusing social media during times of crises don't seem to be justified.
 
-The study is not aligned with several recent studies on this particular research question or closely related research questions.
-This includes a study by @bendauAssociationsCOVID19Related2021, which showed negative relations between social media and well-being.
-However, note that @bendauAssociationsCOVID19Related2021 analyzed cross-sectional data on a between-person level, while not controlling for third variables, which is not optimal for investigating causal effects.
-At the same time, this study is well-aligned with recent studies and meta-analyses analyzing the effects of social media use from a more general perspective or that adopted a somewhat different angle.
-These studies have found that the effects of various types of social media use on several well-being indicators are small at best, often too small to matter [@meierComputerMediatedCommunicationSocial2020; @orbenTeenagersScreensSocial2020; @fergusonThisMetaanalysisScreen2021], which echoes the results obtained here.
+On the one hand, the results are not aligned with several recent studies analyzing similar or closely related research questions.
+This includes a study by @bendauAssociationsCOVID19Related2021, which showed negative relations between social media and well-being [but see @bradleyStressMoodSmartphone2021; or @sewallObjectivelyMeasuredDigital2021].
+However, note that @bendauAssociationsCOVID19Related2021 analyzed cross-sectional data on a between-person level while not controlling for third variables, which is not optimal for investigating causal effects.
+On the other hand, the results are well-aligned with recent studies and meta-analyses analyzing the effects of social media use from a more general perspective or from a somewhat different angle. 
+These studies have found that the effects of various types of social media use on several well-being indicators are small at best, often too small to matter [@meierComputermediatedCommunicationSocial2020a; @orbenTeenagersScreensSocial2020; @fergusonThisMetaanalysisScreen2021], which echoes the results obtained here.
+
+If anything, two preliminary and subtle trends can be observed.
+First, of all the three COVID-19 related social media activities, people who read about the pandemic more than others showed slightly decreased levels of positive affect, and people who actively posted about the pandemic more than others showed slightly increased levels of negative affect.
+On the other hand, however, people who posted more about COVID-19 also showed slightly higher levels of positive affects, so taken together the results are ambivalent.
+Second, in terms of media channels, using Twitter more than usual was related to slightly decreased levels of life satisfaction.
+Twitter is considered to have more negative affordances and tonality as compared to other networks such as Instagram [@waterlooNormsOnlineExpressions2018], which might help explain the results.
+Instagram, on the other hand, was related to slightly increased levels of life satisfaction.
+To speculate, the often-criticized positivity bias on Instagram might have been somewhat beneficial in times of the pandemic.
+That said, all these effects were still very small and arguably too small to matter. 
+But future research might elaborate on these specific relations to probe their stability and relevance.
+
+Finally, another interesting observation is that life satisfaction was remarkably stable.
+Hence, even in times of a pandemic, it seems that such broad assessment of life vary only mildly.
+This supports the hypothesis that life satisfaction seems to be determined largely by stable factors such as one's genes [@brownEasyHappinessPie2019].
 
@@
 
-The current study analyzed whether changes in media use were related to changes in well-being, while controlling for several potential confounds. 
-Together, this allows for a good perspective on potential causality.
+The current study analyzed whether changes in media use were related to changes in well-being, while controlling for several potential confounders. 
+Together, this allows for an improved perspective on assessing causality.
 That said, causality necessitates temporal order, and the cause needs to precede the effect.
 Regarding media use, such effects often happen immediately or shortly after use, necessitating intervals in the hours, minutes, or even seconds.
-In many cases only experience sampling studies that ask users in the very moment can produce such knowledge.
+In many cases only experience sampling studies asking users in the very moment can produce such knowledge.
 However, even then we don't know for certain if we actually measured the right interval.
+Effects depend on the intensity of use or the length of the interval, and to borrow the words from @rohrerTheseAreNot2021, there is no such thing as "the" effect of social media use on well-being.
 Hence, to document how effects unfold, future research needs to employ different study designs probing different time lags.
-In addition, more thought needs to be invested in what relevant stable and varying factors should be selected as control variables.
+In addition, more thought needs to be invested in what relevant stable and varying factors we should include as control variables, and I hope this study provides a first step into this direction.
 
-Although I had already reduced the predefined SESOIs to be less conservative, potentially they were still too large.
+Although I had already reduced the predefined SESOIs to be less conservative, they were potentially still too large.
 Media use is only one aspect of several factors that simultaneously affect well-being.
-Is it realistic that extremely changing only _one_ of these aspects (e.g., by completely stopping the use of social media) should already manifest in a detectable change in well-being?
-Or would it make more sense to expect that if people regularly start doing say _two_ activities (e.g.
-regularly exercising _and_ establishing a reading habit) together should show perceivable improvements to well-being?
-In other words, the beneficial effect of a particular activity is large enough if people feel a difference when implementing two of those activities. 
-Practically, this would imply a SESOI half as large as I have defined here, that is _b_ = |.15| for well-being and _b_ = |.075|. 
-In this case, this would not make a difference, as even with these more liberal thresholds all but one effect would still be completely in the null region.
-However, at all events future research needs to start a thorough discussion on what effect sizes are considered meaningful and relevant, and with this study I hope to provide some helpful input and first concrete guidelines.
+Is it really realistic to expect that extremely changing only _one_ of these aspects should manifest in a detectable change in well-being?
+Or would it make more sense to expect that thoroughly committing to say _two_ activities (e.g.
+regularly exercising _and_ establishing a reading habit) should then cause a detectable improvement in well-being?
+Practically, this would imply a SESOI half as large as I have defined here, namely _b_ = |.15| for well-being and _b_ = |.075| for affect.
+In the case of this study, however, reducing the SESOI would not even make a big difference, as also with these more liberal thresholds all but three effect would still be completely in the null region, and no effect would be outside of the null region.
+However, at all events I encourage future research to start a thorough conversation on what effect sizes are considered meaningful and what not.
+Again, with this study I hope to provide some first input and guidelines.
 
@@
 Measuring well-being with self-reports is adequate, because it by definition requires introspection.
-However, it would be preferable to measure social media use objectively, because people cannot reliably estimate their use.
-That said, objective measures often cannot capture the content or the motivation of the use, and only very complicated tools that record the content that was used (such as the Screenome project) might produce such data.
-Unfortunately, such procedures introduce other problems, for example related to privacy.
-Hence, for this type of research question it seems necessary still to use self-reported measures.
+However, it would be preferable to measure social media use objectively, because people cannot reliably estimate their use [@scharkowAccuracySelfreportedInternet2016].
+That said, objective measures often cannot capture the content or the motivation of the use, and only very complicated tools recording the actual content (such as the Screenome project) might produce such data.
+Unfortunately, such procedures introduce other problems, especially related to privacy.
+Hence, for this type of research question it still seems necessary to use self-reported measures.
 
 Because the data were collected in a single country, the generalizability of the results is limited.
-The results potentially apply primarily to the more Western sphere, and might not hold true in other cultures, especially cultures with a different media landscape or alternative social media channels.
-That said, because this is a comparatively large study representative of a country's entire population, and because several waves were collected across a large time span, the results should be at least as generalizable as other typical empirical studies collected in the social sciences.
+The results apply primarily to the more Western sphere, and might not hold true in other cultures, especially cultures with a different media landscape or alternative social media channels.
+That said, because this is a comparatively large study representative of an entire country, and because several waves were collected across a large time span, the results should be at least as generalizable as other typical empirical studies collected in the social sciences.
 
@@
 
-In this study, COVID-19 related social media use did not causally affect several indicators of well-being, including life satisfaction, positive affect, and negative affect.
-However, factors other than social media use were more meaningfully related to well-being, such as physical health, employment, or believing that one is in control of one's life.
-If it's the aim to improve well-being, it might hence be fruitful not to focus on social media but to address other, potentially more pressing societal problems related to inequality, safe spaces, or health care.
+In this study, COVID-19 related social media use didn't meaningfully affect several indicators of well-being, including life satisfaction, positive affect, and negative affect.
+However, factors other than social media use were meaningfully related to well-being, such as physical health, exercise, satisfaction with democracy, or believing that one is in control of one's life.
+If it's our aim to improve well-being, it might hence be more fruitful not to focus so much on social media but to address other, more pertinent societal problems related to health care, regular exercise, or a functioning democratic system.
 
@@
 
-All the stimuli, presentation materials, analysis scripts, and a reproducible version of the manuscript can be found on the companion website (https://xmtra.github.io/Austrian_Corona_Panel_Project/index.html).
+All the stimuli, presentation materials, analysis scripts, and a reproducible version of the manuscript can be found on the companion website (https://XMtRA.github.io/Austrian_Corona_Panel_Project).
 
@@
 The data can only be used for scientific purposes.
+
+# Acknowledgements
+
+I would like to thank BLINDED for providing valuable feedback on this manuscript.

### 2021-12-13 manuscript.tex
- updated paper for submission

--- 2021-11-05 manuscript.tex
@@
 \hypersetup{
-  pdftitle={No meaningful effects of COVID-19 related social media use on well-being},
+  pdftitle={Analyzing the effects of COVID-19 related social media use on well-being},
   pdflang={en-EN},
-  pdfkeywords={COVID-19, Coronavirus, well-being, affect, life satisfaction, social media use, news use, communication, random effects within between model, panel study, longitudinal.},
+  pdfkeywords={COVID-19, Coronavirus, well-being, affect, life satisfaction, social media use, news use, communication, random effects within between model, panel study, longitudinal study.},
   hidelinks,
@@
 \makeatother
-\keywords{COVID-19, Coronavirus, well-being, affect, life satisfaction, social media use, news use, communication, random effects within between model, panel study, longitudinal.}
+\keywords{COVID-19, Coronavirus, well-being, affect, life satisfaction, social media use, news use, communication, random effects within between model, panel study, longitudinal study.}
 \usepackage{lineno}
@@
 
-\title{No meaningful effects of COVID-19 related social media use on well-being}
-\author{Blinded for review\textsuperscript{1}}
+\title{Analyzing the effects of COVID-19 related social media use on well-being}
+\author{Name blinded for review\textsuperscript{1}}
 \date{}
@@
 
-Blinded for review.
-
-Correspondence concerning this article should be addressed to Blinded for review, Blinded for review. E-mail: Blinded for review
+Correspondence concerning this article should be addressed to Name blinded for review, . E-mail:
 
@@
 
-\affiliation{\vspace{0.5cm}\textsuperscript{1} Blinded for review}
+\affiliation{\vspace{0.5cm}\textsuperscript{1} }
 
 \abstract{
-In times of crisis such as the Corona pandemic, it is important that citizens stay informed about recent events, the latest political decisions, or mandatory protection measures.
+In times of crisis such as the Corona pandemic citizens need to stay informed about recent events, the latest political decisions, or mandatory protection measures.
 To this end, many people use various types of media, and increasingly social media.
-However, because social media are particularly engaging, some find it hard to disconnect and cannot stop `doomscrolling.' In this preregistered study, I investigate whether using social media for COVID-19 related topics might put personal well-being at risk.
+However, because social media are particularly engaging, some find it hard to disconnect and cannot stop `doomscrolling.' In this preregistered study, I investigate whether using social media for COVID-19 related reasons affects psychological well-being.
 To answer this question I analyzed data from the Austrian Corona Panel Project, which consists of 24 waves with overall 3,018 participants.
-I ran three random effects cross lagged panel models, controlling for several stable and varying covariates.
-Results showed that the effects of various types of COVID-19 related social media use on several indicators of well-being were very small, arguably too small to matter.
-The findings suggest fears that social media use during times of crisis might be detrimental for well-being are likely to be unfounded.
+I ran three random effects within between models, controlling for several stable and varying confounders.
+Results showed that the effects of COVID-19 related social media use on well-being were very small, arguably too small to matter.
+The findings suggest that fears that social media use during times of crisis impairs well-being are likely to be unfounded.
 }
@@
 During the COVID-19 pandemic,
-numerous events unfolded in quick succession.
-Several open questions emerged.
+numerous events unfolded in quick succession and several open questions emerged.
 How dangerous is the virus?
 Is it spreading in my region?
-How is it transmitted?
-How can I protect myself?
-Because for many it was (and continues to be) a matter of life or death, citizens had to stay informed regarding the latest developments.
-Governments around the world implemented safety measures, ranging from wearing masks, keeping physical distance, to complete lockdowns.
-In this extraordinary situation, people used media excessively to attain information, and especially social media were at an all time high (Statista, 2021).
-A new phenomenon termed ``doomscrolling'' emerged:
-People could not stop using social media to learn about COVID-19 related news.
+How is it transmitted, and how can I protect myself?
+Because for many it was (and still is) a matter of life or death, people aimed to stay informed regarding the latest developments.
+Governments around the world implemented safety measures, such as wearing masks, keeping physical distance, or enforcing lockdowns.
+In this extraordinary situation, many people used media excessively to attain information, and especially social media were at an all time high (Statista, 2021).
 
-Several people reported that they were glued to their screens and could not pursue other relevant activities such as working, taking a break, or even child-care (Klein, 2021).
-As doomscrolling increased it became doubtful whether such a surge in social media use could still be considered useful and adaptive, or whether it created an additional and new psychological danger for users' mental health (Sandstrom, Buchanan, Aknin, \& Lotun, 2021).
-These concerns seem justified: A study with 6,233 people from Germany found that ``{[}f{]}requency, duration and diversity of media exposure were positively associated with more symptoms of depression and unspecific and COVID-19 specific anxiety'' (Bendau et al., 2021), which implies negative consequences for well-being.
+Some people actually couldn't stop using social media to learn about COVID-19 related news.
+A new phenomenon emerged, termed ``doomscrolling'':
+Users were glued to their screens and found it hard to pursue other relevant activities such as working, taking a break, or looking after their children (Klein, 2021).
+As doomscrolling increased it became doubtful whether such social media is helpful, or whether it created an additional burden on mental health (Sandstrom, Buchanan, Aknin, \& Lotun, 2021).
+These concerns seem justified:
+A study with 6,233 people from Germany conducted during the pandemic found that ``{[}f{]}requency, duration and diversity of media exposure were positively associated with more symptoms of depression'' (Bendau et al., 2021, p.
+283).
 
-As a result, with this study I want to build on this research and investigate the question whether COVID-19 related social media use during the pandemic affected well-being.
-To this end, I analyze a large-scale panel study from the Austrian Corona Panel Project (Kittel et al., 2020).
-The panel consists of 24 waves and has an overall sample size of 3018.
-At least 1,500 participants took part per wave, and the sample is representative of the Austrian population.
-The panel study collected a large number of variables.
-Being able to control for many confounding third variables, both stable and varying, and together with the longitudinal design, this creates a unique opportunity to investigate the causal effects of COVID-19 related social media use on well-being.
+As a result, with this study I want to build on this research and investigate whether or not COVID-19 related social media use affected well-being during the pandemic.
+To this end, I analyzed a large-scale panel study from the Austrian Corona Panel Project (Kittel et al., 2020).
+The panel consists of 24 waves and has an overall sample size of 3,018 participants.
+The panel study collected a large number of psychological and demographic variables.
+I explicitly aimed to investigate the causal effects of COVID-19 related social media use on well-being.
 
-\hypertarget{defining-well-being-and-media-use}{%
-\subsection{Defining Well-being and Media Use}\label{defining-well-being-and-media-use}}
+\hypertarget{understanding-well-being-and-media-use}{%
+\subsection{Understanding Well-being and Media Use}\label{understanding-well-being-and-media-use}}
 
-The underlying theories that guided the selection of variables for this study are the two-continua model of mental health (Greenspoon \& Saklofske, 2001) and the hierarchical taxonomy of computer-mediated medation (Meier \& Reinecke, 2020).
-According to the two-continua model, mental health consists of two dimensions: psychopathology and well-being.
-Well-being can be further differentiated into subjective and psychological well-being (Diener, Lucas, \& Oishi, 2018).
-Whereas subjective well-being emphasizes hedonic aspects such as happiness and joy, psychological well-being focuses on eudaimonic aspects like fulfillment and meaning.
+The underlying theories that guided the selection of variables for my analysis are the two-continua model of mental health (Greenspoon \& Saklofske, 2001) and the hierarchical taxonomy of computer-mediated communication (Meier \& Reinecke, 2020).
+According to the two-continua model, mental health consists of (a) psychopathology and (b) well-being.
+Well-being can be differentiated into subjective and psychological well-being (Diener, Lucas, \& Oishi, 2018).
+Whereas subjective well-being emphasizes hedonic aspects such as happiness and joy, psychological well-being addresses eudaimonic aspects such as fulfillment and meaning.
 Subjective well-being is primarily about achieving positive affect and avoiding negative affect.
 One of the most prominent indicators of well-being is life satisfaction.
-In my view, life satisfaction should be considered a meta concept that combines psychological and subjective well-being, because it represents a general appraisal of one's life.
+In my view, because it represents a general appraisal of one's life, life satisfaction is best thought of as a meta concept that combines psychological and subjective well-being.
 Notably, life satisfaction is stable and fluctuates only little, whereas it's the exact opposite for affect (Dienlin \& Johannes, 2020).
-To capture well-being in this study, I thus build on life satisfaction, positive affect, and negative affect.
+To capture well-being in this study I thus build on life satisfaction, positive affect, and negative affect.
 Together, this should provide an encompassing perspective on potential media effects.
@@
 The hierarchical taxonomy of computer-mediated communication differentiates six levels of how people engage with digital technology.
-First, the device (e.g., smartphone); second, the type of application (e.g., social networking site); third, the branded application (e.g., Twitter); fourth, the feature (e.g., status post); fifth, the interaction (e.g., one-to-many); sixth, the message (e.g., content) (Meier \& Reinecke, 2020).
-Whereas the first four levels are focused on the \emph{channel}, the last two address the \emph{type} of communication.
-To measure social media use for consumption of COVID-19 related news and topics, I here employ both the channel and the communication perspective, which together should provide a more nuanced understanding of communication.
+First, the device (e.g., smartphone); second, the type of application (e.g., social networking site); third, the branded application (e.g., Twitter); fourth, the feature (e.g., status post); fifth, the interaction (e.g., one-to-many); and sixth, the message (e.g., content) (Meier \& Reinecke, 2020).
+Whereas the first four levels focus on the communication \emph{channel}, the last two address the communication \emph{type}.
+To measure social media use for the consumption of COVID-19 related news and topics, I here employ both the channel and the type of communication perspective, which together provides a nuanced understanding of communication.
 
-First, I will investigate how different types of communication affect well-being.
-Specifically, I will differentiate between active and passive use.
-I will distinguish (a) \emph{reading} COVID-19 related social media news (passive), \emph{posting} content regarding COVID (active), and \emph{liking and sharing} COVID-19 related posts (both active and passive).
-Second, I will analyze also how using the most prominent branded applications affect well-being, and whether this effect changes across applications.
-According to Meier and Reinecke (2020), branded apps are separate entities with different effects.
-Therefore, Twitter might have a different effect than WhatsApp because of their different affordances.
-For example, Waterloo, Baumgartner, Peter, and Valkenburg (2018) found that it is more adequate to express negative emotions on WhatsApp than on Twitter or on Instagram.
+First, I investigate how different types of communication affect well-being.
+Specifically, I differentiate between active and passive use.
+I distinguish (a) \emph{reading} (passive), (b) \emph{posting} (active), and (c) \emph{liking and sharing} COVID-19 related posts (both active and passive).
+Second, I analyze how using the most prominent branded applications affects well-being, and whether this effect changes across applications.
+Branded apps are separate entities with potentially divergent effects.
+Twitter might have a different effect as compared to WhatsApp because of their respective affordances.
+For example, Waterloo, Baumgartner, Peter, and Valkenburg (2018) found that it's more adequate to express negative emotions on WhatsApp than on Twitter or on Instagram.
 The branded applications investigated here are Facebook, Twitter, Instagram, WhatsApp, and YouTube.
-Worth noting, this study is not about \emph{general} social media during times of COVID, but on social media use \emph{focused} on COVID-19 related news and interactions.
-Examples of such media use include posting about the pandemic or retweeting COVID-19 related posts.
+Worth noting, this study is not about \emph{general} social media use during times of COVID, but on social media use \emph{focused} on COVID-19 related content.
+Examples of such media use include posting thoughts about the pandemic or retweeting COVID-19 related news.
 
@@
 
-So far, there is only little empirical research on COVID-19 related social media use and well-being.
+So far, there is only little empirical research on COVID-19 related social media use on well-being.
 In their study on the relations between media use and mental health during the pandemic, Bendau et al.
 (2021) found that people who used social media as a primary source of information reported on average ``significantly more unspecific anxiety and depression {[}{]} and significantly more specific COVID-19 related anxiety symptoms'' (p.~288).
-Hence, this might hint at potential negative effects of social media as news use.
-However, note that this finding comes from a between-person relation stemming from cross-sectional data (see below).
-We, therefore, don't know whether the differences in mental health and well-being are due to social media use or other third variables, such as age, health, or education.
-Eden, Johnson, Reinecke, and Grady (2020) analyzed the media use of 425 US college students during the first wave of the pandemic, and found both positive and negative relations with well-being.
+Eden, Johnson, Reinecke, and Grady (2020) analyzed the media use of 425 US college students during the first wave of the pandemic, finding both positive and negative relations with well-being.
 In a sample of 312 respondents collected via Amazon Mechanical Turk, Choi and Choung (2021) reported that people who used media to attain information were more lonely and less satisfied with their lives.
 Stainback, Hearne, and Trieu (2020) analyzed a large-scale study with 11,537 respondents from the US and found that increased COVID-19 media consumption was related to more psychological distress.
+A four-wave panel study with 384 young adults from the U.S.
+analyzed the effects of general digital technology use---objectively measured via screenshots of screen-time applications---on mental health, separating within- and between-person relations (Sewall, Goldstein, \& Rosen, 2021).
+The results showed that digital technology did not have any significant effects on mental health (for a similar study with comparable results, see Bradley \& Howard, 2021).
+Together, the literature is mixed, with a slight focus on the negative effects of social media as news use (see also Dörnemann, Boenisch, Schommer, Winkelhorst, \& Wingen, 2021; Liu \& Tong, 2020; Riehm et al., 2020).
+However, note that all of these findings represent between-person relations stemming from cross-sectional data (see below).
+We therefore don't know whether the differences in mental health and well-being are due to social media use or due to other third variables, such as age, health, employment, or education.
 
-On the other hand, the question of whether and how \emph{general} social media use affect well-being is well-researched.
+The question of whether and how social media use affects well-being \emph{in general}, on the other hand, is well-researched.
 This also holds true for the different types of communication such as active or passive use.
-A meta review (i.e., an analysis of meta-analyses), found that the relation between social media use and well-being is likely in the negative spectrum but very small, potentially too small to matter (Meier \& Reinecke, 2020).
-What determines whether an effect is considered trivial or small?
-If we refer to standardized effect sizes, according to Cohen (1992) small effect sizes start at \emph{r} = .10.
+A meta review (i.e., an analysis of meta-analyses) found that the relation between social media use and well-being is likely in the negative spectrum but very small (Meier \& Reinecke, 2020)---potentially too small to matter.
+What determines whether or not an effect is considered small or trivial?
+As a starting point, we could refer to standardized effect sizes.
+According to Cohen (1992), small effect sizes start at \emph{r} = .10.
 And indeed, several if not most of the current meta-analyses find effect sizes below that threshold (Ferguson et al., 2021; Huang, 2017; Meier \& Reinecke, 2020).
 
-These overviews are well aligned with several individual new studies that employed advanced methods (Keresteš \& Štulhofer, 2020; Orben, Dienlin, \& Przybylski, 2019; Przybylski, Nguyen, Law, \& Weinstein, 2021; Schemer, Masur, Geiß, Müller, \& Schäfer, 2021).
-For example, Beyens, Pouwels, Driel, Keijsers, and Valkenburg (2021) reported that although for some users (roughly one quarter) the effects were negative, for almost the same number of users they were positive, while for the majority the effects were neutral.
-In conclusion, most effects are somewhere between trivial and small.
-I therefore think it's most convincing to expect trivial to small effects also in the case of COVID-19 related social media use.
+These overviews are well aligned with several individual studies employing advanced methods (Keresteš \& Štulhofer, 2020; Orben, Dienlin, \& Przybylski, 2019; Przybylski, Nguyen, Law, \& Weinstein, 2021; Schemer, Masur, Geiß, Müller, \& Schäfer, 2021).
+For example, Beyens, Pouwels, Driel, Keijsers, and Valkenburg (2021) reported that although for some users (roughly one quarter) the effects of social media use on well-being were negative, for almost the same number of users they were positive, while for the rest the effects were neutral.
+In conclusion, most effects are likely somewhere between trivial and small.
+I therefore expect that also in the case of COVID-19 related social media use effects will be trivial to small.
 
-From a theoretical perspective, how could we explain how COVID-19 related social media use might affect well-being?
-In what follows, I outline potential arguments as to why the effect might be positive or negative, direct or indirect.
-In advance, there does not seem to be a clear winner, and it's likely that both positive and negative effects are equally strong.
+From a theoretical perspective, how could we explain whether COVID-19 related social media use might affect well-being?
+In what follows, I outline potential arguments as to why the effect might be positive or negative, direct or indirect, or nonexistent.
+In advance, there doesn't seem to be a clear winner, and it's likely that both positive and negative effects cancel each other out.
 
-First, one could assume a \emph{direct} negative effect on well-being, and especially on positive or negative affect, which is more volatile and fluctuating.
-Dangers, inequalities, corruption---these were the headlines across many countries worldwide.
-If one learns about such things, the initial reaction might be shock, fear, or dismay.
-Repeatedly consuming such news might be depressing, perhaps even changing some general perspectives on life, without any further mediating processes.
-That said, because not all news was negative, and because many people showed solidarity and compassion, there was also positive and potentially uplifting news, potentially compensating for the negative news.
+First, one could assume a \emph{direct} negative effect on well-being, and especially on positive or negative affect, which are more volatile and fluctuating.
+Dangers, inequalities, corruption---these were the headlines during the pandemic across many countries worldwide.
+If one learns about such events, the initial reaction might be shock, fear, or dismay.
+Consuming such news can be depressing (Dörnemann, Boenisch, Schommer, Winkelhorst, \& Wingen, 2021), perhaps even changing some general perspectives on life.
+That said, because not all news was negative, and because many people showed solidarity and compassion, there was also positive and uplifting content, potentially compensating for the negative effects (Dörnemann, Boenisch, Schommer, Winkelhorst, \& Wingen, 2021).
 
@@
 For example, during the pandemic social media use was at an all-time high in the US (Statista, 2021).
-As has been expressed by many before, it is most likely that moderate social media use is not detrimental (Orben, 2020).
-Overuse, however, is more critical, and several studies have shown more pronounced negative effects for extreme users (Przybylski \& Weinstein, 2017).
-To explain, overuse likely impairs well-being if it replaces other meaningful or functional activities such as meeting others, working, actively relaxing, or exercising.
-So if a society collectively overuses social media, there is potential for negative effects.
+In general, as has been expressed by many before, it is most likely that moderate social media use is not detrimental (Orben, 2020).
+Overuse, however, might be more critical, and several studies have shown more pronounced negative effects for extreme users (Przybylski \& Weinstein, 2017).
+To explain, overuse likely impairs well-being if it replaces meaningful or functional activities such as meeting others, working, actively relaxing, or exercising.
+So if a society collectively overuses social media when doomscrolling, there is potential for negative effects.
 
-On the other hand, one can make the case that overuse can also be beneficial, especially in times of a pandemic, even if the use is mainly COVID-19 related.
-Exchanging COVID-19 related messages with friends via WhatsApp might replace the in-person contact one would have otherwise, but which is literally impossible at that time.
-In situations where meaningful and functional activities are explicitly or implicitly prohibited, using social media to exchange about COVID-19 related topics might not be the worst idea.
-In fact, given that nowadays a large number of experts, scientists, and politicians converse directly on social media, one can get first-hand information on the current developments.
+On the other hand, one can make the case that overuse might also be beneficial, especially in times of a pandemic---even if the use is mainly COVID-19 related.
+Exchanging COVID-19 related messages with friends via WhatsApp might replace the in-person contact one would have otherwise, but which is literally impossible at the time.
+In situations where meaningful and functional activities are prohibited, using social media to exchange about COVID-19 related topics might not be the worst idea.
+Besides, given that nowadays a large number of experts, scientists, and politicians converse directly on social media, one can get first-hand high quality information on current developments.
 
-Together, the strongest argument seems to be that \emph{in general} the effects of social media on well-being are, on average, small at best.
-Because this study looks at only \emph{one part} of social media use---namely, COVID-19 related interactions---it is even more focused, and the overall potential of the effects should diminish even further.
+Together, the strongest argument to me is that \emph{in general} the effects of social media on well-being are, on average, small at best.
+Because this study only looks at \emph{one part} of social media use---namely, COVID-19 related interactions---it is very focused, diminishing the overall potential of the effects even further.
 Whether or not using social media for COVID-19 related aspects is detrimental during a pandemic is also not entirely clear.
-Therefore, I expect to find that COVID-19 related communication on social media will not affect well-being in a meaningful or relevant way.
+Therefore, I expect that COVID-19 related communication on social media doesn't affect well-being in a meaningful or relevant way.
 
@@
 Testing this hypothesis, however, is not trivial.
-First, in contrast to most hypotheses typically posited in the social sciences, it implicitly contains an effect size, a so-called smallest effect size of interest (SESOI).
+First, in contrast to most hypotheses typically posited in the social sciences it implicitly contains an effect size, a so-called smallest effect size of interest (SESOI).
 Effectively testing this hypothesis necessitates defining what's considered a ``trivial effect size'' and what's not.
@@
 However, standardized effect sizes should only be a first step toward evaluating an effect's relevance (Baguley, 2009).
-Standardized effect sizes are determined by a sample's variance.\footnote{Consider the effect size Cohen's \emph{d}: The mean's of the two groups that are to be compared are subtracted from one another and then divided by the sample's standard deviation (Cohen, 1992).
-Hence, if there is more deviation/variance in a sample, the effect size decreases, even if the difference of the group's means stays the same.}
-However, this is problematic: The question of whether or not social media use affects a particular person in a relevant way should not depend on the variance in the sample in which that person's data were collected.
+Standardized effect sizes are determined by a sample's variance,\footnote{Consider the effect size Cohen's \emph{d}: The mean's of the two groups that are to be compared are subtracted from one another and then divided by the sample's standard deviation (Cohen, 1992).
+Hence, if there is more deviation/variance in a sample, the effect size decreases, even if the difference of the group's means stays the same.} which is problematic:
+The question of whether or not social media use affects a particular person in a relevant way should not depend on the variance in the sample in which that person's data were collected.
 Instead, it should depend on absolute criteria.
-What could be a minimally interesting, a nontrivial effect?
+
+What could be a minimally interesting, nontrivial effect?
 Because this is a normative and ultimately philosophical question, there can never be a clear, single, or unanimous answer.
-In the end, it is a personal question.
-However, it is still both necessary and fruitful to try to answer it.
-I suggest the following SESOI for this research question:
+However, it is still necessary and helpful to try provide such a plausible benchmark.
+I therefore suggest the following SESOI for this research question:
 
-If a heavy user of COVID-19 related social media news suddenly \emph{completely stops} using social media, this should have a \emph{noticeable} impact on their overall well-being.
+\begin{quote}
+SESOI: If a heavy user of COVID-19 related social media news suddenly \emph{stops} using social media altogether, this should have a \emph{noticeable} impact on their overall well-being.
+\end{quote}
 
-What would this mean practically and how could it be operationalized?
+What does this mean practically and how can it be operationalized?
 In this study, COVID-19 related social media use was measured on a 5-point scale, ranging from 1 = \emph{never} to 5 = \emph{several times a day}.
-Thus, the example from above would imply that a change of four units in social media use (e.g., a complete stop) should correspond to a noticeable change in well-being.
+Thus, a change of four units in social media use (e.g., a complete stop) should correspond to a noticeable change in well-being.
 But what's a noticeable change in well-being?
-According to Norman, Sloan, and Wyrwich (2003), people can reliably differentiate between \emph{seven} levels of satisfaction with health.
-So we could state that a four unit change in social media use should result in a one unit change in life satisfaction, if satisfaction is measured on a 7-point scale.
-Statistically, in a regression \emph{b} estimates the change in the dependent variable if the independent variable increases by one point.
-Transferred to our example, we would hence expect a SESOI of \emph{b} = 1 / 4 = .25.
-
-In this study, life satisfaction was measured on an 11-point scale, which is more than the seven degrees people can reliably differentiate.
-We hence need to transform a 1-point change on a 7-point scale to its equivalent on an 11-point scale, which is 11 / 7 = 1.57.
-Hence, we now expect our effect to be at least as large as \emph{b} = .25 * 1.57 = 0.39.
-For affect, which was measured on a 5-point scale, our SESOI would therefore be \emph{b} = 1/4 * 5/7 = 0.18.
-Because we are agnostic as to whether the effects are positively or negatively nontrivial, the null region will include negative and positive effects (in Bayesian terms, this is called the region of practical equivalence {[}ROPE{]}).
-In order not to exaggerate precision, and to be a bit less conservative, these numbers will be reduced to nearby thresholds.
-Note that other researchers also decreased or recommended decreasing thresholds for effect sizes when analyzing within-person or cumulative effects (Beyens, Pouwels, Driel, Keijsers, and Valkenburg (2021); Funder and Ozer (2019)).
-Together, this leads to an indifference region ranging from \emph{b} = -.30 to \emph{b} = .30 for life satisfaction, and \emph{b} = -.15 to \emph{b} = .15 for positive and negative affect.
+According to Norman, Sloan, and Wyrwich (2003), people can reliably distinguish \emph{seven} levels of satisfaction with health.
+So if satisfaction is measured on a 7-point scale, we would state that a four unit change in social media use should result in a one unit change in life satisfaction.
+(For more information, see Methods section ``Inference Criteria.'')
 
@@
 The hypothesis explicitly states a causal effect.
-In non-experimental studies, longitudinal designs help investigate causality.
-Using longitudinal designs alone, however, is not sufficient for correct causal statements.
-In addition, we also need to control for third variables, and importantly also for \emph{varying} third variables.
+In non-experimental studies, longitudinal designs can help investigate causality.
+Using longitudinal designs alone, however, is not sufficient for establishing correct causal statements (Rohrer \& Murayama, 2021).
+In addition, we for example also need to control for confounding third variables, and importantly also for \emph{varying} third variables.
 
-Consider the following example.
-Imagine that a person suddenly starts using social media much more than usual, and then at the same time also becomes less satisfied with their life.
-After some time, use recedes again, whereas life satisfaction returns to prior levels.
-If this happens to several people at the same time, in a longitudinal study we would then observe a significant effect of social media use on life satisfaction.
-However, it could also be the case that during the study there was a major exogenous event (say, a pandemic), which is why a large part of the working population lost their jobs.
-Hence, the causal effect reported above was likely confounded, because in reality it was the pandemic that caused both social media use to rise and life satisfaction to plummet.
+To illustrate, consider the following example.
+Imagine that a person suddenly starts using social media much more than usual, and then after some time becomes less satisfied with their life.
+Eventually, use and life satisfaction return to prior levels.
+If this happens to several people at the same time, in a longitudinal study we could then observe a significant effect of social media use on life satisfaction.
+However, it could also be the case that during the study there was a major exogenous event (say, a pandemic), which caused large parts of the working population to loose their jobs.
+Hence, the causal effect reported above was confounded, because in reality it was the pandemic that caused both social media use to rise and life satisfaction to go down.
 
-Thus, only if we can control for \emph{all} potential varying third variables, can we correctly estimate causality without bias (Rohrer, 2018).
-Obviously, we can never be entirely sure that we have included all varying covariates, which makes absolute statements regarding causality impossible.
-In addition, when determining the overall causal effect, we should \emph{not} control for mediating variables (Rohrer, 2018).
-Doing so would bias and distort our assessment of the role of social media use.
-Complicating matters, often it is unclear if a variable is a mediator or a confounder.\footnote{In addition, there also exist colliders, which I don't discuss here and which complicate the issue even further (Rohrer, 2018).}
-However, when controlling for relevant variables (that aren't mediators), we can be much more certain that we measured causality correctly.
-The aim should therefore be to collect as many \emph{relevant} varying and nonvarying third variables as possible, while knowing that absolute certainty regarding causality cannot be reached.
+Thus, only when controlling for \emph{all} relevant confounders, we can correctly estimate causality without bias (Rohrer, 2018).
+Obviously, we can never be entirely sure to have included all confounders, which makes absolute statements regarding causality virtually impossible.
+In addition, when determining the overall causal effect, we need to make sure \emph{not} to control for mediating variables (Rohrer, 2018), for doing so would bias our assessment of the causal effect.
+Complicating matters further, it is often unclear if a variable is a mediator or a confounder.\footnote{In addition, there also exist colliders, which I don't discuss here and which complicate the issue even further (Rohrer, 2018).}
+However, despite all these caveats, when controlling for relevant variables (that aren't mediators), we can be much more certain that we measured causality correctly.
+The aim should therefore be to collect as many varying and nonvarying confounders as possible (which I believe is seldom done in our field), while knowing that absolute certainty regarding causality cannot be reached.
 
-When looking for suitable control candidates, ideally we select variables that affect both media use and well-being.
+When searching for suitable candidates for confounders, we should look for variables that affect both media use and well-being.
 Controlling for these factors isolates the actual effect of social media use on well-being.
-We can also control for variables that affect \emph{only} well-being.
-However, in doing so not much is gained or lost because the effects of social media use would remain virtually the same (Kline, 2016).\footnote{But see McElreath (2021).}
+We can also control for variables that affect only social media use or well-being.
+However, in doing so not much is gained or lost, because the effects of social media use would remain virtually the same (Kline, 2016; but see McElreath, 2021).
 
-In this study, I hence plan to control for the following variables, which either have already been shown to affect both social media use and well-being, or which are very likely to do so, and which are likely not mediators:
+In this study, I hence plan to control for the following variables, which either have already been shown to affect both social media use and well-being or which are likely to do so, and which also aren't mediators:
 gender, age, education, Austria country of birth, Austria country of birth of parents, text-based news consumption, video-based news consumption, residency Vienna, household size, health, living space, access to garden, access to balcony, employment, work hours per week, being in home-office, household income, outdoor activities, satisfaction with democracy, disposition to take risks, and locus of control.
-I will not control for variables such as trust in institutions or media, because these variables might be influenced by social media use to a meaningful extent.
+I will not control for variables such as trust in institutions or trust in media, because these variables might be influenced by social media use to a meaningful extent.
 
-Next to including covariates, it is now increasingly understood that causal effects need to be analyzed from an internal, within-person perspective.
-If a specific person changes their media diet, we need to measure how this affects \emph{their} well-being.
-Between-person comparison from cross-sectional data, where participants are interviewed only once, cannot provide such insights.
-In this study, I will hence differentiate between within-person effects and between-person relations.
-As explicated in the hypothesis, to test the hypothesis I will thus consider only the within-person effects.
+Next to including covariates, it's now increasingly understood that causal effects should be analyzed from an internal, within-person perspective (Hamaker, 2014).
+If a specific person changes their media diet, we need to measure how this behavior affects \emph{their own} well-being.
+Between-person comparisons from cross-sectional data, where participants are interviewed only once, cannot provide such insights.
+In this study, I hence differentiate between-person relations from within-person effects.
+And as explicated above, to test the hypothesis I thus consider only the within-person effects.
 
 Finally, one precondition of causality is temporal order.
-The cause needs to precede the outcome.
-Finding the right interval between causes and effects is crucial.
-For example, if we want to measure the effect of alcohol consumption on driving performance, it makes a big difference if driving performance is measured one minute, one hour, one day, or one week after consumption.
-Finding the right interval is difficult.
-If variables are more stable, longer intervals are needed, and if they fluctuate a lot, shorter intervals are necessary.
-In the case of well-being, to analyze affect we need shorter intervals, while for life satisfaction longer ones.
-Still, choosing the right interval is challenging, because especially short intervals are practically hard to implement and often require experience sampling measures (also known as in situ measurement or ambulant assessment).
+The cause needs to precede the effect.
+Finding the right interval between cause and effect is crucial.
+For example, if we want to understand the effect of alcohol consumption on driving performance, it makes a big difference if driving performance is measured one minute, one hour, one day, or one week after consumption.
+If variables are stable, longer intervals are needed; if they fluctuate, shorter intervals.
+In the case of well-being, we need shorter intervals for affect and longer ones for life satisfaction.
+Still, choosing the right interval is challenging, because especially short intervals are hard to implement in practice and often require advanced methods such as experience sampling (also known as in situ measurement or ambulant assessment) (Schnauber-Stockmann \& Karnowski, 2020).
 
-In this study, I will adopt an intermediate perspective.
-I will analyze if, when a person changes their social media diet, are there simultaneous changes in well-being?
-When additionally controlling for both stable and varying covariates, we can then be more sure that the effect is indeed causal.
-Similar approaches were implemented already by several other studies (Johannes, Dienlin, Bakhshi, \& Przybylski, 2021; Scharkow, Mangold, Stier, \& Breuer, 2020) and they considered one of the best practices to analyze causality (Bell, Fairbrother, \& Jones, 2019).
+In this study, I therefore adopt an intermediate perspective.
+I analyze if when a person changes their social media diet, are there \emph{simultaneous} changes in their well-being?
+When additionally controlling for both stable and varying confounders, we can then be more sure that the effect is indeed causal.
 
@@
 The (anonymous) preregistration can be accessed here: \url{https://osf.io/87b24/?view_only=b2289b6fec214fa88ee75a18d45c18f3}.
-Because in this study I analyze data from an already existing large-scale data set, all of these steps were preregistered prior to accessing the data.
+Because in this study I analyzed data from an already existing large-scale data set, all of these steps were done prior to accessing the data.
 The preregistration was designed on the basis of the panel documentation online (Kittel et al., 2020).
-In some cases the analyses could not be executed as originally planned, for example because some properties of the variables became apparent only upon data analysis.
-The most relevant deviations are reported below, and a complete list of all changes can be found in the online companion website (\url{https://xmtra.github.io/Austrian_Corona_Panel_Project/index.html}).
+In some cases I couldn't execute the analyses as I had originally planned, for example because some properties of the variables only became apparent when inspecting the actual data.
+The most relevant deviations are reported below, and a complete list of all changes can be found in the online \href{https://XMtRA.github.io/Austrian_Corona_Panel_Project}{companion website} (\url{https://XMtRA.github.io/Austrian_Corona_Panel_Project}).
 
@@
 The study was conducted between March 2020 and October 2021.
-It contains 26 waves, and at the time of writing, the first 24 waves were available for download.
+It contains 26 waves, and at the time of writing the first 24 waves were available for download.
 Each wave consists of at least 1,500 respondents.
@@
 The sample was representative of the Austrian population in terms of age, gender, region/state, municipality size, and educational level.
-In order to participate in the study, the respondents needed to be Austrian residents and had to be at least 14 years old.
+In order to participate in the study, the respondents needed to be Austrian residents and had to be at least 14 years of age.
+The study received IRB approval from the University of Vienna.
 The average age was 42 years, 49 percent were male, 14 percent had a University degree, and 5 percent were currently unemployed.
 
+\hypertarget{inference-criteria}{%
+\subsection{Inference Criteria}\label{inference-criteria}}
+
 Because the data were analyzed post-hoc, no a-priori sample size planning on the basis of power analyses was conducted.
-The sample is very large, and it is hence likely well-equipped reliably to detect also small effects.
-In addition, because such large samples easily generate significant \emph{p}-values even for very small effects, it helps that to test hypothesis this study uses a smallest effect size of interest-approach.
-To test the hypotheses, I will use the interval testing approach as proposed by Dienes (2014).
-On the basis of the SESOI, I will define a null region.
-For well-being, the null region will be between \emph{b} = -.30 and \emph{b} = .30.
-For example, if the 95\% confidence interval lies completely within the null-region (e.g., \emph{b} = .15, {[}95\% CI: .05, .25{]}), the hypothesis that the effect is only trivial is supported.
-If the effects interval and the null region overlap (e.g., \emph{b} = .25, {[}95\% CI: .15, .35{]}), the hypothesis is not supported and the results are considered inconclusive, whereas a meaningful negative effect is rejected.
-If the confidence falls completely outside of the null-region (e.g., \emph{b} = .40, {[}95\% CI: .35, .45{]}), the hypothesis is rejected and the existence of a meaningful positive effect is supported.
+The sample is large, and it is hence well-equipped reliably to detect also small effects.
+In addition, because such large samples easily generate significant \emph{p}-values even for very small effects, it helps that the hypotheses were tested with a smallest effect size of interest-approach.
+To this end, I adopted the interval testing approach as proposed by Dienes (2014).
+On the basis of the SESOI, I then defined a null region.
+In what follows, I explain how I determined the SESOI and the null region.
+
+In this study, life satisfaction was measured on an 11-point scale.
+If people can reliably differentiate 7 levels as mentioned above, this corresponds to 11 / 7 = 1.57 unit change on an 11-point scale.
+Hence, a four-point change in media use (e.g., a complete stop) should result in a 1.57-point change in life satisfaction.
+In a statistical regression analysis, \emph{b} estimates the change in the dependent variable if the independent variable increases by one point.
+We would therefore expect a SESOI of \emph{b} = 1.57 / 4 = 0.39.
+For affect, which was measured on a 5-point scale, our SESOI would be \emph{b} = 0.71 / 4 = 0.18.
+Because we're agnostic as to whether the effects are positive or negative, the null region includes negative and positive effects.
+Finally, in order not to exaggerate precision and to be less conservative, these numbers are reduced to nearby thresholds.\footnote{Note that other researchers also decreased or recommended decreasing thresholds for effect sizes when analyzing within-person or cumulative effects (Beyens, Pouwels, Driel, Keijsers, \& Valkenburg, 2021; Funder \& Ozer, 2019).}
+Together, this leads to a null region ranging from \emph{b} = -.30 to \emph{b} = .30 for life satisfaction, and \emph{b} = -.15 to \emph{b} = .15 for positive and negative affect.
+
+Let's briefly illustrate what this means in practice.
+If the 95\% confidence interval falls completely within the null-region (e.g., \emph{b} = .20, {[}95\% CI: .15, .25{]}), the hypothesis that the effect is trivial is supported.
+If the confidence interval and the null region overlap (e.g., \emph{b} = .30, {[}95\% CI: .25, .35{]}), the hypothesis is not supported and the results are considered inconclusive, while a meaningful negative effect is rejected.
+If the confidence interval falls completely outside of the null-region (e.g., \emph{b} = .40, {[}95\% CI: .35, .45{]}), the hypothesis is rejected and the existence of a meaningful positive effect is supported.
 
@@
 Three models were run, one for each dependent variable.
-Predictors included all types and channels of social media communication separated into within and between-person factors, as well as stable and varying covariates.
+The data were hierarchical, and responses were separately nested in participants and waves (i.e., participants and waves were implemented as random effects).
+Nesting in participants allowed to separate between-person relations from within-person effects.
+Nesting in waves allowed to control for general exogenous developments, such as general decreases in well-being in the population, for example due to lockdown measures.
+Thus, there was no need additionally to control for specific phases or measures of the lockdown.
+Predictors were modeled as fixed effects.
+They included social media communication types and channels, separated into within and between-person factors, as well as stable and varying covariates.
 All predictors were included simultaneously and in each of the three models.
-The data were hierarchical, and responses were separately nested in participants and waves (put differently, participants and waves were implemented as random effects).
-Nesting in participants allowed to separate between-person relations from within-person effects.
-Nesting in waves allowed to control for general exogenous developments, such as general decrease in well-being in the population, for example due to lockdown measures.
-Thus, there was no need additionally to control for specific phases or measures of the lockdown.
 
-For more information on the analyses, for example a complete documentation of the models and their results, see \href{https://xmtra.github.io/Austrian_Corona_Panel_Project/index.html}{companion website}.
+The factorial validity of the scales were tested with confirmatory factor analyses (CFA).
+Because Mardia's test showed that the assumption of multivariate normality was violated, I used the more robust Satorra-Bentler scaled and mean-adjusted test statistic (MLM) as estimator.
+To avoid overfitting, I tested the scales on more liberal fit criteria (CFI \textgreater{} .90, TLI \textgreater{} .90, RMSEA \textless. .10, SRMR \textless{} .10) (Kline, 2016).
+Finally, REWB-models cannot model latent variables.
+To increase precision, I therefore exported factor scores from the CFAs for positive and negative affect.
+Respondents who answered less than 50\% of all questions were removed.
+The remaining missing responses were imputed using predictive mean matching.
+
+For more information on the analyses, a complete documentation of the models and results, see \href{https://XMtRA.github.io/Austrian_Corona_Panel_Project}{companion website}.
 
@@
 
-In what follows, I briefly list all variables that were collected.
+In what follows, I list all the variables that I analyzed.
 For the variables' means, range, and variance, see Table \ref{tab:descriptives}.
-For a complete list of all items and item characteristics, see \href{https://xmtra.github.io/Austrian_Corona_Panel_Project/index.html}{companion website}.
+For a complete list of all items and item characteristics, see \href{https://XMtRA.github.io/Austrian_Corona_Panel_Project}{companion website}.
 
@@
 The response options were 1 (\emph{never}), 2 (\emph{on some days}), 3 (\emph{several times per week}), 4 (\emph{almost every day}), and 5 (\emph{daily}).
+The scale showed good factorial fit, \(\chi^2\)(46) = 65.30, \emph{p} = .032, CFI = 1.00, RMSEA = .02, 90\% CI {[}.01, .03{]}, SRMR = .01.
+Reliability was high, \(\omega\) = .85.
 
@@
 The response options were 1 (\emph{never}), 2 (\emph{on some days}), 3 (\emph{several times per week}), 4 (\emph{almost every day}), and 5 (\emph{daily}).
+The scale showed good factorial fit, \(\chi^2\)(331) = 3138.37, \emph{p} \textless{} .001, CFI = .97, RMSEA = .08, 90\% CI {[}.07, .08{]}, SRMR = .03.
+Reliability was high, \(\omega\) = .89.
 
@@
 
-COVID-19 related social media use focused on types of communication was measured with three variables (a) reading, (b) liking and sharing, and (c) posting.
+COVID-19 related social media use focused on communication types was measured with the three dimensions of (a) reading, (b) liking and sharing, and (c) posting.
 The general introductory question was ``How often during the last week have you engaged in the following activities on social media?''
-The three items read as follows:
-``Reading the posts of others with content on the Coronavirus''; ``When seeing posts on the Coronavirus, I clicked `like,' `share' or `retweet'\,''; ``I myself wrote posts on the Coronavirus on social media.''
+The three items were ``Reading the posts of others with content on the Coronavirus,'' ``When seeing posts on the Coronavirus, I clicked `like,' `share' or `retweet','' ``I myself wrote posts on the Coronavirus on social media.''
 Answer options were 1 (\emph{several times per day}), 2 (\emph{daily}), 3 (\emph{several times per week}), 4 (\emph{weekly}), 5 (\emph{never}).
@@
 Social media use was measured for all participants on waves 1, 2, 8, 17, and 23.
-Freshly recruited respondents always answered the questions on social media use.
+Freshly recruited respondents always answered all questions on social media use.
 
@@
 Well-being &  &  &  & \\
-\ \ \ Life satisfaction & 1.68 & 6.39 & 6.79 & 6.59\\
-\ \ \ Positive affect & 0.57 & 3.05 & 3.29 & 3.15\\
+\ \ \ Life satisfaction & 1.68 & 6.38 & 6.81 & 6.59\\
+\ \ \ Positive affect & 0.57 & 3.05 & 3.28 & 3.15\\
 \ \ \ Negative affect & 0.39 & 1.66 & 1.81 & 1.73\\
 Social media use &  &  &  & \\
-\ \ \ Read & 1.03 & 2.09 & 2.92 & 2.42\\
-\ \ \ Like \& share & 0.86 & 1.61 & 1.92 & 1.78\\
-\ \ \ Posting & 0.63 & 1.33 & 1.47 & 1.39\\
+\ \ \ Read & 1.03 & 2.10 & 2.92 & 2.43\\
+\ \ \ Like \& share & 0.87 & 1.61 & 1.92 & 1.78\\
+\ \ \ Posting & 0.63 & 1.33 & 1.46 & 1.39\\
 Social media channel &  &  &  & \\
 \ \ \ Facebook & 0.96 & 2.34 & 2.68 & 2.45\\
-\ \ \ Twitter & 0.52 & 1.16 & 1.72 & 1.36\\
-\ \ \ Instagram & 0.82 & 1.85 & 2.65 & 2.09\\
-\ \ \ WhatsApp & 1.23 & 2.29 & 2.62 & 2.46\\
-\ \ \ YouTube & 0.88 & 1.77 & 2.32 & 2.01\\
+\ \ \ Twitter & 0.52 & 1.16 & 1.72 & 1.37\\
+\ \ \ Instagram & 0.83 & 1.84 & 2.66 & 2.09\\
+\ \ \ WhatsApp & 1.23 & 2.28 & 2.62 & 2.45\\
+\ \ \ YouTube & 0.88 & 1.77 & 2.30 & 2.00\\
 \bottomrule
@@
 The effects of COVID-19 related social media use were controlled for the following stable variables:
-(a) gender (answer options: female, male, diverse), (b) age, (c) education (ten options), (d) Austria country of birth (yes/no), (e) Austria parents' country of birth (no parent, one parent, both parents).
-I originally planned to implement other variables as varying covariates.
+(a) gender (female, male, diverse), (b) age, (c) education (ten options), (d) Austria country of birth (yes/no), (e) Austria parents' country of birth (no parent, one parent, both parents).
+I originally planned to implement additional variables as varying covariates.
 However, because they were not measured often enough or not at the time when social media use was measured, I implemented them as stable variables using their average values across all waves.
-This includes (a) text-based media news consumption (five degrees), (b) video-based media news consumption (five degrees), (c) residency is Vienna (yes/no), (d) self-reported physical health (five degrees), (e) living space (in squaremeter), (f) access to balcony (yes/no), (g) access to garden (yes/no), (h) employment (nine options), (i) disposition to take risks (eleven degrees), and (j) locus of control (five degrees).
-I controlled also for the following varying covariates: (a) five items measuring outdoor activities such as sport or meeting friends (five degrees), and (b) satisfaction with democracy (five degrees).
+This includes (a) text-based media news consumption (five degrees), (b) video-based media news consumption (five degrees), (c) residency is Vienna (yes/no), (d) self-reported physical health (five degrees), (e) living space (eleven options), (f) access to balcony (yes/no), (g) access to garden (yes/no), (h) employment (nine options), (i) disposition to take risks (eleven degrees), and (j) locus of control (five degrees).
+I also controlled for the following varying covariates: (a) five items measuring outdoor activities such as sport or meeting friends (five degrees), and (b) satisfaction with democracy (five degrees).
 Because it lead to too much attrition in the sample, I did not control for (a) household size, (b) work hours per week, (c) home office, (d) household income.
@@
 
-\hypertarget{preregistered-analyses}{%
-\subsection{Preregistered Analyses}\label{preregistered-analyses}}
-
-When looking at the variables from a descriptive perspective, we see that all well-being measures did not change substantially across the different waves of data collection.
+First, when looking at the variables from a descriptive perspective, we see that all well-being measures did not change substantially across the different waves of data collection.
 COVID-19 related media use, however, decreased slightly at the beginning of the study and remained stable after approximately six waves.
 The initial decrease might be explained by the fact that the collection of data began at the end of March 2020, hence approximately three months after the pandemic began.
-It could be that after an initial uptick, COVID-19 related social media use was already declining at that time, returning to more normal levels.
+It could be that after an initial uptick, COVID-19 related social media use was already declining at the time, returning to more normal levels.
 
@@
 
-The study's hypothesis was that the effects of social media use on well-being would be trivial.
-Regarding the effects of different \emph{types} of communication---that is, reading vs.~sharing vs.~posting---all within-person effects fell completely within the a-priori defined SESOIs (see Figure \ref{fig:res-activity}).
-For example, respondents who used social media more frequently than usual to read about COVID-19 related topics did not show a simultaneous change in life satisfaction (\emph{b} = 0.04 {[}95\% CI -0.02, 0.09{]}).
-All confidence intervals include zero; hence, all effects were also statistically non-significant.
+\hypertarget{preregistered-analyses}{%
+\subsection{Preregistered Analyses}\label{preregistered-analyses}}
+
+The study's main hypothesis was that the effects of social media use on well-being would be trivial.
+Regarding the effects of different communication \emph{types}---that is, reading vs.~sharing vs.~posting---all within-person effects fell completely within the a-priori defined null region (see Figure \ref{fig:res-activity}).
+For example, respondents who used social media more frequently than usual to read about COVID-19 related topics did not show a simultaneous change in life satisfaction (\emph{b} = 0.05 {[}95\% CI -0.01, 0.1{]}).
+All confidence intervals included zero; hence, all effects were also statistically non-significant.
 As a result, the hypothesis was supported for all COVID-19 related types of social media communication.
@@
 \includegraphics[width=\textwidth]{figures/fig_results_activity} \caption{The effects of various types of social media use on three indicators of well-being.
-Effects are controlled for a large number of covariates (see text).
-The SESOI was |0.30| for life satisfaction and |0.15| for affects.
-Hence, none of the reported effects are considered meaningful.}\label{fig:res-activity}
+The black estimates show the effects controlled for a large number of covariates (see text; preregistered); the grey estimates are without control variables (exploratory).
+The SESOI was b = |0.30| for life satisfaction and b = |0.15| for affect.
+Hence, all of the reported effects are not considered meaningful.}\label{fig:res-activity}
 \end{figure}
 
-Regarding between-person relations, about which no hypotheses were formulated, only one effect didn't include zero.
-Respondents who across all waves used social media more frequently than others to write COVID-19 related posts reported higher levels of negative affect than others (\emph{b} = 0.08 {[}95\% CI 0.04, 0.12{]}).
-The effect was still completely inside of the null region, hence not considered large enough to be practically relevant.
+Regarding between-person relations, about which no hypotheses were formulated, only three effects did not include zero.
+Respondents who across all waves used social media more frequently than others to read about COVID-19 related posts reported slightly lower levels of positive affect than others (\emph{b} = -0.03 {[}95\% CI \textgreater{} -0.01, -0.06{]}).
+Respondents who across all waves used social media more frequently than others to write COVID-19 related posts reported higher levels of negative affect than others (\emph{b} = 0.06 {[}95\% CI 0.09, 0.03{]}).
+Interestingly, respondents who across all waves used social media more frequently than others to write COVID-19 related posts also reported slightly higher levels of positive affect than others (\emph{b} = 0.05 {[}95\% CI 0.09, \textless{} 0.01{]}).
+However, note that the effect were still completely inside of the null region, hence not large enough to be considered practically relevant.
+
+Note that when comparing the results with and without control variables, the results differed.
+For example, on the between-person level, one effect stopped being significant if controlled for additional variables.
+Actively posting on social media was significantly (though not meaningfully) related to decreased life satisfaction.
+However, when controlling for potential confounders, the effect became virtually zero (see Figure \ref{fig:res-channels}).
 
@@
 \includegraphics[width=\textwidth]{figures/fig_results_channel} \caption{The effects of using various social media applications on three indicators of well-being.
-Effects are controlled for a large number of covariates (see text).
-The SESOI was |0.30| for life satisfaction and |0.15| for affects.
-Hence, none of the reported effects are considered meaningful.}\label{fig:res-channels}
+The black estimates show the effects controlled for a large number of covariates (see text; preregistered); the grey estimates are without control variables (exploratory).
+The SESOI was b = |0.30| for life satisfaction and b = |0.15| for affect.
+Hence, all of the reported effects are not considered meaningful.}\label{fig:res-channels}
 \end{figure}
 
-Regarding the COVID-19 related use of social media \emph{channels}, the results were very comparable (see Figure \ref{fig:res-channels}).
+Regarding the COVID-19 related use of social media \emph{channels}, the results were comparable (see Figure \ref{fig:res-channels}).
 Changes in the frequency of using different social media channels to attain information regarding COVID-19 were unrelated to meaningful changes in well-being.
-For example, respondents who used Facebook more frequently than usual to learn about COVID-19 did not show a simultaneous change in well-being (\emph{b} = -0.04 {[}95\% CI -0.1, 0.02{]}).
+For example, respondents who used Facebook more frequently than usual to learn about COVID-19 did not show a simultaneous change in well-being (\emph{b} = -0.05 {[}95\% CI -0.11, 0.01{]}).
 Only two effects differed substantially from zero.
-Respondents who used Instagram more frequently than usual to attain COVID-19 related news reported slightly higher simultaneous levels of life satisfaction than usual (\emph{b} = 0.08 {[}95\% CI \textless{} 0.01, 0.15{]}).
-Respondents who used Twitter more frequently than usual to attain COVID-19 related news reported somewhat lower simultaneous levels of life satisfaction than usual (\emph{b} = -0.15 {[}95\% CI -0.27, -0.04{]}).
-However, both effects were still completely inside of the null region, hence not considered large enough to be meaningful.
+Respondents who used Instagram more frequently than usual to attain COVID-19 related news reported slightly \emph{higher} levels of life satisfaction than usual (\emph{b} = 0.09 {[}95\% CI 0.02, 0.16{]}).
+Respondents who used Twitter more frequently than usual to attain COVID-19 related news reported slightly \emph{lower} levels of life satisfaction than usual (\emph{b} = -0.12 {[}95\% CI -0.23, -0.02{]}).
+However, both effects were still completely inside of the null region, hence not large enough to be considered meaningful.
 In sum, the hypothesis was supported also for the COVID-19 related use of important social media channels.
@@
 Only one relation did not include zero, was hence statistically significant.
-Respondents who across all waves used YouTube more frequently than others for COVID-19 related reasons reported marginally higher levels of negative affect (\emph{b} = 0.03 {[}95\% CI \textless{} 0.01, 0.05{]}).
-However, please note that this effect again was not considered large enough to be practically relevant.
+Respondents who across all waves used YouTube more frequently than others for COVID-19 related reasons reported marginally \emph{higher} levels of life satisfaction (\emph{b} = 0.08 {[}95\% CI \textless{} 0.01, 0.16{]}).
+However, please note that this effect again was not large enough to be considered practically relevant.
+
+Again, note that when comparing the results with and without control variables, the results differed.
+Especially on the between-person level, altogether five effects stopped being significant if they were controlled for additional variables.
+For example, using Instagram was significantly (though not meaningfully) related to increased life satisfaction.
+However, when controlling for additional covariates, the effect became virtually zero (see Figure \ref{fig:res-channels}).
 
@@
 
-In what follows, I briefly report also the results of some covariates.
-Several variables showed larger associations with well-being.
-Note because each variable has a different scaling, we would again need to define a SESOI for each variable, which cannot be implemented here.
-But note that several variables would likely fall outside of such a SESOI.
-This includes for example internal locus of control, health, or employment.
-For a brief overview, see Figure \ref{fig:res-control}.
+In what follows, I briefly report some exploratory analyses that weren't preregistered.
+First, to contextualize the results reported above and to see if the results included any meaningful effects at all, I also looked at the effect sizes of selected (cherry-picked) covariates.
+Because each variable had different response options, we would actually need to define a SESOI for each variable, which for reasons of scope I cannot implement here.
+Therefore, I report the results of the standardized scales,
+which allows for a better comparison across the differently scaled variables.
+For what it's worth, as a rough estimate for the SESOI we can build on the typical convention that small effects start at \emph{r} = \textbar.10\textbar.
+The results showed that several effects fell outside of the SESOI, were hence considered meaningful.
+This includes for example internal locus of control, health, satisfaction with democracy, or exercising.
+For an overview, see Figure \ref{fig:res-control}.
 
 \begin{figure}
-\includegraphics[width=\textwidth]{figures/fig_results_control} \caption{Results of selected covariates.}\label{fig:res-control}
+\includegraphics[width=\textwidth]{figures/fig_results_control_std} \caption{Results of selected covariates.
+All variables were standardize except 'Male'  and ‘Employed in public service', because there were measured on a binary scale.}\label{fig:res-control}
 \end{figure}
+
+To find out whether my inferences were robust across legitimate (though arguably inferior) alternative analyses, I reran the analyses also using standardized estimates, mean scores instead of factor scores, and with a data set where missing data were not imputed.
+The results were virtually the same.
+For example, all standardized COVID-19 related types of social media use or channels were not significantly larger than a SESOI of \(\beta\) = \textbar.10\textbar.
+The additional analyses are reported in the \href{https://XMtRA.github.io/Austrian_Corona_Panel_Project/analyses_additional.html}{companion website}.
 
@@
 In this study I analyzed the effects of COVID-19 related social media use on well-being.
-The data come from a panel study with 24 waves and is representative of the Austrian population.
-A random effects model, which separated between person relations from within-person effects and which controlled for several third variables, showed that within-person effects were trivial.
-People who used social media more than usual to learn about COVID-19 did not show changes in their well-being.
+The data come from a panel study with 24 waves and are representative of the Austrian population.
+In a random effects model I separated between person relations from within-person effects and controlled for a large number of both stable and varying covariates, thereby aiming to assess causality.
+The results showed that within-person effects were trivial.
+People who used social media more than usual to learn about COVID-19 didn't show meaningful changes in their well-being.
 
-The results imply that COVID-19 related social media use does not seem to be particularly relevant for people's well-being.
-Other factors among the third variables that were measured revealed larger effects or relations, implying that well-being is rather determined by aspects such as health, employment, or locus of control.
-According to this study, popular fears that ``doomscrolling'' or overusing social media during times of crises do not seem to be justified.
+The results imply that COVID-19 related social media use doesn't seem to be particularly relevant for well-being.
+Other factors among the third variables that were measured revealed larger effects or relations, suggesting that well-being is determined by alternative aspects such as health, satisfaction with democracy, locus of control, or exercising.
+According to this study, popular fears that ``doomscrolling'' or overusing social media during times of crises don't seem to be justified.
 
-The study is not aligned with several recent studies on this particular research question or closely related research questions.
-This includes a study by Bendau et al. (2021), which showed negative relations between social media and well-being.
-However, note that Bendau et al. (2021) analyzed cross-sectional data on a between-person level, while not controlling for third variables, which is not optimal for investigating causal effects.
-At the same time, this study is well-aligned with recent studies and meta-analyses analyzing the effects of social media use from a more general perspective or that adopted a somewhat different angle.
+On the one hand, the results are not aligned with several recent studies analyzing similar or closely related research questions.
+This includes a study by Bendau et al. (2021), which showed negative relations between social media and well-being (but see Bradley \& Howard, 2021; or Sewall, Goldstein, \& Rosen, 2021).
+However, note that Bendau et al. (2021) analyzed cross-sectional data on a between-person level while not controlling for third variables, which is not optimal for investigating causal effects.
+On the other hand, the results are well-aligned with recent studies and meta-analyses analyzing the effects of social media use from a more general perspective or from a somewhat different angle.
 These studies have found that the effects of various types of social media use on several well-being indicators are small at best, often too small to matter (Ferguson et al., 2021; Meier \& Reinecke, 2020; Orben, 2020), which echoes the results obtained here.
+
+If anything, two preliminary and subtle trends can be observed.
+First, of all the three COVID-19 related social media activities, people who read about the pandemic more than others showed slightly decreased levels of positive affect, and people who actively posted about the pandemic more than others showed slightly increased levels of negative affect.
+On the other hand, however, people who posted more about COVID-19 also showed slightly higher levels of positive affects, so taken together the results are ambivalent.
+Second, in terms of media channels, using Twitter more than usual was related to slightly decreased levels of life satisfaction.
+Twitter is considered to have more negative affordances and tonality as compared to other networks such as Instagram (Waterloo, Baumgartner, Peter, \& Valkenburg, 2018), which might help explain the results.
+Instagram, on the other hand, was related to slightly increased levels of life satisfaction.
+To speculate, the often-criticized positivity bias on Instagram might have been somewhat beneficial in times of the pandemic.
+That said, all these effects were still very small and arguably too small to matter.
+But future research might elaborate on these specific relations to probe their stability and relevance.
+
+Finally, another interesting observation is that life satisfaction was remarkably stable.
+Hence, even in times of a pandemic, it seems that such broad assessment of life vary only mildly.
+This supports the hypothesis that life satisfaction seems to be determined largely by stable factors such as one's genes (Brown \& Rohrer, 2019).
 
@@
 
-The current study analyzed whether changes in media use were related to changes in well-being, while controlling for several potential confounds.
-Together, this allows for a good perspective on potential causality.
+The current study analyzed whether changes in media use were related to changes in well-being, while controlling for several potential confounders.
+Together, this allows for an improved perspective on assessing causality.
 That said, causality necessitates temporal order, and the cause needs to precede the effect.
 Regarding media use, such effects often happen immediately or shortly after use, necessitating intervals in the hours, minutes, or even seconds.
-In many cases only experience sampling studies that ask users in the very moment can produce such knowledge.
+In many cases only experience sampling studies asking users in the very moment can produce such knowledge.
 However, even then we don't know for certain if we actually measured the right interval.
+Effects depend on the intensity of use or the length of the interval, and to borrow the words from Rohrer and Murayama (2021), there is no such thing as ``the'' effect of social media use on well-being.
 Hence, to document how effects unfold, future research needs to employ different study designs probing different time lags.
-In addition, more thought needs to be invested in what relevant stable and varying factors should be selected as control variables.
+In addition, more thought needs to be invested in what relevant stable and varying factors we should include as control variables, and I hope this study provides a first step into this direction.
 
-Although I had already reduced the predefined SESOIs to be less conservative, potentially they were still too large.
+Although I had already reduced the predefined SESOIs to be less conservative, they were potentially still too large.
 Media use is only one aspect of several factors that simultaneously affect well-being.
-Is it realistic that extremely changing only \emph{one} of these aspects (e.g., by completely stopping the use of social media) should already manifest in a detectable change in well-being?
-Or would it make more sense to expect that if people regularly start doing say \emph{two} activities (e.g.~regularly exercising \emph{and} establishing a reading habit) together should show perceivable improvements to well-being?
-In other words, the beneficial effect of a particular activity is large enough if people feel a difference when implementing two of those activities.
-Practically, this would imply a SESOI half as large as I have defined here, that is \emph{b} = \textbar.15\textbar{} for well-being and \emph{b} = \textbar.075\textbar.
-In this case, this would not make a difference, as even with these more liberal thresholds all but one effect would still be completely in the null region.
-However, at all events future research needs to start a thorough discussion on what effect sizes are considered meaningful and relevant, and with this study I hope to provide some helpful input and first concrete guidelines.
+Is it really realistic to expect that extremely changing only \emph{one} of these aspects should manifest in a detectable change in well-being?
+Or would it make more sense to expect that thoroughly committing to say \emph{two} activities (e.g.~regularly exercising \emph{and} establishing a reading habit) should then cause a detectable improvement in well-being?
+Practically, this would imply a SESOI half as large as I have defined here, namely \emph{b} = \textbar.15\textbar{} for well-being and \emph{b} = \textbar.075\textbar{} for affect.
+In the case of this study, however, reducing the SESOI would not even make a big difference, as also with these more liberal thresholds all but three effect would still be completely in the null region, and no effect would be outside of the null region.
+However, at all events I encourage future research to start a thorough conversation on what effect sizes are considered meaningful and what not.
+Again, with this study I hope to provide some first input and guidelines.
 
@@
 Measuring well-being with self-reports is adequate, because it by definition requires introspection.
-However, it would be preferable to measure social media use objectively, because people cannot reliably estimate their use.
-That said, objective measures often cannot capture the content or the motivation of the use, and only very complicated tools that record the content that was used (such as the Screenome project) might produce such data.
-Unfortunately, such procedures introduce other problems, for example related to privacy.
-Hence, for this type of research question it seems necessary still to use self-reported measures.
+However, it would be preferable to measure social media use objectively, because people cannot reliably estimate their use (Scharkow, 2016).
+That said, objective measures often cannot capture the content or the motivation of the use, and only very complicated tools recording the actual content (such as the Screenome project) might produce such data.
+Unfortunately, such procedures introduce other problems, especially related to privacy.
+Hence, for this type of research question it still seems necessary to use self-reported measures.
 
 Because the data were collected in a single country, the generalizability of the results is limited.
-The results potentially apply primarily to the more Western sphere, and might not hold true in other cultures, especially cultures with a different media landscape or alternative social media channels.
-That said, because this is a comparatively large study representative of a country's entire population, and because several waves were collected across a large time span, the results should be at least as generalizable as other typical empirical studies collected in the social sciences.
+The results apply primarily to the more Western sphere, and might not hold true in other cultures, especially cultures with a different media landscape or alternative social media channels.
+That said, because this is a comparatively large study representative of an entire country, and because several waves were collected across a large time span, the results should be at least as generalizable as other typical empirical studies collected in the social sciences.
 
@@
 (2021).
-Social {Media} {Use} and {Adolescents}' {Well}-{Being}: {Developing} a {Typology} of {Person}-{Specific} {Effect} {Patterns}.
+Social media use and adolescents' well-being: {Developing} a typology of person-specific effect patterns.
 \emph{Communication Research}.
 \url{https://doi.org/10.31234/osf.io/ftygp}
+
+\leavevmode\hypertarget{ref-bradleyStressMoodSmartphone2021}{}%
+Bradley, A., \& Howard, A. (2021). \emph{Stress, mood, and smartphone use in {University} students: {A} 12-week longitudinal study}. OSF Preprints. \url{https://doi.org/10.31219/osf.io/frvpb}
+
+\leavevmode\hypertarget{ref-brownEasyHappinessPie2019}{}%
+Brown, N.
+J.
+L., \& Rohrer, J.
+M.
+(2019).
+Easy as (happiness) pie?
+{A} critical evaluation of a popular model of the determinants of well-being.
+\emph{Journal of Happiness Studies}.
+\url{https://doi.org/10.1007/s10902-019-00128-4}
 
@@
 
+\leavevmode\hypertarget{ref-dornemannHowGoodBad2021}{}%
+Dörnemann, A., Boenisch, N., Schommer, L., Winkelhorst, L., \& Wingen, T.
+(2021).
+\emph{How do good and bad news impact mood during the {Covid}-19 pandemic?
+{The} role of similarity}.
+OSF Preprints.
+\url{https://doi.org/10.31219/osf.io/sy2kd}
+
 \leavevmode\hypertarget{ref-edenMediaCopingCOVID192020}{}%
@@
 (2020).
-Media for {Coping} {During} {COVID}-19 {Social} {Distancing}: {Stress}, {Anxiety}, and {Psychological} {Well}-{Being}.
+Media for coping during {COVID}-19 social distancing: {Stress}, anxiety, and psychological well-being.
 \emph{Frontiers in Psychology}, \emph{11}, 577639.
@@
 
+\leavevmode\hypertarget{ref-hamakerWhyResearchersShould2014}{}%
+Hamaker, E.
+L.
+(2014).
+Why researchers should think "within-person": {A} paradigmatic rationale.
+In M.
+R.
+Mehl, T.
+S.
+Conner, \& M.
+Csikszentmihalyi (Eds.), \emph{Handbook of research methods for studying daily life} (Paperback ed.).
+New York, NY: Guilford.
+
 \leavevmode\hypertarget{ref-huangTimeSpentSocial2017}{}%
@@
 \url{https://doi.org/10.1089/cyber.2016.0758}
-
-\leavevmode\hypertarget{ref-johannesNoEffectDifferent2021}{}%
-Johannes, N., Dienlin, T., Bakhshi, H., \& Przybylski, A. K. (2021). \emph{No effect of different types of media on well-being}. PsyArXiv. \url{https://doi.org/10.31234/osf.io/zgb5y}
 
@@
 
+\leavevmode\hypertarget{ref-liuRelationOfficialWhatsAppdistributed2020}{}%
+Liu, J.
+C.
+J., \& Tong, E.
+M.
+W.
+(2020).
+The relation between official {WhatsApp}-distributed {COVID}-19 news exposure and psychological symptoms: Cross-sectional survey study.
+\emph{Journal of Medical Internet Research}, \emph{22}(9), e22142.
+\url{https://doi.org/10.2196/22142}
+
 \leavevmode\hypertarget{ref-mcelreathYesterdayClass2021}{}%
@@
 
-\leavevmode\hypertarget{ref-meierComputerMediatedCommunicationSocial2020}{}%
+\leavevmode\hypertarget{ref-meierComputermediatedCommunicationSocial2020a}{}%
 Meier, A., \& Reinecke, L.
 (2020).
-Computer-{Mediated} {Communication}, {Social} {Media}, and {Mental} {Health}: {A} {Conceptual} and {Empirical} {Meta}-{Review}.
+Computer-mediated communication, social media, and mental health: {A} conceptual and empirical meta-review.
 \emph{Communication Research}, 009365022095822.
@@
 (2021).
-Does {Taking} a {Short} {Break} from {Social} {Media} {Have} a {Positive} {Effect} on {Well}-being?
-{Evidence} from {Three} {Preregistered} {Field} {Experiments}.
+Does taking a short break from social media have a positive effect on well-being?
+{Evidence} from three preregistered field experiments.
 \emph{Journal of Technology in Behavioral Science}, \emph{6}(3), 507--514.
@@
 Przybylski, A. K., \& Weinstein, N. (2017). A large-scale test of the {Goldilocks} hypothesis. \emph{Psychological Science}, \emph{28}(2), 204--215. \url{https://doi.org/10.1177/0956797616678438}
+
+\leavevmode\hypertarget{ref-riehmAssociationsMediaExposure2020}{}%
+Riehm, K.
+E., Holingue, C., Kalb, L.
+G., Bennett, D., Kapteyn, A., Jiang, Q., \ldots{} Thrul, J.
+(2020).
+Associations between media exposure and mental distress among {U}.{S}.
+Adults at the beginning of the {COVID}-19 pandemic.
+\emph{American Journal of Preventive Medicine}, \emph{59}(5), 630--638.
+\url{https://doi.org/10.1016/j.amepre.2020.06.008}
 
@@
 
+\leavevmode\hypertarget{ref-rohrerTheseAreNot2021}{}%
+Rohrer, J.
+M., \& Murayama, K.
+(2021).
+\emph{These are not the effects you are looking for: {Causality} and the within-/between-person distinction in longitudinal data analysis} {[}Preprint{]}.
+PsyArXiv.
+\url{https://doi.org/10.31234/osf.io/tg4vj}
+
 \leavevmode\hypertarget{ref-sandstromDoomscrollingCOVIDNews2021}{}%
@@
 
-\leavevmode\hypertarget{ref-scharkowHowSocialNetwork2020}{}%
-Scharkow, M., Mangold, F., Stier, S., \& Breuer, J.
-(2020).
-How social network sites and other online intermediaries increase exposure to news.
-\emph{Proceedings of the National Academy of Sciences}, \emph{117}(6), 2761--2763.
-\url{https://doi.org/10.1073/pnas.1918279117}
+\leavevmode\hypertarget{ref-scharkowAccuracySelfreportedInternet2016}{}%
+Scharkow, M.
+(2016).
+The accuracy of self-reported {Internet} use---{A} validation study using client log data.
+\emph{Communication Methods and Measures}, \emph{10}(1), 13--27.
+\url{https://doi.org/10.1080/19312458.2015.1118446}
 
@@
 \url{https://doi.org/10.1093/jcmc/zmaa014}
+
+\leavevmode\hypertarget{ref-schnauber-stockmannMobileDevicesTools2020}{}%
+Schnauber-Stockmann, A., \& Karnowski, V.
+(2020).
+Mobile devices as tools for media and communication research: {A} scoping review on collecting self-report data in repeated measurement designs.
+\emph{Communication Methods and Measures}, \emph{14}(3), 145--164.
+\url{https://doi.org/10.1080/19312458.2020.1784402}
+
+\leavevmode\hypertarget{ref-sewallObjectivelyMeasuredDigital2021}{}%
+Sewall, C.
+J.
+R., Goldstein, T.
+R., \& Rosen, D.
+(2021).
+Objectively measured digital technology use during the {COVID}-19 pandemic: {Impact} on depression, anxiety, and suicidal ideation among young adults.
+\emph{Journal of Affective Disorders}, \emph{288}, 145--147.
+\url{https://doi.org/10.1016/j.jad.2021.04.008}
 
@@
 
-All the stimuli, presentation materials, analysis scripts, and a reproducible version of the manuscript can be found on the companion website (\url{https://xmtra.github.io/Austrian_Corona_Panel_Project/index.html}).
+All the stimuli, presentation materials, analysis scripts, and a reproducible version of the manuscript can be found on the companion website (\url{https://XMtRA.github.io/Austrian_Corona_Panel_Project}).
 
@@
 
+\hypertarget{acknowledgements}{%
+\section{Acknowledgements}\label{acknowledgements}}
+
+I would like to thank BLINDED for providing valuable feedback on this manuscript.
+
 \end{document}

### 2022-04-11 analyses.RmD
- deleted data

@@
 set.seed(170819)
-knitr::opts_chunk$set(warning=F, echo=T, error=F, message=F, cache = T)
+knitr::opts_chunk$set(warning=F, echo=T, error=F, message=F, cache = F)
 options(digits = 3, width = 250, scipen=999)
@@
 # define packages
-packages <- c("brms", "devtools", "GGally", "ggplot2", 
+packages <- c("broom.mixed", "brms", "devtools", "GGally", "ggplot2", 
               "gridExtra", "kableExtra", "knitr", "lavaan", "lme4", 
-              "magrittr", "mice", "mvnormalTest", 
+              "magrittr", "mice", #"mvnormalTest", 
               "PerFit", "psych", "quanteda.textstats", "semTools", "tidyverse")
@@
 ## Custom functions
+
+### find_var
 
@@
 
+### get_specs
+
 Extract characteristics from fitted lmer models.
@@
 ```
+
+### get_dat
 
@@
 ```
+
+### get_rel
 
@@
 ```
+
+### make_graph
 
@@
 
+### get_dat_res
+
 Get data of lmer objects for results.
@@
 # get data
-get_dat_res <- function(data_aff_neg, data_aff_pos, data_life_sat, type){
-
+get_dat_res <- function(data_aff_neg, data_aff_pos, data_life_sat, type, analysis = NULL){
+
+  if(isTRUE(class(data_aff_neg) == "lmerModLmerTest")) {
          dat_fig_results <- 
@@
              broom.mixed::tidy(data_aff_pos, conf.int = T) %>% 
-    mutate(dv = "aff_pos")
-  ) %>% 
+               mutate(dv = "aff_pos")) %>% 
            rbind(
              broom.mixed::tidy(data_life_sat, conf.int = T) %>% 
-    mutate(dv = "life_sat")
-  ) %>% 
+               mutate(dv = "life_sat"))
+         } else{ 
+           dat_fig_results <- 
+             data_aff_neg %>% 
+             mutate(dv = "aff_neg") %>% 
+             rbind(data_aff_pos %>% 
+                   mutate(dv = "aff_pos")) %>% 
+             rbind(data_life_sat %>% 
+                     mutate(dv = "life_sat")) %>% 
+           rename(conf.low = `2.5 %`, conf.high = `97.5 %`)
+           }
+  dat_fig_results %<>% 
   mutate(
@@
                 levels = c("life_sat", "aff_pos", "aff_neg"),
-                labels = c("Life satisfaction", "Positive affect", "Negative affect"))
+                labels = c("Life satisfaction", "Positive affect", "Negative affect")),
+    analysis = analysis
   ) %>% 
-  select(dv, iv, type, estimate, conf.low, conf.high, p.value)
+  select(dv, iv, type, analysis, estimate, conf.low, conf.high, p.value)
 
@@
     )
-  } else if(type == "channel"){
+  } else if(type == "channels"){
     dat_fig_results %<>%
@@
 
+### make_graph_res
+
 Make graph of effects
@@
 ```{r}
-make_graph_res <- function(data_1, data_2 = NULL, sesoi = NULL){
-  ggplot(data_1, aes(x = estimate, y = iv)) + 
-  scale_color_manual(values = c("grey40", "deepskyblue")) +
+make_graph_res <- function(data, sesoi = NULL, legend = TRUE){
+  ggplot(data, aes(x = estimate, y = iv)) + 
+  scale_color_manual(values = c("grey40", "deepskyblue", "darkcyan", "cornflowerblue", "aquamarine")) +
   geom_vline(xintercept = 0, lwd = .75, colour = "darkgrey") + 
-  geom_errorbarh(aes(xmin = conf.low, xmax = conf.high),
-                lwd = .75, height = .2) +
-  geom_point(size = 2) +
-  {if(!is.null(data_2)) geom_errorbarh(aes(xmin = conf.low, xmax = conf.high), data_2,
-                lwd = .75, height = .2, color = "grey",
-                position = position_nudge(y = -.3)
-                )} +
-  {if(!is.null(data_2)) geom_point(data = data_2, size = 2, color = "grey",
-             position = position_nudge(y = -.3)
-             )} +
-  {if(isTRUE(sesoi == "est")) geom_vline(data=filter(data_1, dv=="Life satisfaction"), aes(xintercept=-.3), colour="darkgrey", linetype = "dashed")} + 
-    {if(isTRUE(sesoi == "est")) geom_vline(data=filter(data_1, dv=="Life satisfaction"), aes(xintercept=.3), colour="darkgrey", linetype = "dashed")} + 
-  {if(isTRUE(sesoi == "est")) geom_vline(data=filter(data_1, dv!="Life satisfaction"), aes(xintercept=-.15), colour="darkgrey", linetype = "dashed")} +
-    {if(isTRUE(sesoi == "est")) geom_vline(data=filter(data_1, dv!="Life satisfaction"), aes(xintercept=.15), colour="darkgrey", linetype = "dashed")} +
+  geom_errorbarh(aes(xmin = conf.low, xmax = conf.high, color = analysis),
+                lwd = .75, height = .2, position = position_dodge(.5)) +
+  geom_point(aes(color = analysis), size = 2, position = position_dodge(.5)) +
+  {if(isTRUE(sesoi == "est")) geom_vline(data=filter(data, dv=="Life satisfaction"), aes(xintercept=-.3), colour="darkgrey", linetype = "dashed")} + 
+    {if(isTRUE(sesoi == "est")) geom_vline(data=filter(data, dv=="Life satisfaction"), aes(xintercept=.3), colour="darkgrey", linetype = "dashed")} + 
+  {if(isTRUE(sesoi == "est")) geom_vline(data=filter(data, dv!="Life satisfaction"), aes(xintercept=-.15), colour="darkgrey", linetype = "dashed")} +
+    {if(isTRUE(sesoi == "est")) geom_vline(data=filter(data, dv!="Life satisfaction"), aes(xintercept=.15), colour="darkgrey", linetype = "dashed")} +
   {if(isTRUE(sesoi == "std")) geom_vline(aes(xintercept=.1), colour="darkgrey", linetype = "dashed")} + 
-  {if(isTRUE(sesoi == "std")) geom_vline(aes(xintercept=-.1), colour="darkgrey", linetype = "dashed")} +   theme(
+  {if(isTRUE(sesoi == "std")) geom_vline(aes(xintercept=-.1), colour="darkgrey", linetype = "dashed")} +   
+    theme(
     axis.title.y = element_blank(),
@@
     ) +
+  {if(!isTRUE(legend)) theme(legend.position = "none")} + 
+  guides(colour = guide_legend(reverse=T)) +
   facet_grid(rows = vars(type),
@@
 ```
+
+### get_fs
 
@@
 
-```{r eval=TRUE}
+```{r eval=TRUE, cache=TRUE}
 vars_used <- c("life_sat", "aff_pos_1", "aff_pos_2", "aff_pos_3", "aff_neg_1", "aff_neg_2", "aff_neg_3", "aff_neg_4", "aff_neg_5", "aff_neg_6", "soc_med_read", "soc_med_like_share", "soc_med_post", "soc_med_fb", "soc_med_ig", "soc_med_wa", "soc_med_yt", "soc_med_tw", "age", "male", "born_aus", "born_aus_prnts", "edu_fac", "employment_fac", "health", "res_vienna", "acc_bal", "acc_gar", "home_sqm", "med_txt_kro", "med_txt_sta", "med_txt_pre", "med_txt_oes", "med_txt_kur", "med_txt_slz", "med_txt_son", "med_vid_orf", "med_vid_pri", "risk_prop", "loc_cntrl_int_1", "loc_cntrl_int_2", "loc_cntrl_int_3", "loc_cntrl_int_4", "act_wrk", "act_spo", "act_frn", "act_sho", "act_pet", "sat_dem")
@@
 
-# init <- mice(select(d_long_test_filt, maxit = 0))
-# pred_ma <- init$predictorMatrix
-# pred_ma[, vars_excl] <- 0
-d_long_imp <- mice(d_long_filt, method = "pmm", m = 1, maxit = 30, 
+d_long_imp <- mice(d_long_filt, 
+                   method = "pmm", # use predictive mean matching
+                   m = 1, maxit = 30, # only 1 imputation
                    where = incl_ma,
-              # predictorMatrix = pred_ma, 
                    seed = 180719, print = FALSE)
 d_long_imp <- mice::complete(d_long_imp)
-
 write_csv(d_long_imp, "data/data_imputed.csv")
+```
+
+```{r eval=TRUE, cache=TRUE}
+# impute missing data with multiple imputation
+d_long_mim_mice <- mice(d_long_filt, 
+                   method = "pmm", # use predictive mean matching
+                   m = 5, maxit = 5, # 5 imputations, 5 iterations
+                   where = incl_ma,
+                   seed = 180719, print = FALSE)
+d_long_mim <- mice::complete(d_long_mim_mice, action = "long", include = TRUE)
 ```
@@
 
+And now the same for the multiple imputed data
+
+```{r}
+d_long_mim %<>%
+  mutate(aff_pos_m = rowMeans(select(., aff_pos_1 : aff_pos_3)),
+         aff_neg_m = rowMeans(select(., aff_neg_1 : aff_neg_6)),
+         loc_cntrl_int_m = rowMeans(select(., loc_cntrl_int_1 : loc_cntrl_int_4)))
+
+d_long_mim %<>% 
+  group_by(id) %>% 
+  mutate_at(vars(health:loc_cntrl_int_4, loc_cntrl_int_m), funs(b = mean(., na.rm = TRUE), w = . - mean(., na.rm = TRUE))) %>% 
+  ungroup()
+
+d_long_mim_mice <- as.mids(d_long_mim)
+```
+
 # Measures
@@
 
-Assumption of multivariate normal distributino was violated; hence, robust estimator will be used.
+Assumption of multivariate normal distribution was violated; hence, robust estimator will be used.
 
@@
 # get data
-dat_fig_results_activity <- get_dat_res(model_aff_neg_lmer, model_aff_pos_lmer, model_life_sat_lmer, type = "activity")
-dat_fig_results_activity_nco <- get_dat_res(model_aff_neg_lmer_nco, model_aff_pos_lmer_nco, model_life_sat_lmer_nco, type = "activity")
+dat_fig_results_activity <- get_dat_res(model_aff_neg_lmer, model_aff_pos_lmer, model_life_sat_lmer, 
+                                        type = "activity", analysis = "regular")
+dat_fig_results_activity_nco <- get_dat_res(model_aff_neg_lmer_nco, model_aff_pos_lmer_nco, model_life_sat_lmer_nco, 
+                                            type = "activity", analysis = "no controls")
+dat_fig_results_activity <- rbind(
+  dat_fig_results_activity, 
+  dat_fig_results_activity_nco
+)
 
@@
 fig_results_activity <- make_graph_res(
-  data_1 = dat_fig_results_activity,
-  data_2 = dat_fig_results_activity_nco,
+  data = dat_fig_results_activity,
   sesoi = "est"
@@
 # get data
-dat_fig_results_channel <- get_dat_res(model_aff_neg_lmer, model_aff_pos_lmer, model_life_sat_lmer, type = "channel")
-dat_fig_results_channel_nco <- get_dat_res(model_aff_neg_lmer_nco, model_aff_pos_lmer_nco, model_life_sat_lmer_nco, type = "channel")
+dat_fig_results_channels <- get_dat_res(model_aff_neg_lmer, model_aff_pos_lmer, model_life_sat_lmer, 
+                                        type = "channels", analysis = "regular")
+dat_fig_results_channels_nco <- get_dat_res(model_aff_neg_lmer_nco, model_aff_pos_lmer_nco, model_life_sat_lmer_nco, 
+                                            type = "channels", analysis = "no controls")
+dat_fig_results_channels <- rbind(
+  dat_fig_results_channels, 
+  dat_fig_results_channels_nco
+)
 
 # make figure
-fig_results_channel <- make_graph_res(
-  data_1 = dat_fig_results_channel,
-  data_2 = dat_fig_results_channel_nco,
+fig_results_channels <- make_graph_res(
+  data = dat_fig_results_channels,
   sesoi = "est"
   )
-fig_results_channel
-
-# save figure
-ggsave("figures/fig_results_channel.pdf", 
+fig_results_channels
+
+# safe figure
+ggsave("figures/fig_results_channels.pdf", 
        width = 7, height = 4,
-       plot = fig_results_channel)
+       plot = fig_results_channels)
 ```
@@
 fig_results_control <- make_graph_res(
-  get_dat_res(model_aff_neg_lmer, model_aff_pos_lmer, model_life_sat_lmer, type = "control")
+  get_dat_res(model_aff_neg_lmer, model_aff_pos_lmer, model_life_sat_lmer, analysis = "exploratory", type = "control"),
+  legend = FALSE
   )

### 2022-04-11 analyses_additional.Rmd
- deleted data

---
title: "Additional analyses"
output:
  html_document:
    toc: TRUE
    toc_float:
      collapsed: FALSE
    toc_depth: 2
    code_folding: hide
bibliography      : 
  - "bibliography/r-references.bib"
  - "bibliography/references.bib"
  # pdf_document:
  #   toc: TRUE
---

```{r knitr-setup, include=F}
set.seed(170819)
knitr::opts_chunk$set(warning=F, echo=T, error=F, message=F, cache = F)
options(digits = 3, width = 200, scipen=999)
```

Here you can find the code and the results of all additional analyses. To see the code, click on button "Code". Alternatively, you can download the rmd file from the github repo.

# Set-up

Load packages, and make repo replicable.

```{r r-setup, results='hide'}
# install packages
# devtools::install_github("https://github.com/tdienlin/td@v.0.0.2.5")

# define packages
packages <- c("broom.mixed", "brms", "devtools", "GGally", "ggplot2", 
              "gridExtra", "kableExtra", "knitr", "lavaan", "lme4", 
              "magrittr", "mice", "mvnormalTest", 
              "PerFit", "psych", "quanteda.textstats", "semTools", "tidyverse")

# load packages
lapply(c(packages, "td"), library, character.only = TRUE)

# load workspace
load("data/workspace_1.RData")
```

# Standardized results

First, create standardized dataset.

```{r}
d_long_std <- d_long_imp %>% 
  mutate(across(c(-id, -wave, -gender, -male, -born_aus, -born_aus_prnts, -edu_fac, -employment_fac, -res_vienna, -acc_bal, -acc_gar), scale))
```

Then let's rerun the analyses.

## Life satisfaction

```{r}
model_life_sat_lmer_std <- lmerTest::lmer(life_sat ~
                                        (1 | id) + (1 | wave) + 
                                        soc_med_read_w + soc_med_like_share_w + soc_med_post_w  + 
                                        soc_med_fb_w + soc_med_ig_w + soc_med_wa_w + soc_med_yt_w + soc_med_tw_w +
                                        soc_med_read_b + soc_med_like_share_b + soc_med_post_b + 
                                        soc_med_fb_b + soc_med_ig_b + soc_med_wa_b + soc_med_yt_b + soc_med_tw_b +  
                                        age + male + born_aus + born_aus_prnts + edu_fac + employment_fac + health_b + 
                                        res_vienna + acc_bal + acc_gar + home_sqm + 
                                        med_txt_kro_b + med_txt_sta_b + med_txt_pre_b + med_txt_oes_b + med_txt_kur_b + med_txt_slz_b + med_txt_son_b + 
                                        med_vid_orf_b + med_vid_pri_b + 
                                        risk_prop_b + loc_cntrl_int_m_b + 
                                        act_wrk_w + act_spo_w + act_frn_w + act_sho_w + act_pet_w + 
                                        act_wrk_b + act_spo_b + act_frn_b + act_sho_b + act_pet_b +
                                        sat_dem_w + sat_dem_b 
                                      , data = d_long_std)
summary(model_life_sat_lmer_std)
```

## Positive affect

```{r}
model_aff_pos_lmer_std <- lmerTest::lmer(aff_pos_fs ~ 
                                        (1 | id) + (1 | wave) + 
                                        soc_med_read_w + soc_med_like_share_w + soc_med_post_w + 
                                        soc_med_fb_w + soc_med_ig_w + soc_med_wa_w + soc_med_yt_w + soc_med_tw_w +
                                        soc_med_read_b + soc_med_like_share_b + soc_med_post_b + 
                                        soc_med_fb_b + soc_med_ig_b + soc_med_wa_b + soc_med_yt_b + soc_med_tw_b +  
                                        age + male + born_aus + born_aus_prnts + edu_fac + employment_fac + health_b + 
                                        res_vienna + acc_bal + acc_gar + home_sqm + 
                                        med_txt_kro_b + med_txt_sta_b + med_txt_pre_b + med_txt_oes_b + med_txt_kur_b + med_txt_slz_b + med_txt_son_b + 
                                        med_vid_orf_b + med_vid_pri_b + 
                                        risk_prop_b + loc_cntrl_int_m_b + 
                                        act_wrk_w + act_spo_w + act_frn_w + act_sho_w + act_pet_w + 
                                        act_wrk_b + act_spo_b + act_frn_b + act_sho_b + act_pet_b +
                                        sat_dem_w + sat_dem_b 
                                      , data = d_long_imp)
summary(model_aff_pos_lmer_std)
```

## Negative Affect

```{r}
model_aff_neg_lmer_std <- lmerTest::lmer(aff_neg_fs ~ 
                                        (1 | id) + (1 | wave) + 
                                        soc_med_read_w + soc_med_like_share_w + soc_med_post_w + 
                                        soc_med_fb_w + soc_med_ig_w + soc_med_wa_w + soc_med_yt_w + soc_med_tw_w +
                                        soc_med_read_b + soc_med_like_share_b + soc_med_post_b + 
                                        soc_med_fb_b + soc_med_ig_b + soc_med_wa_b + soc_med_yt_b + soc_med_tw_b +  
                                        age + male + born_aus + born_aus_prnts + edu_fac + employment_fac + health_b + 
                                        res_vienna + acc_bal + acc_gar + home_sqm + 
                                        med_txt_kro_b + med_txt_sta_b + med_txt_pre_b + med_txt_oes_b + med_txt_kur_b + med_txt_slz_b + med_txt_son_b + 
                                        med_vid_orf_b + med_vid_pri_b + 
                                        risk_prop_b + loc_cntrl_int_m_b + 
                                        act_wrk_w + act_spo_w + act_frn_w + act_sho_w + act_pet_w + 
                                        act_wrk_b + act_spo_b + act_frn_b + act_sho_b + act_pet_b +
                                        sat_dem_w + sat_dem_b 
                                      , data = d_long_imp)
summary(model_aff_neg_lmer_std)
```

# Results with mean scores

Instead of using factor scores, in what follows I report the results with mean scores of affect. Note that the results for life satisfaction were the same, as it was measured with a single item.

## Positive affect

```{r}
model_aff_pos_lmer_m <- lmerTest::lmer(aff_pos_m ~ 
                                        (1 | id) + (1 | wave) + 
                                        soc_med_read_w + soc_med_like_share_w + soc_med_post_w + 
                                        soc_med_fb_w + soc_med_ig_w + soc_med_wa_w + soc_med_yt_w + soc_med_tw_w +
                                        soc_med_read_b + soc_med_like_share_b + soc_med_post_b + 
                                        soc_med_fb_b + soc_med_ig_b + soc_med_wa_b + soc_med_yt_b + soc_med_tw_b +  
                                        age + male + born_aus + born_aus_prnts + edu_fac + employment_fac + health_b + 
                                        res_vienna + acc_bal + acc_gar + home_sqm + 
                                        med_txt_kro_b + med_txt_sta_b + med_txt_pre_b + med_txt_oes_b + med_txt_kur_b + med_txt_slz_b + med_txt_son_b + 
                                        med_vid_orf_b + med_vid_pri_b + 
                                        risk_prop_b + loc_cntrl_int_m_b + 
                                        act_wrk_w + act_spo_w + act_frn_w + act_sho_w + act_pet_w + 
                                        act_wrk_b + act_spo_b + act_frn_b + act_sho_b + act_pet_b +
                                        sat_dem_w + sat_dem_b 
                                      , data = d_long_imp)
summary(model_aff_pos_lmer_m)
```

The results differed only slightly, and all inferences remained the same.

## Negative Affect

```{r}
model_aff_neg_lmer_m <- lmerTest::lmer(aff_neg_m ~ 
                                        (1 | id) + (1 | wave) + 
                                        soc_med_read_w + soc_med_like_share_w + soc_med_post_w + 
                                        soc_med_fb_w + soc_med_ig_w + soc_med_wa_w + soc_med_yt_w + soc_med_tw_w +
                                        soc_med_read_b + soc_med_like_share_b + soc_med_post_b + 
                                        soc_med_fb_b + soc_med_ig_b + soc_med_wa_b + soc_med_yt_b + soc_med_tw_b +  
                                        age + male + born_aus + born_aus_prnts + edu_fac + employment_fac + health_b + 
                                        res_vienna + acc_bal + acc_gar + home_sqm + 
                                        med_txt_kro_b + med_txt_sta_b + med_txt_pre_b + med_txt_oes_b + med_txt_kur_b + med_txt_slz_b + med_txt_son_b + 
                                        med_vid_orf_b + med_vid_pri_b + 
                                        risk_prop_b + loc_cntrl_int_m_b + 
                                        act_wrk_w + act_spo_w + act_frn_w + act_sho_w + act_pet_w + 
                                        act_wrk_b + act_spo_b + act_frn_b + act_sho_b + act_pet_b +
                                        sat_dem_w + sat_dem_b 
                                      , data = d_long_imp)
summary(model_aff_neg_lmer_m)
```

The results differed only slightly, and all inferences remained the same.

# Results without imputed data

We first need to export factor scores for variables without imputed data

```{r}
model <- "
aff_pos =~ a1*aff_pos_1 + a2*aff_pos_2 + a3*aff_pos_3
"
cfa_aff_pos <- cfa(model, d_long, group = "wave", estimator = "MLM")
d_long$aff_pos_fs <- get_fs(cfa_aff_pos)

model <- "
aff_neg =~ a1*aff_neg_1 + a2*aff_neg_2 + a3*aff_neg_3 + a4*aff_neg_4 + a5*aff_neg_5 + a6*aff_neg_6
"
cfa_aff_neg <- cfa(model, d_long, group = "wave", estimator = "MLM")
d_long$aff_neg_fs <- get_fs(cfa_aff_neg)
```

## Life satisfaction

```{r}
model_life_sat_lmer_noi <- lmerTest::lmer(life_sat ~
                                        (1 | id) + (1 | wave) + 
                                        soc_med_read_w + soc_med_like_share_w + soc_med_post_w  + 
                                        soc_med_fb_w + soc_med_ig_w + soc_med_wa_w + soc_med_yt_w + soc_med_tw_w +
                                        soc_med_read_b + soc_med_like_share_b + soc_med_post_b + 
                                        soc_med_fb_b + soc_med_ig_b + soc_med_wa_b + soc_med_yt_b + soc_med_tw_b +  
                                        age + male + born_aus + born_aus_prnts + edu_fac + employment_fac + health_b + 
                                        res_vienna + acc_bal + acc_gar + home_sqm + 
                                        med_txt_kro_b + med_txt_sta_b + med_txt_pre_b + med_txt_oes_b + med_txt_kur_b + med_txt_slz_b + med_txt_son_b + 
                                        med_vid_orf_b + med_vid_pri_b + 
                                        risk_prop_b + loc_cntrl_int_m_b + 
                                        act_wrk_w + act_spo_w + act_frn_w + act_sho_w + act_pet_w + 
                                        act_wrk_b + act_spo_b + act_frn_b + act_sho_b + act_pet_b +
                                        sat_dem_w + sat_dem_b 
                                      , data = d_long)
summary(model_life_sat_lmer_noi)
```

## Positive Affect

```{r}
model_aff_pos_lmer_noi <- lmerTest::lmer(aff_pos_fs ~ 
                                        (1 | id) + (1 | wave) + 
                                        soc_med_read_w + soc_med_like_share_w + soc_med_post_w + 
                                        soc_med_fb_w + soc_med_ig_w + soc_med_wa_w + soc_med_yt_w + soc_med_tw_w +
                                        soc_med_read_b + soc_med_like_share_b + soc_med_post_b + 
                                        soc_med_fb_b + soc_med_ig_b + soc_med_wa_b + soc_med_yt_b + soc_med_tw_b +  
                                        age + male + born_aus + born_aus_prnts + edu_fac + employment_fac + health_b + 
                                        res_vienna + acc_bal + acc_gar + home_sqm + 
                                        med_txt_kro_b + med_txt_sta_b + med_txt_pre_b + med_txt_oes_b + med_txt_kur_b + med_txt_slz_b + med_txt_son_b + 
                                        med_vid_orf_b + med_vid_pri_b + 
                                        risk_prop_b + loc_cntrl_int_m_b + 
                                        act_wrk_w + act_spo_w + act_frn_w + act_sho_w + act_pet_w + 
                                        act_wrk_b + act_spo_b + act_frn_b + act_sho_b + act_pet_b +
                                        sat_dem_w + sat_dem_b 
                                      , data = d_long)
summary(model_aff_pos_lmer_noi)
```

## Negative Affect

```{r}
model_aff_neg_lmer_noi <- lmerTest::lmer(aff_neg_fs ~ 
                                        (1 | id) + (1 | wave) + 
                                        soc_med_read_w + soc_med_like_share_w + soc_med_post_w + 
                                        soc_med_fb_w + soc_med_ig_w + soc_med_wa_w + soc_med_yt_w + soc_med_tw_w +
                                        soc_med_read_b + soc_med_like_share_b + soc_med_post_b + 
                                        soc_med_fb_b + soc_med_ig_b + soc_med_wa_b + soc_med_yt_b + soc_med_tw_b +  
                                        age + male + born_aus + born_aus_prnts + edu_fac + employment_fac + health_b + 
                                        res_vienna + acc_bal + acc_gar + home_sqm + 
                                        med_txt_kro_b + med_txt_sta_b + med_txt_pre_b + med_txt_oes_b + med_txt_kur_b + med_txt_slz_b + med_txt_son_b + 
                                        med_vid_orf_b + med_vid_pri_b + 
                                        risk_prop_b + loc_cntrl_int_m_b + 
                                        act_wrk_w + act_spo_w + act_frn_w + act_sho_w + act_pet_w + 
                                        act_wrk_b + act_spo_b + act_frn_b + act_sho_b + act_pet_b +
                                        sat_dem_w + sat_dem_b 
                                      , data = d_long)
summary(model_aff_neg_lmer_noi)
```

# Results with multiple imputation
## Life satisfaction

```{r}
model_life_sat_lmer_mim <- with(d_long_mim_mice, exp =  
                                lmerTest::lmer(life_sat ~
                                        (1 | id) + (1 | wave) + 
                                        soc_med_read_w + soc_med_like_share_w + soc_med_post_w  + 
                                        soc_med_fb_w + soc_med_ig_w + soc_med_wa_w + soc_med_yt_w + soc_med_tw_w +
                                        soc_med_read_b + soc_med_like_share_b + soc_med_post_b + 
                                        soc_med_fb_b + soc_med_ig_b + soc_med_wa_b + soc_med_yt_b + soc_med_tw_b +  
                                        age + male + born_aus + born_aus_prnts + edu_fac + employment_fac + health_b + 
                                        res_vienna + acc_bal + acc_gar + home_sqm + 
                                        med_txt_kro_b + med_txt_sta_b + med_txt_pre_b + med_txt_oes_b + med_txt_kur_b + med_txt_slz_b + med_txt_son_b + 
                                        med_vid_orf_b + med_vid_pri_b + 
                                        risk_prop_b + loc_cntrl_int_m_b + 
                                        act_wrk_w + act_spo_w + act_frn_w + act_sho_w + act_pet_w + 
                                        act_wrk_b + act_spo_b + act_frn_b + act_sho_b + act_pet_b +
                                        sat_dem_w + sat_dem_b 
                                      ))
model_life_sat_lmer_mim <- summary(pool(model_life_sat_lmer_mim), conf.int = TRUE)
```

## Positive Affect

For simplicity, note that we use mean-scores and not factors scores here.

```{r}
model_aff_pos_lmer_mim <- with(d_long_mim_mice, exp =  
                           lmerTest::lmer(aff_pos_m ~ 
                                        (1 | id) + (1 | wave) + 
                                        soc_med_read_w + soc_med_like_share_w + soc_med_post_w + 
                                        soc_med_fb_w + soc_med_ig_w + soc_med_wa_w + soc_med_yt_w + soc_med_tw_w +
                                        soc_med_read_b + soc_med_like_share_b + soc_med_post_b + 
                                        soc_med_fb_b + soc_med_ig_b + soc_med_wa_b + soc_med_yt_b + soc_med_tw_b +  
                                        age + male + born_aus + born_aus_prnts + edu_fac + employment_fac + health_b + 
                                        res_vienna + acc_bal + acc_gar + home_sqm + 
                                        med_txt_kro_b + med_txt_sta_b + med_txt_pre_b + med_txt_oes_b + med_txt_kur_b + med_txt_slz_b + med_txt_son_b + 
                                        med_vid_orf_b + med_vid_pri_b + 
                                        risk_prop_b + loc_cntrl_int_m_b + 
                                        act_wrk_w + act_spo_w + act_frn_w + act_sho_w + act_pet_w + 
                                        act_wrk_b + act_spo_b + act_frn_b + act_sho_b + act_pet_b +
                                        sat_dem_w + sat_dem_b)
)
model_aff_pos_lmer_mim <- summary(pool(model_aff_pos_lmer_mim), conf.int = TRUE)
```

## Negative Affect

For simplicity, note that we use mean-scores and not factors scores here.

```{r}
model_aff_neg_lmer_mim <- with(d_long_mim_mice, exp =  
                               lmerTest::lmer(aff_neg_m ~ 
                                        (1 | id) + (1 | wave) + 
                                        soc_med_read_w + soc_med_like_share_w + soc_med_post_w + 
                                        soc_med_fb_w + soc_med_ig_w + soc_med_wa_w + soc_med_yt_w + soc_med_tw_w +
                                        soc_med_read_b + soc_med_like_share_b + soc_med_post_b + 
                                        soc_med_fb_b + soc_med_ig_b + soc_med_wa_b + soc_med_yt_b + soc_med_tw_b +  
                                        age + male + born_aus + born_aus_prnts + edu_fac + employment_fac + health_b + 
                                        res_vienna + acc_bal + acc_gar + home_sqm + 
                                        med_txt_kro_b + med_txt_sta_b + med_txt_pre_b + med_txt_oes_b + med_txt_kur_b + med_txt_slz_b + med_txt_son_b + 
                                        med_vid_orf_b + med_vid_pri_b + 
                                        risk_prop_b + loc_cntrl_int_m_b + 
                                        act_wrk_w + act_spo_w + act_frn_w + act_sho_w + act_pet_w + 
                                        act_wrk_b + act_spo_b + act_frn_b + act_sho_b + act_pet_b +
                                        sat_dem_w + sat_dem_b)
)
model_aff_neg_lmer_mim <- summary(pool(model_aff_neg_lmer_mim), conf.int = TRUE)
```

# Figures
## Standardized results

First, for activities.

```{r}
# make figure
fig_results_activity_std <- make_graph_res(
  get_dat_res(model_aff_neg_lmer_std, model_aff_pos_lmer_std, model_life_sat_lmer_std, 
              type = "activity", analysis = "standardized"),
  sesoi = "std",
  legend = FALSE
  )
fig_results_activity_std

# safe figure
ggsave("figures/fig_results_activity_std.pdf", 
       width = 7, height = 4,
       plot = fig_results_activity_std)
```

Then for channels.

```{r}
# make figure
fig_results_channels_std <- make_graph_res(
  get_dat_res(model_aff_neg_lmer_std, model_aff_pos_lmer_std, model_life_sat_lmer_std, type = "channels", analysis = "standardized"),
  sesoi = "std",
  legend
  )
fig_results_channels_std

# save figure
ggsave("figures/fig_results_channels_std.pdf", 
       width = 7, height = 4,
       plot = fig_results_channels_std)
```

Select control variables and make graph.

```{r}
# make figure
fig_results_control_std <- make_graph_res(
  get_dat_res(model_aff_neg_lmer_std, model_aff_pos_lmer_std, model_life_sat_lmer_std, 
              type = "control", analysis = "standardized"),
  sesoi = "std"
  )
fig_results_control_std

# save figure
ggsave("figures/fig_results_control_std.pdf", 
       width = 7, height = 4,
       plot = fig_results_control_std)
```

## Comparison of unstandardized methods

In what follows, see figure with all results combined. Only nonstandardized results are presented to use the same scaling.

First, for activities.

```{r}
# get data
dat_fig_results_activity <- get_dat_res(model_aff_neg_lmer, model_aff_pos_lmer, model_life_sat_lmer, type = "activity", analysis = "1. regular")
dat_fig_results_activity_nco <- get_dat_res(model_aff_neg_lmer_nco, model_aff_pos_lmer_nco, model_life_sat_lmer_nco, type = "activity", analysis = "2. no covars")
dat_fig_results_activity_m <- get_dat_res(model_aff_neg_lmer_m, model_aff_pos_lmer_m, model_life_sat_lmer, type = "activity", analysis = "3. mean scores")
dat_fig_results_activity_noi <- get_dat_res(model_aff_neg_lmer_noi, model_aff_pos_lmer_noi, model_life_sat_lmer_noi, type = "activity", analysis = "4. no imp")
dat_fig_results_activity_mim <- get_dat_res(model_aff_neg_lmer_mim, model_aff_pos_lmer_mim, model_life_sat_lmer_mim, type = "activity", analysis = "5. mult imp")

dat_fig_results_activity <- rbind(
  dat_fig_results_activity,
  dat_fig_results_activity_nco,
  dat_fig_results_activity_m,
  dat_fig_results_activity_noi,
  dat_fig_results_activity_mim
)
# make figure
fig_results_activity <- make_graph_res(
  dat_fig_results_activity,
  sesoi = "est"
  )
fig_results_activity

# safe figure
ggsave("figures/fig_results_activity.pdf", 
       width = 7, height = 4,
       plot = fig_results_activity)
```

Then for channels.

```{r}
# get data
dat_fig_results_channels <- get_dat_res(model_aff_neg_lmer, model_aff_pos_lmer, model_life_sat_lmer, type = "channels", analysis = "1. regular")
dat_fig_results_channels_nco <- get_dat_res(model_aff_neg_lmer_nco, model_aff_pos_lmer_nco, model_life_sat_lmer_nco, type = "channels", analysis = "2. no covars")
dat_fig_results_channels_m <- get_dat_res(model_aff_neg_lmer_m, model_aff_pos_lmer_m, model_life_sat_lmer, type = "channels", analysis = "3. mean scores")
dat_fig_results_channels_noi <- get_dat_res(model_aff_neg_lmer_noi, model_aff_pos_lmer_noi, model_life_sat_lmer_noi, type = "channels", analysis = "4. no imp")
dat_fig_results_channels_mim <- get_dat_res(model_aff_neg_lmer_mim, model_aff_pos_lmer_mim, model_life_sat_lmer_mim, type = "channels", analysis = "5. mult imp")

dat_fig_results_channels <- rbind(
  dat_fig_results_channels,
  dat_fig_results_channels_nco,
  dat_fig_results_channels_m,
  dat_fig_results_channels_noi,
  dat_fig_results_channels_mim
)
# make figure
fig_results_channels <- make_graph_res(
  dat_fig_results_channels,
  sesoi = "est"
  )
fig_results_channels

# safe figure
ggsave("figures/fig_results_channels.pdf", 
       width = 7, height = 4,
       plot = fig_results_channels)
```

Save results so that we can extract them in the manuscript.

```{r}
save.image("data/workspace_2.RData")
```

### 2022-04-11 manuscript.Rmd
- deleted data

@@
 
-abstract          : "In times of crisis such as the Corona pandemic citizens need to stay informed about recent events, the latest political decisions, or mandatory protection measures.
+abstract          : "In times of crisis such as the Corona pandemic citizens need to stay informed about recent events, political decisions, or mandatory protection measures.
 To this end, many people use various types of media, and increasingly social media.
@@
 The findings suggest that fears that social media use during times of crisis impairs well-being are likely to be unfounded."
-keywords          : "COVID-19, Coronavirus, well-being, affect, life satisfaction, social media use, news use, communication, random effects within between model, panel study, longitudinal study."
+keywords          : "COVID-19, well-being, social media, news use, panel study."
 
@@
 
-floatsintext      : yes
+floatsintext      : no
+endnotes          : yes
 figurelist        : no
@@
 classoption       : "man"
-output            : papaja::apa6_pdf
+# output            : papaja::apa6_pdf
+output            : papaja::apa6_docx
 header-includes:
+  - \usepackage{endnotes}
   - \setlength{\parskip}{0em}
@@
 # devtools::install_github("https://github.com/tdienlin/td@v.0.0.2.5")
-# devtools::install_github("https://github.com/crsh/papaja@devel")
+# 1devtools::install_github("https://github.com/crsh/papaja@devel")
 
@@
 Users were glued to their screens and found it hard to pursue other relevant activities such as working, taking a break, or looking after their children [@kleinDarklySoothingCompulsion2021].
-As doomscrolling increased it became doubtful whether such social media is helpful, or whether it created an additional burden on mental health [@sandstromDoomscrollingCOVIDNews2021].
+It was asked whether using social media for COVID-19 related reasons is helpful, or whether it creates an additional burden on mental health [@sandstromDoomscrollingCOVIDNews2021].
 These concerns seem justified: 
@@
 
-First, I investigate how different types of communication affect well-being.
-Specifically, I differentiate between active and passive use.
-I distinguish (a) _reading_ (passive), (b) _posting_ (active), and (c) _liking and sharing_ COVID-19 related posts (both active and passive).
+First, I investigate how well-being is affected by different types of communication affect, namely active and passive use.
+Defining what constitutes active and what passive use is not always clear, and different understandings are currently discussed [@ellisonWhyWeDon2020, @meierComputermediatedCommunicationSocial2020a].
+Reading is generally considered as passive and writing as active, while there are also specific behaviors such as such liking or sharing content that fall somewhere in-between [@meierDoesPassiveSocial2022].
+In this study, I hence distinguish (a) _reading_ (passive), (b) _posting_ (active), and (c) _liking and sharing_ COVID-19 related posts (both active and passive).
 Second, I analyze how using the most prominent branded applications affects well-being, and whether this effect changes across applications.
@@
 
-These overviews are well aligned with several individual studies employing advanced methods [@orbenSocialMediaEnduring2019; @przybylskiDoesTakingShort2021a; @schemerImpactInternetSocial2021; @kerestesAdolescentsOnlineSocial2020].
+It is often asked if social media use can become addictive. 
+Users sometimes express this fear themselves [@yangCanWatchingOnline2021]. 
+A recently published meta-analysis found that the two most prominent measures of addiction, the Bergen Facebook Addiction Scale and the Bergen Social Media Addiction Scale, have only small relations to well-being [@duradoniWellbeingSocialMedia2020].
+That said, the general idea of labeling excessive social and new media use as addiction was criticized, arguing that social and new media represent new regular behaviors that should not be pathologized [@galerHowMuchToo2018; @vanrooijWeakScientificBasis2018].
+
+Several individual studies employing advanced methods found smalls relations between social media use and well-being [@orbenSocialMediaEnduring2019; @przybylskiDoesTakingShort2021a; @schemerImpactInternetSocial2021; @kerestesAdolescentsOnlineSocial2020].
 For example, @beyensSocialMediaUse2021 reported that although for some users (roughly one quarter) the effects of social media use on well-being were negative, for almost the same number of users they were positive, while for the rest the effects were neutral.
-In conclusion, most effects are likely somewhere between trivial and small.
-I therefore expect that also in the case of COVID-19 related social media use effects will be trivial to small.
+This finding is aligned the Differential Susceptibility to Media Effects Model, which suggests that although there is substantial variation of media effects for individual users, the average effects reported in the literature are often small [@valkenburgDifferentialSusceptibilityMedia2013].
 
-From a theoretical perspective, how could we explain whether COVID-19 related social media use might affect well-being?
-In what follows, I outline potential arguments as to why the effect might be positive or negative, direct or indirect, or nonexistent.
-In advance, there doesn't seem to be a clear winner, and it's likely that both positive and negative effects cancel each other out.
+<!-- In conclusion, most effects are likely somewhere between trivial and small. -->
+<!-- I therefore expect that also in the case of COVID-19 related social media use effects will be trivial to small. -->
 
-First, one could assume a _direct_ negative effect on well-being, and especially on positive or negative affect, which are more volatile and fluctuating.
+From a theoretical perspective, how could we explain whether well-being might be affected by COVID-19 related social media use in particular?
+<!-- In what follows, I outline potential arguments as to why the effect might be positive or negative, direct or indirect, or nonexistent. -->
+<!-- In advance, there doesn't seem to be a clear winner, and it's likely that both positive and negative effects cancel each other out. -->
+According to the set-point model of subjective well-being, well-being is surprisingly stable [@lykkenHappinessWhatStudies1999].
+Although specific events such as marriage or salary can increase well-being, after some time it routinely returns to prior levels, which are mostly determined genetically [@sheldonStabilityHappinessTheories2014].
+Only very specific very negative events such as unemployment, disability, or death can cause long-term decreases in well-being [@lucasAdaptationSetpointModel2007], and specific positive factors such as XXX to long-term increase [].
+
+Can media use be such a negative or positive factor? 
+<!-- Although the set-point theory has been criticized lately because events such as marriage, unemployment, or death can indeed lead to long-term changes, it is still widely maintained that well-being is very stable and hence hard to be influenced  -->
+
+one could assume a _direct_ negative effect on well-being, and especially on positive or negative affect, which are more volatile and fluctuating.
 Dangers, inequalities, corruption---these were the headlines during the pandemic across many countries worldwide.
@@
 That said, because not all news was negative, and because many people showed solidarity and compassion, there was also positive and uplifting content, potentially compensating for the negative effects [@dornemannHowGoodBad2021].
+A study with 2.057 respondents from Italy reported that during the pandemic virtual community and social connectedness potentially even increased [@guazziniSecondWaveAnalysis2022].
 <!-- However, in light of a worldwide pandemic with millions of deaths, the negative direct effect seems more plausible. -->
@@
 
-The sample was representative of the Austrian population in terms of age, gender, region/state, municipality size, and educational level. 
+Achieved via quota sampling, the sample matched the Austrian population in terms of age, gender, region/state, municipality size, and educational level.
 In order to participate in the study, the respondents needed to be Austrian residents and had to be at least 14 years of age. 
-The study received IRB approval from the University of Vienna.
+Ethical review and approval was not required for the study in accordance with the local legislation and institutional requirements. 
+The participants provided their written informed consent to participate in this study.
 The average age was `r mean(2021 - d_wide$year_birth, na.rm = T) %>% round(0)` years, `r mean(d_wide$male, na.rm = T) %>% round(., 2) * 100` percent were male, `r (d_wide %>% filter(.$edu_fac %in% c("State college", "Bachelor", "Master", "PhD"))) %>% nrow() %>% divide_by(sum(table(d_wide$edu_fac))) %>% multiply_by(100) %>% round(0)` percent had a University degree, and `r (table(d_wide$employment_fac)["Unemployed"] / sum(table(d_wide$employment_fac))) %>% unname() %>% multiply_by(100) %>% round(., digits = 0)` percent were currently unemployed.
@@
 
-Life satisfaction was measured with the item "Taken everything together, how satisfied are you currently with your life?".
-The response options ranged from 0 (_extremely unsatisfied_) to 10 (_extremely satisfied_).
+Life satisfaction was measured with the item "All things considered, how satisfied are you with your life as a whole nowadays?" from the European Social Survey [@europeansocialsurveyESS9Edition20182021].
+The response options ranged from 0 (_extremely dissatisfied_) to 10 (_extremely satisfied_).
 
-To capture positive affect, respondents were asked how often in the last week they felt (a) calm and relaxed, (b) happy, and (c) full of energy.
+To capture positive affect, respondents were asked how often in the last week they felt (a) calm and relaxed, (b) happy, and (c) full of energy [@worldhealthorganizationWellbeingMeasuresPrimary1998].
 The response options were 1 (_never_), 2 (_on some days_), 3 (_several times per week_), 4 (_almost every day_), and 5 (_daily_).
@@
 
-For negative affect, respondents were asked how often in the last week they felt (a) lonely, (b) aggravated, (c) so depressed, that nothing could lift you up, (d) very nervous, (e) anxious, and (h) glum and sad.
+For negative affect, respondents were asked how often in the last week they felt (a) lonely, (b) aggravated, (c) so depressed, that nothing could lift you up, (d) very nervous, (e) anxious, and (h) glum and sad [@worldhealthorganizationWellbeingMeasuresPrimary1998].
 The response options were 1 (_never_), 2 (_on some days_), 3 (_several times per week_), 4 (_almost every day_), and 5 (_daily_).
@@
 
-COVID-19 related social media use focused on communication types was measured with the three dimensions of (a) reading, (b) liking and sharing, and (c) posting.
+COVID-19 related social media use focused on communication types was measured with the three dimensions of (a) reading, (b) liking and sharing, and (c) posting from @wagnerAUTNESOnlinePanel2018, adapted for the context of this study.
 The general introductory question was "How often during the last week have you engaged in the following activities on social media?"
@@
 
-COVID-19 related social media use focused on channels was measured with five variables.
+COVID-19 related social media use focused on channels was measured with five variables from @wagnerAUTNESOnlinePanel2018, adapted for this study. 
 The general introductory question was "How often in the last week have you followed information related to the Corona-crisis on the following social media?"
@@
 Freshly recruited respondents always answered all questions on social media use.
-
-```{r descriptives, cache=F}
-apa_table(tab_desc_dat,
-          stub_indents = list(
-            "Well-being" = c(1:3),
-            "Social media use" = c(4:6),
-            "Social media channel" = c(7:11)),
-          caption = "Descriptives of the main variables."
-          )
-```
 
@@
 
-First, when looking at the variables from a descriptive perspective, we see that all well-being measures did not change substantially across the different waves of data collection.
+First, when looking at the variables from a descriptive perspective (Figure 1), we see that all well-being measures did not change substantially across the different waves of data collection.
 COVID-19 related media use, however, decreased slightly at the beginning of the study and remained stable after approximately six waves.
@@
 It could be that after an initial uptick, COVID-19 related social media use was already declining at the time, returning to more normal levels.
-
-```{r fig-desc, fig.cap="Development of well-being and media use measures across the pandemic.
-Values obtained from mixed effect models, with participants and waves as grouping factors and without additional predictors.", out.width = "\\textwidth", cache=F}
-knitr::include_graphics("figures/fig_descriptives.pdf")
-```
 
@@
 
-```{r res-activity, fig.cap="The effects of various types of social media use on three indicators of well-being.
-The black estimates show the effects controlled for a large number of covariates (see text; preregistered); the grey estimates are without control variables (exploratory).
-The SESOI was b = |0.30| for life satisfaction and b = |0.15| for affect.
-Hence, all of the reported effects are not considered meaningful.", out.width = "\\textwidth", cache=F}
-knitr::include_graphics("figures/fig_results_activity.pdf")
-```
-
 Regarding between-person relations, about which no hypotheses were formulated, only three effects did not include zero.
@@
 However, when controlling for potential confounders, the effect became virtually zero (see Figure \@ref(fig:res-channels)).
-
-```{r res-channels, fig.cap="The effects of using various social media applications on three indicators of well-being.
-The black estimates show the effects controlled for a large number of covariates (see text; preregistered); the grey estimates are without control variables (exploratory).
-The SESOI was b = |0.30| for life satisfaction and b = |0.15| for affect.
-Hence, all of the reported effects are not considered meaningful.", out.width = "\\textwidth", cache=F}
-knitr::include_graphics("figures/fig_results_channel.pdf")
-```
 
@@
 
-```{r res-control, fig.cap="Results of selected covariates.
-All variables were standardize except 'Male'  and ‘Employed in public service', because there were measured on a binary scale.", out.width = "\\textwidth", cache=F}
-knitr::include_graphics("figures/fig_results_control_std.pdf")
-```
-
 To find out whether my inferences were robust across legitimate (though arguably inferior) alternative analyses, I reran the analyses also using standardized estimates, mean scores instead of factor scores, and with a data set where missing data were not imputed.
@@
 In this study I analyzed the effects of COVID-19 related social media use on well-being.
-The data come from a panel study with 24 waves and are representative of the Austrian population.
+The data come from a panel study with 24 waves and are largely representative of the Austrian population.
 In a random effects model I separated between person relations from within-person effects and controlled for a large number of both stable and varying covariates, thereby aiming to assess causality.
@@
 The results apply primarily to the more Western sphere, and might not hold true in other cultures, especially cultures with a different media landscape or alternative social media channels.
-That said, because this is a comparatively large study representative of an entire country, and because several waves were collected across a large time span, the results should be at least as generalizable as other typical empirical studies collected in the social sciences.
+That said, because this is a comparatively large study largely representative of an entire country, and because several waves were collected across a large time span, the results should be at least as generalizable as other typical empirical studies collected in the social sciences.
 
@@
 
+```{r descriptives, cache=F}
+apa_table(tab_desc_dat,
+          stub_indents = list(
+            "Well-being" = c(1:3),
+            "Social media use" = c(4:6),
+            "Social media channel" = c(7:11)),
+          caption = "Descriptives of the main variables."
+          )
+```
+
+\newpage
+
+```{r fig-desc, fig.cap="Development of well-being and media use measures across the pandemic.
+Values obtained from mixed effect models, with participants and waves as grouping factors and without additional predictors.", out.width = "\\textwidth", cache=F}
+knitr::include_graphics("figures/fig_descriptives.pdf")
+```
+
+\newpage
+
+```{r res-activity, fig.cap="The effects of various types of social media use on three indicators of well-being.
+The black estimates show the effects controlled for a large number of covariates (see text; preregistered); the grey estimates are without control variables (exploratory).
+The SESOI was b = |0.30| for life satisfaction and b = |0.15| for affect.
+Hence, all of the reported effects are not considered meaningful.", out.width = "\\textwidth", cache=F}
+knitr::include_graphics("figures/fig_results_activity.pdf")
+```
+
+\newpage
+
+```{r res-channels, fig.cap="The effects of using various social media applications on three indicators of well-being.
+The black estimates show the effects controlled for a large number of covariates (see text; preregistered); the grey estimates are without control variables (exploratory).
+The SESOI was b = |0.30| for life satisfaction and b = |0.15| for affect.
+Hence, all of the reported effects are not considered meaningful.", out.width = "\\textwidth", cache=F}
+knitr::include_graphics("figures/fig_results_channel.pdf")
+```
+
+\newpage
+
+```{r res-control, fig.cap="Results of selected covariates.
+All variables were standardize except 'Male'  and ‘Employed in public service', because there were measured on a binary scale.", out.width = "\\textwidth", cache=F}
+knitr::include_graphics("figures/fig_results_control_std.pdf")
+```
+
+\newpage
+
 # Competing Interests

### 2022-04-11 analyses.RmD
- updated figures

@@
 lapply(c(packages, "td"), library, character.only = TRUE)
+```
+
+## Session info
+
+Provide session info to make results reproducible.
+
+```{r}
+sessionInfo()
 ```
@@
                 labels = c("Life satisfaction", "Positive affect", "Negative affect")),
-    analysis = analysis
+    Analysis = analysis
   ) %>% 
-  select(dv, iv, type, analysis, estimate, conf.low, conf.high, p.value)
+  select(dv, iv, type, Analysis, estimate, conf.low, conf.high, p.value)
 
@@
   ggplot(data, aes(x = estimate, y = iv)) + 
-  scale_color_manual(values = c("grey40", "deepskyblue", "darkcyan", "cornflowerblue", "aquamarine")) +
+  scale_color_manual(values = c("black", "grey60", "grey75", "deepskyblue", "cornflowerblue", "darkcyan", "aquamarine")) +
   geom_vline(xintercept = 0, lwd = .75, colour = "darkgrey") + 
-  geom_errorbarh(aes(xmin = conf.low, xmax = conf.high, color = analysis),
-                lwd = .75, height = .2, position = position_dodge(.5)) +
-  geom_point(aes(color = analysis), size = 2, position = position_dodge(.5)) +
+  geom_errorbarh(aes(xmin = conf.low, xmax = conf.high, color = Analysis),
+                lwd = .75, height = .2, position = position_dodge(-.7)) +
+  geom_point(aes(color = Analysis), size = 2, position = position_dodge(-.7)) +
   {if(isTRUE(sesoi == "est")) geom_vline(data=filter(data, dv=="Life satisfaction"), aes(xintercept=-.3), colour="darkgrey", linetype = "dashed")} + 
@@
   {if(!isTRUE(legend)) theme(legend.position = "none")} + 
-  guides(colour = guide_legend(reverse=T)) +
+  # guides(colour = guide_legend(reverse=T)) +
   facet_grid(rows = vars(type),
@@
 
-Determine amount of missingness per respondent per wave. Because calculation takes long, this chunk is not executed but loaded from memory. To execute, set eval == TRUE.
+Determine amount of missingness per respondent per wave. Because calculation takes long, this chunk is loaded from memory.
 
@@
 
-Load imputed data from disc.
-
-```{r}
+```{r eval=FALSE}
+# to further increase speed, you can also load the imputed data from the directory
 d_long_imp <- read_csv("data/data_imputed.csv")
@@
 # Analyses
-## Life satisfaction
-
-First, without control variables.
-
-```{r}
-model_life_sat_lmer_nco <- lmerTest::lmer(life_sat ~ 
-                            (1 | id) + (1 | wave) + 
-                            soc_med_read_w + soc_med_like_share_w + soc_med_post_w + 
-                            soc_med_fb_w + soc_med_ig_w + soc_med_wa_w + soc_med_yt_w + soc_med_tw_w +
-                            soc_med_read_b + soc_med_like_share_b + soc_med_post_b + 
-                            soc_med_fb_b + soc_med_ig_b + soc_med_wa_b + soc_med_yt_b + soc_med_tw_b,
-                data = d_long_imp)
-summary(model_life_sat_lmer_nco)
-```
-
-Now with control variables.
+## Preregistered
+### Life satisfaction
 
@@
 
-## Positive Affect
-
-First, without control variables.
-
-```{r}
-model_aff_pos_lmer_nco <- lmerTest::lmer(aff_pos_fs ~ 
-                                        (1 | id) + (1 | wave) + 
-                                        soc_med_read_w + soc_med_like_share_w + soc_med_post_w + 
-                                        soc_med_fb_w + soc_med_ig_w + soc_med_wa_w + soc_med_yt_w + soc_med_tw_w +
-                                        soc_med_read_b + soc_med_like_share_b + soc_med_post_b + 
-                                        soc_med_fb_b + soc_med_ig_b + soc_med_wa_b + soc_med_yt_b + soc_med_tw_b
-                                      , data = d_long_imp)
-summary(model_aff_pos_lmer_nco)
-```
-
-Now with control variables.
+### Positive Affect
 
@@
 
-## Negative Affect
-
-First, without control variables.
-
-```{r}
-model_aff_neg_lmer_nco <- lmerTest::lmer(aff_neg_fs ~ 
-                                        (1 | id) + (1 | wave) + 
-                                        soc_med_read_w + soc_med_like_share_w + soc_med_post_w + 
-                                        soc_med_fb_w + soc_med_ig_w + soc_med_wa_w + soc_med_yt_w + soc_med_tw_w +
-                                        soc_med_read_b + soc_med_like_share_b + soc_med_post_b + 
-                                        soc_med_fb_b + soc_med_ig_b + soc_med_wa_b + soc_med_yt_b + soc_med_tw_b
-                                      , data = d_long_imp)
-summary(model_aff_neg_lmer_nco)
-```
-
-Then, with control variables.
+### Negative Affect
 
@@
 
+## Without control variables
+### Life Satisfaction
+
+```{r}
+model_life_sat_lmer_nco <- lmerTest::lmer(life_sat ~ 
+                            (1 | id) + (1 | wave) + 
+                            soc_med_read_w + soc_med_like_share_w + soc_med_post_w + 
+                            soc_med_fb_w + soc_med_ig_w + soc_med_wa_w + soc_med_yt_w + soc_med_tw_w +
+                            soc_med_read_b + soc_med_like_share_b + soc_med_post_b + 
+                            soc_med_fb_b + soc_med_ig_b + soc_med_wa_b + soc_med_yt_b + soc_med_tw_b,
+                data = d_long_imp)
+summary(model_life_sat_lmer_nco)
+```
+
+### Positive Affect
+
+```{r}
+model_aff_pos_lmer_nco <- lmerTest::lmer(aff_pos_fs ~ 
+                                        (1 | id) + (1 | wave) + 
+                                        soc_med_read_w + soc_med_like_share_w + soc_med_post_w + 
+                                        soc_med_fb_w + soc_med_ig_w + soc_med_wa_w + soc_med_yt_w + soc_med_tw_w +
+                                        soc_med_read_b + soc_med_like_share_b + soc_med_post_b + 
+                                        soc_med_fb_b + soc_med_ig_b + soc_med_wa_b + soc_med_yt_b + soc_med_tw_b
+                                      , data = d_long_imp)
+summary(model_aff_pos_lmer_nco)
+```
+
+### Negative Affect
+
+```{r}
+model_aff_neg_lmer_nco <- lmerTest::lmer(aff_neg_fs ~ 
+                                        (1 | id) + (1 | wave) + 
+                                        soc_med_read_w + soc_med_like_share_w + soc_med_post_w + 
+                                        soc_med_fb_w + soc_med_ig_w + soc_med_wa_w + soc_med_yt_w + soc_med_tw_w +
+                                        soc_med_read_b + soc_med_like_share_b + soc_med_post_b + 
+                                        soc_med_fb_b + soc_med_ig_b + soc_med_wa_b + soc_med_yt_b + soc_med_tw_b
+                                      , data = d_long_imp)
+summary(model_aff_neg_lmer_nco)
+```
+
+## Standardized results
+
+First, create standardized dataset.
+
+```{r}
+d_long_std <- d_long_imp %>% 
+  mutate(across(c(-id, -wave, -gender, -male, -born_aus, -born_aus_prnts, -edu_fac, -employment_fac, -res_vienna, -acc_bal, -acc_gar), scale))
+```
+
+Then let's rerun the analyses.
+
+### Life satisfaction
+
+```{r}
+model_life_sat_lmer_std <- lmerTest::lmer(life_sat ~
+                                        (1 | id) + (1 | wave) + 
+                                        soc_med_read_w + soc_med_like_share_w + soc_med_post_w  + 
+                                        soc_med_fb_w + soc_med_ig_w + soc_med_wa_w + soc_med_yt_w + soc_med_tw_w +
+                                        soc_med_read_b + soc_med_like_share_b + soc_med_post_b + 
+                                        soc_med_fb_b + soc_med_ig_b + soc_med_wa_b + soc_med_yt_b + soc_med_tw_b +  
+                                        age + male + born_aus + born_aus_prnts + edu_fac + employment_fac + health_b + 
+                                        res_vienna + acc_bal + acc_gar + home_sqm + 
+                                        med_txt_kro_b + med_txt_sta_b + med_txt_pre_b + med_txt_oes_b + med_txt_kur_b + med_txt_slz_b + med_txt_son_b + 
+                                        med_vid_orf_b + med_vid_pri_b + 
+                                        risk_prop_b + loc_cntrl_int_m_b + 
+                                        act_wrk_w + act_spo_w + act_frn_w + act_sho_w + act_pet_w + 
+                                        act_wrk_b + act_spo_b + act_frn_b + act_sho_b + act_pet_b +
+                                        sat_dem_w + sat_dem_b 
+                                      , data = d_long_std)
+summary(model_life_sat_lmer_std)
+```
+
+### Positive affect
+
+```{r}
+model_aff_pos_lmer_std <- lmerTest::lmer(aff_pos_fs ~ 
+                                        (1 | id) + (1 | wave) + 
+                                        soc_med_read_w + soc_med_like_share_w + soc_med_post_w + 
+                                        soc_med_fb_w + soc_med_ig_w + soc_med_wa_w + soc_med_yt_w + soc_med_tw_w +
+                                        soc_med_read_b + soc_med_like_share_b + soc_med_post_b + 
+                                        soc_med_fb_b + soc_med_ig_b + soc_med_wa_b + soc_med_yt_b + soc_med_tw_b +  
+                                        age + male + born_aus + born_aus_prnts + edu_fac + employment_fac + health_b + 
+                                        res_vienna + acc_bal + acc_gar + home_sqm + 
+                                        med_txt_kro_b + med_txt_sta_b + med_txt_pre_b + med_txt_oes_b + med_txt_kur_b + med_txt_slz_b + med_txt_son_b + 
+                                        med_vid_orf_b + med_vid_pri_b + 
+                                        risk_prop_b + loc_cntrl_int_m_b + 
+                                        act_wrk_w + act_spo_w + act_frn_w + act_sho_w + act_pet_w + 
+                                        act_wrk_b + act_spo_b + act_frn_b + act_sho_b + act_pet_b +
+                                        sat_dem_w + sat_dem_b 
+                                      , data = d_long_imp)
+summary(model_aff_pos_lmer_std)
+```
+
+### Negative Affect
+
+```{r}
+model_aff_neg_lmer_std <- lmerTest::lmer(aff_neg_fs ~ 
+                                        (1 | id) + (1 | wave) + 
+                                        soc_med_read_w + soc_med_like_share_w + soc_med_post_w + 
+                                        soc_med_fb_w + soc_med_ig_w + soc_med_wa_w + soc_med_yt_w + soc_med_tw_w +
+                                        soc_med_read_b + soc_med_like_share_b + soc_med_post_b + 
+                                        soc_med_fb_b + soc_med_ig_b + soc_med_wa_b + soc_med_yt_b + soc_med_tw_b +  
+                                        age + male + born_aus + born_aus_prnts + edu_fac + employment_fac + health_b + 
+                                        res_vienna + acc_bal + acc_gar + home_sqm + 
+                                        med_txt_kro_b + med_txt_sta_b + med_txt_pre_b + med_txt_oes_b + med_txt_kur_b + med_txt_slz_b + med_txt_son_b + 
+                                        med_vid_orf_b + med_vid_pri_b + 
+                                        risk_prop_b + loc_cntrl_int_m_b + 
+                                        act_wrk_w + act_spo_w + act_frn_w + act_sho_w + act_pet_w + 
+                                        act_wrk_b + act_spo_b + act_frn_b + act_sho_b + act_pet_b +
+                                        sat_dem_w + sat_dem_b 
+                                      , data = d_long_imp)
+summary(model_aff_neg_lmer_std)
+```
+
 # Figures
 
-First, for activities.
+## Activities
 
@@
                                             type = "activity", analysis = "no controls")
+dat_fig_results_activity_std <- get_dat_res(model_aff_neg_lmer_std, model_aff_pos_lmer_std, model_life_sat_lmer_std, 
+                                            type = "activity", analysis = "standardized")
+
 dat_fig_results_activity <- rbind(
   dat_fig_results_activity, 
-  dat_fig_results_activity_nco
-)
+  dat_fig_results_activity_nco,
+  dat_fig_results_activity_std
+) %>% 
+  mutate(Analysis = factor(Analysis, levels = c("regular", "no controls", "standardized")))
 
@@
 
-Then for channels.
+## Channels
 
@@
                                             type = "channels", analysis = "no controls")
+dat_fig_results_channels_std <- get_dat_res(model_aff_neg_lmer_std, model_aff_pos_lmer_std, model_life_sat_lmer_std, 
+                                            type = "channels", analysis = "standardized")
+
 dat_fig_results_channels <- rbind(
   dat_fig_results_channels, 
-  dat_fig_results_channels_nco
-)
+  dat_fig_results_channels_nco,
+  dat_fig_results_channels_std
+) %>% 
+  mutate(Analysis = factor(Analysis, levels = c("regular", "no controls", "standardized")))
 
@@
 
-Select control variables and make graph.
+## Selected control variables
+
+Standardized, to allow for better comparison across differently scaled variables.
 
@@
 # make figure
-fig_results_control <- make_graph_res(
-  get_dat_res(model_aff_neg_lmer, model_aff_pos_lmer, model_life_sat_lmer, analysis = "exploratory", type = "control"),
+fig_results_control_std <- make_graph_res(
+  get_dat_res(model_aff_neg_lmer_std, model_aff_pos_lmer_std, model_life_sat_lmer_std, 
+              type = "control", analysis = "standardized"),
+  sesoi = "std",
   legend = FALSE
   )
-fig_results_control
+fig_results_control_std
 
 # save figure
-ggsave("figures/fig_results_control.pdf", 
+ggsave("figures/fig_results_control_std.pdf", 
        width = 7, height = 4,
-       plot = fig_results_control)
+       plot = fig_results_control_std)
 ```

### 2022-04-11 analyses_additional.Rmd
- updated figures

@@
 
-Here you can find the code and the results of all additional analyses. To see the code, click on button "Code". Alternatively, you can download the rmd file from the github repo.
+ - Here you can find the code and the results of all additional analyses. 
+ - To see the results, I recommend you directly go down to the Figures section. 
+ - But if you're interested in how the analyses were executed, just tag along.
+ - To see the code, click on button "Code". Alternatively, you can download the rmd file from the github repo.
+ - To execute the code, first download the actual data from AUSSDA, then run analysis.rmd, then this file
 
@@
 
-Load packages, and make repo replicable.
+Load packages.
 
@@
 load("data/workspace_1.RData")
-```
-
-# Standardized results
-
-First, create standardized dataset.
-
-```{r}
-d_long_std <- d_long_imp %>% 
-  mutate(across(c(-id, -wave, -gender, -male, -born_aus, -born_aus_prnts, -edu_fac, -employment_fac, -res_vienna, -acc_bal, -acc_gar), scale))
-```
-
-Then let's rerun the analyses.
-
-## Life satisfaction
-
-```{r}
-model_life_sat_lmer_std <- lmerTest::lmer(life_sat ~
-                                        (1 | id) + (1 | wave) + 
-                                        soc_med_read_w + soc_med_like_share_w + soc_med_post_w  + 
-                                        soc_med_fb_w + soc_med_ig_w + soc_med_wa_w + soc_med_yt_w + soc_med_tw_w +
-                                        soc_med_read_b + soc_med_like_share_b + soc_med_post_b + 
-                                        soc_med_fb_b + soc_med_ig_b + soc_med_wa_b + soc_med_yt_b + soc_med_tw_b +  
-                                        age + male + born_aus + born_aus_prnts + edu_fac + employment_fac + health_b + 
-                                        res_vienna + acc_bal + acc_gar + home_sqm + 
-                                        med_txt_kro_b + med_txt_sta_b + med_txt_pre_b + med_txt_oes_b + med_txt_kur_b + med_txt_slz_b + med_txt_son_b + 
-                                        med_vid_orf_b + med_vid_pri_b + 
-                                        risk_prop_b + loc_cntrl_int_m_b + 
-                                        act_wrk_w + act_spo_w + act_frn_w + act_sho_w + act_pet_w + 
-                                        act_wrk_b + act_spo_b + act_frn_b + act_sho_b + act_pet_b +
-                                        sat_dem_w + sat_dem_b 
-                                      , data = d_long_std)
-summary(model_life_sat_lmer_std)
-```
-
-## Positive affect
-
-```{r}
-model_aff_pos_lmer_std <- lmerTest::lmer(aff_pos_fs ~ 
-                                        (1 | id) + (1 | wave) + 
-                                        soc_med_read_w + soc_med_like_share_w + soc_med_post_w + 
-                                        soc_med_fb_w + soc_med_ig_w + soc_med_wa_w + soc_med_yt_w + soc_med_tw_w +
-                                        soc_med_read_b + soc_med_like_share_b + soc_med_post_b + 
-                                        soc_med_fb_b + soc_med_ig_b + soc_med_wa_b + soc_med_yt_b + soc_med_tw_b +  
-                                        age + male + born_aus + born_aus_prnts + edu_fac + employment_fac + health_b + 
-                                        res_vienna + acc_bal + acc_gar + home_sqm + 
-                                        med_txt_kro_b + med_txt_sta_b + med_txt_pre_b + med_txt_oes_b + med_txt_kur_b + med_txt_slz_b + med_txt_son_b + 
-                                        med_vid_orf_b + med_vid_pri_b + 
-                                        risk_prop_b + loc_cntrl_int_m_b + 
-                                        act_wrk_w + act_spo_w + act_frn_w + act_sho_w + act_pet_w + 
-                                        act_wrk_b + act_spo_b + act_frn_b + act_sho_b + act_pet_b +
-                                        sat_dem_w + sat_dem_b 
-                                      , data = d_long_imp)
-summary(model_aff_pos_lmer_std)
-```
-
-## Negative Affect
-
-```{r}
-model_aff_neg_lmer_std <- lmerTest::lmer(aff_neg_fs ~ 
-                                        (1 | id) + (1 | wave) + 
-                                        soc_med_read_w + soc_med_like_share_w + soc_med_post_w + 
-                                        soc_med_fb_w + soc_med_ig_w + soc_med_wa_w + soc_med_yt_w + soc_med_tw_w +
-                                        soc_med_read_b + soc_med_like_share_b + soc_med_post_b + 
-                                        soc_med_fb_b + soc_med_ig_b + soc_med_wa_b + soc_med_yt_b + soc_med_tw_b +  
-                                        age + male + born_aus + born_aus_prnts + edu_fac + employment_fac + health_b + 
-                                        res_vienna + acc_bal + acc_gar + home_sqm + 
-                                        med_txt_kro_b + med_txt_sta_b + med_txt_pre_b + med_txt_oes_b + med_txt_kur_b + med_txt_slz_b + med_txt_son_b + 
-                                        med_vid_orf_b + med_vid_pri_b + 
-                                        risk_prop_b + loc_cntrl_int_m_b + 
-                                        act_wrk_w + act_spo_w + act_frn_w + act_sho_w + act_pet_w + 
-                                        act_wrk_b + act_spo_b + act_frn_b + act_sho_b + act_pet_b +
-                                        sat_dem_w + sat_dem_b 
-                                      , data = d_long_imp)
-summary(model_aff_neg_lmer_std)
 ```
@@
 # Figures
-## Standardized results
 
-First, for activities.
+In what follows, see figure with all results combined.
 
-```{r}
-# make figure
-fig_results_activity_std <- make_graph_res(
-  get_dat_res(model_aff_neg_lmer_std, model_aff_pos_lmer_std, model_life_sat_lmer_std, 
-              type = "activity", analysis = "standardized"),
-  sesoi = "std",
-  legend = FALSE
-  )
-fig_results_activity_std
-
-# safe figure
-ggsave("figures/fig_results_activity_std.pdf", 
-       width = 7, height = 4,
-       plot = fig_results_activity_std)
-```
-
-Then for channels.
-
-```{r}
-# make figure
-fig_results_channels_std <- make_graph_res(
-  get_dat_res(model_aff_neg_lmer_std, model_aff_pos_lmer_std, model_life_sat_lmer_std, type = "channels", analysis = "standardized"),
-  sesoi = "std",
-  legend
-  )
-fig_results_channels_std
-
-# save figure
-ggsave("figures/fig_results_channels_std.pdf", 
-       width = 7, height = 4,
-       plot = fig_results_channels_std)
-```
-
-Select control variables and make graph.
-
-```{r}
-# make figure
-fig_results_control_std <- make_graph_res(
-  get_dat_res(model_aff_neg_lmer_std, model_aff_pos_lmer_std, model_life_sat_lmer_std, 
-              type = "control", analysis = "standardized"),
-  sesoi = "std"
-  )
-fig_results_control_std
-
-# save figure
-ggsave("figures/fig_results_control_std.pdf", 
-       width = 7, height = 4,
-       plot = fig_results_control_std)
-```
-
-## Comparison of unstandardized methods
-
-In what follows, see figure with all results combined. Only nonstandardized results are presented to use the same scaling.
-
-First, for activities.
+## Activities
 

### 2022-04-12 analyses.RmD
- implented reviewer feedback

@@
                 lwd = .75, height = .2, position = position_dodge(-.7)) +
-  geom_point(aes(color = Analysis), size = 2, position = position_dodge(-.7)) +
+  geom_point(aes(color = Analysis), size = 1.5, position = position_dodge(-.7)) +
   {if(isTRUE(sesoi == "est")) geom_vline(data=filter(data, dv=="Life satisfaction"), aes(xintercept=-.3), colour="darkgrey", linetype = "dashed")} + 
@@
 
-ggsave("figures/fig_descriptives.pdf", 
-       width = 10, height = 6,
+ggsave("figures/fig_descriptives.png", 
+       width = 10, height = 5,
        plot = fig_desc)
@@
 
-```{r}
+```{r warning=FALSE}
 # get data
@@
 # safe figure
-ggsave("figures/fig_results_activity.pdf", 
+ggsave("figures/fig_results_activity.png", 
        width = 7, height = 4,
@@
 # safe figure
-ggsave("figures/fig_results_channels.pdf", 
+ggsave("figures/fig_results_channels.png", 
        width = 7, height = 4,
@@
 
-```{r}
+```{r warning=FALSE}
 # make figure
@@
 # save figure
-ggsave("figures/fig_results_control_std.pdf", 
+ggsave("figures/fig_results_control_std.png", 
        width = 7, height = 4,

### 2022-04-12 analyses_additional.Rmd
- implented reviewer feedback

@@
 # make figure
-fig_results_activity <- make_graph_res(
+fig_results_activity_comparison <- make_graph_res(
   dat_fig_results_activity,
@@
   )
-fig_results_activity
+fig_results_activity_comparison
 
 # safe figure
-ggsave("figures/fig_results_activity.pdf", 
+ggsave("figures/fig_results_activity_comparison.pdf", 
        width = 7, height = 4,
@@
 # make figure
-fig_results_channels <- make_graph_res(
+fig_results_channels_comparison <- make_graph_res(
   dat_fig_results_channels,
@@
   )
-fig_results_channels
+fig_results_channels_comparison
 
 # safe figure
-ggsave("figures/fig_results_channels.pdf", 
+ggsave("figures/fig_results_channels_comparison.pdf", 
        width = 7, height = 4,

### 2022-04-12 manuscript.Rmd
- implented reviewer feedback

@@
 To this end, many people use various types of media, and increasingly social media.
-However, because social media are particularly engaging, some find it hard to disconnect and cannot stop 'doomscrolling'.
+However, because social media are particularly engaging, some find it hard to disconnect.
 In this preregistered study, I investigate whether using social media for COVID-19 related reasons affects psychological well-being.
-To answer this question I analyzed data from the Austrian Corona Panel Project, which consists of 24 waves with overall 3,018 participants.
+To answer this question I analyzed data from the Austrian Corona Panel Project, which consists of 3,018 participants.
+Well-being was measured at each wave, and communication at five waves.
 I ran three random effects within between models, controlling for several stable and varying confounders.
 Results showed that the effects of COVID-19 related social media use on well-being were very small, arguably too small to matter.
-The findings suggest that fears that social media use during times of crisis impairs well-being are likely to be unfounded."
+Fears that social media use during times of crisis critically impairs well-being are likely to be unfounded."
 keywords          : "COVID-19, well-being, social media, news use, panel study."
@@
 classoption       : "man"
-# output            : papaja::apa6_pdf
+# output            :
+#   papaja::apa6_pdf:
+#     latex_engine  : xelatex
 output            : papaja::apa6_docx
@@
 # devtools::install_github("https://github.com/tdienlin/td@v.0.0.2.5")
-# 1devtools::install_github("https://github.com/crsh/papaja@devel")
+# devtools::install_github("https://github.com/crsh/papaja@devel")
 
 # load packages
-packages <- c("devtools", "english", "ggplot2", "kableExtra", 
+packages <- c("devtools", "english", "ggplot2", "gridExtra", "kableExtra", 
               "lavaan", "lme4", "lmerTest",
@@
 
-## Effects of Social Media on Well-Being
+## Theorizing Social Media Effects on Well-Being
 
-So far, there is only little empirical research on COVID-19 related social media use on well-being.
+<!-- In conclusion, most effects are likely somewhere between trivial and small. -->
+<!-- I therefore expect that also in the case of COVID-19 related social media use effects will be trivial to small. -->
+
+From a theoretical perspective, how could we explain whether COVID-19 related social media use might affect well-being?
+<!-- In what follows, I outline potential arguments as to why the effect might be positive or negative, direct or indirect, or nonexistent. -->
+According to the set-point model of subjective well-being, well-being is surprisingly stable [@lykkenHappinessWhatStudies1999].
+Although specific events such as marriage or salary can have significant effects, after some time well-being routinely returns to prior levels, which are mostly determined genetically [@sheldonStabilityHappinessTheories2014].
+Only very specific events such as unemployment, disability, or death can cause long-term decreases in well-being [@lucasAdaptationSetpointModel2007].
+So although well-being can change this does not happen easily.
+<!-- Although the set-point theory has been criticized lately because events such as marriage, unemployment, or death can indeed lead to long-term changes, it is still widely maintained that well-being is very stable and hence hard to be influenced  -->
+
+Can media use be such a negative or positive factor?
+In advance, there doesn't seem to be a clear winner, and it's likely that both positive and negative effects cancel each other out.
+Empirically, social media use---on average---does not have a strong effect on well-being [@meierComputermediatedCommunicationSocial2020a].
+According to the Differential Susceptibility to Media Effects Model [@valkenburgDifferentialSusceptibilityMedia2013], the effects of media use differ across individuals. 
+Whereas for some media are beneficial, for others they are harmful.
+On average, however, effects are often small or negligible.
+Also when focusing on individual users, social media have both positive and negative effects on well-being [@buchiDigitalWellbeingTheory2021].
+They can impair well-being when causing embarrassment, stress, or disinformation, and they can improve well-being when providing connectedness, information, or entertainment [@buchiDigitalWellbeingTheory2021].
+
+Two of the most prominent media effect theories argue against strong negative impacts.
+First, uses and gratifications theory states that people explicitly and rationally chose specific media because of their respective benefits. 
+If those benefits don't exist, they invest their time elsewhere.
+And social media offer ample benefits. 
+Most prominently, they help find relevant information, maintain and foster relationships, express one's personality, and entertain oneself [@pelletierOneSizeDoesn2020].
+The second major theory is mood management theory [@zillmannMoodManagementCommunication1988].
+Users implicitly learn what media help them balance their mood and affect.
+For example, when bored people use social media to entertain themselves.
+Using experience sampling of well-being and logs of social media use, a study with 82 participants from Italy found that after episodes of social media use, levels of positive affect increased for a short time [@marcianoDynamicsAdolescentsSmartphone2022].
+
+But people also misjudge media affects, often being overly optimistic [@metzgerComparativeOptimismPrivacy2017].
+And precisely because social media have so many positive consequences, one can ask if this where the actual problem lies.
+In other words, social media aren't problematic because they're inherently bad, but rather because they're too good.
+And as with many other things, there can be too much of a good thing.
+It is therefore often asked whether social media are addictive, and users sometimes express this fear themselves [@yangCanWatchingOnline2021]. 
+However, a recently published meta-analysis found that the two most prominent measures of addiction, the Bergen Facebook Addiction Scale and the Bergen Social Media Addiction Scale, have only small relations to well-being [@duradoniWellbeingSocialMedia2020].
+In addition, the general idea of labeling excessive social and new media use as addiction was criticized, arguing that social and new media represent new regular behaviors that should not be pathologized [@galerHowMuchToo2018; @vanrooijWeakScientificBasis2018].
+
+Because effects can differ across situations, I now briefly focus on the effects of COVID-19 related social media use specifically.
+First, one could assume a _direct_ negative effect on well-being, and especially on positive or negative affect, which are more volatile and fluctuating.
+Dangers, inequalities, corruption---these were the headlines during the pandemic across many countries worldwide.
+If one learns about such events, the initial reaction might be shock, fear, or dismay.
+Consuming such news can be depressing [@dornemannHowGoodBad2021], perhaps even changing some general perspectives on life.
+<!-- So, just like being hit by a hammer hurts and we don't need any "mediating mechanism", this could be the case here as well. -->
+That said, because not all news was negative, and because many people showed solidarity and compassion, there was also positive and uplifting content, potentially compensating for the negative effects [@dornemannHowGoodBad2021].
+A study with 2.057 respondents from Italy reported that during the pandemic virtual community and social connectedness even increased during the pandemic [@guazziniSecondWaveAnalysis2022].
+<!-- However, in light of a worldwide pandemic with millions of deaths, the negative direct effect seems more plausible. -->
+
+There could also be _indirect_ effects.
+When browsing social media for Covid-19 related news, many users reported being captivated to such an extent that they could not stop using social media [@kleinDarklySoothingCompulsion2021].
+During the pandemic social media use was at an all-time high in the US [@statistaAverageDailyTime2021].
+It is most likely that moderate social media use is not detrimental [@orbenTeenagersScreensSocial2020].
+Overuse, however, might be more critical, and several studies have shown more pronounced negative effects for extreme users [@przybylskiLargescaleTestGoldilocks2017].
+To explain, overuse could impair well-being if it replaces meaningful or functional activities such as meeting others, working, actively relaxing, or exercising.
+So if a society collectively overuses social media during a pandemic, there might be potential for negative effects.
+<!-- If overuse replaces such activities it's reasonable to assume that it's also detrimental. -->
+On the other hand, one can make the case that overuse might also be beneficial, especially in times of a pandemic---even if the use is mainly COVID-19 related.
+Exchanging COVID-19 related messages with friends via WhatsApp might replace the in-person contact one would have otherwise, but which is literally impossible at the time.
+In situations where meaningful and functional activities are prohibited, using social media to exchange about COVID-19 related topics might not be the worst idea.
+Besides, given that nowadays a large number of experts, scientists, and politicians converse directly on social media, one can get first-hand high quality information on current developments.
+<!-- On the other hand, there is of course also much disinformation, and "bingeing" on COVID-19 fake news might also pose risks for impaired well-being. -->
+
+To summarize, from a theoretical perspective it is most likely that the average effects of social media use on well-being are negligible. 
+Building on established theories from Communication, it is not particularly likely that effects are either profoundly negative or strongly positive.
+
+## Empirical Studies on Social Media Effects
+
+So far, there is only little empirical research on how well-being is affected by COVID-19 related social media.
 In their study on the relations between media use and mental health during the pandemic, @bendauAssociationsCOVID19Related2021 found that people who used social media as a primary source of information reported on average "significantly more unspecific anxiety and depression [] and significantly more specific COVID-19 related anxiety symptoms" (p.
@@
 
-It is often asked if social media use can become addictive. 
-Users sometimes express this fear themselves [@yangCanWatchingOnline2021]. 
-A recently published meta-analysis found that the two most prominent measures of addiction, the Bergen Facebook Addiction Scale and the Bergen Social Media Addiction Scale, have only small relations to well-being [@duradoniWellbeingSocialMedia2020].
-That said, the general idea of labeling excessive social and new media use as addiction was criticized, arguing that social and new media represent new regular behaviors that should not be pathologized [@galerHowMuchToo2018; @vanrooijWeakScientificBasis2018].
-
 Several individual studies employing advanced methods found smalls relations between social media use and well-being [@orbenSocialMediaEnduring2019; @przybylskiDoesTakingShort2021a; @schemerImpactInternetSocial2021; @kerestesAdolescentsOnlineSocial2020].
 For example, @beyensSocialMediaUse2021 reported that although for some users (roughly one quarter) the effects of social media use on well-being were negative, for almost the same number of users they were positive, while for the rest the effects were neutral.
-This finding is aligned the Differential Susceptibility to Media Effects Model, which suggests that although there is substantial variation of media effects for individual users, the average effects reported in the literature are often small [@valkenburgDifferentialSusceptibilityMedia2013].
+This finding is aligned the Differential Susceptibility to Media Effects Model.
+Although there is substantial variation of media effects for individual users, the average effects reported in the literature are often small [@valkenburgDifferentialSusceptibilityMedia2013].
 
-<!-- In conclusion, most effects are likely somewhere between trivial and small. -->
-<!-- I therefore expect that also in the case of COVID-19 related social media use effects will be trivial to small. -->
+<!-- Together, the strongest argument to me is that _in general_ the effects of social media on well-being are, on average, small at best.  -->
+<!-- Because this study only looks at _one part_ of social media use---namely, COVID-19 related interactions---it is very focused, diminishing the overall potential of the effects even further. -->
+<!-- Whether or not using social media for COVID-19 related aspects is detrimental during a pandemic is also not entirely clear. -->
+<!-- At least, there does not seem to be a case where we should expect a clear deviation to the negative, despite anecdotes of doomscrolling. -->
 
-From a theoretical perspective, how could we explain whether well-being might be affected by COVID-19 related social media use in particular?
-<!-- In what follows, I outline potential arguments as to why the effect might be positive or negative, direct or indirect, or nonexistent. -->
-<!-- In advance, there doesn't seem to be a clear winner, and it's likely that both positive and negative effects cancel each other out. -->
-According to the set-point model of subjective well-being, well-being is surprisingly stable [@lykkenHappinessWhatStudies1999].
-Although specific events such as marriage or salary can increase well-being, after some time it routinely returns to prior levels, which are mostly determined genetically [@sheldonStabilityHappinessTheories2014].
-Only very specific very negative events such as unemployment, disability, or death can cause long-term decreases in well-being [@lucasAdaptationSetpointModel2007], and specific positive factors such as XXX to long-term increase [].
-
-Can media use be such a negative or positive factor? 
-<!-- Although the set-point theory has been criticized lately because events such as marriage, unemployment, or death can indeed lead to long-term changes, it is still widely maintained that well-being is very stable and hence hard to be influenced  -->
-
-one could assume a _direct_ negative effect on well-being, and especially on positive or negative affect, which are more volatile and fluctuating.
-Dangers, inequalities, corruption---these were the headlines during the pandemic across many countries worldwide.
-If one learns about such events, the initial reaction might be shock, fear, or dismay.
-Consuming such news can be depressing [@dornemannHowGoodBad2021], perhaps even changing some general perspectives on life.
-<!-- So, just like being hit by a hammer hurts and we don't need any "mediating mechanism", this could be the case here as well. -->
-That said, because not all news was negative, and because many people showed solidarity and compassion, there was also positive and uplifting content, potentially compensating for the negative effects [@dornemannHowGoodBad2021].
-A study with 2.057 respondents from Italy reported that during the pandemic virtual community and social connectedness potentially even increased [@guazziniSecondWaveAnalysis2022].
-<!-- However, in light of a worldwide pandemic with millions of deaths, the negative direct effect seems more plausible. -->
-
-There could also be _indirect_ effects.
-When doomscrolling, users are captivated to such an extent that they cannot stop using social media.
-For example, during the pandemic social media use was at an all-time high in the US [@statistaAverageDailyTime2021].
-In general, as has been expressed by many before, it is most likely that moderate social media use is not detrimental [@orbenTeenagersScreensSocial2020].
-Overuse, however, might be more critical, and several studies have shown more pronounced negative effects for extreme users [@przybylskiLargescaleTestGoldilocks2017].
-To explain, overuse likely impairs well-being if it replaces meaningful or functional activities such as meeting others, working, actively relaxing, or exercising.
-So if a society collectively overuses social media when doomscrolling, there is potential for negative effects.
-<!-- If overuse replaces such activities it's reasonable to assume that it's also detrimental. -->
-
-On the other hand, one can make the case that overuse might also be beneficial, especially in times of a pandemic---even if the use is mainly COVID-19 related.
-Exchanging COVID-19 related messages with friends via WhatsApp might replace the in-person contact one would have otherwise, but which is literally impossible at the time.
-In situations where meaningful and functional activities are prohibited, using social media to exchange about COVID-19 related topics might not be the worst idea.
-Besides, given that nowadays a large number of experts, scientists, and politicians converse directly on social media, one can get first-hand high quality information on current developments.
-<!-- On the other hand, there is of course also much disinformation, and "bingeing" on COVID-19 fake news might also pose risks for impaired well-being. -->
-
-Together, the strongest argument to me is that _in general_ the effects of social media on well-being are, on average, small at best.
-Because this study only looks at _one part_ of social media use---namely, COVID-19 related interactions---it is very focused, diminishing the overall potential of the effects even further.
-Whether or not using social media for COVID-19 related aspects is detrimental during a pandemic is also not entirely clear.
-<!-- At least, there does not seem to be a case where we should expect a clear deviation to the negative, despite anecdotes of doomscrolling. -->
-Therefore, I expect that COVID-19 related communication on social media doesn't affect well-being in a meaningful or relevant way.
+In conclusion, in light of the theoretical considerations and the empirical studies presented above, I expect that COVID-19 related communication on social media doesn't affect well-being in a meaningful or relevant way.
 
@@
 gender, age, education, Austria country of birth, Austria country of birth of parents, text-based news consumption, video-based news consumption, residency Vienna, household size, health, living space, access to garden, access to balcony, employment, work hours per week, being in home-office, household income, outdoor activities, satisfaction with democracy, disposition to take risks, and locus of control.
-I will not control for variables such as trust in institutions or trust in media, because these variables might be influenced by social media use to a meaningful extent.
+<!-- I will not control for variables such as trust in institutions or trust in media, because these variables might be influenced by social media use to a meaningful extent. -->
+The dataset includes many other variables that one could also potentially control for, and I invite interested readers to download the and explore potential interesting relationships.
 
@@
 
-In this study, I therefore adopt an intermediate perspective.
-I analyze if when a person changes their social media diet, are there _simultaneous_ changes in their well-being?
+In this study, I analyze how using social media during the last week affected their positive and negative affect during the last week.
+More precisely, if people during the last week used COVID-19 related social media more than the usually do, did they feel better or worse during that week than they usually do?
+Regarding life satisfaction, I implemented a longer interval.
+Namely, if people during the last week used COVID-19 related social media more than the usually do, did this affect if they are now more or less satisfied with their lives than they usually are?
+I hence analyze if when a person changes their social media diet, are there (a) _simultaneous_ changes in their affect and (b) _subsequent_ changes in their life satisfaction?
 When additionally controlling for both stable and varying confounders, we can then be more sure that the effect is indeed causal.
-<!-- Similar approaches were already implemented by several other studies [@johannesNoEffectDifferent2021; @scharkowHowSocialNetwork2020] and they are considered a best practice to analyze causality [@bellFixedRandomEffects2019].
--->
+Similar approaches were implemented by other studies [@johannesNoEffectDifferent2022; @scharkowHowSocialNetwork2020], and they are considered a best practice to analyze causality [@bellFixedRandomEffects2019].
 
@@
 If the confidence interval falls completely outside of the null-region (e.g., _b_ = .40, [95% CI: .35, .45]), the hypothesis is rejected and the existence of a meaningful positive effect is supported.
+For an illustration, see Figure \@ref(fig:sesoi)).
 
@@
 
-For more information on the analyses, a complete documentation of the models and results, see [companion website](https://XMtRA.github.io/Austrian_Corona_Panel_Project).
+For more information on the analyses, a complete documentation of the models and results, additional analyses (for example using multiple imputation or no imputation), see [companion website](https://XMtRA.github.io/Austrian_Corona_Panel_Project).
 
@@
 In what follows, I list all the variables that I analyzed.
-For the variables' means, range, and variance, see Table \@ref(tab:descriptives).
+For the variables' means, range, and variance, see Table \@ref(tab:tab-descriptives).
 For a complete list of all items and item characteristics, see [companion website](https://XMtRA.github.io/Austrian_Corona_Panel_Project).
@@
 
-COVID-19 related social media use focused on communication types was measured with the three dimensions of (a) reading, (b) liking and sharing, and (c) posting from @wagnerAUTNESOnlinePanel2018, adapted for the context of this study.
+COVID-19 related social media use focused on communication types was measured with the three dimensions of (a) reading, (b) liking and sharing, and (c) posting.
+The items come from @wagnerAUTNESOnlinePanel2018 and were adapted for the context of this study.
 The general introductory question was "How often during the last week have you engaged in the following activities on social media?"
@@
 
-First, when looking at the variables from a descriptive perspective (Figure 1), we see that all well-being measures did not change substantially across the different waves of data collection.
+First, when looking at the variables from a descriptive perspective (Figure \@ref(fig:descriptives))), we see that all well-being measures did not change substantially across the different waves of data collection.
 COVID-19 related media use, however, decreased slightly at the beginning of the study and remained stable after approximately six waves.
@@
 posting---all within-person effects fell completely within the a-priori defined null region (see Figure \@ref(fig:res-activity)).
-For example, respondents who used social media more frequently than usual to read about COVID-19 related topics did not show a simultaneous change in life satisfaction (_b_ = `r dat_fig_results_activity %>% filter(dv == "Life satisfaction" & iv == "Reading" & type == "within") %>% select(estimate) %>% round(2)` [95% CI `r dat_fig_results_activity %>% filter(dv == "Life satisfaction" & iv == "Reading" & type == "within") %>% select(conf.low) %>% round(2)`, `r dat_fig_results_activity %>% filter(dv == "Life satisfaction" & iv == "Reading" & type == "within") %>% select(conf.high) %>% round(2)`]).
+For example, respondents who used social media more frequently than usual to read about COVID-19 related topics did not show a simultaneous change in life satisfaction (_b_ = `r dat_fig_results_activity %>% filter(dv == "Life satisfaction" & iv == "Reading" & type == "within" & Analysis == "1.
+regular") %>% select(estimate) %>% round(2)` [95% CI `r dat_fig_results_activity %>% filter(dv == "Life satisfaction" & iv == "Reading" & type == "within" & Analysis == "1.
+regular") %>% select(conf.low) %>% round(2)`, `r dat_fig_results_activity %>% filter(dv == "Life satisfaction" & iv == "Reading" & type == "within" & Analysis == "1.
+regular") %>% select(conf.high) %>% round(2)`]).
 All confidence intervals included zero; hence, all effects were also statistically non-significant.
@@
 Regarding between-person relations, about which no hypotheses were formulated, only three effects did not include zero.
-Respondents who across all waves used social media more frequently than others to read about COVID-19 related posts reported slightly lower levels of positive affect than others (_b_ `r dat_fig_results_activity %>% filter(dv == "Positive affect" & iv == "Reading" & type == "between") %>% select(estimate) %>% my_round("b_txt")` [95% CI `r dat_fig_results_activity %>% filter(dv == "Positive affect" & iv == "Reading" & type == "between") %>% select(conf.high) %>% my_round("b")`, `r dat_fig_results_activity %>% filter(dv == "Positive affect" & iv == "Reading" & type == "between") %>% select(conf.low) %>% my_round("b")`]).
-Respondents who across all waves used social media more frequently than others to write COVID-19 related posts reported higher levels of negative affect than others (_b_ `r dat_fig_results_activity %>% filter(dv == "Negative affect" & iv == "Posting" & type == "between") %>% select(estimate) %>% my_round("b_txt")` [95% CI `r dat_fig_results_activity %>% filter(dv == "Negative affect" & iv == "Posting" & type == "between") %>% select(conf.high) %>% my_round("b")`, `r dat_fig_results_activity %>% filter(dv == "Negative affect" & iv == "Posting" & type == "between") %>% select(conf.low) %>% my_round("b")`]).
-Interestingly, respondents who across all waves used social media more frequently than others to write COVID-19 related posts also reported slightly higher levels of positive affect than others (_b_ `r dat_fig_results_activity %>% filter(dv == "Positive affect" & iv == "Posting" & type == "between") %>% select(estimate) %>% my_round("b_txt")` [95% CI `r dat_fig_results_activity %>% filter(dv == "Positive affect" & iv == "Posting" & type == "between") %>% select(conf.high) %>% my_round("b")`, `r dat_fig_results_activity %>% filter(dv == "Positive affect" & iv == "Posting" & type == "between") %>% select(conf.low) %>% my_round("b")`]).
+Respondents who across all waves used social media more frequently than others to read about COVID-19 related posts reported slightly lower levels of positive affect than others (_b_ `r dat_fig_results_activity %>% filter(dv == "Positive affect" & iv == "Reading" & type == "between" & Analysis == "1.
+regular") %>% select(estimate) %>% my_round("b_txt")` [95% CI `r dat_fig_results_activity %>% filter(dv == "Positive affect" & iv == "Reading" & type == "between" & Analysis == "1.
+regular") %>% select(conf.high) %>% my_round("b")`, `r dat_fig_results_activity %>% filter(dv == "Positive affect" & iv == "Reading" & type == "between" & Analysis == "1.
+regular") %>% select(conf.low) %>% my_round("b")`]).
+Respondents who across all waves used social media more frequently than others to write COVID-19 related posts reported higher levels of negative affect than others (_b_ `r dat_fig_results_activity %>% filter(dv == "Negative affect" & iv == "Posting" & type == "between" & Analysis == "1.
+regular") %>% select(estimate) %>% my_round("b_txt")` [95% CI `r dat_fig_results_activity %>% filter(dv == "Negative affect" & iv == "Posting" & type == "between" & Analysis == "1.
+regular") %>% select(conf.high) %>% my_round("b")`, `r dat_fig_results_activity %>% filter(dv == "Negative affect" & iv == "Posting" & type == "between" & Analysis == "1.
+regular") %>% select(conf.low) %>% my_round("b")`]).
+Interestingly, respondents who across all waves used social media more frequently than others to write COVID-19 related posts also reported slightly higher levels of positive affect than others (_b_ `r dat_fig_results_activity %>% filter(dv == "Positive affect" & iv == "Posting" & type == "between" & Analysis == "1.
+regular") %>% select(estimate) %>% my_round("b_txt")` [95% CI `r dat_fig_results_activity %>% filter(dv == "Positive affect" & iv == "Posting" & type == "between" & Analysis == "1.
+regular") %>% select(conf.high) %>% my_round("b")`, `r dat_fig_results_activity %>% filter(dv == "Positive affect" & iv == "Posting" & type == "between" & Analysis == "1.
+regular") %>% select(conf.low) %>% my_round("b")`]).
 However, note that the effect were still completely inside of the null region, hence not large enough to be considered practically relevant.
@@
 Changes in the frequency of using different social media channels to attain information regarding COVID-19 were unrelated to meaningful changes in well-being.
-For example, respondents who used Facebook more frequently than usual to learn about COVID-19 did not show a simultaneous change in well-being (_b_ = `r dat_fig_results_channel %>% filter(dv == "Life satisfaction" & iv == "Facebook" & type == "within") %>% select(estimate) %>% round(2)` [95% CI `r dat_fig_results_channel %>% filter(dv == "Life satisfaction" & iv == "Facebook" & type == "within") %>% select(conf.low) %>% round(2)`, `r dat_fig_results_channel %>% filter(dv == "Life satisfaction" & iv == "Facebook" & type == "within") %>% select(conf.high) %>% round(2)`]).
+For example, respondents who used Facebook more frequently than usual to learn about COVID-19 did not show a simultaneous change in well-being (_b_ = `r dat_fig_results_channels %>% filter(dv == "Life satisfaction" & iv == "Facebook" & type == "within" & Analysis == "1.
+regular") %>% select(estimate) %>% round(2)` [95% CI `r dat_fig_results_channels %>% filter(dv == "Life satisfaction" & iv == "Facebook" & type == "within" & Analysis == "1.
+regular") %>% select(conf.low) %>% round(2)`, `r dat_fig_results_channels %>% filter(dv == "Life satisfaction" & iv == "Facebook" & type == "within" & Analysis == "1.
+regular") %>% select(conf.high) %>% round(2)`]).
 Only two effects differed substantially from zero.
-Respondents who used Instagram more frequently than usual to attain COVID-19 related news reported slightly _higher_ levels of life satisfaction than usual (_b_ = `r dat_fig_results_channel %>% filter(dv == "Life satisfaction" & iv == "Instagram" & type == "within") %>% select(estimate) %>% my_round("b")` [95% CI `r dat_fig_results_channel %>% filter(dv == "Life satisfaction" & iv == "Instagram" & type == "within") %>% select(conf.low) %>% my_round("b")`, `r dat_fig_results_channel %>% filter(dv == "Life satisfaction" & iv == "Instagram" & type == "within") %>% select(conf.high) %>% my_round("b")`]).
-Respondents who used Twitter more frequently than usual to attain COVID-19 related news reported slightly _lower_ levels of life satisfaction than usual (_b_ = `r dat_fig_results_channel %>% filter(dv == "Life satisfaction" & iv == "Twitter" & type == "within") %>% select(estimate) %>% round(2)` [95% CI `r dat_fig_results_channel %>% filter(dv == "Life satisfaction" & iv == "Twitter" & type == "within") %>% select(conf.low) %>% round(2)`, `r dat_fig_results_channel %>% filter(dv == "Life satisfaction" & iv == "Twitter" & type == "within") %>% select(conf.high) %>% my_round("b")`]).
+Respondents who used Instagram more frequently than usual to attain COVID-19 related news reported slightly _higher_ levels of life satisfaction than usual (_b_ = `r dat_fig_results_channels %>% filter(dv == "Life satisfaction" & iv == "Instagram" & type == "within" & Analysis == "1.
+regular") %>% select(estimate) %>% my_round("b")` [95% CI `r dat_fig_results_channels %>% filter(dv == "Life satisfaction" & iv == "Instagram" & type == "within" & Analysis == "1.
+regular") %>% select(conf.low) %>% my_round("b")`, `r dat_fig_results_channels %>% filter(dv == "Life satisfaction" & iv == "Instagram" & type == "within" & Analysis == "1.
+regular") %>% select(conf.high) %>% my_round("b")`]).
+Respondents who used Twitter more frequently than usual to attain COVID-19 related news reported slightly _lower_ levels of life satisfaction than usual (_b_ = `r dat_fig_results_channels %>% filter(dv == "Life satisfaction" & iv == "Twitter" & type == "within" & Analysis == "1.
+regular") %>% select(estimate) %>% round(2)` [95% CI `r dat_fig_results_channels %>% filter(dv == "Life satisfaction" & iv == "Twitter" & type == "within" & Analysis == "1.
+regular") %>% select(conf.low) %>% round(2)`, `r dat_fig_results_channels %>% filter(dv == "Life satisfaction" & iv == "Twitter" & type == "within" & Analysis == "1.
+regular") %>% select(conf.high) %>% my_round("b")`]).
 However, both effects were still completely inside of the null region, hence not large enough to be considered meaningful.
@@
 Only one relation did not include zero, was hence statistically significant. 
-Respondents who across all waves used YouTube more frequently than others for COVID-19 related reasons reported marginally _higher_ levels of life satisfaction (_b_ = `r dat_fig_results_channel %>% filter(dv == "Life satisfaction" & iv == "YouTube" & type == "between") %>% select(estimate) %>% round(2)` [95% CI `r dat_fig_results_channel %>% filter(dv == "Life satisfaction" & iv == "YouTube" & type == "between") %>% select(conf.low) %>% my_round("b_txt")`, `r dat_fig_results_channel %>% filter(dv == "Life satisfaction" & iv == "YouTube" & type == "between") %>% select(conf.high) %>% my_round("b")`]).
+Respondents who across all waves used YouTube more frequently than others for COVID-19 related reasons reported marginally _higher_ levels of life satisfaction (_b_ = `r dat_fig_results_channels %>% filter(dv == "Life satisfaction" & iv == "YouTube" & type == "between" & Analysis == "1.
+regular") %>% select(estimate) %>% round(2)` [95% CI `r dat_fig_results_channels %>% filter(dv == "Life satisfaction" & iv == "YouTube" & type == "between" & Analysis == "1.
+regular") %>% select(conf.low) %>% my_round("b_txt")`, `r dat_fig_results_channels %>% filter(dv == "Life satisfaction" & iv == "YouTube" & type == "between" & Analysis == "1.
+regular") %>% select(conf.high) %>% my_round("b")`]).
 However, please note that this effect again was not large enough to be considered practically relevant.
@@
 If anything, two preliminary and subtle trends can be observed.
-First, of all the three COVID-19 related social media activities, people who read about the pandemic more than others showed slightly decreased levels of positive affect, and people who actively posted about the pandemic more than others showed slightly increased levels of negative affect.
-On the other hand, however, people who posted more about COVID-19 also showed slightly higher levels of positive affects, so taken together the results are ambivalent.
-Second, in terms of media channels, using Twitter more than usual was related to slightly decreased levels of life satisfaction.
+First, in terms of media channels, using Twitter more than usual was related to slightly decreased levels of life satisfaction.
 Twitter is considered to have more negative affordances and tonality as compared to other networks such as Instagram [@waterlooNormsOnlineExpressions2018], which might help explain the results.
@@
 To speculate, the often-criticized positivity bias on Instagram might have been somewhat beneficial in times of the pandemic.
-That said, all these effects were still very small and arguably too small to matter. 
+The critique that the positivity bias necessarily leads to envy and negative feelings is unrealistic and myopic, because positive content can also inspire and motivate users [@meierInstagramInspirationHow2020].
+Second, of all the three COVID-19 related social media activities, people who read about the pandemic more than others showed slightly decreased levels of positive affect, and people who actively posted about the pandemic more than others showed slightly increased levels of negative affect.
+On the other hand, however, people who posted more about COVID-19 also showed slightly higher levels of positive affects, so taken together the results are ambivalent.
 But future research might elaborate on these specific relations to probe their stability and relevance.
 
-Finally, another interesting observation is that life satisfaction was remarkably stable.
-Hence, even in times of a pandemic, it seems that such broad assessment of life vary only mildly.
-This supports the hypothesis that life satisfaction seems to be determined largely by stable factors such as one's genes [@brownEasyHappinessPie2019].
+<!-- Finally, another interesting observation is that life satisfaction was remarkably stable. -->
+<!-- Hence, even in times of a pandemic, it seems that such broad assessment of life vary only mildly. -->
+<!-- This supports the hypothesis that life satisfaction seems to be determined largely by stable factors such as one's genes [@brownEasyHappinessPie2019]. -->
 
@@
 Measuring well-being with self-reports is adequate, because it by definition requires introspection.
-However, it would be preferable to measure social media use objectively, because people cannot reliably estimate their use [@scharkowAccuracySelfreportedInternet2016].
+However, it would be preferable to measure social media use objectively, as people cannot reliably estimate their use [@scharkowAccuracySelfreportedInternet2016].
 That said, objective measures often cannot capture the content or the motivation of the use, and only very complicated tools recording the actual content (such as the Screenome project) might produce such data.
 Unfortunately, such procedures introduce other problems, especially related to privacy.
-Hence, for this type of research question it still seems necessary to use self-reported measures.
+Hence, for this type of research question it still seems necessary to use self-reported measures, and in many cases they can still be very informative [@verbeijSelfreportedMeasuresSocial2021].
 
@@
 
-```{r descriptives, cache=F}
+```{r tab-descriptives, cache=F}
 apa_table(tab_desc_dat,
@@
 
-```{r fig-desc, fig.cap="Development of well-being and media use measures across the pandemic.
-Values obtained from mixed effect models, with participants and waves as grouping factors and without additional predictors.", out.width = "\\textwidth", cache=F}
-knitr::include_graphics("figures/fig_descriptives.pdf")
+```{r sesoi, warning=F, message=F, fig.cap="Illustration of how confidence intervals are used to test a null region.
+Here, a trivial effect of social media use on life satisfaction is defined as ranging from b = -.25 to b = .25", fig.width=6, fig.height=2}
+# make table with data
+d_plot <- tribble(
+  ~name, ~type, ~Effect, ~ll, ~ul,
+  "1. Accept trivial effect", "Decision Rule", -.02, -.12, .08, 
+  "2. Reject trivial effect", "Decision Rule", -.37, -.47, -.27,
+  "3. Reject positive effect", "Decision Rule", -.22, -.27, -.17,
+  "4. Suspend judgement", "Decision Rule", -.1, -.30, .10,
+) %>% 
+  mutate(
+    name = factor(name, levels = name),
+    name = fct_rev(name)
+         )
+
+# design plot
+plot <- ggplot(d_plot, aes(y = name, x = Effect)) +
+  geom_vline(xintercept = 0, color = "darkgrey") + 
+  geom_vline(xintercept = -.25, color = "darkgrey", linetype = "dashed") +
+  geom_vline(xintercept = .25, color = "darkgrey", linetype = "dashed") +
+  geom_point(aes(Effect)) +
+  geom_errorbarh(aes(xmin = ll, xmax = ul), height = .5) +
+  theme(axis.title.y = element_blank()) +
+  labs(x = "Effect Size (b)",
+       caption = "Smallest effect size of interest: b = |.25|
+                  Null region: b = -.25, .25")
+
+# ggsave("figures/figure_intervals.png", height = 3.5)
+plot
+```
+
+\newpage
+
+```{r descriptives, fig.cap="Development of well-being and media use measures across the pandemic.
+Values obtained from mixed effect models, with participants and waves as grouping factors and without additional predictors.", out.width = "\\textwidth", cache=F, warning=F, message=F, prompt=FALSE}
+# print(fig_desc)
+knitr::include_graphics("figures/fig_descriptives.png")
 ```
@@
 The SESOI was b = |0.30| for life satisfaction and b = |0.15| for affect.
-Hence, all of the reported effects are not considered meaningful.", out.width = "\\textwidth", cache=F}
-knitr::include_graphics("figures/fig_results_activity.pdf")
+Hence, all of the reported effects are not considered meaningful.", out.width = "\\textwidth", cache=F, warning=F, message=F, fig.width=6, fig.height=3}
+fig_results_activity
+# knitr::include_graphics("figures/fig_results_activity.png")
 ```
@@
 The SESOI was b = |0.30| for life satisfaction and b = |0.15| for affect.
-Hence, all of the reported effects are not considered meaningful.", out.width = "\\textwidth", cache=F}
-knitr::include_graphics("figures/fig_results_channel.pdf")
+Hence, all of the reported effects are not considered meaningful.", out.width = "\\textwidth", cache=F, warning=F, message=F, fig.width=6, fig.height=4}
+# knitr::include_graphics("figures/fig_results_channels.png")
+fig_results_channels
 ```
@@
 ```{r res-control, fig.cap="Results of selected covariates.
-All variables were standardize except 'Male'  and ‘Employed in public service', because there were measured on a binary scale.", out.width = "\\textwidth", cache=F}
-knitr::include_graphics("figures/fig_results_control_std.pdf")
+All variables were standardize except 'Male'  and ‘Employed in public service', because there were measured on a binary scale.", out.width = "\\textwidth", cache=F, warning=F, message=F, fig.width=6, fig.height=3}
+fig_results_control_std
+# knitr::include_graphics("figures/fig_results_control_std.png")
 ```

### 2022-04-12 manuscript.tex
- implented reviewer feedback

@@
   english,
-  man,mask,floatsintext]{apa6}
+  man,mask]{apa6}
+\title{Analyzing the effects of COVID-19 related social media use on well-being}
+\author{Name blinded for review\textsuperscript{1}}
+\date{}
+
 \usepackage{amsmath,amssymb}
 \usepackage{lmodern}
-\usepackage{ifxetex,ifluatex}
-\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
+\usepackage{iftex}
+\ifPDFTeX
   \usepackage[T1]{fontenc}
@@
   pdftitle={Analyzing the effects of COVID-19 related social media use on well-being},
+  pdfauthor={Name blinded for review1},
   pdflang={en-EN},
-  pdfkeywords={COVID-19, Coronavirus, well-being, affect, life satisfaction, social media use, news use, communication, random effects within between model, panel study, longitudinal study.},
+  pdfkeywords={COVID-19, well-being, social media, news use, panel study.},
   hidelinks,
@@
 \fi
+\newlength{\cslhangindent}
+\setlength{\cslhangindent}{1.5em}
+\newlength{\csllabelwidth}
+\setlength{\csllabelwidth}{3em}
+\newlength{\cslentryspacingunit} % times entry-spacing
+\setlength{\cslentryspacingunit}{\parskip}
+\newenvironment{CSLReferences}[2] % #1 hanging-ident, #2 entry spacing
+ {% don't indent paragraphs
+  \setlength{\parindent}{0pt}
+  % turn on hanging indent if param 1 is 1
+  \ifodd #1
+  \let\oldpar\par
+  \def\par{\hangindent=\cslhangindent\oldpar}
+  \fi
+  % set entry spacing
+  \setlength{\parskip}{#2\cslentryspacingunit}
+ }%
+ {}
+\usepackage{calc}
+\newcommand{\CSLBlock}[1]{#1\hfill\break}
+\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\csllabelwidth}{#1}}
+\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth - \csllabelwidth}{#1}\break}
+\newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}
 % Manuscript styling
@@
 % \newenvironment{ltable}
-%   {\begin{landscape}\begin{center}\begin{threeparttable}}
-%   {\end{threeparttable}\end{center}\end{landscape}}
-\newenvironment{lltable}{\begin{landscape}\begin{center}\begin{ThreePartTable}}{\end{ThreePartTable}\end{center}\end{landscape}}
+%   {\begin{landscape}\centering\begin{threeparttable}}
+%   {\end{threeparttable}\end{landscape}}
+\newenvironment{lltable}{\begin{landscape}\centering\begin{ThreePartTable}}{\end{ThreePartTable}\end{landscape}}
 
@@
 % \setlength{\parskip}{0pt plus 0pt minus 0pt}
-
-% Overwrite redefinition of paragraph and subparagraph by the default LaTeX template
-% See https://github.com/crsh/papaja/issues/292
-\makeatletter
-\renewcommand{\paragraph}{\@startsection{paragraph}{4}{\parindent}%
-  {0\baselineskip \@plus 0.2ex \@minus 0.2ex}%
-  {-1em}%
-  {\normalfont\normalsize\bfseries\itshape\typesectitle}}
-
-\renewcommand{\subparagraph}[1]{\@startsection{subparagraph}{5}{1em}%
-  {0\baselineskip \@plus 0.2ex \@minus 0.2ex}%
-  {-\z@\relax}%
-  {\normalfont\normalsize\itshape\hspace{\parindent}{#1}\textit{\addperi}}{\relax}}
-\makeatother
 
@@
 \makeatother
-\keywords{COVID-19, Coronavirus, well-being, affect, life satisfaction, social media use, news use, communication, random effects within between model, panel study, longitudinal study.}
+\shorttitle{COVID-19 related social media use and well-being}
+\keywords{COVID-19, well-being, social media, news use, panel study.}
+\DeclareDelayedFloatFlavor{ThreePartTable}{table}
+\DeclareDelayedFloatFlavor{lltable}{table}
+\DeclareDelayedFloatFlavor*{longtable}{table}
+\makeatletter
+\renewcommand{\efloat@iwrite}[1]{\immediate\expandafter\protected@write\csname efloat@post#1\endcsname{}}
+\makeatother
 \usepackage{lineno}
@@
 \usepackage{csquotes}
+\usepackage{endnotes}
 \setlength{\parskip}{0em}
@@
 \note{\clearpage}
-\ifxetex
+\ifXeTeX
   % Load polyglossia as late as possible: uses bidi with RTL langages (e.g. Hebrew, Arabic)
@@
 \fi
-\ifluatex
+\ifLuaTeX
   \usepackage{selnolig}  % disable illegal ligatures
 \fi
-\newlength{\cslhangindent}
-\setlength{\cslhangindent}{1.5em}
-\newlength{\csllabelwidth}
-\setlength{\csllabelwidth}{3em}
-\newenvironment{CSLReferences}[2] % #1 hanging-ident, #2 entry spacing
- {% don't indent paragraphs
-  \setlength{\parindent}{0pt}
-  % turn on hanging indent if param 1 is 1
-  \ifodd #1 \everypar{\setlength{\hangindent}{\cslhangindent}}\ignorespaces\fi
-  % set entry spacing
-  \ifnum #2 > 0
-  \setlength{\parskip}{#2\baselineskip}
-  \fi
- }%
- {}
-\usepackage{calc}
-\newcommand{\CSLBlock}[1]{#1\hfill\break}
-\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\csllabelwidth}{#1}}
-\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth - \csllabelwidth}{#1}\break}
-\newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}
-
-\title{Analyzing the effects of COVID-19 related social media use on well-being}
-\author{Name blinded for review\textsuperscript{1}}
-\date{}
-
-\shorttitle{COVID-19 related social media use and well-being}
-
-\authornote{
-
-Correspondence concerning this article should be addressed to Name blinded for review, . E-mail:
-
-}
 
@@
 \abstract{
-In times of crisis such as the Corona pandemic citizens need to stay informed about recent events, the latest political decisions, or mandatory protection measures.
+In times of crisis such as the Corona pandemic citizens need to stay informed about recent events, political decisions, or mandatory protection measures.
 To this end, many people use various types of media, and increasingly social media.
-However, because social media are particularly engaging, some find it hard to disconnect and cannot stop `doomscrolling.' In this preregistered study, I investigate whether using social media for COVID-19 related reasons affects psychological well-being.
-To answer this question I analyzed data from the Austrian Corona Panel Project, which consists of 24 waves with overall 3,018 participants.
+However, because social media are particularly engaging, some find it hard to disconnect.
+In this preregistered study, I investigate whether using social media for COVID-19 related reasons affects psychological well-being.
+To answer this question I analyzed data from the Austrian Corona Panel Project, which consists of 3,018 participants.
+Well-being was measured at each wave, and communication at five waves.
 I ran three random effects within between models, controlling for several stable and varying confounders.
 Results showed that the effects of COVID-19 related social media use on well-being were very small, arguably too small to matter.
-The findings suggest that fears that social media use during times of crisis impairs well-being are likely to be unfounded.
+Fears that social media use during times of crisis critically impairs well-being are likely to be unfounded.
 }
@@
 Users were glued to their screens and found it hard to pursue other relevant activities such as working, taking a break, or looking after their children (Klein, 2021).
-As doomscrolling increased it became doubtful whether such social media is helpful, or whether it created an additional burden on mental health (Sandstrom, Buchanan, Aknin, \& Lotun, 2021).
+It was asked whether using social media for COVID-19 related reasons is helpful, or whether it creates an additional burden on mental health (Sandstrom, Buchanan, Aknin, \& Lotun, 2021).
 These concerns seem justified:
@@
 
-First, I investigate how different types of communication affect well-being.
-Specifically, I differentiate between active and passive use.
-I distinguish (a) \emph{reading} (passive), (b) \emph{posting} (active), and (c) \emph{liking and sharing} COVID-19 related posts (both active and passive).
+First, I investigate how well-being is affected by different types of communication affect, namely active and passive use.
+Defining what constitutes active and what passive use is not always clear, and different understandings are currently discussed Meier \& Reinecke (2020).
+Reading is generally considered as passive and writing as active, while there are also specific behaviors such as such liking or sharing content that fall somewhere in-between (Meier \& Krause, 2022).
+In this study, I hence distinguish (a) \emph{reading} (passive), (b) \emph{posting} (active), and (c) \emph{liking and sharing} COVID-19 related posts (both active and passive).
 Second, I analyze how using the most prominent branded applications affects well-being, and whether this effect changes across applications.
@@
 
-\hypertarget{effects-of-social-media-on-well-being}{%
-\subsection{Effects of Social Media on Well-Being}\label{effects-of-social-media-on-well-being}}
-
-So far, there is only little empirical research on COVID-19 related social media use on well-being.
+\hypertarget{theorizing-social-media-effects-on-well-being}{%
+\subsection{Theorizing Social Media Effects on Well-Being}\label{theorizing-social-media-effects-on-well-being}}
+
+From a theoretical perspective, how could we explain whether COVID-19 related social media use might affect well-being?
+According to the set-point model of subjective well-being, well-being is surprisingly stable (Lykken, 1999).
+Although specific events such as marriage or salary can have significant effects, after some time well-being routinely returns to prior levels, which are mostly determined genetically (Sheldon \& Lucas, 2014).
+Only very specific events such as unemployment, disability, or death can cause long-term decreases in well-being (Lucas, 2007).
+So although well-being can change this does not happen easily.
+
+Can media use be such a negative or positive factor?
+In advance, there doesn't seem to be a clear winner, and it's likely that both positive and negative effects cancel each other out.
+Empirically, social media use---on average---does not have a strong effect on well-being (Meier \& Reinecke, 2020).
+According to the different susceptibility of media effects model (Valkenburg \& Peter, 2013), the effects of media use differ across individuals.
+Whereas for some media are beneficial, for others they are harmful.
+On average, however, effects are often small or negligible.
+Also when focusing on individual users, social media have both positive and negative effects on well-being (Büchi, 2021).
+They can impair well-being when causing embarrassment, stress, or disinformation, and they can improve well-being when providing connectedness, information, or entertainment (Büchi, 2021).
+
+Two of the most prominent media effect theories argue against strong negative impacts.
+First, uses and gratifications theory states that people explicitly and rationally chose specific media because of their respective benefits.
+If those benefits don't exist, they invest their time elsewhere.
+And social media offer ample benefits.
+Most prominently, they help find relevant information, maintain and foster relationships, express one's personality, and entertain oneself (Pelletier, Krallman, Adams, \& Hancock, 2020).
+The second major theory is mood management theory (Zillmann, 1988).
+Users implicitly learn what media help them balance their mood and affect.
+For example, when bored people use social media to entertain themselves.
+Using experience sampling of well-being and logs of social media use, a study with 82 participants from Italy found that after episodes of social media use, levels of positive affect increased for a short time (Marciano, Driver, Schulz, \& Camerini, 2022).
+
+Precisely because social media have so many positive consequences, one can ask if this is the main problem.
+In other words, social media aren't problematic because they're inherently bad, but rather because they're too good.
+And as with many other things, there can be too much of a good thing.
+It is therefore often asked whether social media are addictive, and users sometimes express this fear themselves (Yang, Griffiths, Yan, \& Xu, 2021).
+However, a recently published meta-analysis found that the two most prominent measures of addiction, the Bergen Facebook Addiction Scale and the Bergen Social Media Addiction Scale, have only small relations to well-being (Duradoni, Innocenti, \& Guazzini, 2020).
+In addition, the general idea of labeling excessive social and new media use as addiction was criticized, arguing that social and new media represent new regular behaviors that should not be pathologized (Galer, 2018; van Rooij et al., 2018).
+
+Because effects can differ across situations, I now briefly focus on the effects of COVID-19 related social media use specifically.
+First, one could assume a \emph{direct} negative effect on well-being, and especially on positive or negative affect, which are more volatile and fluctuating.
+Dangers, inequalities, corruption---these were the headlines during the pandemic across many countries worldwide.
+If one learns about such events, the initial reaction might be shock, fear, or dismay.
+Consuming such news can be depressing (Dörnemann, Boenisch, Schommer, Winkelhorst, \& Wingen, 2021), perhaps even changing some general perspectives on life.
+That said, because not all news was negative, and because many people showed solidarity and compassion, there was also positive and uplifting content, potentially compensating for the negative effects (Dörnemann et al., 2021).
+A study with 2.057 respondents from Italy reported that during the pandemic virtual community and social connectedness even increased during the pandemic (Guazzini, Pesce, Marotta, \& Duradoni, 2022).
+
+There could also be \emph{indirect} effects.
+When browsing social media for Covid-19 related news, many users reported being captivated to such an extent that they could not stop using social media (Klein, 2021).
+During the pandemic social media use was at an all-time high in the US (Statista, 2021).
+It is most likely that moderate social media use is not detrimental (Orben, 2020).
+Overuse, however, might be more critical, and several studies have shown more pronounced negative effects for extreme users (Przybylski \& Weinstein, 2017).
+To explain, overuse could impair well-being if it replaces meaningful or functional activities such as meeting others, working, actively relaxing, or exercising.
+So if a society collectively overuses social media during a pandemic, there might be potential for negative effects.
+On the other hand, one can make the case that overuse might also be beneficial, especially in times of a pandemic---even if the use is mainly COVID-19 related.
+Exchanging COVID-19 related messages with friends via WhatsApp might replace the in-person contact one would have otherwise, but which is literally impossible at the time.
+In situations where meaningful and functional activities are prohibited, using social media to exchange about COVID-19 related topics might not be the worst idea.
+Besides, given that nowadays a large number of experts, scientists, and politicians converse directly on social media, one can get first-hand high quality information on current developments.
+
+To summarize, from a theoretical perspective it is most likely that the average effects of social media use on well-being are negligible.
+Building on established theories from Communication, it is not particularly likely that effects are either profoundly negative or strongly positive.
+
+\hypertarget{empirical-studies-on-social-media-effects}{%
+\subsection{Empirical Studies on Social Media Effects}\label{empirical-studies-on-social-media-effects}}
+
+So far, there is only little empirical research on how well-being is affected by COVID-19 related social media.
 In their study on the relations between media use and mental health during the pandemic, Bendau et al.
@@
 The results showed that digital technology did not have any significant effects on mental health (for a similar study with comparable results, see Bradley \& Howard, 2021).
-Together, the literature is mixed, with a slight focus on the negative effects of social media as news use (see also Dörnemann, Boenisch, Schommer, Winkelhorst, \& Wingen, 2021; Liu \& Tong, 2020; Riehm et al., 2020).
+Together, the literature is mixed, with a slight focus on the negative effects of social media as news use (see also Dörnemann et al., 2021; Liu \& Tong, 2020; Riehm et al., 2020).
 However, note that all of these findings represent between-person relations stemming from cross-sectional data (see below).
@@
 
-These overviews are well aligned with several individual studies employing advanced methods (Keresteš \& Štulhofer, 2020; Orben, Dienlin, \& Przybylski, 2019; Przybylski, Nguyen, Law, \& Weinstein, 2021; Schemer, Masur, Geiß, Müller, \& Schäfer, 2021).
-For example, Beyens, Pouwels, Driel, Keijsers, and Valkenburg (2021) reported that although for some users (roughly one quarter) the effects of social media use on well-being were negative, for almost the same number of users they were positive, while for the rest the effects were neutral.
-In conclusion, most effects are likely somewhere between trivial and small.
-I therefore expect that also in the case of COVID-19 related social media use effects will be trivial to small.
-
-From a theoretical perspective, how could we explain whether COVID-19 related social media use might affect well-being?
-In what follows, I outline potential arguments as to why the effect might be positive or negative, direct or indirect, or nonexistent.
-In advance, there doesn't seem to be a clear winner, and it's likely that both positive and negative effects cancel each other out.
-
-First, one could assume a \emph{direct} negative effect on well-being, and especially on positive or negative affect, which are more volatile and fluctuating.
-Dangers, inequalities, corruption---these were the headlines during the pandemic across many countries worldwide.
-If one learns about such events, the initial reaction might be shock, fear, or dismay.
-Consuming such news can be depressing (Dörnemann, Boenisch, Schommer, Winkelhorst, \& Wingen, 2021), perhaps even changing some general perspectives on life.
-That said, because not all news was negative, and because many people showed solidarity and compassion, there was also positive and uplifting content, potentially compensating for the negative effects (Dörnemann, Boenisch, Schommer, Winkelhorst, \& Wingen, 2021).
-
-There could also be \emph{indirect} effects.
-When doomscrolling, users are captivated to such an extent that they cannot stop using social media.
-For example, during the pandemic social media use was at an all-time high in the US (Statista, 2021).
-In general, as has been expressed by many before, it is most likely that moderate social media use is not detrimental (Orben, 2020).
-Overuse, however, might be more critical, and several studies have shown more pronounced negative effects for extreme users (Przybylski \& Weinstein, 2017).
-To explain, overuse likely impairs well-being if it replaces meaningful or functional activities such as meeting others, working, actively relaxing, or exercising.
-So if a society collectively overuses social media when doomscrolling, there is potential for negative effects.
-
-On the other hand, one can make the case that overuse might also be beneficial, especially in times of a pandemic---even if the use is mainly COVID-19 related.
-Exchanging COVID-19 related messages with friends via WhatsApp might replace the in-person contact one would have otherwise, but which is literally impossible at the time.
-In situations where meaningful and functional activities are prohibited, using social media to exchange about COVID-19 related topics might not be the worst idea.
-Besides, given that nowadays a large number of experts, scientists, and politicians converse directly on social media, one can get first-hand high quality information on current developments.
-
-Together, the strongest argument to me is that \emph{in general} the effects of social media on well-being are, on average, small at best.
-Because this study only looks at \emph{one part} of social media use---namely, COVID-19 related interactions---it is very focused, diminishing the overall potential of the effects even further.
-Whether or not using social media for COVID-19 related aspects is detrimental during a pandemic is also not entirely clear.
-Therefore, I expect that COVID-19 related communication on social media doesn't affect well-being in a meaningful or relevant way.
+Several individual studies employing advanced methods found smalls relations between social media use and well-being (Keresteš \& Štulhofer, 2020; Orben, Dienlin, \& Przybylski, 2019; Przybylski, Nguyen, Law, \& Weinstein, 2021; Schemer, Masur, Geiß, Müller, \& Schäfer, 2021).
+For example, Beyens, Pouwels, van Driel, Keijsers, and Valkenburg (2021) reported that although for some users (roughly one quarter) the effects of social media use on well-being were negative, for almost the same number of users they were positive, while for the rest the effects were neutral.
+This finding is aligned the Differential Susceptibility to Media Effects Model.
+Although there is substantial variation of media effects for individual users, the average effects reported in the literature are often small (Valkenburg \& Peter, 2013).
+
+In conclusion, in light of the theoretical considerations and the empirical studies presented above, I expect that COVID-19 related communication on social media doesn't affect well-being in a meaningful or relevant way.
 
@@
 
-The sample was representative of the Austrian population in terms of age, gender, region/state, municipality size, and educational level.
+Achieved via quota sampling, the sample matched the Austrian population in terms of age, gender, region/state, municipality size, and educational level.
 In order to participate in the study, the respondents needed to be Austrian residents and had to be at least 14 years of age.
-The study received IRB approval from the University of Vienna.
+Ethical review and approval was not required for the study in accordance with the local legislation and institutional requirements.
+The participants provided their written informed consent to participate in this study.
 The average age was 42 years, 49 percent were male, 14 percent had a University degree, and 5 percent were currently unemployed.
@@
 Because we're agnostic as to whether the effects are positive or negative, the null region includes negative and positive effects.
-Finally, in order not to exaggerate precision and to be less conservative, these numbers are reduced to nearby thresholds.\footnote{Note that other researchers also decreased or recommended decreasing thresholds for effect sizes when analyzing within-person or cumulative effects (Beyens, Pouwels, Driel, Keijsers, \& Valkenburg, 2021; Funder \& Ozer, 2019).}
+Finally, in order not to exaggerate precision and to be less conservative, these numbers are reduced to nearby thresholds.\footnote{Note that other researchers also decreased or recommended decreasing thresholds for effect sizes when analyzing within-person or cumulative effects (Beyens et al., 2021; Funder \& Ozer, 2019).}
 Together, this leads to a null region ranging from \emph{b} = -.30 to \emph{b} = .30 for life satisfaction, and \emph{b} = -.15 to \emph{b} = .15 for positive and negative affect.
@@
 If the confidence interval falls completely outside of the null-region (e.g., \emph{b} = .40, {[}95\% CI: .35, .45{]}), the hypothesis is rejected and the existence of a meaningful positive effect is supported.
+For an illustration, see Figure \ref{fig:sesoi}).
 
@@
 In what follows, I list all the variables that I analyzed.
-For the variables' means, range, and variance, see Table \ref{tab:descriptives}.
+For the variables' means, range, and variance, see Table \ref{tab:tab-descriptives}.
 For a complete list of all items and item characteristics, see \href{https://XMtRA.github.io/Austrian_Corona_Panel_Project}{companion website}.
@@
 
-Life satisfaction was measured with the item ``Taken everything together, how satisfied are you currently with your life?''
-The response options ranged from 0 (\emph{extremely unsatisfied}) to 10 (\emph{extremely satisfied}).
-
-To capture positive affect, respondents were asked how often in the last week they felt (a) calm and relaxed, (b) happy, and (c) full of energy.
+Life satisfaction was measured with the item ``All things considered, how satisfied are you with your life as a whole nowadays?'' from the European Social Survey (European Social Survey, 2021).
+The response options ranged from 0 (\emph{extremely dissatisfied}) to 10 (\emph{extremely satisfied}).
+
+To capture positive affect, respondents were asked how often in the last week they felt (a) calm and relaxed, (b) happy, and (c) full of energy (World Health Organization, 1998).
 The response options were 1 (\emph{never}), 2 (\emph{on some days}), 3 (\emph{several times per week}), 4 (\emph{almost every day}), and 5 (\emph{daily}).
@@
 
-For negative affect, respondents were asked how often in the last week they felt (a) lonely, (b) aggravated, (c) so depressed, that nothing could lift you up, (d) very nervous, (e) anxious, and (h) glum and sad.
+For negative affect, respondents were asked how often in the last week they felt (a) lonely, (b) aggravated, (c) so depressed, that nothing could lift you up, (d) very nervous, (e) anxious, and (h) glum and sad (World Health Organization, 1998).
 The response options were 1 (\emph{never}), 2 (\emph{on some days}), 3 (\emph{several times per week}), 4 (\emph{almost every day}), and 5 (\emph{daily}).
@@
 
-COVID-19 related social media use focused on communication types was measured with the three dimensions of (a) reading, (b) liking and sharing, and (c) posting.
+COVID-19 related social media use focused on communication types was measured with the three dimensions of (a) reading, (b) liking and sharing, and (c) posting from Wagner et al.
+(2018), adapted for the context of this study.
 The general introductory question was ``How often during the last week have you engaged in the following activities on social media?''
@@
 
-COVID-19 related social media use focused on channels was measured with five variables.
+COVID-19 related social media use focused on channels was measured with five variables from Wagner et al. (2018), adapted for this study.
 The general introductory question was ``How often in the last week have you followed information related to the Corona-crisis on the following social media?''
@@
 
+\hypertarget{control-variables}{%
+\subsubsection{Control variables}\label{control-variables}}
+
+The effects of COVID-19 related social media use were controlled for the following stable variables:
+(a) gender (female, male, diverse), (b) age, (c) education (ten options), (d) Austria country of birth (yes/no), (e) Austria parents' country of birth (no parent, one parent, both parents).
+I originally planned to implement additional variables as varying covariates.
+However, because they were not measured often enough or not at the time when social media use was measured, I implemented them as stable variables using their average values across all waves.
+This includes (a) text-based media news consumption (five degrees), (b) video-based media news consumption (five degrees), (c) residency is Vienna (yes/no), (d) self-reported physical health (five degrees), (e) living space (eleven options), (f) access to balcony (yes/no), (g) access to garden (yes/no), (h) employment (nine options), (i) disposition to take risks (eleven degrees), and (j) locus of control (five degrees).
+I also controlled for the following varying covariates: (a) five items measuring outdoor activities such as sport or meeting friends (five degrees), and (b) satisfaction with democracy (five degrees).
+Because it lead to too much attrition in the sample, I did not control for (a) household size, (b) work hours per week, (c) home office, (d) household income.
+
+\hypertarget{results}{%
+\section{Results}\label{results}}
+
+First, when looking at the variables from a descriptive perspective (Figure \ref{fig:descriptives})), we see that all well-being measures did not change substantially across the different waves of data collection.
+COVID-19 related media use, however, decreased slightly at the beginning of the study and remained stable after approximately six waves.
+The initial decrease might be explained by the fact that the collection of data began at the end of March 2020, hence approximately three months after the pandemic began.
+It could be that after an initial uptick, COVID-19 related social media use was already declining at the time, returning to more normal levels.
+
+\hypertarget{preregistered-analyses}{%
+\subsection{Preregistered Analyses}\label{preregistered-analyses}}
+
+The study's main hypothesis was that the effects of social media use on well-being would be trivial.
+Regarding the effects of different communication \emph{types}---that is, reading vs.~sharing vs.~posting---all within-person effects fell completely within the a-priori defined null region (see Figure \ref{fig:res-activity}).
+For example, respondents who used social media more frequently than usual to read about COVID-19 related topics did not show a simultaneous change in life satisfaction (\emph{b} = 0.05 {[}95\% CI -0.01, 0.1{]}).
+All confidence intervals included zero; hence, all effects were also statistically non-significant.
+As a result, the hypothesis was supported for all COVID-19 related types of social media communication.
+
+Regarding between-person relations, about which no hypotheses were formulated, only three effects did not include zero.
+Respondents who across all waves used social media more frequently than others to read about COVID-19 related posts reported slightly lower levels of positive affect than others (\emph{b} = -0.03 {[}95\% CI \textgreater{} -0.01, -0.06{]}).
+Respondents who across all waves used social media more frequently than others to write COVID-19 related posts reported higher levels of negative affect than others (\emph{b} = 0.06 {[}95\% CI 0.09, 0.03{]}).
+Interestingly, respondents who across all waves used social media more frequently than others to write COVID-19 related posts also reported slightly higher levels of positive affect than others (\emph{b} = 0.05 {[}95\% CI 0.09, \textless{} 0.01{]}).
+However, note that the effect were still completely inside of the null region, hence not large enough to be considered practically relevant.
+
+Note that when comparing the results with and without control variables, the results differed.
+For example, on the between-person level, one effect stopped being significant if controlled for additional variables.
+Actively posting on social media was significantly (though not meaningfully) related to decreased life satisfaction.
+However, when controlling for potential confounders, the effect became virtually zero (see Figure \ref{fig:res-channels}).
+
+Regarding the COVID-19 related use of social media \emph{channels}, the results were comparable (see Figure \ref{fig:res-channels}).
+Changes in the frequency of using different social media channels to attain information regarding COVID-19 were unrelated to meaningful changes in well-being.
+For example, respondents who used Facebook more frequently than usual to learn about COVID-19 did not show a simultaneous change in well-being (\emph{b} = -0.05 {[}95\% CI -0.11, 0.01{]}).
+Only two effects differed substantially from zero.
+Respondents who used Instagram more frequently than usual to attain COVID-19 related news reported slightly \emph{higher} levels of life satisfaction than usual (\emph{b} = 0.09 {[}95\% CI 0.02, 0.16{]}).
+Respondents who used Twitter more frequently than usual to attain COVID-19 related news reported slightly \emph{lower} levels of life satisfaction than usual (\emph{b} = -0.12 {[}95\% CI -0.23, -0.02{]}).
+However, both effects were still completely inside of the null region, hence not large enough to be considered meaningful.
+In sum, the hypothesis was supported also for the COVID-19 related use of important social media channels.
+
+In terms of between-person relations---which, again, weren't included in the hypotheses---no relations crossed the null region or fell outside of it.
+Only one relation did not include zero, was hence statistically significant.
+Respondents who across all waves used YouTube more frequently than others for COVID-19 related reasons reported marginally \emph{higher} levels of life satisfaction (\emph{b} = 0.08 {[}95\% CI \textless{} 0.01, 0.16{]}).
+However, please note that this effect again was not large enough to be considered practically relevant.
+
+Again, note that when comparing the results with and without control variables, the results differed.
+Especially on the between-person level, altogether five effects stopped being significant if they were controlled for additional variables.
+For example, using Instagram was significantly (though not meaningfully) related to increased life satisfaction.
+However, when controlling for additional covariates, the effect became virtually zero (see Figure \ref{fig:res-channels}).
+
+\hypertarget{exploratory-analyses}{%
+\subsection{Exploratory Analyses}\label{exploratory-analyses}}
+
+In what follows, I briefly report some exploratory analyses that weren't preregistered.
+First, to contextualize the results reported above and to see if the results included any meaningful effects at all, I also looked at the effect sizes of selected (cherry-picked) covariates.
+Because each variable had different response options, we would actually need to define a SESOI for each variable, which for reasons of scope I cannot implement here.
+Therefore, I report the results of the standardized scales,
+which allows for a better comparison across the differently scaled variables.
+For what it's worth, as a rough estimate for the SESOI we can build on the typical convention that small effects start at \emph{r} = \textbar.10\textbar.
+The results showed that several effects fell outside of the SESOI, were hence considered meaningful.
+This includes for example internal locus of control, health, satisfaction with democracy, or exercising.
+For an overview, see Figure \ref{fig:res-control}.
+
+To find out whether my inferences were robust across legitimate (though arguably inferior) alternative analyses, I reran the analyses also using standardized estimates, mean scores instead of factor scores, and with a data set where missing data were not imputed.
+The results were virtually the same.
+For example, all standardized COVID-19 related types of social media use or channels were not significantly larger than a SESOI of \(\beta\) = \textbar.10\textbar.
+The additional analyses are reported in the \href{https://XMtRA.github.io/Austrian_Corona_Panel_Project/analyses_additional.html}{companion website}.
+
+\hypertarget{discussion}{%
+\section{Discussion}\label{discussion}}
+
+In this study I analyzed the effects of COVID-19 related social media use on well-being.
+The data come from a panel study with 24 waves and are largely representative of the Austrian population.
+In a random effects model I separated between person relations from within-person effects and controlled for a large number of both stable and varying covariates, thereby aiming to assess causality.
+The results showed that within-person effects were trivial.
+People who used social media more than usual to learn about COVID-19 didn't show meaningful changes in their well-being.
+
+The results imply that COVID-19 related social media use doesn't seem to be particularly relevant for well-being.
+Other factors among the third variables that were measured revealed larger effects or relations, suggesting that well-being is determined by alternative aspects such as health, satisfaction with democracy, locus of control, or exercising.
+According to this study, popular fears that ``doomscrolling'' or overusing social media during times of crises don't seem to be justified.
+
+On the one hand, the results are not aligned with several recent studies analyzing similar or closely related research questions.
+This includes a study by Bendau et al. (2021), which showed negative relations between social media and well-being (but see Bradley \& Howard, 2021; or Sewall et al., 2021).
+However, note that Bendau et al. (2021) analyzed cross-sectional data on a between-person level while not controlling for third variables, which is not optimal for investigating causal effects.
+On the other hand, the results are well-aligned with recent studies and meta-analyses analyzing the effects of social media use from a more general perspective or from a somewhat different angle.
+These studies have found that the effects of various types of social media use on several well-being indicators are small at best, often too small to matter (Ferguson et al., 2021; Meier \& Reinecke, 2020; Orben, 2020), which echoes the results obtained here.
+
+If anything, two preliminary and subtle trends can be observed.
+First, of all the three COVID-19 related social media activities, people who read about the pandemic more than others showed slightly decreased levels of positive affect, and people who actively posted about the pandemic more than others showed slightly increased levels of negative affect.
+On the other hand, however, people who posted more about COVID-19 also showed slightly higher levels of positive affects, so taken together the results are ambivalent.
+Second, in terms of media channels, using Twitter more than usual was related to slightly decreased levels of life satisfaction.
+Twitter is considered to have more negative affordances and tonality as compared to other networks such as Instagram (Waterloo et al., 2018), which might help explain the results.
+Instagram, on the other hand, was related to slightly increased levels of life satisfaction.
+To speculate, the often-criticized positivity bias on Instagram might have been somewhat beneficial in times of the pandemic.
+That said, all these effects were still very small and arguably too small to matter.
+But future research might elaborate on these specific relations to probe their stability and relevance.
+
+Finally, another interesting observation is that life satisfaction was remarkably stable.
+Hence, even in times of a pandemic, it seems that such broad assessment of life vary only mildly.
+This supports the hypothesis that life satisfaction seems to be determined largely by stable factors such as one's genes (Brown \& Rohrer, 2019).
+
+\hypertarget{limitations}{%
+\subsection{Limitations}\label{limitations}}
+
+The current study analyzed whether changes in media use were related to changes in well-being, while controlling for several potential confounders.
+Together, this allows for an improved perspective on assessing causality.
+That said, causality necessitates temporal order, and the cause needs to precede the effect.
+Regarding media use, such effects often happen immediately or shortly after use, necessitating intervals in the hours, minutes, or even seconds.
+In many cases only experience sampling studies asking users in the very moment can produce such knowledge.
+However, even then we don't know for certain if we actually measured the right interval.
+Effects depend on the intensity of use or the length of the interval, and to borrow the words from Rohrer and Murayama (2021), there is no such thing as ``the'' effect of social media use on well-being.
+Hence, to document how effects unfold, future research needs to employ different study designs probing different time lags.
+In addition, more thought needs to be invested in what relevant stable and varying factors we should include as control variables, and I hope this study provides a first step into this direction.
+
+Although I had already reduced the predefined SESOIs to be less conservative, they were potentially still too large.
+Media use is only one aspect of several factors that simultaneously affect well-being.
+Is it really realistic to expect that extremely changing only \emph{one} of these aspects should manifest in a detectable change in well-being?
+Or would it make more sense to expect that thoroughly committing to say \emph{two} activities (e.g.~regularly exercising \emph{and} establishing a reading habit) should then cause a detectable improvement in well-being?
+Practically, this would imply a SESOI half as large as I have defined here, namely \emph{b} = \textbar.15\textbar{} for well-being and \emph{b} = \textbar.075\textbar{} for affect.
+In the case of this study, however, reducing the SESOI would not even make a big difference, as also with these more liberal thresholds all but three effect would still be completely in the null region, and no effect would be outside of the null region.
+However, at all events I encourage future research to start a thorough conversation on what effect sizes are considered meaningful and what not.
+Again, with this study I hope to provide some first input and guidelines.
+
+Both media use and well-being were measured using self-reports.
+Measuring well-being with self-reports is adequate, because it by definition requires introspection.
+However, it would be preferable to measure social media use objectively, because people cannot reliably estimate their use (Scharkow, 2016).
+That said, objective measures often cannot capture the content or the motivation of the use, and only very complicated tools recording the actual content (such as the Screenome project) might produce such data.
+Unfortunately, such procedures introduce other problems, especially related to privacy.
+Hence, for this type of research question it still seems necessary to use self-reported measures.
+
+Because the data were collected in a single country, the generalizability of the results is limited.
+The results apply primarily to the more Western sphere, and might not hold true in other cultures, especially cultures with a different media landscape or alternative social media channels.
+That said, because this is a comparatively large study largely representative of an entire country, and because several waves were collected across a large time span, the results should be at least as generalizable as other typical empirical studies collected in the social sciences.
+
+\hypertarget{conclusion}{%
+\subsection{Conclusion}\label{conclusion}}
+
+In this study, COVID-19 related social media use didn't meaningfully affect several indicators of well-being, including life satisfaction, positive affect, and negative affect.
+However, factors other than social media use were meaningfully related to well-being, such as physical health, exercise, satisfaction with democracy, or believing that one is in control of one's life.
+If it's our aim to improve well-being, it might hence be more fruitful not to focus so much on social media but to address other, more pertinent societal problems related to health care, regular exercise, or a functioning democratic system.
+
+\newpage
+
+\hypertarget{references}{%
+\section{References}\label{references}}
+
+\hypertarget{refs}{}
+\begin{CSLReferences}{1}{0}
+\leavevmode\vadjust pre{\hypertarget{ref-baguleyStandardizedSimpleEffect2009}{}}%
+Baguley, T. (2009). Standardized or simple effect size: {What} should be reported? \emph{British Journal of Psychology}, \emph{100}(3), 603--617. \url{https://doi.org/10.1348/000712608X377117}
+
+\leavevmode\vadjust pre{\hypertarget{ref-bellFixedRandomEffects2019}{}}%
+Bell, A., Fairbrother, M., \& Jones, K.
+(2019).
+Fixed and random effects models: Making an informed choice.
+\emph{Quality \& Quantity}, \emph{53}(2), 1051--1074.
+\url{https://doi.org/10.1007/s11135-018-0802-x}
+
+\leavevmode\vadjust pre{\hypertarget{ref-bendauAssociationsCOVID19Related2021}{}}%
+Bendau, A., Petzold, M.
+B., Pyrkosch, L., Mascarell Maricic, L., Betzler, F., Rogoll, J., \ldots{} Plag, J.
+(2021).
+Associations between {COVID-19} related media consumption and symptoms of anxiety, depression and {COVID-19} related fear in the general population in {Germany}.
+\emph{European Archives of Psychiatry and Clinical Neuroscience}, \emph{271}(2), 283--291.
+\url{https://doi.org/10.1007/s00406-020-01171-6}
+
+\leavevmode\vadjust pre{\hypertarget{ref-beyensSocialMediaUse2021}{}}%
+Beyens, I., Pouwels, J.
+L., van Driel, I.
+I., Keijsers, L., \& Valkenburg, P.
+M.
+(2021).
+Social media use and adolescents' well-being: {Developing} a typology of person-specific effect patterns.
+\emph{Communication Research}.
+\url{https://doi.org/10.1177/00936502211038196}
+
+\leavevmode\vadjust pre{\hypertarget{ref-bradleyStressMoodSmartphone2021}{}}%
+Bradley, A., \& Howard, A. (2021). Stress, mood, and smartphone use in {University} students: {A} 12-week longitudinal study. \url{https://doi.org/10.31219/osf.io/frvpb}
+
+\leavevmode\vadjust pre{\hypertarget{ref-brownEasyHappinessPie2019}{}}%
+Brown, N.
+J.
+L., \& Rohrer, J.
+M.
+(2019).
+Easy as (happiness) pie?
+{A} critical evaluation of a popular model of the determinants of well-being.
+\emph{Journal of Happiness Studies}.
+\url{https://doi.org/10.1007/s10902-019-00128-4}
+
+\leavevmode\vadjust pre{\hypertarget{ref-buchiDigitalWellbeingTheory2021}{}}%
+Büchi, M. (2021). Digital well-being theory and research. \emph{New Media \& Society}, 146144482110568. \url{https://doi.org/10.1177/14614448211056851}
+
+\leavevmode\vadjust pre{\hypertarget{ref-choiMediatedCommunicationMatters2021}{}}%
+Choi, M., \& Choung, H.
+(2021).
+Mediated communication matters during the {COVID-19} pandemic: {The} use of interpersonal and masspersonal media and psychological well-being.
+\emph{Journal of Social and Personal Relationships}, \emph{38}(8), 2397--2418.
+\url{https://doi.org/10.1177/02654075211029378}
+
+\leavevmode\vadjust pre{\hypertarget{ref-cohenPowerPrimer1992}{}}%
+Cohen, J. (1992). A power primer. \emph{Psychological Bulletin}, \emph{112}(1), 155--159. \url{https://doi.org/10.1037/0033-2909.112.1.155}
+
+\leavevmode\vadjust pre{\hypertarget{ref-dienerAdvancesOpenQuestions2018}{}}%
+Diener, E., Lucas, R.
+E., \& Oishi, S.
+(2018).
+Advances and open questions in the science of subjective well-being.
+\emph{Collabra: Psychology}, \emph{4}(1), 15.
+\url{https://doi.org/10.1525/collabra.115}
+
+\leavevmode\vadjust pre{\hypertarget{ref-dienesUsingBayesGet2014}{}}%
+Dienes, Z. (2014). Using {Bayes} to get the most out of non-significant results. \emph{Frontiers in Psychology}, \emph{5}. \url{https://doi.org/10.3389/fpsyg.2014.00781}
+
+\leavevmode\vadjust pre{\hypertarget{ref-dienlinImpactDigitalTechnology2020}{}}%
+Dienlin, T., \& Johannes, N.
+(2020).
+The impact of digital technology use on adolescent well-being.
+\emph{Dialogues in Clinical Neuroscience}, \emph{22}(2), 135--142.
+\url{https://doi.org/doi:10.31887/DCNS.2020.22.2/tdienlin}
+
+\leavevmode\vadjust pre{\hypertarget{ref-dornemannHowGoodBad2021}{}}%
+Dörnemann, A., Boenisch, N., Schommer, L., Winkelhorst, L., \& Wingen, T.
+(2021).
+How do good and bad news impact mood during the {Covid-19} pandemic?
+{The} role of similarity.
+\url{https://doi.org/10.31219/osf.io/sy2kd}
+
+\leavevmode\vadjust pre{\hypertarget{ref-duradoniWellbeingSocialMedia2020}{}}%
+Duradoni, M., Innocenti, F., \& Guazzini, A.
+(2020).
+Well-being and social media: {A} systematic review of {Bergen Addiction Scales}.
+\emph{Future Internet}, \emph{12}(2, 2), 24.
+\url{https://doi.org/10.3390/fi12020024}
+
+\leavevmode\vadjust pre{\hypertarget{ref-edenMediaCopingCOVID192020}{}}%
+Eden, A.
+L., Johnson, B.
+K., Reinecke, L., \& Grady, S.
+M.
+(2020).
+Media for coping during {COVID-19} social distancing: {Stress}, anxiety, and psychological well-being.
+\emph{Frontiers in Psychology}, \emph{11}, 577639.
+\url{https://doi.org/10.3389/fpsyg.2020.577639}
+
+\leavevmode\vadjust pre{\hypertarget{ref-ellisonWhyWeDon2020}{}}%
+Ellison, N.
+B., Triệu, P., Schoenebeck, S., Brewer, R., \& Israni, A.
+(2020).
+Why we don't click: {Interrogating} the relationship between viewing and clicking in social media contexts by exploring the {``{Non-Click}.''} \emph{Journal of Computer-Mediated Communication}, \emph{25}(6), 402--426.
+\url{https://doi.org/10.1093/jcmc/zmaa013}
+
+\leavevmode\vadjust pre{\hypertarget{ref-europeansocialsurveyESS9Edition20182021}{}}%
+European Social Survey. (2021). {ESS9} edition 3.1 - 2018 {Documentation Report}. Retrieved from \url{https://www.europeansocialsurvey.org/docs/round9/survey/ESS9_data_documentation_report_e03_1.pdf}
+
+\leavevmode\vadjust pre{\hypertarget{ref-fergusonThisMetaanalysisScreen2021}{}}%
+Ferguson, C.
+J., Kaye, L.
+K., Branley-Bell, D., Markey, P., Ivory, J.
+D., Klisanin, D., \ldots{} Wilson, J.
+(2021).
+Like this meta-analysis: {Screen} media and mental health.
+\emph{Professional Psychology: Research and Practice}.
+\url{https://doi.org/10.1037/pro0000426}
+
+\leavevmode\vadjust pre{\hypertarget{ref-funderEvaluatingEffectSize2019}{}}%
+Funder, D.
+C., \& Ozer, D.
+J.
+(2019).
+Evaluating effect size in psychological research: {Sense} and nonsense.
+\emph{Advances in Methods and Practices in Psychological Science}, \emph{2}(2), 156--168.
+\url{https://doi.org/10.1177/2515245919847202}
+
+\leavevmode\vadjust pre{\hypertarget{ref-galerHowMuchToo2018}{}}%
+Galer, S.
+S.
+(2018, January 19).
+How much is {`too much time'} on social media?
+Retrieved April 4, 2022, from \url{https://www.bbc.com/future/article/20180118-how-much-is-too-much-time-on-social-media}
+
+\leavevmode\vadjust pre{\hypertarget{ref-greenspoonIntegrationSubjectiveWellbeing2001}{}}%
+Greenspoon, P.
+J., \& Saklofske, D.
+H.
+(2001).
+Toward an integration of subjective well-being and psychopathology.
+\emph{Social Indicators Research}, \emph{54}(1), 81--108.
+\url{https://doi.org/10.1023/A:1007219227883}
+
+\leavevmode\vadjust pre{\hypertarget{ref-guazziniSecondWaveAnalysis2022}{}}%
+Guazzini, A., Pesce, A., Marotta, L., \& Duradoni, M.
+(2022).
+Through the second wave: {Analysis} of the psychological and perceptive changes in the {Italian} population during the {COVID-19} pandemic.
+\emph{International Journal of Environmental Research and Public Health}, \emph{19}(3), 1635.
+\url{https://doi.org/10.3390/ijerph19031635}
+
+\leavevmode\vadjust pre{\hypertarget{ref-hamakerWhyResearchersShould2014}{}}%
+Hamaker, E.
+L.
+(2014).
+Why researchers should think "within-person": {A} paradigmatic rationale.
+In M.
+R.
+Mehl, T.
+S.
+Conner, \& M.
+Csikszentmihalyi (Eds.), \emph{Handbook of research methods for studying daily life} (Paperback ed.).
+{New York, NY}: {Guilford}.
+
+\leavevmode\vadjust pre{\hypertarget{ref-huangTimeSpentSocial2017}{}}%
+Huang, C.
+(2017).
+Time spent on social network sites and psychological well-being: {A} meta-analysis.
+\emph{Cyberpsychology, Behavior and Social Networking}, \emph{20}(6), 346--354.
+\url{https://doi.org/10.1089/cyber.2016.0758}
+
+\leavevmode\vadjust pre{\hypertarget{ref-kerestesAdolescentsOnlineSocial2020}{}}%
+Keresteš, G., \& Štulhofer, A.
+(2020).
+Adolescents' online social network use and life satisfaction: {A} latent growth curve modeling approach.
+\emph{Computers in Human Behavior}, \emph{104}, 106187.
+\url{https://doi.org/10.1016/j.chb.2019.106187}
+
+\leavevmode\vadjust pre{\hypertarget{ref-kittelAustrianCoronaPanel2020}{}}%
+Kittel, B., Kritzinger, S., Boomgaarden, H., Prainsack, B., Eberl, J.-M., Kalleitner, F., \ldots{} Schlogl, L.
+(2020).
+\emph{Austrian {Corona Panel Project} ({SUF} edition)} {[}Data set{]}.
+{AUSSDA}.
+\url{https://doi.org/10.11587/28KQNS}
+
+\leavevmode\vadjust pre{\hypertarget{ref-kittelAustrianCoronaPanel2021}{}}%
+Kittel, B., Kritzinger, S., Boomgaarden, H., Prainsack, B., Eberl, J.-M., Kalleitner, F., \ldots{} Schlogl, L.
+(2021).
+The {Austrian Corona Panel Project}: Monitoring individual and societal dynamics amidst the {COVID-19} crisis.
+\emph{European Political Science}, \emph{20}(2), 318--344.
+\url{https://doi.org/10.1057/s41304-020-00294-7}
+
+\leavevmode\vadjust pre{\hypertarget{ref-kleinDarklySoothingCompulsion2021}{}}%
+Klein, J.
+(2021, March 3).
+The darkly soothing compulsion of 'doomscrolling'.
+Retrieved November 1, 2021, from \url{https://www.bbc.com/worklife/article/20210226-the-darkly-soothing-compulsion-of-doomscrolling}
+
+\leavevmode\vadjust pre{\hypertarget{ref-klinePrinciplesPracticeStructural2016}{}}%
+Kline, R. B. (2016). \emph{Principles and practice of structural equation modeling} (4th ed.). {New York, NY}: {The Guilford Press}.
+
+\leavevmode\vadjust pre{\hypertarget{ref-liuRelationOfficialWhatsAppdistributed2020}{}}%
+Liu, J.
+C.
+J., \& Tong, E.
+M.
+W.
+(2020).
+The relation between official {WhatsApp-distributed COVID-19} news exposure and psychological symptoms: Cross-sectional survey study.
+\emph{Journal of Medical Internet Research}, \emph{22}(9), e22142.
+\url{https://doi.org/10.2196/22142}
+
+\leavevmode\vadjust pre{\hypertarget{ref-lucasAdaptationSetpointModel2007}{}}%
+Lucas, R.
+E.
+(2007).
+Adaptation and the set-point model of subjective well-being.
+\emph{Current Directions in Psychological Science}, \emph{16}(2), 75--79.
+\url{https://doi.org/10.1111/j.1467-8721.2007.00479.x}
+
+\leavevmode\vadjust pre{\hypertarget{ref-lykkenHappinessWhatStudies1999}{}}%
+Lykken, D. T. (1999). \emph{Happiness: {What} studies on twins show us about nature, nurture, and the happiness set-point}. {New York, N.Y.}: {Golden Books}.
+
+\leavevmode\vadjust pre{\hypertarget{ref-marcianoDynamicsAdolescentsSmartphone2022}{}}%
+Marciano, L., Driver, C.
+C., Schulz, P.
+J., \& Camerini, A.-L.
+(2022).
+Dynamics of adolescents' smartphone use and well-being are positive but ephemeral.
+\emph{Scientific Reports}, \emph{12}(1), 1316.
+\url{https://doi.org/10.1038/s41598-022-05291-y}
+
+\leavevmode\vadjust pre{\hypertarget{ref-mcelreathYesterdayClass2021}{}}%
+McElreath, R. (2021, January 28). Yesterday in class, ... {[}Tweet{]}. Retrieved November 5, 2021, from \url{https://twitter.com/rlmcelreath/status/1354786005996482563}
+
+\leavevmode\vadjust pre{\hypertarget{ref-meierDoesPassiveSocial2022}{}}%
+Meier, A., \& Krause, H.-V. (2022). \emph{Does passive social media use harm well-being? {An} adversarial review} (preprint). {PsyArXiv}. \url{https://doi.org/10.31234/osf.io/nvbwh}
+
+\leavevmode\vadjust pre{\hypertarget{ref-meierComputermediatedCommunicationSocial2020a}{}}%
+Meier, A., \& Reinecke, L.
+(2020).
+Computer-mediated communication, social media, and mental health: {A} conceptual and empirical meta-review.
+\emph{Communication Research}, 009365022095822.
+\url{https://doi.org/10.1177/0093650220958224}
+
+\leavevmode\vadjust pre{\hypertarget{ref-normanInterpretationChangesHealthrelated2003}{}}%
+Norman, G., Sloan, J., \& Wyrwich, K.
+(2003).
+Interpretation of changes in health-related quality of life: {The} remarkable universality of half a standard deviation.
+\emph{Medical Care}, \emph{41}(5), 582--592.
+Retrieved from \url{Retrieved\%20from\%20http://www.jstor.org/stable/3768017}
+
+\leavevmode\vadjust pre{\hypertarget{ref-orbenTeenagersScreensSocial2020}{}}%
+Orben, A.
+(2020).
+Teenagers, screens and social media: A narrative review of reviews and key studies.
+\emph{Social Psychiatry and Psychiatric Epidemiology}, \emph{55}(4), 407--414.
+\url{https://doi.org/10.1007/s00127-019-01825-4}
+
+\leavevmode\vadjust pre{\hypertarget{ref-orbenSocialMediaEnduring2019}{}}%
+Orben, A., Dienlin, T., \& Przybylski, A.
+K.
+(2019).
+Social media's enduring effect on adolescent life satisfaction.
+\emph{Proceedings of the National Academy of Sciences of the United States of America}, \emph{116}(21), 10226--10228.
+\url{https://doi.org/10.1073/pnas.1902058116}
+
+\leavevmode\vadjust pre{\hypertarget{ref-pelletierOneSizeDoesn2020}{}}%
+Pelletier, M.
+J., Krallman, A., Adams, F.
+G., \& Hancock, T.
+(2020).
+One size doesn't fit all: A uses and gratifications analysis of social media platforms.
+\emph{Journal of Research in Interactive Marketing}, \emph{14}(2), 269--284.
+\url{https://doi.org/10.1108/JRIM-10-2019-0159}
+
+\leavevmode\vadjust pre{\hypertarget{ref-przybylskiDoesTakingShort2021a}{}}%
+Przybylski, A.
+K., Nguyen, T.
+T., Law, W., \& Weinstein, N.
+(2021).
+Does taking a short break from social media have a positive effect on well-being?
+{Evidence} from three preregistered field experiments.
+\emph{Journal of Technology in Behavioral Science}, \emph{6}(3), 507--514.
+\url{https://doi.org/10.1007/s41347-020-00189-w}
+
+\leavevmode\vadjust pre{\hypertarget{ref-przybylskiLargescaleTestGoldilocks2017}{}}%
+Przybylski, A. K., \& Weinstein, N. (2017). A large-scale test of the {Goldilocks} hypothesis. \emph{Psychological Science}, \emph{28}(2), 204--215. \url{https://doi.org/10.1177/0956797616678438}
+
+\leavevmode\vadjust pre{\hypertarget{ref-riehmAssociationsMediaExposure2020}{}}%
+Riehm, K.
+E., Holingue, C., Kalb, L.
+G., Bennett, D., Kapteyn, A., Jiang, Q., \ldots{} Thrul, J.
+(2020).
+Associations between media exposure and mental distress among {U}.{S}.
+Adults at the beginning of the {COVID-19} pandemic.
+\emph{American Journal of Preventive Medicine}, \emph{59}(5), 630--638.
+\url{https://doi.org/10.1016/j.amepre.2020.06.008}
+
+\leavevmode\vadjust pre{\hypertarget{ref-rohrerThinkingClearlyCorrelations2018}{}}%
+Rohrer, J.
+M.
+(2018).
+Thinking clearly about correlations and causation: {Graphical} causal models for observational data.
+\emph{Advances in Methods and Practices in Psychological Science}, \emph{24}(2), 251524591774562.
+\url{https://doi.org/10.1177/2515245917745629}
+
+\leavevmode\vadjust pre{\hypertarget{ref-rohrerTheseAreNot2021}{}}%
+Rohrer, J.
+M., \& Murayama, K.
+(2021).
+\emph{These are not the effects you are looking for: {Causality} and the within-/between-person distinction in longitudinal data analysis} (preprint).
+{PsyArXiv}.
+\url{https://doi.org/10.31234/osf.io/tg4vj}
+
+\leavevmode\vadjust pre{\hypertarget{ref-sandstromDoomscrollingCOVIDNews2021}{}}%
+Sandstrom, G., Buchanan, K., Aknin, L., \& Lotun, S.
+(2021, October 22).
+Doomscrolling {COVID} news takes an emotional toll -- here's how to make your social media a happier place.
+Retrieved November 3, 2021, from \url{http://theconversation.com/doomscrolling-covid-news-takes-an-emotional-toll-heres-how-to-make-your-social-media-a-happier-place-170342}
+
+\leavevmode\vadjust pre{\hypertarget{ref-scharkowAccuracySelfreportedInternet2016}{}}%
+Scharkow, M.
+(2016).
+The accuracy of self-reported {Internet} use---{A} validation study using client log data.
+\emph{Communication Methods and Measures}, \emph{10}(1), 13--27.
+\url{https://doi.org/10.1080/19312458.2015.1118446}
+
+\leavevmode\vadjust pre{\hypertarget{ref-schemerImpactInternetSocial2021}{}}%
+Schemer, C., Masur, P.
+K., Geiß, S., Müller, P., \& Schäfer, S.
+(2021).
+The impact of {Internet} and social media use on well-being: {A} longitudinal analysis of adolescents across nine years.
+\emph{Journal of Computer-Mediated Communication}, \emph{26}(1), 1--21.
+\url{https://doi.org/10.1093/jcmc/zmaa014}
+
+\leavevmode\vadjust pre{\hypertarget{ref-schnauber-stockmannMobileDevicesTools2020}{}}%
+Schnauber-Stockmann, A., \& Karnowski, V.
+(2020).
+Mobile devices as tools for media and communication research: {A} scoping review on collecting self-report data in repeated measurement designs.
+\emph{Communication Methods and Measures}, \emph{14}(3), 145--164.
+\url{https://doi.org/10.1080/19312458.2020.1784402}
+
+\leavevmode\vadjust pre{\hypertarget{ref-sewallObjectivelyMeasuredDigital2021}{}}%
+Sewall, C.
+J.
+R., Goldstein, T.
+R., \& Rosen, D.
+(2021).
+Objectively measured digital technology use during the {COVID-19} pandemic: {Impact} on depression, anxiety, and suicidal ideation among young adults.
+\emph{Journal of Affective Disorders}, \emph{288}, 145--147.
+\url{https://doi.org/10.1016/j.jad.2021.04.008}
+
+\leavevmode\vadjust pre{\hypertarget{ref-sheldonStabilityHappinessTheories2014}{}}%
+Sheldon, K. M., \& Lucas, R. E. (2014). \emph{Stability of happiness: Theories and evidence on whether happiness can change}. Retrieved from \url{http://site.ebrary.com/id/10891875}
+
+\leavevmode\vadjust pre{\hypertarget{ref-stainbackCOVID1924News2020}{}}%
+Stainback, K., Hearne, B.
+N., \& Trieu, M.
+M.
+(2020).
+{COVID-19} and the 24/7 {News Cycle}: {Does COVID-19 News Exposure Affect Mental Health}?
+\emph{Socius}, \emph{6}, 2378023120969339.
+\url{https://doi.org/10.1177/2378023120969339}
+
+\leavevmode\vadjust pre{\hypertarget{ref-statistaAverageDailyTime2021}{}}%
+Statista.
+(2021, May 21).
+Average daily time spent on social networks by users in the {United States} from 2018 to 2022.
+Retrieved from \url{https://www.statista.com/statistics/1018324/us-users-daily-social-media-minutes/}
+
+\leavevmode\vadjust pre{\hypertarget{ref-valkenburgDifferentialSusceptibilityMedia2013}{}}%
+Valkenburg, P. M., \& Peter, J. (2013). The differential susceptibility to media effects model. \emph{Journal of Communication}, \emph{63}(2), 221--243. \url{https://doi.org/10.1111/jcom.12024}
+
+\leavevmode\vadjust pre{\hypertarget{ref-vanrooijWeakScientificBasis2018}{}}%
+van Rooij, A.
+J., Ferguson, C.
+J., Colder Carras, M., Kardefelt-Winther, D., Shi, J., Aarseth, E., \ldots{} Przybylski, A.
+K.
+(2018).
+A weak scientific basis for gaming disorder: {Let} us err on the side of caution.
+\emph{Journal of Behavioral Addictions}, \emph{7}(1), 1--9.
+\url{https://doi.org/10.1556/2006.7.2018.19}
+
+\leavevmode\vadjust pre{\hypertarget{ref-wagnerAUTNESOnlinePanel2018}{}}%
+Wagner, M., Aichholzer, J., Eberl, J.-M., Meyer, T.
+M., Berk, N., Büttner, N., \ldots{} Müller, W.
+C.
+(2018).
+\emph{{AUTNES Online Panel Study} 2017 ({SUF} edition)} {[}Data set{]}.
+{AUSSDA}.
+\url{https://doi.org/10.11587/I7QIYJ}
+
+\leavevmode\vadjust pre{\hypertarget{ref-waterlooNormsOnlineExpressions2018}{}}%
+Waterloo, S.
+F., Baumgartner, S.
+E., Peter, J., \& Valkenburg, P.
+M.
+(2018).
+Norms of online expressions of emotion: {Comparing Facebook}, {Twitter}, {Instagram}, and {WhatsApp}.
+\emph{New Media \& Society}, \emph{20}(5), 1813--1831.
+\url{https://doi.org/10.1177/1461444817707349}
+
+\leavevmode\vadjust pre{\hypertarget{ref-worldhealthorganizationWellbeingMeasuresPrimary1998}{}}%
+World Health Organization. (1998). Wellbeing measures in primary health care/{The Depcare Project}.
+
+\leavevmode\vadjust pre{\hypertarget{ref-yangCanWatchingOnline2021}{}}%
+Yang, Z., Griffiths, M.
+D., Yan, Z., \& Xu, W.
+(2021).
+Can watching online videos be addictive?
+{A} qualitative exploration of online video watching among {Chinese} young adults.
+\emph{International Journal of Environmental Research and Public Health}, \emph{18}(14, 14), 7247.
+\url{https://doi.org/10.3390/ijerph18147247}
+
+\leavevmode\vadjust pre{\hypertarget{ref-zillmannMoodManagementCommunication1988}{}}%
+Zillmann, D. (1988). Mood {Management Through Communication Choices}. \emph{American Behavioral Scientist}, \emph{31}(3), 327340. \url{https://doi.org/10.1177/000276488031003005}
+
+\end{CSLReferences}
+
+\newpage
+
 \begin{table}[tbp]
@@
 
-\caption{\label{tab:descriptives}Descriptives of the main variables.}
+\caption{\label{tab:tab-descriptives}Descriptives of the main variables.}
 
@@
 
-\hypertarget{control-variables}{%
-\subsubsection{Control variables}\label{control-variables}}
-
-The effects of COVID-19 related social media use were controlled for the following stable variables:
-(a) gender (female, male, diverse), (b) age, (c) education (ten options), (d) Austria country of birth (yes/no), (e) Austria parents' country of birth (no parent, one parent, both parents).
-I originally planned to implement additional variables as varying covariates.
-However, because they were not measured often enough or not at the time when social media use was measured, I implemented them as stable variables using their average values across all waves.
-This includes (a) text-based media news consumption (five degrees), (b) video-based media news consumption (five degrees), (c) residency is Vienna (yes/no), (d) self-reported physical health (five degrees), (e) living space (eleven options), (f) access to balcony (yes/no), (g) access to garden (yes/no), (h) employment (nine options), (i) disposition to take risks (eleven degrees), and (j) locus of control (five degrees).
-I also controlled for the following varying covariates: (a) five items measuring outdoor activities such as sport or meeting friends (five degrees), and (b) satisfaction with democracy (five degrees).
-Because it lead to too much attrition in the sample, I did not control for (a) household size, (b) work hours per week, (c) home office, (d) household income.
-
-\hypertarget{results}{%
-\section{Results}\label{results}}
-
-First, when looking at the variables from a descriptive perspective, we see that all well-being measures did not change substantially across the different waves of data collection.
-COVID-19 related media use, however, decreased slightly at the beginning of the study and remained stable after approximately six waves.
-The initial decrease might be explained by the fact that the collection of data began at the end of March 2020, hence approximately three months after the pandemic began.
-It could be that after an initial uptick, COVID-19 related social media use was already declining at the time, returning to more normal levels.
+\newpage
+
+\begin{figure}
+\centering
+\includegraphics{manuscript_files/figure-latex/sesoi-1.pdf}
+\caption{\label{fig:sesoi}Illustration of how confidence intervals are used to test a null region.
+Here, a trivial effect of social media use on life satisfaction is defined as ranging from b = -.25 to b = .25}
+\end{figure}
+
+\newpage
 
@@
 \includegraphics[width=\textwidth]{figures/fig_descriptives} \caption{Development of well-being and media use measures across the pandemic.
-Values obtained from mixed effect models, with participants and waves as grouping factors and without additional predictors.}\label{fig:fig-desc}
+Values obtained from mixed effect models, with participants and waves as grouping factors and without additional predictors.}\label{fig:descriptives}
 \end{figure}
 
-\hypertarget{preregistered-analyses}{%
-\subsection{Preregistered Analyses}\label{preregistered-analyses}}
-
-The study's main hypothesis was that the effects of social media use on well-being would be trivial.
-Regarding the effects of different communication \emph{types}---that is, reading vs.~sharing vs.~posting---all within-person effects fell completely within the a-priori defined null region (see Figure \ref{fig:res-activity}).
-For example, respondents who used social media more frequently than usual to read about COVID-19 related topics did not show a simultaneous change in life satisfaction (\emph{b} = 0.05 {[}95\% CI -0.01, 0.1{]}).
-All confidence intervals included zero; hence, all effects were also statistically non-significant.
-As a result, the hypothesis was supported for all COVID-19 related types of social media communication.
+\newpage
 
@@
 
-Regarding between-person relations, about which no hypotheses were formulated, only three effects did not include zero.
-Respondents who across all waves used social media more frequently than others to read about COVID-19 related posts reported slightly lower levels of positive affect than others (\emph{b} = -0.03 {[}95\% CI \textgreater{} -0.01, -0.06{]}).
-Respondents who across all waves used social media more frequently than others to write COVID-19 related posts reported higher levels of negative affect than others (\emph{b} = 0.06 {[}95\% CI 0.09, 0.03{]}).
-Interestingly, respondents who across all waves used social media more frequently than others to write COVID-19 related posts also reported slightly higher levels of positive affect than others (\emph{b} = 0.05 {[}95\% CI 0.09, \textless{} 0.01{]}).
-However, note that the effect were still completely inside of the null region, hence not large enough to be considered practically relevant.
-
-Note that when comparing the results with and without control variables, the results differed.
-For example, on the between-person level, one effect stopped being significant if controlled for additional variables.
-Actively posting on social media was significantly (though not meaningfully) related to decreased life satisfaction.
-However, when controlling for potential confounders, the effect became virtually zero (see Figure \ref{fig:res-channels}).
+\newpage
 
 \begin{figure}
-\includegraphics[width=\textwidth]{figures/fig_results_channel} \caption{The effects of using various social media applications on three indicators of well-being.
+\includegraphics[width=\textwidth]{figures/fig_results_channels} \caption{The effects of using various social media applications on three indicators of well-being.
 The black estimates show the effects controlled for a large number of covariates (see text; preregistered); the grey estimates are without control variables (exploratory).
@@
 
-Regarding the COVID-19 related use of social media \emph{channels}, the results were comparable (see Figure \ref{fig:res-channels}).
-Changes in the frequency of using different social media channels to attain information regarding COVID-19 were unrelated to meaningful changes in well-being.
-For example, respondents who used Facebook more frequently than usual to learn about COVID-19 did not show a simultaneous change in well-being (\emph{b} = -0.05 {[}95\% CI -0.11, 0.01{]}).
-Only two effects differed substantially from zero.
-Respondents who used Instagram more frequently than usual to attain COVID-19 related news reported slightly \emph{higher} levels of life satisfaction than usual (\emph{b} = 0.09 {[}95\% CI 0.02, 0.16{]}).
-Respondents who used Twitter more frequently than usual to attain COVID-19 related news reported slightly \emph{lower} levels of life satisfaction than usual (\emph{b} = -0.12 {[}95\% CI -0.23, -0.02{]}).
-However, both effects were still completely inside of the null region, hence not large enough to be considered meaningful.
-In sum, the hypothesis was supported also for the COVID-19 related use of important social media channels.
-
-In terms of between-person relations---which, again, weren't included in the hypotheses---no relations crossed the null region or fell outside of it.
-Only one relation did not include zero, was hence statistically significant.
-Respondents who across all waves used YouTube more frequently than others for COVID-19 related reasons reported marginally \emph{higher} levels of life satisfaction (\emph{b} = 0.08 {[}95\% CI \textless{} 0.01, 0.16{]}).
-However, please note that this effect again was not large enough to be considered practically relevant.
-
-Again, note that when comparing the results with and without control variables, the results differed.
-Especially on the between-person level, altogether five effects stopped being significant if they were controlled for additional variables.
-For example, using Instagram was significantly (though not meaningfully) related to increased life satisfaction.
-However, when controlling for additional covariates, the effect became virtually zero (see Figure \ref{fig:res-channels}).
-
-\hypertarget{exploratory-analyses}{%
-\subsection{Exploratory Analyses}\label{exploratory-analyses}}
-
-In what follows, I briefly report some exploratory analyses that weren't preregistered.
-First, to contextualize the results reported above and to see if the results included any meaningful effects at all, I also looked at the effect sizes of selected (cherry-picked) covariates.
-Because each variable had different response options, we would actually need to define a SESOI for each variable, which for reasons of scope I cannot implement here.
-Therefore, I report the results of the standardized scales,
-which allows for a better comparison across the differently scaled variables.
-For what it's worth, as a rough estimate for the SESOI we can build on the typical convention that small effects start at \emph{r} = \textbar.10\textbar.
-The results showed that several effects fell outside of the SESOI, were hence considered meaningful.
-This includes for example internal locus of control, health, satisfaction with democracy, or exercising.
-For an overview, see Figure \ref{fig:res-control}.
+\newpage
 
@@
 
-To find out whether my inferences were robust across legitimate (though arguably inferior) alternative analyses, I reran the analyses also using standardized estimates, mean scores instead of factor scores, and with a data set where missing data were not imputed.
-The results were virtually the same.
-For example, all standardized COVID-19 related types of social media use or channels were not significantly larger than a SESOI of \(\beta\) = \textbar.10\textbar.
-The additional analyses are reported in the \href{https://XMtRA.github.io/Austrian_Corona_Panel_Project/analyses_additional.html}{companion website}.
-
-\hypertarget{discussion}{%
-\section{Discussion}\label{discussion}}
-
-In this study I analyzed the effects of COVID-19 related social media use on well-being.
-The data come from a panel study with 24 waves and are representative of the Austrian population.
-In a random effects model I separated between person relations from within-person effects and controlled for a large number of both stable and varying covariates, thereby aiming to assess causality.
-The results showed that within-person effects were trivial.
-People who used social media more than usual to learn about COVID-19 didn't show meaningful changes in their well-being.
-
-The results imply that COVID-19 related social media use doesn't seem to be particularly relevant for well-being.
-Other factors among the third variables that were measured revealed larger effects or relations, suggesting that well-being is determined by alternative aspects such as health, satisfaction with democracy, locus of control, or exercising.
-According to this study, popular fears that ``doomscrolling'' or overusing social media during times of crises don't seem to be justified.
-
-On the one hand, the results are not aligned with several recent studies analyzing similar or closely related research questions.
-This includes a study by Bendau et al. (2021), which showed negative relations between social media and well-being (but see Bradley \& Howard, 2021; or Sewall, Goldstein, \& Rosen, 2021).
-However, note that Bendau et al. (2021) analyzed cross-sectional data on a between-person level while not controlling for third variables, which is not optimal for investigating causal effects.
-On the other hand, the results are well-aligned with recent studies and meta-analyses analyzing the effects of social media use from a more general perspective or from a somewhat different angle.
-These studies have found that the effects of various types of social media use on several well-being indicators are small at best, often too small to matter (Ferguson et al., 2021; Meier \& Reinecke, 2020; Orben, 2020), which echoes the results obtained here.
-
-If anything, two preliminary and subtle trends can be observed.
-First, of all the three COVID-19 related social media activities, people who read about the pandemic more than others showed slightly decreased levels of positive affect, and people who actively posted about the pandemic more than others showed slightly increased levels of negative affect.
-On the other hand, however, people who posted more about COVID-19 also showed slightly higher levels of positive affects, so taken together the results are ambivalent.
-Second, in terms of media channels, using Twitter more than usual was related to slightly decreased levels of life satisfaction.
-Twitter is considered to have more negative affordances and tonality as compared to other networks such as Instagram (Waterloo, Baumgartner, Peter, \& Valkenburg, 2018), which might help explain the results.
-Instagram, on the other hand, was related to slightly increased levels of life satisfaction.
-To speculate, the often-criticized positivity bias on Instagram might have been somewhat beneficial in times of the pandemic.
-That said, all these effects were still very small and arguably too small to matter.
-But future research might elaborate on these specific relations to probe their stability and relevance.
-
-Finally, another interesting observation is that life satisfaction was remarkably stable.
-Hence, even in times of a pandemic, it seems that such broad assessment of life vary only mildly.
-This supports the hypothesis that life satisfaction seems to be determined largely by stable factors such as one's genes (Brown \& Rohrer, 2019).
-
-\hypertarget{limitations}{%
-\subsection{Limitations}\label{limitations}}
-
-The current study analyzed whether changes in media use were related to changes in well-being, while controlling for several potential confounders.
-Together, this allows for an improved perspective on assessing causality.
-That said, causality necessitates temporal order, and the cause needs to precede the effect.
-Regarding media use, such effects often happen immediately or shortly after use, necessitating intervals in the hours, minutes, or even seconds.
-In many cases only experience sampling studies asking users in the very moment can produce such knowledge.
-However, even then we don't know for certain if we actually measured the right interval.
-Effects depend on the intensity of use or the length of the interval, and to borrow the words from Rohrer and Murayama (2021), there is no such thing as ``the'' effect of social media use on well-being.
-Hence, to document how effects unfold, future research needs to employ different study designs probing different time lags.
-In addition, more thought needs to be invested in what relevant stable and varying factors we should include as control variables, and I hope this study provides a first step into this direction.
-
-Although I had already reduced the predefined SESOIs to be less conservative, they were potentially still too large.
-Media use is only one aspect of several factors that simultaneously affect well-being.
-Is it really realistic to expect that extremely changing only \emph{one} of these aspects should manifest in a detectable change in well-being?
-Or would it make more sense to expect that thoroughly committing to say \emph{two} activities (e.g.~regularly exercising \emph{and} establishing a reading habit) should then cause a detectable improvement in well-being?
-Practically, this would imply a SESOI half as large as I have defined here, namely \emph{b} = \textbar.15\textbar{} for well-being and \emph{b} = \textbar.075\textbar{} for affect.
-In the case of this study, however, reducing the SESOI would not even make a big difference, as also with these more liberal thresholds all but three effect would still be completely in the null region, and no effect would be outside of the null region.
-However, at all events I encourage future research to start a thorough conversation on what effect sizes are considered meaningful and what not.
-Again, with this study I hope to provide some first input and guidelines.
-
-Both media use and well-being were measured using self-reports.
-Measuring well-being with self-reports is adequate, because it by definition requires introspection.
-However, it would be preferable to measure social media use objectively, because people cannot reliably estimate their use (Scharkow, 2016).
-That said, objective measures often cannot capture the content or the motivation of the use, and only very complicated tools recording the actual content (such as the Screenome project) might produce such data.
-Unfortunately, such procedures introduce other problems, especially related to privacy.
-Hence, for this type of research question it still seems necessary to use self-reported measures.
-
-Because the data were collected in a single country, the generalizability of the results is limited.
-The results apply primarily to the more Western sphere, and might not hold true in other cultures, especially cultures with a different media landscape or alternative social media channels.
-That said, because this is a comparatively large study representative of an entire country, and because several waves were collected across a large time span, the results should be at least as generalizable as other typical empirical studies collected in the social sciences.
-
-\hypertarget{conclusion}{%
-\subsection{Conclusion}\label{conclusion}}
-
-In this study, COVID-19 related social media use didn't meaningfully affect several indicators of well-being, including life satisfaction, positive affect, and negative affect.
-However, factors other than social media use were meaningfully related to well-being, such as physical health, exercise, satisfaction with democracy, or believing that one is in control of one's life.
-If it's our aim to improve well-being, it might hence be more fruitful not to focus so much on social media but to address other, more pertinent societal problems related to health care, regular exercise, or a functioning democratic system.
-
-\newpage
-
-\hypertarget{references}{%
-\section{References}\label{references}}
-
-\hypertarget{refs}{}
-\begin{CSLReferences}{1}{0}
-\leavevmode\hypertarget{ref-baguleyStandardizedSimpleEffect2009}{}%
-Baguley, T. (2009). Standardized or simple effect size: {What} should be reported? \emph{British Journal of Psychology}, \emph{100}(3), 603--617. \url{https://doi.org/10.1348/000712608X377117}
-
-\leavevmode\hypertarget{ref-bellFixedRandomEffects2019}{}%
-Bell, A., Fairbrother, M., \& Jones, K.
-(2019).
-Fixed and random effects models: Making an informed choice.
-\emph{Quality \& Quantity}, \emph{53}(2), 1051--1074.
-\url{https://doi.org/10.1007/s11135-018-0802-x}
-
-\leavevmode\hypertarget{ref-bendauAssociationsCOVID19Related2021}{}%
-Bendau, A., Petzold, M.
-B., Pyrkosch, L., Mascarell Maricic, L., Betzler, F., Rogoll, J., \ldots{} Plag, J.
-(2021).
-Associations between {COVID}-19 related media consumption and symptoms of anxiety, depression and {COVID}-19 related fear in the general population in {Germany}.
-\emph{European Archives of Psychiatry and Clinical Neuroscience}, \emph{271}(2), 283--291.
-\url{https://doi.org/10.1007/s00406-020-01171-6}
-
-\leavevmode\hypertarget{ref-beyensSocialMediaUse2021}{}%
-Beyens, I., Pouwels, J.
-L., Driel, I.
-I.
-van, Keijsers, L., \& Valkenburg, P.
-M.
-(2021).
-Social media use and adolescents' well-being: {Developing} a typology of person-specific effect patterns.
-\emph{Communication Research}.
-\url{https://doi.org/10.31234/osf.io/ftygp}
-
-\leavevmode\hypertarget{ref-bradleyStressMoodSmartphone2021}{}%
-Bradley, A., \& Howard, A. (2021). \emph{Stress, mood, and smartphone use in {University} students: {A} 12-week longitudinal study}. OSF Preprints. \url{https://doi.org/10.31219/osf.io/frvpb}
-
-\leavevmode\hypertarget{ref-brownEasyHappinessPie2019}{}%
-Brown, N.
-J.
-L., \& Rohrer, J.
-M.
-(2019).
-Easy as (happiness) pie?
-{A} critical evaluation of a popular model of the determinants of well-being.
-\emph{Journal of Happiness Studies}.
-\url{https://doi.org/10.1007/s10902-019-00128-4}
-
-\leavevmode\hypertarget{ref-choiMediatedCommunicationMatters2021}{}%
-Choi, M., \& Choung, H.
-(2021).
-Mediated communication matters during the {COVID}-19 pandemic: {The} use of interpersonal and masspersonal media and psychological well-being.
-\emph{Journal of Social and Personal Relationships}, \emph{38}(8), 2397--2418.
-\url{https://doi.org/10.1177/02654075211029378}
-
-\leavevmode\hypertarget{ref-cohenPowerPrimer1992}{}%
-Cohen, J. (1992). A power primer. \emph{Psychological Bulletin}, \emph{112}(1), 155--159. \url{https://doi.org/10.1037/0033-2909.112.1.155}
-
-\leavevmode\hypertarget{ref-dienerAdvancesOpenQuestions2018}{}%
-Diener, E., Lucas, R.
-E., \& Oishi, S.
-(2018).
-Advances and open questions in the science of subjective well-being.
-\emph{Collabra: Psychology}, \emph{4}(1), 15.
-\url{https://doi.org/10.1525/collabra.115}
-
-\leavevmode\hypertarget{ref-dienesUsingBayesGet2014}{}%
-Dienes, Z. (2014). Using {Bayes} to get the most out of non-significant results. \emph{Frontiers in Psychology}, \emph{5}. \url{https://doi.org/10.3389/fpsyg.2014.00781}
-
-\leavevmode\hypertarget{ref-dienlinImpactDigitalTechnology2020}{}%
-Dienlin, T., \& Johannes, N.
-(2020).
-The impact of digital technology use on adolescent well-being.
-\emph{Dialogues in Clinical Neuroscience}, \emph{22}(2), 135--142.
-\url{https://doi.org/doi:10.31887/DCNS.2020.22.2/tdienlin}
-
-\leavevmode\hypertarget{ref-dornemannHowGoodBad2021}{}%
-Dörnemann, A., Boenisch, N., Schommer, L., Winkelhorst, L., \& Wingen, T.
-(2021).
-\emph{How do good and bad news impact mood during the {Covid}-19 pandemic?
-{The} role of similarity}.
-OSF Preprints.
-\url{https://doi.org/10.31219/osf.io/sy2kd}
-
-\leavevmode\hypertarget{ref-edenMediaCopingCOVID192020}{}%
-Eden, A.
-L., Johnson, B.
-K., Reinecke, L., \& Grady, S.
-M.
-(2020).
-Media for coping during {COVID}-19 social distancing: {Stress}, anxiety, and psychological well-being.
-\emph{Frontiers in Psychology}, \emph{11}, 577639.
-\url{https://doi.org/10.3389/fpsyg.2020.577639}
-
-\leavevmode\hypertarget{ref-fergusonThisMetaanalysisScreen2021}{}%
-Ferguson, C.
-J., Kaye, L.
-K., Branley-Bell, D., Markey, P., Ivory, J.
-D., Klisanin, D., \ldots{} Wilson, J.
-(2021).
-Like this meta-analysis: {Screen} media and mental health.
-\emph{Professional Psychology: Research and Practice}.
-\url{https://doi.org/10.1037/pro0000426}
-
-\leavevmode\hypertarget{ref-funderEvaluatingEffectSize2019}{}%
-Funder, D.
-C., \& Ozer, D.
-J.
-(2019).
-Evaluating effect size in psychological research: {Sense} and nonsense.
-\emph{Advances in Methods and Practices in Psychological Science}, \emph{2}(2), 156--168.
-\url{https://doi.org/10.1177/2515245919847202}
-
-\leavevmode\hypertarget{ref-greenspoonIntegrationSubjectiveWellbeing2001}{}%
-Greenspoon, P.
-J., \& Saklofske, D.
-H.
-(2001).
-Toward an integration of subjective well-being and psychopathology.
-\emph{Social Indicators Research}, \emph{54}(1), 81--108.
-\url{https://doi.org/10.1023/A:1007219227883}
-
-\leavevmode\hypertarget{ref-hamakerWhyResearchersShould2014}{}%
-Hamaker, E.
-L.
-(2014).
-Why researchers should think "within-person": {A} paradigmatic rationale.
-In M.
-R.
-Mehl, T.
-S.
-Conner, \& M.
-Csikszentmihalyi (Eds.), \emph{Handbook of research methods for studying daily life} (Paperback ed.).
-New York, NY: Guilford.
-
-\leavevmode\hypertarget{ref-huangTimeSpentSocial2017}{}%
-Huang, C.
-(2017).
-Time spent on social network sites and psychological well-being: {A} meta-analysis.
-\emph{Cyberpsychology, Behavior and Social Networking}, \emph{20}(6), 346--354.
-\url{https://doi.org/10.1089/cyber.2016.0758}
-
-\leavevmode\hypertarget{ref-kerestesAdolescentsOnlineSocial2020}{}%
-Keresteš, G., \& Štulhofer, A.
-(2020).
-Adolescents' online social network use and life satisfaction: {A} latent growth curve modeling approach.
-\emph{Computers in Human Behavior}, \emph{104}, 106187.
-\url{https://doi.org/10.1016/j.chb.2019.106187}
-
-\leavevmode\hypertarget{ref-kittelAustrianCoronaPanel2020}{}%
-Kittel, B., Kritzinger, S., Boomgaarden, H., Prainsack, B., Eberl, J.-M., Kalleitner, F., \ldots{} Schlogl, L.
-(2020).
-\emph{Austrian {Corona} {Panel} {Project} ({SUF} edition)}.
-AUSSDA.
-\url{https://doi.org/10.11587/28KQNS}
-
-\leavevmode\hypertarget{ref-kittelAustrianCoronaPanel2021}{}%
-Kittel, B., Kritzinger, S., Boomgaarden, H., Prainsack, B., Eberl, J.-M., Kalleitner, F., \ldots{} Schlogl, L.
-(2021).
-The {Austrian} {Corona} {Panel} {Project}: Monitoring individual and societal dynamics amidst the {COVID}-19 crisis.
-\emph{European Political Science}, \emph{20}(2), 318--344.
-\url{https://doi.org/10.1057/s41304-020-00294-7}
-
-\leavevmode\hypertarget{ref-kleinDarklySoothingCompulsion2021}{}%
-Klein, J. (2021). \emph{The darkly soothing compulsion of 'doomscrolling'}. Retrieved from \url{https://www.bbc.com/worklife/article/20210226-the-darkly-soothing-compulsion-of-doomscrolling}
-
-\leavevmode\hypertarget{ref-klinePrinciplesPracticeStructural2016}{}%
-Kline, R. B. (2016). \emph{Principles and practice of structural equation modeling} (4th ed.). New York, NY: The Guilford Press.
-
-\leavevmode\hypertarget{ref-liuRelationOfficialWhatsAppdistributed2020}{}%
-Liu, J.
-C.
-J., \& Tong, E.
-M.
-W.
-(2020).
-The relation between official {WhatsApp}-distributed {COVID}-19 news exposure and psychological symptoms: Cross-sectional survey study.
-\emph{Journal of Medical Internet Research}, \emph{22}(9), e22142.
-\url{https://doi.org/10.2196/22142}
-
-\leavevmode\hypertarget{ref-mcelreathYesterdayClass2021}{}%
-McElreath, R. (2021). Yesterday in class, ... {[}Tweet{]}. Retrieved from \url{https://twitter.com/rlmcelreath/status/1354786005996482563}
-
-\leavevmode\hypertarget{ref-meierComputermediatedCommunicationSocial2020a}{}%
-Meier, A., \& Reinecke, L.
-(2020).
-Computer-mediated communication, social media, and mental health: {A} conceptual and empirical meta-review.
-\emph{Communication Research}, 009365022095822.
-\url{https://doi.org/10.1177/0093650220958224}
-
-\leavevmode\hypertarget{ref-normanInterpretationChangesHealthrelated2003}{}%
-Norman, G., Sloan, J., \& Wyrwich, K.
-(2003).
-Interpretation of changes in health-related quality of life: {The} remarkable universality of half a standard deviation.
-\emph{Medical Care}, \emph{41}(5), 582--592.
-Retrieved from \url{Retrieved\%20from\%20http://www.jstor.org/stable/3768017}
-
-\leavevmode\hypertarget{ref-orbenTeenagersScreensSocial2020}{}%
-Orben, A.
-(2020).
-Teenagers, screens and social media: A narrative review of reviews and key studies.
-\emph{Social Psychiatry and Psychiatric Epidemiology}, \emph{55}(4), 407--414.
-\url{https://doi.org/10.1007/s00127-019-01825-4}
-
-\leavevmode\hypertarget{ref-orbenSocialMediaEnduring2019}{}%
-Orben, A., Dienlin, T., \& Przybylski, A.
-K.
-(2019).
-Social media's enduring effect on adolescent life satisfaction.
-\emph{Proceedings of the National Academy of Sciences of the United States of America}, \emph{116}(21), 10226--10228.
-\url{https://doi.org/10.1073/pnas.1902058116}
-
-\leavevmode\hypertarget{ref-przybylskiDoesTakingShort2021a}{}%
-Przybylski, A.
-K., Nguyen, T.
-T., Law, W., \& Weinstein, N.
-(2021).
-Does taking a short break from social media have a positive effect on well-being?
-{Evidence} from three preregistered field experiments.
-\emph{Journal of Technology in Behavioral Science}, \emph{6}(3), 507--514.
-\url{https://doi.org/10.1007/s41347-020-00189-w}
-
-\leavevmode\hypertarget{ref-przybylskiLargescaleTestGoldilocks2017}{}%
-Przybylski, A. K., \& Weinstein, N. (2017). A large-scale test of the {Goldilocks} hypothesis. \emph{Psychological Science}, \emph{28}(2), 204--215. \url{https://doi.org/10.1177/0956797616678438}
-
-\leavevmode\hypertarget{ref-riehmAssociationsMediaExposure2020}{}%
-Riehm, K.
-E., Holingue, C., Kalb, L.
-G., Bennett, D., Kapteyn, A., Jiang, Q., \ldots{} Thrul, J.
-(2020).
-Associations between media exposure and mental distress among {U}.{S}.
-Adults at the beginning of the {COVID}-19 pandemic.
-\emph{American Journal of Preventive Medicine}, \emph{59}(5), 630--638.
-\url{https://doi.org/10.1016/j.amepre.2020.06.008}
-
-\leavevmode\hypertarget{ref-rohrerThinkingClearlyCorrelations2018}{}%
-Rohrer, J.
-M.
-(2018).
-Thinking clearly about correlations and causation: {Graphical} causal models for observational data.
-\emph{Advances in Methods and Practices in Psychological Science}, \emph{24}(2), 251524591774562.
-\url{https://doi.org/10.1177/2515245917745629}
-
-\leavevmode\hypertarget{ref-rohrerTheseAreNot2021}{}%
-Rohrer, J.
-M., \& Murayama, K.
-(2021).
-\emph{These are not the effects you are looking for: {Causality} and the within-/between-person distinction in longitudinal data analysis} {[}Preprint{]}.
-PsyArXiv.
-\url{https://doi.org/10.31234/osf.io/tg4vj}
-
-\leavevmode\hypertarget{ref-sandstromDoomscrollingCOVIDNews2021}{}%
-Sandstrom, G., Buchanan, K., Aknin, L., \& Lotun, S.
-(2021).
-Doomscrolling {COVID} news takes an emotional toll -- here's how to make your social media a happier place.
-Retrieved from \url{http://theconversation.com/doomscrolling-covid-news-takes-an-emotional-toll-heres-how-to-make-your-social-media-a-happier-place-170342}
-
-\leavevmode\hypertarget{ref-scharkowAccuracySelfreportedInternet2016}{}%
-Scharkow, M.
-(2016).
-The accuracy of self-reported {Internet} use---{A} validation study using client log data.
-\emph{Communication Methods and Measures}, \emph{10}(1), 13--27.
-\url{https://doi.org/10.1080/19312458.2015.1118446}
-
-\leavevmode\hypertarget{ref-schemerImpactInternetSocial2021}{}%
-Schemer, C., Masur, P.
-K., Geiß, S., Müller, P., \& Schäfer, S.
-(2021).
-The {Impact} of {Internet} and {Social} {Media} {Use} on {Well}-{Being}: {A} {Longitudinal} {Analysis} of {Adolescents} {Across} {Nine} {Years}.
-\emph{Journal of Computer-Mediated Communication}, \emph{26}(1), 1--21.
-\url{https://doi.org/10.1093/jcmc/zmaa014}
-
-\leavevmode\hypertarget{ref-schnauber-stockmannMobileDevicesTools2020}{}%
-Schnauber-Stockmann, A., \& Karnowski, V.
-(2020).
-Mobile devices as tools for media and communication research: {A} scoping review on collecting self-report data in repeated measurement designs.
-\emph{Communication Methods and Measures}, \emph{14}(3), 145--164.
-\url{https://doi.org/10.1080/19312458.2020.1784402}
-
-\leavevmode\hypertarget{ref-sewallObjectivelyMeasuredDigital2021}{}%
-Sewall, C.
-J.
-R., Goldstein, T.
-R., \& Rosen, D.
-(2021).
-Objectively measured digital technology use during the {COVID}-19 pandemic: {Impact} on depression, anxiety, and suicidal ideation among young adults.
-\emph{Journal of Affective Disorders}, \emph{288}, 145--147.
-\url{https://doi.org/10.1016/j.jad.2021.04.008}
-
-\leavevmode\hypertarget{ref-stainbackCOVID1924News2020}{}%
-Stainback, K., Hearne, B.
-N., \& Trieu, M.
-M.
-(2020).
-{COVID}-19 and the 24/7 {News} {Cycle}: {Does} {COVID}-19 {News} {Exposure} {Affect} {Mental} {Health}?
-\emph{Socius}, \emph{6}, 2378023120969339.
-\url{https://doi.org/10.1177/2378023120969339}
-
-\leavevmode\hypertarget{ref-statistaAverageDailyTime2021}{}%
-Statista.
-(2021).
-\emph{Average daily time spent on social networks by users in the {United} {States} from 2018 to 2022}.
-Retrieved from \url{https://www.statista.com/statistics/1018324/us-users-daily-social-media-minutes/}
-
-\leavevmode\hypertarget{ref-waterlooNormsOnlineExpressions2018}{}%
-Waterloo, S.
-F., Baumgartner, S.
-E., Peter, J., \& Valkenburg, P.
-M.
-(2018).
-Norms of online expressions of emotion: {Comparing} {Facebook}, {Twitter}, {Instagram}, and {WhatsApp}.
-\emph{New Media \& Society}, \emph{20}(5), 1813--1831.
-\url{https://doi.org/10.1177/1461444817707349}
-
-\end{CSLReferences}
-
 \newpage

### 2022-04-12 manuscript.Rmd
Implementing updated to manuscript

@@
 However, because social media are particularly engaging, some find it hard to disconnect.
-In this preregistered study, I investigate whether using social media for COVID-19 related reasons affects psychological well-being.
+In this preregistered study, I investigated whether using social media for COVID-19 related reasons affected psychological well-being.
 To answer this question I analyzed data from the Austrian Corona Panel Project, which consists of 3,018 participants.
-Well-being was measured at each wave, and communication at five waves.
+Well-being was measured at all 24 waves, and communication at five specific waves.
 I ran three random effects within between models, controlling for several stable and varying confounders.
 Results showed that the effects of COVID-19 related social media use on well-being were very small, arguably too small to matter.
-Fears that social media use during times of crisis critically impairs well-being are likely to be unfounded."
+Fears that social media use during times of crisis critically impairs well-being are not supported."
 keywords          : "COVID-19, well-being, social media, news use, panel study."
@@
 How is it transmitted, and how can I protect myself?
-Because for many it was (and still is) a matter of life or death, people aimed to stay informed regarding the latest developments.
+Because for many it was (and at the time of writing still is) a matter of life or death, people aimed to stay informed regarding the latest developments.
 Governments around the world implemented safety measures, such as wearing masks, keeping physical distance, or enforcing lockdowns.
-In this extraordinary situation, many people used media excessively to attain information, and especially social media were at an all time high [@statistaAverageDailyTime2021].
+In this extraordinary situation, many people heavily relied on media to attain relevant information, and especially social media were at an all time high [@statistaAverageDailyTime2021].
 
@@
 Whereas the first four levels focus on the communication _channel_, the last two address the communication _type_.
-To measure social media use for the consumption of COVID-19 related news and topics, I here employ both the channel and the type of communication perspective, which together provides a nuanced understanding of communication.
+To measure social media use for the consumption of COVID-19 related news and topics, I here employ both the channel and the type of communication perspective, which together should offer a nuanced understanding of communication.
 
 First, I investigate how well-being is affected by different types of communication affect, namely active and passive use.
-Defining what constitutes active and what passive use is not always clear, and different understandings are currently discussed [@ellisonWhyWeDon2020, @meierComputermediatedCommunicationSocial2020a].
+Defining what constitutes active and what passive use is not always clear, and different understandings are currently discussed [@ellisonWhyWeDon2020; @meierComputermediatedCommunicationSocial2020a].
 Reading is generally considered as passive and writing as active, while there are also specific behaviors such as such liking or sharing content that fall somewhere in-between [@meierDoesPassiveSocial2022].
@@
 
-From a theoretical perspective, how could we explain whether COVID-19 related social media use might affect well-being?
+From a theoretical perspective, how could we explain whether COVID-19 related social media use affects well-being?
 <!-- In what follows, I outline potential arguments as to why the effect might be positive or negative, direct or indirect, or nonexistent. -->
@@
 Although specific events such as marriage or salary can have significant effects, after some time well-being routinely returns to prior levels, which are mostly determined genetically [@sheldonStabilityHappinessTheories2014].
-Only very specific events such as unemployment, disability, or death can cause long-term decreases in well-being [@lucasAdaptationSetpointModel2007].
-So although well-being can change this does not happen easily.
+Only very specific events and factors such as unemployment, disability, or death can cause long-term changes in well-being [@lucasAdaptationSetpointModel2007].
+<!-- So although well-being can change this does not happen easily. -->
 <!-- Although the set-point theory has been criticized lately because events such as marriage, unemployment, or death can indeed lead to long-term changes, it is still widely maintained that well-being is very stable and hence hard to be influenced  -->
 
-Can media use be such a negative or positive factor?
+Can media use be such a factor?
 In advance, there doesn't seem to be a clear winner, and it's likely that both positive and negative effects cancel each other out.
-Empirically, social media use---on average---does not have a strong effect on well-being [@meierComputermediatedCommunicationSocial2020a].
+Empirically, social media use on average oes not have a strong effect on well-being [@meierComputermediatedCommunicationSocial2020a].
 According to the Differential Susceptibility to Media Effects Model [@valkenburgDifferentialSusceptibilityMedia2013], the effects of media use differ across individuals. 
@@
 On average, however, effects are often small or negligible.
-Also when focusing on individual users, social media have both positive and negative effects on well-being [@buchiDigitalWellbeingTheory2021].
+But also when focusing on individual users, social media can have both positive and negative effects on well-being [@buchiDigitalWellbeingTheory2021].
 They can impair well-being when causing embarrassment, stress, or disinformation, and they can improve well-being when providing connectedness, information, or entertainment [@buchiDigitalWellbeingTheory2021].
@@
 First, uses and gratifications theory states that people explicitly and rationally chose specific media because of their respective benefits. 
-If those benefits don't exist, they invest their time elsewhere.
-And social media offer ample benefits. 
+If those benefits don't exist, they will spend their time elsewhere.
+And social media offer copiuos benefits. 
 Most prominently, they help find relevant information, maintain and foster relationships, express one's personality, and entertain oneself [@pelletierOneSizeDoesn2020].
-The second major theory is mood management theory [@zillmannMoodManagementCommunication1988].
+The second major theory that needs to be listed here is mood management theory [@zillmannMoodManagementCommunication1988].
 Users implicitly learn what media help them balance their mood and affect.
 For example, when bored people use social media to entertain themselves.
-Using experience sampling of well-being and logs of social media use, a study with 82 participants from Italy found that after episodes of social media use, levels of positive affect increased for a short time [@marcianoDynamicsAdolescentsSmartphone2022].
+Using experience sampling of well-being and logs of social media use, a study with 82 participants from Italy found that after episodes of social media use, levels of positive affect increased significantly [@marcianoDynamicsAdolescentsSmartphone2022].
 
-But people also misjudge media affects, often being overly optimistic [@metzgerComparativeOptimismPrivacy2017].
-And precisely because social media have so many positive consequences, one can ask if this where the actual problem lies.
+But people can also misjudge media affects and are often overly optimistic [@metzgerComparativeOptimismPrivacy2017].
+And precisely because social media have so many positive consequences, one can ask if this is not where the actual problem lies.
 In other words, social media aren't problematic because they're inherently bad, but rather because they're too good.
@@
 However, a recently published meta-analysis found that the two most prominent measures of addiction, the Bergen Facebook Addiction Scale and the Bergen Social Media Addiction Scale, have only small relations to well-being [@duradoniWellbeingSocialMedia2020].
-In addition, the general idea of labeling excessive social and new media use as addiction was criticized, arguing that social and new media represent new regular behaviors that should not be pathologized [@galerHowMuchToo2018; @vanrooijWeakScientificBasis2018].
+In addition, the general idea of labeling excessive social and new media use as addiction was criticized, arguing that social media represent new regular behaviors that should not be pathologized [@galerHowMuchToo2018; @vanrooijWeakScientificBasis2018].
 
-Because effects can differ across situations, I now briefly focus on the effects of COVID-19 related social media use specifically.
+Because media effects can differ across situations [@valkenburgDifferentialSusceptibilityMedia2013], I now briefly focus on the effects of COVID-19 related social media use specifically.
 First, one could assume a _direct_ negative effect on well-being, and especially on positive or negative affect, which are more volatile and fluctuating.
@@
 During the pandemic social media use was at an all-time high in the US [@statistaAverageDailyTime2021].
-It is most likely that moderate social media use is not detrimental [@orbenTeenagersScreensSocial2020].
-Overuse, however, might be more critical, and several studies have shown more pronounced negative effects for extreme users [@przybylskiLargescaleTestGoldilocks2017].
+Although it is most likely that moderate social media use is not detrimental [@orbenTeenagersScreensSocial2020], overuse, however, might be more critical, and several studies have shown more pronounced negative effects for extreme users [@przybylskiLargescaleTestGoldilocks2017].
 To explain, overuse could impair well-being if it replaces meaningful or functional activities such as meeting others, working, actively relaxing, or exercising.
@@
 <!-- If overuse replaces such activities it's reasonable to assume that it's also detrimental. -->
-On the other hand, one can make the case that overuse might also be beneficial, especially in times of a pandemic---even if the use is mainly COVID-19 related.
+On the other hand, one can make the case that overuse might even be beneficial, especially in times of a pandemic.
 Exchanging COVID-19 related messages with friends via WhatsApp might replace the in-person contact one would have otherwise, but which is literally impossible at the time.
@@
 To summarize, from a theoretical perspective it is most likely that the average effects of social media use on well-being are negligible. 
-Building on established theories from Communication, it is not particularly likely that effects are either profoundly negative or strongly positive.
+Building on established theories from Communication, we would not assume that effects are either profoundly negative or strongly positive.
 
@@
 
-Several individual studies employing advanced methods found smalls relations between social media use and well-being [@orbenSocialMediaEnduring2019; @przybylskiDoesTakingShort2021a; @schemerImpactInternetSocial2021; @kerestesAdolescentsOnlineSocial2020].
+Finally, also several individual studies employing advanced methods found smalls relations between social media use and well-being [@orbenSocialMediaEnduring2019; @przybylskiDoesTakingShort2021a; @schemerImpactInternetSocial2021; @kerestesAdolescentsOnlineSocial2020].
 For example, @beyensSocialMediaUse2021 reported that although for some users (roughly one quarter) the effects of social media use on well-being were negative, for almost the same number of users they were positive, while for the rest the effects were neutral.
-This finding is aligned the Differential Susceptibility to Media Effects Model.
-Although there is substantial variation of media effects for individual users, the average effects reported in the literature are often small [@valkenburgDifferentialSusceptibilityMedia2013].
+This finding is aligned the Differential Susceptibility to Media Effects Model: 
+Although there is substantial _variation_ of media effects for individual users, the _average_ effects reported in the literature are often small [@valkenburgDifferentialSusceptibilityMedia2013].
 
@@
 
-In conclusion, in light of the theoretical considerations and the empirical studies presented above, I expect that COVID-19 related communication on social media doesn't affect well-being in a meaningful or relevant way.
+In conclusion, in light of the theoretical considerations and empirical studies presented above, I expect that COVID-19 related communication on social media doesn't affect well-being in a meaningful or relevant way.
 
@@
 <!-- (I'll additionally include variables that likely affect only well-being, also to obtain a comparison benchmark for social media effects): -->
-gender, age, education, Austria country of birth, Austria country of birth of parents, text-based news consumption, video-based news consumption, residency Vienna, household size, health, living space, access to garden, access to balcony, employment, work hours per week, being in home-office, household income, outdoor activities, satisfaction with democracy, disposition to take risks, and locus of control.
+gender, age, education, Austria country of birth, Austria country of birth of parents, text-based news consumption, video-based news consumption, residency Vienna, household size, health, living space, access to garden, access to balcony, employment, work hours per week, being in home-office, household income, outdoor activities, satisfaction with democracy, disposition to take risks, and locus of control.[^explore]
 <!-- I will not control for variables such as trust in institutions or trust in media, because these variables might be influenced by social media use to a meaningful extent. -->
-The dataset includes many other variables that one could also potentially control for, and I invite interested readers to download the and explore potential interesting relationships.
+
+[^explore]: The dataset includes many other variables that one could also potentially control for, and I invite interested readers to download the and explore potential interesting relationships.
 
@@
 
-In this study, I analyze how using social media during the last week affected their positive and negative affect during the last week.
+In this study, I hence analyze how using social media during the last week affected their positive and negative affect during the same week.
 More precisely, if people during the last week used COVID-19 related social media more than the usually do, did they feel better or worse during that week than they usually do?
 Regarding life satisfaction, I implemented a longer interval.
-Namely, if people during the last week used COVID-19 related social media more than the usually do, did this affect if they are now more or less satisfied with their lives than they usually are?
+Namely, if people during the last week used COVID-19 related social media more than the usually do, were they at the end of the week more or less satisfied with their lives than they usually are?
 I hence analyze if when a person changes their social media diet, are there (a) _simultaneous_ changes in their affect and (b) _subsequent_ changes in their life satisfaction?
 When additionally controlling for both stable and varying confounders, we can then be more sure that the effect is indeed causal.
-Similar approaches were implemented by other studies [@johannesNoEffectDifferent2022; @scharkowHowSocialNetwork2020], and they are considered a best practice to analyze causality [@bellFixedRandomEffects2019].
+Similar approaches were implemented by other studies [@johannesNoEffectDifferent2022; @scharkowHowSocialNetwork2020], and they are considered a best practice approach toward analyzing causality [@bellFixedRandomEffects2019].
 
@@
 It contains 26 waves, and at the time of writing the first 24 waves were available for download.
+Between March 2020 and July 2020, the intervals were weekly, and afterward the intervals were monthly.
 Each wave consists of at least 1,500 respondents.

### 2022-04-12 manuscript.Rmd
- last edits before revision

@@
 classoption       : "man"
-# output            :
-#   papaja::apa6_pdf:
-#     latex_engine  : xelatex
-output            : papaja::apa6_docx
+output            :
+  papaja::apa6_pdf:
+    latex_engine  : xelatex
+# output            : papaja::apa6_docx
 header-includes:
@@
 Two of the most prominent media effect theories argue against strong negative impacts.
-First, uses and gratifications theory states that people explicitly and rationally chose specific media because of their respective benefits. 
+First, uses and gratifications theory states that people explicitly and rationally chose specific media because of their respective benefits [@katzUsesGratificationsResearch1973]. 
 If those benefits don't exist, they will spend their time elsewhere.
@@
 If variables are stable, longer intervals are needed; if they fluctuate, shorter intervals.
-In the case of well-being, we need shorter intervals for affect and longer ones for life satisfaction.
+In the case of well-being, we need shorter intervals for affect and longer ones for life satisfaction [@dienlinImpactDigitalTechnology2020].
 Still, choosing the right interval is challenging, because especially short intervals are hard to implement in practice and often require advanced methods such as experience sampling (also known as in situ measurement or ambulant assessment) [@schnauber-stockmannMobileDevicesTools2020].
@@
 Because the data were analyzed post-hoc, no a-priori sample size planning on the basis of power analyses was conducted.
-The sample is large, and it is hence well-equipped reliably to detect also small effects.
+The sample is large, and it is hence well-equipped to reliably detect also small effects.
 <!-- , which is why no exact post hoc power analysis were conducted. -->
@@
 
-First, when looking at the variables from a descriptive perspective (Figure \@ref(fig:descriptives))), we see that all well-being measures did not change substantially across the different waves of data collection.
+First, when looking at the variables from a descriptive perspective (Figure \@ref(fig:descriptives)), we see that all well-being measures did not change substantially across the different waves of data collection.
 COVID-19 related media use, however, decreased slightly at the beginning of the study and remained stable after approximately six waves.
@@
 The results were virtually the same.
-For example, all standardized COVID-19 related types of social media use or channels were not significantly larger than a SESOI of $\beta$ = |.10|.
-The additional analyses are reported in the [companion website](https://XMtRA.github.io/Austrian_Corona_Panel_Project/analyses_additional.html).
+For example, all within-person standardized COVID-19 related types of social media use or channels were significantly smaller than $\beta$ = |.05|, again supporting that effects were tiny.
+The additional analyses are reported in Figure 3 and Figure 4, and in the [companion website](https://XMtRA.github.io/Austrian_Corona_Panel_Project/analyses_additional.html).
 
@@
 Other factors among the third variables that were measured revealed larger effects or relations, suggesting that well-being is determined by alternative aspects such as health, satisfaction with democracy, locus of control, or exercising.
-According to this study, popular fears that "doomscrolling" or overusing social media during times of crises don't seem to be justified.
+According to this study, popular fears that "doomscrolling" or overusing social media during times of crises are not supported.
 
@@
 However, note that @bendauAssociationsCOVID19Related2021 analyzed cross-sectional data on a between-person level while not controlling for third variables, which is not optimal for investigating causal effects.
-On the other hand, the results are well-aligned with recent studies and meta-analyses analyzing the effects of social media use from a more general perspective or from a somewhat different angle. 
+On the other hand, the results are well-aligned with many seminal Communication theories such as the uses and gratifications model [@katzUsesGratificationsResearch1973].
+If effects were indeed profoundly negative on average, then people likely wouldn't spend so much time on social media.
+Likewise, recent studies and meta-analyses analyzing the effects of social media use from a more general perspective or from a somewhat different angle. 
 These studies have found that the effects of various types of social media use on several well-being indicators are small at best, often too small to matter [@meierComputermediatedCommunicationSocial2020a; @orbenTeenagersScreensSocial2020; @fergusonThisMetaanalysisScreen2021], which echoes the results obtained here.
@@
 To speculate, the often-criticized positivity bias on Instagram might have been somewhat beneficial in times of the pandemic.
-The critique that the positivity bias necessarily leads to envy and negative feelings is unrealistic and myopic, because positive content can also inspire and motivate users [@meierInstagramInspirationHow2020].
+The critique that the positivity bias necessarily leads to envy and negative feelings is unrealistic and myopic, because positive content can also inspire and motivate users [@meierInstagramInspirationHow2020], which could be especially helpful in times of lockdown and home-office.
 Second, of all the three COVID-19 related social media activities, people who read about the pandemic more than others showed slightly decreased levels of positive affect, and people who actively posted about the pandemic more than others showed slightly increased levels of negative affect.
@@
 The current study analyzed whether changes in media use were related to changes in well-being, while controlling for several potential confounders. 
-Together, this allows for an improved perspective on assessing causality.
+Together, this allowed for an improved perspective on assessing causality.
 That said, causality necessitates temporal order, and the cause needs to precede the effect.

### 2022-04-12 manuscript.tex
- last edits before revision

@@
 \documentclass[
-  english,
   man,mask]{apa6}
-\title{Analyzing the effects of COVID-19 related social media use on well-being}
-\author{Name blinded for review\textsuperscript{1}}
-\date{}
-
 \usepackage{amsmath,amssymb}
@@
   pdftitle={Analyzing the effects of COVID-19 related social media use on well-being},
-  pdfauthor={Name blinded for review1},
   pdflang={en-EN},
@@
 \newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}
+\ifLuaTeX
+\usepackage[bidi=basic]{babel}
+\else
+\usepackage[bidi=default]{babel}
+\fi
+\babelprovide[main,import]{english}
+% get rid of language-specific shorthands (see #6817):
+\let\LanguageShortHands\languageshorthands
+\def\languageshorthands#1{}
 % Manuscript styling
@@
 % \setlength{\parskip}{0pt plus 0pt minus 0pt}
+
+% Overwrite redefinition of paragraph and subparagraph by the default LaTeX template
+% See https://github.com/crsh/papaja/issues/292
+\makeatletter
+\renewcommand{\paragraph}{\@startsection{paragraph}{4}{\parindent}%
+  {0\baselineskip \@plus 0.2ex \@minus 0.2ex}%
+  {-1em}%
+  {\normalfont\normalsize\bfseries\itshape\typesectitle}}
+
+\renewcommand{\subparagraph}[1]{\@startsection{subparagraph}{5}{1em}%
+  {0\baselineskip \@plus 0.2ex \@minus 0.2ex}%
+  {-\z@\relax}%
+  {\normalfont\normalsize\itshape\hspace{\parindent}{#1}\textit{\addperi}}{\relax}}
+\makeatother
 
@@
 \makeatother
-\shorttitle{COVID-19 related social media use and well-being}
+
+\usepackage{xpatch}
+\makeatletter
+\xapptocmd\appendix
+  {\xapptocmd\section
+    {\addcontentsline{toc}{section}{\appendixname\ifoneappendix\else~\theappendix\fi\\: #1}}
+    {}{\InnerPatchFailed}%
+  }
+{}{\PatchFailed}
 \keywords{COVID-19, well-being, social media, news use, panel study.}
@@
 \note{\clearpage}
-\ifXeTeX
-  % Load polyglossia as late as possible: uses bidi with RTL langages (e.g. Hebrew, Arabic)
-  \usepackage{polyglossia}
-  \setmainlanguage[]{english}
-\else
-  \usepackage[main=english]{babel}
-% get rid of language-specific shorthands (see #6817):
-\let\LanguageShortHands\languageshorthands
-\def\languageshorthands#1{}
-\fi
 \ifLuaTeX
@@
 
+\title{Analyzing the effects of COVID-19 related social media use on well-being}
+\author{Name blinded for review\textsuperscript{1}}
+\date{}
+
+\shorttitle{COVID-19 related social media use and well-being}
+
+\authornote{
+
+Correspondence concerning this article should be addressed to Name blinded for review, . E-mail:
+
+}
+
 \affiliation{\vspace{0.5cm}\textsuperscript{1} }
 
-\abstract{
+\abstract{%
 In times of crisis such as the Corona pandemic citizens need to stay informed about recent events, political decisions, or mandatory protection measures.
@@
 However, because social media are particularly engaging, some find it hard to disconnect.
-In this preregistered study, I investigate whether using social media for COVID-19 related reasons affects psychological well-being.
+In this preregistered study, I investigated whether using social media for COVID-19 related reasons affected psychological well-being.
 To answer this question I analyzed data from the Austrian Corona Panel Project, which consists of 3,018 participants.
-Well-being was measured at each wave, and communication at five waves.
+Well-being was measured at all 24 waves, and communication at five specific waves.
 I ran three random effects within between models, controlling for several stable and varying confounders.
 Results showed that the effects of COVID-19 related social media use on well-being were very small, arguably too small to matter.
-Fears that social media use during times of crisis critically impairs well-being are likely to be unfounded.
+Fears that social media use during times of crisis critically impairs well-being are not supported.
 }
@@
 How is it transmitted, and how can I protect myself?
-Because for many it was (and still is) a matter of life or death, people aimed to stay informed regarding the latest developments.
+Because for many it was (and at the time of writing still is) a matter of life or death, people aimed to stay informed regarding the latest developments.
 Governments around the world implemented safety measures, such as wearing masks, keeping physical distance, or enforcing lockdowns.
-In this extraordinary situation, many people used media excessively to attain information, and especially social media were at an all time high (Statista, 2021).
+In this extraordinary situation, many people heavily relied on media to attain relevant information, and especially social media were at an all time high (Statista, 2021).
 
@@
 Whereas the first four levels focus on the communication \emph{channel}, the last two address the communication \emph{type}.
-To measure social media use for the consumption of COVID-19 related news and topics, I here employ both the channel and the type of communication perspective, which together provides a nuanced understanding of communication.
+To measure social media use for the consumption of COVID-19 related news and topics, I here employ both the channel and the type of communication perspective, which together should offer a nuanced understanding of communication.
 
 First, I investigate how well-being is affected by different types of communication affect, namely active and passive use.
-Defining what constitutes active and what passive use is not always clear, and different understandings are currently discussed Meier \& Reinecke (2020).
+Defining what constitutes active and what passive use is not always clear, and different understandings are currently discussed (Ellison, Triẹû, Schoenebeck, Brewer, \& Israni, 2020; Meier \& Reinecke, 2020).
 Reading is generally considered as passive and writing as active, while there are also specific behaviors such as such liking or sharing content that fall somewhere in-between (Meier \& Krause, 2022).
@@
 
-From a theoretical perspective, how could we explain whether COVID-19 related social media use might affect well-being?
+From a theoretical perspective, how could we explain whether COVID-19 related social media use affects well-being?
 According to the set-point model of subjective well-being, well-being is surprisingly stable (Lykken, 1999).
 Although specific events such as marriage or salary can have significant effects, after some time well-being routinely returns to prior levels, which are mostly determined genetically (Sheldon \& Lucas, 2014).
-Only very specific events such as unemployment, disability, or death can cause long-term decreases in well-being (Lucas, 2007).
-So although well-being can change this does not happen easily.
-
-Can media use be such a negative or positive factor?
+Only very specific events and factors such as unemployment, disability, or death can cause long-term changes in well-being (Lucas, 2007).
+
+Can media use be such a factor?
 In advance, there doesn't seem to be a clear winner, and it's likely that both positive and negative effects cancel each other out.
-Empirically, social media use---on average---does not have a strong effect on well-being (Meier \& Reinecke, 2020).
-According to the different susceptibility of media effects model (Valkenburg \& Peter, 2013), the effects of media use differ across individuals.
+Empirically, social media use on average oes not have a strong effect on well-being (Meier \& Reinecke, 2020).
+According to the Differential Susceptibility to Media Effects Model (Valkenburg \& Peter, 2013), the effects of media use differ across individuals.
 Whereas for some media are beneficial, for others they are harmful.
 On average, however, effects are often small or negligible.
-Also when focusing on individual users, social media have both positive and negative effects on well-being (Büchi, 2021).
+But also when focusing on individual users, social media can have both positive and negative effects on well-being (Büchi, 2021).
 They can impair well-being when causing embarrassment, stress, or disinformation, and they can improve well-being when providing connectedness, information, or entertainment (Büchi, 2021).
@@
 Two of the most prominent media effect theories argue against strong negative impacts.
-First, uses and gratifications theory states that people explicitly and rationally chose specific media because of their respective benefits.
-If those benefits don't exist, they invest their time elsewhere.
-And social media offer ample benefits.
+First, uses and gratifications theory states that people explicitly and rationally chose specific media because of their respective benefits (Katz, Blumler, \& Gurevitch, 1973).
+If those benefits don't exist, they will spend their time elsewhere.
+And social media offer copiuos benefits.
 Most prominently, they help find relevant information, maintain and foster relationships, express one's personality, and entertain oneself (Pelletier, Krallman, Adams, \& Hancock, 2020).
-The second major theory is mood management theory (Zillmann, 1988).
+The second major theory that needs to be listed here is mood management theory (Zillmann, 1988).
 Users implicitly learn what media help them balance their mood and affect.
 For example, when bored people use social media to entertain themselves.
-Using experience sampling of well-being and logs of social media use, a study with 82 participants from Italy found that after episodes of social media use, levels of positive affect increased for a short time (Marciano, Driver, Schulz, \& Camerini, 2022).
-
-Precisely because social media have so many positive consequences, one can ask if this is the main problem.
+Using experience sampling of well-being and logs of social media use, a study with 82 participants from Italy found that after episodes of social media use, levels of positive affect increased significantly (Marciano, Driver, Schulz, \& Camerini, 2022).
+
+But people can also misjudge media affects and are often overly optimistic (Metzger \& Suh, 2017).
+And precisely because social media have so many positive consequences, one can ask if this is not where the actual problem lies.
 In other words, social media aren't problematic because they're inherently bad, but rather because they're too good.
@@
 However, a recently published meta-analysis found that the two most prominent measures of addiction, the Bergen Facebook Addiction Scale and the Bergen Social Media Addiction Scale, have only small relations to well-being (Duradoni, Innocenti, \& Guazzini, 2020).
-In addition, the general idea of labeling excessive social and new media use as addiction was criticized, arguing that social and new media represent new regular behaviors that should not be pathologized (Galer, 2018; van Rooij et al., 2018).
-
-Because effects can differ across situations, I now briefly focus on the effects of COVID-19 related social media use specifically.
+In addition, the general idea of labeling excessive social and new media use as addiction was criticized, arguing that social media represent new regular behaviors that should not be pathologized (Galer, 2018; van Rooij et al., 2018).
+
+Because media effects can differ across situations (Valkenburg \& Peter, 2013), I now briefly focus on the effects of COVID-19 related social media use specifically.
 First, one could assume a \emph{direct} negative effect on well-being, and especially on positive or negative affect, which are more volatile and fluctuating.
@@
 During the pandemic social media use was at an all-time high in the US (Statista, 2021).
-It is most likely that moderate social media use is not detrimental (Orben, 2020).
-Overuse, however, might be more critical, and several studies have shown more pronounced negative effects for extreme users (Przybylski \& Weinstein, 2017).
+Although it is most likely that moderate social media use is not detrimental (Orben, 2020), overuse, however, might be more critical, and several studies have shown more pronounced negative effects for extreme users (Przybylski \& Weinstein, 2017).
 To explain, overuse could impair well-being if it replaces meaningful or functional activities such as meeting others, working, actively relaxing, or exercising.
 So if a society collectively overuses social media during a pandemic, there might be potential for negative effects.
-On the other hand, one can make the case that overuse might also be beneficial, especially in times of a pandemic---even if the use is mainly COVID-19 related.
+On the other hand, one can make the case that overuse might even be beneficial, especially in times of a pandemic.
 Exchanging COVID-19 related messages with friends via WhatsApp might replace the in-person contact one would have otherwise, but which is literally impossible at the time.
@@
 To summarize, from a theoretical perspective it is most likely that the average effects of social media use on well-being are negligible.
-Building on established theories from Communication, it is not particularly likely that effects are either profoundly negative or strongly positive.
+Building on established theories from Communication, we would not assume that effects are either profoundly negative or strongly positive.
 
@@
 
-Several individual studies employing advanced methods found smalls relations between social media use and well-being (Keresteš \& Štulhofer, 2020; Orben, Dienlin, \& Przybylski, 2019; Przybylski, Nguyen, Law, \& Weinstein, 2021; Schemer, Masur, Geiß, Müller, \& Schäfer, 2021).
+Finally, also several individual studies employing advanced methods found smalls relations between social media use and well-being (Keresteš \& Štulhofer, 2020; Orben, Dienlin, \& Przybylski, 2019; Przybylski, Nguyen, Law, \& Weinstein, 2021; Schemer, Masur, Geiß, Müller, \& Schäfer, 2021).
 For example, Beyens, Pouwels, van Driel, Keijsers, and Valkenburg (2021) reported that although for some users (roughly one quarter) the effects of social media use on well-being were negative, for almost the same number of users they were positive, while for the rest the effects were neutral.
-This finding is aligned the Differential Susceptibility to Media Effects Model.
-Although there is substantial variation of media effects for individual users, the average effects reported in the literature are often small (Valkenburg \& Peter, 2013).
-
-In conclusion, in light of the theoretical considerations and the empirical studies presented above, I expect that COVID-19 related communication on social media doesn't affect well-being in a meaningful or relevant way.
+This finding is aligned the Differential Susceptibility to Media Effects Model:
+Although there is substantial \emph{variation} of media effects for individual users, the \emph{average} effects reported in the literature are often small (Valkenburg \& Peter, 2013).
+
+In conclusion, in light of the theoretical considerations and empirical studies presented above, I expect that COVID-19 related communication on social media doesn't affect well-being in a meaningful or relevant way.
 
@@
 In this study, I hence plan to control for the following variables, which either have already been shown to affect both social media use and well-being or which are likely to do so, and which also aren't mediators:
-gender, age, education, Austria country of birth, Austria country of birth of parents, text-based news consumption, video-based news consumption, residency Vienna, household size, health, living space, access to garden, access to balcony, employment, work hours per week, being in home-office, household income, outdoor activities, satisfaction with democracy, disposition to take risks, and locus of control.
-I will not control for variables such as trust in institutions or trust in media, because these variables might be influenced by social media use to a meaningful extent.
+gender, age, education, Austria country of birth, Austria country of birth of parents, text-based news consumption, video-based news consumption, residency Vienna, household size, health, living space, access to garden, access to balcony, employment, work hours per week, being in home-office, household income, outdoor activities, satisfaction with democracy, disposition to take risks, and locus of control.\footnote{The dataset includes many other variables that one could also potentially control for, and I invite interested readers to download the and explore potential interesting relationships.}
 
@@
 If variables are stable, longer intervals are needed; if they fluctuate, shorter intervals.
-In the case of well-being, we need shorter intervals for affect and longer ones for life satisfaction.
+In the case of well-being, we need shorter intervals for affect and longer ones for life satisfaction (Dienlin \& Johannes, 2020).
 Still, choosing the right interval is challenging, because especially short intervals are hard to implement in practice and often require advanced methods such as experience sampling (also known as in situ measurement or ambulant assessment) (Schnauber-Stockmann \& Karnowski, 2020).
 
-In this study, I therefore adopt an intermediate perspective.
-I analyze if when a person changes their social media diet, are there \emph{simultaneous} changes in their well-being?
+In this study, I hence analyze how using social media during the last week affected their positive and negative affect during the same week.
+More precisely, if people during the last week used COVID-19 related social media more than the usually do, did they feel better or worse during that week than they usually do?
+Regarding life satisfaction, I implemented a longer interval.
+Namely, if people during the last week used COVID-19 related social media more than the usually do, were they at the end of the week more or less satisfied with their lives than they usually are?
+I hence analyze if when a person changes their social media diet, are there (a) \emph{simultaneous} changes in their affect and (b) \emph{subsequent} changes in their life satisfaction?
 When additionally controlling for both stable and varying confounders, we can then be more sure that the effect is indeed causal.
+Similar approaches were implemented by other studies (Johannes, Dienlin, Bakhshi, \& Przybylski, 2022; Scharkow, Mangold, Stier, \& Breuer, 2020), and they are considered a best practice approach toward analyzing causality (Bell, Fairbrother, \& Jones, 2019).
 
@@
 Because the data were analyzed post-hoc, no a-priori sample size planning on the basis of power analyses was conducted.
-The sample is large, and it is hence well-equipped reliably to detect also small effects.
+The sample is large, and it is hence well-equipped to reliably detect also small effects.
 In addition, because such large samples easily generate significant \emph{p}-values even for very small effects, it helps that the hypotheses were tested with a smallest effect size of interest-approach.
@@
 
-The hypothesis was analyzed using mixed effects models, namely random effect within-between models (REWB)(Bell, Fairbrother, \& Jones, 2019).
+The hypothesis was analyzed using mixed effects models, namely random effect within-between models (REWB)(Bell et al., 2019).
 Three models were run, one for each dependent variable.
@@
 
-For more information on the analyses, a complete documentation of the models and results, see \href{https://XMtRA.github.io/Austrian_Corona_Panel_Project}{companion website}.
+For more information on the analyses, a complete documentation of the models and results, additional analyses (for example using multiple imputation or no imputation), see \href{https://XMtRA.github.io/Austrian_Corona_Panel_Project}{companion website}.
 
@@
 
-COVID-19 related social media use focused on communication types was measured with the three dimensions of (a) reading, (b) liking and sharing, and (c) posting from Wagner et al.
-(2018), adapted for the context of this study.
+COVID-19 related social media use focused on communication types was measured with the three dimensions of (a) reading, (b) liking and sharing, and (c) posting.
+The items come from Wagner et al. (2018) and were adapted for the context of this study.
 The general introductory question was ``How often during the last week have you engaged in the following activities on social media?''
-The three items were ``Reading the posts of others with content on the Coronavirus,'' ``When seeing posts on the Coronavirus, I clicked `like,' `share' or `retweet','' ``I myself wrote posts on the Coronavirus on social media.''
+The three items were ``Reading the posts of others with content on the Coronavirus'', ``When seeing posts on the Coronavirus, I clicked `like', `share' or `retweet'\,'', ``I myself wrote posts on the Coronavirus on social media.''
 Answer options were 1 (\emph{several times per day}), 2 (\emph{daily}), 3 (\emph{several times per week}), 4 (\emph{weekly}), 5 (\emph{never}).
@@
 
-First, when looking at the variables from a descriptive perspective (Figure \ref{fig:descriptives})), we see that all well-being measures did not change substantially across the different waves of data collection.
+First, when looking at the variables from a descriptive perspective (Figure \ref{fig:descriptives}), we see that all well-being measures did not change substantially across the different waves of data collection.
 COVID-19 related media use, however, decreased slightly at the beginning of the study and remained stable after approximately six waves.
@@
 The results were virtually the same.
-For example, all standardized COVID-19 related types of social media use or channels were not significantly larger than a SESOI of \(\beta\) = \textbar.10\textbar.
-The additional analyses are reported in the \href{https://XMtRA.github.io/Austrian_Corona_Panel_Project/analyses_additional.html}{companion website}.
+For example, all within-person standardized COVID-19 related types of social media use or channels were significantly smaller than \(\beta\) = \textbar.05\textbar, again supporting that effects were tiny.
+The additional analyses are reported in Figure 3 and Figure 4, and in the \href{https://XMtRA.github.io/Austrian_Corona_Panel_Project/analyses_additional.html}{companion website}.
 
@@
 Other factors among the third variables that were measured revealed larger effects or relations, suggesting that well-being is determined by alternative aspects such as health, satisfaction with democracy, locus of control, or exercising.
-According to this study, popular fears that ``doomscrolling'' or overusing social media during times of crises don't seem to be justified.
+According to this study, popular fears that ``doomscrolling'' or overusing social media during times of crises are not supported.
 
@@
 However, note that Bendau et al. (2021) analyzed cross-sectional data on a between-person level while not controlling for third variables, which is not optimal for investigating causal effects.
-On the other hand, the results are well-aligned with recent studies and meta-analyses analyzing the effects of social media use from a more general perspective or from a somewhat different angle.
+On the other hand, the results are well-aligned with many seminal Communication theories such as the uses and gratifications model (Katz et al., 1973).
+If effects were indeed profoundly negative on average, then people likely wouldn't spend so much time on social media.
+Likewise, recent studies and meta-analyses analyzing the effects of social media use from a more general perspective or from a somewhat different angle.
 These studies have found that the effects of various types of social media use on several well-being indicators are small at best, often too small to matter (Ferguson et al., 2021; Meier \& Reinecke, 2020; Orben, 2020), which echoes the results obtained here.
@@
 If anything, two preliminary and subtle trends can be observed.
-First, of all the three COVID-19 related social media activities, people who read about the pandemic more than others showed slightly decreased levels of positive affect, and people who actively posted about the pandemic more than others showed slightly increased levels of negative affect.
-On the other hand, however, people who posted more about COVID-19 also showed slightly higher levels of positive affects, so taken together the results are ambivalent.
-Second, in terms of media channels, using Twitter more than usual was related to slightly decreased levels of life satisfaction.
+First, in terms of media channels, using Twitter more than usual was related to slightly decreased levels of life satisfaction.
 Twitter is considered to have more negative affordances and tonality as compared to other networks such as Instagram (Waterloo et al., 2018), which might help explain the results.
@@
 To speculate, the often-criticized positivity bias on Instagram might have been somewhat beneficial in times of the pandemic.
-That said, all these effects were still very small and arguably too small to matter.
+The critique that the positivity bias necessarily leads to envy and negative feelings is unrealistic and myopic, because positive content can also inspire and motivate users (Meier, Gilbert, Börner, \& Possler, 2020), which could be especially helpful in times of lockdown and home-office.
+Second, of all the three COVID-19 related social media activities, people who read about the pandemic more than others showed slightly decreased levels of positive affect, and people who actively posted about the pandemic more than others showed slightly increased levels of negative affect.
+On the other hand, however, people who posted more about COVID-19 also showed slightly higher levels of positive affects, so taken together the results are ambivalent.
 But future research might elaborate on these specific relations to probe their stability and relevance.
-
-Finally, another interesting observation is that life satisfaction was remarkably stable.
-Hence, even in times of a pandemic, it seems that such broad assessment of life vary only mildly.
-This supports the hypothesis that life satisfaction seems to be determined largely by stable factors such as one's genes (Brown \& Rohrer, 2019).
 
@@
 The current study analyzed whether changes in media use were related to changes in well-being, while controlling for several potential confounders.
-Together, this allows for an improved perspective on assessing causality.
+Together, this allowed for an improved perspective on assessing causality.
 That said, causality necessitates temporal order, and the cause needs to precede the effect.
@@
 Measuring well-being with self-reports is adequate, because it by definition requires introspection.
-However, it would be preferable to measure social media use objectively, because people cannot reliably estimate their use (Scharkow, 2016).
+However, it would be preferable to measure social media use objectively, as people cannot reliably estimate their use (Scharkow, 2016).
 That said, objective measures often cannot capture the content or the motivation of the use, and only very complicated tools recording the actual content (such as the Screenome project) might produce such data.
 Unfortunately, such procedures introduce other problems, especially related to privacy.
-Hence, for this type of research question it still seems necessary to use self-reported measures.
+Hence, for this type of research question it still seems necessary to use self-reported measures, and in many cases they can still be very informative (Verbeij, Pouwels, Beyens, \& Valkenburg, 2021).
 
@@
 \leavevmode\vadjust pre{\hypertarget{ref-bradleyStressMoodSmartphone2021}{}}%
-Bradley, A., \& Howard, A. (2021). Stress, mood, and smartphone use in {University} students: {A} 12-week longitudinal study. \url{https://doi.org/10.31219/osf.io/frvpb}
-
-\leavevmode\vadjust pre{\hypertarget{ref-brownEasyHappinessPie2019}{}}%
-Brown, N.
-J.
-L., \& Rohrer, J.
-M.
-(2019).
-Easy as (happiness) pie?
-{A} critical evaluation of a popular model of the determinants of well-being.
-\emph{Journal of Happiness Studies}.
-\url{https://doi.org/10.1007/s10902-019-00128-4}
+Bradley, A., \& Howard, A. (2021). \emph{Stress, mood, and smartphone use in {University} students: {A} 12-week longitudinal study}. {OSF Preprints}. \url{https://doi.org/10.31219/osf.io/frvpb}
 
@@
 (2021).
-How do good and bad news impact mood during the {Covid-19} pandemic?
-{The} role of similarity.
+\emph{How do good and bad news impact mood during the {Covid-19} pandemic?
+{The} role of similarity}.
+{OSF Preprints}.
 \url{https://doi.org/10.31219/osf.io/sy2kd}
@@
 Well-being and social media: {A} systematic review of {Bergen Addiction Scales}.
-\emph{Future Internet}, \emph{12}(2, 2), 24.
+\emph{Future Internet}, \emph{12}(2), 24.
 \url{https://doi.org/10.3390/fi12020024}
@@
 Ellison, N.
-B., Triệu, P., Schoenebeck, S., Brewer, R., \& Israni, A.
+B., Triẹû, P., Schoenebeck, S., Brewer, R., \& Israni, A.
 (2020).
@@
 \leavevmode\vadjust pre{\hypertarget{ref-europeansocialsurveyESS9Edition20182021}{}}%
-European Social Survey. (2021). {ESS9} edition 3.1 - 2018 {Documentation Report}. Retrieved from \url{https://www.europeansocialsurvey.org/docs/round9/survey/ESS9_data_documentation_report_e03_1.pdf}
+European Social Survey. (2021). \emph{{ESS9} edition 3.1 - 2018 {Documentation Report}}.
 
@@
 \leavevmode\vadjust pre{\hypertarget{ref-galerHowMuchToo2018}{}}%
-Galer, S.
-S.
-(2018, January 19).
-How much is {`too much time'} on social media?
-Retrieved April 4, 2022, from \url{https://www.bbc.com/future/article/20180118-how-much-is-too-much-time-on-social-media}
+Galer, S. S. (2018). \emph{How much is {``too much time''} on social media?} https://www.bbc.com/future/article/20180118-how-much-is-too-much-time-on-social-media.
 
@@
 
+\leavevmode\vadjust pre{\hypertarget{ref-johannesNoEffectDifferent2022}{}}%
+Johannes, N., Dienlin, T., Bakhshi, H., \& Przybylski, A.
+K.
+(2022).
+No effect of different types of media on well-being.
+\emph{Scientific Reports}, \emph{12}(1), 61.
+\url{https://doi.org/10.1038/s41598-021-03218-7}
+
+\leavevmode\vadjust pre{\hypertarget{ref-katzUsesGratificationsResearch1973}{}}%
+Katz, E., Blumler, J. G., \& Gurevitch, M. (1973). Uses and {Gratifications Research}. \emph{Public Opinion Quarterly}, \emph{37}(4), 509. \url{https://doi.org/10.1086/268109}
+
 \leavevmode\vadjust pre{\hypertarget{ref-kerestesAdolescentsOnlineSocial2020}{}}%
@@
 (2020).
-\emph{Austrian {Corona Panel Project} ({SUF} edition)} {[}Data set{]}.
+\emph{Austrian {Corona Panel Project} ({SUF} edition)}.
 {AUSSDA}.
@@
 \leavevmode\vadjust pre{\hypertarget{ref-kleinDarklySoothingCompulsion2021}{}}%
-Klein, J.
-(2021, March 3).
-The darkly soothing compulsion of 'doomscrolling'.
-Retrieved November 1, 2021, from \url{https://www.bbc.com/worklife/article/20210226-the-darkly-soothing-compulsion-of-doomscrolling}
+Klein, J. (2021). \emph{The darkly soothing compulsion of 'doomscrolling'}. https://www.bbc.com/worklife/article/20210226-the-darkly-soothing-compulsion-of-doomscrolling.
 
 \leavevmode\vadjust pre{\hypertarget{ref-klinePrinciplesPracticeStructural2016}{}}%
-Kline, R. B. (2016). \emph{Principles and practice of structural equation modeling} (4th ed.). {New York, NY}: {The Guilford Press}.
+Kline, R. B. (2016). \emph{Principles and practice of structural equation modeling} (Fourth). {New York, NY}: {The Guilford Press}.
 
@@
 \leavevmode\vadjust pre{\hypertarget{ref-mcelreathYesterdayClass2021}{}}%
-McElreath, R. (2021, January 28). Yesterday in class, ... {[}Tweet{]}. Retrieved November 5, 2021, from \url{https://twitter.com/rlmcelreath/status/1354786005996482563}
+McElreath, R. (2021). Yesterday in class, ... {[}Tweet{]}.
+
+\leavevmode\vadjust pre{\hypertarget{ref-meierInstagramInspirationHow2020}{}}%
+Meier, A., Gilbert, A., Börner, S., \& Possler, D.
+(2020).
+Instagram inspiration: {How} upward comparison on social network sites can contribute to well-being.
+\emph{Journal of Communication}, \emph{70}(5), 721--743.
+\url{https://doi.org/10.1093/joc/jqaa025}
 
 \leavevmode\vadjust pre{\hypertarget{ref-meierDoesPassiveSocial2022}{}}%
-Meier, A., \& Krause, H.-V. (2022). \emph{Does passive social media use harm well-being? {An} adversarial review} (preprint). {PsyArXiv}. \url{https://doi.org/10.31234/osf.io/nvbwh}
+Meier, A., \& Krause, H.-V. (2022). \emph{Does passive social media use harm well-being? {An} adversarial review} {[}Preprint{]}. {PsyArXiv}. \url{https://doi.org/10.31234/osf.io/nvbwh}
 
@@
 \url{https://doi.org/10.1177/0093650220958224}
+
+\leavevmode\vadjust pre{\hypertarget{ref-metzgerComparativeOptimismPrivacy2017}{}}%
+Metzger, M. J., \& Suh, J. J. (2017). Comparative optimism about privacy risks on {Facebook}. \emph{Journal of Communication}, \emph{67}(2), 203--232. \url{https://doi.org/10.1111/jcom.12290}
 
@@
 \emph{Medical Care}, \emph{41}(5), 582--592.
-Retrieved from \url{Retrieved\%20from\%20http://www.jstor.org/stable/3768017}
 
@@
 (2021).
-\emph{These are not the effects you are looking for: {Causality} and the within-/between-person distinction in longitudinal data analysis} (preprint).
+\emph{These are not the effects you are looking for: {Causality} and the within-/between-person distinction in longitudinal data analysis} {[}Preprint{]}.
 {PsyArXiv}.
@@
 Sandstrom, G., Buchanan, K., Aknin, L., \& Lotun, S.
-(2021, October 22).
-Doomscrolling {COVID} news takes an emotional toll -- here's how to make your social media a happier place.
-Retrieved November 3, 2021, from \url{http://theconversation.com/doomscrolling-covid-news-takes-an-emotional-toll-heres-how-to-make-your-social-media-a-happier-place-170342}
+(2021).
+Doomscrolling {COVID} news takes an emotional toll \textendash{} here's how to make your social media a happier place.
+http://theconversation.com/doomscrolling-covid-news-takes-an-emotional-toll-heres-how-to-make-your-social-media-a-happier-place-170342.
 
@@
 (2016).
-The accuracy of self-reported {Internet} use---{A} validation study using client log data.
+The accuracy of self-reported {Internet} use\textemdash{{A}} validation study using client log data.
 \emph{Communication Methods and Measures}, \emph{10}(1), 13--27.
 \url{https://doi.org/10.1080/19312458.2015.1118446}
+
+\leavevmode\vadjust pre{\hypertarget{ref-scharkowHowSocialNetwork2020}{}}%
+Scharkow, M., Mangold, F., Stier, S., \& Breuer, J.
+(2020).
+How social network sites and other online intermediaries increase exposure to news.
+\emph{Proceedings of the National Academy of Sciences}, \emph{117}(6), 2761--2763.
+\url{https://doi.org/10.1073/pnas.1918279117}
 
@@
 \leavevmode\vadjust pre{\hypertarget{ref-sheldonStabilityHappinessTheories2014}{}}%
-Sheldon, K. M., \& Lucas, R. E. (2014). \emph{Stability of happiness: Theories and evidence on whether happiness can change}. Retrieved from \url{http://site.ebrary.com/id/10891875}
+Sheldon, K. M., \& Lucas, R. E. (2014). \emph{Stability of happiness: Theories and evidence on whether happiness can change}.
 
@@
 Statista.
-(2021, May 21).
-Average daily time spent on social networks by users in the {United States} from 2018 to 2022.
-Retrieved from \url{https://www.statista.com/statistics/1018324/us-users-daily-social-media-minutes/}
+(2021).
+\emph{Average daily time spent on social networks by users in the {United States} from 2018 to 2022}.
+https://www.statista.com/statistics/1018324/us-users-daily-social-media-minutes/.
 
@@
 
+\leavevmode\vadjust pre{\hypertarget{ref-verbeijSelfreportedMeasuresSocial2021}{}}%
+Verbeij, T., Pouwels, J.
+L., Beyens, I., \& Valkenburg, P.
+M.
+(2021).
+\emph{Self-reported measures of social media use show high predictive validity} {[}Preprint{]}.
+{PsyArXiv}.
+\url{https://doi.org/10.31234/osf.io/c9bj7}
+
 \leavevmode\vadjust pre{\hypertarget{ref-wagnerAUTNESOnlinePanel2018}{}}%
@@
 (2018).
-\emph{{AUTNES Online Panel Study} 2017 ({SUF} edition)} {[}Data set{]}.
+\emph{{AUTNES Online Panel Study} 2017 ({SUF} edition)}.
 {AUSSDA}.
@@
 \leavevmode\vadjust pre{\hypertarget{ref-worldhealthorganizationWellbeingMeasuresPrimary1998}{}}%
-World Health Organization. (1998). Wellbeing measures in primary health care/{The Depcare Project}.
+World Health Organization. (1998). \emph{Wellbeing measures in primary health care/{The Depcare Project}}.
 
@@
 {A} qualitative exploration of online video watching among {Chinese} young adults.
-\emph{International Journal of Environmental Research and Public Health}, \emph{18}(14, 14), 7247.
+\emph{International Journal of Environmental Research and Public Health}, \emph{18}(14), 7247.
 \url{https://doi.org/10.3390/ijerph18147247}
@@
 \begin{figure}
-\includegraphics[width=\textwidth]{figures/fig_results_activity} \caption{The effects of various types of social media use on three indicators of well-being.
+\includegraphics[width=\textwidth]{manuscript_files/figure-latex/res-activity-1} \caption{The effects of various types of social media use on three indicators of well-being.
 The black estimates show the effects controlled for a large number of covariates (see text; preregistered); the grey estimates are without control variables (exploratory).
@@
 \begin{figure}
-\includegraphics[width=\textwidth]{figures/fig_results_channels} \caption{The effects of using various social media applications on three indicators of well-being.
+\includegraphics[width=\textwidth]{manuscript_files/figure-latex/res-channels-1} \caption{The effects of using various social media applications on three indicators of well-being.
 The black estimates show the effects controlled for a large number of covariates (see text; preregistered); the grey estimates are without control variables (exploratory).
@@
 \begin{figure}
-\includegraphics[width=\textwidth]{figures/fig_results_control_std} \caption{Results of selected covariates.
+\includegraphics[width=\textwidth]{manuscript_files/figure-latex/res-control-1} \caption{Results of selected covariates.
 All variables were standardize except 'Male'  and ‘Employed in public service', because there were measured on a binary scale.}\label{fig:res-control}

### 2022-07-25 analyses.RmD
- major updates for revision - included multiple imputation - changed theory - include new data-set - extended discussion

--- 2022-04-11 analyses.RmD
@@
     code_folding: hide
+    df_print: paged
 bibliography      : 
@@
 set.seed(170819)
-knitr::opts_chunk$set(warning=F, echo=T, error=F, message=F, cache = F)
-options(digits = 3, width = 250, scipen=999)
-```
-
-Here you can find the code and the results of all analyses. To see the code, click on button "Code". Alternatively, you can download the rmd file from the github repo.
+knitr::opts_chunk$set(warning=F, echo=T, error=F, message=F, cache=T, rows.print=20)
+options(digits=2, width=260, scipen=999)
+```
+
+ - Here you can find the code and the results of all analyses. 
+ - To see the code, click on button "Code". A
+ - Alternatively, you can download the rmd file from the github repo.
+ - The data are hosted on aussda. Please download them separately, to rerun the analyses (https://doi.org/10.11587/28KQNS). 
+ - Note that this paper uses an advance version consisting of 32 waves, which is at the time of writing not officially published.
 
@@
 
-```{r r-setup, results='hide', cache=FALSE}
+```{r r-setup, results='hide', cache=F}
 # install packages
@@
 
+## Session info
+
+Provide session info to make results reproducible.
+
+```{r}
+sessionInfo()
+```
+
 ## Custom functions
@@
     filter(Label == name) %>%
-    select(variable) %>% 
+    select(Variable) %>%
     unlist() %>%
@@
 )
+# find_var <- function(name, data = doc)(
+#   # finds the variables names for an item for each wave
+#     data %>% 
+#     filter(`Variable names (DE)` == name) %>% 
+#     select(`VARNAME (CA)`) %>% 
+#     unlist() %>% 
+#     set_names(sub("\\_.*", "", .))
+# )
 ```
@@
     dat_fig_results %<>%
-      filter(iv %in% c("male", "health", "loc_cntrl_int_m", "employment_facPublic service", "act_spo", "sat_dem")) %>% 
-      mutate(iv = factor(.$iv, levels = c("male", "health", "loc_cntrl_int_m", "employment_facPublic service", "act_spo", "sat_dem"), 
-                         labels = c("Male", "Health", "Internal locus of control", "Employed public service", "Sport", "Satisfaction democracy")))
+      filter(iv %in% c("male", "health", "loc_cntrl_int_m", #"employment_facPublic service", 
+                       "act_spo", "sat_dem")) %>% 
+      mutate(iv = factor(.$iv, levels = c("male", "health", "loc_cntrl_int_m", #"employment_facPublic service", 
+                                          "act_spo", "sat_dem"), 
+                         labels = c("Male", "Health", "Internal locus of control", #"Employed public service", 
+                                    "Sport", "Satisfaction democracy")))
   }
@@
 ```{r}
-make_graph_res <- function(data, sesoi = NULL, legend = TRUE){
+make_graph_res <- function(data, sesoi = NULL, legend = TRUE, facet = "comm", title = NULL){
   ggplot(data, aes(x = estimate, y = iv)) + 
-  scale_color_manual(values = c("grey40", "deepskyblue", "darkcyan", "cornflowerblue", "aquamarine")) +
+  scale_color_manual(values = c("black", "grey75", "darkcyan", "deepskyblue", "cornflowerblue", "darkcyan", "aquamarine")) +
   geom_vline(xintercept = 0, lwd = .75, colour = "darkgrey") + 
-  geom_errorbarh(aes(xmin = conf.low, xmax = conf.high, color = analysis),
-                lwd = .75, height = .2, position = position_dodge(.5)) +
-  geom_point(aes(color = analysis), size = 2, position = position_dodge(.5)) +
+  geom_errorbarh(aes(xmin = conf.low, xmax = conf.high, color = Analysis),
+                lwd = .75, height = .2, position = position_dodge(-.7)) +
+  geom_point(aes(color = Analysis), size = 1.5, position = position_dodge(-.7)) +
     {if(isTRUE(sesoi == "est")) geom_vline(data=filter(data, dv=="Life satisfaction"), aes(xintercept=-.3), colour="darkgrey", linetype = "dashed")} + 
@@
     text = element_text(size = 12),
-    axis.text.y = element_text(hjust = 0)
+    axis.text.y = element_text(hjust = 1)
     ) +
+  {if(!is.null(title)) ggtitle(title)} +
   {if(!isTRUE(legend)) theme(legend.position = "none")} + 
-  guides(colour = guide_legend(reverse=T)) +
+  # guides(colour = guide_legend(reverse=T)) +
+    {if(facet == "comm")
+      facet_grid(rows = vars(comm),
+             cols = vars(dv),
+             scales = "free",
+             space = "free_y")
+    } + 
+    {if(facet == "analysis")
       facet_grid(rows = vars(type),
@@
              space = "free_y")
+    }
 }
@@
 # Data Wrangling
-
-```{r}
-doc <- read_csv("documentation/10094_vi_de_v4_0.csv")
-d_raw <- read_csv("data/10094_da_de_v4_0.csv")
-```
+## Load data
+
+```{r}
+# doc <- read_csv("documentation/10094_vi_de_v4_0.csv")
+doc <- read_csv2("data/Variable_List_w1-34.csv")
+# doc <- read_csv("documentation/W1-32_ACPP_Variablelist_V20220601.csv")
+# d_raw <- read_csv("data/10094_da_de_v4_0.csv")
+# d_raw_test <- read_csv("data/10094_da_de_v4_0.csv")
+d_raw <- read_csv("data/W1-32_ACPP_V20220601.csv")
+```
+
+## Find names
 
@@
 loc_cntrl_int_4 <- find_var("Psychologie: Schicksal")
-```
-
-Select variables.
+trst_media <- find_var("Vertrauen: ORF")
+trst_police <- find_var("Vertrauen: Polizei")
+trst_media <- find_var("Vertrauen: Parlament")
+trst_hlthsec <- find_var("Vertrauen: Gesundheitswesen")
+trst_gov <- find_var("Vertrauen: Bundesregierung")
+trst_army <- find_var("Vertrauen: Bundesheer")
+corona_pos <- c(
+  find_var("Corona-Diagnose: Respondent"),
+  find_var("Corona-Diagnose: Monat")
+)
+```
+
+## Select variables
 
@@
     loc_cntrl_int_3 = all_of(loc_cntrl_int_3),
-    loc_cntrl_int_4 = all_of(loc_cntrl_int_4)
+    loc_cntrl_int_4 = all_of(loc_cntrl_int_4),
+    trst_media = all_of(trst_media),
+    trst_police = all_of(trst_police),
+    trst_media = all_of(trst_media),
+    trst_hlthsec = all_of(trst_hlthsec),
+    trst_gov = all_of(trst_gov),
+    trst_army = all_of(trst_army),
+    corona_pos = all_of(corona_pos)
   )
@@
 ```{r}
-doc_selected <- filter(doc, variable %in% c(
+doc_selected <- filter(doc, Variable %in% c(
   "RESPID",
@@
   loc_cntrl_int_3[1],
-  loc_cntrl_int_4[1]
+  loc_cntrl_int_4[1],
+  trst_media[1],
+  trst_police[1],
+  trst_media[1],
+  trst_hlthsec[1],
+  trst_gov[1],
+  trst_army[1],
+  corona_pos[1]
   ))
-```
-
-Recode variables.
+write.csv(doc_selected, "data/variables_selected.csv")
+```
+
+## Recode variables
 
@@
 
-Make data from wide to long.
-
-```{r}
+## Make long data format
+
+```{r cache=T}
 d_long <-
@@
   pivot_longer(
-    cols = health...W1:loc_cntrl_int_4...W9,
+    cols = health...W1:corona_pos...W32,
     names_to = "item",
@@
 
-Recode values.
+## Recode values
 
@@
 
-Order dataset. First id, then wave.
+## Order dataset
 
@@
   arrange(id, wave)
-```
-
-# Impute data
-
-Determine amount of missingness per respondent per wave. Because calculation takes long, this chunk is not executed but loaded from memory. To execute, set eval == TRUE.
-
-```{r eval=TRUE, cache=TRUE}
-vars_used <- c("life_sat", "aff_pos_1", "aff_pos_2", "aff_pos_3", "aff_neg_1", "aff_neg_2", "aff_neg_3", "aff_neg_4", "aff_neg_5", "aff_neg_6", "soc_med_read", "soc_med_like_share", "soc_med_post", "soc_med_fb", "soc_med_ig", "soc_med_wa", "soc_med_yt", "soc_med_tw", "age", "male", "born_aus", "born_aus_prnts", "edu_fac", "employment_fac", "health", "res_vienna", "acc_bal", "acc_gar", "home_sqm", "med_txt_kro", "med_txt_sta", "med_txt_pre", "med_txt_oes", "med_txt_kur", "med_txt_slz", "med_txt_son", "med_vid_orf", "med_vid_pri", "risk_prop", "loc_cntrl_int_1", "loc_cntrl_int_2", "loc_cntrl_int_3", "loc_cntrl_int_4", "act_wrk", "act_spo", "act_frn", "act_sho", "act_pet", "sat_dem")
-
-# filter respondents with more than 50% missing data
-d_long_filt <- 
+
+# create backup bc comp takes long
+d_long_backup <- d_long
+```
+
+## Impute data
+
+Next we impute data. Note that the actual imputation below is deactivated and loaded from memory to save time.
+
+### Prep
+
+Determine amount of missingness per respondent per wave. Because calculation takes long, this chunk is loaded from memory.
+
+```{r}
+vars_used <- c("life_sat", "aff_pos_1", "aff_pos_2", "aff_pos_3", "aff_neg_1", "aff_neg_2", "aff_neg_3", "aff_neg_4", "aff_neg_5", "aff_neg_6", "soc_med_read", "soc_med_like_share", "soc_med_post", "soc_med_fb", "soc_med_ig", "soc_med_wa", "soc_med_yt", "soc_med_tw", "age", "male", "born_aus", "born_aus_prnts", "edu_fac", "employment_fac", "health", "res_vienna", "acc_bal", "acc_gar", "home_sqm", "med_txt_kro", "med_txt_sta", "med_txt_pre", "med_txt_oes", "med_txt_kur", "med_txt_slz", "med_txt_son", "med_vid_orf", "med_vid_pri", "risk_prop", "loc_cntrl_int_1", "loc_cntrl_int_2", "loc_cntrl_int_3", "loc_cntrl_int_4", "act_wrk", "act_spo", "act_frn", "act_sho", "act_pet", "sat_dem", "trst_media", "trst_police", "trst_media", "trst_hlthsec", "trst_gov", "trst_army", "corona_pos") # only include vars measured at _all_ points
+```
+
+Filter respondents with more than 50% missing data -- only needed for analyses as originally preregistered (see additional analyses).
+
+```{r eval=T}
+d_long_50 <- 
   d_long %>% 
@@
   filter(na_perc < .5)
-
-# impute missing data
-# exclude social media use data, because they were measured only on selected waves and would lead to the imputation of too much missing data
-vars_excl <- c("na_perc", "soc_med_read", "soc_med_like_share", "soc_med_post", "soc_med_fb", "soc_med_ig", "soc_med_wa", "soc_med_yt", "soc_med_tw")
-incl_ma <- d_long_filt %>% 
+write_csv(d_long_50, "data/data_50.csv")
+```
+
+Exclude social media use data, because they were measured only on selected waves.
+
+```{r}
+vars_excl <- c("soc_med_read", "soc_med_like_share", "soc_med_post", "soc_med_fb", "soc_med_ig", "soc_med_wa", "soc_med_yt", "soc_med_tw")
+incl_ma <- d_long %>% 
   mutate(across(.cols = everything(), .fns = is.na))
@@
 
-d_long_imp <- mice(d_long_filt, 
+# now also for data where 50% missing excluded
+incl_ma_50 <- d_long_50 %>% 
+  mutate(across(.cols = everything(), .fns = is.na))
+incl_ma_50[vars_excl] <- FALSE
+```
+
+### Single imputation
+
+For analyses as originally preregistered (see additional analyses), hence only participants who provided more than 50% of all data.
+
+```{r imputation-single-50, eval=F, cache=T}
+d_long_50_imp <- mice(d_long_50, 
+                   method = "pmm", # use predictive mean matching
+                   m = 1, maxit = 30, # only 1 imputation
+                   where = incl_ma_50,
+                   seed = 180719, print = FALSE)
+d_long_50_imp <- mice::complete(d_long_50_imp)
+write_csv(d_long_50_imp, "data/data_50_imputed.csv")
+```
+
+Because CFAs aren't evaluated with multiple imputation, impute single data-set. Missing data was imputed for everyone.
+
+```{r imputation-single-100, eval=F, cache=T}
+d_long_100_imp <- mice(d_long, 
                    method = "pmm", # use predictive mean matching
@@
                    seed = 180719, print = FALSE)
-d_long_imp <- mice::complete(d_long_imp)
-write_csv(d_long_imp, "data/data_imputed.csv")
-```
-
-```{r eval=TRUE, cache=TRUE}
+d_long_100_imp <- mice::complete(d_long_100_imp)
+write_csv(d_long_100_imp, "data/data_100_imputed.csv")
+```
+
+### Multiple Imputation
+
+Now, multiple imputation of data-sets (5 data sets, 5 iterations), used for the final analyses.
+
+```{r imputation-mult, eval=F, cache=T}
 # impute missing data with multiple imputation
-d_long_mim_mice <- mice(d_long_filt, 
+d_long_mim_mice <- mice(d_long, 
                    method = "pmm", # use predictive mean matching
@@
 d_long_mim <- mice::complete(d_long_mim_mice, action = "long", include = TRUE)
-```
-
-Load imputed data from disc.
-
-```{r}
-d_long_imp <- read_csv("data/data_imputed.csv")
-```
-
-# Create within-between data
-
-Create means for scales
-
-```{r}
-d_long_imp %<>%
+write_csv(d_long_mim, "data/data_100_mim.csv")
+```
+
+```{r eval=T}
+# to further increase speed, you can also load the imputed data from the directory
+# d_long_imp <- read_csv("data/data_imputed.csv")
+# d_long_mim <- read_csv("data/data_imputed_mim.csv")
+# d_long_imp <- read_csv("data/data_imputed_50.csv")  # data where people with >50% missing data were filtered
+# d_long_mim <- read_csv("data/data_imputed_mim_50.csv")  # data where people with >50% missing data were filtered
+d_long_50_imp <- read_csv("data/data_50_imputed.csv") # data where all people were included
+d_long_100_imp <- read_csv("data/data_100_imputed.csv") # data where all people were included
+d_long_100_mim <- read_csv("data/data_100_mim.csv") # data where all people were included
+```
+
+## Create within-between data
+
+First, create means for scales. Then, create mean values for each person (between), and unique changes per wave (within).
+
+```{r}
+d_long_mim <-
+  d_long_100_mim %>% 
   mutate(aff_pos_m = rowMeans(select(., aff_pos_1 : aff_pos_3)),
@@
          loc_cntrl_int_m = rowMeans(select(., loc_cntrl_int_1 : loc_cntrl_int_4)))
-```
-
-Create mean values for each person (between), and unique changes per wave (within).
-
-```{r}
-d_long_imp %<>% 
+
+d_long_mim %<>%
   group_by(id) %>% 
-  mutate_at(vars(health:loc_cntrl_int_4, loc_cntrl_int_m), funs(b = mean(., na.rm = TRUE), w = . - mean(., na.rm = TRUE))) %>% 
+  mutate_at(vars(health:corona_pos, loc_cntrl_int_m), funs(b = mean(., na.rm = TRUE), w = . - mean(., na.rm = TRUE))) %>% 
   ungroup()
-```
-
-And because later we want to rerun the results without imputed data, now the same for data without imputation.
-
-```{r}
-d_long %<>%
+
+d_long_mim_mice <- as.mids(d_long_mim)
+```
+
+```{r}
+d_long_100_imp <-
+  d_long_100_imp %>% 
   mutate(aff_pos_m = rowMeans(select(., aff_pos_1 : aff_pos_3)),
@@
 
-d_long %<>% 
+d_long_100_imp %<>%
   group_by(id) %>% 
-  mutate_at(vars(health:loc_cntrl_int_4, loc_cntrl_int_m), funs(b = mean(., na.rm = TRUE), w = . - mean(., na.rm = TRUE))) %>% 
+  mutate_at(vars(health:corona_pos, loc_cntrl_int_m), funs(b = mean(., na.rm = TRUE), w = . - mean(., na.rm = TRUE))) %>% 
   ungroup()
@@
 
-And now the same for the multiple imputed data
-
-```{r}
-d_long_mim %<>%
+And because later we want to rerun the results as originally preregistered, we'll now do the same for analyses without filtered data.
+
+```{r}
+d_long_50_imp <-
+  d_long_50_imp %>% 
   mutate(aff_pos_m = rowMeans(select(., aff_pos_1 : aff_pos_3)),
@@
 
-d_long_mim %<>% 
+d_long_50_imp %<>% 
   group_by(id) %>% 
-  mutate_at(vars(health:loc_cntrl_int_4, loc_cntrl_int_m), funs(b = mean(., na.rm = TRUE), w = . - mean(., na.rm = TRUE))) %>% 
+  mutate_at(vars(health:corona_pos, loc_cntrl_int_m), 
+            funs(b = mean(., na.rm = TRUE), w = . - mean(., na.rm = TRUE))) %>% 
   ungroup()
-
-d_long_mim_mice <- as.mids(d_long_mim)
+```
+
+And because later we want to rerun the results without imputed data, now the same for data without imputation.
+
+```{r}
+d_long_50 %<>% 
+  mutate(aff_pos_m = rowMeans(select(., aff_pos_1 : aff_pos_3)),
+         aff_neg_m = rowMeans(select(., aff_neg_1 : aff_neg_6)),
+         loc_cntrl_int_m = rowMeans(select(., loc_cntrl_int_1 : loc_cntrl_int_4)))
+
+d_long_50 %<>% 
+  group_by(id) %>% 
+  mutate_at(vars(health:corona_pos, loc_cntrl_int_m), funs(b = mean(., na.rm = TRUE), w = . - mean(., na.rm = TRUE))) %>% 
+  ungroup()
+```
+
+And now the same for the standardized data set.
+
+```{r}
+# first standardize
+d_long_mim_std <-
+  d_long_100_mim %>% 
+  group_by(.imp) %>% 
+  mutate_at(vars(gender:corona_pos, -id, -wave, -gender, -male, -born_aus, -born_aus_prnts, -edu_fac, -employment_fac, -res_vienna, -acc_bal, -acc_gar), ~ c(scale(.))) %>% 
+  ungroup()
+
+d_long_mim_std %<>%
+  mutate(aff_pos_m = rowMeans(select(., aff_pos_1 : aff_pos_3)),
+         aff_neg_m = rowMeans(select(., aff_neg_1 : aff_neg_6)),
+         loc_cntrl_int_m = rowMeans(select(., loc_cntrl_int_1 : loc_cntrl_int_4)))
+
+d_long_mim_std %<>%
+  group_by(id) %>%
+  mutate_at(vars(health:corona_pos, loc_cntrl_int_m), funs(b = mean(., na.rm = TRUE), w = . - mean(., na.rm = TRUE))) %>%
+  ungroup()
+
+d_long_mim_mice_std <- as.mids(d_long_mim_std)
 ```
@@
 ```{r}
-model_life_sat <- lmer(life_sat ~ (1 | id) + (1 | wave), d_long_imp)
+model_life_sat <- lmer(life_sat ~ (1 | id) + (1 | wave), d_long_100_imp)
 summary(model_life_sat)
@@
 
-```{r eval=FALSE}
-mardia(d_long_imp %>% select(aff_pos_1, aff_pos_2, aff_pos_3))
+```{r eval=F}
+mardia(d_long_100_imp %>% select(aff_pos_1, aff_pos_2, aff_pos_3))
 ```
@@
 "
-cfa_aff_pos <- cfa(model, d_long_imp, group = "wave", estimator = "MLM")
+cfa_aff_pos <- cfa(model, d_long_100_imp, group = "wave", estimator = "MLM")
+cfa_aff_pos_50 <- cfa(model, d_long_50_imp, group = "wave", estimator = "MLM")
 summary(cfa_aff_pos, standardized = TRUE, fit = TRUE, estimates = FALSE)
@@
 The data fit the model very well, `r td::fit_txt(cfa_aff_pos)`.
-Let's next inspect reliability
+Let's next inspect reliability.
 
@@
 rel_aff_pos <- get_rel(cfa_aff_pos)
-rel_aff_pos
-```
-
-The average reliability across all waves was omega = `r rel_aff_pos`, hence good.
+```
+
+The average reliability across all waves was omega = `r rel_aff_pos[[1]]`, hence good.
 
@@
 ```{r}
-model_aff_pos <- lmer(aff_pos_m ~ (1 | id) + (1 | wave), d_long_imp)
+model_aff_pos <- lmer(aff_pos_m ~ (1 | id) + (1 | wave), d_long_100_imp)
 summary(model_aff_pos)
@@
 ```{r}
-d_long_imp$aff_pos_fs <- get_fs(cfa_aff_pos)
+d_long_50_imp$aff_pos_fs <- get_fs(cfa_aff_pos_50)
 ```
@@
 
-Negative affect was measured as a scale. We hence inspect factorial validty using CFA.
-
-We first test assumption of multivariate normality. (Warning, as these analyses take long and might require additional working memory, they're deactived currently.)
+Negative affect was measured as a scale. We hence inspect factorial validity using CFA.
+
+We first test assumption of multivariate normality. (Warning, as these analyses take long and require much working memory, they're deactivated currently.)
 
 ```{r eval=FALSE}
-mardia(d_long_imp %>% select(aff_neg_1, aff_neg_2, aff_neg_3, aff_neg_4, aff_neg_5, aff_neg_6))
-```
-
-Assumption of multivariate normal distributino was violated; hence, robust estimator will be used.
+mardia(d_long_100_imp %>% select(aff_neg_1, aff_neg_2, aff_neg_3, aff_neg_4, aff_neg_5, aff_neg_6))
+```
+
+Assumption of multivariate normal distribution was violated; hence, robust estimator will be used.
 
@@
 "
-cfa_aff_neg <- cfa(model, d_long_imp, group = "wave", estimator = "MLM")
+cfa_aff_neg <- cfa(model, d_long_100_imp, group = "wave", estimator = "MLM")
+cfa_aff_neg_50 <- cfa(model, d_long_50_imp, group = "wave", estimator = "MLM")
 summary(cfa_aff_neg, standardized = TRUE, fit = TRUE, estimates = FALSE)
@@
 rel_aff_neg <- get_rel(cfa_aff_neg)
-rel_aff_neg
-```
-
-The average reliability across all waves was omega = `r rel_aff_neg`, hence good.
-
-```{r}
-model_aff_neg <- lmer(aff_neg_m ~ (1 | id) + (1 | wave), d_long_imp)
+```
+
+The average reliability across all waves was omega = `r rel_aff_neg[[1]]`, hence good.
+
+```{r}
+model_aff_neg <- lmer(aff_neg_m ~ (1 | id) + (1 | wave), d_long_100_imp)
 summary(model_aff_neg)
@@
 ```{r}
-d_long_imp$aff_neg_fs <- get_fs(cfa_aff_neg)
+d_long_50_imp$aff_neg_fs <- get_fs(cfa_aff_neg_50)
 ```
@@
 ## Social media use
+
+Social media use (and channels) were measured at waves 1, 2, 8, 17, 23, 28, and for everyone who was newly recruited during the study at the first wave.
+
 ### Reading
@@
 ```{r}
-model_soc_med_read <- lmer(soc_med_read ~ (1 | id) + (1 | wave), d_long_imp)
+model_soc_med_read <- lmer(soc_med_read ~ (1 | id) + (1 | wave), d_long_100_imp)
 summary(model_soc_med_read)
@@
 ```{r}
-model_soc_med_like_share <- lmer(soc_med_like_share ~ (1 | id) + (1 | wave), d_long_imp)
+model_soc_med_like_share <- lmer(soc_med_like_share ~ (1 | id) + (1 | wave), d_long_100_imp)
 summary(model_soc_med_like_share)
@@
 ```{r}
-model_soc_med_post <- lmer(soc_med_post ~ (1 | id) + (1 | wave), d_long_imp)
+model_soc_med_post <- lmer(soc_med_post ~ (1 | id) + (1 | wave), d_long_100_imp)
 summary(model_soc_med_post)
@@
 ```{r}
-model_soc_med_fb <- lmer(soc_med_fb ~ (1 | id) + (1 | wave), d_long_imp)
+model_soc_med_fb <- lmer(soc_med_fb ~ (1 | id) + (1 | wave), d_long_100_imp)
 summary(model_soc_med_fb)
@@
 ```{r}
-model_soc_med_tw <- lmer(soc_med_tw ~ (1 | id) + (1 | wave), d_long_imp)
+model_soc_med_tw <- lmer(soc_med_tw ~ (1 | id) + (1 | wave), d_long_100_imp)
 summary(model_soc_med_tw)
@@
 ```{r}
-soc_med_tw_m <- d_long_imp %>% group_by(wave) %>% summarise(value = mean(soc_med_tw, na.rm = T))
+soc_med_tw_m <- d_long_100_imp %>% group_by(wave) %>% summarise(value = mean(soc_med_tw, na.rm = T))
 soc_med_tw_m
@@
 ```{r}
-model_soc_med_ig <- lmer(soc_med_ig ~ (1 | id) + (1 | wave), d_long_imp)
+model_soc_med_ig <- lmer(soc_med_ig ~ (1 | id) + (1 | wave), d_long_100_imp)
 summary(model_soc_med_ig)
@@
 ```{r}
-model_soc_med_wa <- lmer(soc_med_wa ~ (1 | id) + (1 | wave), d_long_imp)
+model_soc_med_wa <- lmer(soc_med_wa ~ (1 | id) + (1 | wave), d_long_100_imp)
 summary(model_soc_med_wa)
@@
 ```{r}
-model_soc_med_yt <- lmer(soc_med_yt ~ (1 | id) + (1 | wave), d_long_imp)
+model_soc_med_yt <- lmer(soc_med_yt ~ (1 | id) + (1 | wave), d_long_100_imp)
 summary(model_soc_med_yt)
@@
 
-The only other variable that was measured as a scale was Locus of Control. Below I hence report the scale's factorial validity.
-
-```{r}
+The only other variable that was measured as a scale was Locus of Control.
+Below I hence report the scale's factorial validity.
+Waves for which not a sufficient number of respondents took part were excluded.
+
+```{r eval=T}
 model <- "
@@
 "
-cfa_loc_cntrl_int <- cfa(model, filter(d_long_imp, wave != 20), group = "wave")
+# cfa_loc_cntrl_int <- cfa(model, filter(d_long_imp, wave != 20), group = "wave")
+cfa_loc_cntrl_int <- cfa(model, filter(d_long_100_imp, wave != 20,  wave != 26, wave != 27, wave != 29, wave != 31, wave != 32), group = "wave")
 summary(cfa_loc_cntrl_int, standardized = TRUE, fit = TRUE, estimates = FALSE)
@@
 The data fit the model very well, `r td::fit_txt(cfa_loc_cntrl_int)`.
+
+The model didn't converge, likely for reasons of complexity, which is why the fit indices cannot be reported here.
 
@@
     "YouTube" = get_specs(model_soc_med_yt)
-    )
-kable(tab_desc_dat)
+    ) %>%
+  as.data.frame()
+tab_desc_dat
 ```
@@
 # Analyses
-## Life satisfaction
-
-First, without control variables.
-
-```{r}
-model_life_sat_lmer_nco <- lmerTest::lmer(life_sat ~ 
-                            (1 | id) + (1 | wave) + 
-                            soc_med_read_w + soc_med_like_share_w + soc_med_post_w + 
-                            soc_med_fb_w + soc_med_ig_w + soc_med_wa_w + soc_med_yt_w + soc_med_tw_w +
-                            soc_med_read_b + soc_med_like_share_b + soc_med_post_b + 
-                            soc_med_fb_b + soc_med_ig_b + soc_med_wa_b + soc_med_yt_b + soc_med_tw_b,
-                data = d_long_imp)
-summary(model_life_sat_lmer_nco)
-```
-
-Now with control variables.
-
-```{r}
-model_life_sat_lmer <- lmerTest::lmer(life_sat ~
+
+Before running the analyses, let's briefly check multicollinearity.
+
+```{r}
+d_long_100_imp %>% 
+  filter(wave == 1) %>% 
+  select(soc_med_read, soc_med_post, soc_med_like_share) %>% 
+  cor(use = "pairwise.complete.obs")
+```
+
+Shows that communication types aren't collinear.
+
+```{r}
+d_long_100_imp %>% 
+  filter(wave == 1) %>% 
+  select(soc_med_fb, soc_med_ig, soc_med_tw, soc_med_wa, soc_med_yt) %>% 
+  cor(use = "pairwise.complete.obs")
+```
+
+Shows that communication channels aren't collinear.
+
+## Publication
+### Life satisfaction
+
+```{r}
+model_life_sat_lmer_pub <- with(d_long_mim_mice, exp =  
+                                lmerTest::lmer(life_sat ~
                                         (1 | id) + (1 | wave) + 
@@
                                         soc_med_fb_b + soc_med_ig_b + soc_med_wa_b + soc_med_yt_b + soc_med_tw_b +  
-                                        age + male + born_aus + born_aus_prnts + edu_fac + employment_fac + health_b + 
-                                        res_vienna + acc_bal + acc_gar + home_sqm + 
-                                        # hh_income_b + # large dropout (900 obs)
-                                        # hh_adults + # too much dropout
-                                        # hh_child18 + hh_child17 + hh_child14 + hh_child5 + hh_child2 + hh_oldfam + hh_outfam + hh_partner + 
-                                        # home_owner + # too much dropout
+                                        age + male + born_aus + born_aus_prnts + edu_fac + employment_fac +
+                                        res_vienna + acc_bal + acc_gar + home_sqm + corona_pos +
+                                        hh_income_b +
+                                        hh_adults +
+                                        hh_child18 + hh_child17 + hh_child14 + hh_child5 + hh_child2 + hh_oldfam + hh_outfam + hh_partner +
+                                        home_owner +
                                         med_txt_kro_b + med_txt_sta_b + med_txt_pre_b + med_txt_oes_b + med_txt_kur_b + med_txt_slz_b + med_txt_son_b + 
@@
                                         act_wrk_b + act_spo_b + act_frn_b + act_sho_b + act_pet_b +
-                                        sat_dem_w + sat_dem_b 
-                                      , data = d_long_imp)
-summary(model_life_sat_lmer)
-```
-
-## Positive Affect
-
-First, without control variables.
-
-```{r}
-model_aff_pos_lmer_nco <- lmerTest::lmer(aff_pos_fs ~ 
+                                        sat_dem_w + sat_dem_b + 
+                                        health_w + health_b
+                                      ))
+model_life_sat_lmer_pub <- summary(pool(model_life_sat_lmer_pub), conf.int = TRUE)
+model_life_sat_lmer_pub %>% 
+  as.data.frame %>% 
+  select(term, estimate, `2.5 %`, `97.5 %`, p.value) %>% 
+  mutate(p.value = td::my_round(p.value, "p"))
+```
+
+### Positive Affect
+
+```{r}
+model_aff_pos_lmer_pub <- with(d_long_mim_mice, exp =  
+                           lmerTest::lmer(aff_pos_m ~ 
+                                        (1 | id) + (1 | wave) + 
+                                        soc_med_read_w + soc_med_like_share_w + soc_med_post_w  + 
+                                        soc_med_fb_w + soc_med_ig_w + soc_med_wa_w + soc_med_yt_w + soc_med_tw_w +
+                                        soc_med_read_b + soc_med_like_share_b + soc_med_post_b + 
+                                        soc_med_fb_b + soc_med_ig_b + soc_med_wa_b + soc_med_yt_b + soc_med_tw_b +  
+                                        age + male + born_aus + born_aus_prnts + edu_fac + employment_fac +
+                                        res_vienna + acc_bal + acc_gar + home_sqm + corona_pos +
+                                        hh_income_b +
+                                        hh_adults +
+                                        hh_child18 + hh_child17 + hh_child14 + hh_child5 + hh_child2 + hh_oldfam + hh_outfam + hh_partner +
+                                        home_owner +
+                                        med_txt_kro_b + med_txt_sta_b + med_txt_pre_b + med_txt_oes_b + med_txt_kur_b + med_txt_slz_b + med_txt_son_b + 
+                                        med_vid_orf_b + med_vid_pri_b + 
+                                        risk_prop_b + loc_cntrl_int_m_b + 
+                                        act_wrk_w + act_spo_w + act_frn_w + act_sho_w + act_pet_w + 
+                                        act_wrk_b + act_spo_b + act_frn_b + act_sho_b + act_pet_b +
+                                        sat_dem_w + sat_dem_b + 
+                                        health_w + health_b
+                                        ))
+model_aff_pos_lmer_pub <- summary(pool(model_aff_pos_lmer_pub), conf.int = TRUE)
+model_aff_pos_lmer_pub %>% 
+  as.data.frame %>% 
+  select(term, estimate, `2.5 %`, `97.5 %`, p.value) %>% 
+  mutate(p.value = td::my_round(p.value, "p"))
+```
+
+### Negative Affect
+
+```{r}
+model_aff_neg_lmer_pub <- with(d_long_mim_mice, exp =  
+                               lmerTest::lmer(aff_neg_m ~ 
+                                        (1 | id) + (1 | wave) + 
+                                        soc_med_read_w + soc_med_like_share_w + soc_med_post_w  + 
+                                        soc_med_fb_w + soc_med_ig_w + soc_med_wa_w + soc_med_yt_w + soc_med_tw_w +
+                                        soc_med_read_b + soc_med_like_share_b + soc_med_post_b + 
+                                        soc_med_fb_b + soc_med_ig_b + soc_med_wa_b + soc_med_yt_b + soc_med_tw_b +  
+                                        age + male + born_aus + born_aus_prnts + edu_fac + employment_fac +
+                                        res_vienna + acc_bal + acc_gar + home_sqm + corona_pos +
+                                        hh_income_b +
+                                        hh_adults +
+                                        hh_child18 + hh_child17 + hh_child14 + hh_child5 + hh_child2 + hh_oldfam + hh_outfam + hh_partner +
+                                        home_owner +
+                                        med_txt_kro_b + med_txt_sta_b + med_txt_pre_b + med_txt_oes_b + med_txt_kur_b + med_txt_slz_b + med_txt_son_b + 
+                                        med_vid_orf_b + med_vid_pri_b + 
+                                        risk_prop_b + loc_cntrl_int_m_b + 
+                                        act_wrk_w + act_spo_w + act_frn_w + act_sho_w + act_pet_w + 
+                                        act_wrk_b + act_spo_b + act_frn_b + act_sho_b + act_pet_b +
+                                        sat_dem_w + sat_dem_b + 
+                                        health_w + health_b
+                                        ))
+model_aff_neg_lmer_pub <- summary(pool(model_aff_neg_lmer_pub), conf.int = TRUE)
+model_aff_neg_lmer_pub %>% 
+  as.data.frame %>% 
+  select(term, estimate, `2.5 %`, `97.5 %`, p.value) %>% 
+  mutate(p.value = td::my_round(p.value, "p"))
+```
+
+## Without control variables
+### Life satisfaction
+
+```{r}
+model_life_sat_lmer_nco <- with(d_long_mim_mice, exp =  
+                                lmerTest::lmer(life_sat ~
                                         (1 | id) + (1 | wave) + 
@@
                                         soc_med_fb_b + soc_med_ig_b + soc_med_wa_b + soc_med_yt_b + soc_med_tw_b
-                                      , data = d_long_imp)
-summary(model_aff_pos_lmer_nco)
-```
-
-Now with control variables.
-
-```{r}
-model_aff_pos_lmer <- lmerTest::lmer(aff_pos_fs ~ 
+                                      ))
+model_life_sat_lmer_nco <- summary(pool(model_life_sat_lmer_nco), conf.int = TRUE)
+model_life_sat_lmer_nco %>% 
+  as.data.frame %>% 
+  select(term, estimate, `2.5 %`, `97.5 %`, p.value) %>% 
+  mutate(p.value = td::my_round(p.value, "p"))
+```
+
+### Positive Affect
+
+```{r}
+model_aff_pos_lmer_nco <- with(d_long_mim_mice, exp =  
+                           lmerTest::lmer(aff_pos_m ~ 
+                                        (1 | id) + (1 | wave) + 
+                                        soc_med_read_w + soc_med_like_share_w + soc_med_post_w  + 
+                                        soc_med_fb_w + soc_med_ig_w + soc_med_wa_w + soc_med_yt_w + soc_med_tw_w +
+                                        soc_med_read_b + soc_med_like_share_b + soc_med_post_b + 
+                                        soc_med_fb_b + soc_med_ig_b + soc_med_wa_b + soc_med_yt_b + soc_med_tw_b
+                                        ))
+model_aff_pos_lmer_nco <- summary(pool(model_aff_pos_lmer_nco), conf.int = TRUE)
+model_aff_pos_lmer_nco %>% 
+  as.data.frame %>% 
+  select(term, estimate, `2.5 %`, `97.5 %`, p.value) %>% 
+  mutate(p.value = td::my_round(p.value, "p"))
+```
+
+### Negative Affect
+
+```{r}
+model_aff_neg_lmer_nco <- with(d_long_mim_mice, exp =  
+                               lmerTest::lmer(aff_neg_m ~ 
+                                        (1 | id) + (1 | wave) + 
+                                        soc_med_read_w + soc_med_like_share_w + soc_med_post_w  + 
+                                        soc_med_fb_w + soc_med_ig_w + soc_med_wa_w + soc_med_yt_w + soc_med_tw_w +
+                                        soc_med_read_b + soc_med_like_share_b + soc_med_post_b + 
+                                        soc_med_fb_b + soc_med_ig_b + soc_med_wa_b + soc_med_yt_b + soc_med_tw_b
+                                        ))
+model_aff_neg_lmer_nco <- summary(pool(model_aff_neg_lmer_nco), conf.int = TRUE)
+model_aff_neg_lmer_nco %>% 
+  as.data.frame %>% 
+  select(term, estimate, `2.5 %`, `97.5 %`, p.value) %>% 
+  mutate(p.value = td::my_round(p.value, "p"))
+```
+
+## Standardized results
+
+Then let's rerun the analyses.
+
+### Life satisfaction
+
+```{r}
+model_life_sat_lmer_std <- with(d_long_mim_mice_std, exp =  
+                                lmerTest::lmer(life_sat ~
                                         (1 | id) + (1 | wave) + 
@@
                                         soc_med_fb_b + soc_med_ig_b + soc_med_wa_b + soc_med_yt_b + soc_med_tw_b +  
-                                        age + male + born_aus + born_aus_prnts + edu_fac + employment_fac + health_b + 
-                                        res_vienna + acc_bal + acc_gar + home_sqm + 
+                                        age + male + born_aus + born_aus_prnts + edu_fac + employment_fac +
+                                        res_vienna + acc_bal + acc_gar + home_sqm + corona_pos +
+                                        hh_income_b +
+                                        hh_adults +
+                                        hh_child18 + hh_child17 + hh_child14 + hh_child5 + hh_child2 + hh_oldfam + hh_outfam + hh_partner +
+                                        home_owner +
                                         med_txt_kro_b + med_txt_sta_b + med_txt_pre_b + med_txt_oes_b + med_txt_kur_b + med_txt_slz_b + med_txt_son_b + 
@@
                                         act_wrk_b + act_spo_b + act_frn_b + act_sho_b + act_pet_b +
-                                        sat_dem_w + sat_dem_b 
-                                      , data = d_long_imp)
-summary(model_aff_pos_lmer)
-```
-
-## Negative Affect
-
-First, without control variables.
-
-```{r}
-model_aff_neg_lmer_nco <- lmerTest::lmer(aff_neg_fs ~ 
-                                        (1 | id) + (1 | wave) + 
-                                        soc_med_read_w + soc_med_like_share_w + soc_med_post_w + 
-                                        soc_med_fb_w + soc_med_ig_w + soc_med_wa_w + soc_med_yt_w + soc_med_tw_w +
-                                        soc_med_read_b + soc_med_like_share_b + soc_med_post_b + 
-                                        soc_med_fb_b + soc_med_ig_b + soc_med_wa_b + soc_med_yt_b + soc_med_tw_b
-                                      , data = d_long_imp)
-summary(model_aff_neg_lmer_nco)
-```
-
-Then, with control variables.
-
-```{r}
-model_aff_neg_lmer <- lmerTest::lmer(aff_neg_fs ~ 
+                                        sat_dem_w + sat_dem_b + 
+                                        health_w + health_b
+                                      ))
+model_life_sat_lmer_std <- summary(pool(model_life_sat_lmer_std), conf.int = TRUE)
+model_life_sat_lmer_std %>% 
+  as.data.frame %>% 
+  select(term, estimate, `2.5 %`, `97.5 %`, p.value) %>% 
+  mutate(p.value = td::my_round(p.value, "p"))
+```
+
+### Positive Affect
+
+```{r}
+model_aff_pos_lmer_std <- with(d_long_mim_mice_std, exp =  
+                           lmerTest::lmer(aff_pos_m ~ 
                                         (1 | id) + (1 | wave) + 
@@
                                         act_wrk_b + act_spo_b + act_frn_b + act_sho_b + act_pet_b +
-                                        sat_dem_w + sat_dem_b 
-                                      , data = d_long_imp)
-summary(model_aff_neg_lmer)
+                                        sat_dem_w + sat_dem_b + 
+                                        health_w + health_b
+                                        ))
+model_aff_pos_lmer_std <- summary(pool(model_aff_pos_lmer_std), conf.int = TRUE)
+model_aff_pos_lmer_std %>% 
+  as.data.frame %>% 
+  select(term, estimate, `2.5 %`, `97.5 %`, p.value) %>% 
+  mutate(p.value = td::my_round(p.value, "p"))
+```
+
+### Negative Affect
+
+```{r}
+model_aff_neg_lmer_std <- with(d_long_mim_mice_std, exp =  
+                               lmerTest::lmer(aff_neg_m ~ 
+                                        (1 | id) + (1 | wave) + 
+                                        soc_med_read_w + soc_med_like_share_w + soc_med_post_w  + 
+                                        soc_med_fb_w + soc_med_ig_w + soc_med_wa_w + soc_med_yt_w + soc_med_tw_w +
+                                        soc_med_read_b + soc_med_like_share_b + soc_med_post_b + 
+                                        soc_med_fb_b + soc_med_ig_b + soc_med_wa_b + soc_med_yt_b + soc_med_tw_b +  
+                                        age + male + born_aus + born_aus_prnts + edu_fac + employment_fac +
+                                        res_vienna + acc_bal + acc_gar + home_sqm + corona_pos +
+                                        hh_income_b +
+                                        hh_adults +
+                                        hh_child18 + hh_child17 + hh_child14 + hh_child5 + hh_child2 + hh_oldfam + hh_outfam + hh_partner +
+                                        home_owner +
+                                        med_txt_kro_b + med_txt_sta_b + med_txt_pre_b + med_txt_oes_b + med_txt_kur_b + med_txt_slz_b + med_txt_son_b + 
+                                        med_vid_orf_b + med_vid_pri_b + 
+                                        risk_prop_b + loc_cntrl_int_m_b + 
+                                        act_wrk_w + act_spo_w + act_frn_w + act_sho_w + act_pet_w + 
+                                        act_wrk_b + act_spo_b + act_frn_b + act_sho_b + act_pet_b +
+                                        sat_dem_w + sat_dem_b + 
+                                        health_w + health_b
+                                        ))
+model_aff_neg_lmer_std <- summary(pool(model_aff_neg_lmer_std), conf.int = TRUE)
+model_aff_neg_lmer_std %>% 
+  as.data.frame %>% 
+  select(term, estimate, `2.5 %`, `97.5 %`, p.value) %>% 
+  mutate(p.value = td::my_round(p.value, "p"))
 ```
@@
 
-First, for activities.
-
-```{r}
+## Within
+
+```{r warning=FALSE}
 # get data
-dat_fig_results_activity <- get_dat_res(model_aff_neg_lmer, model_aff_pos_lmer, model_life_sat_lmer, 
-                                        type = "activity", analysis = "regular")
+dat_fig_results_activity_pub <- get_dat_res(model_aff_neg_lmer_pub, model_aff_pos_lmer_pub, model_life_sat_lmer_pub, 
+                                        type = "activity", analysis = "Regular")
+dat_fig_results_channel_pub <- get_dat_res(model_aff_neg_lmer_pub, model_aff_pos_lmer_pub, model_life_sat_lmer_pub, 
+                                        type = "channels", analysis = "Regular")
+dat_fig_results_within_pub <- rbind(cbind(dat_fig_results_activity_pub, comm = "Activity"), 
+                                cbind(dat_fig_results_channel_pub, comm = "Channel")) %>% 
+  filter(type == "within")
+
 dat_fig_results_activity_nco <- get_dat_res(model_aff_neg_lmer_nco, model_aff_pos_lmer_nco, model_life_sat_lmer_nco, 
-                                            type = "activity", analysis = "no controls")
-dat_fig_results_activity <- rbind(
-  dat_fig_results_activity, 
-  dat_fig_results_activity_nco
+                                        type = "activity", analysis = "No controls")
+dat_fig_results_channel_nco <- get_dat_res(model_aff_neg_lmer_nco, model_aff_pos_lmer_nco, model_life_sat_lmer_nco, 
+                                        type = "channels", analysis = "No controls")
+dat_fig_results_within_nco <- rbind(cbind(dat_fig_results_activity_nco, comm = "Activity"), 
+                                cbind(dat_fig_results_channel_nco, comm = "Channel")) %>% 
+  filter(type == "within")
+
+dat_fig_results_activity_std <- get_dat_res(model_aff_neg_lmer_std, model_aff_pos_lmer_std, model_life_sat_lmer_std, 
+                                        type = "activity", analysis = "standardized")
+dat_fig_results_channel_std <- get_dat_res(model_aff_neg_lmer_std, model_aff_pos_lmer_std, model_life_sat_lmer_std, 
+                                        type = "channels", analysis = "standardized")
+dat_fig_results_within_std <- rbind(cbind(dat_fig_results_activity_std, comm = "Activity"), 
+                                cbind(dat_fig_results_channel_std, comm = "Channel")) %>% 
+  filter(type == "within")
+
+dat_fig_results_within <- rbind(
+  dat_fig_results_within_pub, 
+  dat_fig_results_within_nco,
+  dat_fig_results_within_std
+) %>% 
+  mutate(Analysis = factor(Analysis, levels = c("Regular", "No controls", "standardized")))
+
+# make figure
+fig_results_within <- make_graph_res(
+  data = filter(dat_fig_results_within, Analysis != "standardized"),
+  sesoi = "est",
+  facet = "comm"
+  # , title = "Within-person effects of COVID-19 related social media use on well-being"
   )
+fig_results_within
+
+# safe figure
+ggsave("figures/fig_results_within.png", 
+       width = 7, height = 4,
+       plot = fig_results_within)
+```
+
+## Between
+
+```{r warning=FALSE}
+# get data
+dat_fig_results_activity_pub <- get_dat_res(model_aff_neg_lmer_pub, model_aff_pos_lmer_pub, model_life_sat_lmer_pub, 
+                                        type = "activity", analysis = "Regular")
+dat_fig_results_channel_pub <- get_dat_res(model_aff_neg_lmer_pub, model_aff_pos_lmer_pub, model_life_sat_lmer_pub, 
+                                        type = "channels", analysis = "Regular")
+dat_fig_results_between_pub <- rbind(cbind(dat_fig_results_activity_pub, comm = "Activity"), 
+                                cbind(dat_fig_results_channel_pub, comm = "Channel")) %>% 
+  filter(type == "between")
+
+dat_fig_results_activity_nco <- get_dat_res(model_aff_neg_lmer_nco, model_aff_pos_lmer_nco, model_life_sat_lmer_nco, 
+                                        type = "activity", analysis = "No controls")
+dat_fig_results_channel_nco <- get_dat_res(model_aff_neg_lmer_nco, model_aff_pos_lmer_nco, model_life_sat_lmer_nco, 
+                                        type = "channels", analysis = "No controls")
+dat_fig_results_between_nco <- rbind(cbind(dat_fig_results_activity_nco, comm = "Activity"), 
+                                cbind(dat_fig_results_channel_nco, comm = "Channel")) %>% 
+  filter(type == "between")
+
+dat_fig_results_activity_std <- get_dat_res(model_aff_neg_lmer_std, model_aff_pos_lmer_std, model_life_sat_lmer_std, 
+                                        type = "activity", analysis = "standardized")
+dat_fig_results_channel_std <- get_dat_res(model_aff_neg_lmer_std, model_aff_pos_lmer_std, model_life_sat_lmer_std, 
+                                        type = "channels", analysis = "standardized")
+dat_fig_results_between_std <- rbind(cbind(dat_fig_results_activity_std, comm = "Activity"), 
+                                cbind(dat_fig_results_channel_std, comm = "Channel")) %>% 
+  filter(type == "between")
+
+dat_fig_results_between <- rbind(
+  dat_fig_results_between_pub, 
+  dat_fig_results_between_nco,
+  dat_fig_results_between_std
+) %>% 
+  mutate(Analysis = factor(Analysis, levels = c("Regular", "No controls", "standardized")))
 
 # make figure
-fig_results_activity <- make_graph_res(
-  data = dat_fig_results_activity,
+fig_results_between <- make_graph_res(
+  data = filter(dat_fig_results_between, Analysis != "standardized"),
   sesoi = "est"
+  # , title = "Between-person relations of COVID-19 related social media use and well-being"
   )
-fig_results_activity
+fig_results_between
 
 # safe figure
-ggsave("figures/fig_results_activity.pdf", 
+ggsave("figures/fig_results_between.png", 
        width = 7, height = 4,
-       plot = fig_results_activity)
-```
-
-Then for channels.
-
-```{r}
-# get data
-dat_fig_results_channels <- get_dat_res(model_aff_neg_lmer, model_aff_pos_lmer, model_life_sat_lmer, 
-                                        type = "channels", analysis = "regular")
-dat_fig_results_channels_nco <- get_dat_res(model_aff_neg_lmer_nco, model_aff_pos_lmer_nco, model_life_sat_lmer_nco, 
-                                            type = "channels", analysis = "no controls")
-dat_fig_results_channels <- rbind(
-  dat_fig_results_channels, 
-  dat_fig_results_channels_nco
+       plot = fig_results_between)
+```
+
+## Selected control variables
+
+Standardized, to allow for better comparison across differently scaled variables.
+
+```{r warning=F, cache=T}
+# make figure
+fig_results_control_std <- make_graph_res(
+  get_dat_res(model_aff_neg_lmer_std, model_aff_pos_lmer_std, model_life_sat_lmer_std, 
+              type = "control", analysis = "standardized"),
+  sesoi = "std",
+  legend = FALSE,
+  facet = "analysis"
+  # , title = "Results of selected covariates"
   )
-
-# make figure
-fig_results_channels <- make_graph_res(
-  data = dat_fig_results_channels,
-  sesoi = "est"
-  )
-fig_results_channels
-
-# safe figure
-ggsave("figures/fig_results_channels.pdf", 
+fig_results_control_std
+
+# save figure
+ggsave("figures/fig_results_control_std.png", 
        width = 7, height = 4,
-       plot = fig_results_channels)
-```
-
-Select control variables and make graph.
-
-```{r}
-# make figure
-fig_results_control <- make_graph_res(
-  get_dat_res(model_aff_neg_lmer, model_aff_pos_lmer, model_life_sat_lmer, analysis = "exploratory", type = "control"),
-  legend = FALSE
-  )
-fig_results_control
-
-# save figure
-ggsave("figures/fig_results_control.pdf", 
-       width = 7, height = 4,
-       plot = fig_results_control)
+       plot = fig_results_control_std)
+```
+
+# Tables
+
+Table of within-person effects
+
+```{r}
+tab_within <- dat_fig_results_within
+
+tab_within_std <-
+  tab_within %>%  
+  filter(Analysis == "standardized") %>% 
+  select(std = estimate)
+
+tab_within <-
+  tab_within %>% 
+  filter(Analysis == "Regular") %>% 
+  cbind(tab_within_std) %>% 
+  select(dv, iv, estimate, conf.low, conf.high, std, p.value) %>% 
+  mutate(p.value = td::my_round(p.value, "p")) %>% 
+  arrange(dv)
+
+tab_within
 ```

### 2022-07-25 analyses_additional.Rmd
- major updates for revision - included multiple imputation - changed theory - include new data-set - extended discussion

--- 2022-04-11 analyses_additional.Rmd
@@
 set.seed(170819)
-knitr::opts_chunk$set(warning=F, echo=T, error=F, message=F, cache = F)
-options(digits = 3, width = 200, scipen=999)
+knitr::opts_chunk$set(warning=F, echo=T, error=F, message=F, cache=T, max.row=20)
+options(digits=2, width=260, scipen=999)
 ```
@@
  - Here you can find the code and the results of all additional analyses. 
- - To see the results, I recommend you directly go down to the Figures section. 
+ - To see a good overview of the results, I recommend you directly go down to the Figures section. 
  - But if you're interested in how the analyses were executed, just tag along.
  - To see the code, click on button "Code". Alternatively, you can download the rmd file from the github repo.
- - To execute the code, first download the actual data from AUSSDA, then run analysis.rmd, then this file
+ - To execute the code, first download the actual data from AUSSDA (https://doi.org/10.11587/28KQNS), save as .csv, then run analysis.rmd, then this file
+ - Note that this paper uses an advance version consisting of 32 waves, which is at the time of writing not officially published.
 
@@
 
-# Results with mean scores
+# With additional covariates
 
-Instead of using factor scores, in what follows I report the results with mean scores of affect. Note that the results for life satisfaction were the same, as it was measured with a single item.
+These models include additional covariates that weren't preregistered. These are:
 
-## Positive affect
+- Trust in ORF
+- Trust in police
+- Trust in parliament
+- Trust in health sector
+- Trust in government
+- Trust in army
+
+## Life satisfaction
 
 ```{r}
-model_aff_pos_lmer_m <- lmerTest::lmer(aff_pos_m ~ 
+model_life_sat_lmer_add <- with(d_long_mim_mice, exp =  
+                                lmerTest::lmer(life_sat ~
                                         (1 | id) + (1 | wave) + 
@@
                                         act_wrk_b + act_spo_b + act_frn_b + act_sho_b + act_pet_b +
-                                        sat_dem_w + sat_dem_b 
-                                      , data = d_long_imp)
-summary(model_aff_pos_lmer_m)
+                                        sat_dem_w + sat_dem_b + 
+                                        health_w + health_b +
+                                        trst_media_w + trst_police_w + trst_media_w + trst_hlthsec_w + trst_gov_w + trst_army_w +
+                                        trst_media_b + trst_police_b + trst_media_b + trst_hlthsec_b + trst_gov_b + trst_army_b
+                                      ))
+model_life_sat_lmer_add <- summary(pool(model_life_sat_lmer_add), conf.int = TRUE)
+model_life_sat_lmer_add %>% 
+  as.data.frame %>% 
+  select(term, estimate, `2.5 %`, `97.5 %`, p.value) %>% 
+  mutate(p.value = td::my_round(p.value, "p"))
 ```
 
-The results differed only slightly, and all inferences remained the same.
-
-## Negative Affect
+## Positive Affect
 
 ```{r}
-model_aff_neg_lmer_m <- lmerTest::lmer(aff_neg_m ~ 
+model_aff_pos_lmer_add <- with(d_long_mim_mice, exp =  
+                           lmerTest::lmer(aff_pos_m ~ 
                                         (1 | id) + (1 | wave) + 
@@
                                         act_wrk_b + act_spo_b + act_frn_b + act_sho_b + act_pet_b +
-                                        sat_dem_w + sat_dem_b 
-                                      , data = d_long_imp)
-summary(model_aff_neg_lmer_m)
+                                        sat_dem_w + sat_dem_b + 
+                                        health_w + health_b +
+                                        trst_media_w + trst_police_w + trst_media_w + trst_hlthsec_w + trst_gov_w + trst_army_w +
+                                        trst_media_b + trst_police_b + trst_media_b + trst_hlthsec_b + trst_gov_b + trst_army_b
+                                        ))
+model_aff_pos_lmer_add <- summary(pool(model_aff_pos_lmer_add), conf.int = TRUE)
+model_aff_pos_lmer_add %>% 
+  as.data.frame %>% 
+  select(term, estimate, `2.5 %`, `97.5 %`, p.value) %>% 
+  mutate(p.value = td::my_round(p.value, "p"))
 ```
 
-The results differed only slightly, and all inferences remained the same.
+## Negative Affect
 
-# Results without imputed data
+```{r}
+model_aff_neg_lmer_add <- with(d_long_mim_mice, exp =  
+                               lmerTest::lmer(aff_neg_m ~ 
+                                        (1 | id) + (1 | wave) + 
+                                        soc_med_read_w + soc_med_like_share_w + soc_med_post_w  + 
+                                        soc_med_fb_w + soc_med_ig_w + soc_med_wa_w + soc_med_yt_w + soc_med_tw_w +
+                                        soc_med_read_b + soc_med_like_share_b + soc_med_post_b + 
+                                        soc_med_fb_b + soc_med_ig_b + soc_med_wa_b + soc_med_yt_b + soc_med_tw_b +  
+                                        age + male + born_aus + born_aus_prnts + edu_fac + employment_fac +
+                                        res_vienna + acc_bal + acc_gar + home_sqm + corona_pos +
+                                        hh_income_b +
+                                        hh_adults +
+                                        hh_child18 + hh_child17 + hh_child14 + hh_child5 + hh_child2 + hh_oldfam + hh_outfam + hh_partner +
+                                        home_owner +
+                                        med_txt_kro_b + med_txt_sta_b + med_txt_pre_b + med_txt_oes_b + med_txt_kur_b + med_txt_slz_b + med_txt_son_b + 
+                                        med_vid_orf_b + med_vid_pri_b + 
+                                        risk_prop_b + loc_cntrl_int_m_b + 
+                                        act_wrk_w + act_spo_w + act_frn_w + act_sho_w + act_pet_w + 
+                                        act_wrk_b + act_spo_b + act_frn_b + act_sho_b + act_pet_b +
+                                        sat_dem_w + sat_dem_b + 
+                                        health_w + health_b +
+                                        trst_media_w + trst_police_w + trst_media_w + trst_hlthsec_w + trst_gov_w + trst_army_w +
+                                        trst_media_b + trst_police_b + trst_media_b + trst_hlthsec_b + trst_gov_b + trst_army_b
+                                        ))
+model_aff_neg_lmer_add <- summary(pool(model_aff_neg_lmer_add), conf.int = TRUE)
+model_aff_neg_lmer_add %>% 
+  as.data.frame %>% 
+  select(term, estimate, `2.5 %`, `97.5 %`, p.value) %>% 
+  mutate(p.value = td::my_round(p.value, "p"))
+```
 
-We first need to export factor scores for variables without imputed data
+# Preregistered
+
+Several changes were introduced during peer review (multiple imputation, imputation of all participants, inclusion of all variables).
+In what follows, please find the results as originally planned (but with some necessary deviations; for example not preregistered all variables could be included).
+
+ - Use data-set in which participants with >50% data were removed. 
+ - Also use factor scores (possible without multiple imputation).
+
+ Because no multiple imputation was used, output looks differently.
+
+## Life satisfaction
+
+```{r}
+model_life_sat_lmer_pre <- lmerTest::lmer(life_sat ~
+                                        (1 | id) + (1 | wave) + 
+                                        soc_med_read_w + soc_med_like_share_w + soc_med_post_w  + 
+                                        soc_med_fb_w + soc_med_ig_w + soc_med_wa_w + soc_med_yt_w + soc_med_tw_w +
+                                        soc_med_read_b + soc_med_like_share_b + soc_med_post_b + 
+                                        soc_med_fb_b + soc_med_ig_b + soc_med_wa_b + soc_med_yt_b + soc_med_tw_b +  
+                                        age + male + born_aus + born_aus_prnts + edu_fac + employment_fac +
+                                        res_vienna + acc_bal + acc_gar + home_sqm + corona_pos +
+                                        med_txt_kro_b + med_txt_sta_b + med_txt_pre_b + med_txt_oes_b + med_txt_kur_b + med_txt_slz_b + med_txt_son_b + 
+                                        med_vid_orf_b + med_vid_pri_b + 
+                                        risk_prop_b + loc_cntrl_int_m_b + 
+                                        act_wrk_w + act_spo_w + act_frn_w + act_sho_w + act_pet_w + 
+                                        act_wrk_b + act_spo_b + act_frn_b + act_sho_b + act_pet_b +
+                                        sat_dem_w + sat_dem_b + 
+                                        health_w + health_b 
+                                      , data = d_long_50_imp)
+summary(model_life_sat_lmer_pre)
+```
+
+## Positive Affect
+
+```{r}
+model_aff_pos_lmer_pre <- lmerTest::lmer(aff_pos_fs ~ 
+                                        (1 | id) + (1 | wave) + 
+                                        soc_med_read_w + soc_med_like_share_w + soc_med_post_w  + 
+                                        soc_med_fb_w + soc_med_ig_w + soc_med_wa_w + soc_med_yt_w + soc_med_tw_w +
+                                        soc_med_read_b + soc_med_like_share_b + soc_med_post_b + 
+                                        soc_med_fb_b + soc_med_ig_b + soc_med_wa_b + soc_med_yt_b + soc_med_tw_b +  
+                                        age + male + born_aus + born_aus_prnts + edu_fac + employment_fac +
+                                        res_vienna + acc_bal + acc_gar + home_sqm + corona_pos +
+                                        med_txt_kro_b + med_txt_sta_b + med_txt_pre_b + med_txt_oes_b + med_txt_kur_b + med_txt_slz_b + med_txt_son_b + 
+                                        med_vid_orf_b + med_vid_pri_b + 
+                                        risk_prop_b + loc_cntrl_int_m_b + 
+                                        act_wrk_w + act_spo_w + act_frn_w + act_sho_w + act_pet_w + 
+                                        act_wrk_b + act_spo_b + act_frn_b + act_sho_b + act_pet_b +
+                                        sat_dem_w + sat_dem_b + 
+                                        health_w + health_b 
+                                      , data = d_long_50_imp)
+summary(model_aff_pos_lmer_pre)
+```
+
+## Negative Affect
+
+```{r}
+model_aff_neg_lmer_pre <- lmerTest::lmer(aff_neg_fs ~ 
+                                        (1 | id) + (1 | wave) + 
+                                        soc_med_read_w + soc_med_like_share_w + soc_med_post_w  + 
+                                        soc_med_fb_w + soc_med_ig_w + soc_med_wa_w + soc_med_yt_w + soc_med_tw_w +
+                                        soc_med_read_b + soc_med_like_share_b + soc_med_post_b + 
+                                        soc_med_fb_b + soc_med_ig_b + soc_med_wa_b + soc_med_yt_b + soc_med_tw_b +  
+                                        age + male + born_aus + born_aus_prnts + edu_fac + employment_fac +
+                                        res_vienna + acc_bal + acc_gar + home_sqm + corona_pos +
+                                        med_txt_kro_b + med_txt_sta_b + med_txt_pre_b + med_txt_oes_b + med_txt_kur_b + med_txt_slz_b + med_txt_son_b + 
+                                        med_vid_orf_b + med_vid_pri_b + 
+                                        risk_prop_b + loc_cntrl_int_m_b + 
+                                        act_wrk_w + act_spo_w + act_frn_w + act_sho_w + act_pet_w + 
+                                        act_wrk_b + act_spo_b + act_frn_b + act_sho_b + act_pet_b +
+                                        sat_dem_w + sat_dem_b + 
+                                        health_w + health_b
+                                      , data = d_long_50_imp)
+summary(model_aff_neg_lmer_pre)
+```
+
+# Preregistered without imputation
+
+Same as before, but now don't use any imputed data whatsoever. Was preregistered as additional analysis to provide comparison.
+
+We first need to export factor scores for variables without imputed data. Note that I don't report results of mean scores to save space, as results would be highly similar.
 
@@
 "
-cfa_aff_pos <- cfa(model, d_long, group = "wave", estimator = "MLM")
-d_long$aff_pos_fs <- get_fs(cfa_aff_pos)
+cfa_aff_pos <- cfa(model, d_long_50, group = "wave", estimator = "MLM")
+d_long_50$aff_pos_fs <- get_fs(cfa_aff_pos)
 
@@
 "
-cfa_aff_neg <- cfa(model, d_long, group = "wave", estimator = "MLM")
-d_long$aff_neg_fs <- get_fs(cfa_aff_neg)
+cfa_aff_neg <- cfa(model, d_long_50, group = "wave", estimator = "MLM")
+d_long_50$aff_neg_fs <- get_fs(cfa_aff_neg)
 ```
@@
                                         soc_med_fb_b + soc_med_ig_b + soc_med_wa_b + soc_med_yt_b + soc_med_tw_b +  
-                                        age + male + born_aus + born_aus_prnts + edu_fac + employment_fac + health_b + 
-                                        res_vienna + acc_bal + acc_gar + home_sqm + 
+                                        age + male + born_aus + born_aus_prnts + edu_fac + employment_fac +
+                                        res_vienna + acc_bal + acc_gar + home_sqm + corona_pos +
                                         med_txt_kro_b + med_txt_sta_b + med_txt_pre_b + med_txt_oes_b + med_txt_kur_b + med_txt_slz_b + med_txt_son_b + 
@@
                                         act_wrk_b + act_spo_b + act_frn_b + act_sho_b + act_pet_b +
-                                        sat_dem_w + sat_dem_b 
-                                      , data = d_long)
+                                        sat_dem_w + sat_dem_b + 
+                                        health_w + health_b 
+                                      , data = d_long_50)
 summary(model_life_sat_lmer_noi)
@@
                                         act_wrk_b + act_spo_b + act_frn_b + act_sho_b + act_pet_b +
-                                        sat_dem_w + sat_dem_b 
-                                      , data = d_long)
+                                        sat_dem_w + sat_dem_b + 
+                                        health_w + health_b 
+                                      , data = d_long_50)
 summary(model_aff_pos_lmer_noi)
@@
                                         act_wrk_b + act_spo_b + act_frn_b + act_sho_b + act_pet_b +
-                                        sat_dem_w + sat_dem_b 
-                                      , data = d_long)
+                                        sat_dem_w + sat_dem_b + 
+                                        health_w + health_b
+                                      , data = d_long_50)
 summary(model_aff_neg_lmer_noi)
-```
-
-# Results with multiple imputation
-## Life satisfaction
-
-```{r}
-model_life_sat_lmer_mim <- with(d_long_mim_mice, exp =  
-                                lmerTest::lmer(life_sat ~
-                                        (1 | id) + (1 | wave) + 
-                                        soc_med_read_w + soc_med_like_share_w + soc_med_post_w  + 
-                                        soc_med_fb_w + soc_med_ig_w + soc_med_wa_w + soc_med_yt_w + soc_med_tw_w +
-                                        soc_med_read_b + soc_med_like_share_b + soc_med_post_b + 
-                                        soc_med_fb_b + soc_med_ig_b + soc_med_wa_b + soc_med_yt_b + soc_med_tw_b +  
-                                        age + male + born_aus + born_aus_prnts + edu_fac + employment_fac + health_b + 
-                                        res_vienna + acc_bal + acc_gar + home_sqm + 
-                                        med_txt_kro_b + med_txt_sta_b + med_txt_pre_b + med_txt_oes_b + med_txt_kur_b + med_txt_slz_b + med_txt_son_b + 
-                                        med_vid_orf_b + med_vid_pri_b + 
-                                        risk_prop_b + loc_cntrl_int_m_b + 
-                                        act_wrk_w + act_spo_w + act_frn_w + act_sho_w + act_pet_w + 
-                                        act_wrk_b + act_spo_b + act_frn_b + act_sho_b + act_pet_b +
-                                        sat_dem_w + sat_dem_b 
-                                      ))
-model_life_sat_lmer_mim <- summary(pool(model_life_sat_lmer_mim), conf.int = TRUE)
-```
-
-## Positive Affect
-
-For simplicity, note that we use mean-scores and not factors scores here.
-
-```{r}
-model_aff_pos_lmer_mim <- with(d_long_mim_mice, exp =  
-                           lmerTest::lmer(aff_pos_m ~ 
-                                        (1 | id) + (1 | wave) + 
-                                        soc_med_read_w + soc_med_like_share_w + soc_med_post_w + 
-                                        soc_med_fb_w + soc_med_ig_w + soc_med_wa_w + soc_med_yt_w + soc_med_tw_w +
-                                        soc_med_read_b + soc_med_like_share_b + soc_med_post_b + 
-                                        soc_med_fb_b + soc_med_ig_b + soc_med_wa_b + soc_med_yt_b + soc_med_tw_b +  
-                                        age + male + born_aus + born_aus_prnts + edu_fac + employment_fac + health_b + 
-                                        res_vienna + acc_bal + acc_gar + home_sqm + 
-                                        med_txt_kro_b + med_txt_sta_b + med_txt_pre_b + med_txt_oes_b + med_txt_kur_b + med_txt_slz_b + med_txt_son_b + 
-                                        med_vid_orf_b + med_vid_pri_b + 
-                                        risk_prop_b + loc_cntrl_int_m_b + 
-                                        act_wrk_w + act_spo_w + act_frn_w + act_sho_w + act_pet_w + 
-                                        act_wrk_b + act_spo_b + act_frn_b + act_sho_b + act_pet_b +
-                                        sat_dem_w + sat_dem_b)
-)
-model_aff_pos_lmer_mim <- summary(pool(model_aff_pos_lmer_mim), conf.int = TRUE)
-```
-
-## Negative Affect
-
-For simplicity, note that we use mean-scores and not factors scores here.
-
-```{r}
-model_aff_neg_lmer_mim <- with(d_long_mim_mice, exp =  
-                               lmerTest::lmer(aff_neg_m ~ 
-                                        (1 | id) + (1 | wave) + 
-                                        soc_med_read_w + soc_med_like_share_w + soc_med_post_w + 
-                                        soc_med_fb_w + soc_med_ig_w + soc_med_wa_w + soc_med_yt_w + soc_med_tw_w +
-                                        soc_med_read_b + soc_med_like_share_b + soc_med_post_b + 
-                                        soc_med_fb_b + soc_med_ig_b + soc_med_wa_b + soc_med_yt_b + soc_med_tw_b +  
-                                        age + male + born_aus + born_aus_prnts + edu_fac + employment_fac + health_b + 
-                                        res_vienna + acc_bal + acc_gar + home_sqm + 
-                                        med_txt_kro_b + med_txt_sta_b + med_txt_pre_b + med_txt_oes_b + med_txt_kur_b + med_txt_slz_b + med_txt_son_b + 
-                                        med_vid_orf_b + med_vid_pri_b + 
-                                        risk_prop_b + loc_cntrl_int_m_b + 
-                                        act_wrk_w + act_spo_w + act_frn_w + act_sho_w + act_pet_w + 
-                                        act_wrk_b + act_spo_b + act_frn_b + act_sho_b + act_pet_b +
-                                        sat_dem_w + sat_dem_b)
-)
-model_aff_neg_lmer_mim <- summary(pool(model_aff_neg_lmer_mim), conf.int = TRUE)
 ```
@@
 
-In what follows, see figure with all results combined.
+ - In what follows, see figure with all results combined. 
+ - Results show that different models are very comparable. 
+ - The model without control variables stands a bit out, showing that controlling for additional varying variables is important.
+ - The preregistered model without imputation also has larger confidence intervals, which was to be expected given fewer data.
 
-## Activities
+## Within
 
-```{r}
+```{r warning=F, fig.height=5}
 # get data
-dat_fig_results_activity <- get_dat_res(model_aff_neg_lmer, model_aff_pos_lmer, model_life_sat_lmer, type = "activity", analysis = "1. regular")
-dat_fig_results_activity_nco <- get_dat_res(model_aff_neg_lmer_nco, model_aff_pos_lmer_nco, model_life_sat_lmer_nco, type = "activity", analysis = "2. no covars")
-dat_fig_results_activity_m <- get_dat_res(model_aff_neg_lmer_m, model_aff_pos_lmer_m, model_life_sat_lmer, type = "activity", analysis = "3. mean scores")
-dat_fig_results_activity_noi <- get_dat_res(model_aff_neg_lmer_noi, model_aff_pos_lmer_noi, model_life_sat_lmer_noi, type = "activity", analysis = "4. no imp")
-dat_fig_results_activity_mim <- get_dat_res(model_aff_neg_lmer_mim, model_aff_pos_lmer_mim, model_life_sat_lmer_mim, type = "activity", analysis = "5. mult imp")
+dat_fig_results_activity_pub <- get_dat_res(
+  model_aff_neg_lmer_pub, model_aff_pos_lmer_pub, model_life_sat_lmer_pub, 
+  type = "activity", analysis = "1. preregistered")
 
-dat_fig_results_activity <- rbind(
-  dat_fig_results_activity,
-  dat_fig_results_activity_nco,
-  dat_fig_results_activity_m,
-  dat_fig_results_activity_noi,
-  dat_fig_results_activity_mim
+dat_fig_results_channel_pub <- get_dat_res(
+  model_aff_neg_lmer_pub, model_aff_pos_lmer_pub, model_life_sat_lmer_pub, 
+  type = "channels", analysis = "1. preregistered")
+
+dat_fig_results_within_pub <- rbind(
+  cbind(dat_fig_results_activity_pub, comm = "activity"), 
+  cbind(dat_fig_results_channel_pub, comm = "channel")) %>% 
+  filter(type == "within")
+
+dat_fig_results_activity_nco <- get_dat_res(
+  model_aff_neg_lmer_nco, model_aff_pos_lmer_nco, model_life_sat_lmer_nco, 
+  type = "activity", analysis = "2. no covars")
+
+dat_fig_results_channel_nco <- get_dat_res(
+  model_aff_neg_lmer_nco, model_aff_pos_lmer_nco, model_life_sat_lmer_nco, 
+  type = "channels", analysis = "2. no covars")
+
+dat_fig_results_within_nco <- rbind(
+  cbind(dat_fig_results_activity_nco, comm = "activity"), 
+  cbind(dat_fig_results_channel_nco, comm = "channel")) %>% 
+  filter(type == "within")
+
+dat_fig_results_activity_add <- get_dat_res(
+  model_aff_neg_lmer_add, model_aff_pos_lmer_add, model_life_sat_lmer_add, 
+  type = "activity", analysis = "3. additional covars")
+
+dat_fig_results_channel_add <- get_dat_res(
+  model_aff_neg_lmer_add, model_aff_pos_lmer_add, model_life_sat_lmer_add, 
+  type = "channels", analysis = "3. additional covars")
+
+dat_fig_results_within_add <- rbind(
+  cbind(dat_fig_results_activity_add, comm = "activity"), 
+  cbind(dat_fig_results_channel_add, comm = "channel")) %>% 
+  filter(type == "within")
+
+dat_fig_results_activity_pre <- get_dat_res(
+  model_aff_neg_lmer_pre, model_aff_pos_lmer_pre, model_life_sat_lmer_pre, 
+  type = "activity", analysis = "4. preregistered")
+
+dat_fig_results_channel_pre <- get_dat_res(
+  model_aff_neg_lmer_pre, model_aff_pos_lmer_pre, model_life_sat_lmer_pre, 
+  type = "channels", analysis = "4. preregistered")
+
+dat_fig_results_within_pre <- rbind(
+  cbind(dat_fig_results_activity_pre, comm = "activity"), 
+  cbind(dat_fig_results_channel_pre, comm = "channel")) %>% 
+  filter(type == "within")
+
+dat_fig_results_activity_noi <- get_dat_res(
+  model_aff_neg_lmer_noi, model_aff_pos_lmer_noi, model_life_sat_lmer_noi, 
+  type = "activity", analysis = "5. prereg no imp")
+
+dat_fig_results_channel_noi <- get_dat_res(
+  model_aff_neg_lmer_noi, model_aff_pos_lmer_noi, model_life_sat_lmer_noi, 
+  type = "channels", analysis = "5. prereg no imp")
+
+dat_fig_results_within_noi <- rbind(
+  cbind(dat_fig_results_activity_noi, comm = "activity"), 
+  cbind(dat_fig_results_channel_noi, comm = "channel")) %>% 
+  filter(type == "within")
+
+dat_fig_results_within_comp <- rbind(
+  dat_fig_results_within_pub,
+  dat_fig_results_within_nco,
+  dat_fig_results_within_add,
+  dat_fig_results_within_pre,
+  dat_fig_results_within_noi
 )
 # make figure
-fig_results_activity <- make_graph_res(
-  dat_fig_results_activity,
-  sesoi = "est"
+fig_results_within_comp <- make_graph_res(
+  dat_fig_results_within_comp,
+  sesoi = "est",
+  facet = "comm"
   )
-fig_results_activity
+fig_results_within_comp
 
 # safe figure
-ggsave("figures/fig_results_activity.pdf", 
-       width = 7, height = 4,
-       plot = fig_results_activity)
+ggsave("figures/fig_results_within_comparison.pdf", 
+       width = 7, height = 5,
+       plot = fig_results_within_comp)
 ```
 
-## Channels
+## Between
 
-```{r}
+```{r warning=FALSE, fig.height=5}
 # get data
-dat_fig_results_channels <- get_dat_res(model_aff_neg_lmer, model_aff_pos_lmer, model_life_sat_lmer, type = "channels", analysis = "1. regular")
-dat_fig_results_channels_nco <- get_dat_res(model_aff_neg_lmer_nco, model_aff_pos_lmer_nco, model_life_sat_lmer_nco, type = "channels", analysis = "2. no covars")
-dat_fig_results_channels_m <- get_dat_res(model_aff_neg_lmer_m, model_aff_pos_lmer_m, model_life_sat_lmer, type = "channels", analysis = "3. mean scores")
-dat_fig_results_channels_noi <- get_dat_res(model_aff_neg_lmer_noi, model_aff_pos_lmer_noi, model_life_sat_lmer_noi, type = "channels", analysis = "4. no imp")
-dat_fig_results_channels_mim <- get_dat_res(model_aff_neg_lmer_mim, model_aff_pos_lmer_mim, model_life_sat_lmer_mim, type = "channels", analysis = "5. mult imp")
+dat_fig_results_activity_pub <- get_dat_res(
+  model_aff_neg_lmer_pub, model_aff_pos_lmer_pub, model_life_sat_lmer_pub, 
+  type = "activity", analysis = "1. preregistered")
 
-dat_fig_results_channels <- rbind(
-  dat_fig_results_channels,
-  dat_fig_results_channels_nco,
-  dat_fig_results_channels_m,
-  dat_fig_results_channels_noi,
-  dat_fig_results_channels_mim
+dat_fig_results_channel_pub <- get_dat_res(
+  model_aff_neg_lmer_pub, model_aff_pos_lmer_pub, model_life_sat_lmer_pub, 
+  type = "channels", analysis = "1. preregistered")
+
+dat_fig_results_between_pub <- rbind(
+  cbind(dat_fig_results_activity_pub, comm = "activity"), 
+  cbind(dat_fig_results_channel_pub, comm = "channel")) %>% 
+  filter(type == "between")
+
+dat_fig_results_activity_nco <- get_dat_res(
+  model_aff_neg_lmer_nco, model_aff_pos_lmer_nco, model_life_sat_lmer_nco, 
+  type = "activity", analysis = "2. no covars")
+
+dat_fig_results_channel_nco <- get_dat_res(
+  model_aff_neg_lmer_nco, model_aff_pos_lmer_nco, model_life_sat_lmer_nco, 
+  type = "channels", analysis = "2. no covars")
+
+dat_fig_results_between_nco <- rbind(
+  cbind(dat_fig_results_activity_nco, comm = "activity"), 
+  cbind(dat_fig_results_channel_nco, comm = "channel")) %>% 
+  filter(type == "between")
+
+dat_fig_results_activity_add <- get_dat_res(
+  model_aff_neg_lmer_add, model_aff_pos_lmer_add, model_life_sat_lmer_add, 
+  type = "activity", analysis = "3. additional covars")
+
+dat_fig_results_channel_add <- get_dat_res(
+  model_aff_neg_lmer_add, model_aff_pos_lmer_add, model_life_sat_lmer_add, 
+  type = "channels", analysis = "3. additional covars")
+
+dat_fig_results_between_add <- rbind(
+  cbind(dat_fig_results_activity_add, comm = "activity"), 
+  cbind(dat_fig_results_channel_add, comm = "channel")) %>% 
+  filter(type == "between")
+
+dat_fig_results_activity_pre <- get_dat_res(
+  model_aff_neg_lmer_pre, model_aff_pos_lmer_pre, model_life_sat_lmer_pre, 
+  type = "activity", analysis = "4. preregistered")
+
+dat_fig_results_channel_pre <- get_dat_res(
+  model_aff_neg_lmer_pre, model_aff_pos_lmer_pre, model_life_sat_lmer_pre, 
+  type = "channels", analysis = "4. preregistered")
+
+dat_fig_results_between_pre <- rbind(
+  cbind(dat_fig_results_activity_pre, comm = "activity"), 
+  cbind(dat_fig_results_channel_pre, comm = "channel")) %>% 
+  filter(type == "between")
+
+dat_fig_results_activity_noi <- get_dat_res(
+  model_aff_neg_lmer_noi, model_aff_pos_lmer_noi, model_life_sat_lmer_noi, 
+  type = "activity", analysis = "5. prereg no imp")
+
+dat_fig_results_channel_noi <- get_dat_res(
+  model_aff_neg_lmer_noi, model_aff_pos_lmer_noi, model_life_sat_lmer_noi, 
+  type = "channels", analysis = "5. prereg no imp")
+
+dat_fig_results_between_noi <- rbind(
+  cbind(dat_fig_results_activity_noi, comm = "activity"), 
+  cbind(dat_fig_results_channel_noi, comm = "channel")) %>% 
+  filter(type == "between")
+
+dat_fig_results_between_comp <- rbind(
+  dat_fig_results_between_pub,
+  dat_fig_results_between_nco,
+  dat_fig_results_between_add,
+  dat_fig_results_between_pre,
+  dat_fig_results_between_noi
 )
 # make figure
-fig_results_channels <- make_graph_res(
-  dat_fig_results_channels,
-  sesoi = "est"
+fig_results_between_comp <- make_graph_res(
+  dat_fig_results_between_comp,
+  sesoi = "est",
+  facet = "comm"
   )
-fig_results_channels
+fig_results_between_comp
 
 # safe figure
-ggsave("figures/fig_results_channels.pdf", 
-       width = 7, height = 4,
-       plot = fig_results_channels)
+ggsave("figures/fig_results_between_comparison.pdf", 
+       width = 7, height = 5,
+       plot = fig_results_between_comp)
 ```
 
-Save results so that we can extract them in the manuscript.
-
-```{r}
+```{r eval=F}
 save.image("data/workspace_2.RData")

### 2022-07-25 index.Rmd
- major updates for revision - included multiple imputation - changed theory - include new data-set - extended discussion

@@
 ---
-title: "Analyzing the effects of COVID-19 related social media use on well-being"
+title: "The effects of COVID-19 related social media use on well-being"
 ---
 
-On this website, you can find the companion material for the paper _Analyzing the effects of COVID-19 related social media use on well-being_.
+On this website, you can find the companion material for the paper _The effects of COVID-19 related social media use on well-being_.
 

### 2022-07-25 manuscript.Rmd
- major updates for revision - included multiple imputation - changed theory - include new data-set - extended discussion

@@
 ---
-title             : "Analyzing the effects of COVID-19 related social media use on well-being"
+title             : "The effects of COVID-19 related social media use on well-being"
 shorttitle        : "COVID-19 related social media use and well-being"
@@
 
-abstract          : "In times of crisis such as the Corona pandemic citizens need to stay informed about recent events, political decisions, or mandatory protection measures.
+abstract          : "In times of crisis such as the COVID-19 pandemic citizens need to stay informed about recent events, political decisions, or mandatory protection measures.
 To this end, many people use various types of media, and increasingly social media.
@@
 In this preregistered study, I investigated whether using social media for COVID-19 related reasons affected psychological well-being.
-To answer this question I analyzed data from the Austrian Corona Panel Project, which consists of 3,018 participants.
-Well-being was measured at all 24 waves, and communication at five specific waves.
+To answer this question I analyzed data from the Austrian Corona Panel Project, which consists of 3,485 participants.
+Well-being was measured at all 32 waves, and communication at six specific waves.
 I ran three random effects within between models, controlling for several stable and varying confounders.
@@
 
-documentclass     : "apa6"
-classoption       : "man"
-output            :
-  papaja::apa6_pdf:
-    latex_engine  : xelatex
+header-includes:
+  - |
+    \makeatletter
+    \renewcommand{\paragraph}{\@startsection{paragraph}{4}{\parindent}%
+      {0\baselineskip \@plus 0.2ex \@minus 0.2ex}%
+      {-1em}%
+      {\normalfont\normalsize\bfseries\typesectitle}}
+
+    \renewcommand{\subparagraph}[1]{\@startsection{subparagraph}{5}{1em}%
+      {0\baselineskip \@plus 0.2ex \@minus 0.2ex}%
+      {-\z@\relax}%
+      {\normalfont\normalsize\bfseries\itshape\hspace{\parindent}{#1}\textit{\addperi}}{\relax}}
+    \makeatother
+    \usepackage{endnotes}
+    \setlength{\parskip}{0em}
+    \raggedbottom
+    \note{\clearpage}
+
+csl               : "`r system.file('rmd', 'apa7.csl', package = 'papaja')`"
+documentclass     : "apa7"
+# output            :
+#   papaja::apa6_pdf:
+#     latex_engine  : xelatex
 # output            : papaja::apa6_docx
-header-includes:
-  - \usepackage{endnotes}
-  - \setlength{\parskip}{0em}
-  - \raggedbottom
-  - \note{\clearpage}
+output            : html_document
 ---
@@
 ```{r analysis-preferences, include=F, cache=F}
-# Seed for random number generation
-set.seed(42)
+set.seed(170819) # Seed for random number generation
 knitr::opts_chunk$set(cache.extra = knitr::rand_seed,
@@
               "lavaan", "lme4", "lmerTest",
-              "magrittr", "semPlot", "tidyverse")
-lapply(c(packages, "papaja", "td"), library, character.only = TRUE)
+              "magrittr", "papaja", "semPlot", "tidyverse")
+lapply(c(packages, "td"), library, character.only = TRUE)
 
 # load workspace created from analyses.rmd to use results
-load("data/workspace_2.RData")
+load("data/workspace_1.RData")
 ```
@@
 To this end, I analyzed a large-scale panel study from the Austrian Corona Panel Project [@kittelAustrianCoronaPanel2020]. 
-The panel consists of 24 waves and has an overall sample size of `r nrow(d_wide) %>% prettyNum(big.mark = ",")` participants.
+The panel consists of **32** waves and has an overall sample size of **`r nrow(d_wide) %>% prettyNum(big.mark = ",")` participants**.
 <!-- At least 1,500 participants took part per wave, and the sample is representative of the Austrian population. -->
@@
 Whereas the first four levels focus on the communication _channel_, the last two address the communication _type_.
-To measure social media use for the consumption of COVID-19 related news and topics, I here employ both the channel and the type of communication perspective, which together should offer a nuanced understanding of communication.
+**Differentiating different communication channels and communication types is important, because the effects on well-being likely differ across communication channels and communication types.**
+**Whereas active social media use such as chatting is routinely linked to improved well-being, passive use such as reading is considered more negative [@dienlinImpactDigitalTechnology2020].**
+Similarly, branded apps are separate entities with potentially divergent effects. 
+For example, @waterlooNormsOnlineExpressions2018 found that it's more adequate to express negative emotions on WhatsApp than on Twitter or on Instagram. 
+**Especially during a pandemic, it makes sense to analyze if users engage with COVID-19 related content on Instagram, where communication is more positive, or on Facebook, where communication is more critical.**
 
+**In this study, to measure the effects of social media use focused on COVID-19 related news and topics, I adopt both the channel and the type of communication perspective.**
+**Together, this should offer a nuanced and comprehensive understanding of communication.**
 First, I investigate how well-being is affected by different types of communication affect, namely active and passive use.
@@
 Second, I analyze how using the most prominent branded applications affects well-being, and whether this effect changes across applications.
-Branded apps are separate entities with potentially divergent effects. 
-Twitter might have a different effect as compared to WhatsApp because of their respective affordances.
-For example, @waterlooNormsOnlineExpressions2018 found that it's more adequate to express negative emotions on WhatsApp than on Twitter or on Instagram. 
-The branded applications investigated here are Facebook, Twitter, Instagram, WhatsApp, and YouTube.
-Worth noting, this study is not about _general_ social media use during times of COVID, but on social media use _focused_ on COVID-19 related content. 
-Examples of such media use include posting thoughts about the pandemic or retweeting COVID-19 related news.
+The branded applications investigated here are Facebook, Twitter, Instagram, WhatsApp, and YouTube---which were, at the time of writing, the most relevant social media apps.
+
+**Worth noting, this study is not about _general_ social media use during times of COVID, but on social media use _focused_ on COVID-19 related content.** 
+**For example, posting thoughts about the pandemic, reading posts and comments, or retweeting COVID-19 related news.**
 
@@
 In advance, there doesn't seem to be a clear winner, and it's likely that both positive and negative effects cancel each other out.
-Empirically, social media use on average oes not have a strong effect on well-being [@meierComputermediatedCommunicationSocial2020a].
-According to the Differential Susceptibility to Media Effects Model [@valkenburgDifferentialSusceptibilityMedia2013], the effects of media use differ across individuals. 
+Empirically, social media use on average does not have a strong effect on well-being [@meierComputermediatedCommunicationSocial2020a].
+According to the Differential Susceptibility to Media Effects Model [@valkenburgDifferentialSusceptibilityMedia2013], the effects of media use differ across individuals **and content**.
 Whereas for some media are beneficial, for others they are harmful.
+**Whereas some content provides opportunities (education, advice), other content creates risks (misinformation, hate). [@livingstoneEuropeanResearchChildren2018].**
+Social media can impair well-being when causing embarrassment, stress, or disinformation, and they can improve well-being when providing connectedness, information, or entertainment [@buchiDigitalWellbeingTheory2021].
 On average, however, effects are often small or negligible.
-But also when focusing on individual users, social media can have both positive and negative effects on well-being [@buchiDigitalWellbeingTheory2021].
-They can impair well-being when causing embarrassment, stress, or disinformation, and they can improve well-being when providing connectedness, information, or entertainment [@buchiDigitalWellbeingTheory2021].
 
-Two of the most prominent media effect theories argue against strong negative impacts.
-First, uses and gratifications theory states that people explicitly and rationally chose specific media because of their respective benefits [@katzUsesGratificationsResearch1973]. 
-If those benefits don't exist, they will spend their time elsewhere.
-And social media offer copiuos benefits. 
-Most prominently, they help find relevant information, maintain and foster relationships, express one's personality, and entertain oneself [@pelletierOneSizeDoesn2020].
-The second major theory that needs to be listed here is mood management theory [@zillmannMoodManagementCommunication1988].
-Users implicitly learn what media help them balance their mood and affect.
-For example, when bored people use social media to entertain themselves.
-Using experience sampling of well-being and logs of social media use, a study with 82 participants from Italy found that after episodes of social media use, levels of positive affect increased significantly [@marcianoDynamicsAdolescentsSmartphone2022].
+**Two of the most prominent media effect theories argue (indirectly) against strong average negative impacts.**
+**According to mood management theory [@zillmannMoodManagementCommunication1988], using media substantially affects people's moods.**
+**Effects can be stimulating or overwhelming, relaxing or boring.**
+**After some time, users implicitly learn what media help them balance their mood and affect according to their own situational needs [@zillmannMoodManagementCommunication1988].**
+**Those media that eventually become part of one's media repertoire are hence, on average, beneficial for users and their moods.**
+**Using experience sampling of well-being and logs of social media use, a study with 82 participants from Italy found that after episodes of social media use, levels of positive affect increased significantly [@marcianoDynamicsAdolescentsSmartphone2022].**
+
+**While mood management theory considers media use mainly driven by implicit learning experiences, uses and gratifications theory upholds that the process is more explicit and rational [@katzUsesGratificationsResearch1973].**
+**Users select those media that they expect to have a desired effect, for example on mood, knowledge, or entertainment.** 
+**If those beneficial media effects do not exist or if they are not expected, people will spend their time elsewhere.**
+**And social media offer several beneficial effects, explaining why they are used that much.**
+**They help find relevant information, maintain and foster relationships, express one's personality, and entertain oneself [@pelletierOneSizeDoesn2020].**
+**In conclusion, because people spend so much time on social media consuming COVID-19 related content, according to both mood management theory and uses and gratifications theory we wouldn't expect to find strong average negative effects.**
 
@@
 
-Because media effects can differ across situations [@valkenburgDifferentialSusceptibilityMedia2013], I now briefly focus on the effects of COVID-19 related social media use specifically.
+**Because media effects can differ across users, situations, and content [@livingstoneEuropeanResearchChildren2018; @valkenburgDifferentialSusceptibilityMedia2013], I now briefly focus on the effects of COVID-19 related social media use specifically.**
 First, one could assume a _direct_ negative effect on well-being, and especially on positive or negative affect, which are more volatile and fluctuating.
@@
 A study with 2.057 respondents from Italy reported that during the pandemic virtual community and social connectedness even increased during the pandemic [@guazziniSecondWaveAnalysis2022].
+**In Finland, in a sample of 735 participants, levels of loneliness did not decrease during the pandemic, and people who engaged more on social media showed less loneliness [@latikkaLonelinessPsychologicalDistress2022].**
 <!-- However, in light of a worldwide pandemic with millions of deaths, the negative direct effect seems more plausible. -->
@@
 To explain, overuse could impair well-being if it replaces meaningful or functional activities such as meeting others, working, actively relaxing, or exercising.
-So if a society collectively overuses social media during a pandemic, there might be potential for negative effects.
+**Another potentially negative mechanism at play are problematic social comparison processes.
+During the pandemic, several users shared how they successfully dealt with challenges such as physical distancing.
+In a study with 1131 residents from Wuhan in China [@yuePassiveSocialMedia2022], people who spent more time in quarantine also spent more time on social media.
+Those, who spent more time on social media also engaged in more upward social comparison, which was related to increased levels of stress.**
+<!-- Therefore, if a society collectively overuses social media during a pandemic, there is potential for negative effects. -->
 <!-- If overuse replaces such activities it's reasonable to assume that it's also detrimental. -->
-On the other hand, one can make the case that overuse might even be beneficial, especially in times of a pandemic.
+
+On the other hand, one can make the case that using social media for COVID-19 related reasons might even be beneficial, especially in times of a pandemic.
 Exchanging COVID-19 related messages with friends via WhatsApp might replace the in-person contact one would have otherwise, but which is literally impossible at the time.
@@
 In conclusion, in light of the theoretical considerations and empirical studies presented above, I expect that COVID-19 related communication on social media doesn't affect well-being in a meaningful or relevant way.
+**This general hypothesis will be analyzed specifically for the communication types of (a) time spent reading, (b) liking and sharing, and (c) actively posting COVID-19 related content.**
+**In addition, I will analyze how well-being is influenced by spending time on five prominent social media apps, including (a) Facebook, (b) Instagram, (c) Twitter, (d) WhatsApp, and (e) YouTube.** 
+**Three different well-being indicators will be differentiated: life satisfaction, positive affect, and negative affect.**
 
@@
 Using longitudinal designs alone, however, is not sufficient for establishing correct causal statements [@rohrerTheseAreNot2021].
-In addition, we for example also need to control for confounding third variables, and importantly also for _varying_ third variables.
+In addition, we for example also need to control for confounding third variables.
+**Importantly, when analyzing longitudinal (within-person) relationships and effects, it is important to control for _varying_ third variables.**
+**Non-varying third variables only help control non-varying (between-person) relations.**
 
@@
 However, despite all these caveats, when controlling for relevant variables (that aren't mediators), we can be much more certain that we measured causality correctly.
-The aim should therefore be to collect as many varying and nonvarying confounders as possible (which I believe is seldom done in our field), while knowing that absolute certainty regarding causality cannot be reached.
+The aim should therefore be to collect as many varying and non-varying confounders as possible (which I believe is seldom done in our field), while knowing that absolute certainty regarding causality cannot be reached.
 
@@
 
-[^explore]: The dataset includes many other variables that one could also potentially control for, and I invite interested readers to download the and explore potential interesting relationships.
+[^explore]: The data-set includes many other variables that one could also potentially control for, and I invite interested readers to download the and explore potential interesting relationships.
 
@@
 If variables are stable, longer intervals are needed; if they fluctuate, shorter intervals.
-In the case of well-being, we need shorter intervals for affect and longer ones for life satisfaction [@dienlinImpactDigitalTechnology2020].
-Still, choosing the right interval is challenging, because especially short intervals are hard to implement in practice and often require advanced methods such as experience sampling (also known as in situ measurement or ambulant assessment) [@schnauber-stockmannMobileDevicesTools2020].
-
-In this study, I hence analyze how using social media during the last week affected their positive and negative affect during the same week.
-More precisely, if people during the last week used COVID-19 related social media more than the usually do, did they feel better or worse during that week than they usually do?
+In the case of well-being, we need shorter intervals for **the more fluctuating** positive and negative affect and longer ones for **the more stable** life satisfaction [@dienlinImpactDigitalTechnology2020].
+**Using social media can have instant effects on mood [@marcianoDynamicsAdolescentsSmartphone2022].**
+**Effects on life satisfaction often take longer to manifest, for example because media use leads to actual changes in specific behaviors, which then affect life satisfaction [@dienlinDisplacementReinforcementReciprocity2017].**
+Choosing the right interval is challenging, because especially short intervals are hard to implement in practice, often requiring advanced methods such as experience sampling (also known as in situ measurement or ambulant assessment) [@schnauber-stockmannMobileDevicesTools2020].
+In this study, I hence analyze how using social media during the last week affected positive and negative affect during the same week.
+In other words, if people during the last week engaged in more COVID-19 related social media use than the usually do, did they feel better or worse during that week than they usually do?
 Regarding life satisfaction, I implemented a longer interval.
-Namely, if people during the last week used COVID-19 related social media more than the usually do, were they at the end of the week more or less satisfied with their lives than they usually are?
+If people during the last week used COVID-19 related social media more than the usually do, were they at the end of the week more or less satisfied with their lives than they usually are?
 I hence analyze if when a person changes their social media diet, are there (a) _simultaneous_ changes in their affect and (b) _subsequent_ changes in their life satisfaction?
-When additionally controlling for both stable and varying confounders, we can then be more sure that the effect is indeed causal.
+**These relations will be controlled for varying confounders, which fosters a causal interpretation.**
 Similar approaches were implemented by other studies [@johannesNoEffectDifferent2022; @scharkowHowSocialNetwork2020], and they are considered a best practice approach toward analyzing causality [@bellFixedRandomEffects2019].
@@
 
-The data come from the Austrian Corona Panel Project [@kittelAustrianCoronaPanel2021].
-The study was conducted between March 2020 and October 2021.
-It contains 26 waves, and at the time of writing the first 24 waves were available for download.
+**The data come from the Austrian Corona Panel Project [@kittelAustrianCoronaPanel2021], which is a large-scale standalone panel study. **
+**The data are hosted on AUSSDA and are publicly available here: https://doi.org/10.11587/28KQNS.**
+**At the time of writing, the official website included a data-set consisting of 24 waves. **
+**For the analyses presented here, I was able to receive an advance version consisting of all 32 waves.**
+**The study was conducted between March 2020 and June 2022, and data collection is now officially finished.**
+**It contains 32 waves.**
 Between March 2020 and July 2020, the intervals were weekly, and afterward the intervals were monthly.
@@
 Because Mardia’s test showed that the assumption of multivariate normality was violated, I used the more robust Satorra-Bentler scaled and mean-adjusted test statistic (MLM) as estimator.
-To avoid overfitting, I tested the scales on more liberal fit criteria (CFI > .90, TLI > .90, RMSEA <. .10, SRMR < .10) [@klinePrinciplesPracticeStructural2016].
-Finally, REWB-models cannot model latent variables.
-To increase precision, I therefore exported factor scores from the CFAs for positive and negative affect.
-Respondents who answered less than 50% of all questions were removed.
-The remaining missing responses were imputed using predictive mean matching.
+To avoid over-fitting, I tested the scales on more liberal fit criteria (CFI > .90, TLI > .90, RMSEA <. .10, SRMR < .10) [@klinePrinciplesPracticeStructural2016].
+<!-- Finally, REWB-models cannot model latent variables. -->
+<!-- To increase precision, I therefore exported factor scores from the CFAs for positive and negative affect. -->
+Mean scores were used for positive and negative affect.
+<!-- Respondents who answered less than 50% of all questions were removed. -->
+**Missing responses were imputed using multiple imputation with predictive mean matching (five iterations, five data-sets).**
+**Predictive mean matching allowed to impute also categorical variables.**
+**All variables were imputed except the media use measures, as they were not collected on each wave.**
+**All variables included in the analyses presented here were used to impute missing data.**
+**For the main analyses, results were pooled across all five data-sets.**
 
@@
 
-Social media use was measured for all participants on waves 1, 2, 8, 17, and 23. 
+Social media use was measured for all participants on waves 1, 2, 8, 17, 23, **and 28**.
 Freshly recruited respondents always answered all questions on social media use.
@@
 
-First, when looking at the variables from a descriptive perspective (Figure \@ref(fig:descriptives)), we see that all well-being measures did not change substantially across the different waves of data collection.
+First, when looking at the variables from a descriptive perspective (Figure \@ref(fig:fig-descriptives)), we see that all well-being measures did not change substantially across the different waves of data collection.
 COVID-19 related media use, however, decreased slightly at the beginning of the study and remained stable after approximately six waves.
@@
 sharing vs.
-posting---all within-person effects fell completely within the a-priori defined null region (see Figure \@ref(fig:res-activity)).
-For example, respondents who used social media more frequently than usual to read about COVID-19 related topics did not show a simultaneous change in life satisfaction (_b_ = `r dat_fig_results_activity %>% filter(dv == "Life satisfaction" & iv == "Reading" & type == "within" & Analysis == "1.
-regular") %>% select(estimate) %>% round(2)` [95% CI `r dat_fig_results_activity %>% filter(dv == "Life satisfaction" & iv == "Reading" & type == "within" & Analysis == "1.
-regular") %>% select(conf.low) %>% round(2)`, `r dat_fig_results_activity %>% filter(dv == "Life satisfaction" & iv == "Reading" & type == "within" & Analysis == "1.
-regular") %>% select(conf.high) %>% round(2)`]).
-All confidence intervals included zero; hence, all effects were also statistically non-significant.
+posting---all within-person effects fell completely within the a-priori defined null region (see Figure \@ref(fig:fig-between)).
+For example, respondents who used social media more frequently than usual to read about COVID-19 related topics did not show a simultaneous change in life satisfaction (_b_ = `r dat_fig_results_within %>% filter(dv == "Life satisfaction" & iv == "Reading" & type == "within" & Analysis == "Regular") %>% select(estimate) %>% round(2)` [95% CI `r dat_fig_results_within %>% filter(dv == "Life satisfaction" & iv == "Reading" & type == "within" & Analysis == "Regular") %>% select(conf.low) %>% round(2)`, `r dat_fig_results_within %>% filter(dv == "Life satisfaction" & iv == "Reading" & type == "within" & Analysis == "Regular") %>% select(conf.high) %>% round(2)`]).
 As a result, the hypothesis was supported for all COVID-19 related types of social media communication.
+**However, two effects were statistically significantly different from zero.**
+**Users who wrote more COVID-19 related posts than usual were also slightly less satisfied with their lives as usual (_b_ = `r dat_fig_results_within %>% filter(dv == "Life satisfaction" & iv == "Posting" & type == "within" & Analysis == "Regular") %>% select(estimate) %>% round(2)` [95% CI `r dat_fig_results_within %>% filter(dv == "Life satisfaction" & iv == "Posting" & type == "within" & Analysis == "Regular") %>% select(conf.low) %>% round(2)`, `r dat_fig_results_within %>% filter(dv == "Life satisfaction" & iv == "Posting" & type == "within" & Analysis == "Regular") %>% select(conf.high) %>% round(2)`]).**
+**Users who wrote more COVID-19 related posts than usual also experienced slightly more negative affect than usual (_b_ = `r dat_fig_results_within %>% filter(dv == "Negative affect" & iv == "Posting" & type == "within" & Analysis == "Regular") %>% select(estimate) %>% round(2)` [95% CI `r dat_fig_results_within %>% filter(dv == "Negative affect" & iv == "Posting" & type == "within" & Analysis == "Regular") %>% select(conf.low) %>% round(2)`, `r dat_fig_results_within %>% filter(dv == "Negative affect" & iv == "Posting" & type == "within" & Analysis == "Regular") %>% select(conf.high) %>% round(2)`]).**
+**There was a small and statistically nonsignificant trend that reading COVID-19 related content slightly increased life satisfaction (_b_ = `r dat_fig_results_within %>% filter(dv == "Life satisfaction" & iv == "Reading" & type == "within" & Analysis == "Regular") %>% select(estimate) %>% round(2)` [95% CI `r dat_fig_results_within %>% filter(dv == "Life satisfaction" & iv == "Reading" & type == "within" & Analysis == "Regular") %>% select(conf.low) %>% round(2)`, `r dat_fig_results_within %>% filter(dv == "Life satisfaction" & iv == "Reading" & type == "within" & Analysis == "Regular") %>% select(conf.high) %>% round(2)`], _p_ = `r dat_fig_results_within %>% filter(dv == "Life satisfaction" & iv == "Reading" & type == "within" & Analysis == "Regular") %>% select(p.value) %>% my_round("p")`).**
+**At the same time, there was also a small and statistically nonsignificant trend that reading COVID-19 related content decreased positive affect (_b_ = `r dat_fig_results_within %>% filter(dv == "Positive affect" & iv == "Reading" & type == "within" & Analysis == "Regular") %>% select(estimate) %>% round(2)` [95% CI `r dat_fig_results_within %>% filter(dv == "Positive affect" & iv == "Reading" & type == "within" & Analysis == "Regular") %>% select(conf.low) %>% round(2)`, `r dat_fig_results_within %>% filter(dv == "Positive affect" & iv == "Reading" & type == "within" & Analysis == "Regular") %>% select(conf.high) %>% round(2)`], _p_ = `r dat_fig_results_within %>% filter(dv == "Positive affect" & iv == "Reading" & type == "within" & Analysis == "Regular") %>% select(p.value) %>% my_round("p")`).**
 
-Regarding between-person relations, about which no hypotheses were formulated, only three effects did not include zero.
-Respondents who across all waves used social media more frequently than others to read about COVID-19 related posts reported slightly lower levels of positive affect than others (_b_ `r dat_fig_results_activity %>% filter(dv == "Positive affect" & iv == "Reading" & type == "between" & Analysis == "1.
-regular") %>% select(estimate) %>% my_round("b_txt")` [95% CI `r dat_fig_results_activity %>% filter(dv == "Positive affect" & iv == "Reading" & type == "between" & Analysis == "1.
-regular") %>% select(conf.high) %>% my_round("b")`, `r dat_fig_results_activity %>% filter(dv == "Positive affect" & iv == "Reading" & type == "between" & Analysis == "1.
-regular") %>% select(conf.low) %>% my_round("b")`]).
-Respondents who across all waves used social media more frequently than others to write COVID-19 related posts reported higher levels of negative affect than others (_b_ `r dat_fig_results_activity %>% filter(dv == "Negative affect" & iv == "Posting" & type == "between" & Analysis == "1.
-regular") %>% select(estimate) %>% my_round("b_txt")` [95% CI `r dat_fig_results_activity %>% filter(dv == "Negative affect" & iv == "Posting" & type == "between" & Analysis == "1.
-regular") %>% select(conf.high) %>% my_round("b")`, `r dat_fig_results_activity %>% filter(dv == "Negative affect" & iv == "Posting" & type == "between" & Analysis == "1.
-regular") %>% select(conf.low) %>% my_round("b")`]).
-Interestingly, respondents who across all waves used social media more frequently than others to write COVID-19 related posts also reported slightly higher levels of positive affect than others (_b_ `r dat_fig_results_activity %>% filter(dv == "Positive affect" & iv == "Posting" & type == "between" & Analysis == "1.
-regular") %>% select(estimate) %>% my_round("b_txt")` [95% CI `r dat_fig_results_activity %>% filter(dv == "Positive affect" & iv == "Posting" & type == "between" & Analysis == "1.
-regular") %>% select(conf.high) %>% my_round("b")`, `r dat_fig_results_activity %>% filter(dv == "Positive affect" & iv == "Posting" & type == "between" & Analysis == "1.
-regular") %>% select(conf.low) %>% my_round("b")`]).
-However, note that the effect were still completely inside of the null region, hence not large enough to be considered practically relevant.
+Regarding the COVID-19 related use of social media _channels_, the results were comparable (see Figure \@ref(fig:fig-within)).
+Changes in the frequency of using different social media channels to attain information regarding COVID-19 were unrelated to meaningful changes in well-being.
+For example, respondents who used Facebook more frequently than usual to learn about COVID-19 did not show a simultaneous change in well-being (_b_ = `r dat_fig_results_within %>% filter(dv == "Life satisfaction" & iv == "Facebook" & type == "within" & Analysis == "Regular") %>% select(estimate) %>% round(2)` [95% CI `r dat_fig_results_within %>% filter(dv == "Life satisfaction" & iv == "Facebook" & type == "within" & Analysis == "Regular") %>% select(conf.low) %>% round(2)`, `r dat_fig_results_within %>% filter(dv == "Life satisfaction" & iv == "Facebook" & type == "within" & Analysis == "Regular") %>% select(conf.high) %>% round(2)`]).
+In sum, the hypothesis was supported also for the COVID-19 related use of important social media channels.
+**However, two effects differed substantially from zero.**
+**Respondents who used Instagram more frequently than usual to attain COVID-19 related news reported slightly _lower_ levels of negative affect than usual (_b_ = `r dat_fig_results_within %>% filter(dv == "Negative affect" & iv == "Instagram" & type == "within" & Analysis == "Regular") %>% select(estimate) %>% my_round("b")` [95% CI `r dat_fig_results_within %>% filter(dv == "Negative affect" & iv == "Instagram" & type == "within" & Analysis == "Regular") %>% select(conf.low) %>% my_round("b")`, `r dat_fig_results_within %>% filter(dv == "Negative affect" & iv == "Instagram" & type == "within" & Analysis == "Regular") %>% select(conf.high) %>% my_round("b")`]).**
+**Respondents who used YouTube more frequently than usual to attain COVID-19 related news reported slightly _higher_ levels of negative affect than usual (_b_ = `r dat_fig_results_within %>% filter(dv == "Negative affect" & iv == "YouTube" & type == "within" & Analysis == "Regular") %>% select(estimate) %>% round(2)` [95% CI `r dat_fig_results_within %>% filter(dv == "Negative affect" & iv == "YouTube" & type == "within" & Analysis == "Regular") %>% select(conf.low) %>% round(2)`, `r dat_fig_results_within %>% filter(dv == "Negative affect" & iv == "YouTube" & type == "within" & Analysis == "Regular") %>% select(conf.high) %>% my_round("b")`]).**
+**However, both effects were still completely inside of the null region, hence not large enough to be considered meaningful.**
+
+**For an overview of all within-person effects, see Table \@ref(tab:tab-within) and Figure \@ref(fig:fig-within).**
+
+## Exploratory Analyses
+
+In what follows, I briefly report some exploratory analyses that weren't preregistered.
+
+### Between-person relations
+
+For between-person relations, no a-priori hypotheses were formulated.
+Results showed that no relation crossed or was completely outside of the SESOI.
+Four relations were statistically significant.
+**Respondents who across all waves used social media more frequently than others to read about COVID-19 related posts reported slightly lower levels of positive affect than others (_b_ `r dat_fig_results_between %>% filter(dv == "Positive affect" & iv == "Reading" & type == "between" & Analysis == "Regular") %>% select(estimate) %>% my_round("b_txt")` [95% CI `r dat_fig_results_between %>% filter(dv == "Positive affect" & iv == "Reading" & type == "between" & Analysis == "Regular") %>% select(conf.high) %>% my_round("b")`, `r dat_fig_results_between %>% filter(dv == "Positive affect" & iv == "Reading" & type == "between" & Analysis == "Regular") %>% select(conf.low) %>% my_round("b")`]).**
+**Respondents who across all waves used social media more frequently than others to write COVID-19 related posts reported sligthly higher levels of negative affect than others (_b_ `r dat_fig_results_between %>% filter(dv == "Negative affect" & iv == "Posting" & type == "between" & Analysis == "Regular") %>% select(estimate) %>% my_round("b_txt")` [95% CI `r dat_fig_results_between %>% filter(dv == "Negative affect" & iv == "Posting" & type == "between" & Analysis == "Regular") %>% select(conf.high) %>% my_round("b")`, `r dat_fig_results_between %>% filter(dv == "Negative affect" & iv == "Posting" & type == "between" & Analysis == "Regular") %>% select(conf.low) %>% my_round("b")`]).**
+**At the same time, respondents who across all waves used social media more frequently than others to write COVID-19 related posts also reported slightly higher levels of positive affect (_b_ `r dat_fig_results_between %>% filter(dv == "Positive affect" & iv == "Posting" & type == "between" & Analysis == "Regular") %>% select(estimate) %>% my_round("b_txt")` [95% CI `r dat_fig_results_between %>% filter(dv == "Positive affect" & iv == "Posting" & type == "between" & Analysis == "Regular") %>% select(conf.high) %>% my_round("b")`, `r dat_fig_results_between %>% filter(dv == "Positive affect" & iv == "Posting" & type == "between" & Analysis == "Regular") %>% select(conf.low) %>% my_round("b")`]).**
+**Finally, respondents who across all waves used YouTube more frequently than others also reported slightly higher levels of life satisfaction than others (_b_ `r dat_fig_results_between %>% filter(dv == "Life satisfaction" & iv == "YouTube" & type == "between" & Analysis == "Regular") %>% select(estimate) %>% my_round("b_txt")` [95% CI `r dat_fig_results_between %>% filter(dv == "Life satisfaction" & iv == "YouTube" & type == "between" & Analysis == "Regular") %>% select(conf.high) %>% my_round("b")`, `r dat_fig_results_between %>% filter(dv == "Life satisfaction" & iv == "YouTube" & type == "between" & Analysis == "Regular") %>% select(conf.low) %>% my_round("b")`]).**
+<!-- However, note that the effect were still completely inside of the null region, hence not large enough to be considered practically relevant. -->
 
@@
 Actively posting on social media was significantly (though not meaningfully) related to decreased life satisfaction.
-However, when controlling for potential confounders, the effect became virtually zero (see Figure \@ref(fig:res-channels)).
+However, when controlling for potential confounders, the effect became virtually zero.
 
-Regarding the COVID-19 related use of social media _channels_, the results were comparable (see Figure \@ref(fig:res-channels)).
-Changes in the frequency of using different social media channels to attain information regarding COVID-19 were unrelated to meaningful changes in well-being.
-For example, respondents who used Facebook more frequently than usual to learn about COVID-19 did not show a simultaneous change in well-being (_b_ = `r dat_fig_results_channels %>% filter(dv == "Life satisfaction" & iv == "Facebook" & type == "within" & Analysis == "1.
-regular") %>% select(estimate) %>% round(2)` [95% CI `r dat_fig_results_channels %>% filter(dv == "Life satisfaction" & iv == "Facebook" & type == "within" & Analysis == "1.
-regular") %>% select(conf.low) %>% round(2)`, `r dat_fig_results_channels %>% filter(dv == "Life satisfaction" & iv == "Facebook" & type == "within" & Analysis == "1.
-regular") %>% select(conf.high) %>% round(2)`]).
-Only two effects differed substantially from zero.
-Respondents who used Instagram more frequently than usual to attain COVID-19 related news reported slightly _higher_ levels of life satisfaction than usual (_b_ = `r dat_fig_results_channels %>% filter(dv == "Life satisfaction" & iv == "Instagram" & type == "within" & Analysis == "1.
-regular") %>% select(estimate) %>% my_round("b")` [95% CI `r dat_fig_results_channels %>% filter(dv == "Life satisfaction" & iv == "Instagram" & type == "within" & Analysis == "1.
-regular") %>% select(conf.low) %>% my_round("b")`, `r dat_fig_results_channels %>% filter(dv == "Life satisfaction" & iv == "Instagram" & type == "within" & Analysis == "1.
-regular") %>% select(conf.high) %>% my_round("b")`]).
-Respondents who used Twitter more frequently than usual to attain COVID-19 related news reported slightly _lower_ levels of life satisfaction than usual (_b_ = `r dat_fig_results_channels %>% filter(dv == "Life satisfaction" & iv == "Twitter" & type == "within" & Analysis == "1.
-regular") %>% select(estimate) %>% round(2)` [95% CI `r dat_fig_results_channels %>% filter(dv == "Life satisfaction" & iv == "Twitter" & type == "within" & Analysis == "1.
-regular") %>% select(conf.low) %>% round(2)`, `r dat_fig_results_channels %>% filter(dv == "Life satisfaction" & iv == "Twitter" & type == "within" & Analysis == "1.
-regular") %>% select(conf.high) %>% my_round("b")`]).
-However, both effects were still completely inside of the null region, hence not large enough to be considered meaningful.
-In sum, the hypothesis was supported also for the COVID-19 related use of important social media channels.
+For an overview of all between-person relations, see Figure \@ref(fig:fig-between).
 
-In terms of between-person relations---which, again, weren't included in the hypotheses---no relations crossed the null region or fell outside of it.
-Only one relation did not include zero, was hence statistically significant. 
-Respondents who across all waves used YouTube more frequently than others for COVID-19 related reasons reported marginally _higher_ levels of life satisfaction (_b_ = `r dat_fig_results_channels %>% filter(dv == "Life satisfaction" & iv == "YouTube" & type == "between" & Analysis == "1.
-regular") %>% select(estimate) %>% round(2)` [95% CI `r dat_fig_results_channels %>% filter(dv == "Life satisfaction" & iv == "YouTube" & type == "between" & Analysis == "1.
-regular") %>% select(conf.low) %>% my_round("b_txt")`, `r dat_fig_results_channels %>% filter(dv == "Life satisfaction" & iv == "YouTube" & type == "between" & Analysis == "1.
-regular") %>% select(conf.high) %>% my_round("b")`]).
-However, please note that this effect again was not large enough to be considered practically relevant.
+### Covariates
 
-Again, note that when comparing the results with and without control variables, the results differed.
-Especially on the between-person level, altogether five effects stopped being significant if they were controlled for additional variables.
-For example, using Instagram was significantly (though not meaningfully) related to increased life satisfaction.
-However, when controlling for additional covariates, the effect became virtually zero (see Figure \@ref(fig:res-channels)).
-
-## Exploratory Analyses
-
-In what follows, I briefly report some exploratory analyses that weren't preregistered.
-First, to contextualize the results reported above and to see if the results included any meaningful effects at all, I also looked at the effect sizes of selected (cherry-picked) covariates.
+To contextualize the results reported above and to see if the results included any meaningful effects at all, I also looked at the effect sizes of the covariates.
 Because each variable had different response options, we would actually need to define a SESOI for each variable, which for reasons of scope I cannot implement here. 
@@
 which allows for a better comparison across the differently scaled variables.
-For what it's worth, as a rough estimate for the SESOI we can build on the typical convention that small effects start at _r_ = |.10|.
-The results showed that several effects fell outside of the SESOI, were hence considered meaningful.
+As a rough estimate for the SESOI we can build on the typical convention that small effects start at _r_ = |.10|.
+The results showed that several effects crossed or fell outside of the SESOI, were hence considered meaningful.
 This includes for example internal locus of control, health, satisfaction with democracy, or exercising.
-For an overview, see Figure \@ref(fig:res-control).
+For an overview, see Figure \@ref(fig:fig-control).
+
+### Robustness-check
 
@@
 The results were virtually the same.
-For example, all within-person standardized COVID-19 related types of social media use or channels were significantly smaller than $\beta$ = |.05|, again supporting that effects were tiny.
-The additional analyses are reported in Figure 3 and Figure 4, and in the [companion website](https://XMtRA.github.io/Austrian_Corona_Panel_Project/analyses_additional.html).
+For example, all within-person standardized COVID-19 related types of social media use or channels were significantly smaller than $\beta$ = |.05|, again supporting that effects were negligible.
+The results of the standardized analyses are reported in Table \@ref(tab:tab-within).
+The additional analyses are reported on the [companion website](https://XMtRA.github.io/Austrian_Corona_Panel_Project/analyses_additional.html).
 
@@
 In this study I analyzed the effects of COVID-19 related social media use on well-being.
-The data come from a panel study with 24 waves and are largely representative of the Austrian population.
+The data come from a panel study with 32 waves and are largely representative of the Austrian population.
 In a random effects model I separated between person relations from within-person effects and controlled for a large number of both stable and varying covariates, thereby aiming to assess causality.
-The results showed that within-person effects were trivial.
+The results showed that some statistically significant negative within-person effects existed, but that they were **very small and likely** trivial.
 People who used social media more than usual to learn about COVID-19 didn't show meaningful changes in their well-being.
@@
 The results imply that COVID-19 related social media use doesn't seem to be particularly relevant for well-being.
-Other factors among the third variables that were measured revealed larger effects or relations, suggesting that well-being is determined by alternative aspects such as health, satisfaction with democracy, locus of control, or exercising.
-According to this study, popular fears that "doomscrolling" or overusing social media during times of crises are not supported.
+Other factors among the third variables that were measured revealed larger effects or relations, suggesting that well-being is rather determined by alternative aspects such as health, satisfaction with democracy, locus of control, or exercising.
+According to this study, popular fears that "doomscrolling" or overusing social media during times of crises is detrimental are not supported.
+
+**That said, several preliminary and subtle trends can be observed.**
+**First, overall the results do suggest that effects of COVID-19 related social media use on well-being rather tend to take place in the negative as opposed to the positive spectrum.**
+**For example, people who wrote more COVID-19 related posts than usual reported slightly lower levels of life satisfaction than usual.**
+**Similarly, people who wrote more COVID-19 related posts than usual also reported slightly more negative affect.**
+**When writing posts and comments on social media, people explicitly and more deeply engage with COVID-19 related content.**
+**Unfortunately, in general online tonality is often more extreme, negative, or aggressive, which potentially affects their authors.**
+**Note that I controlled for whether or not participants had an corona infection during a specific wave, which rules out the potential explanation that having an infection was the root cause of increased communication and reduced well-being.**
+
+**The potential explanation that tonality might be a relevant factor at play here is also supported by the second trend.**
+**People, who during the pandemic spend more time on Instagram than usual, also experienced less negative affect than usual.**
+**Instagram is well-known for its positivity bias [@waterlooNormsOnlineExpressions2018].**
+**Content is generally post positive, uplifting, and (self-)flattering.**
+**It seems the often-criticized positivity bias on Instagram might have been somewhat beneficial in times of the pandemic.**
+**The critique that the positivity bias necessarily leads to envy and negative feelings is one-sided, because positive content can also inspire and motivate users [@meierInstagramInspirationHow2020], which could be especially helpful in times of lockdown and home-office.**
+**To provide a concrete example, Instagram was successfully used as an interactive communication channel for first year students to have a better start into their new degree, effectively complementing alternative learning platform tools [@yeTurningInformationDissipation2020].**
+
+**Similarly, people who spent more time on YouTube than usual also reported slightly more negative affect than usual.** 
+**Communication on YouTube is often found to be more negative and less polite compared to other SNSs [@halpernSocialMediaCatalyst2013].**
+**YouTube is also routinely linked to mis- and disinformation.**
+**Of the 69 most viewed videos on YouTube on COVID-19, 19 (27.5%) contained nonfactual information [@liYouTubeSourceInformation2020].**
+**Consuming more negative and misleading information might hence be a potential explanation for the slightly increased levels of negative affect.**
 
@@
 However, note that @bendauAssociationsCOVID19Related2021 analyzed cross-sectional data on a between-person level while not controlling for third variables, which is not optimal for investigating causal effects.
-On the other hand, the results are well-aligned with many seminal Communication theories such as the uses and gratifications model [@katzUsesGratificationsResearch1973].
+On the other hand, the results are well-aligned with **mood management theory [@zillmannMoodManagementCommunication1988]** or the uses and gratifications approach [@katzUsesGratificationsResearch1973].
 If effects were indeed profoundly negative on average, then people likely wouldn't spend so much time on social media.
@@
 
-If anything, two preliminary and subtle trends can be observed.
-First, in terms of media channels, using Twitter more than usual was related to slightly decreased levels of life satisfaction.
-Twitter is considered to have more negative affordances and tonality as compared to other networks such as Instagram [@waterlooNormsOnlineExpressions2018], which might help explain the results.
-Instagram, on the other hand, was related to slightly increased levels of life satisfaction.
-To speculate, the often-criticized positivity bias on Instagram might have been somewhat beneficial in times of the pandemic.
-The critique that the positivity bias necessarily leads to envy and negative feelings is unrealistic and myopic, because positive content can also inspire and motivate users [@meierInstagramInspirationHow2020], which could be especially helpful in times of lockdown and home-office.
-Second, of all the three COVID-19 related social media activities, people who read about the pandemic more than others showed slightly decreased levels of positive affect, and people who actively posted about the pandemic more than others showed slightly increased levels of negative affect.
-On the other hand, however, people who posted more about COVID-19 also showed slightly higher levels of positive affects, so taken together the results are ambivalent.
-But future research might elaborate on these specific relations to probe their stability and relevance.
+**The results showed that it makes sense to analyze different communication types and communication channels, and that active and passive communication showed different results.**
+**Liking and sharing content did not show any within-person effects. **
+**Such rather low-key active behaviors do not seem to affect well-being at all.**
+**Regarding passive use, reading COVID-19 related posts is more ambivalent; results showed some weak trends towards a positive effect on life satisfaction, but a negative affect on mood.**
+**It might be that reading and informing oneself about COVID-19 on social media might be helpful in the long term, but more negative for short-term affect.**
+**Finally, proactively engaging via writing posts, the most active form of communication analyzed here, showed only negative effects on well-being.**
+**Together, the results support the findings from @valkenburgAssociationsActivePassive2022, who could not confirm the claim that active use is good, versus passive is bad.**
+**Focusing on communication channels, YouTube seems to be more negative, whereas Instagram is likely more positive.**
+**Again, these are only very small effects.**
+**Future research might elaborate on these specific relations to probe their stability and relevance.**
+
+**From a more political and societal perspective, I believe the results imply that it does makes sense critically to reflect upon COVID-19 related social media use.** 
+**In terms of concrete recommendations, on average it might be slightly beneficial to post less actively about COVID-19 on social media and to spent less time on YouTube.**
+**Potential positive effects, however, will for many users likely not be noticeable.**
+**Results allow to reject a positive effect: Writing more posts on social media will likely not increase well-being.**
+**At all events, engaging in COVID 19-related social media use should, on average, not be a cause for concern.**
+
+<!-- First, in terms of media channels, using Twitter more than usual was related to slightly decreased levels of life satisfaction. -->
+<!-- Twitter is considered to have more negative affordances and tonality as compared to other networks such as Instagram [@waterlooNormsOnlineExpressions2018], which might help explain the results.
+-->
+<!-- Second, of all the three COVID-19 related social media activities, people who read about the pandemic more than others showed slightly decreased levels of positive affect, and people who actively posted about the pandemic more than others showed slightly increased levels of negative affect.
+-->
+<!-- On the other hand, however, people who posted more about COVID-19 also showed slightly higher levels of positive affects, so taken together the results are ambivalent. -->
 
@@
 Together, this allowed for an improved perspective on assessing causality.
-That said, causality necessitates temporal order, and the cause needs to precede the effect.
-Regarding media use, such effects often happen immediately or shortly after use, necessitating intervals in the hours, minutes, or even seconds.
+**However, the opposite effect is also plausible, namely that well-being affects media use [@zillmannMoodManagementCommunication1988].**
+**While controlling for potential confounders can support claims of causality, the procedure implemented here cannot prove causality.**
+Causality necessitates temporal order, and the cause needs to precede the effect.
+**The challenge is that** regarding media use, such effects often happen immediately or shortly after use, necessitating intervals in the hours, minutes, or even seconds.
 In many cases only experience sampling studies asking users in the very moment can produce such knowledge.
 However, even then we don't know for certain if we actually measured the right interval.
-Effects depend on the intensity of use or the length of the interval, and to borrow the words from @rohrerTheseAreNot2021, there is no such thing as "the" effect of social media use on well-being.
+Effects depend on the intensity of use or the length of the interval.
+To borrow the words from @rohrerTheseAreNot2021, there is no such thing as "the" effect of social media use on well-being.
 Hence, to document how effects unfold, future research needs to employ different study designs probing different time lags.
@@
 In this study, COVID-19 related social media use didn't meaningfully affect several indicators of well-being, including life satisfaction, positive affect, and negative affect.
-However, factors other than social media use were meaningfully related to well-being, such as physical health, exercise, satisfaction with democracy, or believing that one is in control of one's life.
-If it's our aim to improve well-being, it might hence be more fruitful not to focus so much on social media but to address other, more pertinent societal problems related to health care, regular exercise, or a functioning democratic system.
+**If people wrote more COVID-19 related posts than usual, or if they spent less time on Instagram and more time on YouTube, very small but statistically significant effects were found.**
+Notably, however, factors other than social media use were more meaningfully related to well-being, such as physical health, exercise, satisfaction with democracy, or believing that one is in control of one's life.
+If it's our aim to improve well-being **during a pandemic**, it might hence be more fruitful not to focus so much on social media but to address other, more pertinent societal problems related to health care, regular exercise, or a functioning democratic system.
 
@@
 
-```{r tab-descriptives, cache=F}
+```{r tab-descriptives, cache=F, warning=F, message=F, prompt=F}
 apa_table(tab_desc_dat,
@@
 
-```{r sesoi, warning=F, message=F, fig.cap="Illustration of how confidence intervals are used to test a null region.
-Here, a trivial effect of social media use on life satisfaction is defined as ranging from b = -.25 to b = .25", fig.width=6, fig.height=2}
+```{r tab-within, cache=F, warning=F, message=F, prompt=F}
+tab_within %<>%
+  select(Predictor = iv,
+         b = estimate,
+         `Lower level` = conf.low,
+         `Higher level` = conf.high,
+         beta = "std",
+         p = "p.value")
+apa_table(tab_within,
+          align = c("l", "r", "r", "r", "r", "r"),
+          stub_indents = list(
+            "Life satisfaction" = c(1:8),
+            "Positive affect" = c(9:16),
+            "Negative affect" = c(17:24)),
+          caption = "Overview of all within-person effects.",
+          col_spanners = list(`Confidence interval` = c(3, 4)),
+          font_size = "footnotesize"
+          )
+```
+
+\newpage
+
+```{r sesoi, warning=F, message=F, fig.cap="Using confidence intervals to test a null region.
+Note.
+Here, a trivial effect of social media use on life satisfaction is defined as ranging from b = -.25 to b = .25", fig.width=6, fig.height=2, warning=F, message=F, prompt=F}
 # make table with data
@@
 
-```{r descriptives, fig.cap="Development of well-being and media use measures across the pandemic.
-Values obtained from mixed effect models, with participants and waves as grouping factors and without additional predictors.", out.width = "\\textwidth", cache=F, warning=F, message=F, prompt=FALSE}
-# print(fig_desc)
+```{r fig-descriptives, fig.cap="Well-being and media use across the 32 waves.
+Note.
+Values obtained from mixed effect models, with participants and waves as grouping factors and without additional predictors.", out.width = "\\textwidth", cache=F, warning=F, message=F, prompt=F}
+# fig_desc
 knitr::include_graphics("figures/fig_descriptives.png")
@@
 
-```{r res-activity, fig.cap="The effects of various types of social media use on three indicators of well-being.
+```{r fig-within, fig.cap="Within-person effects of COVID-19 related social media use on well-being.
+Note.
 The black estimates show the effects controlled for a large number of covariates (see text; preregistered); the grey estimates are without control variables (exploratory).
 The SESOI was b = |0.30| for life satisfaction and b = |0.15| for affect.
-Hence, all of the reported effects are not considered meaningful.", out.width = "\\textwidth", cache=F, warning=F, message=F, fig.width=6, fig.height=3}
-fig_results_activity
-# knitr::include_graphics("figures/fig_results_activity.png")
+Hence, all of the reported effects are not considered meaningful.", out.width = "\\textwidth", cache=F, warning=F, message=F, fig.width=6, fig.height=4}
+fig_results_within
+# knitr::include_graphics("figures/fig_results_within.png")
 ```
@@
 
-```{r res-channels, fig.cap="The effects of using various social media applications on three indicators of well-being.
+```{r fig-between, fig.cap="Between-person relations between COVID-19 related social media use and well-being.
+Note.
 The black estimates show the effects controlled for a large number of covariates (see text; preregistered); the grey estimates are without control variables (exploratory).
@@
 Hence, all of the reported effects are not considered meaningful.", out.width = "\\textwidth", cache=F, warning=F, message=F, fig.width=6, fig.height=4}
-# knitr::include_graphics("figures/fig_results_channels.png")
-fig_results_channels
+fig_results_between
+# knitr::include_graphics("figures/fig_results_between.png")
 ```
@@
 
-```{r res-control, fig.cap="Results of selected covariates.
-All variables were standardize except 'Male'  and ‘Employed in public service', because there were measured on a binary scale.", out.width = "\\textwidth", cache=F, warning=F, message=F, fig.width=6, fig.height=3}
+```{r fig-control, fig.cap = "Results of selected covariates. Note. All variables standardized except 'Male'.", out.width = "\\textwidth", cache=F, warning=F, message=F, fig.width=6, fig.height=4}
 fig_results_control_std

### 2022-07-25 manuscript.tex
- major updates for revision - included multiple imputation - changed theory - include new data-set - extended discussion

@@
 \documentclass[
-  man,mask]{apa6}
+  man,mask]{apa7}
 \usepackage{amsmath,amssymb}
@@
 \usepackage{xcolor}
-\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
-\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
-\hypersetup{
-  pdftitle={Analyzing the effects of COVID-19 related social media use on well-being},
-  pdflang={en-EN},
-  pdfkeywords={COVID-19, well-being, social media, news use, panel study.},
-  hidelinks,
-  pdfcreator={LaTeX via pandoc}}
-\urlstyle{same} % disable monospaced font for URLs
 \usepackage{graphicx}
@@
 \usepackage{csquotes}
+\makeatletter
+\renewcommand{\paragraph}{\@startsection{paragraph}{4}{\parindent}%
+  {0\baselineskip \@plus 0.2ex \@minus 0.2ex}%
+  {-1em}%
+  {\normalfont\normalsize\bfseries\typesectitle}}
+
+\renewcommand{\subparagraph}[1]{\@startsection{subparagraph}{5}{1em}%
+  {0\baselineskip \@plus 0.2ex \@minus 0.2ex}%
+  {-\z@\relax}%
+  {\normalfont\normalsize\bfseries\itshape\hspace{\parindent}{#1}\textit{\addperi}}{\relax}}
+\makeatother
 \usepackage{endnotes}
@@
 \note{\clearpage}
+
 \ifLuaTeX
@@
 \fi
-
-\title{Analyzing the effects of COVID-19 related social media use on well-being}
+\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
+\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
+\urlstyle{same} % disable monospaced font for URLs
+\hypersetup{
+  pdftitle={The effects of COVID-19 related social media use on well-being},
+  pdflang={en-EN},
+  pdfkeywords={COVID-19, well-being, social media, news use, panel study.},
+  hidelinks,
+  pdfcreator={LaTeX via pandoc}}
+
+\title{The effects of COVID-19 related social media use on well-being}
 \author{Name blinded for review\textsuperscript{1}}
@@
 \abstract{%
-In times of crisis such as the Corona pandemic citizens need to stay informed about recent events, political decisions, or mandatory protection measures.
+In times of crisis such as the COVID-19 pandemic citizens need to stay informed about recent events, political decisions, or mandatory protection measures.
 To this end, many people use various types of media, and increasingly social media.
@@
 Users were glued to their screens and found it hard to pursue other relevant activities such as working, taking a break, or looking after their children (Klein, 2021).
-It was asked whether using social media for COVID-19 related reasons is helpful, or whether it creates an additional burden on mental health (Sandstrom, Buchanan, Aknin, \& Lotun, 2021).
+It was asked whether using social media for COVID-19 related reasons is helpful, or whether it creates an additional burden on mental health (Sandstrom et al., 2021).
 These concerns seem justified:
@@
 To this end, I analyzed a large-scale panel study from the Austrian Corona Panel Project (Kittel et al., 2020).
-The panel consists of 24 waves and has an overall sample size of 3,018 participants.
+The panel consists of \textbf{32} waves and has an overall sample size of \textbf{3,485 participants}.
 The panel study collected a large number of psychological and demographic variables.
@@
 According to the two-continua model, mental health consists of (a) psychopathology and (b) well-being.
-Well-being can be differentiated into subjective and psychological well-being (Diener, Lucas, \& Oishi, 2018).
+Well-being can be differentiated into subjective and psychological well-being (Diener et al., 2018).
 Whereas subjective well-being emphasizes hedonic aspects such as happiness and joy, psychological well-being addresses eudaimonic aspects such as fulfillment and meaning.
@@
 Whereas the first four levels focus on the communication \emph{channel}, the last two address the communication \emph{type}.
-To measure social media use for the consumption of COVID-19 related news and topics, I here employ both the channel and the type of communication perspective, which together should offer a nuanced understanding of communication.
-
+\textbf{Differentiating different communication channels and communication types is important, because the effects on well-being likely differ across communication channels and communication types.}
+\textbf{Whereas active social media use such as chatting is routinely linked to improved well-being, passive use such as reading is considered more negative (Dienlin \& Johannes, 2020).}
+Similarly, branded apps are separate entities with potentially divergent effects.
+For example, Waterloo et al. (2018) found that it's more adequate to express negative emotions on WhatsApp than on Twitter or on Instagram.
+\textbf{Especially during a pandemic, it makes sense to analyze if users engage with COVID-19 related content on Instagram, where communication is more positive, or on Facebook, where communication is more critical.}
+
+\textbf{In this study, to measure the effects of social media use focused on COVID-19 related news and topics, I adopt both the channel and the type of communication perspective.}
+\textbf{Together, this should offer a nuanced and comprehensive understanding of communication.}
 First, I investigate how well-being is affected by different types of communication affect, namely active and passive use.
-Defining what constitutes active and what passive use is not always clear, and different understandings are currently discussed (Ellison, Triẹû, Schoenebeck, Brewer, \& Israni, 2020; Meier \& Reinecke, 2020).
+Defining what constitutes active and what passive use is not always clear, and different understandings are currently discussed (Ellison et al., 2020; Meier \& Reinecke, 2020).
 Reading is generally considered as passive and writing as active, while there are also specific behaviors such as such liking or sharing content that fall somewhere in-between (Meier \& Krause, 2022).
@@
 Second, I analyze how using the most prominent branded applications affects well-being, and whether this effect changes across applications.
-Branded apps are separate entities with potentially divergent effects.
-Twitter might have a different effect as compared to WhatsApp because of their respective affordances.
-For example, Waterloo, Baumgartner, Peter, and Valkenburg (2018) found that it's more adequate to express negative emotions on WhatsApp than on Twitter or on Instagram.
-The branded applications investigated here are Facebook, Twitter, Instagram, WhatsApp, and YouTube.
-Worth noting, this study is not about \emph{general} social media use during times of COVID, but on social media use \emph{focused} on COVID-19 related content.
-Examples of such media use include posting thoughts about the pandemic or retweeting COVID-19 related news.
+The branded applications investigated here are Facebook, Twitter, Instagram, WhatsApp, and YouTube---which were, at the time of writing, the most relevant social media apps.
+
+\textbf{Worth noting, this study is not about \emph{general} social media use during times of COVID, but on social media use \emph{focused} on COVID-19 related content.}
+\textbf{For example, posting thoughts about the pandemic, reading posts and comments, or retweeting COVID-19 related news.}
 
@@
 In advance, there doesn't seem to be a clear winner, and it's likely that both positive and negative effects cancel each other out.
-Empirically, social media use on average oes not have a strong effect on well-being (Meier \& Reinecke, 2020).
-According to the Differential Susceptibility to Media Effects Model (Valkenburg \& Peter, 2013), the effects of media use differ across individuals.
+Empirically, social media use on average does not have a strong effect on well-being (Meier \& Reinecke, 2020).
+According to the Differential Susceptibility to Media Effects Model (Valkenburg \& Peter, 2013), the effects of media use differ across individuals \textbf{and content}.
 Whereas for some media are beneficial, for others they are harmful.
+\textbf{Whereas some content provides opportunities (education, advice), other content creates risks (misinformation, hate). (Livingstone et al., 2018).}
+Social media can impair well-being when causing embarrassment, stress, or disinformation, and they can improve well-being when providing connectedness, information, or entertainment (Büchi, 2021).
 On average, however, effects are often small or negligible.
-But also when focusing on individual users, social media can have both positive and negative effects on well-being (Büchi, 2021).
-They can impair well-being when causing embarrassment, stress, or disinformation, and they can improve well-being when providing connectedness, information, or entertainment (Büchi, 2021).
-
-Two of the most prominent media effect theories argue against strong negative impacts.
-First, uses and gratifications theory states that people explicitly and rationally chose specific media because of their respective benefits (Katz, Blumler, \& Gurevitch, 1973).
-If those benefits don't exist, they will spend their time elsewhere.
-And social media offer copiuos benefits.
-Most prominently, they help find relevant information, maintain and foster relationships, express one's personality, and entertain oneself (Pelletier, Krallman, Adams, \& Hancock, 2020).
-The second major theory that needs to be listed here is mood management theory (Zillmann, 1988).
-Users implicitly learn what media help them balance their mood and affect.
-For example, when bored people use social media to entertain themselves.
-Using experience sampling of well-being and logs of social media use, a study with 82 participants from Italy found that after episodes of social media use, levels of positive affect increased significantly (Marciano, Driver, Schulz, \& Camerini, 2022).
+
+\textbf{Two of the most prominent media effect theories argue (indirectly) against strong average negative impacts.}
+\textbf{According to mood management theory (Zillmann, 1988), using media substantially affects people's moods.}
+\textbf{Effects can be stimulating or overwhelming, relaxing or boring.}
+\textbf{After some time, users implicitly learn what media help them balance their mood and affect according to their own situational needs (Zillmann, 1988).}
+\textbf{Those media that eventually become part of one's media repertoire are hence, on average, beneficial for users and their moods.}
+\textbf{Using experience sampling of well-being and logs of social media use, a study with 82 participants from Italy found that after episodes of social media use, levels of positive affect increased significantly (Marciano et al., 2022).}
+
+\textbf{While mood management theory considers media use mainly driven by implicit learning experiences, uses and gratifications theory upholds that the process is more explicit and rational (Katz et al., 1973).}
+\textbf{Users select those media that they expect to have a desired effect, for example on mood, knowledge, or entertainment.}
+\textbf{If those beneficial media effects do not exist or if they are not expected, people will spend their time elsewhere.}
+\textbf{And social media offer several beneficial effects, explaining why they are used that much.}
+\textbf{They help find relevant information, maintain and foster relationships, express one's personality, and entertain oneself (Pelletier et al., 2020).}
+\textbf{In conclusion, because people spend so much time on social media consuming COVID-19 related content, according to both mood management theory and uses and gratifications theory we wouldn't expect to find strong average negative effects.}
 
@@
 And as with many other things, there can be too much of a good thing.
-It is therefore often asked whether social media are addictive, and users sometimes express this fear themselves (Yang, Griffiths, Yan, \& Xu, 2021).
-However, a recently published meta-analysis found that the two most prominent measures of addiction, the Bergen Facebook Addiction Scale and the Bergen Social Media Addiction Scale, have only small relations to well-being (Duradoni, Innocenti, \& Guazzini, 2020).
+It is therefore often asked whether social media are addictive, and users sometimes express this fear themselves (Yang et al., 2021).
+However, a recently published meta-analysis found that the two most prominent measures of addiction, the Bergen Facebook Addiction Scale and the Bergen Social Media Addiction Scale, have only small relations to well-being (Duradoni et al., 2020).
 In addition, the general idea of labeling excessive social and new media use as addiction was criticized, arguing that social media represent new regular behaviors that should not be pathologized (Galer, 2018; van Rooij et al., 2018).
 
-Because media effects can differ across situations (Valkenburg \& Peter, 2013), I now briefly focus on the effects of COVID-19 related social media use specifically.
+\textbf{Because media effects can differ across users, situations, and content (Livingstone et al., 2018; Valkenburg \& Peter, 2013), I now briefly focus on the effects of COVID-19 related social media use specifically.}
 First, one could assume a \emph{direct} negative effect on well-being, and especially on positive or negative affect, which are more volatile and fluctuating.
@@
 If one learns about such events, the initial reaction might be shock, fear, or dismay.
-Consuming such news can be depressing (Dörnemann, Boenisch, Schommer, Winkelhorst, \& Wingen, 2021), perhaps even changing some general perspectives on life.
+Consuming such news can be depressing (Dörnemann et al., 2021), perhaps even changing some general perspectives on life.
 That said, because not all news was negative, and because many people showed solidarity and compassion, there was also positive and uplifting content, potentially compensating for the negative effects (Dörnemann et al., 2021).
-A study with 2.057 respondents from Italy reported that during the pandemic virtual community and social connectedness even increased during the pandemic (Guazzini, Pesce, Marotta, \& Duradoni, 2022).
+A study with 2.057 respondents from Italy reported that during the pandemic virtual community and social connectedness even increased during the pandemic (Guazzini et al., 2022).
+\textbf{In Finland, in a sample of 735 participants, levels of loneliness did not decrease during the pandemic, and people who engaged more on social media showed less loneliness (Latikka et al., 2022).}
 
@@
 To explain, overuse could impair well-being if it replaces meaningful or functional activities such as meeting others, working, actively relaxing, or exercising.
-So if a society collectively overuses social media during a pandemic, there might be potential for negative effects.
-On the other hand, one can make the case that overuse might even be beneficial, especially in times of a pandemic.
+\textbf{Another potentially negative mechanism at play are problematic social comparison processes.
+During the pandemic, several users shared how they successfully dealt with challenges such as physical distancing.
+In a study with 1131 residents from Wuhan in China (Yue et al., 2022), people who spent more time in quarantine also spent more time on social media.
+Those, who spent more time on social media also engaged in more upward social comparison, which was related to increased levels of stress.}
+
+On the other hand, one can make the case that using social media for COVID-19 related reasons might even be beneficial, especially in times of a pandemic.
 Exchanging COVID-19 related messages with friends via WhatsApp might replace the in-person contact one would have otherwise, but which is literally impossible at the time.
@@
 (2021) found that people who used social media as a primary source of information reported on average ``significantly more unspecific anxiety and depression {[}{]} and significantly more specific COVID-19 related anxiety symptoms'' (p.~288).
-Eden, Johnson, Reinecke, and Grady (2020) analyzed the media use of 425 US college students during the first wave of the pandemic, finding both positive and negative relations with well-being.
+Eden et al. (2020) analyzed the media use of 425 US college students during the first wave of the pandemic, finding both positive and negative relations with well-being.
 In a sample of 312 respondents collected via Amazon Mechanical Turk, Choi and Choung (2021) reported that people who used media to attain information were more lonely and less satisfied with their lives.
-Stainback, Hearne, and Trieu (2020) analyzed a large-scale study with 11,537 respondents from the US and found that increased COVID-19 media consumption was related to more psychological distress.
+Stainback et al. (2020) analyzed a large-scale study with 11,537 respondents from the US and found that increased COVID-19 media consumption was related to more psychological distress.
 A four-wave panel study with 384 young adults from the U.S.
-analyzed the effects of general digital technology use---objectively measured via screenshots of screen-time applications---on mental health, separating within- and between-person relations (Sewall, Goldstein, \& Rosen, 2021).
+analyzed the effects of general digital technology use---objectively measured via screenshots of screen-time applications---on mental health, separating within- and between-person relations (Sewall et al., 2021).
 The results showed that digital technology did not have any significant effects on mental health (for a similar study with comparable results, see Bradley \& Howard, 2021).
@@
 
-Finally, also several individual studies employing advanced methods found smalls relations between social media use and well-being (Keresteš \& Štulhofer, 2020; Orben, Dienlin, \& Przybylski, 2019; Przybylski, Nguyen, Law, \& Weinstein, 2021; Schemer, Masur, Geiß, Müller, \& Schäfer, 2021).
-For example, Beyens, Pouwels, van Driel, Keijsers, and Valkenburg (2021) reported that although for some users (roughly one quarter) the effects of social media use on well-being were negative, for almost the same number of users they were positive, while for the rest the effects were neutral.
+Finally, also several individual studies employing advanced methods found smalls relations between social media use and well-being (Keresteš \& Štulhofer, 2020; Orben et al., 2019; Przybylski et al., 2021; Schemer et al., 2021).
+For example, Beyens et al.
+(2021) reported that although for some users (roughly one quarter) the effects of social media use on well-being were negative, for almost the same number of users they were positive, while for the rest the effects were neutral.
 This finding is aligned the Differential Susceptibility to Media Effects Model:
@@
 In conclusion, in light of the theoretical considerations and empirical studies presented above, I expect that COVID-19 related communication on social media doesn't affect well-being in a meaningful or relevant way.
+\textbf{This general hypothesis will be analyzed specifically for the communication types of (a) time spent reading, (b) liking and sharing, and (c) actively posting COVID-19 related content.}
+\textbf{In addition, I will analyze how well-being is influenced by spending time on five prominent social media apps, including (a) Facebook, (b) Instagram, (c) Twitter, (d) WhatsApp, and (e) YouTube.}
+\textbf{Three different well-being indicators will be differentiated: life satisfaction, positive affect, and negative affect.}
 
@@
 But what's a noticeable change in well-being?
-According to Norman, Sloan, and Wyrwich (2003), people can reliably distinguish \emph{seven} levels of satisfaction with health.
+According to Norman et al. (2003), people can reliably distinguish \emph{seven} levels of satisfaction with health.
 So if satisfaction is measured on a 7-point scale, we would state that a four unit change in social media use should result in a one unit change in life satisfaction.
@@
 Using longitudinal designs alone, however, is not sufficient for establishing correct causal statements (Rohrer \& Murayama, 2021).
-In addition, we for example also need to control for confounding third variables, and importantly also for \emph{varying} third variables.
+In addition, we for example also need to control for confounding third variables.
+\textbf{Importantly, when analyzing longitudinal (within-person) relationships and effects, it is important to control for \emph{varying} third variables.}
+\textbf{Non-varying third variables only help control non-varying (between-person) relations.}
 
@@
 In this study, I hence plan to control for the following variables, which either have already been shown to affect both social media use and well-being or which are likely to do so, and which also aren't mediators:
-gender, age, education, Austria country of birth, Austria country of birth of parents, text-based news consumption, video-based news consumption, residency Vienna, household size, health, living space, access to garden, access to balcony, employment, work hours per week, being in home-office, household income, outdoor activities, satisfaction with democracy, disposition to take risks, and locus of control.\footnote{The dataset includes many other variables that one could also potentially control for, and I invite interested readers to download the and explore potential interesting relationships.}
+gender, age, education, Austria country of birth, Austria country of birth of parents, text-based news consumption, video-based news consumption, residency Vienna, household size, health, living space, access to garden, access to balcony, employment, work hours per week, being in home-office, household income, outdoor activities, satisfaction with democracy, disposition to take risks, and locus of control.\footnote{The data-set includes many other variables that one could also potentially control for, and I invite interested readers to download the and explore potential interesting relationships.}
 
@@
 If variables are stable, longer intervals are needed; if they fluctuate, shorter intervals.
-In the case of well-being, we need shorter intervals for affect and longer ones for life satisfaction (Dienlin \& Johannes, 2020).
-Still, choosing the right interval is challenging, because especially short intervals are hard to implement in practice and often require advanced methods such as experience sampling (also known as in situ measurement or ambulant assessment) (Schnauber-Stockmann \& Karnowski, 2020).
-
-In this study, I hence analyze how using social media during the last week affected their positive and negative affect during the same week.
-More precisely, if people during the last week used COVID-19 related social media more than the usually do, did they feel better or worse during that week than they usually do?
+In the case of well-being, we need shorter intervals for \textbf{the more fluctuating} positive and negative affect and longer ones for \textbf{the more stable} life satisfaction (Dienlin \& Johannes, 2020).
+\textbf{Using social media can have instant effects on mood (Marciano et al., 2022).}
+\textbf{Effects on life satisfaction often take longer to manifest, for example because media use leads to actual changes in specific behaviors, which then affect life satisfaction (Dienlin et al., 2017).}
+Choosing the right interval is challenging, because especially short intervals are hard to implement in practice, often requiring advanced methods such as experience sampling (also known as in situ measurement or ambulant assessment) (Schnauber-Stockmann \& Karnowski, 2020).
+In this study, I hence analyze how using social media during the last week affected positive and negative affect during the same week.
+In other words, if people during the last week engaged in more COVID-19 related social media use than the usually do, did they feel better or worse during that week than they usually do?
 Regarding life satisfaction, I implemented a longer interval.
-Namely, if people during the last week used COVID-19 related social media more than the usually do, were they at the end of the week more or less satisfied with their lives than they usually are?
+If people during the last week used COVID-19 related social media more than the usually do, were they at the end of the week more or less satisfied with their lives than they usually are?
 I hence analyze if when a person changes their social media diet, are there (a) \emph{simultaneous} changes in their affect and (b) \emph{subsequent} changes in their life satisfaction?
-When additionally controlling for both stable and varying confounders, we can then be more sure that the effect is indeed causal.
-Similar approaches were implemented by other studies (Johannes, Dienlin, Bakhshi, \& Przybylski, 2022; Scharkow, Mangold, Stier, \& Breuer, 2020), and they are considered a best practice approach toward analyzing causality (Bell, Fairbrother, \& Jones, 2019).
+\textbf{These relations will be controlled for varying confounders, which fosters a causal interpretation.}
+Similar approaches were implemented by other studies (Johannes et al., 2022; Scharkow et al., 2020), and they are considered a best practice approach toward analyzing causality (Bell et al., 2019).
 
@@
 
-The data come from the Austrian Corona Panel Project (Kittel et al., 2021).
-The study was conducted between March 2020 and October 2021.
-It contains 26 waves, and at the time of writing the first 24 waves were available for download.
+\textbf{The data come from the Austrian Corona Panel Project (Kittel et al., 2021), which is a large-scale standalone panel study. }
+\textbf{The data are hosted on AUSSDA and are publicly available here: \url{https://doi.org/10.11587/28KQNS}.}
+\textbf{At the time of writing, the official website included a data-set consisting of 24 waves. }
+\textbf{For the analyses presented here, I was able to receive an advance version consisting of all 32 waves.}
+\textbf{The study was conducted between March 2020 and June 2022, and data collection is now officially finished.}
+\textbf{It contains 32 waves.}
 Between March 2020 and July 2020, the intervals were weekly, and afterward the intervals were monthly.
 Each wave consists of at least 1,500 respondents.
-The overall sample size was \emph{N} = 3,018, and 72,432 observations were collected.
+The overall sample size was \emph{N} = 3,485, and 111,520 observations were collected.
 Panel mortality was compensated through a continuous acquisition of new participants.
@@
 The participants provided their written informed consent to participate in this study.
-The average age was 42 years, 49 percent were male, 14 percent had a University degree, and 5 percent were currently unemployed.
+The average age was 41 years, 49 percent were male, 14 percent had a University degree, and 5 percent were currently unemployed.
 
@@
 Because Mardia's test showed that the assumption of multivariate normality was violated, I used the more robust Satorra-Bentler scaled and mean-adjusted test statistic (MLM) as estimator.
-To avoid overfitting, I tested the scales on more liberal fit criteria (CFI \textgreater{} .90, TLI \textgreater{} .90, RMSEA \textless. .10, SRMR \textless{} .10) (Kline, 2016).
-Finally, REWB-models cannot model latent variables.
-To increase precision, I therefore exported factor scores from the CFAs for positive and negative affect.
-Respondents who answered less than 50\% of all questions were removed.
-The remaining missing responses were imputed using predictive mean matching.
+To avoid over-fitting, I tested the scales on more liberal fit criteria (CFI \textgreater{} .90, TLI \textgreater{} .90, RMSEA \textless. .10, SRMR \textless{} .10) (Kline, 2016).
+Mean scores were used for positive and negative affect.
+\textbf{Missing responses were imputed using multiple imputation with predictive mean matching (five iterations, five data-sets).}
+\textbf{Predictive mean matching allowed to impute also categorical variables.}
+\textbf{All variables were imputed except the media use measures, as they were not collected on each wave.}
+\textbf{All variables included in the analyses presented here were used to impute missing data.}
+\textbf{For the main analyses, results were pooled across all five data-sets.}
 
@@
 The response options were 1 (\emph{never}), 2 (\emph{on some days}), 3 (\emph{several times per week}), 4 (\emph{almost every day}), and 5 (\emph{daily}).
-The scale showed good factorial fit, \(\chi^2\)(46) = 65.30, \emph{p} = .032, CFI = 1.00, RMSEA = .02, 90\% CI {[}.01, .03{]}, SRMR = .01.
+The scale showed good factorial fit, \(\chi^2\)(62) = 79.27, \emph{p} = .069, CFI = 1.00, RMSEA = .01, 90\% CI {[}\textless{} .01, .02{]}, SRMR = .01.
 Reliability was high, \(\omega\) = .85.
@@
 The response options were 1 (\emph{never}), 2 (\emph{on some days}), 3 (\emph{several times per week}), 4 (\emph{almost every day}), and 5 (\emph{daily}).
-The scale showed good factorial fit, \(\chi^2\)(331) = 3138.37, \emph{p} \textless{} .001, CFI = .97, RMSEA = .08, 90\% CI {[}.07, .08{]}, SRMR = .03.
+The scale showed good factorial fit, \(\chi^2\)(443) = 3990.32, \emph{p} \textless{} .001, CFI = .98, RMSEA = .07, 90\% CI {[}.07, .08{]}, SRMR = .03.
 Reliability was high, \(\omega\) = .89.
@@
 
-Social media use was measured for all participants on waves 1, 2, 8, 17, and 23.
+Social media use was measured for all participants on waves 1, 2, 8, 17, 23, \textbf{and 28}.
 Freshly recruited respondents always answered all questions on social media use.
@@
 
-First, when looking at the variables from a descriptive perspective (Figure \ref{fig:descriptives}), we see that all well-being measures did not change substantially across the different waves of data collection.
+First, when looking at the variables from a descriptive perspective (Figure \ref{fig:fig-descriptives}), we see that all well-being measures did not change substantially across the different waves of data collection.
 COVID-19 related media use, however, decreased slightly at the beginning of the study and remained stable after approximately six waves.
@@
 The study's main hypothesis was that the effects of social media use on well-being would be trivial.
-Regarding the effects of different communication \emph{types}---that is, reading vs.~sharing vs.~posting---all within-person effects fell completely within the a-priori defined null region (see Figure \ref{fig:res-activity}).
-For example, respondents who used social media more frequently than usual to read about COVID-19 related topics did not show a simultaneous change in life satisfaction (\emph{b} = 0.05 {[}95\% CI -0.01, 0.1{]}).
-All confidence intervals included zero; hence, all effects were also statistically non-significant.
+Regarding the effects of different communication \emph{types}---that is, reading vs.~sharing vs.~posting---all within-person effects fell completely within the a-priori defined null region (see Figure \ref{fig:fig-between}).
+For example, respondents who used social media more frequently than usual to read about COVID-19 related topics did not show a simultaneous change in life satisfaction (\emph{b} = 0.04 {[}95\% CI -0.01, 0.09{]}).
 As a result, the hypothesis was supported for all COVID-19 related types of social media communication.
-
-Regarding between-person relations, about which no hypotheses were formulated, only three effects did not include zero.
-Respondents who across all waves used social media more frequently than others to read about COVID-19 related posts reported slightly lower levels of positive affect than others (\emph{b} = -0.03 {[}95\% CI \textgreater{} -0.01, -0.06{]}).
-Respondents who across all waves used social media more frequently than others to write COVID-19 related posts reported higher levels of negative affect than others (\emph{b} = 0.06 {[}95\% CI 0.09, 0.03{]}).
-Interestingly, respondents who across all waves used social media more frequently than others to write COVID-19 related posts also reported slightly higher levels of positive affect than others (\emph{b} = 0.05 {[}95\% CI 0.09, \textless{} 0.01{]}).
+\textbf{However, two effects were statistically significantly different from zero.}
+\textbf{Users who wrote more COVID-19 related posts than usual were also slightly less satisfied with their lives as usual (\emph{b} = -0.13 {[}95\% CI -0.21, -0.05{]}).}
+\textbf{Users who wrote more COVID-19 related posts than usual also experienced slightly more negative affect than usual (\emph{b} = 0.03 {[}95\% CI 0.01, 0.05{]}).}
+There was a small (statistically nonsignificant) trend that reading COVID-19 related content slightly increased life satisfaction (\emph{b} = 0.04 {[}95\% CI -0.01, 0.09{]}, \emph{p} = .078).
+At the same time, there was also a small (statistically nonsignificant) trend that reading COVID-19 related content decreased positive affect (\emph{b} = -0.02 {[}95\% CI -0.03, 0{]}, \emph{p} = .078).
+
+Regarding the COVID-19 related use of social media \emph{channels}, the results were comparable (see Figure \ref{fig:fig-within}).
+Changes in the frequency of using different social media channels to attain information regarding COVID-19 were unrelated to meaningful changes in well-being.
+For example, respondents who used Facebook more frequently than usual to learn about COVID-19 did not show a simultaneous change in well-being (\emph{b} = -0.04 {[}95\% CI -0.09, 0.02{]}).
+In sum, the hypothesis was supported also for the COVID-19 related use of important social media channels.
+\textbf{However, two effects differed substantially from zero.}
+\textbf{Respondents who used Instagram more frequently than usual to attain COVID-19 related news reported slightly \emph{lower} levels of negative affect than usual (\emph{b} = -0.02 {[}95\% CI -0.03, \textgreater{} -0.01{]}).}
+\textbf{Respondents who used YouTube more frequently than usual to attain COVID-19 related news reported slightly \emph{higher} levels of negative affect than usual (\emph{b} = 0.02 {[}95\% CI 0, 0.03{]}).}
+\textbf{However, both effects were still completely inside of the null region, hence not large enough to be considered meaningful.}
+
+For an overview of all within-person effects, see Table \ref{tab:tab-within} and Figure \ref{fig:fig-within}.
+
+\hypertarget{exploratory-analyses}{%
+\subsection{Exploratory Analyses}\label{exploratory-analyses}}
+
+In what follows, I briefly report some exploratory analyses that weren't preregistered.
+
+\hypertarget{between-person-relations}{%
+\subsubsection{Between-person relations}\label{between-person-relations}}
+
+Regarding between-person relations, about which no hypotheses were formulated, again no relation crossed or was completely outside of the SESOI.
+Four relations were statistically significant.
+Respondents who across all waves used social media more frequently than others to read about COVID-19 related posts reported slightly lower levels of positive affect than others (\emph{b} = -0.05 {[}95\% CI -0.02, -0.08{]}).
+Respondents who across all waves used social media more frequently than others to write COVID-19 related posts reported higher levels of negative affect than others (\emph{b} = 0.06 {[}95\% CI 0.10, 0.03{]}).
+Interestingly, respondents who across all waves used social media more frequently than others to write COVID-19 related posts also reported higher levels of positive affect (\emph{b} = 0.06 {[}95\% CI 0.11, 0.01{]}).
+Finally, respondents who across all waves used YouTube more frequently than others also reported slightly higher levels of life satisfaction than others (\emph{b} = 0.09 {[}95\% CI 0.16, 0.02{]}).
 However, note that the effect were still completely inside of the null region, hence not large enough to be considered practically relevant.
@@
 Actively posting on social media was significantly (though not meaningfully) related to decreased life satisfaction.
-However, when controlling for potential confounders, the effect became virtually zero (see Figure \ref{fig:res-channels}).
-
-Regarding the COVID-19 related use of social media \emph{channels}, the results were comparable (see Figure \ref{fig:res-channels}).
-Changes in the frequency of using different social media channels to attain information regarding COVID-19 were unrelated to meaningful changes in well-being.
-For example, respondents who used Facebook more frequently than usual to learn about COVID-19 did not show a simultaneous change in well-being (\emph{b} = -0.05 {[}95\% CI -0.11, 0.01{]}).
-Only two effects differed substantially from zero.
-Respondents who used Instagram more frequently than usual to attain COVID-19 related news reported slightly \emph{higher} levels of life satisfaction than usual (\emph{b} = 0.09 {[}95\% CI 0.02, 0.16{]}).
-Respondents who used Twitter more frequently than usual to attain COVID-19 related news reported slightly \emph{lower} levels of life satisfaction than usual (\emph{b} = -0.12 {[}95\% CI -0.23, -0.02{]}).
-However, both effects were still completely inside of the null region, hence not large enough to be considered meaningful.
-In sum, the hypothesis was supported also for the COVID-19 related use of important social media channels.
-
-In terms of between-person relations---which, again, weren't included in the hypotheses---no relations crossed the null region or fell outside of it.
-Only one relation did not include zero, was hence statistically significant.
-Respondents who across all waves used YouTube more frequently than others for COVID-19 related reasons reported marginally \emph{higher} levels of life satisfaction (\emph{b} = 0.08 {[}95\% CI \textless{} 0.01, 0.16{]}).
-However, please note that this effect again was not large enough to be considered practically relevant.
-
-Again, note that when comparing the results with and without control variables, the results differed.
-Especially on the between-person level, altogether five effects stopped being significant if they were controlled for additional variables.
-For example, using Instagram was significantly (though not meaningfully) related to increased life satisfaction.
-However, when controlling for additional covariates, the effect became virtually zero (see Figure \ref{fig:res-channels}).
-
-\hypertarget{exploratory-analyses}{%
-\subsection{Exploratory Analyses}\label{exploratory-analyses}}
-
-In what follows, I briefly report some exploratory analyses that weren't preregistered.
-First, to contextualize the results reported above and to see if the results included any meaningful effects at all, I also looked at the effect sizes of selected (cherry-picked) covariates.
+However, when controlling for potential confounders, the effect became virtually zero.
+
+For an overview of all between-person relations, see Figure \ref{fig:fig-between}.
+
+\hypertarget{covariates}{%
+\subsubsection{Covariates}\label{covariates}}
+
+To contextualize the results reported above and to see if the results included any meaningful effects at all, I also looked at the effect sizes of the covariates.
 Because each variable had different response options, we would actually need to define a SESOI for each variable, which for reasons of scope I cannot implement here.
@@
 which allows for a better comparison across the differently scaled variables.
-For what it's worth, as a rough estimate for the SESOI we can build on the typical convention that small effects start at \emph{r} = \textbar.10\textbar.
-The results showed that several effects fell outside of the SESOI, were hence considered meaningful.
+As a rough estimate for the SESOI we can build on the typical convention that small effects start at \emph{r} = \textbar.10\textbar.
+The results showed that several effects crossed or fell outside of the SESOI, were hence considered meaningful.
 This includes for example internal locus of control, health, satisfaction with democracy, or exercising.
-For an overview, see Figure \ref{fig:res-control}.
+For an overview, see Figure \ref{fig:fig-control}.
+
+\hypertarget{robustness-check}{%
+\subsubsection{Robustness-check}\label{robustness-check}}
 
@@
 In this study I analyzed the effects of COVID-19 related social media use on well-being.
-The data come from a panel study with 24 waves and are largely representative of the Austrian population.
+The data come from a panel study with 32 waves and are largely representative of the Austrian population.
 In a random effects model I separated between person relations from within-person effects and controlled for a large number of both stable and varying covariates, thereby aiming to assess causality.
-The results showed that within-person effects were trivial.
+The results showed that some statistically significant negative within-person effects existed, but that they were \textbf{very small and likely} trivial.
 People who used social media more than usual to learn about COVID-19 didn't show meaningful changes in their well-being.
@@
 The results imply that COVID-19 related social media use doesn't seem to be particularly relevant for well-being.
-Other factors among the third variables that were measured revealed larger effects or relations, suggesting that well-being is determined by alternative aspects such as health, satisfaction with democracy, locus of control, or exercising.
-According to this study, popular fears that ``doomscrolling'' or overusing social media during times of crises are not supported.
+Other factors among the third variables that were measured revealed larger effects or relations, suggesting that well-being is rather determined by alternative aspects such as health, satisfaction with democracy, locus of control, or exercising.
+According to this study, popular fears that ``doomscrolling'' or overusing social media during times of crises is detrimental are not supported.
+
+\textbf{That said, several preliminary and subtle trends can be observed.}
+\textbf{First, overall the results do suggest that effects of COVID-19 related social media use on well-being rather tend to take place in the negative as opposed to the positive spectrum.}
+\textbf{For example, people who wrote more COVID-19 related posts than usual reported slightly lower levels of life satisfaction than usual.}
+\textbf{Similarly, people who wrote more COVID-19 related posts than usual also reported slightly more negative affect.}
+\textbf{When writing posts and comments on social media, people explicitly and more deeply engage with COVID-19 related content.}
+\textbf{Unfortunately, in general online tonality is often more extreme, negative, or aggressive, which potentially affects their authors.}
+\textbf{Note that I controlled for whether or not participants had an corona infection during a specific wave, which rules out the potential explanation that having an infection was the root cause of increased communication and reduced well-being.}
+
+\textbf{The potential explanation that tonality might be a relevant factor at play here is also supported by the second trend.}
+\textbf{People, who during the pandemic spend more time on Instagram than usual, also experienced less negative affect than usual.}
+\textbf{Instagram is well-known for its positivity bias (Waterloo et al., 2018).}
+\textbf{Content is generally post positive, uplifting, and (self-)flattering.}
+\textbf{It seems the often-criticized positivity bias on Instagram might have been somewhat beneficial in times of the pandemic.}
+\textbf{The critique that the positivity bias necessarily leads to envy and negative feelings is one-sided, because positive content can also inspire and motivate users (Meier et al., 2020), which could be especially helpful in times of lockdown and home-office.}
+\textbf{To provide a concrete example, Instagram was successfully used as an interactive communication channel for first year students to have a better start into their new degree, effectively complementing alternative learning platform tools (Ye et al., 2020).}
+
+\textbf{Similarly, people who spent more time on YouTube than usual also reported slightly more negative affect than usual.}
+\textbf{Communication on YouTube is often found to be more negative and less polite compared to other SNSs (Halpern \& Gibbs, 2013).}
+\textbf{YouTube is also routinely linked to mis- and disinformation.}
+\textbf{Of the 69 most viewed videos on YouTube on COVID-19, 19 (27.5\%) contained nonfactual information (Li et al., 2020).}
+\textbf{Consuming more negative and misleading information might hence be a potential explanation for the slightly increased levels of negative affect.}
 
@@
 However, note that Bendau et al. (2021) analyzed cross-sectional data on a between-person level while not controlling for third variables, which is not optimal for investigating causal effects.
-On the other hand, the results are well-aligned with many seminal Communication theories such as the uses and gratifications model (Katz et al., 1973).
+On the other hand, the results are well-aligned with \textbf{mood management theory (Zillmann, 1988)} or the uses and gratifications approach (Katz et al., 1973).
 If effects were indeed profoundly negative on average, then people likely wouldn't spend so much time on social media.
@@
 
-If anything, two preliminary and subtle trends can be observed.
-First, in terms of media channels, using Twitter more than usual was related to slightly decreased levels of life satisfaction.
-Twitter is considered to have more negative affordances and tonality as compared to other networks such as Instagram (Waterloo et al., 2018), which might help explain the results.
-Instagram, on the other hand, was related to slightly increased levels of life satisfaction.
-To speculate, the often-criticized positivity bias on Instagram might have been somewhat beneficial in times of the pandemic.
-The critique that the positivity bias necessarily leads to envy and negative feelings is unrealistic and myopic, because positive content can also inspire and motivate users (Meier, Gilbert, Börner, \& Possler, 2020), which could be especially helpful in times of lockdown and home-office.
-Second, of all the three COVID-19 related social media activities, people who read about the pandemic more than others showed slightly decreased levels of positive affect, and people who actively posted about the pandemic more than others showed slightly increased levels of negative affect.
-On the other hand, however, people who posted more about COVID-19 also showed slightly higher levels of positive affects, so taken together the results are ambivalent.
-But future research might elaborate on these specific relations to probe their stability and relevance.
+\textbf{The results showed that it makes sense to analyze different communication types and communication channels, and that active and passive communication showed different results.}
+\textbf{Liking and sharing content did not show any within-person effects. }
+\textbf{Such rather low-key active behaviors do not seem to affect well-being at all.}
+\textbf{Regarding passive use, reading COVID-19 related posts is more ambivalent; results showed some weak trends towards a positive effect on life satisfaction, but a negative affect on mood.}
+\textbf{It might be that reading and informing oneself about COVID-19 on social media might be helpful in the long term, but more negative for short-term affect.}
+\textbf{Finally, proactively engaging via writing posts, the most active form of communication analyzed here, showed only negative effects on well-being.}
+\textbf{Together, the results support the findings from Valkenburg et al. (2022), who could not confirm the claim that active use is good, versus passive is bad.}
+\textbf{Focusing on communication channels, YouTube seems to be more negative, whereas Instagram is likely more positive.}
+\textbf{Again, these are only very small effects.}
+\textbf{Future research might elaborate on these specific relations to probe their stability and relevance.}
+
+\textbf{From a more political and societal perspective, I believe the results imply that it does makes sense critically to reflect upon COVID-19 related social media use.}
+\textbf{In terms of concrete recommendations, on average it might be slightly beneficial to post less actively about COVID-19 on social media and to spent less time on YouTube.}
+\textbf{Potential positive effects, however, will for many users likely not be noticeable.}
+\textbf{Results allow to reject a positive effect: Writing more posts on social media will likely not increase well-being.}
+\textbf{At all events, engaging in COVID 19-related social media use should, on average, not be a cause for concern.}
 
@@
 Together, this allowed for an improved perspective on assessing causality.
-That said, causality necessitates temporal order, and the cause needs to precede the effect.
-Regarding media use, such effects often happen immediately or shortly after use, necessitating intervals in the hours, minutes, or even seconds.
+\textbf{However, the opposite effect is also plausible, namely that well-being affects media use (Zillmann, 1988).}
+\textbf{While controlling for potential confounders can support claims of causality, the procedure implemented here cannot prove causality.}
+Causality necessitates temporal order, and the cause needs to precede the effect.
+\textbf{The challenge is that} regarding media use, such effects often happen immediately or shortly after use, necessitating intervals in the hours, minutes, or even seconds.
 In many cases only experience sampling studies asking users in the very moment can produce such knowledge.
 However, even then we don't know for certain if we actually measured the right interval.
-Effects depend on the intensity of use or the length of the interval, and to borrow the words from Rohrer and Murayama (2021), there is no such thing as ``the'' effect of social media use on well-being.
+Effects depend on the intensity of use or the length of the interval.
+To borrow the words from Rohrer and Murayama (2021), there is no such thing as ``the'' effect of social media use on well-being.
 Hence, to document how effects unfold, future research needs to employ different study designs probing different time lags.
@@
 Unfortunately, such procedures introduce other problems, especially related to privacy.
-Hence, for this type of research question it still seems necessary to use self-reported measures, and in many cases they can still be very informative (Verbeij, Pouwels, Beyens, \& Valkenburg, 2021).
+Hence, for this type of research question it still seems necessary to use self-reported measures, and in many cases they can still be very informative (Verbeij et al., 2021).
 
@@
 In this study, COVID-19 related social media use didn't meaningfully affect several indicators of well-being, including life satisfaction, positive affect, and negative affect.
-However, factors other than social media use were meaningfully related to well-being, such as physical health, exercise, satisfaction with democracy, or believing that one is in control of one's life.
-If it's our aim to improve well-being, it might hence be more fruitful not to focus so much on social media but to address other, more pertinent societal problems related to health care, regular exercise, or a functioning democratic system.
+\textbf{If people wrote more COVID-19 related posts than usual, or if they spent less time on Instagram and more time on YouTube, very small but statistically significant effects were found.}
+Notably, however, factors other than social media use were more meaningfully related to well-being, such as physical health, exercise, satisfaction with democracy, or believing that one is in control of one's life.
+If it's our aim to improve well-being \textbf{during a pandemic}, it might hence be more fruitful not to focus so much on social media but to address other, more pertinent societal problems related to health care, regular exercise, or a functioning democratic system.
 
@@
 Bendau, A., Petzold, M.
-B., Pyrkosch, L., Mascarell Maricic, L., Betzler, F., Rogoll, J., \ldots{} Plag, J.
+B., Pyrkosch, L., Mascarell Maricic, L., Betzler, F., Rogoll, J., Große, J., Ströhle, A., \& Plag, J.
 (2021).
@@
 \leavevmode\vadjust pre{\hypertarget{ref-bradleyStressMoodSmartphone2021}{}}%
-Bradley, A., \& Howard, A. (2021). \emph{Stress, mood, and smartphone use in {University} students: {A} 12-week longitudinal study}. {OSF Preprints}. \url{https://doi.org/10.31219/osf.io/frvpb}
+Bradley, A., \& Howard, A. (2021). \emph{Stress, mood, and smartphone use in {University} students: {A} 12-week longitudinal study}. \url{https://doi.org/10.31219/osf.io/frvpb}
 
@@
 
+\leavevmode\vadjust pre{\hypertarget{ref-dienlinDisplacementReinforcementReciprocity2017}{}}%
+Dienlin, T., Masur, P.
+K., \& Trepte, S.
+(2017).
+Displacement or reinforcement?
+{The} reciprocity of {FtF}, {IM}, and {SNS} communication and their effects on loneliness and life satisfaction.
+\emph{Journal of Computer-Mediated Communication}, \emph{22}(2), 71--87.
+\url{https://doi.org/10.1111/jcc4.12183}
+
 \leavevmode\vadjust pre{\hypertarget{ref-dornemannHowGoodBad2021}{}}%
@@
 {The} role of similarity}.
-{OSF Preprints}.
 \url{https://doi.org/10.31219/osf.io/sy2kd}
@@
 Well-being and social media: {A} systematic review of {Bergen Addiction Scales}.
-\emph{Future Internet}, \emph{12}(2), 24.
+\emph{Future Internet}, \emph{12}(2, 2), 24.
 \url{https://doi.org/10.3390/fi12020024}
@@
 Ellison, N.
-B., Triẹû, P., Schoenebeck, S., Brewer, R., \& Israni, A.
+B., Triệu, P., Schoenebeck, S., Brewer, R., \& Israni, A.
 (2020).
@@
 \leavevmode\vadjust pre{\hypertarget{ref-europeansocialsurveyESS9Edition20182021}{}}%
-European Social Survey. (2021). \emph{{ESS9} edition 3.1 - 2018 {Documentation Report}}.
+European Social Survey. (2021). \emph{{ESS9} edition 3.1 - 2018 {Documentation Report}}. \url{https://www.europeansocialsurvey.org/docs/round9/survey/ESS9_data_documentation_report_e03_1.pdf}
 
@@
 K., Branley-Bell, D., Markey, P., Ivory, J.
-D., Klisanin, D., \ldots{} Wilson, J.
+D., Klisanin, D., Elson, M., Smyth, M., Hogg, J.
+L., McDonnell, D., Nichols, D., Siddiqui, S., Gregerson, M., \& Wilson, J.
 (2021).
@@
 \leavevmode\vadjust pre{\hypertarget{ref-galerHowMuchToo2018}{}}%
-Galer, S. S. (2018). \emph{How much is {``too much time''} on social media?} https://www.bbc.com/future/article/20180118-how-much-is-too-much-time-on-social-media.
+Galer, S. S. (2018, January 19). \emph{How much is {``too much time''} on social media?} \url{https://www.bbc.com/future/article/20180118-how-much-is-too-much-time-on-social-media}
 
@@
 \url{https://doi.org/10.3390/ijerph19031635}
+
+\leavevmode\vadjust pre{\hypertarget{ref-halpernSocialMediaCatalyst2013}{}}%
+Halpern, D., \& Gibbs, J.
+(2013).
+Social media as a catalyst for online deliberation?
+{Exploring} the affordances of {Facebook} and {YouTube} for political expression.
+\emph{Computers in Human Behavior}, \emph{29}(3), 1159--1168.
+\url{https://doi.org/10.1016/j.chb.2012.10.008}
 
@@
 Csikszentmihalyi (Eds.), \emph{Handbook of research methods for studying daily life} (Paperback ed.).
-{New York, NY}: {Guilford}.
+{Guilford}.
 
@@
 No effect of different types of media on well-being.
-\emph{Scientific Reports}, \emph{12}(1), 61.
+\emph{Scientific Reports}, \emph{12}(1, 1), 61.
 \url{https://doi.org/10.1038/s41598-021-03218-7}
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-kittelAustrianCoronaPanel2020}{}}%
-Kittel, B., Kritzinger, S., Boomgaarden, H., Prainsack, B., Eberl, J.-M., Kalleitner, F., \ldots{} Schlogl, L.
-(2020).
-\emph{Austrian {Corona Panel Project} ({SUF} edition)}.
-{AUSSDA}.
-\url{https://doi.org/10.11587/28KQNS}
-
 \leavevmode\vadjust pre{\hypertarget{ref-kittelAustrianCoronaPanel2021}{}}%
-Kittel, B., Kritzinger, S., Boomgaarden, H., Prainsack, B., Eberl, J.-M., Kalleitner, F., \ldots{} Schlogl, L.
+Kittel, B., Kritzinger, S., Boomgaarden, H., Prainsack, B., Eberl, J.-M., Kalleitner, F., Lebernegg, N.
+S., Partheymüller, J., Plescia, C., Schiestl, D.
+W., \& Schlogl, L.
 (2021).
@@
 
+\leavevmode\vadjust pre{\hypertarget{ref-kittelAustrianCoronaPanel2020}{}}%
+Kittel, B., Kritzinger, S., Boomgaarden, H., Prainsack, B., Eberl, J.-M., Kalleitner, F., Lebernegg, N.
+S., Partheymüller, J., Plescia, C., Schiestl, D.
+W., \& Schlogl, L.
+(2020).
+\emph{Austrian {Corona Panel Project} ({SUF} edition)} {[}Data set{]}.
+{AUSSDA}.
+\url{https://doi.org/10.11587/28KQNS}
+
 \leavevmode\vadjust pre{\hypertarget{ref-kleinDarklySoothingCompulsion2021}{}}%
-Klein, J. (2021). \emph{The darkly soothing compulsion of 'doomscrolling'}. https://www.bbc.com/worklife/article/20210226-the-darkly-soothing-compulsion-of-doomscrolling.
+Klein, J. (2021, March 3). \emph{The darkly soothing compulsion of 'doomscrolling'}. \url{https://www.bbc.com/worklife/article/20210226-the-darkly-soothing-compulsion-of-doomscrolling}
 
 \leavevmode\vadjust pre{\hypertarget{ref-klinePrinciplesPracticeStructural2016}{}}%
-Kline, R. B. (2016). \emph{Principles and practice of structural equation modeling} (Fourth). {New York, NY}: {The Guilford Press}.
+Kline, R. B. (2016). \emph{Principles and practice of structural equation modeling} (4th ed.). {The Guilford Press}.
+
+\leavevmode\vadjust pre{\hypertarget{ref-latikkaLonelinessPsychologicalDistress2022}{}}%
+Latikka, R., Koivula, A., Oksa, R., Savela, N., \& Oksanen, A.
+(2022).
+Loneliness and psychological distress before and during the {COVID-19} pandemic: {Relationships} with social media identity bubbles.
+\emph{Social Science \& Medicine}, \emph{293}, 114674.
+\url{https://doi.org/10.1016/j.socscimed.2021.114674}
+
+\leavevmode\vadjust pre{\hypertarget{ref-liYouTubeSourceInformation2020}{}}%
+Li, H.
+O.-Y., Bailey, A., Huynh, D., \& Chan, J.
+(2020).
+{YouTube} as a source of information on {COVID-19}: A pandemic of misinformation?
+\emph{BMJ Global Health}, \emph{5}(5), e002604.
+\url{https://doi.org/10.1136/bmjgh-2020-002604}
 
@@
 \url{https://doi.org/10.2196/22142}
+
+\leavevmode\vadjust pre{\hypertarget{ref-livingstoneEuropeanResearchChildren2018}{}}%
+Livingstone, S., Mascheroni, G., \& Staksrud, E.
+(2018).
+European research on children's internet use: {Assessing} the past and anticipating the future.
+\emph{New Media \& Society}, \emph{20}(3), 1103--1122.
+\url{https://doi.org/10.1177/1461444816685930}
 
@@
 \leavevmode\vadjust pre{\hypertarget{ref-lykkenHappinessWhatStudies1999}{}}%
-Lykken, D. T. (1999). \emph{Happiness: {What} studies on twins show us about nature, nurture, and the happiness set-point}. {New York, N.Y.}: {Golden Books}.
+Lykken, D. T. (1999). \emph{Happiness: {What} studies on twins show us about nature, nurture, and the happiness set-point}. {Golden Books}.
 
@@
 \leavevmode\vadjust pre{\hypertarget{ref-mcelreathYesterdayClass2021}{}}%
-McElreath, R. (2021). Yesterday in class, ... {[}Tweet{]}.
+McElreath, R. (2021, January 28). \emph{Yesterday in class, ...} {[}Tweet{]}. {@rlmcelreath}. \url{https://twitter.com/rlmcelreath/status/1354786005996482563}
 
@@
 \emph{Medical Care}, \emph{41}(5), 582--592.
+\href{Retrieved\%20from\%20http://www.jstor.org/stable/3768017}{Retrieved from http://www.jstor.org/stable/3768017}
 
@@
 E., Holingue, C., Kalb, L.
-G., Bennett, D., Kapteyn, A., Jiang, Q., \ldots{} Thrul, J.
+G., Bennett, D., Kapteyn, A., Jiang, Q., Veldhuis, C.
+B., Johnson, R.
+M., Fallin, M.
+D., Kreuter, F., Stuart, E.
+A., \& Thrul, J.
 (2020).
@@
 Sandstrom, G., Buchanan, K., Aknin, L., \& Lotun, S.
-(2021).
-Doomscrolling {COVID} news takes an emotional toll \textendash{} here's how to make your social media a happier place.
-http://theconversation.com/doomscrolling-covid-news-takes-an-emotional-toll-heres-how-to-make-your-social-media-a-happier-place-170342.
+(2021, October 22).
+\emph{Doomscrolling {COVID} news takes an emotional toll -- here's how to make your social media a happier place}.
+{The Conversation}.
+\url{http://theconversation.com/doomscrolling-covid-news-takes-an-emotional-toll-heres-how-to-make-your-social-media-a-happier-place-170342}
 
@@
 (2016).
-The accuracy of self-reported {Internet} use\textemdash{{A}} validation study using client log data.
+The accuracy of self-reported {Internet} use---{A} validation study using client log data.
 \emph{Communication Methods and Measures}, \emph{10}(1), 13--27.
@@
 \leavevmode\vadjust pre{\hypertarget{ref-sheldonStabilityHappinessTheories2014}{}}%
-Sheldon, K. M., \& Lucas, R. E. (2014). \emph{Stability of happiness: Theories and evidence on whether happiness can change}.
+Sheldon, K. M., \& Lucas, R. E. (2014). \emph{Stability of happiness: Theories and evidence on whether happiness can change}. \url{http://site.ebrary.com/id/10891875}
 
@@
 Statista.
-(2021).
+(2021, May 21).
 \emph{Average daily time spent on social networks by users in the {United States} from 2018 to 2022}.
-https://www.statista.com/statistics/1018324/us-users-daily-social-media-minutes/.
+\url{https://www.statista.com/statistics/1018324/us-users-daily-social-media-minutes/}
 
@@
 Valkenburg, P. M., \& Peter, J. (2013). The differential susceptibility to media effects model. \emph{Journal of Communication}, \emph{63}(2), 221--243. \url{https://doi.org/10.1111/jcom.12024}
+
+\leavevmode\vadjust pre{\hypertarget{ref-valkenburgAssociationsActivePassive2022}{}}%
+Valkenburg, P.
+M., van Driel, I.
+I., \& Beyens, I.
+(2022).
+The associations of active and passive social media use with well-being: {A} critical scoping review.
+\emph{New Media \& Society}, \emph{24}(2), 530--549.
+\url{https://doi.org/10.1177/14614448211065425}
 
@@
 J., Ferguson, C.
-J., Colder Carras, M., Kardefelt-Winther, D., Shi, J., Aarseth, E., \ldots{} Przybylski, A.
+J., Colder Carras, M., Kardefelt-Winther, D., Shi, J., Aarseth, E., Bean, A.
+M., Bergmark, K.
+H., Brus, A., Coulson, M., Deleuze, J., Dullur, P., Dunkels, E., Edman, J., Elson, M., Etchells, P.
+J., Fiskaali, A., Granic, I., Jansz, J., \ldots{} Przybylski, A.
 K.
@@
 Wagner, M., Aichholzer, J., Eberl, J.-M., Meyer, T.
-M., Berk, N., Büttner, N., \ldots{} Müller, W.
+M., Berk, N., Büttner, N., Boomgaarden, H., Kritzinger, S., \& Müller, W.
 C.
 (2018).
-\emph{{AUTNES Online Panel Study} 2017 ({SUF} edition)}.
+\emph{{AUTNES Online Panel Study} 2017 ({SUF} edition)} {[}Data set{]}.
 {AUSSDA}.
@@
 {A} qualitative exploration of online video watching among {Chinese} young adults.
-\emph{International Journal of Environmental Research and Public Health}, \emph{18}(14), 7247.
+\emph{International Journal of Environmental Research and Public Health}, \emph{18}(14, 14), 7247.
 \url{https://doi.org/10.3390/ijerph18147247}
+
+\leavevmode\vadjust pre{\hypertarget{ref-yeTurningInformationDissipation2020}{}}%
+Ye, S., Hartmann, R.
+W., Söderström, M., Amin, M.
+A., Skillinghaug, B., Schembri, L.
+S., \& Odell, L.
+R.
+(2020).
+Turning information dissipation into dissemination: {Instagram} as a communication enhancing tool during the {COVID-19} pandemic and beyond.
+\emph{Journal of Chemical Education}, \emph{97}(9), 3217--3222.
+\url{https://doi.org/10.1021/acs.jchemed.0c00724}
+
+\leavevmode\vadjust pre{\hypertarget{ref-yuePassiveSocialMedia2022}{}}%
+Yue, Z., Zhang, R., \& Xiao, J.
+(2022).
+Passive social media use and psychological well-being during the {COVID-19} pandemic: {The} role of social comparison and emotion regulation.
+\emph{Computers in Human Behavior}, \emph{127}, 107050.
+\url{https://doi.org/10.1016/j.chb.2021.107050}
 
@@
 Well-being &  &  &  & \\
-\ \ \ Life satisfaction & 1.68 & 6.38 & 6.81 & 6.59\\
-\ \ \ Positive affect & 0.57 & 3.05 & 3.28 & 3.15\\
-\ \ \ Negative affect & 0.39 & 1.66 & 1.81 & 1.73\\
+\ \ \ Life satisfaction & 1.67 & 6.29 & 6.77 & 6.54\\
+\ \ \ Positive affect & 0.58 & 3.05 & 3.29 & 3.14\\
+\ \ \ Negative affect & 0.41 & 1.71 & 1.86 & 1.79\\
 Social media use &  &  &  & \\
-\ \ \ Read & 1.03 & 2.10 & 2.92 & 2.43\\
-\ \ \ Like \& share & 0.87 & 1.61 & 1.92 & 1.78\\
-\ \ \ Posting & 0.63 & 1.33 & 1.46 & 1.39\\
+\ \ \ Read & 1.02 & 2.01 & 2.93 & 2.37\\
+\ \ \ Like \& share & 0.85 & 1.62 & 1.96 & 1.77\\
+\ \ \ Posting & 0.62 & 1.32 & 1.66 & 1.43\\
 Social media channel &  &  &  & \\
-\ \ \ Facebook & 0.96 & 2.34 & 2.68 & 2.45\\
-\ \ \ Twitter & 0.52 & 1.16 & 1.72 & 1.37\\
-\ \ \ Instagram & 0.83 & 1.84 & 2.66 & 2.09\\
-\ \ \ WhatsApp & 1.23 & 2.28 & 2.62 & 2.45\\
-\ \ \ YouTube & 0.88 & 1.77 & 2.30 & 2.00\\
+\ \ \ Facebook & 0.97 & 2.22 & 2.67 & 2.40\\
+\ \ \ Twitter & 0.53 & 1.16 & 2.19 & 1.43\\
+\ \ \ Instagram & 0.84 & 1.90 & 2.51 & 2.12\\
+\ \ \ WhatsApp & 1.23 & 2.31 & 2.63 & 2.44\\
+\ \ \ YouTube & 0.88 & 1.80 & 2.24 & 2.01\\
 \bottomrule
 \end{tabular}
+
+\end{threeparttable}
+\end{center}
+
+\end{table}
+
+\newpage
+
+\begin{table}[tbp]
+
+\begin{center}
+\begin{threeparttable}
+
+\caption{\label{tab:tab-within}Overview of all within-person effects.}
+
+\footnotesize{
+
+\begin{tabular}{lrrrrr}
+\toprule
+ &  & \multicolumn{2}{c}{Confidence interval}  &  &\\
+\cmidrule(r){3-4}
+Predictor & \multicolumn{1}{c}{b} & \multicolumn{1}{c}{Lower level} & \multicolumn{1}{c}{Higher level} & \multicolumn{1}{c}{beta} & \multicolumn{1}{c}{p}\\
+\midrule
+Life satisfaction &  &  &  &  & \\
+\ \ \ Reading & 0.04 & -0.01 & 0.09 & 0.03 & .078\\
+\ \ \ Liking \& Sharing & 0.01 & -0.05 & 0.07 & 0.01 & .676\\
+\ \ \ Posting & -0.13 & -0.21 & -0.05 & -0.05 & .002\\
+\ \ \ Facebook & -0.04 & -0.09 & 0.02 & -0.03 & .167\\
+\ \ \ Instagram & 0.05 & -0.01 & 0.11 & 0.03 & .103\\
+\ \ \ WhatsApp & -0.01 & -0.05 & 0.04 & -0.01 & .735\\
+\ \ \ YouTube & 0.02 & -0.04 & 0.08 & 0.01 & .579\\
+\ \ \ Twitter & -0.07 & -0.16 & 0.02 & -0.03 & .133\\
+Positive affect &  &  &  &  & \\
+\ \ \ Reading & -0.02 & -0.03 & 0.00 & -0.02 & .078\\
+\ \ \ Liking \& Sharing & 0.00 & -0.02 & 0.02 & 0.00 & .975\\
+\ \ \ Posting & -0.02 & -0.05 & 0.01 & -0.02 & .150\\
+\ \ \ Facebook & 0.01 & -0.01 & 0.02 & 0.01 & .554\\
+\ \ \ Instagram & 0.00 & -0.02 & 0.03 & 0.01 & .670\\
+\ \ \ WhatsApp & 0.00 & -0.02 & 0.01 & 0.00 & .893\\
+\ \ \ YouTube & 0.01 & -0.01 & 0.03 & 0.02 & .183\\
+\ \ \ Twitter & 0.02 & -0.02 & 0.05 & 0.01 & .335\\
+Negative affect &  &  &  &  & \\
+\ \ \ Reading & 0.00 & -0.01 & 0.01 & 0.00 & .790\\
+\ \ \ Liking \& Sharing & 0.01 & -0.01 & 0.02 & 0.01 & .281\\
+\ \ \ Posting & 0.03 & 0.01 & 0.05 & 0.02 & .008\\
+\ \ \ Facebook & 0.00 & -0.01 & 0.01 & 0.00 & .913\\
+\ \ \ Instagram & -0.02 & -0.03 & 0.00 & -0.02 & .047\\
+\ \ \ WhatsApp & 0.00 & -0.01 & 0.01 & 0.00 & .651\\
+\ \ \ YouTube & 0.02 & 0.00 & 0.03 & 0.02 & .031\\
+\ \ \ Twitter & 0.02 & -0.01 & 0.04 & 0.02 & .137\\
+\bottomrule
+\end{tabular}
+
+}
 
@@
 \includegraphics{manuscript_files/figure-latex/sesoi-1.pdf}
-\caption{\label{fig:sesoi}Illustration of how confidence intervals are used to test a null region.
-Here, a trivial effect of social media use on life satisfaction is defined as ranging from b = -.25 to b = .25}
+\caption{\label{fig:sesoi}Using confidence intervals to test a null region. Note. Here, a trivial effect of social media use on life satisfaction is defined as ranging from b = -.25 to b = .25}
 \end{figure}
@@
 \begin{figure}
-\includegraphics[width=\textwidth]{figures/fig_descriptives} \caption{Development of well-being and media use measures across the pandemic.
-Values obtained from mixed effect models, with participants and waves as grouping factors and without additional predictors.}\label{fig:descriptives}
+\includegraphics[width=\textwidth]{figures/fig_descriptives} \caption{Well-being and media use across the 32 waves.
+Note.
+Values obtained from mixed effect models, with participants and waves as grouping factors and without additional predictors.}\label{fig:fig-descriptives}
 \end{figure}
@@
 \begin{figure}
-\includegraphics[width=\textwidth]{manuscript_files/figure-latex/res-activity-1} \caption{The effects of various types of social media use on three indicators of well-being.
+\includegraphics[width=\textwidth]{manuscript_files/figure-latex/fig-within-1} \caption{Within-person effects of COVID-19 related social media use on well-being.
+Note.
 The black estimates show the effects controlled for a large number of covariates (see text; preregistered); the grey estimates are without control variables (exploratory).
 The SESOI was b = |0.30| for life satisfaction and b = |0.15| for affect.
-Hence, all of the reported effects are not considered meaningful.}\label{fig:res-activity}
+Hence, all of the reported effects are not considered meaningful.}\label{fig:fig-within}
 \end{figure}
@@
 \begin{figure}
-\includegraphics[width=\textwidth]{manuscript_files/figure-latex/res-channels-1} \caption{The effects of using various social media applications on three indicators of well-being.
+\includegraphics[width=\textwidth]{manuscript_files/figure-latex/fig-between-1} \caption{Between-person relations between COVID-19 related social media use and well-being.
+Note.
 The black estimates show the effects controlled for a large number of covariates (see text; preregistered); the grey estimates are without control variables (exploratory).
 The SESOI was b = |0.30| for life satisfaction and b = |0.15| for affect.
-Hence, all of the reported effects are not considered meaningful.}\label{fig:res-channels}
+Hence, all of the reported effects are not considered meaningful.}\label{fig:fig-between}
 \end{figure}
@@
 \begin{figure}
-\includegraphics[width=\textwidth]{manuscript_files/figure-latex/res-control-1} \caption{Results of selected covariates.
-All variables were standardize except 'Male'  and ‘Employed in public service', because there were measured on a binary scale.}\label{fig:res-control}
+\includegraphics[width=\textwidth]{manuscript_files/figure-latex/fig-control-1} \caption{Results of selected covariates. Note. All variables standardized except 'Male'.}\label{fig:fig-control}
 \end{figure}

### 2022-07-25 preregistration_changes.Rmd
- Status of paper when submitted to ICA
- major updates for revision - included multiple imputation - changed theory - include new data-set - extended discussion

@@
 ---
-title: "Items: English"
+title: "Preregistration: Changes"
 output:
@@
 
-## Analyses
+## Imputation
 
-I planned to conduct also the following analyses. I the time of writing, I simply hadn't yet found the time to do so.
+Reviewer feedback required changes regarding the imputation strategy
 
-	- Impute missing values
-	- Report CFAs for scales
-	- Run analyses using factor scores
+- Before: predictive mean matching with n = 1 data-sets; now: multiple imputation (also predictive mean matching) with n = 5 data-sets.
+- Before: remove participants with > 50% missing data; now: include all participants
+- Before: remove variables leading to much attrition; now: include all variables
+
+## Estimation
+
+Positive affect and negative affect were modeled as mean scores and not as factor scores. The reason was that it wasn't possible to extract factor scores for the data set with multiple imputation.

### 2022-07-25 title page/title_page.Rmd
- major updates for revision - included multiple imputation - changed theory - include new data-set - extended discussion

---
title             : "The effects of COVID-19 related social media use on well-being"
shorttitle        : "COVID-19 related social media use and well-being"

author: 
  - name          : "Tobias Dienlin"
    affiliation   : "1"
    corresponding : yes
    address       : "University of Vienna, Department of Communication, 1090 Vienna, Austria"
    email         : "tobias.dienlin@univie.ac.at"

affiliation:
  - id            : "1"
    institution   : "University of Vienna"

authornote: | 
  Tobias Dienlin, Department of Communication, University of Vienna.

abstract          : "In times of crisis such as the Corona pandemic citizens need to stay informed about recent events, political decisions, or mandatory protection measures. To this end, many people use various types of media, and increasingly social media. However, because social media are particularly engaging, some find it hard to disconnect. In this preregistered study, I investigated whether using social media for COVID-19 related reasons affected psychological well-being. To answer this question I analyzed data from the Austrian Corona Panel Project, which consists of 3,018 participants. Well-being was measured at all 24 waves, and communication at five specific waves. I ran three random effects within between models, controlling for several stable and varying confounders. Results showed that the effects of COVID-19 related social media use on well-being were very small, arguably too small to matter. Fears that social media use during times of crisis critically impairs well-being are not supported."
keywords          : "COVID-19, Coronavirus, well-being, affect, life satisfaction, social media use, news use, communication, random effects within between model, panel study, longitudinal study."

bibliography      : 
  - "bibliography/r-references.bib"
  - "bibliography/references.bib"

floatsintext      : yes
figurelist        : no
tablelist         : no
footnotelist      : no
linenumbers       : yes
mask              : no
draft             : no

documentclass     : "apa6"
classoption       : "man"
output            : papaja::apa6_pdf
# output            : papaja::apa6_word
header-includes:
  - \setlength{\parskip}{0em}
  - \raggedbottom
  # - \note{\clearpage}
---

```{r analysis-preferences, include=F, cache=F}
# Seed for random number generation
set.seed(42)
knitr::opts_chunk$set(cache.extra = knitr::rand_seed,
                      cache=T, echo=F
                      )
```

```{r setup, include=F, cache=F}

# install necessary packages
# devtools::install_github("https://github.com/tdienlin/td@v.0.0.2.5")
# devtools::install_github("https://github.com/crsh/papaja@devel")

# load packages
packages <- c("devtools", "english", "ggplot2", "kableExtra", 
              "lavaan", "lme4", "lmerTest",
              "magrittr", "semPlot", "tidyverse")
lapply(c(packages, "papaja", "td"), library, character.only = TRUE)

# load workspace created from analyses.rmd to use results
load("data/workspace_2.RData")
```

### 2022-07-25 title page/title_page.tex
- major updates for revision - included multiple imputation - changed theory - include new data-set - extended discussion

--- 2022-04-12 manuscript.tex
@@
 \documentclass[
-  english,
-  man,mask]{apa6}
-\title{Analyzing the effects of COVID-19 related social media use on well-being}
-\author{Name blinded for review\textsuperscript{1}}
-\date{}
-
+  man,floatsintext]{apa6}
 \usepackage{amsmath,amssymb}
@@
   pdftitle={Analyzing the effects of COVID-19 related social media use on well-being},
-  pdfauthor={Name blinded for review1},
+  pdfauthor={Tobias Dienlin1},
   pdflang={en-EN},
-  pdfkeywords={COVID-19, well-being, social media, news use, panel study.},
+  pdfkeywords={COVID-19, Coronavirus, well-being, affect, life satisfaction, social media use, news use, communication, random effects within between model, panel study, longitudinal study.},
   hidelinks,
@@
 \fi
-\newlength{\cslhangindent}
-\setlength{\cslhangindent}{1.5em}
-\newlength{\csllabelwidth}
-\setlength{\csllabelwidth}{3em}
-\newlength{\cslentryspacingunit} % times entry-spacing
-\setlength{\cslentryspacingunit}{\parskip}
-\newenvironment{CSLReferences}[2] % #1 hanging-ident, #2 entry spacing
- {% don't indent paragraphs
-  \setlength{\parindent}{0pt}
-  % turn on hanging indent if param 1 is 1
-  \ifodd #1
-  \let\oldpar\par
-  \def\par{\hangindent=\cslhangindent\oldpar}
+\ifLuaTeX
+\usepackage[bidi=basic]{babel}
+\else
+\usepackage[bidi=default]{babel}
 \fi
-  % set entry spacing
-  \setlength{\parskip}{#2\cslentryspacingunit}
- }%
- {}
-\usepackage{calc}
-\newcommand{\CSLBlock}[1]{#1\hfill\break}
-\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\csllabelwidth}{#1}}
-\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth - \csllabelwidth}{#1}\break}
-\newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}
+\babelprovide[main,import]{english}
+% get rid of language-specific shorthands (see #6817):
+\let\LanguageShortHands\languageshorthands
+\def\languageshorthands#1{}
 % Manuscript styling
@@
 
+% Overwrite redefinition of paragraph and subparagraph by the default LaTeX template
+% See https://github.com/crsh/papaja/issues/292
+\makeatletter
+\renewcommand{\paragraph}{\@startsection{paragraph}{4}{\parindent}%
+  {0\baselineskip \@plus 0.2ex \@minus 0.2ex}%
+  {-1em}%
+  {\normalfont\normalsize\bfseries\itshape\typesectitle}}
+
+\renewcommand{\subparagraph}[1]{\@startsection{subparagraph}{5}{1em}%
+  {0\baselineskip \@plus 0.2ex \@minus 0.2ex}%
+  {-\z@\relax}%
+  {\normalfont\normalsize\itshape\hspace{\parindent}{#1}\textit{\addperi}}{\relax}}
+\makeatother
+
 % \usepackage{etoolbox}
@@
 \makeatother
-\shorttitle{COVID-19 related social media use and well-being}
-\keywords{COVID-19, well-being, social media, news use, panel study.}
-\DeclareDelayedFloatFlavor{ThreePartTable}{table}
-\DeclareDelayedFloatFlavor{lltable}{table}
-\DeclareDelayedFloatFlavor*{longtable}{table}
+
+\usepackage{xpatch}
 \makeatletter
-\renewcommand{\efloat@iwrite}[1]{\immediate\expandafter\protected@write\csname efloat@post#1\endcsname{}}
-\makeatother
+\xapptocmd\appendix
+  {\xapptocmd\section
+    {\addcontentsline{toc}{section}{\appendixname\ifoneappendix\else~\theappendix\fi\\: #1}}
+    {}{\InnerPatchFailed}%
+  }
+{}{\PatchFailed}
+\keywords{COVID-19, Coronavirus, well-being, affect, life satisfaction, social media use, news use, communication, random effects within between model, panel study, longitudinal study.}
 \usepackage{lineno}
@@
 \usepackage{csquotes}
-\usepackage{endnotes}
 \setlength{\parskip}{0em}
 \raggedbottom
-\note{\clearpage}
-\ifXeTeX
-  % Load polyglossia as late as possible: uses bidi with RTL langages (e.g. Hebrew, Arabic)
-  \usepackage{polyglossia}
-  \setmainlanguage[]{english}
-\else
-  \usepackage[main=english]{babel}
-% get rid of language-specific shorthands (see #6817):
-\let\LanguageShortHands\languageshorthands
-\def\languageshorthands#1{}
-\fi
 \ifLuaTeX
@@
 
-\affiliation{\vspace{0.5cm}\textsuperscript{1} }
+\title{Analyzing the effects of COVID-19 related social media use on well-being}
+\author{Tobias Dienlin\textsuperscript{1}}
+\date{}
 
-\abstract{
+\shorttitle{COVID-19 related social media use and well-being}
+
+\authornote{
+
+Tobias Dienlin, Department of Communication, University of Vienna.
+
+Correspondence concerning this article should be addressed to Tobias Dienlin, University of Vienna, Department of Communication, 1090 Vienna, Austria.
+E-mail: \href{mailto:tobias.dienlin@univie.ac.at}{\nolinkurl{tobias.dienlin@univie.ac.at}}
+
+}
+
+\affiliation{\vspace{0.5cm}\textsuperscript{1} University of Vienna}
+
+\abstract{%
 In times of crisis such as the Corona pandemic citizens need to stay informed about recent events, political decisions, or mandatory protection measures.
@@
 
-During the COVID-19 pandemic,
-numerous events unfolded in quick succession and several open questions emerged.
-How dangerous is the virus?
-Is it spreading in my region?
-How is it transmitted, and how can I protect myself?
-Because for many it was (and still is) a matter of life or death, people aimed to stay informed regarding the latest developments.
-Governments around the world implemented safety measures, such as wearing masks, keeping physical distance, or enforcing lockdowns.
-In this extraordinary situation, many people used media excessively to attain information, and especially social media were at an all time high (Statista, 2021).
-
-Some people actually couldn't stop using social media to learn about COVID-19 related news.
-A new phenomenon emerged, termed ``doomscrolling'':
-Users were glued to their screens and found it hard to pursue other relevant activities such as working, taking a break, or looking after their children (Klein, 2021).
-It was asked whether using social media for COVID-19 related reasons is helpful, or whether it creates an additional burden on mental health (Sandstrom, Buchanan, Aknin, \& Lotun, 2021).
-These concerns seem justified:
-A study with 6,233 people from Germany conducted during the pandemic found that ``{[}f{]}requency, duration and diversity of media exposure were positively associated with more symptoms of depression'' (Bendau et al., 2021, p.
-283).
-
-As a result, with this study I want to build on this research and investigate whether or not COVID-19 related social media use affected well-being during the pandemic.
-To this end, I analyzed a large-scale panel study from the Austrian Corona Panel Project (Kittel et al., 2020).
-The panel consists of 24 waves and has an overall sample size of 3,018 participants.
-The panel study collected a large number of psychological and demographic variables.
-I explicitly aimed to investigate the causal effects of COVID-19 related social media use on well-being.
-
-\hypertarget{understanding-well-being-and-media-use}{%
-\subsection{Understanding Well-being and Media Use}\label{understanding-well-being-and-media-use}}
-
-The underlying theories that guided the selection of variables for my analysis are the two-continua model of mental health (Greenspoon \& Saklofske, 2001) and the hierarchical taxonomy of computer-mediated communication (Meier \& Reinecke, 2020).
-According to the two-continua model, mental health consists of (a) psychopathology and (b) well-being.
-Well-being can be differentiated into subjective and psychological well-being (Diener, Lucas, \& Oishi, 2018).
-Whereas subjective well-being emphasizes hedonic aspects such as happiness and joy, psychological well-being addresses eudaimonic aspects such as fulfillment and meaning.
-Subjective well-being is primarily about achieving positive affect and avoiding negative affect.
-One of the most prominent indicators of well-being is life satisfaction.
-In my view, because it represents a general appraisal of one's life, life satisfaction is best thought of as a meta concept that combines psychological and subjective well-being.
-Notably, life satisfaction is stable and fluctuates only little, whereas it's the exact opposite for affect (Dienlin \& Johannes, 2020).
-To capture well-being in this study I thus build on life satisfaction, positive affect, and negative affect.
-Together, this should provide an encompassing perspective on potential media effects.
-
-The hierarchical taxonomy of computer-mediated communication differentiates six levels of how people engage with digital technology.
-First, the device (e.g., smartphone); second, the type of application (e.g., social networking site); third, the branded application (e.g., Twitter); fourth, the feature (e.g., status post); fifth, the interaction (e.g., one-to-many); and sixth, the message (e.g., content) (Meier \& Reinecke, 2020).
-Whereas the first four levels focus on the communication \emph{channel}, the last two address the communication \emph{type}.
-To measure social media use for the consumption of COVID-19 related news and topics, I here employ both the channel and the type of communication perspective, which together provides a nuanced understanding of communication.
-
-First, I investigate how well-being is affected by different types of communication affect, namely active and passive use.
-Defining what constitutes active and what passive use is not always clear, and different understandings are currently discussed Meier \& Reinecke (2020).
-Reading is generally considered as passive and writing as active, while there are also specific behaviors such as such liking or sharing content that fall somewhere in-between (Meier \& Krause, 2022).
-In this study, I hence distinguish (a) \emph{reading} (passive), (b) \emph{posting} (active), and (c) \emph{liking and sharing} COVID-19 related posts (both active and passive).
-Second, I analyze how using the most prominent branded applications affects well-being, and whether this effect changes across applications.
-Branded apps are separate entities with potentially divergent effects.
-Twitter might have a different effect as compared to WhatsApp because of their respective affordances.
-For example, Waterloo, Baumgartner, Peter, and Valkenburg (2018) found that it's more adequate to express negative emotions on WhatsApp than on Twitter or on Instagram.
-The branded applications investigated here are Facebook, Twitter, Instagram, WhatsApp, and YouTube.
-Worth noting, this study is not about \emph{general} social media use during times of COVID, but on social media use \emph{focused} on COVID-19 related content.
-Examples of such media use include posting thoughts about the pandemic or retweeting COVID-19 related news.
-
-\hypertarget{theorizing-social-media-effects-on-well-being}{%
-\subsection{Theorizing Social Media Effects on Well-Being}\label{theorizing-social-media-effects-on-well-being}}
-
-From a theoretical perspective, how could we explain whether COVID-19 related social media use might affect well-being?
-According to the set-point model of subjective well-being, well-being is surprisingly stable (Lykken, 1999).
-Although specific events such as marriage or salary can have significant effects, after some time well-being routinely returns to prior levels, which are mostly determined genetically (Sheldon \& Lucas, 2014).
-Only very specific events such as unemployment, disability, or death can cause long-term decreases in well-being (Lucas, 2007).
-So although well-being can change this does not happen easily.
-
-Can media use be such a negative or positive factor?
-In advance, there doesn't seem to be a clear winner, and it's likely that both positive and negative effects cancel each other out.
-Empirically, social media use---on average---does not have a strong effect on well-being (Meier \& Reinecke, 2020).
-According to the different susceptibility of media effects model (Valkenburg \& Peter, 2013), the effects of media use differ across individuals.
-Whereas for some media are beneficial, for others they are harmful.
-On average, however, effects are often small or negligible.
-Also when focusing on individual users, social media have both positive and negative effects on well-being (Büchi, 2021).
-They can impair well-being when causing embarrassment, stress, or disinformation, and they can improve well-being when providing connectedness, information, or entertainment (Büchi, 2021).
-
-Two of the most prominent media effect theories argue against strong negative impacts.
-First, uses and gratifications theory states that people explicitly and rationally chose specific media because of their respective benefits.
-If those benefits don't exist, they invest their time elsewhere.
-And social media offer ample benefits.
-Most prominently, they help find relevant information, maintain and foster relationships, express one's personality, and entertain oneself (Pelletier, Krallman, Adams, \& Hancock, 2020).
-The second major theory is mood management theory (Zillmann, 1988).
-Users implicitly learn what media help them balance their mood and affect.
-For example, when bored people use social media to entertain themselves.
-Using experience sampling of well-being and logs of social media use, a study with 82 participants from Italy found that after episodes of social media use, levels of positive affect increased for a short time (Marciano, Driver, Schulz, \& Camerini, 2022).
-
-Precisely because social media have so many positive consequences, one can ask if this is the main problem.
-In other words, social media aren't problematic because they're inherently bad, but rather because they're too good.
-And as with many other things, there can be too much of a good thing.
-It is therefore often asked whether social media are addictive, and users sometimes express this fear themselves (Yang, Griffiths, Yan, \& Xu, 2021).
-However, a recently published meta-analysis found that the two most prominent measures of addiction, the Bergen Facebook Addiction Scale and the Bergen Social Media Addiction Scale, have only small relations to well-being (Duradoni, Innocenti, \& Guazzini, 2020).
-In addition, the general idea of labeling excessive social and new media use as addiction was criticized, arguing that social and new media represent new regular behaviors that should not be pathologized (Galer, 2018; van Rooij et al., 2018).
-
-Because effects can differ across situations, I now briefly focus on the effects of COVID-19 related social media use specifically.
-First, one could assume a \emph{direct} negative effect on well-being, and especially on positive or negative affect, which are more volatile and fluctuating.
-Dangers, inequalities, corruption---these were the headlines during the pandemic across many countries worldwide.
-If one learns about such events, the initial reaction might be shock, fear, or dismay.
-Consuming such news can be depressing (Dörnemann, Boenisch, Schommer, Winkelhorst, \& Wingen, 2021), perhaps even changing some general perspectives on life.
-That said, because not all news was negative, and because many people showed solidarity and compassion, there was also positive and uplifting content, potentially compensating for the negative effects (Dörnemann et al., 2021).
-A study with 2.057 respondents from Italy reported that during the pandemic virtual community and social connectedness even increased during the pandemic (Guazzini, Pesce, Marotta, \& Duradoni, 2022).
-
-There could also be \emph{indirect} effects.
-When browsing social media for Covid-19 related news, many users reported being captivated to such an extent that they could not stop using social media (Klein, 2021).
-During the pandemic social media use was at an all-time high in the US (Statista, 2021).
-It is most likely that moderate social media use is not detrimental (Orben, 2020).
-Overuse, however, might be more critical, and several studies have shown more pronounced negative effects for extreme users (Przybylski \& Weinstein, 2017).
-To explain, overuse could impair well-being if it replaces meaningful or functional activities such as meeting others, working, actively relaxing, or exercising.
-So if a society collectively overuses social media during a pandemic, there might be potential for negative effects.
-On the other hand, one can make the case that overuse might also be beneficial, especially in times of a pandemic---even if the use is mainly COVID-19 related.
-Exchanging COVID-19 related messages with friends via WhatsApp might replace the in-person contact one would have otherwise, but which is literally impossible at the time.
-In situations where meaningful and functional activities are prohibited, using social media to exchange about COVID-19 related topics might not be the worst idea.
-Besides, given that nowadays a large number of experts, scientists, and politicians converse directly on social media, one can get first-hand high quality information on current developments.
-
-To summarize, from a theoretical perspective it is most likely that the average effects of social media use on well-being are negligible.
-Building on established theories from Communication, it is not particularly likely that effects are either profoundly negative or strongly positive.
-
-\hypertarget{empirical-studies-on-social-media-effects}{%
-\subsection{Empirical Studies on Social Media Effects}\label{empirical-studies-on-social-media-effects}}
-
-So far, there is only little empirical research on how well-being is affected by COVID-19 related social media.
-In their study on the relations between media use and mental health during the pandemic, Bendau et al.
-(2021) found that people who used social media as a primary source of information reported on average ``significantly more unspecific anxiety and depression {[}{]} and significantly more specific COVID-19 related anxiety symptoms'' (p.~288).
-Eden, Johnson, Reinecke, and Grady (2020) analyzed the media use of 425 US college students during the first wave of the pandemic, finding both positive and negative relations with well-being.
-In a sample of 312 respondents collected via Amazon Mechanical Turk, Choi and Choung (2021) reported that people who used media to attain information were more lonely and less satisfied with their lives.
-Stainback, Hearne, and Trieu (2020) analyzed a large-scale study with 11,537 respondents from the US and found that increased COVID-19 media consumption was related to more psychological distress.
-A four-wave panel study with 384 young adults from the U.S.
-analyzed the effects of general digital technology use---objectively measured via screenshots of screen-time applications---on mental health, separating within- and between-person relations (Sewall, Goldstein, \& Rosen, 2021).
-The results showed that digital technology did not have any significant effects on mental health (for a similar study with comparable results, see Bradley \& Howard, 2021).
-Together, the literature is mixed, with a slight focus on the negative effects of social media as news use (see also Dörnemann et al., 2021; Liu \& Tong, 2020; Riehm et al., 2020).
-However, note that all of these findings represent between-person relations stemming from cross-sectional data (see below).
-We therefore don't know whether the differences in mental health and well-being are due to social media use or due to other third variables, such as age, health, employment, or education.
-
-The question of whether and how social media use affects well-being \emph{in general}, on the other hand, is well-researched.
-This also holds true for the different types of communication such as active or passive use.
-A meta review (i.e., an analysis of meta-analyses) found that the relation between social media use and well-being is likely in the negative spectrum but very small (Meier \& Reinecke, 2020)---potentially too small to matter.
-What determines whether or not an effect is considered small or trivial?
-As a starting point, we could refer to standardized effect sizes.
-According to Cohen (1992), small effect sizes start at \emph{r} = .10.
-And indeed, several if not most of the current meta-analyses find effect sizes below that threshold (Ferguson et al., 2021; Huang, 2017; Meier \& Reinecke, 2020).
-
-Several individual studies employing advanced methods found smalls relations between social media use and well-being (Keresteš \& Štulhofer, 2020; Orben, Dienlin, \& Przybylski, 2019; Przybylski, Nguyen, Law, \& Weinstein, 2021; Schemer, Masur, Geiß, Müller, \& Schäfer, 2021).
-For example, Beyens, Pouwels, van Driel, Keijsers, and Valkenburg (2021) reported that although for some users (roughly one quarter) the effects of social media use on well-being were negative, for almost the same number of users they were positive, while for the rest the effects were neutral.
-This finding is aligned the Differential Susceptibility to Media Effects Model.
-Although there is substantial variation of media effects for individual users, the average effects reported in the literature are often small (Valkenburg \& Peter, 2013).
-
-In conclusion, in light of the theoretical considerations and the empirical studies presented above, I expect that COVID-19 related communication on social media doesn't affect well-being in a meaningful or relevant way.
-
-\begin{quote}
-Hypothesis: The within-person effects of all types of COVID-19 related social media use on all types of well-being indicators---while controlling for several stable and varying covariates such as sociodemographic variables and psychological dispositions---will be trivial.
-\end{quote}
-
-\hypertarget{current-study}{%
-\section{Current Study}\label{current-study}}
-
-\hypertarget{smallest-effect-size-of-interest}{%
-\subsection{Smallest Effect Size of Interest}\label{smallest-effect-size-of-interest}}
-
-Testing this hypothesis, however, is not trivial.
-First, in contrast to most hypotheses typically posited in the social sciences it implicitly contains an effect size, a so-called smallest effect size of interest (SESOI).
-Effectively testing this hypothesis necessitates defining what's considered a ``trivial effect size'' and what's not.
-Above I already referred to standardized effect sizes.
-However, standardized effect sizes should only be a first step toward evaluating an effect's relevance (Baguley, 2009).
-Standardized effect sizes are determined by a sample's variance,\footnote{Consider the effect size Cohen's \emph{d}: The mean's of the two groups that are to be compared are subtracted from one another and then divided by the sample's standard deviation (Cohen, 1992).
-Hence, if there is more deviation/variance in a sample, the effect size decreases, even if the difference of the group's means stays the same.} which is problematic:
-The question of whether or not social media use affects a particular person in a relevant way should not depend on the variance in the sample in which that person's data were collected.
-Instead, it should depend on absolute criteria.
-
-What could be a minimally interesting, nontrivial effect?
-Because this is a normative and ultimately philosophical question, there can never be a clear, single, or unanimous answer.
-However, it is still necessary and helpful to try provide such a plausible benchmark.
-I therefore suggest the following SESOI for this research question:
-
-\begin{quote}
-SESOI: If a heavy user of COVID-19 related social media news suddenly \emph{stops} using social media altogether, this should have a \emph{noticeable} impact on their overall well-being.
-\end{quote}
-
-What does this mean practically and how can it be operationalized?
-In this study, COVID-19 related social media use was measured on a 5-point scale, ranging from 1 = \emph{never} to 5 = \emph{several times a day}.
-Thus, a change of four units in social media use (e.g., a complete stop) should correspond to a noticeable change in well-being.
-But what's a noticeable change in well-being?
-According to Norman, Sloan, and Wyrwich (2003), people can reliably distinguish \emph{seven} levels of satisfaction with health.
-So if satisfaction is measured on a 7-point scale, we would state that a four unit change in social media use should result in a one unit change in life satisfaction.
-(For more information, see Methods section ``Inference Criteria.'')
-
-\hypertarget{causality}{%
-\subsection{Causality}\label{causality}}
-
-The hypothesis explicitly states a causal effect.
-In non-experimental studies, longitudinal designs can help investigate causality.
-Using longitudinal designs alone, however, is not sufficient for establishing correct causal statements (Rohrer \& Murayama, 2021).
-In addition, we for example also need to control for confounding third variables, and importantly also for \emph{varying} third variables.
-
-To illustrate, consider the following example.
-Imagine that a person suddenly starts using social media much more than usual, and then after some time becomes less satisfied with their life.
-Eventually, use and life satisfaction return to prior levels.
-If this happens to several people at the same time, in a longitudinal study we could then observe a significant effect of social media use on life satisfaction.
-However, it could also be the case that during the study there was a major exogenous event (say, a pandemic), which caused large parts of the working population to loose their jobs.
-Hence, the causal effect reported above was confounded, because in reality it was the pandemic that caused both social media use to rise and life satisfaction to go down.
-
-Thus, only when controlling for \emph{all} relevant confounders, we can correctly estimate causality without bias (Rohrer, 2018).
-Obviously, we can never be entirely sure to have included all confounders, which makes absolute statements regarding causality virtually impossible.
-In addition, when determining the overall causal effect, we need to make sure \emph{not} to control for mediating variables (Rohrer, 2018), for doing so would bias our assessment of the causal effect.
-Complicating matters further, it is often unclear if a variable is a mediator or a confounder.\footnote{In addition, there also exist colliders, which I don't discuss here and which complicate the issue even further (Rohrer, 2018).}
-However, despite all these caveats, when controlling for relevant variables (that aren't mediators), we can be much more certain that we measured causality correctly.
-The aim should therefore be to collect as many varying and nonvarying confounders as possible (which I believe is seldom done in our field), while knowing that absolute certainty regarding causality cannot be reached.
-
-When searching for suitable candidates for confounders, we should look for variables that affect both media use and well-being.
-Controlling for these factors isolates the actual effect of social media use on well-being.
-We can also control for variables that affect only social media use or well-being.
-However, in doing so not much is gained or lost, because the effects of social media use would remain virtually the same (Kline, 2016; but see McElreath, 2021).
-
-In this study, I hence plan to control for the following variables, which either have already been shown to affect both social media use and well-being or which are likely to do so, and which also aren't mediators:
-gender, age, education, Austria country of birth, Austria country of birth of parents, text-based news consumption, video-based news consumption, residency Vienna, household size, health, living space, access to garden, access to balcony, employment, work hours per week, being in home-office, household income, outdoor activities, satisfaction with democracy, disposition to take risks, and locus of control.
-I will not control for variables such as trust in institutions or trust in media, because these variables might be influenced by social media use to a meaningful extent.
-
-Next to including covariates, it's now increasingly understood that causal effects should be analyzed from an internal, within-person perspective (Hamaker, 2014).
-If a specific person changes their media diet, we need to measure how this behavior affects \emph{their own} well-being.
-Between-person comparisons from cross-sectional data, where participants are interviewed only once, cannot provide such insights.
-In this study, I hence differentiate between-person relations from within-person effects.
-And as explicated above, to test the hypothesis I thus consider only the within-person effects.
-
-Finally, one precondition of causality is temporal order.
-The cause needs to precede the effect.
-Finding the right interval between cause and effect is crucial.
-For example, if we want to understand the effect of alcohol consumption on driving performance, it makes a big difference if driving performance is measured one minute, one hour, one day, or one week after consumption.
-If variables are stable, longer intervals are needed; if they fluctuate, shorter intervals.
-In the case of well-being, we need shorter intervals for affect and longer ones for life satisfaction.
-Still, choosing the right interval is challenging, because especially short intervals are hard to implement in practice and often require advanced methods such as experience sampling (also known as in situ measurement or ambulant assessment) (Schnauber-Stockmann \& Karnowski, 2020).
-
-In this study, I therefore adopt an intermediate perspective.
-I analyze if when a person changes their social media diet, are there \emph{simultaneous} changes in their well-being?
-When additionally controlling for both stable and varying confounders, we can then be more sure that the effect is indeed causal.
-
-\hypertarget{method}{%
-\section{Method}\label{method}}
-
-In this section I describe the preregistration and how I determined the sample size, data exclusions, the analyses, and all measures in the study.
-
-\hypertarget{preregistration}{%
-\subsection{Preregistration}\label{preregistration}}
-
-The hypotheses, the sample, the measures, the analyses, and the inference criteria (SESOI, p-value) were preregistered on the Open Science Framework.
-The (anonymous) preregistration can be accessed here: \url{https://osf.io/87b24/?view_only=b2289b6fec214fa88ee75a18d45c18f3}.
-Because in this study I analyzed data from an already existing large-scale data set, all of these steps were done prior to accessing the data.
-The preregistration was designed on the basis of the panel documentation online (Kittel et al., 2020).
-In some cases I couldn't execute the analyses as I had originally planned, for example because some properties of the variables only became apparent when inspecting the actual data.
-The most relevant deviations are reported below, and a complete list of all changes can be found in the online \href{https://XMtRA.github.io/Austrian_Corona_Panel_Project}{companion website} (\url{https://XMtRA.github.io/Austrian_Corona_Panel_Project}).
-
-\hypertarget{sample}{%
-\subsection{Sample}\label{sample}}
-
-The data come from the Austrian Corona Panel Project (Kittel et al., 2021).
-The study was conducted between March 2020 and October 2021.
-It contains 26 waves, and at the time of writing the first 24 waves were available for download.
-Each wave consists of at least 1,500 respondents.
-The overall sample size was \emph{N} = 3,018, and 72,432 observations were collected.
-Panel mortality was compensated through a continuous acquisition of new participants.
-All respondents needed to have access to the internet (via computer or mobile devices such as smartphones or tablets).
-They were sampled from a pre-existing online access panel provided by the company Marketagent, Austria.
-Respondents were asked and incentivized with 180 credit points to participate in each wave of the panel.
-
-Achieved via quota sampling, the sample matched the Austrian population in terms of age, gender, region/state, municipality size, and educational level.
-In order to participate in the study, the respondents needed to be Austrian residents and had to be at least 14 years of age.
-Ethical review and approval was not required for the study in accordance with the local legislation and institutional requirements.
-The participants provided their written informed consent to participate in this study.
-The average age was 42 years, 49 percent were male, 14 percent had a University degree, and 5 percent were currently unemployed.
-
-\hypertarget{inference-criteria}{%
-\subsection{Inference Criteria}\label{inference-criteria}}
-
-Because the data were analyzed post-hoc, no a-priori sample size planning on the basis of power analyses was conducted.
-The sample is large, and it is hence well-equipped reliably to detect also small effects.
-In addition, because such large samples easily generate significant \emph{p}-values even for very small effects, it helps that the hypotheses were tested with a smallest effect size of interest-approach.
-To this end, I adopted the interval testing approach as proposed by Dienes (2014).
-On the basis of the SESOI, I then defined a null region.
-In what follows, I explain how I determined the SESOI and the null region.
-
-In this study, life satisfaction was measured on an 11-point scale.
-If people can reliably differentiate 7 levels as mentioned above, this corresponds to 11 / 7 = 1.57 unit change on an 11-point scale.
-Hence, a four-point change in media use (e.g., a complete stop) should result in a 1.57-point change in life satisfaction.
-In a statistical regression analysis, \emph{b} estimates the change in the dependent variable if the independent variable increases by one point.
-We would therefore expect a SESOI of \emph{b} = 1.57 / 4 = 0.39.
-For affect, which was measured on a 5-point scale, our SESOI would be \emph{b} = 0.71 / 4 = 0.18.
-Because we're agnostic as to whether the effects are positive or negative, the null region includes negative and positive effects.
-Finally, in order not to exaggerate precision and to be less conservative, these numbers are reduced to nearby thresholds.\footnote{Note that other researchers also decreased or recommended decreasing thresholds for effect sizes when analyzing within-person or cumulative effects (Beyens et al., 2021; Funder \& Ozer, 2019).}
-Together, this leads to a null region ranging from \emph{b} = -.30 to \emph{b} = .30 for life satisfaction, and \emph{b} = -.15 to \emph{b} = .15 for positive and negative affect.
-
-Let's briefly illustrate what this means in practice.
-If the 95\% confidence interval falls completely within the null-region (e.g., \emph{b} = .20, {[}95\% CI: .15, .25{]}), the hypothesis that the effect is trivial is supported.
-If the confidence interval and the null region overlap (e.g., \emph{b} = .30, {[}95\% CI: .25, .35{]}), the hypothesis is not supported and the results are considered inconclusive, while a meaningful negative effect is rejected.
-If the confidence interval falls completely outside of the null-region (e.g., \emph{b} = .40, {[}95\% CI: .35, .45{]}), the hypothesis is rejected and the existence of a meaningful positive effect is supported.
-For an illustration, see Figure \ref{fig:sesoi}).
-
-\hypertarget{data-analysis}{%
-\subsection{Data Analysis}\label{data-analysis}}
-
-The hypothesis was analyzed using mixed effects models, namely random effect within-between models (REWB)(Bell, Fairbrother, \& Jones, 2019).
-Three models were run, one for each dependent variable.
-The data were hierarchical, and responses were separately nested in participants and waves (i.e., participants and waves were implemented as random effects).
-Nesting in participants allowed to separate between-person relations from within-person effects.
-Nesting in waves allowed to control for general exogenous developments, such as general decreases in well-being in the population, for example due to lockdown measures.
-Thus, there was no need additionally to control for specific phases or measures of the lockdown.
-Predictors were modeled as fixed effects.
-They included social media communication types and channels, separated into within and between-person factors, as well as stable and varying covariates.
-All predictors were included simultaneously and in each of the three models.
-
-The factorial validity of the scales were tested with confirmatory factor analyses (CFA).
-Because Mardia's test showed that the assumption of multivariate normality was violated, I used the more robust Satorra-Bentler scaled and mean-adjusted test statistic (MLM) as estimator.
-To avoid overfitting, I tested the scales on more liberal fit criteria (CFI \textgreater{} .90, TLI \textgreater{} .90, RMSEA \textless. .10, SRMR \textless{} .10) (Kline, 2016).
-Finally, REWB-models cannot model latent variables.
-To increase precision, I therefore exported factor scores from the CFAs for positive and negative affect.
-Respondents who answered less than 50\% of all questions were removed.
-The remaining missing responses were imputed using predictive mean matching.
-
-For more information on the analyses, a complete documentation of the models and results, see \href{https://XMtRA.github.io/Austrian_Corona_Panel_Project}{companion website}.
-
-\hypertarget{measures}{%
-\subsection{Measures}\label{measures}}
-
-In what follows, I list all the variables that I analyzed.
-For the variables' means, range, and variance, see Table \ref{tab:tab-descriptives}.
-For a complete list of all items and item characteristics, see \href{https://XMtRA.github.io/Austrian_Corona_Panel_Project}{companion website}.
-
-\hypertarget{well-being}{%
-\subsubsection{Well-being}\label{well-being}}
-
-Life satisfaction was measured with the item ``All things considered, how satisfied are you with your life as a whole nowadays?'' from the European Social Survey (European Social Survey, 2021).
-The response options ranged from 0 (\emph{extremely dissatisfied}) to 10 (\emph{extremely satisfied}).
-
-To capture positive affect, respondents were asked how often in the last week they felt (a) calm and relaxed, (b) happy, and (c) full of energy (World Health Organization, 1998).
-The response options were 1 (\emph{never}), 2 (\emph{on some days}), 3 (\emph{several times per week}), 4 (\emph{almost every day}), and 5 (\emph{daily}).
-The scale showed good factorial fit, \(\chi^2\)(46) = 65.30, \emph{p} = .032, CFI = 1.00, RMSEA = .02, 90\% CI {[}.01, .03{]}, SRMR = .01.
-Reliability was high, \(\omega\) = .85.
-
-For negative affect, respondents were asked how often in the last week they felt (a) lonely, (b) aggravated, (c) so depressed, that nothing could lift you up, (d) very nervous, (e) anxious, and (h) glum and sad (World Health Organization, 1998).
-The response options were 1 (\emph{never}), 2 (\emph{on some days}), 3 (\emph{several times per week}), 4 (\emph{almost every day}), and 5 (\emph{daily}).
-The scale showed good factorial fit, \(\chi^2\)(331) = 3138.37, \emph{p} \textless{} .001, CFI = .97, RMSEA = .08, 90\% CI {[}.07, .08{]}, SRMR = .03.
-Reliability was high, \(\omega\) = .89.
-
-All three variables were measured on each wave.
-
-\hypertarget{covid-19-related-social-media-use}{%
-\subsubsection{COVID-19 related social media use}\label{covid-19-related-social-media-use}}
-
-COVID-19 related social media use focused on communication types was measured with the three dimensions of (a) reading, (b) liking and sharing, and (c) posting from Wagner et al.
-(2018), adapted for the context of this study.
-The general introductory question was ``How often during the last week have you engaged in the following activities on social media?''
-The three items were ``Reading the posts of others with content on the Coronavirus,'' ``When seeing posts on the Coronavirus, I clicked `like,' `share' or `retweet','' ``I myself wrote posts on the Coronavirus on social media.''
-Answer options were 1 (\emph{several times per day}), 2 (\emph{daily}), 3 (\emph{several times per week}), 4 (\emph{weekly}), 5 (\emph{never}).
-The items were inverted for the analyses.
-
-COVID-19 related social media use focused on channels was measured with five variables from Wagner et al. (2018), adapted for this study.
-The general introductory question was ``How often in the last week have you followed information related to the Corona-crisis on the following social media?''
-The five items were (a) Facebook, (b) Twitter, (c) Instagram, (d) Youtube, and (e) WhatsApp.
-Again, the answer options were 1 (\emph{several times per day}), 2 (\emph{daily}), 3 (\emph{several times per week}), 4 (\emph{weekly}), 5 (\emph{never}).
-Again, the items were inverted for the analyses.
-
-Social media use was measured for all participants on waves 1, 2, 8, 17, and 23.
-Freshly recruited respondents always answered all questions on social media use.
-
-\hypertarget{control-variables}{%
-\subsubsection{Control variables}\label{control-variables}}
-
-The effects of COVID-19 related social media use were controlled for the following stable variables:
-(a) gender (female, male, diverse), (b) age, (c) education (ten options), (d) Austria country of birth (yes/no), (e) Austria parents' country of birth (no parent, one parent, both parents).
-I originally planned to implement additional variables as varying covariates.
-However, because they were not measured often enough or not at the time when social media use was measured, I implemented them as stable variables using their average values across all waves.
-This includes (a) text-based media news consumption (five degrees), (b) video-based media news consumption (five degrees), (c) residency is Vienna (yes/no), (d) self-reported physical health (five degrees), (e) living space (eleven options), (f) access to balcony (yes/no), (g) access to garden (yes/no), (h) employment (nine options), (i) disposition to take risks (eleven degrees), and (j) locus of control (five degrees).
-I also controlled for the following varying covariates: (a) five items measuring outdoor activities such as sport or meeting friends (five degrees), and (b) satisfaction with democracy (five degrees).
-Because it lead to too much attrition in the sample, I did not control for (a) household size, (b) work hours per week, (c) home office, (d) household income.
-
-\hypertarget{results}{%
-\section{Results}\label{results}}
-
-First, when looking at the variables from a descriptive perspective (Figure \ref{fig:descriptives})), we see that all well-being measures did not change substantially across the different waves of data collection.
-COVID-19 related media use, however, decreased slightly at the beginning of the study and remained stable after approximately six waves.
-The initial decrease might be explained by the fact that the collection of data began at the end of March 2020, hence approximately three months after the pandemic began.
-It could be that after an initial uptick, COVID-19 related social media use was already declining at the time, returning to more normal levels.
-
-\hypertarget{preregistered-analyses}{%
-\subsection{Preregistered Analyses}\label{preregistered-analyses}}
-
-The study's main hypothesis was that the effects of social media use on well-being would be trivial.
-Regarding the effects of different communication \emph{types}---that is, reading vs.~sharing vs.~posting---all within-person effects fell completely within the a-priori defined null region (see Figure \ref{fig:res-activity}).
-For example, respondents who used social media more frequently than usual to read about COVID-19 related topics did not show a simultaneous change in life satisfaction (\emph{b} = 0.05 {[}95\% CI -0.01, 0.1{]}).
-All confidence intervals included zero; hence, all effects were also statistically non-significant.
-As a result, the hypothesis was supported for all COVID-19 related types of social media communication.
-
-Regarding between-person relations, about which no hypotheses were formulated, only three effects did not include zero.
-Respondents who across all waves used social media more frequently than others to read about COVID-19 related posts reported slightly lower levels of positive affect than others (\emph{b} = -0.03 {[}95\% CI \textgreater{} -0.01, -0.06{]}).
-Respondents who across all waves used social media more frequently than others to write COVID-19 related posts reported higher levels of negative affect than others (\emph{b} = 0.06 {[}95\% CI 0.09, 0.03{]}).
-Interestingly, respondents who across all waves used social media more frequently than others to write COVID-19 related posts also reported slightly higher levels of positive affect than others (\emph{b} = 0.05 {[}95\% CI 0.09, \textless{} 0.01{]}).
-However, note that the effect were still completely inside of the null region, hence not large enough to be considered practically relevant.
-
-Note that when comparing the results with and without control variables, the results differed.
-For example, on the between-person level, one effect stopped being significant if controlled for additional variables.
-Actively posting on social media was significantly (though not meaningfully) related to decreased life satisfaction.
-However, when controlling for potential confounders, the effect became virtually zero (see Figure \ref{fig:res-channels}).
-
-Regarding the COVID-19 related use of social media \emph{channels}, the results were comparable (see Figure \ref{fig:res-channels}).
-Changes in the frequency of using different social media channels to attain information regarding COVID-19 were unrelated to meaningful changes in well-being.
-For example, respondents who used Facebook more frequently than usual to learn about COVID-19 did not show a simultaneous change in well-being (\emph{b} = -0.05 {[}95\% CI -0.11, 0.01{]}).
-Only two effects differed substantially from zero.
-Respondents who used Instagram more frequently than usual to attain COVID-19 related news reported slightly \emph{higher} levels of life satisfaction than usual (\emph{b} = 0.09 {[}95\% CI 0.02, 0.16{]}).
-Respondents who used Twitter more frequently than usual to attain COVID-19 related news reported slightly \emph{lower} levels of life satisfaction than usual (\emph{b} = -0.12 {[}95\% CI -0.23, -0.02{]}).
-However, both effects were still completely inside of the null region, hence not large enough to be considered meaningful.
-In sum, the hypothesis was supported also for the COVID-19 related use of important social media channels.
-
-In terms of between-person relations---which, again, weren't included in the hypotheses---no relations crossed the null region or fell outside of it.
-Only one relation did not include zero, was hence statistically significant.
-Respondents who across all waves used YouTube more frequently than others for COVID-19 related reasons reported marginally \emph{higher} levels of life satisfaction (\emph{b} = 0.08 {[}95\% CI \textless{} 0.01, 0.16{]}).
-However, please note that this effect again was not large enough to be considered practically relevant.
-
-Again, note that when comparing the results with and without control variables, the results differed.
-Especially on the between-person level, altogether five effects stopped being significant if they were controlled for additional variables.
-For example, using Instagram was significantly (though not meaningfully) related to increased life satisfaction.
-However, when controlling for additional covariates, the effect became virtually zero (see Figure \ref{fig:res-channels}).
-
-\hypertarget{exploratory-analyses}{%
-\subsection{Exploratory Analyses}\label{exploratory-analyses}}
-
-In what follows, I briefly report some exploratory analyses that weren't preregistered.
-First, to contextualize the results reported above and to see if the results included any meaningful effects at all, I also looked at the effect sizes of selected (cherry-picked) covariates.
-Because each variable had different response options, we would actually need to define a SESOI for each variable, which for reasons of scope I cannot implement here.
-Therefore, I report the results of the standardized scales,
-which allows for a better comparison across the differently scaled variables.
-For what it's worth, as a rough estimate for the SESOI we can build on the typical convention that small effects start at \emph{r} = \textbar.10\textbar.
-The results showed that several effects fell outside of the SESOI, were hence considered meaningful.
-This includes for example internal locus of control, health, satisfaction with democracy, or exercising.
-For an overview, see Figure \ref{fig:res-control}.
-
-To find out whether my inferences were robust across legitimate (though arguably inferior) alternative analyses, I reran the analyses also using standardized estimates, mean scores instead of factor scores, and with a data set where missing data were not imputed.
-The results were virtually the same.
-For example, all standardized COVID-19 related types of social media use or channels were not significantly larger than a SESOI of \(\beta\) = \textbar.10\textbar.
-The additional analyses are reported in the \href{https://XMtRA.github.io/Austrian_Corona_Panel_Project/analyses_additional.html}{companion website}.
-
-\hypertarget{discussion}{%
-\section{Discussion}\label{discussion}}
-
-In this study I analyzed the effects of COVID-19 related social media use on well-being.
-The data come from a panel study with 24 waves and are largely representative of the Austrian population.
-In a random effects model I separated between person relations from within-person effects and controlled for a large number of both stable and varying covariates, thereby aiming to assess causality.
-The results showed that within-person effects were trivial.
-People who used social media more than usual to learn about COVID-19 didn't show meaningful changes in their well-being.
-
-The results imply that COVID-19 related social media use doesn't seem to be particularly relevant for well-being.
-Other factors among the third variables that were measured revealed larger effects or relations, suggesting that well-being is determined by alternative aspects such as health, satisfaction with democracy, locus of control, or exercising.
-According to this study, popular fears that ``doomscrolling'' or overusing social media during times of crises don't seem to be justified.
-
-On the one hand, the results are not aligned with several recent studies analyzing similar or closely related research questions.
-This includes a study by Bendau et al. (2021), which showed negative relations between social media and well-being (but see Bradley \& Howard, 2021; or Sewall et al., 2021).
-However, note that Bendau et al. (2021) analyzed cross-sectional data on a between-person level while not controlling for third variables, which is not optimal for investigating causal effects.
-On the other hand, the results are well-aligned with recent studies and meta-analyses analyzing the effects of social media use from a more general perspective or from a somewhat different angle.
-These studies have found that the effects of various types of social media use on several well-being indicators are small at best, often too small to matter (Ferguson et al., 2021; Meier \& Reinecke, 2020; Orben, 2020), which echoes the results obtained here.
-
-If anything, two preliminary and subtle trends can be observed.
-First, of all the three COVID-19 related social media activities, people who read about the pandemic more than others showed slightly decreased levels of positive affect, and people who actively posted about the pandemic more than others showed slightly increased levels of negative affect.
-On the other hand, however, people who posted more about COVID-19 also showed slightly higher levels of positive affects, so taken together the results are ambivalent.
-Second, in terms of media channels, using Twitter more than usual was related to slightly decreased levels of life satisfaction.
-Twitter is considered to have more negative affordances and tonality as compared to other networks such as Instagram (Waterloo et al., 2018), which might help explain the results.
-Instagram, on the other hand, was related to slightly increased levels of life satisfaction.
-To speculate, the often-criticized positivity bias on Instagram might have been somewhat beneficial in times of the pandemic.
-That said, all these effects were still very small and arguably too small to matter.
-But future research might elaborate on these specific relations to probe their stability and relevance.
-
-Finally, another interesting observation is that life satisfaction was remarkably stable.
-Hence, even in times of a pandemic, it seems that such broad assessment of life vary only mildly.
-This supports the hypothesis that life satisfaction seems to be determined largely by stable factors such as one's genes (Brown \& Rohrer, 2019).
-
-\hypertarget{limitations}{%
-\subsection{Limitations}\label{limitations}}
-
-The current study analyzed whether changes in media use were related to changes in well-being, while controlling for several potential confounders.
-Together, this allows for an improved perspective on assessing causality.
-That said, causality necessitates temporal order, and the cause needs to precede the effect.
-Regarding media use, such effects often happen immediately or shortly after use, necessitating intervals in the hours, minutes, or even seconds.
-In many cases only experience sampling studies asking users in the very moment can produce such knowledge.
-However, even then we don't know for certain if we actually measured the right interval.
-Effects depend on the intensity of use or the length of the interval, and to borrow the words from Rohrer and Murayama (2021), there is no such thing as ``the'' effect of social media use on well-being.
-Hence, to document how effects unfold, future research needs to employ different study designs probing different time lags.
-In addition, more thought needs to be invested in what relevant stable and varying factors we should include as control variables, and I hope this study provides a first step into this direction.
-
-Although I had already reduced the predefined SESOIs to be less conservative, they were potentially still too large.
-Media use is only one aspect of several factors that simultaneously affect well-being.
-Is it really realistic to expect that extremely changing only \emph{one} of these aspects should manifest in a detectable change in well-being?
-Or would it make more sense to expect that thoroughly committing to say \emph{two} activities (e.g.~regularly exercising \emph{and} establishing a reading habit) should then cause a detectable improvement in well-being?
-Practically, this would imply a SESOI half as large as I have defined here, namely \emph{b} = \textbar.15\textbar{} for well-being and \emph{b} = \textbar.075\textbar{} for affect.
-In the case of this study, however, reducing the SESOI would not even make a big difference, as also with these more liberal thresholds all but three effect would still be completely in the null region, and no effect would be outside of the null region.
-However, at all events I encourage future research to start a thorough conversation on what effect sizes are considered meaningful and what not.
-Again, with this study I hope to provide some first input and guidelines.
-
-Both media use and well-being were measured using self-reports.
-Measuring well-being with self-reports is adequate, because it by definition requires introspection.
-However, it would be preferable to measure social media use objectively, because people cannot reliably estimate their use (Scharkow, 2016).
-That said, objective measures often cannot capture the content or the motivation of the use, and only very complicated tools recording the actual content (such as the Screenome project) might produce such data.
-Unfortunately, such procedures introduce other problems, especially related to privacy.
-Hence, for this type of research question it still seems necessary to use self-reported measures.
-
-Because the data were collected in a single country, the generalizability of the results is limited.
-The results apply primarily to the more Western sphere, and might not hold true in other cultures, especially cultures with a different media landscape or alternative social media channels.
-That said, because this is a comparatively large study largely representative of an entire country, and because several waves were collected across a large time span, the results should be at least as generalizable as other typical empirical studies collected in the social sciences.
-
-\hypertarget{conclusion}{%
-\subsection{Conclusion}\label{conclusion}}
-
-In this study, COVID-19 related social media use didn't meaningfully affect several indicators of well-being, including life satisfaction, positive affect, and negative affect.
-However, factors other than social media use were meaningfully related to well-being, such as physical health, exercise, satisfaction with democracy, or believing that one is in control of one's life.
-If it's our aim to improve well-being, it might hence be more fruitful not to focus so much on social media but to address other, more pertinent societal problems related to health care, regular exercise, or a functioning democratic system.
-
-\newpage
-
-\hypertarget{references}{%
-\section{References}\label{references}}
-
-\hypertarget{refs}{}
-\begin{CSLReferences}{1}{0}
-\leavevmode\vadjust pre{\hypertarget{ref-baguleyStandardizedSimpleEffect2009}{}}%
-Baguley, T. (2009). Standardized or simple effect size: {What} should be reported? \emph{British Journal of Psychology}, \emph{100}(3), 603--617. \url{https://doi.org/10.1348/000712608X377117}
-
-\leavevmode\vadjust pre{\hypertarget{ref-bellFixedRandomEffects2019}{}}%
-Bell, A., Fairbrother, M., \& Jones, K.
-(2019).
-Fixed and random effects models: Making an informed choice.
-\emph{Quality \& Quantity}, \emph{53}(2), 1051--1074.
-\url{https://doi.org/10.1007/s11135-018-0802-x}
-
-\leavevmode\vadjust pre{\hypertarget{ref-bendauAssociationsCOVID19Related2021}{}}%
-Bendau, A., Petzold, M.
-B., Pyrkosch, L., Mascarell Maricic, L., Betzler, F., Rogoll, J., \ldots{} Plag, J.
-(2021).
-Associations between {COVID-19} related media consumption and symptoms of anxiety, depression and {COVID-19} related fear in the general population in {Germany}.
-\emph{European Archives of Psychiatry and Clinical Neuroscience}, \emph{271}(2), 283--291.
-\url{https://doi.org/10.1007/s00406-020-01171-6}
-
-\leavevmode\vadjust pre{\hypertarget{ref-beyensSocialMediaUse2021}{}}%
-Beyens, I., Pouwels, J.
-L., van Driel, I.
-I., Keijsers, L., \& Valkenburg, P.
-M.
-(2021).
-Social media use and adolescents' well-being: {Developing} a typology of person-specific effect patterns.
-\emph{Communication Research}.
-\url{https://doi.org/10.1177/00936502211038196}
-
-\leavevmode\vadjust pre{\hypertarget{ref-bradleyStressMoodSmartphone2021}{}}%
-Bradley, A., \& Howard, A. (2021). Stress, mood, and smartphone use in {University} students: {A} 12-week longitudinal study. \url{https://doi.org/10.31219/osf.io/frvpb}
-
-\leavevmode\vadjust pre{\hypertarget{ref-brownEasyHappinessPie2019}{}}%
-Brown, N.
-J.
-L., \& Rohrer, J.
-M.
-(2019).
-Easy as (happiness) pie?
-{A} critical evaluation of a popular model of the determinants of well-being.
-\emph{Journal of Happiness Studies}.
-\url{https://doi.org/10.1007/s10902-019-00128-4}
-
-\leavevmode\vadjust pre{\hypertarget{ref-buchiDigitalWellbeingTheory2021}{}}%
-Büchi, M. (2021). Digital well-being theory and research. \emph{New Media \& Society}, 146144482110568. \url{https://doi.org/10.1177/14614448211056851}
-
-\leavevmode\vadjust pre{\hypertarget{ref-choiMediatedCommunicationMatters2021}{}}%
-Choi, M., \& Choung, H.
-(2021).
-Mediated communication matters during the {COVID-19} pandemic: {The} use of interpersonal and masspersonal media and psychological well-being.
-\emph{Journal of Social and Personal Relationships}, \emph{38}(8), 2397--2418.
-\url{https://doi.org/10.1177/02654075211029378}
-
-\leavevmode\vadjust pre{\hypertarget{ref-cohenPowerPrimer1992}{}}%
-Cohen, J. (1992). A power primer. \emph{Psychological Bulletin}, \emph{112}(1), 155--159. \url{https://doi.org/10.1037/0033-2909.112.1.155}
-
-\leavevmode\vadjust pre{\hypertarget{ref-dienerAdvancesOpenQuestions2018}{}}%
-Diener, E., Lucas, R.
-E., \& Oishi, S.
-(2018).
-Advances and open questions in the science of subjective well-being.
-\emph{Collabra: Psychology}, \emph{4}(1), 15.
-\url{https://doi.org/10.1525/collabra.115}
-
-\leavevmode\vadjust pre{\hypertarget{ref-dienesUsingBayesGet2014}{}}%
-Dienes, Z. (2014). Using {Bayes} to get the most out of non-significant results. \emph{Frontiers in Psychology}, \emph{5}. \url{https://doi.org/10.3389/fpsyg.2014.00781}
-
-\leavevmode\vadjust pre{\hypertarget{ref-dienlinImpactDigitalTechnology2020}{}}%
-Dienlin, T., \& Johannes, N.
-(2020).
-The impact of digital technology use on adolescent well-being.
-\emph{Dialogues in Clinical Neuroscience}, \emph{22}(2), 135--142.
-\url{https://doi.org/doi:10.31887/DCNS.2020.22.2/tdienlin}
-
-\leavevmode\vadjust pre{\hypertarget{ref-dornemannHowGoodBad2021}{}}%
-Dörnemann, A., Boenisch, N., Schommer, L., Winkelhorst, L., \& Wingen, T.
-(2021).
-How do good and bad news impact mood during the {Covid-19} pandemic?
-{The} role of similarity.
-\url{https://doi.org/10.31219/osf.io/sy2kd}
-
-\leavevmode\vadjust pre{\hypertarget{ref-duradoniWellbeingSocialMedia2020}{}}%
-Duradoni, M., Innocenti, F., \& Guazzini, A.
-(2020).
-Well-being and social media: {A} systematic review of {Bergen Addiction Scales}.
-\emph{Future Internet}, \emph{12}(2, 2), 24.
-\url{https://doi.org/10.3390/fi12020024}
-
-\leavevmode\vadjust pre{\hypertarget{ref-edenMediaCopingCOVID192020}{}}%
-Eden, A.
-L., Johnson, B.
-K., Reinecke, L., \& Grady, S.
-M.
-(2020).
-Media for coping during {COVID-19} social distancing: {Stress}, anxiety, and psychological well-being.
-\emph{Frontiers in Psychology}, \emph{11}, 577639.
-\url{https://doi.org/10.3389/fpsyg.2020.577639}
-
-\leavevmode\vadjust pre{\hypertarget{ref-ellisonWhyWeDon2020}{}}%
-Ellison, N.
-B., Triệu, P., Schoenebeck, S., Brewer, R., \& Israni, A.
-(2020).
-Why we don't click: {Interrogating} the relationship between viewing and clicking in social media contexts by exploring the {``{Non-Click}.''} \emph{Journal of Computer-Mediated Communication}, \emph{25}(6), 402--426.
-\url{https://doi.org/10.1093/jcmc/zmaa013}
-
-\leavevmode\vadjust pre{\hypertarget{ref-europeansocialsurveyESS9Edition20182021}{}}%
-European Social Survey. (2021). {ESS9} edition 3.1 - 2018 {Documentation Report}. Retrieved from \url{https://www.europeansocialsurvey.org/docs/round9/survey/ESS9_data_documentation_report_e03_1.pdf}
-
-\leavevmode\vadjust pre{\hypertarget{ref-fergusonThisMetaanalysisScreen2021}{}}%
-Ferguson, C.
-J., Kaye, L.
-K., Branley-Bell, D., Markey, P., Ivory, J.
-D., Klisanin, D., \ldots{} Wilson, J.
-(2021).
-Like this meta-analysis: {Screen} media and mental health.
-\emph{Professional Psychology: Research and Practice}.
-\url{https://doi.org/10.1037/pro0000426}
-
-\leavevmode\vadjust pre{\hypertarget{ref-funderEvaluatingEffectSize2019}{}}%
-Funder, D.
-C., \& Ozer, D.
-J.
-(2019).
-Evaluating effect size in psychological research: {Sense} and nonsense.
-\emph{Advances in Methods and Practices in Psychological Science}, \emph{2}(2), 156--168.
-\url{https://doi.org/10.1177/2515245919847202}
-
-\leavevmode\vadjust pre{\hypertarget{ref-galerHowMuchToo2018}{}}%
-Galer, S.
-S.
-(2018, January 19).
-How much is {`too much time'} on social media?
-Retrieved April 4, 2022, from \url{https://www.bbc.com/future/article/20180118-how-much-is-too-much-time-on-social-media}
-
-\leavevmode\vadjust pre{\hypertarget{ref-greenspoonIntegrationSubjectiveWellbeing2001}{}}%
-Greenspoon, P.
-J., \& Saklofske, D.
-H.
-(2001).
-Toward an integration of subjective well-being and psychopathology.
-\emph{Social Indicators Research}, \emph{54}(1), 81--108.
-\url{https://doi.org/10.1023/A:1007219227883}
-
-\leavevmode\vadjust pre{\hypertarget{ref-guazziniSecondWaveAnalysis2022}{}}%
-Guazzini, A., Pesce, A., Marotta, L., \& Duradoni, M.
-(2022).
-Through the second wave: {Analysis} of the psychological and perceptive changes in the {Italian} population during the {COVID-19} pandemic.
-\emph{International Journal of Environmental Research and Public Health}, \emph{19}(3), 1635.
-\url{https://doi.org/10.3390/ijerph19031635}
-
-\leavevmode\vadjust pre{\hypertarget{ref-hamakerWhyResearchersShould2014}{}}%
-Hamaker, E.
-L.
-(2014).
-Why researchers should think "within-person": {A} paradigmatic rationale.
-In M.
-R.
-Mehl, T.
-S.
-Conner, \& M.
-Csikszentmihalyi (Eds.), \emph{Handbook of research methods for studying daily life} (Paperback ed.).
-{New York, NY}: {Guilford}.
-
-\leavevmode\vadjust pre{\hypertarget{ref-huangTimeSpentSocial2017}{}}%
-Huang, C.
-(2017).
-Time spent on social network sites and psychological well-being: {A} meta-analysis.
-\emph{Cyberpsychology, Behavior and Social Networking}, \emph{20}(6), 346--354.
-\url{https://doi.org/10.1089/cyber.2016.0758}
-
-\leavevmode\vadjust pre{\hypertarget{ref-kerestesAdolescentsOnlineSocial2020}{}}%
-Keresteš, G., \& Štulhofer, A.
-(2020).
-Adolescents' online social network use and life satisfaction: {A} latent growth curve modeling approach.
-\emph{Computers in Human Behavior}, \emph{104}, 106187.
-\url{https://doi.org/10.1016/j.chb.2019.106187}
-
-\leavevmode\vadjust pre{\hypertarget{ref-kittelAustrianCoronaPanel2020}{}}%
-Kittel, B., Kritzinger, S., Boomgaarden, H., Prainsack, B., Eberl, J.-M., Kalleitner, F., \ldots{} Schlogl, L.
-(2020).
-\emph{Austrian {Corona Panel Project} ({SUF} edition)} {[}Data set{]}.
-{AUSSDA}.
-\url{https://doi.org/10.11587/28KQNS}
-
-\leavevmode\vadjust pre{\hypertarget{ref-kittelAustrianCoronaPanel2021}{}}%
-Kittel, B., Kritzinger, S., Boomgaarden, H., Prainsack, B., Eberl, J.-M., Kalleitner, F., \ldots{} Schlogl, L.
-(2021).
-The {Austrian Corona Panel Project}: Monitoring individual and societal dynamics amidst the {COVID-19} crisis.
-\emph{European Political Science}, \emph{20}(2), 318--344.
-\url{https://doi.org/10.1057/s41304-020-00294-7}
-
-\leavevmode\vadjust pre{\hypertarget{ref-kleinDarklySoothingCompulsion2021}{}}%
-Klein, J.
-(2021, March 3).
-The darkly soothing compulsion of 'doomscrolling'.
-Retrieved November 1, 2021, from \url{https://www.bbc.com/worklife/article/20210226-the-darkly-soothing-compulsion-of-doomscrolling}
-
-\leavevmode\vadjust pre{\hypertarget{ref-klinePrinciplesPracticeStructural2016}{}}%
-Kline, R. B. (2016). \emph{Principles and practice of structural equation modeling} (4th ed.). {New York, NY}: {The Guilford Press}.
-
-\leavevmode\vadjust pre{\hypertarget{ref-liuRelationOfficialWhatsAppdistributed2020}{}}%
-Liu, J.
-C.
-J., \& Tong, E.
-M.
-W.
-(2020).
-The relation between official {WhatsApp-distributed COVID-19} news exposure and psychological symptoms: Cross-sectional survey study.
-\emph{Journal of Medical Internet Research}, \emph{22}(9), e22142.
-\url{https://doi.org/10.2196/22142}
-
-\leavevmode\vadjust pre{\hypertarget{ref-lucasAdaptationSetpointModel2007}{}}%
-Lucas, R.
-E.
-(2007).
-Adaptation and the set-point model of subjective well-being.
-\emph{Current Directions in Psychological Science}, \emph{16}(2), 75--79.
-\url{https://doi.org/10.1111/j.1467-8721.2007.00479.x}
-
-\leavevmode\vadjust pre{\hypertarget{ref-lykkenHappinessWhatStudies1999}{}}%
-Lykken, D. T. (1999). \emph{Happiness: {What} studies on twins show us about nature, nurture, and the happiness set-point}. {New York, N.Y.}: {Golden Books}.
-
-\leavevmode\vadjust pre{\hypertarget{ref-marcianoDynamicsAdolescentsSmartphone2022}{}}%
-Marciano, L., Driver, C.
-C., Schulz, P.
-J., \& Camerini, A.-L.
-(2022).
-Dynamics of adolescents' smartphone use and well-being are positive but ephemeral.
-\emph{Scientific Reports}, \emph{12}(1), 1316.
-\url{https://doi.org/10.1038/s41598-022-05291-y}
-
-\leavevmode\vadjust pre{\hypertarget{ref-mcelreathYesterdayClass2021}{}}%
-McElreath, R. (2021, January 28). Yesterday in class, ... {[}Tweet{]}. Retrieved November 5, 2021, from \url{https://twitter.com/rlmcelreath/status/1354786005996482563}
-
-\leavevmode\vadjust pre{\hypertarget{ref-meierDoesPassiveSocial2022}{}}%
-Meier, A., \& Krause, H.-V. (2022). \emph{Does passive social media use harm well-being? {An} adversarial review} (preprint). {PsyArXiv}. \url{https://doi.org/10.31234/osf.io/nvbwh}
-
-\leavevmode\vadjust pre{\hypertarget{ref-meierComputermediatedCommunicationSocial2020a}{}}%
-Meier, A., \& Reinecke, L.
-(2020).
-Computer-mediated communication, social media, and mental health: {A} conceptual and empirical meta-review.
-\emph{Communication Research}, 009365022095822.
-\url{https://doi.org/10.1177/0093650220958224}
-
-\leavevmode\vadjust pre{\hypertarget{ref-normanInterpretationChangesHealthrelated2003}{}}%
-Norman, G., Sloan, J., \& Wyrwich, K.
-(2003).
-Interpretation of changes in health-related quality of life: {The} remarkable universality of half a standard deviation.
-\emph{Medical Care}, \emph{41}(5), 582--592.
-Retrieved from \url{Retrieved\%20from\%20http://www.jstor.org/stable/3768017}
-
-\leavevmode\vadjust pre{\hypertarget{ref-orbenTeenagersScreensSocial2020}{}}%
-Orben, A.
-(2020).
-Teenagers, screens and social media: A narrative review of reviews and key studies.
-\emph{Social Psychiatry and Psychiatric Epidemiology}, \emph{55}(4), 407--414.
-\url{https://doi.org/10.1007/s00127-019-01825-4}
-
-\leavevmode\vadjust pre{\hypertarget{ref-orbenSocialMediaEnduring2019}{}}%
-Orben, A., Dienlin, T., \& Przybylski, A.
-K.
-(2019).
-Social media's enduring effect on adolescent life satisfaction.
-\emph{Proceedings of the National Academy of Sciences of the United States of America}, \emph{116}(21), 10226--10228.
-\url{https://doi.org/10.1073/pnas.1902058116}
-
-\leavevmode\vadjust pre{\hypertarget{ref-pelletierOneSizeDoesn2020}{}}%
-Pelletier, M.
-J., Krallman, A., Adams, F.
-G., \& Hancock, T.
-(2020).
-One size doesn't fit all: A uses and gratifications analysis of social media platforms.
-\emph{Journal of Research in Interactive Marketing}, \emph{14}(2), 269--284.
-\url{https://doi.org/10.1108/JRIM-10-2019-0159}
-
-\leavevmode\vadjust pre{\hypertarget{ref-przybylskiDoesTakingShort2021a}{}}%
-Przybylski, A.
-K., Nguyen, T.
-T., Law, W., \& Weinstein, N.
-(2021).
-Does taking a short break from social media have a positive effect on well-being?
-{Evidence} from three preregistered field experiments.
-\emph{Journal of Technology in Behavioral Science}, \emph{6}(3), 507--514.
-\url{https://doi.org/10.1007/s41347-020-00189-w}
-
-\leavevmode\vadjust pre{\hypertarget{ref-przybylskiLargescaleTestGoldilocks2017}{}}%
-Przybylski, A. K., \& Weinstein, N. (2017). A large-scale test of the {Goldilocks} hypothesis. \emph{Psychological Science}, \emph{28}(2), 204--215. \url{https://doi.org/10.1177/0956797616678438}
-
-\leavevmode\vadjust pre{\hypertarget{ref-riehmAssociationsMediaExposure2020}{}}%
-Riehm, K.
-E., Holingue, C., Kalb, L.
-G., Bennett, D., Kapteyn, A., Jiang, Q., \ldots{} Thrul, J.
-(2020).
-Associations between media exposure and mental distress among {U}.{S}.
-Adults at the beginning of the {COVID-19} pandemic.
-\emph{American Journal of Preventive Medicine}, \emph{59}(5), 630--638.
-\url{https://doi.org/10.1016/j.amepre.2020.06.008}
-
-\leavevmode\vadjust pre{\hypertarget{ref-rohrerThinkingClearlyCorrelations2018}{}}%
-Rohrer, J.
-M.
-(2018).
-Thinking clearly about correlations and causation: {Graphical} causal models for observational data.
-\emph{Advances in Methods and Practices in Psychological Science}, \emph{24}(2), 251524591774562.
-\url{https://doi.org/10.1177/2515245917745629}
-
-\leavevmode\vadjust pre{\hypertarget{ref-rohrerTheseAreNot2021}{}}%
-Rohrer, J.
-M., \& Murayama, K.
-(2021).
-\emph{These are not the effects you are looking for: {Causality} and the within-/between-person distinction in longitudinal data analysis} (preprint).
-{PsyArXiv}.
-\url{https://doi.org/10.31234/osf.io/tg4vj}
-
-\leavevmode\vadjust pre{\hypertarget{ref-sandstromDoomscrollingCOVIDNews2021}{}}%
-Sandstrom, G., Buchanan, K., Aknin, L., \& Lotun, S.
-(2021, October 22).
-Doomscrolling {COVID} news takes an emotional toll -- here's how to make your social media a happier place.
-Retrieved November 3, 2021, from \url{http://theconversation.com/doomscrolling-covid-news-takes-an-emotional-toll-heres-how-to-make-your-social-media-a-happier-place-170342}
-
-\leavevmode\vadjust pre{\hypertarget{ref-scharkowAccuracySelfreportedInternet2016}{}}%
-Scharkow, M.
-(2016).
-The accuracy of self-reported {Internet} use---{A} validation study using client log data.
-\emph{Communication Methods and Measures}, \emph{10}(1), 13--27.
-\url{https://doi.org/10.1080/19312458.2015.1118446}
-
-\leavevmode\vadjust pre{\hypertarget{ref-schemerImpactInternetSocial2021}{}}%
-Schemer, C., Masur, P.
-K., Geiß, S., Müller, P., \& Schäfer, S.
-(2021).
-The impact of {Internet} and social media use on well-being: {A} longitudinal analysis of adolescents across nine years.
-\emph{Journal of Computer-Mediated Communication}, \emph{26}(1), 1--21.
-\url{https://doi.org/10.1093/jcmc/zmaa014}
-
-\leavevmode\vadjust pre{\hypertarget{ref-schnauber-stockmannMobileDevicesTools2020}{}}%
-Schnauber-Stockmann, A., \& Karnowski, V.
-(2020).
-Mobile devices as tools for media and communication research: {A} scoping review on collecting self-report data in repeated measurement designs.
-\emph{Communication Methods and Measures}, \emph{14}(3), 145--164.
-\url{https://doi.org/10.1080/19312458.2020.1784402}
-
-\leavevmode\vadjust pre{\hypertarget{ref-sewallObjectivelyMeasuredDigital2021}{}}%
-Sewall, C.
-J.
-R., Goldstein, T.
-R., \& Rosen, D.
-(2021).
-Objectively measured digital technology use during the {COVID-19} pandemic: {Impact} on depression, anxiety, and suicidal ideation among young adults.
-\emph{Journal of Affective Disorders}, \emph{288}, 145--147.
-\url{https://doi.org/10.1016/j.jad.2021.04.008}
-
-\leavevmode\vadjust pre{\hypertarget{ref-sheldonStabilityHappinessTheories2014}{}}%
-Sheldon, K. M., \& Lucas, R. E. (2014). \emph{Stability of happiness: Theories and evidence on whether happiness can change}. Retrieved from \url{http://site.ebrary.com/id/10891875}
-
-\leavevmode\vadjust pre{\hypertarget{ref-stainbackCOVID1924News2020}{}}%
-Stainback, K., Hearne, B.
-N., \& Trieu, M.
-M.
-(2020).
-{COVID-19} and the 24/7 {News Cycle}: {Does COVID-19 News Exposure Affect Mental Health}?
-\emph{Socius}, \emph{6}, 2378023120969339.
-\url{https://doi.org/10.1177/2378023120969339}
-
-\leavevmode\vadjust pre{\hypertarget{ref-statistaAverageDailyTime2021}{}}%
-Statista.
-(2021, May 21).
-Average daily time spent on social networks by users in the {United States} from 2018 to 2022.
-Retrieved from \url{https://www.statista.com/statistics/1018324/us-users-daily-social-media-minutes/}
-
-\leavevmode\vadjust pre{\hypertarget{ref-valkenburgDifferentialSusceptibilityMedia2013}{}}%
-Valkenburg, P. M., \& Peter, J. (2013). The differential susceptibility to media effects model. \emph{Journal of Communication}, \emph{63}(2), 221--243. \url{https://doi.org/10.1111/jcom.12024}
-
-\leavevmode\vadjust pre{\hypertarget{ref-vanrooijWeakScientificBasis2018}{}}%
-van Rooij, A.
-J., Ferguson, C.
-J., Colder Carras, M., Kardefelt-Winther, D., Shi, J., Aarseth, E., \ldots{} Przybylski, A.
-K.
-(2018).
-A weak scientific basis for gaming disorder: {Let} us err on the side of caution.
-\emph{Journal of Behavioral Addictions}, \emph{7}(1), 1--9.
-\url{https://doi.org/10.1556/2006.7.2018.19}
-
-\leavevmode\vadjust pre{\hypertarget{ref-wagnerAUTNESOnlinePanel2018}{}}%
-Wagner, M., Aichholzer, J., Eberl, J.-M., Meyer, T.
-M., Berk, N., Büttner, N., \ldots{} Müller, W.
-C.
-(2018).
-\emph{{AUTNES Online Panel Study} 2017 ({SUF} edition)} {[}Data set{]}.
-{AUSSDA}.
-\url{https://doi.org/10.11587/I7QIYJ}
-
-\leavevmode\vadjust pre{\hypertarget{ref-waterlooNormsOnlineExpressions2018}{}}%
-Waterloo, S.
-F., Baumgartner, S.
-E., Peter, J., \& Valkenburg, P.
-M.
-(2018).
-Norms of online expressions of emotion: {Comparing Facebook}, {Twitter}, {Instagram}, and {WhatsApp}.
-\emph{New Media \& Society}, \emph{20}(5), 1813--1831.
-\url{https://doi.org/10.1177/1461444817707349}
-
-\leavevmode\vadjust pre{\hypertarget{ref-worldhealthorganizationWellbeingMeasuresPrimary1998}{}}%
-World Health Organization. (1998). Wellbeing measures in primary health care/{The Depcare Project}.
-
-\leavevmode\vadjust pre{\hypertarget{ref-yangCanWatchingOnline2021}{}}%
-Yang, Z., Griffiths, M.
-D., Yan, Z., \& Xu, W.
-(2021).
-Can watching online videos be addictive?
-{A} qualitative exploration of online video watching among {Chinese} young adults.
-\emph{International Journal of Environmental Research and Public Health}, \emph{18}(14, 14), 7247.
-\url{https://doi.org/10.3390/ijerph18147247}
-
-\leavevmode\vadjust pre{\hypertarget{ref-zillmannMoodManagementCommunication1988}{}}%
-Zillmann, D. (1988). Mood {Management Through Communication Choices}. \emph{American Behavioral Scientist}, \emph{31}(3), 327340. \url{https://doi.org/10.1177/000276488031003005}
-
-\end{CSLReferences}
-
-\newpage
-
-\begin{table}[tbp]
-
-\begin{center}
-\begin{threeparttable}
-
-\caption{\label{tab:tab-descriptives}Descriptives of the main variables.}
-
-\begin{tabular}{lllll}
-\toprule
- & \multicolumn{1}{c}{sd} & \multicolumn{1}{c}{min} & \multicolumn{1}{c}{max} & \multicolumn{1}{c}{mean}\\
-\midrule
-Well-being &  &  &  & \\
-\ \ \ Life satisfaction & 1.68 & 6.38 & 6.81 & 6.59\\
-\ \ \ Positive affect & 0.57 & 3.05 & 3.28 & 3.15\\
-\ \ \ Negative affect & 0.39 & 1.66 & 1.81 & 1.73\\
-Social media use &  &  &  & \\
-\ \ \ Read & 1.03 & 2.10 & 2.92 & 2.43\\
-\ \ \ Like \& share & 0.87 & 1.61 & 1.92 & 1.78\\
-\ \ \ Posting & 0.63 & 1.33 & 1.46 & 1.39\\
-Social media channel &  &  &  & \\
-\ \ \ Facebook & 0.96 & 2.34 & 2.68 & 2.45\\
-\ \ \ Twitter & 0.52 & 1.16 & 1.72 & 1.37\\
-\ \ \ Instagram & 0.83 & 1.84 & 2.66 & 2.09\\
-\ \ \ WhatsApp & 1.23 & 2.28 & 2.62 & 2.45\\
-\ \ \ YouTube & 0.88 & 1.77 & 2.30 & 2.00\\
-\bottomrule
-\end{tabular}
-
-\end{threeparttable}
-\end{center}
-
-\end{table}
-
-\newpage
-
-\begin{figure}
-\centering
-\includegraphics{manuscript_files/figure-latex/sesoi-1.pdf}
-\caption{\label{fig:sesoi}Illustration of how confidence intervals are used to test a null region.
-Here, a trivial effect of social media use on life satisfaction is defined as ranging from b = -.25 to b = .25}
-\end{figure}
-
-\newpage
-
-\begin{figure}
-\includegraphics[width=\textwidth]{figures/fig_descriptives} \caption{Development of well-being and media use measures across the pandemic.
-Values obtained from mixed effect models, with participants and waves as grouping factors and without additional predictors.}\label{fig:descriptives}
-\end{figure}
-
-\newpage
-
-\begin{figure}
-\includegraphics[width=\textwidth]{figures/fig_results_activity} \caption{The effects of various types of social media use on three indicators of well-being.
-The black estimates show the effects controlled for a large number of covariates (see text; preregistered); the grey estimates are without control variables (exploratory).
-The SESOI was b = |0.30| for life satisfaction and b = |0.15| for affect.
-Hence, all of the reported effects are not considered meaningful.}\label{fig:res-activity}
-\end{figure}
-
-\newpage
-
-\begin{figure}
-\includegraphics[width=\textwidth]{figures/fig_results_channels} \caption{The effects of using various social media applications on three indicators of well-being.
-The black estimates show the effects controlled for a large number of covariates (see text; preregistered); the grey estimates are without control variables (exploratory).
-The SESOI was b = |0.30| for life satisfaction and b = |0.15| for affect.
-Hence, all of the reported effects are not considered meaningful.}\label{fig:res-channels}
-\end{figure}
-
-\newpage
-
-\begin{figure}
-\includegraphics[width=\textwidth]{figures/fig_results_control_std} \caption{Results of selected covariates.
-All variables were standardize except 'Male'  and ‘Employed in public service', because there were measured on a binary scale.}\label{fig:res-control}
-\end{figure}
-
-\newpage
-
-\hypertarget{competing-interests}{%
-\section{Competing Interests}\label{competing-interests}}
-
-I declare no competing interests.
-
-\hypertarget{supplementary-material}{%
-\section{Supplementary Material}\label{supplementary-material}}
-
-All the stimuli, presentation materials, analysis scripts, and a reproducible version of the manuscript can be found on the companion website (\url{https://XMtRA.github.io/Austrian_Corona_Panel_Project}).
-
-\hypertarget{data-accessibility-statement}{%
-\section{Data Accessibility Statement}\label{data-accessibility-statement}}
-
-The data are shared on AUSSDA, see \url{https://doi.org/10.11587/28KQNS}.
-The data can only be used for scientific purposes.
-
-\hypertarget{acknowledgements}{%
-\section{Acknowledgements}\label{acknowledgements}}
-
-I would like to thank BLINDED for providing valuable feedback on this manuscript.
-
 \end{document}

### 2022-07-25 manuscript.Rmd
- corrected typos

@@
 Governments around the world implemented safety measures, such as wearing masks, keeping physical distance, or enforcing lockdowns.
-In this extraordinary situation, many people heavily relied on media to attain relevant information, and especially social media were at an all time high [@statistaAverageDailyTime2021].
+In this extraordinary situation, many people heavily relied on media to obtain relevant information, and especially social media were at an all time high [@statistaAverageDailyTime2021].
 
@@
 **Together, this should offer a nuanced and comprehensive understanding of communication.**
-First, I investigate how well-being is affected by different types of communication affect, namely active and passive use.
+First, I investigate how well-being is affected by different types of communication, namely active and passive use.
 Defining what constitutes active and what passive use is not always clear, and different understandings are currently discussed [@ellisonWhyWeDon2020; @meierComputermediatedCommunicationSocial2020a].
-Reading is generally considered as passive and writing as active, while there are also specific behaviors such as such liking or sharing content that fall somewhere in-between [@meierDoesPassiveSocial2022].
+Reading is generally considered as passive and writing as active, while there are also specific behaviors such as liking or sharing content that fall somewhere in-between [@meierDoesPassiveSocial2022].
 In this study, I hence distinguish (a) _reading_ (passive), (b) _posting_ (active), and (c) _liking and sharing_ COVID-19 related posts (both active and passive).
@@
 
-But people can also misjudge media affects and are often overly optimistic [@metzgerComparativeOptimismPrivacy2017].
+But people can also misjudge media effects and are often overly optimistic [@metzgerComparativeOptimismPrivacy2017].
 And precisely because social media have so many positive consequences, one can ask if this is not where the actual problem lies.
@@
 There could also be _indirect_ effects.
-When browsing social media for Covid-19 related news, many users reported being captivated to such an extent that they could not stop using social media [@kleinDarklySoothingCompulsion2021].
+When browsing social media for COVID-19 related news, many users reported being captivated to such an extent that they could not stop using social media [@kleinDarklySoothingCompulsion2021].
 During the pandemic social media use was at an all-time high in the US [@statistaAverageDailyTime2021].
@@
 For example, @beyensSocialMediaUse2021 reported that although for some users (roughly one quarter) the effects of social media use on well-being were negative, for almost the same number of users they were positive, while for the rest the effects were neutral.
-This finding is aligned the Differential Susceptibility to Media Effects Model: 
+This finding is aligned with the Differential Susceptibility to Media Effects Model: 
 Although there is substantial _variation_ of media effects for individual users, the _average_ effects reported in the literature are often small [@valkenburgDifferentialSusceptibilityMedia2013].
@@
 <!-- In the end, it is a personal question. -->
-However, it is still necessary and helpful to try provide such a plausible benchmark. 
+However, it is still necessary and helpful to try to provide such a plausible benchmark. 
 I therefore suggest the following SESOI for this research question:
@@
 
-Thus, only when controlling for _all_ relevant confounders, we can correctly estimate causality without bias [@rohrerThinkingClearlyCorrelations2018].
+Thus, only when controlling for _all_ relevant confounders, can we correctly estimate causality without bias [@rohrerThinkingClearlyCorrelations2018].
 Obviously, we can never be entirely sure to have included all confounders, which makes absolute statements regarding causality virtually impossible.
@@
 In this study, I hence analyze how using social media during the last week affected positive and negative affect during the same week.
-In other words, if people during the last week engaged in more COVID-19 related social media use than the usually do, did they feel better or worse during that week than they usually do?
+In other words, if people during the last week engaged in more COVID-19 related social media use than they usually do, did they feel better or worse during that week than they usually do?
 Regarding life satisfaction, I implemented a longer interval.
-If people during the last week used COVID-19 related social media more than the usually do, were they at the end of the week more or less satisfied with their lives than they usually are?
+If people during the last week used COVID-19 related social media more than they usually do, were they at the end of the week more or less satisfied with their lives than they usually are?
 I hence analyze if when a person changes their social media diet, are there (a) _simultaneous_ changes in their affect and (b) _subsequent_ changes in their life satisfaction?
@@
 Because the data were analyzed post-hoc, no a-priori sample size planning on the basis of power analyses was conducted.
-The sample is large, and it is hence well-equipped to reliably detect also small effects.
+The sample is large, and it is hence well-equipped to reliably detect small effects.
 <!-- , which is why no exact post hoc power analysis were conducted. -->
@@
 <!-- Respondents who answered less than 50% of all questions were removed. -->
-**Missing responses were imputed using multiple imputation with predictive mean matching (five iterations, five data-sets).**
-**Predictive mean matching allowed to impute also categorical variables.**
+**Missing responses were imputed using multiple imputation with predictive mean matching (five iterations, five data-sets), including categorical variables.**
 **All variables were imputed except the media use measures, as they were not collected on each wave.**
@@
 **Users who wrote more COVID-19 related posts than usual also experienced slightly more negative affect than usual (_b_ = `r dat_fig_results_within %>% filter(dv == "Negative affect" & iv == "Posting" & type == "within" & Analysis == "Regular") %>% select(estimate) %>% round(2)` [95% CI `r dat_fig_results_within %>% filter(dv == "Negative affect" & iv == "Posting" & type == "within" & Analysis == "Regular") %>% select(conf.low) %>% round(2)`, `r dat_fig_results_within %>% filter(dv == "Negative affect" & iv == "Posting" & type == "within" & Analysis == "Regular") %>% select(conf.high) %>% round(2)`]).**
-**There was a small and statistically nonsignificant trend that reading COVID-19 related content slightly increased life satisfaction (_b_ = `r dat_fig_results_within %>% filter(dv == "Life satisfaction" & iv == "Reading" & type == "within" & Analysis == "Regular") %>% select(estimate) %>% round(2)` [95% CI `r dat_fig_results_within %>% filter(dv == "Life satisfaction" & iv == "Reading" & type == "within" & Analysis == "Regular") %>% select(conf.low) %>% round(2)`, `r dat_fig_results_within %>% filter(dv == "Life satisfaction" & iv == "Reading" & type == "within" & Analysis == "Regular") %>% select(conf.high) %>% round(2)`], _p_ = `r dat_fig_results_within %>% filter(dv == "Life satisfaction" & iv == "Reading" & type == "within" & Analysis == "Regular") %>% select(p.value) %>% my_round("p")`).**
-**At the same time, there was also a small and statistically nonsignificant trend that reading COVID-19 related content decreased positive affect (_b_ = `r dat_fig_results_within %>% filter(dv == "Positive affect" & iv == "Reading" & type == "within" & Analysis == "Regular") %>% select(estimate) %>% round(2)` [95% CI `r dat_fig_results_within %>% filter(dv == "Positive affect" & iv == "Reading" & type == "within" & Analysis == "Regular") %>% select(conf.low) %>% round(2)`, `r dat_fig_results_within %>% filter(dv == "Positive affect" & iv == "Reading" & type == "within" & Analysis == "Regular") %>% select(conf.high) %>% round(2)`], _p_ = `r dat_fig_results_within %>% filter(dv == "Positive affect" & iv == "Reading" & type == "within" & Analysis == "Regular") %>% select(p.value) %>% my_round("p")`).**
+**There was a small and statistically non-significant trend that reading COVID-19 related content slightly increased life satisfaction (_b_ = `r dat_fig_results_within %>% filter(dv == "Life satisfaction" & iv == "Reading" & type == "within" & Analysis == "Regular") %>% select(estimate) %>% round(2)` [95% CI `r dat_fig_results_within %>% filter(dv == "Life satisfaction" & iv == "Reading" & type == "within" & Analysis == "Regular") %>% select(conf.low) %>% round(2)`, `r dat_fig_results_within %>% filter(dv == "Life satisfaction" & iv == "Reading" & type == "within" & Analysis == "Regular") %>% select(conf.high) %>% round(2)`], _p_ = `r dat_fig_results_within %>% filter(dv == "Life satisfaction" & iv == "Reading" & type == "within" & Analysis == "Regular") %>% select(p.value) %>% my_round("p")`).**
+**At the same time, there was also a small and statistically non-significant trend that reading COVID-19 related content decreased positive affect (_b_ = `r dat_fig_results_within %>% filter(dv == "Positive affect" & iv == "Reading" & type == "within" & Analysis == "Regular") %>% select(estimate) %>% round(2)` [95% CI `r dat_fig_results_within %>% filter(dv == "Positive affect" & iv == "Reading" & type == "within" & Analysis == "Regular") %>% select(conf.low) %>% round(2)`, `r dat_fig_results_within %>% filter(dv == "Positive affect" & iv == "Reading" & type == "within" & Analysis == "Regular") %>% select(conf.high) %>% round(2)`], _p_ = `r dat_fig_results_within %>% filter(dv == "Positive affect" & iv == "Reading" & type == "within" & Analysis == "Regular") %>% select(p.value) %>% my_round("p")`).**
 
@@
 **Respondents who across all waves used social media more frequently than others to read about COVID-19 related posts reported slightly lower levels of positive affect than others (_b_ `r dat_fig_results_between %>% filter(dv == "Positive affect" & iv == "Reading" & type == "between" & Analysis == "Regular") %>% select(estimate) %>% my_round("b_txt")` [95% CI `r dat_fig_results_between %>% filter(dv == "Positive affect" & iv == "Reading" & type == "between" & Analysis == "Regular") %>% select(conf.high) %>% my_round("b")`, `r dat_fig_results_between %>% filter(dv == "Positive affect" & iv == "Reading" & type == "between" & Analysis == "Regular") %>% select(conf.low) %>% my_round("b")`]).**
-**Respondents who across all waves used social media more frequently than others to write COVID-19 related posts reported sligthly higher levels of negative affect than others (_b_ `r dat_fig_results_between %>% filter(dv == "Negative affect" & iv == "Posting" & type == "between" & Analysis == "Regular") %>% select(estimate) %>% my_round("b_txt")` [95% CI `r dat_fig_results_between %>% filter(dv == "Negative affect" & iv == "Posting" & type == "between" & Analysis == "Regular") %>% select(conf.high) %>% my_round("b")`, `r dat_fig_results_between %>% filter(dv == "Negative affect" & iv == "Posting" & type == "between" & Analysis == "Regular") %>% select(conf.low) %>% my_round("b")`]).**
+**Respondents who across all waves used social media more frequently than others to write COVID-19 related posts reported slightly higher levels of negative affect than others (_b_ `r dat_fig_results_between %>% filter(dv == "Negative affect" & iv == "Posting" & type == "between" & Analysis == "Regular") %>% select(estimate) %>% my_round("b_txt")` [95% CI `r dat_fig_results_between %>% filter(dv == "Negative affect" & iv == "Posting" & type == "between" & Analysis == "Regular") %>% select(conf.high) %>% my_round("b")`, `r dat_fig_results_between %>% filter(dv == "Negative affect" & iv == "Posting" & type == "between" & Analysis == "Regular") %>% select(conf.low) %>% my_round("b")`]).**
 **At the same time, respondents who across all waves used social media more frequently than others to write COVID-19 related posts also reported slightly higher levels of positive affect (_b_ `r dat_fig_results_between %>% filter(dv == "Positive affect" & iv == "Posting" & type == "between" & Analysis == "Regular") %>% select(estimate) %>% my_round("b_txt")` [95% CI `r dat_fig_results_between %>% filter(dv == "Positive affect" & iv == "Posting" & type == "between" & Analysis == "Regular") %>% select(conf.high) %>% my_round("b")`, `r dat_fig_results_between %>% filter(dv == "Positive affect" & iv == "Posting" & type == "between" & Analysis == "Regular") %>% select(conf.low) %>% my_round("b")`]).**
@@
 **Unfortunately, in general online tonality is often more extreme, negative, or aggressive, which potentially affects their authors.**
-**Note that I controlled for whether or not participants had an corona infection during a specific wave, which rules out the potential explanation that having an infection was the root cause of increased communication and reduced well-being.**
+**Note that I controlled for whether or not participants had a COVID-19 infection during a specific wave, which rules out the potential explanation that having an infection was the root cause of increased communication and reduced well-being.**
 
 **The potential explanation that tonality might be a relevant factor at play here is also supported by the second trend.**
-**People, who during the pandemic spend more time on Instagram than usual, also experienced less negative affect than usual.**
+**People, who during the pandemic spent more time on Instagram than usual, also experienced less negative affect than usual.**
 **Instagram is well-known for its positivity bias [@waterlooNormsOnlineExpressions2018].**
@@
 **Such rather low-key active behaviors do not seem to affect well-being at all.**
-**Regarding passive use, reading COVID-19 related posts is more ambivalent; results showed some weak trends towards a positive effect on life satisfaction, but a negative affect on mood.**
+**Regarding passive use, reading COVID-19 related posts is more ambivalent; results showed some weak trends towards a positive effect on life satisfaction, but a negative effect on mood.**
 **It might be that reading and informing oneself about COVID-19 on social media might be helpful in the long term, but more negative for short-term affect.**
@@
 
-**From a more political and societal perspective, I believe the results imply that it does makes sense critically to reflect upon COVID-19 related social media use.** 
-**In terms of concrete recommendations, on average it might be slightly beneficial to post less actively about COVID-19 on social media and to spent less time on YouTube.**
+**From a more political and societal perspective, the results imply that it can make sense to critically reflect upon COVID-19 related social media use.** 
+**In terms of concrete recommendations, on average it might be slightly beneficial to post less actively about COVID-19 on social media and to spend less time on YouTube.**
 **Potential positive effects, however, will for many users likely not be noticeable.**
-**Results allow to reject a positive effect: Writing more posts on social media will likely not increase well-being.**
+**Results allow us to reject a positive effect: Writing more posts on social media will likely not increase well-being.**
 **At all events, engaging in COVID 19-related social media use should, on average, not be a cause for concern.**
@@
 **The challenge is that** regarding media use, such effects often happen immediately or shortly after use, necessitating intervals in the hours, minutes, or even seconds.
-In many cases only experience sampling studies asking users in the very moment can produce such knowledge.
+In many cases only experience sampling studies asking users at the very moment can produce such knowledge.
 However, even then we don't know for certain if we actually measured the right interval.
@@
 Practically, this would imply a SESOI half as large as I have defined here, namely _b_ = |.15| for well-being and _b_ = |.075| for affect.
-In the case of this study, however, reducing the SESOI would not even make a big difference, as also with these more liberal thresholds all but three effect would still be completely in the null region, and no effect would be outside of the null region.
+In the case of this study, however, reducing the SESOI would not even make a big difference, as also with these more liberal thresholds all but three effects would still be completely in the null region, and no effect would be outside of the null region.
 However, at all events I encourage future research to start a thorough conversation on what effect sizes are considered meaningful and what not.

### 2022-07-25 manuscript.Rmd
- updates to theory

@@
 Is it spreading in my region? 
-How is it transmitted, and how can I protect myself?
+How is it transmitted and how can I protect myself?
 Because for many it was (and at the time of writing still is) a matter of life or death, people aimed to stay informed regarding the latest developments.
@@
 Some people actually couldn't stop using social media to learn about COVID-19 related news.
-A new phenomenon emerged, termed "doomscrolling":
+A new phenomenon termed "doomscrolling" emerged:
 Users were glued to their screens and found it hard to pursue other relevant activities such as working, taking a break, or looking after their children [@kleinDarklySoothingCompulsion2021].
-It was asked whether using social media for COVID-19 related reasons is helpful, or whether it creates an additional burden on mental health [@sandstromDoomscrollingCOVIDNews2021].
+It was increasingly asked whether using social media for COVID-19 related reasons is helpful or whether it creates an additional burden on mental health [@sandstromDoomscrollingCOVIDNews2021].
 These concerns seem justified: 
@@
 
-As a result, with this study I want to build on this research and investigate whether or not COVID-19 related social media use affected well-being during the pandemic.
-To this end, I analyzed a large-scale panel study from the Austrian Corona Panel Project [@kittelAustrianCoronaPanel2020]. 
+As a result, with this study I want to build on this research and investigate whether or not COVID-19 related social media use meaningfully affected well-being during the pandemic.
+To this end, I analyzed a large-scale **stand-alone** panel study from the Austrian Corona Panel Project [@kittelAustrianCoronaPanel2020]. 
 The panel consists of **32** waves and has an overall sample size of **`r nrow(d_wide) %>% prettyNum(big.mark = ",")` participants**.
@@
 
-The underlying theories that guided the selection of variables for my analysis are the two-continua model of mental health [@greenspoonIntegrationSubjectiveWellbeing2001] and the hierarchical taxonomy of computer-mediated communication [@meierComputermediatedCommunicationSocial2020a].
+**Two underlying theories guided the selection of variables for my analysis, the two-continua model of mental health [@greenspoonIntegrationSubjectiveWellbeing2001] and the hierarchical taxonomy of computer-mediated communication [@meierComputermediatedCommunicationSocial2020a].**
 According to the two-continua model, mental health consists of (a) psychopathology and (b) well-being.
@@
 Whereas subjective well-being emphasizes hedonic aspects such as happiness and joy, psychological well-being addresses eudaimonic aspects such as fulfillment and meaning.
-Subjective well-being is primarily about achieving positive affect and avoiding negative affect.
+Subjective well-being is primarily about achieving positive and avoiding negative affect.
 One of the most prominent indicators of well-being is life satisfaction.
-In my view, because it represents a general appraisal of one's life, life satisfaction is best thought of as a meta concept that combines psychological and subjective well-being.
+In my view, because it represents a general appraisal of one's life, life satisfaction is best thought of as a meta concept combining psychological and subjective well-being.
 Notably, life satisfaction is stable and fluctuates only little, whereas it's the exact opposite for affect [@dienlinImpactDigitalTechnology2020].
@@
 **Differentiating different communication channels and communication types is important, because the effects on well-being likely differ across communication channels and communication types.**
-**Whereas active social media use such as chatting is routinely linked to improved well-being, passive use such as reading is considered more negative [@dienlinImpactDigitalTechnology2020].**
-Similarly, branded apps are separate entities with potentially divergent effects. 
-For example, @waterlooNormsOnlineExpressions2018 found that it's more adequate to express negative emotions on WhatsApp than on Twitter or on Instagram. 
+**Whereas active social media use such as chatting is routinely linked to improved well-being, passive use such as reading is often considered more negative [@dienlinImpactDigitalTechnology2020].**
+**Similarly, branded apps are separate communication entities with potentially divergent effects and affordances.**
+**For example, @waterlooNormsOnlineExpressions2018 found that it's more adequate to express negative emotions on WhatsApp than on Twitter or on Instagram.**
 **Especially during a pandemic, it makes sense to analyze if users engage with COVID-19 related content on Instagram, where communication is more positive, or on Facebook, where communication is more critical.**
+**First studies suggest that Instagram use was indeed more beneficial for well-being than Facebook use [@masciantonioDonPutAll2021].**
 

### 2022-07-25 manuscript.Rmd
- language edits

@@
 As a result, with this study I want to build on this research and investigate whether or not COVID-19 related social media use meaningfully affected well-being during the pandemic.
-To this end, I analyzed a large-scale **stand-alone** panel study from the Austrian Corona Panel Project [@kittelAustrianCoronaPanel2020]. 
+To this end, I analyzed a large-scale **standalone** panel study from the Austrian Corona Panel Project [@kittelAustrianCoronaPanel2020]. 
 The panel consists of **32** waves and has an overall sample size of **`r nrow(d_wide) %>% prettyNum(big.mark = ",")` participants**.
@@
 
-**Two underlying theories guided the selection of variables for my analysis, the two-continua model of mental health [@greenspoonIntegrationSubjectiveWellbeing2001] and the hierarchical taxonomy of computer-mediated communication [@meierComputermediatedCommunicationSocial2020a].**
+**Two underlying theories guided the selection of variables for this study, namely the two-continua model of mental health [@greenspoonIntegrationSubjectiveWellbeing2001] and the hierarchical taxonomy of computer-mediated communication [@meierComputermediatedCommunicationSocial2020a].**
 According to the two-continua model, mental health consists of (a) psychopathology and (b) well-being.
@@
 Whereas the first four levels focus on the communication _channel_, the last two address the communication _type_.
-**Differentiating different communication channels and communication types is important, because the effects on well-being likely differ across communication channels and communication types.**
+**Distinguishing different communication channels and communication types is important, because the effects on well-being likely differ across communication channels and communication types.**
 **Whereas active social media use such as chatting is routinely linked to improved well-being, passive use such as reading is often considered more negative [@dienlinImpactDigitalTechnology2020].**
@@
 **Especially during a pandemic, it makes sense to analyze if users engage with COVID-19 related content on Instagram, where communication is more positive, or on Facebook, where communication is more critical.**
-**First studies suggest that Instagram use was indeed more beneficial for well-being than Facebook use [@masciantonioDonPutAll2021].**
+**First studies suggest that during the pandemic Instagram use was indeed more beneficial for well-being than Facebook use [@masciantonioDonPutAll2021].**
+**In this study, to measure the effects of social media use focused on COVID-19 related news and topics, I adopt both the channel and the type of communication perspective.**
+<!-- **Together, this should offer a nuanced and comprehensive understanding of communication.** -->
 
-**In this study, to measure the effects of social media use focused on COVID-19 related news and topics, I adopt both the channel and the type of communication perspective.**
-**Together, this should offer a nuanced and comprehensive understanding of communication.**
-First, I investigate how well-being is affected by different types of communication, namely active and passive use.
+Specifically, I investigate how well-being is affected by different types of communication, namely active and passive use.
 Defining what constitutes active and what passive use is not always clear, and different understandings are currently discussed [@ellisonWhyWeDon2020; @meierComputermediatedCommunicationSocial2020a].
-Reading is generally considered as passive and writing as active, while there are also specific behaviors such as liking or sharing content that fall somewhere in-between [@meierDoesPassiveSocial2022].
-In this study, I hence distinguish (a) _reading_ (passive), (b) _posting_ (active), and (c) _liking and sharing_ COVID-19 related posts (both active and passive).
-Second, I analyze how using the most prominent branded applications affects well-being, and whether this effect changes across applications.
-The branded applications investigated here are Facebook, Twitter, Instagram, WhatsApp, and YouTube---which were, at the time of writing, the most relevant social media apps.
+Reading is generally considered as passive and writing as active use, while there are also specific behaviors falling somewhere in-between such as liking or sharing content [@meierDoesPassiveSocial2022].
+**In this study, I hence distinguish (a) reading (passive), (b) posting (active), and (c) liking and sharing COVID-19 related posts (both active and passive).**
+**Second, I analyze how using the most prominent branded applications affects well-being, and whether this effect changes across applications.**
+**The branded applications investigated here are Facebook, Twitter, Instagram, WhatsApp, and YouTube---which were, at the time of writing, the most relevant social media apps.**
 
 **Worth noting, this study is not about _general_ social media use during times of COVID, but on social media use _focused_ on COVID-19 related content.** 
-**For example, posting thoughts about the pandemic, reading posts and comments, or retweeting COVID-19 related news.**
+**This, for example, includes posting thoughts about the pandemic, reading posts and comments, or retweeting COVID-19 related news.**
 
@@
 
-From a theoretical perspective, how could we explain whether COVID-19 related social media use affects well-being?
+From a theoretical perspective, how could we explain whether and how the various types of COVID-19 related social media use affects well-being?
 <!-- In what follows, I outline potential arguments as to why the effect might be positive or negative, direct or indirect, or nonexistent. -->
@@
 Empirically, social media use on average does not have a strong effect on well-being [@meierComputermediatedCommunicationSocial2020a].
-According to the Differential Susceptibility to Media Effects Model [@valkenburgDifferentialSusceptibilityMedia2013], the effects of media use differ across individuals **and content**.
+According to the Differential Susceptibility to Media Effects Model [@valkenburgDifferentialSusceptibilityMedia2013], the effects of media use differ across individuals **and types of content**.
 Whereas for some media are beneficial, for others they are harmful.
-**Whereas some content provides opportunities (education, advice), other content creates risks (misinformation, hate). [@livingstoneEuropeanResearchChildren2018].**
+**Whereas some content mostly provides opportunities (education, advice), other content rather creates risks (misinformation, hate) [@livingstoneEuropeanResearchChildren2018].**
 Social media can impair well-being when causing embarrassment, stress, or disinformation, and they can improve well-being when providing connectedness, information, or entertainment [@buchiDigitalWellbeingTheory2021].
@@
 
-**Two of the most prominent media effect theories argue (indirectly) against strong average negative impacts.**
-**According to mood management theory [@zillmannMoodManagementCommunication1988], using media substantially affects people's moods.**
-**Effects can be stimulating or overwhelming, relaxing or boring.**
+**Two prominent media effect theories argue (mostly implicitly) against strong average negative effects.**
+**According to mood management theory [@zillmannMoodManagementCommunication1988], using media can substantially affect people's moods.**
+**Use can be stimulating or overwhelming, relaxing or boring.**
 **After some time, users implicitly learn what media help them balance their mood and affect according to their own situational needs [@zillmannMoodManagementCommunication1988].**
-**Those media that eventually become part of one's media repertoire are hence, on average, beneficial for users and their moods.**
+**Those media that eventually become part of one's media repertoire are hence, on average, beneficial for users to regulate their mood.**
 **Using experience sampling of well-being and logs of social media use, a study with 82 participants from Italy found that after episodes of social media use, levels of positive affect increased significantly [@marcianoDynamicsAdolescentsSmartphone2022].**
@@
 **They help find relevant information, maintain and foster relationships, express one's personality, and entertain oneself [@pelletierOneSizeDoesn2020].**
-**In conclusion, because people spend so much time on social media consuming COVID-19 related content, according to both mood management theory and uses and gratifications theory we wouldn't expect to find strong average negative effects.**
+**In conclusion, because people spend so much time on social media consuming COVID-19 related content, according to both mood management theory and uses and gratifications theory this is indirect proof that average effects on well-being are likely not particularly negative.**
 
 But people can also misjudge media effects and are often overly optimistic [@metzgerComparativeOptimismPrivacy2017].
-And precisely because social media have so many positive consequences, one can ask if this is not where the actual problem lies.
-In other words, social media aren't problematic because they're inherently bad, but rather because they're too good.
-And as with many other things, there can be too much of a good thing.
-It is therefore often asked whether social media are addictive, and users sometimes express this fear themselves [@yangCanWatchingOnline2021]. 
+Precisely because social media have so many positive consequences, one can ask if this is not where the actual problem lies.
+In other words, social media might not problematic because they are inherently bad, but rather because they are too good.
+As with many other things, there can be too much of a good thing.
+It is therefore often asked whether social media can become addictive, and users sometimes express this fear themselves [@yangCanWatchingOnline2021]. 
 However, a recently published meta-analysis found that the two most prominent measures of addiction, the Bergen Facebook Addiction Scale and the Bergen Social Media Addiction Scale, have only small relations to well-being [@duradoniWellbeingSocialMedia2020].
@@
 That said, because not all news was negative, and because many people showed solidarity and compassion, there was also positive and uplifting content, potentially compensating for the negative effects [@dornemannHowGoodBad2021].
-A study with 2.057 respondents from Italy reported that during the pandemic virtual community and social connectedness even increased during the pandemic [@guazziniSecondWaveAnalysis2022].
-**In Finland, in a sample of 735 participants, levels of loneliness did not decrease during the pandemic, and people who engaged more on social media showed less loneliness [@latikkaLonelinessPsychologicalDistress2022].**
+A study with 2.057 respondents from Italy reported that during the pandemic virtual community and social connectedness even increased [@guazziniSecondWaveAnalysis2022].
+**In a study with 735 participants from Finland, levels of loneliness did not decrease during the pandemic, and people who engaged more on social media experienced less loneliness [@latikkaLonelinessPsychologicalDistress2022].**
 <!-- However, in light of a worldwide pandemic with millions of deaths, the negative direct effect seems more plausible. -->
 
-There could also be _indirect_ effects.
+Second, there could also be _indirect_ effects.
 When browsing social media for COVID-19 related news, many users reported being captivated to such an extent that they could not stop using social media [@kleinDarklySoothingCompulsion2021].
@@
 During the pandemic, several users shared how they successfully dealt with challenges such as physical distancing.
-In a study with 1131 residents from Wuhan in China [@yuePassiveSocialMedia2022], people who spent more time in quarantine also spent more time on social media.
+In a study with 1,131 residents from Wuhan in China [@yuePassiveSocialMedia2022], people who spent more time in quarantine also spent more time on social media.
 Those, who spent more time on social media also engaged in more upward social comparison, which was related to increased levels of stress.**
@@
 
-To summarize, from a theoretical perspective it is most likely that the average effects of social media use on well-being are negligible. 
-Building on established theories from Communication, we would not assume that effects are either profoundly negative or strongly positive.
+To summarize, it seems that from a theoretical perspective it is most likely that the average effects of social media use on well-being are negligible. 
+Building on established theories from Communication and current empirical findings, we would not assume that effects are either profoundly negative or strongly positive.
 
@@
 
-So far, there is only little empirical research on how well-being is affected by COVID-19 related social media.
+So far, there is still comparatively little empirical research on how well-being is affected by COVID-19 related social media use.
 In their study on the relations between media use and mental health during the pandemic, @bendauAssociationsCOVID19Related2021 found that people who used social media as a primary source of information reported on average "significantly more unspecific anxiety and depression [] and significantly more specific COVID-19 related anxiety symptoms" (p.
@@
 analyzed the effects of general digital technology use---objectively measured via screenshots of screen-time applications---on mental health, separating within- and between-person relations [@sewallObjectivelyMeasuredDigital2021].
-The results showed that digital technology did not have any significant effects on mental health [for a similar study with comparable results, see @bradleyStressMoodSmartphone2021].
+The results showed that digital technology did not have significant effects on mental health [for a similar study with comparable results, see @bradleyStressMoodSmartphone2021].
 Together, the literature is mixed, with a slight focus on the negative effects of social media as news use [see also @dornemannHowGoodBad2021; @liuRelationOfficialWhatsAppdistributed2020; @riehmAssociationsMediaExposure2020].
-However, note that all of these findings represent between-person relations stemming from cross-sectional data (see below).
-We therefore don't know whether the differences in mental health and well-being are due to social media use or due to other third variables, such as age, health, employment, or education.
+<!-- However, note that all of these findings represent between-person relations stemming from cross-sectional data (see below). -->
+<!-- We therefore don't know whether the differences in mental health and well-being are due to social media use or due to other third variables, such as age, health, employment, or education. -->
 
@@
 This also holds true for the different types of communication such as active or passive use.
-A meta review (i.e., an analysis of meta-analyses) found that the relation between social media use and well-being is likely in the negative spectrum but very small [@meierComputermediatedCommunicationSocial2020a]---potentially too small to matter.
-What determines whether or not an effect is considered small or trivial?
+A meta review (i.e., an analysis of meta-analyses) found that the relation between social media use and well-being is likely in the negative spectrum but very small, potentially too small to matter [@meierComputermediatedCommunicationSocial2020a].
+What determines whether or not an effect should be considered small or trivial?
 As a starting point, we could refer to standardized effect sizes.
@@
 
-Finally, also several individual studies employing advanced methods found smalls relations between social media use and well-being [@orbenSocialMediaEnduring2019; @przybylskiDoesTakingShort2021a; @schemerImpactInternetSocial2021; @kerestesAdolescentsOnlineSocial2020].
-For example, @beyensSocialMediaUse2021 reported that although for some users (roughly one quarter) the effects of social media use on well-being were negative, for almost the same number of users they were positive, while for the rest the effects were neutral.
+Finally, several individual studies employing advanced methods found smalls relations between social media use and well-being [@orbenSocialMediaEnduring2019; @przybylskiDoesTakingShort2021a; @schemerImpactInternetSocial2021; @kerestesAdolescentsOnlineSocial2020].
+For example, @beyensSocialMediaUse2021 reported that although for roughly one quarter of all users the effects of social media use on well-being were negative, for almost the same number of users they were positive, while for the rest the effects were neutral.
 This finding is aligned with the Differential Susceptibility to Media Effects Model: 
@@
 In conclusion, in light of the theoretical considerations and empirical studies presented above, I expect that COVID-19 related communication on social media doesn't affect well-being in a meaningful or relevant way.
-**This general hypothesis will be analyzed specifically for the communication types of (a) time spent reading, (b) liking and sharing, and (c) actively posting COVID-19 related content.**
-**In addition, I will analyze how well-being is influenced by spending time on five prominent social media apps, including (a) Facebook, (b) Instagram, (c) Twitter, (d) WhatsApp, and (e) YouTube.** 
-**Three different well-being indicators will be differentiated: life satisfaction, positive affect, and negative affect.**
 
 > Hypothesis: The within-person effects of all types of COVID-19 related social media use on all types of well-being indicators---while controlling for several stable and varying covariates such as sociodemographic variables and psychological dispositions---will be trivial.
+
+**In this study, this general hypothesis will be analyzed specifically for the three communication types of (a) time spent reading, (b) liking and sharing, and (c) actively posting COVID-19 related content.**
+**In addition, I will analyze how well-being is influenced by spending time on five prominent social media apps, including (a) Facebook, (b) Instagram, (c) Twitter, (d) WhatsApp, and (e) YouTube.** 
+**Three different well-being indicators will be distinguished: life satisfaction, positive affect, and negative affect.**
 
@@
 But what's a noticeable change in well-being?
-According to @normanInterpretationChangesHealthrelated2003, people can reliably distinguish _seven_ levels of satisfaction with health.
+According to @normanInterpretationChangesHealthrelated2003, people can reliably distinguish seven levels of satisfaction with health.
 So if satisfaction is measured on a 7-point scale, we would state that a four unit change in social media use should result in a one unit change in life satisfaction. 
@@
 **Importantly, when analyzing longitudinal (within-person) relationships and effects, it is important to control for _varying_ third variables.**
-**Non-varying third variables only help control non-varying (between-person) relations.**
+**Non-varying third variables can only help control non-varying (between-person) relations.**
 
@@
 We can also control for variables that affect only social media use or well-being. 
-However, in doing so not much is gained or lost, because the effects of social media use would remain virtually the same [@klinePrinciplesPracticeStructural2016; but see @mcelreathYesterdayClass2021].
+However, in doing so not much is gained, because the effects of social media use would remain virtually the same [@klinePrinciplesPracticeStructural2016; but see @mcelreathYesterdayClass2021].
 
@@
 Next to including covariates, it's now increasingly understood that causal effects should be analyzed from an internal, within-person perspective [@hamakerWhyResearchersShould2014].
-If a specific person changes their media diet, we need to measure how this behavior affects _their own_ well-being.
+If a specific person changes their media diet, we need to measure how this behavior affects their own well-being.
 Between-person comparisons from cross-sectional data, where participants are interviewed only once, cannot provide such insights.
@@
 <!-- Finding the right interval is difficult. -->
-If variables are stable, longer intervals are needed; if they fluctuate, shorter intervals.
-In the case of well-being, we need shorter intervals for **the more fluctuating** positive and negative affect and longer ones for **the more stable** life satisfaction [@dienlinImpactDigitalTechnology2020].
+**If variables are stable, longer intervals are needed; if they fluctuate, shorter intervals.**
+**In the case of well-being, we need shorter intervals for the more fluctuating positive and negative affect, and longer ones for the more stable life satisfaction [@dienlinImpactDigitalTechnology2020].**
 **Using social media can have instant effects on mood [@marcianoDynamicsAdolescentsSmartphone2022].**
-**Effects on life satisfaction often take longer to manifest, for example because media use leads to actual changes in specific behaviors, which then affect life satisfaction [@dienlinDisplacementReinforcementReciprocity2017].**
+**Effects on life satisfaction often take longer to manifest, for example because media use leads to actual changes in specific behaviors, which then in turn affect life satisfaction [@dienlinDisplacementReinforcementReciprocity2017].**
 Choosing the right interval is challenging, because especially short intervals are hard to implement in practice, often requiring advanced methods such as experience sampling (also known as in situ measurement or ambulant assessment) [@schnauber-stockmannMobileDevicesTools2020].
@@
 **The data are hosted on AUSSDA and are publicly available here: https://doi.org/10.11587/28KQNS.**
-**At the time of writing, the official website included a data-set consisting of 24 waves. **
-**For the analyses presented here, I was able to receive an advance version consisting of all 32 waves.**
-**The study was conducted between March 2020 and June 2022, and data collection is now officially finished.**
-**It contains 32 waves.**
-Between March 2020 and July 2020, the intervals were weekly, and afterward the intervals were monthly.
+**At the time of writing, the official website featured a data-set consisting of 24 waves.**
+**For the analyses presented here, I received an advance version consisting of all 32 waves.**
+**The study was conducted between March 2020 and June 2022, and data collection is now finished.**
+Between March 2020 and July 2020, the intervals between waves were weekly, and afterward the intervals were monthly.
 Each wave consists of at least 1,500 respondents.
@@
 To this end, I adopted the interval testing approach as proposed by @dienesUsingBayesGet2014.
-On the basis of the SESOI, I then defined a null region.
+On the basis of the SESOI, I defined a null region.
 In what follows, I explain how I determined the SESOI and the null region.
@@
 Let's briefly illustrate what this means in practice.
-If the 95% confidence interval falls completely within the null-region (e.g., _b_ = .20, [95% CI: .15, .25]), the hypothesis that the effect is trivial is supported.
-If the confidence interval and the null region overlap (e.g., _b_ = .30, [95% CI: .25, .35]), the hypothesis is not supported and the results are considered inconclusive, while a meaningful negative effect is rejected.
-If the confidence interval falls completely outside of the null-region (e.g., _b_ = .40, [95% CI: .35, .45]), the hypothesis is rejected and the existence of a meaningful positive effect is supported.
+If the 95% confidence interval falls completely within the null-region (e.g., _b_ = -.02, [95% CI: -.12, .08]), the hypothesis that the effect is trivial is supported.
+If the confidence interval and the null region overlap (e.g., _b_ = -.22, [95% CI: -.27, -.17]), the hypothesis is not supported and the results are considered inconclusive, while a meaningful negative effect is rejected.
+If the confidence interval falls completely outside of the null-region (e.g., _b_ = -.40, [95% CI: -.45, -.35]), the hypothesis is rejected and the existence of a meaningful positive effect is supported.
 For an illustration, see Figure \@ref(fig:sesoi)).
@@
 Social media use was measured for all participants on waves 1, 2, 8, 17, 23, **and 28**.
-Freshly recruited respondents always answered all questions on social media use.
+Freshly recruited respondents always answered all questions on COVID 19-related social media use.
 
@@
 The effects of COVID-19 related social media use were controlled for the following stable variables: 
-(a) gender (female, male, diverse), (b) age, (c) education (ten options), (d) Austria country of birth (yes/no), (e) Austria parents' country of birth (no parent, one parent, both parents).
+(a) gender (female, male, diverse), (b) age, (c) education (ten options), (d) Austria country of birth (yes/no), (e) Austria parents' country of birth (no parent, one parent, both parents), (f) household size, (g) work hours per week, (h) home office, and (i) household income.
 I originally planned to implement additional variables as varying covariates.
 However, because they were not measured often enough or not at the time when social media use was measured, I implemented them as stable variables using their average values across all waves.
-This includes (a) text-based media news consumption (five degrees), (b) video-based media news consumption (five degrees), (c) residency is Vienna (yes/no), (d) self-reported physical health (five degrees), (e) living space (eleven options), (f) access to balcony (yes/no), (g) access to garden (yes/no), (h) employment (nine options), (i) disposition to take risks (eleven degrees), and (j) locus of control (five degrees).
-I also controlled for the following varying covariates: (a) five items measuring outdoor activities such as sport or meeting friends (five degrees), and (b) satisfaction with democracy (five degrees).
-Because it lead to too much attrition in the sample, I did not control for (a) household size, (b) work hours per week, (c) home office, (d) household income.
+This includes (a) text-based media news consumption (five degrees), (b) video-based media news consumption (five degrees), (c) residency is Vienna (yes/no), (d) living space (eleven options), (e) access to balcony (yes/no), (f) access to garden (yes/no), (g) employment (nine options), (h) disposition to take risks (eleven degrees), and (i) locus of control (five degrees).
+I also controlled for the following varying covariates: (a) five items measuring outdoor activities such as sport or meeting friends (five degrees), (b) satisfaction with democracy (five degrees), (c) self-reported physical health (five degrees), and (d) whether participants contracted COVID-19 since the last wave.
 
@@
 The study's main hypothesis was that the effects of social media use on well-being would be trivial.
-Regarding the effects of different communication _types_---that is, reading vs.
+Regarding the effects of different communication types---that is, reading vs.
 sharing vs.
 posting---all within-person effects fell completely within the a-priori defined null region (see Figure \@ref(fig:fig-between)).
-For example, respondents who used social media more frequently than usual to read about COVID-19 related topics did not show a simultaneous change in life satisfaction (_b_ = `r dat_fig_results_within %>% filter(dv == "Life satisfaction" & iv == "Reading" & type == "within" & Analysis == "Regular") %>% select(estimate) %>% round(2)` [95% CI `r dat_fig_results_within %>% filter(dv == "Life satisfaction" & iv == "Reading" & type == "within" & Analysis == "Regular") %>% select(conf.low) %>% round(2)`, `r dat_fig_results_within %>% filter(dv == "Life satisfaction" & iv == "Reading" & type == "within" & Analysis == "Regular") %>% select(conf.high) %>% round(2)`]).
+For example, respondents who used social media more frequently than usual to like or share COVID-19 related content did not show a simultaneous change in negative affect (_b_ = `r dat_fig_results_within %>% filter(dv == "Life satisfaction" & iv == "Liking & Sharing" & type == "within" & Analysis == "Regular") %>% select(estimate) %>% round(2)` [95% CI `r dat_fig_results_within %>% filter(dv == "Life satisfaction" & iv == "Liking & Sharing" & type == "within" & Analysis == "Regular") %>% select(conf.low) %>% round(2)`, `r dat_fig_results_within %>% filter(dv == "Life satisfaction" & iv == "Liking & Sharing" & type == "within" & Analysis == "Regular") %>% select(conf.high) %>% round(2)`]).
 As a result, the hypothesis was supported for all COVID-19 related types of social media communication.
+
 **However, two effects were statistically significantly different from zero.**
-**Users who wrote more COVID-19 related posts than usual were also slightly less satisfied with their lives as usual (_b_ = `r dat_fig_results_within %>% filter(dv == "Life satisfaction" & iv == "Posting" & type == "within" & Analysis == "Regular") %>% select(estimate) %>% round(2)` [95% CI `r dat_fig_results_within %>% filter(dv == "Life satisfaction" & iv == "Posting" & type == "within" & Analysis == "Regular") %>% select(conf.low) %>% round(2)`, `r dat_fig_results_within %>% filter(dv == "Life satisfaction" & iv == "Posting" & type == "within" & Analysis == "Regular") %>% select(conf.high) %>% round(2)`]).**
+**Users who wrote more COVID-19 related posts than usual were also slightly less satisfied with their lives than usual (_b_ = `r dat_fig_results_within %>% filter(dv == "Life satisfaction" & iv == "Posting" & type == "within" & Analysis == "Regular") %>% select(estimate) %>% round(2)` [95% CI `r dat_fig_results_within %>% filter(dv == "Life satisfaction" & iv == "Posting" & type == "within" & Analysis == "Regular") %>% select(conf.low) %>% round(2)`, `r dat_fig_results_within %>% filter(dv == "Life satisfaction" & iv == "Posting" & type == "within" & Analysis == "Regular") %>% select(conf.high) %>% round(2)`]).**
 **Users who wrote more COVID-19 related posts than usual also experienced slightly more negative affect than usual (_b_ = `r dat_fig_results_within %>% filter(dv == "Negative affect" & iv == "Posting" & type == "within" & Analysis == "Regular") %>% select(estimate) %>% round(2)` [95% CI `r dat_fig_results_within %>% filter(dv == "Negative affect" & iv == "Posting" & type == "within" & Analysis == "Regular") %>% select(conf.low) %>% round(2)`, `r dat_fig_results_within %>% filter(dv == "Negative affect" & iv == "Posting" & type == "within" & Analysis == "Regular") %>% select(conf.high) %>% round(2)`]).**
-**There was a small and statistically non-significant trend that reading COVID-19 related content slightly increased life satisfaction (_b_ = `r dat_fig_results_within %>% filter(dv == "Life satisfaction" & iv == "Reading" & type == "within" & Analysis == "Regular") %>% select(estimate) %>% round(2)` [95% CI `r dat_fig_results_within %>% filter(dv == "Life satisfaction" & iv == "Reading" & type == "within" & Analysis == "Regular") %>% select(conf.low) %>% round(2)`, `r dat_fig_results_within %>% filter(dv == "Life satisfaction" & iv == "Reading" & type == "within" & Analysis == "Regular") %>% select(conf.high) %>% round(2)`], _p_ = `r dat_fig_results_within %>% filter(dv == "Life satisfaction" & iv == "Reading" & type == "within" & Analysis == "Regular") %>% select(p.value) %>% my_round("p")`).**
+**There was also a small and statistically non-significant trend that reading COVID-19 related content slightly increased life satisfaction (_b_ = `r dat_fig_results_within %>% filter(dv == "Life satisfaction" & iv == "Reading" & type == "within" & Analysis == "Regular") %>% select(estimate) %>% round(2)` [95% CI `r dat_fig_results_within %>% filter(dv == "Life satisfaction" & iv == "Reading" & type == "within" & Analysis == "Regular") %>% select(conf.low) %>% round(2)`, `r dat_fig_results_within %>% filter(dv == "Life satisfaction" & iv == "Reading" & type == "within" & Analysis == "Regular") %>% select(conf.high) %>% round(2)`], _p_ = `r dat_fig_results_within %>% filter(dv == "Life satisfaction" & iv == "Reading" & type == "within" & Analysis == "Regular") %>% select(p.value) %>% my_round("p")`).**
 **At the same time, there was also a small and statistically non-significant trend that reading COVID-19 related content decreased positive affect (_b_ = `r dat_fig_results_within %>% filter(dv == "Positive affect" & iv == "Reading" & type == "within" & Analysis == "Regular") %>% select(estimate) %>% round(2)` [95% CI `r dat_fig_results_within %>% filter(dv == "Positive affect" & iv == "Reading" & type == "within" & Analysis == "Regular") %>% select(conf.low) %>% round(2)`, `r dat_fig_results_within %>% filter(dv == "Positive affect" & iv == "Reading" & type == "within" & Analysis == "Regular") %>% select(conf.high) %>% round(2)`], _p_ = `r dat_fig_results_within %>% filter(dv == "Positive affect" & iv == "Reading" & type == "within" & Analysis == "Regular") %>% select(p.value) %>% my_round("p")`).**
 
-Regarding the COVID-19 related use of social media _channels_, the results were comparable (see Figure \@ref(fig:fig-within)).
+Regarding the COVID-19 related use of social media channels, the results were comparable (see Figure \@ref(fig:fig-within)).
 Changes in the frequency of using different social media channels to attain information regarding COVID-19 were unrelated to meaningful changes in well-being.
@@
 In sum, the hypothesis was supported also for the COVID-19 related use of important social media channels.
-**However, two effects differed substantially from zero.**
-**Respondents who used Instagram more frequently than usual to attain COVID-19 related news reported slightly _lower_ levels of negative affect than usual (_b_ = `r dat_fig_results_within %>% filter(dv == "Negative affect" & iv == "Instagram" & type == "within" & Analysis == "Regular") %>% select(estimate) %>% my_round("b")` [95% CI `r dat_fig_results_within %>% filter(dv == "Negative affect" & iv == "Instagram" & type == "within" & Analysis == "Regular") %>% select(conf.low) %>% my_round("b")`, `r dat_fig_results_within %>% filter(dv == "Negative affect" & iv == "Instagram" & type == "within" & Analysis == "Regular") %>% select(conf.high) %>% my_round("b")`]).**
-**Respondents who used YouTube more frequently than usual to attain COVID-19 related news reported slightly _higher_ levels of negative affect than usual (_b_ = `r dat_fig_results_within %>% filter(dv == "Negative affect" & iv == "YouTube" & type == "within" & Analysis == "Regular") %>% select(estimate) %>% round(2)` [95% CI `r dat_fig_results_within %>% filter(dv == "Negative affect" & iv == "YouTube" & type == "within" & Analysis == "Regular") %>% select(conf.low) %>% round(2)`, `r dat_fig_results_within %>% filter(dv == "Negative affect" & iv == "YouTube" & type == "within" & Analysis == "Regular") %>% select(conf.high) %>% my_round("b")`]).**
-**However, both effects were still completely inside of the null region, hence not large enough to be considered meaningful.**
+
+**That said, two effects differed substantially from zero.**
+**Respondents who used Instagram more frequently than usual to attain COVID-19 related news reported slightly lower levels of negative affect than usual (_b_ = `r dat_fig_results_within %>% filter(dv == "Negative affect" & iv == "Instagram" & type == "within" & Analysis == "Regular") %>% select(estimate) %>% my_round("b")` [95% CI `r dat_fig_results_within %>% filter(dv == "Negative affect" & iv == "Instagram" & type == "within" & Analysis == "Regular") %>% select(conf.low) %>% my_round("b")`, `r dat_fig_results_within %>% filter(dv == "Negative affect" & iv == "Instagram" & type == "within" & Analysis == "Regular") %>% select(conf.high) %>% my_round("b")`]).**
+**Respondents who used YouTube more frequently than usual to attain COVID-19 related news reported slightly higher levels of negative affect than usual (_b_ = `r dat_fig_results_within %>% filter(dv == "Negative affect" & iv == "YouTube" & type == "within" & Analysis == "Regular") %>% select(estimate) %>% round(2)` [95% CI `r dat_fig_results_within %>% filter(dv == "Negative affect" & iv == "YouTube" & type == "within" & Analysis == "Regular") %>% select(conf.low) %>% my_round("b")`, `r dat_fig_results_within %>% filter(dv == "Negative affect" & iv == "YouTube" & type == "within" & Analysis == "Regular") %>% select(conf.high) %>% my_round("b")`]).**
+**However, both effects were still completely inside of the null region, hence likely not large enough to be considered meaningful.**
 
@@
 Four relations were statistically significant.
-**Respondents who across all waves used social media more frequently than others to read about COVID-19 related posts reported slightly lower levels of positive affect than others (_b_ `r dat_fig_results_between %>% filter(dv == "Positive affect" & iv == "Reading" & type == "between" & Analysis == "Regular") %>% select(estimate) %>% my_round("b_txt")` [95% CI `r dat_fig_results_between %>% filter(dv == "Positive affect" & iv == "Reading" & type == "between" & Analysis == "Regular") %>% select(conf.high) %>% my_round("b")`, `r dat_fig_results_between %>% filter(dv == "Positive affect" & iv == "Reading" & type == "between" & Analysis == "Regular") %>% select(conf.low) %>% my_round("b")`]).**
-**Respondents who across all waves used social media more frequently than others to write COVID-19 related posts reported slightly higher levels of negative affect than others (_b_ `r dat_fig_results_between %>% filter(dv == "Negative affect" & iv == "Posting" & type == "between" & Analysis == "Regular") %>% select(estimate) %>% my_round("b_txt")` [95% CI `r dat_fig_results_between %>% filter(dv == "Negative affect" & iv == "Posting" & type == "between" & Analysis == "Regular") %>% select(conf.high) %>% my_round("b")`, `r dat_fig_results_between %>% filter(dv == "Negative affect" & iv == "Posting" & type == "between" & Analysis == "Regular") %>% select(conf.low) %>% my_round("b")`]).**
-**At the same time, respondents who across all waves used social media more frequently than others to write COVID-19 related posts also reported slightly higher levels of positive affect (_b_ `r dat_fig_results_between %>% filter(dv == "Positive affect" & iv == "Posting" & type == "between" & Analysis == "Regular") %>% select(estimate) %>% my_round("b_txt")` [95% CI `r dat_fig_results_between %>% filter(dv == "Positive affect" & iv == "Posting" & type == "between" & Analysis == "Regular") %>% select(conf.high) %>% my_round("b")`, `r dat_fig_results_between %>% filter(dv == "Positive affect" & iv == "Posting" & type == "between" & Analysis == "Regular") %>% select(conf.low) %>% my_round("b")`]).**
-**Finally, respondents who across all waves used YouTube more frequently than others also reported slightly higher levels of life satisfaction than others (_b_ `r dat_fig_results_between %>% filter(dv == "Life satisfaction" & iv == "YouTube" & type == "between" & Analysis == "Regular") %>% select(estimate) %>% my_round("b_txt")` [95% CI `r dat_fig_results_between %>% filter(dv == "Life satisfaction" & iv == "YouTube" & type == "between" & Analysis == "Regular") %>% select(conf.high) %>% my_round("b")`, `r dat_fig_results_between %>% filter(dv == "Life satisfaction" & iv == "YouTube" & type == "between" & Analysis == "Regular") %>% select(conf.low) %>% my_round("b")`]).**
+**Respondents who across all waves used social media more frequently than others to read about COVID-19 related posts reported slightly lower levels of positive affect than others (_b_ `r dat_fig_results_between %>% filter(dv == "Positive affect" & iv == "Reading" & type == "between" & Analysis == "Regular") %>% select(estimate) %>% my_round("b_txt")` [95% CI `r dat_fig_results_between %>% filter(dv == "Positive affect" & iv == "Reading" & type == "between" & Analysis == "Regular") %>% select(conf.low) %>% my_round("b")`, `r dat_fig_results_between %>% filter(dv == "Positive affect" & iv == "Reading" & type == "between" & Analysis == "Regular") %>% select(conf.high) %>% my_round("b")`]).**
+**Respondents who across all waves used social media more frequently than others to write COVID-19 related posts reported slightly higher levels of negative affect than others (_b_ `r dat_fig_results_between %>% filter(dv == "Negative affect" & iv == "Posting" & type == "between" & Analysis == "Regular") %>% select(estimate) %>% my_round("b_txt")` [95% CI `r dat_fig_results_between %>% filter(dv == "Negative affect" & iv == "Posting" & type == "between" & Analysis == "Regular") %>% select(conf.low) %>% my_round("b")`, `r dat_fig_results_between %>% filter(dv == "Negative affect" & iv == "Posting" & type == "between" & Analysis == "Regular") %>% select(conf.high) %>% my_round("b")`]).**
+**At the same time, respondents who across all waves used social media more frequently than others to write COVID-19 related posts also reported slightly higher levels of positive affect (_b_ `r dat_fig_results_between %>% filter(dv == "Positive affect" & iv == "Posting" & type == "between" & Analysis == "Regular") %>% select(estimate) %>% my_round("b_txt")` [95% CI `r dat_fig_results_between %>% filter(dv == "Positive affect" & iv == "Posting" & type == "between" & Analysis == "Regular") %>% select(conf.low) %>% my_round("b")`, `r dat_fig_results_between %>% filter(dv == "Positive affect" & iv == "Posting" & type == "between" & Analysis == "Regular") %>% select(conf.high) %>% my_round("b")`]).**
+**Finally, respondents who across all waves used YouTube more frequently than others also reported slightly higher levels of life satisfaction than others (_b_ `r dat_fig_results_between %>% filter(dv == "Life satisfaction" & iv == "YouTube" & type == "between" & Analysis == "Regular") %>% select(estimate) %>% my_round("b_txt")` [95% CI `r dat_fig_results_between %>% filter(dv == "Life satisfaction" & iv == "YouTube" & type == "between" & Analysis == "Regular") %>% select(conf.low) %>% my_round("b")`, `r dat_fig_results_between %>% filter(dv == "Life satisfaction" & iv == "YouTube" & type == "between" & Analysis == "Regular") %>% select(conf.high) %>% my_round("b")`]).**
 <!-- However, note that the effect were still completely inside of the null region, hence not large enough to be considered practically relevant. -->
@@
 For example, on the between-person level, one effect stopped being significant if controlled for additional variables.
-Actively posting on social media was significantly (though not meaningfully) related to decreased life satisfaction.
+Initially, actively posting on social media was significantly (though not meaningfully) related to decreased life satisfaction.
 However, when controlling for potential confounders, the effect became virtually zero.
@@
 Therefore, I report the results of the standardized scales,
-which allows for a better comparison across the differently scaled variables.
+which also allows for a better comparison across the differently scaled variables.
 As a rough estimate for the SESOI we can build on the typical convention that small effects start at _r_ = |.10|.
@@
 
-To find out whether my inferences were robust across legitimate (though arguably inferior) alternative analyses, I reran the analyses also using standardized estimates, mean scores instead of factor scores, and with a data set where missing data were not imputed.
+To find out whether the inferences were robust across plausible (though arguably inferior) alternative analyses, I reran the analyses also using standardized estimates, additional covariates including trust in media or government, single imputation, and with a data set where missing data were not imputed.
 The results were virtually the same.
@@
 The data come from a panel study with 32 waves and are largely representative of the Austrian population.
-In a random effects model I separated between person relations from within-person effects and controlled for a large number of both stable and varying covariates, thereby aiming to assess causality.
-The results showed that some statistically significant negative within-person effects existed, but that they were **very small and likely** trivial.
+In a random effects model I separated between person relations from within-person effects. 
+I controlled for a large number of both stable and varying covariates, aiming to assess causality.
+**The results showed that some statistically significant negative within-person effects existed, but that they were very small and likely negligible.**
 People who used social media more than usual to learn about COVID-19 didn't show meaningful changes in their well-being.
@@
 **Similarly, people who wrote more COVID-19 related posts than usual also reported slightly more negative affect.**
-**When writing posts and comments on social media, people explicitly and more deeply engage with COVID-19 related content.**
-**Unfortunately, in general online tonality is often more extreme, negative, or aggressive, which potentially affects their authors.**
-**Note that I controlled for whether or not participants had a COVID-19 infection during a specific wave, which rules out the potential explanation that having an infection was the root cause of increased communication and reduced well-being.**
+**As a potential explanation, when writing posts and comments on social media people explicitly and more deeply engage with COVID-19 related content.**
+**The tonality on social media is often extreme, negative, or aggressive, which potentially affects their authors.**
+**And because I controlled for whether or not participants had a COVID-19 infection during a specific wave, we can rule out the potential explanation that having an infection was the root cause of increased communication and reduced well-being.**
 
-**The potential explanation that tonality might be a relevant factor at play here is also supported by the second trend.**
+**The hypothesis that tonality might be a relevant factor at play is also supported by the second trend.**
 **People, who during the pandemic spent more time on Instagram than usual, also experienced less negative affect than usual.**
 **Instagram is well-known for its positivity bias [@waterlooNormsOnlineExpressions2018].**
-**Content is generally post positive, uplifting, and (self-)flattering.**
-**It seems the often-criticized positivity bias on Instagram might have been somewhat beneficial in times of the pandemic.**
+**Content is generally more positive, uplifting, and (self-)flattering.**
+**It seems the much maligned positivity bias on Instagram might have been somewhat beneficial in times of the pandemic.**
 **The critique that the positivity bias necessarily leads to envy and negative feelings is one-sided, because positive content can also inspire and motivate users [@meierInstagramInspirationHow2020], which could be especially helpful in times of lockdown and home-office.**
-**To provide a concrete example, Instagram was successfully used as an interactive communication channel for first year students to have a better start into their new degree, effectively complementing alternative learning platform tools [@yeTurningInformationDissipation2020].**
+**To provide a concrete example, during the pandemic Instagram was successfully used as an interactive communication channel for first year students to have a better start into their new degree, effectively complementing alternative learning platform tools [@yeTurningInformationDissipation2020].**
 
@@
 
-On the one hand, the results are not aligned with several recent studies analyzing similar or closely related research questions.
-This includes a study by @bendauAssociationsCOVID19Related2021, which showed negative relations between social media and well-being [but see @bradleyStressMoodSmartphone2021; or @sewallObjectivelyMeasuredDigital2021].
-However, note that @bendauAssociationsCOVID19Related2021 analyzed cross-sectional data on a between-person level while not controlling for third variables, which is not optimal for investigating causal effects.
-On the other hand, the results are well-aligned with **mood management theory [@zillmannMoodManagementCommunication1988]** or the uses and gratifications approach [@katzUsesGratificationsResearch1973].
-If effects were indeed profoundly negative on average, then people likely wouldn't spend so much time on social media.
-Likewise, recent studies and meta-analyses analyzing the effects of social media use from a more general perspective or from a somewhat different angle. 
-These studies have found that the effects of various types of social media use on several well-being indicators are small at best, often too small to matter [@meierComputermediatedCommunicationSocial2020a; @orbenTeenagersScreensSocial2020; @fergusonThisMetaanalysisScreen2021], which echoes the results obtained here.
-
 **The results showed that it makes sense to analyze different communication types and communication channels, and that active and passive communication showed different results.**
@@
 **Such rather low-key active behaviors do not seem to affect well-being at all.**
-**Regarding passive use, reading COVID-19 related posts is more ambivalent; results showed some weak trends towards a positive effect on life satisfaction, but a negative effect on mood.**
-**It might be that reading and informing oneself about COVID-19 on social media might be helpful in the long term, but more negative for short-term affect.**
+**Regarding passive use, reading COVID-19 related posts is more ambivalent.**
+**Results showed some weak trends towards a positive effect on life satisfaction, but a negative effect on mood.**
+**It might be that reading and informing oneself about COVID-19 on social media is helpful in the long run, but more negative for short-term affect.**
 **Finally, proactively engaging via writing posts, the most active form of communication analyzed here, showed only negative effects on well-being.**
-**Together, the results support the findings from @valkenburgAssociationsActivePassive2022, who could not confirm the claim that active use is good, versus passive is bad.**
+**The results support the findings from @valkenburgAssociationsActivePassive2022, who also could not confirm the claim active use is good and passive use is bad.**
 **Focusing on communication channels, YouTube seems to be more negative, whereas Instagram is likely more positive.**
@@
 
+Taken together, on the one hand the results are not aligned with several recent studies analyzing similar or closely related research questions.
+This includes a study by @bendauAssociationsCOVID19Related2021, which showed negative relations between social media and well-being [but see @bradleyStressMoodSmartphone2021; or @sewallObjectivelyMeasuredDigital2021].
+However, note that @bendauAssociationsCOVID19Related2021 analyzed cross-sectional data on a between-person level while not controlling for third variables, which is not optimal for investigating causal effects.
+On the other hand, the results are well-aligned with **mood management theory [@zillmannMoodManagementCommunication1988]** or the uses and gratifications approach [@katzUsesGratificationsResearch1973].
+If effects were indeed profoundly negative on average, then people likely wouldn't spend so much time on social media engaging with COVID-19 content.
+Likewise, recent studies and meta-analyses analyzed the effects of social media use from a more general perspective or from a somewhat different angle. 
+These studies have found that the effects of various types of social media use on several well-being indicators are small at best, often too small to matter [@meierComputermediatedCommunicationSocial2020a; @orbenTeenagersScreensSocial2020; @fergusonThisMetaanalysisScreen2021], which echoes the results obtained here.
+
 **From a more political and societal perspective, the results imply that it can make sense to critically reflect upon COVID-19 related social media use.** 
-**In terms of concrete recommendations, on average it might be slightly beneficial to post less actively about COVID-19 on social media and to spend less time on YouTube.**
-**Potential positive effects, however, will for many users likely not be noticeable.**
+**On average, it might be slightly beneficial to post less actively about COVID-19 on social media and to spend less time on YouTube.**
+**Spending time on Instagram seems to be okay.**
+**The potentially resulting positive effects, however, will for many users likely not be noticeable.**
 **Results allow us to reject a positive effect: Writing more posts on social media will likely not increase well-being.**
-**At all events, engaging in COVID 19-related social media use should, on average, not be a cause for concern.**
+**At all events, engaging in COVID 19-related social media use should, on average, not be a major cause for concern.**
 
@@
 Together, this allowed for an improved perspective on assessing causality.
-**However, the opposite effect is also plausible, namely that well-being affects media use [@zillmannMoodManagementCommunication1988].**
+**However, the opposite effect is still also plausible, namely that well-being affected media use [@zillmannMoodManagementCommunication1988].**
 **While controlling for potential confounders can support claims of causality, the procedure implemented here cannot prove causality.**
@@
 To borrow the words from @rohrerTheseAreNot2021, there is no such thing as "the" effect of social media use on well-being.
-Hence, to document how effects unfold, future research needs to employ different study designs probing different time lags.
+Hence, to document how effects unfold, future research needs to employ different study designs probing different intervals.
 In addition, more thought needs to be invested in what relevant stable and varying factors we should include as control variables, and I hope this study provides a first step into this direction.
@@
 regularly exercising _and_ establishing a reading habit) should then cause a detectable improvement in well-being?
-Practically, this would imply a SESOI half as large as I have defined here, namely _b_ = |.15| for well-being and _b_ = |.075| for affect.
-In the case of this study, however, reducing the SESOI would not even make a big difference, as also with these more liberal thresholds all but three effects would still be completely in the null region, and no effect would be outside of the null region.
+Practically, this would imply a SESOI half as large as I have defined here, namely _b_ = |.15| for life satisfaction and _b_ = |.075| for affect.
+In the case of this study, however, reducing the SESOI would not even make a big difference, as also with these more liberal thresholds all but two effects would still be completely in the null region, and no effect would be outside of the null region.
 However, at all events I encourage future research to start a thorough conversation on what effect sizes are considered meaningful and what not.
@@
 Note.
-Here, a trivial effect of social media use on life satisfaction is defined as ranging from b = -.25 to b = .25", fig.width=6, fig.height=2, warning=F, message=F, prompt=F}
+Here, a trivial effect of social media use on life satisfaction is defined as ranging from b = -.30 to b = .30", fig.width=6, fig.height=2, warning=F, message=F, prompt=F}
 # make table with data
@@
   "1. Accept trivial effect", "Decision Rule", -.02, -.12, .08, 
-  "2. Reject trivial effect", "Decision Rule", -.37, -.47, -.27,
+  "2. Reject trivial effect", "Decision Rule", -.40, -.45, -.35,
   "3. Reject positive effect", "Decision Rule", -.22, -.27, -.17,
-  "4. Suspend judgement", "Decision Rule", -.1, -.30, .10,
+  "4. Suspend judgement", "Decision Rule", -.1, -.32, .08,
 ) %>% 
@@
   geom_vline(xintercept = 0, color = "darkgrey") + 
-  geom_vline(xintercept = -.25, color = "darkgrey", linetype = "dashed") +
-  geom_vline(xintercept = .25, color = "darkgrey", linetype = "dashed") +
+  geom_vline(xintercept = -.3, color = "darkgrey", linetype = "dashed") +
+  geom_vline(xintercept = .3, color = "darkgrey", linetype = "dashed") +
   geom_point(aes(Effect)) +
@@
   labs(x = "Effect Size (b)",
-       caption = "Smallest effect size of interest: b = |.25|
-                  Null region: b = -.25, .25")
+       caption = "Smallest effect size of interest: b = |.30|
+                  Null region: b = -.30, .30")
 

### 2022-11-17 manuscript.Rmd
- small changes likely bc different mac was used.

@@
 documentclass     : "apa7"
-# output            :
-#   papaja::apa6_pdf:
-#     latex_engine  : xelatex
+output            :
+  papaja::apa6_pdf:
+    latex_engine  : xelatex
 # output            : papaja::apa6_docx
-output            : html_document
+# output            : html_document
 ---

### 2022-11-17 manuscript.tex
- small changes likely bc different mac was used.

@@
 Is it spreading in my region?
-How is it transmitted, and how can I protect myself?
+How is it transmitted and how can I protect myself?
 Because for many it was (and at the time of writing still is) a matter of life or death, people aimed to stay informed regarding the latest developments.
 Governments around the world implemented safety measures, such as wearing masks, keeping physical distance, or enforcing lockdowns.
-In this extraordinary situation, many people heavily relied on media to attain relevant information, and especially social media were at an all time high (Statista, 2021).
+In this extraordinary situation, many people heavily relied on media to obtain relevant information, and especially social media were at an all time high (Statista, 2021).
 
 Some people actually couldn't stop using social media to learn about COVID-19 related news.
-A new phenomenon emerged, termed ``doomscrolling'':
+A new phenomenon termed ``doomscrolling'' emerged:
 Users were glued to their screens and found it hard to pursue other relevant activities such as working, taking a break, or looking after their children (Klein, 2021).
-It was asked whether using social media for COVID-19 related reasons is helpful, or whether it creates an additional burden on mental health (Sandstrom et al., 2021).
+It was increasingly asked whether using social media for COVID-19 related reasons is helpful or whether it creates an additional burden on mental health (Sandstrom et al., 2021).
 These concerns seem justified:
@@
 
-As a result, with this study I want to build on this research and investigate whether or not COVID-19 related social media use affected well-being during the pandemic.
-To this end, I analyzed a large-scale panel study from the Austrian Corona Panel Project (Kittel et al., 2020).
+As a result, with this study I want to build on this research and investigate whether or not COVID-19 related social media use meaningfully affected well-being during the pandemic.
+To this end, I analyzed a large-scale \textbf{standalone} panel study from the Austrian Corona Panel Project (Kittel et al., 2020).
 The panel consists of \textbf{32} waves and has an overall sample size of \textbf{3,485 participants}.
@@
 
-The underlying theories that guided the selection of variables for my analysis are the two-continua model of mental health (Greenspoon \& Saklofske, 2001) and the hierarchical taxonomy of computer-mediated communication (Meier \& Reinecke, 2020).
+\textbf{Two underlying theories guided the selection of variables for this study, namely the two-continua model of mental health (Greenspoon \& Saklofske, 2001) and the hierarchical taxonomy of computer-mediated communication (Meier \& Reinecke, 2020).}
 According to the two-continua model, mental health consists of (a) psychopathology and (b) well-being.
@@
 Whereas subjective well-being emphasizes hedonic aspects such as happiness and joy, psychological well-being addresses eudaimonic aspects such as fulfillment and meaning.
-Subjective well-being is primarily about achieving positive affect and avoiding negative affect.
+Subjective well-being is primarily about achieving positive and avoiding negative affect.
 One of the most prominent indicators of well-being is life satisfaction.
-In my view, because it represents a general appraisal of one's life, life satisfaction is best thought of as a meta concept that combines psychological and subjective well-being.
+In my view, because it represents a general appraisal of one's life, life satisfaction is best thought of as a meta concept combining psychological and subjective well-being.
 Notably, life satisfaction is stable and fluctuates only little, whereas it's the exact opposite for affect (Dienlin \& Johannes, 2020).
@@
 Whereas the first four levels focus on the communication \emph{channel}, the last two address the communication \emph{type}.
-\textbf{Differentiating different communication channels and communication types is important, because the effects on well-being likely differ across communication channels and communication types.}
-\textbf{Whereas active social media use such as chatting is routinely linked to improved well-being, passive use such as reading is considered more negative (Dienlin \& Johannes, 2020).}
-Similarly, branded apps are separate entities with potentially divergent effects.
-For example, Waterloo et al. (2018) found that it's more adequate to express negative emotions on WhatsApp than on Twitter or on Instagram.
+\textbf{Distinguishing different communication channels and communication types is important, because the effects on well-being likely differ across communication channels and communication types.}
+\textbf{Whereas active social media use such as chatting is routinely linked to improved well-being, passive use such as reading is often considered more negative (Dienlin \& Johannes, 2020).}
+\textbf{Similarly, branded apps are separate communication entities with potentially divergent effects and affordances.}
+\textbf{For example, Waterloo et al. (2018) found that it's more adequate to express negative emotions on WhatsApp than on Twitter or on Instagram.}
 \textbf{Especially during a pandemic, it makes sense to analyze if users engage with COVID-19 related content on Instagram, where communication is more positive, or on Facebook, where communication is more critical.}
-
+\textbf{First studies suggest that during the pandemic Instagram use was indeed more beneficial for well-being than Facebook use (Masciantonio et al., 2021).}
 \textbf{In this study, to measure the effects of social media use focused on COVID-19 related news and topics, I adopt both the channel and the type of communication perspective.}
-\textbf{Together, this should offer a nuanced and comprehensive understanding of communication.}
-First, I investigate how well-being is affected by different types of communication affect, namely active and passive use.
+
+Specifically, I investigate how well-being is affected by different types of communication, namely active and passive use.
 Defining what constitutes active and what passive use is not always clear, and different understandings are currently discussed (Ellison et al., 2020; Meier \& Reinecke, 2020).
-Reading is generally considered as passive and writing as active, while there are also specific behaviors such as such liking or sharing content that fall somewhere in-between (Meier \& Krause, 2022).
-In this study, I hence distinguish (a) \emph{reading} (passive), (b) \emph{posting} (active), and (c) \emph{liking and sharing} COVID-19 related posts (both active and passive).
-Second, I analyze how using the most prominent branded applications affects well-being, and whether this effect changes across applications.
-The branded applications investigated here are Facebook, Twitter, Instagram, WhatsApp, and YouTube---which were, at the time of writing, the most relevant social media apps.
+Reading is generally considered as passive and writing as active use, while there are also specific behaviors falling somewhere in-between such as liking or sharing content (Meier \& Krause, 2022).
+\textbf{In this study, I hence distinguish (a) reading (passive), (b) posting (active), and (c) liking and sharing COVID-19 related posts (both active and passive).}
+\textbf{Second, I analyze how using the most prominent branded applications affects well-being, and whether this effect changes across applications.}
+\textbf{The branded applications investigated here are Facebook, Twitter, Instagram, WhatsApp, and YouTube---which were, at the time of writing, the most relevant social media apps.}
 
 \textbf{Worth noting, this study is not about \emph{general} social media use during times of COVID, but on social media use \emph{focused} on COVID-19 related content.}
-\textbf{For example, posting thoughts about the pandemic, reading posts and comments, or retweeting COVID-19 related news.}
+\textbf{This, for example, includes posting thoughts about the pandemic, reading posts and comments, or retweeting COVID-19 related news.}
 
@@
 
-From a theoretical perspective, how could we explain whether COVID-19 related social media use affects well-being?
+From a theoretical perspective, how could we explain whether and how the various types of COVID-19 related social media use affects well-being?
 According to the set-point model of subjective well-being, well-being is surprisingly stable (Lykken, 1999).
@@
 Empirically, social media use on average does not have a strong effect on well-being (Meier \& Reinecke, 2020).
-According to the Differential Susceptibility to Media Effects Model (Valkenburg \& Peter, 2013), the effects of media use differ across individuals \textbf{and content}.
+According to the Differential Susceptibility to Media Effects Model (Valkenburg \& Peter, 2013), the effects of media use differ across individuals \textbf{and types of content}.
 Whereas for some media are beneficial, for others they are harmful.
-\textbf{Whereas some content provides opportunities (education, advice), other content creates risks (misinformation, hate). (Livingstone et al., 2018).}
+\textbf{Whereas some content mostly provides opportunities (education, advice), other content rather creates risks (misinformation, hate) (Livingstone et al., 2018).}
 Social media can impair well-being when causing embarrassment, stress, or disinformation, and they can improve well-being when providing connectedness, information, or entertainment (Büchi, 2021).
@@
 
-\textbf{Two of the most prominent media effect theories argue (indirectly) against strong average negative impacts.}
-\textbf{According to mood management theory (Zillmann, 1988), using media substantially affects people's moods.}
-\textbf{Effects can be stimulating or overwhelming, relaxing or boring.}
+\textbf{Two prominent media effect theories argue (mostly implicitly) against strong average negative effects.}
+\textbf{According to mood management theory (Zillmann, 1988), using media can substantially affect people's moods.}
+\textbf{Use can be stimulating or overwhelming, relaxing or boring.}
 \textbf{After some time, users implicitly learn what media help them balance their mood and affect according to their own situational needs (Zillmann, 1988).}
-\textbf{Those media that eventually become part of one's media repertoire are hence, on average, beneficial for users and their moods.}
+\textbf{Those media that eventually become part of one's media repertoire are hence, on average, beneficial for users to regulate their mood.}
 \textbf{Using experience sampling of well-being and logs of social media use, a study with 82 participants from Italy found that after episodes of social media use, levels of positive affect increased significantly (Marciano et al., 2022).}
@@
 \textbf{They help find relevant information, maintain and foster relationships, express one's personality, and entertain oneself (Pelletier et al., 2020).}
-\textbf{In conclusion, because people spend so much time on social media consuming COVID-19 related content, according to both mood management theory and uses and gratifications theory we wouldn't expect to find strong average negative effects.}
-
-But people can also misjudge media affects and are often overly optimistic (Metzger \& Suh, 2017).
-And precisely because social media have so many positive consequences, one can ask if this is not where the actual problem lies.
-In other words, social media aren't problematic because they're inherently bad, but rather because they're too good.
-And as with many other things, there can be too much of a good thing.
-It is therefore often asked whether social media are addictive, and users sometimes express this fear themselves (Yang et al., 2021).
+\textbf{In conclusion, because people spend so much time on social media consuming COVID-19 related content, according to both mood management theory and uses and gratifications theory this is indirect proof that average effects on well-being are likely not particularly negative.}
+
+But people can also misjudge media effects and are often overly optimistic (Metzger \& Suh, 2017).
+Precisely because social media have so many positive consequences, one can ask if this is not where the actual problem lies.
+In other words, social media might not problematic because they are inherently bad, but rather because they are too good.
+As with many other things, there can be too much of a good thing.
+It is therefore often asked whether social media can become addictive, and users sometimes express this fear themselves (Yang et al., 2021).
 However, a recently published meta-analysis found that the two most prominent measures of addiction, the Bergen Facebook Addiction Scale and the Bergen Social Media Addiction Scale, have only small relations to well-being (Duradoni et al., 2020).
@@
 That said, because not all news was negative, and because many people showed solidarity and compassion, there was also positive and uplifting content, potentially compensating for the negative effects (Dörnemann et al., 2021).
-A study with 2.057 respondents from Italy reported that during the pandemic virtual community and social connectedness even increased during the pandemic (Guazzini et al., 2022).
-\textbf{In Finland, in a sample of 735 participants, levels of loneliness did not decrease during the pandemic, and people who engaged more on social media showed less loneliness (Latikka et al., 2022).}
-
-There could also be \emph{indirect} effects.
-When browsing social media for Covid-19 related news, many users reported being captivated to such an extent that they could not stop using social media (Klein, 2021).
+A study with 2.057 respondents from Italy reported that during the pandemic virtual community and social connectedness even increased (Guazzini et al., 2022).
+\textbf{In a study with 735 participants from Finland, levels of loneliness did not decrease during the pandemic, and people who engaged more on social media experienced less loneliness (Latikka et al., 2022).}
+
+Second, there could also be \emph{indirect} effects.
+When browsing social media for COVID-19 related news, many users reported being captivated to such an extent that they could not stop using social media (Klein, 2021).
 During the pandemic social media use was at an all-time high in the US (Statista, 2021).
@@
 During the pandemic, several users shared how they successfully dealt with challenges such as physical distancing.
-In a study with 1131 residents from Wuhan in China (Yue et al., 2022), people who spent more time in quarantine also spent more time on social media.
+In a study with 1,131 residents from Wuhan in China (Yue et al., 2022), people who spent more time in quarantine also spent more time on social media.
 Those, who spent more time on social media also engaged in more upward social comparison, which was related to increased levels of stress.}
@@
 
-To summarize, from a theoretical perspective it is most likely that the average effects of social media use on well-being are negligible.
-Building on established theories from Communication, we would not assume that effects are either profoundly negative or strongly positive.
+To summarize, it seems that from a theoretical perspective it is most likely that the average effects of social media use on well-being are negligible.
+Building on established theories from Communication and current empirical findings, we would not assume that effects are either profoundly negative or strongly positive.
 
@@
 
-So far, there is only little empirical research on how well-being is affected by COVID-19 related social media.
+So far, there is still comparatively little empirical research on how well-being is affected by COVID-19 related social media use.
 In their study on the relations between media use and mental health during the pandemic, Bendau et al.
@@
 analyzed the effects of general digital technology use---objectively measured via screenshots of screen-time applications---on mental health, separating within- and between-person relations (Sewall et al., 2021).
-The results showed that digital technology did not have any significant effects on mental health (for a similar study with comparable results, see Bradley \& Howard, 2021).
+The results showed that digital technology did not have significant effects on mental health (for a similar study with comparable results, see Bradley \& Howard, 2021).
 Together, the literature is mixed, with a slight focus on the negative effects of social media as news use (see also Dörnemann et al., 2021; Liu \& Tong, 2020; Riehm et al., 2020).
-However, note that all of these findings represent between-person relations stemming from cross-sectional data (see below).
-We therefore don't know whether the differences in mental health and well-being are due to social media use or due to other third variables, such as age, health, employment, or education.
 
@@
 This also holds true for the different types of communication such as active or passive use.
-A meta review (i.e., an analysis of meta-analyses) found that the relation between social media use and well-being is likely in the negative spectrum but very small (Meier \& Reinecke, 2020)---potentially too small to matter.
-What determines whether or not an effect is considered small or trivial?
+A meta review (i.e., an analysis of meta-analyses) found that the relation between social media use and well-being is likely in the negative spectrum but very small, potentially too small to matter (Meier \& Reinecke, 2020).
+What determines whether or not an effect should be considered small or trivial?
 As a starting point, we could refer to standardized effect sizes.
@@
 
-Finally, also several individual studies employing advanced methods found smalls relations between social media use and well-being (Keresteš \& Štulhofer, 2020; Orben et al., 2019; Przybylski et al., 2021; Schemer et al., 2021).
+Finally, several individual studies employing advanced methods found smalls relations between social media use and well-being (Keresteš \& Štulhofer, 2020; Orben et al., 2019; Przybylski et al., 2021; Schemer et al., 2021).
 For example, Beyens et al.
-(2021) reported that although for some users (roughly one quarter) the effects of social media use on well-being were negative, for almost the same number of users they were positive, while for the rest the effects were neutral.
-This finding is aligned the Differential Susceptibility to Media Effects Model:
+(2021) reported that although for roughly one quarter of all users the effects of social media use on well-being were negative, for almost the same number of users they were positive, while for the rest the effects were neutral.
+This finding is aligned with the Differential Susceptibility to Media Effects Model:
 Although there is substantial \emph{variation} of media effects for individual users, the \emph{average} effects reported in the literature are often small (Valkenburg \& Peter, 2013).
@@
 In conclusion, in light of the theoretical considerations and empirical studies presented above, I expect that COVID-19 related communication on social media doesn't affect well-being in a meaningful or relevant way.
-\textbf{This general hypothesis will be analyzed specifically for the communication types of (a) time spent reading, (b) liking and sharing, and (c) actively posting COVID-19 related content.}
-\textbf{In addition, I will analyze how well-being is influenced by spending time on five prominent social media apps, including (a) Facebook, (b) Instagram, (c) Twitter, (d) WhatsApp, and (e) YouTube.}
-\textbf{Three different well-being indicators will be differentiated: life satisfaction, positive affect, and negative affect.}
 
@@
 \end{quote}
+
+\textbf{In this study, this general hypothesis will be analyzed specifically for the three communication types of (a) time spent reading, (b) liking and sharing, and (c) actively posting COVID-19 related content.}
+\textbf{In addition, I will analyze how well-being is influenced by spending time on five prominent social media apps, including (a) Facebook, (b) Instagram, (c) Twitter, (d) WhatsApp, and (e) YouTube.}
+\textbf{Three different well-being indicators will be distinguished: life satisfaction, positive affect, and negative affect.}
 
@@
 Because this is a normative and ultimately philosophical question, there can never be a clear, single, or unanimous answer.
-However, it is still necessary and helpful to try provide such a plausible benchmark.
+However, it is still necessary and helpful to try to provide such a plausible benchmark.
 I therefore suggest the following SESOI for this research question:
@@
 But what's a noticeable change in well-being?
-According to Norman et al. (2003), people can reliably distinguish \emph{seven} levels of satisfaction with health.
+According to Norman et al. (2003), people can reliably distinguish seven levels of satisfaction with health.
 So if satisfaction is measured on a 7-point scale, we would state that a four unit change in social media use should result in a one unit change in life satisfaction.
@@
 \textbf{Importantly, when analyzing longitudinal (within-person) relationships and effects, it is important to control for \emph{varying} third variables.}
-\textbf{Non-varying third variables only help control non-varying (between-person) relations.}
+\textbf{Non-varying third variables can only help control non-varying (between-person) relations.}
 
@@
 
-Thus, only when controlling for \emph{all} relevant confounders, we can correctly estimate causality without bias (Rohrer, 2018).
+Thus, only when controlling for \emph{all} relevant confounders, can we correctly estimate causality without bias (Rohrer, 2018).
 Obviously, we can never be entirely sure to have included all confounders, which makes absolute statements regarding causality virtually impossible.
@@
 We can also control for variables that affect only social media use or well-being.
-However, in doing so not much is gained or lost, because the effects of social media use would remain virtually the same (Kline, 2016; but see McElreath, 2021).
+However, in doing so not much is gained, because the effects of social media use would remain virtually the same (Kline, 2016; but see McElreath, 2021).
 
@@
 Next to including covariates, it's now increasingly understood that causal effects should be analyzed from an internal, within-person perspective (Hamaker, 2014).
-If a specific person changes their media diet, we need to measure how this behavior affects \emph{their own} well-being.
+If a specific person changes their media diet, we need to measure how this behavior affects their own well-being.
 Between-person comparisons from cross-sectional data, where participants are interviewed only once, cannot provide such insights.
@@
 For example, if we want to understand the effect of alcohol consumption on driving performance, it makes a big difference if driving performance is measured one minute, one hour, one day, or one week after consumption.
-If variables are stable, longer intervals are needed; if they fluctuate, shorter intervals.
-In the case of well-being, we need shorter intervals for \textbf{the more fluctuating} positive and negative affect and longer ones for \textbf{the more stable} life satisfaction (Dienlin \& Johannes, 2020).
+\textbf{If variables are stable, longer intervals are needed; if they fluctuate, shorter intervals.}
+\textbf{In the case of well-being, we need shorter intervals for the more fluctuating positive and negative affect, and longer ones for the more stable life satisfaction (Dienlin \& Johannes, 2020).}
 \textbf{Using social media can have instant effects on mood (Marciano et al., 2022).}
-\textbf{Effects on life satisfaction often take longer to manifest, for example because media use leads to actual changes in specific behaviors, which then affect life satisfaction (Dienlin et al., 2017).}
+\textbf{Effects on life satisfaction often take longer to manifest, for example because media use leads to actual changes in specific behaviors, which then in turn affect life satisfaction (Dienlin et al., 2017).}
 Choosing the right interval is challenging, because especially short intervals are hard to implement in practice, often requiring advanced methods such as experience sampling (also known as in situ measurement or ambulant assessment) (Schnauber-Stockmann \& Karnowski, 2020).
 In this study, I hence analyze how using social media during the last week affected positive and negative affect during the same week.
-In other words, if people during the last week engaged in more COVID-19 related social media use than the usually do, did they feel better or worse during that week than they usually do?
+In other words, if people during the last week engaged in more COVID-19 related social media use than they usually do, did they feel better or worse during that week than they usually do?
 Regarding life satisfaction, I implemented a longer interval.
-If people during the last week used COVID-19 related social media more than the usually do, were they at the end of the week more or less satisfied with their lives than they usually are?
+If people during the last week used COVID-19 related social media more than they usually do, were they at the end of the week more or less satisfied with their lives than they usually are?
 I hence analyze if when a person changes their social media diet, are there (a) \emph{simultaneous} changes in their affect and (b) \emph{subsequent} changes in their life satisfaction?
@@
 \textbf{The data are hosted on AUSSDA and are publicly available here: \url{https://doi.org/10.11587/28KQNS}.}
-\textbf{At the time of writing, the official website included a data-set consisting of 24 waves. }
-\textbf{For the analyses presented here, I was able to receive an advance version consisting of all 32 waves.}
-\textbf{The study was conducted between March 2020 and June 2022, and data collection is now officially finished.}
-\textbf{It contains 32 waves.}
-Between March 2020 and July 2020, the intervals were weekly, and afterward the intervals were monthly.
+\textbf{At the time of writing, the official website featured a data-set consisting of 24 waves.}
+\textbf{For the analyses presented here, I received an advance version consisting of all 32 waves.}
+\textbf{The study was conducted between March 2020 and June 2022, and data collection is now finished.}
+Between March 2020 and July 2020, the intervals between waves were weekly, and afterward the intervals were monthly.
 Each wave consists of at least 1,500 respondents.
@@
 Because the data were analyzed post-hoc, no a-priori sample size planning on the basis of power analyses was conducted.
-The sample is large, and it is hence well-equipped to reliably detect also small effects.
+The sample is large, and it is hence well-equipped to reliably detect small effects.
 In addition, because such large samples easily generate significant \emph{p}-values even for very small effects, it helps that the hypotheses were tested with a smallest effect size of interest-approach.
 To this end, I adopted the interval testing approach as proposed by Dienes (2014).
-On the basis of the SESOI, I then defined a null region.
+On the basis of the SESOI, I defined a null region.
 In what follows, I explain how I determined the SESOI and the null region.
@@
 Let's briefly illustrate what this means in practice.
-If the 95\% confidence interval falls completely within the null-region (e.g., \emph{b} = .20, {[}95\% CI: .15, .25{]}), the hypothesis that the effect is trivial is supported.
-If the confidence interval and the null region overlap (e.g., \emph{b} = .30, {[}95\% CI: .25, .35{]}), the hypothesis is not supported and the results are considered inconclusive, while a meaningful negative effect is rejected.
-If the confidence interval falls completely outside of the null-region (e.g., \emph{b} = .40, {[}95\% CI: .35, .45{]}), the hypothesis is rejected and the existence of a meaningful positive effect is supported.
+If the 95\% confidence interval falls completely within the null-region (e.g., \emph{b} = -.02, {[}95\% CI: -.12, .08{]}), the hypothesis that the effect is trivial is supported.
+If the confidence interval and the null region overlap (e.g., \emph{b} = -.22, {[}95\% CI: -.27, -.17{]}), the hypothesis is not supported and the results are considered inconclusive, while a meaningful negative effect is rejected.
+If the confidence interval falls completely outside of the null-region (e.g., \emph{b} = -.40, {[}95\% CI: -.45, -.35{]}), the hypothesis is rejected and the existence of a meaningful positive effect is supported.
 For an illustration, see Figure \ref{fig:sesoi}).
@@
 Mean scores were used for positive and negative affect.
-\textbf{Missing responses were imputed using multiple imputation with predictive mean matching (five iterations, five data-sets).}
-\textbf{Predictive mean matching allowed to impute also categorical variables.}
+\textbf{Missing responses were imputed using multiple imputation with predictive mean matching (five iterations, five data-sets), including categorical variables.}
 \textbf{All variables were imputed except the media use measures, as they were not collected on each wave.}
@@
 Social media use was measured for all participants on waves 1, 2, 8, 17, 23, \textbf{and 28}.
-Freshly recruited respondents always answered all questions on social media use.
+Freshly recruited respondents always answered all questions on COVID 19-related social media use.
 
@@
 The effects of COVID-19 related social media use were controlled for the following stable variables:
-(a) gender (female, male, diverse), (b) age, (c) education (ten options), (d) Austria country of birth (yes/no), (e) Austria parents' country of birth (no parent, one parent, both parents).
+(a) gender (female, male, diverse), (b) age, (c) education (ten options), (d) Austria country of birth (yes/no), (e) Austria parents' country of birth (no parent, one parent, both parents), (f) household size, (g) work hours per week, (h) home office, and (i) household income.
 I originally planned to implement additional variables as varying covariates.
 However, because they were not measured often enough or not at the time when social media use was measured, I implemented them as stable variables using their average values across all waves.
-This includes (a) text-based media news consumption (five degrees), (b) video-based media news consumption (five degrees), (c) residency is Vienna (yes/no), (d) self-reported physical health (five degrees), (e) living space (eleven options), (f) access to balcony (yes/no), (g) access to garden (yes/no), (h) employment (nine options), (i) disposition to take risks (eleven degrees), and (j) locus of control (five degrees).
-I also controlled for the following varying covariates: (a) five items measuring outdoor activities such as sport or meeting friends (five degrees), and (b) satisfaction with democracy (five degrees).
-Because it lead to too much attrition in the sample, I did not control for (a) household size, (b) work hours per week, (c) home office, (d) household income.
+This includes (a) text-based media news consumption (five degrees), (b) video-based media news consumption (five degrees), (c) residency is Vienna (yes/no), (d) living space (eleven options), (e) access to balcony (yes/no), (f) access to garden (yes/no), (g) employment (nine options), (h) disposition to take risks (eleven degrees), and (i) locus of control (five degrees).
+I also controlled for the following varying covariates: (a) five items measuring outdoor activities such as sport or meeting friends (five degrees), (b) satisfaction with democracy (five degrees), (c) self-reported physical health (five degrees), and (d) whether participants contracted COVID-19 since the last wave.
 
@@
 The study's main hypothesis was that the effects of social media use on well-being would be trivial.
-Regarding the effects of different communication \emph{types}---that is, reading vs.~sharing vs.~posting---all within-person effects fell completely within the a-priori defined null region (see Figure \ref{fig:fig-between}).
-For example, respondents who used social media more frequently than usual to read about COVID-19 related topics did not show a simultaneous change in life satisfaction (\emph{b} = 0.04 {[}95\% CI -0.01, 0.09{]}).
+Regarding the effects of different communication types---that is, reading vs.~sharing vs.~posting---all within-person effects fell completely within the a-priori defined null region (see Figure \ref{fig:fig-between}).
+For example, respondents who used social media more frequently than usual to like or share COVID-19 related content did not show a simultaneous change in negative affect (\emph{b} = 0.01 {[}95\% CI -0.05, 0.07{]}).
 As a result, the hypothesis was supported for all COVID-19 related types of social media communication.
+
 \textbf{However, two effects were statistically significantly different from zero.}
-\textbf{Users who wrote more COVID-19 related posts than usual were also slightly less satisfied with their lives as usual (\emph{b} = -0.13 {[}95\% CI -0.21, -0.05{]}).}
+\textbf{Users who wrote more COVID-19 related posts than usual were also slightly less satisfied with their lives than usual (\emph{b} = -0.13 {[}95\% CI -0.21, -0.05{]}).}
 \textbf{Users who wrote more COVID-19 related posts than usual also experienced slightly more negative affect than usual (\emph{b} = 0.03 {[}95\% CI 0.01, 0.05{]}).}
-There was a small (statistically nonsignificant) trend that reading COVID-19 related content slightly increased life satisfaction (\emph{b} = 0.04 {[}95\% CI -0.01, 0.09{]}, \emph{p} = .078).
-At the same time, there was also a small (statistically nonsignificant) trend that reading COVID-19 related content decreased positive affect (\emph{b} = -0.02 {[}95\% CI -0.03, 0{]}, \emph{p} = .078).
-
-Regarding the COVID-19 related use of social media \emph{channels}, the results were comparable (see Figure \ref{fig:fig-within}).
+\textbf{There was also a small and statistically non-significant trend that reading COVID-19 related content slightly increased life satisfaction (\emph{b} = 0.04 {[}95\% CI -0.01, 0.09{]}, \emph{p} = .078).}
+\textbf{At the same time, there was also a small and statistically non-significant trend that reading COVID-19 related content decreased positive affect (\emph{b} = -0.02 {[}95\% CI -0.03, 0{]}, \emph{p} = .078).}
+
+Regarding the COVID-19 related use of social media channels, the results were comparable (see Figure \ref{fig:fig-within}).
 Changes in the frequency of using different social media channels to attain information regarding COVID-19 were unrelated to meaningful changes in well-being.
@@
 In sum, the hypothesis was supported also for the COVID-19 related use of important social media channels.
-\textbf{However, two effects differed substantially from zero.}
-\textbf{Respondents who used Instagram more frequently than usual to attain COVID-19 related news reported slightly \emph{lower} levels of negative affect than usual (\emph{b} = -0.02 {[}95\% CI -0.03, \textgreater{} -0.01{]}).}
-\textbf{Respondents who used YouTube more frequently than usual to attain COVID-19 related news reported slightly \emph{higher} levels of negative affect than usual (\emph{b} = 0.02 {[}95\% CI 0, 0.03{]}).}
-\textbf{However, both effects were still completely inside of the null region, hence not large enough to be considered meaningful.}
-
-For an overview of all within-person effects, see Table \ref{tab:tab-within} and Figure \ref{fig:fig-within}.
+
+\textbf{That said, two effects differed substantially from zero.}
+\textbf{Respondents who used Instagram more frequently than usual to attain COVID-19 related news reported slightly lower levels of negative affect than usual (\emph{b} = -0.02 {[}95\% CI -0.03, \textgreater{} -0.01{]}).}
+\textbf{Respondents who used YouTube more frequently than usual to attain COVID-19 related news reported slightly higher levels of negative affect than usual (\emph{b} = 0.02 {[}95\% CI \textless{} 0.01, 0.03{]}).}
+\textbf{However, both effects were still completely inside of the null region, hence likely not large enough to be considered meaningful.}
+
+\textbf{For an overview of all within-person effects, see Table \ref{tab:tab-within} and Figure \ref{fig:fig-within}.}
 
@@
 
-Regarding between-person relations, about which no hypotheses were formulated, again no relation crossed or was completely outside of the SESOI.
+For between-person relations, no a-priori hypotheses were formulated.
+Results showed that no relation crossed or was completely outside of the SESOI.
 Four relations were statistically significant.
-Respondents who across all waves used social media more frequently than others to read about COVID-19 related posts reported slightly lower levels of positive affect than others (\emph{b} = -0.05 {[}95\% CI -0.02, -0.08{]}).
-Respondents who across all waves used social media more frequently than others to write COVID-19 related posts reported higher levels of negative affect than others (\emph{b} = 0.06 {[}95\% CI 0.10, 0.03{]}).
-Interestingly, respondents who across all waves used social media more frequently than others to write COVID-19 related posts also reported higher levels of positive affect (\emph{b} = 0.06 {[}95\% CI 0.11, 0.01{]}).
-Finally, respondents who across all waves used YouTube more frequently than others also reported slightly higher levels of life satisfaction than others (\emph{b} = 0.09 {[}95\% CI 0.16, 0.02{]}).
-However, note that the effect were still completely inside of the null region, hence not large enough to be considered practically relevant.
+\textbf{Respondents who across all waves used social media more frequently than others to read about COVID-19 related posts reported slightly lower levels of positive affect than others (\emph{b} = -0.05 {[}95\% CI -0.08, -0.02{]}).}
+\textbf{Respondents who across all waves used social media more frequently than others to write COVID-19 related posts reported slightly higher levels of negative affect than others (\emph{b} = 0.06 {[}95\% CI 0.03, 0.10{]}).}
+\textbf{At the same time, respondents who across all waves used social media more frequently than others to write COVID-19 related posts also reported slightly higher levels of positive affect (\emph{b} = 0.06 {[}95\% CI 0.01, 0.11{]}).}
+\textbf{Finally, respondents who across all waves used YouTube more frequently than others also reported slightly higher levels of life satisfaction than others (\emph{b} = 0.09 {[}95\% CI 0.02, 0.16{]}).}
 
@@
 Therefore, I report the results of the standardized scales,
-which allows for a better comparison across the differently scaled variables.
+which also allows for a better comparison across the differently scaled variables.
 As a rough estimate for the SESOI we can build on the typical convention that small effects start at \emph{r} = \textbar.10\textbar.
@@
 
-To find out whether my inferences were robust across legitimate (though arguably inferior) alternative analyses, I reran the analyses also using standardized estimates, mean scores instead of factor scores, and with a data set where missing data were not imputed.
+To find out whether the inferences were robust across plausible (though arguably inferior) alternative analyses, I reran the analyses also using standardized estimates, additional covariates including trust in media or government, single imputation, and with a data set where missing data were not imputed.
 The results were virtually the same.
-For example, all within-person standardized COVID-19 related types of social media use or channels were significantly smaller than \(\beta\) = \textbar.05\textbar, again supporting that effects were tiny.
-The additional analyses are reported in Figure 3 and Figure 4, and in the \href{https://XMtRA.github.io/Austrian_Corona_Panel_Project/analyses_additional.html}{companion website}.
+For example, all within-person standardized COVID-19 related types of social media use or channels were significantly smaller than \(\beta\) = \textbar.05\textbar, again supporting that effects were negligible.
+The results of the standardized analyses are reported in Table \ref{tab:tab-within}.
+The additional analyses are reported on the \href{https://XMtRA.github.io/Austrian_Corona_Panel_Project/analyses_additional.html}{companion website}.
 
@@
 The data come from a panel study with 32 waves and are largely representative of the Austrian population.
-In a random effects model I separated between person relations from within-person effects and controlled for a large number of both stable and varying covariates, thereby aiming to assess causality.
-The results showed that some statistically significant negative within-person effects existed, but that they were \textbf{very small and likely} trivial.
+In a random effects model I separated between person relations from within-person effects.
+I controlled for a large number of both stable and varying covariates, aiming to assess causality.
+\textbf{The results showed that some statistically significant negative within-person effects existed, but that they were very small and likely negligible.}
 People who used social media more than usual to learn about COVID-19 didn't show meaningful changes in their well-being.
@@
 \textbf{Similarly, people who wrote more COVID-19 related posts than usual also reported slightly more negative affect.}
-\textbf{When writing posts and comments on social media, people explicitly and more deeply engage with COVID-19 related content.}
-\textbf{Unfortunately, in general online tonality is often more extreme, negative, or aggressive, which potentially affects their authors.}
-\textbf{Note that I controlled for whether or not participants had an corona infection during a specific wave, which rules out the potential explanation that having an infection was the root cause of increased communication and reduced well-being.}
-
-\textbf{The potential explanation that tonality might be a relevant factor at play here is also supported by the second trend.}
-\textbf{People, who during the pandemic spend more time on Instagram than usual, also experienced less negative affect than usual.}
+\textbf{As a potential explanation, when writing posts and comments on social media people explicitly and more deeply engage with COVID-19 related content.}
+\textbf{The tonality on social media is often extreme, negative, or aggressive, which potentially affects their authors.}
+\textbf{And because I controlled for whether or not participants had a COVID-19 infection during a specific wave, we can rule out the potential explanation that having an infection was the root cause of increased communication and reduced well-being.}
+
+\textbf{The hypothesis that tonality might be a relevant factor at play is also supported by the second trend.}
+\textbf{People, who during the pandemic spent more time on Instagram than usual, also experienced less negative affect than usual.}
 \textbf{Instagram is well-known for its positivity bias (Waterloo et al., 2018).}
-\textbf{Content is generally post positive, uplifting, and (self-)flattering.}
-\textbf{It seems the often-criticized positivity bias on Instagram might have been somewhat beneficial in times of the pandemic.}
+\textbf{Content is generally more positive, uplifting, and (self-)flattering.}
+\textbf{It seems the much maligned positivity bias on Instagram might have been somewhat beneficial in times of the pandemic.}
 \textbf{The critique that the positivity bias necessarily leads to envy and negative feelings is one-sided, because positive content can also inspire and motivate users (Meier et al., 2020), which could be especially helpful in times of lockdown and home-office.}
-\textbf{To provide a concrete example, Instagram was successfully used as an interactive communication channel for first year students to have a better start into their new degree, effectively complementing alternative learning platform tools (Ye et al., 2020).}
+\textbf{To provide a concrete example, during the pandemic Instagram was successfully used as an interactive communication channel for first year students to have a better start into their new degree, effectively complementing alternative learning platform tools (Ye et al., 2020).}
 
@@
 
-On the one hand, the results are not aligned with several recent studies analyzing similar or closely related research questions.
+\textbf{The results showed that it makes sense to analyze different communication types and communication channels, and that active and passive communication showed different results.}
+\textbf{Liking and sharing content did not show any within-person effects. }
+\textbf{Such rather low-key active behaviors do not seem to affect well-being at all.}
+\textbf{Regarding passive use, reading COVID-19 related posts is more ambivalent.}
+\textbf{Results showed some weak trends towards a positive effect on life satisfaction, but a negative effect on mood.}
+\textbf{It might be that reading and informing oneself about COVID-19 on social media is helpful in the long run, but more negative for short-term affect.}
+\textbf{Finally, proactively engaging via writing posts, the most active form of communication analyzed here, showed only negative effects on well-being.}
+\textbf{The results support the findings from Valkenburg et al. (2022), who also could not confirm the claim active use is good and passive use is bad.}
+\textbf{Focusing on communication channels, YouTube seems to be more negative, whereas Instagram is likely more positive.}
+\textbf{Again, these are only very small effects.}
+\textbf{Future research might elaborate on these specific relations to probe their stability and relevance.}
+
+Taken together, on the one hand the results are not aligned with several recent studies analyzing similar or closely related research questions.
 This includes a study by Bendau et al. (2021), which showed negative relations between social media and well-being (but see Bradley \& Howard, 2021; or Sewall et al., 2021).
@@
 On the other hand, the results are well-aligned with \textbf{mood management theory (Zillmann, 1988)} or the uses and gratifications approach (Katz et al., 1973).
-If effects were indeed profoundly negative on average, then people likely wouldn't spend so much time on social media.
-Likewise, recent studies and meta-analyses analyzing the effects of social media use from a more general perspective or from a somewhat different angle.
+If effects were indeed profoundly negative on average, then people likely wouldn't spend so much time on social media engaging with COVID-19 content.
+Likewise, recent studies and meta-analyses analyzed the effects of social media use from a more general perspective or from a somewhat different angle.
 These studies have found that the effects of various types of social media use on several well-being indicators are small at best, often too small to matter (Ferguson et al., 2021; Meier \& Reinecke, 2020; Orben, 2020), which echoes the results obtained here.
 
-\textbf{The results showed that it makes sense to analyze different communication types and communication channels, and that active and passive communication showed different results.}
-\textbf{Liking and sharing content did not show any within-person effects. }
-\textbf{Such rather low-key active behaviors do not seem to affect well-being at all.}
-\textbf{Regarding passive use, reading COVID-19 related posts is more ambivalent; results showed some weak trends towards a positive effect on life satisfaction, but a negative affect on mood.}
-\textbf{It might be that reading and informing oneself about COVID-19 on social media might be helpful in the long term, but more negative for short-term affect.}
-\textbf{Finally, proactively engaging via writing posts, the most active form of communication analyzed here, showed only negative effects on well-being.}
-\textbf{Together, the results support the findings from Valkenburg et al. (2022), who could not confirm the claim that active use is good, versus passive is bad.}
-\textbf{Focusing on communication channels, YouTube seems to be more negative, whereas Instagram is likely more positive.}
-\textbf{Again, these are only very small effects.}
-\textbf{Future research might elaborate on these specific relations to probe their stability and relevance.}
-
-\textbf{From a more political and societal perspective, I believe the results imply that it does makes sense critically to reflect upon COVID-19 related social media use.}
-\textbf{In terms of concrete recommendations, on average it might be slightly beneficial to post less actively about COVID-19 on social media and to spent less time on YouTube.}
-\textbf{Potential positive effects, however, will for many users likely not be noticeable.}
-\textbf{Results allow to reject a positive effect: Writing more posts on social media will likely not increase well-being.}
-\textbf{At all events, engaging in COVID 19-related social media use should, on average, not be a cause for concern.}
+\textbf{From a more political and societal perspective, the results imply that it can make sense to critically reflect upon COVID-19 related social media use.}
+\textbf{On average, it might be slightly beneficial to post less actively about COVID-19 on social media and to spend less time on YouTube.}
+\textbf{Spending time on Instagram seems to be okay.}
+\textbf{The potentially resulting positive effects, however, will for many users likely not be noticeable.}
+\textbf{Results allow us to reject a positive effect: Writing more posts on social media will likely not increase well-being.}
+\textbf{At all events, engaging in COVID 19-related social media use should, on average, not be a major cause for concern.}
 
@@
 Together, this allowed for an improved perspective on assessing causality.
-\textbf{However, the opposite effect is also plausible, namely that well-being affects media use (Zillmann, 1988).}
+\textbf{However, the opposite effect is still also plausible, namely that well-being affected media use (Zillmann, 1988).}
 \textbf{While controlling for potential confounders can support claims of causality, the procedure implemented here cannot prove causality.}
@@
 \textbf{The challenge is that} regarding media use, such effects often happen immediately or shortly after use, necessitating intervals in the hours, minutes, or even seconds.
-In many cases only experience sampling studies asking users in the very moment can produce such knowledge.
+In many cases only experience sampling studies asking users at the very moment can produce such knowledge.
 However, even then we don't know for certain if we actually measured the right interval.
@@
 To borrow the words from Rohrer and Murayama (2021), there is no such thing as ``the'' effect of social media use on well-being.
-Hence, to document how effects unfold, future research needs to employ different study designs probing different time lags.
+Hence, to document how effects unfold, future research needs to employ different study designs probing different intervals.
 In addition, more thought needs to be invested in what relevant stable and varying factors we should include as control variables, and I hope this study provides a first step into this direction.
@@
 Or would it make more sense to expect that thoroughly committing to say \emph{two} activities (e.g.~regularly exercising \emph{and} establishing a reading habit) should then cause a detectable improvement in well-being?
-Practically, this would imply a SESOI half as large as I have defined here, namely \emph{b} = \textbar.15\textbar{} for well-being and \emph{b} = \textbar.075\textbar{} for affect.
-In the case of this study, however, reducing the SESOI would not even make a big difference, as also with these more liberal thresholds all but three effect would still be completely in the null region, and no effect would be outside of the null region.
+Practically, this would imply a SESOI half as large as I have defined here, namely \emph{b} = \textbar.15\textbar{} for life satisfaction and \emph{b} = \textbar.075\textbar{} for affect.
+In the case of this study, however, reducing the SESOI would not even make a big difference, as also with these more liberal thresholds all but two effects would still be completely in the null region, and no effect would be outside of the null region.
 However, at all events I encourage future research to start a thorough conversation on what effect sizes are considered meaningful and what not.
@@
 \leavevmode\vadjust pre{\hypertarget{ref-bradleyStressMoodSmartphone2021}{}}%
-Bradley, A., \& Howard, A. (2021). \emph{Stress, mood, and smartphone use in {University} students: {A} 12-week longitudinal study}. \url{https://doi.org/10.31219/osf.io/frvpb}
+Bradley, A., \& Howard, A. (2021). \emph{Stress, mood, and smartphone use in {University} students: {A} 12-week longitudinal study}. {OSF Preprints}. \url{https://doi.org/10.31219/osf.io/frvpb}
 
@@
 {The} role of similarity}.
+{OSF Preprints}.
 \url{https://doi.org/10.31219/osf.io/sy2kd}
@@
 \leavevmode\vadjust pre{\hypertarget{ref-europeansocialsurveyESS9Edition20182021}{}}%
-European Social Survey. (2021). \emph{{ESS9} edition 3.1 - 2018 {Documentation Report}}. \url{https://www.europeansocialsurvey.org/docs/round9/survey/ESS9_data_documentation_report_e03_1.pdf}
+European Social Survey. (2021). \emph{{ESS9} edition 3.1 - 2018 {Documentation Report}}.
 
@@
 \leavevmode\vadjust pre{\hypertarget{ref-galerHowMuchToo2018}{}}%
-Galer, S. S. (2018, January 19). \emph{How much is {``too much time''} on social media?} \url{https://www.bbc.com/future/article/20180118-how-much-is-too-much-time-on-social-media}
+Galer, S. S. (2018). \emph{How much is {``too much time''} on social media?} https://www.bbc.com/future/article/20180118-how-much-is-too-much-time-on-social-media.
 
@@
 No effect of different types of media on well-being.
-\emph{Scientific Reports}, \emph{12}(1, 1), 61.
+\emph{Scientific Reports}, \emph{12}(1), 61.
 \url{https://doi.org/10.1038/s41598-021-03218-7}
@@
 \leavevmode\vadjust pre{\hypertarget{ref-kleinDarklySoothingCompulsion2021}{}}%
-Klein, J. (2021, March 3). \emph{The darkly soothing compulsion of 'doomscrolling'}. \url{https://www.bbc.com/worklife/article/20210226-the-darkly-soothing-compulsion-of-doomscrolling}
+Klein, J. (2021). \emph{The darkly soothing compulsion of 'doomscrolling'}. https://www.bbc.com/worklife/article/20210226-the-darkly-soothing-compulsion-of-doomscrolling.
 
 \leavevmode\vadjust pre{\hypertarget{ref-klinePrinciplesPracticeStructural2016}{}}%
-Kline, R. B. (2016). \emph{Principles and practice of structural equation modeling} (4th ed.). {The Guilford Press}.
+Kline, R. B. (2016). \emph{Principles and practice of structural equation modeling} (Fourth). {The Guilford Press}.
 
@@
 
+\leavevmode\vadjust pre{\hypertarget{ref-masciantonioDonPutAll2021}{}}%
+Masciantonio, A., Bourguignon, D., Bouchat, P., Balty, M., \& Rimé, B.
+(2021).
+Don't put all social network sites in one basket: {Facebook}, {Instagram}, {Twitter}, {TikTok}, and their relations with well-being during the {COVID-19} pandemic.
+\emph{PloS One}, \emph{16}(3), e0248384.
+\url{https://doi.org/10.1371/journal.pone.0248384}
+
 \leavevmode\vadjust pre{\hypertarget{ref-mcelreathYesterdayClass2021}{}}%
-McElreath, R. (2021, January 28). \emph{Yesterday in class, ...} {[}Tweet{]}. {@rlmcelreath}. \url{https://twitter.com/rlmcelreath/status/1354786005996482563}
+McElreath, R. (2021). Yesterday in class, ... {[}Tweet{]}. In \emph{@rlmcelreath}.
 
@@
 \emph{Medical Care}, \emph{41}(5), 582--592.
-\href{Retrieved\%20from\%20http://www.jstor.org/stable/3768017}{Retrieved from http://www.jstor.org/stable/3768017}
 
@@
 Sandstrom, G., Buchanan, K., Aknin, L., \& Lotun, S.
-(2021, October 22).
-\emph{Doomscrolling {COVID} news takes an emotional toll -- here's how to make your social media a happier place}.
-{The Conversation}.
-\url{http://theconversation.com/doomscrolling-covid-news-takes-an-emotional-toll-heres-how-to-make-your-social-media-a-happier-place-170342}
+(2021).
+Doomscrolling {COVID} news takes an emotional toll \textendash{} here's how to make your social media a happier place.
+In \emph{The Conversation}.
+http://theconversation.com/doomscrolling-covid-news-takes-an-emotional-toll-heres-how-to-make-your-social-media-a-happier-place-170342.
 
@@
 (2016).
-The accuracy of self-reported {Internet} use---{A} validation study using client log data.
+The accuracy of self-reported {Internet} use\textemdash{{A}} validation study using client log data.
 \emph{Communication Methods and Measures}, \emph{10}(1), 13--27.
@@
 \leavevmode\vadjust pre{\hypertarget{ref-sheldonStabilityHappinessTheories2014}{}}%
-Sheldon, K. M., \& Lucas, R. E. (2014). \emph{Stability of happiness: Theories and evidence on whether happiness can change}. \url{http://site.ebrary.com/id/10891875}
+Sheldon, K. M., \& Lucas, R. E. (2014). \emph{Stability of happiness: Theories and evidence on whether happiness can change}.
 
@@
 Statista.
-(2021, May 21).
+(2021).
 \emph{Average daily time spent on social networks by users in the {United States} from 2018 to 2022}.
-\url{https://www.statista.com/statistics/1018324/us-users-daily-social-media-minutes/}
+https://www.statista.com/statistics/1018324/us-users-daily-social-media-minutes/.
 
@@
 \includegraphics{manuscript_files/figure-latex/sesoi-1.pdf}
-\caption{\label{fig:sesoi}Using confidence intervals to test a null region. Note. Here, a trivial effect of social media use on life satisfaction is defined as ranging from b = -.25 to b = .25}
+\caption{\label{fig:sesoi}Using confidence intervals to test a null region. Note. Here, a trivial effect of social media use on life satisfaction is defined as ranging from b = -.30 to b = .30}
 \end{figure}

### 2023-01-30 manuscript.Rmd
- thorough update to paper - shortened introduction - added additional exploratory analyses - e.g., analysis of between person vars

@@
 ---
-title             : "The effects of COVID-19 related social media use on well-being"
-shorttitle        : "COVID-19 related social media use and well-being"
+title             : "COVID-19 related social media use slightly decreases well-being"
+shorttitle        : "COVID-19 related social media use & well-being"
 
@@
 
-abstract          : "In times of crisis such as the COVID-19 pandemic citizens need to stay informed about recent events, political decisions, or mandatory protection measures.
-To this end, many people use various types of media, and increasingly social media.
-However, because social media are particularly engaging, some find it hard to disconnect.
-In this preregistered study, I investigated whether using social media for COVID-19 related reasons affected psychological well-being.
-To answer this question I analyzed data from the Austrian Corona Panel Project, which consists of 3,485 participants.
-Well-being was measured at all 32 waves, and communication at six specific waves.
-I ran three random effects within between models, controlling for several stable and varying confounders.
-Results showed that the effects of COVID-19 related social media use on well-being were very small, arguably too small to matter.
-Fears that social media use during times of crisis critically impairs well-being are not supported."
+abstract          : "In times of crisis such as the COVID-19 pandemic, citizens need to stay informed about recent political events.
+To this end, people increasingly use social media.
+However, being particularly engaging, many find it hard to disconnect from social media, especially during times of crisis.
+Using data from the Austrian Corona Panel Project consisting of 3,485 participants from 32 waves, controlling for several stable and varying confounders, the results showed that COVID-19 related social media use reduced facets of well-being.
+However, the effects were very small.
+Other factors such as health, income, exercise, or internal locus of control showed larger and more meaningful effects."
 keywords          : "COVID-19, well-being, social media, news use, panel study."
@@
 
-floatsintext      : no
+floatsintext      : yes
 endnotes          : yes
@@
 # load packages
-packages <- c("devtools", "english", "ggplot2", "gridExtra", "kableExtra", 
+packages <- c("devtools", "english", "GGally", "ggplot2", "gridExtra", "kableExtra", 
               "lavaan", "lme4", "lmerTest",
-              "magrittr", "papaja", "semPlot", "tidyverse")
+              "magrittr", "papaja", "tidyverse")
 lapply(c(packages, "td"), library, character.only = TRUE)
@@
 # load workspace created from analyses.rmd to use results
-load("data/workspace_1.RData")
+load("data/workspace_2.RData")
 ```
 
-During the COVID-19 pandemic, 
-numerous events unfolded in quick succession and several open questions emerged.
+During the COVID-19 pandemic it was critical to stay informed regarding the latest developments.
 How dangerous is the virus? 
-Is it spreading in my region? 
-How is it transmitted and how can I protect myself?
-Because for many it was (and at the time of writing still is) a matter of life or death, people aimed to stay informed regarding the latest developments.
-Governments around the world implemented safety measures, such as wearing masks, keeping physical distance, or enforcing lockdowns.
-In this extraordinary situation, many people heavily relied on media to obtain relevant information, and especially social media were at an all time high [@statistaAverageDailyTime2021].
+In what region is it spreading? 
+How is it transmitted? What are the current safety regulations?
+To obtain relevant information, many people heavily relied on social media, which were at an all time high [@statistaAverageDailyTime2021].
 
-Some people actually couldn't stop using social media to learn about COVID-19 related news.
-A new phenomenon termed "doomscrolling" emerged:
-Users were glued to their screens and found it hard to pursue other relevant activities such as working, taking a break, or looking after their children [@kleinDarklySoothingCompulsion2021].
-It was increasingly asked whether using social media for COVID-19 related reasons is helpful or whether it creates an additional burden on mental health [@sandstromDoomscrollingCOVIDNews2021].
-These concerns seem justified: 
-A study with 6,233 people from Germany conducted during the pandemic found that "[f]requency, duration and diversity of media exposure were positively associated with more symptoms of depression" [@bendauAssociationsCOVID19Related2021, p.
-283].
-
-As a result, with this study I want to build on this research and investigate whether or not COVID-19 related social media use meaningfully affected well-being during the pandemic.
-To this end, I analyzed a large-scale **standalone** panel study from the Austrian Corona Panel Project [@kittelAustrianCoronaPanel2020]. 
-The panel consists of **32** waves and has an overall sample size of **`r nrow(d_wide) %>% prettyNum(big.mark = ",")` participants**.
-<!-- At least 1,500 participants took part per wave, and the sample is representative of the Austrian population. -->
-The panel study collected a large number of psychological and demographic variables.
-<!-- Being able to control for many confounding third variables, both stable and varying, together with the longitudinal design, this creates a unique opportunity  -->
-I explicitly aimed to investigate the causal effects of COVID-19 related social media use on well-being.
+Some actually could not stop using social media to learn about COVID-19 related news.
+A new phenomenon termed "doomscrolling" emerged.
+Many users were glued to their screens and found it hard to pursue other relevant activities such as working, taking a break, or even looking after their children [@kleinDarklySoothingCompulsion2021].
+In the media it was hence increasingly asked whether using social media for COVID-19 related reasons would, next to all other stressors, create an additional burden on mental health [@sandstromDoomscrollingCOVIDNews2021].
+Although research has begun addressing this question [@bendauAssociationsCOVID19Related2021; @bradleyStressMoodSmartphone2021; @choiMediatedCommunicationMatters2021; @dornemannHowGoodBad2021; @edenMediaCopingCOVID192020; @guazziniSecondWaveAnalysis2022; @latikkaLonelinessPsychologicalDistress2022; @liuRelationOfficialWhatsAppdistributed2020; @riehmAssociationsMediaExposure2020; @sewallObjectivelyMeasuredDigital2021; @stainbackCOVID1924News2020; @yuePassiveSocialMedia2022], it still largely unknown if COVID-19 related social media use during the pandemic has had a meaningful impact on well-being.
+By analyzing large-scale panel data from Austria and by focusing on intra-individual causal effects, this study aims to answer to this question.
 
@@
 
-**Two underlying theories guided the selection of variables for this study, namely the two-continua model of mental health [@greenspoonIntegrationSubjectiveWellbeing2001] and the hierarchical taxonomy of computer-mediated communication [@meierComputermediatedCommunicationSocial2020a].**
-According to the two-continua model, mental health consists of (a) psychopathology and (b) well-being.
-<!-- Because the aim of this study is better to understand typical users and everyday contexts, my focus will be on well-being. -->
-Well-being can be differentiated into subjective and psychological well-being [@dienerAdvancesOpenQuestions2018].
-Whereas subjective well-being emphasizes hedonic aspects such as happiness and joy, psychological well-being addresses eudaimonic aspects such as fulfillment and meaning.
-Subjective well-being is primarily about achieving positive and avoiding negative affect.
-One of the most prominent indicators of well-being is life satisfaction.
-In my view, because it represents a general appraisal of one's life, life satisfaction is best thought of as a meta concept combining psychological and subjective well-being.
-Notably, life satisfaction is stable and fluctuates only little, whereas it's the exact opposite for affect [@dienlinImpactDigitalTechnology2020].
-<!-- Everyone can feel bad on any given day, but whether or not one feels generally comfortable in one's life rather changes across years. -->
-To capture well-being in this study I thus build on life satisfaction, positive affect, and negative affect.
-Together, this should provide an encompassing perspective on potential media effects.
+<!-- Two underlying theories guided the selection of variables for this study, namely the two-continua model of mental health [@greenspoonIntegrationSubjectiveWellbeing2001] and the hierarchical taxonomy of computer-mediated communication [@meierComputermediatedCommunicationSocial2020a].
+-->
+<!-- According to the two-continua model, mental health consists of (a) psychopathology and (b) well-being. -->
+<!-- <!-- Because the aim of this study is better to understand typical users and everyday contexts, my focus will be on well-being. -->
+<!-- Well-being can be differentiated into subjective and psychological well-being [@dienerAdvancesOpenQuestions2018]. -->
+<!-- Whereas subjective well-being emphasizes hedonic aspects such as happiness and joy, psychological well-being addresses eudaimonic aspects such as fulfillment and meaning. -->
+<!-- Subjective well-being is primarily about achieving positive and avoiding negative affect. -->
+<!-- One of the most prominent indicators of well-being is life satisfaction. -->
+<!-- In my view, because it represents a general appraisal of one's life, life satisfaction is best thought of as a meta concept combining psychological and subjective well-being. -->
+<!-- Notably, life satisfaction is stable and fluctuates only little, whereas it's the exact opposite for affect [@dienlinImpactDigitalTechnology2020]. -->
+<!-- <!-- Everyone can feel bad on any given day, but whether or not one feels generally comfortable in one's life rather changes across years. -->
+<!-- To capture well-being in this study I thus build on life satisfaction, positive affect, and negative affect. -->
+<!-- Together, this should provide an encompassing perspective on potential media effects. -->
 
-The hierarchical taxonomy of computer-mediated communication differentiates six levels of how people engage with digital technology.
-First, the device (e.g., smartphone); second, the type of application (e.g., social networking site); third, the branded application (e.g., Twitter); fourth, the feature (e.g., status post); fifth, the interaction (e.g., one-to-many); and sixth, the message (e.g., content) [@meierComputermediatedCommunicationSocial2020a].
-Whereas the first four levels focus on the communication _channel_, the last two address the communication _type_.
-**Distinguishing different communication channels and communication types is important, because the effects on well-being likely differ across communication channels and communication types.**
-**Whereas active social media use such as chatting is routinely linked to improved well-being, passive use such as reading is often considered more negative [@dienlinImpactDigitalTechnology2020].**
-**Similarly, branded apps are separate communication entities with potentially divergent effects and affordances.**
-**For example, @waterlooNormsOnlineExpressions2018 found that it's more adequate to express negative emotions on WhatsApp than on Twitter or on Instagram.**
-**Especially during a pandemic, it makes sense to analyze if users engage with COVID-19 related content on Instagram, where communication is more positive, or on Facebook, where communication is more critical.**
-**First studies suggest that during the pandemic Instagram use was indeed more beneficial for well-being than Facebook use [@masciantonioDonPutAll2021].**
-**In this study, to measure the effects of social media use focused on COVID-19 related news and topics, I adopt both the channel and the type of communication perspective.**
-<!-- **Together, this should offer a nuanced and comprehensive understanding of communication.** -->
+<!-- The hierarchical taxonomy of computer-mediated communication differentiates six levels of how people engage with digital technology. -->
+<!-- First, the device (e.g., smartphone); second, the type of application (e.g., social networking site); third, the branded application (e.g., Twitter); fourth, the feature (e.g., status post); fifth, the interaction (e.g., one-to-many); and sixth, the message (e.g., content) [@meierComputermediatedCommunicationSocial2020a].
+-->
+<!-- Whereas the first four levels focus on the communication _channel_, the last two address the communication _type_. -->
+<!-- Distinguishing different communication channels and communication types is important, because the effects on well-being likely differ across communication channels and communication types. -->
+<!-- Whereas active social media use such as chatting is routinely linked to improved well-being, passive use such as reading is often considered more negative [@dienlinImpactDigitalTechnology2020].
+-->
+<!-- Similarly, branded apps are separate communication entities with potentially divergent effects and affordances. -->
+<!-- For example, @waterlooNormsOnlineExpressions2018 found that it's more adequate to express negative emotions on WhatsApp than on Twitter or on Instagram. -->
+<!-- Especially during a pandemic, it makes sense to analyze if users engage with COVID-19 related content on Instagram, where communication is more positive, or on Facebook, where communication is more critical.
+-->
+<!-- First studies suggest that during the pandemic Instagram use was indeed more beneficial for well-being than Facebook use [@masciantonioDonPutAll2021]. -->
+<!-- In this study, to measure the effects of social media use focused on COVID-19 related news and topics, I adopt both the channel and the type of communication perspective. -->
+<!-- Together, this should offer a nuanced and comprehensive understanding of communication. -->
 
-Specifically, I investigate how well-being is affected by different types of communication, namely active and passive use.
-Defining what constitutes active and what passive use is not always clear, and different understandings are currently discussed [@ellisonWhyWeDon2020; @meierComputermediatedCommunicationSocial2020a].
-Reading is generally considered as passive and writing as active use, while there are also specific behaviors falling somewhere in-between such as liking or sharing content [@meierDoesPassiveSocial2022].
-**In this study, I hence distinguish (a) reading (passive), (b) posting (active), and (c) liking and sharing COVID-19 related posts (both active and passive).**
-**Second, I analyze how using the most prominent branded applications affects well-being, and whether this effect changes across applications.**
-**The branded applications investigated here are Facebook, Twitter, Instagram, WhatsApp, and YouTube---which were, at the time of writing, the most relevant social media apps.**
+This study investigates how different _facets_ of well-being are affected by different _types_ and different _channels_ of communication [@meierComputermediatedCommunicationSocial2020a].
+<!-- , namely active and passive use [@ellisonWhyWeDon2020; @meierComputermediatedCommunicationSocial2020a]. -->
+<!-- Defining what constitutes active and what passive use is not always clear, and different understandings are currently discussed [@ellisonWhyWeDon2020; @meierComputermediatedCommunicationSocial2020a].
+-->
+<!-- Reading is generally considered as passive and writing as active use, while there are also specific behaviors falling somewhere in-between such as liking or sharing content [@meierDoesPassiveSocial2022].
+-->
+<!-- In this study,  -->
+Building on the typology of subjective well-being [@dienerAdvancesOpenQuestions2018], three different well-being facets are analyzed: life satisfaction, positive affect, and negative affect.
+Because effects of social media depend on how they are used [@verduynPassiveFacebookUsage2015], I further distinguish three types of use and five popular channels. 
+The types of use include reading, posting, and liking and sharing COVID-19 related posts.
+In doing so, this study analyzes social media use focused on COVID-19 related content, which includes posting thoughts about the pandemic, reading posts and comments, or retweeting COVID-19 related news.
+The five channels to be investigated are Facebook, Twitter, Instagram, WhatsApp, and YouTube, which at the time ranked among the most popular social media services in Austria.
 
-**Worth noting, this study is not about _general_ social media use during times of COVID, but on social media use _focused_ on COVID-19 related content.** 
-**This, for example, includes posting thoughts about the pandemic, reading posts and comments, or retweeting COVID-19 related news.**
+<!-- ## Theorizing Social Media Effects on Well-Being -->
 
-## Theorizing Social Media Effects on Well-Being
+<!-- <!-- In conclusion, most effects are likely somewhere between trivial and small. -->
+<!-- <!-- I therefore expect that also in the case of COVID-19 related social media use effects will be trivial to small. -->
 
-<!-- In conclusion, most effects are likely somewhere between trivial and small. -->
-<!-- I therefore expect that also in the case of COVID-19 related social media use effects will be trivial to small. -->
-
-From a theoretical perspective, how could we explain whether and how the various types of COVID-19 related social media use affects well-being?
+How could the various types and channels of COVID-19 related social media use affect well-being?
 <!-- In what follows, I outline potential arguments as to why the effect might be positive or negative, direct or indirect, or nonexistent. -->
-According to the set-point model of subjective well-being, well-being is surprisingly stable [@lykkenHappinessWhatStudies1999].
-Although specific events such as marriage or salary can have significant effects, after some time well-being routinely returns to prior levels, which are mostly determined genetically [@sheldonStabilityHappinessTheories2014].
-Only very specific events and factors such as unemployment, disability, or death can cause long-term changes in well-being [@lucasAdaptationSetpointModel2007].
+According to the set-point theory, well-being is surprisingly stable [@lykkenHappinessWhatStudies1999].
+Although specific events such as marriage or salary can have significant impacts on well-being, in most cases effects are only short-term, with well-being after some time routinely returning to prior levels [@sheldonStabilityHappinessTheories2014].
+Only very specific factors such as unemployment, disability, or death can cause long-term changes in well-being [@lucasAdaptationSetpointModel2007].
 <!-- So although well-being can change this does not happen easily. -->
 <!-- Although the set-point theory has been criticized lately because events such as marriage, unemployment, or death can indeed lead to long-term changes, it is still widely maintained that well-being is very stable and hence hard to be influenced  -->
+Can media use be such a factor?
+Current literature overviews suggest no: 
+Although social media use on average decreases well-being, the effects are small [@meierComputermediatedCommunicationSocial2020a].
+In addition, the effects of media use differ across individuals and types of content [@valkenburgDifferentialSusceptibilityMedia2013].
+Whereas for some users social media are more beneficial, for others they are more harmful.
+For example, in one study it was estimated that roughly one quarter of all users experienced negative effects, another quarter positive effects, while for the rest the effects were neutral [@beyensSocialMediaUse2021].
+This finding is aligned with the Differential Susceptibility to Media Effects Model: 
+<!-- This finding is aligned with the Differential Susceptibility to Media Effects Model:  -->
+Although there is substantial _variation_ of media effects for individual users, there are both negative and positive aspects, leading to small effects on average [@valkenburgDifferentialSusceptibilityMedia2013]---as confirmed by several literature overviews [@dienlinImpactDigitalTechnology2020; @huangTimeSpentSocial2017; @meierComputermediatedCommunicationSocial2020a; @orbenTeenagersScreensSocial2020].
 
-Can media use be such a factor?
-In advance, there doesn't seem to be a clear winner, and it's likely that both positive and negative effects cancel each other out.
-Empirically, social media use on average does not have a strong effect on well-being [@meierComputermediatedCommunicationSocial2020a].
-According to the Differential Susceptibility to Media Effects Model [@valkenburgDifferentialSusceptibilityMedia2013], the effects of media use differ across individuals **and types of content**.
-Whereas for some media are beneficial, for others they are harmful.
-**Whereas some content mostly provides opportunities (education, advice), other content rather creates risks (misinformation, hate) [@livingstoneEuropeanResearchChildren2018].**
-Social media can impair well-being when causing embarrassment, stress, or disinformation, and they can improve well-being when providing connectedness, information, or entertainment [@buchiDigitalWellbeingTheory2021].
-On average, however, effects are often small or negligible.
+Why are the effects of social media use on well-being small on average?
+Using Social media is ambivalent.
+It can impair well-being when causing embarrassment, stress, or disinformation, and it can improve well-being when providing connectedness, information, or entertainment [@buchiDigitalWellbeingTheory2021].
+Although social media are often associated with negative outcomes, there are a large number of positive effects, which either offset or at least significantly reduce negative outcomes.
+Positive outcomes encompass finding relevant information; maintaining and fostering relationships; expressing one's personality; or entertaining oneself [e.g., @pelletierOneSizeDoesn2020].
+Negative effects include distraction; displacement of other meaningful activities; consumption of shallow, misleading or challenging content; or negative social comparisons [e.g., @meierDoesPassiveSocial2022].
+People are much more likely to use media that offer plentiful benefits [@katzUsesGratificationsResearch1973].
+Users implicitly learn what media help them regulate their mood and thereby well-being [@zillmannMoodManagementCommunication1988].
+Given the ubiquitous use of social media, the many benefits they offer, plus the general and implicit competence to use media that foster mood management, it is hence unlikely that the effects of social media use on well-being are profusely negative.
 
-**Two prominent media effect theories argue (mostly implicitly) against strong average negative effects.**
-**According to mood management theory [@zillmannMoodManagementCommunication1988], using media can substantially affect people's moods.**
-**Use can be stimulating or overwhelming, relaxing or boring.**
-**After some time, users implicitly learn what media help them balance their mood and affect according to their own situational needs [@zillmannMoodManagementCommunication1988].**
-**Those media that eventually become part of one's media repertoire are hence, on average, beneficial for users to regulate their mood.**
-**Using experience sampling of well-being and logs of social media use, a study with 82 participants from Italy found that after episodes of social media use, levels of positive affect increased significantly [@marcianoDynamicsAdolescentsSmartphone2022].**
+<!-- Two prominent media effect theories argue (mostly implicitly) against strong average negative effects. -->
+<!-- According to mood management theory [@zillmannMoodManagementCommunication1988], using media can substantially affect people's moods. -->
+<!-- Use can be stimulating or overwhelming, relaxing or boring. -->
+<!-- After some time, users implicitly learn what media help them balance their mood and affect according to their own situational needs [@zillmannMoodManagementCommunication1988]. -->
+<!-- Those media that eventually become part of one's media repertoire are hence, on average, beneficial for users to regulate their mood. -->
+<!-- Using experience sampling of well-being and logs of social media use, a study with 82 participants from Italy found that after episodes of social media use, levels of positive affect increased significantly [@marcianoDynamicsAdolescentsSmartphone2022].
+-->
 
-**While mood management theory considers media use mainly driven by implicit learning experiences, uses and gratifications theory upholds that the process is more explicit and rational [@katzUsesGratificationsResearch1973].**
-**Users select those media that they expect to have a desired effect, for example on mood, knowledge, or entertainment.** 
-**If those beneficial media effects do not exist or if they are not expected, people will spend their time elsewhere.**
-**And social media offer several beneficial effects, explaining why they are used that much.**
-**They help find relevant information, maintain and foster relationships, express one's personality, and entertain oneself [@pelletierOneSizeDoesn2020].**
-**In conclusion, because people spend so much time on social media consuming COVID-19 related content, according to both mood management theory and uses and gratifications theory this is indirect proof that average effects on well-being are likely not particularly negative.**
+<!-- While mood management theory considers media use mainly driven by implicit learning experiences, uses and gratifications theory upholds that the process is more explicit and rational [@katzUsesGratificationsResearch1973].
+-->
+<!-- Users select those media that they expect to have a desired effect, for example on mood, knowledge, or entertainment.  -->
+<!-- If those beneficial media effects do not exist or if they are not expected, people will spend their time elsewhere. -->
+<!-- And social media offer several beneficial effects, explaining why they are used that much. -->
+<!-- They help find relevant information, maintain and foster relationships, express one's personality, and entertain oneself [@pelletierOneSizeDoesn2020]. -->
+<!-- In conclusion, because people spend so much time on social media consuming COVID-19 related content, according to both mood management theory and uses and gratifications theory this is indirect proof that average effects on well-being are likely not particularly negative.
+-->
 
-But people can also misjudge media effects and are often overly optimistic [@metzgerComparativeOptimismPrivacy2017].
-Precisely because social media have so many positive consequences, one can ask if this is not where the actual problem lies.
-In other words, social media might not problematic because they are inherently bad, but rather because they are too good.
-As with many other things, there can be too much of a good thing.
-It is therefore often asked whether social media can become addictive, and users sometimes express this fear themselves [@yangCanWatchingOnline2021]. 
-However, a recently published meta-analysis found that the two most prominent measures of addiction, the Bergen Facebook Addiction Scale and the Bergen Social Media Addiction Scale, have only small relations to well-being [@duradoniWellbeingSocialMedia2020].
-In addition, the general idea of labeling excessive social and new media use as addiction was criticized, arguing that social media represent new regular behaviors that should not be pathologized [@galerHowMuchToo2018; @vanrooijWeakScientificBasis2018].
+<!-- But people can also misjudge media effects and are often overly optimistic [@metzgerComparativeOptimismPrivacy2017]. -->
+<!-- Precisely because social media have so many positive consequences, one can ask if this is not where the actual problem lies. -->
+<!-- In other words, social media might not problematic because they are inherently bad, but rather because they are too good. -->
+<!-- As with many other things, there can be too much of a good thing. -->
+<!-- It is therefore often asked whether social media can become addictive, and users sometimes express this fear themselves [@yangCanWatchingOnline2021].  -->
+<!-- However, a recently published meta-analysis found that the two most prominent measures of addiction, the Bergen Facebook Addiction Scale and the Bergen Social Media Addiction Scale, have only small relations to well-being [@duradoniWellbeingSocialMedia2020].
+-->
+<!-- In addition, the general idea of labeling excessive social and new media use as addiction was criticized, arguing that social media represent new regular behaviors that should not be pathologized [@galerHowMuchToo2018; @vanrooijWeakScientificBasis2018].
+-->
 
-**Because media effects can differ across users, situations, and content [@livingstoneEuropeanResearchChildren2018; @valkenburgDifferentialSusceptibilityMedia2013], I now briefly focus on the effects of COVID-19 related social media use specifically.**
-First, one could assume a _direct_ negative effect on well-being, and especially on positive or negative affect, which are more volatile and fluctuating.
-Dangers, inequalities, corruption---these were the headlines during the pandemic across many countries worldwide.
-If one learns about such events, the initial reaction might be shock, fear, or dismay.
-Consuming such news can be depressing [@dornemannHowGoodBad2021], perhaps even changing some general perspectives on life.
-<!-- So, just like being hit by a hammer hurts and we don't need any "mediating mechanism", this could be the case here as well. -->
-That said, because not all news was negative, and because many people showed solidarity and compassion, there was also positive and uplifting content, potentially compensating for the negative effects [@dornemannHowGoodBad2021].
-A study with 2.057 respondents from Italy reported that during the pandemic virtual community and social connectedness even increased [@guazziniSecondWaveAnalysis2022].
-**In a study with 735 participants from Finland, levels of loneliness did not decrease during the pandemic, and people who engaged more on social media experienced less loneliness [@latikkaLonelinessPsychologicalDistress2022].**
+<!-- Because media effects can differ across users, situations, and content [@livingstoneEuropeanResearchChildren2018; @valkenburgDifferentialSusceptibilityMedia2013], I now briefly focus on the effects of COVID-19 related social media use specifically.
+-->
+<!-- First, one could assume a _direct_ negative effect on well-being, and especially on positive or negative affect, which are more volatile and fluctuating. -->
+<!-- Dangers, inequalities, corruption---these were the headlines during the pandemic across many countries worldwide. -->
+<!-- If one learns about such events, the initial reaction might be shock, fear, or dismay. -->
+<!-- Consuming such news can be depressing [@dornemannHowGoodBad2021], perhaps even changing some general perspectives on life. -->
+<!-- <!-- So, just like being hit by a hammer hurts and we don't need any "mediating mechanism", this could be the case here as well. -->
+<!-- That said, because not all news was negative, and because many people showed solidarity and compassion, there was also positive and uplifting content, potentially compensating for the negative effects [@dornemannHowGoodBad2021].
+-->
+<!-- A study with 2.057 respondents from Italy reported that during the pandemic virtual community and social connectedness even increased [@guazziniSecondWaveAnalysis2022]. -->
+
+<!-- In a study with 735 participants from Finland, levels of loneliness did not decrease during the pandemic, and people who engaged more on social media experienced less loneliness [@latikkaLonelinessPsychologicalDistress2022].
+-->
 <!-- However, in light of a worldwide pandemic with millions of deaths, the negative direct effect seems more plausible. -->
 
-Second, there could also be _indirect_ effects.
-When browsing social media for COVID-19 related news, many users reported being captivated to such an extent that they could not stop using social media [@kleinDarklySoothingCompulsion2021].
-During the pandemic social media use was at an all-time high in the US [@statistaAverageDailyTime2021].
-Although it is most likely that moderate social media use is not detrimental [@orbenTeenagersScreensSocial2020], overuse, however, might be more critical, and several studies have shown more pronounced negative effects for extreme users [@przybylskiLargescaleTestGoldilocks2017].
-To explain, overuse could impair well-being if it replaces meaningful or functional activities such as meeting others, working, actively relaxing, or exercising.
-**Another potentially negative mechanism at play are problematic social comparison processes.
-During the pandemic, several users shared how they successfully dealt with challenges such as physical distancing.
-In a study with 1,131 residents from Wuhan in China [@yuePassiveSocialMedia2022], people who spent more time in quarantine also spent more time on social media.
-Those, who spent more time on social media also engaged in more upward social comparison, which was related to increased levels of stress.**
+<!-- Second, there could also be _indirect_ effects. -->
+<!-- Although it is most likely that moderate social media use is not detrimental [@orbenTeenagersScreensSocial2020], overuse, however, might be more critical, and several studies have shown more pronounced negative effects for extreme users [@przybylskiLargescaleTestGoldilocks2017].
+-->
+<!-- To explain, overuse could impair well-being if it replaces meaningful or functional activities such as meeting others, working, actively relaxing, or exercising. -->
+<!-- Another potentially negative mechanism at play are problematic social comparison processes.  -->
 <!-- Therefore, if a society collectively overuses social media during a pandemic, there is potential for negative effects. -->
@@
 
-On the other hand, one can make the case that using social media for COVID-19 related reasons might even be beneficial, especially in times of a pandemic.
-Exchanging COVID-19 related messages with friends via WhatsApp might replace the in-person contact one would have otherwise, but which is literally impossible at the time.
-In situations where meaningful and functional activities are prohibited, using social media to exchange about COVID-19 related topics might not be the worst idea.
-Besides, given that nowadays a large number of experts, scientists, and politicians converse directly on social media, one can get first-hand high quality information on current developments.
-<!-- On the other hand, there is of course also much disinformation, and "bingeing" on COVID-19 fake news might also pose risks for impaired well-being. -->
+<!-- On the other hand, one can make the case that using social media for COVID-19 related reasons might even be beneficial, especially in times of a pandemic. -->
+<!-- Exchanging COVID-19 related messages with friends via WhatsApp might replace the in-person contact one would have otherwise, but which is literally impossible at the time. -->
+<!-- In situations where meaningful and functional activities are prohibited, using social media to exchange about COVID-19 related topics might not be the worst idea. -->
+<!-- Besides, given that nowadays a large number of experts, scientists, and politicians converse directly on social media, one can get first-hand high quality information on current developments. -->
+<!-- <!-- On the other hand, there is of course also much disinformation, and "bingeing" on COVID-19 fake news might also pose risks for impaired well-being. -->
 
-To summarize, it seems that from a theoretical perspective it is most likely that the average effects of social media use on well-being are negligible. 
-Building on established theories from Communication and current empirical findings, we would not assume that effects are either profoundly negative or strongly positive.
+<!-- To summarize, it seems that from a theoretical perspective it is most likely that the average effects of social media use on well-being are negligible.  -->
+<!-- Building on established theories from Communication and current empirical findings, we would not assume that effects are either profoundly negative or strongly positive. -->
 
-## Empirical Studies on Social Media Effects
+<!-- ## Empirical Studies on Social Media Effects -->
 
-So far, there is still comparatively little empirical research on how well-being is affected by COVID-19 related social media use.
-In their study on the relations between media use and mental health during the pandemic, @bendauAssociationsCOVID19Related2021 found that people who used social media as a primary source of information reported on average "significantly more unspecific anxiety and depression [] and significantly more specific COVID-19 related anxiety symptoms" (p.
-288).
-@edenMediaCopingCOVID192020 analyzed the media use of 425 US college students during the first wave of the pandemic, finding both positive and negative relations with well-being.
-In a sample of 312 respondents collected via Amazon Mechanical Turk, @choiMediatedCommunicationMatters2021 reported that people who used media to attain information were more lonely and less satisfied with their lives.
-@stainbackCOVID1924News2020 analyzed a large-scale study with 11,537 respondents from the US and found that increased COVID-19 media consumption was related to more psychological distress.
+There is still little empirical research on how well-being is affected by social media use that is focused on COVID-19 specifically.
+Results are mixed.
+When browsing social media for COVID-19 related news, many users reported being captivated to such an extent they could not stop using social media [@kleinDarklySoothingCompulsion2021].
+<!-- During the pandemic social media use was at an all-time high in the US [@statistaAverageDailyTime2021], for example to cope with the required physical distancing measures. -->
+In a study with 1,131 residents from Wuhan in China, people who spent more time in quarantine also spent more time on social media [@yuePassiveSocialMedia2022]. 
+Those who spent more time on social media also engaged in more upward social comparison, which was related to increased levels of stress.
+People who used social media as a primary source of information reported on average "significantly more unspecific anxiety and depression [...] and significantly more specific COVID-19 related anxiety symptoms" [@bendauAssociationsCOVID19Related2021, p.
+288].
+@edenMediaCopingCOVID192020 analyzed how 425 US college students used media during the first wave of the pandemic, finding both positive and negative relations with well-being.
+<!-- In a sample of 312 respondents collected via Amazon Mechanical Turk, people who used media to attain information were more lonely and less satisfied with their lives [@choiMediatedCommunicationMatters2021].
+-->
+In a large-scale study with 11,537 respondents from the US, increased COVID-19-related media consumption was related to more psychological distress [@stainbackCOVID1924News2020].
 A four-wave panel study with 384 young adults from the U.S.
-analyzed the effects of general digital technology use---objectively measured via screenshots of screen-time applications---on mental health, separating within- and between-person relations [@sewallObjectivelyMeasuredDigital2021].
-The results showed that digital technology did not have significant effects on mental health [for a similar study with comparable results, see @bradleyStressMoodSmartphone2021].
-Together, the literature is mixed, with a slight focus on the negative effects of social media as news use [see also @dornemannHowGoodBad2021; @liuRelationOfficialWhatsAppdistributed2020; @riehmAssociationsMediaExposure2020].
+analyzed the effects of general digital technology use during the pandemic on mental health, finding that digital technology did not have significant effects on mental health [@sewallObjectivelyMeasuredDigital2021; for a similar study with comparable results, see @bradleyStressMoodSmartphone2021].
+A study with 2.057 respondents from Italy reported that during the pandemic virtual community and social connectedness increased [@guazziniSecondWaveAnalysis2022].
+In a study with 735 participants from Finland, levels of loneliness were stable during the pandemic, and people who engaged more on social media experienced less loneliness [@latikkaLonelinessPsychologicalDistress2022].
+Together, the literature is mixed, showing both positive and negative affects of social media use focused on or during COVID-19 on well-being [see also @dornemannHowGoodBad2021; @liuRelationOfficialWhatsAppdistributed2020; @riehmAssociationsMediaExposure2020].
 <!-- However, note that all of these findings represent between-person relations stemming from cross-sectional data (see below). -->
@@
 
-The question of whether and how social media use affects well-being _in general_, on the other hand, is well-researched.
-This also holds true for the different types of communication such as active or passive use.
-A meta review (i.e., an analysis of meta-analyses) found that the relation between social media use and well-being is likely in the negative spectrum but very small, potentially too small to matter [@meierComputermediatedCommunicationSocial2020a].
-What determines whether or not an effect should be considered small or trivial?
-As a starting point, we could refer to standardized effect sizes.
-According to @cohenPowerPrimer1992, small effect sizes start at _r_ = .10.
-And indeed, several if not most of the current meta-analyses find effect sizes below that threshold [@huangTimeSpentSocial2017; @meierComputermediatedCommunicationSocial2020a; @fergusonThisMetaanalysisScreen2021].
-
-Finally, several individual studies employing advanced methods found smalls relations between social media use and well-being [@orbenSocialMediaEnduring2019; @przybylskiDoesTakingShort2021a; @schemerImpactInternetSocial2021; @kerestesAdolescentsOnlineSocial2020].
-For example, @beyensSocialMediaUse2021 reported that although for roughly one quarter of all users the effects of social media use on well-being were negative, for almost the same number of users they were positive, while for the rest the effects were neutral.
-This finding is aligned with the Differential Susceptibility to Media Effects Model: 
-Although there is substantial _variation_ of media effects for individual users, the _average_ effects reported in the literature are often small [@valkenburgDifferentialSusceptibilityMedia2013].
+<!-- The question of whether and how social media use affects well-being _in general_, on the other hand, is well-researched. -->
+<!-- This also holds true for the different types of communication such as active or passive use. -->
+<!-- What determines whether or not an effect should be considered small or trivial? -->
+<!-- As a starting point, we could refer to standardized effect sizes. -->
+<!-- According to @cohenPowerPrimer1992, small effect sizes start at _r_ = .10. -->
+<!-- Finally, several individual studies employing advanced methods found smalls relations between social media use and well-being [@orbenSocialMediaEnduring2019; @przybylskiDoesTakingShort2021a; @schemerImpactInternetSocial2021; @kerestesAdolescentsOnlineSocial2020].
+-->
+<!-- Although for roughly one quarter of all users the effects of social media use on well-being were negative, for almost the same number of users they were positive, while for the rest the effects were neutral [@beyensSocialMediaUse2021].
+-->
+<!-- This finding is aligned with the Differential Susceptibility to Media Effects Model:  -->
+<!-- Although there is substantial _variation_ of media effects for individual users, the _average_ effects reported in the literature are often small [@valkenburgDifferentialSusceptibilityMedia2013].
+-->
 
@@
 
-In conclusion, in light of the theoretical considerations and empirical studies presented above, I expect that COVID-19 related communication on social media doesn't affect well-being in a meaningful or relevant way.
+In conclusion, these mixed empirical results, together with the observation that social media effects on well-being are very small in general, I expect that COVID-19 related communication on social media does not affect well-being in a meaningful or relevant way.
+
+<!-- A meta review (i.e., an analysis of meta-analyses) found that the relation between social media use and well-being is likely in the negative spectrum but very small, potentially too small to matter [@meierComputermediatedCommunicationSocial2020a].
+-->
+<!-- And indeed, several if not most of the current meta-analyses find small to trivial effects [@huangTimeSpentSocial2017; @meierComputermediatedCommunicationSocial2020a; @fergusonThisMetaanalysisScreen2021].
+-->
 
@@
 
-**In this study, this general hypothesis will be analyzed specifically for the three communication types of (a) time spent reading, (b) liking and sharing, and (c) actively posting COVID-19 related content.**
-**In addition, I will analyze how well-being is influenced by spending time on five prominent social media apps, including (a) Facebook, (b) Instagram, (c) Twitter, (d) WhatsApp, and (e) YouTube.** 
-**Three different well-being indicators will be distinguished: life satisfaction, positive affect, and negative affect.**
+<!-- In this study, this general hypothesis will be analyzed specifically for the three communication types of (a) time spent reading, (b) liking and sharing, and (c) actively posting COVID-19 related content.
+-->
+<!-- In addition, I will analyze how well-being is influenced by spending time on five prominent social media apps, including (a) Facebook, (b) Instagram, (c) Twitter, (d) WhatsApp, and (e) YouTube.
+-->
 
-# Current Study
+# Method
+## Preregistration
+
+The hypotheses, the sample, the measures, the analyses, and the inference criteria (SESOI, _p_-value) were preregistered on the Open Science Framework, accessible here: https://osf.io/87b24/?view_only=b2289b6fec214fa88ee75a18d45c18f3.
+Because in this study I analyze data from an already existing large-scale data set, the preregistration was done prior to accessing the data.
+The preregistration was designed on the basis of the panel documentation online [@kittelAustrianCoronaPanel2020].
+In some cases, it was impossible to execute the analyses as I had originally planned, for example because some properties of the variables only became apparent when seeing the actual data.
+The most relevant deviations are reported below, and a complete list of all changes can be found in the online [companion website](https://XMtRA.github.io/Austrian_Corona_Panel_Project) (https://XMtRA.github.io/Austrian_Corona_Panel_Project).
+
+## Sample
+
+The data come from the Austrian Corona Panel Project [@kittelAustrianCoronaPanel2021], which is a large-scale standalone panel study. 
+The data are hosted on AUSSDA, are publicly available here (https://doi.org/10.11587/28KQNS), and consist of 32 waves.
+The study was conducted between March 2020 and June 2022, and data collection is now finished.
+Between March 2020 and July 2020, the intervals between waves were weekly and afterward monthly.
+Each wave consists of at least 1,500 respondents.
+The sample size was _N_ = `r d_raw %>% nrow() %>% prettyNum(big.mark = ",")`, with overall `r d_long %>% nrow() %>% prettyNum(big.mark = ",")` observations.
+Panel mortality was compensated through a continuous re-acquisition of new participants. 
+Participants were sampled from a pre-existing online access panel provided by the company Marketagent, Austria.
+Panel members were incentivized with 180 credit points for each wave of the study.
+
+Achieved via quota sampling, the sample matched the Austrian population in terms of age, gender, region/state, municipality size, and educational level.
+In order to participate in the study, the respondents needed to be Austrian residents and had to be at least 14 years of age. 
+All respondents needed to have access to the internet (via computer or mobile devices such as smartphones or tablets). 
+Ethical review and approval was not required for the study in accordance with the local legislation and institutional requirements. 
+The participants provided their written informed consent to participate in this study.
+The average age was `r mean(2021 - d_wide$year_birth, na.rm = T) %>% round(0)` years, `r mean(d_wide$male, na.rm = T) %>% round(., 2) * 100` percent were male, `r (d_wide %>% filter(.$edu_fac %in% c("State college", "Bachelor", "Master", "PhD"))) %>% nrow() %>% divide_by(sum(table(d_wide$edu_fac))) %>% multiply_by(100) %>% round(0)` percent had a University degree, and `r (table(d_wide$employment_fac)["Unemployed"] / sum(table(d_wide$employment_fac))) %>% unname() %>% multiply_by(100) %>% round(., digits = 0)` percent were currently unemployed.
+
+<!-- # Current Study -->
 ## Smallest Effect Size of Interest
+<!-- ## Inference Criteria -->
 
-Testing this hypothesis, however, is not trivial.
-First, in contrast to most hypotheses typically posited in the social sciences it implicitly contains an effect size, a so-called smallest effect size of interest (SESOI).
-Effectively testing this hypothesis necessitates defining what's considered a "trivial effect size" and what's not.
-Above I already referred to standardized effect sizes.
-However, standardized effect sizes should only be a first step toward evaluating an effect's relevance [@baguleyStandardizedSimpleEffect2009].
-Standardized effect sizes are determined by a sample's variance,[^cohensd] which is problematic: 
-The question of whether or not social media use affects a particular person in a relevant way should not depend on the variance in the sample in which that person's data were collected.
-Instead, it should depend on absolute criteria.
-
-What could be a minimally interesting, nontrivial effect?
-Because this is a normative and ultimately philosophical question, there can never be a clear, single, or unanimous answer. 
+Testing the hypothesis necessitates defining what is considered a "trivial effect size". 
+<!-- Effectively testing this hypothesis  -->
+<!-- Above I already referred to standardized effect sizes. -->
+<!-- However, standardized effect sizes should only be a first step toward evaluating an effect's relevance [@baguleyStandardizedSimpleEffect2009]. -->
+<!-- Standardized effect sizes are determined by a sample's variance,[^cohensd] which is problematic:  -->
+<!-- The question of whether or not social media use affects a particular person in a relevant way should not depend on the variance in the sample in which that person's data were collected. -->
+<!-- Instead, it should depend on absolute criteria. -->
+<!-- What could be a minimally interesting, nontrivial effect? -->
+Being a normative question, finding a clear, single, or unanimous answer is impossible.
 <!-- In the end, it is a personal question. -->
-However, it is still necessary and helpful to try to provide such a plausible benchmark. 
-I therefore suggest the following SESOI for this research question:
-
-[^cohensd]: Consider the effect size Cohen's _d_: The mean's of the two groups that are to be compared are subtracted from one another and then divided by the sample's standard deviation [@cohenPowerPrimer1992].
-Hence, if there is more deviation/variance in a sample, the effect size decreases, even if the difference of the group's means stays the same.
+However, it is still necessary and helpful to work toward a plausible benchmark. 
+I suggest the following SESOI for this research question:
 
@@
 Thus, a change of four units in social media use (e.g., a complete stop) should correspond to a noticeable change in well-being.
-But what's a noticeable change in well-being?
+What is a noticeable change in well-being?
 According to @normanInterpretationChangesHealthrelated2003, people can reliably distinguish seven levels of satisfaction with health.
-So if satisfaction is measured on a 7-point scale, we would state that a four unit change in social media use should result in a one unit change in life satisfaction. 
-(For more information, see Methods section "Inference Criteria.")
-
-## Causality
-
-The hypothesis explicitly states a causal effect.
-In non-experimental studies, longitudinal designs can help investigate causality.
-Using longitudinal designs alone, however, is not sufficient for establishing correct causal statements [@rohrerTheseAreNot2021].
-In addition, we for example also need to control for confounding third variables.
-**Importantly, when analyzing longitudinal (within-person) relationships and effects, it is important to control for _varying_ third variables.**
-**Non-varying third variables can only help control non-varying (between-person) relations.**
-
-To illustrate, consider the following example.
-Imagine that a person suddenly starts using social media much more than usual, and then after some time  becomes less satisfied with their life.
-Eventually, use and life satisfaction return to prior levels.
-If this happens to several people at the same time, in a longitudinal study we could then observe a significant effect of social media use on life satisfaction.
-However, it could also be the case that during the study there was a major exogenous event (say, a pandemic), which caused large parts of the working population to loose their jobs.
-Hence, the causal effect reported above was confounded, because in reality it was the pandemic that caused both social media use to rise and life satisfaction to go down.
-
-Thus, only when controlling for _all_ relevant confounders, can we correctly estimate causality without bias [@rohrerThinkingClearlyCorrelations2018].
-Obviously, we can never be entirely sure to have included all confounders, which makes absolute statements regarding causality virtually impossible.
-In addition, when determining the overall causal effect, we need to make sure _not_ to control for mediating variables [@rohrerThinkingClearlyCorrelations2018], for doing so would bias our assessment of the causal effect.
-Complicating matters further, it is often unclear if a variable is a mediator or a confounder.[^collider]
-However, despite all these caveats, when controlling for relevant variables (that aren't mediators), we can be much more certain that we measured causality correctly.
-The aim should therefore be to collect as many varying and non-varying confounders as possible (which I believe is seldom done in our field), while knowing that absolute certainty regarding causality cannot be reached.
-
-[^collider]: In addition, there also exist colliders, which I don't discuss here and which complicate the issue even further [@rohrerThinkingClearlyCorrelations2018].
-
-When searching for suitable candidates for confounders, we should look for variables that affect both media use and well-being. 
-Controlling for these factors isolates the actual effect of social media use on well-being.
-We can also control for variables that affect only social media use or well-being. 
-However, in doing so not much is gained, because the effects of social media use would remain virtually the same [@klinePrinciplesPracticeStructural2016; but see @mcelreathYesterdayClass2021].
-
-In this study, I hence plan to control for the following variables, which either have already been shown to affect both social media use and well-being or which are likely to do so, and which also aren't mediators:
-<!-- (I'll additionally include variables that likely affect only well-being, also to obtain a comparison benchmark for social media effects): -->
-gender, age, education, Austria country of birth, Austria country of birth of parents, text-based news consumption, video-based news consumption, residency Vienna, household size, health, living space, access to garden, access to balcony, employment, work hours per week, being in home-office, household income, outdoor activities, satisfaction with democracy, disposition to take risks, and locus of control.[^explore]
-<!-- I will not control for variables such as trust in institutions or trust in media, because these variables might be influenced by social media use to a meaningful extent. -->
-
-[^explore]: The data-set includes many other variables that one could also potentially control for, and I invite interested readers to download the and explore potential interesting relationships.
-
-Next to including covariates, it's now increasingly understood that causal effects should be analyzed from an internal, within-person perspective [@hamakerWhyResearchersShould2014].
-If a specific person changes their media diet, we need to measure how this behavior affects their own well-being.
-Between-person comparisons from cross-sectional data, where participants are interviewed only once, cannot provide such insights.
-In this study, I hence differentiate between-person relations from within-person effects. 
-And as explicated above, to test the hypothesis I thus consider only the within-person effects.
-
-Finally, one precondition of causality is temporal order. 
-The cause needs to precede the effect.
-Finding the right interval between cause and effect is crucial.
-For example, if we want to understand the effect of alcohol consumption on driving performance, it makes a big difference if driving performance is measured one minute, one hour, one day, or one week after consumption.
-<!-- Finding the right interval is difficult. -->
-**If variables are stable, longer intervals are needed; if they fluctuate, shorter intervals.**
-**In the case of well-being, we need shorter intervals for the more fluctuating positive and negative affect, and longer ones for the more stable life satisfaction [@dienlinImpactDigitalTechnology2020].**
-**Using social media can have instant effects on mood [@marcianoDynamicsAdolescentsSmartphone2022].**
-**Effects on life satisfaction often take longer to manifest, for example because media use leads to actual changes in specific behaviors, which then in turn affect life satisfaction [@dienlinDisplacementReinforcementReciprocity2017].**
-Choosing the right interval is challenging, because especially short intervals are hard to implement in practice, often requiring advanced methods such as experience sampling (also known as in situ measurement or ambulant assessment) [@schnauber-stockmannMobileDevicesTools2020].
-In this study, I hence analyze how using social media during the last week affected positive and negative affect during the same week.
-In other words, if people during the last week engaged in more COVID-19 related social media use than they usually do, did they feel better or worse during that week than they usually do?
-Regarding life satisfaction, I implemented a longer interval.
-If people during the last week used COVID-19 related social media more than they usually do, were they at the end of the week more or less satisfied with their lives than they usually are?
-I hence analyze if when a person changes their social media diet, are there (a) _simultaneous_ changes in their affect and (b) _subsequent_ changes in their life satisfaction?
-**These relations will be controlled for varying confounders, which fosters a causal interpretation.**
-Similar approaches were implemented by other studies [@johannesNoEffectDifferent2022; @scharkowHowSocialNetwork2020], and they are considered a best practice approach toward analyzing causality [@bellFixedRandomEffects2019].
-
-# Method
-
-In this section I describe the preregistration and how I determined the sample size, data exclusions, the analyses, and all measures in the study.
-
-## Preregistration
-
-The hypotheses, the sample, the measures, the analyses, and the inference criteria (SESOI, p-value) were preregistered on the Open Science Framework. 
-The (anonymous) preregistration can be accessed here: https://osf.io/87b24/?view_only=b2289b6fec214fa88ee75a18d45c18f3.
-Because in this study I analyzed data from an already existing large-scale data set, all of these steps were done prior to accessing the data.
-The preregistration was designed on the basis of the panel documentation online [@kittelAustrianCoronaPanel2020].
-In some cases I couldn't execute the analyses as I had originally planned, for example because some properties of the variables only became apparent when inspecting the actual data.
-The most relevant deviations are reported below, and a complete list of all changes can be found in the online [companion website](https://XMtRA.github.io/Austrian_Corona_Panel_Project) (https://XMtRA.github.io/Austrian_Corona_Panel_Project).
-
-## Sample
-
-**The data come from the Austrian Corona Panel Project [@kittelAustrianCoronaPanel2021], which is a large-scale standalone panel study. **
-**The data are hosted on AUSSDA and are publicly available here: https://doi.org/10.11587/28KQNS.**
-**At the time of writing, the official website featured a data-set consisting of 24 waves.**
-**For the analyses presented here, I received an advance version consisting of all 32 waves.**
-**The study was conducted between March 2020 and June 2022, and data collection is now finished.**
-Between March 2020 and July 2020, the intervals between waves were weekly, and afterward the intervals were monthly.
-Each wave consists of at least 1,500 respondents.
-The overall sample size was _N_ = `r d_raw %>% nrow() %>% prettyNum(big.mark = ",")`, and `r d_long %>% nrow() %>% prettyNum(big.mark = ",")` observations were collected.
-Panel mortality was compensated through a continuous acquisition of new participants. 
-All respondents needed to have access to the internet (via computer or mobile devices such as smartphones or tablets).
-They were sampled from a pre-existing online access panel provided by the company Marketagent, Austria.
-Respondents were asked and incentivized with 180 credit points to participate in each wave of the panel.
-
-Achieved via quota sampling, the sample matched the Austrian population in terms of age, gender, region/state, municipality size, and educational level.
-In order to participate in the study, the respondents needed to be Austrian residents and had to be at least 14 years of age. 
-Ethical review and approval was not required for the study in accordance with the local legislation and institutional requirements. 
-The participants provided their written informed consent to participate in this study.
-The average age was `r mean(2021 - d_wide$year_birth, na.rm = T) %>% round(0)` years, `r mean(d_wide$male, na.rm = T) %>% round(., 2) * 100` percent were male, `r (d_wide %>% filter(.$edu_fac %in% c("State college", "Bachelor", "Master", "PhD"))) %>% nrow() %>% divide_by(sum(table(d_wide$edu_fac))) %>% multiply_by(100) %>% round(0)` percent had a University degree, and `r (table(d_wide$employment_fac)["Unemployed"] / sum(table(d_wide$employment_fac))) %>% unname() %>% multiply_by(100) %>% round(., digits = 0)` percent were currently unemployed.
-
-## Inference Criteria
-
-Because the data were analyzed post-hoc, no a-priori sample size planning on the basis of power analyses was conducted.
-The sample is large, and it is hence well-equipped to reliably detect small effects.
-<!-- , which is why no exact post hoc power analysis were conducted. -->
-In addition, because such large samples easily generate significant _p_-values even for very small effects, it helps that the hypotheses were tested with a smallest effect size of interest-approach.
-To this end, I adopted the interval testing approach as proposed by @dienesUsingBayesGet2014.
-On the basis of the SESOI, I defined a null region.
-In what follows, I explain how I determined the SESOI and the null region.
+So if satisfaction is measured on a 7-point scale, a four unit change in social media use should result in a one unit change in life satisfaction. 
+<!-- Because the data were analyzed post-hoc, no a-priori sample size planning on the basis of power analyses was conducted. -->
+<!-- The sample is large, and it is hence well-equipped to reliably detect small effects. -->
+<!-- <!-- , which is why no exact post hoc power analysis were conducted. -->
+<!-- In addition, because such large samples easily generate significant _p_-values even for very small effects, it helps that the hypotheses were tested with a smallest effect size of interest-approach.
+-->
+<!-- To this end, I adopted the interval testing approach as proposed by @dienesUsingBayesGet2014. -->
+<!-- On the basis of the SESOI, I defined a null region. -->
+<!-- In what follows, I explain how I determined the SESOI and the null region. -->
 
 In this study, life satisfaction was measured on an 11-point scale.
-If people can reliably differentiate 7 levels as mentioned above, this corresponds to 11 / 7 = `r round(11/7, 2)` unit change on an 11-point scale.
+If people can reliably differentiate 7 levels, this corresponds to 11 / 7 = `r round(11/7, 2)` unit change on an 11-point scale.
 Hence, a four-point change in media use (e.g., a complete stop) should result in a `r round(11/7, 2)`-point change in life satisfaction.
 In a statistical regression analysis, _b_ estimates the change in the dependent variable if the independent variable increases by one point.
-We would therefore expect a SESOI of _b_ = `r round(11/7, 2)` / 4 = `r round(11/7*.25, 2)`.
-For affect, which was measured on a 5-point scale, our SESOI would be _b_ = `r round(5 / 7, 2)` / 4 = `r round(5 / 7 * .25, 2)`.
-Because we're agnostic as to whether the effects are positive or negative, the null region includes negative and positive effects.
+For life satisfaction, we would therefore define a SESOI of _b_ = `r round(11/7, 2)` / 4 = `r round(11/7*.25, 2)`.
+For positive or negative affect, which was measured on a 5-point scale, our SESOI would be _b_ = `r round(5 / 7, 2)` / 4 = `r round(5 / 7 * .25, 2)`.
+Because we are agnostic as to whether the effects are positive or negative, the null region includes both negative and positive effects.
 Finally, in order not to exaggerate precision and to be less conservative, these numbers are reduced to nearby thresholds.[^rounding] 
@@
 <!-- As explained above, for well-being the null region was between _b_ = -.30 and _b_ = .30 (i.e., _b_ = |.30|). -->
-Let's briefly illustrate what this means in practice.
-If the 95% confidence interval falls completely within the null-region (e.g., _b_ = -.02, [95% CI: -.12, .08]), the hypothesis that the effect is trivial is supported.
+The hypothesis is analyzed using the interval testing approach as proposed by @dienesUsingBayesGet2014.
+To illustrate, if the 95% confidence interval falls completely within the null-region (e.g., _b_ = -.02, [95% CI: -.12, .08]), the hypothesis that the effect is trivial is supported.
 If the confidence interval and the null region overlap (e.g., _b_ = -.22, [95% CI: -.27, -.17]), the hypothesis is not supported and the results are considered inconclusive, while a meaningful negative effect is rejected.
 If the confidence interval falls completely outside of the null-region (e.g., _b_ = -.40, [95% CI: -.45, -.35]), the hypothesis is rejected and the existence of a meaningful positive effect is supported.
-For an illustration, see Figure \@ref(fig:sesoi)).
-
-<!-- Responses were individually checked for patterns such as straight-lining or missing of inverted items. -->
-<!-- X clear cases were removed. -->
-
-## Data Analysis
-
-The hypothesis was analyzed using mixed effects models, namely random effect within-between models (REWB)[@bellFixedRandomEffects2019]. 
-Three models were run, one for each dependent variable.
-The data were hierarchical, and responses were separately nested in participants and waves (i.e., participants and waves were implemented as random effects).
-Nesting in participants allowed to separate between-person relations from within-person effects.
-Nesting in waves allowed to control for general exogenous developments, such as general decreases in well-being in the population, for example due to lockdown measures.
-Thus, there was no need additionally to control for specific phases or measures of the lockdown.
-Predictors were modeled as fixed effects.
-They included social media communication types and channels, separated into within and between-person factors, as well as stable and varying covariates.
-All predictors were included simultaneously and in each of the three models.
-
-The factorial validity of the scales were tested with confirmatory factor analyses (CFA).
-Because Mardia’s test showed that the assumption of multivariate normality was violated, I used the more robust Satorra-Bentler scaled and mean-adjusted test statistic (MLM) as estimator.
-To avoid over-fitting, I tested the scales on more liberal fit criteria (CFI > .90, TLI > .90, RMSEA <. .10, SRMR < .10) [@klinePrinciplesPracticeStructural2016].
-<!-- Finally, REWB-models cannot model latent variables. -->
-<!-- To increase precision, I therefore exported factor scores from the CFAs for positive and negative affect. -->
-Mean scores were used for positive and negative affect.
-<!-- Respondents who answered less than 50% of all questions were removed. -->
-**Missing responses were imputed using multiple imputation with predictive mean matching (five iterations, five data-sets), including categorical variables.**
-**All variables were imputed except the media use measures, as they were not collected on each wave.**
-**All variables included in the analyses presented here were used to impute missing data.**
-**For the main analyses, results were pooled across all five data-sets.**
-
-For more information on the analyses, a complete documentation of the models and results, additional analyses (for example using multiple imputation or no imputation), see [companion website](https://XMtRA.github.io/Austrian_Corona_Panel_Project).
-
-## Measures
-
-In what follows, I list all the variables that I analyzed.
-For the variables' means, range, and variance, see Table \@ref(tab:tab-descriptives).
-For a complete list of all items and item characteristics, see [companion website](https://XMtRA.github.io/Austrian_Corona_Panel_Project).
-
-### Well-being
-
-Life satisfaction was measured with the item "All things considered, how satisfied are you with your life as a whole nowadays?" from the European Social Survey [@europeansocialsurveyESS9Edition20182021].
-The response options ranged from 0 (_extremely dissatisfied_) to 10 (_extremely satisfied_).
-
-To capture positive affect, respondents were asked how often in the last week they felt (a) calm and relaxed, (b) happy, and (c) full of energy [@worldhealthorganizationWellbeingMeasuresPrimary1998].
-The response options were 1 (_never_), 2 (_on some days_), 3 (_several times per week_), 4 (_almost every day_), and 5 (_daily_).
-The scale showed good factorial fit, `r fit_txt(cfa_aff_pos)`.
-Reliability was high, $\omega$ = `r rel_aff_pos %>% my_round("std")`.
-
-For negative affect, respondents were asked how often in the last week they felt (a) lonely, (b) aggravated, (c) so depressed, that nothing could lift you up, (d) very nervous, (e) anxious, and (h) glum and sad [@worldhealthorganizationWellbeingMeasuresPrimary1998].
-The response options were 1 (_never_), 2 (_on some days_), 3 (_several times per week_), 4 (_almost every day_), and 5 (_daily_).
-The scale showed good factorial fit, `r fit_txt(cfa_aff_neg)`.
-Reliability was high, $\omega$ = `r rel_aff_neg %>% my_round("std")`.
-
-All three variables were measured on each wave.
-
-### COVID-19 related social media use
-
-COVID-19 related social media use focused on communication types was measured with the three dimensions of (a) reading, (b) liking and sharing, and (c) posting.
-The items come from @wagnerAUTNESOnlinePanel2018 and were adapted for the context of this study.
-The general introductory question was "How often during the last week have you engaged in the following activities on social media?"
-The three items were "Reading the posts of others with content on the Coronavirus", "When seeing posts on the Coronavirus, I clicked 'like', 'share' or 'retweet'", "I myself wrote posts on the Coronavirus on social media."
-Answer options were 1 (_several times per day_), 2 (_daily_), 3 (_several times per week_), 4 (_weekly_), 5 (_never_).
-The items were inverted for the analyses.
-
-COVID-19 related social media use focused on channels was measured with five variables from @wagnerAUTNESOnlinePanel2018, adapted for this study. 
-The general introductory question was "How often in the last week have you followed information related to the Corona-crisis on the following social media?"
-The five items were (a) Facebook, (b) Twitter, (c) Instagram, (d) Youtube, and (e) WhatsApp.
-Again, the answer options were 1 (_several times per day_), 2 (_daily_), 3 (_several times per week_), 4 (_weekly_), 5 (_never_).
-Again, the items were inverted for the analyses.
-
-Social media use was measured for all participants on waves 1, 2, 8, 17, 23, **and 28**.
-Freshly recruited respondents always answered all questions on COVID 19-related social media use.
-
-### Control variables
-
-The effects of COVID-19 related social media use were controlled for the following stable variables: 
-(a) gender (female, male, diverse), (b) age, (c) education (ten options), (d) Austria country of birth (yes/no), (e) Austria parents' country of birth (no parent, one parent, both parents), (f) household size, (g) work hours per week, (h) home office, and (i) household income.
-I originally planned to implement additional variables as varying covariates.
-However, because they were not measured often enough or not at the time when social media use was measured, I implemented them as stable variables using their average values across all waves.
-This includes (a) text-based media news consumption (five degrees), (b) video-based media news consumption (five degrees), (c) residency is Vienna (yes/no), (d) living space (eleven options), (e) access to balcony (yes/no), (f) access to garden (yes/no), (g) employment (nine options), (h) disposition to take risks (eleven degrees), and (i) locus of control (five degrees).
-I also controlled for the following varying covariates: (a) five items measuring outdoor activities such as sport or meeting friends (five degrees), (b) satisfaction with democracy (five degrees), (c) self-reported physical health (five degrees), and (d) whether participants contracted COVID-19 since the last wave.
-
-# Results
-
-First, when looking at the variables from a descriptive perspective (Figure \@ref(fig:fig-descriptives)), we see that all well-being measures did not change substantially across the different waves of data collection.
-COVID-19 related media use, however, decreased slightly at the beginning of the study and remained stable after approximately six waves.
-The initial decrease might be explained by the fact that the collection of data began at the end of March 2020, hence approximately three months after the pandemic began.
-It could be that after an initial uptick, COVID-19 related social media use was already declining at the time, returning to more normal levels.
-
-## Preregistered Analyses
-
-The study's main hypothesis was that the effects of social media use on well-being would be trivial.
-Regarding the effects of different communication types---that is, reading vs.
-sharing vs.
-posting---all within-person effects fell completely within the a-priori defined null region (see Figure \@ref(fig:fig-between)).
-For example, respondents who used social media more frequently than usual to like or share COVID-19 related content did not show a simultaneous change in negative affect (_b_ = `r dat_fig_results_within %>% filter(dv == "Life satisfaction" & iv == "Liking & Sharing" & type == "within" & Analysis == "Regular") %>% select(estimate) %>% round(2)` [95% CI `r dat_fig_results_within %>% filter(dv == "Life satisfaction" & iv == "Liking & Sharing" & type == "within" & Analysis == "Regular") %>% select(conf.low) %>% round(2)`, `r dat_fig_results_within %>% filter(dv == "Life satisfaction" & iv == "Liking & Sharing" & type == "within" & Analysis == "Regular") %>% select(conf.high) %>% round(2)`]).
-As a result, the hypothesis was supported for all COVID-19 related types of social media communication.
-
-**However, two effects were statistically significantly different from zero.**
-**Users who wrote more COVID-19 related posts than usual were also slightly less satisfied with their lives than usual (_b_ = `r dat_fig_results_within %>% filter(dv == "Life satisfaction" & iv == "Posting" & type == "within" & Analysis == "Regular") %>% select(estimate) %>% round(2)` [95% CI `r dat_fig_results_within %>% filter(dv == "Life satisfaction" & iv == "Posting" & type == "within" & Analysis == "Regular") %>% select(conf.low) %>% round(2)`, `r dat_fig_results_within %>% filter(dv == "Life satisfaction" & iv == "Posting" & type == "within" & Analysis == "Regular") %>% select(conf.high) %>% round(2)`]).**
-**Users who wrote more COVID-19 related posts than usual also experienced slightly more negative affect than usual (_b_ = `r dat_fig_results_within %>% filter(dv == "Negative affect" & iv == "Posting" & type == "within" & Analysis == "Regular") %>% select(estimate) %>% round(2)` [95% CI `r dat_fig_results_within %>% filter(dv == "Negative affect" & iv == "Posting" & type == "within" & Analysis == "Regular") %>% select(conf.low) %>% round(2)`, `r dat_fig_results_within %>% filter(dv == "Negative affect" & iv == "Posting" & type == "within" & Analysis == "Regular") %>% select(conf.high) %>% round(2)`]).**
-**There was also a small and statistically non-significant trend that reading COVID-19 related content slightly increased life satisfaction (_b_ = `r dat_fig_results_within %>% filter(dv == "Life satisfaction" & iv == "Reading" & type == "within" & Analysis == "Regular") %>% select(estimate) %>% round(2)` [95% CI `r dat_fig_results_within %>% filter(dv == "Life satisfaction" & iv == "Reading" & type == "within" & Analysis == "Regular") %>% select(conf.low) %>% round(2)`, `r dat_fig_results_within %>% filter(dv == "Life satisfaction" & iv == "Reading" & type == "within" & Analysis == "Regular") %>% select(conf.high) %>% round(2)`], _p_ = `r dat_fig_results_within %>% filter(dv == "Life satisfaction" & iv == "Reading" & type == "within" & Analysis == "Regular") %>% select(p.value) %>% my_round("p")`).**
-**At the same time, there was also a small and statistically non-significant trend that reading COVID-19 related content decreased positive affect (_b_ = `r dat_fig_results_within %>% filter(dv == "Positive affect" & iv == "Reading" & type == "within" & Analysis == "Regular") %>% select(estimate) %>% round(2)` [95% CI `r dat_fig_results_within %>% filter(dv == "Positive affect" & iv == "Reading" & type == "within" & Analysis == "Regular") %>% select(conf.low) %>% round(2)`, `r dat_fig_results_within %>% filter(dv == "Positive affect" & iv == "Reading" & type == "within" & Analysis == "Regular") %>% select(conf.high) %>% round(2)`], _p_ = `r dat_fig_results_within %>% filter(dv == "Positive affect" & iv == "Reading" & type == "within" & Analysis == "Regular") %>% select(p.value) %>% my_round("p")`).**
-
-Regarding the COVID-19 related use of social media channels, the results were comparable (see Figure \@ref(fig:fig-within)).
-Changes in the frequency of using different social media channels to attain information regarding COVID-19 were unrelated to meaningful changes in well-being.
-For example, respondents who used Facebook more frequently than usual to learn about COVID-19 did not show a simultaneous change in well-being (_b_ = `r dat_fig_results_within %>% filter(dv == "Life satisfaction" & iv == "Facebook" & type == "within" & Analysis == "Regular") %>% select(estimate) %>% round(2)` [95% CI `r dat_fig_results_within %>% filter(dv == "Life satisfaction" & iv == "Facebook" & type == "within" & Analysis == "Regular") %>% select(conf.low) %>% round(2)`, `r dat_fig_results_within %>% filter(dv == "Life satisfaction" & iv == "Facebook" & type == "within" & Analysis == "Regular") %>% select(conf.high) %>% round(2)`]).
-In sum, the hypothesis was supported also for the COVID-19 related use of important social media channels.
-
-**That said, two effects differed substantially from zero.**
-**Respondents who used Instagram more frequently than usual to attain COVID-19 related news reported slightly lower levels of negative affect than usual (_b_ = `r dat_fig_results_within %>% filter(dv == "Negative affect" & iv == "Instagram" & type == "within" & Analysis == "Regular") %>% select(estimate) %>% my_round("b")` [95% CI `r dat_fig_results_within %>% filter(dv == "Negative affect" & iv == "Instagram" & type == "within" & Analysis == "Regular") %>% select(conf.low) %>% my_round("b")`, `r dat_fig_results_within %>% filter(dv == "Negative affect" & iv == "Instagram" & type == "within" & Analysis == "Regular") %>% select(conf.high) %>% my_round("b")`]).**
-**Respondents who used YouTube more frequently than usual to attain COVID-19 related news reported slightly higher levels of negative affect than usual (_b_ = `r dat_fig_results_within %>% filter(dv == "Negative affect" & iv == "YouTube" & type == "within" & Analysis == "Regular") %>% select(estimate) %>% round(2)` [95% CI `r dat_fig_results_within %>% filter(dv == "Negative affect" & iv == "YouTube" & type == "within" & Analysis == "Regular") %>% select(conf.low) %>% my_round("b")`, `r dat_fig_results_within %>% filter(dv == "Negative affect" & iv == "YouTube" & type == "within" & Analysis == "Regular") %>% select(conf.high) %>% my_round("b")`]).**
-**However, both effects were still completely inside of the null region, hence likely not large enough to be considered meaningful.**
-
-**For an overview of all within-person effects, see Table \@ref(tab:tab-within) and Figure \@ref(fig:fig-within).**
-
-## Exploratory Analyses
-
-In what follows, I briefly report some exploratory analyses that weren't preregistered.
-
-### Between-person relations
-
-For between-person relations, no a-priori hypotheses were formulated.
-Results showed that no relation crossed or was completely outside of the SESOI.
-Four relations were statistically significant.
-**Respondents who across all waves used social media more frequently than others to read about COVID-19 related posts reported slightly lower levels of positive affect than others (_b_ `r dat_fig_results_between %>% filter(dv == "Positive affect" & iv == "Reading" & type == "between" & Analysis == "Regular") %>% select(estimate) %>% my_round("b_txt")` [95% CI `r dat_fig_results_between %>% filter(dv == "Positive affect" & iv == "Reading" & type == "between" & Analysis == "Regular") %>% select(conf.low) %>% my_round("b")`, `r dat_fig_results_between %>% filter(dv == "Positive affect" & iv == "Reading" & type == "between" & Analysis == "Regular") %>% select(conf.high) %>% my_round("b")`]).**
-**Respondents who across all waves used social media more frequently than others to write COVID-19 related posts reported slightly higher levels of negative affect than others (_b_ `r dat_fig_results_between %>% filter(dv == "Negative affect" & iv == "Posting" & type == "between" & Analysis == "Regular") %>% select(estimate) %>% my_round("b_txt")` [95% CI `r dat_fig_results_between %>% filter(dv == "Negative affect" & iv == "Posting" & type == "between" & Analysis == "Regular") %>% select(conf.low) %>% my_round("b")`, `r dat_fig_results_between %>% filter(dv == "Negative affect" & iv == "Posting" & type == "between" & Analysis == "Regular") %>% select(conf.high) %>% my_round("b")`]).**
-**At the same time, respondents who across all waves used social media more frequently than others to write COVID-19 related posts also reported slightly higher levels of positive affect (_b_ `r dat_fig_results_between %>% filter(dv == "Positive affect" & iv == "Posting" & type == "between" & Analysis == "Regular") %>% select(estimate) %>% my_round("b_txt")` [95% CI `r dat_fig_results_between %>% filter(dv == "Positive affect" & iv == "Posting" & type == "between" & Analysis == "Regular") %>% select(conf.low) %>% my_round("b")`, `r dat_fig_results_between %>% filter(dv == "Positive affect" & iv == "Posting" & type == "between" & Analysis == "Regular") %>% select(conf.high) %>% my_round("b")`]).**
-**Finally, respondents who across all waves used YouTube more frequently than others also reported slightly higher levels of life satisfaction than others (_b_ `r dat_fig_results_between %>% filter(dv == "Life satisfaction" & iv == "YouTube" & type == "between" & Analysis == "Regular") %>% select(estimate) %>% my_round("b_txt")` [95% CI `r dat_fig_results_between %>% filter(dv == "Life satisfaction" & iv == "YouTube" & type == "between" & Analysis == "Regular") %>% select(conf.low) %>% my_round("b")`, `r dat_fig_results_between %>% filter(dv == "Life satisfaction" & iv == "YouTube" & type == "between" & Analysis == "Regular") %>% select(conf.high) %>% my_round("b")`]).**
-<!-- However, note that the effect were still completely inside of the null region, hence not large enough to be considered practically relevant. -->
-
-Note that when comparing the results with and without control variables, the results differed.
-For example, on the between-person level, one effect stopped being significant if controlled for additional variables.
-Initially, actively posting on social media was significantly (though not meaningfully) related to decreased life satisfaction.
-However, when controlling for potential confounders, the effect became virtually zero.
-
-For an overview of all between-person relations, see Figure \@ref(fig:fig-between).
-
-### Covariates
-
-To contextualize the results reported above and to see if the results included any meaningful effects at all, I also looked at the effect sizes of the covariates.
-Because each variable had different response options, we would actually need to define a SESOI for each variable, which for reasons of scope I cannot implement here. 
-Therefore, I report the results of the standardized scales,
-which also allows for a better comparison across the differently scaled variables.
-As a rough estimate for the SESOI we can build on the typical convention that small effects start at _r_ = |.10|.
-The results showed that several effects crossed or fell outside of the SESOI, were hence considered meaningful.
-This includes for example internal locus of control, health, satisfaction with democracy, or exercising.
-For an overview, see Figure \@ref(fig:fig-control).
-
-### Robustness-check
-
-To find out whether the inferences were robust across plausible (though arguably inferior) alternative analyses, I reran the analyses also using standardized estimates, additional covariates including trust in media or government, single imputation, and with a data set where missing data were not imputed.
-The results were virtually the same.
-For example, all within-person standardized COVID-19 related types of social media use or channels were significantly smaller than $\beta$ = |.05|, again supporting that effects were negligible.
-The results of the standardized analyses are reported in Table \@ref(tab:tab-within).
-The additional analyses are reported on the [companion website](https://XMtRA.github.io/Austrian_Corona_Panel_Project/analyses_additional.html).
-
-# Discussion
-
-In this study I analyzed the effects of COVID-19 related social media use on well-being.
-The data come from a panel study with 32 waves and are largely representative of the Austrian population.
-In a random effects model I separated between person relations from within-person effects. 
-I controlled for a large number of both stable and varying covariates, aiming to assess causality.
-**The results showed that some statistically significant negative within-person effects existed, but that they were very small and likely negligible.**
-People who used social media more than usual to learn about COVID-19 didn't show meaningful changes in their well-being.
-
-The results imply that COVID-19 related social media use doesn't seem to be particularly relevant for well-being.
-Other factors among the third variables that were measured revealed larger effects or relations, suggesting that well-being is rather determined by alternative aspects such as health, satisfaction with democracy, locus of control, or exercising.
-According to this study, popular fears that "doomscrolling" or overusing social media during times of crises is detrimental are not supported.
-
-**That said, several preliminary and subtle trends can be observed.**
-**First, overall the results do suggest that effects of COVID-19 related social media use on well-being rather tend to take place in the negative as opposed to the positive spectrum.**
-**For example, people who wrote more COVID-19 related posts than usual reported slightly lower levels of life satisfaction than usual.**
-**Similarly, people who wrote more COVID-19 related posts than usual also reported slightly more negative affect.**
-**As a potential explanation, when writing posts and comments on social media people explicitly and more deeply engage with COVID-19 related content.**
-**The tonality on social media is often extreme, negative, or aggressive, which potentially affects their authors.**
-**And because I controlled for whether or not participants had a COVID-19 infection during a specific wave, we can rule out the potential explanation that having an infection was the root cause of increased communication and reduced well-being.**
-
-**The hypothesis that tonality might be a relevant factor at play is also supported by the second trend.**
-**People, who during the pandemic spent more time on Instagram than usual, also experienced less negative affect than usual.**
-**Instagram is well-known for its positivity bias [@waterlooNormsOnlineExpressions2018].**
-**Content is generally more positive, uplifting, and (self-)flattering.**
-**It seems the much maligned positivity bias on Instagram might have been somewhat beneficial in times of the pandemic.**
-**The critique that the positivity bias necessarily leads to envy and negative feelings is one-sided, because positive content can also inspire and motivate users [@meierInstagramInspirationHow2020], which could be especially helpful in times of lockdown and home-office.**
-**To provide a concrete example, during the pandemic Instagram was successfully used as an interactive communication channel for first year students to have a better start into their new degree, effectively complementing alternative learning platform tools [@yeTurningInformationDissipation2020].**
-
-**Similarly, people who spent more time on YouTube than usual also reported slightly more negative affect than usual.** 
-**Communication on YouTube is often found to be more negative and less polite compared to other SNSs [@halpernSocialMediaCatalyst2013].**
-**YouTube is also routinely linked to mis- and disinformation.**
-**Of the 69 most viewed videos on YouTube on COVID-19, 19 (27.5%) contained nonfactual information [@liYouTubeSourceInformation2020].**
-**Consuming more negative and misleading information might hence be a potential explanation for the slightly increased levels of negative affect.**
-
-**The results showed that it makes sense to analyze different communication types and communication channels, and that active and passive communication showed different results.**
-**Liking and sharing content did not show any within-person effects. **
-**Such rather low-key active behaviors do not seem to affect well-being at all.**
-**Regarding passive use, reading COVID-19 related posts is more ambivalent.**
-**Results showed some weak trends towards a positive effect on life satisfaction, but a negative effect on mood.**
-**It might be that reading and informing oneself about COVID-19 on social media is helpful in the long run, but more negative for short-term affect.**
-**Finally, proactively engaging via writing posts, the most active form of communication analyzed here, showed only negative effects on well-being.**
-**The results support the findings from @valkenburgAssociationsActivePassive2022, who also could not confirm the claim active use is good and passive use is bad.**
-**Focusing on communication channels, YouTube seems to be more negative, whereas Instagram is likely more positive.**
-**Again, these are only very small effects.**
-**Future research might elaborate on these specific relations to probe their stability and relevance.**
-
-Taken together, on the one hand the results are not aligned with several recent studies analyzing similar or closely related research questions.
-This includes a study by @bendauAssociationsCOVID19Related2021, which showed negative relations between social media and well-being [but see @bradleyStressMoodSmartphone2021; or @sewallObjectivelyMeasuredDigital2021].
-However, note that @bendauAssociationsCOVID19Related2021 analyzed cross-sectional data on a between-person level while not controlling for third variables, which is not optimal for investigating causal effects.
-On the other hand, the results are well-aligned with **mood management theory [@zillmannMoodManagementCommunication1988]** or the uses and gratifications approach [@katzUsesGratificationsResearch1973].
-If effects were indeed profoundly negative on average, then people likely wouldn't spend so much time on social media engaging with COVID-19 content.
-Likewise, recent studies and meta-analyses analyzed the effects of social media use from a more general perspective or from a somewhat different angle. 
-These studies have found that the effects of various types of social media use on several well-being indicators are small at best, often too small to matter [@meierComputermediatedCommunicationSocial2020a; @orbenTeenagersScreensSocial2020; @fergusonThisMetaanalysisScreen2021], which echoes the results obtained here.
-
-**From a more political and societal perspective, the results imply that it can make sense to critically reflect upon COVID-19 related social media use.** 
-**On average, it might be slightly beneficial to post less actively about COVID-19 on social media and to spend less time on YouTube.**
-**Spending time on Instagram seems to be okay.**
-**The potentially resulting positive effects, however, will for many users likely not be noticeable.**
-**Results allow us to reject a positive effect: Writing more posts on social media will likely not increase well-being.**
-**At all events, engaging in COVID 19-related social media use should, on average, not be a major cause for concern.**
-
-<!-- First, in terms of media channels, using Twitter more than usual was related to slightly decreased levels of life satisfaction. -->
-<!-- Twitter is considered to have more negative affordances and tonality as compared to other networks such as Instagram [@waterlooNormsOnlineExpressions2018], which might help explain the results.
--->
-<!-- Second, of all the three COVID-19 related social media activities, people who read about the pandemic more than others showed slightly decreased levels of positive affect, and people who actively posted about the pandemic more than others showed slightly increased levels of negative affect.
--->
-<!-- On the other hand, however, people who posted more about COVID-19 also showed slightly higher levels of positive affects, so taken together the results are ambivalent. -->
-
-<!-- Finally, another interesting observation is that life satisfaction was remarkably stable. -->
-<!-- Hence, even in times of a pandemic, it seems that such broad assessment of life vary only mildly. -->
-<!-- This supports the hypothesis that life satisfaction seems to be determined largely by stable factors such as one's genes [@brownEasyHappinessPie2019]. -->
-
-## Limitations
-
-The current study analyzed whether changes in media use were related to changes in well-being, while controlling for several potential confounders. 
-Together, this allowed for an improved perspective on assessing causality.
-**However, the opposite effect is still also plausible, namely that well-being affected media use [@zillmannMoodManagementCommunication1988].**
-**While controlling for potential confounders can support claims of causality, the procedure implemented here cannot prove causality.**
-Causality necessitates temporal order, and the cause needs to precede the effect.
-**The challenge is that** regarding media use, such effects often happen immediately or shortly after use, necessitating intervals in the hours, minutes, or even seconds.
-In many cases only experience sampling studies asking users at the very moment can produce such knowledge.
-However, even then we don't know for certain if we actually measured the right interval.
-Effects depend on the intensity of use or the length of the interval.
-To borrow the words from @rohrerTheseAreNot2021, there is no such thing as "the" effect of social media use on well-being.
-Hence, to document how effects unfold, future research needs to employ different study designs probing different intervals.
-In addition, more thought needs to be invested in what relevant stable and varying factors we should include as control variables, and I hope this study provides a first step into this direction.
-
-Although I had already reduced the predefined SESOIs to be less conservative, they were potentially still too large.
-Media use is only one aspect of several factors that simultaneously affect well-being.
-Is it really realistic to expect that extremely changing only _one_ of these aspects should manifest in a detectable change in well-being?
-Or would it make more sense to expect that thoroughly committing to say _two_ activities (e.g.
-regularly exercising _and_ establishing a reading habit) should then cause a detectable improvement in well-being?
-Practically, this would imply a SESOI half as large as I have defined here, namely _b_ = |.15| for life satisfaction and _b_ = |.075| for affect.
-In the case of this study, however, reducing the SESOI would not even make a big difference, as also with these more liberal thresholds all but two effects would still be completely in the null region, and no effect would be outside of the null region.
-However, at all events I encourage future research to start a thorough conversation on what effect sizes are considered meaningful and what not.
-Again, with this study I hope to provide some first input and guidelines.
-
-Both media use and well-being were measured using self-reports.
-Measuring well-being with self-reports is adequate, because it by definition requires introspection.
-However, it would be preferable to measure social media use objectively, as people cannot reliably estimate their use [@scharkowAccuracySelfreportedInternet2016].
-That said, objective measures often cannot capture the content or the motivation of the use, and only very complicated tools recording the actual content (such as the Screenome project) might produce such data.
-Unfortunately, such procedures introduce other problems, especially related to privacy.
-Hence, for this type of research question it still seems necessary to use self-reported measures, and in many cases they can still be very informative [@verbeijSelfreportedMeasuresSocial2021].
-
-Because the data were collected in a single country, the generalizability of the results is limited.
-The results apply primarily to the more Western sphere, and might not hold true in other cultures, especially cultures with a different media landscape or alternative social media channels.
-That said, because this is a comparatively large study largely representative of an entire country, and because several waves were collected across a large time span, the results should be at least as generalizable as other typical empirical studies collected in the social sciences.
-
-<!-- Social media use was measured with an ordinal variable, however in the analyses it was treated as a numerical one. -->
-<!-- If treated as an ordinal one, it would have been necessary to analyze four different contrasts for each media measure, which plus the differentiation between between and within factor would have produced eight different measures, we would have made the model exceedingly complex.
--->
-
-## Conclusion
-
-In this study, COVID-19 related social media use didn't meaningfully affect several indicators of well-being, including life satisfaction, positive affect, and negative affect.
-**If people wrote more COVID-19 related posts than usual, or if they spent less time on Instagram and more time on YouTube, very small but statistically significant effects were found.**
-Notably, however, factors other than social media use were more meaningfully related to well-being, such as physical health, exercise, satisfaction with democracy, or believing that one is in control of one's life.
-If it's our aim to improve well-being **during a pandemic**, it might hence be more fruitful not to focus so much on social media but to address other, more pertinent societal problems related to health care, regular exercise, or a functioning democratic system.
-
-\newpage
-
-# References
-
-<!-- \begingroup -->
-<!-- \setlength{\parindent}{-0.5in} -->
-<!-- \setlength{\leftskip}{0.5in} -->
-<div id = "refs"></div>
-<!-- \endgroup -->
-
-\newpage
-
-```{r tab-descriptives, cache=F, warning=F, message=F, prompt=F}
-apa_table(tab_desc_dat,
-          stub_indents = list(
-            "Well-being" = c(1:3),
-            "Social media use" = c(4:6),
-            "Social media channel" = c(7:11)),
-          caption = "Descriptives of the main variables."
-          )
-```
-
-\newpage
-
-```{r tab-within, cache=F, warning=F, message=F, prompt=F}
-tab_within %<>%
-  select(Predictor = iv,
-         b = estimate,
-         `Lower level` = conf.low,
-         `Higher level` = conf.high,
-         beta = "std",
-         p = "p.value")
-apa_table(tab_within,
-          align = c("l", "r", "r", "r", "r", "r"),
-          stub_indents = list(
-            "Life satisfaction" = c(1:8),
-            "Positive affect" = c(9:16),
-            "Negative affect" = c(17:24)),
-          caption = "Overview of all within-person effects.",
-          col_spanners = list(`Confidence interval` = c(3, 4)),
-          font_size = "footnotesize"
-          )
-```
-
-\newpage
+For an illustration, see Figure \@ref(fig:sesoi).
 
 ```{r sesoi, warning=F, message=F, fig.cap="Using confidence intervals to test a null region.
-Note.
-Here, a trivial effect of social media use on life satisfaction is defined as ranging from b = -.30 to b = .30", fig.width=6, fig.height=2, warning=F, message=F, prompt=F}
+In this study, a trivial effect of social media use on life satisfaction is defined as ranging from b = -.30 to b = .30", fig.width=6, fig.height=2, warning=F, message=F, prompt=F}
 # make table with data
@@
 
-\newpage
+<!-- Responses were individually checked for patterns such as straight-lining or missing of inverted items. -->
+<!-- X clear cases were removed. -->
+
+## Data Analysis
+### Causality
+
+<!-- The hypothesis explicitly states a causal effect. -->
+<!-- In non-experimental studies, longitudinal designs can help investigate causality. -->
+When using longitudinal designs to analyze causality, it is important to (a) focus on within-person effects [@hamakerWhyResearchersShould2014]; to (b) control for confounders [@rohrerTheseAreNot2021]; and to (c) test a plausible interval between measures [@dormannOptimalTimeLags2015].
+<!-- To illustrate, consider the following example. -->
+<!-- Imagine that a person suddenly starts using social media much more than usual, and then after some time  becomes less satisfied with their life. -->
+<!-- Eventually, use and life satisfaction return to prior levels. -->
+<!-- If this happens to several people at the same time, in a longitudinal study we could then observe a significant effect of social media use on life satisfaction. -->
+<!-- However, it could also be the case that during the study there was a major exogenous event (say, a pandemic), which caused large parts of the working population to loose their jobs. -->
+<!-- Hence, the causal effect reported above was confounded, because in reality it was the pandemic that caused both social media use to rise and life satisfaction to go down. -->
+<!-- Thus, only when controlling for _all_ relevant confounders, can we correctly estimate causality without bias [@rohrerThinkingClearlyCorrelations2018]. -->
+<!-- Obviously, we can never be entirely sure to have included all confounders, which makes absolute statements regarding causality virtually impossible. -->
+To elaborate, in non-experimental designs it makes much sense to analyze causal effects from an internal, within-person perspective [@hamakerWhyResearchersShould2014].
+If a specific person changes their media diet, we need to measure how this behavior affects their well-being.
+Between-person comparisons from longitudinal data cannot provide such insights [@lucasWhyCrossLaggedPanel2022].
+To test the hypothesis, I thus consider only the within-person effects.
+
+Second, to identify confounders we should control for variables that affect both media use and well-being, which helps isolate the actual effect of social media use on well-being [@rohrerThinkingClearlyCorrelations2018].
+Because we are adopting a within-person perspective, we need to implement _time-varying_ confounders [@rohrerTheseAreNot2021]. 
+And because we are determining the _overall_ causal effect, we need to make sure _not_ to control for mediating variables [@rohrerThinkingClearlyCorrelations2018], for doing so would bias our assessment of the causal effect.
+<!-- Complicating matters further, it is often unclear if a variable is a mediator or a confounder.[^collider] -->
+<!-- However, despite all these caveats, when controlling for relevant variables (that aren't mediators), we can be much more certain that we measured causality correctly. -->
+<!-- The aim should therefore be to collect as many varying and non-varying confounders as possible (which I believe is seldom done in our field), while knowing that absolute certainty regarding causality cannot be reached.
+-->
+<!-- [^collider]: In addition, there also exist colliders, which I don't discuss here and which complicate the issue even further [@rohrerThinkingClearlyCorrelations2018]. -->
+<!-- We can also control for variables that affect only social media use or well-being.  -->
+<!-- However, in doing so not much is gained, because the effects of social media use would remain virtually the same [@klinePrinciplesPracticeStructural2016; but see @mcelreathYesterdayClass2021].
+-->
+In this study, I hence preregistered to control for the following variables, which either have already been shown or are very likely to affect both social media use and well-being, and which also are not mediators:
+<!-- (I'll additionally include variables that likely affect only well-being, also to obtain a comparison benchmark for social media effects): -->
+gender, age, education, Austria country of birth, Austria country of birth of parents, residency Vienna, text-based news consumption, video-based news consumption, household size, health, living space, access to garden, access to balcony, employment, work hours per week, being in home-office, household income, outdoor activities, satisfaction with democracy, disposition to take risks, and locus of control.
+<!-- [^explore] -->
+<!-- I will not control for variables such as trust in institutions or trust in media, because these variables might be influenced by social media use to a meaningful extent. -->
+
+<!-- [^explore]: The data-set includes many other variables that one could also potentially control for, and I invite interested readers to download the and explore potential interesting relationships.
+-->
+
+<!-- Next to including covariates, it is now increasingly understood that  -->
+Finally, one precondition of causality is temporal order and finding a plausible interval [@dormannOptimalTimeLags2015].
+<!-- The cause needs to precede the effect. -->
+<!-- Finding the right interval between cause and effect is crucial. -->
+<!-- For example, if we want to understand the effect of alcohol consumption on driving performance, it makes a big difference if driving performance is measured one minute, one hour, one day, or one week after consumption.
+-->
+<!-- Finding the right interval is difficult. -->
+If variables are stable, longer intervals are needed; if they fluctuate, shorter intervals are required.
+In the case of well-being, we need shorter intervals for the more fluctuating positive and negative affect, and longer ones for the more stable life satisfaction [@dienlinImpactDigitalTechnology2020].
+Using social media can have instant effects on mood [@marcianoDynamicsAdolescentsSmartphone2022].
+Effects on life satisfaction often take longer to manifest, for example because media use leads to actual changes in specific behaviors, which then in turn affect life satisfaction [@dienlinDisplacementReinforcementReciprocity2017].
+<!-- Choosing the right interval is challenging, because especially short intervals are hard to implement in practice, often requiring advanced methods such as experience sampling (also known as in situ measurement or ambulant assessment) [@schnauber-stockmannMobileDevicesTools2020].
+-->
+In this study, I hence analyze how using social media during the last week affected positive and negative affect during the same week.
+In other words, if people during the last week engaged in more COVID-19 related social media use than usual, did they feel better or worse during that week than usual?
+For life satisfaction, I implemented a longer interval.
+If people during the last week used COVID-19 related social media more than they usually do, were they at the end of the week more or less satisfied with their lives than they usually are?
+I hence analyze if when a person changes their social media diet, are there (a) _simultaneous_ changes in their affect and (b) _subsequent_ changes in their life satisfaction?
+These relations will be controlled for varying confounders, which fosters a causal interpretation.
+<!-- Similar approaches were implemented by other studies [@johannesNoEffectDifferent2022; @scharkowHowSocialNetwork2020], and they are considered a best practice approach toward analyzing causality [@bellFixedRandomEffects2019].
+-->
+
+### Statistical model
+
+The hypothesis was analyzed using random effect within-between models [REWB, @bellFixedRandomEffects2019]. 
+Altogether three models were run, one for each dependent variable.
+The data were hierarchical, and responses were separately nested in participants and waves (i.e., participants and waves were implemented as random effects).
+Nesting in participants allowed to separate between-person relations from within-person effects.
+Nesting in waves allowed to control for general exogenous developments, such as general decreases in well-being in the population, for example due to lockdown measures.
+Thus, there was no need additionally to control for specific phases or measures of the lockdown.
+Predictors were modeled as fixed effects.
+They included social media communication types and channels, separated into within and between-person factors, as well as stable and varying covariates.
+All predictors were included simultaneously in each of the three models.
+
+The factorial validity of the scales were tested with confirmatory factor analyses (CFA).
+Because Mardia’s test showed that the assumption of multivariate normality was violated, I used the more robust Satorra-Bentler scaled and mean-adjusted test statistic (MLM) as estimator.
+<!-- To avoid over-fitting, I tested the scales on more liberal fit criteria (CFI > .90, TLI > .90, RMSEA <. .10, SRMR < .10) [@klinePrinciplesPracticeStructural2016]. -->
+<!-- Finally, REWB-models cannot model latent variables. -->
+<!-- To increase precision, I therefore exported factor scores from the CFAs for positive and negative affect. -->
+Mean scores were used for positive and negative affect.
+<!-- Respondents who answered less than 50% of all questions were removed. -->
+Missing responses were imputed using multiple imputation with predictive mean matching (five iterations, five data-sets), including categorical variables.
+All variables were imputed except the social media use measures, as they were not collected on each wave.
+All variables included in the analyses presented here were used to impute missing data.
+<!-- For the main analyses, results were pooled across all five data-sets. -->
+
+### Robustness-checks
+
+To find out whether the inferences were robust across plausible but subordinate analyses, I conducted additional exploratory analyses. 
+I reran the analyses (a) with additional not-preregistered covariates such as trust in media or government, (b) without covariates, (c) with single imputation, and (d) without imputation.
+For more information on the analyses, a complete documentation of the models and results, and all additional analyses, see [companion website](https://XMtRA.github.io/Austrian_Corona_Panel_Project).
+<!-- The additional analyses are reported on the [companion website](https://XMtRA.github.io/Austrian_Corona_Panel_Project/analyses_additional.html). -->
+<!-- For example, all within-person standardized COVID-19 related types of social media use or channels were significantly smaller than $\beta$ = |.05|, again supporting that effects were negligible.
+-->
+
+## Measures
+
+<!-- In what follows, I list all the variables that I analyzed. -->
+For the variables' means, range, and variance, see Table \@ref(tab:tab-descriptives).
+For a complete list of all items and item characteristics, see [companion website](https://XMtRA.github.io/Austrian_Corona_Panel_Project).
+
+### Well-being
+
+Life satisfaction was measured with the item "All things considered, how satisfied are you with your life as a whole nowadays?", which comes from the European Social Survey [@europeansocialsurveyESS9Edition20182021].
+The response options ranged from 0 (_extremely dissatisfied_) to 10 (_extremely satisfied_).
+
+To capture positive affect, respondents were asked how often in the last week they felt (a) calm and relaxed, (b) happy, and (c) full of energy [@worldhealthorganizationWellbeingMeasuresPrimary1998].
+The response options were 1 (_never_), 2 (_on some days_), 3 (_several times per week_), 4 (_almost every day_), and 5 (_daily_).
+The scale showed good factorial fit, `r fit_txt(cfa_aff_pos)`.
+Reliability was high, $\omega$ = `r rel_aff_pos %>% my_round("std")`.
+
+For negative affect, respondents were asked how often in the last week they felt (a) lonely, (b) aggravated, (c) so depressed, that nothing could lift you up, (d) very nervous, (e) anxious, and (h) glum and sad [@worldhealthorganizationWellbeingMeasuresPrimary1998].
+The response options were 1 (_never_), 2 (_on some days_), 3 (_several times per week_), 4 (_almost every day_), and 5 (_daily_).
+The scale showed good factorial fit, `r fit_txt(cfa_aff_neg)`.
+Reliability was high, $\omega$ = `r rel_aff_neg %>% my_round("std")`.
+
+All three variables were measured on each wave.
+
+### COVID-19 related social media use
+
+COVID-19 related social media use focused on communication types was measured with the three dimensions of (a) reading, (b) liking and sharing, and (c) posting.
+The items come from @wagnerAUTNESOnlinePanel2018 and were adapted for the context of this study.
+The general introductory question was "How often during the last week have you engaged in the following activities on social media?".
+The three items were "Reading the posts of others with content on the Coronavirus", "When seeing posts on the Coronavirus, I clicked 'like', 'share' or 'retweet'", "I myself wrote posts on the Coronavirus on social media."
+Answer options were 1 (_several times per day_), 2 (_daily_), 3 (_several times per week_), 4 (_weekly_), 5 (_never_).
+The items were inverted for the analyses.
+
+COVID-19 related social media use focused on channels was measured with five variables from @wagnerAUTNESOnlinePanel2018, adapted for this study. 
+The general introductory question was "How often in the last week have you followed information related to the Corona-crisis on the following social media?"
+The five items were (a) Facebook, (b) Twitter, (c) Instagram, (d) Youtube, and (e) WhatsApp.
+Again, the answer options were 1 (_several times per day_), 2 (_daily_), 3 (_several times per week_), 4 (_weekly_), 5 (_never_).
+Again, the items were inverted for the analyses.
+
+Social media use was measured for all participants on waves 1, 2, 8, 17, 23, and 28.
+Freshly recruited respondents always answered all questions on COVID 19-related social media use.
+For the analyses I hence used all 32 waves.
+
+```{r tab-descriptives, cache=F, warning=F, message=F, prompt=F}
+apa_table(tab_desc_dat,
+          stub_indents = list(
+            "Well-being" = c(1:3),
+            "Social media use" = c(4:6),
+            "Social media channel" = c(7:11)),
+          caption = "Descriptives of the main variables."
+          )
+```
+
+### Control variables
+
+The effects of COVID-19 related social media use were controlled for the following stable variables: 
+gender (female, male, diverse), age, education (ten options), Austria country of birth (yes/no), Austria parents' country of birth (no parent, one parent, both parents), and household size.
+<!-- I originally planned to implement additional variables as varying covariates. -->
+I also controlled for the following varying covariates: five items on current living conditions, including self-reported physical health, whether participants contracted COVID-19 since the last wave, current household income, working in home-office, and overall work hours; nine items measuring use of specific national text-based and video-based news outlets; five items measuring outdoor activities such as exercise or meeting friends; and three more psychological measures including locus of control, disposition to take risks, and satisfaction with democracy.
+
+# Results
+## Descriptive Analyses
+
+Looking at the variables from a descriptive perspective, aligned with set-point theory we can see that the level of all well-being measures were surprisingly stable during data collection (see Figure \@ref(fig:fig-descriptives)).
+COVID-19 related social media use, however, showed changes. 
+Reading, sharing and liking COVID-19 related content decreased substantially (almost one scale point from 3 to 2). 
+Posting about COVID-19 related content stayed the same.
+Using Facebook, WhatsApp, and YouTube for COVID-19 related content decreased. 
+Instagram stayed the same.
+Twitter use slightly increased.
+The general initial decrease could be explained by the fact that the collection of data began at the end of March 2020, hence approximately three months after the pandemic's onset.
+After an initial uptick, COVID-19 related social media use might have already been declining at the time.
 
@@
 
-\newpage
+Using the average values across all waves, which provides a stable picture of the general relations, I next looked at the correlations between social media use and well-being (see Figure \@ref(fig:fig-correlations)).
+Several interesting patterns emerged.
+In general, people who spend more time engaging with COVID-19 related content on social media reported reduced well-being.
+Users who spend more time reading, liking and sharing, and posting COVID-19 related content were less satisfied with their lives.
+They also showed slightly less positive affect.
+This overall negative picture was even more pronounced for negative affect.
+People who engaged more with COVID-19 related content, including all types and channels of communication, reported substantially higher levels of negative affect.
+For example, people who were more likely to post COVID-19 content had much higher levels of negative affect (_r_ = `r tab_cor[tab_cor$term == "Posting", "Affect.negative"]`).
+Note that these results represent between-person correlations, not causal within-person effects.
+
+```{r fig-correlations, fig.cap="Descriptives of the main variables, capturing well-being and social media use with their average values across all waves.
+Upper triangle: correlation coefficients; diagonal: density plots; lower triangle: scatter plots.", out.width = "\\textwidth", cache=F, warning=F, message=F, prompt=F}
+# fig_cor
+knitr::include_graphics("figures/fig_cor.png")
+```
+
+## Preregistered Analyses
+### Social media communication types
+
+The study's main hypothesis was that the causal effects of all types and channels of social media use on all facets of well-being would be trivial.
+Regarding the effects of different communication types (i.e., reading, sharing, of posting about COVID-19 related content), all within-person effects fell completely within the a-priori defined null region (see Figure \@ref(fig:fig-within)).
+For example, respondents who used social media more frequently than usual to like or share COVID-19 related content did not show a simultaneous change in life satisfaction (_b_ = `r data_tab_within %>% filter(dv == "Life satisfaction" & iv == "Liking & Sharing" & Variance == "within" & Analysis == "Publication") %>% select(estimate) %>% round(2)` [95% CI `r data_tab_within %>% filter(dv == "Life satisfaction" & iv == "Liking & Sharing" & Variance == "within" & Analysis == "Publication") %>% select(conf.low) %>% round(2)`, `r data_tab_within %>% filter(dv == "Life satisfaction" & iv == "Liking & Sharing" & Variance == "within" & Analysis == "Publication") %>% select(conf.high) %>% round(2)`]).
+As a result, the hypothesis of trivial effects was supported for all COVID-19 related types of social media communication.
+
+However, three effects stood out, as they were significantly different from zero.
+Users who read more COVID-19 related content than usual reported slightly reduced levels of positive affect (_b_ = `r data_tab_within %>% filter(dv == "Positive affect" & iv == "Reading" & Variance == "within" & Analysis == "Publication") %>% select(estimate) %>% round(2)` [95% CI `r data_tab_within %>% filter(dv == "Positive affect" & iv == "Reading" & Variance == "within" & Analysis == "Publication") %>% select(conf.low) %>% round(2)`, `r data_tab_within %>% filter(dv == "Positive affect" & iv == "Reading" & Variance == "within" & Analysis == "Publication") %>% select(conf.high) %>% round(2)`]).
+Users who liked and shared more COVID-19 related content than usual also experienced slightly more negative affect than usual (_b_ = `r data_tab_within %>% filter(dv == "Negative affect" & iv == "Posting" & Variance == "within" & Analysis == "Publication") %>% select(estimate) %>% round(2)` [95% CI `r data_tab_within %>% filter(dv == "Negative affect" & iv == "Posting" & Variance == "within" & Analysis == "Publication") %>% select(conf.low) %>% round(2)`, `r data_tab_within %>% filter(dv == "Negative affect" & iv == "Posting" & Variance == "within" & Analysis == "Publication") %>% select(conf.high) %>% round(2)`]).
+Users who wrote more COVID-19 related posts than usual also experienced slightly more negative affect than usual (_b_ = `r data_tab_within %>% filter(dv == "Negative affect" & iv == "Posting" & Variance == "within" & Analysis == "Publication") %>% select(estimate) %>% round(2)` [95% CI `r data_tab_within %>% filter(dv == "Negative affect" & iv == "Posting" & Variance == "within" & Analysis == "Publication") %>% select(conf.low) %>% round(2)`, `r data_tab_within %>% filter(dv == "Negative affect" & iv == "Posting" & Variance == "within" & Analysis == "Publication") %>% select(conf.high) %>% round(2)`]).
+
+### Social media communication channels
+
+Regarding the COVID-19 related use of social media channels (i.e., Facebook, Instagram, WhatsApp, YouTube, and Twitter) the results were comparable (see Figure \@ref(fig:fig-within)).
+Changes in the frequency of using different social media channels to attain information regarding COVID-19 were unrelated to meaningful changes in well-being.
+For example, respondents who used Facebook more frequently than usual to learn about COVID-19 did not show a simultaneous change in life satisfaction (_b_ `r data_tab_within %>% filter(dv == "Life satisfaction" & iv == "Facebook" & Variance == "within" & Analysis == "Publication") %>% select(estimate) %>% my_round("b")` [95% CI `r data_tab_within %>% filter(dv == "Life satisfaction" & iv == "Facebook" & Variance == "within" & Analysis == "Publication") %>% select(conf.low) %>% round(2)`, `r data_tab_within %>% filter(dv == "Life satisfaction" & iv == "Facebook" & Variance == "within" & Analysis == "Publication") %>% select(conf.high) %>% round(2)`]).
+In sum, the hypothesis of trivial effects was supported also for the COVID-19 related use of important social media channels.
+
+That said, one effect differed substantially from zero.
+Respondents who used Twitter more frequently than usual to attain COVID-19 related news reported slightly higher levels of negative affect than usual (_b_ = `r data_tab_within %>% filter(dv == "Negative affect" & iv == "Twitter" & Variance == "within" & Analysis == "Publication") %>% select(estimate) %>% round(2)` [95% CI `r data_tab_within %>% filter(dv == "Negative affect" & iv == "Twitter" & Variance == "within" & Analysis == "Publication") %>% select(conf.low) %>% my_round("b")`, `r data_tab_within %>% filter(dv == "Negative affect" & iv == "Twitter" & Variance == "within" & Analysis == "Publication") %>% select(conf.high) %>% my_round("b")`]).
+However, the effect was still completely inside of the null region, hence likely not large enough to be considered meaningful.
+
+For an overview of all within-person effects, see Table \@ref(tab:tab-within) and Figure \@ref(fig:fig-within).
 
@@
 
-\newpage
-
-```{r fig-between, fig.cap="Between-person relations between COVID-19 related social media use and well-being.
-Note.
-The black estimates show the effects controlled for a large number of covariates (see text; preregistered); the grey estimates are without control variables (exploratory).
-The SESOI was b = |0.30| for life satisfaction and b = |0.15| for affect.
-Hence, all of the reported effects are not considered meaningful.", out.width = "\\textwidth", cache=F, warning=F, message=F, fig.width=6, fig.height=4}
-fig_results_between
-# knitr::include_graphics("figures/fig_results_between.png")
+```{r tab-within, cache=F, warning=F, message=F, prompt=F}
+apa_table(tab_within %>% select(-Outcome),
+          align = c("l", "r", "r", "r", "r", "r"),
+          stub_indents = list(
+            "Life satisfaction" = c(1:8),
+            "Positive affect" = c(9:16),
+            "Negative affect" = c(17:24)),
+          caption = "Overview of all within-person effects.",
+          col_spanners = list(`Confidence interval` = c(3, 4)),
+          font_size = "footnotesize"
+          )
 ```
 
-\newpage
+## Exploratory Analyses
 
-```{r fig-control, fig.cap = "Results of selected covariates. Note. All variables standardized except 'Male'.", out.width = "\\textwidth", cache=F, warning=F, message=F, fig.width=6, fig.height=4}
+<!-- In what follows, I briefly report some exploratory analyses that weren't preregistered. -->
+
+<!-- ### Between-person relations -->
+
+<!-- For between-person relations, no a-priori hypotheses were formulated. -->
+<!-- Results showed that no relation crossed or was completely outside of the SESOI. -->
+<!-- Four relations were statistically significant. -->
+<!-- Respondents who across all waves used social media more frequently than others to read about COVID-19 related posts reported slightly lower levels of positive affect than others (_b_ `r dat_fig_results_between %>% filter(dv == "Positive affect" & iv == "Reading" & type == "between" & Analysis == "Regular") %>% select(estimate) %>% my_round("b_txt")` [95% CI `r dat_fig_results_between %>% filter(dv == "Positive affect" & iv == "Reading" & type == "between" & Analysis == "Regular") %>% select(conf.low) %>% my_round("b")`, `r dat_fig_results_between %>% filter(dv == "Positive affect" & iv == "Reading" & type == "between" & Analysis == "Regular") %>% select(conf.high) %>% my_round("b")`]).
+-->
+<!-- Respondents who across all waves used social media more frequently than others to write COVID-19 related posts reported slightly higher levels of negative affect than others (_b_ `r dat_fig_results_between %>% filter(dv == "Negative affect" & iv == "Posting" & type == "between" & Analysis == "Regular") %>% select(estimate) %>% my_round("b_txt")` [95% CI `r dat_fig_results_between %>% filter(dv == "Negative affect" & iv == "Posting" & type == "between" & Analysis == "Regular") %>% select(conf.low) %>% my_round("b")`, `r dat_fig_results_between %>% filter(dv == "Negative affect" & iv == "Posting" & type == "between" & Analysis == "Regular") %>% select(conf.high) %>% my_round("b")`]).
+-->
+<!-- At the same time, respondents who across all waves used social media more frequently than others to write COVID-19 related posts also reported slightly higher levels of positive affect (_b_ `r dat_fig_results_between %>% filter(dv == "Positive affect" & iv == "Posting" & type == "between" & Analysis == "Regular") %>% select(estimate) %>% my_round("b_txt")` [95% CI `r dat_fig_results_between %>% filter(dv == "Positive affect" & iv == "Posting" & type == "between" & Analysis == "Regular") %>% select(conf.low) %>% my_round("b")`, `r dat_fig_results_between %>% filter(dv == "Positive affect" & iv == "Posting" & type == "between" & Analysis == "Regular") %>% select(conf.high) %>% my_round("b")`]).
+-->
+<!-- Finally, respondents who across all waves used YouTube more frequently than others also reported slightly higher levels of life satisfaction than others (_b_ `r dat_fig_results_between %>% filter(dv == "Life satisfaction" & iv == "YouTube" & type == "between" & Analysis == "Regular") %>% select(estimate) %>% my_round("b_txt")` [95% CI `r dat_fig_results_between %>% filter(dv == "Life satisfaction" & iv == "YouTube" & type == "between" & Analysis == "Regular") %>% select(conf.low) %>% my_round("b")`, `r dat_fig_results_between %>% filter(dv == "Life satisfaction" & iv == "YouTube" & type == "between" & Analysis == "Regular") %>% select(conf.high) %>% my_round("b")`]).
+-->
+<!-- However, note that the effect were still completely inside of the null region, hence not large enough to be considered practically relevant. -->
+
+<!-- Note that when comparing the results with and without control variables, the results differed. -->
+<!-- For example, on the between-person level, one effect stopped being significant if controlled for additional variables. -->
+<!-- Initially, actively posting on social media was significantly (though not meaningfully) related to decreased life satisfaction. -->
+<!-- However, when controlling for potential confounders, the effect became virtually zero. -->
+
+<!-- For an overview of all between-person relations, see Figure \@ref(fig:fig-between). -->
+
+<!-- ### Covariates -->
+
+To contextualize the results reported above and to see if the study included any meaningful effects at all, I also looked at the effect sizes of the covariates.
+Because each variable featured different response options, which would require defining a SESOI for each variable, I hence report the results of the standardized scales, which allows for a better comparison across the differently scaled variables.
+Here, we can build on Cohen's convention that small effects begin at _r_ = |.10|.
+
+The results showed that several effects crossed or fell completely outside of the SESOI. 
+They can hence be considered meaningful.
+For example, if physical health decreased, this had a meaningful detrimental impact on life satisfaction ($\beta$ `r data_tab_comp %>% filter(dv == "Life satisfaction" & iv == "Health") %>% select(estimate) %>% my_round("std_txt")` [95% CI `r data_tab_comp %>% filter(dv == "Life satisfaction" & iv == "Health") %>% select(conf.low) %>% my_round("std")`, `r data_tab_comp %>% filter(dv == "Life satisfaction" & iv == "Health") %>% select(conf.high) %>% my_round("std")`]), positive affect ($\beta$ `r data_tab_comp %>% filter(dv == "Positive affect" & iv == "Health") %>% select(estimate) %>% my_round("std_txt")` [95% CI `r data_tab_comp %>% filter(dv == "Positive affect" & iv == "Health") %>% select(conf.low) %>% my_round("std")`, `r data_tab_comp %>% filter(dv == "Positive affect" & iv == "Health") %>% select(conf.high) %>% my_round("std")`]), and negative affect ($\beta$ `r data_tab_comp %>% filter(dv == "Negative affect" & iv == "Health") %>% select(estimate) %>% my_round("std_txt")` [95% CI `r data_tab_comp %>% filter(dv == "Negative affect" & iv == "Health") %>% select(conf.low) %>% my_round("std")`, `r data_tab_comp %>% filter(dv == "Negative affect" & iv == "Health") %>% select(conf.high) %>% my_round("std")`]).
+Spending more time outside to exercise meaningfully increased positive affect ($\beta$ `r data_tab_comp %>% filter(dv == "Positive affect" & iv == "Sports") %>% select(estimate) %>% my_round("std_txt")` [95% CI `r data_tab_comp %>% filter(dv == "Positive affect" & iv == "Sports") %>% select(conf.low) %>% my_round("std")`, `r data_tab_comp %>% filter(dv == "Positive affect" & iv == "Sports") %>% select(conf.high) %>% my_round("std")`]).
+Being more satisfied with the current democratic system meaningfully increased life satisfaction ($\beta$ `r data_tab_comp %>% filter(dv == "Life satisfaction" & iv == "Satisfaction democracy") %>% select(estimate) %>% my_round("std_txt")` [95% CI `r data_tab_comp %>% filter(dv == "Life satisfaction" & iv == "Satisfaction democracy") %>% select(conf.low) %>% my_round("std")`, `r data_tab_comp %>% filter(dv == "Life satisfaction" & iv == "Satisfaction democracy") %>% select(conf.high) %>% my_round("std")`]).
+The strongest aspect affecting well-being was internal locus of control.
+If people felt more in control of their lives, this strongly increased both life satisfaction ($\beta$ `r data_tab_comp %>% filter(dv == "Life satisfaction" & iv == "Internal locus of control") %>% select(estimate) %>% my_round("std_txt")` [95% CI `r data_tab_comp %>% filter(dv == "Life satisfaction" & iv == "Internal locus of control") %>% select(conf.low) %>% my_round("std")`, `r data_tab_comp %>% filter(dv == "Life satisfaction" & iv == "Internal locus of control") %>% select(conf.high) %>% my_round("std")`]) and
+positive affect ($\beta$ `r data_tab_comp %>% filter(dv == "Positive affect" & iv == "Internal locus of control") %>% select(estimate) %>% my_round("std_txt")` [95% CI `r data_tab_comp %>% filter(dv == "Positive affect" & iv == "Internal locus of control") %>% select(conf.low) %>% my_round("std")`, `r data_tab_comp %>% filter(dv == "Positive affect" & iv == "Internal locus of control") %>% select(conf.high) %>% my_round("std")`]),
+while decreasing negative affect ($\beta$ `r data_tab_comp %>% filter(dv == "Negative affect" & iv == "Internal locus of control") %>% select(estimate) %>% my_round("std_txt")` [95% CI `r data_tab_comp %>% filter(dv == "Negative affect" & iv == "Internal locus of control") %>% select(conf.low) %>% my_round("std")`, `r data_tab_comp %>% filter(dv == "Negative affect" & iv == "Internal locus of control") %>% select(conf.high) %>% my_round("std")`]).
+
+For an overview, see Figure \@ref(fig:fig-control).
+
+```{r fig-control, fig.cap = "Results of main variables together with covariates to provide context.
+All variables standardized.
+SESOI: beta = |.10|", out.width = "\\textwidth", cache=F, warning=F, message=F, fig.width=7, fig.height=8}
 fig_results_control_std
@@
 
+# Discussion
+
+Based on a panel study with 32 waves largely representative of the Austrian population, this study analyzed the effects of COVID-19 related social media use on well-being.
+Between person correlation analyses showed that more active users of COVID-19 related content on social media also reported decreased well-being. 
+The within-person relations, which are more informative regarding causal effects, showed a different pattern, however.
+If people consumed more COVID-19 content on social media than usual, this did not meaningfully reduce their well-being.
+Several statistically significant effects were found, but these were very small.
+For example, people who liked and shared more COVID-19 related posts than usual reported slightly higher levels of negative affect.
+Posting more content about COVID-19 than usual slightly increased negative affect.
+People who read more COVID-19 related posts than usual reported slightly decreased positive affect.
+Using Twitter for COVID-19 related content slightly increased negative affect. 
+<!-- was related to Reading COVID-19 related content on social media slightly reduced positive affect. -->
+Again, although all of these within-person effects were statistically significant, they were very small, smaller than the predefined smallest effect size of interest.
+According to the preregistered procedure, they should hence be considered irrelevant. 
+Other factors, for which we would expect to find meaningful effects, such as health or sports, indeed showed substantial and meaningful impacts on well-being. 
+In conclusion, COVID-19 related activity on social media was not a particularly strong influence on peoples' well-being. 
+The results do not support the popular fears that "doomscrolling" or overusing social media during times of crises constitutes a relevant risk for well-being.
+
+<!-- People who used social media more than usual to learn about COVID-19 didn't show meaningful changes in their well-being. -->
+
+<!-- The results imply that COVID-19 related social media use doesn't seem to be particularly relevant for well-being. -->
+<!-- Other factors among the third variables that were measured revealed larger effects or relations, suggesting that well-being is rather determined by alternative aspects such as health, satisfaction with democracy, locus of control, or exercising.
+-->
+
+These specific observations notwithstanding, several general trends can be observed.
+First, overall the results do suggest that effects of COVID-19 related social media use on well-being tend to take place in the negative as opposed to the positive spectrum.
+Although very small, four statistically significant negative results of COVID-19 related social media use on well-being were found.
+Not a single positive effect emerged.
+Also note that in the analyses several control variables were included, ruling out plausible alternative explanations for the negative results.
+For example, it was controlled for whether or not participants contracted a COVID-19 infection during a specific wave.
+Hence, we can rule out the alternative explanation that having an infection was the root cause of increased communication and reduced well-being.
+
+Second, significant outcomes emerged for positive or negative affect, but not for life satisfaction. 
+Life satisfaction was stable and not affected by any type or channel of social media communication.
+The more fluctuating positive and negative affect, however, were affected.
+Liking, sharing, and posting COVID-19 related content, and spending time on Twitter to browse COVID-19 related content, all negatively influenced affect. 
+This is aligned with prior findings which showed that social media can trigger negative affect but does not reduce life satisfaction [@huangTimeSpentSocial2017].
+<!-- Writing posts implies more substantial engagement, and perhaps also negative exchanges with other users, which could explain the results. -->
+<!-- For example, people who wrote more COVID-19 related posts than usual reported slightly lower levels of life satisfaction than usual. -->
+<!-- Similarly, people who wrote more COVID-19 related posts than usual also reported slightly more negative affect. -->
+<!-- As a potential explanation, when writing posts and comments on social media people explicitly and more deeply engage with COVID-19 related content. -->
+Conversations about COVID-19 on social media are often extreme, negative, or aggressive [@fanStigmatizationSocialMedia2020].
+More deeply engaging with this type of content could negatively affect active authors.
+The hypothesis that tonality could explain the negative effects is especially supported by the observation that spending more time on Twitter than usual increased negative affect.
+Communication on Twitter is often found to be more negative and impolite compared to other SNSs [@halpernSocialMediaCatalyst2013].
+Consuming more negative and misleading information could hence explain the (slightly) increased levels of negative affect.
+
+Third, the results show that it makes sense to analyze different communication types and communication channels.
+Reading reduced positive affect, while liking, sharing and commenting increased negative affect.
+Whereas it was often stated that passive use is bad and active use good [@verduynPassiveFacebookUsage2015], this pattern thus did not emerge here.
+Instead, all three types were negative.
+The results hence support the findings from @valkenburgAssociationsActivePassive2022, who also could not confirm the claim that active use is good and that passive use is bad.
+Focusing on communication channels, Twitter seems to be more negative, as has often been observed [@halpernSocialMediaCatalyst2013].
+But, again, all of these effects are very small.
+Future research might elaborate on these specific relations to probe their stability and relevance.
+
+Taken together, the results are hence aligned with the underlying theoretical models and prior empirical results.
+The findings support the Different Susceptibility to Media Effects Theory, which states that effects are generally small and depend on the type of communication.
+The results are well-aligned with mood management theory [@zillmannMoodManagementCommunication1988] or the uses and gratifications approach [@katzUsesGratificationsResearch1973], whose premises preclude particularly negative effects of routine and widespread media consumption.
+If the effects of social media were indeed profoundly negative on average, then people likely would not spend so much time on social media engaging with COVID-19 content.
+Likewise, recent empirical studies and meta-analyses reported rather small negative effects, too.
+Several studies found that the effects of various types of social media use on well-being are small, often too small to matter [@bendauAssociationsCOVID19Related2021; @meierComputermediatedCommunicationSocial2020a; @orbenTeenagersScreensSocial2020; @fergusonThisMetaanalysisScreen2021], echoing the results obtained here.
+
+In light of the overall very small effects, engaging in COVID 19-related social media use should not be a major concern for one's well-being.
+That said, the results still imply that it does make sense to critically reflect upon COVID-19 related social media use. 
+On average, it is likely more beneficial to post less actively about COVID-19 on social media and to spend less time on Twitter consuming COVID-19 related content.
+The results allow us to reject a positive effect: Writing more posts on social media will likely not increase well-being.
+
+<!-- First, in terms of media channels, using Twitter more than usual was related to slightly decreased levels of life satisfaction. -->
+<!-- Twitter is considered to have more negative affordances and tonality as compared to other networks such as Instagram [@waterlooNormsOnlineExpressions2018], which might help explain the results.
+-->
+<!-- Second, of all the three COVID-19 related social media activities, people who read about the pandemic more than others showed slightly decreased levels of positive affect, and people who actively posted about the pandemic more than others showed slightly increased levels of negative affect.
+-->
+<!-- On the other hand, however, people who posted more about COVID-19 also showed slightly higher levels of positive affects, so taken together the results are ambivalent. -->
+
+<!-- Finally, another interesting observation is that life satisfaction was remarkably stable. -->
+<!-- Hence, even in times of a pandemic, it seems that such broad assessment of life vary only mildly. -->
+<!-- This supports the hypothesis that life satisfaction seems to be determined largely by stable factors such as one's genes [@brownEasyHappinessPie2019]. -->
+
+## Limitations
+
+<!-- The current study analyzed whether changes in media use were related to changes in well-being. -->
+Focusing on within-person effects and controlling for several potential confounders, this study provides an improved perspective on assessing causality.
+However, three major challenges remain. 
+First, it could be that the timing of the interval was wrong; it could be that effects unfold in shorter or longer intervals. 
+Next, it could be that there was not sufficient variance in media use and well-being. 
+Without sufficient variability, statistical models cannot detect existing effects.
+Third, it is necessary to control for _all_ relevant confounding third variables. 
+Although this study included are large list of confounders, it could still be that crucial variables were missed  [@zillmannMoodManagementCommunication1988].
+<!-- While controlling for potential confounders can support claims of causality, the procedure implemented here cannot prove causality. -->
+<!-- Causality necessitates temporal order, and the cause needs to precede the effect. -->
+<!-- The challenge is that regarding media use, such effects often happen immediately or shortly after use, necessitating intervals in the hours, minutes, or even seconds. -->
+<!-- In many cases only experience sampling studies asking users at the very moment can produce such knowledge. -->
+<!-- However, even then we don't know for certain if we actually measured the right interval. -->
+<!-- Effects depend on the intensity of use or the length of the interval. -->
+To borrow the words from @rohrerTheseAreNot2021, there is no such thing as "the" effect of social media use on well-being.
+The results reported here are hence contingent on the design I used.
+To document how effects unfold, future research needs to employ different study designs probing different intervals.
+More thought needs to be invested in what factors to control for, and I hope this study provides a first step into this direction.
+
+Although I had already reduced the predefined SESOIs to be less conservative, one could argue they were still too large.
+Media use is only one aspect of several factors that simultaneously affect well-being.
+Is it realistic to expect that changing only _one_ of these aspects should already manifest in a detectable change in well-being?
+Or would it make more sense to expect that thoroughly committing to say _two_ activities (e.g.
+regularly exercising _and_ establishing a reading habit) should then cause a detectable improvement in well-being?
+Practically, this would imply a SESOI half the size defined here, namely _b_ = |.15| for life satisfaction and _b_ = |.075| for affect.
+In the case of this study, however, even halving the SESOI would not make a difference.
+All but one effect would still be completely in the null region, and no effect would fall completely outside of the null region.
+I encourage future research to elaborate on what effect sizes are considered meaningful and what not.
+<!-- Again, with this study I hope to provide some first input and guidelines. -->
+
+Both media use and well-being were measured using self-reports.
+Because assessing well-being necessarily requires introspection, using self-reports for affect and life satisfaction is adequate.
+However, for social media use objective measures are clearly preferable, as people often cannot reliably estimate their use [@scharkowAccuracySelfreportedInternet2016].
+At the same time, most objective measures cannot capture the content or the motivation of use.
+Hence, for the type of research question analyzed here, it still seems necessary to use self-reported measures.
+In many cases they can still be informative [@verbeijSelfreportedMeasuresSocial2021].
+
+Being collected in a single country, the generalizability of the results is limited.
+The results apply primarily to the more Western sphere.
+They might not hold true in other cultures, especially cultures with a different media landscape or alternative social media channels.
+<!-- That said, being a comparatively large study largely representative of one country, with several waves collected across a large time span, the results should be at least as generalizable as other typical empirical studies collected in the social sciences.
+-->
+<!-- Social media use was measured with an ordinal variable, however in the analyses it was treated as a numerical one. -->
+<!-- If treated as an ordinal one, it would have been necessary to analyze four different contrasts for each media measure, which plus the differentiation between between and within factor would have produced eight different measures, we would have made the model exceedingly complex.
+-->
+
+## Conclusion
+
+In this study, COVID-19 related social media use did not meaningfully affect well-being. 
+Very small negative effects were found for writing COVID-19 related posts, sharing COVID-19 related content, and spending more time than usual on Twitter.
+Factors other than social media use, however, were meaningfully related to well-being, including physical health, exercise, satisfaction with democracy, or believing that one is in control of one's life.
+Hence, when trying to improve well-being during a pandemic, instead of focusing on social media it seems more fruitful to address other, more pertinent societal problems related to health care, regular exercise, or a functioning democratic system.
+
 \newpage
+
+# References
+
+<!-- \begingroup -->
+<!-- \setlength{\parindent}{-0.5in} -->
+<!-- \setlength{\leftskip}{0.5in} -->
+<div id = "refs"></div>
+<!-- \endgroup -->
 

### 2023-01-30 manuscript.tex
- thorough update to paper - shortened introduction - added additional exploratory analyses - e.g., analysis of between person vars

--- 2022-04-12 manuscript.tex
@@
 \documentclass[
-  man,mask]{apa6}
+  man,mask,floatsintext]{apa7}
 \usepackage{amsmath,amssymb}
@@
 \keywords{COVID-19, well-being, social media, news use, panel study.}
-\DeclareDelayedFloatFlavor{ThreePartTable}{table}
-\DeclareDelayedFloatFlavor{lltable}{table}
-\DeclareDelayedFloatFlavor*{longtable}{table}
-\makeatletter
-\renewcommand{\efloat@iwrite}[1]{\immediate\expandafter\protected@write\csname efloat@post#1\endcsname{}}
-\makeatother
 \usepackage{lineno}
@@
 \fi
-
-\title{Analyzing the effects of COVID-19 related social media use on well-being}
+\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
+\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
+\urlstyle{same} % disable monospaced font for URLs
+\hypersetup{
+  pdftitle={COVID-19 related social media use slightly decreases well-being},
+  pdflang={en-EN},
+  pdfkeywords={COVID-19, well-being, social media, news use, panel study.},
+  hidelinks,
+  pdfcreator={LaTeX via pandoc}}
+
+\title{COVID-19 related social media use slightly decreases well-being}
 \author{Name blinded for review\textsuperscript{1}}
@@
 
-\shorttitle{COVID-19 related social media use and well-being}
+\shorttitle{COVID-19 related social media use \& well-being}
 
@@
 \abstract{%
-In times of crisis such as the Corona pandemic citizens need to stay informed about recent events, political decisions, or mandatory protection measures.
-To this end, many people use various types of media, and increasingly social media.
-However, because social media are particularly engaging, some find it hard to disconnect.
-In this preregistered study, I investigated whether using social media for COVID-19 related reasons affected psychological well-being.
-To answer this question I analyzed data from the Austrian Corona Panel Project, which consists of 3,018 participants.
-Well-being was measured at all 24 waves, and communication at five specific waves.
-I ran three random effects within between models, controlling for several stable and varying confounders.
-Results showed that the effects of COVID-19 related social media use on well-being were very small, arguably too small to matter.
-Fears that social media use during times of crisis critically impairs well-being are not supported.
+In times of crisis such as the COVID-19 pandemic, citizens need to stay informed about recent political events.
+To this end, people increasingly use social media.
+However, being particularly engaging, many find it hard to disconnect from social media, especially during times of crisis.
+Using data from the Austrian Corona Panel Project consisting of 3,485 participants from 32 waves, controlling for several stable and varying confounders, the results showed that COVID-19 related social media use reduced facets of well-being.
+However, the effects were very small.
+Other factors such as health, income, exercise, or internal locus of control showed larger and more meaningful effects.
 }
@@
 
-During the COVID-19 pandemic,
-numerous events unfolded in quick succession and several open questions emerged.
+During the COVID-19 pandemic it was critical to stay informed regarding the latest developments.
 How dangerous is the virus?
-Is it spreading in my region?
-How is it transmitted, and how can I protect myself?
-Because for many it was (and at the time of writing still is) a matter of life or death, people aimed to stay informed regarding the latest developments.
-Governments around the world implemented safety measures, such as wearing masks, keeping physical distance, or enforcing lockdowns.
-In this extraordinary situation, many people heavily relied on media to attain relevant information, and especially social media were at an all time high (Statista, 2021).
-
-Some people actually couldn't stop using social media to learn about COVID-19 related news.
-A new phenomenon emerged, termed ``doomscrolling'':
-Users were glued to their screens and found it hard to pursue other relevant activities such as working, taking a break, or looking after their children (Klein, 2021).
-It was asked whether using social media for COVID-19 related reasons is helpful, or whether it creates an additional burden on mental health (Sandstrom, Buchanan, Aknin, \& Lotun, 2021).
-These concerns seem justified:
-A study with 6,233 people from Germany conducted during the pandemic found that ``{[}f{]}requency, duration and diversity of media exposure were positively associated with more symptoms of depression'' (Bendau et al., 2021, p.
-283).
-
-As a result, with this study I want to build on this research and investigate whether or not COVID-19 related social media use affected well-being during the pandemic.
-To this end, I analyzed a large-scale panel study from the Austrian Corona Panel Project (Kittel et al., 2020).
-The panel consists of 24 waves and has an overall sample size of 3,018 participants.
-The panel study collected a large number of psychological and demographic variables.
-I explicitly aimed to investigate the causal effects of COVID-19 related social media use on well-being.
+In what region is it spreading?
+How is it transmitted? What are the current safety regulations?
+To obtain relevant information, many people heavily relied on social media, which were at an all time high (Statista, 2021).
+
+Some actually could not stop using social media to learn about COVID-19 related news.
+A new phenomenon termed ``doomscrolling'' emerged.
+Many users were glued to their screens and found it hard to pursue other relevant activities such as working, taking a break, or even looking after their children (Klein, 2021).
+In the media it was hence increasingly asked whether using social media for COVID-19 related reasons would, next to all other stressors, create an additional burden on mental health (Sandstrom et al., 2021).
+Although research has begun addressing this question (Bendau et al., 2021; Bradley \& Howard, 2021; Choi \& Choung, 2021; Dörnemann et al., 2021; Eden et al., 2020; Guazzini et al., 2022; Latikka et al., 2022; Liu \& Tong, 2020; Riehm et al., 2020; Sewall et al., 2021; Stainback et al., 2020; Yue et al., 2022), it still largely unknown if COVID-19 related social media use during the pandemic has had a meaningful impact on well-being.
+By analyzing large-scale panel data from Austria and by focusing on intra-individual causal effects, this study aims to answer to this question.
 
@@
 
-The underlying theories that guided the selection of variables for my analysis are the two-continua model of mental health (Greenspoon \& Saklofske, 2001) and the hierarchical taxonomy of computer-mediated communication (Meier \& Reinecke, 2020).
-According to the two-continua model, mental health consists of (a) psychopathology and (b) well-being.
-Well-being can be differentiated into subjective and psychological well-being (Diener, Lucas, \& Oishi, 2018).
-Whereas subjective well-being emphasizes hedonic aspects such as happiness and joy, psychological well-being addresses eudaimonic aspects such as fulfillment and meaning.
-Subjective well-being is primarily about achieving positive affect and avoiding negative affect.
-One of the most prominent indicators of well-being is life satisfaction.
-In my view, because it represents a general appraisal of one's life, life satisfaction is best thought of as a meta concept that combines psychological and subjective well-being.
-Notably, life satisfaction is stable and fluctuates only little, whereas it's the exact opposite for affect (Dienlin \& Johannes, 2020).
-To capture well-being in this study I thus build on life satisfaction, positive affect, and negative affect.
-Together, this should provide an encompassing perspective on potential media effects.
-
-The hierarchical taxonomy of computer-mediated communication differentiates six levels of how people engage with digital technology.
-First, the device (e.g., smartphone); second, the type of application (e.g., social networking site); third, the branded application (e.g., Twitter); fourth, the feature (e.g., status post); fifth, the interaction (e.g., one-to-many); and sixth, the message (e.g., content) (Meier \& Reinecke, 2020).
-Whereas the first four levels focus on the communication \emph{channel}, the last two address the communication \emph{type}.
-To measure social media use for the consumption of COVID-19 related news and topics, I here employ both the channel and the type of communication perspective, which together should offer a nuanced understanding of communication.
-
-First, I investigate how well-being is affected by different types of communication affect, namely active and passive use.
-Defining what constitutes active and what passive use is not always clear, and different understandings are currently discussed (Ellison, Triẹû, Schoenebeck, Brewer, \& Israni, 2020; Meier \& Reinecke, 2020).
-Reading is generally considered as passive and writing as active, while there are also specific behaviors such as such liking or sharing content that fall somewhere in-between (Meier \& Krause, 2022).
-In this study, I hence distinguish (a) \emph{reading} (passive), (b) \emph{posting} (active), and (c) \emph{liking and sharing} COVID-19 related posts (both active and passive).
-Second, I analyze how using the most prominent branded applications affects well-being, and whether this effect changes across applications.
-Branded apps are separate entities with potentially divergent effects.
-Twitter might have a different effect as compared to WhatsApp because of their respective affordances.
-For example, Waterloo, Baumgartner, Peter, and Valkenburg (2018) found that it's more adequate to express negative emotions on WhatsApp than on Twitter or on Instagram.
-The branded applications investigated here are Facebook, Twitter, Instagram, WhatsApp, and YouTube.
-Worth noting, this study is not about \emph{general} social media use during times of COVID, but on social media use \emph{focused} on COVID-19 related content.
-Examples of such media use include posting thoughts about the pandemic or retweeting COVID-19 related news.
-
-\hypertarget{theorizing-social-media-effects-on-well-being}{%
-\subsection{Theorizing Social Media Effects on Well-Being}\label{theorizing-social-media-effects-on-well-being}}
-
-From a theoretical perspective, how could we explain whether COVID-19 related social media use affects well-being?
-According to the set-point model of subjective well-being, well-being is surprisingly stable (Lykken, 1999).
-Although specific events such as marriage or salary can have significant effects, after some time well-being routinely returns to prior levels, which are mostly determined genetically (Sheldon \& Lucas, 2014).
-Only very specific events and factors such as unemployment, disability, or death can cause long-term changes in well-being (Lucas, 2007).
-
+This study investigates how different \emph{facets} of well-being are affected by different \emph{types} and different \emph{channels} of communication (Meier \& Reinecke, 2020).
+Building on the typology of subjective well-being (Diener et al., 2018), three different well-being facets are analyzed: life satisfaction, positive affect, and negative affect.
+Because effects of social media depend on how they are used (Verduyn et al., 2015), I further distinguish three types of use and five popular channels.
+The types of use include reading, posting, and liking and sharing COVID-19 related posts.
+In doing so, this study analyzes social media use focused on COVID-19 related content, which includes posting thoughts about the pandemic, reading posts and comments, or retweeting COVID-19 related news.
+The five channels to be investigated are Facebook, Twitter, Instagram, WhatsApp, and YouTube, which at the time ranked among the most popular social media services in Austria.
+
+How could the various types and channels of COVID-19 related social media use affect well-being?
+According to the set-point theory, well-being is surprisingly stable (Lykken, 1999).
+Although specific events such as marriage or salary can have significant impacts on well-being, in most cases effects are only short-term, with well-being after some time routinely returning to prior levels (Sheldon \& Lucas, 2014).
+Only very specific factors such as unemployment, disability, or death can cause long-term changes in well-being (Lucas, 2007).
 Can media use be such a factor?
-In advance, there doesn't seem to be a clear winner, and it's likely that both positive and negative effects cancel each other out.
-Empirically, social media use on average oes not have a strong effect on well-being (Meier \& Reinecke, 2020).
-According to the Differential Susceptibility to Media Effects Model (Valkenburg \& Peter, 2013), the effects of media use differ across individuals.
-Whereas for some media are beneficial, for others they are harmful.
-On average, however, effects are often small or negligible.
-But also when focusing on individual users, social media can have both positive and negative effects on well-being (Büchi, 2021).
-They can impair well-being when causing embarrassment, stress, or disinformation, and they can improve well-being when providing connectedness, information, or entertainment (Büchi, 2021).
-
-Two of the most prominent media effect theories argue against strong negative impacts.
-First, uses and gratifications theory states that people explicitly and rationally chose specific media because of their respective benefits (Katz, Blumler, \& Gurevitch, 1973).
-If those benefits don't exist, they will spend their time elsewhere.
-And social media offer copiuos benefits.
-Most prominently, they help find relevant information, maintain and foster relationships, express one's personality, and entertain oneself (Pelletier, Krallman, Adams, \& Hancock, 2020).
-The second major theory that needs to be listed here is mood management theory (Zillmann, 1988).
-Users implicitly learn what media help them balance their mood and affect.
-For example, when bored people use social media to entertain themselves.
-Using experience sampling of well-being and logs of social media use, a study with 82 participants from Italy found that after episodes of social media use, levels of positive affect increased significantly (Marciano, Driver, Schulz, \& Camerini, 2022).
-
-But people can also misjudge media affects and are often overly optimistic (Metzger \& Suh, 2017).
-And precisely because social media have so many positive consequences, one can ask if this is not where the actual problem lies.
-In other words, social media aren't problematic because they're inherently bad, but rather because they're too good.
-And as with many other things, there can be too much of a good thing.
-It is therefore often asked whether social media are addictive, and users sometimes express this fear themselves (Yang, Griffiths, Yan, \& Xu, 2021).
-However, a recently published meta-analysis found that the two most prominent measures of addiction, the Bergen Facebook Addiction Scale and the Bergen Social Media Addiction Scale, have only small relations to well-being (Duradoni, Innocenti, \& Guazzini, 2020).
-In addition, the general idea of labeling excessive social and new media use as addiction was criticized, arguing that social media represent new regular behaviors that should not be pathologized (Galer, 2018; van Rooij et al., 2018).
-
-Because media effects can differ across situations (Valkenburg \& Peter, 2013), I now briefly focus on the effects of COVID-19 related social media use specifically.
-First, one could assume a \emph{direct} negative effect on well-being, and especially on positive or negative affect, which are more volatile and fluctuating.
-Dangers, inequalities, corruption---these were the headlines during the pandemic across many countries worldwide.
-If one learns about such events, the initial reaction might be shock, fear, or dismay.
-Consuming such news can be depressing (Dörnemann, Boenisch, Schommer, Winkelhorst, \& Wingen, 2021), perhaps even changing some general perspectives on life.
-That said, because not all news was negative, and because many people showed solidarity and compassion, there was also positive and uplifting content, potentially compensating for the negative effects (Dörnemann et al., 2021).
-A study with 2.057 respondents from Italy reported that during the pandemic virtual community and social connectedness even increased during the pandemic (Guazzini, Pesce, Marotta, \& Duradoni, 2022).
-
-There could also be \emph{indirect} effects.
-When browsing social media for Covid-19 related news, many users reported being captivated to such an extent that they could not stop using social media (Klein, 2021).
-During the pandemic social media use was at an all-time high in the US (Statista, 2021).
-Although it is most likely that moderate social media use is not detrimental (Orben, 2020), overuse, however, might be more critical, and several studies have shown more pronounced negative effects for extreme users (Przybylski \& Weinstein, 2017).
-To explain, overuse could impair well-being if it replaces meaningful or functional activities such as meeting others, working, actively relaxing, or exercising.
-So if a society collectively overuses social media during a pandemic, there might be potential for negative effects.
-On the other hand, one can make the case that overuse might even be beneficial, especially in times of a pandemic.
-Exchanging COVID-19 related messages with friends via WhatsApp might replace the in-person contact one would have otherwise, but which is literally impossible at the time.
-In situations where meaningful and functional activities are prohibited, using social media to exchange about COVID-19 related topics might not be the worst idea.
-Besides, given that nowadays a large number of experts, scientists, and politicians converse directly on social media, one can get first-hand high quality information on current developments.
-
-To summarize, from a theoretical perspective it is most likely that the average effects of social media use on well-being are negligible.
-Building on established theories from Communication, we would not assume that effects are either profoundly negative or strongly positive.
-
-\hypertarget{empirical-studies-on-social-media-effects}{%
-\subsection{Empirical Studies on Social Media Effects}\label{empirical-studies-on-social-media-effects}}
-
-So far, there is only little empirical research on how well-being is affected by COVID-19 related social media.
-In their study on the relations between media use and mental health during the pandemic, Bendau et al.
-(2021) found that people who used social media as a primary source of information reported on average ``significantly more unspecific anxiety and depression {[}{]} and significantly more specific COVID-19 related anxiety symptoms'' (p.~288).
-Eden, Johnson, Reinecke, and Grady (2020) analyzed the media use of 425 US college students during the first wave of the pandemic, finding both positive and negative relations with well-being.
-In a sample of 312 respondents collected via Amazon Mechanical Turk, Choi and Choung (2021) reported that people who used media to attain information were more lonely and less satisfied with their lives.
-Stainback, Hearne, and Trieu (2020) analyzed a large-scale study with 11,537 respondents from the US and found that increased COVID-19 media consumption was related to more psychological distress.
+Current literature overviews suggest no:
+Although social media use on average decreases well-being, the effects are small (Meier \& Reinecke, 2020).
+In addition, the effects of media use differ across individuals and types of content (Valkenburg \& Peter, 2013).
+Whereas for some users social media are more beneficial, for others they are more harmful.
+For example, in one study it was estimated that roughly one quarter of all users experienced negative effects, another quarter positive effects, while for the rest the effects were neutral (Beyens et al., 2021).
+This finding is aligned with the Differential Susceptibility to Media Effects Model:
+Although there is substantial \emph{variation} of media effects for individual users, there are both negative and positive aspects, leading to small effects on average (Valkenburg \& Peter, 2013)---as confirmed by several literature overviews (Dienlin \& Johannes, 2020; Huang, 2017; Meier \& Reinecke, 2020; Orben, 2020).
+
+Why are the effects of social media use on well-being small on average?
+Using Social media is ambivalent.
+It can impair well-being when causing embarrassment, stress, or disinformation, and it can improve well-being when providing connectedness, information, or entertainment (Büchi, 2021).
+Although social media are often associated with negative outcomes, there are a large number of positive effects, which either offset or at least significantly reduce negative outcomes.
+Positive outcomes encompass finding relevant information; maintaining and fostering relationships; expressing one's personality; or entertaining oneself (e.g., Pelletier et al., 2020).
+Negative effects include distraction; displacement of other meaningful activities; consumption of shallow, misleading or challenging content; or negative social comparisons (e.g., Meier \& Krause, 2022).
+People are much more likely to use media that offer plentiful benefits (Katz et al., 1973).
+Users implicitly learn what media help them regulate their mood and thereby well-being (Zillmann, 1988).
+Given the ubiquitous use of social media, the many benefits they offer, plus the general and implicit competence to use media that foster mood management, it is hence unlikely that the effects of social media use on well-being are profusely negative.
+
+There is still little empirical research on how well-being is affected by social media use that is focused on COVID-19 specifically.
+Results are mixed.
+When browsing social media for COVID-19 related news, many users reported being captivated to such an extent they could not stop using social media (Klein, 2021).
+In a study with 1,131 residents from Wuhan in China, people who spent more time in quarantine also spent more time on social media (Yue et al., 2022).
+Those who spent more time on social media also engaged in more upward social comparison, which was related to increased levels of stress.
+People who used social media as a primary source of information reported on average ``significantly more unspecific anxiety and depression {[}\ldots{]} and significantly more specific COVID-19 related anxiety symptoms'' (Bendau et al., 2021, p.
+288).
+Eden et al. (2020) analyzed how 425 US college students used media during the first wave of the pandemic, finding both positive and negative relations with well-being.
+In a large-scale study with 11,537 respondents from the US, increased COVID-19-related media consumption was related to more psychological distress (Stainback et al., 2020).
 A four-wave panel study with 384 young adults from the U.S.
-analyzed the effects of general digital technology use---objectively measured via screenshots of screen-time applications---on mental health, separating within- and between-person relations (Sewall, Goldstein, \& Rosen, 2021).
-The results showed that digital technology did not have any significant effects on mental health (for a similar study with comparable results, see Bradley \& Howard, 2021).
-Together, the literature is mixed, with a slight focus on the negative effects of social media as news use (see also Dörnemann et al., 2021; Liu \& Tong, 2020; Riehm et al., 2020).
-However, note that all of these findings represent between-person relations stemming from cross-sectional data (see below).
-We therefore don't know whether the differences in mental health and well-being are due to social media use or due to other third variables, such as age, health, employment, or education.
-
-The question of whether and how social media use affects well-being \emph{in general}, on the other hand, is well-researched.
-This also holds true for the different types of communication such as active or passive use.
-A meta review (i.e., an analysis of meta-analyses) found that the relation between social media use and well-being is likely in the negative spectrum but very small (Meier \& Reinecke, 2020)---potentially too small to matter.
-What determines whether or not an effect is considered small or trivial?
-As a starting point, we could refer to standardized effect sizes.
-According to Cohen (1992), small effect sizes start at \emph{r} = .10.
-And indeed, several if not most of the current meta-analyses find effect sizes below that threshold (Ferguson et al., 2021; Huang, 2017; Meier \& Reinecke, 2020).
-
-Finally, also several individual studies employing advanced methods found smalls relations between social media use and well-being (Keresteš \& Štulhofer, 2020; Orben, Dienlin, \& Przybylski, 2019; Przybylski, Nguyen, Law, \& Weinstein, 2021; Schemer, Masur, Geiß, Müller, \& Schäfer, 2021).
-For example, Beyens, Pouwels, van Driel, Keijsers, and Valkenburg (2021) reported that although for some users (roughly one quarter) the effects of social media use on well-being were negative, for almost the same number of users they were positive, while for the rest the effects were neutral.
-This finding is aligned the Differential Susceptibility to Media Effects Model:
-Although there is substantial \emph{variation} of media effects for individual users, the \emph{average} effects reported in the literature are often small (Valkenburg \& Peter, 2013).
-
-In conclusion, in light of the theoretical considerations and empirical studies presented above, I expect that COVID-19 related communication on social media doesn't affect well-being in a meaningful or relevant way.
+analyzed the effects of general digital technology use during the pandemic on mental health, finding that digital technology did not have significant effects on mental health (Sewall et al., 2021; for a similar study with comparable results, see Bradley \& Howard, 2021).
+A study with 2.057 respondents from Italy reported that during the pandemic virtual community and social connectedness increased (Guazzini et al., 2022).
+In a study with 735 participants from Finland, levels of loneliness were stable during the pandemic, and people who engaged more on social media experienced less loneliness (Latikka et al., 2022).
+Together, the literature is mixed, showing both positive and negative affects of social media use focused on or during COVID-19 on well-being (see also Dörnemann et al., 2021; Liu \& Tong, 2020; Riehm et al., 2020).
+
+In conclusion, these mixed empirical results, together with the observation that social media effects on well-being are very small in general, I expect that COVID-19 related communication on social media does not affect well-being in a meaningful or relevant way.
 
@@
 
-\hypertarget{current-study}{%
-\section{Current Study}\label{current-study}}
+\hypertarget{method}{%
+\section{Method}\label{method}}
+
+\hypertarget{preregistration}{%
+\subsection{Preregistration}\label{preregistration}}
+
+The hypotheses, the sample, the measures, the analyses, and the inference criteria (SESOI, \emph{p}-value) were preregistered on the Open Science Framework, accessible here: \url{https://osf.io/87b24/?view_only=b2289b6fec214fa88ee75a18d45c18f3}.
+Because in this study I analyze data from an already existing large-scale data set, the preregistration was done prior to accessing the data.
+The preregistration was designed on the basis of the panel documentation online (Kittel et al., 2020).
+In some cases, it was impossible to execute the analyses as I had originally planned, for example because some properties of the variables only became apparent when seeing the actual data.
+The most relevant deviations are reported below, and a complete list of all changes can be found in the online \href{https://XMtRA.github.io/Austrian_Corona_Panel_Project}{companion website} (\url{https://XMtRA.github.io/Austrian_Corona_Panel_Project}).
+
+\hypertarget{sample}{%
+\subsection{Sample}\label{sample}}
+
+The data come from the Austrian Corona Panel Project (Kittel et al., 2021), which is a large-scale standalone panel study.
+The data are hosted on AUSSDA, are publicly available here (\url{https://doi.org/10.11587/28KQNS}), and consist of 32 waves.
+The study was conducted between March 2020 and June 2022, and data collection is now finished.
+Between March 2020 and July 2020, the intervals between waves were weekly and afterward monthly.
+Each wave consists of at least 1,500 respondents.
+The sample size was \emph{N} = 3,485, with overall 111,520 observations.
+Panel mortality was compensated through a continuous re-acquisition of new participants.
+Participants were sampled from a pre-existing online access panel provided by the company Marketagent, Austria.
+Panel members were incentivized with 180 credit points for each wave of the study.
+
+Achieved via quota sampling, the sample matched the Austrian population in terms of age, gender, region/state, municipality size, and educational level.
+In order to participate in the study, the respondents needed to be Austrian residents and had to be at least 14 years of age.
+All respondents needed to have access to the internet (via computer or mobile devices such as smartphones or tablets).
+Ethical review and approval was not required for the study in accordance with the local legislation and institutional requirements.
+The participants provided their written informed consent to participate in this study.
+The average age was 41 years, 49 percent were male, 14 percent had a University degree, and 5 percent were currently unemployed.
 
@@
 
-Testing this hypothesis, however, is not trivial.
-First, in contrast to most hypotheses typically posited in the social sciences it implicitly contains an effect size, a so-called smallest effect size of interest (SESOI).
-Effectively testing this hypothesis necessitates defining what's considered a ``trivial effect size'' and what's not.
-Above I already referred to standardized effect sizes.
-However, standardized effect sizes should only be a first step toward evaluating an effect's relevance (Baguley, 2009).
-Standardized effect sizes are determined by a sample's variance,\footnote{Consider the effect size Cohen's \emph{d}: The mean's of the two groups that are to be compared are subtracted from one another and then divided by the sample's standard deviation (Cohen, 1992).
-Hence, if there is more deviation/variance in a sample, the effect size decreases, even if the difference of the group's means stays the same.} which is problematic:
-The question of whether or not social media use affects a particular person in a relevant way should not depend on the variance in the sample in which that person's data were collected.
-Instead, it should depend on absolute criteria.
-
-What could be a minimally interesting, nontrivial effect?
-Because this is a normative and ultimately philosophical question, there can never be a clear, single, or unanimous answer.
-However, it is still necessary and helpful to try provide such a plausible benchmark.
-I therefore suggest the following SESOI for this research question:
+Testing the hypothesis necessitates defining what is considered a ``trivial effect size''.
+Being a normative question, finding a clear, single, or unanimous answer is impossible.
+However, it is still necessary and helpful to work toward a plausible benchmark.
+I suggest the following SESOI for this research question:
 
@@
 Thus, a change of four units in social media use (e.g., a complete stop) should correspond to a noticeable change in well-being.
-But what's a noticeable change in well-being?
-According to Norman, Sloan, and Wyrwich (2003), people can reliably distinguish \emph{seven} levels of satisfaction with health.
-So if satisfaction is measured on a 7-point scale, we would state that a four unit change in social media use should result in a one unit change in life satisfaction.
-(For more information, see Methods section ``Inference Criteria.'')
-
-\hypertarget{causality}{%
-\subsection{Causality}\label{causality}}
-
-The hypothesis explicitly states a causal effect.
-In non-experimental studies, longitudinal designs can help investigate causality.
-Using longitudinal designs alone, however, is not sufficient for establishing correct causal statements (Rohrer \& Murayama, 2021).
-In addition, we for example also need to control for confounding third variables, and importantly also for \emph{varying} third variables.
-
-To illustrate, consider the following example.
-Imagine that a person suddenly starts using social media much more than usual, and then after some time becomes less satisfied with their life.
-Eventually, use and life satisfaction return to prior levels.
-If this happens to several people at the same time, in a longitudinal study we could then observe a significant effect of social media use on life satisfaction.
-However, it could also be the case that during the study there was a major exogenous event (say, a pandemic), which caused large parts of the working population to loose their jobs.
-Hence, the causal effect reported above was confounded, because in reality it was the pandemic that caused both social media use to rise and life satisfaction to go down.
-
-Thus, only when controlling for \emph{all} relevant confounders, we can correctly estimate causality without bias (Rohrer, 2018).
-Obviously, we can never be entirely sure to have included all confounders, which makes absolute statements regarding causality virtually impossible.
-In addition, when determining the overall causal effect, we need to make sure \emph{not} to control for mediating variables (Rohrer, 2018), for doing so would bias our assessment of the causal effect.
-Complicating matters further, it is often unclear if a variable is a mediator or a confounder.\footnote{In addition, there also exist colliders, which I don't discuss here and which complicate the issue even further (Rohrer, 2018).}
-However, despite all these caveats, when controlling for relevant variables (that aren't mediators), we can be much more certain that we measured causality correctly.
-The aim should therefore be to collect as many varying and nonvarying confounders as possible (which I believe is seldom done in our field), while knowing that absolute certainty regarding causality cannot be reached.
-
-When searching for suitable candidates for confounders, we should look for variables that affect both media use and well-being.
-Controlling for these factors isolates the actual effect of social media use on well-being.
-We can also control for variables that affect only social media use or well-being.
-However, in doing so not much is gained or lost, because the effects of social media use would remain virtually the same (Kline, 2016; but see McElreath, 2021).
-
-In this study, I hence plan to control for the following variables, which either have already been shown to affect both social media use and well-being or which are likely to do so, and which also aren't mediators:
-gender, age, education, Austria country of birth, Austria country of birth of parents, text-based news consumption, video-based news consumption, residency Vienna, household size, health, living space, access to garden, access to balcony, employment, work hours per week, being in home-office, household income, outdoor activities, satisfaction with democracy, disposition to take risks, and locus of control.\footnote{The dataset includes many other variables that one could also potentially control for, and I invite interested readers to download the and explore potential interesting relationships.}
-
-Next to including covariates, it's now increasingly understood that causal effects should be analyzed from an internal, within-person perspective (Hamaker, 2014).
-If a specific person changes their media diet, we need to measure how this behavior affects \emph{their own} well-being.
-Between-person comparisons from cross-sectional data, where participants are interviewed only once, cannot provide such insights.
-In this study, I hence differentiate between-person relations from within-person effects.
-And as explicated above, to test the hypothesis I thus consider only the within-person effects.
-
-Finally, one precondition of causality is temporal order.
-The cause needs to precede the effect.
-Finding the right interval between cause and effect is crucial.
-For example, if we want to understand the effect of alcohol consumption on driving performance, it makes a big difference if driving performance is measured one minute, one hour, one day, or one week after consumption.
-If variables are stable, longer intervals are needed; if they fluctuate, shorter intervals.
-In the case of well-being, we need shorter intervals for affect and longer ones for life satisfaction (Dienlin \& Johannes, 2020).
-Still, choosing the right interval is challenging, because especially short intervals are hard to implement in practice and often require advanced methods such as experience sampling (also known as in situ measurement or ambulant assessment) (Schnauber-Stockmann \& Karnowski, 2020).
-
-In this study, I hence analyze how using social media during the last week affected their positive and negative affect during the same week.
-More precisely, if people during the last week used COVID-19 related social media more than the usually do, did they feel better or worse during that week than they usually do?
-Regarding life satisfaction, I implemented a longer interval.
-Namely, if people during the last week used COVID-19 related social media more than the usually do, were they at the end of the week more or less satisfied with their lives than they usually are?
-I hence analyze if when a person changes their social media diet, are there (a) \emph{simultaneous} changes in their affect and (b) \emph{subsequent} changes in their life satisfaction?
-When additionally controlling for both stable and varying confounders, we can then be more sure that the effect is indeed causal.
-Similar approaches were implemented by other studies (Johannes, Dienlin, Bakhshi, \& Przybylski, 2022; Scharkow, Mangold, Stier, \& Breuer, 2020), and they are considered a best practice approach toward analyzing causality (Bell, Fairbrother, \& Jones, 2019).
-
-\hypertarget{method}{%
-\section{Method}\label{method}}
-
-In this section I describe the preregistration and how I determined the sample size, data exclusions, the analyses, and all measures in the study.
-
-\hypertarget{preregistration}{%
-\subsection{Preregistration}\label{preregistration}}
-
-The hypotheses, the sample, the measures, the analyses, and the inference criteria (SESOI, p-value) were preregistered on the Open Science Framework.
-The (anonymous) preregistration can be accessed here: \url{https://osf.io/87b24/?view_only=b2289b6fec214fa88ee75a18d45c18f3}.
-Because in this study I analyzed data from an already existing large-scale data set, all of these steps were done prior to accessing the data.
-The preregistration was designed on the basis of the panel documentation online (Kittel et al., 2020).
-In some cases I couldn't execute the analyses as I had originally planned, for example because some properties of the variables only became apparent when inspecting the actual data.
-The most relevant deviations are reported below, and a complete list of all changes can be found in the online \href{https://XMtRA.github.io/Austrian_Corona_Panel_Project}{companion website} (\url{https://XMtRA.github.io/Austrian_Corona_Panel_Project}).
-
-\hypertarget{sample}{%
-\subsection{Sample}\label{sample}}
-
-The data come from the Austrian Corona Panel Project (Kittel et al., 2021).
-The study was conducted between March 2020 and October 2021.
-It contains 26 waves, and at the time of writing the first 24 waves were available for download.
-Between March 2020 and July 2020, the intervals were weekly, and afterward the intervals were monthly.
-Each wave consists of at least 1,500 respondents.
-The overall sample size was \emph{N} = 3,018, and 72,432 observations were collected.
-Panel mortality was compensated through a continuous acquisition of new participants.
-All respondents needed to have access to the internet (via computer or mobile devices such as smartphones or tablets).
-They were sampled from a pre-existing online access panel provided by the company Marketagent, Austria.
-Respondents were asked and incentivized with 180 credit points to participate in each wave of the panel.
-
-Achieved via quota sampling, the sample matched the Austrian population in terms of age, gender, region/state, municipality size, and educational level.
-In order to participate in the study, the respondents needed to be Austrian residents and had to be at least 14 years of age.
-Ethical review and approval was not required for the study in accordance with the local legislation and institutional requirements.
-The participants provided their written informed consent to participate in this study.
-The average age was 42 years, 49 percent were male, 14 percent had a University degree, and 5 percent were currently unemployed.
-
-\hypertarget{inference-criteria}{%
-\subsection{Inference Criteria}\label{inference-criteria}}
-
-Because the data were analyzed post-hoc, no a-priori sample size planning on the basis of power analyses was conducted.
-The sample is large, and it is hence well-equipped to reliably detect also small effects.
-In addition, because such large samples easily generate significant \emph{p}-values even for very small effects, it helps that the hypotheses were tested with a smallest effect size of interest-approach.
-To this end, I adopted the interval testing approach as proposed by Dienes (2014).
-On the basis of the SESOI, I then defined a null region.
-In what follows, I explain how I determined the SESOI and the null region.
+What is a noticeable change in well-being?
+According to Norman et al. (2003), people can reliably distinguish seven levels of satisfaction with health.
+So if satisfaction is measured on a 7-point scale, a four unit change in social media use should result in a one unit change in life satisfaction.
 
 In this study, life satisfaction was measured on an 11-point scale.
-If people can reliably differentiate 7 levels as mentioned above, this corresponds to 11 / 7 = 1.57 unit change on an 11-point scale.
+If people can reliably differentiate 7 levels, this corresponds to 11 / 7 = 1.57 unit change on an 11-point scale.
 Hence, a four-point change in media use (e.g., a complete stop) should result in a 1.57-point change in life satisfaction.
 In a statistical regression analysis, \emph{b} estimates the change in the dependent variable if the independent variable increases by one point.
-We would therefore expect a SESOI of \emph{b} = 1.57 / 4 = 0.39.
-For affect, which was measured on a 5-point scale, our SESOI would be \emph{b} = 0.71 / 4 = 0.18.
-Because we're agnostic as to whether the effects are positive or negative, the null region includes negative and positive effects.
+For life satisfaction, we would therefore define a SESOI of \emph{b} = 1.57 / 4 = 0.39.
+For positive or negative affect, which was measured on a 5-point scale, our SESOI would be \emph{b} = 0.71 / 4 = 0.18.
+Because we are agnostic as to whether the effects are positive or negative, the null region includes both negative and positive effects.
 Finally, in order not to exaggerate precision and to be less conservative, these numbers are reduced to nearby thresholds.\footnote{Note that other researchers also decreased or recommended decreasing thresholds for effect sizes when analyzing within-person or cumulative effects (Beyens et al., 2021; Funder \& Ozer, 2019).}
@@
 
-Let's briefly illustrate what this means in practice.
-If the 95\% confidence interval falls completely within the null-region (e.g., \emph{b} = .20, {[}95\% CI: .15, .25{]}), the hypothesis that the effect is trivial is supported.
-If the confidence interval and the null region overlap (e.g., \emph{b} = .30, {[}95\% CI: .25, .35{]}), the hypothesis is not supported and the results are considered inconclusive, while a meaningful negative effect is rejected.
-If the confidence interval falls completely outside of the null-region (e.g., \emph{b} = .40, {[}95\% CI: .35, .45{]}), the hypothesis is rejected and the existence of a meaningful positive effect is supported.
-For an illustration, see Figure \ref{fig:sesoi}).
+The hypothesis is analyzed using the interval testing approach as proposed by Dienes (2014).
+To illustrate, if the 95\% confidence interval falls completely within the null-region (e.g., \emph{b} = -.02, {[}95\% CI: -.12, .08{]}), the hypothesis that the effect is trivial is supported.
+If the confidence interval and the null region overlap (e.g., \emph{b} = -.22, {[}95\% CI: -.27, -.17{]}), the hypothesis is not supported and the results are considered inconclusive, while a meaningful negative effect is rejected.
+If the confidence interval falls completely outside of the null-region (e.g., \emph{b} = -.40, {[}95\% CI: -.45, -.35{]}), the hypothesis is rejected and the existence of a meaningful positive effect is supported.
+For an illustration, see Figure \ref{fig:sesoi}.
+
+\begin{figure}
+\centering
+\includegraphics{manuscript_files/figure-latex/sesoi-1.pdf}
+\caption{\label{fig:sesoi}Using confidence intervals to test a null region. In this study, a trivial effect of social media use on life satisfaction is defined as ranging from b = -.30 to b = .30}
+\end{figure}
 
@@
 
-The hypothesis was analyzed using mixed effects models, namely random effect within-between models (REWB)(Bell et al., 2019).
-Three models were run, one for each dependent variable.
+\hypertarget{causality}{%
+\subsubsection{Causality}\label{causality}}
+
+When using longitudinal designs to analyze causality, it is important to (a) focus on within-person effects (Hamaker, 2014); to (b) control for confounders (Rohrer \& Murayama, 2021); and to (c) test a plausible interval between measures (Dormann \& Griffin, 2015).
+To elaborate, in non-experimental designs it makes much sense to analyze causal effects from an internal, within-person perspective (Hamaker, 2014).
+If a specific person changes their media diet, we need to measure how this behavior affects their well-being.
+Between-person comparisons from longitudinal data cannot provide such insights (Lucas, 2022).
+To test the hypothesis, I thus consider only the within-person effects.
+
+Second, to identify confounders we should control for variables that affect both media use and well-being, which helps isolate the actual effect of social media use on well-being (Rohrer, 2018).
+Because we are adopting a within-person perspective, we need to implement \emph{time-varying} confounders (Rohrer \& Murayama, 2021).
+And because we are determining the \emph{overall} causal effect, we need to make sure \emph{not} to control for mediating variables (Rohrer, 2018), for doing so would bias our assessment of the causal effect.
+In this study, I hence preregistered to control for the following variables, which either have already been shown or are very likely to affect both social media use and well-being, and which also are not mediators:
+gender, age, education, Austria country of birth, Austria country of birth of parents, residency Vienna, text-based news consumption, video-based news consumption, household size, health, living space, access to garden, access to balcony, employment, work hours per week, being in home-office, household income, outdoor activities, satisfaction with democracy, disposition to take risks, and locus of control.
+
+Finally, one precondition of causality is temporal order and finding a plausible interval (Dormann \& Griffin, 2015).
+If variables are stable, longer intervals are needed; if they fluctuate, shorter intervals are required.
+In the case of well-being, we need shorter intervals for the more fluctuating positive and negative affect, and longer ones for the more stable life satisfaction (Dienlin \& Johannes, 2020).
+Using social media can have instant effects on mood (Marciano et al., 2022).
+Effects on life satisfaction often take longer to manifest, for example because media use leads to actual changes in specific behaviors, which then in turn affect life satisfaction (Dienlin et al., 2017).
+In this study, I hence analyze how using social media during the last week affected positive and negative affect during the same week.
+In other words, if people during the last week engaged in more COVID-19 related social media use than usual, did they feel better or worse during that week than usual?
+For life satisfaction, I implemented a longer interval.
+If people during the last week used COVID-19 related social media more than they usually do, were they at the end of the week more or less satisfied with their lives than they usually are?
+I hence analyze if when a person changes their social media diet, are there (a) \emph{simultaneous} changes in their affect and (b) \emph{subsequent} changes in their life satisfaction?
+These relations will be controlled for varying confounders, which fosters a causal interpretation.
+
+\hypertarget{statistical-model}{%
+\subsubsection{Statistical model}\label{statistical-model}}
+
+The hypothesis was analyzed using random effect within-between models (REWB, Bell et al., 2019).
+Altogether three models were run, one for each dependent variable.
 The data were hierarchical, and responses were separately nested in participants and waves (i.e., participants and waves were implemented as random effects).
@@
 They included social media communication types and channels, separated into within and between-person factors, as well as stable and varying covariates.
-All predictors were included simultaneously and in each of the three models.
+All predictors were included simultaneously in each of the three models.
 
@@
 Because Mardia's test showed that the assumption of multivariate normality was violated, I used the more robust Satorra-Bentler scaled and mean-adjusted test statistic (MLM) as estimator.
-To avoid overfitting, I tested the scales on more liberal fit criteria (CFI \textgreater{} .90, TLI \textgreater{} .90, RMSEA \textless. .10, SRMR \textless{} .10) (Kline, 2016).
-Finally, REWB-models cannot model latent variables.
-To increase precision, I therefore exported factor scores from the CFAs for positive and negative affect.
-Respondents who answered less than 50\% of all questions were removed.
-The remaining missing responses were imputed using predictive mean matching.
-
-For more information on the analyses, a complete documentation of the models and results, additional analyses (for example using multiple imputation or no imputation), see \href{https://XMtRA.github.io/Austrian_Corona_Panel_Project}{companion website}.
+Mean scores were used for positive and negative affect.
+Missing responses were imputed using multiple imputation with predictive mean matching (five iterations, five data-sets), including categorical variables.
+All variables were imputed except the social media use measures, as they were not collected on each wave.
+All variables included in the analyses presented here were used to impute missing data.
+
+\hypertarget{robustness-checks}{%
+\subsubsection{Robustness-checks}\label{robustness-checks}}
+
+To find out whether the inferences were robust across plausible but subordinate analyses, I conducted additional exploratory analyses.
+I reran the analyses (a) with additional not-preregistered covariates such as trust in media or government, (b) without covariates, (c) with single imputation, and (d) without imputation.
+For more information on the analyses, a complete documentation of the models and results, and all additional analyses, see \href{https://XMtRA.github.io/Austrian_Corona_Panel_Project}{companion website}.
 
@@
 
-In what follows, I list all the variables that I analyzed.
 For the variables' means, range, and variance, see Table \ref{tab:tab-descriptives}.
@@
 
-Life satisfaction was measured with the item ``All things considered, how satisfied are you with your life as a whole nowadays?'' from the European Social Survey (European Social Survey, 2021).
+Life satisfaction was measured with the item ``All things considered, how satisfied are you with your life as a whole nowadays?'', which comes from the European Social Survey (European Social Survey, 2021).
 The response options ranged from 0 (\emph{extremely dissatisfied}) to 10 (\emph{extremely satisfied}).
@@
 The response options were 1 (\emph{never}), 2 (\emph{on some days}), 3 (\emph{several times per week}), 4 (\emph{almost every day}), and 5 (\emph{daily}).
-The scale showed good factorial fit, \(\chi^2\)(46) = 65.30, \emph{p} = .032, CFI = 1.00, RMSEA = .02, 90\% CI {[}.01, .03{]}, SRMR = .01.
+The scale showed good factorial fit, \(\chi^2\)(62) = 67.03, \emph{p} = .309, CFI = 1.00, RMSEA \textless{} .01, 90\% CI {[}\textless{} .01, .02{]}, SRMR = .01.
 Reliability was high, \(\omega\) = .85.
@@
 The response options were 1 (\emph{never}), 2 (\emph{on some days}), 3 (\emph{several times per week}), 4 (\emph{almost every day}), and 5 (\emph{daily}).
-The scale showed good factorial fit, \(\chi^2\)(331) = 3138.37, \emph{p} \textless{} .001, CFI = .97, RMSEA = .08, 90\% CI {[}.07, .08{]}, SRMR = .03.
-Reliability was high, \(\omega\) = .89.
+The scale showed good factorial fit, \(\chi^2\)(443) = 3810.53, \emph{p} \textless{} .001, CFI = .98, RMSEA = .07, 90\% CI {[}.07, .08{]}, SRMR = .03.
+Reliability was high, \(\omega\) = .91.
 
@@
 The items come from Wagner et al. (2018) and were adapted for the context of this study.
-The general introductory question was ``How often during the last week have you engaged in the following activities on social media?''
+The general introductory question was ``How often during the last week have you engaged in the following activities on social media?''.
 The three items were ``Reading the posts of others with content on the Coronavirus'', ``When seeing posts on the Coronavirus, I clicked `like', `share' or `retweet'\,'', ``I myself wrote posts on the Coronavirus on social media.''
@@
 
-Social media use was measured for all participants on waves 1, 2, 8, 17, and 23.
-Freshly recruited respondents always answered all questions on social media use.
+Social media use was measured for all participants on waves 1, 2, 8, 17, 23, and 28.
+Freshly recruited respondents always answered all questions on COVID 19-related social media use.
+For the analyses I hence used all 32 waves.
+
+\begin{table}[tbp]
+
+\begin{center}
+\begin{threeparttable}
+
+\caption{\label{tab:tab-descriptives}Descriptives of the main variables.}
+
+\begin{tabular}{lllll}
+\toprule
+ & \multicolumn{1}{c}{sd} & \multicolumn{1}{c}{min} & \multicolumn{1}{c}{max} & \multicolumn{1}{c}{mean}\\
+\midrule
+Well-being &  &  &  & \\
+\ \ \ Life satisfaction & 2.22 & 6.37 & 6.59 & 6.51\\
+\ \ \ Positive affect & 0.94 & 3.11 & 3.22 & 3.16\\
+\ \ \ Negative affect & 0.76 & 1.74 & 1.87 & 1.79\\
+Social media use &  &  &  & \\
+\ \ \ Read & 1.39 & 2.02 & 2.87 & 2.37\\
+\ \ \ Like \& share & 1.19 & 1.56 & 1.95 & 1.74\\
+\ \ \ Posting & 0.89 & 1.37 & 1.42 & 1.40\\
+Social media channel &  &  &  & \\
+\ \ \ Facebook & 1.60 & 2.21 & 2.74 & 2.43\\
+\ \ \ Twitter & 0.95 & 1.32 & 1.45 & 1.38\\
+\ \ \ Instagram & 1.33 & 2.04 & 2.05 & 2.05\\
+\ \ \ WhatsApp & 1.67 & 2.32 & 2.65 & 2.46\\
+\ \ \ YouTube & 1.28 & 1.90 & 2.02 & 1.96\\
+\bottomrule
+\end{tabular}
+
+\end{threeparttable}
+\end{center}
+
+\end{table}
 
@@
 The effects of COVID-19 related social media use were controlled for the following stable variables:
-(a) gender (female, male, diverse), (b) age, (c) education (ten options), (d) Austria country of birth (yes/no), (e) Austria parents' country of birth (no parent, one parent, both parents).
-I originally planned to implement additional variables as varying covariates.
-However, because they were not measured often enough or not at the time when social media use was measured, I implemented them as stable variables using their average values across all waves.
-This includes (a) text-based media news consumption (five degrees), (b) video-based media news consumption (five degrees), (c) residency is Vienna (yes/no), (d) self-reported physical health (five degrees), (e) living space (eleven options), (f) access to balcony (yes/no), (g) access to garden (yes/no), (h) employment (nine options), (i) disposition to take risks (eleven degrees), and (j) locus of control (five degrees).
-I also controlled for the following varying covariates: (a) five items measuring outdoor activities such as sport or meeting friends (five degrees), and (b) satisfaction with democracy (five degrees).
-Because it lead to too much attrition in the sample, I did not control for (a) household size, (b) work hours per week, (c) home office, (d) household income.
+gender (female, male, diverse), age, education (ten options), Austria country of birth (yes/no), Austria parents' country of birth (no parent, one parent, both parents), and household size.
+I also controlled for the following varying covariates: five items on current living conditions, including self-reported physical health, whether participants contracted COVID-19 since the last wave, current household income, working in home-office, and overall work hours; nine items measuring use of specific national text-based and video-based news outlets; five items measuring outdoor activities such as exercise or meeting friends; and three more psychological measures including locus of control, disposition to take risks, and satisfaction with democracy.
 
@@
 
-First, when looking at the variables from a descriptive perspective (Figure \ref{fig:descriptives}), we see that all well-being measures did not change substantially across the different waves of data collection.
-COVID-19 related media use, however, decreased slightly at the beginning of the study and remained stable after approximately six waves.
-The initial decrease might be explained by the fact that the collection of data began at the end of March 2020, hence approximately three months after the pandemic began.
-It could be that after an initial uptick, COVID-19 related social media use was already declining at the time, returning to more normal levels.
+\hypertarget{descriptive-analyses}{%
+\subsection{Descriptive Analyses}\label{descriptive-analyses}}
+
+Looking at the variables from a descriptive perspective, aligned with set-point theory we can see that the level of all well-being measures were surprisingly stable during data collection (see Figure \ref{fig:fig-descriptives}).
+COVID-19 related social media use, however, showed changes.
+Reading, sharing and liking COVID-19 related content decreased substantially (almost one scale point from 3 to 2).
+Posting about COVID-19 related content stayed the same.
+Using Facebook, WhatsApp, and YouTube for COVID-19 related content decreased.
+Instagram stayed the same.
+Twitter use slightly increased.
+The general initial decrease could be explained by the fact that the collection of data began at the end of March 2020, hence approximately three months after the pandemic's onset.
+After an initial uptick, COVID-19 related social media use might have already been declining at the time.
+
+\begin{figure}
+\includegraphics[width=\textwidth]{figures/fig_descriptives} \caption{Well-being and media use across the 32 waves.
+Note.
+Values obtained from mixed effect models, with participants and waves as grouping factors and without additional predictors.}\label{fig:fig-descriptives}
+\end{figure}
+
+Using the average values across all waves, which provides a stable picture of the general relations, I next looked at the correlations between social media use and well-being (see Figure \ref{fig:fig-correlations}).
+Several interesting patterns emerged.
+In general, people who spend more time engaging with COVID-19 related content on social media reported reduced well-being.
+Users who spend more time reading, liking and sharing, and posting COVID-19 related content were less satisfied with their lives.
+They also showed slightly less positive affect.
+This overall negative picture was even more pronounced for negative affect.
+People who engaged more with COVID-19 related content, including all types and channels of communication, reported substantially higher levels of negative affect.
+For example, people who were more likely to post COVID-19 content had much higher levels of negative affect (\emph{r} = .56).
+Note that these results represent between-person correlations, not causal within-person effects.
+
+\begin{figure}
+\includegraphics[width=\textwidth]{figures/fig_cor} \caption{Descriptives of the main variables, capturing well-being and social media use with their average values across all waves.
+Upper triangle: correlation coefficients; diagonal: density plots; lower triangle: scatter plots.}\label{fig:fig-correlations}
+\end{figure}
 
@@
 
-The study's main hypothesis was that the effects of social media use on well-being would be trivial.
-Regarding the effects of different communication \emph{types}---that is, reading vs.~sharing vs.~posting---all within-person effects fell completely within the a-priori defined null region (see Figure \ref{fig:res-activity}).
-For example, respondents who used social media more frequently than usual to read about COVID-19 related topics did not show a simultaneous change in life satisfaction (\emph{b} = 0.05 {[}95\% CI -0.01, 0.1{]}).
-All confidence intervals included zero; hence, all effects were also statistically non-significant.
-As a result, the hypothesis was supported for all COVID-19 related types of social media communication.
-
-Regarding between-person relations, about which no hypotheses were formulated, only three effects did not include zero.
-Respondents who across all waves used social media more frequently than others to read about COVID-19 related posts reported slightly lower levels of positive affect than others (\emph{b} = -0.03 {[}95\% CI \textgreater{} -0.01, -0.06{]}).
-Respondents who across all waves used social media more frequently than others to write COVID-19 related posts reported higher levels of negative affect than others (\emph{b} = 0.06 {[}95\% CI 0.09, 0.03{]}).
-Interestingly, respondents who across all waves used social media more frequently than others to write COVID-19 related posts also reported slightly higher levels of positive affect than others (\emph{b} = 0.05 {[}95\% CI 0.09, \textless{} 0.01{]}).
-However, note that the effect were still completely inside of the null region, hence not large enough to be considered practically relevant.
-
-Note that when comparing the results with and without control variables, the results differed.
-For example, on the between-person level, one effect stopped being significant if controlled for additional variables.
-Actively posting on social media was significantly (though not meaningfully) related to decreased life satisfaction.
-However, when controlling for potential confounders, the effect became virtually zero (see Figure \ref{fig:res-channels}).
-
-Regarding the COVID-19 related use of social media \emph{channels}, the results were comparable (see Figure \ref{fig:res-channels}).
+\hypertarget{social-media-communication-types}{%
+\subsubsection{Social media communication types}\label{social-media-communication-types}}
+
+The study's main hypothesis was that the causal effects of all types and channels of social media use on all facets of well-being would be trivial.
+Regarding the effects of different communication types (i.e., reading, sharing, of posting about COVID-19 related content), all within-person effects fell completely within the a-priori defined null region (see Figure \ref{fig:fig-within}).
+For example, respondents who used social media more frequently than usual to like or share COVID-19 related content did not show a simultaneous change in life satisfaction (\emph{b} = -0.02 {[}95\% CI -0.08, 0.03{]}).
+As a result, the hypothesis of trivial effects was supported for all COVID-19 related types of social media communication.
+
+However, three effects stood out, as they were significantly different from zero.
+Users who read more COVID-19 related content than usual reported slightly reduced levels of positive affect (\emph{b} = -0.03 {[}95\% CI -0.05, -0.01{]}).
+Users who liked and shared more COVID-19 related content than usual also experienced slightly more negative affect than usual (\emph{b} = 0.05 {[}95\% CI 0.02, 0.08{]}).
+Users who wrote more COVID-19 related posts than usual also experienced slightly more negative affect than usual (\emph{b} = 0.05 {[}95\% CI 0.02, 0.08{]}).
+
+\hypertarget{social-media-communication-channels}{%
+\subsubsection{Social media communication channels}\label{social-media-communication-channels}}
+
+Regarding the COVID-19 related use of social media channels (i.e., Facebook, Instagram, WhatsApp, YouTube, and Twitter) the results were comparable (see Figure \ref{fig:fig-within}).
 Changes in the frequency of using different social media channels to attain information regarding COVID-19 were unrelated to meaningful changes in well-being.
-For example, respondents who used Facebook more frequently than usual to learn about COVID-19 did not show a simultaneous change in well-being (\emph{b} = -0.05 {[}95\% CI -0.11, 0.01{]}).
-Only two effects differed substantially from zero.
-Respondents who used Instagram more frequently than usual to attain COVID-19 related news reported slightly \emph{higher} levels of life satisfaction than usual (\emph{b} = 0.09 {[}95\% CI 0.02, 0.16{]}).
-Respondents who used Twitter more frequently than usual to attain COVID-19 related news reported slightly \emph{lower} levels of life satisfaction than usual (\emph{b} = -0.12 {[}95\% CI -0.23, -0.02{]}).
-However, both effects were still completely inside of the null region, hence not large enough to be considered meaningful.
-In sum, the hypothesis was supported also for the COVID-19 related use of important social media channels.
-
-In terms of between-person relations---which, again, weren't included in the hypotheses---no relations crossed the null region or fell outside of it.
-Only one relation did not include zero, was hence statistically significant.
-Respondents who across all waves used YouTube more frequently than others for COVID-19 related reasons reported marginally \emph{higher} levels of life satisfaction (\emph{b} = 0.08 {[}95\% CI \textless{} 0.01, 0.16{]}).
-However, please note that this effect again was not large enough to be considered practically relevant.
-
-Again, note that when comparing the results with and without control variables, the results differed.
-Especially on the between-person level, altogether five effects stopped being significant if they were controlled for additional variables.
-For example, using Instagram was significantly (though not meaningfully) related to increased life satisfaction.
-However, when controlling for additional covariates, the effect became virtually zero (see Figure \ref{fig:res-channels}).
+For example, respondents who used Facebook more frequently than usual to learn about COVID-19 did not show a simultaneous change in life satisfaction (\emph{b} \textless{} 0.01 {[}95\% CI -0.04, 0.04{]}).
+In sum, the hypothesis of trivial effects was supported also for the COVID-19 related use of important social media channels.
+
+That said, one effect differed substantially from zero.
+Respondents who used Twitter more frequently than usual to attain COVID-19 related news reported slightly higher levels of negative affect than usual (\emph{b} = 0.01 {[}95\% CI \textless{} 0.01, 0.03{]}).
+However, the effect was still completely inside of the null region, hence likely not large enough to be considered meaningful.
+
+For an overview of all within-person effects, see Table \ref{tab:tab-within} and Figure \ref{fig:fig-within}.
+
+\begin{figure}
+\includegraphics[width=\textwidth]{manuscript_files/figure-latex/fig-within-1} \caption{Within-person effects of COVID-19 related social media use on well-being.
+Note.
+The black estimates show the effects controlled for a large number of covariates (see text; preregistered); the grey estimates are without control variables (exploratory).
+The SESOI was b = |0.30| for life satisfaction and b = |0.15| for affect.
+Hence, all of the reported effects are not considered meaningful.}\label{fig:fig-within}
+\end{figure}
+
+\begin{table}[tbp]
+
+\begin{center}
+\begin{threeparttable}
+
+\caption{\label{tab:tab-within}Overview of all within-person effects.}
+
+\footnotesize{
+
+\begin{tabular}{lrrrrr}
+\toprule
+ &  & \multicolumn{2}{c}{Confidence interval}  &  &\\
+\cmidrule(r){3-4}
+Predictor & \multicolumn{1}{c}{b} & \multicolumn{1}{c}{Lower} & \multicolumn{1}{c}{Higher} & \multicolumn{1}{c}{beta} & \multicolumn{1}{c}{p}\\
+\midrule
+Life satisfaction &  &  &  &  & \\
+\ \ \ Reading & 0.03 & -0.01 & 0.07 & 0.02 & .091\\
+\ \ \ Liking \& Sharing & -0.02 & -0.08 & 0.03 & -0.01 & .299\\
+\ \ \ Posting & -0.05 & -0.11 & 0.01 & -0.02 & .114\\
+\ \ \ Facebook & 0.00 & -0.04 & 0.04 & 0.00 & .964\\
+\ \ \ Instagram & 0.01 & -0.01 & 0.03 & 0.01 & .226\\
+\ \ \ WhatsApp & 0.00 & -0.03 & 0.02 & 0.00 & .664\\
+\ \ \ YouTube & 0.02 & -0.03 & 0.06 & 0.01 & .402\\
+\ \ \ Twitter & -0.02 & -0.09 & 0.05 & -0.01 & .420\\
+Positive affect &  &  &  &  & \\
+\ \ \ Reading & -0.03 & -0.05 & -0.01 & -0.04 & .011\\
+\ \ \ Liking \& Sharing & 0.01 & -0.02 & 0.04 & 0.01 & .535\\
+\ \ \ Posting & 0.01 & -0.02 & 0.05 & 0.01 & .404\\
+\ \ \ Facebook & -0.01 & -0.04 & 0.02 & -0.01 & .581\\
+\ \ \ Instagram & 0.00 & -0.02 & 0.02 & 0.00 & .704\\
+\ \ \ WhatsApp & 0.01 & -0.02 & 0.04 & 0.01 & .444\\
+\ \ \ YouTube & 0.00 & -0.04 & 0.04 & 0.00 & .904\\
+\ \ \ Twitter & 0.01 & 0.00 & 0.02 & 0.01 & .129\\
+Negative affect &  &  &  &  & \\
+\ \ \ Reading & 0.00 & -0.01 & 0.01 & 0.00 & .971\\
+\ \ \ Liking \& Sharing & 0.02 & 0.00 & 0.03 & 0.02 & .019\\
+\ \ \ Posting & 0.05 & 0.02 & 0.08 & 0.04 & .009\\
+\ \ \ Facebook & -0.01 & -0.02 & 0.01 & -0.01 & .226\\
+\ \ \ Instagram & 0.00 & -0.01 & 0.01 & 0.00 & .371\\
+\ \ \ WhatsApp & 0.00 & -0.01 & 0.02 & 0.01 & .562\\
+\ \ \ YouTube & 0.01 & -0.01 & 0.04 & 0.02 & .250\\
+\ \ \ Twitter & 0.01 & 0.00 & 0.03 & 0.01 & .034\\
+\bottomrule
+\end{tabular}
+
+}
+
+\end{threeparttable}
+\end{center}
+
+\end{table}
 
@@
 
-In what follows, I briefly report some exploratory analyses that weren't preregistered.
-First, to contextualize the results reported above and to see if the results included any meaningful effects at all, I also looked at the effect sizes of selected (cherry-picked) covariates.
-Because each variable had different response options, we would actually need to define a SESOI for each variable, which for reasons of scope I cannot implement here.
-Therefore, I report the results of the standardized scales,
-which allows for a better comparison across the differently scaled variables.
-For what it's worth, as a rough estimate for the SESOI we can build on the typical convention that small effects start at \emph{r} = \textbar.10\textbar.
-The results showed that several effects fell outside of the SESOI, were hence considered meaningful.
-This includes for example internal locus of control, health, satisfaction with democracy, or exercising.
-For an overview, see Figure \ref{fig:res-control}.
-
-To find out whether my inferences were robust across legitimate (though arguably inferior) alternative analyses, I reran the analyses also using standardized estimates, mean scores instead of factor scores, and with a data set where missing data were not imputed.
-The results were virtually the same.
-For example, all within-person standardized COVID-19 related types of social media use or channels were significantly smaller than \(\beta\) = \textbar.05\textbar, again supporting that effects were tiny.
-The additional analyses are reported in Figure 3 and Figure 4, and in the \href{https://XMtRA.github.io/Austrian_Corona_Panel_Project/analyses_additional.html}{companion website}.
+To contextualize the results reported above and to see if the study included any meaningful effects at all, I also looked at the effect sizes of the covariates.
+Because each variable featured different response options, which would require defining a SESOI for each variable, I hence report the results of the standardized scales, which allows for a better comparison across the differently scaled variables.
+Here, we can build on Cohen's convention that small effects begin at \emph{r} = \textbar.10\textbar.
+
+The results showed that several effects crossed or fell completely outside of the SESOI.
+They can hence be considered meaningful.
+For example, if physical health decreased, this had a meaningful detrimental impact on life satisfaction (\(\beta\) = .18 {[}95\% CI .16, .20{]}), positive affect (\(\beta\) = .17 {[}95\% CI .16, .18{]}), and negative affect (\(\beta\) = -.18 {[}95\% CI -.19, -.17{]}).
+Spending more time outside to exercise meaningfully increased positive affect (\(\beta\) = .12 {[}95\% CI .11, .13{]}).
+Being more satisfied with the current democratic system meaningfully increased life satisfaction (\(\beta\) = .13 {[}95\% CI .12, .14{]}).
+The strongest aspect affecting well-being was internal locus of control.
+If people felt more in control of their lives, this strongly increased both life satisfaction (\(\beta\) = .31 {[}95\% CI .28, .34{]}) and
+positive affect (\(\beta\) = .27 {[}95\% CI .25, .29{]}),
+while decreasing negative affect (\(\beta\) = -.28 {[}95\% CI -.30, -.26{]}).
+
+For an overview, see Figure \ref{fig:fig-control}.
+
+\begin{figure}
+\includegraphics[width=\textwidth]{manuscript_files/figure-latex/fig-control-1} \caption{Results of main variables together with covariates to provide context.
+All variables standardized.
+SESOI: beta = |.10|}\label{fig:fig-control}
+\end{figure}
 
@@
 
-In this study I analyzed the effects of COVID-19 related social media use on well-being.
-The data come from a panel study with 24 waves and are largely representative of the Austrian population.
-In a random effects model I separated between person relations from within-person effects and controlled for a large number of both stable and varying covariates, thereby aiming to assess causality.
-The results showed that within-person effects were trivial.
-People who used social media more than usual to learn about COVID-19 didn't show meaningful changes in their well-being.
-
-The results imply that COVID-19 related social media use doesn't seem to be particularly relevant for well-being.
-Other factors among the third variables that were measured revealed larger effects or relations, suggesting that well-being is determined by alternative aspects such as health, satisfaction with democracy, locus of control, or exercising.
-According to this study, popular fears that ``doomscrolling'' or overusing social media during times of crises are not supported.
-
-On the one hand, the results are not aligned with several recent studies analyzing similar or closely related research questions.
-This includes a study by Bendau et al. (2021), which showed negative relations between social media and well-being (but see Bradley \& Howard, 2021; or Sewall et al., 2021).
-However, note that Bendau et al. (2021) analyzed cross-sectional data on a between-person level while not controlling for third variables, which is not optimal for investigating causal effects.
-On the other hand, the results are well-aligned with many seminal Communication theories such as the uses and gratifications model (Katz et al., 1973).
-If effects were indeed profoundly negative on average, then people likely wouldn't spend so much time on social media.
-Likewise, recent studies and meta-analyses analyzing the effects of social media use from a more general perspective or from a somewhat different angle.
-These studies have found that the effects of various types of social media use on several well-being indicators are small at best, often too small to matter (Ferguson et al., 2021; Meier \& Reinecke, 2020; Orben, 2020), which echoes the results obtained here.
-
-If anything, two preliminary and subtle trends can be observed.
-First, in terms of media channels, using Twitter more than usual was related to slightly decreased levels of life satisfaction.
-Twitter is considered to have more negative affordances and tonality as compared to other networks such as Instagram (Waterloo et al., 2018), which might help explain the results.
-Instagram, on the other hand, was related to slightly increased levels of life satisfaction.
-To speculate, the often-criticized positivity bias on Instagram might have been somewhat beneficial in times of the pandemic.
-The critique that the positivity bias necessarily leads to envy and negative feelings is unrealistic and myopic, because positive content can also inspire and motivate users (Meier, Gilbert, Börner, \& Possler, 2020), which could be especially helpful in times of lockdown and home-office.
-Second, of all the three COVID-19 related social media activities, people who read about the pandemic more than others showed slightly decreased levels of positive affect, and people who actively posted about the pandemic more than others showed slightly increased levels of negative affect.
-On the other hand, however, people who posted more about COVID-19 also showed slightly higher levels of positive affects, so taken together the results are ambivalent.
-But future research might elaborate on these specific relations to probe their stability and relevance.
+Based on a panel study with 32 waves largely representative of the Austrian population, this study analyzed the effects of COVID-19 related social media use on well-being.
+Between person correlation analyses showed that more active users of COVID-19 related content on social media also reported decreased well-being.
+The within-person relations, which are more informative regarding causal effects, showed a different pattern, however.
+If people consumed more COVID-19 content on social media than usual, this did not meaningfully reduce their well-being.
+Several statistically significant effects were found, but these were very small.
+For example, people who liked and shared more COVID-19 related posts than usual reported slightly higher levels of negative affect.
+Posting more content about COVID-19 than usual slightly increased negative affect.
+People who read more COVID-19 related posts than usual reported slightly decreased positive affect.
+Using Twitter for COVID-19 related content slightly increased negative affect.
+Again, although all of these within-person effects were statistically significant, they were very small, smaller than the predefined smallest effect size of interest.
+According to the preregistered procedure, they should hence be considered irrelevant.
+Other factors, for which we would expect to find meaningful effects, such as health or sports, indeed showed substantial and meaningful impacts on well-being.
+In conclusion, COVID-19 related activity on social media was not a particularly strong influence on peoples' well-being.
+The results do not support the popular fears that ``doomscrolling'' or overusing social media during times of crises constitutes a relevant risk for well-being.
+
+These specific observations notwithstanding, several general trends can be observed.
+First, overall the results do suggest that effects of COVID-19 related social media use on well-being tend to take place in the negative as opposed to the positive spectrum.
+Although very small, four statistically significant negative results of COVID-19 related social media use on well-being were found.
+Not a single positive effect emerged.
+Also note that in the analyses several control variables were included, ruling out plausible alternative explanations for the negative results.
+For example, it was controlled for whether or not participants contracted a COVID-19 infection during a specific wave.
+Hence, we can rule out the alternative explanation that having an infection was the root cause of increased communication and reduced well-being.
+
+Second, significant outcomes emerged for positive or negative affect, but not for life satisfaction.
+Life satisfaction was stable and not affected by any type or channel of social media communication.
+The more fluctuating positive and negative affect, however, were affected.
+Liking, sharing, and posting COVID-19 related content, and spending time on Twitter to browse COVID-19 related content, all negatively influenced affect.
+This is aligned with prior findings which showed that social media can trigger negative affect but does not reduce life satisfaction (Huang, 2017).
+Conversations about COVID-19 on social media are often extreme, negative, or aggressive (Fan et al., 2020).
+More deeply engaging with this type of content could negatively affect active authors.
+The hypothesis that tonality could explain the negative effects is especially supported by the observation that spending more time on Twitter than usual increased negative affect.
+Communication on Twitter is often found to be more negative and impolite compared to other SNSs (Halpern \& Gibbs, 2013).
+Consuming more negative and misleading information could hence explain the (slightly) increased levels of negative affect.
+
+Third, the results show that it makes sense to analyze different communication types and communication channels.
+Reading reduced positive affect, while liking, sharing and commenting increased negative affect.
+Whereas it was often stated that passive use is bad and active use good (Verduyn et al., 2015), this pattern thus did not emerge here.
+Instead, all three types were negative.
+The results hence support the findings from Valkenburg et al. (2022), who also could not confirm the claim that active use is good and that passive use is bad.
+Focusing on communication channels, Twitter seems to be more negative, as has often been observed (Halpern \& Gibbs, 2013).
+But, again, all of these effects are very small.
+Future research might elaborate on these specific relations to probe their stability and relevance.
+
+Taken together, the results are hence aligned with the underlying theoretical models and prior empirical results.
+The findings support the Different Susceptibility to Media Effects Theory, which states that effects are generally small and depend on the type of communication.
+The results are well-aligned with mood management theory (Zillmann, 1988) or the uses and gratifications approach (Katz et al., 1973), whose premises preclude particularly negative effects of routine and widespread media consumption.
+If the effects of social media were indeed profoundly negative on average, then people likely would not spend so much time on social media engaging with COVID-19 content.
+Likewise, recent empirical studies and meta-analyses reported rather small negative effects, too.
+Several studies found that the effects of various types of social media use on well-being are small, often too small to matter (Bendau et al., 2021; Ferguson et al., 2021; Meier \& Reinecke, 2020; Orben, 2020), echoing the results obtained here.
+
+In light of the overall very small effects, engaging in COVID 19-related social media use should not be a major concern for one's well-being.
+That said, the results still imply that it does make sense to critically reflect upon COVID-19 related social media use.
+On average, it is likely more beneficial to post less actively about COVID-19 on social media and to spend less time on Twitter consuming COVID-19 related content.
+The results allow us to reject a positive effect: Writing more posts on social media will likely not increase well-being.
 
@@
 
-The current study analyzed whether changes in media use were related to changes in well-being, while controlling for several potential confounders.
-Together, this allowed for an improved perspective on assessing causality.
-That said, causality necessitates temporal order, and the cause needs to precede the effect.
-Regarding media use, such effects often happen immediately or shortly after use, necessitating intervals in the hours, minutes, or even seconds.
-In many cases only experience sampling studies asking users in the very moment can produce such knowledge.
-However, even then we don't know for certain if we actually measured the right interval.
-Effects depend on the intensity of use or the length of the interval, and to borrow the words from Rohrer and Murayama (2021), there is no such thing as ``the'' effect of social media use on well-being.
-Hence, to document how effects unfold, future research needs to employ different study designs probing different time lags.
-In addition, more thought needs to be invested in what relevant stable and varying factors we should include as control variables, and I hope this study provides a first step into this direction.
-
-Although I had already reduced the predefined SESOIs to be less conservative, they were potentially still too large.
+Focusing on within-person effects and controlling for several potential confounders, this study provides an improved perspective on assessing causality.
+However, three major challenges remain.
+First, it could be that the timing of the interval was wrong; it could be that effects unfold in shorter or longer intervals.
+Next, it could be that there was not sufficient variance in media use and well-being.
+Without sufficient variability, statistical models cannot detect existing effects.
+Third, it is necessary to control for \emph{all} relevant confounding third variables.
+Although this study included are large list of confounders, it could still be that crucial variables were missed (Zillmann, 1988).
+To borrow the words from Rohrer and Murayama (2021), there is no such thing as ``the'' effect of social media use on well-being.
+The results reported here are hence contingent on the design I used.
+To document how effects unfold, future research needs to employ different study designs probing different intervals.
+More thought needs to be invested in what factors to control for, and I hope this study provides a first step into this direction.
+
+Although I had already reduced the predefined SESOIs to be less conservative, one could argue they were still too large.
 Media use is only one aspect of several factors that simultaneously affect well-being.
-Is it really realistic to expect that extremely changing only \emph{one} of these aspects should manifest in a detectable change in well-being?
+Is it realistic to expect that changing only \emph{one} of these aspects should already manifest in a detectable change in well-being?
 Or would it make more sense to expect that thoroughly committing to say \emph{two} activities (e.g.~regularly exercising \emph{and} establishing a reading habit) should then cause a detectable improvement in well-being?
-Practically, this would imply a SESOI half as large as I have defined here, namely \emph{b} = \textbar.15\textbar{} for well-being and \emph{b} = \textbar.075\textbar{} for affect.
-In the case of this study, however, reducing the SESOI would not even make a big difference, as also with these more liberal thresholds all but three effect would still be completely in the null region, and no effect would be outside of the null region.
-However, at all events I encourage future research to start a thorough conversation on what effect sizes are considered meaningful and what not.
-Again, with this study I hope to provide some first input and guidelines.
+Practically, this would imply a SESOI half the size defined here, namely \emph{b} = \textbar.15\textbar{} for life satisfaction and \emph{b} = \textbar.075\textbar{} for affect.
+In the case of this study, however, even halving the SESOI would not make a difference.
+All but one effect would still be completely in the null region, and no effect would fall completely outside of the null region.
+I encourage future research to elaborate on what effect sizes are considered meaningful and what not.
 
 Both media use and well-being were measured using self-reports.
-Measuring well-being with self-reports is adequate, because it by definition requires introspection.
-However, it would be preferable to measure social media use objectively, as people cannot reliably estimate their use (Scharkow, 2016).
-That said, objective measures often cannot capture the content or the motivation of the use, and only very complicated tools recording the actual content (such as the Screenome project) might produce such data.
-Unfortunately, such procedures introduce other problems, especially related to privacy.
-Hence, for this type of research question it still seems necessary to use self-reported measures, and in many cases they can still be very informative (Verbeij, Pouwels, Beyens, \& Valkenburg, 2021).
-
-Because the data were collected in a single country, the generalizability of the results is limited.
-The results apply primarily to the more Western sphere, and might not hold true in other cultures, especially cultures with a different media landscape or alternative social media channels.
-That said, because this is a comparatively large study largely representative of an entire country, and because several waves were collected across a large time span, the results should be at least as generalizable as other typical empirical studies collected in the social sciences.
+Because assessing well-being necessarily requires introspection, using self-reports for affect and life satisfaction is adequate.
+However, for social media use objective measures are clearly preferable, as people often cannot reliably estimate their use (Scharkow, 2016).
+At the same time, most objective measures cannot capture the content or the motivation of use.
+Hence, for the type of research question analyzed here, it still seems necessary to use self-reported measures.
+In many cases they can still be informative (Verbeij et al., 2021).
+
+Being collected in a single country, the generalizability of the results is limited.
+The results apply primarily to the more Western sphere.
+They might not hold true in other cultures, especially cultures with a different media landscape or alternative social media channels.
 
@@
 
-In this study, COVID-19 related social media use didn't meaningfully affect several indicators of well-being, including life satisfaction, positive affect, and negative affect.
-However, factors other than social media use were meaningfully related to well-being, such as physical health, exercise, satisfaction with democracy, or believing that one is in control of one's life.
-If it's our aim to improve well-being, it might hence be more fruitful not to focus so much on social media but to address other, more pertinent societal problems related to health care, regular exercise, or a functioning democratic system.
+In this study, COVID-19 related social media use did not meaningfully affect well-being.
+Very small negative effects were found for writing COVID-19 related posts, sharing COVID-19 related content, and spending more time than usual on Twitter.
+Factors other than social media use, however, were meaningfully related to well-being, including physical health, exercise, satisfaction with democracy, or believing that one is in control of one's life.
+Hence, when trying to improve well-being during a pandemic, instead of focusing on social media it seems more fruitful to address other, more pertinent societal problems related to health care, regular exercise, or a functioning democratic system.
 
@@
 \begin{CSLReferences}{1}{0}
-\leavevmode\vadjust pre{\hypertarget{ref-baguleyStandardizedSimpleEffect2009}{}}%
-Baguley, T. (2009). Standardized or simple effect size: {What} should be reported? \emph{British Journal of Psychology}, \emph{100}(3), 603--617. \url{https://doi.org/10.1348/000712608X377117}
-
 \leavevmode\vadjust pre{\hypertarget{ref-bellFixedRandomEffects2019}{}}%
@@
 \url{https://doi.org/10.1177/02654075211029378}
-
-\leavevmode\vadjust pre{\hypertarget{ref-cohenPowerPrimer1992}{}}%
-Cohen, J. (1992). A power primer. \emph{Psychological Bulletin}, \emph{112}(1), 155--159. \url{https://doi.org/10.1037/0033-2909.112.1.155}
 
@@
 
+\leavevmode\vadjust pre{\hypertarget{ref-dienlinDisplacementReinforcementReciprocity2017}{}}%
+Dienlin, T., Masur, P.
+K., \& Trepte, S.
+(2017).
+Displacement or reinforcement?
+{The} reciprocity of {FtF}, {IM}, and {SNS} communication and their effects on loneliness and life satisfaction.
+\emph{Journal of Computer-Mediated Communication}, \emph{22}(2), 71--87.
+\url{https://doi.org/10.1111/jcc4.12183}
+
+\leavevmode\vadjust pre{\hypertarget{ref-dormannOptimalTimeLags2015}{}}%
+Dormann, C., \& Griffin, M. A. (2015). Optimal time lags in panel studies. \emph{Psychological Methods}, \emph{20}(4), 489--505. \url{https://doi.org/10.1037/met0000041}
+
 \leavevmode\vadjust pre{\hypertarget{ref-dornemannHowGoodBad2021}{}}%
@@
 {The} role of similarity}.
-{OSF Preprints}.
 \url{https://doi.org/10.31219/osf.io/sy2kd}
-
-\leavevmode\vadjust pre{\hypertarget{ref-duradoniWellbeingSocialMedia2020}{}}%
-Duradoni, M., Innocenti, F., \& Guazzini, A.
-(2020).
-Well-being and social media: {A} systematic review of {Bergen Addiction Scales}.
-\emph{Future Internet}, \emph{12}(2), 24.
-\url{https://doi.org/10.3390/fi12020024}
 
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-ellisonWhyWeDon2020}{}}%
-Ellison, N.
-B., Triẹû, P., Schoenebeck, S., Brewer, R., \& Israni, A.
+\leavevmode\vadjust pre{\hypertarget{ref-europeansocialsurveyESS9Edition20182021}{}}%
+European Social Survey. (2021). \emph{{ESS9} edition 3.1 - 2018 {Documentation Report}}. \url{https://www.europeansocialsurvey.org/docs/round9/survey/ESS9_data_documentation_report_e03_1.pdf}
+
+\leavevmode\vadjust pre{\hypertarget{ref-fanStigmatizationSocialMedia2020}{}}%
+Fan, L., Yu, H., \& Yin, Z.
 (2020).
-Why we don't click: {Interrogating} the relationship between viewing and clicking in social media contexts by exploring the {``{Non-Click}.''} \emph{Journal of Computer-Mediated Communication}, \emph{25}(6), 402--426.
-\url{https://doi.org/10.1093/jcmc/zmaa013}
-
-\leavevmode\vadjust pre{\hypertarget{ref-europeansocialsurveyESS9Edition20182021}{}}%
-European Social Survey. (2021). \emph{{ESS9} edition 3.1 - 2018 {Documentation Report}}.
+Stigmatization in social media: {Documenting} and analyzing hate speech for {\textsc{COVID}} ‐19 on {Twitter}.
+\emph{Proceedings of the Association for Information Science and Technology}, \emph{57}(1).
+\url{https://doi.org/10.1002/pra2.313}
 
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-galerHowMuchToo2018}{}}%
-Galer, S. S. (2018). \emph{How much is {``too much time''} on social media?} https://www.bbc.com/future/article/20180118-how-much-is-too-much-time-on-social-media.
-
-\leavevmode\vadjust pre{\hypertarget{ref-greenspoonIntegrationSubjectiveWellbeing2001}{}}%
-Greenspoon, P.
-J., \& Saklofske, D.
-H.
-(2001).
-Toward an integration of subjective well-being and psychopathology.
-\emph{Social Indicators Research}, \emph{54}(1), 81--108.
-\url{https://doi.org/10.1023/A:1007219227883}
-
 \leavevmode\vadjust pre{\hypertarget{ref-guazziniSecondWaveAnalysis2022}{}}%
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-johannesNoEffectDifferent2022}{}}%
-Johannes, N., Dienlin, T., Bakhshi, H., \& Przybylski, A.
-K.
-(2022).
-No effect of different types of media on well-being.
-\emph{Scientific Reports}, \emph{12}(1), 61.
-\url{https://doi.org/10.1038/s41598-021-03218-7}
-
 \leavevmode\vadjust pre{\hypertarget{ref-katzUsesGratificationsResearch1973}{}}%
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-kerestesAdolescentsOnlineSocial2020}{}}%
-Keresteš, G., \& Štulhofer, A.
-(2020).
-Adolescents' online social network use and life satisfaction: {A} latent growth curve modeling approach.
-\emph{Computers in Human Behavior}, \emph{104}, 106187.
-\url{https://doi.org/10.1016/j.chb.2019.106187}
-
-\leavevmode\vadjust pre{\hypertarget{ref-kittelAustrianCoronaPanel2020}{}}%
-Kittel, B., Kritzinger, S., Boomgaarden, H., Prainsack, B., Eberl, J.-M., Kalleitner, F., \ldots{} Schlogl, L.
-(2020).
-\emph{Austrian {Corona Panel Project} ({SUF} edition)}.
-{AUSSDA}.
-\url{https://doi.org/10.11587/28KQNS}
-
 \leavevmode\vadjust pre{\hypertarget{ref-kittelAustrianCoronaPanel2021}{}}%
-Kittel, B., Kritzinger, S., Boomgaarden, H., Prainsack, B., Eberl, J.-M., Kalleitner, F., \ldots{} Schlogl, L.
+Kittel, B., Kritzinger, S., Boomgaarden, H., Prainsack, B., Eberl, J.-M., Kalleitner, F., Lebernegg, N.
+S., Partheymüller, J., Plescia, C., Schiestl, D.
+W., \& Schlogl, L.
 (2021).
@@
 
+\leavevmode\vadjust pre{\hypertarget{ref-kittelAustrianCoronaPanel2020}{}}%
+Kittel, B., Kritzinger, S., Boomgaarden, H., Prainsack, B., Eberl, J.-M., Kalleitner, F., Lebernegg, N.
+S., Partheymüller, J., Plescia, C., Schiestl, D.
+W., \& Schlogl, L.
+(2020).
+\emph{Austrian {Corona Panel Project} ({SUF} edition)} {[}Data set{]}.
+{AUSSDA}.
+\url{https://doi.org/10.11587/28KQNS}
+
 \leavevmode\vadjust pre{\hypertarget{ref-kleinDarklySoothingCompulsion2021}{}}%
-Klein, J. (2021). \emph{The darkly soothing compulsion of 'doomscrolling'}. https://www.bbc.com/worklife/article/20210226-the-darkly-soothing-compulsion-of-doomscrolling.
-
-\leavevmode\vadjust pre{\hypertarget{ref-klinePrinciplesPracticeStructural2016}{}}%
-Kline, R. B. (2016). \emph{Principles and practice of structural equation modeling} (Fourth). {New York, NY}: {The Guilford Press}.
+Klein, J. (2021, March 3). \emph{The darkly soothing compulsion of 'doomscrolling'}. \url{https://www.bbc.com/worklife/article/20210226-the-darkly-soothing-compulsion-of-doomscrolling}
+
+\leavevmode\vadjust pre{\hypertarget{ref-latikkaLonelinessPsychologicalDistress2022}{}}%
+Latikka, R., Koivula, A., Oksa, R., Savela, N., \& Oksanen, A.
+(2022).
+Loneliness and psychological distress before and during the {COVID-19} pandemic: {Relationships} with social media identity bubbles.
+\emph{Social Science \& Medicine}, \emph{293}, 114674.
+\url{https://doi.org/10.1016/j.socscimed.2021.114674}
 
@@
 
+\leavevmode\vadjust pre{\hypertarget{ref-lucasWhyCrossLaggedPanel2022}{}}%
+Lucas, R. E. (2022). \emph{Why the {Cross-Lagged Panel Model} is {Almost Never} the {Right Choice}} {[}Preprint{]}. {PsyArXiv}. \url{https://doi.org/10.31234/osf.io/pkec7}
+
 \leavevmode\vadjust pre{\hypertarget{ref-lykkenHappinessWhatStudies1999}{}}%
-Lykken, D. T. (1999). \emph{Happiness: {What} studies on twins show us about nature, nurture, and the happiness set-point}. {New York, N.Y.}: {Golden Books}.
+Lykken, D. T. (1999). \emph{Happiness: {What} studies on twins show us about nature, nurture, and the happiness set-point}. {Golden Books}.
 
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-mcelreathYesterdayClass2021}{}}%
-McElreath, R. (2021). Yesterday in class, ... {[}Tweet{]}.
-
-\leavevmode\vadjust pre{\hypertarget{ref-meierInstagramInspirationHow2020}{}}%
-Meier, A., Gilbert, A., Börner, S., \& Possler, D.
-(2020).
-Instagram inspiration: {How} upward comparison on social network sites can contribute to well-being.
-\emph{Journal of Communication}, \emph{70}(5), 721--743.
-\url{https://doi.org/10.1093/joc/jqaa025}
-
 \leavevmode\vadjust pre{\hypertarget{ref-meierDoesPassiveSocial2022}{}}%
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-metzgerComparativeOptimismPrivacy2017}{}}%
-Metzger, M. J., \& Suh, J. J. (2017). Comparative optimism about privacy risks on {Facebook}. \emph{Journal of Communication}, \emph{67}(2), 203--232. \url{https://doi.org/10.1111/jcom.12290}
-
 \leavevmode\vadjust pre{\hypertarget{ref-normanInterpretationChangesHealthrelated2003}{}}%
@@
 \url{https://doi.org/10.1007/s00127-019-01825-4}
-
-\leavevmode\vadjust pre{\hypertarget{ref-orbenSocialMediaEnduring2019}{}}%
-Orben, A., Dienlin, T., \& Przybylski, A.
-K.
-(2019).
-Social media's enduring effect on adolescent life satisfaction.
-\emph{Proceedings of the National Academy of Sciences of the United States of America}, \emph{116}(21), 10226--10228.
-\url{https://doi.org/10.1073/pnas.1902058116}
 
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-przybylskiDoesTakingShort2021a}{}}%
-Przybylski, A.
-K., Nguyen, T.
-T., Law, W., \& Weinstein, N.
-(2021).
-Does taking a short break from social media have a positive effect on well-being?
-{Evidence} from three preregistered field experiments.
-\emph{Journal of Technology in Behavioral Science}, \emph{6}(3), 507--514.
-\url{https://doi.org/10.1007/s41347-020-00189-w}
-
-\leavevmode\vadjust pre{\hypertarget{ref-przybylskiLargescaleTestGoldilocks2017}{}}%
-Przybylski, A. K., \& Weinstein, N. (2017). A large-scale test of the {Goldilocks} hypothesis. \emph{Psychological Science}, \emph{28}(2), 204--215. \url{https://doi.org/10.1177/0956797616678438}
-
 \leavevmode\vadjust pre{\hypertarget{ref-riehmAssociationsMediaExposure2020}{}}%
@@
 (2021).
-\emph{These are not the effects you are looking for: {Causality} and the within-/between-person distinction in longitudinal data analysis} {[}Preprint{]}.
+\emph{These are not the effects you are looking for: {Causality} and the within-/between-person distinction in longitudinal data analysis}.
 {PsyArXiv}.
@@
 (2016).
-The accuracy of self-reported {Internet} use\textemdash{{A}} validation study using client log data.
+The accuracy of self-reported {Internet} use---{A} validation study using client log data.
 \emph{Communication Methods and Measures}, \emph{10}(1), 13--27.
 \url{https://doi.org/10.1080/19312458.2015.1118446}
-
-\leavevmode\vadjust pre{\hypertarget{ref-scharkowHowSocialNetwork2020}{}}%
-Scharkow, M., Mangold, F., Stier, S., \& Breuer, J.
-(2020).
-How social network sites and other online intermediaries increase exposure to news.
-\emph{Proceedings of the National Academy of Sciences}, \emph{117}(6), 2761--2763.
-\url{https://doi.org/10.1073/pnas.1918279117}
-
-\leavevmode\vadjust pre{\hypertarget{ref-schemerImpactInternetSocial2021}{}}%
-Schemer, C., Masur, P.
-K., Geiß, S., Müller, P., \& Schäfer, S.
-(2021).
-The impact of {Internet} and social media use on well-being: {A} longitudinal analysis of adolescents across nine years.
-\emph{Journal of Computer-Mediated Communication}, \emph{26}(1), 1--21.
-\url{https://doi.org/10.1093/jcmc/zmaa014}
-
-\leavevmode\vadjust pre{\hypertarget{ref-schnauber-stockmannMobileDevicesTools2020}{}}%
-Schnauber-Stockmann, A., \& Karnowski, V.
-(2020).
-Mobile devices as tools for media and communication research: {A} scoping review on collecting self-report data in repeated measurement designs.
-\emph{Communication Methods and Measures}, \emph{14}(3), 145--164.
-\url{https://doi.org/10.1080/19312458.2020.1784402}
 
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-vanrooijWeakScientificBasis2018}{}}%
-van Rooij, A.
-J., Ferguson, C.
-J., Colder Carras, M., Kardefelt-Winther, D., Shi, J., Aarseth, E., \ldots{} Przybylski, A.
-K.
-(2018).
-A weak scientific basis for gaming disorder: {Let} us err on the side of caution.
-\emph{Journal of Behavioral Addictions}, \emph{7}(1), 1--9.
-\url{https://doi.org/10.1556/2006.7.2018.19}
+\leavevmode\vadjust pre{\hypertarget{ref-valkenburgAssociationsActivePassive2022}{}}%
+Valkenburg, P.
+M., van Driel, I.
+I., \& Beyens, I.
+(2022).
+The associations of active and passive social media use with well-being: {A} critical scoping review.
+\emph{New Media \& Society}, \emph{24}(2), 530--549.
+\url{https://doi.org/10.1177/14614448211065425}
 
@@
 
+\leavevmode\vadjust pre{\hypertarget{ref-verduynPassiveFacebookUsage2015}{}}%
+Verduyn, P., Lee, D.
+S., Park, J., Shablack, H., Orvell, A., Bayer, J., Ybarra, O., Jonides, J., \& Kross, E.
+(2015).
+Passive {Facebook} usage undermines affective well-being: {Experimental} and longitudinal evidence.
+\emph{Journal of Experimental Psychology: General}, \emph{144}(2), 480--488.
+\url{https://doi.org/10.1037/xge0000057}
+
 \leavevmode\vadjust pre{\hypertarget{ref-wagnerAUTNESOnlinePanel2018}{}}%
 Wagner, M., Aichholzer, J., Eberl, J.-M., Meyer, T.
-M., Berk, N., Büttner, N., \ldots{} Müller, W.
+M., Berk, N., Büttner, N., Boomgaarden, H., Kritzinger, S., \& Müller, W.
 C.
 (2018).
-\emph{{AUTNES Online Panel Study} 2017 ({SUF} edition)}.
+\emph{{AUTNES Online Panel Study} 2017 ({SUF} edition)} {[}Data set{]}.
 {AUSSDA}.
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-waterlooNormsOnlineExpressions2018}{}}%
-Waterloo, S.
-F., Baumgartner, S.
-E., Peter, J., \& Valkenburg, P.
-M.
-(2018).
-Norms of online expressions of emotion: {Comparing Facebook}, {Twitter}, {Instagram}, and {WhatsApp}.
-\emph{New Media \& Society}, \emph{20}(5), 1813--1831.
-\url{https://doi.org/10.1177/1461444817707349}
-
 \leavevmode\vadjust pre{\hypertarget{ref-worldhealthorganizationWellbeingMeasuresPrimary1998}{}}%
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-yangCanWatchingOnline2021}{}}%
-Yang, Z., Griffiths, M.
-D., Yan, Z., \& Xu, W.
-(2021).
-Can watching online videos be addictive?
-{A} qualitative exploration of online video watching among {Chinese} young adults.
-\emph{International Journal of Environmental Research and Public Health}, \emph{18}(14), 7247.
-\url{https://doi.org/10.3390/ijerph18147247}
+\leavevmode\vadjust pre{\hypertarget{ref-yuePassiveSocialMedia2022}{}}%
+Yue, Z., Zhang, R., \& Xiao, J.
+(2022).
+Passive social media use and psychological well-being during the {COVID-19} pandemic: {The} role of social comparison and emotion regulation.
+\emph{Computers in Human Behavior}, \emph{127}, 107050.
+\url{https://doi.org/10.1016/j.chb.2021.107050}
 
@@
 
-\newpage
-
-\begin{table}[tbp]
-
-\begin{center}
-\begin{threeparttable}
-
-\caption{\label{tab:tab-descriptives}Descriptives of the main variables.}
-
-\begin{tabular}{lllll}
-\toprule
- & \multicolumn{1}{c}{sd} & \multicolumn{1}{c}{min} & \multicolumn{1}{c}{max} & \multicolumn{1}{c}{mean}\\
-\midrule
-Well-being &  &  &  & \\
-\ \ \ Life satisfaction & 1.68 & 6.38 & 6.81 & 6.59\\
-\ \ \ Positive affect & 0.57 & 3.05 & 3.28 & 3.15\\
-\ \ \ Negative affect & 0.39 & 1.66 & 1.81 & 1.73\\
-Social media use &  &  &  & \\
-\ \ \ Read & 1.03 & 2.10 & 2.92 & 2.43\\
-\ \ \ Like \& share & 0.87 & 1.61 & 1.92 & 1.78\\
-\ \ \ Posting & 0.63 & 1.33 & 1.46 & 1.39\\
-Social media channel &  &  &  & \\
-\ \ \ Facebook & 0.96 & 2.34 & 2.68 & 2.45\\
-\ \ \ Twitter & 0.52 & 1.16 & 1.72 & 1.37\\
-\ \ \ Instagram & 0.83 & 1.84 & 2.66 & 2.09\\
-\ \ \ WhatsApp & 1.23 & 2.28 & 2.62 & 2.45\\
-\ \ \ YouTube & 0.88 & 1.77 & 2.30 & 2.00\\
-\bottomrule
-\end{tabular}
-
-\end{threeparttable}
-\end{center}
-
-\end{table}
-
-\newpage
-
-\begin{figure}
-\centering
-\includegraphics{manuscript_files/figure-latex/sesoi-1.pdf}
-\caption{\label{fig:sesoi}Illustration of how confidence intervals are used to test a null region.
-Here, a trivial effect of social media use on life satisfaction is defined as ranging from b = -.25 to b = .25}
-\end{figure}
-
-\newpage
-
-\begin{figure}
-\includegraphics[width=\textwidth]{figures/fig_descriptives} \caption{Development of well-being and media use measures across the pandemic.
-Values obtained from mixed effect models, with participants and waves as grouping factors and without additional predictors.}\label{fig:descriptives}
-\end{figure}
-
-\newpage
-
-\begin{figure}
-\includegraphics[width=\textwidth]{manuscript_files/figure-latex/res-activity-1} \caption{The effects of various types of social media use on three indicators of well-being.
-The black estimates show the effects controlled for a large number of covariates (see text; preregistered); the grey estimates are without control variables (exploratory).
-The SESOI was b = |0.30| for life satisfaction and b = |0.15| for affect.
-Hence, all of the reported effects are not considered meaningful.}\label{fig:res-activity}
-\end{figure}
-
-\newpage
-
-\begin{figure}
-\includegraphics[width=\textwidth]{manuscript_files/figure-latex/res-channels-1} \caption{The effects of using various social media applications on three indicators of well-being.
-The black estimates show the effects controlled for a large number of covariates (see text; preregistered); the grey estimates are without control variables (exploratory).
-The SESOI was b = |0.30| for life satisfaction and b = |0.15| for affect.
-Hence, all of the reported effects are not considered meaningful.}\label{fig:res-channels}
-\end{figure}
-
-\newpage
-
-\begin{figure}
-\includegraphics[width=\textwidth]{manuscript_files/figure-latex/res-control-1} \caption{Results of selected covariates.
-All variables were standardize except 'Male'  and ‘Employed in public service', because there were measured on a binary scale.}\label{fig:res-control}
-\end{figure}
-
-\newpage
-
 \hypertarget{competing-interests}{%

### 2023-01-30 manuscript.Rmd
- updated html

@@
     latex_engine  : xelatex
+# documentclass     : "apa6"
 # output            : papaja::apa6_docx

### 2023-07-11 analyses.RmD
Major updates Implemented feedback

--- 2022-04-12 analyses.RmD
@@
   - "bibliography/references.bib"
-  # pdf_document:
-  #   toc: TRUE
 ---
@@
 set.seed(170819)
-knitr::opts_chunk$set(warning=F, echo=T, error=F, message=F, cache = F)
-options(digits = 3, width = 250, scipen=999)
+
+knitr::opts_chunk$set(
+  warning=F, 
+  echo=T, 
+  error=F, 
+  message=F, 
+  cache=T, 
+  rows.print=20
+  )
+
+options(
+  digits=2, 
+  width=260, 
+  scipen=999
+  )
 ```
 
-Here you can find the code and the results of all analyses. To see the code, click on button "Code". Alternatively, you can download the rmd file from the github repo.
+ - Here you can find the code and the results of all main analyses. 
+ - To see the code, click on button "Code".
+ - Alternatively, you can download the rmd file from the github repo.
 
@@
 
-Load packages.
+Load packages & workspace.
 
-```{r r-setup, results='hide', cache=FALSE}
+```{r r-setup, results='hide', cache=F}
 # install packages
-# devtools::install_github("https://github.com/tdienlin/td@v.0.0.2.5")
+## devtools::install_github("https://github.com/tdienlin/td@v.0.0.2.5")
 
 # define packages
-packages <- c("broom.mixed", "brms", "devtools", "GGally", "ggplot2", 
-              "gridExtra", "kableExtra", "knitr", "lavaan", "lme4", 
-              "magrittr", "mice", #"mvnormalTest", 
-              "PerFit", "psych", "quanteda.textstats", "semTools", "tidyverse")
+library(broom.mixed)
+library(brms)
+library(corrr)
+library(devtools)
+library(GGally)
+library(ggplot2)
+library(gridExtra)
+library(kableExtra)
+library(knitr)
+library(lavaan)
+library(lme4)
+library(magrittr)
+library(mice)
+library(PerFit)
+library(performance)
+library(psych)
+library(quanteda.textstats)
+library(scales)
+library(semTools)
+library(tidyverse)
+library(td)
 
-# load packages
-lapply(c(packages, "td"), library, character.only = TRUE)
-```
-
-## Session info
-
-Provide session info to make results reproducible.
-
-```{r}
-sessionInfo()
-```
-
-## Custom functions
-
-### find_var
-
-Find variables in dataset.
-
-```{r}
-find_var <- function(name, data = doc)(
-  # finds the variables names for an item for each wave
-    data %>% 
-    filter(Label == name) %>% 
-    select(variable) %>% 
-    unlist() %>% 
-    set_names(sub("\\_.*", "", .))
-)
-```
-
-### get_specs
-
-Extract characteristics from fitted lmer models.
-
-```{r}
-get_specs <- function(model){
-    # Get mean, max, and min values
-
-    dat <- coefficients(model)$wave
-    specs <- data.frame(
-    sd = attr(VarCorr(model), "sc"),
-    min = min(dat),
-    max = max(dat),
-    mean = mean(dat$`(Intercept)`)
-  )
-}
-```
-
-### get_dat
-
-Get data from fitted lmer objects for descriptives.
-
-```{r}
-get_dat <- function(model){
-  coefficients(model)$wave %>% 
-  tibble::rownames_to_column("wave") %>% 
-  rename(value = "(Intercept)") %>% 
-  mutate(wave = as.integer(.$wave),
-         value = as.numeric(.$value)) 
-} 
-```
-
-### get_rel
-
-Determine average reliability for measures across waves
-
-```{r}
-get_rel <- function(data, waves=24){
-
-  # extract average reliability from lavaan fitted cfa with several groups
-  reliability(data) %>% 
-  unlist() %>% 
-  matrix(5, waves) %>% 
-  as.data.frame() %>% 
-  set_rownames(c("alpha", "omega", "omega2", "omega3", "avevar")) %>% 
-  summarise(omega = rowMeans(.["omega",])) %>% 
-  return()
-}
-```
-
-### make_graph
-
-Make graph of variables' development.
-
-```{r}
-make_graph <- function(model, title, ll, ul, line = FALSE, labels = FALSE, lmer=TRUE, legend=TRUE){
-
-    if(isTRUE(lmer)){
-      dat <- get_dat(model)
-    } else{
-    dat <- model
-  }
-
-  graph <- 
-    ggplot(dat, aes(wave, value, color = dimension)) +
-    {if(legend) theme(
-      legend.position = c(0.85, 0.8),
-      legend.title = element_blank()
-      )} + 
-    {if(!legend) theme(
-      legend.position = "none"
-    )} +
-    theme(axis.title.x=element_blank()) +
-    coord_cartesian(ylim = c(ll, ul)) +
-    ggtitle(title) +
-    {if(line) geom_smooth(se = FALSE, method = 'loess')} +
-    # geom_point() +
-    scale_color_manual(values=c("dodgerblue3", "deepskyblue2", "magenta2", "green2", "red"))
-  graph
-}
-```
-
-### get_dat_res
-
-Get data of lmer objects for results.
-
-```{r}
-# get data
-get_dat_res <- function(data_aff_neg, data_aff_pos, data_life_sat, type, analysis = NULL){
-
-  if(isTRUE(class(data_aff_neg) == "lmerModLmerTest")) {
-         dat_fig_results <- 
-           broom.mixed::tidy(data_aff_neg, conf.int = T) %>% 
-           mutate(dv = "aff_neg") %>% 
-           rbind(
-             broom.mixed::tidy(data_aff_pos, conf.int = T) %>% 
-               mutate(dv = "aff_pos")) %>% 
-           rbind(
-             broom.mixed::tidy(data_life_sat, conf.int = T) %>% 
-               mutate(dv = "life_sat"))
-         } else{ 
-           dat_fig_results <- 
-             data_aff_neg %>% 
-             mutate(dv = "aff_neg") %>% 
-             rbind(data_aff_pos %>% 
-                   mutate(dv = "aff_pos")) %>% 
-             rbind(data_life_sat %>% 
-                     mutate(dv = "life_sat")) %>% 
-           rename(conf.low = `2.5 %`, conf.high = `97.5 %`)
-           }
-  dat_fig_results %<>% 
-  mutate(
-    type = ifelse(grepl(".*_w\\>", .$term), "within", "between"),
-    iv = gsub("_(w|b)\\>", "", .$term)
-         ) %>%
-  mutate(
-    type = factor(.$type, 
-                  levels = c("within", "between")),
-    dv = factor(.$dv, 
-                levels = c("life_sat", "aff_pos", "aff_neg"),
-                labels = c("Life satisfaction", "Positive affect", "Negative affect")),
-    Analysis = analysis
-  ) %>% 
-  select(dv, iv, type, Analysis, estimate, conf.low, conf.high, p.value)
-
-  # select Social Media type of activity
-  if(type == "activity") {
-    dat_fig_results %<>%
-    filter(iv %in% c("soc_med_read", "soc_med_like_share", "soc_med_post")) %>% 
-    mutate(
-      iv = factor(.$iv, 
-                  levels = c("soc_med_post", "soc_med_like_share", "soc_med_read"), 
-                  labels = c("Posting", "Liking & Sharing", "Reading"))
-    )
-  } else if(type == "channels"){
-    dat_fig_results %<>%
-    filter(iv %in% c("soc_med_fb", "soc_med_ig", "soc_med_wa", "soc_med_yt", "soc_med_tw")) %>% 
-    mutate(iv = factor(.$iv, levels = c("soc_med_yt", "soc_med_wa", "soc_med_ig", "soc_med_tw", "soc_med_fb"), labels = c("YouTube", "WhatsApp", "Instagram", "Twitter", "Facebook")))
-  } else if(type == "control"){
-    dat_fig_results %<>%
-      filter(iv %in% c("male", "health", "loc_cntrl_int_m", "employment_facPublic service", "act_spo", "sat_dem")) %>% 
-      mutate(iv = factor(.$iv, levels = c("male", "health", "loc_cntrl_int_m", "employment_facPublic service", "act_spo", "sat_dem"), 
-                         labels = c("Male", "Health", "Internal locus of control", "Employed public service", "Sport", "Satisfaction democracy")))
-  }
-  return(dat_fig_results)
-}
-
-```
-
-### make_graph_res
-
-Make graph of effects
-
-```{r}
-make_graph_res <- function(data, sesoi = NULL, legend = TRUE){
-  ggplot(data, aes(x = estimate, y = iv)) + 
-  scale_color_manual(values = c("black", "grey60", "grey75", "deepskyblue", "cornflowerblue", "darkcyan", "aquamarine")) +
-  geom_vline(xintercept = 0, lwd = .75, colour = "darkgrey") + 
-  geom_errorbarh(aes(xmin = conf.low, xmax = conf.high, color = Analysis),
-                lwd = .75, height = .2, position = position_dodge(-.7)) +
-  geom_point(aes(color = Analysis), size = 1.5, position = position_dodge(-.7)) +
-  {if(isTRUE(sesoi == "est")) geom_vline(data=filter(data, dv=="Life satisfaction"), aes(xintercept=-.3), colour="darkgrey", linetype = "dashed")} + 
-    {if(isTRUE(sesoi == "est")) geom_vline(data=filter(data, dv=="Life satisfaction"), aes(xintercept=.3), colour="darkgrey", linetype = "dashed")} + 
-  {if(isTRUE(sesoi == "est")) geom_vline(data=filter(data, dv!="Life satisfaction"), aes(xintercept=-.15), colour="darkgrey", linetype = "dashed")} +
-    {if(isTRUE(sesoi == "est")) geom_vline(data=filter(data, dv!="Life satisfaction"), aes(xintercept=.15), colour="darkgrey", linetype = "dashed")} +
-  {if(isTRUE(sesoi == "std")) geom_vline(aes(xintercept=.1), colour="darkgrey", linetype = "dashed")} + 
-  {if(isTRUE(sesoi == "std")) geom_vline(aes(xintercept=-.1), colour="darkgrey", linetype = "dashed")} +   
-    theme(
-    axis.title.y = element_blank(),
-    plot.title = element_text(hjust = .5),
-    panel.spacing = unit(.9, "lines"),
-    text = element_text(size = 12),
-    axis.text.y = element_text(hjust = 0)
-    ) +
-  {if(!isTRUE(legend)) theme(legend.position = "none")} + 
-  # guides(colour = guide_legend(reverse=T)) +
-  facet_grid(rows = vars(type),
-             cols = vars(dv),
-             scales = "free",
-             space = "free_y") 
-}
-```
-
-### get_fs
-
-Extract factor scores.
-
-```{r}
-get_fs <- function(model) {
-  # the aim is to get factor scores on original scaling
-  # hence export factor values for all items
-  # then compute average
-  dat_ov <- lavaan::lavPredict(model, type = "ov", assemble = TRUE) %>% 
-    mutate(fs = rowMeans(.[1:3]))
-  return(dat_ov$fs)
-}
-```
-
-# Data Wrangling
-
-```{r}
-doc <- read_csv("documentation/10094_vi_de_v4_0.csv")
-d_raw <- read_csv("data/10094_da_de_v4_0.csv")
-```
-
-Identify the names of each item for each wave.
-
-```{r}
-# identify variable names of items, using custom function
-health <- find_var("Gesundheitszustand") # wsn't collected at wave 21
-aff_neg_1 <- find_var("Depressivitaet: einsam")
-aff_neg_2 <- find_var("Depressivitaet: aergerlich")
-aff_neg_3 <- find_var("Depressivitaet: so niedergeschlagen")
-aff_neg_4 <- find_var("Depressivitaet: sehr nervoes")
-aff_neg_5 <- find_var("Depressivitaet: aengstlich")
-aff_neg_6 <- find_var("Depressivitaet: bedrueckt und traurig")
-aff_pos_1 <- find_var("Depressivitaet: ruhig und gelassen")
-aff_pos_2 <- find_var("Depressivitaet: gluecklich")
-aff_pos_3 <- find_var("Depressivitaet: voller Energie")
-act_wrk <- find_var("Zuhause verlassen: Arbeit")
-act_spo <- find_var("Zuhause verlassen: Sport")
-act_frn <- find_var("Zuhause verlassen: Freunde oder Verwandte treffen")
-act_sho <- find_var("Zuhause verlassen: Essen einkaufen")
-act_pet <- find_var("Zuhause verlassen: Haustier ausfuehren")
-sat_dem <- find_var("Demokratiezufriedenheit: Oesterreich")
-work_h <- find_var("Arbeitsstunden: Jetzt pro Woche")
-work_homeoff <- find_var("Aenderung berufliche Situation: Home-Office")
-hh_income <- find_var("Aktuelles Haushaltseinkommen")
-med_txt_kro <- find_var("Mediennutzung: Kronen Zeitung oder www.krone.at")
-med_txt_sta <- find_var("Mediennutzung: Der Standard oder derstandard.at")
-med_txt_pre <- find_var("Mediennutzung: Die Presse oder diepresse.com")
-med_txt_oes <- find_var("Mediennutzung: Oesterreich oder oe24.at")
-med_txt_kur <- find_var("Mediennutzung: Kurier oder kurier.at")
-med_txt_slz <- find_var("Mediennutzung: Salzburger Nachrichten oder salzburg.at")
-med_txt_son <- find_var("Mediennutzung: Sonstige oesterreichische Tageszeitungen")
-med_vid_orf <- find_var("Mediennutzung: ORF (Nachrichten)")
-med_vid_pri <- find_var("Mediennutzung: Privatfernsehen (Nachrichten)")
-soc_med_fb <- find_var("Soziale Medien: Facebook")
-soc_med_tw <- find_var("Soziale Medien: Twitter")
-soc_med_ig <- find_var("Soziale Medien: Instagram")
-soc_med_yt <- find_var("Soziale Medien: Youtube")
-soc_med_wa <- find_var("Soziale Medien: WhatsApp")
-soc_med_read <- find_var("Soziale Medien Aktivitaet: Postings zu Corona anderer lesen")
-soc_med_like_share <- find_var("Soziale Medien Aktivitaet: Postings liken, teilen oder retweeten")
-soc_med_post <- find_var("Soziale Medien Aktivitaet: selber Postings zu Corona verfassen")
-life_sat <- find_var("Lebenszufriedenheit")
-risk_prop <- find_var("Risikobereitschaft")
-loc_cntrl_int_1 <- find_var("Psychologie: habe Leben selbst in der Hand")
-loc_cntrl_int_2 <- find_var("Psychologie: Belohnung durch Anstrengung")
-loc_cntrl_int_3 <- find_var("Psychologie: Fremdbestimmung")
-loc_cntrl_int_4 <- find_var("Psychologie: Schicksal")
-```
-
-Select variables.
-
-```{r}
-d_wide <- d_raw %>% 
-  select(
-    id = RESPID,
-    gender = SD_GENDER,
-    acc_bal = SD_ACCESS_BALCONY,
-    acc_gar = SD_ACCESS_GARDEN,
-    year_birth = SD_BIRTHYEAR,
-    born_aus = SD_BORN_AUSTRIA,
-    born_aus_prnts = SD_MIGR_BACKGR,
-    county = SD_BULA,
-    edu = SD_EDU,
-    employment = SD_EMPLSTATUS_FEB2020,
-    hh_adults = SD_HH_ADULTS,
-    hh_child18 = SD_HH_CHILD18,
-    hh_child17 = SD_HH_TEENS,
-    hh_child14 = SD_HH_CHILD14,
-    hh_child5 = SD_HH_CHILD5,
-    hh_child2 = SD_HH_CHILD2,
-    hh_oldfam = SD_HH_OLDERFAM,
-    hh_outfam = SD_HH_OUTERFAM,
-    hh_partner = SD_HH_PARTNER,
-    # hh_income = SD_HHINCOME_FEB2020,
-    home_sqm = SD_HOME_SQM,
-    home_owner = SD_HOMEOWNER,
-    # work_h = SD_WORKHOURS_FEB2020,
-    health = all_of(health),
-    life_sat = all_of(life_sat),
-    aff_neg_1 = all_of(aff_neg_1),
-    aff_neg_2 = all_of(aff_neg_2),
-    aff_neg_3 = all_of(aff_neg_3),
-    aff_neg_4 = all_of(aff_neg_4),
-    aff_neg_5 = all_of(aff_neg_5),
-    aff_neg_6 = all_of(aff_neg_6),
-    aff_pos_1 = all_of(aff_pos_1),
-    aff_pos_2 = all_of(aff_pos_2),
-    aff_pos_3 = all_of(aff_pos_3),
-    act_wrk = all_of(act_wrk),
-    act_spo = all_of(act_spo),
-    act_frn = all_of(act_frn),
-    act_sho = all_of(act_sho),
-    act_pet = all_of(act_pet),
-    sat_dem = all_of(sat_dem),
-    sat_dem = all_of(sat_dem),
-    work_h = all_of(work_h),
-    work_homeoff = all_of(work_homeoff),
-    hh_income = all_of(hh_income),
-    med_txt_kro = all_of(med_txt_kro),
-    med_txt_sta = all_of(med_txt_sta),
-    med_txt_pre = all_of(med_txt_pre),
-    med_txt_oes = all_of(med_txt_oes),
-    med_txt_kur = all_of(med_txt_kur),
-    med_txt_slz = all_of(med_txt_slz),
-    med_txt_son = all_of(med_txt_son),
-    med_vid_orf = all_of(med_vid_orf),
-    med_vid_pri = all_of(med_vid_pri),
-    soc_med_fb = all_of(soc_med_fb),
-    soc_med_tw = all_of(soc_med_tw),
-    soc_med_ig = all_of(soc_med_ig),
-    soc_med_yt = all_of(soc_med_yt),
-    soc_med_wa = all_of(soc_med_wa),
-    soc_med_like_share = all_of(soc_med_like_share),
-    soc_med_read = all_of(soc_med_read),
-    soc_med_post = all_of(soc_med_post),
-    risk_prop = all_of(risk_prop),
-    loc_cntrl_int_1 = all_of(loc_cntrl_int_1),
-    loc_cntrl_int_2 = all_of(loc_cntrl_int_2),
-    loc_cntrl_int_3 = all_of(loc_cntrl_int_3),
-    loc_cntrl_int_4 = all_of(loc_cntrl_int_4)
-  )
-```
-
-Make new documentation with selected variables.
-
-```{r}
-doc_selected <- filter(doc, variable %in% c(
-  "RESPID",
-  "SD_GENDER",
-  "SD_ACCESS_BALCONY",
-  "SD_ACCESS_GARDEN",
-  "SD_BIRTHYEAR",
-  "SD_BORN_AUSTRIA",
-  "SD_MIGR_BACKGR",
-  "SD_BULA",
-  "SD_EDU",
-  "SD_EMPLSTATUS_FEB2020",
-  "SD_HH_ADULTS",
-  "SD_HH_CHILD18",
-  "SD_HH_TEENS",
-  "SD_HH_CHILD14",
-  "SD_HH_CHILD5",
-  "SD_HH_CHILD2",
-  "SD_HH_OLDERFAM",
-  "SD_HH_OUTERFAM",
-  "SD_HH_PARTNER",
-  "SD_HHINCOME_FEB2020",
-  "SD_HOME_SQM",
-  "SD_HOMEOWNER",
-  "SD_WORKHOURS_FEB2020",
-  health[1],
-  life_sat[1],
-  aff_neg_1[1],
-  aff_neg_2[1],
-  aff_neg_3[1],
-  aff_neg_4[1],
-  aff_neg_5[1],
-  aff_neg_6[1],
-  aff_pos_1[1],
-  aff_pos_2[1],
-  aff_pos_3[1],
-  act_wrk[1],
-  act_spo[1],
-  act_frn[1],
-  act_sho[1],
-  act_pet[1],
-  sat_dem[1],
-  work_h[1],
-  work_homeoff[1],
-  hh_income[1],
-  med_txt_kro[1],
-  med_txt_sta[1],
-  med_txt_pre[1],
-  med_txt_oes[1],
-  med_txt_kur[1],
-  med_txt_slz[1],
-  med_txt_son[1],
-  med_vid_orf[1],
-  med_vid_pri[1],
-  soc_med_fb[1],
-  soc_med_tw[1],
-  soc_med_ig[1],
-  soc_med_yt[1],
-  soc_med_wa[1],
-  soc_med_like_share[1],
-  soc_med_post[1],
-  soc_med_read[1],
-  risk_prop[1],
-  loc_cntrl_int_1[1],
-  loc_cntrl_int_2[1],
-  loc_cntrl_int_3[1],
-  loc_cntrl_int_4[1]
-  ))
-```
-
-Recode variables.
-
-```{r}
-d_wide %<>%
-  mutate_at(vars(everything(.)), funs(na_if(., 88))) %>%
-  mutate_at(vars(everything(.)), funs(na_if(., 99))) %>%
-  mutate(
-    male = 2 - .$gender,
-    age = 2021 - .$year_birth,
-    res_vienna = recode(.$county, `8` = 1L, .default = 0L,),
-    born_aus = 2 - .$born_aus,
-    home_owner = 2 - .$home_owner,
-    employment_fac = factor(.$employment, 
-                           labels = c("Unemployed",
-                                      "Industrie",
-                                      "Public service",
-                                      "Self-employed",
-                                      "Retired",
-                                      "Housekeeping",
-                                      "Student",
-                                      "Incapacitated",
-                                      "Parental Leave"),
-                            levels = c(4, 1:3, 5:8, 10) # make unemployment reference cat
-                           ),
-    edu_fac = factor(.$edu,
-                     labels = c("No degree", 
-                                "Middle school",
-                                "Vocational school",
-                                "Technical school",
-                                "High school",
-                                "Applied high school",
-                                "State college",
-                                "Bachelor",
-                                "Master",
-                                "PhD")
-                     ))
-```
-
-Make data from wide to long.
-
-```{r}
-d_long <-
-  d_wide %>% 
-  pivot_longer(
-    cols = health...W1:loc_cntrl_int_4...W9,
-    names_to = "item",
-    values_to = "value"
-  ) %>% 
-  separate(item, c("item", "wave"), sep = "\\.\\.\\.", extra = "merge") %>% 
-  pivot_wider(names_from = "item", values_from = "value")
-```
-
-Recode values.
-
-```{r}
-# recode such that higher values imply more strength / align with wording
-d_long %<>% 
-  mutate_at(vars(med_txt_kro:med_vid_pri, health, sat_dem, soc_med_fb:soc_med_post), 
-            funs(recode(., `1` = 5L, `2` = 4L, `3` = 3L, `4` = 2L, `5` = 1L))) %>% 
-  mutate_at(vars(loc_cntrl_int_1:loc_cntrl_int_4), 
-            funs(recode(., `1` = 4L, `2` = 3L, `3` = 2L, `4` = 1L))) %>% 
-  mutate_at(vars(born_aus_prnts), 
-            funs(recode(., `3` = 0L, `2` = 2L, `1` = 1L)))
-
-# recode inverted items
-d_long %<>%
-  mutate_at(vars(loc_cntrl_int_3, loc_cntrl_int_4), 
-            funs(recode(., `1` = 4L, `2` = 3L, `3` = 2L, `4` = 1L)))
-
-# recode other
-d_long %<>%
-  mutate(
-    wave = gsub("W", "", .$wave) %>% as.integer(),
-    id = as.integer(id)
-  )
-```
-
-Order dataset. First id, then wave.
-
-```{r}
-d_long %<>%
-  arrange(id, wave)
-```
-
-# Impute data
-
-Determine amount of missingness per respondent per wave. Because calculation takes long, this chunk is loaded from memory.
-
-```{r eval=TRUE, cache=TRUE}
-vars_used <- c("life_sat", "aff_pos_1", "aff_pos_2", "aff_pos_3", "aff_neg_1", "aff_neg_2", "aff_neg_3", "aff_neg_4", "aff_neg_5", "aff_neg_6", "soc_med_read", "soc_med_like_share", "soc_med_post", "soc_med_fb", "soc_med_ig", "soc_med_wa", "soc_med_yt", "soc_med_tw", "age", "male", "born_aus", "born_aus_prnts", "edu_fac", "employment_fac", "health", "res_vienna", "acc_bal", "acc_gar", "home_sqm", "med_txt_kro", "med_txt_sta", "med_txt_pre", "med_txt_oes", "med_txt_kur", "med_txt_slz", "med_txt_son", "med_vid_orf", "med_vid_pri", "risk_prop", "loc_cntrl_int_1", "loc_cntrl_int_2", "loc_cntrl_int_3", "loc_cntrl_int_4", "act_wrk", "act_spo", "act_frn", "act_sho", "act_pet", "sat_dem")
-
-# filter respondents with more than 50% missing data
-d_long_filt <- 
-  d_long %>% 
-  mutate(na_perc = rowSums(is.na(select(., all_of(vars_used))) / ncol(select(., all_of(vars_used))))) %>% 
-  filter(na_perc < .5)
-
-# impute missing data
-# exclude social media use data, because they were measured only on selected waves and would lead to the imputation of too much missing data
-vars_excl <- c("na_perc", "soc_med_read", "soc_med_like_share", "soc_med_post", "soc_med_fb", "soc_med_ig", "soc_med_wa", "soc_med_yt", "soc_med_tw")
-incl_ma <- d_long_filt %>% 
-  mutate(across(.cols = everything(), .fns = is.na))
-incl_ma[vars_excl] <- FALSE
-
-d_long_imp <- mice(d_long_filt, 
-                   method = "pmm", # use predictive mean matching
-                   m = 1, maxit = 30, # only 1 imputation
-                   where = incl_ma,
-                   seed = 180719, print = FALSE)
-d_long_imp <- mice::complete(d_long_imp)
-write_csv(d_long_imp, "data/data_imputed.csv")
-```
-
-```{r eval=TRUE, cache=TRUE}
-# impute missing data with multiple imputation
-d_long_mim_mice <- mice(d_long_filt, 
-                   method = "pmm", # use predictive mean matching
-                   m = 5, maxit = 5, # 5 imputations, 5 iterations
-                   where = incl_ma,
-                   seed = 180719, print = FALSE)
-d_long_mim <- mice::complete(d_long_mim_mice, action = "long", include = TRUE)
-```
-
-```{r eval=FALSE}
-# to further increase speed, you can also load the imputed data from the directory
-d_long_imp <- read_csv("data/data_imputed.csv")
-```
-
-# Create within-between data
-
-Create means for scales
-
-```{r}
-d_long_imp %<>%
-  mutate(aff_pos_m = rowMeans(select(., aff_pos_1 : aff_pos_3)),
-         aff_neg_m = rowMeans(select(., aff_neg_1 : aff_neg_6)),
-         loc_cntrl_int_m = rowMeans(select(., loc_cntrl_int_1 : loc_cntrl_int_4)))
-```
-
-Create mean values for each person (between), and unique changes per wave (within).
-
-```{r}
-d_long_imp %<>% 
-  group_by(id) %>% 
-  mutate_at(vars(health:loc_cntrl_int_4, loc_cntrl_int_m), funs(b = mean(., na.rm = TRUE), w = . - mean(., na.rm = TRUE))) %>% 
-  ungroup()
-```
-
-And because later we want to rerun the results without imputed data, now the same for data without imputation.
-
-```{r}
-d_long %<>%
-  mutate(aff_pos_m = rowMeans(select(., aff_pos_1 : aff_pos_3)),
-         aff_neg_m = rowMeans(select(., aff_neg_1 : aff_neg_6)),
-         loc_cntrl_int_m = rowMeans(select(., loc_cntrl_int_1 : loc_cntrl_int_4)))
-
-d_long %<>% 
-  group_by(id) %>% 
-  mutate_at(vars(health:loc_cntrl_int_4, loc_cntrl_int_m), funs(b = mean(., na.rm = TRUE), w = . - mean(., na.rm = TRUE))) %>% 
-  ungroup()
-```
-
-And now the same for the multiple imputed data
-
-```{r}
-d_long_mim %<>%
-  mutate(aff_pos_m = rowMeans(select(., aff_pos_1 : aff_pos_3)),
-         aff_neg_m = rowMeans(select(., aff_neg_1 : aff_neg_6)),
-         loc_cntrl_int_m = rowMeans(select(., loc_cntrl_int_1 : loc_cntrl_int_4)))
-
-d_long_mim %<>% 
-  group_by(id) %>% 
-  mutate_at(vars(health:loc_cntrl_int_4, loc_cntrl_int_m), funs(b = mean(., na.rm = TRUE), w = . - mean(., na.rm = TRUE))) %>% 
-  ungroup()
-
-d_long_mim_mice <- as.mids(d_long_mim)
+# load workspace
+load("data/workspace_1.RData")
 ```
@@
 # Measures
+
+Let's first inspect the individual measures, how they develop over time. For positive and negative affect, we also look at their factor structure, as they're measured with multiple items.
+
 ## Well-being
@@
 
+Let's inspect the development of life satisfaction across the study. We nest reponses inside participants and waves to get better results.
+
 ```{r}
-model_life_sat <- lmer(life_sat ~ (1 | id) + (1 | wave), d_long_imp)
-summary(model_life_sat)
+fit_life_sat <- lmer(life_sat ~ (1 | id) + (1 | wave), d_long_100_imp)
+summary(fit_life_sat)
+dat_fig_life_sat <- data.frame(type = "Life satisfaction", dimension = "Life satisfaction", 
+                               get_dat(fit_life_sat))
+make_graph(dat_fig_life_sat, "Life Satisfaction", 1, 10)
 ```
@@
 
-Positive affect was measured as a scale. We hence first inspect factorial validity using CFA.
+Let's next expect the development across waves.
 
-We first test assumption of multivariate normality.
+```{r}
+model_aff_pos <- lmer(aff_pos_m ~ (1 | id) + (1 | wave), d_long_100_imp)
+summary(model_aff_pos)
+dat_fig_aff_pos <- data.frame(type = "Affect", dimension = "Positive", 
+                              get_dat(model_aff_pos))
+make_graph(dat_fig_aff_pos, "Positive Affect", 1, 5)
+```
 
-```{r eval=FALSE}
-mardia(d_long_imp %>% select(aff_pos_1, aff_pos_2, aff_pos_3))
+Positive affect was measured as a scale. We hence also inspect factorial validity using CFA.
+
+We first test assumption of multivariate normality. We focus on wave 1 here.
+
+```{r}
+d_long_100_imp %>%
+  filter(wave == 1) %>%
+  select(aff_pos_1, aff_pos_2, aff_pos_3) %>% 
+  mardia()
 ```
@@
 "
-cfa_aff_pos <- cfa(model, d_long_imp, group = "wave", estimator = "MLM")
+cfa_aff_pos <- cfa(model, d_long_100_imp, group = "wave", estimator = "MLM")
 summary(cfa_aff_pos, standardized = TRUE, fit = TRUE, estimates = FALSE)
@@
 rel_aff_pos <- get_rel(cfa_aff_pos)
-rel_aff_pos
 ```
 
-The average reliability across all waves was omega = `r rel_aff_pos`, hence good.
+The average reliability across all waves was omega = `r rel_aff_pos[[1]]`, hence good.
 
-Let's next expect the development across waves.
+Let's now export factor scores for results reported in additional analyses.
 
 ```{r}
-model_aff_pos <- lmer(aff_pos_m ~ (1 | id) + (1 | wave), d_long_imp)
-summary(model_aff_pos)
-```
+# with imputed data
+cfa_aff_pos_50 <- cfa(model, d_long_50_imp, group = "wave", estimator = "MLM")
+d_long_50_imp$aff_pos_fs <- get_fs(cfa_aff_pos_50)
 
-Let's now export factor scores.
-
-```{r}
-d_long_imp$aff_pos_fs <- get_fs(cfa_aff_pos)
+# without imputed data
+cfa_aff_pos <- cfa(model, d_long_50, group = "wave", estimator = "MLM")
+d_long_50$aff_pos_fs <- get_fs(cfa_aff_pos)
 ```
@@
 
-Negative affect was measured as a scale. We hence inspect factorial validty using CFA.
-
-We first test assumption of multivariate normality. (Warning, as these analyses take long and might require additional working memory, they're deactived currently.)
-
-```{r eval=FALSE}
-mardia(d_long_imp %>% select(aff_neg_1, aff_neg_2, aff_neg_3, aff_neg_4, aff_neg_5, aff_neg_6))
+```{r}
+model_aff_neg <- lmer(aff_neg_m ~ (1 | id) + (1 | wave), d_long_100_imp)
+summary(model_aff_neg)
+dat_fig_aff_neg <- data.frame(type = "Affect", dimension = "Negative", get_dat(model_aff_neg))
+make_graph(dat_fig_aff_neg, "Negative Affect", 1, 5)
 ```
 
-Assumption of multivariate normal distributino was violated; hence, robust estimator will be used.
+Negative affect was measured as a scale. We hence inspect factorial validity using CFA.
+
+We first test assumption of multivariate normality. We'll focus on wave 1 here.
+
+```{r eval=T}
+d_long_100_imp %>%
+  filter(wave == 1) %>% 
+  select(aff_neg_1, aff_neg_2, aff_neg_3, aff_neg_4, aff_neg_5, aff_neg_6) %>% 
+  mardia()
+```
+
+Assumption of multivariate normal distribution was violated; hence, robust estimator will be used.
 
@@
 "
-cfa_aff_neg <- cfa(model, d_long_imp, group = "wave", estimator = "MLM")
+cfa_aff_neg <- cfa(model, d_long_100_imp, group = "wave", estimator = "MLM")
 summary(cfa_aff_neg, standardized = TRUE, fit = TRUE, estimates = FALSE)
@@
 rel_aff_neg <- get_rel(cfa_aff_neg)
-rel_aff_neg
 ```
 
-The average reliability across all waves was omega = `r rel_aff_neg`, hence good.
+The average reliability across all waves was omega = `r rel_aff_neg[[1]]`, hence good.
 
 ```{r}
-model_aff_neg <- lmer(aff_neg_m ~ (1 | id) + (1 | wave), d_long_imp)
+model_aff_neg <- lmer(aff_neg_m ~ (1 | id) + (1 | wave), d_long_100_imp)
 summary(model_aff_neg)
@@
 
-Let's now export factor scores.
+Let's now export factor scores, necessary for results reported in additional analyses.
 
 ```{r}
-d_long_imp$aff_neg_fs <- get_fs(cfa_aff_neg)
+# with imputed data
+cfa_aff_neg_50 <- cfa(model, d_long_50_imp, group = "wave", estimator = "MLM")
+d_long_50_imp$aff_neg_fs <- get_fs(cfa_aff_neg_50)
+
+# without imputed data
+cfa_aff_neg <- cfa(model, d_long_50, group = "wave", estimator = "MLM")
+d_long_50$aff_neg_fs <- get_fs(cfa_aff_neg)
 ```
@@
 ```{r}
-model_soc_med_read <- lmer(soc_med_read ~ (1 | id) + (1 | wave), d_long_imp)
+model_soc_med_read <- lmer(soc_med_read ~ (1 | id) + (1 | wave), d_long_100_imp)
 summary(model_soc_med_read)
+dat_fig_soc_med_read <- data.frame(type = "Social media use", dimension = "Reading", 
+                                   get_dat(model_soc_med_read))
+make_graph(dat_fig_soc_med_read, "Social Media Reading", 1, 5)
 ```
@@
 ```{r}
-model_soc_med_like_share <- lmer(soc_med_like_share ~ (1 | id) + (1 | wave), d_long_imp)
+model_soc_med_like_share <- lmer(soc_med_like_share ~ (1 | id) + (1 | wave), d_long_100_imp)
 summary(model_soc_med_like_share)
+dat_fig_soc_med_like_share <- data.frame(type = "Social media use", dimension = "Liking & Sharing",
+                                         get_dat(model_soc_med_like_share))
+make_graph(dat_fig_soc_med_like_share, "Social Media Liking and Sharing", 1, 5)
 ```
@@
 ```{r}
-model_soc_med_post <- lmer(soc_med_post ~ (1 | id) + (1 | wave), d_long_imp)
+model_soc_med_post <- lmer(soc_med_post ~ (1 | id) + (1 | wave), d_long_100_imp)
 summary(model_soc_med_post)
+dat_fig_soc_med_post <- data.frame(type = "Social media use", dimension = "Posting", 
+                                   get_dat(model_soc_med_post))
+make_graph(dat_fig_soc_med_post, "Social Media Posting", 1, 5)
 ```
@@
 ```{r}
-model_soc_med_fb <- lmer(soc_med_fb ~ (1 | id) + (1 | wave), d_long_imp)
+model_soc_med_fb <- lmer(soc_med_fb ~ (1 | id) + (1 | wave), d_long_100_imp)
 summary(model_soc_med_fb)
+dat_fig_soc_med_fb <- data.frame(type = "Social media channel", dimension = "Facebook", 
+                                 get_dat(model_soc_med_fb))
+make_graph(dat_fig_soc_med_fb, "Social Media Facebook", 1, 5)
 ```
@@
 ```{r}
-model_soc_med_tw <- lmer(soc_med_tw ~ (1 | id) + (1 | wave), d_long_imp)
+model_soc_med_tw <- lmer(soc_med_tw ~ (1 | id) + (1 | wave), d_long_100_imp)
 summary(model_soc_med_tw)
+dat_fig_soc_med_tw <- data.frame(type = "Social media channel", dimension = "Twitter", 
+                                 get_dat(model_soc_med_tw))
+make_graph(dat_fig_soc_med_tw, "Social Media Twitter", 1, 5)
 ```
@@
 ```{r}
-soc_med_tw_m <- d_long_imp %>% group_by(wave) %>% summarise(value = mean(soc_med_tw, na.rm = T))
+soc_med_tw_m <- 
+  d_long_100_imp %>% 
+  group_by(wave) %>% 
+  summarise(value = mean(soc_med_tw, na.rm = T))
 soc_med_tw_m
@@
 ```{r}
-model_soc_med_ig <- lmer(soc_med_ig ~ (1 | id) + (1 | wave), d_long_imp)
+model_soc_med_ig <- lmer(soc_med_ig ~ (1 | id) + (1 | wave), d_long_100_imp)
 summary(model_soc_med_ig)
+dat_fig_soc_med_ig <- data.frame(type = "Social media channel", dimension = "Instagram", 
+                                 get_dat(model_soc_med_ig))
+make_graph(dat_fig_soc_med_ig, "Social Media Instagram", 1, 5)
 ```
@@
 ```{r}
-model_soc_med_wa <- lmer(soc_med_wa ~ (1 | id) + (1 | wave), d_long_imp)
+model_soc_med_wa <- lmer(soc_med_wa ~ (1 | id) + (1 | wave), d_long_100_imp)
 summary(model_soc_med_wa)
+dat_fig_soc_med_wa <- data.frame(type = "Social media channel", dimension = "WhatsApp", 
+                                 get_dat(model_soc_med_wa))
+make_graph(dat_fig_soc_med_wa, "Social Media WhatsApp", 1, 5)
 ```
@@
 ```{r}
-model_soc_med_yt <- lmer(soc_med_yt ~ (1 | id) + (1 | wave), d_long_imp)
+model_soc_med_yt <- lmer(soc_med_yt ~ (1 | id) + (1 | wave), d_long_100_imp)
 summary(model_soc_med_yt)
+dat_fig_soc_med_yt <- data.frame(type = "Social media channel", dimension = "YouTube", 
+                                 get_dat(model_soc_med_yt))
+make_graph(dat_fig_soc_med_yt, "Social Media YouTube", 1, 5)
 ```
@@
 
-The only other variable that was measured as a scale was Locus of Control. Below I hence report the scale's factorial validity.
+The only other variable that was measured as a scale was Locus of Control.
+Below I hence report the scale's factorial validity.
+Waves for which not a sufficient number of respondents took part were excluded.
 
-```{r}
+```{r eval=T}
 model <- "
@@
 "
-cfa_loc_cntrl_int <- cfa(model, filter(d_long_imp, wave != 20), group = "wave")
+cfa_loc_cntrl_int <- cfa(model, 
+                         filter(d_long_100_imp, wave != 11, wave != 20,  wave != 26, wave != 27, wave != 29, wave != 31, wave != 32),
+                         # d_long_100_imp,
+                         group = "wave")
 summary(cfa_loc_cntrl_int, standardized = TRUE, fit = TRUE, estimates = FALSE)
@@
 
-## Figure
+## Table
 
-Display the results in a combined figure.
-
-```{r}
-fig_desc_dat <- data.frame(
-  rbind(
-    data.frame(type = "Life satisfaction", dimension = "Life satisfaction", get_dat(model_life_sat)),
-    data.frame(type = "Affect", dimension = "Positive", get_dat(model_aff_pos)),
-    data.frame(type = "Affect", dimension = "Negative", get_dat(model_aff_neg)),
-    data.frame(type = "Social media use", dimension = "Reading", get_dat(model_soc_med_read)),
-    data.frame(type = "Social media use", dimension = "Liking & Sharing", get_dat(model_soc_med_like_share)),
-    data.frame(type = "Social media use", dimension = "Posting", get_dat(model_soc_med_post)),
-    data.frame(type = "Social media channel", dimension = "Facebook", get_dat(model_soc_med_fb)),
-    data.frame(type = "Social media channel", dimension = "Twitter", get_dat(model_soc_med_tw)),
-    data.frame(type = "Social media channel", dimension = "Instagram", get_dat(model_soc_med_ig)),
-    data.frame(type = "Social media channel", dimension = "WhatsApp", get_dat(model_soc_med_wa)),
-    data.frame(type = "Social media channel", dimension = "YouTube", get_dat(model_soc_med_yt))
-  ) %>% 
-    mutate(
-      type = factor(.$type, levels = c("Life satisfaction", "Affect", "Social media use", "Social media channel")),
-      dimension = factor(.$dimension, levels = c("Life satisfaction", "Positive", "Negative", "Reading", "Liking & Sharing", "Posting", "Facebook", "Twitter", "Instagram", "WhatsApp", "YouTube"))
-      )
-)
-
-fig_desc_life_sat <- make_graph(
-  fig_desc_dat %>% filter(type == "Life satisfaction"), 
-  title = "Life satisfaction", 
-  ll = 0, ul = 10, 
-  lmer = FALSE, 
-  line = TRUE,
-  legend = FALSE
-  )
-
-fig_desc_aff <- make_graph(
-  fig_desc_dat %>% filter(type == "Affect"), 
-  title = "Affect", 
-  ll = 1, ul = 5, 
-  lmer = FALSE, 
-  line = TRUE
-  )
-
-fig_desc_soc_med_use <- make_graph(
-  fig_desc_dat %>% filter(type == "Social media use"), 
-  title = "Social media use", 
-  ll = 1, ul = 5, 
-  lmer = FALSE, 
-  line = TRUE
-  )
-
-fig_desc_soc_med_channel <- make_graph(
-  fig_desc_dat %>% filter(type == "Social media channel"), 
-  title = "Social media channel", 
-  ll = 1, ul = 5, 
-  lmer = FALSE, 
-  line = TRUE
-  )
-
-fig_desc <- grid.arrange(fig_desc_life_sat, fig_desc_aff, 
-                         fig_desc_soc_med_use, fig_desc_soc_med_channel,
-                         nrow = 2, ncol = 2)
-
-ggsave("figures/fig_descriptives.png", 
-       width = 10, height = 5,
-       plot = fig_desc)
-```
-
-## Table
+Table with descriptives of main variables.
 
@@
 tab_desc_dat <- rbind(
-    "Life satisfaction" = get_specs(model_life_sat),
+    "Life satisfaction" = get_specs(fit_life_sat),
     "Positive affect" = get_specs(model_aff_pos),
@@
     "YouTube" = get_specs(model_soc_med_yt)
+    ) %>%
+  as.data.frame()
+tab_desc_dat
+```
+
+## Figure
+
+Display the developments of all variables in a combined figure.
+
+```{r}
+fig_desc_dat <- data.frame(
+  rbind(
+    dat_fig_life_sat,
+    dat_fig_aff_pos,
+    dat_fig_aff_neg,
+    dat_fig_soc_med_read,
+    dat_fig_soc_med_like_share,
+    dat_fig_soc_med_post,
+    dat_fig_soc_med_fb,
+    dat_fig_soc_med_tw,
+    dat_fig_soc_med_ig,
+    dat_fig_soc_med_wa,
+    dat_fig_soc_med_yt
+  ) %>% 
+    mutate(
+      type = factor(.$type, levels = c("Life satisfaction", "Affect", "Social media use", "Social media channel")),
+      dimension = factor(.$dimension, levels = c("Life satisfaction", "Positive", "Negative", "Reading", "Liking & Sharing", "Posting", "Facebook", "Twitter", "Instagram", "WhatsApp", "YouTube"))
       )
-kable(tab_desc_dat)
+)
+
+fig_desc_life_sat <- make_graph(
+  fig_desc_dat %>% filter(type == "Life satisfaction"), 
+  title = "Life satisfaction", 
+  ll = 0, ul = 10, 
+  lmer = FALSE, 
+  line = TRUE,
+  legend = FALSE,
+  points = FALSE
+  )
+
+fig_desc_aff <- make_graph(
+  fig_desc_dat %>% filter(type == "Affect"), 
+  title = "Affect", 
+  ll = 1, ul = 5, 
+  lmer = FALSE, 
+  line = TRUE,
+  points = FALSE,
+  legend = TRUE
+  )
+
+fig_desc_soc_med_use <- make_graph(
+  fig_desc_dat %>% filter(type == "Social media use"), 
+  title = "Social media use", 
+  ll = 1, ul = 5, 
+  lmer = FALSE, 
+  line = TRUE,
+  points = FALSE,
+  legend = TRUE
+  )
+
+fig_desc_soc_med_channel <- make_graph(
+  fig_desc_dat %>% filter(type == "Social media channel"), 
+  title = "Social media channel", 
+  ll = 1, ul = 5, 
+  lmer = FALSE, 
+  line = TRUE,
+  points = FALSE,
+  legend = TRUE
+  )
+
+fig_desc <- grid.arrange(fig_desc_life_sat, fig_desc_aff, 
+                         fig_desc_soc_med_use, fig_desc_soc_med_channel,
+                         nrow = 2, ncol = 2)
+
+ggsave("figures/fig_descriptives.png", 
+       width = 10, height = 5,
+       plot = fig_desc)
 ```
@@
 # Analyses
-## Preregistered
+
+## Multicollinearity
+
+Before running the analyses, let's briefly check zero-order correlation matrix, to get general picture and also for potential multicollinearity. We use variables from T1.
+
+Multicollinearity will then also be checked explicitly in each analysis.
+
+```{r eval=T}
+d_long_100_imp %>% 
+  filter(wave == 1) %>% 
+  select(life_sat, aff_neg_m, aff_pos_m,
+         soc_med_read, soc_med_post, soc_med_like_share, 
+         soc_med_fb, soc_med_ig, soc_med_tw, soc_med_wa, soc_med_yt,
+         health, corona_pos, work_h, work_homeoff, hh_income, med_txt_kro, med_txt_sta, med_txt_pre, med_txt_oes, med_txt_kur, med_txt_slz, med_txt_son, med_vid_orf, med_vid_pri, med_txt_kro, med_txt_sta, med_txt_pre, med_txt_oes, med_txt_kur, med_txt_slz, med_txt_son, med_vid_orf, med_vid_pri, act_wrk, act_spo, act_frn, act_sho, act_pet, risk_prop, loc_cntrl_int_m, sat_dem) %>%
+  cor(use = "pairwise.complete.obs") %>% 
+  as.data.frame()
+```
+
+## Correlations
+
+Let's also briefly look at bivariate relations between the types and channels of social media use and the well-being facets.
+
+### Table
+
+```{r}
+dat_cor <- 
+  d_long_100_mim %>% 
+  filter(wave == 1) %>%  # we can use wave 1 only, because mean values are the same across waves
+  select(`Life satis-\nfaction` = life_sat_b, 
+         `Affect\npositive` = aff_pos_m_b, 
+         `Affect\nnegative` = aff_neg_m_b,
+         `Reading` = soc_med_read_b, 
+         `Posting` = soc_med_post_b, 
+         `Like &\nshare` = soc_med_like_share_b, 
+         `Facebook` = soc_med_fb_b, 
+         `Instagram` = soc_med_ig_b, 
+         `Twitter` = soc_med_tw_b, 
+         `WhatsApp` = soc_med_wa_b, 
+         `YouTube` = soc_med_yt_b)
+  # cor() %>% 
+  # as.data.frame()
+  # correlate() %>% 
+  # fashion()
+
+tab_cor <- 
+  dat_cor %>% 
+  correlate() %>% 
+  fashion() %T>%
+  print()
+```
+
+### Figure
+
+```{r}
+int_breaks <- function(x, n = 4) {
+  l <- pretty(x, n)
+  l[abs(l %% 1) < .Machine$double.eps ^ 0.5] 
+}
+
+fig_cor <- 
+  dat_cor %>% 
+  ggpairs(
+    upper = list(continuous = cor_plot),
+    lower = list(continuous = wrap("points", alpha = 0.3, size=0.1), 
+              combo = wrap("dot", alpha = 0.3, size=0.1)),
+    progress = FALSE
+    ) + 
+  scale_x_continuous(breaks = int_breaks) +
+  theme_bw()
+
+fig_cor
+ggsave("figures/fig_cor.png", width = 8, height = 8)
+```
+
+## Publication
 ### Life satisfaction
@@
 ```{r}
-model_life_sat_lmer <- lmerTest::lmer(life_sat ~
+model_life_sat_lmer_pub <- "
+  life_sat ~ 
     (1 | id) + (1 | wave) + 
@@
     soc_med_fb_b + soc_med_ig_b + soc_med_wa_b + soc_med_yt_b + soc_med_tw_b +  
-                                        age + male + born_aus + born_aus_prnts + edu_fac + employment_fac + health_b + 
+    age + male + born_aus + born_aus_prnts + edu_fac + employment_fac +
     res_vienna + acc_bal + acc_gar + home_sqm + 
-                                        # hh_income_b + # large dropout (900 obs)
-                                        # hh_adults + # too much dropout
-                                        # hh_child18 + hh_child17 + hh_child14 + hh_child5 + hh_child2 + hh_oldfam + hh_outfam + hh_partner + 
-                                        # home_owner + # too much dropout
-                                        med_txt_kro_b + med_txt_sta_b + med_txt_pre_b + med_txt_oes_b + med_txt_kur_b + med_txt_slz_b + med_txt_son_b + 
-                                        med_vid_orf_b + med_vid_pri_b + 
-                                        risk_prop_b + loc_cntrl_int_m_b + 
+    corona_pos_b + corona_pos_w +
+    work_h_b + work_h_w +
+    work_homeoff_b +  work_homeoff_w +
+    hh_income_b + hh_income_w +
+    hh_adults + hh_child18 + hh_child17 + hh_child14 + hh_child5 + hh_child2 + 
+    hh_oldfam + hh_outfam + hh_partner +
+    home_owner +
+    risk_prop_b + risk_prop_w + 
     act_wrk_w + act_spo_w + act_frn_w + act_sho_w + act_pet_w + 
     act_wrk_b + act_spo_b + act_frn_b + act_sho_b + act_pet_b +
-                                        sat_dem_w + sat_dem_b 
-                                      , data = d_long_imp)
-summary(model_life_sat_lmer)
+    health_w + health_b +
+    loc_cntrl_int_m_w + loc_cntrl_int_m_b
+"
+```
+
+Let's first inspect multicollinearity.
+
+```{r}
+check_collinearity(lmerTest::lmer(model_life_sat_lmer_pub, d_long_100_imp))
+```
+
+No within-person predictor shows increased values for multicollinearity. Hence results straightforward.
+
+Let's next inspect results of within-person predictors. Only parts of model shown to save space.
+
+```{r}
+fit_life_sat_lmer_pub <- with(d_long_100_mim_mice, exp = lmerTest::lmer(model_life_sat_lmer_pub))
+fit_life_sat_lmer_pub <- summary(pool(fit_life_sat_lmer_pub), conf.int = TRUE)
+print_res(fit_life_sat_lmer_pub)
 ```
@@
 ```{r}
-model_aff_pos_lmer <- lmerTest::lmer(aff_pos_fs ~ 
+model_aff_pos_lmer_pub <- "
+  aff_pos_m ~ 
     (1 | id) + (1 | wave) + 
@@
     soc_med_fb_b + soc_med_ig_b + soc_med_wa_b + soc_med_yt_b + soc_med_tw_b +  
-                                        age + male + born_aus + born_aus_prnts + edu_fac + employment_fac + health_b + 
+    age + male + born_aus + born_aus_prnts + edu_fac + employment_fac +
     res_vienna + acc_bal + acc_gar + home_sqm + 
-                                        med_txt_kro_b + med_txt_sta_b + med_txt_pre_b + med_txt_oes_b + med_txt_kur_b + med_txt_slz_b + med_txt_son_b + 
-                                        med_vid_orf_b + med_vid_pri_b + 
-                                        risk_prop_b + loc_cntrl_int_m_b + 
+    corona_pos_b + corona_pos_w +
+    work_h_b + work_h_w +
+    work_homeoff_b +  work_homeoff_w +
+    hh_income_b + hh_income_w +
+    hh_adults + hh_child18 + hh_child17 + hh_child14 + hh_child5 + hh_child2 + 
+    hh_oldfam + hh_outfam + hh_partner +
+    home_owner +
+    risk_prop_b + risk_prop_w + 
     act_wrk_w + act_spo_w + act_frn_w + act_sho_w + act_pet_w + 
     act_wrk_b + act_spo_b + act_frn_b + act_sho_b + act_pet_b +
-                                        sat_dem_w + sat_dem_b 
-                                      , data = d_long_imp)
-summary(model_aff_pos_lmer)
+    health_w + health_b +
+    loc_cntrl_int_m_w + loc_cntrl_int_m_b
+"
+```
+
+Let's first inspect multicollinearity.
+
+```{r}
+check_collinearity(lmerTest::lmer(model_aff_pos_lmer_pub, d_long_100_imp))
+```
+
+No within-person predictors show multicollinear relations.
+
+In what follows, the results of within-person predictors.
+
+```{r}
+fit_aff_pos_lmer_pub <- with(d_long_100_mim_mice, exp = lmerTest::lmer(model_aff_pos_lmer_pub))
+fit_aff_pos_lmer_pub <- summary(pool(fit_aff_pos_lmer_pub), conf.int = TRUE)
+print_res(fit_aff_pos_lmer_pub)
 ```
@@
 ```{r}
-model_aff_neg_lmer <- lmerTest::lmer(aff_neg_fs ~ 
+model_aff_neg_lmer_pub <- "
+  aff_neg_m ~ 
     (1 | id) + (1 | wave) + 
@@
     soc_med_fb_b + soc_med_ig_b + soc_med_wa_b + soc_med_yt_b + soc_med_tw_b +  
-                                        age + male + born_aus + born_aus_prnts + edu_fac + employment_fac + health_b + 
+    age + male + born_aus + born_aus_prnts + edu_fac + employment_fac +
     res_vienna + acc_bal + acc_gar + home_sqm + 
-                                        med_txt_kro_b + med_txt_sta_b + med_txt_pre_b + med_txt_oes_b + med_txt_kur_b + med_txt_slz_b + med_txt_son_b + 
-                                        med_vid_orf_b + med_vid_pri_b + 
-                                        risk_prop_b + loc_cntrl_int_m_b + 
+    corona_pos_b + corona_pos_w +
+    work_h_b + work_h_w +
+    work_homeoff_b +  work_homeoff_w +
+    hh_income_b + hh_income_w +
+    hh_adults + hh_child18 + hh_child17 + hh_child14 + hh_child5 + hh_child2 + 
+    hh_oldfam + hh_outfam + hh_partner +
+    home_owner +
+    risk_prop_b + risk_prop_w + 
     act_wrk_w + act_spo_w + act_frn_w + act_sho_w + act_pet_w + 
     act_wrk_b + act_spo_b + act_frn_b + act_sho_b + act_pet_b +
-                                        sat_dem_w + sat_dem_b 
-                                      , data = d_long_imp)
-summary(model_aff_neg_lmer)
+    health_w + health_b +
+    loc_cntrl_int_m_w + loc_cntrl_int_m_b
+"
 ```
 
-## Without control variables
-### Life Satisfaction
+Let's inspect multicollinearity.
 
 ```{r}
-model_life_sat_lmer_nco <- lmerTest::lmer(life_sat ~ 
-                            (1 | id) + (1 | wave) + 
-                            soc_med_read_w + soc_med_like_share_w + soc_med_post_w + 
-                            soc_med_fb_w + soc_med_ig_w + soc_med_wa_w + soc_med_yt_w + soc_med_tw_w +
-                            soc_med_read_b + soc_med_like_share_b + soc_med_post_b + 
-                            soc_med_fb_b + soc_med_ig_b + soc_med_wa_b + soc_med_yt_b + soc_med_tw_b,
-                data = d_long_imp)
-summary(model_life_sat_lmer_nco)
+check_collinearity(lmerTest::lmer(model_aff_neg_lmer_pub, d_long_100_imp))
+```
+
+No within-person predictors show multicollinear relations.
+
+Here are the results for the within-person predictors.
+
+```{r}
+fit_aff_neg_lmer_pub <- with(d_long_100_mim_mice, exp = lmerTest::lmer(model_aff_neg_lmer_pub))
+fit_aff_neg_lmer_pub <- summary(pool(fit_aff_neg_lmer_pub), conf.int = TRUE)
+print_res(fit_aff_neg_lmer_pub)
+```
+
+## Standardized results
+
+Then let's report also the standardized results. Helps compare effect sizes across differently scaled predictors.
+
+### Life satisfaction
+
+```{r}
+fit_life_sat_lmer_std <- with(d_long_100_mim_mice_std, exp = lmerTest::lmer(model_life_sat_lmer_pub))
+fit_life_sat_lmer_std <- summary(pool(fit_life_sat_lmer_std), conf.int = TRUE)
+print_res(fit_life_sat_lmer_std)
 ```
@@
 ```{r}
-model_aff_pos_lmer_nco <- lmerTest::lmer(aff_pos_fs ~ 
-                                        (1 | id) + (1 | wave) + 
-                                        soc_med_read_w + soc_med_like_share_w + soc_med_post_w + 
-                                        soc_med_fb_w + soc_med_ig_w + soc_med_wa_w + soc_med_yt_w + soc_med_tw_w +
-                                        soc_med_read_b + soc_med_like_share_b + soc_med_post_b + 
-                                        soc_med_fb_b + soc_med_ig_b + soc_med_wa_b + soc_med_yt_b + soc_med_tw_b
-                                      , data = d_long_imp)
-summary(model_aff_pos_lmer_nco)
+fit_aff_pos_lmer_std <- with(d_long_100_mim_mice_std, exp = lmerTest::lmer(model_aff_pos_lmer_pub))
+fit_aff_pos_lmer_std <- summary(pool(fit_aff_pos_lmer_std), conf.int = TRUE)
+print_res(fit_aff_pos_lmer_std)
 ```
@@
 ```{r}
-model_aff_neg_lmer_nco <- lmerTest::lmer(aff_neg_fs ~ 
-                                        (1 | id) + (1 | wave) + 
-                                        soc_med_read_w + soc_med_like_share_w + soc_med_post_w + 
-                                        soc_med_fb_w + soc_med_ig_w + soc_med_wa_w + soc_med_yt_w + soc_med_tw_w +
-                                        soc_med_read_b + soc_med_like_share_b + soc_med_post_b + 
-                                        soc_med_fb_b + soc_med_ig_b + soc_med_wa_b + soc_med_yt_b + soc_med_tw_b
-                                      , data = d_long_imp)
-summary(model_aff_neg_lmer_nco)
+fit_aff_neg_lmer_std <- with(d_long_100_mim_mice_std, exp = lmerTest::lmer(model_aff_neg_lmer_pub))
+fit_aff_neg_lmer_std <- summary(pool(fit_aff_neg_lmer_std), conf.int = TRUE)
+print_res(fit_aff_neg_lmer_std)
 ```
 
-## Standardized results
+## Figures
 
-First, create standardized dataset.
+Let's visualize results. First, results of unstandardized predictors.
+
+```{r warning=F, cache=F}
+# get dat
+data_tab_within <- rbind(
+  get_dat_res(fit_aff_neg_lmer_pub, fit_aff_pos_lmer_pub, fit_life_sat_lmer_pub, 
+              type = "Activity", variance = "within", analysis = "Publication"),
+  get_dat_res(fit_aff_neg_lmer_pub, fit_aff_pos_lmer_pub, fit_life_sat_lmer_pub, 
+              type = "Channels", variance = "within", analysis = "Publication")
+  )
+
+# make fig
+fig_results_within <- make_graph_res(
+  data = data_tab_within,
+  sesoi = "est",
+  legend = FALSE
+  , facet = "type"
+  # , title = "Results of selected covariates"
+  )
+fig_results_within
+
+# save figure
+ggsave("figures/fig_results_within.png", 
+       width = 7, height = 4,
+       plot = fig_results_within)
+```
+
+Let's next visualize standardized predictors, to allow for better comparison across differently scaled variables.
+
+```{r warning=F, cache=F, fig.height=7}
+# make figure
+data_tab_comp_std <- rbind(
+  get_dat_res(fit_aff_neg_lmer_std, fit_aff_pos_lmer_std, fit_life_sat_lmer_std, 
+              type = "Activity", variance = "within", analysis = "standardized"),
+  get_dat_res(fit_aff_neg_lmer_std, fit_aff_pos_lmer_std, fit_life_sat_lmer_std, 
+              type = "Channels", variance = "within", analysis = "standardized"),
+  get_dat_res(fit_aff_neg_lmer_std, fit_aff_pos_lmer_std, fit_life_sat_lmer_std, 
+              type = "News\nuse", variance = "within", analysis = "standardized"),
+  get_dat_res(fit_aff_neg_lmer_std, fit_aff_pos_lmer_std, fit_life_sat_lmer_std, 
+              type = "Living\nconditions", variance = "within", analysis = "standardized"),
+  get_dat_res(fit_aff_neg_lmer_std, fit_aff_pos_lmer_std, fit_life_sat_lmer_std, 
+              type = "Outdoor\nactivities", variance = "within", analysis = "standardized"),
+  get_dat_res(fit_aff_neg_lmer_std, fit_aff_pos_lmer_std, fit_life_sat_lmer_std, 
+              type = "Psycho-\nlogy", variance = "within", analysis = "standardized")
+  )
+
+fig_results_comp_std <- make_graph_res(
+  data = data_tab_comp_std,
+  sesoi = "std",
+  legend = FALSE
+  , facet = "type"
+  # , title = "Results of selected covariates"
+  )
+fig_results_comp_std
+
+# save figure
+ggsave("figures/fig_results_comp_std.png", 
+       width = 7, height = 7,
+       plot = fig_results_comp_std)
+```
+
+## Table
+
+Let's extract results for a table of within-person effects.
+
+```{r cache=F, eval=T}
+tab_within <-
+  data_tab_comp_std %>% 
+  filter(Type %in% c("Activity", "Channels")) %>% 
+  select(std = estimate) %>% 
+  cbind(data_tab_within) %>% 
+  mutate(p.value = td::my_round(p.value, "p")) %>% 
+  arrange(dv) %>% 
+  select(Outcome = dv,
+         Predictor = iv,
+         b = estimate,
+         `Lower` = conf.low,
+         `Higher` = conf.high,
+         beta = "std",
+         p = "p.value")
+
+tab_within
+```
+
+# Save workspace
+
+First remove large objects
 
 ```{r}
-d_long_std <- d_long_imp %>% 
-  mutate(across(c(-id, -wave, -gender, -male, -born_aus, -born_aus_prnts, -edu_fac, -employment_fac, -res_vienna, -acc_bal, -acc_gar), scale))
+rm(fig_cor)
 ```
 
-Then let's rerun the analyses.
-
-### Life satisfaction
-
-```{r}
-model_life_sat_lmer_std <- lmerTest::lmer(life_sat ~
-                                        (1 | id) + (1 | wave) + 
-                                        soc_med_read_w + soc_med_like_share_w + soc_med_post_w  + 
-                                        soc_med_fb_w + soc_med_ig_w + soc_med_wa_w + soc_med_yt_w + soc_med_tw_w +
-                                        soc_med_read_b + soc_med_like_share_b + soc_med_post_b + 
-                                        soc_med_fb_b + soc_med_ig_b + soc_med_wa_b + soc_med_yt_b + soc_med_tw_b +  
-                                        age + male + born_aus + born_aus_prnts + edu_fac + employment_fac + health_b + 
-                                        res_vienna + acc_bal + acc_gar + home_sqm + 
-                                        med_txt_kro_b + med_txt_sta_b + med_txt_pre_b + med_txt_oes_b + med_txt_kur_b + med_txt_slz_b + med_txt_son_b + 
-                                        med_vid_orf_b + med_vid_pri_b + 
-                                        risk_prop_b + loc_cntrl_int_m_b + 
-                                        act_wrk_w + act_spo_w + act_frn_w + act_sho_w + act_pet_w + 
-                                        act_wrk_b + act_spo_b + act_frn_b + act_sho_b + act_pet_b +
-                                        sat_dem_w + sat_dem_b 
-                                      , data = d_long_std)
-summary(model_life_sat_lmer_std)
+```{r cache=F}
+save.image("data/workspace_2.RData")
 ```
-
-### Positive affect
-
-```{r}
-model_aff_pos_lmer_std <- lmerTest::lmer(aff_pos_fs ~ 
-                                        (1 | id) + (1 | wave) + 
-                                        soc_med_read_w + soc_med_like_share_w + soc_med_post_w + 
-                                        soc_med_fb_w + soc_med_ig_w + soc_med_wa_w + soc_med_yt_w + soc_med_tw_w +
-                                        soc_med_read_b + soc_med_like_share_b + soc_med_post_b + 
-                                        soc_med_fb_b + soc_med_ig_b + soc_med_wa_b + soc_med_yt_b + soc_med_tw_b +  
-                                        age + male + born_aus + born_aus_prnts + edu_fac + employment_fac + health_b + 
-                                        res_vienna + acc_bal + acc_gar + home_sqm + 
-                                        med_txt_kro_b + med_txt_sta_b + med_txt_pre_b + med_txt_oes_b + med_txt_kur_b + med_txt_slz_b + med_txt_son_b + 
-                                        med_vid_orf_b + med_vid_pri_b + 
-                                        risk_prop_b + loc_cntrl_int_m_b + 
-                                        act_wrk_w + act_spo_w + act_frn_w + act_sho_w + act_pet_w + 
-                                        act_wrk_b + act_spo_b + act_frn_b + act_sho_b + act_pet_b +
-                                        sat_dem_w + sat_dem_b 
-                                      , data = d_long_imp)
-summary(model_aff_pos_lmer_std)
-```
-
-### Negative Affect
-
-```{r}
-model_aff_neg_lmer_std <- lmerTest::lmer(aff_neg_fs ~ 
-                                        (1 | id) + (1 | wave) + 
-                                        soc_med_read_w + soc_med_like_share_w + soc_med_post_w + 
-                                        soc_med_fb_w + soc_med_ig_w + soc_med_wa_w + soc_med_yt_w + soc_med_tw_w +
-                                        soc_med_read_b + soc_med_like_share_b + soc_med_post_b + 
-                                        soc_med_fb_b + soc_med_ig_b + soc_med_wa_b + soc_med_yt_b + soc_med_tw_b +  
-                                        age + male + born_aus + born_aus_prnts + edu_fac + employment_fac + health_b + 
-                                        res_vienna + acc_bal + acc_gar + home_sqm + 
-                                        med_txt_kro_b + med_txt_sta_b + med_txt_pre_b + med_txt_oes_b + med_txt_kur_b + med_txt_slz_b + med_txt_son_b + 
-                                        med_vid_orf_b + med_vid_pri_b + 
-                                        risk_prop_b + loc_cntrl_int_m_b + 
-                                        act_wrk_w + act_spo_w + act_frn_w + act_sho_w + act_pet_w + 
-                                        act_wrk_b + act_spo_b + act_frn_b + act_sho_b + act_pet_b +
-                                        sat_dem_w + sat_dem_b 
-                                      , data = d_long_imp)
-summary(model_aff_neg_lmer_std)
-```
-
-# Figures
-
-## Activities
-
-```{r warning=FALSE}
-# get data
-dat_fig_results_activity <- get_dat_res(model_aff_neg_lmer, model_aff_pos_lmer, model_life_sat_lmer, 
-                                        type = "activity", analysis = "regular")
-dat_fig_results_activity_nco <- get_dat_res(model_aff_neg_lmer_nco, model_aff_pos_lmer_nco, model_life_sat_lmer_nco, 
-                                            type = "activity", analysis = "no controls")
-dat_fig_results_activity_std <- get_dat_res(model_aff_neg_lmer_std, model_aff_pos_lmer_std, model_life_sat_lmer_std, 
-                                            type = "activity", analysis = "standardized")
-
-dat_fig_results_activity <- rbind(
-  dat_fig_results_activity, 
-  dat_fig_results_activity_nco,
-  dat_fig_results_activity_std
-) %>% 
-  mutate(Analysis = factor(Analysis, levels = c("regular", "no controls", "standardized")))
-
-# make figure
-fig_results_activity <- make_graph_res(
-  data = dat_fig_results_activity,
-  sesoi = "est"
-  )
-fig_results_activity
-
-# safe figure
-ggsave("figures/fig_results_activity.png", 
-       width = 7, height = 4,
-       plot = fig_results_activity)
-```
-
-## Channels
-
-```{r warning=FALSE}
-# get data
-dat_fig_results_channels <- get_dat_res(model_aff_neg_lmer, model_aff_pos_lmer, model_life_sat_lmer, 
-                                        type = "channels", analysis = "regular")
-dat_fig_results_channels_nco <- get_dat_res(model_aff_neg_lmer_nco, model_aff_pos_lmer_nco, model_life_sat_lmer_nco, 
-                                            type = "channels", analysis = "no controls")
-dat_fig_results_channels_std <- get_dat_res(model_aff_neg_lmer_std, model_aff_pos_lmer_std, model_life_sat_lmer_std, 
-                                            type = "channels", analysis = "standardized")
-
-dat_fig_results_channels <- rbind(
-  dat_fig_results_channels, 
-  dat_fig_results_channels_nco,
-  dat_fig_results_channels_std
-) %>% 
-  mutate(Analysis = factor(Analysis, levels = c("regular", "no controls", "standardized")))
-
-# make figure
-fig_results_channels <- make_graph_res(
-  data = dat_fig_results_channels,
-  sesoi = "est"
-  )
-fig_results_channels
-
-# safe figure
-ggsave("figures/fig_results_channels.png", 
-       width = 7, height = 4,
-       plot = fig_results_channels)
-```
-
-## Selected control variables
-
-Standardized, to allow for better comparison across differently scaled variables.
-
-```{r warning=FALSE}
-# make figure
-fig_results_control_std <- make_graph_res(
-  get_dat_res(model_aff_neg_lmer_std, model_aff_pos_lmer_std, model_life_sat_lmer_std, 
-              type = "control", analysis = "standardized"),
-  sesoi = "std",
-  legend = FALSE
-  )
-fig_results_control_std
-
-# save figure
-ggsave("figures/fig_results_control_std.png", 
-       width = 7, height = 4,
-       plot = fig_results_control_std)
-```
-
-```{r}
-save.image("data/workspace_1.RData")
-```

### 2023-07-11 analyses_additional.Rmd
Major updates Implemented feedback

--- 2022-04-11 analyses_additional.Rmd
@@
  - Here you can find the code and the results of all additional analyses. 
- - To see the results, I recommend you directly go down to the Figures section. 
- - But if you're interested in how the analyses were executed, just tag along.
+ - To see a good overview of the results, I recommend you directly go down to the Figures section. 
+ - But if you're interested in how the analyses were executed, see below.
  - To see the code, click on button "Code". Alternatively, you can download the rmd file from the github repo.
- - To execute the code, first download the actual data from AUSSDA, then run analysis.rmd, then this file
 
@@
 
-```{r r-setup, results='hide'}
+```{r r-setup, results='hide', cache=F}
 # install packages
@@
 # define packages
-packages <- c("broom.mixed", "brms", "devtools", "GGally", "ggplot2", 
-              "gridExtra", "kableExtra", "knitr", "lavaan", "lme4", 
-              "magrittr", "mice", "mvnormalTest", 
-              "PerFit", "psych", "quanteda.textstats", "semTools", "tidyverse")
-
-# load packages
-lapply(c(packages, "td"), library, character.only = TRUE)
+library(broom.mixed)
+library(brms)
+library(devtools)
+library(GGally)
+library(ggplot2)
+library(gridExtra)
+library(kableExtra)
+library(knitr)
+library(lavaan)
+library(lme4)
+library(magrittr)
+library(mice)
+library(mvnormalTest)
+library(PerFit)
+library(psych)
+library(quanteda.textstats)
+library(semTools)
+library(tidyverse)
+library(td)
 
 # load workspace
-load("data/workspace_1.RData")
+load("data/workspace_2.RData")
 ```
 
-# Results with mean scores
+# Robustness analyses
+## With additional covariates
 
-Instead of using factor scores, in what follows I report the results with mean scores of affect. Note that the results for life satisfaction were the same, as it was measured with a single item.
+These models include additional covariates that weren't preregistered and that weren't included, mainly because they are likely to be mediators. These are:
 
-## Positive affect
+- Trust in ORF
+- Trust in police
+- Trust in parliament
+- Trust in health sector
+- Trust in government
+- Trust in army
+
+ - Media use: Kronen Zeitung oder www.krone.at
+ - Media use: Der Standard oder derstandard.at
+ - Media use: Die Presse oder diepresse.com
+ - Media use: Oesterreich oder oe24.at
+ - Media use: Kurier oder kurier.at
+ - Media use: Salzburger Nachrichten oder salzburg.at
+ - Media use: Sonstige oesterreichische Tageszeitungen
+ - Media use: ORF (Nachrichten)
+ - Media use: Privatfernsehen (Nachrichten)
+
+### Life satisfaction
 
 ```{r}
-model_aff_pos_lmer_m <- lmerTest::lmer(aff_pos_m ~ 
+model_life_sat_lmer_add <- "
+  life_sat ~ 
     (1 | id) + (1 | wave) + 
@@
     soc_med_fb_b + soc_med_ig_b + soc_med_wa_b + soc_med_yt_b + soc_med_tw_b +  
-                                        age + male + born_aus + born_aus_prnts + edu_fac + employment_fac + health_b + 
+    age + male + born_aus + born_aus_prnts + edu_fac + employment_fac +
     res_vienna + acc_bal + acc_gar + home_sqm + 
+    corona_pos_b + corona_pos_w +
+    work_h_b + work_h_w +
+    work_homeoff_b +  work_homeoff_w +
+    hh_income_b + hh_income_w +
+    hh_adults + hh_child18 + hh_child17 + hh_child14 + hh_child5 + hh_child2 + 
+    hh_oldfam + hh_outfam + hh_partner +
+    home_owner +
     med_txt_kro_b + med_txt_sta_b + med_txt_pre_b + med_txt_oes_b + med_txt_kur_b + med_txt_slz_b + med_txt_son_b + 
     med_vid_orf_b + med_vid_pri_b + 
-                                        risk_prop_b + loc_cntrl_int_m_b + 
+    med_txt_kro_w + med_txt_sta_w + med_txt_pre_w + med_txt_oes_w + med_txt_kur_w + med_txt_slz_w + med_txt_son_w + 
+    med_vid_orf_w + med_vid_pri_w + 
+    risk_prop_b + risk_prop_w + 
     act_wrk_w + act_spo_w + act_frn_w + act_sho_w + act_pet_w + 
     act_wrk_b + act_spo_b + act_frn_b + act_sho_b + act_pet_b +
-                                        sat_dem_w + sat_dem_b 
-                                      , data = d_long_imp)
-summary(model_aff_pos_lmer_m)
+    sat_dem_w + sat_dem_b + 
+    health_w + health_b +
+    loc_cntrl_int_m_w + loc_cntrl_int_m_b +
+    trst_media_w + trst_police_w + trst_media_w + trst_hlthsec_w + trst_gov_w + trst_army_w + 
+    trst_media_b + trst_police_b + trst_media_b + trst_hlthsec_b + trst_gov_b + trst_army_b
+"
+
+fit_life_sat_lmer_add <- with(d_long_100_mim_mice, 
+                              exp = lmerTest::lmer(model_life_sat_lmer_add))
+fit_life_sat_lmer_add <- summary(pool(fit_life_sat_lmer_add), conf.int = TRUE)
+print_res(fit_life_sat_lmer_add)
 ```
 
-The results differed only slightly, and all inferences remained the same.
-
-## Negative Affect
+### Positive Affect
 
 ```{r}
-model_aff_neg_lmer_m <- lmerTest::lmer(aff_neg_m ~ 
+model_aff_pos_lmer_add <- "
+  aff_pos_m ~ 
     (1 | id) + (1 | wave) + 
@@
     soc_med_fb_b + soc_med_ig_b + soc_med_wa_b + soc_med_yt_b + soc_med_tw_b +  
-                                        age + male + born_aus + born_aus_prnts + edu_fac + employment_fac + health_b + 
+    age + male + born_aus + born_aus_prnts + edu_fac + employment_fac +
     res_vienna + acc_bal + acc_gar + home_sqm + 
+    corona_pos_b + corona_pos_w +
+    work_h_b + work_h_w +
+    work_homeoff_b +  work_homeoff_w +
+    hh_income_b + hh_income_w +
+    hh_adults + hh_child18 + hh_child17 + hh_child14 + hh_child5 + hh_child2 + 
+    hh_oldfam + hh_outfam + hh_partner +
+    home_owner +
     med_txt_kro_b + med_txt_sta_b + med_txt_pre_b + med_txt_oes_b + med_txt_kur_b + med_txt_slz_b + med_txt_son_b + 
     med_vid_orf_b + med_vid_pri_b + 
-                                        risk_prop_b + loc_cntrl_int_m_b + 
+    med_txt_kro_w + med_txt_sta_w + med_txt_pre_w + med_txt_oes_w + med_txt_kur_w + med_txt_slz_w + med_txt_son_w + 
+    med_vid_orf_w + med_vid_pri_w + 
+    risk_prop_b + risk_prop_w + 
     act_wrk_w + act_spo_w + act_frn_w + act_sho_w + act_pet_w + 
     act_wrk_b + act_spo_b + act_frn_b + act_sho_b + act_pet_b +
-                                        sat_dem_w + sat_dem_b 
-                                      , data = d_long_imp)
-summary(model_aff_neg_lmer_m)
+    sat_dem_w + sat_dem_b + 
+    health_w + health_b +
+    loc_cntrl_int_m_w + loc_cntrl_int_m_b +
+    trst_media_w + trst_police_w + trst_media_w + trst_hlthsec_w + trst_gov_w + trst_army_w + 
+    trst_media_b + trst_police_b + trst_media_b + trst_hlthsec_b + trst_gov_b + trst_army_b
+"
+
+fit_aff_pos_lmer_add <- with(d_long_100_mim_mice, 
+                              exp = lmerTest::lmer(model_aff_pos_lmer_add))
+fit_aff_pos_lmer_add <- summary(pool(fit_aff_pos_lmer_add), conf.int = TRUE)
+print_res(fit_aff_pos_lmer_add)
 ```
 
-The results differed only slightly, and all inferences remained the same.
-
-# Results without imputed data
-
-We first need to export factor scores for variables without imputed data
+### Negative Affect
 
 ```{r}
-model <- "
-aff_pos =~ a1*aff_pos_1 + a2*aff_pos_2 + a3*aff_pos_3
-"
-cfa_aff_pos <- cfa(model, d_long, group = "wave", estimator = "MLM")
-d_long$aff_pos_fs <- get_fs(cfa_aff_pos)
-
-model <- "
-aff_neg =~ a1*aff_neg_1 + a2*aff_neg_2 + a3*aff_neg_3 + a4*aff_neg_4 + a5*aff_neg_5 + a6*aff_neg_6
-"
-cfa_aff_neg <- cfa(model, d_long, group = "wave", estimator = "MLM")
-d_long$aff_neg_fs <- get_fs(cfa_aff_neg)
-```
-
-## Life satisfaction
-
-```{r}
-model_life_sat_lmer_noi <- lmerTest::lmer(life_sat ~
+model_aff_neg_lmer_add <- "
+  aff_neg_m ~ 
     (1 | id) + (1 | wave) + 
@@
     soc_med_fb_b + soc_med_ig_b + soc_med_wa_b + soc_med_yt_b + soc_med_tw_b +  
-                                        age + male + born_aus + born_aus_prnts + edu_fac + employment_fac + health_b + 
+    age + male + born_aus + born_aus_prnts + edu_fac + employment_fac +
     res_vienna + acc_bal + acc_gar + home_sqm + 
+    corona_pos_b + corona_pos_w +
+    work_h_b + work_h_w +
+    work_homeoff_b +  work_homeoff_w +
+    hh_income_b + hh_income_w +
+    hh_adults + hh_child18 + hh_child17 + hh_child14 + hh_child5 + hh_child2 + 
+    hh_oldfam + hh_outfam + hh_partner +
+    home_owner +
     med_txt_kro_b + med_txt_sta_b + med_txt_pre_b + med_txt_oes_b + med_txt_kur_b + med_txt_slz_b + med_txt_son_b + 
     med_vid_orf_b + med_vid_pri_b + 
-                                        risk_prop_b + loc_cntrl_int_m_b + 
+    med_txt_kro_w + med_txt_sta_w + med_txt_pre_w + med_txt_oes_w + med_txt_kur_w + med_txt_slz_w + med_txt_son_w + 
+    med_vid_orf_w + med_vid_pri_w + 
+    risk_prop_b + risk_prop_w + 
     act_wrk_w + act_spo_w + act_frn_w + act_sho_w + act_pet_w + 
     act_wrk_b + act_spo_b + act_frn_b + act_sho_b + act_pet_b +
-                                        sat_dem_w + sat_dem_b 
-                                      , data = d_long)
-summary(model_life_sat_lmer_noi)
+    sat_dem_w + sat_dem_b + 
+    health_w + health_b +
+    loc_cntrl_int_m_w + loc_cntrl_int_m_b +
+    trst_media_w + trst_police_w + trst_media_w + trst_hlthsec_w + trst_gov_w + trst_army_w + 
+    trst_media_b + trst_police_b + trst_media_b + trst_hlthsec_b + trst_gov_b + trst_army_b
+"
+
+fit_aff_neg_lmer_add <- with(d_long_100_mim_mice, 
+                              exp = lmerTest::lmer(model_aff_neg_lmer_add))
+fit_aff_neg_lmer_add <- summary(pool(fit_aff_neg_lmer_add), conf.int = TRUE)
+print_res(fit_aff_neg_lmer_add)
 ```
 
-## Positive Affect
+## Without control variables
+
+Let's inspect how results change when control variables are omitted.
+
+### Life satisfaction
 
 ```{r}
-model_aff_pos_lmer_noi <- lmerTest::lmer(aff_pos_fs ~ 
+model_life_sat_lmer_nco <- "
+  life_sat ~ 
+    (1 | id) + (1 | wave) + 
+    soc_med_read_w + soc_med_like_share_w + soc_med_post_w  + 
+    soc_med_fb_w + soc_med_ig_w + soc_med_wa_w + soc_med_yt_w + soc_med_tw_w +
+    soc_med_read_b + soc_med_like_share_b + soc_med_post_b + 
+    soc_med_fb_b + soc_med_ig_b + soc_med_wa_b + soc_med_yt_b + soc_med_tw_b
+"
+fit_life_sat_lmer_nco <- with(d_long_100_mim_mice, exp = lmerTest::lmer(model_life_sat_lmer_nco))
+fit_life_sat_lmer_nco <- summary(pool(fit_life_sat_lmer_nco), conf.int = TRUE)
+print_res(fit_life_sat_lmer_nco)
+```
+
+### Positive Affect
+
+```{r}
+model_aff_pos_lmer_nco <- "
+  aff_pos_m ~ 
+    (1 | id) + (1 | wave) + 
+    soc_med_read_w + soc_med_like_share_w + soc_med_post_w  + 
+    soc_med_fb_w + soc_med_ig_w + soc_med_wa_w + soc_med_yt_w + soc_med_tw_w +
+    soc_med_read_b + soc_med_like_share_b + soc_med_post_b + 
+    soc_med_fb_b + soc_med_ig_b + soc_med_wa_b + soc_med_yt_b + soc_med_tw_b
+"
+fit_aff_pos_lmer_nco <- with(d_long_100_mim_mice, exp = lmerTest::lmer(model_aff_pos_lmer_nco))
+fit_aff_pos_lmer_nco <- summary(pool(fit_aff_pos_lmer_nco), conf.int = TRUE)
+print_res(fit_aff_pos_lmer_nco)
+```
+
+### Negative Affect
+
+```{r}
+model_aff_neg_lmer_nco <- "
+  aff_neg_m ~ 
+    (1 | id) + (1 | wave) + 
+    soc_med_read_w + soc_med_like_share_w + soc_med_post_w  + 
+    soc_med_fb_w + soc_med_ig_w + soc_med_wa_w + soc_med_yt_w + soc_med_tw_w +
+    soc_med_read_b + soc_med_like_share_b + soc_med_post_b + 
+    soc_med_fb_b + soc_med_ig_b + soc_med_wa_b + soc_med_yt_b + soc_med_tw_b
+"
+fit_aff_neg_lmer_nco <- with(d_long_100_mim_mice, exp = lmerTest::lmer(model_aff_neg_lmer_nco))
+fit_aff_neg_lmer_nco <- summary(pool(fit_aff_neg_lmer_nco), conf.int = TRUE)
+print_res(fit_aff_neg_lmer_nco)
+```
+
+## Preregistered
+
+Some changes were introduced because of feedback from colleagues (multiple imputation, imputation of all participants, inclusion of all variables).
+In what follows, please find the results as originally planned (but with some necessary deviations; for example not all preregistered variables could be included).
+
+ - Use data-set in which participants with >50% data were removed. 
+ - Also use factor scores (possible without multiple imputation).
+
+Note that because here no multiple imputation was used, output looks differently.
+
+### Life satisfaction
+
+```{r}
+model_life_sat_lmer_pre <- "
+  life_sat ~ 
     (1 | id) + (1 | wave) + 
@@
     soc_med_fb_b + soc_med_ig_b + soc_med_wa_b + soc_med_yt_b + soc_med_tw_b +  
-                                        age + male + born_aus + born_aus_prnts + edu_fac + employment_fac + health_b + 
+    age + male + born_aus + born_aus_prnts + edu_fac + employment_fac +
     res_vienna + acc_bal + acc_gar + home_sqm + 
+    work_h_b + work_h_w +
+    work_homeoff_b +  work_homeoff_w +
+    hh_income_b + hh_income_w +
+    hh_adults + hh_child18 + hh_child17 + hh_child14 + hh_child5 + hh_child2 + 
+    hh_oldfam + hh_outfam + hh_partner +
+    home_owner +
     med_txt_kro_b + med_txt_sta_b + med_txt_pre_b + med_txt_oes_b + med_txt_kur_b + med_txt_slz_b + med_txt_son_b + 
     med_vid_orf_b + med_vid_pri_b + 
-                                        risk_prop_b + loc_cntrl_int_m_b + 
+    med_txt_kro_w + med_txt_sta_w + med_txt_pre_w + med_txt_oes_w + med_txt_kur_w + med_txt_slz_w + med_txt_son_w + 
+    med_vid_orf_w + med_vid_pri_w + 
+    risk_prop_b + risk_prop_w + 
     act_wrk_w + act_spo_w + act_frn_w + act_sho_w + act_pet_w + 
     act_wrk_b + act_spo_b + act_frn_b + act_sho_b + act_pet_b +
-                                        sat_dem_w + sat_dem_b 
-                                      , data = d_long)
-summary(model_aff_pos_lmer_noi)
+    sat_dem_w + sat_dem_b + 
+    health_w + health_b +
+    loc_cntrl_int_m_w + loc_cntrl_int_m_b
+    "
+fit_life_sat_lmer_pre <- lmerTest::lmer(model_life_sat_lmer_pre, data = d_long_50_imp)
+print_res(broom.mixed::tidy(fit_life_sat_lmer_pre, conf.int = T), imputation = FALSE)
 ```
 
-## Negative Affect
+### Positive Affect
 
 ```{r}
-model_aff_neg_lmer_noi <- lmerTest::lmer(aff_neg_fs ~ 
+model_aff_pos_lmer_pre <- "
+  aff_pos_fs ~ 
     (1 | id) + (1 | wave) + 
@@
     soc_med_fb_b + soc_med_ig_b + soc_med_wa_b + soc_med_yt_b + soc_med_tw_b +  
-                                        age + male + born_aus + born_aus_prnts + edu_fac + employment_fac + health_b + 
+    age + male + born_aus + born_aus_prnts + edu_fac + employment_fac +
     res_vienna + acc_bal + acc_gar + home_sqm + 
+    work_h_b + work_h_w +
+    work_homeoff_b +  work_homeoff_w +
+    hh_income_b + hh_income_w +
+    hh_adults + hh_child18 + hh_child17 + hh_child14 + hh_child5 + hh_child2 + 
+    hh_oldfam + hh_outfam + hh_partner +
+    home_owner +
     med_txt_kro_b + med_txt_sta_b + med_txt_pre_b + med_txt_oes_b + med_txt_kur_b + med_txt_slz_b + med_txt_son_b + 
     med_vid_orf_b + med_vid_pri_b + 
-                                        risk_prop_b + loc_cntrl_int_m_b + 
+    med_txt_kro_w + med_txt_sta_w + med_txt_pre_w + med_txt_oes_w + med_txt_kur_w + med_txt_slz_w + med_txt_son_w + 
+    med_vid_orf_w + med_vid_pri_w + 
+    risk_prop_b + risk_prop_w + 
     act_wrk_w + act_spo_w + act_frn_w + act_sho_w + act_pet_w + 
     act_wrk_b + act_spo_b + act_frn_b + act_sho_b + act_pet_b +
-                                        sat_dem_w + sat_dem_b 
-                                      , data = d_long)
-summary(model_aff_neg_lmer_noi)
+    sat_dem_w + sat_dem_b + 
+    health_w + health_b +
+    loc_cntrl_int_m_w + loc_cntrl_int_m_b
+    "
+fit_aff_pos_lmer_pre <- lmerTest::lmer(model_aff_pos_lmer_pre, data = d_long_50_imp)
+print_res(broom.mixed::tidy(fit_aff_pos_lmer_pre, conf.int = T), imputation = FALSE)
 ```
 
-# Results with multiple imputation
-## Life satisfaction
+### Negative Affect
 
 ```{r}
-model_life_sat_lmer_mim <- with(d_long_mim_mice, exp =  
-                                lmerTest::lmer(life_sat ~
+model_aff_neg_lmer_pre <- "
+  aff_neg_fs ~ 
     (1 | id) + (1 | wave) + 
@@
     soc_med_fb_b + soc_med_ig_b + soc_med_wa_b + soc_med_yt_b + soc_med_tw_b +  
-                                        age + male + born_aus + born_aus_prnts + edu_fac + employment_fac + health_b + 
+    age + male + born_aus + born_aus_prnts + edu_fac + employment_fac +
     res_vienna + acc_bal + acc_gar + home_sqm + 
+    work_h_b + work_h_w +
+    work_homeoff_b +  work_homeoff_w +
+    hh_income_b + hh_income_w +
+    hh_adults + hh_child18 + hh_child17 + hh_child14 + hh_child5 + hh_child2 + 
+    hh_oldfam + hh_outfam + hh_partner +
+    home_owner +
     med_txt_kro_b + med_txt_sta_b + med_txt_pre_b + med_txt_oes_b + med_txt_kur_b + med_txt_slz_b + med_txt_son_b + 
     med_vid_orf_b + med_vid_pri_b + 
-                                        risk_prop_b + loc_cntrl_int_m_b + 
+    med_txt_kro_w + med_txt_sta_w + med_txt_pre_w + med_txt_oes_w + med_txt_kur_w + med_txt_slz_w + med_txt_son_w + 
+    med_vid_orf_w + med_vid_pri_w + 
+    risk_prop_b + risk_prop_w + 
     act_wrk_w + act_spo_w + act_frn_w + act_sho_w + act_pet_w + 
     act_wrk_b + act_spo_b + act_frn_b + act_sho_b + act_pet_b +
-                                        sat_dem_w + sat_dem_b 
-                                      ))
-model_life_sat_lmer_mim <- summary(pool(model_life_sat_lmer_mim), conf.int = TRUE)
+    sat_dem_w + sat_dem_b + 
+    health_w + health_b +
+    loc_cntrl_int_m_w + loc_cntrl_int_m_b
+    "
+fit_aff_neg_lmer_pre <- lmerTest::lmer(model_aff_neg_lmer_pre, data = d_long_50_imp)
+print_res(broom.mixed::tidy(fit_aff_neg_lmer_pre, conf.int = T), imputation = FALSE)
 ```
 
-## Positive Affect
+## Without imputation
 
-For simplicity, note that we use mean-scores and not factors scores here.
+Final analysis, but now don't use any imputed data whatsoever. Was preregistered as additional analysis to provide comparison.
+
+### Life satisfaction
 
 ```{r}
-model_aff_pos_lmer_mim <- with(d_long_mim_mice, exp =  
-                           lmerTest::lmer(aff_pos_m ~ 
+model_life_sat_lmer_noi <- "
+  life_sat ~
     (1 | id) + (1 | wave) + 
@@
     soc_med_fb_b + soc_med_ig_b + soc_med_wa_b + soc_med_yt_b + soc_med_tw_b +  
-                                        age + male + born_aus + born_aus_prnts + edu_fac + employment_fac + health_b + 
+    age + male + born_aus + born_aus_prnts + edu_fac + employment_fac +
     res_vienna + acc_bal + acc_gar + home_sqm + 
-                                        med_txt_kro_b + med_txt_sta_b + med_txt_pre_b + med_txt_oes_b + med_txt_kur_b + med_txt_slz_b + med_txt_son_b + 
-                                        med_vid_orf_b + med_vid_pri_b + 
-                                        risk_prop_b + loc_cntrl_int_m_b + 
+    corona_pos_b + corona_pos_w +
+    work_h_b + work_h_w +
+    work_homeoff_b +  work_homeoff_w +
+    hh_income_b + hh_income_w +
+    hh_adults + hh_child18 + hh_child17 + hh_child14 + hh_child5 + hh_child2 + 
+    hh_oldfam + hh_outfam + hh_partner +
+    home_owner +
+    risk_prop_b + risk_prop_w + 
     act_wrk_w + act_spo_w + act_frn_w + act_sho_w + act_pet_w + 
     act_wrk_b + act_spo_b + act_frn_b + act_sho_b + act_pet_b +
-                                        sat_dem_w + sat_dem_b)
-)
-model_aff_pos_lmer_mim <- summary(pool(model_aff_pos_lmer_mim), conf.int = TRUE)
+    health_w + health_b +
+    loc_cntrl_int_m_w + loc_cntrl_int_m_b
+"
+fit_life_sat_lmer_noi <- lmerTest::lmer(model_life_sat_lmer_noi, data = d_long_50)
+print_res(broom.mixed::tidy(fit_life_sat_lmer_noi, conf.int = T), imputation = FALSE)
 ```
 
-## Negative Affect
-
-For simplicity, note that we use mean-scores and not factors scores here.
+### Positive Affect
 
 ```{r}
-model_aff_neg_lmer_mim <- with(d_long_mim_mice, exp =  
-                               lmerTest::lmer(aff_neg_m ~ 
+model_aff_pos_lmer_noi <- "
+  aff_pos_fs ~ 
     (1 | id) + (1 | wave) + 
@@
     soc_med_fb_b + soc_med_ig_b + soc_med_wa_b + soc_med_yt_b + soc_med_tw_b +  
-                                        age + male + born_aus + born_aus_prnts + edu_fac + employment_fac + health_b + 
+    age + male + born_aus + born_aus_prnts + edu_fac + employment_fac +
     res_vienna + acc_bal + acc_gar + home_sqm + 
-                                        med_txt_kro_b + med_txt_sta_b + med_txt_pre_b + med_txt_oes_b + med_txt_kur_b + med_txt_slz_b + med_txt_son_b + 
-                                        med_vid_orf_b + med_vid_pri_b + 
-                                        risk_prop_b + loc_cntrl_int_m_b + 
+    corona_pos_b + corona_pos_w +
+    work_h_b + work_h_w +
+    work_homeoff_b +  work_homeoff_w +
+    hh_income_b + hh_income_w +
+    hh_adults + hh_child18 + hh_child17 + hh_child14 + hh_child5 + hh_child2 + 
+    hh_oldfam + hh_outfam + hh_partner +
+    home_owner +
+    risk_prop_b + risk_prop_w + 
     act_wrk_w + act_spo_w + act_frn_w + act_sho_w + act_pet_w + 
     act_wrk_b + act_spo_b + act_frn_b + act_sho_b + act_pet_b +
-                                        sat_dem_w + sat_dem_b)
-)
-model_aff_neg_lmer_mim <- summary(pool(model_aff_neg_lmer_mim), conf.int = TRUE)
+    health_w + health_b +
+    loc_cntrl_int_m_w + loc_cntrl_int_m_b
+"
+fit_aff_pos_lmer_noi <- lmerTest::lmer(model_aff_pos_lmer_noi, data = d_long_50)
+print_res(broom.mixed::tidy(fit_aff_pos_lmer_noi, conf.int = T), imputation = FALSE)
 ```
 
-# Figures
-
-In what follows, see figure with all results combined.
-
-## Activities
+### Negative Affect
 
 ```{r}
+model_aff_neg_lmer_noi <- "
+  aff_neg_fs ~ 
+    (1 | id) + (1 | wave) + 
+    soc_med_read_w + soc_med_like_share_w + soc_med_post_w  + 
+    soc_med_fb_w + soc_med_ig_w + soc_med_wa_w + soc_med_yt_w + soc_med_tw_w +
+    soc_med_read_b + soc_med_like_share_b + soc_med_post_b + 
+    soc_med_fb_b + soc_med_ig_b + soc_med_wa_b + soc_med_yt_b + soc_med_tw_b +  
+    age + male + born_aus + born_aus_prnts + edu_fac + employment_fac +
+    res_vienna + acc_bal + acc_gar + home_sqm + 
+    corona_pos_b + corona_pos_w +
+    work_h_b + work_h_w +
+    work_homeoff_b +  work_homeoff_w +
+    hh_income_b + hh_income_w +
+    hh_adults + hh_child18 + hh_child17 + hh_child14 + hh_child5 + hh_child2 + 
+    hh_oldfam + hh_outfam + hh_partner +
+    home_owner +
+    risk_prop_b + risk_prop_w + 
+    act_wrk_w + act_spo_w + act_frn_w + act_sho_w + act_pet_w + 
+    act_wrk_b + act_spo_b + act_frn_b + act_sho_b + act_pet_b +
+    health_w + health_b +
+    loc_cntrl_int_m_w + loc_cntrl_int_m_b
+"
+fit_aff_neg_lmer_noi <- lmerTest::lmer(model_aff_neg_lmer_noi, data = d_long_50)
+print_res(broom.mixed::tidy(fit_aff_neg_lmer_noi, conf.int = T), imputation = FALSE)
+```
+
+## Figure
+
+ - In what follows, see figure with all results combined. 
+ - Results show that different models are comparable. 
+ - The model without control variables stands out a bit, showing that controlling for additional varying variables makes a difference.
+ - The preregistered model without imputation also has larger confidence intervals, which was to be expected given fewer data.
+
+```{r figure, warning=F, fig.height=6}
 # get data
-dat_fig_results_activity <- get_dat_res(model_aff_neg_lmer, model_aff_pos_lmer, model_life_sat_lmer, type = "activity", analysis = "1. regular")
-dat_fig_results_activity_nco <- get_dat_res(model_aff_neg_lmer_nco, model_aff_pos_lmer_nco, model_life_sat_lmer_nco, type = "activity", analysis = "2. no covars")
-dat_fig_results_activity_m <- get_dat_res(model_aff_neg_lmer_m, model_aff_pos_lmer_m, model_life_sat_lmer, type = "activity", analysis = "3. mean scores")
-dat_fig_results_activity_noi <- get_dat_res(model_aff_neg_lmer_noi, model_aff_pos_lmer_noi, model_life_sat_lmer_noi, type = "activity", analysis = "4. no imp")
-dat_fig_results_activity_mim <- get_dat_res(model_aff_neg_lmer_mim, model_aff_pos_lmer_mim, model_life_sat_lmer_mim, type = "activity", analysis = "5. mult imp")
+dat_fig_results_activity_pub <- get_dat_res(
+  fit_aff_neg_lmer_pub, fit_aff_pos_lmer_pub, fit_life_sat_lmer_pub, 
+  type = "Activity", analysis = "1. Final, preferred", variance = "within")
 
-dat_fig_results_activity <- rbind(
-  dat_fig_results_activity,
+dat_fig_results_channel_pub <- get_dat_res(
+  fit_aff_neg_lmer_pub, fit_aff_pos_lmer_pub, fit_life_sat_lmer_pub, 
+  type = "Channels", analysis = "1. Final, preferred", variance = "within")
+
+dat_fig_results_activity_add <- get_dat_res(
+  fit_aff_neg_lmer_add, fit_aff_pos_lmer_add, fit_life_sat_lmer_add, 
+  type = "Activity", analysis = "2. Additional covariates", variance = "within")
+
+dat_fig_results_channel_add <- get_dat_res(
+  fit_aff_neg_lmer_add, fit_aff_pos_lmer_add, fit_life_sat_lmer_add, 
+  type = "Channels", analysis = "2. Additional covariates", variance = "within")
+
+dat_fig_results_activity_nco <- get_dat_res(
+  fit_aff_neg_lmer_nco, fit_aff_pos_lmer_nco, fit_life_sat_lmer_nco, 
+  type = "Activity", analysis = "3. No covariates", variance = "within")
+
+dat_fig_results_channel_nco <- get_dat_res(
+  fit_aff_neg_lmer_nco, fit_aff_pos_lmer_nco, fit_life_sat_lmer_nco, 
+  type = "Channels", analysis = "3. No covariates", variance = "within")
+
+dat_fig_results_activity_pre <- get_dat_res(
+  fit_aff_neg_lmer_pre, fit_aff_pos_lmer_pre, fit_life_sat_lmer_pre, 
+  type = "Activity", analysis = "4. As preregistered", variance = "within")
+
+dat_fig_results_channel_pre <- get_dat_res(
+  fit_aff_neg_lmer_pre, fit_aff_pos_lmer_pre, fit_life_sat_lmer_pre, 
+  type = "Channels", analysis = "4. As preregistered", variance = "within")
+
+dat_fig_results_activity_noi <- get_dat_res(
+  fit_aff_neg_lmer_noi, fit_aff_pos_lmer_noi, fit_life_sat_lmer_noi, 
+  type = "Activity", analysis = "5. No imputation", variance = "within")
+
+dat_fig_results_channel_noi <- get_dat_res(
+  fit_aff_neg_lmer_noi, fit_aff_pos_lmer_noi, fit_life_sat_lmer_noi, 
+  type = "Channels", analysis = "5. No imputation", variance = "within")
+
+dat_fig_results_comp <- rbind(
+  dat_fig_results_activity_pub, 
+  dat_fig_results_channel_pub,
   dat_fig_results_activity_nco, 
-  dat_fig_results_activity_m,
+  dat_fig_results_channel_nco,
+  dat_fig_results_activity_add, 
+  dat_fig_results_channel_add,
+  dat_fig_results_activity_pre, 
+  dat_fig_results_channel_pre,
   dat_fig_results_activity_noi, 
-  dat_fig_results_activity_mim
+  dat_fig_results_channel_noi
 )
+
 # make figure
-fig_results_activity <- make_graph_res(
-  dat_fig_results_activity,
-  sesoi = "est"
+fig_results_comp <- make_graph_res(
+  dat_fig_results_comp,
+  sesoi = "est",
+  facet = "type"
   )
-fig_results_activity
+fig_results_comp
 
 # safe figure
-ggsave("figures/fig_results_activity.pdf", 
-       width = 7, height = 4,
-       plot = fig_results_activity)
+ggsave("figures/fig_results_comp.png", 
+       width = 7, height = 7,
+       plot = fig_results_comp)
 ```
 
-## Channels
+# Exploratory analyses
+## 1 Month interval
+
+Let's now inspect effects of media use on well-being some time later (i.e., 1 month or 4 months).
+This is especially relevant for life-satisfaction, which is more stable.
+
+Let's first format data to introduce lags.
 
 ```{r}
-# get data
-dat_fig_results_channels <- get_dat_res(model_aff_neg_lmer, model_aff_pos_lmer, model_life_sat_lmer, type = "channels", analysis = "1. regular")
-dat_fig_results_channels_nco <- get_dat_res(model_aff_neg_lmer_nco, model_aff_pos_lmer_nco, model_life_sat_lmer_nco, type = "channels", analysis = "2. no covars")
-dat_fig_results_channels_m <- get_dat_res(model_aff_neg_lmer_m, model_aff_pos_lmer_m, model_life_sat_lmer, type = "channels", analysis = "3. mean scores")
-dat_fig_results_channels_noi <- get_dat_res(model_aff_neg_lmer_noi, model_aff_pos_lmer_noi, model_life_sat_lmer_noi, type = "channels", analysis = "4. no imp")
-dat_fig_results_channels_mim <- get_dat_res(model_aff_neg_lmer_mim, model_aff_pos_lmer_mim, model_life_sat_lmer_mim, type = "channels", analysis = "5. mult imp")
+# define waves with 1 month lag to well-being
+waves_predictors_1m <- c(5, 11, 18, 24, 29)
 
-dat_fig_results_channels <- rbind(
-  dat_fig_results_channels,
-  dat_fig_results_channels_nco,
-  dat_fig_results_channels_m,
-  dat_fig_results_channels_noi,
-  dat_fig_results_channels_mim
+# define waves where media use was measured
+waves_media_use <- c(1, 8, 17, 23, 28)
+
+waves_include_1m <- c(waves_media_use, waves_predictors_1m)
+
+d_long_100_mim_mice_1m <- 
+  d_long_100_mim_mice %>% 
+  complete(
+    'long', 
+    include = TRUE
+    ) %>%
+  filter(
+    wave %in% waves_include_1m
+    ) %>% 
+  group_by(id) %>% 
+  mutate(
+    life_sat = lead(life_sat, n = 1, default = NA),
+    aff_pos_m = lead(aff_pos_m, n = 1, default = NA), 
+    aff_neg_m = lead(aff_neg_m, n = 1, default = NA),
+    ) %>%
+  ungroup() %>% 
+  as.mids() %>% 
+  # select waves used for analyses
+  filter(
+    wave %in% waves_predictors_1m
   )
-# make figure
-fig_results_channels <- make_graph_res(
-  dat_fig_results_channels,
-  sesoi = "est"
+
+waves_predictors_4m <- c(13, 15, 21, 25, 32)
+waves_include_4m <- c(waves_media_use, waves_predictors_4m)
+
+d_long_100_mim_mice_4m <- 
+  d_long_100_mim_mice %>% 
+  complete(
+    'long', 
+    include = TRUE
+    ) %>%
+  filter(
+    wave %in% waves_include_4m
+    ) %>% 
+  group_by(id) %>% 
+  mutate(
+    life_sat = lead(life_sat, n = 1, default = NA),
+    aff_pos_m = lead(aff_pos_m, n = 1, default = NA), 
+    aff_neg_m = lead(aff_neg_m, n = 1, default = NA),
+    ) %>%
+  ungroup() %>% 
+  as.mids() %>% 
+  # select waves used for analyses
+  filter(
+    wave %in% waves_predictors_4m
   )
-fig_results_channels
 
-# safe figure
-ggsave("figures/fig_results_channels.pdf", 
-       width = 7, height = 4,
-       plot = fig_results_channels)
 ```
 
-Save results so that we can extract them in the manuscript.
+#### Life satisfaction
 
 ```{r}
-save.image("data/workspace_2.RData")
+fit_life_sat_lmer_1m <- with(d_long_100_mim_mice_1m, exp = lmerTest::lmer(model_life_sat_lmer_pub))
+fit_life_sat_lmer_1m <- summary(pool(fit_life_sat_lmer_1m), conf.int = TRUE)
+print_res(fit_life_sat_lmer_1m)
 ```
+
+We find no significant effects.
+
+#### Positive Affect
+
+```{r}
+fit_aff_pos_lmer_1m <- with(d_long_100_mim_mice_1m, exp = lmerTest::lmer(model_aff_pos_lmer_pub))
+fit_aff_pos_lmer_1m <- summary(pool(fit_aff_pos_lmer_1m), conf.int = TRUE)
+print_res(fit_aff_pos_lmer_1m)
+```
+
+We find no significant effects.
+
+#### Negative Affect
+
+```{r}
+fit_aff_neg_lmer_1m <- with(d_long_100_mim_mice_1m, exp = lmerTest::lmer(model_aff_neg_lmer_pub))
+fit_aff_neg_lmer_1m <- summary(pool(fit_aff_neg_lmer_1m), conf.int = TRUE)
+print_res(fit_aff_neg_lmer_1m)
+```
+
+We find no significant effects.
+
+## 4 Months Interval
+
+Let's not see if we find effects looking at longer intervals.
+
+### Life satisfaction
+
+```{r}
+fit_life_sat_lmer_4m <- with(d_long_100_mim_mice_4m, exp = lmerTest::lmer(model_life_sat_lmer_pub))
+fit_life_sat_lmer_4m <- summary(pool(fit_life_sat_lmer_4m), conf.int = TRUE)
+print_res(fit_life_sat_lmer_4m)
+```
+
+We find no significant effects.
+
+### Positive Affect
+
+```{r}
+fit_aff_pos_lmer_4m <- with(d_long_100_mim_mice_4m, exp = lmerTest::lmer(model_aff_pos_lmer_pub))
+fit_aff_pos_lmer_4m <- summary(pool(fit_aff_pos_lmer_4m), conf.int = TRUE)
+print_res(fit_aff_pos_lmer_4m)
+```
+
+We find no significant effects.
+
+### Negative Affect
+
+```{r}
+fit_aff_neg_lmer_4m <- with(d_long_100_mim_mice_4m, exp = lmerTest::lmer(model_aff_neg_lmer_pub))
+fit_aff_neg_lmer_4m <- summary(pool(fit_aff_neg_lmer_4m), conf.int = TRUE)
+print_res(fit_aff_neg_lmer_4m)
+```
+
+We find no significant effects.
+
+## Gender
+
+Let's next see if effects differ for males and females.
+
+### Life satisfaction
+
+```{r}
+model_life_sat_lmer_male <- "
+  life_sat ~ 
+    (1 | id) + (1 | wave) + 
+    soc_med_read_w * male + soc_med_like_share_w * male + soc_med_post_w * male + 
+    soc_med_fb_w * male + soc_med_ig_w * male + soc_med_wa_w * male + soc_med_yt_w * male + soc_med_tw_w * male +
+    soc_med_read_b + soc_med_like_share_b + soc_med_post_b + 
+    soc_med_fb_b + soc_med_ig_b + soc_med_wa_b + soc_med_yt_b + soc_med_tw_b +  
+    age + born_aus + born_aus_prnts + edu_fac + employment_fac +
+    res_vienna + acc_bal + acc_gar + home_sqm + 
+    corona_pos_b + corona_pos_w +
+    work_h_b + work_h_w +
+    work_homeoff_b +  work_homeoff_w +
+    hh_income_b + hh_income_w +
+    hh_adults + hh_child18 + hh_child17 + hh_child14 + hh_child5 + hh_child2 + 
+    hh_oldfam + hh_outfam + hh_partner +
+    home_owner +
+    risk_prop_b + risk_prop_w + 
+    act_wrk_w + act_spo_w + act_frn_w + act_sho_w + act_pet_w + 
+    act_wrk_b + act_spo_b + act_frn_b + act_sho_b + act_pet_b +
+    health_w + health_b +
+    loc_cntrl_int_m_w + loc_cntrl_int_m_b
+"
+
+fit_life_sat_lmer_male <- with(d_long_100_mim_mice, exp = lmerTest::lmer(model_life_sat_lmer_male))
+fit_life_sat_lmer_male <- summary(pool(fit_life_sat_lmer_male), conf.int = TRUE)
+print_res(fit_life_sat_lmer_male)
+```
+
+Interaction effects are insignificant, showing that effects don't differ across genders.
+
+### Positive Affect
+
+```{r}
+model_aff_pos_lmer_male <- "
+  aff_pos_m ~ 
+    (1 | id) + (1 | wave) + 
+    soc_med_read_w * male + soc_med_like_share_w * male + soc_med_post_w * male + 
+    soc_med_fb_w * male + soc_med_ig_w * male + soc_med_wa_w * male + soc_med_yt_w * male + soc_med_tw_w * male +
+    soc_med_read_b + soc_med_like_share_b + soc_med_post_b + 
+    soc_med_fb_b + soc_med_ig_b + soc_med_wa_b + soc_med_yt_b + soc_med_tw_b +  
+    age + born_aus + born_aus_prnts + edu_fac + employment_fac +
+    res_vienna + acc_bal + acc_gar + home_sqm + 
+    corona_pos_b + corona_pos_w +
+    work_h_b + work_h_w +
+    work_homeoff_b +  work_homeoff_w +
+    hh_income_b + hh_income_w +
+    hh_adults + hh_child18 + hh_child17 + hh_child14 + hh_child5 + hh_child2 + 
+    hh_oldfam + hh_outfam + hh_partner +
+    home_owner +
+    risk_prop_b + risk_prop_w + 
+    act_wrk_w + act_spo_w + act_frn_w + act_sho_w + act_pet_w + 
+    act_wrk_b + act_spo_b + act_frn_b + act_sho_b + act_pet_b +
+    health_w + health_b +
+    loc_cntrl_int_m_w + loc_cntrl_int_m_b
+"
+
+fit_aff_pos_lmer_male <- with(d_long_100_mim_mice, exp = lmerTest::lmer(model_aff_pos_lmer_male))
+fit_aff_pos_lmer_male <- summary(pool(fit_aff_pos_lmer_male), conf.int = TRUE)
+print_res(fit_aff_pos_lmer_male)
+```
+
+Interaction effects are insignificant, showing that effects don't differ across genders.
+
+### Negative Affect
+
+```{r}
+model_aff_neg_lmer_male <- "
+  aff_neg_m ~ 
+    (1 | id) + (1 | wave) + 
+    soc_med_read_w * male + soc_med_like_share_w * male + soc_med_post_w * male + 
+    soc_med_fb_w * male + soc_med_ig_w * male + soc_med_wa_w * male + soc_med_yt_w * male + soc_med_tw_w * male +
+    soc_med_read_b + soc_med_like_share_b + soc_med_post_b + 
+    soc_med_fb_b + soc_med_ig_b + soc_med_wa_b + soc_med_yt_b + soc_med_tw_b +  
+    age + born_aus + born_aus_prnts + edu_fac + employment_fac +
+    res_vienna + acc_bal + acc_gar + home_sqm + 
+    corona_pos_b + corona_pos_w +
+    work_h_b + work_h_w +
+    work_homeoff_b +  work_homeoff_w +
+    hh_income_b + hh_income_w +
+    hh_adults + hh_child18 + hh_child17 + hh_child14 + hh_child5 + hh_child2 + 
+    hh_oldfam + hh_outfam + hh_partner +
+    home_owner +
+    risk_prop_b + risk_prop_w + 
+    act_wrk_w + act_spo_w + act_frn_w + act_sho_w + act_pet_w + 
+    act_wrk_b + act_spo_b + act_frn_b + act_sho_b + act_pet_b +
+    health_w + health_b +
+    loc_cntrl_int_m_w + loc_cntrl_int_m_b
+"
+
+fit_aff_neg_lmer_male <- with(d_long_100_mim_mice, exp = lmerTest::lmer(model_aff_neg_lmer_male))
+fit_aff_neg_lmer_male <- summary(pool(fit_aff_neg_lmer_male), conf.int = TRUE)
+print_res(fit_aff_neg_lmer_male)
+```
+
+Interaction effects are insignificant, showing that effects don't differ across genders.
+
+## Age
+
+Let's next inspect if effects differ across age. We'll look at generations (Gen Z, Gen Y, ...).
+Let's first transform data to make age categories.
+
+```{r}
+d_long_100_mim_mice <-
+  d_long_100_mim_mice %>% 
+  complete(
+    'long', 
+    include = TRUE
+    ) %>%
+  mutate(
+    age_gen = ifelse(
+      age > 2022 - 1946, 
+      "Silent",
+      ifelse(
+        age > 2022 - 1965, 
+        "Boomer",
+        ifelse(
+          age > 2022 - 1981, 
+          "Gen X",
+          ifelse(
+            age > 2022 - 1997, 
+            "Gen Y",
+            "Gen Z"
+            )
+          )
+        )
+      )
+    ) %>% 
+  mutate(
+    age_gen = factor(
+      age_gen,
+      c("Gen X", "Gen Z", "Gen Y", "Boomer", "Silent")
+    )
+  ) %>% 
+  as.mids()
+
+d_long_100_mim_mice %>% 
+  complete(
+    'long', 
+    include = TRUE
+    ) %>%
+  select(age_gen) %>% 
+  table()
+```
+
+### Life satisfaction
+
+```{r}
+model_life_sat_lmer_age <- "
+  life_sat ~ 
+    (1 | id) + (1 | wave) + 
+    soc_med_read_w * age_gen + soc_med_like_share_w * age_gen + soc_med_post_w * age_gen + 
+    soc_med_fb_w * age_gen + soc_med_ig_w * age_gen + soc_med_wa_w * age_gen + soc_med_yt_w * age_gen + soc_med_tw_w * age_gen +
+    soc_med_read_b + soc_med_like_share_b + soc_med_post_b + 
+    soc_med_fb_b + soc_med_ig_b + soc_med_wa_b + soc_med_yt_b + soc_med_tw_b +  
+    male + born_aus + born_aus_prnts + edu_fac + employment_fac +
+    res_vienna + acc_bal + acc_gar + home_sqm + 
+    corona_pos_b + corona_pos_w +
+    work_h_b + work_h_w +
+    work_homeoff_b +  work_homeoff_w +
+    hh_income_b + hh_income_w +
+    hh_adults + hh_child18 + hh_child17 + hh_child14 + hh_child5 + hh_child2 + 
+    hh_oldfam + hh_outfam + hh_partner +
+    home_owner +
+    risk_prop_b + risk_prop_w + 
+    act_wrk_w + act_spo_w + act_frn_w + act_sho_w + act_pet_w + 
+    act_wrk_b + act_spo_b + act_frn_b + act_sho_b + act_pet_b +
+    health_w + health_b +
+    loc_cntrl_int_m_w + loc_cntrl_int_m_b
+"
+
+fit_life_sat_lmer_age <- with(d_long_100_mim_mice, exp = lmerTest::lmer(model_life_sat_lmer_age))
+fit_life_sat_lmer_age <- summary(pool(fit_life_sat_lmer_age), conf.int = TRUE)
+print_res(fit_life_sat_lmer_age) %>% 
+  filter(grepl(".*:.*", term))
+```
+
+Interaction effects are insignificant, showing that effects don't differ across generations.
+
+### Positive Affect
+
+```{r}
+model_aff_pos_lmer_age <- "
+  aff_pos_m ~ 
+    (1 | id) + (1 | wave) + 
+    soc_med_read_w * age_gen + soc_med_like_share_w * age_gen + soc_med_post_w * age_gen + 
+    soc_med_fb_w * age_gen + soc_med_ig_w * age_gen + soc_med_wa_w * age_gen + soc_med_yt_w * age_gen + soc_med_tw_w * age_gen +
+    soc_med_read_b + soc_med_like_share_b + soc_med_post_b + 
+    soc_med_fb_b + soc_med_ig_b + soc_med_wa_b + soc_med_yt_b + soc_med_tw_b +  
+    male + born_aus + born_aus_prnts + edu_fac + employment_fac +
+    res_vienna + acc_bal + acc_gar + home_sqm + 
+    corona_pos_b + corona_pos_w +
+    work_h_b + work_h_w +
+    work_homeoff_b +  work_homeoff_w +
+    hh_income_b + hh_income_w +
+    hh_adults + hh_child18 + hh_child17 + hh_child14 + hh_child5 + hh_child2 + 
+    hh_oldfam + hh_outfam + hh_partner +
+    home_owner +
+    risk_prop_b + risk_prop_w + 
+    act_wrk_w + act_spo_w + act_frn_w + act_sho_w + act_pet_w + 
+    act_wrk_b + act_spo_b + act_frn_b + act_sho_b + act_pet_b +
+    health_w + health_b +
+    loc_cntrl_int_m_w + loc_cntrl_int_m_b
+"
+
+fit_aff_pos_lmer_age <- with(d_long_100_mim_mice, exp = lmerTest::lmer(model_aff_pos_lmer_age))
+fit_aff_pos_lmer_age <- summary(pool(fit_aff_pos_lmer_age), conf.int = TRUE)
+print_res(fit_aff_pos_lmer_age) %>% 
+  filter(grepl(".*:.*", term))
+```
+
+Interaction effects are insignificant, showing that effects don't differ across generations.
+
+### Negative Affect
+
+```{r}
+model_aff_neg_lmer_age <- "
+  aff_neg_m ~ 
+    (1 | id) + (1 | wave) + 
+    soc_med_read_w * age_gen + soc_med_like_share_w * age_gen + soc_med_post_w * age_gen + 
+    soc_med_fb_w * age_gen + soc_med_ig_w * age_gen + soc_med_wa_w * age_gen + soc_med_yt_w * age_gen + soc_med_tw_w * age_gen +
+    soc_med_read_b + soc_med_like_share_b + soc_med_post_b + 
+    soc_med_fb_b + soc_med_ig_b + soc_med_wa_b + soc_med_yt_b + soc_med_tw_b +  
+    male + born_aus + born_aus_prnts + edu_fac + employment_fac +
+    res_vienna + acc_bal + acc_gar + home_sqm + 
+    corona_pos_b + corona_pos_w +
+    work_h_b + work_h_w +
+    work_homeoff_b +  work_homeoff_w +
+    hh_income_b + hh_income_w +
+    hh_adults + hh_child18 + hh_child17 + hh_child14 + hh_child5 + hh_child2 + 
+    hh_oldfam + hh_outfam + hh_partner +
+    home_owner +
+    risk_prop_b + risk_prop_w + 
+    act_wrk_w + act_spo_w + act_frn_w + act_sho_w + act_pet_w + 
+    act_wrk_b + act_spo_b + act_frn_b + act_sho_b + act_pet_b +
+    health_w + health_b +
+    loc_cntrl_int_m_w + loc_cntrl_int_m_b
+"
+
+fit_aff_neg_lmer_age <- with(d_long_100_mim_mice, exp = lmerTest::lmer(model_aff_neg_lmer_age))
+fit_aff_neg_lmer_age <- summary(pool(fit_aff_neg_lmer_age), conf.int = TRUE)
+fit_aff_neg_lmer_age_tab <- print_res(fit_aff_neg_lmer_age)
+```
+
+```{r}
+fit_aff_neg_lmer_age_tab %>% 
+  filter(grepl(".*:.*", term))
+```
+
+Interaction effects are insignificant, showing that effects don't differ across generations.
+However, there's one exception: Gen Z differs significantly from Gen X in terms of negative affect experienced when posting more than usual COVID-19 related content.
+
+```{r eval=T, cache=F}
+save.image("data/workspace_3.RData")
+```

### 2023-07-11 analyses_setup.Rmd
Major updates Implemented feedback

---
title: "Analyses Setup"
output:
  html_document:
    toc: TRUE
    toc_float:
      collapsed: FALSE
    toc_depth: 2
    code_folding: hide
    df_print: paged
bibliography      : 
  - "bibliography/r-references.bib"
  - "bibliography/references.bib"
---

```{r knitr-setup, include=F}
set.seed(170819)

knitr::opts_chunk$set(
  warning = F, 
  echo = T, 
  error = F, 
  message = F, 
  cache = T, 
  rows.print = 20,
  cache.lazy = FALSE
  )

options(
  digits = 2, 
  width = 260, 
  scipen = 999,
  max.print = 10000
  )
```

- Here you can find the set-up for the analyses and the results.
- Includes custom functions, data-wrangling, and data imputation. 
- To see the code, click on button "Code".
- Alternatively, you can download the rmd file from the github repo.
- The data are hosted on aussda. Please download them separately, to rerun the analyses (https://doi.org/10.11587/28KQNS). 
- Note that this paper uses a scientific use file, which is at the time of writing not officially published.
- Download the .sav file and save it in the folder "data"

# Set-up
## Load packages

```{r r-setup, results='hide', cache=F}
# install packages
# devtools::install_github("https://github.com/tdienlin/td@v.0.0.2.5")

# define packages
library(broom.mixed)
library(brms)
library(devtools)
library(GGally)
library(ggplot2)
library(gridExtra)
library(haven)
library(kableExtra)
library(knitr)
library(lavaan)
library(lme4)
library(magrittr)
library(mice)
library(mvnormalTest)
library(PerFit)
library(performance)
library(psych)
library(quanteda.textstats)
library(readxl)
library(semTools)
library(td)
library(tidyverse)
# unload("MASS")
```

## Session info

Provide session info to make results reproducible.

```{r}
sessionInfo()
```

# Custom functions

## find_var

Find variables in dataset.

```{r}
find_var <- function(name, data = doc)(
  # finds the variables names for an item for each wave

  doc %>% 
    filter(Label == name) %>%
    dplyr::select(Variable) %>%
    unlist() %>%
    set_names(sub("\\_.*", "", .))
)
```

## get_specs

Extract characteristics from fitted lmer models.

```{r}
get_specs <- function(model){
  # Get mean, max, and min values

  dat <- coefficients(model)$wave
  specs <- data.frame(
    sd = attr(VarCorr(model), "sc"),
    min = min(dat),
    max = max(dat),
    mean = mean(dat$`(Intercept)`)
  )
}
```

## get_dat

Get data from fitted lmer objects for descriptives.

```{r}
get_dat <- function(model){
  coefficients(model)$wave %>% 
    tibble::rownames_to_column("wave") %>% 
    rename(value = "(Intercept)") %>% 
    mutate(wave = as.integer(.$wave),
           value = as.numeric(.$value)) 
} 
```

## get_rel

Determine average reliability for measures across waves

```{r}
get_rel <- function(data, waves=24){

  # extract average reliability from lavaan fitted cfa with several groups
  reliability(data) %>% 
    unlist() %>% 
    matrix(5, waves) %>% 
    as.data.frame() %>% 
    set_rownames(c("alpha", "omega", "omega2", "omega3", "avevar")) %>% 
    summarise(omega = rowMeans(.["omega",])) %>% 
    return()
}
```

## make_graph

Make graph of variables' development.

```{r}
make_graph <- function(model, title, ll, ul, line = TRUE, points = TRUE, labels = FALSE, lmer=FALSE, legend=FALSE){

    if(isTRUE(lmer)){
      dat <- get_dat(model)
    } else{
    dat <- model
  }

  graph <- 
    ggplot(dat, aes(wave, value, color = dimension)) +
    {if(legend) theme(
      legend.position = c(0.85, 0.8),
      legend.title = element_blank()
      )} + 
    {if(!legend) theme(
      legend.position = "none"
    )} +
    theme(axis.title.x=element_blank()) +
    coord_cartesian(ylim = c(ll, ul)) +
    ggtitle(title) +
    {if(line) geom_smooth(se = FALSE, method = 'loess')} +
    {if(points) geom_point()} +
    scale_color_manual(values=c("dodgerblue3", "deepskyblue2", "magenta2", "green2", "red"))
  graph
}
```

## print_res

```{r}
# print results for online material
print_res <- function(object, imputation = TRUE){

  if(isTRUE(imputation)){
    object %>% 
      as.data.frame %>% 
      select(term, estimate, `2.5 %`, `97.5 %`, p.value) %>% 
      mutate(p.value = td::my_round(p.value, "p"))
    }
  else{
    object %>% 
      as.data.frame %>% 
      select(term, estimate, `2.5 %` = conf.low, `97.5 %` = conf.high, p.value) %>% 
      mutate(p.value = td::my_round(p.value, "p"))
  }
}
```

## get_dat_res

Get data of lmer objects for specific results.

```{r}
# get data
get_dat_res <- function(data_aff_neg, data_aff_pos, data_life_sat, variance = "within", type = "channels", analysis = "Standardized"){

  if(isTRUE(class(data_aff_neg) == "lmerModLmerTest")) {
         dat_fig_results <- 
           broom.mixed::tidy(data_aff_neg, conf.int = T) %>% 
           mutate(dv = "aff_neg") %>% 
           rbind(
             broom.mixed::tidy(data_aff_pos, conf.int = T) %>% 
               mutate(dv = "aff_pos")) %>% 
           rbind(
             broom.mixed::tidy(data_life_sat, conf.int = T) %>% 
               mutate(dv = "life_sat"))
         } else{ 
           dat_fig_results <- 
             data_aff_neg %>% 
             mutate(dv = "aff_neg") %>% 
             rbind(data_aff_pos %>% 
                   mutate(dv = "aff_pos")) %>% 
             rbind(data_life_sat %>% 
                     mutate(dv = "life_sat")) %>% 
           rename(conf.low = `2.5 %`, conf.high = `97.5 %`)
           }
  dat_fig_results %<>% 
  mutate(
    Variance = ifelse(grepl(".*_w\\>", .$term), "within", "between"),
    iv = gsub("_(w|b)\\>", "", .$term)
         ) %>%
  mutate(
    Variance = factor(.$Variance, 
                  levels = c("within", "between")),
    dv = factor(.$dv, 
                levels = c("life_sat", "aff_pos", "aff_neg"),
                labels = c("Life satisfaction", "Positive affect", "Negative affect")),
    Type = type,
    Analysis = analysis
  ) %>% 
  select(dv, iv, Variance, Type, Analysis, estimate, conf.low, conf.high, p.value) %>% 
    filter(Variance == variance)

  # select Social Media type of activity
  if(type == "Activity") {
    dat_fig_results %<>%
    filter(iv %in% c("soc_med_read", "soc_med_like_share", "soc_med_post")) %>% 
    mutate(
      iv = factor(.$iv, 
                  levels = c("soc_med_post", "soc_med_like_share", "soc_med_read"), 
                  labels = c("Posting", "Liking & Sharing", "Reading"))
    )
  } else if(type == "Channels"){
    dat_fig_results %<>%
    filter(iv %in% c("soc_med_fb", "soc_med_ig", "soc_med_wa", "soc_med_yt", "soc_med_tw")) %>% 
    mutate(iv = factor(.$iv, levels = c("soc_med_yt", "soc_med_wa", "soc_med_ig", "soc_med_tw", "soc_med_fb"), labels = c("YouTube", "WhatsApp", "Instagram", "Twitter", "Facebook")))
  } else if(type == "Controls"){
    dat_fig_results %<>%
      filter(iv %in% c("male", "health", "loc_cntrl_int_m",
                       "act_spo", "sat_dem", "corona_pos")) %>% 
      mutate(iv = factor(.$iv, levels = c("male", "health", "loc_cntrl_int_m",
                                          "act_spo", "sat_dem", "corona_pos"), 
                         labels = c("Male", "Health", "Internal locus of control",
                                    "Sport", "Satisfaction democracy", "Corona positive")))
  } else if(type == "Living\nconditions"){
    dat_fig_results %<>%
      filter(iv %in% c("health", "corona_pos", "work_h", "work_homeoff", "hh_income")) %>% 
      mutate(iv = factor(.$iv, levels = c("health", "corona_pos", "work_h", "work_homeoff", "hh_income"),
                         labels = c("Health", "Corona positive", "Work hours", "Homeoffice", "Household income")))
  } else if(type == "News\nuse"){
    dat_fig_results %<>%
      filter(iv %in% c("med_txt_kro", "med_txt_sta", "med_txt_pre", "med_txt_oes", "med_txt_kur", "med_txt_slz", "med_txt_son", "med_vid_orf", "med_vid_pri")) %>% 
      mutate(iv = factor(.$iv, levels = c("med_txt_kro", "med_txt_sta", "med_txt_pre", "med_txt_oes", "med_txt_kur", "med_txt_slz", "med_txt_son", "med_vid_orf", "med_vid_pri"),
                         labels = c("Krone", "Der Standard", "Die Presse", "Österreich", "Kurier", "Salzb. Nachrichten", "Other", "ORF", "Private news")))
  } else if(type == "Outdoor\nactivities"){
    dat_fig_results %<>%
      filter(iv %in% c("act_wrk", "act_spo", "act_frn", "act_sho", "act_pet")) %>% 
      mutate(iv = factor(.$iv, levels = c("act_wrk", "act_spo", "act_frn", "act_sho", "act_pet"),
                         labels = c("Working", "Sports", "Friends", "Shopping", "With pets")))
  } else if(type == "Psycho-\nlogy"){
    dat_fig_results %<>%
      filter(iv %in% c("risk_prop", "loc_cntrl_int_m", "sat_dem")) %>% 
      mutate(iv = factor(.$iv, levels = c("risk_prop", "loc_cntrl_int_m", "sat_dem"),
                         labels = c("Risk taking", "Internal locus of control", "Satisfaction democracy")))
  }
  return(dat_fig_results)
}
```

## make_graph_res

Make graph of effects

```{r}
make_graph_res <- function(data, sesoi = NULL, legend = TRUE, facet = "type", title = NULL){
  ggplot(data, aes(x = estimate, y = iv)) + 
  scale_color_manual(values = c("black", "grey75", "darkcyan", "deepskyblue", "cornflowerblue", "darkcyan", "aquamarine")) +
  geom_vline(xintercept = 0, lwd = .75, colour = "darkgrey") + 
  geom_errorbarh(aes(xmin = conf.low, xmax = conf.high, color = Analysis),
                lwd = .75, height = .2, position = position_dodge(-.7)) +
  geom_point(aes(color = Analysis), size = 1.5, position = position_dodge(-.7)) +
    {if(isTRUE(sesoi == "est")) geom_vline(data=filter(data, dv=="Life satisfaction"), aes(xintercept=-.3), colour="darkgrey", linetype = "dashed")} + 
    {if(isTRUE(sesoi == "est")) geom_vline(data=filter(data, dv=="Life satisfaction"), aes(xintercept=.3), colour="darkgrey", linetype = "dashed")} + 
    {if(isTRUE(sesoi == "est")) geom_vline(data=filter(data, dv!="Life satisfaction"), aes(xintercept=-.15), colour="darkgrey", linetype = "dashed")} +
    {if(isTRUE(sesoi == "est")) geom_vline(data=filter(data, dv!="Life satisfaction"), aes(xintercept=.15), colour="darkgrey", linetype = "dashed")} +
    {if(isTRUE(sesoi == "std")) geom_vline(aes(xintercept=.1), colour="darkgrey", linetype = "dashed")} + 
    {if(isTRUE(sesoi == "std")) geom_vline(aes(xintercept=-.1), colour="darkgrey", linetype = "dashed")} +   
  theme(
    axis.title.y = element_blank(),
    plot.title = element_text(hjust = .5),
    panel.spacing = unit(.9, "lines"),
    text = element_text(size = 12),
    axis.text.y = element_text(hjust = 1)
    ) +
  {if(!is.null(title)) ggtitle(title)} +
  {if(!isTRUE(legend)) theme(legend.position = "none")} + 
  # guides(colour = guide_legend(reverse=T)) +
    {if(facet == "type")
      facet_grid(rows = vars(Type),
             cols = vars(dv),
             scales = "free",
             space = "free_y")
    } + 
    {if(facet == "analysis")
      facet_grid(rows = vars(Variance),
             cols = vars(dv),
             scales = "free",
             space = "free_y")
    }
}
```

## get_fs

Extract factor scores.

```{r}
get_fs <- function(model) {
  # the aim is to get factor scores on original scaling
  # hence export factor values for all items
  # then compute average
  dat_ov <- lavaan::lavPredict(model, type = "ov", assemble = TRUE) %>% 
    mutate(fs = rowMeans(.[1:3]))
  return(dat_ov$fs)
}
```

# Data Wrangling
## Load data

```{r load-data}
d_spss_en <- read_spss("data/W1-34_ACPP_V20230303_EN.SAV")
d_spss <- read_spss("data/W1-34_ACPP_V20230303.SAV")
write.csv(d_spss, file = "data/W1-34_ACPP_V20230303_EN.csv")
d_raw <- read_csv("data/W1-34_ACPP_V20230303_EN.csv")
```

## Documentation

Make file with documentation about variables

```{r}
# get labels
doc_raw <- map(d_spss, attr, "label")

# reformat
doc_raw <- 
  data.frame(
    Variable = names(doc),
    Label = unlist(doc, use.names = FALSE)
  )

# define waves where media use was measured
waves_media_use <- c(1, 8, 17, 23, 28)
```

## Find names

Identify the names of each item for each wave.

```{r get-var-names}
# identify variable names of items, using custom function
health <- find_var("Gesundheitszustand") # wsn't collected at wave 21
aff_neg_1 <- find_var("Depressivitaet: Einsam")
aff_neg_2 <- find_var("Depressivitaet: Aergerlich")
aff_neg_3 <- find_var("Depressivitaet: So niedergeschlagen")
aff_neg_4 <- find_var("Depressivitaet: Sehr nervoes")
aff_neg_5 <- find_var("Depressivitaet: Aengstlich")
aff_neg_6 <- find_var("Depressivitaet: Bedrueckt und traurig")
aff_pos_1 <- find_var("Depressivitaet: Ruhig und gelassen")
aff_pos_2 <- find_var("Depressivitaet: Gluecklich")
aff_pos_3 <- find_var("Depressivitaet: Voller Energie")
act_wrk <- find_var("Zuhause verlassen: Arbeit")
act_spo <- find_var("Zuhause verlassen: Sport")
act_frn <- find_var("Zuhause verlassen: Freunde oder Verwandte treffen")
act_sho <- find_var("Zuhause verlassen: Essen einkaufen")
act_pet <- find_var("Zuhause verlassen: Haustier ausfuehren")
sat_dem <- find_var("Demokratiezufriedenheit: Oesterreich")
work_h <- find_var("Arbeitsstunden: Jetzt pro Woche")
work_homeoff <- find_var("Aenderung berufliche Situation: Home-Office")
hh_income <- find_var("Aktuelles Haushaltseinkommen")
med_txt_kro <- find_var("Mediennutzung: Kronen Zeitung oder www.krone.at")
med_txt_sta <- find_var("Mediennutzung: Der Standard oder derstandard.at")
med_txt_pre <- find_var("Mediennutzung: Die Presse oder diepresse.com")
med_txt_oes <- find_var("Mediennutzung: Oesterreich oder oe24.at")
med_txt_kur <- find_var("Mediennutzung: Kurier oder kurier.at")
med_txt_slz <- find_var("Mediennutzung: Salzburger Nachrichten oder salzburg.at")
med_txt_son <- find_var("Mediennutzung: Sonstige oesterreichische Tageszeitungen")
med_vid_orf <- find_var("Mediennutzung: ORF (Nachrichten)")
med_vid_pri <- find_var("Mediennutzung: Privatfernsehen (Nachrichten)")
soc_med_fb <- find_var("Soziale Medien: Facebook")
soc_med_tw <- find_var("Soziale Medien: Twitter")
soc_med_ig <- find_var("Soziale Medien: Instagram")
soc_med_yt <- find_var("Soziale Medien: Youtube")
soc_med_wa <- find_var("Soziale Medien: WhatsApp")
soc_med_read <- find_var("Soziale Medien Aktivitaet: Postings zu Corona anderer lesen")
soc_med_like_share <- find_var("Soziale Medien Aktivitaet: Postings liken, teilen oder retweeten")
soc_med_post <- find_var("Soziale Medien Aktivitaet: Selber Postings zu Corona verfassen")
life_sat <- find_var("Lebenszufriedenheit")
risk_prop <- find_var("Risikobereitschaft")
loc_cntrl_int_1 <- find_var("Psychologie: Habe Leben selbst in der Hand")
loc_cntrl_int_2 <- find_var("Psychologie: Belohnung durch Anstrengung")
loc_cntrl_int_3 <- find_var("Psychologie: Fremdbestimmung")
loc_cntrl_int_4 <- find_var("Psychologie: Schicksal")
trst_media <- find_var("Vertrauen: ORF")
trst_police <- find_var("Vertrauen: Polizei")
trst_media <- find_var("Vertrauen: Parlament")
trst_hlthsec <- find_var("Vertrauen: Gesundheitswesen")
trst_gov <- find_var("Vertrauen: Bundesregierung")
trst_army <- find_var("Vertrauen: Bundesheer")
corona_pos <- c(
  find_var("Corona-Diagnose: Respondent"),
  find_var("Corona-Diagnose: Monat")
)
```

## Select variables

```{r select-vars}
d_wide <- d_raw %>% 
  select(
    id = RESPID,
    gender = SD_GENDER,
    acc_bal = SD_ACCESS_BALCONY,
    acc_gar = SD_ACCESS_GARDEN,
    year_birth = SD_BIRTHYEAR,
    born_aus = SD_BORN_AUSTRIA,
    born_aus_prnts = SD_MIGR_BACKGR,
    county = SD_BULA,
    edu = SD_EDU,
    employment = SD_EMPLSTATUS_FEB2020,
    hh_adults = SD_HH_ADULTS,
    hh_child18 = SD_HH_CHILD18,
    hh_child17 = SD_HH_TEENS,
    hh_child14 = SD_HH_CHILD14,
    hh_child5 = SD_HH_CHILD5,
    hh_child2 = SD_HH_CHILD2,
    hh_oldfam = SD_HH_OLDERFAM,
    hh_outfam = SD_HH_OUTERFAM,
    hh_partner = SD_HH_PARTNER,
    # hh_income = SD_HHINCOME_FEB2020,
    home_sqm = SD_HOME_SQM,
    home_owner = SD_HOMEOWNER,
    # work_h = SD_WORKHOURS_FEB2020,
    health = all_of(health),
    life_sat = all_of(life_sat),
    aff_neg_1 = all_of(aff_neg_1),
    aff_neg_2 = all_of(aff_neg_2),
    aff_neg_3 = all_of(aff_neg_3),
    aff_neg_4 = all_of(aff_neg_4),
    aff_neg_5 = all_of(aff_neg_5),
    aff_neg_6 = all_of(aff_neg_6),
    aff_pos_1 = all_of(aff_pos_1),
    aff_pos_2 = all_of(aff_pos_2),
    aff_pos_3 = all_of(aff_pos_3),
    act_wrk = all_of(act_wrk),
    act_spo = all_of(act_spo),
    act_frn = all_of(act_frn),
    act_sho = all_of(act_sho),
    act_pet = all_of(act_pet),
    sat_dem = all_of(sat_dem),
    sat_dem = all_of(sat_dem),
    work_h = all_of(work_h),
    work_homeoff = all_of(work_homeoff),
    hh_income = all_of(hh_income),
    med_txt_kro = all_of(med_txt_kro),
    med_txt_sta = all_of(med_txt_sta),
    med_txt_pre = all_of(med_txt_pre),
    med_txt_oes = all_of(med_txt_oes),
    med_txt_kur = all_of(med_txt_kur),
    med_txt_slz = all_of(med_txt_slz),
    med_txt_son = all_of(med_txt_son),
    med_vid_orf = all_of(med_vid_orf),
    med_vid_pri = all_of(med_vid_pri),
    soc_med_fb = all_of(soc_med_fb),
    soc_med_tw = all_of(soc_med_tw),
    soc_med_ig = all_of(soc_med_ig),
    soc_med_yt = all_of(soc_med_yt),
    soc_med_wa = all_of(soc_med_wa),
    soc_med_like_share = all_of(soc_med_like_share),
    soc_med_read = all_of(soc_med_read),
    soc_med_post = all_of(soc_med_post),
    risk_prop = all_of(risk_prop),
    loc_cntrl_int_1 = all_of(loc_cntrl_int_1),
    loc_cntrl_int_2 = all_of(loc_cntrl_int_2),
    loc_cntrl_int_3 = all_of(loc_cntrl_int_3),
    loc_cntrl_int_4 = all_of(loc_cntrl_int_4),
    trst_media = all_of(trst_media),
    trst_police = all_of(trst_police),
    trst_media = all_of(trst_media),
    trst_hlthsec = all_of(trst_hlthsec),
    trst_gov = all_of(trst_gov),
    trst_army = all_of(trst_army),
    corona_pos = all_of(corona_pos),
    wave_part_1 = W1_PANELIST,
    wave_part_8 = W8_PANELIST,
    wave_part_17 = W17_PANELIST,
    wave_part_23 = W23_PANELIST,
    wave_part_28 = W28_PANELIST
  )
```

Make new documentation with selected variables.

```{r make-doc}
doc_selected <- filter(doc_raw, Variable %in% c(
  "RESPID",
  "SD_GENDER",
  "SD_ACCESS_BALCONY",
  "SD_ACCESS_GARDEN",
  "SD_BIRTHYEAR",
  "SD_BORN_AUSTRIA",
  "SD_MIGR_BACKGR",
  "SD_BULA",
  "SD_EDU",
  "SD_EMPLSTATUS_FEB2020",
  "SD_HH_ADULTS",
  "SD_HH_CHILD18",
  "SD_HH_TEENS",
  "SD_HH_CHILD14",
  "SD_HH_CHILD5",
  "SD_HH_CHILD2",
  "SD_HH_OLDERFAM",
  "SD_HH_OUTERFAM",
  "SD_HH_PARTNER",
  "SD_HHINCOME_FEB2020",
  "SD_HOME_SQM",
  "SD_HOMEOWNER",
  "SD_WORKHOURS_FEB2020",
  health,
  life_sat,
  aff_neg_1,
  aff_neg_2,
  aff_neg_3,
  aff_neg_4,
  aff_neg_5,
  aff_neg_6,
  aff_pos_1,
  aff_pos_2,
  aff_pos_3,
  act_wrk,
  act_spo,
  act_frn,
  act_sho,
  act_pet,
  sat_dem,
  work_h,
  work_homeoff,
  hh_income,
  med_txt_kro,
  med_txt_sta,
  med_txt_pre,
  med_txt_oes,
  med_txt_kur,
  med_txt_slz,
  med_txt_son,
  med_vid_orf,
  med_vid_pri,
  soc_med_fb,
  soc_med_tw,
  soc_med_ig,
  soc_med_yt,
  soc_med_wa,
  soc_med_like_share,
  soc_med_post,
  soc_med_read,
  risk_prop,
  loc_cntrl_int_1,
  loc_cntrl_int_2,
  loc_cntrl_int_3,
  loc_cntrl_int_4,
  trst_media,
  trst_police,
  trst_media,
  trst_hlthsec,
  trst_gov,
  trst_army,
  corona_pos
))
write.csv(doc_selected, "data/used_variables.csv")
```

## Recode variables

```{r recode-vars}
d_wide %<>%
  mutate_at(vars(everything(.)), funs(na_if(., 88))) %>%
  mutate_at(vars(everything(.)), funs(na_if(., 99))) %>%
  mutate(
    male = 2 - .$gender,
    age = 2021 - .$year_birth,
    res_vienna = recode(.$county, `8` = 1L, .default = 0L,),
    born_aus = 2 - .$born_aus,
    home_owner = 2 - .$home_owner,
    employment_fac = factor(.$employment, 
                            labels = c("Unemployed",
                                       "Industrie",
                                       "Public service",
                                       "Self-employed",
                                       "Retired",
                                       "Housekeeping",
                                       "Student",
                                       "Incapacitated",
                                       "Parental Leave"),
                            levels = c(4, 1:3, 5:8, 10) # make unemployment reference cat
    ),
    edu_fac = factor(.$edu,
                     labels = c("No degree", 
                                "Middle school",
                                "Vocational school",
                                "Technical school",
                                "High school",
                                "Applied high school",
                                "State college",
                                "Bachelor",
                                "Master",
                                "PhD")
    ))
```

## Make long data format

```{r long-format, cache=T}
d_long <-
  d_wide %>% 
  pivot_longer(
    cols = health...W1:corona_pos...W34,
    names_to = "item",
    values_to = "value"
  ) %>% 
  separate(item, c("item", "wave"), sep = "\\.\\.\\.", extra = "merge") %>% 
  pivot_wider(names_from = "item", values_from = "value")
```

## Recode values

```{r recode-vars-2}
# recode such that higher values imply more strength / align with wording
d_long %<>% 
  mutate_at(
    vars(
      med_txt_kro:med_vid_pri, 
      health, 
      sat_dem, 
      soc_med_fb:soc_med_post
      ), 
    funs(recode(., `1` = 5L, `2` = 4L, `3` = 3L, `4` = 2L, `5` = 1L))) %>% 
  mutate_at(
    vars(
      loc_cntrl_int_1:loc_cntrl_int_4
      ), 
    funs(recode(., `1` = 4L, `2` = 3L, `3` = 2L, `4` = 1L))) %>% 
  mutate_at(
    vars(born_aus_prnts), 
    funs(recode(., `3` = 0L, `2` = 2L, `1` = 1L))
    )

# recode inverted items
d_long %<>%
  mutate_at(vars(loc_cntrl_int_3, loc_cntrl_int_4), 
            funs(recode(., `1` = 4L, `2` = 3L, `3` = 2L, `4` = 1L)))

# recode other
d_long %<>%
  mutate(
    wave = gsub("W", "", .$wave) %>% as.integer(),
    id = as.integer(id)
  )
```

## Order dataset

```{r ordering}
d_long %<>%
  arrange(id, wave)
```

# Impute data

Next we impute data. Note that the actual imputation below is deactivated and loaded from memory to save time.

## Prep

Determine amount of missingness per respondent per wave.

```{r}
vars_used <- c("life_sat", "aff_pos_1", "aff_pos_2", "aff_pos_3", "aff_neg_1", "aff_neg_2", "aff_neg_3", "aff_neg_4", "aff_neg_5", "aff_neg_6", "soc_med_read", "soc_med_like_share", "soc_med_post", "soc_med_fb", "soc_med_ig", "soc_med_wa", "soc_med_yt", "soc_med_tw", "age", "male", "born_aus", "born_aus_prnts", "edu_fac", "employment_fac", "health", "res_vienna", "acc_bal", "acc_gar", "home_sqm", "med_txt_kro", "med_txt_sta", "med_txt_pre", "med_txt_oes", "med_txt_kur", "med_txt_slz", "med_txt_son", "med_vid_orf", "med_vid_pri", "risk_prop", "loc_cntrl_int_1", "loc_cntrl_int_2", "loc_cntrl_int_3", "loc_cntrl_int_4", "act_wrk", "act_spo", "act_frn", "act_sho", "act_pet", "sat_dem", "trst_media", "trst_police", "trst_media", "trst_hlthsec", "trst_gov", "trst_army", "corona_pos") # only include vars measured at _all_ points
```

Filter respondents with more than 50% missing data -- only needed for analyses as originally preregistered (see additional analyses).

```{r eval=T}
d_long_50 <- 
  d_long %>% 
  mutate(na_perc = rowSums(is.na(select(., all_of(vars_used))) / ncol(select(., all_of(vars_used))))) %>% 
  filter(na_perc < .5)
write_csv(d_long_50, "data/data_50.csv")
```

Exclude social media use data, because they were measured only on selected waves.

```{r}
vars_excl <- c("soc_med_read", "soc_med_like_share", "soc_med_post", "soc_med_fb", "soc_med_ig", "soc_med_wa", "soc_med_yt", "soc_med_tw")
incl_ma <- d_long %>% 
  mutate(across(.cols = everything(), .fns = is.na))
incl_ma[vars_excl] <- FALSE

# now also for data where 50% missing excluded
incl_ma_50 <- d_long_50 %>% 
  mutate(across(.cols = everything(), .fns = is.na))
incl_ma_50[vars_excl] <- FALSE
```

## Single imputation

For analyses as originally preregistered (see additional analyses), hence using only participants who provided more than 50% of all data.

```{r imputation-single-50, eval=T, cache=T}
d_long_50_imp <- mice(d_long_50, 
                      method = "pmm", # use predictive mean matching
                      m = 1, maxit = 30, # only 1 imputation
                      # where = incl_ma_50,
                      seed = 180719, print = FALSE)
d_long_50_imp <- mice::complete(d_long_50_imp)
write_csv(d_long_50_imp, "data/data_50_imputed.csv")
```

Because CFAs aren't evaluated with multiple imputation, impute single data-set. Missing data was imputed for everyone.

```{r imputation-single-100, eval=T, cache=T}
d_long_100_imp <- mice(d_long, 
                   method = "pmm", # use predictive mean matching
                   m = 1, # only 1 imputation
                   maxit = 30, 
                   # where = incl_ma,
                   seed = 180719, print = FALSE)
d_long_100_imp <- mice::complete(d_long_100_imp)
write_csv(d_long_100_imp, "data/data_100_imputed.csv")
```

## Multiple Imputation

Now, multiple imputation of data-sets (5 data sets, 20 iterations), used for the final analyses.

```{r imputation-mult, eval=T, cache=T}
# impute missing data with multiple imputation
d_long_100_mim_mice <- mice(d_long, 
                   method = "pmm", # use predictive mean matching
                   m = 20, # 5 imputations, 
                   maxit = 5, # 5 iterations
                   # where = incl_ma,
                   seed = 180719, print = FALSE)
d_long_100_mim <- mice::complete(d_long_100_mim_mice, action = "long", include = TRUE)
write_csv(d_long_100_mim, "data/data_100_mim.csv")
```

## Load data

To increase speed, you can also load the imputed data from the directory

```{r eval=T}
d_long_50_imp <- read_csv("data/data_50_imputed.csv") # data where people with >50% missing data were filtered
d_long_100_imp <- read_csv("data/data_100_imputed.csv") # data where all people were included
d_long_100_mim <- read_csv("data/data_100_mim.csv") # data where all people were included
```

# Create within-between data

First, create means for scales. Then, create mean values for each person (between), and unique changes per wave (within).

```{r within-between-100-mim}
d_long_100_mim <-
  d_long_100_mim %>% 
  mutate(aff_pos_m = rowMeans(select(., aff_pos_1 : aff_pos_3)),
         aff_neg_m = rowMeans(select(., aff_neg_1 : aff_neg_6)),
         loc_cntrl_int_m = rowMeans(select(., loc_cntrl_int_1 : loc_cntrl_int_4)))

d_long_100_mim %<>%
  group_by(id) %>% 
  mutate_at(vars(health:loc_cntrl_int_m), funs(b = mean(., na.rm = TRUE), w = . - mean(., na.rm = TRUE))) %>% 
  ungroup()

d_long_100_mim_mice <- as.mids(d_long_100_mim)
```

```{r within-between-100-imp}
d_long_100_imp <-
  d_long_100_imp %>% 
  mutate(aff_pos_m = rowMeans(select(., aff_pos_1 : aff_pos_3)),
         aff_neg_m = rowMeans(select(., aff_neg_1 : aff_neg_6)),
         loc_cntrl_int_m = rowMeans(select(., loc_cntrl_int_1 : loc_cntrl_int_4)))

d_long_100_imp %<>%
  group_by(id) %>% 
  mutate_at(vars(health:corona_pos, loc_cntrl_int_m), funs(b = mean(., na.rm = TRUE), w = . - mean(., na.rm = TRUE))) %>% 
  ungroup()
```

And now the same for the standardized data set.

```{r within-between-100-imp-std}
d_long_100_mim_std <-
  d_long_100_mim %>% 
  group_by(.imp) %>% 
  mutate_at(vars(gender:corona_pos, -id, -wave, -gender, -male, -born_aus, -born_aus_prnts, -edu_fac, -employment_fac, -res_vienna, -acc_bal, -acc_gar), ~ c(scale(.))) %>% 
  ungroup()

d_long_100_mim_std %<>%
  mutate(aff_pos_m = rowMeans(select(., aff_pos_1 : aff_pos_3)),
         aff_neg_m = rowMeans(select(., aff_neg_1 : aff_neg_6)),
         loc_cntrl_int_m = rowMeans(select(., loc_cntrl_int_1 : loc_cntrl_int_4)))

d_long_100_mim_std %<>%
  group_by(id) %>%
  mutate_at(vars(health:corona_pos, loc_cntrl_int_m), funs(b = mean(., na.rm = TRUE), w = . - mean(., na.rm = TRUE))) %>%
  ungroup()

d_long_100_mim_mice_std <- as.mids(d_long_100_mim_std)
```

And because later we want to rerun the results as originally preregistered, we'll now do the same for analyses without filtered data.

```{r within-between-50-imp}
d_long_50_imp <-
  d_long_50_imp %>% 
  mutate(aff_pos_m = rowMeans(select(., aff_pos_1 : aff_pos_3)),
         aff_neg_m = rowMeans(select(., aff_neg_1 : aff_neg_6)),
         loc_cntrl_int_m = rowMeans(select(., loc_cntrl_int_1 : loc_cntrl_int_4)))

d_long_50_imp %<>% 
  group_by(id) %>% 
  mutate_at(vars(health:corona_pos, loc_cntrl_int_m), 
            funs(b = mean(., na.rm = TRUE), w = . - mean(., na.rm = TRUE))) %>% 
  ungroup()
```

And because later we want to rerun the results without imputed data, now the same for data without imputation.

```{r within-between-50}
d_long_50 %<>% 
  mutate(aff_pos_m = rowMeans(select(., aff_pos_1 : aff_pos_3)),
         aff_neg_m = rowMeans(select(., aff_neg_1 : aff_neg_6)),
         loc_cntrl_int_m = rowMeans(select(., loc_cntrl_int_1 : loc_cntrl_int_4)))

d_long_50 %<>% 
  group_by(id) %>% 
  mutate_at(vars(health:corona_pos, loc_cntrl_int_m), funs(b = mean(., na.rm = TRUE), w = . - mean(., na.rm = TRUE))) %>% 
  ungroup()
```

# Save workspace

```{r cache=F}
save.image("data/workspace_1.RData")
```

### 2023-07-11 index.Rmd
Major updates Implemented feedback

@@
 ---
-title: "The effects of COVID-19 related social media use on well-being"
+title: "Effects of COVID-19 related social media use on well-being"
 ---
 
-On this website, you can find the companion material for the paper _The effects of COVID-19 related social media use on well-being_.
+On this website, you can find the companion material for the paper _Effects of COVID-19 related social media use on well-being_.
 
@@
 - The paper, which you can read [in your browser](manuscript.html) or download as [pdf](manuscript.pdf). 
+- The [set-up](analyses_setup.html) for the analyses, including session information, custom functions, data wrangling, and data imputation.
 - The detailed code and summaries of the [analyses](analyses.html)

### 2023-07-11 manuscript.Rmd
Major updates Implemented feedback

--- 2023-01-30 manuscript.Rmd
@@
 ---
-title             : "COVID-19 related social media use slightly decreases well-being"
-shorttitle        : "COVID-19 related social media use & well-being"
+title             : "Effects of COVID-19 related social media use on well-being"
+shorttitle        : "Effects of COVID-19 related social media use on well-being"
 
@@
 To this end, people increasingly use social media.
-However, being particularly engaging, many find it hard to disconnect from social media, especially during times of crisis.
-Using data from the Austrian Corona Panel Project consisting of 3,485 participants from 32 waves, controlling for several stable and varying confounders, the results showed that COVID-19 related social media use reduced facets of well-being.
-However, the effects were very small.
-Other factors such as health, income, exercise, or internal locus of control showed larger and more meaningful effects."
+However, because social media are particularly engaging, many find it hard to disconnect, especially during times of crisis.
+Using data from the Austrian Corona Panel Project consisting of 3,485 participants from 34 waves, controlling for several stable and varying confounders, the results showed that COVID-19 related social media use did not meaningfully reduce well-being.
+Other factors such as health, income, exercise, or internal locus of control showed larger and meaningful effects."
 keywords          : "COVID-19, well-being, social media, news use, panel study."
@@
 
-header-includes:
-  - |
-    \makeatletter
-    \renewcommand{\paragraph}{\@startsection{paragraph}{4}{\parindent}%
-      {0\baselineskip \@plus 0.2ex \@minus 0.2ex}%
-      {-1em}%
-      {\normalfont\normalsize\bfseries\typesectitle}}
-
-    \renewcommand{\subparagraph}[1]{\@startsection{subparagraph}{5}{1em}%
-      {0\baselineskip \@plus 0.2ex \@minus 0.2ex}%
-      {-\z@\relax}%
-      {\normalfont\normalsize\bfseries\itshape\hspace{\parindent}{#1}\textit{\addperi}}{\relax}}
-    \makeatother
-    \usepackage{endnotes}
-    \setlength{\parskip}{0em}
-    \raggedbottom
-    \note{\clearpage}
-
-csl               : "`r system.file('rmd', 'apa7.csl', package = 'papaja')`"
-documentclass     : "apa7"
-output            :
-  papaja::apa6_pdf:
-    latex_engine  : xelatex
+# header-includes:
+#   - |
+#     \makeatletter
+#     \renewcommand{\paragraph}{\@startsection{paragraph}{4}{\parindent}%
+#       {0\baselineskip \@plus 0.2ex \@minus 0.2ex}%
+#       {-1em}%
+#       {\normalfont\normalsize\bfseries\typesectitle}}
+# 
+#     \renewcommand{\subparagraph}[1]{\@startsection{subparagraph}{5}{1em}%
+#       {0\baselineskip \@plus 0.2ex \@minus 0.2ex}%
+#       {-\z@\relax}%
+#       {\normalfont\normalsize\bfseries\itshape\hspace{\parindent}{#1}\textit{\addperi}}{\relax}}
+#     \makeatother
+#     \usepackage{endnotes}
+#     \setlength{\parskip}{0em}
+#     \raggedbottom
+#     \note{\clearpage}
+#     \usepackage{setspace}
+#     \AtBeginEnvironment{tabular}{\singlespacing}
+#     \AtBeginEnvironment{lltable}{\singlespacing}
+#     \AtBeginEnvironment{tablenotes}{\doublespacing}
+#     \captionsetup[table]{font={stretch=1.5}}
+#     \captionsetup[figure]{font={stretch=1.5}}
+# 
+# csl               : "`r system.file('rmd', 'apa7.csl', package = 'papaja')`"
+# documentclass     : "apa7"
+# output            :
+#   papaja::apa6_pdf:
+#     latex_engine  : xelatex
 # output            : papaja::apa6_docx
-# output            : html_document
+output            : html_document
 ---
 
-```{r analysis-preferences, include=F, cache=F}
+```{r set-up, include=F, cache=F}
 set.seed(170819) # Seed for random number generation
-knitr::opts_chunk$set(cache.extra = knitr::rand_seed,
-                      cache=T, echo=F)
-```
 
-```{r setup, include=F, cache=F}
+knitr::opts_chunk$set(
+  cache.extra = knitr::rand_seed,
+  cache = T, 
+  echo = F, 
+  error = F,
+  cache.lazy = FALSE
+  )
 
@@
 # load packages
-packages <- c("devtools", "english", "GGally", "ggplot2", "gridExtra", "kableExtra", 
-              "lavaan", "lme4", "lmerTest",
-              "magrittr", "papaja", "tidyverse")
-lapply(c(packages, "td"), library, character.only = TRUE)
+library(devtools)
+library(english)
+library(GGally)
+library(ggplot2)
+library(gridExtra)
+library(kableExtra)
+library(lavaan)
+library(lme4)
+library(lmerTest)
+library(magrittr)
+library(papaja)
+library(tidyverse)
+library(td)
+```
 
+```{r load-data, include=F, cache=T}
 # load workspace created from analyses.rmd to use results
-load("data/workspace_2.RData")
+load("data/workspace_3.RData")
 ```
@@
 How is it transmitted? What are the current safety regulations?
-To obtain relevant information, many people heavily relied on social media, which were at an all time high [@statistaAverageDailyTime2021].
-
+To obtain relevant information, many people heavily relied on social media, with use being at an all time high [@statistaAverageDailyTime2021].
 Some actually could not stop using social media to learn about COVID-19 related news.
-A new phenomenon termed "doomscrolling" emerged.
+A new phenomenon termed "doomscrolling" emerged [@sharmaDarkEndTunnel2022].
 Many users were glued to their screens and found it hard to pursue other relevant activities such as working, taking a break, or even looking after their children [@kleinDarklySoothingCompulsion2021].
 In the media it was hence increasingly asked whether using social media for COVID-19 related reasons would, next to all other stressors, create an additional burden on mental health [@sandstromDoomscrollingCOVIDNews2021].
-Although research has begun addressing this question [@bendauAssociationsCOVID19Related2021; @bradleyStressMoodSmartphone2021; @choiMediatedCommunicationMatters2021; @dornemannHowGoodBad2021; @edenMediaCopingCOVID192020; @guazziniSecondWaveAnalysis2022; @latikkaLonelinessPsychologicalDistress2022; @liuRelationOfficialWhatsAppdistributed2020; @riehmAssociationsMediaExposure2020; @sewallObjectivelyMeasuredDigital2021; @stainbackCOVID1924News2020; @yuePassiveSocialMedia2022], it still largely unknown if COVID-19 related social media use during the pandemic has had a meaningful impact on well-being.
-By analyzing large-scale panel data from Austria and by focusing on intra-individual causal effects, this study aims to answer to this question.
+Although research has begun addressing this question
+[e.g., @bendauAssociationsCOVID19Related2021; @edenMediaCopingCOVID192020; @sewallObjectivelyMeasuredDigital2021],
+<!-- [@bendauAssociationsCOVID19Related2021; @bradleyStressMoodSmartphone2021; @choiMediatedCommunicationMatters2021; @dornemannHowGoodBad2021; @edenMediaCopingCOVID192020; @guazziniSecondWaveAnalysis2022; @latikkaLonelinessPsychologicalDistress2022; @liuRelationOfficialWhatsAppdistributed2020; @riehmAssociationsMediaExposure2020; @sewallObjectivelyMeasuredDigital2021; @stainbackCOVID1924News2020; @yuePassiveSocialMedia2022] -->
+it still largely unknown if COVID-19 related social media use during the pandemic has had a meaningful impact on well-being.
+This study hence aims to 1) reveal the effect of the different types and channels of social media use on individual well-being, 2) provide generalizable and robust results by analyzing a large-scale longitudinal data-set with 34 waves, and 3) determine the within-person causal effects by analyzing how changes in social media use lead to changes in well-being.
 
@@
 Because effects of social media depend on how they are used [@verduynPassiveFacebookUsage2015], I further distinguish three types of use and five popular channels. 
-The types of use include reading, posting, and liking and sharing COVID-19 related posts.
-In doing so, this study analyzes social media use focused on COVID-19 related content, which includes posting thoughts about the pandemic, reading posts and comments, or retweeting COVID-19 related news.
+The types of use include reading, liking and sharing, and posting COVID-19 related content.
+In doing so, this study analyzes social media use focused on COVID-19 related content, which includes posting thoughts about the pandemic, reading posts and comments, or retweeting and liking COVID-19 related news.
+Liking and sharing are combined as they both represent low-threshold, platform-ingrained, easily quantifiable interactions.
 The five channels to be investigated are Facebook, Twitter, Instagram, WhatsApp, and YouTube, which at the time ranked among the most popular social media services in Austria.
 
-<!-- ## Theorizing Social Media Effects on Well-Being -->
+## Social Media Effects on Well-Being
 
@@
 
-How could the various types and channels of COVID-19 related social media use affect well-being?
 <!-- In what follows, I outline potential arguments as to why the effect might be positive or negative, direct or indirect, or nonexistent. -->
-According to the set-point theory, well-being is surprisingly stable [@lykkenHappinessWhatStudies1999].
-Although specific events such as marriage or salary can have significant impacts on well-being, in most cases effects are only short-term, with well-being after some time routinely returning to prior levels [@sheldonStabilityHappinessTheories2014].
+How easily can well-being be affected by external influences?
+In general, according to the set-point theory, well-being is surprisingly stable [@lykkenHappinessWhatStudies1999].
+Although specific events such as marriage or salary increase can have significant impacts on well-being, in most cases effects are only short-term, with well-being after some time returning to prior levels [@sheldonStabilityHappinessTheories2014].
 Only very specific factors such as unemployment, disability, or death can cause long-term changes in well-being [@lucasAdaptationSetpointModel2007].
-<!-- So although well-being can change this does not happen easily. -->
+So although well-being can be affected by external events and factors, this does not happen easily.
 <!-- Although the set-point theory has been criticized lately because events such as marriage, unemployment, or death can indeed lead to long-term changes, it is still widely maintained that well-being is very stable and hence hard to be influenced  -->
-Can media use be such a factor?
-Current literature overviews suggest no: 
-Although social media use on average decreases well-being, the effects are small [@meierComputermediatedCommunicationSocial2020a].
-In addition, the effects of media use differ across individuals and types of content [@valkenburgDifferentialSusceptibilityMedia2013].
+
+Can social media use be such a factor?
+Current literature overviews suggest that social media use on average does seem to decrease well-being [@meierComputermediatedCommunicationSocial2020a].
+However, for most well-being outcomes, such as life satisfaction, general well-being, or loneliness, the effects are small [@meierComputermediatedCommunicationSocial2020a].
+These small findings can be explained with the differential susceptibility of media effects model [@valkenburgDifferentialSusceptibilityMedia2013], which states that there is substantial _variation_ of media effects for individual users.
 Whereas for some users social media are more beneficial, for others they are more harmful.
+On average, however, and this is central for this study here, effects tend to be small [@valkenburgDifferentialSusceptibilityMedia2013].
 For example, in one study it was estimated that roughly one quarter of all users experienced negative effects, another quarter positive effects, while for the rest the effects were neutral [@beyensSocialMediaUse2021].
-This finding is aligned with the Differential Susceptibility to Media Effects Model: 
-<!-- This finding is aligned with the Differential Susceptibility to Media Effects Model:  -->
-Although there is substantial _variation_ of media effects for individual users, there are both negative and positive aspects, leading to small effects on average [@valkenburgDifferentialSusceptibilityMedia2013]---as confirmed by several literature overviews [@dienlinImpactDigitalTechnology2020; @huangTimeSpentSocial2017; @meierComputermediatedCommunicationSocial2020a; @orbenTeenagersScreensSocial2020].
+Whether or not effects are positive or negative depend on (a) dispositional factors (e.g., personality, temperament, gender), (b) developmental factors (e.g., age, developmental tasks), (c) and social factors (e.g., environment, norms, upbringing).
+Finally, effects depend also on the content that is consumed.
+If the content is aligned with dispositions, developmental capacities, and converging contexts, effects tend to be stronger [@valkenburgDifferentialSusceptibilityMedia2013].
 
 Why are the effects of social media use on well-being small on average?
-Using Social media is ambivalent.
-It can impair well-being when causing embarrassment, stress, or disinformation, and it can improve well-being when providing connectedness, information, or entertainment [@buchiDigitalWellbeingTheory2021].
-Although social media are often associated with negative outcomes, there are a large number of positive effects, which either offset or at least significantly reduce negative outcomes.
-Positive outcomes encompass finding relevant information; maintaining and fostering relationships; expressing one's personality; or entertaining oneself [e.g., @pelletierOneSizeDoesn2020].
-Negative effects include distraction; displacement of other meaningful activities; consumption of shallow, misleading or challenging content; or negative social comparisons [e.g., @meierDoesPassiveSocial2022].
-People are much more likely to use media that offer plentiful benefits [@katzUsesGratificationsResearch1973].
-Users implicitly learn what media help them regulate their mood and thereby well-being [@zillmannMoodManagementCommunication1988].
-Given the ubiquitous use of social media, the many benefits they offer, plus the general and implicit competence to use media that foster mood management, it is hence unlikely that the effects of social media use on well-being are profusely negative.
-
-<!-- Two prominent media effect theories argue (mostly implicitly) against strong average negative effects. -->
-<!-- According to mood management theory [@zillmannMoodManagementCommunication1988], using media can substantially affect people's moods. -->
-<!-- Use can be stimulating or overwhelming, relaxing or boring. -->
-<!-- After some time, users implicitly learn what media help them balance their mood and affect according to their own situational needs [@zillmannMoodManagementCommunication1988]. -->
-<!-- Those media that eventually become part of one's media repertoire are hence, on average, beneficial for users to regulate their mood. -->
-<!-- Using experience sampling of well-being and logs of social media use, a study with 82 participants from Italy found that after episodes of social media use, levels of positive affect increased significantly [@marcianoDynamicsAdolescentsSmartphone2022].
+Two prominent media effect theories argue implicitly against strong average negative effects.
+First, according to mood management theory [@zillmannMoodManagementCommunication1988], using media can affect people's moods.
+Use can be stimulating or overwhelming, relaxing or boring.
+After some time, users implicitly learn which media help them balance their mood and affect according to their own situational needs [@zillmannMoodManagementCommunication1988].
+Those media that eventually become part of one's media repertoire hence, on average, tend to be beneficial for users to regulate their mood [@marcianoDynamicsAdolescentsSmartphone2022]. 
+In conclusion, if a certain medium is used frequently, mood-management theory argues that it is likely not detrimental for well-being.
+<!-- For example, using experience sampling of well-being and logs of social media use, a study with 82 participants from Italy found that after episodes of social media use, levels of positive affect increased significantly [@marcianoDynamicsAdolescentsSmartphone2022].
+-->
+On the other hand, although mood management theory suggests that effects should not be overly negative, it could be that although short-term effects are positive, long-term effects are negative. 
+<!-- it is still possible that people misjudge media effects. -->
+<!-- Users are often overly optimistic [@metzgerComparativeOptimismPrivacy2017], focusing on short term benefits while ignoring long-term costs.  -->
+Precisely because social media have so many positive consequences in the short run, this might cause problem in the long run---for example, because social media use displaces other potentially more meaningful and/or relaxing activities [@hallSocialMediaUse2022].
+<!-- In other words, social media might not problematic because they are inherently bad, but rather because they are too good. -->
+As with many other things, there can be too much of a good thing.
+It is therefore often asked whether social media can become addictive, and users sometimes express this fear themselves [@yangCanWatchingOnline2021].
+However, a recently published meta-analysis found that the two most prominent measures of addiction, the Bergen Facebook Addiction Scale and the Bergen Social Media Addiction Scale, have only small relations to well-being [@duradoniWellbeingSocialMedia2020].
+In addition, the general idea of labeling excessive social and new media use as addiction was criticized, arguing that social media use represent a new regular behavior that should not be pathologized [@galerHowMuchToo2018; @vanrooijWeakScientificBasis2018].
+<!-- In conclusion, there are several plausible theoretical mechanisms as to why effects might be negative or positive, explaining why on average effects might not be particularly negative or positive.
 -->
 
-<!-- While mood management theory considers media use mainly driven by implicit learning experiences, uses and gratifications theory upholds that the process is more explicit and rational [@katzUsesGratificationsResearch1973].
+Second, while mood management theory considers media use mainly driven by implicit learning experiences, uses and gratifications theory upholds that the process is more explicit and rational [@katzUsesGratificationsResearch1973].
+Users select those media that they expect to have a desired effect, for example on mood, knowledge, or entertainment.
+If those beneficial media effects are missing, people will spend their time elsewhere.
+And social media, in general, offer several beneficial effects, explaining why they are used that much.
+They help find relevant information, maintain and foster relationships, express one's personality, and entertain oneself [@pelletierOneSizeDoesn2020].
+In conclusion, because people spend so much time on social media consuming COVID-19 related content, according to both mood management theory and uses and gratifications theory this indirectly suggest that average effects on well-being are likely not particularly negative.
+
+## Social Media During COVID-19
+
+If we look at COVID-19 related use more specifically, how could the various types and channels of COVID-19 related social media use affect well-being?
+Several uses and gratifications exist, which help explain why people used social media frequently during the pandemic.
+<!-- , outlining mechanisms of potential positive effects.  -->
+<!-- At the same time, there are also several plausible reasons as to why COVID-19 related social media use is beneficial for well-being. -->
+Access to information and resources: 
+Despite incorrect information, social media provide a vast platform for disseminating accurate and timely information about COVID-19 [@johnhopkinsuniversityCOVID19Map2023]. 
+It allows individuals to stay informed about the latest updates, guidelines, and recommendations from reputable health organizations and experts. 
+Access to reliable information can help people make informed decisions, alleviate uncertainties, and feel empowered during the pandemic.
+Educational opportunities: 
+Social media can be a valuable educational tool, with various organizations and experts sharing informative content about COVID-19 [@worldhealthorganizationCoronavirusDiseaseCOVID192023]. 
+This can include explanatory videos, infographics, and articles that help individuals understand the virus, its transmission, prevention strategies, and vaccination information. 
+Access to educational resources on social media can empower individuals to make informed decisions regarding their health.
+Community support and solidarity: 
+Social media platforms enable individuals to connect with others who are experiencing similar challenges during the pandemic [@guazziniSecondWaveAnalysis2022]. 
+Sharing experiences, offering support, and expressing solidarity can foster a sense of community and reduce feelings of isolation. 
+Engaging in online communities and support groups can provide emotional support and create a network of like-minded individuals.
+Promoting mental health and well-being: 
+Many mental health organizations and professionals utilize social media to share tips, strategies, and resources for maintaining mental well-being during the pandemic [@twitterTalkingMentalHealthAwarenessTwitter2020].
+They provide guidance on coping mechanisms, stress reduction techniques, and self-care practices. 
+Engaging with such content might help individuals prioritize their mental health and develop resilience during challenging times.
+Encouraging positive behaviors: 
+Social media campaigns and initiatives can promote positive behaviors related to COVID-19, such as mask-wearing, physical distancing, hand hygiene, and vaccination [@atheyDigitalPublicHealth2023; @huntSocialMediaBased2022].
+Public health organizations and influencers leverage the power of social media to spread awareness and encourage responsible actions, contributing to public health efforts and fostering a sense of collective responsibility.
+<!-- Amplifying diverse voices and perspectives:  -->
+<!-- Social media platforms provide a space for diverse voices and perspectives to be heard [@brunsAreFilterBubbles2019].  -->
+<!-- This allows individuals to access a range of viewpoints, experiences, and cultural practices related to COVID-19.  -->
+<!-- Exposure to diverse perspectives might promote empathy, understanding, and a more comprehensive understanding of the pandemic's impact on different communities. -->
+
+On the other hand, the effects might be negative, perhaps best explained by the following five mechanisms.
+Misinformation and rumors: 
+Social media platforms can easily spread false or misleading information about COVID-19 [@liYouTubeSourceInformation2020]. 
+Due to the ease of sharing and the lack of fact-checking, inaccurate information can go viral and might cause confusion, anxiety, and panic among users.
+Overwhelm and anxiety: 
+Constant exposure to COVID-19-related content on social media can lead to information overload and contribute to heightened anxiety levels [@fanInformationOverloadWellbeing2021]. 
+The rapid spread of news, updates, and opinions can be overwhelming and might exacerbate existing stress or fears about the pandemic [@sharmaDarkEndTunnel2022].
+Confirmation bias and echo chambers: 
+Social media algorithms are designed to show users content that aligns with their interests and beliefs, potentially contributing to echo chambers [@cinelliEchoChamberEffect2021]. 
+This might lead to confirmation bias, where individuals are more likely to be exposed to information that confirms their preexisting beliefs. 
+In the context of COVID-19, this might reinforce false or misleading information and prevent individuals from considering alternative perspectives.
+Cyberbullying and harassment: 
+Social media platforms are known for fostering negativity, with users sometimes engaging in cyberbullying and harassment [@giumettiCyberbullyingSocialMedia2022]. 
+Discussions around COVID-19 can become heated and polarized, leading to personal attacks and online conflicts. 
+Such experiences threaten mental well-being and might contribute to feelings of distress and isolation.
+Social comparison and fear of missing out (FOMO): 
+Social media often showcase the highlights and accomplishments of others, encouraging social comparison [@przybylskiMotivationalEmotionalBehavioral2013]. 
+During a pandemic, seeing posts about others' successes or seemingly perfect lives might intensify feelings of inadequacy or FOMO, especially when individuals are unable to participate in similar activities due to restrictions or personal circumstances [@sharmaDarkEndTunnel2022].
+<!-- Reduced face-to-face interaction:  -->
+<!-- Spending excessive time on social media to learn about COVID-19 can lead to a decrease in face-to-face interactions and real-life connections [but see @hallSocialMediaUse2022].  -->
+This might impact mental health, as most humans have an innate need for social support and physical presence, which online interactions might not fully replicate.
+
+<!-- Using Social media is ambivalent. -->
+<!-- It can impair well-being when causing embarrassment, stress, or disinformation, and it can improve well-being when providing connectedness, information, or entertainment [@buchiDigitalWellbeingTheory2021].
 -->
-<!-- Users select those media that they expect to have a desired effect, for example on mood, knowledge, or entertainment.  -->
-<!-- If those beneficial media effects do not exist or if they are not expected, people will spend their time elsewhere. -->
-<!-- And social media offer several beneficial effects, explaining why they are used that much. -->
-<!-- They help find relevant information, maintain and foster relationships, express one's personality, and entertain oneself [@pelletierOneSizeDoesn2020]. -->
-<!-- In conclusion, because people spend so much time on social media consuming COVID-19 related content, according to both mood management theory and uses and gratifications theory this is indirect proof that average effects on well-being are likely not particularly negative.
+<!-- Although social media are often associated with negative outcomes, there are a large number of positive effects, which either offset or at least significantly reduce negative outcomes. -->
+<!-- Positive outcomes encompass finding relevant information; maintaining and fostering relationships; expressing one's personality; or entertaining oneself [e.g., @pelletierOneSizeDoesn2020]. -->
+<!-- Negative effects include distraction; displacement of other meaningful activities; consumption of shallow, misleading or challenging content; or negative social comparisons [e.g., @meierDoesPassiveSocial2022].
+-->
+<!-- People are much more likely to use media that offer plentiful benefits [@katzUsesGratificationsResearch1973]. -->
+<!-- Users implicitly learn what media help them regulate their mood and thereby well-being [@zillmannMoodManagementCommunication1988]. -->
+<!-- Given the ubiquitous use of social media, the many benefits they offer, plus the general and implicit competence to use media that foster mood management, it is hence unlikely that the effects of social media use on well-being are profusely negative.
 -->
 
-<!-- But people can also misjudge media effects and are often overly optimistic [@metzgerComparativeOptimismPrivacy2017]. -->
-<!-- Precisely because social media have so many positive consequences, one can ask if this is not where the actual problem lies. -->
-<!-- In other words, social media might not problematic because they are inherently bad, but rather because they are too good. -->
-<!-- As with many other things, there can be too much of a good thing. -->
-<!-- It is therefore often asked whether social media can become addictive, and users sometimes express this fear themselves [@yangCanWatchingOnline2021].  -->
-<!-- However, a recently published meta-analysis found that the two most prominent measures of addiction, the Bergen Facebook Addiction Scale and the Bergen Social Media Addiction Scale, have only small relations to well-being [@duradoniWellbeingSocialMedia2020].
+<!-- As outline above,   -->
+There is still little empirical research on how well-being is affected by social media use that is focused on COVID-19 specifically.
+Echoing the theoretical rationales outlines above, studies have yielded mixed results.
+Some studies found negative effects, indicating that excessive social media use for COVID-19 news led to compulsive behavior and increased stress levels, particularly due to upward social comparison [@kleinDarklySoothingCompulsion2021; @stainbackCOVID1924News2020; @yuePassiveSocialMedia2022].
+Individuals who relied on social media as their primary information source reported higher levels of anxiety and depression symptoms [@bendauAssociationsCOVID19Related2021]. 
+Similarly, increased COVID-19-related media consumption was associated with higher psychological distress.
+On the other hand, some studies reported positive outcomes. 
+Certain individuals experienced increased virtual community and social connectedness during the pandemic through social media, which contributed to their well-being [@guazziniSecondWaveAnalysis2022]. 
+Additionally, engaging more on social media was associated with reduced feelings of loneliness [@latikkaLonelinessPsychologicalDistress2022].
+A couple of studies reported mostly neutral effects of social media use on well-being indicators [@bradleyStressMoodSmartphone2021; @edenMediaCopingCOVID192020; @sewallObjectivelyMeasuredDigital2021].
+Overall, the literature demonstrates a mixed picture, highlighting both positive and negative effects of social media use focused on or during COVID-19 on well-being [see also @bradleyStressMoodSmartphone2021; @dornemannHowGoodBad2021; @liuRelationOfficialWhatsAppdistributed2020; @riehmAssociationsMediaExposure2020; @sewallObjectivelyMeasuredDigital2021].
+
+<!-- It is important to consider individual differences, usage patterns, and other factors that may influence the impact of social media on well-being during this challenging period. -->
+
+<!-- ## Empirical Studies on Social Media Effects -->
+
+<!-- There is still little empirical research on how well-being is affected by social media use that is focused on COVID-19 specifically. -->
+<!-- Results are mixed. -->
+<!-- When browsing social media for COVID-19 related news, many users reported being captivated to such an extent they could not stop using social media [@kleinDarklySoothingCompulsion2021]. -->
+<!-- During the pandemic social media use was at an all-time high in the US [@statistaAverageDailyTime2021], for example to cope with the required physical distancing measures. -->
+<!-- In a study with 1,131 residents from Wuhan in China, people who spent more time in quarantine also spent more time on social media [@yuePassiveSocialMedia2022].  -->
+<!-- Those who spent more time on social media also engaged in more upward social comparison, which was related to increased levels of stress. -->
+<!-- People who used social media as a primary source of information reported on average "significantly more unspecific anxiety and depression [...] and significantly more specific COVID-19 related anxiety symptoms" [@bendauAssociationsCOVID19Related2021, p.
+288].
 -->
-<!-- In addition, the general idea of labeling excessive social and new media use as addiction was criticized, arguing that social media represent new regular behaviors that should not be pathologized [@galerHowMuchToo2018; @vanrooijWeakScientificBasis2018].
+<!-- @edenMediaCopingCOVID192020 analyzed how 425 US college students used media during the first wave of the pandemic, finding both positive and negative relations with well-being. -->
+<!-- In a sample of 312 respondents collected via Amazon Mechanical Turk, people who used media to attain information were more lonely and less satisfied with their lives [@choiMediatedCommunicationMatters2021].
 -->
+<!-- In a large-scale study with 11,537 respondents from the US, increased COVID-19-related media consumption was related to more psychological distress [@stainbackCOVID1924News2020]. -->
+<!-- A four-wave panel study with 384 young adults from the U.S.
+analyzed the effects of general digital technology use during the pandemic on mental health, finding that digital technology did not have significant effects on mental health [@sewallObjectivelyMeasuredDigital2021; for a similar study with comparable results, see @bradleyStressMoodSmartphone2021].
+-->
+<!-- A study with 2.057 respondents from Italy reported that during the pandemic virtual community and social connectedness increased [@guazziniSecondWaveAnalysis2022]. -->
+<!-- In a study with 735 participants from Finland, levels of loneliness were stable during the pandemic, and people who engaged more on social media experienced less loneliness [@latikkaLonelinessPsychologicalDistress2022].
+-->
+<!-- Together, the literature is mixed, showing both positive and negative affects of social media use focused on or during COVID-19 on well-being [see also @dornemannHowGoodBad2021; @liuRelationOfficialWhatsAppdistributed2020; @riehmAssociationsMediaExposure2020].
+-->
+
+In conclusion, given these mixed empirical results, together with the observation that social media effects on well-being are very small in general, and that several plausible theoretical mechanisms exist for both positive and negative effects, I expect that COVID-19 related communication on social media should not be decidedly positive or negative.
+It seems most likely that both positive and negative coexist, but that on average using social media for COVID-19 related reasons should not have substantial effects on well-being.
+
+> Hypothesis: The within-person effects of all measures of COVID-19 related social media use (types: reading, liking and sharing, posting; channels: Twitter, Instagram, Facebook, YouTube, WhatsApp) on all measures of well-being indicators (positive affect, negative affect, life satisfaction)---while controlling for several stable and varying covariates such as sociodemographic variables and psychological dispositions (see below)---will be trivial.
+
+<!-- Although the average effect of use on well-being is likely very small, it might be that specific subgroups exhibit larger effects. -->
+
+<!-- > Research question: How do the effects of COVID-19 related social media use on well-being depend on gender, age, immigration background, education, and urban living context? -->
 
@@
 
-<!-- ## Empirical Studies on Social Media Effects -->
-
-There is still little empirical research on how well-being is affected by social media use that is focused on COVID-19 specifically.
-Results are mixed.
-When browsing social media for COVID-19 related news, many users reported being captivated to such an extent they could not stop using social media [@kleinDarklySoothingCompulsion2021].
-<!-- During the pandemic social media use was at an all-time high in the US [@statistaAverageDailyTime2021], for example to cope with the required physical distancing measures. -->
-In a study with 1,131 residents from Wuhan in China, people who spent more time in quarantine also spent more time on social media [@yuePassiveSocialMedia2022]. 
-Those who spent more time on social media also engaged in more upward social comparison, which was related to increased levels of stress.
-People who used social media as a primary source of information reported on average "significantly more unspecific anxiety and depression [...] and significantly more specific COVID-19 related anxiety symptoms" [@bendauAssociationsCOVID19Related2021, p.
-288].
-@edenMediaCopingCOVID192020 analyzed how 425 US college students used media during the first wave of the pandemic, finding both positive and negative relations with well-being.
-<!-- In a sample of 312 respondents collected via Amazon Mechanical Turk, people who used media to attain information were more lonely and less satisfied with their lives [@choiMediatedCommunicationMatters2021].
--->
-In a large-scale study with 11,537 respondents from the US, increased COVID-19-related media consumption was related to more psychological distress [@stainbackCOVID1924News2020].
-A four-wave panel study with 384 young adults from the U.S.
-analyzed the effects of general digital technology use during the pandemic on mental health, finding that digital technology did not have significant effects on mental health [@sewallObjectivelyMeasuredDigital2021; for a similar study with comparable results, see @bradleyStressMoodSmartphone2021].
-A study with 2.057 respondents from Italy reported that during the pandemic virtual community and social connectedness increased [@guazziniSecondWaveAnalysis2022].
-In a study with 735 participants from Finland, levels of loneliness were stable during the pandemic, and people who engaged more on social media experienced less loneliness [@latikkaLonelinessPsychologicalDistress2022].
-Together, the literature is mixed, showing both positive and negative affects of social media use focused on or during COVID-19 on well-being [see also @dornemannHowGoodBad2021; @liuRelationOfficialWhatsAppdistributed2020; @riehmAssociationsMediaExposure2020].
 <!-- However, note that all of these findings represent between-person relations stemming from cross-sectional data (see below). -->
@@
 
-In conclusion, these mixed empirical results, together with the observation that social media effects on well-being are very small in general, I expect that COVID-19 related communication on social media does not affect well-being in a meaningful or relevant way.
-
 <!-- A meta review (i.e., an analysis of meta-analyses) found that the relation between social media use and well-being is likely in the negative spectrum but very small, potentially too small to matter [@meierComputermediatedCommunicationSocial2020a].
@@
 -->
-
-> Hypothesis: The within-person effects of all types of COVID-19 related social media use on all types of well-being indicators---while controlling for several stable and varying covariates such as sociodemographic variables and psychological dispositions---will be trivial.
 
@@
 The data come from the Austrian Corona Panel Project [@kittelAustrianCoronaPanel2021], which is a large-scale standalone panel study. 
-The data are hosted on AUSSDA, are publicly available here (https://doi.org/10.11587/28KQNS), and consist of 32 waves.
-The study was conducted between March 2020 and June 2022, and data collection is now finished.
-Between March 2020 and July 2020, the intervals between waves were weekly and afterward monthly.
-Each wave consists of at least 1,500 respondents.
-The sample size was _N_ = `r d_raw %>% nrow() %>% prettyNum(big.mark = ",")`, with overall `r d_long %>% nrow() %>% prettyNum(big.mark = ",")` observations.
-Panel mortality was compensated through a continuous re-acquisition of new participants. 
+The data are hosted on AUSSDA, are publicly available here (https://doi.org/10.11587/28KQNS), and consist of 34 waves.
 Participants were sampled from a pre-existing online access panel provided by the company Marketagent, Austria.
 Panel members were incentivized with 180 credit points for each wave of the study.
+The study was conducted between March 2020 and February 2023.
+Between March 2020 and July 2020, the intervals between waves were weekly, until May 2022 (wave 32) monthly, and afterward after 5 months.
+Each wave consists of at least 1,500 respondents.
+Panel mortality was compensated through a continuous re-acquisition of new participants. 
+The sample size was _N_ = `r d_raw %>% nrow() %>% prettyNum(big.mark = ",")`, with overall `r d_long %>% nrow() %>% prettyNum(big.mark = ",")` observations.
+For an overview of the study set-up, see Figure \@ref(fig:study-desc).
+
+```{r study-desc, warning=F, message=F, fig.cap="Overview of study set-up", out.width = "1\\textwidth", warning=F, message=F, prompt=F}
+knitr::include_graphics("figures/fig_study_description.png")
+```
 
@@
 <!-- Instead, it should depend on absolute criteria. -->
-<!-- What could be a minimally interesting, nontrivial effect? -->
+To this end, we need to define a so-called smallest effect size of interest (SESOI) [@lakensEquivalenceTestingPsychological2018].
+A trivial effect would then need to be smaller than the SESOI (see below).
+What could be a minimally interesting, nontrivial effect?
 Being a normative question, finding a clear, single, or unanimous answer is impossible.
@@
 The hypothesis is analyzed using the interval testing approach as proposed by @dienesUsingBayesGet2014.
-To illustrate, if the 95% confidence interval falls completely within the null-region (e.g., _b_ = -.02, [95% CI: -.12, .08]), the hypothesis that the effect is trivial is supported.
-If the confidence interval and the null region overlap (e.g., _b_ = -.22, [95% CI: -.27, -.17]), the hypothesis is not supported and the results are considered inconclusive, while a meaningful negative effect is rejected.
-If the confidence interval falls completely outside of the null-region (e.g., _b_ = -.40, [95% CI: -.45, -.35]), the hypothesis is rejected and the existence of a meaningful positive effect is supported.
+To illustrate, let us consider the case of life satisfaction [SESOI: -.30 : +.30]. 
+If the 95% confidence interval falls completely within the null-region (e.g., _b_ = -.05, [95% CI: -.15, .05]), the hypothesis that the effect is trivial is supported.
+If the confidence interval falls completely outside of the null-region (e.g., _b_ = -.40, [95% CI: -.45, -.35]), the hypothesis is rejected and the existence of a meaningful negative effect is supported.
+If the confidence interval and the null region overlap (e.g., _b_ = -.30, [95% CI: -.35, -.25]), the hypothesis is not supported and the results are considered inconclusive, while a meaningful positive effect is rejected.
+If the confidence interval exceeds both sides of the null region (e.g., _b_ = -.025, [95% CI: -.40, .35]), the hypothesis is not supported and judgement is suspended.
 For an illustration, see Figure \@ref(fig:sesoi).
@@
 ```{r sesoi, warning=F, message=F, fig.cap="Using confidence intervals to test a null region.
-In this study, a trivial effect of social media use on life satisfaction is defined as ranging from b = -.30 to b = .30", fig.width=6, fig.height=2, warning=F, message=F, prompt=F}
+In this study, a trivial effect of social media use on life satisfaction is defined as ranging from b = -.30 to b = .30.
+Figure adapted from Dienes (2014).", fig.width=6, fig.height=2, warning=F, message=F, prompt=F}
 # make table with data
@@
   "2. Reject trivial effect", "Decision Rule", -.40, -.45, -.35,
-  "3. Reject positive effect", "Decision Rule", -.22, -.27, -.17,
-  "4. Suspend judgement", "Decision Rule", -.1, -.32, .08,
+  "3. Reject positive effect", "Decision Rule", -.25, -.15, -.35,
+  "4. Suspend judgement", "Decision Rule", -.025, -.40, .35,
 ) %>% 
@@
 <!-- Obviously, we can never be entirely sure to have included all confounders, which makes absolute statements regarding causality virtually impossible. -->
-To elaborate, in non-experimental designs it makes much sense to analyze causal effects from an internal, within-person perspective [@hamakerWhyResearchersShould2014].
+First, in non-experimental designs it makes much sense to analyze causal effects from an internal, within-person perspective [@hamakerWhyResearchersShould2014].
 If a specific person changes their media diet, we need to measure how this behavior affects their well-being.
@@
 
-Second, to identify confounders we should control for variables that affect both media use and well-being, which helps isolate the actual effect of social media use on well-being [@rohrerThinkingClearlyCorrelations2018].
+Second, to identify confounders we should control for variables that affect both media use and well-being, which helps isolate the actual effect [@rohrerThinkingClearlyCorrelations2018].
 Because we are adopting a within-person perspective, we need to implement _time-varying_ confounders [@rohrerTheseAreNot2021]. 
@@
 -->
-In this study, I hence preregistered to control for the following variables, which either have already been shown or are very likely to affect both social media use and well-being, and which also are not mediators:
+In this study, I hence preregistered to control for the following variables, which either have already been shown or are likely to affect both social media use and well-being, and which also are not mediators:
 <!-- (I'll additionally include variables that likely affect only well-being, also to obtain a comparison benchmark for social media effects): -->
-gender, age, education, Austria country of birth, Austria country of birth of parents, residency Vienna, text-based news consumption, video-based news consumption, household size, health, living space, access to garden, access to balcony, employment, work hours per week, being in home-office, household income, outdoor activities, satisfaction with democracy, disposition to take risks, and locus of control.
+gender, age, education, Austria country of birth, Austria country of birth of parents, residency Vienna, text-based news consumption, video-based news consumption, household size, health, living space, access to garden, access to balcony, employment, work hours per week, being in home-office, household income, outdoor activities, disposition to take risks, and locus of control [@egerStatisticalMetaanalysisWellbeing2015; @wardOriginsHappinessEvidence2016].
 <!-- [^explore] -->
@@
 <!-- Finding the right interval is difficult. -->
-If variables are stable, longer intervals are needed; if they fluctuate, shorter intervals are required.
+If variables are stable, longer intervals are needed; if they fluctuate, shorter intervals.
 In the case of well-being, we need shorter intervals for the more fluctuating positive and negative affect, and longer ones for the more stable life satisfaction [@dienlinImpactDigitalTechnology2020].
-Using social media can have instant effects on mood [@marcianoDynamicsAdolescentsSmartphone2022].
-Effects on life satisfaction often take longer to manifest, for example because media use leads to actual changes in specific behaviors, which then in turn affect life satisfaction [@dienlinDisplacementReinforcementReciprocity2017].
+Whereas using social media can have instant effects on mood [@marcianoDynamicsAdolescentsSmartphone2022], effects on life satisfaction often take longer to manifest.
+For example, because media use leads to actual changes in specific behaviors, which then in turn affect life satisfaction [@dienlinDisplacementReinforcementReciprocity2017].
 <!-- Choosing the right interval is challenging, because especially short intervals are hard to implement in practice, often requiring advanced methods such as experience sampling (also known as in situ measurement or ambulant assessment) [@schnauber-stockmannMobileDevicesTools2020].
 -->
-In this study, I hence analyze how using social media during the last week affected positive and negative affect during the same week.
+
+In this study, I hence analyze how changes in using social media _during the last week_ affected changes in positive and negative affect _during the same week_.
 In other words, if people during the last week engaged in more COVID-19 related social media use than usual, did they feel better or worse during that week than usual?
 For life satisfaction, I implemented a longer interval.
-If people during the last week used COVID-19 related social media more than they usually do, were they at the end of the week more or less satisfied with their lives than they usually are?
-I hence analyze if when a person changes their social media diet, are there (a) _simultaneous_ changes in their affect and (b) _subsequent_ changes in their life satisfaction?
-These relations will be controlled for varying confounders, which fosters a causal interpretation.
+If people _during the last week_ used COVID-19 related social media more than they usually do, were they _at the end of the week_ more or less satisfied with their lives than they usually are?
+This way it is analyzed if when a person changes their social media diet, are there (a) _simultaneous_ changes in their affect and (b) _subsequent_ changes in their life satisfaction?
+For the main analyses, the interval is implemented via the wording of the items (see below), not by using lagged measures coming from prior waves waves. 
+In additional analyses, I also tested how media use affects well-being one month or four months later.
+<!-- This significantly increases sample size, such that all respondents who answered the media measures at least two times can be implemented. -->
+All analyses will be controlled for varying confounders (see below), which fosters a causal interpretation.
 <!-- Similar approaches were implemented by other studies [@johannesNoEffectDifferent2022; @scharkowHowSocialNetwork2020], and they are considered a best practice approach toward analyzing causality [@bellFixedRandomEffects2019].
@@
 The data were hierarchical, and responses were separately nested in participants and waves (i.e., participants and waves were implemented as random effects).
-Nesting in participants allowed to separate between-person relations from within-person effects.
-Nesting in waves allowed to control for general exogenous developments, such as general decreases in well-being in the population, for example due to lockdown measures.
+Nesting in participants accounts for the longitudinal design. 
+Nesting in waves controls for general exogenous developments, such as general decreases in well-being in the population, for example due to lockdown measures.
 Thus, there was no need additionally to control for specific phases or measures of the lockdown.
@@
 They included social media communication types and channels, separated into within and between-person factors, as well as stable and varying covariates.
+Between-person predictors are the predictors centered on the grand mean; within-person predictors are the predictors centered on the person's mean.
+Between-person predictors (which, measuring relations, are not of particular interest in this study) represent how the mean of one respondent differs from the mean of all the other respondents.
+The within-person predictors represent how much a person at one specific wave differs from their own mean.
+For example, we could find that on Wave 3 a person used social media more than usual, while also experiencing more negative affect than usual.
 All predictors were included simultaneously in each of the three models.
@@
 <!-- Respondents who answered less than 50% of all questions were removed. -->
-Missing responses were imputed using multiple imputation with predictive mean matching (five iterations, five data-sets), including categorical variables.
+Missing responses were imputed using multiple imputation with predictive mean matching (five iterations, 30 data-sets), including categorical variables.
 All variables were imputed except the social media use measures, as they were not collected on each wave.
 All variables included in the analyses presented here were used to impute missing data.
-<!-- For the main analyses, results were pooled across all five data-sets. -->
+For the main analyses, results were pooled across all thirty data-sets.
 
-### Robustness-checks
-
-To find out whether the inferences were robust across plausible but subordinate analyses, I conducted additional exploratory analyses. 
+To contextualize the results, I conducted additional exploratory analyses. 
 I reran the analyses (a) with additional not-preregistered covariates such as trust in media or government, (b) without covariates, (c) with single imputation, and (d) without imputation.
@@
 
-Social media use was measured for all participants on waves 1, 2, 8, 17, 23, and 28.
+Social media use was measured for all participants on waves 1, 2, 8, 17, 23, and 28 (see Figure 1).
 Freshly recruited respondents always answered all questions on COVID 19-related social media use.
-For the analyses I hence used all 32 waves.
+Because new respondents always provided data on media use, it was possible to include these data into the analyses. 
+Hence, for the main analyses data from all 34 waves were used.
+In the additional analyses I tested longer intervals, namely if changes in social media use were associated with changes in well-being either one month of four months later.
+For these analyzes I used the predictors from waves 1, 2, 8, 17, 23, and 28, to see if they predicted changes in well-being either one month or four months later.
 
@@
           caption = "Descriptives of the main variables."
+          , font_size = "small"
           )
@@
 <!-- I originally planned to implement additional variables as varying covariates. -->
-I also controlled for the following varying covariates: five items on current living conditions, including self-reported physical health, whether participants contracted COVID-19 since the last wave, current household income, working in home-office, and overall work hours; nine items measuring use of specific national text-based and video-based news outlets; five items measuring outdoor activities such as exercise or meeting friends; and three more psychological measures including locus of control, disposition to take risks, and satisfaction with democracy.
+I also controlled for the following varying covariates: five items on current living conditions, including self-reported physical health, whether participants contracted COVID-19 since the last wave, current household income, working in home-office, and overall work hours; nine items measuring use of specific national text-based and video-based news outlets; five items measuring outdoor activities such as exercise or meeting friends; and two more psychological measures including locus of control and disposition to take risks.
 
@@
 Posting about COVID-19 related content stayed the same.
-Using Facebook, WhatsApp, and YouTube for COVID-19 related content decreased. 
-Instagram stayed the same.
-Twitter use slightly increased.
+Using Facebook and WhatsApp for COVID-19 related content decreased. 
+Instagram, YouTube, and Twitter stayed the same.
 The general initial decrease could be explained by the fact that the collection of data began at the end of March 2020, hence approximately three months after the pandemic's onset.
@@
 
-```{r fig-descriptives, fig.cap="Well-being and media use across the 32 waves.
+```{r fig-descriptives, fig.cap="Well-being and media use across the 34 waves.
 Note.
@@
 Upper triangle: correlation coefficients; diagonal: density plots; lower triangle: scatter plots.", out.width = "\\textwidth", cache=F, warning=F, message=F, prompt=F}
-# fig_cor
 knitr::include_graphics("figures/fig_cor.png")
@@
 
-However, three effects stood out, as they were significantly different from zero.
+However, several effects stood out, as statistically they were significantly different from zero.
 Users who read more COVID-19 related content than usual reported slightly reduced levels of positive affect (_b_ = `r data_tab_within %>% filter(dv == "Positive affect" & iv == "Reading" & Variance == "within" & Analysis == "Publication") %>% select(estimate) %>% round(2)` [95% CI `r data_tab_within %>% filter(dv == "Positive affect" & iv == "Reading" & Variance == "within" & Analysis == "Publication") %>% select(conf.low) %>% round(2)`, `r data_tab_within %>% filter(dv == "Positive affect" & iv == "Reading" & Variance == "within" & Analysis == "Publication") %>% select(conf.high) %>% round(2)`]).
 Users who liked and shared more COVID-19 related content than usual also experienced slightly more negative affect than usual (_b_ = `r data_tab_within %>% filter(dv == "Negative affect" & iv == "Posting" & Variance == "within" & Analysis == "Publication") %>% select(estimate) %>% round(2)` [95% CI `r data_tab_within %>% filter(dv == "Negative affect" & iv == "Posting" & Variance == "within" & Analysis == "Publication") %>% select(conf.low) %>% round(2)`, `r data_tab_within %>% filter(dv == "Negative affect" & iv == "Posting" & Variance == "within" & Analysis == "Publication") %>% select(conf.high) %>% round(2)`]).
-Users who wrote more COVID-19 related posts than usual also experienced slightly more negative affect than usual (_b_ = `r data_tab_within %>% filter(dv == "Negative affect" & iv == "Posting" & Variance == "within" & Analysis == "Publication") %>% select(estimate) %>% round(2)` [95% CI `r data_tab_within %>% filter(dv == "Negative affect" & iv == "Posting" & Variance == "within" & Analysis == "Publication") %>% select(conf.low) %>% round(2)`, `r data_tab_within %>% filter(dv == "Negative affect" & iv == "Posting" & Variance == "within" & Analysis == "Publication") %>% select(conf.high) %>% round(2)`]).
+Posting COVID-19 related content affected all types of well-being. 
+Users who wrote more COVID-19 related posts than usual also reported slightly less life satisfaction than usual (_b_ = `r data_tab_within %>% filter(dv == "Life satisfaction" & iv == "Posting" & Variance == "within" & Analysis == "Publication") %>% select(estimate) %>% round(2)` [95% CI `r data_tab_within %>% filter(dv == "Life satisfaction" & iv == "Posting" & Variance == "within" & Analysis == "Publication") %>% select(conf.low) %>% round(2)`, `r data_tab_within %>% filter(dv == "Life satisfaction" & iv == "Posting" & Variance == "within" & Analysis == "Publication") %>% select(conf.high) %>% round(2)`]) and slightly more negative affect than usual (_b_ = `r data_tab_within %>% filter(dv == "Negative affect" & iv == "Posting" & Variance == "within" & Analysis == "Publication") %>% select(estimate) %>% round(2)` [95% CI `r data_tab_within %>% filter(dv == "Negative affect" & iv == "Posting" & Variance == "within" & Analysis == "Publication") %>% select(conf.low) %>% round(2)`, `r data_tab_within %>% filter(dv == "Negative affect" & iv == "Posting" & Variance == "within" & Analysis == "Publication") %>% select(conf.high) %>% round(2)`]).
+Interestingly, however, users who wrote more COVID-19 related posts than usual also experienced slightly _higher_ levels of positive affect than usual (_b_ = `r data_tab_within %>% filter(dv == "Positive affect" & iv == "Posting" & Variance == "within" & Analysis == "Publication") %>% select(estimate) %>% round(2)` [95% CI `r data_tab_within %>% filter(dv == "Positive affect" & iv == "Posting" & Variance == "within" & Analysis == "Publication") %>% select(conf.low) %>% round(2)`, `r data_tab_within %>% filter(dv == "Positive affect" & iv == "Posting" & Variance == "within" & Analysis == "Publication") %>% select(conf.high) %>% round(2)`]).
 
@@
 
-That said, one effect differed substantially from zero.
+That said, two effects differed statistically from zero.
 Respondents who used Twitter more frequently than usual to attain COVID-19 related news reported slightly higher levels of negative affect than usual (_b_ = `r data_tab_within %>% filter(dv == "Negative affect" & iv == "Twitter" & Variance == "within" & Analysis == "Publication") %>% select(estimate) %>% round(2)` [95% CI `r data_tab_within %>% filter(dv == "Negative affect" & iv == "Twitter" & Variance == "within" & Analysis == "Publication") %>% select(conf.low) %>% my_round("b")`, `r data_tab_within %>% filter(dv == "Negative affect" & iv == "Twitter" & Variance == "within" & Analysis == "Publication") %>% select(conf.high) %>% my_round("b")`]).
-However, the effect was still completely inside of the null region, hence likely not large enough to be considered meaningful.
+Likewise, respondents who used YouTube more frequently than usual for COVID-19 related issues reported slightly higher levels of negative affect than usual (_b_ = `r data_tab_within %>% filter(dv == "Negative affect" & iv == "YouTube" & Variance == "within" & Analysis == "Publication") %>% select(estimate) %>% round(2)` [95% CI `r data_tab_within %>% filter(dv == "Negative affect" & iv == "YouTube" & Variance == "within" & Analysis == "Publication") %>% select(conf.low) %>% my_round("b")`, `r data_tab_within %>% filter(dv == "Negative affect" & iv == "YouTube" & Variance == "within" & Analysis == "Publication") %>% select(conf.high) %>% my_round("b")`]).
+However, both effects were still completely inside of the null region, hence likely not large enough to be considered meaningful.
 
@@
 
-```{r fig-within, fig.cap="Within-person effects of COVID-19 related social media use on well-being.
+```{r fig-within, fig.cap="Unstandardized within-person effects of COVID-19 related social media use on well-being.
 Note.
-The black estimates show the effects controlled for a large number of covariates (see text; preregistered); the grey estimates are without control variables (exploratory).
 The SESOI was b = |0.30| for life satisfaction and b = |0.15| for affect.
-Hence, all of the reported effects are not considered meaningful.", out.width = "\\textwidth", cache=F, warning=F, message=F, fig.width=6, fig.height=4}
+Hence, all of the reported effects are not considered large enough to be meaningful.", cache=F, warning=F, message=F, fig.width=6, fig.height=4}
 fig_results_within
@@
           caption = "Overview of all within-person effects.",
-          col_spanners = list(`Confidence interval` = c(3, 4)),
-          font_size = "footnotesize"
+          col_spanners = list(`Confidence interval` = c(3, 4))
+          , font_size = "footnotesize"
           )
@@
 
-<!-- ### Between-person relations -->
-
-<!-- For between-person relations, no a-priori hypotheses were formulated. -->
-<!-- Results showed that no relation crossed or was completely outside of the SESOI. -->
-<!-- Four relations were statistically significant. -->
-<!-- Respondents who across all waves used social media more frequently than others to read about COVID-19 related posts reported slightly lower levels of positive affect than others (_b_ `r dat_fig_results_between %>% filter(dv == "Positive affect" & iv == "Reading" & type == "between" & Analysis == "Regular") %>% select(estimate) %>% my_round("b_txt")` [95% CI `r dat_fig_results_between %>% filter(dv == "Positive affect" & iv == "Reading" & type == "between" & Analysis == "Regular") %>% select(conf.low) %>% my_round("b")`, `r dat_fig_results_between %>% filter(dv == "Positive affect" & iv == "Reading" & type == "between" & Analysis == "Regular") %>% select(conf.high) %>% my_round("b")`]).
--->
-<!-- Respondents who across all waves used social media more frequently than others to write COVID-19 related posts reported slightly higher levels of negative affect than others (_b_ `r dat_fig_results_between %>% filter(dv == "Negative affect" & iv == "Posting" & type == "between" & Analysis == "Regular") %>% select(estimate) %>% my_round("b_txt")` [95% CI `r dat_fig_results_between %>% filter(dv == "Negative affect" & iv == "Posting" & type == "between" & Analysis == "Regular") %>% select(conf.low) %>% my_round("b")`, `r dat_fig_results_between %>% filter(dv == "Negative affect" & iv == "Posting" & type == "between" & Analysis == "Regular") %>% select(conf.high) %>% my_round("b")`]).
--->
-<!-- At the same time, respondents who across all waves used social media more frequently than others to write COVID-19 related posts also reported slightly higher levels of positive affect (_b_ `r dat_fig_results_between %>% filter(dv == "Positive affect" & iv == "Posting" & type == "between" & Analysis == "Regular") %>% select(estimate) %>% my_round("b_txt")` [95% CI `r dat_fig_results_between %>% filter(dv == "Positive affect" & iv == "Posting" & type == "between" & Analysis == "Regular") %>% select(conf.low) %>% my_round("b")`, `r dat_fig_results_between %>% filter(dv == "Positive affect" & iv == "Posting" & type == "between" & Analysis == "Regular") %>% select(conf.high) %>% my_round("b")`]).
--->
-<!-- Finally, respondents who across all waves used YouTube more frequently than others also reported slightly higher levels of life satisfaction than others (_b_ `r dat_fig_results_between %>% filter(dv == "Life satisfaction" & iv == "YouTube" & type == "between" & Analysis == "Regular") %>% select(estimate) %>% my_round("b_txt")` [95% CI `r dat_fig_results_between %>% filter(dv == "Life satisfaction" & iv == "YouTube" & type == "between" & Analysis == "Regular") %>% select(conf.low) %>% my_round("b")`, `r dat_fig_results_between %>% filter(dv == "Life satisfaction" & iv == "YouTube" & type == "between" & Analysis == "Regular") %>% select(conf.high) %>% my_round("b")`]).
--->
-<!-- However, note that the effect were still completely inside of the null region, hence not large enough to be considered practically relevant. -->
-
-<!-- Note that when comparing the results with and without control variables, the results differed. -->
-<!-- For example, on the between-person level, one effect stopped being significant if controlled for additional variables. -->
-<!-- Initially, actively posting on social media was significantly (though not meaningfully) related to decreased life satisfaction. -->
-<!-- However, when controlling for potential confounders, the effect became virtually zero. -->
-
-<!-- For an overview of all between-person relations, see Figure \@ref(fig:fig-between). -->
-
-<!-- ### Covariates -->
-
 To contextualize the results reported above and to see if the study included any meaningful effects at all, I also looked at the effect sizes of the covariates.
@@
 
-The results showed that several effects crossed or fell completely outside of the SESOI. 
-They can hence be considered meaningful.
-For example, if physical health decreased, this had a meaningful detrimental impact on life satisfaction ($\beta$ `r data_tab_comp %>% filter(dv == "Life satisfaction" & iv == "Health") %>% select(estimate) %>% my_round("std_txt")` [95% CI `r data_tab_comp %>% filter(dv == "Life satisfaction" & iv == "Health") %>% select(conf.low) %>% my_round("std")`, `r data_tab_comp %>% filter(dv == "Life satisfaction" & iv == "Health") %>% select(conf.high) %>% my_round("std")`]), positive affect ($\beta$ `r data_tab_comp %>% filter(dv == "Positive affect" & iv == "Health") %>% select(estimate) %>% my_round("std_txt")` [95% CI `r data_tab_comp %>% filter(dv == "Positive affect" & iv == "Health") %>% select(conf.low) %>% my_round("std")`, `r data_tab_comp %>% filter(dv == "Positive affect" & iv == "Health") %>% select(conf.high) %>% my_round("std")`]), and negative affect ($\beta$ `r data_tab_comp %>% filter(dv == "Negative affect" & iv == "Health") %>% select(estimate) %>% my_round("std_txt")` [95% CI `r data_tab_comp %>% filter(dv == "Negative affect" & iv == "Health") %>% select(conf.low) %>% my_round("std")`, `r data_tab_comp %>% filter(dv == "Negative affect" & iv == "Health") %>% select(conf.high) %>% my_round("std")`]).
-Spending more time outside to exercise meaningfully increased positive affect ($\beta$ `r data_tab_comp %>% filter(dv == "Positive affect" & iv == "Sports") %>% select(estimate) %>% my_round("std_txt")` [95% CI `r data_tab_comp %>% filter(dv == "Positive affect" & iv == "Sports") %>% select(conf.low) %>% my_round("std")`, `r data_tab_comp %>% filter(dv == "Positive affect" & iv == "Sports") %>% select(conf.high) %>% my_round("std")`]).
-Being more satisfied with the current democratic system meaningfully increased life satisfaction ($\beta$ `r data_tab_comp %>% filter(dv == "Life satisfaction" & iv == "Satisfaction democracy") %>% select(estimate) %>% my_round("std_txt")` [95% CI `r data_tab_comp %>% filter(dv == "Life satisfaction" & iv == "Satisfaction democracy") %>% select(conf.low) %>% my_round("std")`, `r data_tab_comp %>% filter(dv == "Life satisfaction" & iv == "Satisfaction democracy") %>% select(conf.high) %>% my_round("std")`]).
+The results showed that several effects crossed or fell completely outside of the SESOI, and can hence be considered meaningful.
+For example, if physical health decreased, this had a meaningful detrimental impact on life satisfaction ($\beta$ `r data_tab_comp_std %>% filter(dv == "Life satisfaction" & iv == "Health") %>% select(estimate) %>% my_round("std_txt")` [95% CI `r data_tab_comp_std %>% filter(dv == "Life satisfaction" & iv == "Health") %>% select(conf.low) %>% my_round("std")`, `r data_tab_comp_std %>% filter(dv == "Life satisfaction" & iv == "Health") %>% select(conf.high) %>% my_round("std")`]), positive affect ($\beta$ `r data_tab_comp_std %>% filter(dv == "Positive affect" & iv == "Health") %>% select(estimate) %>% my_round("std_txt")` [95% CI `r data_tab_comp_std %>% filter(dv == "Positive affect" & iv == "Health") %>% select(conf.low) %>% my_round("std")`, `r data_tab_comp_std %>% filter(dv == "Positive affect" & iv == "Health") %>% select(conf.high) %>% my_round("std")`]), and negative affect ($\beta$ `r data_tab_comp_std %>% filter(dv == "Negative affect" & iv == "Health") %>% select(estimate) %>% my_round("std_txt")` [95% CI `r data_tab_comp_std %>% filter(dv == "Negative affect" & iv == "Health") %>% select(conf.low) %>% my_round("std")`, `r data_tab_comp_std %>% filter(dv == "Negative affect" & iv == "Health") %>% select(conf.high) %>% my_round("std")`]).
+Spending more time outside to exercise meaningfully increased positive affect ($\beta$ `r data_tab_comp_std %>% filter(dv == "Positive affect" & iv == "Sports") %>% select(estimate) %>% my_round("std_txt")` [95% CI `r data_tab_comp_std %>% filter(dv == "Positive affect" & iv == "Sports") %>% select(conf.low) %>% my_round("std")`, `r data_tab_comp_std %>% filter(dv == "Positive affect" & iv == "Sports") %>% select(conf.high) %>% my_round("std")`]).
+<!-- Being more satisfied with the current democratic system meaningfully increased life satisfaction ($\beta$ `r data_tab_comp_std %>% filter(dv == "Life satisfaction" & iv == "Satisfaction democracy") %>% select(estimate) %>% my_round("std_txt")` [95% CI `r data_tab_comp_std %>% filter(dv == "Life satisfaction" & iv == "Satisfaction democracy") %>% select(conf.low) %>% my_round("std")`, `r data_tab_comp_std %>% filter(dv == "Life satisfaction" & iv == "Satisfaction democracy") %>% select(conf.high) %>% my_round("std")`]).
+-->
 The strongest aspect affecting well-being was internal locus of control.
-If people felt more in control of their lives, this strongly increased both life satisfaction ($\beta$ `r data_tab_comp %>% filter(dv == "Life satisfaction" & iv == "Internal locus of control") %>% select(estimate) %>% my_round("std_txt")` [95% CI `r data_tab_comp %>% filter(dv == "Life satisfaction" & iv == "Internal locus of control") %>% select(conf.low) %>% my_round("std")`, `r data_tab_comp %>% filter(dv == "Life satisfaction" & iv == "Internal locus of control") %>% select(conf.high) %>% my_round("std")`]) and
-positive affect ($\beta$ `r data_tab_comp %>% filter(dv == "Positive affect" & iv == "Internal locus of control") %>% select(estimate) %>% my_round("std_txt")` [95% CI `r data_tab_comp %>% filter(dv == "Positive affect" & iv == "Internal locus of control") %>% select(conf.low) %>% my_round("std")`, `r data_tab_comp %>% filter(dv == "Positive affect" & iv == "Internal locus of control") %>% select(conf.high) %>% my_round("std")`]),
-while decreasing negative affect ($\beta$ `r data_tab_comp %>% filter(dv == "Negative affect" & iv == "Internal locus of control") %>% select(estimate) %>% my_round("std_txt")` [95% CI `r data_tab_comp %>% filter(dv == "Negative affect" & iv == "Internal locus of control") %>% select(conf.low) %>% my_round("std")`, `r data_tab_comp %>% filter(dv == "Negative affect" & iv == "Internal locus of control") %>% select(conf.high) %>% my_round("std")`]).
-
+If people felt more in control of their lives, this strongly increased both life satisfaction ($\beta$ `r data_tab_comp_std %>% filter(dv == "Life satisfaction" & iv == "Internal locus of control") %>% select(estimate) %>% my_round("std_txt")` [95% CI `r data_tab_comp_std %>% filter(dv == "Life satisfaction" & iv == "Internal locus of control") %>% select(conf.low) %>% my_round("std")`, `r data_tab_comp_std %>% filter(dv == "Life satisfaction" & iv == "Internal locus of control") %>% select(conf.high) %>% my_round("std")`]) and
+positive affect ($\beta$ `r data_tab_comp_std %>% filter(dv == "Positive affect" & iv == "Internal locus of control") %>% select(estimate) %>% my_round("std_txt")` [95% CI `r data_tab_comp_std %>% filter(dv == "Positive affect" & iv == "Internal locus of control") %>% select(conf.low) %>% my_round("std")`, `r data_tab_comp_std %>% filter(dv == "Positive affect" & iv == "Internal locus of control") %>% select(conf.high) %>% my_round("std")`]),
+while decreasing negative affect ($\beta$ `r data_tab_comp_std %>% filter(dv == "Negative affect" & iv == "Internal locus of control") %>% select(estimate) %>% my_round("std_txt")` [95% CI `r data_tab_comp_std %>% filter(dv == "Negative affect" & iv == "Internal locus of control") %>% select(conf.low) %>% my_round("std")`, `r data_tab_comp_std %>% filter(dv == "Negative affect" & iv == "Internal locus of control") %>% select(conf.high) %>% my_round("std")`]).
 For an overview, see Figure \@ref(fig:fig-control).
@@
 All variables standardized.
-SESOI: beta = |.10|", out.width = "\\textwidth", cache=F, warning=F, message=F, fig.width=7, fig.height=8}
-fig_results_control_std
-# knitr::include_graphics("figures/fig_results_control_std.png")
+SESOI: beta = |.10|", cache=F, warning=F, message=F, fig.width=7, fig.height=8, out.width="\\textwidth"}
+# fig_results_comp_std
+knitr::include_graphics("figures/fig_results_comp_std.png")
 ```
+
+Because life satisfaction is more stable than affect, the effects of communication might materialize some time later. 
+I hence also tested the effects across the longer intervals of one month and four months.
+Results showed that all effects disappeared. 
+No effect remained significant, implying that at least in this case in this case effects take place on a shorter interval.
+
+Finally, as suggested by the differential susceptibility of media effects model, media effects can depend on dispositional factors, developmental stages, or cultural norms [@valkenburgDifferentialSusceptibilityMedia2013], such as gender and age [@orbenWindowsDevelopmentalSensitivity2022].
+I hence reran the analyses, differentiating effects for boys and girls and for age cohorts.
+The results showed that effects did not differ across genders.
+The effects also did not depend on age. 
+However, one effect stood out and was significant.
+Compared to the middle age category Generation X, results showed that if users from Generation Z posted more COVID-19 content than usual this lead to significantly more negative affect ($\beta$ `r fit_aff_neg_lmer_age_tab %>% filter(term == "age_genGen Z:soc_med_post_w") %>% select(estimate) %>% my_round("std_txt")` [95% CI `r fit_aff_neg_lmer_age_tab %>% filter(term == "age_genGen Z:soc_med_post_w") %>% select("2.5 %") %>% my_round("std")`, `r fit_aff_neg_lmer_age_tab %>% filter(term == "age_genGen Z:soc_med_post_w") %>% select("97.5 %") %>% my_round("std")`]).
+<!-- Although this might just be a fluke finding, the low p-value suggests that it might actually be real. -->
 
@@
 
-Based on a panel study with 32 waves largely representative of the Austrian population, this study analyzed the effects of COVID-19 related social media use on well-being.
+Based on a panel study with 34 waves largely representative of the Austrian population, this study analyzed the effects of COVID-19 related social media use on well-being.
 Between person correlation analyses showed that more active users of COVID-19 related content on social media also reported decreased well-being. 
-The within-person relations, which are more informative regarding causal effects, showed a different pattern, however.
+For example, respondents who read more COVID-19 related content than others reported slightly lower levels of life satisfaction, somewhat lower levels of positive affect, and substantially higher levels of negative affect than others.
+To see if these between person correlations would translate to within-person effects, I analyzed if changes in a person's media use led to changes in their well-being.
+The within-person relations showed a different pattern.
 If people consumed more COVID-19 content on social media than usual, this did not meaningfully reduce their well-being.
-Several statistically significant effects were found, but these were very small.
-For example, people who liked and shared more COVID-19 related posts than usual reported slightly higher levels of negative affect.
-Posting more content about COVID-19 than usual slightly increased negative affect.
-People who read more COVID-19 related posts than usual reported slightly decreased positive affect.
-Using Twitter for COVID-19 related content slightly increased negative affect. 
+Although several statistically significant effects were found, these were very small.
+For example, people who read more COVID-19 related posts than usual reported slightly decreased positive affect.
+People who liked and shared more COVID-19 related posts than usual reported slightly higher levels of negative affect.
+Posting more content about COVID-19 than usual slightly decreased life satisfaction, while increasing both negative affect and positive affect.
+Using Twitter for COVID-19 related content slightly increased negative affect, as did YouTube.
 <!-- was related to Reading COVID-19 related content on social media slightly reduced positive affect. -->
@@
 According to the preregistered procedure, they should hence be considered irrelevant. 
-Other factors, for which we would expect to find meaningful effects, such as health or sports, indeed showed substantial and meaningful impacts on well-being. 
+Additional analyses revealed that other factors, for which we would expect to find meaningful effects, such as health or sports, indeed showed substantial and meaningful impacts on well-being.
+In addition, when testing for the longer intervals of one month and four months, again no meaningful effects were found.
 In conclusion, COVID-19 related activity on social media was not a particularly strong influence on peoples' well-being. 
-The results do not support the popular fears that "doomscrolling" or overusing social media during times of crises constitutes a relevant risk for well-being.
+The results do not support the popular fears that "doomscrolling" or overusing social media during times of crises constitutes a prominent risk for well-being.
 
@@
 First, overall the results do suggest that effects of COVID-19 related social media use on well-being tend to take place in the negative as opposed to the positive spectrum.
-Although very small, four statistically significant negative results of COVID-19 related social media use on well-being were found.
-Not a single positive effect emerged.
+Although very small, five statistically significant negative results of COVID-19 related social media use on well-being were found.
+Only one positive effect emerged.
 Also note that in the analyses several control variables were included, ruling out plausible alternative explanations for the negative results.
-For example, it was controlled for whether or not participants contracted a COVID-19 infection during a specific wave.
+For example, it was controlled for as to whether or not participants contracted a COVID-19 infection during a specific wave.
 Hence, we can rule out the alternative explanation that having an infection was the root cause of increased communication and reduced well-being.
 
-Second, significant outcomes emerged for positive or negative affect, but not for life satisfaction. 
-Life satisfaction was stable and not affected by any type or channel of social media communication.
-The more fluctuating positive and negative affect, however, were affected.
-Liking, sharing, and posting COVID-19 related content, and spending time on Twitter to browse COVID-19 related content, all negatively influenced affect. 
-This is aligned with prior findings which showed that social media can trigger negative affect but does not reduce life satisfaction [@huangTimeSpentSocial2017].
+Second, six significant outcomes emerged for positive or negative affect, but only one for life satisfaction. 
+Life satisfaction is more stable and not that easily affected by any type or channel of social media communication.
+The more fluctuating positive and negative affect, however, were affected (albeit only slightly).
+Liking, sharing, and posting COVID-19 related content, and spending more time on Twitter and YouTube to browse COVID-19 related content, all slightly negatively influenced affect. 
+This is aligned with prior findings which showed that social media use can trigger negative affect, but that it is less likely to determine life satisfaction [@huangTimeSpentSocial2017].
 <!-- Writing posts implies more substantial engagement, and perhaps also negative exchanges with other users, which could explain the results. -->
@@
 More deeply engaging with this type of content could negatively affect active authors.
-The hypothesis that tonality could explain the negative effects is especially supported by the observation that spending more time on Twitter than usual increased negative affect.
-Communication on Twitter is often found to be more negative and impolite compared to other SNSs [@halpernSocialMediaCatalyst2013].
+The hypothesis that tonality could explain the negative effects is especially supported by the observation that spending more time on Twitter and YouTube than usual increased negative affect.
+Communication on both channels is often found to be negative and impolite [e.g., @muellerTwitterMadeMe2022], also when compared to other SNSs [@halpernSocialMediaCatalyst2013].
 Consuming more negative and misleading information could hence explain the (slightly) increased levels of negative affect.
@@
 Third, the results show that it makes sense to analyze different communication types and communication channels.
-Reading reduced positive affect, while liking, sharing and commenting increased negative affect.
-Whereas it was often stated that passive use is bad and active use good [@verduynPassiveFacebookUsage2015], this pattern thus did not emerge here.
-Instead, all three types were negative.
-The results hence support the findings from @valkenburgAssociationsActivePassive2022, who also could not confirm the claim that active use is good and that passive use is bad.
-Focusing on communication channels, Twitter seems to be more negative, as has often been observed [@halpernSocialMediaCatalyst2013].
+Reading slightly reduced positive affect, while liking, sharing, and posting slightly increased negative affect.
+Interestingly, posting COVID-19 related comment slightly increased negative affect, while at the same time it also slightly increased positive affect.
+Posting content is often met with strong reaction, both positive by means of likes and negative by means of critical comments. 
+Overall, though, posting led to slightly reduced levels of life satisfaction.
+In conclusion, whereas it was often stated that passive use is bad and active use good [@verduynPassiveFacebookUsage2015], this pattern was only partially found here.
+<!-- Instead, all three types were negative. -->
+The results are aligned with the findings from @valkenburgAssociationsActivePassive2022, who could not confirm that active use is good and that passive use is bad.
+Focusing on communication channels, Twitter and YouTube seem to be more negative, as has often been observed [@halpernSocialMediaCatalyst2013], while Instagram, WhatsApp, and Facebook were neutral.
 But, again, all of these effects are very small.
@@
 Taken together, the results are hence aligned with the underlying theoretical models and prior empirical results.
-The findings support the Different Susceptibility to Media Effects Theory, which states that effects are generally small and depend on the type of communication.
-The results are well-aligned with mood management theory [@zillmannMoodManagementCommunication1988] or the uses and gratifications approach [@katzUsesGratificationsResearch1973], whose premises preclude particularly negative effects of routine and widespread media consumption.
-If the effects of social media were indeed profoundly negative on average, then people likely would not spend so much time on social media engaging with COVID-19 content.
-Likewise, recent empirical studies and meta-analyses reported rather small negative effects, too.
+The findings support the differential susceptibility of media effects model [@valkenburgDifferentialSusceptibilityMedia2013], such that effects are generally small and that they depend on the type and channel of communication.
+Additional analyses did not reveal that effects depended on gender.
+Age also large did not play a significant moderation role, but effects of posting COVID-19 related content were found to be more negative for Generation Z.
+Indeed, it has often been argued that effects of social media use are more negative for Gen Z than for prior generations, and this finding can be seen a further tentative support for this hypothesis. 
+From a broader perspective, the results are well-aligned with mood management theory [@zillmannMoodManagementCommunication1988] and the uses and gratifications approach [@katzUsesGratificationsResearch1973], whose premises preclude particularly negative effects of routine and widespread media consumption.
+Both theories posit that if the effects of social media were indeed profoundly negative on average, then people likely would not spend so much time on social media engaging with COVID-19 content.
+Finally, recent empirical studies and meta-analyses reported rather small negative effects, too.
 Several studies found that the effects of various types of social media use on well-being are small, often too small to matter [@bendauAssociationsCOVID19Related2021; @meierComputermediatedCommunicationSocial2020a; @orbenTeenagersScreensSocial2020; @fergusonThisMetaanalysisScreen2021], echoing the results obtained here.
 
-In light of the overall very small effects, engaging in COVID 19-related social media use should not be a major concern for one's well-being.
-That said, the results still imply that it does make sense to critically reflect upon COVID-19 related social media use. 
-On average, it is likely more beneficial to post less actively about COVID-19 on social media and to spend less time on Twitter consuming COVID-19 related content.
-The results allow us to reject a positive effect: Writing more posts on social media will likely not increase well-being.
+<!-- The results allow us to reject a positive effect: Writing more posts on social media will likely not increase well-being. -->
 
@@
 Focusing on within-person effects and controlling for several potential confounders, this study provides an improved perspective on assessing causality.
-However, three major challenges remain. 
-First, it could be that the timing of the interval was wrong; it could be that effects unfold in shorter or longer intervals. 
-Next, it could be that there was not sufficient variance in media use and well-being. 
-Without sufficient variability, statistical models cannot detect existing effects.
-Third, it is necessary to control for _all_ relevant confounding third variables. 
-Although this study included are large list of confounders, it could still be that crucial variables were missed  [@zillmannMoodManagementCommunication1988].
+However, several challenges remain. 
+<!-- First, it could be that the timing of the interval was wrong; it could be that effects unfold in shorter or longer intervals.  -->
+<!-- Next, it could be that there was not sufficient variance in media use and well-being.  -->
+<!-- Without sufficient variability, statistical models cannot detect existing effects. -->
+<!-- Third,  -->
+In order to correctly establish causality in non-experimental designs, it is necessary to control for _all_ relevant confounding third variables [@rohrerThinkingClearlyCorrelations2018]. 
+Although this study included are large list of confounders, it could still be that crucial variables were missed.
+<!-- [@zillmannMoodManagementCommunication1988]. -->
 <!-- While controlling for potential confounders can support claims of causality, the procedure implemented here cannot prove causality. -->
@@
 <!-- Effects depend on the intensity of use or the length of the interval. -->
-To borrow the words from @rohrerTheseAreNot2021, there is no such thing as "the" effect of social media use on well-being.
-The results reported here are hence contingent on the design I used.
-To document how effects unfold, future research needs to employ different study designs probing different intervals.
-More thought needs to be invested in what factors to control for, and I hope this study provides a first step into this direction.
+<!-- To borrow the words from @rohrerTheseAreNot2021, there is no such thing as "the" effect of social media use on well-being. -->
+<!-- The results reported here are hence contingent on the design I used. -->
+<!-- To document how effects unfold, future research needs to employ different study designs probing different intervals. -->
+More thought needs to be invested in which factors to control for and, equally important, for which factors not to control for.
+I hope this study provides a first step into this direction.
 
@@
 
+<!-- That said, the results still imply that it does make sense to critically reflect upon COVID-19 related social media use.  -->
+<!-- On average, it seems more beneficial to post less actively about COVID-19 on social media and to spend less time on Twitter and YouTube consuming COVID-19 related content. -->
+
 In this study, COVID-19 related social media use did not meaningfully affect well-being. 
@@
 Factors other than social media use, however, were meaningfully related to well-being, including physical health, exercise, satisfaction with democracy, or believing that one is in control of one's life.
-Hence, when trying to improve well-being during a pandemic, instead of focusing on social media it seems more fruitful to address other, more pertinent societal problems related to health care, regular exercise, or a functioning democratic system.
+In light of the overall very small effects, engaging in COVID 19-related social media use should not be considered a major concern for one's well-being.
+Hence, when trying to improve well-being during a pandemic, instead of focusing on social media it seems more fruitful to address other, more pertinent societal problems related to health care, regular exercise, or psychological resilience.
 

### 2023-07-11 manuscript.tex
Major updates Implemented feedback

@@
 \usepackage{amsmath,amssymb}
-\usepackage{lmodern}
 \usepackage{iftex}
@@
 \else % if luatex or xetex
-  \usepackage{unicode-math}
+  \usepackage{unicode-math} % this also loads fontspec
   \defaultfontfeatures{Scale=MatchLowercase}
   \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
+\fi
+\usepackage{lmodern}
+\ifPDFTeX\else
+  % xetex/luatex font selection
 \fi
@@
 \note{\clearpage}
+\usepackage{setspace}
+\AtBeginEnvironment{tabular}{\singlespacing}
+\AtBeginEnvironment{lltable}{\singlespacing}
+\AtBeginEnvironment{tablenotes}{\doublespacing}
+\captionsetup[table]{font={stretch=1.5}}
+\captionsetup[figure]{font={stretch=1.5}}
 
@@
 \IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
-\urlstyle{same} % disable monospaced font for URLs
+\urlstyle{same}
 \hypersetup{
-  pdftitle={COVID-19 related social media use slightly decreases well-being},
+  pdftitle={Effects of COVID-19 related social media use on well-being},
   pdflang={en-EN},
@@
 
-\title{COVID-19 related social media use slightly decreases well-being}
+\title{Effects of COVID-19 related social media use on well-being}
 \author{Name blinded for review\textsuperscript{1}}
@@
 
-\shorttitle{COVID-19 related social media use \& well-being}
+\shorttitle{Effects of COVID-19 related social media use on well-being}
 
@@
 To this end, people increasingly use social media.
-However, being particularly engaging, many find it hard to disconnect from social media, especially during times of crisis.
-Using data from the Austrian Corona Panel Project consisting of 3,485 participants from 32 waves, controlling for several stable and varying confounders, the results showed that COVID-19 related social media use reduced facets of well-being.
-However, the effects were very small.
-Other factors such as health, income, exercise, or internal locus of control showed larger and more meaningful effects.
+However, because social media are particularly engaging, many find it hard to disconnect, especially during times of crisis.
+Using data from the Austrian Corona Panel Project consisting of 3,485 participants from 34 waves, controlling for several stable and varying confounders, the results showed that COVID-19 related social media use did not meaningfully reduce well-being.
+Other factors such as health, income, exercise, or internal locus of control showed larger and meaningful effects.
 }
@@
 How is it transmitted? What are the current safety regulations?
-To obtain relevant information, many people heavily relied on social media, which were at an all time high (Statista, 2021).
-
+To obtain relevant information, many people heavily relied on social media, with use being at an all time high (Statista, 2021).
 Some actually could not stop using social media to learn about COVID-19 related news.
-A new phenomenon termed ``doomscrolling'' emerged.
+A new phenomenon termed ``doomscrolling'' emerged (Sharma et al., 2022).
 Many users were glued to their screens and found it hard to pursue other relevant activities such as working, taking a break, or even looking after their children (Klein, 2021).
 In the media it was hence increasingly asked whether using social media for COVID-19 related reasons would, next to all other stressors, create an additional burden on mental health (Sandstrom et al., 2021).
-Although research has begun addressing this question (Bendau et al., 2021; Bradley \& Howard, 2021; Choi \& Choung, 2021; Dörnemann et al., 2021; Eden et al., 2020; Guazzini et al., 2022; Latikka et al., 2022; Liu \& Tong, 2020; Riehm et al., 2020; Sewall et al., 2021; Stainback et al., 2020; Yue et al., 2022), it still largely unknown if COVID-19 related social media use during the pandemic has had a meaningful impact on well-being.
-By analyzing large-scale panel data from Austria and by focusing on intra-individual causal effects, this study aims to answer to this question.
+Although research has begun addressing this question
+(e.g., Bendau et al., 2021; Eden et al., 2020; Sewall et al., 2021),
+it still largely unknown if COVID-19 related social media use during the pandemic has had a meaningful impact on well-being.
+This study hence aims to 1) reveal the effect of the different types and channels of social media use on individual well-being, 2) provide generalizable and robust results by analyzing a large-scale longitudinal data-set with 34 waves, and 3) determine the within-person causal effects by analyzing how changes in social media use lead to changes in well-being.
 
@@
 Because effects of social media depend on how they are used (Verduyn et al., 2015), I further distinguish three types of use and five popular channels.
-The types of use include reading, posting, and liking and sharing COVID-19 related posts.
-In doing so, this study analyzes social media use focused on COVID-19 related content, which includes posting thoughts about the pandemic, reading posts and comments, or retweeting COVID-19 related news.
+The types of use include reading, liking and sharing, and posting COVID-19 related content.
+In doing so, this study analyzes social media use focused on COVID-19 related content, which includes posting thoughts about the pandemic, reading posts and comments, or retweeting and liking COVID-19 related news.
+Liking and sharing are combined as they both represent low-threshold, platform-ingrained, easily quantifiable interactions.
 The five channels to be investigated are Facebook, Twitter, Instagram, WhatsApp, and YouTube, which at the time ranked among the most popular social media services in Austria.
 
-How could the various types and channels of COVID-19 related social media use affect well-being?
-According to the set-point theory, well-being is surprisingly stable (Lykken, 1999).
-Although specific events such as marriage or salary can have significant impacts on well-being, in most cases effects are only short-term, with well-being after some time routinely returning to prior levels (Sheldon \& Lucas, 2014).
+\hypertarget{social-media-effects-on-well-being}{%
+\subsection{Social Media Effects on Well-Being}\label{social-media-effects-on-well-being}}
+
+How easily can well-being be affected by external influences?
+In general, according to the set-point theory, well-being is surprisingly stable (Lykken, 1999).
+Although specific events such as marriage or salary increase can have significant impacts on well-being, in most cases effects are only short-term, with well-being after some time returning to prior levels (Sheldon \& Lucas, 2014).
 Only very specific factors such as unemployment, disability, or death can cause long-term changes in well-being (Lucas, 2007).
-Can media use be such a factor?
-Current literature overviews suggest no:
-Although social media use on average decreases well-being, the effects are small (Meier \& Reinecke, 2020).
-In addition, the effects of media use differ across individuals and types of content (Valkenburg \& Peter, 2013).
+So although well-being can be affected by external events and factors, this does not happen easily.
+
+Can social media use be such a factor?
+Current literature overviews suggest that social media use on average does seem to decrease well-being (Meier \& Reinecke, 2020).
+However, for most well-being outcomes, such as life satisfaction, general well-being, or loneliness, the effects are small (Meier \& Reinecke, 2020).
+These small findings can be explained with the differential susceptibility of media effects model (Valkenburg \& Peter, 2013), which states that there is substantial \emph{variation} of media effects for individual users.
 Whereas for some users social media are more beneficial, for others they are more harmful.
+On average, however, and this is central for this study here, effects tend to be small (Valkenburg \& Peter, 2013).
 For example, in one study it was estimated that roughly one quarter of all users experienced negative effects, another quarter positive effects, while for the rest the effects were neutral (Beyens et al., 2021).
-This finding is aligned with the Differential Susceptibility to Media Effects Model:
-Although there is substantial \emph{variation} of media effects for individual users, there are both negative and positive aspects, leading to small effects on average (Valkenburg \& Peter, 2013)---as confirmed by several literature overviews (Dienlin \& Johannes, 2020; Huang, 2017; Meier \& Reinecke, 2020; Orben, 2020).
+Whether or not effects are positive or negative depend on (a) dispositional factors (e.g., personality, temperament, gender), (b) developmental factors (e.g., age, developmental tasks), (c) and social factors (e.g., environment, norms, upbringing).
+Finally, effects depend also on the content that is consumed.
+If the content is aligned with dispositions, developmental capacities, and converging contexts, effects tend to be stronger (Valkenburg \& Peter, 2013).
 
 Why are the effects of social media use on well-being small on average?
-Using Social media is ambivalent.
-It can impair well-being when causing embarrassment, stress, or disinformation, and it can improve well-being when providing connectedness, information, or entertainment (Büchi, 2021).
-Although social media are often associated with negative outcomes, there are a large number of positive effects, which either offset or at least significantly reduce negative outcomes.
-Positive outcomes encompass finding relevant information; maintaining and fostering relationships; expressing one's personality; or entertaining oneself (e.g., Pelletier et al., 2020).
-Negative effects include distraction; displacement of other meaningful activities; consumption of shallow, misleading or challenging content; or negative social comparisons (e.g., Meier \& Krause, 2022).
-People are much more likely to use media that offer plentiful benefits (Katz et al., 1973).
-Users implicitly learn what media help them regulate their mood and thereby well-being (Zillmann, 1988).
-Given the ubiquitous use of social media, the many benefits they offer, plus the general and implicit competence to use media that foster mood management, it is hence unlikely that the effects of social media use on well-being are profusely negative.
+Two prominent media effect theories argue implicitly against strong average negative effects.
+First, according to mood management theory (Zillmann, 1988), using media can affect people's moods.
+Use can be stimulating or overwhelming, relaxing or boring.
+After some time, users implicitly learn which media help them balance their mood and affect according to their own situational needs (Zillmann, 1988).
+Those media that eventually become part of one's media repertoire hence, on average, tend to be beneficial for users to regulate their mood (Marciano et al., 2022).
+In conclusion, if a certain medium is used frequently, mood-management theory argues that it is likely not detrimental for well-being.
+On the other hand, although mood management theory suggests that effects should not be overly negative, it could be that although short-term effects are positive, long-term effects are negative.
+Precisely because social media have so many positive consequences in the short run, this might cause problem in the long run---for example, because social media use displaces other potentially more meaningful and/or relaxing activities (Hall \& Liu, 2022).
+As with many other things, there can be too much of a good thing.
+It is therefore often asked whether social media can become addictive, and users sometimes express this fear themselves (Yang et al., 2021).
+However, a recently published meta-analysis found that the two most prominent measures of addiction, the Bergen Facebook Addiction Scale and the Bergen Social Media Addiction Scale, have only small relations to well-being (Duradoni et al., 2020).
+In addition, the general idea of labeling excessive social and new media use as addiction was criticized, arguing that social media use represent a new regular behavior that should not be pathologized (Galer, 2018; van Rooij et al., 2018).
+
+Second, while mood management theory considers media use mainly driven by implicit learning experiences, uses and gratifications theory upholds that the process is more explicit and rational (Katz et al., 1973).
+Users select those media that they expect to have a desired effect, for example on mood, knowledge, or entertainment.
+If those beneficial media effects are missing, people will spend their time elsewhere.
+And social media, in general, offer several beneficial effects, explaining why they are used that much.
+They help find relevant information, maintain and foster relationships, express one's personality, and entertain oneself (Pelletier et al., 2020).
+In conclusion, because people spend so much time on social media consuming COVID-19 related content, according to both mood management theory and uses and gratifications theory this indirectly suggest that average effects on well-being are likely not particularly negative.
+
+\hypertarget{social-media-during-covid-19}{%
+\subsection{Social Media During COVID-19}\label{social-media-during-covid-19}}
+
+If we look at COVID-19 related use more specifically, how could the various types and channels of COVID-19 related social media use affect well-being?
+Several uses and gratifications exist, which help explain why people used social media frequently during the pandemic.
+Access to information and resources:
+Despite incorrect information, social media provide a vast platform for disseminating accurate and timely information about COVID-19 (John Hopkins University, 2023).
+It allows individuals to stay informed about the latest updates, guidelines, and recommendations from reputable health organizations and experts.
+Access to reliable information can help people make informed decisions, alleviate uncertainties, and feel empowered during the pandemic.
+Educational opportunities:
+Social media can be a valuable educational tool, with various organizations and experts sharing informative content about COVID-19 (World Health Organization, 2023).
+This can include explanatory videos, infographics, and articles that help individuals understand the virus, its transmission, prevention strategies, and vaccination information.
+Access to educational resources on social media can empower individuals to make informed decisions regarding their health.
+Community support and solidarity:
+Social media platforms enable individuals to connect with others who are experiencing similar challenges during the pandemic (Guazzini et al., 2022).
+Sharing experiences, offering support, and expressing solidarity can foster a sense of community and reduce feelings of isolation.
+Engaging in online communities and support groups can provide emotional support and create a network of like-minded individuals.
+Promoting mental health and well-being:
+Many mental health organizations and professionals utilize social media to share tips, strategies, and resources for maintaining mental well-being during the pandemic (Twitter, 2020).
+They provide guidance on coping mechanisms, stress reduction techniques, and self-care practices.
+Engaging with such content might help individuals prioritize their mental health and develop resilience during challenging times.
+Encouraging positive behaviors:
+Social media campaigns and initiatives can promote positive behaviors related to COVID-19, such as mask-wearing, physical distancing, hand hygiene, and vaccination (Athey et al., 2023; Hunt et al., 2022).
+Public health organizations and influencers leverage the power of social media to spread awareness and encourage responsible actions, contributing to public health efforts and fostering a sense of collective responsibility.
+
+On the other hand, the effects might be negative, perhaps best explained by the following five mechanisms.
+Misinformation and rumors:
+Social media platforms can easily spread false or misleading information about COVID-19 (Li et al., 2020).
+Due to the ease of sharing and the lack of fact-checking, inaccurate information can go viral and might cause confusion, anxiety, and panic among users.
+Overwhelm and anxiety:
+Constant exposure to COVID-19-related content on social media can lead to information overload and contribute to heightened anxiety levels (J. Fan \& Smith, 2021).
+The rapid spread of news, updates, and opinions can be overwhelming and might exacerbate existing stress or fears about the pandemic (Sharma et al., 2022).
+Confirmation bias and echo chambers:
+Social media algorithms are designed to show users content that aligns with their interests and beliefs, potentially contributing to echo chambers (Cinelli et al., 2021).
+This might lead to confirmation bias, where individuals are more likely to be exposed to information that confirms their preexisting beliefs.
+In the context of COVID-19, this might reinforce false or misleading information and prevent individuals from considering alternative perspectives.
+Cyberbullying and harassment:
+Social media platforms are known for fostering negativity, with users sometimes engaging in cyberbullying and harassment (Giumetti \& Kowalski, 2022).
+Discussions around COVID-19 can become heated and polarized, leading to personal attacks and online conflicts.
+Such experiences threaten mental well-being and might contribute to feelings of distress and isolation.
+Social comparison and fear of missing out (FOMO):
+Social media often showcase the highlights and accomplishments of others, encouraging social comparison (Przybylski et al., 2013).
+During a pandemic, seeing posts about others' successes or seemingly perfect lives might intensify feelings of inadequacy or FOMO, especially when individuals are unable to participate in similar activities due to restrictions or personal circumstances (Sharma et al., 2022).
+This might impact mental health, as most humans have an innate need for social support and physical presence, which online interactions might not fully replicate.
 
 There is still little empirical research on how well-being is affected by social media use that is focused on COVID-19 specifically.
-Results are mixed.
-When browsing social media for COVID-19 related news, many users reported being captivated to such an extent they could not stop using social media (Klein, 2021).
-In a study with 1,131 residents from Wuhan in China, people who spent more time in quarantine also spent more time on social media (Yue et al., 2022).
-Those who spent more time on social media also engaged in more upward social comparison, which was related to increased levels of stress.
-People who used social media as a primary source of information reported on average ``significantly more unspecific anxiety and depression {[}\ldots{]} and significantly more specific COVID-19 related anxiety symptoms'' (Bendau et al., 2021, p.
-288).
-Eden et al. (2020) analyzed how 425 US college students used media during the first wave of the pandemic, finding both positive and negative relations with well-being.
-In a large-scale study with 11,537 respondents from the US, increased COVID-19-related media consumption was related to more psychological distress (Stainback et al., 2020).
-A four-wave panel study with 384 young adults from the U.S.
-analyzed the effects of general digital technology use during the pandemic on mental health, finding that digital technology did not have significant effects on mental health (Sewall et al., 2021; for a similar study with comparable results, see Bradley \& Howard, 2021).
-A study with 2.057 respondents from Italy reported that during the pandemic virtual community and social connectedness increased (Guazzini et al., 2022).
-In a study with 735 participants from Finland, levels of loneliness were stable during the pandemic, and people who engaged more on social media experienced less loneliness (Latikka et al., 2022).
-Together, the literature is mixed, showing both positive and negative affects of social media use focused on or during COVID-19 on well-being (see also Dörnemann et al., 2021; Liu \& Tong, 2020; Riehm et al., 2020).
-
-In conclusion, these mixed empirical results, together with the observation that social media effects on well-being are very small in general, I expect that COVID-19 related communication on social media does not affect well-being in a meaningful or relevant way.
+Echoing the theoretical rationales outlines above, studies have yielded mixed results.
+Some studies found negative effects, indicating that excessive social media use for COVID-19 news led to compulsive behavior and increased stress levels, particularly due to upward social comparison (Klein, 2021; Stainback et al., 2020; Yue et al., 2022).
+Individuals who relied on social media as their primary information source reported higher levels of anxiety and depression symptoms (Bendau et al., 2021).
+Similarly, increased COVID-19-related media consumption was associated with higher psychological distress.
+On the other hand, some studies reported positive outcomes.
+Certain individuals experienced increased virtual community and social connectedness during the pandemic through social media, which contributed to their well-being (Guazzini et al., 2022).
+Additionally, engaging more on social media was associated with reduced feelings of loneliness (Latikka et al., 2022).
+A couple of studies reported mostly neutral effects of social media use on well-being indicators (Bradley \& Howard, 2021; Eden et al., 2020; Sewall et al., 2021).
+Overall, the literature demonstrates a mixed picture, highlighting both positive and negative effects of social media use focused on or during COVID-19 on well-being (see also Bradley \& Howard, 2021; Dörnemann et al., 2021; Liu \& Tong, 2020; Riehm et al., 2020; Sewall et al., 2021).
+
+In conclusion, given these mixed empirical results, together with the observation that social media effects on well-being are very small in general, and that several plausible theoretical mechanisms exist for both positive and negative effects, I expect that COVID-19 related communication on social media should not be decidedly positive or negative.
+It seems most likely that both positive and negative coexist, but that on average using social media for COVID-19 related reasons should not have substantial effects on well-being.
 
 \begin{quote}
-Hypothesis: The within-person effects of all types of COVID-19 related social media use on all types of well-being indicators---while controlling for several stable and varying covariates such as sociodemographic variables and psychological dispositions---will be trivial.
+Hypothesis: The within-person effects of all measures of COVID-19 related social media use (types: reading, liking and sharing, posting; channels: Twitter, Instagram, Facebook, YouTube, WhatsApp) on all measures of well-being indicators (positive affect, negative affect, life satisfaction)---while controlling for several stable and varying covariates such as sociodemographic variables and psychological dispositions (see below)---will be trivial.
 \end{quote}
@@
 The data come from the Austrian Corona Panel Project (Kittel et al., 2021), which is a large-scale standalone panel study.
-The data are hosted on AUSSDA, are publicly available here (\url{https://doi.org/10.11587/28KQNS}), and consist of 32 waves.
-The study was conducted between March 2020 and June 2022, and data collection is now finished.
-Between March 2020 and July 2020, the intervals between waves were weekly and afterward monthly.
-Each wave consists of at least 1,500 respondents.
-The sample size was \emph{N} = 3,485, with overall 111,520 observations.
-Panel mortality was compensated through a continuous re-acquisition of new participants.
+The data are hosted on AUSSDA, are publicly available here (\url{https://doi.org/10.11587/28KQNS}), and consist of 34 waves.
 Participants were sampled from a pre-existing online access panel provided by the company Marketagent, Austria.
 Panel members were incentivized with 180 credit points for each wave of the study.
+The study was conducted between March 2020 and February 2023.
+Between March 2020 and July 2020, the intervals between waves were weekly, until May 2022 (wave 32) monthly, and afterward after 5 months.
+Each wave consists of at least 1,500 respondents.
+Panel mortality was compensated through a continuous re-acquisition of new participants.
+The sample size was \emph{N} = 3,641, with overall 123,794 observations.
+For an overview of the study set-up, see Figure \ref{fig:study-desc}.
+
+\begin{figure}
+\includegraphics[width=1\textwidth]{figures/fig_study_description} \caption{Overview of study set-up}\label{fig:study-desc}
+\end{figure}
 
@@
 The participants provided their written informed consent to participate in this study.
-The average age was 41 years, 49 percent were male, 14 percent had a University degree, and 5 percent were currently unemployed.
+The average age was 40 years, 49 percent were male, 14 percent had a University degree, and 5 percent were currently unemployed.
 
@@
 Testing the hypothesis necessitates defining what is considered a ``trivial effect size''.
+To this end, we need to define a so-called smallest effect size of interest (SESOI) (Lakens et al., 2018).
+A trivial effect would then need to be smaller than the SESOI (see below).
+What could be a minimally interesting, nontrivial effect?
 Being a normative question, finding a clear, single, or unanimous answer is impossible.
@@
 The hypothesis is analyzed using the interval testing approach as proposed by Dienes (2014).
-To illustrate, if the 95\% confidence interval falls completely within the null-region (e.g., \emph{b} = -.02, {[}95\% CI: -.12, .08{]}), the hypothesis that the effect is trivial is supported.
-If the confidence interval and the null region overlap (e.g., \emph{b} = -.22, {[}95\% CI: -.27, -.17{]}), the hypothesis is not supported and the results are considered inconclusive, while a meaningful negative effect is rejected.
-If the confidence interval falls completely outside of the null-region (e.g., \emph{b} = -.40, {[}95\% CI: -.45, -.35{]}), the hypothesis is rejected and the existence of a meaningful positive effect is supported.
+To illustrate, let us consider the case of life satisfaction {[}SESOI: -.30 : +.30{]}.
+If the 95\% confidence interval falls completely within the null-region (e.g., \emph{b} = -.05, {[}95\% CI: -.15, .05{]}), the hypothesis that the effect is trivial is supported.
+If the confidence interval falls completely outside of the null-region (e.g., \emph{b} = -.40, {[}95\% CI: -.45, -.35{]}), the hypothesis is rejected and the existence of a meaningful negative effect is supported.
+If the confidence interval and the null region overlap (e.g., \emph{b} = -.30, {[}95\% CI: -.35, -.25{]}), the hypothesis is not supported and the results are considered inconclusive, while a meaningful positive effect is rejected.
+If the confidence interval exceeds both sides of the null region (e.g., \emph{b} = -.025, {[}95\% CI: -.40, .35{]}), the hypothesis is not supported and judgement is suspended.
 For an illustration, see Figure \ref{fig:sesoi}.
@@
 \includegraphics{manuscript_files/figure-latex/sesoi-1.pdf}
-\caption{\label{fig:sesoi}Using confidence intervals to test a null region. In this study, a trivial effect of social media use on life satisfaction is defined as ranging from b = -.30 to b = .30}
+\caption{\label{fig:sesoi}Using confidence intervals to test a null region.
+In this study, a trivial effect of social media use on life satisfaction is defined as ranging from b = -.30 to b = .30.
+Figure adapted from Dienes (2014).}
 \end{figure}
@@
 When using longitudinal designs to analyze causality, it is important to (a) focus on within-person effects (Hamaker, 2014); to (b) control for confounders (Rohrer \& Murayama, 2021); and to (c) test a plausible interval between measures (Dormann \& Griffin, 2015).
-To elaborate, in non-experimental designs it makes much sense to analyze causal effects from an internal, within-person perspective (Hamaker, 2014).
+First, in non-experimental designs it makes much sense to analyze causal effects from an internal, within-person perspective (Hamaker, 2014).
 If a specific person changes their media diet, we need to measure how this behavior affects their well-being.
@@
 
-Second, to identify confounders we should control for variables that affect both media use and well-being, which helps isolate the actual effect of social media use on well-being (Rohrer, 2018).
+Second, to identify confounders we should control for variables that affect both media use and well-being, which helps isolate the actual effect (Rohrer, 2018).
 Because we are adopting a within-person perspective, we need to implement \emph{time-varying} confounders (Rohrer \& Murayama, 2021).
 And because we are determining the \emph{overall} causal effect, we need to make sure \emph{not} to control for mediating variables (Rohrer, 2018), for doing so would bias our assessment of the causal effect.
-In this study, I hence preregistered to control for the following variables, which either have already been shown or are very likely to affect both social media use and well-being, and which also are not mediators:
-gender, age, education, Austria country of birth, Austria country of birth of parents, residency Vienna, text-based news consumption, video-based news consumption, household size, health, living space, access to garden, access to balcony, employment, work hours per week, being in home-office, household income, outdoor activities, satisfaction with democracy, disposition to take risks, and locus of control.
+In this study, I hence preregistered to control for the following variables, which either have already been shown or are likely to affect both social media use and well-being, and which also are not mediators:
+gender, age, education, Austria country of birth, Austria country of birth of parents, residency Vienna, text-based news consumption, video-based news consumption, household size, health, living space, access to garden, access to balcony, employment, work hours per week, being in home-office, household income, outdoor activities, disposition to take risks, and locus of control (Eger \& Maridal, 2015; Ward et al., 2016).
 
 Finally, one precondition of causality is temporal order and finding a plausible interval (Dormann \& Griffin, 2015).
-If variables are stable, longer intervals are needed; if they fluctuate, shorter intervals are required.
+If variables are stable, longer intervals are needed; if they fluctuate, shorter intervals.
 In the case of well-being, we need shorter intervals for the more fluctuating positive and negative affect, and longer ones for the more stable life satisfaction (Dienlin \& Johannes, 2020).
-Using social media can have instant effects on mood (Marciano et al., 2022).
-Effects on life satisfaction often take longer to manifest, for example because media use leads to actual changes in specific behaviors, which then in turn affect life satisfaction (Dienlin et al., 2017).
-In this study, I hence analyze how using social media during the last week affected positive and negative affect during the same week.
+Whereas using social media can have instant effects on mood (Marciano et al., 2022), effects on life satisfaction often take longer to manifest.
+For example, because media use leads to actual changes in specific behaviors, which then in turn affect life satisfaction (Dienlin et al., 2017).
+
+In this study, I hence analyze how changes in using social media \emph{during the last week} affected changes in positive and negative affect \emph{during the same week}.
 In other words, if people during the last week engaged in more COVID-19 related social media use than usual, did they feel better or worse during that week than usual?
 For life satisfaction, I implemented a longer interval.
-If people during the last week used COVID-19 related social media more than they usually do, were they at the end of the week more or less satisfied with their lives than they usually are?
-I hence analyze if when a person changes their social media diet, are there (a) \emph{simultaneous} changes in their affect and (b) \emph{subsequent} changes in their life satisfaction?
-These relations will be controlled for varying confounders, which fosters a causal interpretation.
+If people \emph{during the last week} used COVID-19 related social media more than they usually do, were they \emph{at the end of the week} more or less satisfied with their lives than they usually are?
+This way it is analyzed if when a person changes their social media diet, are there (a) \emph{simultaneous} changes in their affect and (b) \emph{subsequent} changes in their life satisfaction?
+For the main analyses, the interval is implemented via the wording of the items (see below), not by using lagged measures coming from prior waves waves.
+In additional analyses, I also tested how media use affects well-being one month or four months later.
+All analyses will be controlled for varying confounders (see below), which fosters a causal interpretation.
 
@@
 The data were hierarchical, and responses were separately nested in participants and waves (i.e., participants and waves were implemented as random effects).
-Nesting in participants allowed to separate between-person relations from within-person effects.
-Nesting in waves allowed to control for general exogenous developments, such as general decreases in well-being in the population, for example due to lockdown measures.
+Nesting in participants accounts for the longitudinal design.
+Nesting in waves controls for general exogenous developments, such as general decreases in well-being in the population, for example due to lockdown measures.
 Thus, there was no need additionally to control for specific phases or measures of the lockdown.
@@
 Mean scores were used for positive and negative affect.
-Missing responses were imputed using multiple imputation with predictive mean matching (five iterations, five data-sets), including categorical variables.
+Missing responses were imputed using multiple imputation with predictive mean matching (five iterations, 30 data-sets), including categorical variables.
 All variables were imputed except the social media use measures, as they were not collected on each wave.
 All variables included in the analyses presented here were used to impute missing data.
-
-\hypertarget{robustness-checks}{%
-\subsubsection{Robustness-checks}\label{robustness-checks}}
-
-To find out whether the inferences were robust across plausible but subordinate analyses, I conducted additional exploratory analyses.
+For the main analyses, results were pooled across all thirty data-sets.
+
+To contextualize the results, I conducted additional exploratory analyses.
 I reran the analyses (a) with additional not-preregistered covariates such as trust in media or government, (b) without covariates, (c) with single imputation, and (d) without imputation.
@@
 The response options were 1 (\emph{never}), 2 (\emph{on some days}), 3 (\emph{several times per week}), 4 (\emph{almost every day}), and 5 (\emph{daily}).
-The scale showed good factorial fit, \(\chi^2\)(62) = 67.03, \emph{p} = .309, CFI = 1.00, RMSEA \textless{} .01, 90\% CI {[}\textless{} .01, .02{]}, SRMR = .01.
+The scale showed good factorial fit, \(\chi^2\)(66) = 69.42, \emph{p} = .363, CFI = 1.00, RMSEA \textless{} .01, 90\% CI {[}\textless{} .01, .02{]}, SRMR = .01.
 Reliability was high, \(\omega\) = .85.
@@
 The response options were 1 (\emph{never}), 2 (\emph{on some days}), 3 (\emph{several times per week}), 4 (\emph{almost every day}), and 5 (\emph{daily}).
-The scale showed good factorial fit, \(\chi^2\)(443) = 3810.53, \emph{p} \textless{} .001, CFI = .98, RMSEA = .07, 90\% CI {[}.07, .08{]}, SRMR = .03.
+The scale showed good factorial fit, \(\chi^2\)(471) = 4012.14, \emph{p} \textless{} .001, CFI = .98, RMSEA = .07, 90\% CI {[}.07, .08{]}, SRMR = .03.
 Reliability was high, \(\omega\) = .91.
@@
 
-Social media use was measured for all participants on waves 1, 2, 8, 17, 23, and 28.
+Social media use was measured for all participants on waves 1, 2, 8, 17, 23, and 28 (see Figure 1).
 Freshly recruited respondents always answered all questions on COVID 19-related social media use.
-For the analyses I hence used all 32 waves.
+Because new respondents always provided data on media use, it was possible to include these data into the analyses.
+Hence, for the main analyses data from all 34 waves were used.
+In the additional analyses I tested longer intervals, namely if changes in social media use were associated with changes in well-being either one month of four months later.
+For these analyzes I used the predictors from waves 1, 2, 8, 17, 23, and 28, to see if they predicted changes in well-being either one month or four months later.
 
@@
 \caption{\label{tab:tab-descriptives}Descriptives of the main variables.}
+
+\small{
 
@@
 Well-being &  &  &  & \\
-\ \ \ Life satisfaction & 2.22 & 6.37 & 6.59 & 6.51\\
-\ \ \ Positive affect & 0.94 & 3.11 & 3.22 & 3.16\\
-\ \ \ Negative affect & 0.76 & 1.74 & 1.87 & 1.79\\
+\ \ \ Life satisfaction & 2.23 & 6.32 & 6.60 & 6.49\\
+\ \ \ Positive affect & 0.94 & 3.09 & 3.22 & 3.16\\
+\ \ \ Negative affect & 0.77 & 1.75 & 1.86 & 1.81\\
 Social media use &  &  &  & \\
-\ \ \ Read & 1.39 & 2.02 & 2.87 & 2.37\\
-\ \ \ Like \& share & 1.19 & 1.56 & 1.95 & 1.74\\
-\ \ \ Posting & 0.89 & 1.37 & 1.42 & 1.40\\
+\ \ \ Read & 1.38 & 1.92 & 2.88 & 2.35\\
+\ \ \ Like \& share & 1.19 & 1.54 & 1.94 & 1.74\\
+\ \ \ Posting & 0.89 & 1.36 & 1.42 & 1.40\\
 Social media channel &  &  &  & \\
-\ \ \ Facebook & 1.60 & 2.21 & 2.74 & 2.43\\
-\ \ \ Twitter & 0.95 & 1.32 & 1.45 & 1.38\\
-\ \ \ Instagram & 1.33 & 2.04 & 2.05 & 2.05\\
-\ \ \ WhatsApp & 1.67 & 2.32 & 2.65 & 2.46\\
-\ \ \ YouTube & 1.28 & 1.90 & 2.02 & 1.96\\
+\ \ \ Facebook & 1.58 & 2.02 & 2.72 & 2.37\\
+\ \ \ Twitter & 0.95 & 1.36 & 1.43 & 1.39\\
+\ \ \ Instagram & 1.34 & 2.00 & 2.08 & 2.05\\
+\ \ \ WhatsApp & 1.66 & 2.27 & 2.60 & 2.46\\
+\ \ \ YouTube & 1.28 & 1.91 & 1.98 & 1.95\\
 \bottomrule
@@
 
+}
+
 \end{threeparttable}
@@
 gender (female, male, diverse), age, education (ten options), Austria country of birth (yes/no), Austria parents' country of birth (no parent, one parent, both parents), and household size.
-I also controlled for the following varying covariates: five items on current living conditions, including self-reported physical health, whether participants contracted COVID-19 since the last wave, current household income, working in home-office, and overall work hours; nine items measuring use of specific national text-based and video-based news outlets; five items measuring outdoor activities such as exercise or meeting friends; and three more psychological measures including locus of control, disposition to take risks, and satisfaction with democracy.
+I also controlled for the following varying covariates: five items on current living conditions, including self-reported physical health, whether participants contracted COVID-19 since the last wave, current household income, working in home-office, and overall work hours; nine items measuring use of specific national text-based and video-based news outlets; five items measuring outdoor activities such as exercise or meeting friends; and two more psychological measures including locus of control and disposition to take risks.
 
@@
 Posting about COVID-19 related content stayed the same.
-Using Facebook, WhatsApp, and YouTube for COVID-19 related content decreased.
-Instagram stayed the same.
-Twitter use slightly increased.
+Using Facebook and WhatsApp for COVID-19 related content decreased.
+Instagram, YouTube, and Twitter stayed the same.
 The general initial decrease could be explained by the fact that the collection of data began at the end of March 2020, hence approximately three months after the pandemic's onset.
@@
 \begin{figure}
-\includegraphics[width=\textwidth]{figures/fig_descriptives} \caption{Well-being and media use across the 32 waves.
+\includegraphics[width=\textwidth]{figures/fig_descriptives} \caption{Well-being and media use across the 34 waves.
 Note.
@@
 People who engaged more with COVID-19 related content, including all types and channels of communication, reported substantially higher levels of negative affect.
-For example, people who were more likely to post COVID-19 content had much higher levels of negative affect (\emph{r} = .56).
+For example, people who were more likely to post COVID-19 content had much higher levels of negative affect (\emph{r} = .61).
 Note that these results represent between-person correlations, not causal within-person effects.
@@
 Regarding the effects of different communication types (i.e., reading, sharing, of posting about COVID-19 related content), all within-person effects fell completely within the a-priori defined null region (see Figure \ref{fig:fig-within}).
-For example, respondents who used social media more frequently than usual to like or share COVID-19 related content did not show a simultaneous change in life satisfaction (\emph{b} = -0.02 {[}95\% CI -0.08, 0.03{]}).
+For example, respondents who used social media more frequently than usual to like or share COVID-19 related content did not show a simultaneous change in life satisfaction (\emph{b} = -0.02 {[}95\% CI -0.06, 0.01{]}).
 As a result, the hypothesis of trivial effects was supported for all COVID-19 related types of social media communication.
 
-However, three effects stood out, as they were significantly different from zero.
-Users who read more COVID-19 related content than usual reported slightly reduced levels of positive affect (\emph{b} = -0.03 {[}95\% CI -0.05, -0.01{]}).
-Users who liked and shared more COVID-19 related content than usual also experienced slightly more negative affect than usual (\emph{b} = 0.05 {[}95\% CI 0.02, 0.08{]}).
-Users who wrote more COVID-19 related posts than usual also experienced slightly more negative affect than usual (\emph{b} = 0.05 {[}95\% CI 0.02, 0.08{]}).
+However, several effects stood out, as statistically they were significantly different from zero.
+Users who read more COVID-19 related content than usual reported slightly reduced levels of positive affect (\emph{b} = -0.03 {[}95\% CI -0.05, -0.02{]}).
+Users who liked and shared more COVID-19 related content than usual also experienced slightly more negative affect than usual (\emph{b} = 0.05 {[}95\% CI 0.04, 0.07{]}).
+Posting COVID-19 related content affected all types of well-being.
+Users who wrote more COVID-19 related posts than usual also reported slightly less life satisfaction than usual (\emph{b} = -0.04 {[}95\% CI -0.08, -0.01{]}) and slightly more negative affect than usual (\emph{b} = 0.05 {[}95\% CI 0.04, 0.07{]}).
+Interestingly, however, users who wrote more COVID-19 related posts than usual also experienced slightly \emph{higher} levels of positive affect than usual (\emph{b} = 0.02 {[}95\% CI 0.01, 0.04{]}).
 
@@
 Changes in the frequency of using different social media channels to attain information regarding COVID-19 were unrelated to meaningful changes in well-being.
-For example, respondents who used Facebook more frequently than usual to learn about COVID-19 did not show a simultaneous change in life satisfaction (\emph{b} \textless{} 0.01 {[}95\% CI -0.04, 0.04{]}).
+For example, respondents who used Facebook more frequently than usual to learn about COVID-19 did not show a simultaneous change in life satisfaction (\emph{b} -0.01 {[}95\% CI -0.04, 0.02{]}).
 In sum, the hypothesis of trivial effects was supported also for the COVID-19 related use of important social media channels.
 
-That said, one effect differed substantially from zero.
-Respondents who used Twitter more frequently than usual to attain COVID-19 related news reported slightly higher levels of negative affect than usual (\emph{b} = 0.01 {[}95\% CI \textless{} 0.01, 0.03{]}).
-However, the effect was still completely inside of the null region, hence likely not large enough to be considered meaningful.
+That said, two effects differed statistically from zero.
+Respondents who used Twitter more frequently than usual to attain COVID-19 related news reported slightly higher levels of negative affect than usual (\emph{b} = 0.02 {[}95\% CI 0.01, 0.04{]}).
+Likewise, respondents who used YouTube more frequently than usual for COVID-19 related issues reported slightly higher levels of negative affect than usual (\emph{b} = 0.01 {[}95\% CI \textless{} 0.01, 0.02{]}).
+However, both effects were still completely inside of the null region, hence likely not large enough to be considered meaningful.
 
@@
 \begin{figure}
-\includegraphics[width=\textwidth]{manuscript_files/figure-latex/fig-within-1} \caption{Within-person effects of COVID-19 related social media use on well-being.
+\centering
+\includegraphics{manuscript_files/figure-latex/fig-within-1.pdf}
+\caption{\label{fig:fig-within}Unstandardized within-person effects of COVID-19 related social media use on well-being.
 Note.
-The black estimates show the effects controlled for a large number of covariates (see text; preregistered); the grey estimates are without control variables (exploratory).
-The SESOI was b = |0.30| for life satisfaction and b = |0.15| for affect.
-Hence, all of the reported effects are not considered meaningful.}\label{fig:fig-within}
+The SESOI was b = \textbar0.30\textbar{} for life satisfaction and b = \textbar0.15\textbar{} for affect.
+Hence, all of the reported effects are not considered large enough to be meaningful.}
 \end{figure}
@@
 Life satisfaction &  &  &  &  & \\
-\ \ \ Reading & 0.03 & -0.01 & 0.07 & 0.02 & .091\\
-\ \ \ Liking \& Sharing & -0.02 & -0.08 & 0.03 & -0.01 & .299\\
-\ \ \ Posting & -0.05 & -0.11 & 0.01 & -0.02 & .114\\
-\ \ \ Facebook & 0.00 & -0.04 & 0.04 & 0.00 & .964\\
-\ \ \ Instagram & 0.01 & -0.01 & 0.03 & 0.01 & .226\\
-\ \ \ WhatsApp & 0.00 & -0.03 & 0.02 & 0.00 & .664\\
-\ \ \ YouTube & 0.02 & -0.03 & 0.06 & 0.01 & .402\\
-\ \ \ Twitter & -0.02 & -0.09 & 0.05 & -0.01 & .420\\
+\ \ \ Reading & 0.01 & -0.03 & 0.05 & 0.01 & .639\\
+\ \ \ Liking \& Sharing & -0.02 & -0.06 & 0.01 & -0.01 & .227\\
+\ \ \ Posting & -0.04 & -0.08 & -0.01 & -0.02 & .025\\
+\ \ \ Facebook & -0.01 & -0.04 & 0.02 & -0.01 & .527\\
+\ \ \ Instagram & 0.03 & -0.01 & 0.07 & 0.02 & .149\\
+\ \ \ WhatsApp & 0.00 & -0.04 & 0.04 & 0.00 & .917\\
+\ \ \ YouTube & 0.01 & -0.03 & 0.04 & 0.00 & .713\\
+\ \ \ Twitter & -0.01 & -0.05 & 0.03 & 0.00 & .503\\
 Positive affect &  &  &  &  & \\
-\ \ \ Reading & -0.03 & -0.05 & -0.01 & -0.04 & .011\\
-\ \ \ Liking \& Sharing & 0.01 & -0.02 & 0.04 & 0.01 & .535\\
-\ \ \ Posting & 0.01 & -0.02 & 0.05 & 0.01 & .404\\
-\ \ \ Facebook & -0.01 & -0.04 & 0.02 & -0.01 & .581\\
-\ \ \ Instagram & 0.00 & -0.02 & 0.02 & 0.00 & .704\\
-\ \ \ WhatsApp & 0.01 & -0.02 & 0.04 & 0.01 & .444\\
-\ \ \ YouTube & 0.00 & -0.04 & 0.04 & 0.00 & .904\\
-\ \ \ Twitter & 0.01 & 0.00 & 0.02 & 0.01 & .129\\
+\ \ \ Reading & -0.03 & -0.05 & -0.02 & -0.04 & < .001\\
+\ \ \ Liking \& Sharing & 0.01 & -0.01 & 0.02 & 0.01 & .508\\
+\ \ \ Posting & 0.02 & 0.01 & 0.04 & 0.02 & .003\\
+\ \ \ Facebook & 0.00 & -0.02 & 0.01 & 0.00 & .671\\
+\ \ \ Instagram & -0.01 & -0.02 & 0.01 & -0.01 & .390\\
+\ \ \ WhatsApp & 0.01 & -0.01 & 0.02 & 0.01 & .374\\
+\ \ \ YouTube & 0.00 & -0.01 & 0.02 & 0.00 & .686\\
+\ \ \ Twitter & 0.01 & 0.00 & 0.03 & 0.01 & .130\\
 Negative affect &  &  &  &  & \\
-\ \ \ Reading & 0.00 & -0.01 & 0.01 & 0.00 & .971\\
-\ \ \ Liking \& Sharing & 0.02 & 0.00 & 0.03 & 0.02 & .019\\
-\ \ \ Posting & 0.05 & 0.02 & 0.08 & 0.04 & .009\\
-\ \ \ Facebook & -0.01 & -0.02 & 0.01 & -0.01 & .226\\
-\ \ \ Instagram & 0.00 & -0.01 & 0.01 & 0.00 & .371\\
-\ \ \ WhatsApp & 0.00 & -0.01 & 0.02 & 0.01 & .562\\
-\ \ \ YouTube & 0.01 & -0.01 & 0.04 & 0.02 & .250\\
-\ \ \ Twitter & 0.01 & 0.00 & 0.03 & 0.01 & .034\\
+\ \ \ Reading & 0.00 & -0.01 & 0.02 & 0.00 & .747\\
+\ \ \ Liking \& Sharing & 0.02 & 0.00 & 0.04 & 0.02 & .022\\
+\ \ \ Posting & 0.05 & 0.04 & 0.07 & 0.05 & < .001\\
+\ \ \ Facebook & 0.00 & -0.01 & 0.01 & 0.00 & .710\\
+\ \ \ Instagram & 0.00 & -0.02 & 0.01 & 0.00 & .654\\
+\ \ \ WhatsApp & 0.00 & -0.01 & 0.01 & 0.01 & .417\\
+\ \ \ YouTube & 0.01 & 0.00 & 0.02 & 0.02 & .011\\
+\ \ \ Twitter & 0.02 & 0.01 & 0.04 & 0.02 & .008\\
 \bottomrule
@@
 
-The results showed that several effects crossed or fell completely outside of the SESOI.
-They can hence be considered meaningful.
-For example, if physical health decreased, this had a meaningful detrimental impact on life satisfaction (\(\beta\) = .18 {[}95\% CI .16, .20{]}), positive affect (\(\beta\) = .17 {[}95\% CI .16, .18{]}), and negative affect (\(\beta\) = -.18 {[}95\% CI -.19, -.17{]}).
-Spending more time outside to exercise meaningfully increased positive affect (\(\beta\) = .12 {[}95\% CI .11, .13{]}).
-Being more satisfied with the current democratic system meaningfully increased life satisfaction (\(\beta\) = .13 {[}95\% CI .12, .14{]}).
+The results showed that several effects crossed or fell completely outside of the SESOI, and can hence be considered meaningful.
+For example, if physical health decreased, this had a meaningful detrimental impact on life satisfaction (\(\beta\) = .19 {[}95\% CI .18, .20{]}), positive affect (\(\beta\) = .18 {[}95\% CI .17, .19{]}), and negative affect (\(\beta\) = -.19 {[}95\% CI -.20, -.18{]}).
+Spending more time outside to exercise meaningfully increased positive affect (\(\beta\) = .12 {[}95\% CI .11, .14{]}).
 The strongest aspect affecting well-being was internal locus of control.
-If people felt more in control of their lives, this strongly increased both life satisfaction (\(\beta\) = .31 {[}95\% CI .28, .34{]}) and
-positive affect (\(\beta\) = .27 {[}95\% CI .25, .29{]}),
-while decreasing negative affect (\(\beta\) = -.28 {[}95\% CI -.30, -.26{]}).
-
+If people felt more in control of their lives, this strongly increased both life satisfaction (\(\beta\) = .33 {[}95\% CI .31, .35{]}) and
+positive affect (\(\beta\) = .28 {[}95\% CI .27, .30{]}),
+while decreasing negative affect (\(\beta\) = -.29 {[}95\% CI -.31, -.27{]}).
 For an overview, see Figure \ref{fig:fig-control}.
@@
 \begin{figure}
-\includegraphics[width=\textwidth]{manuscript_files/figure-latex/fig-control-1} \caption{Results of main variables together with covariates to provide context.
+\includegraphics[width=\textwidth]{figures/fig_results_comp_std} \caption{Results of main variables together with covariates to provide context.
 All variables standardized.
@@
 
+Because life satisfaction is more stable than affect, the effects of communication might materialize some time later.
+I hence also tested the effects across the longer intervals of one month and four months.
+Results showed that all effects disappeared.
+No effect remained significant, implying that at least in this case in this case effects take place on a shorter interval.
+
+Finally, as suggested by the differential susceptibility of media effects model, media effects can depend on dispositional factors, developmental stages, or cultural norms (Valkenburg \& Peter, 2013), such as gender and age (Orben et al., 2022).
+I hence reran the analyses, differentiating effects for boys and girls and for age cohorts.
+The results showed that effects did not differ across genders.
+The effects also did not depend on age.
+However, one effect stood out and was significant.
+Compared to the middle age category Generation X, results showed that if users from Generation Z posted more COVID-19 content than usual this lead to significantly more negative affect (\(\beta\) = .04 {[}95\% CI .01, .06{]}).
+
 \hypertarget{discussion}{%
@@
 
-Based on a panel study with 32 waves largely representative of the Austrian population, this study analyzed the effects of COVID-19 related social media use on well-being.
+Based on a panel study with 34 waves largely representative of the Austrian population, this study analyzed the effects of COVID-19 related social media use on well-being.
 Between person correlation analyses showed that more active users of COVID-19 related content on social media also reported decreased well-being.
-The within-person relations, which are more informative regarding causal effects, showed a different pattern, however.
+For example, respondents who read more COVID-19 related content than others reported slightly lower levels of life satisfaction, somewhat lower levels of positive affect, and substantially higher levels of negative affect than others.
+To see if these between person correlations would translate to within-person effects, I analyzed if changes in a person's media use led to changes in their well-being.
+The within-person relations showed a different pattern.
 If people consumed more COVID-19 content on social media than usual, this did not meaningfully reduce their well-being.
-Several statistically significant effects were found, but these were very small.
-For example, people who liked and shared more COVID-19 related posts than usual reported slightly higher levels of negative affect.
-Posting more content about COVID-19 than usual slightly increased negative affect.
-People who read more COVID-19 related posts than usual reported slightly decreased positive affect.
-Using Twitter for COVID-19 related content slightly increased negative affect.
+Although several statistically significant effects were found, these were very small.
+For example, people who read more COVID-19 related posts than usual reported slightly decreased positive affect.
+People who liked and shared more COVID-19 related posts than usual reported slightly higher levels of negative affect.
+Posting more content about COVID-19 than usual slightly decreased life satisfaction, while increasing both negative affect and positive affect.
+Using Twitter for COVID-19 related content slightly increased negative affect, as did YouTube.
 Again, although all of these within-person effects were statistically significant, they were very small, smaller than the predefined smallest effect size of interest.
 According to the preregistered procedure, they should hence be considered irrelevant.
-Other factors, for which we would expect to find meaningful effects, such as health or sports, indeed showed substantial and meaningful impacts on well-being.
+Additional analyses revealed that other factors, for which we would expect to find meaningful effects, such as health or sports, indeed showed substantial and meaningful impacts on well-being.
+In addition, when testing for the longer intervals of one month and four months, again no meaningful effects were found.
 In conclusion, COVID-19 related activity on social media was not a particularly strong influence on peoples' well-being.
-The results do not support the popular fears that ``doomscrolling'' or overusing social media during times of crises constitutes a relevant risk for well-being.
+The results do not support the popular fears that ``doomscrolling'' or overusing social media during times of crises constitutes a prominent risk for well-being.
 
@@
 First, overall the results do suggest that effects of COVID-19 related social media use on well-being tend to take place in the negative as opposed to the positive spectrum.
-Although very small, four statistically significant negative results of COVID-19 related social media use on well-being were found.
-Not a single positive effect emerged.
+Although very small, five statistically significant negative results of COVID-19 related social media use on well-being were found.
+Only one positive effect emerged.
 Also note that in the analyses several control variables were included, ruling out plausible alternative explanations for the negative results.
-For example, it was controlled for whether or not participants contracted a COVID-19 infection during a specific wave.
+For example, it was controlled for as to whether or not participants contracted a COVID-19 infection during a specific wave.
 Hence, we can rule out the alternative explanation that having an infection was the root cause of increased communication and reduced well-being.
 
-Second, significant outcomes emerged for positive or negative affect, but not for life satisfaction.
-Life satisfaction was stable and not affected by any type or channel of social media communication.
-The more fluctuating positive and negative affect, however, were affected.
-Liking, sharing, and posting COVID-19 related content, and spending time on Twitter to browse COVID-19 related content, all negatively influenced affect.
-This is aligned with prior findings which showed that social media can trigger negative affect but does not reduce life satisfaction (Huang, 2017).
-Conversations about COVID-19 on social media are often extreme, negative, or aggressive (Fan et al., 2020).
+Second, six significant outcomes emerged for positive or negative affect, but only one for life satisfaction.
+Life satisfaction is more stable and not that easily affected by any type or channel of social media communication.
+The more fluctuating positive and negative affect, however, were affected (albeit only slightly).
+Liking, sharing, and posting COVID-19 related content, and spending more time on Twitter and YouTube to browse COVID-19 related content, all slightly negatively influenced affect.
+This is aligned with prior findings which showed that social media use can trigger negative affect, but that it is less likely to determine life satisfaction (Huang, 2017).
+Conversations about COVID-19 on social media are often extreme, negative, or aggressive (L. Fan et al., 2020).
 More deeply engaging with this type of content could negatively affect active authors.
-The hypothesis that tonality could explain the negative effects is especially supported by the observation that spending more time on Twitter than usual increased negative affect.
-Communication on Twitter is often found to be more negative and impolite compared to other SNSs (Halpern \& Gibbs, 2013).
+The hypothesis that tonality could explain the negative effects is especially supported by the observation that spending more time on Twitter and YouTube than usual increased negative affect.
+Communication on both channels is often found to be negative and impolite (e.g., Mueller \& Saeltzer, 2022), also when compared to other SNSs (Halpern \& Gibbs, 2013).
 Consuming more negative and misleading information could hence explain the (slightly) increased levels of negative affect.
@@
 Third, the results show that it makes sense to analyze different communication types and communication channels.
-Reading reduced positive affect, while liking, sharing and commenting increased negative affect.
-Whereas it was often stated that passive use is bad and active use good (Verduyn et al., 2015), this pattern thus did not emerge here.
-Instead, all three types were negative.
-The results hence support the findings from Valkenburg et al. (2022), who also could not confirm the claim that active use is good and that passive use is bad.
-Focusing on communication channels, Twitter seems to be more negative, as has often been observed (Halpern \& Gibbs, 2013).
+Reading slightly reduced positive affect, while liking, sharing, and posting slightly increased negative affect.
+Interestingly, posting COVID-19 related comment slightly increased negative affect, while at the same time it also slightly increased positive affect.
+Posting content is often met with strong reaction, both positive by means of likes and negative by means of critical comments.
+Overall, though, posting led to slightly reduced levels of life satisfaction.
+In conclusion, whereas it was often stated that passive use is bad and active use good (Verduyn et al., 2015), this pattern was only partially found here.
+The results are aligned with the findings from Valkenburg et al. (2022), who could not confirm that active use is good and that passive use is bad.
+Focusing on communication channels, Twitter and YouTube seem to be more negative, as has often been observed (Halpern \& Gibbs, 2013), while Instagram, WhatsApp, and Facebook were neutral.
 But, again, all of these effects are very small.
@@
 Taken together, the results are hence aligned with the underlying theoretical models and prior empirical results.
-The findings support the Different Susceptibility to Media Effects Theory, which states that effects are generally small and depend on the type of communication.
-The results are well-aligned with mood management theory (Zillmann, 1988) or the uses and gratifications approach (Katz et al., 1973), whose premises preclude particularly negative effects of routine and widespread media consumption.
-If the effects of social media were indeed profoundly negative on average, then people likely would not spend so much time on social media engaging with COVID-19 content.
-Likewise, recent empirical studies and meta-analyses reported rather small negative effects, too.
+The findings support the differential susceptibility of media effects model (Valkenburg \& Peter, 2013), such that effects are generally small and that they depend on the type and channel of communication.
+Additional analyses did not reveal that effects depended on gender.
+Age also large did not play a significant moderation role, but effects of posting COVID-19 related content were found to be more negative for Generation Z.
+Indeed, it has often been argued that effects of social media use are more negative for Gen Z than for prior generations, and this finding can be seen a further tentative support for this hypothesis.
+From a broader perspective, the results are well-aligned with mood management theory (Zillmann, 1988) and the uses and gratifications approach (Katz et al., 1973), whose premises preclude particularly negative effects of routine and widespread media consumption.
+Both theories posit that if the effects of social media were indeed profoundly negative on average, then people likely would not spend so much time on social media engaging with COVID-19 content.
+Finally, recent empirical studies and meta-analyses reported rather small negative effects, too.
 Several studies found that the effects of various types of social media use on well-being are small, often too small to matter (Bendau et al., 2021; Ferguson et al., 2021; Meier \& Reinecke, 2020; Orben, 2020), echoing the results obtained here.
-
-In light of the overall very small effects, engaging in COVID 19-related social media use should not be a major concern for one's well-being.
-That said, the results still imply that it does make sense to critically reflect upon COVID-19 related social media use.
-On average, it is likely more beneficial to post less actively about COVID-19 on social media and to spend less time on Twitter consuming COVID-19 related content.
-The results allow us to reject a positive effect: Writing more posts on social media will likely not increase well-being.
 
@@
 Focusing on within-person effects and controlling for several potential confounders, this study provides an improved perspective on assessing causality.
-However, three major challenges remain.
-First, it could be that the timing of the interval was wrong; it could be that effects unfold in shorter or longer intervals.
-Next, it could be that there was not sufficient variance in media use and well-being.
-Without sufficient variability, statistical models cannot detect existing effects.
-Third, it is necessary to control for \emph{all} relevant confounding third variables.
-Although this study included are large list of confounders, it could still be that crucial variables were missed (Zillmann, 1988).
-To borrow the words from Rohrer and Murayama (2021), there is no such thing as ``the'' effect of social media use on well-being.
-The results reported here are hence contingent on the design I used.
-To document how effects unfold, future research needs to employ different study designs probing different intervals.
-More thought needs to be invested in what factors to control for, and I hope this study provides a first step into this direction.
+However, several challenges remain.
+In order to correctly establish causality in non-experimental designs, it is necessary to control for \emph{all} relevant confounding third variables (Rohrer, 2018).
+Although this study included are large list of confounders, it could still be that crucial variables were missed.
+More thought needs to be invested in which factors to control for and, equally important, for which factors not to control for.
+I hope this study provides a first step into this direction.
 
@@
 \begin{CSLReferences}{1}{0}
+\leavevmode\vadjust pre{\hypertarget{ref-atheyDigitalPublicHealth2023}{}}%
+Athey, S., Grabarz, K., Luca, M., \& Wernerfelt, N.
+(2023).
+Digital public health interventions at scale: {The} impact of social media advertising on beliefs and outcomes related to {COVID} vaccines.
+\emph{Proceedings of the National Academy of Sciences}, \emph{120}(5), e2208110120.
+
 \leavevmode\vadjust pre{\hypertarget{ref-bellFixedRandomEffects2019}{}}%
-Bell, A., Fairbrother, M., \& Jones, K.
-(2019).
-Fixed and random effects models: Making an informed choice.
-\emph{Quality \& Quantity}, \emph{53}(2), 1051--1074.
-\url{https://doi.org/10.1007/s11135-018-0802-x}
+Bell, A., Fairbrother, M., \& Jones, K. (2019). Fixed and random effects models: Making an informed choice. \emph{Quality \& Quantity}, \emph{53}(2), 1051--1074.
 
@@
 \emph{European Archives of Psychiatry and Clinical Neuroscience}, \emph{271}(2), 283--291.
-\url{https://doi.org/10.1007/s00406-020-01171-6}
 
@@
 \emph{Communication Research}.
-\url{https://doi.org/10.1177/00936502211038196}
 
 \leavevmode\vadjust pre{\hypertarget{ref-bradleyStressMoodSmartphone2021}{}}%
-Bradley, A., \& Howard, A. (2021). \emph{Stress, mood, and smartphone use in {University} students: {A} 12-week longitudinal study}. \url{https://doi.org/10.31219/osf.io/frvpb}
-
-\leavevmode\vadjust pre{\hypertarget{ref-buchiDigitalWellbeingTheory2021}{}}%
-Büchi, M. (2021). Digital well-being theory and research. \emph{New Media \& Society}, 146144482110568. \url{https://doi.org/10.1177/14614448211056851}
-
-\leavevmode\vadjust pre{\hypertarget{ref-choiMediatedCommunicationMatters2021}{}}%
-Choi, M., \& Choung, H.
+Bradley, A., \& Howard, A. (2021). \emph{Stress, mood, and smartphone use in {University} students: {A} 12-week longitudinal study}. {OSF Preprints}.
+
+\leavevmode\vadjust pre{\hypertarget{ref-cinelliEchoChamberEffect2021}{}}%
+Cinelli, M., De Francisci Morales, G., Galeazzi, A., Quattrociocchi, W., \& Starnini, M.
 (2021).
-Mediated communication matters during the {COVID-19} pandemic: {The} use of interpersonal and masspersonal media and psychological well-being.
-\emph{Journal of Social and Personal Relationships}, \emph{38}(8), 2397--2418.
-\url{https://doi.org/10.1177/02654075211029378}
+The echo chamber effect on social media.
+\emph{Proceedings of the National Academy of Sciences}, \emph{118}(9), e2023301118.
 
 \leavevmode\vadjust pre{\hypertarget{ref-dienerAdvancesOpenQuestions2018}{}}%
-Diener, E., Lucas, R.
-E., \& Oishi, S.
-(2018).
-Advances and open questions in the science of subjective well-being.
-\emph{Collabra: Psychology}, \emph{4}(1), 15.
-\url{https://doi.org/10.1525/collabra.115}
+Diener, E., Lucas, R. E., \& Oishi, S. (2018). Advances and open questions in the science of subjective well-being. \emph{Collabra: Psychology}, \emph{4}(1), 15.
 
 \leavevmode\vadjust pre{\hypertarget{ref-dienesUsingBayesGet2014}{}}%
-Dienes, Z. (2014). Using {Bayes} to get the most out of non-significant results. \emph{Frontiers in Psychology}, \emph{5}. \url{https://doi.org/10.3389/fpsyg.2014.00781}
+Dienes, Z. (2014). Using {Bayes} to get the most out of non-significant results. \emph{Frontiers in Psychology}, \emph{5}.
 
 \leavevmode\vadjust pre{\hypertarget{ref-dienlinImpactDigitalTechnology2020}{}}%
-Dienlin, T., \& Johannes, N.
-(2020).
-The impact of digital technology use on adolescent well-being.
-\emph{Dialogues in Clinical Neuroscience}, \emph{22}(2), 135--142.
-\url{https://doi.org/doi:10.31887/DCNS.2020.22.2/tdienlin}
+Dienlin, T., \& Johannes, N. (2020). The impact of digital technology use on adolescent well-being. \emph{Dialogues in Clinical Neuroscience}, \emph{22}(2), 135--142.
 
@@
 \emph{Journal of Computer-Mediated Communication}, \emph{22}(2), 71--87.
-\url{https://doi.org/10.1111/jcc4.12183}
 
 \leavevmode\vadjust pre{\hypertarget{ref-dormannOptimalTimeLags2015}{}}%
-Dormann, C., \& Griffin, M. A. (2015). Optimal time lags in panel studies. \emph{Psychological Methods}, \emph{20}(4), 489--505. \url{https://doi.org/10.1037/met0000041}
+Dormann, C., \& Griffin, M. A. (2015). \href{https://www.ncbi.nlm.nih.gov/pubmed/26322999}{Optimal time lags in panel studies}. \emph{Psychological Methods}, \emph{20}(4), 489--505.
 
 \leavevmode\vadjust pre{\hypertarget{ref-dornemannHowGoodBad2021}{}}%
-Dörnemann, A., Boenisch, N., Schommer, L., Winkelhorst, L., \& Wingen, T.
-(2021).
-\emph{How do good and bad news impact mood during the {Covid-19} pandemic?
-{The} role of similarity}.
-\url{https://doi.org/10.31219/osf.io/sy2kd}
+Dörnemann, A., Boenisch, N., Schommer, L., Winkelhorst, L., \& Wingen, T. (2021). \emph{How do good and bad news impact mood during the {Covid-19} pandemic? {The} role of similarity}. {OSF Preprints}.
+
+\leavevmode\vadjust pre{\hypertarget{ref-duradoniWellbeingSocialMedia2020}{}}%
+Duradoni, M., Innocenti, F., \& Guazzini, A. (2020). Well-being and social media: {A} systematic review of {Bergen Addiction Scales}. \emph{Future Internet}, \emph{12}(2), 24.
 
@@
 \emph{Frontiers in Psychology}, \emph{11}, 577639.
-\url{https://doi.org/10.3389/fpsyg.2020.577639}
+
+\leavevmode\vadjust pre{\hypertarget{ref-egerStatisticalMetaanalysisWellbeing2015}{}}%
+Eger, R. J., \& Maridal, J. H. (2015). A statistical meta-analysis of the wellbeing literature. \emph{International Journal of Wellbeing}, \emph{5}(2), 45--74.
 
 \leavevmode\vadjust pre{\hypertarget{ref-europeansocialsurveyESS9Edition20182021}{}}%
-European Social Survey. (2021). \emph{{ESS9} edition 3.1 - 2018 {Documentation Report}}. \url{https://www.europeansocialsurvey.org/docs/round9/survey/ESS9_data_documentation_report_e03_1.pdf}
+European Social Survey. (2021). \emph{{ESS9} edition 3.1 - 2018 {Documentation Report}}.
+
+\leavevmode\vadjust pre{\hypertarget{ref-fanInformationOverloadWellbeing2021}{}}%
+Fan, J., \& Smith, A. P. (2021). Information overload, wellbeing and {COVID-19}: {A} survey in {China}. \emph{Behavioral Sciences}, \emph{11}(5), 62.
 
@@
 (2020).
-Stigmatization in social media: {Documenting} and analyzing hate speech for {\textsc{COVID}} ‐19 on {Twitter}.
+Stigmatization in social media: {Documenting} and analyzing hate speech for {\textsc{COVID}} -19 on {Twitter}.
 \emph{Proceedings of the Association for Information Science and Technology}, \emph{57}(1).
-\url{https://doi.org/10.1002/pra2.313}
 
@@
 \emph{Professional Psychology: Research and Practice}.
-\url{https://doi.org/10.1037/pro0000426}
 
 \leavevmode\vadjust pre{\hypertarget{ref-funderEvaluatingEffectSize2019}{}}%
-Funder, D.
-C., \& Ozer, D.
-J.
-(2019).
-Evaluating effect size in psychological research: {Sense} and nonsense.
-\emph{Advances in Methods and Practices in Psychological Science}, \emph{2}(2), 156--168.
-\url{https://doi.org/10.1177/2515245919847202}
+Funder, D. C., \& Ozer, D. J. (2019). Evaluating effect size in psychological research: {Sense} and nonsense. \emph{Advances in Methods and Practices in Psychological Science}, \emph{2}(2), 156--168.
+
+\leavevmode\vadjust pre{\hypertarget{ref-galerHowMuchToo2018}{}}%
+Galer, S. S. (2018). \emph{How much is {``too much time''} on social media?} https://www.bbc.com/future/article/20180118-how-much-is-too-much-time-on-social-media.
+
+\leavevmode\vadjust pre{\hypertarget{ref-giumettiCyberbullyingSocialMedia2022}{}}%
+Giumetti, G. W., \& Kowalski, R. M. (2022). Cyberbullying via social media and well-being. \emph{Current Opinion in Psychology}, \emph{45}, 101314.
 
@@
 \emph{International Journal of Environmental Research and Public Health}, \emph{19}(3), 1635.
-\url{https://doi.org/10.3390/ijerph19031635}
+
+\leavevmode\vadjust pre{\hypertarget{ref-hallSocialMediaUse2022}{}}%
+Hall, J. A., \& Liu, D. (2022). Social media use, social displacement, and well-being. \emph{Current Opinion in Psychology}, \emph{46}, 101339.
 
@@
 \emph{Computers in Human Behavior}, \emph{29}(3), 1159--1168.
-\url{https://doi.org/10.1016/j.chb.2012.10.008}
 
@@
 (2017).
-Time spent on social network sites and psychological well-being: {A} meta-analysis.
+\href{https://www.ncbi.nlm.nih.gov/pubmed/28622031}{Time spent on social network sites and psychological well-being: {A} meta-analysis}.
 \emph{Cyberpsychology, Behavior and Social Networking}, \emph{20}(6), 346--354.
-\url{https://doi.org/10.1089/cyber.2016.0758}
+
+\leavevmode\vadjust pre{\hypertarget{ref-huntSocialMediaBased2022}{}}%
+Hunt, I.
+de V., Dunn, T., Mahoney, M., Chen, M., Nava, V., \& Linos, E.
+(2022).
+\href{https://www.ncbi.nlm.nih.gov/pmc/articles/PMC9382165}{A social media-based public health campaign encouraging {COVID-19} vaccination across the {United States}}.
+\emph{American Journal of Public Health}, \emph{112}(9), 1253--1256.
+
+\leavevmode\vadjust pre{\hypertarget{ref-johnhopkinsuniversityCOVID19Map2023}{}}%
+John Hopkins University. (2023). {COVID-19 Map}. In \emph{Johns Hopkins Coronavirus Resource Center}. https://coronavirus.jhu.edu/map.html.
 
 \leavevmode\vadjust pre{\hypertarget{ref-katzUsesGratificationsResearch1973}{}}%
-Katz, E., Blumler, J. G., \& Gurevitch, M. (1973). Uses and {Gratifications Research}. \emph{Public Opinion Quarterly}, \emph{37}(4), 509. \url{https://doi.org/10.1086/268109}
+Katz, E., Blumler, J. G., \& Gurevitch, M. (1973). Uses and {Gratifications Research}. \emph{Public Opinion Quarterly}, \emph{37}(4), 509.
+
+\leavevmode\vadjust pre{\hypertarget{ref-kittelAustrianCoronaPanel2020}{}}%
+Kittel, B., Kritzinger, S., Boomgaarden, H., Prainsack, B., Eberl, J.-M., Kalleitner, F., Lebernegg, N.
+S., Partheymüller, J., Plescia, C., Schiestl, D.
+W., \& Schlogl, L.
+(2020).
+\emph{Austrian {Corona Panel Project} ({SUF} edition)}.
+{AUSSDA}.
 
@@
 \emph{European Political Science}, \emph{20}(2), 318--344.
-\url{https://doi.org/10.1057/s41304-020-00294-7}
-
-\leavevmode\vadjust pre{\hypertarget{ref-kittelAustrianCoronaPanel2020}{}}%
-Kittel, B., Kritzinger, S., Boomgaarden, H., Prainsack, B., Eberl, J.-M., Kalleitner, F., Lebernegg, N.
-S., Partheymüller, J., Plescia, C., Schiestl, D.
-W., \& Schlogl, L.
-(2020).
-\emph{Austrian {Corona Panel Project} ({SUF} edition)} {[}Data set{]}.
-{AUSSDA}.
-\url{https://doi.org/10.11587/28KQNS}
 
 \leavevmode\vadjust pre{\hypertarget{ref-kleinDarklySoothingCompulsion2021}{}}%
-Klein, J. (2021, March 3). \emph{The darkly soothing compulsion of 'doomscrolling'}. \url{https://www.bbc.com/worklife/article/20210226-the-darkly-soothing-compulsion-of-doomscrolling}
+Klein, J. (2021). \emph{The darkly soothing compulsion of 'doomscrolling'}. https://www.bbc.com/worklife/article/20210226-the-darkly-soothing-compulsion-of-doomscrolling.
+
+\leavevmode\vadjust pre{\hypertarget{ref-lakensEquivalenceTestingPsychological2018}{}}%
+Lakens, D., Scheel, A.
+M., \& Isager, P.
+M.
+(2018).
+Equivalence testing for psychological research: {A} tutorial.
+\emph{Advances in Methods and Practices in Psychological Science}, \emph{1}(2), 259--269.
 
@@
 \emph{Social Science \& Medicine}, \emph{293}, 114674.
-\url{https://doi.org/10.1016/j.socscimed.2021.114674}
+
+\leavevmode\vadjust pre{\hypertarget{ref-liYouTubeSourceInformation2020}{}}%
+Li, H.
+O.-Y., Bailey, A., Huynh, D., \& Chan, J.
+(2020).
+\href{https://www.ncbi.nlm.nih.gov/pubmed/32409327}{{YouTube} as a source of information on {COVID-19}: A pandemic of misinformation?} \emph{BMJ Global Health}, \emph{5}(5), e002604.
 
@@
 \emph{Journal of Medical Internet Research}, \emph{22}(9), e22142.
-\url{https://doi.org/10.2196/22142}
 
 \leavevmode\vadjust pre{\hypertarget{ref-lucasAdaptationSetpointModel2007}{}}%
-Lucas, R.
-E.
-(2007).
-Adaptation and the set-point model of subjective well-being.
-\emph{Current Directions in Psychological Science}, \emph{16}(2), 75--79.
-\url{https://doi.org/10.1111/j.1467-8721.2007.00479.x}
+Lucas, R. E. (2007). Adaptation and the set-point model of subjective well-being. \emph{Current Directions in Psychological Science}, \emph{16}(2), 75--79.
 
 \leavevmode\vadjust pre{\hypertarget{ref-lucasWhyCrossLaggedPanel2022}{}}%
-Lucas, R. E. (2022). \emph{Why the {Cross-Lagged Panel Model} is {Almost Never} the {Right Choice}} {[}Preprint{]}. {PsyArXiv}. \url{https://doi.org/10.31234/osf.io/pkec7}
+Lucas, R. E. (2022). \emph{Why the {Cross-Lagged Panel Model} is {Almost Never} the {Right Choice}} {[}Preprint{]}. {PsyArXiv}.
 
@@
 \leavevmode\vadjust pre{\hypertarget{ref-marcianoDynamicsAdolescentsSmartphone2022}{}}%
-Marciano, L., Driver, C.
-C., Schulz, P.
-J., \& Camerini, A.-L.
+Marciano, L., Driver, C. C., Schulz, P. J., \& Camerini, A.-L. (2022). Dynamics of adolescents' smartphone use and well-being are positive but ephemeral. \emph{Scientific Reports}, \emph{12}(1), 1316.
+
+\leavevmode\vadjust pre{\hypertarget{ref-meierComputermediatedCommunicationSocial2020a}{}}%
+Meier, A., \& Reinecke, L. (2020). Computer-mediated communication, social media, and mental health: {A} conceptual and empirical meta-review. \emph{Communication Research}, 009365022095822.
+
+\leavevmode\vadjust pre{\hypertarget{ref-muellerTwitterMadeMe2022}{}}%
+Mueller, S.
+D., \& Saeltzer, M.
 (2022).
-Dynamics of adolescents' smartphone use and well-being are positive but ephemeral.
-\emph{Scientific Reports}, \emph{12}(1), 1316.
-\url{https://doi.org/10.1038/s41598-022-05291-y}
-
-\leavevmode\vadjust pre{\hypertarget{ref-meierDoesPassiveSocial2022}{}}%
-Meier, A., \& Krause, H.-V. (2022). \emph{Does passive social media use harm well-being? {An} adversarial review} {[}Preprint{]}. {PsyArXiv}. \url{https://doi.org/10.31234/osf.io/nvbwh}
-
-\leavevmode\vadjust pre{\hypertarget{ref-meierComputermediatedCommunicationSocial2020a}{}}%
-Meier, A., \& Reinecke, L.
-(2020).
-Computer-mediated communication, social media, and mental health: {A} conceptual and empirical meta-review.
-\emph{Communication Research}, 009365022095822.
-\url{https://doi.org/10.1177/0093650220958224}
+Twitter made me do it!
+{Twitter}'s tonal platform incentive and its effect on online campaigning.
+\emph{Information, Communication \& Society}, \emph{25}(9), 1247--1272.
 
@@
 \emph{Medical Care}, \emph{41}(5), 582--592.
-\href{Retrieved\%20from\%20http://www.jstor.org/stable/3768017}{Retrieved from http://www.jstor.org/stable/3768017}
 
 \leavevmode\vadjust pre{\hypertarget{ref-orbenTeenagersScreensSocial2020}{}}%
-Orben, A.
-(2020).
-Teenagers, screens and social media: A narrative review of reviews and key studies.
-\emph{Social Psychiatry and Psychiatric Epidemiology}, \emph{55}(4), 407--414.
-\url{https://doi.org/10.1007/s00127-019-01825-4}
+Orben, A. (2020). Teenagers, screens and social media: A narrative review of reviews and key studies. \emph{Social Psychiatry and Psychiatric Epidemiology}, \emph{55}(4), 407--414.
+
+\leavevmode\vadjust pre{\hypertarget{ref-orbenWindowsDevelopmentalSensitivity2022}{}}%
+Orben, A., Przybylski, A. K., Blakemore, S.-J., \& Kievit, R. A. (2022). Windows of developmental sensitivity to social media. \emph{Nature Communications}, \emph{13}(1), 1649.
 
@@
 \emph{Journal of Research in Interactive Marketing}, \emph{14}(2), 269--284.
-\url{https://doi.org/10.1108/JRIM-10-2019-0159}
+
+\leavevmode\vadjust pre{\hypertarget{ref-przybylskiMotivationalEmotionalBehavioral2013}{}}%
+Przybylski, A.
+K., Murayama, K., DeHaan, C.
+R., \& Gladwell, V.
+(2013).
+Motivational, emotional, and behavioral correlates of fear of missing out.
+\emph{Computers in Human Behavior}, \emph{29}(4), 1841--1848.
 
@@
 (2020).
-Associations between media exposure and mental distress among {U}.{S}.
-Adults at the beginning of the {COVID-19} pandemic.
+\href{https://www.ncbi.nlm.nih.gov/pubmed/33011008}{Associations between media exposure and mental distress among {U}.{S}.
+Adults at the beginning of the {COVID-19} pandemic}.
 \emph{American Journal of Preventive Medicine}, \emph{59}(5), 630--638.
-\url{https://doi.org/10.1016/j.amepre.2020.06.008}
 
@@
 \emph{Advances in Methods and Practices in Psychological Science}, \emph{24}(2), 251524591774562.
-\url{https://doi.org/10.1177/2515245917745629}
 
 \leavevmode\vadjust pre{\hypertarget{ref-rohrerTheseAreNot2021}{}}%
-Rohrer, J.
-M., \& Murayama, K.
-(2021).
-\emph{These are not the effects you are looking for: {Causality} and the within-/between-person distinction in longitudinal data analysis}.
-{PsyArXiv}.
-\url{https://doi.org/10.31234/osf.io/tg4vj}
+Rohrer, J. M., \& Murayama, K. (2021). \emph{These are not the effects you are looking for: {Causality} and the within-/between-person distinction in longitudinal data analysis}. {PsyArXiv}.
 
@@
 Sandstrom, G., Buchanan, K., Aknin, L., \& Lotun, S.
-(2021, October 22).
-\emph{Doomscrolling {COVID} news takes an emotional toll -- here's how to make your social media a happier place}.
-{The Conversation}.
-\url{http://theconversation.com/doomscrolling-covid-news-takes-an-emotional-toll-heres-how-to-make-your-social-media-a-happier-place-170342}
+(2021).
+Doomscrolling {COVID} news takes an emotional toll \textendash{} here's how to make your social media a happier place.
+In \emph{The Conversation}.
+http://theconversation.com/doomscrolling-covid-news-takes-an-emotional-toll-heres-how-to-make-your-social-media-a-happier-place-170342.
 
 \leavevmode\vadjust pre{\hypertarget{ref-scharkowAccuracySelfreportedInternet2016}{}}%
-Scharkow, M.
-(2016).
-The accuracy of self-reported {Internet} use---{A} validation study using client log data.
-\emph{Communication Methods and Measures}, \emph{10}(1), 13--27.
-\url{https://doi.org/10.1080/19312458.2015.1118446}
+Scharkow, M. (2016). The accuracy of self-reported {Internet} use\textemdash{{A}} validation study using client log data. \emph{Communication Methods and Measures}, \emph{10}(1), 13--27.
 
@@
 \emph{Journal of Affective Disorders}, \emph{288}, 145--147.
-\url{https://doi.org/10.1016/j.jad.2021.04.008}
+
+\leavevmode\vadjust pre{\hypertarget{ref-sharmaDarkEndTunnel2022}{}}%
+Sharma, B., Lee, S. S., \& Johnson, B. K. (2022). The dark at the end of the tunnel: {Doomscrolling} on social media newsfeeds. \emph{Technology, Mind, and Behavior}, \emph{3}(1).
 
 \leavevmode\vadjust pre{\hypertarget{ref-sheldonStabilityHappinessTheories2014}{}}%
-Sheldon, K. M., \& Lucas, R. E. (2014). \emph{Stability of happiness: Theories and evidence on whether happiness can change}. \url{http://site.ebrary.com/id/10891875}
+Sheldon, K. M., \& Lucas, R. E. (2014). \emph{Stability of happiness: Theories and evidence on whether happiness can change}.
 
 \leavevmode\vadjust pre{\hypertarget{ref-stainbackCOVID1924News2020}{}}%
-Stainback, K., Hearne, B.
-N., \& Trieu, M.
-M.
-(2020).
-{COVID-19} and the 24/7 {News Cycle}: {Does COVID-19 News Exposure Affect Mental Health}?
-\emph{Socius}, \emph{6}, 2378023120969339.
-\url{https://doi.org/10.1177/2378023120969339}
+Stainback, K., Hearne, B. N., \& Trieu, M. M. (2020). {COVID-19} and the 24/7 {News Cycle}: {Does COVID-19 News Exposure Affect Mental Health}? \emph{Socius}, \emph{6}, 2378023120969339.
 
@@
 Statista.
-(2021, May 21).
+(2021).
 \emph{Average daily time spent on social networks by users in the {United States} from 2018 to 2022}.
-\url{https://www.statista.com/statistics/1018324/us-users-daily-social-media-minutes/}
+https://www.statista.com/statistics/1018324/us-users-daily-social-media-minutes/.
+
+\leavevmode\vadjust pre{\hypertarget{ref-twitterTalkingMentalHealthAwarenessTwitter2020}{}}%
+Twitter. (2020). \emph{Talking \#{MentalHealthAwareness} on {Twitter} with our global partners}. https://blog.twitter.com/en\_us/topics/company/2020/talking-mental-health-awareness-on-twitter.
 
 \leavevmode\vadjust pre{\hypertarget{ref-valkenburgDifferentialSusceptibilityMedia2013}{}}%
-Valkenburg, P. M., \& Peter, J. (2013). The differential susceptibility to media effects model. \emph{Journal of Communication}, \emph{63}(2), 221--243. \url{https://doi.org/10.1111/jcom.12024}
+Valkenburg, P. M., \& Peter, J. (2013). The differential susceptibility to media effects model. \emph{Journal of Communication}, \emph{63}(2), 221--243.
 
@@
 \emph{New Media \& Society}, \emph{24}(2), 530--549.
-\url{https://doi.org/10.1177/14614448211065425}
+
+\leavevmode\vadjust pre{\hypertarget{ref-vanrooijWeakScientificBasis2018}{}}%
+van Rooij, A.
+J., Ferguson, C.
+J., Colder Carras, M., Kardefelt-Winther, D., Shi, J., Aarseth, E., Bean, A.
+M., Bergmark, K.
+H., Brus, A., Coulson, M., Deleuze, J., Dullur, P., Dunkels, E., Edman, J., Elson, M., Etchells, P.
+J., Fiskaali, A., Granic, I., Jansz, J., \ldots{} Przybylski, A.
+K.
+(2018).
+A weak scientific basis for gaming disorder: {Let} us err on the side of caution.
+\emph{Journal of Behavioral Addictions}, \emph{7}(1), 1--9.
 
 \leavevmode\vadjust pre{\hypertarget{ref-verbeijSelfreportedMeasuresSocial2021}{}}%
-Verbeij, T., Pouwels, J.
-L., Beyens, I., \& Valkenburg, P.
-M.
-(2021).
-\emph{Self-reported measures of social media use show high predictive validity} {[}Preprint{]}.
-{PsyArXiv}.
-\url{https://doi.org/10.31234/osf.io/c9bj7}
+Verbeij, T., Pouwels, J. L., Beyens, I., \& Valkenburg, P. M. (2021). \emph{Self-reported measures of social media use show high predictive validity} {[}Preprint{]}. {PsyArXiv}.
 
@@
 (2015).
-Passive {Facebook} usage undermines affective well-being: {Experimental} and longitudinal evidence.
+\href{https://www.ncbi.nlm.nih.gov/pubmed/25706656}{Passive {Facebook} usage undermines affective well-being: {Experimental} and longitudinal evidence}.
 \emph{Journal of Experimental Psychology: General}, \emph{144}(2), 480--488.
-\url{https://doi.org/10.1037/xge0000057}
 
@@
 (2018).
-\emph{{AUTNES Online Panel Study} 2017 ({SUF} edition)} {[}Data set{]}.
+\emph{{AUTNES Online Panel Study} 2017 ({SUF} edition)}.
 {AUSSDA}.
-\url{https://doi.org/10.11587/I7QIYJ}
+
+\leavevmode\vadjust pre{\hypertarget{ref-wardOriginsHappinessEvidence2016}{}}%
+Ward, G., Powdthavee, N., Fleche, S., Layard, R., \& Clark, A.
+(2016).
+Origins of happiness: {Evidence} and policy implications.
+In \emph{CEPR}.
+https://cepr.org/voxeu/columns/origins-happiness-evidence-and-policy-implications.
 
@@
 World Health Organization. (1998). \emph{Wellbeing measures in primary health care/{The Depcare Project}}.
+
+\leavevmode\vadjust pre{\hypertarget{ref-worldhealthorganizationCoronavirusDiseaseCOVID192023}{}}%
+World Health Organization. (2023). \emph{Coronavirus disease ({COVID-19}) \textendash{} {World Health Organization}}. https://www.who.int/emergencies/diseases/novel-coronavirus-2019.
+
+\leavevmode\vadjust pre{\hypertarget{ref-yangCanWatchingOnline2021}{}}%
+Yang, Z., Griffiths, M.
+D., Yan, Z., \& Xu, W.
+(2021).
+Can watching online videos be addictive?
+{A} qualitative exploration of online video watching among {Chinese} young adults.
+\emph{International Journal of Environmental Research and Public Health}, \emph{18}(14), 7247.
 
@@
 \emph{Computers in Human Behavior}, \emph{127}, 107050.
-\url{https://doi.org/10.1016/j.chb.2021.107050}
 
 \leavevmode\vadjust pre{\hypertarget{ref-zillmannMoodManagementCommunication1988}{}}%
-Zillmann, D. (1988). Mood {Management Through Communication Choices}. \emph{American Behavioral Scientist}, \emph{31}(3), 327340. \url{https://doi.org/10.1177/000276488031003005}
+Zillmann, D. (1988). Mood {Management Through Communication Choices}. \emph{American Behavioral Scientist}, \emph{31}(3), 327340.
 

### 2023-07-11 preregistration_changes.Rmd
Major updates Implemented feedback

@@
 
+## Wording Hypotheses
+
+Before: Hypothesis: The within-person effects of all types of COVID-19 related social media use on all types of well-being indicators---while controlling for several stable and varying covariates such as sociodemographic variables and psychological dispositions---will be trivial.
+
+Now: "Hypothesis: The within-person effects of all measures of COVID-19 related social media use (types: reading, liking and sharing, posting; channels: Twitter, Instagram, Facebook, YouTube, WhatsApp) on all measures of well-being indicators (positive affect, negative affect, life satisfaction)---while controlling for several stable and varying covariates such as sociodemographic variables and psychological dispositions (see below)---will be trivial.
+
+Explanation: The wording of the hypotheses was changed slightly.
+Instead of "trivial", it not states "very small".
+The reason was that trivial contains a normative element (irrelevant).
+Instead, I now state "very small", which seems more objective.
+Some might consider very small effects trivial, others not.
+
+In addition, I link to the control variables explicated below.
+
 ## Covariates
@@
 
-This includes (a) text-based media news consumption, (b) video-based media news consumption, (c) residency is Vienna, (d) self-reported physical health, (e) living space (in squaremeter), (f) access to balcony, (g) access to garden, (h) employment status, (i) disposition to take risks, and (j) locus of control.
+This includes residency is Vienna, self-reported physical health, living space (in squaremeter), access to balcony, access to garden, and employment status.
 
-The following variables I originally planned to include as control variables.
-However, their inclusion led to the deletion of too many cases/observations.
-I hence decided against their inclusion.
-This includes (a) household size, (b) work hours per week, (c) home office, (d) household income.
+I additionally planned to control for satisfaction with democracy and other types of media use.
+However, it's likely that they also serve as mediators, which is why I will not include them in the main models.
+They will, however, be reported in the supplementary analyses.
 
@@
 
-Reviewer feedback required changes regarding the imputation strategy
+After having received feedback from colleagues who are experts on missing data analyses, I changed the imputation strategy.
 
@@
 - Before: remove participants with > 50% missing data; now: include all participants
-- Before: remove variables leading to much attrition; now: include all variables
 

### 2023-07-21 manuscript.Rmd
- reduced in length

@@
 It is therefore often asked whether social media can become addictive, and users sometimes express this fear themselves [@yangCanWatchingOnline2021].
-However, a recently published meta-analysis found that the two most prominent measures of addiction, the Bergen Facebook Addiction Scale and the Bergen Social Media Addiction Scale, have only small relations to well-being [@duradoniWellbeingSocialMedia2020].
-In addition, the general idea of labeling excessive social and new media use as addiction was criticized, arguing that social media use represent a new regular behavior that should not be pathologized [@galerHowMuchToo2018; @vanrooijWeakScientificBasis2018].
+<!-- However, a recently published meta-analysis found that the two most prominent measures of addiction, the Bergen Facebook Addiction Scale and the Bergen Social Media Addiction Scale, have only small relations to well-being [@duradoniWellbeingSocialMedia2020].
+-->
+<!-- In addition, the general idea of labeling excessive social and new media use as addiction was criticized, arguing that social media use represent a new regular behavior that should not be pathologized [@galerHowMuchToo2018; @vanrooijWeakScientificBasis2018].
+-->
 <!-- In conclusion, there are several plausible theoretical mechanisms as to why effects might be negative or positive, explaining why on average effects might not be particularly negative or positive.
@@
 <!-- At the same time, there are also several plausible reasons as to why COVID-19 related social media use is beneficial for well-being. -->
-Access to information and resources: 
-Despite incorrect information, social media provide a vast platform for disseminating accurate and timely information about COVID-19 [@johnhopkinsuniversityCOVID19Map2023]. 
-It allows individuals to stay informed about the latest updates, guidelines, and recommendations from reputable health organizations and experts. 
+<!-- Access to information and resources:  -->
+Despite incorrect information, social media provide a vast platform for disseminating _accurate and timely information_ about COVID-19 [@johnhopkinsuniversityCOVID19Map2023]. 
+<!-- It allows individuals to stay informed about the latest updates, guidelines, and recommendations from reputable health organizations and experts.  -->
 Access to reliable information can help people make informed decisions, alleviate uncertainties, and feel empowered during the pandemic.
-Educational opportunities: 
-Social media can be a valuable educational tool, with various organizations and experts sharing informative content about COVID-19 [@worldhealthorganizationCoronavirusDiseaseCOVID192023]. 
-This can include explanatory videos, infographics, and articles that help individuals understand the virus, its transmission, prevention strategies, and vaccination information. 
+<!-- Educational opportunities:  -->
+Social media can be a valuable _educational tool_, with various organizations and experts sharing informative content about COVID-19 [@worldhealthorganizationCoronavirusDiseaseCOVID192023]. 
+<!-- This can include explanatory videos, infographics, and articles that help individuals understand the virus, its transmission, prevention strategies, and vaccination information.  -->
 Access to educational resources on social media can empower individuals to make informed decisions regarding their health.
-Community support and solidarity: 
-Social media platforms enable individuals to connect with others who are experiencing similar challenges during the pandemic [@guazziniSecondWaveAnalysis2022]. 
-Sharing experiences, offering support, and expressing solidarity can foster a sense of community and reduce feelings of isolation. 
+<!-- Community support and solidarity:  -->
+Social media platforms enable individuals to _connect with others_ who are experiencing similar challenges during the pandemic [@guazziniSecondWaveAnalysis2022]. 
+<!-- Sharing experiences, offering support, and expressing solidarity can foster a sense of community and reduce feelings of isolation.  -->
 Engaging in online communities and support groups can provide emotional support and create a network of like-minded individuals.
-Promoting mental health and well-being: 
-Many mental health organizations and professionals utilize social media to share tips, strategies, and resources for maintaining mental well-being during the pandemic [@twitterTalkingMentalHealthAwarenessTwitter2020].
-They provide guidance on coping mechanisms, stress reduction techniques, and self-care practices. 
+<!-- Promoting mental health and well-being:  -->
+Many mental health organizations and professionals utilize social media to share tips, strategies, and resources for _maintaining mental well-being_ during the pandemic [@twitterTalkingMentalHealthAwarenessTwitter2020].
+<!-- They provide guidance on coping mechanisms, stress reduction techniques, and self-care practices.  -->
 Engaging with such content might help individuals prioritize their mental health and develop resilience during challenging times.
-Encouraging positive behaviors: 
-Social media campaigns and initiatives can promote positive behaviors related to COVID-19, such as mask-wearing, physical distancing, hand hygiene, and vaccination [@atheyDigitalPublicHealth2023; @huntSocialMediaBased2022].
+<!-- Encouraging positive behaviors:  -->
+Social media campaigns and initiatives can promote _positive COVID-19 behaviors_, such as mask-wearing, physical distancing, hand hygiene, and vaccination [@atheyDigitalPublicHealth2023; @huntSocialMediaBased2022].
 Public health organizations and influencers leverage the power of social media to spread awareness and encourage responsible actions, contributing to public health efforts and fostering a sense of collective responsibility.
@@
 On the other hand, the effects might be negative, perhaps best explained by the following five mechanisms.
-Misinformation and rumors: 
-Social media platforms can easily spread false or misleading information about COVID-19 [@liYouTubeSourceInformation2020]. 
+<!-- Misinformation and rumors:  -->
+Social media platforms can easily spread _false or misleading information_ about COVID-19 [@liYouTubeSourceInformation2020]. 
 Due to the ease of sharing and the lack of fact-checking, inaccurate information can go viral and might cause confusion, anxiety, and panic among users.
-Overwhelm and anxiety: 
-Constant exposure to COVID-19-related content on social media can lead to information overload and contribute to heightened anxiety levels [@fanInformationOverloadWellbeing2021]. 
+<!-- Overwhelm and anxiety:  -->
+Constant exposure to COVID-19-related content on social media can lead to _information overload_ and contribute to heightened anxiety levels [@fanInformationOverloadWellbeing2021]. 
 The rapid spread of news, updates, and opinions can be overwhelming and might exacerbate existing stress or fears about the pandemic [@sharmaDarkEndTunnel2022].
-Confirmation bias and echo chambers: 
-Social media algorithms are designed to show users content that aligns with their interests and beliefs, potentially contributing to echo chambers [@cinelliEchoChamberEffect2021]. 
-This might lead to confirmation bias, where individuals are more likely to be exposed to information that confirms their preexisting beliefs. 
+<!-- Confirmation bias and echo chambers:  -->
+Social media algorithms are designed to show users content that aligns with their interests and beliefs, potentially contributing to _echo chambers_ [@cinelliEchoChamberEffect2021]. 
+<!-- This might lead to confirmation bias, where individuals are more likely to be exposed to information that confirms their preexisting beliefs.  -->
 In the context of COVID-19, this might reinforce false or misleading information and prevent individuals from considering alternative perspectives.
-Cyberbullying and harassment: 
-Social media platforms are known for fostering negativity, with users sometimes engaging in cyberbullying and harassment [@giumettiCyberbullyingSocialMedia2022]. 
+<!-- Cyberbullying and harassment:  -->
+Social media platforms are known for fostering negativity, with users sometimes engaging in _cyberbullying and harassment_ [@giumettiCyberbullyingSocialMedia2022]. 
 Discussions around COVID-19 can become heated and polarized, leading to personal attacks and online conflicts. 
 Such experiences threaten mental well-being and might contribute to feelings of distress and isolation.
-Social comparison and fear of missing out (FOMO): 
-Social media often showcase the highlights and accomplishments of others, encouraging social comparison [@przybylskiMotivationalEmotionalBehavioral2013]. 
+<!-- Social comparison and fear of missing out (FOMO):  -->
+Social media often showcase the highlights and accomplishments of others, encouraging _social comparison_ [@przybylskiMotivationalEmotionalBehavioral2013]. 
 During a pandemic, seeing posts about others' successes or seemingly perfect lives might intensify feelings of inadequacy or FOMO, especially when individuals are unable to participate in similar activities due to restrictions or personal circumstances [@sharmaDarkEndTunnel2022].
@@
 <!-- Spending excessive time on social media to learn about COVID-19 can lead to a decrease in face-to-face interactions and real-life connections [but see @hallSocialMediaUse2022].  -->
-This might impact mental health, as most humans have an innate need for social support and physical presence, which online interactions might not fully replicate.
+<!-- This might impact mental health, as most humans have an innate need for social support and physical presence, which online interactions might not fully replicate. -->
 
@@
 A couple of studies reported mostly neutral effects of social media use on well-being indicators [@bradleyStressMoodSmartphone2021; @edenMediaCopingCOVID192020; @sewallObjectivelyMeasuredDigital2021].
-Overall, the literature demonstrates a mixed picture, highlighting both positive and negative effects of social media use focused on or during COVID-19 on well-being [see also @bradleyStressMoodSmartphone2021; @dornemannHowGoodBad2021; @liuRelationOfficialWhatsAppdistributed2020; @riehmAssociationsMediaExposure2020; @sewallObjectivelyMeasuredDigital2021].
+Overall, the literature demonstrates a mixed picture, highlighting both positive and negative effects of social media use focused on or during COVID-19 on well-being.
 
@@
 
-The data come from the Austrian Corona Panel Project [@kittelAustrianCoronaPanel2021], which is a large-scale standalone panel study. 
+The data come from the Austrian Corona Panel Project [@kittelAustrianCoronaPanel2020], which is a large-scale standalone panel study. 
 The data are hosted on AUSSDA, are publicly available here (https://doi.org/10.11587/28KQNS), and consist of 34 waves.

### 2023-07-21 manuscript.tex
- reduced in length

@@
 It is therefore often asked whether social media can become addictive, and users sometimes express this fear themselves (Yang et al., 2021).
-However, a recently published meta-analysis found that the two most prominent measures of addiction, the Bergen Facebook Addiction Scale and the Bergen Social Media Addiction Scale, have only small relations to well-being (Duradoni et al., 2020).
-In addition, the general idea of labeling excessive social and new media use as addiction was criticized, arguing that social media use represent a new regular behavior that should not be pathologized (Galer, 2018; van Rooij et al., 2018).
 
@@
 Several uses and gratifications exist, which help explain why people used social media frequently during the pandemic.
-Access to information and resources:
-Despite incorrect information, social media provide a vast platform for disseminating accurate and timely information about COVID-19 (John Hopkins University, 2023).
-It allows individuals to stay informed about the latest updates, guidelines, and recommendations from reputable health organizations and experts.
+Despite incorrect information, social media provide a vast platform for disseminating \emph{accurate and timely information} about COVID-19 (John Hopkins University, 2023).
 Access to reliable information can help people make informed decisions, alleviate uncertainties, and feel empowered during the pandemic.
-Educational opportunities:
-Social media can be a valuable educational tool, with various organizations and experts sharing informative content about COVID-19 (World Health Organization, 2023).
-This can include explanatory videos, infographics, and articles that help individuals understand the virus, its transmission, prevention strategies, and vaccination information.
+Social media can be a valuable \emph{educational tool}, with various organizations and experts sharing informative content about COVID-19 (World Health Organization, 2023).
 Access to educational resources on social media can empower individuals to make informed decisions regarding their health.
-Community support and solidarity:
-Social media platforms enable individuals to connect with others who are experiencing similar challenges during the pandemic (Guazzini et al., 2022).
-Sharing experiences, offering support, and expressing solidarity can foster a sense of community and reduce feelings of isolation.
+Social media platforms enable individuals to \emph{connect with others} who are experiencing similar challenges during the pandemic (Guazzini et al., 2022).
 Engaging in online communities and support groups can provide emotional support and create a network of like-minded individuals.
-Promoting mental health and well-being:
-Many mental health organizations and professionals utilize social media to share tips, strategies, and resources for maintaining mental well-being during the pandemic (Twitter, 2020).
-They provide guidance on coping mechanisms, stress reduction techniques, and self-care practices.
+Many mental health organizations and professionals utilize social media to share tips, strategies, and resources for \emph{maintaining mental well-being} during the pandemic (Twitter, 2020).
 Engaging with such content might help individuals prioritize their mental health and develop resilience during challenging times.
-Encouraging positive behaviors:
-Social media campaigns and initiatives can promote positive behaviors related to COVID-19, such as mask-wearing, physical distancing, hand hygiene, and vaccination (Athey et al., 2023; Hunt et al., 2022).
+Social media campaigns and initiatives can promote \emph{positive COVID-19 behaviors}, such as mask-wearing, physical distancing, hand hygiene, and vaccination (Athey et al., 2023; Hunt et al., 2022).
 Public health organizations and influencers leverage the power of social media to spread awareness and encourage responsible actions, contributing to public health efforts and fostering a sense of collective responsibility.
@@
 On the other hand, the effects might be negative, perhaps best explained by the following five mechanisms.
-Misinformation and rumors:
-Social media platforms can easily spread false or misleading information about COVID-19 (Li et al., 2020).
+Social media platforms can easily spread \emph{false or misleading information} about COVID-19 (Li et al., 2020).
 Due to the ease of sharing and the lack of fact-checking, inaccurate information can go viral and might cause confusion, anxiety, and panic among users.
-Overwhelm and anxiety:
-Constant exposure to COVID-19-related content on social media can lead to information overload and contribute to heightened anxiety levels (J. Fan \& Smith, 2021).
+Constant exposure to COVID-19-related content on social media can lead to \emph{information overload} and contribute to heightened anxiety levels (J. Fan \& Smith, 2021).
 The rapid spread of news, updates, and opinions can be overwhelming and might exacerbate existing stress or fears about the pandemic (Sharma et al., 2022).
-Confirmation bias and echo chambers:
-Social media algorithms are designed to show users content that aligns with their interests and beliefs, potentially contributing to echo chambers (Cinelli et al., 2021).
-This might lead to confirmation bias, where individuals are more likely to be exposed to information that confirms their preexisting beliefs.
+Social media algorithms are designed to show users content that aligns with their interests and beliefs, potentially contributing to \emph{echo chambers} (Cinelli et al., 2021).
 In the context of COVID-19, this might reinforce false or misleading information and prevent individuals from considering alternative perspectives.
-Cyberbullying and harassment:
-Social media platforms are known for fostering negativity, with users sometimes engaging in cyberbullying and harassment (Giumetti \& Kowalski, 2022).
+Social media platforms are known for fostering negativity, with users sometimes engaging in \emph{cyberbullying and harassment} (Giumetti \& Kowalski, 2022).
 Discussions around COVID-19 can become heated and polarized, leading to personal attacks and online conflicts.
 Such experiences threaten mental well-being and might contribute to feelings of distress and isolation.
-Social comparison and fear of missing out (FOMO):
-Social media often showcase the highlights and accomplishments of others, encouraging social comparison (Przybylski et al., 2013).
+Social media often showcase the highlights and accomplishments of others, encouraging \emph{social comparison} (Przybylski et al., 2013).
 During a pandemic, seeing posts about others' successes or seemingly perfect lives might intensify feelings of inadequacy or FOMO, especially when individuals are unable to participate in similar activities due to restrictions or personal circumstances (Sharma et al., 2022).
-This might impact mental health, as most humans have an innate need for social support and physical presence, which online interactions might not fully replicate.
 
@@
 A couple of studies reported mostly neutral effects of social media use on well-being indicators (Bradley \& Howard, 2021; Eden et al., 2020; Sewall et al., 2021).
-Overall, the literature demonstrates a mixed picture, highlighting both positive and negative effects of social media use focused on or during COVID-19 on well-being (see also Bradley \& Howard, 2021; Dörnemann et al., 2021; Liu \& Tong, 2020; Riehm et al., 2020; Sewall et al., 2021).
+Overall, the literature demonstrates a mixed picture, highlighting both positive and negative effects of social media use focused on or during COVID-19 on well-being.
 
@@
 
-The data come from the Austrian Corona Panel Project (Kittel et al., 2021), which is a large-scale standalone panel study.
+The data come from the Austrian Corona Panel Project (Kittel et al., 2020), which is a large-scale standalone panel study.
 The data are hosted on AUSSDA, are publicly available here (\url{https://doi.org/10.11587/28KQNS}), and consist of 34 waves.
@@
 \emph{Proceedings of the National Academy of Sciences}, \emph{120}(5), e2208110120.
+\url{https://doi.org/10.1073/pnas.2208110120}
 
 \leavevmode\vadjust pre{\hypertarget{ref-bellFixedRandomEffects2019}{}}%
-Bell, A., Fairbrother, M., \& Jones, K. (2019). Fixed and random effects models: Making an informed choice. \emph{Quality \& Quantity}, \emph{53}(2), 1051--1074.
+Bell, A., Fairbrother, M., \& Jones, K.
+(2019).
+Fixed and random effects models: Making an informed choice.
+\emph{Quality \& Quantity}, \emph{53}(2), 1051--1074.
+\url{https://doi.org/10.1007/s11135-018-0802-x}
 
@@
 \emph{European Archives of Psychiatry and Clinical Neuroscience}, \emph{271}(2), 283--291.
+\url{https://doi.org/10.1007/s00406-020-01171-6}
 
@@
 \emph{Communication Research}.
+\url{https://doi.org/10.1177/00936502211038196}
 
 \leavevmode\vadjust pre{\hypertarget{ref-bradleyStressMoodSmartphone2021}{}}%
-Bradley, A., \& Howard, A. (2021). \emph{Stress, mood, and smartphone use in {University} students: {A} 12-week longitudinal study}. {OSF Preprints}.
+Bradley, A., \& Howard, A. (2021). \emph{Stress, mood, and smartphone use in {University} students: {A} 12-week longitudinal study}. \url{https://doi.org/10.31219/osf.io/frvpb}
 
@@
 \emph{Proceedings of the National Academy of Sciences}, \emph{118}(9), e2023301118.
+\url{https://doi.org/10.1073/pnas.2023301118}
 
 \leavevmode\vadjust pre{\hypertarget{ref-dienerAdvancesOpenQuestions2018}{}}%
-Diener, E., Lucas, R. E., \& Oishi, S. (2018). Advances and open questions in the science of subjective well-being. \emph{Collabra: Psychology}, \emph{4}(1), 15.
+Diener, E., Lucas, R.
+E., \& Oishi, S.
+(2018).
+Advances and open questions in the science of subjective well-being.
+\emph{Collabra: Psychology}, \emph{4}(1), 15.
+\url{https://doi.org/10.1525/collabra.115}
 
 \leavevmode\vadjust pre{\hypertarget{ref-dienesUsingBayesGet2014}{}}%
-Dienes, Z. (2014). Using {Bayes} to get the most out of non-significant results. \emph{Frontiers in Psychology}, \emph{5}.
+Dienes, Z. (2014). Using {Bayes} to get the most out of non-significant results. \emph{Frontiers in Psychology}, \emph{5}. \url{https://doi.org/10.3389/fpsyg.2014.00781}
 
 \leavevmode\vadjust pre{\hypertarget{ref-dienlinImpactDigitalTechnology2020}{}}%
-Dienlin, T., \& Johannes, N. (2020). The impact of digital technology use on adolescent well-being. \emph{Dialogues in Clinical Neuroscience}, \emph{22}(2), 135--142.
+Dienlin, T., \& Johannes, N.
+(2020).
+The impact of digital technology use on adolescent well-being.
+\emph{Dialogues in Clinical Neuroscience}, \emph{22}(2), 135--142.
+\url{https://doi.org/doi:10.31887/DCNS.2020.22.2/tdienlin}
 
@@
 \emph{Journal of Computer-Mediated Communication}, \emph{22}(2), 71--87.
+\url{https://doi.org/10.1111/jcc4.12183}
 
 \leavevmode\vadjust pre{\hypertarget{ref-dormannOptimalTimeLags2015}{}}%
-Dormann, C., \& Griffin, M. A. (2015). \href{https://www.ncbi.nlm.nih.gov/pubmed/26322999}{Optimal time lags in panel studies}. \emph{Psychological Methods}, \emph{20}(4), 489--505.
-
-\leavevmode\vadjust pre{\hypertarget{ref-dornemannHowGoodBad2021}{}}%
-Dörnemann, A., Boenisch, N., Schommer, L., Winkelhorst, L., \& Wingen, T. (2021). \emph{How do good and bad news impact mood during the {Covid-19} pandemic? {The} role of similarity}. {OSF Preprints}.
-
-\leavevmode\vadjust pre{\hypertarget{ref-duradoniWellbeingSocialMedia2020}{}}%
-Duradoni, M., Innocenti, F., \& Guazzini, A. (2020). Well-being and social media: {A} systematic review of {Bergen Addiction Scales}. \emph{Future Internet}, \emph{12}(2), 24.
+Dormann, C., \& Griffin, M. A. (2015). Optimal time lags in panel studies. \emph{Psychological Methods}, \emph{20}(4), 489--505. \url{https://doi.org/10.1037/met0000041}
 
@@
 \emph{Frontiers in Psychology}, \emph{11}, 577639.
+\url{https://doi.org/10.3389/fpsyg.2020.577639}
 
 \leavevmode\vadjust pre{\hypertarget{ref-egerStatisticalMetaanalysisWellbeing2015}{}}%
-Eger, R. J., \& Maridal, J. H. (2015). A statistical meta-analysis of the wellbeing literature. \emph{International Journal of Wellbeing}, \emph{5}(2), 45--74.
+Eger, R. J., \& Maridal, J. H. (2015). A statistical meta-analysis of the wellbeing literature. \emph{International Journal of Wellbeing}, \emph{5}(2), 45--74. \url{https://doi.org/10.5502/ijw.v5i2.4}
 
 \leavevmode\vadjust pre{\hypertarget{ref-europeansocialsurveyESS9Edition20182021}{}}%
-European Social Survey. (2021). \emph{{ESS9} edition 3.1 - 2018 {Documentation Report}}.
+European Social Survey. (2021). \emph{{ESS9} edition 3.1 - 2018 {Documentation Report}}. \url{https://www.europeansocialsurvey.org/docs/round9/survey/ESS9_data_documentation_report_e03_1.pdf}
 
 \leavevmode\vadjust pre{\hypertarget{ref-fanInformationOverloadWellbeing2021}{}}%
-Fan, J., \& Smith, A. P. (2021). Information overload, wellbeing and {COVID-19}: {A} survey in {China}. \emph{Behavioral Sciences}, \emph{11}(5), 62.
+Fan, J., \& Smith, A. P. (2021). Information overload, wellbeing and {COVID-19}: {A} survey in {China}. \emph{Behavioral Sciences}, \emph{11}(5, 5), 62. \url{https://doi.org/10.3390/bs11050062}
 
@@
 (2020).
-Stigmatization in social media: {Documenting} and analyzing hate speech for {\textsc{COVID}} -19 on {Twitter}.
+Stigmatization in social media: {Documenting} and analyzing hate speech for {\textsc{COVID}} ‐19 on {Twitter}.
 \emph{Proceedings of the Association for Information Science and Technology}, \emph{57}(1).
+\url{https://doi.org/10.1002/pra2.313}
 
@@
 \emph{Professional Psychology: Research and Practice}.
+\url{https://doi.org/10.1037/pro0000426}
 
 \leavevmode\vadjust pre{\hypertarget{ref-funderEvaluatingEffectSize2019}{}}%
-Funder, D. C., \& Ozer, D. J. (2019). Evaluating effect size in psychological research: {Sense} and nonsense. \emph{Advances in Methods and Practices in Psychological Science}, \emph{2}(2), 156--168.
-
-\leavevmode\vadjust pre{\hypertarget{ref-galerHowMuchToo2018}{}}%
-Galer, S. S. (2018). \emph{How much is {``too much time''} on social media?} https://www.bbc.com/future/article/20180118-how-much-is-too-much-time-on-social-media.
+Funder, D.
+C., \& Ozer, D.
+J.
+(2019).
+Evaluating effect size in psychological research: {Sense} and nonsense.
+\emph{Advances in Methods and Practices in Psychological Science}, \emph{2}(2), 156--168.
+\url{https://doi.org/10.1177/2515245919847202}
 
 \leavevmode\vadjust pre{\hypertarget{ref-giumettiCyberbullyingSocialMedia2022}{}}%
-Giumetti, G. W., \& Kowalski, R. M. (2022). Cyberbullying via social media and well-being. \emph{Current Opinion in Psychology}, \emph{45}, 101314.
+Giumetti, G. W., \& Kowalski, R. M. (2022). Cyberbullying via social media and well-being. \emph{Current Opinion in Psychology}, \emph{45}, 101314. \url{https://doi.org/10.1016/j.copsyc.2022.101314}
 
@@
 \emph{International Journal of Environmental Research and Public Health}, \emph{19}(3), 1635.
+\url{https://doi.org/10.3390/ijerph19031635}
 
 \leavevmode\vadjust pre{\hypertarget{ref-hallSocialMediaUse2022}{}}%
-Hall, J. A., \& Liu, D. (2022). Social media use, social displacement, and well-being. \emph{Current Opinion in Psychology}, \emph{46}, 101339.
+Hall, J. A., \& Liu, D. (2022). Social media use, social displacement, and well-being. \emph{Current Opinion in Psychology}, \emph{46}, 101339. \url{https://doi.org/10.1016/j.copsyc.2022.101339}
 
@@
 \emph{Computers in Human Behavior}, \emph{29}(3), 1159--1168.
+\url{https://doi.org/10.1016/j.chb.2012.10.008}
 
@@
 (2017).
-\href{https://www.ncbi.nlm.nih.gov/pubmed/28622031}{Time spent on social network sites and psychological well-being: {A} meta-analysis}.
+Time spent on social network sites and psychological well-being: {A} meta-analysis.
 \emph{Cyberpsychology, Behavior and Social Networking}, \emph{20}(6), 346--354.
+\url{https://doi.org/10.1089/cyber.2016.0758}
 
@@
 (2022).
-\href{https://www.ncbi.nlm.nih.gov/pmc/articles/PMC9382165}{A social media-based public health campaign encouraging {COVID-19} vaccination across the {United States}}.
+A social media‒based public health campaign encouraging {COVID-19} vaccination across the {United States}.
 \emph{American Journal of Public Health}, \emph{112}(9), 1253--1256.
+\url{https://doi.org/10.2105/AJPH.2022.306934}
 
 \leavevmode\vadjust pre{\hypertarget{ref-johnhopkinsuniversityCOVID19Map2023}{}}%
-John Hopkins University. (2023). {COVID-19 Map}. In \emph{Johns Hopkins Coronavirus Resource Center}. https://coronavirus.jhu.edu/map.html.
+John Hopkins University. (2023). \emph{{COVID-19 Map}}. {Johns Hopkins Coronavirus Resource Center}. \url{https://coronavirus.jhu.edu/map.html}
 
 \leavevmode\vadjust pre{\hypertarget{ref-katzUsesGratificationsResearch1973}{}}%
-Katz, E., Blumler, J. G., \& Gurevitch, M. (1973). Uses and {Gratifications Research}. \emph{Public Opinion Quarterly}, \emph{37}(4), 509.
+Katz, E., Blumler, J. G., \& Gurevitch, M. (1973). Uses and {Gratifications Research}. \emph{Public Opinion Quarterly}, \emph{37}(4), 509. \url{https://doi.org/10.1086/268109}
 
@@
 (2020).
-\emph{Austrian {Corona Panel Project} ({SUF} edition)}.
+\emph{Austrian {Corona Panel Project} ({SUF} edition)} {[}Data set{]}.
 {AUSSDA}.
-
-\leavevmode\vadjust pre{\hypertarget{ref-kittelAustrianCoronaPanel2021}{}}%
-Kittel, B., Kritzinger, S., Boomgaarden, H., Prainsack, B., Eberl, J.-M., Kalleitner, F., Lebernegg, N.
-S., Partheymüller, J., Plescia, C., Schiestl, D.
-W., \& Schlogl, L.
-(2021).
-The {Austrian Corona Panel Project}: Monitoring individual and societal dynamics amidst the {COVID-19} crisis.
-\emph{European Political Science}, \emph{20}(2), 318--344.
+\url{https://doi.org/10.11587/28KQNS}
 
 \leavevmode\vadjust pre{\hypertarget{ref-kleinDarklySoothingCompulsion2021}{}}%
-Klein, J. (2021). \emph{The darkly soothing compulsion of 'doomscrolling'}. https://www.bbc.com/worklife/article/20210226-the-darkly-soothing-compulsion-of-doomscrolling.
+Klein, J. (2021, March 3). \emph{The darkly soothing compulsion of 'doomscrolling'}. \url{https://www.bbc.com/worklife/article/20210226-the-darkly-soothing-compulsion-of-doomscrolling}
 
@@
 \emph{Advances in Methods and Practices in Psychological Science}, \emph{1}(2), 259--269.
+\url{https://doi.org/10.1177/2515245918770963}
 
@@
 \emph{Social Science \& Medicine}, \emph{293}, 114674.
+\url{https://doi.org/10.1016/j.socscimed.2021.114674}
 
@@
 (2020).
-\href{https://www.ncbi.nlm.nih.gov/pubmed/32409327}{{YouTube} as a source of information on {COVID-19}: A pandemic of misinformation?} \emph{BMJ Global Health}, \emph{5}(5), e002604.
-
-\leavevmode\vadjust pre{\hypertarget{ref-liuRelationOfficialWhatsAppdistributed2020}{}}%
-Liu, J.
-C.
-J., \& Tong, E.
-M.
-W.
-(2020).
-The relation between official {WhatsApp-distributed COVID-19} news exposure and psychological symptoms: Cross-sectional survey study.
-\emph{Journal of Medical Internet Research}, \emph{22}(9), e22142.
+{YouTube} as a source of information on {COVID-19}: A pandemic of misinformation?
+\emph{BMJ Global Health}, \emph{5}(5), e002604.
+\url{https://doi.org/10.1136/bmjgh-2020-002604}
 
 \leavevmode\vadjust pre{\hypertarget{ref-lucasAdaptationSetpointModel2007}{}}%
-Lucas, R. E. (2007). Adaptation and the set-point model of subjective well-being. \emph{Current Directions in Psychological Science}, \emph{16}(2), 75--79.
+Lucas, R.
+E.
+(2007).
+Adaptation and the set-point model of subjective well-being.
+\emph{Current Directions in Psychological Science}, \emph{16}(2), 75--79.
+\url{https://doi.org/10.1111/j.1467-8721.2007.00479.x}
 
 \leavevmode\vadjust pre{\hypertarget{ref-lucasWhyCrossLaggedPanel2022}{}}%
-Lucas, R. E. (2022). \emph{Why the {Cross-Lagged Panel Model} is {Almost Never} the {Right Choice}} {[}Preprint{]}. {PsyArXiv}.
+Lucas, R. E. (2022). \emph{Why the {Cross-Lagged Panel Model} is {Almost Never} the {Right Choice}} {[}Preprint{]}. {PsyArXiv}. \url{https://doi.org/10.31234/osf.io/pkec7}
 
@@
 \leavevmode\vadjust pre{\hypertarget{ref-marcianoDynamicsAdolescentsSmartphone2022}{}}%
-Marciano, L., Driver, C. C., Schulz, P. J., \& Camerini, A.-L. (2022). Dynamics of adolescents' smartphone use and well-being are positive but ephemeral. \emph{Scientific Reports}, \emph{12}(1), 1316.
+Marciano, L., Driver, C.
+C., Schulz, P.
+J., \& Camerini, A.-L.
+(2022).
+Dynamics of adolescents' smartphone use and well-being are positive but ephemeral.
+\emph{Scientific Reports}, \emph{12}(1), 1316.
+\url{https://doi.org/10.1038/s41598-022-05291-y}
 
 \leavevmode\vadjust pre{\hypertarget{ref-meierComputermediatedCommunicationSocial2020a}{}}%
-Meier, A., \& Reinecke, L. (2020). Computer-mediated communication, social media, and mental health: {A} conceptual and empirical meta-review. \emph{Communication Research}, 009365022095822.
+Meier, A., \& Reinecke, L.
+(2020).
+Computer-mediated communication, social media, and mental health: {A} conceptual and empirical meta-review.
+\emph{Communication Research}, 009365022095822.
+\url{https://doi.org/10.1177/0093650220958224}
 
@@
 \emph{Information, Communication \& Society}, \emph{25}(9), 1247--1272.
+\url{https://doi.org/10.1080/1369118X.2020.1850841}
 
@@
 \emph{Medical Care}, \emph{41}(5), 582--592.
+\href{Retrieved\%20from\%20http://www.jstor.org/stable/3768017}{Retrieved from http://www.jstor.org/stable/3768017}
 
 \leavevmode\vadjust pre{\hypertarget{ref-orbenTeenagersScreensSocial2020}{}}%
-Orben, A. (2020). Teenagers, screens and social media: A narrative review of reviews and key studies. \emph{Social Psychiatry and Psychiatric Epidemiology}, \emph{55}(4), 407--414.
+Orben, A.
+(2020).
+Teenagers, screens and social media: A narrative review of reviews and key studies.
+\emph{Social Psychiatry and Psychiatric Epidemiology}, \emph{55}(4), 407--414.
+\url{https://doi.org/10.1007/s00127-019-01825-4}
 
 \leavevmode\vadjust pre{\hypertarget{ref-orbenWindowsDevelopmentalSensitivity2022}{}}%
-Orben, A., Przybylski, A. K., Blakemore, S.-J., \& Kievit, R. A. (2022). Windows of developmental sensitivity to social media. \emph{Nature Communications}, \emph{13}(1), 1649.
+Orben, A., Przybylski, A.
+K., Blakemore, S.-J., \& Kievit, R.
+A.
+(2022).
+Windows of developmental sensitivity to social media.
+\emph{Nature Communications}, \emph{13}(1), 1649.
+\url{https://doi.org/10.1038/s41467-022-29296-3}
 
@@
 \emph{Journal of Research in Interactive Marketing}, \emph{14}(2), 269--284.
+\url{https://doi.org/10.1108/JRIM-10-2019-0159}
 
@@
 \emph{Computers in Human Behavior}, \emph{29}(4), 1841--1848.
-
-\leavevmode\vadjust pre{\hypertarget{ref-riehmAssociationsMediaExposure2020}{}}%
-Riehm, K.
-E., Holingue, C., Kalb, L.
-G., Bennett, D., Kapteyn, A., Jiang, Q., Veldhuis, C.
-B., Johnson, R.
-M., Fallin, M.
-D., Kreuter, F., Stuart, E.
-A., \& Thrul, J.
-(2020).
-\href{https://www.ncbi.nlm.nih.gov/pubmed/33011008}{Associations between media exposure and mental distress among {U}.{S}.
-Adults at the beginning of the {COVID-19} pandemic}.
-\emph{American Journal of Preventive Medicine}, \emph{59}(5), 630--638.
+\url{https://doi.org/10.1016/j.chb.2013.02.014}
 
@@
 \emph{Advances in Methods and Practices in Psychological Science}, \emph{24}(2), 251524591774562.
+\url{https://doi.org/10.1177/2515245917745629}
 
 \leavevmode\vadjust pre{\hypertarget{ref-rohrerTheseAreNot2021}{}}%
-Rohrer, J. M., \& Murayama, K. (2021). \emph{These are not the effects you are looking for: {Causality} and the within-/between-person distinction in longitudinal data analysis}. {PsyArXiv}.
+Rohrer, J.
+M., \& Murayama, K.
+(2021, October 11).
+\emph{These are not the effects you are looking for: {Causality} and the within-/between-person distinction in longitudinal data analysis}.
+\url{https://doi.org/10.31234/osf.io/tg4vj}
 
@@
 Sandstrom, G., Buchanan, K., Aknin, L., \& Lotun, S.
-(2021).
-Doomscrolling {COVID} news takes an emotional toll \textendash{} here's how to make your social media a happier place.
-In \emph{The Conversation}.
-http://theconversation.com/doomscrolling-covid-news-takes-an-emotional-toll-heres-how-to-make-your-social-media-a-happier-place-170342.
+(2021, October 22).
+\emph{Doomscrolling {COVID} news takes an emotional toll -- here's how to make your social media a happier place}.
+{The Conversation}.
+\url{http://theconversation.com/doomscrolling-covid-news-takes-an-emotional-toll-heres-how-to-make-your-social-media-a-happier-place-170342}
 
 \leavevmode\vadjust pre{\hypertarget{ref-scharkowAccuracySelfreportedInternet2016}{}}%
-Scharkow, M. (2016). The accuracy of self-reported {Internet} use\textemdash{{A}} validation study using client log data. \emph{Communication Methods and Measures}, \emph{10}(1), 13--27.
+Scharkow, M.
+(2016).
+The accuracy of self-reported {Internet} use---{A} validation study using client log data.
+\emph{Communication Methods and Measures}, \emph{10}(1), 13--27.
+\url{https://doi.org/10.1080/19312458.2015.1118446}
 
@@
 \emph{Journal of Affective Disorders}, \emph{288}, 145--147.
+\url{https://doi.org/10.1016/j.jad.2021.04.008}
 
 \leavevmode\vadjust pre{\hypertarget{ref-sharmaDarkEndTunnel2022}{}}%
-Sharma, B., Lee, S. S., \& Johnson, B. K. (2022). The dark at the end of the tunnel: {Doomscrolling} on social media newsfeeds. \emph{Technology, Mind, and Behavior}, \emph{3}(1).
+Sharma, B., Lee, S.
+S., \& Johnson, B.
+K.
+(2022).
+The dark at the end of the tunnel: {Doomscrolling} on social media newsfeeds.
+\emph{Technology, Mind, and Behavior}, \emph{3}(1).
+\url{https://doi.org/10.1037/tmb0000059}
 
 \leavevmode\vadjust pre{\hypertarget{ref-sheldonStabilityHappinessTheories2014}{}}%
-Sheldon, K. M., \& Lucas, R. E. (2014). \emph{Stability of happiness: Theories and evidence on whether happiness can change}.
+Sheldon, K. M., \& Lucas, R. E. (2014). \emph{Stability of happiness: Theories and evidence on whether happiness can change}. \url{http://site.ebrary.com/id/10891875}
 
 \leavevmode\vadjust pre{\hypertarget{ref-stainbackCOVID1924News2020}{}}%
-Stainback, K., Hearne, B. N., \& Trieu, M. M. (2020). {COVID-19} and the 24/7 {News Cycle}: {Does COVID-19 News Exposure Affect Mental Health}? \emph{Socius}, \emph{6}, 2378023120969339.
+Stainback, K., Hearne, B.
+N., \& Trieu, M.
+M.
+(2020).
+{COVID-19} and the 24/7 {News Cycle}: {Does COVID-19 News Exposure Affect Mental Health}?
+\emph{Socius}, \emph{6}, 2378023120969339.
+\url{https://doi.org/10.1177/2378023120969339}
 
@@
 Statista.
-(2021).
+(2021, May 21).
 \emph{Average daily time spent on social networks by users in the {United States} from 2018 to 2022}.
-https://www.statista.com/statistics/1018324/us-users-daily-social-media-minutes/.
+\url{https://www.statista.com/statistics/1018324/us-users-daily-social-media-minutes/}
 
 \leavevmode\vadjust pre{\hypertarget{ref-twitterTalkingMentalHealthAwarenessTwitter2020}{}}%
-Twitter. (2020). \emph{Talking \#{MentalHealthAwareness} on {Twitter} with our global partners}. https://blog.twitter.com/en\_us/topics/company/2020/talking-mental-health-awareness-on-twitter.
+Twitter.
+(2020, June 15).
+\emph{Talking \#{MentalHealthAwareness} on {Twitter} with our global partners}.
+\url{https://blog.twitter.com/en_us/topics/company/2020/talking-mental-health-awareness-on-twitter}
 
 \leavevmode\vadjust pre{\hypertarget{ref-valkenburgDifferentialSusceptibilityMedia2013}{}}%
-Valkenburg, P. M., \& Peter, J. (2013). The differential susceptibility to media effects model. \emph{Journal of Communication}, \emph{63}(2), 221--243.
+Valkenburg, P. M., \& Peter, J. (2013). The differential susceptibility to media effects model. \emph{Journal of Communication}, \emph{63}(2), 221--243. \url{https://doi.org/10.1111/jcom.12024}
 
@@
 \emph{New Media \& Society}, \emph{24}(2), 530--549.
-
-\leavevmode\vadjust pre{\hypertarget{ref-vanrooijWeakScientificBasis2018}{}}%
-van Rooij, A.
-J., Ferguson, C.
-J., Colder Carras, M., Kardefelt-Winther, D., Shi, J., Aarseth, E., Bean, A.
-M., Bergmark, K.
-H., Brus, A., Coulson, M., Deleuze, J., Dullur, P., Dunkels, E., Edman, J., Elson, M., Etchells, P.
-J., Fiskaali, A., Granic, I., Jansz, J., \ldots{} Przybylski, A.
-K.
-(2018).
-A weak scientific basis for gaming disorder: {Let} us err on the side of caution.
-\emph{Journal of Behavioral Addictions}, \emph{7}(1), 1--9.
+\url{https://doi.org/10.1177/14614448211065425}
 
 \leavevmode\vadjust pre{\hypertarget{ref-verbeijSelfreportedMeasuresSocial2021}{}}%
-Verbeij, T., Pouwels, J. L., Beyens, I., \& Valkenburg, P. M. (2021). \emph{Self-reported measures of social media use show high predictive validity} {[}Preprint{]}. {PsyArXiv}.
+Verbeij, T., Pouwels, J.
+L., Beyens, I., \& Valkenburg, P.
+M.
+(2021).
+\emph{Self-reported measures of social media use show high predictive validity} {[}Preprint{]}.
+{PsyArXiv}.
+\url{https://doi.org/10.31234/osf.io/c9bj7}
 
@@
 (2015).
-\href{https://www.ncbi.nlm.nih.gov/pubmed/25706656}{Passive {Facebook} usage undermines affective well-being: {Experimental} and longitudinal evidence}.
+Passive {Facebook} usage undermines affective well-being: {Experimental} and longitudinal evidence.
 \emph{Journal of Experimental Psychology: General}, \emph{144}(2), 480--488.
+\url{https://doi.org/10.1037/xge0000057}
 
@@
 (2018).
-\emph{{AUTNES Online Panel Study} 2017 ({SUF} edition)}.
+\emph{{AUTNES Online Panel Study} 2017 ({SUF} edition)} {[}Data set{]}.
 {AUSSDA}.
+\url{https://doi.org/10.11587/I7QIYJ}
 
@@
 Ward, G., Powdthavee, N., Fleche, S., Layard, R., \& Clark, A.
-(2016).
-Origins of happiness: {Evidence} and policy implications.
-In \emph{CEPR}.
-https://cepr.org/voxeu/columns/origins-happiness-evidence-and-policy-implications.
+(2016, December 12).
+\emph{Origins of happiness: {Evidence} and policy implications}.
+{CEPR}.
+\url{https://cepr.org/voxeu/columns/origins-happiness-evidence-and-policy-implications}
 
@@
 \leavevmode\vadjust pre{\hypertarget{ref-worldhealthorganizationCoronavirusDiseaseCOVID192023}{}}%
-World Health Organization. (2023). \emph{Coronavirus disease ({COVID-19}) \textendash{} {World Health Organization}}. https://www.who.int/emergencies/diseases/novel-coronavirus-2019.
+World Health Organization. (2023). \emph{Coronavirus disease ({COVID-19}) -- {World Health Organization}}. \url{https://www.who.int/emergencies/diseases/novel-coronavirus-2019}
 
@@
 {A} qualitative exploration of online video watching among {Chinese} young adults.
-\emph{International Journal of Environmental Research and Public Health}, \emph{18}(14), 7247.
+\emph{International Journal of Environmental Research and Public Health}, \emph{18}(14, 14), 7247.
+\url{https://doi.org/10.3390/ijerph18147247}
 
@@
 \emph{Computers in Human Behavior}, \emph{127}, 107050.
+\url{https://doi.org/10.1016/j.chb.2021.107050}
 
 \leavevmode\vadjust pre{\hypertarget{ref-zillmannMoodManagementCommunication1988}{}}%
-Zillmann, D. (1988). Mood {Management Through Communication Choices}. \emph{American Behavioral Scientist}, \emph{31}(3), 327340.
+Zillmann, D. (1988). Mood {Management Through Communication Choices}. \emph{American Behavioral Scientist}, \emph{31}(3), 327340. \url{https://doi.org/10.1177/000276488031003005}
 

### 2023-08-08 manuscript.Rmd
- shortened ms

@@
 A new phenomenon termed "doomscrolling" emerged [@sharmaDarkEndTunnel2022].
-Many users were glued to their screens and found it hard to pursue other relevant activities such as working, taking a break, or even looking after their children [@kleinDarklySoothingCompulsion2021].
+Many users were glued to their screens and found it hard to pursue other relevant activities such as working, taking a break, or even looking after their children.
 In the media it was hence increasingly asked whether using social media for COVID-19 related reasons would, next to all other stressors, create an additional burden on mental health [@sandstromDoomscrollingCOVIDNews2021].
@@
 In general, according to the set-point theory, well-being is surprisingly stable [@lykkenHappinessWhatStudies1999].
-Although specific events such as marriage or salary increase can have significant impacts on well-being, in most cases effects are only short-term, with well-being after some time returning to prior levels [@sheldonStabilityHappinessTheories2014].
-Only very specific factors such as unemployment, disability, or death can cause long-term changes in well-being [@lucasAdaptationSetpointModel2007].
+Although specific events such as marriage or salary increase can have significant impacts on well-being, in most cases effects are short-term, with well-being after some time returning to prior levels [@dienerAdvancesOpenQuestions2018].
+Specific factors such as unemployment, disability, or death, however, can cause long-term changes in well-being [@lucasAdaptationSetpointModel2007].
 So although well-being can be affected by external events and factors, this does not happen easily.
@@
 -->
-On the other hand, although mood management theory suggests that effects should not be overly negative, it could be that although short-term effects are positive, long-term effects are negative. 
+<!-- INCLUDE IF SPACE -->
+<!-- On the other hand, although mood management theory suggests that effects should not be overly negative, it could be that although short-term effects are positive, long-term effects are negative.
+-->
+<!-- Precisely because social media have so many positive consequences in the short run, this might cause problem in the long run---for example, because social media use displaces other potentially more meaningful and/or relaxing activities [@hallSocialMediaUse2022].
+-->
+<!-- As with many other things, there can be too much of a good thing. -->
+<!-- It is therefore often asked whether social media can become addictive, with users sometimes expressing this fear themselves. -->
+<!-- END -->
 <!-- it is still possible that people misjudge media effects. -->
 <!-- Users are often overly optimistic [@metzgerComparativeOptimismPrivacy2017], focusing on short term benefits while ignoring long-term costs.  -->
-Precisely because social media have so many positive consequences in the short run, this might cause problem in the long run---for example, because social media use displaces other potentially more meaningful and/or relaxing activities [@hallSocialMediaUse2022].
 <!-- In other words, social media might not problematic because they are inherently bad, but rather because they are too good. -->
-As with many other things, there can be too much of a good thing.
-It is therefore often asked whether social media can become addictive, and users sometimes express this fear themselves [@yangCanWatchingOnline2021].
 <!-- However, a recently published meta-analysis found that the two most prominent measures of addiction, the Bergen Facebook Addiction Scale and the Bergen Social Media Addiction Scale, have only small relations to well-being [@duradoniWellbeingSocialMedia2020].
@@
 <!-- Educational opportunities:  -->
-Social media can be a valuable _educational tool_, with various organizations and experts sharing informative content about COVID-19 [@worldhealthorganizationCoronavirusDiseaseCOVID192023]. 
-<!-- This can include explanatory videos, infographics, and articles that help individuals understand the virus, its transmission, prevention strategies, and vaccination information.  -->
-Access to educational resources on social media can empower individuals to make informed decisions regarding their health.
+<!-- Social media can be a valuable _educational tool_, with various organizations and experts sharing informative content about COVID-19 [@worldhealthorganizationCoronavirusDiseaseCOVID192023].  -->
+<!-- <!-- This can include explanatory videos, infographics, and articles that help individuals understand the virus, its transmission, prevention strategies, and vaccination information.  -->
+<!-- Access to educational resources on social media can empower individuals to make informed decisions regarding their health. -->
 <!-- Community support and solidarity:  -->
@@
 <!-- Confirmation bias and echo chambers:  -->
-Social media algorithms are designed to show users content that aligns with their interests and beliefs, potentially contributing to _echo chambers_ [@cinelliEchoChamberEffect2021]. 
-<!-- This might lead to confirmation bias, where individuals are more likely to be exposed to information that confirms their preexisting beliefs.  -->
-In the context of COVID-19, this might reinforce false or misleading information and prevent individuals from considering alternative perspectives.
+<!-- Social media algorithms are designed to show users content that aligns with their interests and beliefs, potentially contributing to _echo chambers_ [@cinelliEchoChamberEffect2021].  -->
+<!-- <!-- This might lead to confirmation bias, where individuals are more likely to be exposed to information that confirms their preexisting beliefs.  --> -->
+<!-- In the context of COVID-19, this might reinforce false or misleading information and prevent individuals from considering alternative perspectives. -->
 <!-- Cyberbullying and harassment:  -->
-Social media platforms are known for fostering negativity, with users sometimes engaging in _cyberbullying and harassment_ [@giumettiCyberbullyingSocialMedia2022]. 
+Social media platforms are known for fostering negativity, with users sometimes engaging in _cyberbullying and harassment_. 
 Discussions around COVID-19 can become heated and polarized, leading to personal attacks and online conflicts. 
@@
 Echoing the theoretical rationales outlines above, studies have yielded mixed results.
-Some studies found negative effects, indicating that excessive social media use for COVID-19 news led to compulsive behavior and increased stress levels, particularly due to upward social comparison [@kleinDarklySoothingCompulsion2021; @stainbackCOVID1924News2020; @yuePassiveSocialMedia2022].
+Some studies found negative effects, indicating that excessive social media use for COVID-19 news led to compulsive behavior and increased stress levels, particularly due to upward social comparison [@stainbackCOVID1924News2020; @yuePassiveSocialMedia2022].
 Individuals who relied on social media as their primary information source reported higher levels of anxiety and depression symptoms [@bendauAssociationsCOVID19Related2021]. 
@@
 Additionally, engaging more on social media was associated with reduced feelings of loneliness [@latikkaLonelinessPsychologicalDistress2022].
-A couple of studies reported mostly neutral effects of social media use on well-being indicators [@bradleyStressMoodSmartphone2021; @edenMediaCopingCOVID192020; @sewallObjectivelyMeasuredDigital2021].
+Several studies reported mostly neutral effects of social media use on well-being indicators [@edenMediaCopingCOVID192020; @sewallObjectivelyMeasuredDigital2021].
 Overall, the literature demonstrates a mixed picture, highlighting both positive and negative effects of social media use focused on or during COVID-19 on well-being.
@@
 <!-- In non-experimental studies, longitudinal designs can help investigate causality. -->
-When using longitudinal designs to analyze causality, it is important to (a) focus on within-person effects [@hamakerWhyResearchersShould2014]; to (b) control for confounders [@rohrerTheseAreNot2021]; and to (c) test a plausible interval between measures [@dormannOptimalTimeLags2015].
+When using longitudinal designs to analyze causality, it is important to (a) focus on within-person effects [@hamakerWhyResearchersShould2014]; to (b) control for confounders [@rohrerTheseAreNot2023]; and to (c) test a plausible interval between measures [@dormannOptimalTimeLags2015].
 <!-- To illustrate, consider the following example. -->
@@
 If a specific person changes their media diet, we need to measure how this behavior affects their well-being.
-Between-person comparisons from longitudinal data cannot provide such insights [@lucasWhyCrossLaggedPanel2022].
+Between-person comparisons from longitudinal data cannot provide such insights [@hamakerWhyResearchersShould2014].
 To test the hypothesis, I thus consider only the within-person effects.
@@
 Second, to identify confounders we should control for variables that affect both media use and well-being, which helps isolate the actual effect [@rohrerThinkingClearlyCorrelations2018].
-Because we are adopting a within-person perspective, we need to implement _time-varying_ confounders [@rohrerTheseAreNot2021]. 
+Because we are adopting a within-person perspective, we need to implement _time-varying_ confounders [@rohrerTheseAreNot2023]. 
 And because we are determining the _overall_ causal effect, we need to make sure _not_ to control for mediating variables [@rohrerThinkingClearlyCorrelations2018], for doing so would bias our assessment of the causal effect.
@@
 <!-- (I'll additionally include variables that likely affect only well-being, also to obtain a comparison benchmark for social media effects): -->
-gender, age, education, Austria country of birth, Austria country of birth of parents, residency Vienna, text-based news consumption, video-based news consumption, household size, health, living space, access to garden, access to balcony, employment, work hours per week, being in home-office, household income, outdoor activities, disposition to take risks, and locus of control [@egerStatisticalMetaanalysisWellbeing2015; @wardOriginsHappinessEvidence2016].
+gender, age, education, Austria country of birth, Austria country of birth of parents, residency Vienna, text-based news consumption, video-based news consumption, household size, health, living space, access to garden, access to balcony, employment, work hours per week, being in home-office, household income, outdoor activities, disposition to take risks, and locus of control [@egerStatisticalMetaanalysisWellbeing2015].
 <!-- [^explore] -->
@@
 
-Life satisfaction was measured with the item "All things considered, how satisfied are you with your life as a whole nowadays?", which comes from the European Social Survey [@europeansocialsurveyESS9Edition20182021].
+Life satisfaction was measured with the item "All things considered, how satisfied are you with your life as a whole nowadays?", which comes from the European Social Survey.
 The response options ranged from 0 (_extremely dissatisfied_) to 10 (_extremely satisfied_).
@@
 Only one positive effect emerged.
-Also note that in the analyses several control variables were included, ruling out plausible alternative explanations for the negative results.
-For example, it was controlled for as to whether or not participants contracted a COVID-19 infection during a specific wave.
-Hence, we can rule out the alternative explanation that having an infection was the root cause of increased communication and reduced well-being.
+<!-- Also note that in the analyses several control variables were included, ruling out plausible alternative explanations for the negative results. -->
+<!-- For example, it was controlled for as to whether or not participants contracted a COVID-19 infection during a specific wave. -->
+<!-- Hence, we can rule out the alternative explanation that having an infection was the root cause of increased communication and reduced well-being. -->
 
@@
 Liking, sharing, and posting COVID-19 related content, and spending more time on Twitter and YouTube to browse COVID-19 related content, all slightly negatively influenced affect. 
-This is aligned with prior findings which showed that social media use can trigger negative affect, but that it is less likely to determine life satisfaction [@huangTimeSpentSocial2017].
+This is aligned with prior findings which showed that social media use is associated with increased negative affect but not with life satisfaction [@meierComputermediatedCommunicationSocial2020a].
 <!-- Writing posts implies more substantial engagement, and perhaps also negative exchanges with other users, which could explain the results. -->
@@
 The hypothesis that tonality could explain the negative effects is especially supported by the observation that spending more time on Twitter and YouTube than usual increased negative affect.
-Communication on both channels is often found to be negative and impolite [e.g., @muellerTwitterMadeMe2022], also when compared to other SNSs [@halpernSocialMediaCatalyst2013].
+Communication on both channels is often found to be negative and impolite [e.g., @muellerTwitterMadeMe2022].
 Consuming more negative and misleading information could hence explain the (slightly) increased levels of negative affect.
@@
 The results are aligned with the findings from @valkenburgAssociationsActivePassive2022, who could not confirm that active use is good and that passive use is bad.
-Focusing on communication channels, Twitter and YouTube seem to be more negative, as has often been observed [@halpernSocialMediaCatalyst2013], while Instagram, WhatsApp, and Facebook were neutral.
+Focusing on communication channels, Twitter and YouTube seem to be more negative, while Instagram, WhatsApp, and Facebook were neutral.
 But, again, all of these effects are very small.
-Future research might elaborate on these specific relations to probe their stability and relevance.
+<!-- Future research might elaborate on these specific relations to probe their stability and relevance. -->
 
@@
 Both theories posit that if the effects of social media were indeed profoundly negative on average, then people likely would not spend so much time on social media engaging with COVID-19 content.
-Finally, recent empirical studies and meta-analyses reported rather small negative effects, too.
-Several studies found that the effects of various types of social media use on well-being are small, often too small to matter [@bendauAssociationsCOVID19Related2021; @meierComputermediatedCommunicationSocial2020a; @orbenTeenagersScreensSocial2020; @fergusonThisMetaanalysisScreen2021], echoing the results obtained here.
+Finally, recent empirical studies and meta-analyses reported rather small negative effects, too [@meierComputermediatedCommunicationSocial2020a], 
+<!-- Several studies found that the effects of various types of social media use on well-being are small, often too small to matter [@meierComputermediatedCommunicationSocial2020a],  -->
+echoing the results obtained here.
 
@@
 More thought needs to be invested in which factors to control for and, equally important, for which factors not to control for.
-I hope this study provides a first step into this direction.
+<!-- I hope this study provides a first step into this direction. -->
 
@@
 Because assessing well-being necessarily requires introspection, using self-reports for affect and life satisfaction is adequate.
-However, for social media use objective measures are clearly preferable, as people often cannot reliably estimate their use [@scharkowAccuracySelfreportedInternet2016].
-At the same time, most objective measures cannot capture the content or the motivation of use.
-Hence, for the type of research question analyzed here, it still seems necessary to use self-reported measures.
-In many cases they can still be informative [@verbeijSelfreportedMeasuresSocial2021].
+However, for social media use objective measures are preferable, as people often cannot reliably estimate their use [@scharkowAccuracySelfreportedInternet2016].
+<!-- At the same time, most objective measures cannot capture the content or the motivation of use. -->
+<!-- Hence, for the type of research question analyzed here, it still seems necessary to use self-reported measures. -->
+<!-- In many cases they can still be informative [@verbeijSelfreportedMeasuresSocial2021]. -->
 
 Being collected in a single country, the generalizability of the results is limited.
-The results apply primarily to the more Western sphere.
-They might not hold true in other cultures, especially cultures with a different media landscape or alternative social media channels.
+<!-- The results apply primarily to the more Western sphere. -->
+They might not hold true in other cultures, especially non-Western cultures with a different media landscape or alternative social media channels.
 <!-- That said, being a comparatively large study largely representative of one country, with several waves collected across a large time span, the results should be at least as generalizable as other typical empirical studies collected in the social sciences.
@@
 In light of the overall very small effects, engaging in COVID 19-related social media use should not be considered a major concern for one's well-being.
-Hence, when trying to improve well-being during a pandemic, instead of focusing on social media it seems more fruitful to address other, more pertinent societal problems related to health care, regular exercise, or psychological resilience.
+<!-- Hence, when trying to improve well-being during a pandemic, instead of focusing on social media it seems more fruitful to address other, more pertinent societal problems related to health care, regular exercise, or psychological resilience.
+-->
 

### 2023-08-08 manuscript.tex
- shortened ms

@@
 A new phenomenon termed ``doomscrolling'' emerged (Sharma et al., 2022).
-Many users were glued to their screens and found it hard to pursue other relevant activities such as working, taking a break, or even looking after their children (Klein, 2021).
+Many users were glued to their screens and found it hard to pursue other relevant activities such as working, taking a break, or even looking after their children.
 In the media it was hence increasingly asked whether using social media for COVID-19 related reasons would, next to all other stressors, create an additional burden on mental health (Sandstrom et al., 2021).
@@
 In general, according to the set-point theory, well-being is surprisingly stable (Lykken, 1999).
-Although specific events such as marriage or salary increase can have significant impacts on well-being, in most cases effects are only short-term, with well-being after some time returning to prior levels (Sheldon \& Lucas, 2014).
-Only very specific factors such as unemployment, disability, or death can cause long-term changes in well-being (Lucas, 2007).
+Although specific events such as marriage or salary increase can have significant impacts on well-being, in most cases effects are short-term, with well-being after some time returning to prior levels (Diener et al., 2018).
+Specific factors such as unemployment, disability, or death, however, can cause long-term changes in well-being (Lucas, 2007).
 So although well-being can be affected by external events and factors, this does not happen easily.
@@
 In conclusion, if a certain medium is used frequently, mood-management theory argues that it is likely not detrimental for well-being.
-On the other hand, although mood management theory suggests that effects should not be overly negative, it could be that although short-term effects are positive, long-term effects are negative.
-Precisely because social media have so many positive consequences in the short run, this might cause problem in the long run---for example, because social media use displaces other potentially more meaningful and/or relaxing activities (Hall \& Liu, 2022).
-As with many other things, there can be too much of a good thing.
-It is therefore often asked whether social media can become addictive, and users sometimes express this fear themselves (Yang et al., 2021).
 
@@
 Access to reliable information can help people make informed decisions, alleviate uncertainties, and feel empowered during the pandemic.
-Social media can be a valuable \emph{educational tool}, with various organizations and experts sharing informative content about COVID-19 (World Health Organization, 2023).
-Access to educational resources on social media can empower individuals to make informed decisions regarding their health.
 Social media platforms enable individuals to \emph{connect with others} who are experiencing similar challenges during the pandemic (Guazzini et al., 2022).
@@
 The rapid spread of news, updates, and opinions can be overwhelming and might exacerbate existing stress or fears about the pandemic (Sharma et al., 2022).
-Social media algorithms are designed to show users content that aligns with their interests and beliefs, potentially contributing to \emph{echo chambers} (Cinelli et al., 2021).
-In the context of COVID-19, this might reinforce false or misleading information and prevent individuals from considering alternative perspectives.
-Social media platforms are known for fostering negativity, with users sometimes engaging in \emph{cyberbullying and harassment} (Giumetti \& Kowalski, 2022).
+--\textgreater{}
+Social media platforms are known for fostering negativity, with users sometimes engaging in \emph{cyberbullying and harassment}.
 Discussions around COVID-19 can become heated and polarized, leading to personal attacks and online conflicts.
@@
 Echoing the theoretical rationales outlines above, studies have yielded mixed results.
-Some studies found negative effects, indicating that excessive social media use for COVID-19 news led to compulsive behavior and increased stress levels, particularly due to upward social comparison (Klein, 2021; Stainback et al., 2020; Yue et al., 2022).
+Some studies found negative effects, indicating that excessive social media use for COVID-19 news led to compulsive behavior and increased stress levels, particularly due to upward social comparison (Stainback et al., 2020; Yue et al., 2022).
 Individuals who relied on social media as their primary information source reported higher levels of anxiety and depression symptoms (Bendau et al., 2021).
@@
 Additionally, engaging more on social media was associated with reduced feelings of loneliness (Latikka et al., 2022).
-A couple of studies reported mostly neutral effects of social media use on well-being indicators (Bradley \& Howard, 2021; Eden et al., 2020; Sewall et al., 2021).
+Several studies reported mostly neutral effects of social media use on well-being indicators (Eden et al., 2020; Sewall et al., 2021).
 Overall, the literature demonstrates a mixed picture, highlighting both positive and negative effects of social media use focused on or during COVID-19 on well-being.
@@
 
-When using longitudinal designs to analyze causality, it is important to (a) focus on within-person effects (Hamaker, 2014); to (b) control for confounders (Rohrer \& Murayama, 2021); and to (c) test a plausible interval between measures (Dormann \& Griffin, 2015).
+When using longitudinal designs to analyze causality, it is important to (a) focus on within-person effects (Hamaker, 2014); to (b) control for confounders (Rohrer \& Murayama, 2023); and to (c) test a plausible interval between measures (Dormann \& Griffin, 2015).
 First, in non-experimental designs it makes much sense to analyze causal effects from an internal, within-person perspective (Hamaker, 2014).
 If a specific person changes their media diet, we need to measure how this behavior affects their well-being.
-Between-person comparisons from longitudinal data cannot provide such insights (Lucas, 2022).
+Between-person comparisons from longitudinal data cannot provide such insights (Hamaker, 2014).
 To test the hypothesis, I thus consider only the within-person effects.
@@
 Second, to identify confounders we should control for variables that affect both media use and well-being, which helps isolate the actual effect (Rohrer, 2018).
-Because we are adopting a within-person perspective, we need to implement \emph{time-varying} confounders (Rohrer \& Murayama, 2021).
+Because we are adopting a within-person perspective, we need to implement \emph{time-varying} confounders (Rohrer \& Murayama, 2023).
 And because we are determining the \emph{overall} causal effect, we need to make sure \emph{not} to control for mediating variables (Rohrer, 2018), for doing so would bias our assessment of the causal effect.
 In this study, I hence preregistered to control for the following variables, which either have already been shown or are likely to affect both social media use and well-being, and which also are not mediators:
-gender, age, education, Austria country of birth, Austria country of birth of parents, residency Vienna, text-based news consumption, video-based news consumption, household size, health, living space, access to garden, access to balcony, employment, work hours per week, being in home-office, household income, outdoor activities, disposition to take risks, and locus of control (Eger \& Maridal, 2015; Ward et al., 2016).
+gender, age, education, Austria country of birth, Austria country of birth of parents, residency Vienna, text-based news consumption, video-based news consumption, household size, health, living space, access to garden, access to balcony, employment, work hours per week, being in home-office, household income, outdoor activities, disposition to take risks, and locus of control (Eger \& Maridal, 2015).
 
@@
 
-Life satisfaction was measured with the item ``All things considered, how satisfied are you with your life as a whole nowadays?'', which comes from the European Social Survey (European Social Survey, 2021).
+Life satisfaction was measured with the item ``All things considered, how satisfied are you with your life as a whole nowadays?'', which comes from the European Social Survey.
 The response options ranged from 0 (\emph{extremely dissatisfied}) to 10 (\emph{extremely satisfied}).
@@
 Only one positive effect emerged.
-Also note that in the analyses several control variables were included, ruling out plausible alternative explanations for the negative results.
-For example, it was controlled for as to whether or not participants contracted a COVID-19 infection during a specific wave.
-Hence, we can rule out the alternative explanation that having an infection was the root cause of increased communication and reduced well-being.
 
@@
 Liking, sharing, and posting COVID-19 related content, and spending more time on Twitter and YouTube to browse COVID-19 related content, all slightly negatively influenced affect.
-This is aligned with prior findings which showed that social media use can trigger negative affect, but that it is less likely to determine life satisfaction (Huang, 2017).
+This is aligned with prior findings which showed that social media use is associated with increased negative affect but not with life satisfaction (Meier \& Reinecke, 2020).
 Conversations about COVID-19 on social media are often extreme, negative, or aggressive (L. Fan et al., 2020).
@@
 The hypothesis that tonality could explain the negative effects is especially supported by the observation that spending more time on Twitter and YouTube than usual increased negative affect.
-Communication on both channels is often found to be negative and impolite (e.g., Mueller \& Saeltzer, 2022), also when compared to other SNSs (Halpern \& Gibbs, 2013).
+Communication on both channels is often found to be negative and impolite (e.g., Mueller \& Saeltzer, 2022).
 Consuming more negative and misleading information could hence explain the (slightly) increased levels of negative affect.
@@
 The results are aligned with the findings from Valkenburg et al. (2022), who could not confirm that active use is good and that passive use is bad.
-Focusing on communication channels, Twitter and YouTube seem to be more negative, as has often been observed (Halpern \& Gibbs, 2013), while Instagram, WhatsApp, and Facebook were neutral.
+Focusing on communication channels, Twitter and YouTube seem to be more negative, while Instagram, WhatsApp, and Facebook were neutral.
 But, again, all of these effects are very small.
-Future research might elaborate on these specific relations to probe their stability and relevance.
 
@@
 Both theories posit that if the effects of social media were indeed profoundly negative on average, then people likely would not spend so much time on social media engaging with COVID-19 content.
-Finally, recent empirical studies and meta-analyses reported rather small negative effects, too.
-Several studies found that the effects of various types of social media use on well-being are small, often too small to matter (Bendau et al., 2021; Ferguson et al., 2021; Meier \& Reinecke, 2020; Orben, 2020), echoing the results obtained here.
+Finally, recent empirical studies and meta-analyses reported rather small negative effects, too (Meier \& Reinecke, 2020),
+echoing the results obtained here.
 
@@
 More thought needs to be invested in which factors to control for and, equally important, for which factors not to control for.
-I hope this study provides a first step into this direction.
 
@@
 Because assessing well-being necessarily requires introspection, using self-reports for affect and life satisfaction is adequate.
-However, for social media use objective measures are clearly preferable, as people often cannot reliably estimate their use (Scharkow, 2016).
-At the same time, most objective measures cannot capture the content or the motivation of use.
-Hence, for the type of research question analyzed here, it still seems necessary to use self-reported measures.
-In many cases they can still be informative (Verbeij et al., 2021).
+However, for social media use objective measures are preferable, as people often cannot reliably estimate their use (Scharkow, 2016).
 
 Being collected in a single country, the generalizability of the results is limited.
-The results apply primarily to the more Western sphere.
-They might not hold true in other cultures, especially cultures with a different media landscape or alternative social media channels.
+They might not hold true in other cultures, especially non-Western cultures with a different media landscape or alternative social media channels.
 
@@
 In light of the overall very small effects, engaging in COVID 19-related social media use should not be considered a major concern for one's well-being.
-Hence, when trying to improve well-being during a pandemic, instead of focusing on social media it seems more fruitful to address other, more pertinent societal problems related to health care, regular exercise, or psychological resilience.
 
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-bradleyStressMoodSmartphone2021}{}}%
-Bradley, A., \& Howard, A. (2021). \emph{Stress, mood, and smartphone use in {University} students: {A} 12-week longitudinal study}. \url{https://doi.org/10.31219/osf.io/frvpb}
-
-\leavevmode\vadjust pre{\hypertarget{ref-cinelliEchoChamberEffect2021}{}}%
-Cinelli, M., De Francisci Morales, G., Galeazzi, A., Quattrociocchi, W., \& Starnini, M.
-(2021).
-The echo chamber effect on social media.
-\emph{Proceedings of the National Academy of Sciences}, \emph{118}(9), e2023301118.
-\url{https://doi.org/10.1073/pnas.2023301118}
-
 \leavevmode\vadjust pre{\hypertarget{ref-dienerAdvancesOpenQuestions2018}{}}%
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-europeansocialsurveyESS9Edition20182021}{}}%
-European Social Survey. (2021). \emph{{ESS9} edition 3.1 - 2018 {Documentation Report}}. \url{https://www.europeansocialsurvey.org/docs/round9/survey/ESS9_data_documentation_report_e03_1.pdf}
-
 \leavevmode\vadjust pre{\hypertarget{ref-fanInformationOverloadWellbeing2021}{}}%
-Fan, J., \& Smith, A. P. (2021). Information overload, wellbeing and {COVID-19}: {A} survey in {China}. \emph{Behavioral Sciences}, \emph{11}(5, 5), 62. \url{https://doi.org/10.3390/bs11050062}
+Fan, J., \& Smith, A. P. (2021). Information overload, wellbeing and {COVID-19}: {A} survey in {China}. \emph{Behavioral Sciences}, \emph{11}(5), 62. \url{https://doi.org/10.3390/bs11050062}
 
@@
 (2020).
-Stigmatization in social media: {Documenting} and analyzing hate speech for {\textsc{COVID}} ‐19 on {Twitter}.
+Stigmatization in social media: {Documenting} and analyzing hate speech for {\textsc{COVID}} -19 on {Twitter}.
 \emph{Proceedings of the Association for Information Science and Technology}, \emph{57}(1).
 \url{https://doi.org/10.1002/pra2.313}
-
-\leavevmode\vadjust pre{\hypertarget{ref-fergusonThisMetaanalysisScreen2021}{}}%
-Ferguson, C.
-J., Kaye, L.
-K., Branley-Bell, D., Markey, P., Ivory, J.
-D., Klisanin, D., Elson, M., Smyth, M., Hogg, J.
-L., McDonnell, D., Nichols, D., Siddiqui, S., Gregerson, M., \& Wilson, J.
-(2021).
-Like this meta-analysis: {Screen} media and mental health.
-\emph{Professional Psychology: Research and Practice}.
-\url{https://doi.org/10.1037/pro0000426}
 
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-giumettiCyberbullyingSocialMedia2022}{}}%
-Giumetti, G. W., \& Kowalski, R. M. (2022). Cyberbullying via social media and well-being. \emph{Current Opinion in Psychology}, \emph{45}, 101314. \url{https://doi.org/10.1016/j.copsyc.2022.101314}
-
 \leavevmode\vadjust pre{\hypertarget{ref-guazziniSecondWaveAnalysis2022}{}}%
@@
 \url{https://doi.org/10.3390/ijerph19031635}
-
-\leavevmode\vadjust pre{\hypertarget{ref-hallSocialMediaUse2022}{}}%
-Hall, J. A., \& Liu, D. (2022). Social media use, social displacement, and well-being. \emph{Current Opinion in Psychology}, \emph{46}, 101339. \url{https://doi.org/10.1016/j.copsyc.2022.101339}
-
-\leavevmode\vadjust pre{\hypertarget{ref-halpernSocialMediaCatalyst2013}{}}%
-Halpern, D., \& Gibbs, J.
-(2013).
-Social media as a catalyst for online deliberation?
-{Exploring} the affordances of {Facebook} and {YouTube} for political expression.
-\emph{Computers in Human Behavior}, \emph{29}(3), 1159--1168.
-\url{https://doi.org/10.1016/j.chb.2012.10.008}
 
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-huangTimeSpentSocial2017}{}}%
-Huang, C.
-(2017).
-Time spent on social network sites and psychological well-being: {A} meta-analysis.
-\emph{Cyberpsychology, Behavior and Social Networking}, \emph{20}(6), 346--354.
-\url{https://doi.org/10.1089/cyber.2016.0758}
-
 \leavevmode\vadjust pre{\hypertarget{ref-huntSocialMediaBased2022}{}}%
@@
 (2022).
-A social media‒based public health campaign encouraging {COVID-19} vaccination across the {United States}.
+A social media-based public health campaign encouraging {COVID-19} vaccination across the {United States}.
 \emph{American Journal of Public Health}, \emph{112}(9), 1253--1256.
@@
 \leavevmode\vadjust pre{\hypertarget{ref-johnhopkinsuniversityCOVID19Map2023}{}}%
-John Hopkins University. (2023). \emph{{COVID-19 Map}}. {Johns Hopkins Coronavirus Resource Center}. \url{https://coronavirus.jhu.edu/map.html}
+John Hopkins University. (2023). {COVID-19 Map}. In \emph{Johns Hopkins Coronavirus Resource Center}. https://coronavirus.jhu.edu/map.html.
 
@@
 (2020).
-\emph{Austrian {Corona Panel Project} ({SUF} edition)} {[}Data set{]}.
+\emph{Austrian {Corona Panel Project} ({SUF} edition)}.
 {AUSSDA}.
 \url{https://doi.org/10.11587/28KQNS}
-
-\leavevmode\vadjust pre{\hypertarget{ref-kleinDarklySoothingCompulsion2021}{}}%
-Klein, J. (2021, March 3). \emph{The darkly soothing compulsion of 'doomscrolling'}. \url{https://www.bbc.com/worklife/article/20210226-the-darkly-soothing-compulsion-of-doomscrolling}
 
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-lucasWhyCrossLaggedPanel2022}{}}%
-Lucas, R. E. (2022). \emph{Why the {Cross-Lagged Panel Model} is {Almost Never} the {Right Choice}} {[}Preprint{]}. {PsyArXiv}. \url{https://doi.org/10.31234/osf.io/pkec7}
-
 \leavevmode\vadjust pre{\hypertarget{ref-lykkenHappinessWhatStudies1999}{}}%
@@
 \emph{Medical Care}, \emph{41}(5), 582--592.
-\href{Retrieved\%20from\%20http://www.jstor.org/stable/3768017}{Retrieved from http://www.jstor.org/stable/3768017}
-
-\leavevmode\vadjust pre{\hypertarget{ref-orbenTeenagersScreensSocial2020}{}}%
-Orben, A.
-(2020).
-Teenagers, screens and social media: A narrative review of reviews and key studies.
-\emph{Social Psychiatry and Psychiatric Epidemiology}, \emph{55}(4), 407--414.
-\url{https://doi.org/10.1007/s00127-019-01825-4}
 
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-rohrerTheseAreNot2021}{}}%
+\leavevmode\vadjust pre{\hypertarget{ref-rohrerTheseAreNot2023}{}}%
 Rohrer, J.
 M., \& Murayama, K.
-(2021, October 11).
-\emph{These are not the effects you are looking for: {Causality} and the within-/between-person distinction in longitudinal data analysis}.
-\url{https://doi.org/10.31234/osf.io/tg4vj}
+(2023).
+These are not the effects you are looking for: {Causality} and the within-/between-persons distinction in longitudinal data analysis.
+\emph{Advances in Methods and Practices in Psychological Science}, \emph{6}(1), 25152459221140842.
+\url{https://doi.org/10.1177/25152459221140842}
 
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-sheldonStabilityHappinessTheories2014}{}}%
-Sheldon, K. M., \& Lucas, R. E. (2014). \emph{Stability of happiness: Theories and evidence on whether happiness can change}. \url{http://site.ebrary.com/id/10891875}
-
 \leavevmode\vadjust pre{\hypertarget{ref-stainbackCOVID1924News2020}{}}%
@@
 Statista.
-(2021, May 21).
+(2021).
 \emph{Average daily time spent on social networks by users in the {United States} from 2018 to 2022}.
-\url{https://www.statista.com/statistics/1018324/us-users-daily-social-media-minutes/}
+https://www.statista.com/statistics/1018324/us-users-daily-social-media-minutes/.
 
 \leavevmode\vadjust pre{\hypertarget{ref-twitterTalkingMentalHealthAwarenessTwitter2020}{}}%
-Twitter.
-(2020, June 15).
-\emph{Talking \#{MentalHealthAwareness} on {Twitter} with our global partners}.
-\url{https://blog.twitter.com/en_us/topics/company/2020/talking-mental-health-awareness-on-twitter}
+Twitter. (2020). \emph{Talking \#{MentalHealthAwareness} on {Twitter} with our global partners}. https://blog.twitter.com/en\_us/topics/company/2020/talking-mental-health-awareness-on-twitter.
 
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-verbeijSelfreportedMeasuresSocial2021}{}}%
-Verbeij, T., Pouwels, J.
-L., Beyens, I., \& Valkenburg, P.
-M.
-(2021).
-\emph{Self-reported measures of social media use show high predictive validity} {[}Preprint{]}.
-{PsyArXiv}.
-\url{https://doi.org/10.31234/osf.io/c9bj7}
-
 \leavevmode\vadjust pre{\hypertarget{ref-verduynPassiveFacebookUsage2015}{}}%
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-wardOriginsHappinessEvidence2016}{}}%
-Ward, G., Powdthavee, N., Fleche, S., Layard, R., \& Clark, A.
-(2016, December 12).
-\emph{Origins of happiness: {Evidence} and policy implications}.
-{CEPR}.
-\url{https://cepr.org/voxeu/columns/origins-happiness-evidence-and-policy-implications}
-
 \leavevmode\vadjust pre{\hypertarget{ref-worldhealthorganizationWellbeingMeasuresPrimary1998}{}}%
 World Health Organization. (1998). \emph{Wellbeing measures in primary health care/{The Depcare Project}}.
-
-\leavevmode\vadjust pre{\hypertarget{ref-worldhealthorganizationCoronavirusDiseaseCOVID192023}{}}%
-World Health Organization. (2023). \emph{Coronavirus disease ({COVID-19}) -- {World Health Organization}}. \url{https://www.who.int/emergencies/diseases/novel-coronavirus-2019}
-
-\leavevmode\vadjust pre{\hypertarget{ref-yangCanWatchingOnline2021}{}}%
-Yang, Z., Griffiths, M.
-D., Yan, Z., \& Xu, W.
-(2021).
-Can watching online videos be addictive?
-{A} qualitative exploration of online video watching among {Chinese} young adults.
-\emph{International Journal of Environmental Research and Public Health}, \emph{18}(14, 14), 7247.
-\url{https://doi.org/10.3390/ijerph18147247}
 

### 2023-08-24 analyses.RmD
- shortened manuscript for submission

@@
 ```
+
+# SESOI
+
+In what follows, more information on how the SESOI was defined exactly.
+
+If people can reliably differentiate 7 levels, this corresponds to 11 / 7 = `r round(11/7, 2)` unit change on an 11-point scale.
+Hence, a four-point change in media use (e.g., a complete stop) should result in a `r round(11/7, 2)`-point change in life satisfaction.
+In a statistical regression analysis, _b_ estimates the change in the dependent variable if the independent variable increases by one point.
+For life satisfaction, we would therefore define a SESOI of _b_ = `r round(11/7, 2)` / 4 = `r round(11/7*.25, 2)`.
+For positive or negative affect, which was measured on a 5-point scale, our SESOI would be _b_ = `r round(5 / 7, 2)` / 4 = `r round(5 / 7 * .25, 2)`.
+Because we are agnostic as to whether the effects are positive or negative, the null region includes both negative and positive effects.
+Finally, in order not to exaggerate precision and to be less conservative, these numbers are reduced to nearby thresholds.
+Note that other researchers also decreased or recommended decreasing thresholds for effect sizes when analyzing within-person or cumulative effects [@beyensSocialMediaUse2021; @funderEvaluatingEffectSize2019].
 

### 2023-08-24 manuscript.Rmd
- shortened manuscript for submission

@@
 footnotelist      : no
-linenumbers       : yes
+linenumbers       : no
 mask              : yes
@@
 [e.g., @bendauAssociationsCOVID19Related2021; @edenMediaCopingCOVID192020; @sewallObjectivelyMeasuredDigital2021],
-<!-- [@bendauAssociationsCOVID19Related2021; @bradleyStressMoodSmartphone2021; @choiMediatedCommunicationMatters2021; @dornemannHowGoodBad2021; @edenMediaCopingCOVID192020; @guazziniSecondWaveAnalysis2022; @latikkaLonelinessPsychologicalDistress2022; @liuRelationOfficialWhatsAppdistributed2020; @riehmAssociationsMediaExposure2020; @sewallObjectivelyMeasuredDigital2021; @stainbackCOVID1924News2020; @yuePassiveSocialMedia2022] -->
 it still largely unknown if COVID-19 related social media use during the pandemic has had a meaningful impact on well-being.
@@
 
-<!-- Two underlying theories guided the selection of variables for this study, namely the two-continua model of mental health [@greenspoonIntegrationSubjectiveWellbeing2001] and the hierarchical taxonomy of computer-mediated communication [@meierComputermediatedCommunicationSocial2020a].
--->
-<!-- According to the two-continua model, mental health consists of (a) psychopathology and (b) well-being. -->
-<!-- <!-- Because the aim of this study is better to understand typical users and everyday contexts, my focus will be on well-being. -->
-<!-- Well-being can be differentiated into subjective and psychological well-being [@dienerAdvancesOpenQuestions2018]. -->
-<!-- Whereas subjective well-being emphasizes hedonic aspects such as happiness and joy, psychological well-being addresses eudaimonic aspects such as fulfillment and meaning. -->
-<!-- Subjective well-being is primarily about achieving positive and avoiding negative affect. -->
-<!-- One of the most prominent indicators of well-being is life satisfaction. -->
-<!-- In my view, because it represents a general appraisal of one's life, life satisfaction is best thought of as a meta concept combining psychological and subjective well-being. -->
-<!-- Notably, life satisfaction is stable and fluctuates only little, whereas it's the exact opposite for affect [@dienlinImpactDigitalTechnology2020]. -->
-<!-- <!-- Everyone can feel bad on any given day, but whether or not one feels generally comfortable in one's life rather changes across years. -->
-<!-- To capture well-being in this study I thus build on life satisfaction, positive affect, and negative affect. -->
-<!-- Together, this should provide an encompassing perspective on potential media effects. -->
-
-<!-- The hierarchical taxonomy of computer-mediated communication differentiates six levels of how people engage with digital technology. -->
-<!-- First, the device (e.g., smartphone); second, the type of application (e.g., social networking site); third, the branded application (e.g., Twitter); fourth, the feature (e.g., status post); fifth, the interaction (e.g., one-to-many); and sixth, the message (e.g., content) [@meierComputermediatedCommunicationSocial2020a].
--->
-<!-- Whereas the first four levels focus on the communication _channel_, the last two address the communication _type_. -->
-<!-- Distinguishing different communication channels and communication types is important, because the effects on well-being likely differ across communication channels and communication types. -->
-<!-- Whereas active social media use such as chatting is routinely linked to improved well-being, passive use such as reading is often considered more negative [@dienlinImpactDigitalTechnology2020].
--->
-<!-- Similarly, branded apps are separate communication entities with potentially divergent effects and affordances. -->
-<!-- For example, @waterlooNormsOnlineExpressions2018 found that it's more adequate to express negative emotions on WhatsApp than on Twitter or on Instagram. -->
-<!-- Especially during a pandemic, it makes sense to analyze if users engage with COVID-19 related content on Instagram, where communication is more positive, or on Facebook, where communication is more critical.
--->
-<!-- First studies suggest that during the pandemic Instagram use was indeed more beneficial for well-being than Facebook use [@masciantonioDonPutAll2021]. -->
-<!-- In this study, to measure the effects of social media use focused on COVID-19 related news and topics, I adopt both the channel and the type of communication perspective. -->
-<!-- Together, this should offer a nuanced and comprehensive understanding of communication. -->
-
 This study investigates how different _facets_ of well-being are affected by different _types_ and different _channels_ of communication [@meierComputermediatedCommunicationSocial2020a].
-<!-- , namely active and passive use [@ellisonWhyWeDon2020; @meierComputermediatedCommunicationSocial2020a]. -->
-<!-- Defining what constitutes active and what passive use is not always clear, and different understandings are currently discussed [@ellisonWhyWeDon2020; @meierComputermediatedCommunicationSocial2020a].
--->
-<!-- Reading is generally considered as passive and writing as active use, while there are also specific behaviors falling somewhere in-between such as liking or sharing content [@meierDoesPassiveSocial2022].
--->
-<!-- In this study,  -->
 Building on the typology of subjective well-being [@dienerAdvancesOpenQuestions2018], three different well-being facets are analyzed: life satisfaction, positive affect, and negative affect.
@@
 
-<!-- <!-- In conclusion, most effects are likely somewhere between trivial and small. -->
-<!-- <!-- I therefore expect that also in the case of COVID-19 related social media use effects will be trivial to small. -->
-
-<!-- In what follows, I outline potential arguments as to why the effect might be positive or negative, direct or indirect, or nonexistent. -->
 How easily can well-being be affected by external influences?
@@
 So although well-being can be affected by external events and factors, this does not happen easily.
-<!-- Although the set-point theory has been criticized lately because events such as marriage, unemployment, or death can indeed lead to long-term changes, it is still widely maintained that well-being is very stable and hence hard to be influenced  -->
 
@@
 In conclusion, if a certain medium is used frequently, mood-management theory argues that it is likely not detrimental for well-being.
-<!-- For example, using experience sampling of well-being and logs of social media use, a study with 82 participants from Italy found that after episodes of social media use, levels of positive affect increased significantly [@marcianoDynamicsAdolescentsSmartphone2022].
--->
-<!-- INCLUDE IF SPACE -->
-<!-- On the other hand, although mood management theory suggests that effects should not be overly negative, it could be that although short-term effects are positive, long-term effects are negative.
--->
-<!-- Precisely because social media have so many positive consequences in the short run, this might cause problem in the long run---for example, because social media use displaces other potentially more meaningful and/or relaxing activities [@hallSocialMediaUse2022].
--->
-<!-- As with many other things, there can be too much of a good thing. -->
-<!-- It is therefore often asked whether social media can become addictive, with users sometimes expressing this fear themselves. -->
-<!-- END -->
-<!-- it is still possible that people misjudge media effects. -->
-<!-- Users are often overly optimistic [@metzgerComparativeOptimismPrivacy2017], focusing on short term benefits while ignoring long-term costs.  -->
-<!-- In other words, social media might not problematic because they are inherently bad, but rather because they are too good. -->
-<!-- However, a recently published meta-analysis found that the two most prominent measures of addiction, the Bergen Facebook Addiction Scale and the Bergen Social Media Addiction Scale, have only small relations to well-being [@duradoniWellbeingSocialMedia2020].
--->
-<!-- In addition, the general idea of labeling excessive social and new media use as addiction was criticized, arguing that social media use represent a new regular behavior that should not be pathologized [@galerHowMuchToo2018; @vanrooijWeakScientificBasis2018].
--->
-<!-- In conclusion, there are several plausible theoretical mechanisms as to why effects might be negative or positive, explaining why on average effects might not be particularly negative or positive.
--->
 
@@
 Several uses and gratifications exist, which help explain why people used social media frequently during the pandemic.
-<!-- , outlining mechanisms of potential positive effects.  -->
-<!-- At the same time, there are also several plausible reasons as to why COVID-19 related social media use is beneficial for well-being. -->
-<!-- Access to information and resources:  -->
 Despite incorrect information, social media provide a vast platform for disseminating _accurate and timely information_ about COVID-19 [@johnhopkinsuniversityCOVID19Map2023]. 
-<!-- It allows individuals to stay informed about the latest updates, guidelines, and recommendations from reputable health organizations and experts.  -->
 Access to reliable information can help people make informed decisions, alleviate uncertainties, and feel empowered during the pandemic.
-<!-- Educational opportunities:  -->
-<!-- Social media can be a valuable _educational tool_, with various organizations and experts sharing informative content about COVID-19 [@worldhealthorganizationCoronavirusDiseaseCOVID192023].  -->
-<!-- <!-- This can include explanatory videos, infographics, and articles that help individuals understand the virus, its transmission, prevention strategies, and vaccination information.  -->
-<!-- Access to educational resources on social media can empower individuals to make informed decisions regarding their health. -->
-<!-- Community support and solidarity:  -->
 Social media platforms enable individuals to _connect with others_ who are experiencing similar challenges during the pandemic [@guazziniSecondWaveAnalysis2022]. 
-<!-- Sharing experiences, offering support, and expressing solidarity can foster a sense of community and reduce feelings of isolation.  -->
 Engaging in online communities and support groups can provide emotional support and create a network of like-minded individuals.
-<!-- Promoting mental health and well-being:  -->
 Many mental health organizations and professionals utilize social media to share tips, strategies, and resources for _maintaining mental well-being_ during the pandemic [@twitterTalkingMentalHealthAwarenessTwitter2020].
-<!-- They provide guidance on coping mechanisms, stress reduction techniques, and self-care practices.  -->
 Engaging with such content might help individuals prioritize their mental health and develop resilience during challenging times.
-<!-- Encouraging positive behaviors:  -->
 Social media campaigns and initiatives can promote _positive COVID-19 behaviors_, such as mask-wearing, physical distancing, hand hygiene, and vaccination [@atheyDigitalPublicHealth2023; @huntSocialMediaBased2022].
 Public health organizations and influencers leverage the power of social media to spread awareness and encourage responsible actions, contributing to public health efforts and fostering a sense of collective responsibility.
-<!-- Amplifying diverse voices and perspectives:  -->
-<!-- Social media platforms provide a space for diverse voices and perspectives to be heard [@brunsAreFilterBubbles2019].  -->
-<!-- This allows individuals to access a range of viewpoints, experiences, and cultural practices related to COVID-19.  -->
-<!-- Exposure to diverse perspectives might promote empathy, understanding, and a more comprehensive understanding of the pandemic's impact on different communities. -->
 
 On the other hand, the effects might be negative, perhaps best explained by the following five mechanisms.
-<!-- Misinformation and rumors:  -->
 Social media platforms can easily spread _false or misleading information_ about COVID-19 [@liYouTubeSourceInformation2020]. 
 Due to the ease of sharing and the lack of fact-checking, inaccurate information can go viral and might cause confusion, anxiety, and panic among users.
-<!-- Overwhelm and anxiety:  -->
 Constant exposure to COVID-19-related content on social media can lead to _information overload_ and contribute to heightened anxiety levels [@fanInformationOverloadWellbeing2021]. 
 The rapid spread of news, updates, and opinions can be overwhelming and might exacerbate existing stress or fears about the pandemic [@sharmaDarkEndTunnel2022].
-<!-- Confirmation bias and echo chambers:  -->
-<!-- Social media algorithms are designed to show users content that aligns with their interests and beliefs, potentially contributing to _echo chambers_ [@cinelliEchoChamberEffect2021].  -->
-<!-- <!-- This might lead to confirmation bias, where individuals are more likely to be exposed to information that confirms their preexisting beliefs.  --> -->
-<!-- In the context of COVID-19, this might reinforce false or misleading information and prevent individuals from considering alternative perspectives. -->
-<!-- Cyberbullying and harassment:  -->
 Social media platforms are known for fostering negativity, with users sometimes engaging in _cyberbullying and harassment_. 
@@
 Such experiences threaten mental well-being and might contribute to feelings of distress and isolation.
-<!-- Social comparison and fear of missing out (FOMO):  -->
 Social media often showcase the highlights and accomplishments of others, encouraging _social comparison_ [@przybylskiMotivationalEmotionalBehavioral2013]. 
 During a pandemic, seeing posts about others' successes or seemingly perfect lives might intensify feelings of inadequacy or FOMO, especially when individuals are unable to participate in similar activities due to restrictions or personal circumstances [@sharmaDarkEndTunnel2022].
-<!-- Reduced face-to-face interaction:  -->
-<!-- Spending excessive time on social media to learn about COVID-19 can lead to a decrease in face-to-face interactions and real-life connections [but see @hallSocialMediaUse2022].  -->
-<!-- This might impact mental health, as most humans have an innate need for social support and physical presence, which online interactions might not fully replicate. -->
 
-<!-- Using Social media is ambivalent. -->
-<!-- It can impair well-being when causing embarrassment, stress, or disinformation, and it can improve well-being when providing connectedness, information, or entertainment [@buchiDigitalWellbeingTheory2021].
--->
-<!-- Although social media are often associated with negative outcomes, there are a large number of positive effects, which either offset or at least significantly reduce negative outcomes. -->
-<!-- Positive outcomes encompass finding relevant information; maintaining and fostering relationships; expressing one's personality; or entertaining oneself [e.g., @pelletierOneSizeDoesn2020]. -->
-<!-- Negative effects include distraction; displacement of other meaningful activities; consumption of shallow, misleading or challenging content; or negative social comparisons [e.g., @meierDoesPassiveSocial2022].
--->
-<!-- People are much more likely to use media that offer plentiful benefits [@katzUsesGratificationsResearch1973]. -->
-<!-- Users implicitly learn what media help them regulate their mood and thereby well-being [@zillmannMoodManagementCommunication1988]. -->
-<!-- Given the ubiquitous use of social media, the many benefits they offer, plus the general and implicit competence to use media that foster mood management, it is hence unlikely that the effects of social media use on well-being are profusely negative.
--->
-
-<!-- As outline above,   -->
 There is still little empirical research on how well-being is affected by social media use that is focused on COVID-19 specifically.
@@
 
-<!-- ## Empirical Studies on Social Media Effects -->
-
-<!-- There is still little empirical research on how well-being is affected by social media use that is focused on COVID-19 specifically. -->
-<!-- Results are mixed. -->
-<!-- When browsing social media for COVID-19 related news, many users reported being captivated to such an extent they could not stop using social media [@kleinDarklySoothingCompulsion2021]. -->
-<!-- During the pandemic social media use was at an all-time high in the US [@statistaAverageDailyTime2021], for example to cope with the required physical distancing measures. -->
-<!-- In a study with 1,131 residents from Wuhan in China, people who spent more time in quarantine also spent more time on social media [@yuePassiveSocialMedia2022].  -->
-<!-- Those who spent more time on social media also engaged in more upward social comparison, which was related to increased levels of stress. -->
-<!-- People who used social media as a primary source of information reported on average "significantly more unspecific anxiety and depression [...] and significantly more specific COVID-19 related anxiety symptoms" [@bendauAssociationsCOVID19Related2021, p.
-288].
--->
-<!-- @edenMediaCopingCOVID192020 analyzed how 425 US college students used media during the first wave of the pandemic, finding both positive and negative relations with well-being. -->
-<!-- In a sample of 312 respondents collected via Amazon Mechanical Turk, people who used media to attain information were more lonely and less satisfied with their lives [@choiMediatedCommunicationMatters2021].
--->
-<!-- In a large-scale study with 11,537 respondents from the US, increased COVID-19-related media consumption was related to more psychological distress [@stainbackCOVID1924News2020]. -->
-<!-- A four-wave panel study with 384 young adults from the U.S.
-analyzed the effects of general digital technology use during the pandemic on mental health, finding that digital technology did not have significant effects on mental health [@sewallObjectivelyMeasuredDigital2021; for a similar study with comparable results, see @bradleyStressMoodSmartphone2021].
--->
-<!-- A study with 2.057 respondents from Italy reported that during the pandemic virtual community and social connectedness increased [@guazziniSecondWaveAnalysis2022]. -->
-<!-- In a study with 735 participants from Finland, levels of loneliness were stable during the pandemic, and people who engaged more on social media experienced less loneliness [@latikkaLonelinessPsychologicalDistress2022].
--->
-<!-- Together, the literature is mixed, showing both positive and negative affects of social media use focused on or during COVID-19 on well-being [see also @dornemannHowGoodBad2021; @liuRelationOfficialWhatsAppdistributed2020; @riehmAssociationsMediaExposure2020].
--->
-
 In conclusion, given these mixed empirical results, together with the observation that social media effects on well-being are very small in general, and that several plausible theoretical mechanisms exist for both positive and negative effects, I expect that COVID-19 related communication on social media should not be decidedly positive or negative.
@@
 
-<!-- Although the average effect of use on well-being is likely very small, it might be that specific subgroups exhibit larger effects. -->
+## Smallest Effect Size of Interest
 
-<!-- > Research question: How do the effects of COVID-19 related social media use on well-being depend on gender, age, immigration background, education, and urban living context? -->
+Exploring the hypothesis entails establishing the criteria for discerning a 'trivial effect size.' 
+To achieve this, it is necessary to define the smallest effect size of interest (SESOI) [@anvariUsingAnchorbasedMethods2021a]. 
+In this context, a trivial effect should fall below the threshold set by the SESOI (detailed below).
+Determining what constitutes a minimally intriguing, nontrivial effect is a matter with normative implications, making it challenging to arrive at a definitive, singular consensus. 
+Nonetheless, it remains both essential and beneficial to strive for a reasonable benchmark. 
+With that in mind, I propose the following SESOI as a suitable reference point for this study:
 
-<!-- Because media effects can differ across users, situations, and content [@livingstoneEuropeanResearchChildren2018; @valkenburgDifferentialSusceptibilityMedia2013], I now briefly focus on the effects of COVID-19 related social media use specifically.
--->
-<!-- First, one could assume a _direct_ negative effect on well-being, and especially on positive or negative affect, which are more volatile and fluctuating. -->
-<!-- Dangers, inequalities, corruption---these were the headlines during the pandemic across many countries worldwide. -->
-<!-- If one learns about such events, the initial reaction might be shock, fear, or dismay. -->
-<!-- Consuming such news can be depressing [@dornemannHowGoodBad2021], perhaps even changing some general perspectives on life. -->
-<!-- <!-- So, just like being hit by a hammer hurts and we don't need any "mediating mechanism", this could be the case here as well. -->
-<!-- That said, because not all news was negative, and because many people showed solidarity and compassion, there was also positive and uplifting content, potentially compensating for the negative effects [@dornemannHowGoodBad2021].
--->
-<!-- A study with 2.057 respondents from Italy reported that during the pandemic virtual community and social connectedness even increased [@guazziniSecondWaveAnalysis2022]. -->
+> SESOI: If a heavy user of COVID-19 related social media news suddenly _stops_ using social media altogether, this should have a _noticeable_ impact on their overall well-being.
 
-<!-- In a study with 735 participants from Finland, levels of loneliness did not decrease during the pandemic, and people who engaged more on social media experienced less loneliness [@latikkaLonelinessPsychologicalDistress2022].
--->
-<!-- However, in light of a worldwide pandemic with millions of deaths, the negative direct effect seems more plausible. -->
-
-<!-- Second, there could also be _indirect_ effects. -->
-<!-- Although it is most likely that moderate social media use is not detrimental [@orbenTeenagersScreensSocial2020], overuse, however, might be more critical, and several studies have shown more pronounced negative effects for extreme users [@przybylskiLargescaleTestGoldilocks2017].
--->
-<!-- To explain, overuse could impair well-being if it replaces meaningful or functional activities such as meeting others, working, actively relaxing, or exercising. -->
-<!-- Another potentially negative mechanism at play are problematic social comparison processes.  -->
-<!-- Therefore, if a society collectively overuses social media during a pandemic, there is potential for negative effects. -->
-<!-- If overuse replaces such activities it's reasonable to assume that it's also detrimental. -->
-
-<!-- On the other hand, one can make the case that using social media for COVID-19 related reasons might even be beneficial, especially in times of a pandemic. -->
-<!-- Exchanging COVID-19 related messages with friends via WhatsApp might replace the in-person contact one would have otherwise, but which is literally impossible at the time. -->
-<!-- In situations where meaningful and functional activities are prohibited, using social media to exchange about COVID-19 related topics might not be the worst idea. -->
-<!-- Besides, given that nowadays a large number of experts, scientists, and politicians converse directly on social media, one can get first-hand high quality information on current developments. -->
-<!-- <!-- On the other hand, there is of course also much disinformation, and "bingeing" on COVID-19 fake news might also pose risks for impaired well-being. -->
-
-<!-- To summarize, it seems that from a theoretical perspective it is most likely that the average effects of social media use on well-being are negligible.  -->
-<!-- Building on established theories from Communication and current empirical findings, we would not assume that effects are either profoundly negative or strongly positive. -->
-
-<!-- However, note that all of these findings represent between-person relations stemming from cross-sectional data (see below). -->
-<!-- We therefore don't know whether the differences in mental health and well-being are due to social media use or due to other third variables, such as age, health, employment, or education. -->
-
-<!-- The question of whether and how social media use affects well-being _in general_, on the other hand, is well-researched. -->
-<!-- This also holds true for the different types of communication such as active or passive use. -->
-<!-- What determines whether or not an effect should be considered small or trivial? -->
-<!-- As a starting point, we could refer to standardized effect sizes. -->
-<!-- According to @cohenPowerPrimer1992, small effect sizes start at _r_ = .10. -->
-<!-- Finally, several individual studies employing advanced methods found smalls relations between social media use and well-being [@orbenSocialMediaEnduring2019; @przybylskiDoesTakingShort2021a; @schemerImpactInternetSocial2021; @kerestesAdolescentsOnlineSocial2020].
--->
-<!-- Although for roughly one quarter of all users the effects of social media use on well-being were negative, for almost the same number of users they were positive, while for the rest the effects were neutral [@beyensSocialMediaUse2021].
--->
-<!-- This finding is aligned with the Differential Susceptibility to Media Effects Model:  -->
-<!-- Although there is substantial _variation_ of media effects for individual users, the _average_ effects reported in the literature are often small [@valkenburgDifferentialSusceptibilityMedia2013].
--->
-
-<!-- Together, the strongest argument to me is that _in general_ the effects of social media on well-being are, on average, small at best.  -->
-<!-- Because this study only looks at _one part_ of social media use---namely, COVID-19 related interactions---it is very focused, diminishing the overall potential of the effects even further. -->
-<!-- Whether or not using social media for COVID-19 related aspects is detrimental during a pandemic is also not entirely clear. -->
-<!-- At least, there does not seem to be a case where we should expect a clear deviation to the negative, despite anecdotes of doomscrolling. -->
-
-<!-- A meta review (i.e., an analysis of meta-analyses) found that the relation between social media use and well-being is likely in the negative spectrum but very small, potentially too small to matter [@meierComputermediatedCommunicationSocial2020a].
--->
-<!-- And indeed, several if not most of the current meta-analyses find small to trivial effects [@huangTimeSpentSocial2017; @meierComputermediatedCommunicationSocial2020a; @fergusonThisMetaanalysisScreen2021].
--->
-
-<!-- In this study, this general hypothesis will be analyzed specifically for the three communication types of (a) time spent reading, (b) liking and sharing, and (c) actively posting COVID-19 related content.
--->
-<!-- In addition, I will analyze how well-being is influenced by spending time on five prominent social media apps, including (a) Facebook, (b) Instagram, (c) Twitter, (d) WhatsApp, and (e) YouTube.
--->
+What does this mean practically and how can it be operationalized?
+In this study, COVID-19 related social media use was measured on a 5-point scale, ranging from 1 = _never_ to 5 = _several times a day_.
+Thus, a change of four units in social media use (e.g., a complete stop) should correspond to a noticeable change in well-being.
+According to @normanInterpretationChangesHealthrelated2003, people can reliably notice seven levels of change in satisfaction with health.
+So if satisfaction is measured on a 7-point scale, a four unit change in social media use should result in a one unit change in life satisfaction. 
+In this study, life satisfaction was measured on an 11-point scale, and affect on a 5-point scale. 
+Transposed to this scaling, the SESOI for life satisfaction is _b_ =  ±.30, and for positive and negative affect _b_ = ±.15 (for more information, see online supplementary material).
 
@@
 
-<!-- # Current Study -->
-## Smallest Effect Size of Interest
-<!-- ## Inference Criteria -->
+## Data Analysis
 
-Testing the hypothesis necessitates defining what is considered a "trivial effect size". 
-<!-- Effectively testing this hypothesis  -->
-<!-- Above I already referred to standardized effect sizes. -->
-<!-- However, standardized effect sizes should only be a first step toward evaluating an effect's relevance [@baguleyStandardizedSimpleEffect2009]. -->
-<!-- Standardized effect sizes are determined by a sample's variance,[^cohensd] which is problematic:  -->
-<!-- The question of whether or not social media use affects a particular person in a relevant way should not depend on the variance in the sample in which that person's data were collected. -->
-<!-- Instead, it should depend on absolute criteria. -->
-To this end, we need to define a so-called smallest effect size of interest (SESOI) [@lakensEquivalenceTestingPsychological2018].
-A trivial effect would then need to be smaller than the SESOI (see below).
-What could be a minimally interesting, nontrivial effect?
-Being a normative question, finding a clear, single, or unanimous answer is impossible.
-<!-- In the end, it is a personal question. -->
-However, it is still necessary and helpful to work toward a plausible benchmark. 
-I suggest the following SESOI for this research question:
-
-> SESOI: If a heavy user of COVID-19 related social media news suddenly _stops_ using social media altogether, this should have a _noticeable_ impact on their overall well-being.
-
-What does this mean practically and how can it be operationalized?
-In this study, COVID-19 related social media use was measured on a 5-point scale, ranging from 1 = _never_ to 5 = _several times a day_.
-Thus, a change of four units in social media use (e.g., a complete stop) should correspond to a noticeable change in well-being.
-What is a noticeable change in well-being?
-According to @normanInterpretationChangesHealthrelated2003, people can reliably distinguish seven levels of satisfaction with health.
-So if satisfaction is measured on a 7-point scale, a four unit change in social media use should result in a one unit change in life satisfaction. 
-<!-- Because the data were analyzed post-hoc, no a-priori sample size planning on the basis of power analyses was conducted. -->
-<!-- The sample is large, and it is hence well-equipped to reliably detect small effects. -->
-<!-- <!-- , which is why no exact post hoc power analysis were conducted. -->
-<!-- In addition, because such large samples easily generate significant _p_-values even for very small effects, it helps that the hypotheses were tested with a smallest effect size of interest-approach.
--->
-<!-- To this end, I adopted the interval testing approach as proposed by @dienesUsingBayesGet2014. -->
-<!-- On the basis of the SESOI, I defined a null region. -->
-<!-- In what follows, I explain how I determined the SESOI and the null region. -->
-
-In this study, life satisfaction was measured on an 11-point scale.
-If people can reliably differentiate 7 levels, this corresponds to 11 / 7 = `r round(11/7, 2)` unit change on an 11-point scale.
-Hence, a four-point change in media use (e.g., a complete stop) should result in a `r round(11/7, 2)`-point change in life satisfaction.
-In a statistical regression analysis, _b_ estimates the change in the dependent variable if the independent variable increases by one point.
-For life satisfaction, we would therefore define a SESOI of _b_ = `r round(11/7, 2)` / 4 = `r round(11/7*.25, 2)`.
-For positive or negative affect, which was measured on a 5-point scale, our SESOI would be _b_ = `r round(5 / 7, 2)` / 4 = `r round(5 / 7 * .25, 2)`.
-Because we are agnostic as to whether the effects are positive or negative, the null region includes both negative and positive effects.
-Finally, in order not to exaggerate precision and to be less conservative, these numbers are reduced to nearby thresholds.[^rounding] 
-Together, this leads to a null region ranging from _b_ = -.30 to _b_ = .30 for life satisfaction, and _b_ = -.15 to _b_ = .15 for positive and negative affect.
-
-[^rounding]: Note that other researchers also decreased or recommended decreasing thresholds for effect sizes when analyzing within-person or cumulative effects [@beyensSocialMediaUse2021; @funderEvaluatingEffectSize2019].
-
-<!-- As explained above, for well-being the null region was between _b_ = -.30 and _b_ = .30 (i.e., _b_ = |.30|). -->
 The hypothesis is analyzed using the interval testing approach as proposed by @dienesUsingBayesGet2014.
-To illustrate, let us consider the case of life satisfaction [SESOI: -.30 : +.30]. 
+To illustrate, let us consider the case of life satisfaction [SESOI: ±.30]. 
+Here, the null-region lies between -.30 and +.30.
 If the 95% confidence interval falls completely within the null-region (e.g., _b_ = -.05, [95% CI: -.15, .05]), the hypothesis that the effect is trivial is supported.
@@
 
-<!-- Responses were individually checked for patterns such as straight-lining or missing of inverted items. -->
-<!-- X clear cases were removed. -->
-
-## Data Analysis
 ### Causality
 
-<!-- The hypothesis explicitly states a causal effect. -->
-<!-- In non-experimental studies, longitudinal designs can help investigate causality. -->
-When using longitudinal designs to analyze causality, it is important to (a) focus on within-person effects [@hamakerWhyResearchersShould2014]; to (b) control for confounders [@rohrerTheseAreNot2023]; and to (c) test a plausible interval between measures [@dormannOptimalTimeLags2015].
-<!-- To illustrate, consider the following example. -->
-<!-- Imagine that a person suddenly starts using social media much more than usual, and then after some time  becomes less satisfied with their life. -->
-<!-- Eventually, use and life satisfaction return to prior levels. -->
-<!-- If this happens to several people at the same time, in a longitudinal study we could then observe a significant effect of social media use on life satisfaction. -->
-<!-- However, it could also be the case that during the study there was a major exogenous event (say, a pandemic), which caused large parts of the working population to loose their jobs. -->
-<!-- Hence, the causal effect reported above was confounded, because in reality it was the pandemic that caused both social media use to rise and life satisfaction to go down. -->
-<!-- Thus, only when controlling for _all_ relevant confounders, can we correctly estimate causality without bias [@rohrerThinkingClearlyCorrelations2018]. -->
-<!-- Obviously, we can never be entirely sure to have included all confounders, which makes absolute statements regarding causality virtually impossible. -->
-First, in non-experimental designs it makes much sense to analyze causal effects from an internal, within-person perspective [@hamakerWhyResearchersShould2014].
-If a specific person changes their media diet, we need to measure how this behavior affects their well-being.
-Between-person comparisons from longitudinal data cannot provide such insights [@hamakerWhyResearchersShould2014].
-To test the hypothesis, I thus consider only the within-person effects.
+When analyzing causality through longitudinal designs, it is important to address several critical aspects.
+An inherent requirement of analyzing causal effects within non-experimental designs is adopting an internal perspective, focusing on _within-person effects_ [@hamakerWhyResearchersShould2014].
+This entails evaluating how alterations in an individual's media consumption directly impact their well-being.
+Comparisons between individuals, as observed in longitudinal data, lack the granularity needed to understand these causal dynamics [@hamakerWhyResearchersShould2014].
+Consequently, this study exclusively investigates within-person effects to shed light on this intricate relationship.
 
-Second, to identify confounders we should control for variables that affect both media use and well-being, which helps isolate the actual effect [@rohrerThinkingClearlyCorrelations2018].
-Because we are adopting a within-person perspective, we need to implement _time-varying_ confounders [@rohrerTheseAreNot2023]. 
-And because we are determining the _overall_ causal effect, we need to make sure _not_ to control for mediating variables [@rohrerThinkingClearlyCorrelations2018], for doing so would bias our assessment of the causal effect.
-<!-- Complicating matters further, it is often unclear if a variable is a mediator or a confounder.[^collider] -->
-<!-- However, despite all these caveats, when controlling for relevant variables (that aren't mediators), we can be much more certain that we measured causality correctly. -->
-<!-- The aim should therefore be to collect as many varying and non-varying confounders as possible (which I believe is seldom done in our field), while knowing that absolute certainty regarding causality cannot be reached.
--->
-<!-- [^collider]: In addition, there also exist colliders, which I don't discuss here and which complicate the issue even further [@rohrerThinkingClearlyCorrelations2018]. -->
-<!-- We can also control for variables that affect only social media use or well-being.  -->
-<!-- However, in doing so not much is gained, because the effects of social media use would remain virtually the same [@klinePrinciplesPracticeStructural2016; but see @mcelreathYesterdayClass2021].
--->
-In this study, I hence preregistered to control for the following variables, which either have already been shown or are likely to affect both social media use and well-being, and which also are not mediators:
-<!-- (I'll additionally include variables that likely affect only well-being, also to obtain a comparison benchmark for social media effects): -->
-gender, age, education, Austria country of birth, Austria country of birth of parents, residency Vienna, text-based news consumption, video-based news consumption, household size, health, living space, access to garden, access to balcony, employment, work hours per week, being in home-office, household income, outdoor activities, disposition to take risks, and locus of control [@egerStatisticalMetaanalysisWellbeing2015].
-<!-- [^explore] -->
-<!-- I will not control for variables such as trust in institutions or trust in media, because these variables might be influenced by social media use to a meaningful extent. -->
+An essential strategy for isolating the genuine impact of variables is to _control for confounding factors_ that influence both media use and well-being [@rohrerThinkingClearlyCorrelations2018].
+In the context of within-person analysis, this necessitates accounting for time-varying confounders [@rohrerTheseAreNot2023].
+However, a cautious balance must be maintained by not controlling for variables that mediate the relationship [@rohrerThinkingClearlyCorrelations2018], as their inclusion could distort the assessment of the causal effect.
+This study therefore incorporated several control variables.
+These encompass gender, age, education, place of birth and parental place of birth in Austria, Vienna residency, consumption of text-based and video-based news, household characteristics, health status, living space features, employment-related factors, income, outdoor activities, risk propensity, and locus of control [@egerStatisticalMetaanalysisWellbeing2015].
+These variables have demonstrated connections with both social media use and well-being and are not mediators.
 
-<!-- [^explore]: The data-set includes many other variables that one could also potentially control for, and I invite interested readers to download the and explore potential interesting relationships.
--->
+A fundamental prerequisite for establishing causality is determining a _plausible temporal interval_ [@dormannOptimalTimeLags2015].
+For instance, fluctuations in positive and negative affect call for shorter intervals, while the more enduring nature of life satisfaction implies longer intervals [@dienlinImpactDigitalTechnology2020].
 
-<!-- Next to including covariates, it is now increasingly understood that  -->
-Finally, one precondition of causality is temporal order and finding a plausible interval [@dormannOptimalTimeLags2015].
-<!-- The cause needs to precede the effect. -->
-<!-- Finding the right interval between cause and effect is crucial. -->
-<!-- For example, if we want to understand the effect of alcohol consumption on driving performance, it makes a big difference if driving performance is measured one minute, one hour, one day, or one week after consumption.
--->
-<!-- Finding the right interval is difficult. -->
-If variables are stable, longer intervals are needed; if they fluctuate, shorter intervals.
-In the case of well-being, we need shorter intervals for the more fluctuating positive and negative affect, and longer ones for the more stable life satisfaction [@dienlinImpactDigitalTechnology2020].
-Whereas using social media can have instant effects on mood [@marcianoDynamicsAdolescentsSmartphone2022], effects on life satisfaction often take longer to manifest.
-For example, because media use leads to actual changes in specific behaviors, which then in turn affect life satisfaction [@dienlinDisplacementReinforcementReciprocity2017].
-<!-- Choosing the right interval is challenging, because especially short intervals are hard to implement in practice, often requiring advanced methods such as experience sampling (also known as in situ measurement or ambulant assessment) [@schnauber-stockmannMobileDevicesTools2020].
--->
+In this study, I examine the linkage between changes in social media use and changes in affect within the same week.
+Specifically, I investigate if heightened COVID-19-related social media use during a week corresponds with enhanced or diminished affect during that same week.
+Additionally, I consider a longer interval for life satisfaction, examining whether increased COVID-19-related social media use over the course of a week influences one's life satisfaction at the week's end.
+This approach aims to uncover whether changes in social media consumption prompt simultaneous changes in affect and subsequent changes in life satisfaction.
 
-In this study, I hence analyze how changes in using social media _during the last week_ affected changes in positive and negative affect _during the same week_.
-In other words, if people during the last week engaged in more COVID-19 related social media use than usual, did they feel better or worse during that week than usual?
-For life satisfaction, I implemented a longer interval.
-If people _during the last week_ used COVID-19 related social media more than they usually do, were they _at the end of the week_ more or less satisfied with their lives than they usually are?
-This way it is analyzed if when a person changes their social media diet, are there (a) _simultaneous_ changes in their affect and (b) _subsequent_ changes in their life satisfaction?
-For the main analyses, the interval is implemented via the wording of the items (see below), not by using lagged measures coming from prior waves waves. 
-In additional analyses, I also tested how media use affects well-being one month or four months later.
-<!-- This significantly increases sample size, such that all respondents who answered the media measures at least two times can be implemented. -->
-All analyses will be controlled for varying confounders (see below), which fosters a causal interpretation.
-<!-- Similar approaches were implemented by other studies [@johannesNoEffectDifferent2022; @scharkowHowSocialNetwork2020], and they are considered a best practice approach toward analyzing causality [@bellFixedRandomEffects2019].
--->
+Furthermore, the study's main analyses implement this temporal interval through item wording rather than relying on lagged measures from previous waves.
+Supplementary analyses extend this investigation to observe how media use might affect well-being one or four months later.
+In all instances, confounding variables are methodically controlled for, enhancing the study's capacity for causal interpretation.
 
@@
 Because Mardia’s test showed that the assumption of multivariate normality was violated, I used the more robust Satorra-Bentler scaled and mean-adjusted test statistic (MLM) as estimator.
-<!-- To avoid over-fitting, I tested the scales on more liberal fit criteria (CFI > .90, TLI > .90, RMSEA <. .10, SRMR < .10) [@klinePrinciplesPracticeStructural2016]. -->
-<!-- Finally, REWB-models cannot model latent variables. -->
-<!-- To increase precision, I therefore exported factor scores from the CFAs for positive and negative affect. -->
 Mean scores were used for positive and negative affect.
-<!-- Respondents who answered less than 50% of all questions were removed. -->
 Missing responses were imputed using multiple imputation with predictive mean matching (five iterations, 30 data-sets), including categorical variables.
@@
 For more information on the analyses, a complete documentation of the models and results, and all additional analyses, see [companion website](https://XMtRA.github.io/Austrian_Corona_Panel_Project).
-<!-- The additional analyses are reported on the [companion website](https://XMtRA.github.io/Austrian_Corona_Panel_Project/analyses_additional.html). -->
-<!-- For example, all within-person standardized COVID-19 related types of social media use or channels were significantly smaller than $\beta$ = |.05|, again supporting that effects were negligible.
--->
 
@@
 
-<!-- In what follows, I list all the variables that I analyzed. -->
-For the variables' means, range, and variance, see Table \@ref(tab:tab-descriptives).
+<!-- For the variables' means, range, and variance, see Table \@ref(tab:tab-descriptives). -->
 For a complete list of all items and item characteristics, see [companion website](https://XMtRA.github.io/Austrian_Corona_Panel_Project).
@@
 
-```{r tab-descriptives, cache=F, warning=F, message=F, prompt=F}
-apa_table(tab_desc_dat,
-          stub_indents = list(
-            "Well-being" = c(1:3),
-            "Social media use" = c(4:6),
-            "Social media channel" = c(7:11)),
-          caption = "Descriptives of the main variables."
-          , font_size = "small"
-          )
-```
-
 ### Control variables
@@
 gender (female, male, diverse), age, education (ten options), Austria country of birth (yes/no), Austria parents' country of birth (no parent, one parent, both parents), and household size.
-<!-- I originally planned to implement additional variables as varying covariates. -->
 I also controlled for the following varying covariates: five items on current living conditions, including self-reported physical health, whether participants contracted COVID-19 since the last wave, current household income, working in home-office, and overall work hours; nine items measuring use of specific national text-based and video-based news outlets; five items measuring outdoor activities such as exercise or meeting friends; and two more psychological measures including locus of control and disposition to take risks.
@@
 ## Preregistered Analyses
-### Social media communication types
 
 The study's main hypothesis was that the causal effects of all types and channels of social media use on all facets of well-being would be trivial.
-Regarding the effects of different communication types (i.e., reading, sharing, of posting about COVID-19 related content), all within-person effects fell completely within the a-priori defined null region (see Figure \@ref(fig:fig-within)).
+Regarding the effects of different _communication types_ (i.e., reading, sharing, of posting about COVID-19 related content), all within-person effects fell completely within the a-priori defined null region (see Figure \@ref(fig:fig-within)).
 For example, respondents who used social media more frequently than usual to like or share COVID-19 related content did not show a simultaneous change in life satisfaction (_b_ = `r data_tab_within %>% filter(dv == "Life satisfaction" & iv == "Liking & Sharing" & Variance == "within" & Analysis == "Publication") %>% select(estimate) %>% round(2)` [95% CI `r data_tab_within %>% filter(dv == "Life satisfaction" & iv == "Liking & Sharing" & Variance == "within" & Analysis == "Publication") %>% select(conf.low) %>% round(2)`, `r data_tab_within %>% filter(dv == "Life satisfaction" & iv == "Liking & Sharing" & Variance == "within" & Analysis == "Publication") %>% select(conf.high) %>% round(2)`]).
@@
 
-### Social media communication channels
-
-Regarding the COVID-19 related use of social media channels (i.e., Facebook, Instagram, WhatsApp, YouTube, and Twitter) the results were comparable (see Figure \@ref(fig:fig-within)).
+Regarding the COVID-19 related use of _social media channels_ (i.e., Facebook, Instagram, WhatsApp, YouTube, and Twitter) the results were comparable (see Figure \@ref(fig:fig-within)).
 Changes in the frequency of using different social media channels to attain information regarding COVID-19 were unrelated to meaningful changes in well-being.
@@
 
-For an overview of all within-person effects, see Table \@ref(tab:tab-within) and Figure \@ref(fig:fig-within).
+For an overview of all within-person effects, see Figure \@ref(fig:fig-within).
 
@@
 
-```{r tab-within, cache=F, warning=F, message=F, prompt=F}
-apa_table(tab_within %>% select(-Outcome),
-          align = c("l", "r", "r", "r", "r", "r"),
-          stub_indents = list(
-            "Life satisfaction" = c(1:8),
-            "Positive affect" = c(9:16),
-            "Negative affect" = c(17:24)),
-          caption = "Overview of all within-person effects.",
-          col_spanners = list(`Confidence interval` = c(3, 4))
-          , font_size = "footnotesize"
-          )
-```
-
 ## Exploratory Analyses
-
-<!-- In what follows, I briefly report some exploratory analyses that weren't preregistered. -->
 
@@
 Spending more time outside to exercise meaningfully increased positive affect ($\beta$ `r data_tab_comp_std %>% filter(dv == "Positive affect" & iv == "Sports") %>% select(estimate) %>% my_round("std_txt")` [95% CI `r data_tab_comp_std %>% filter(dv == "Positive affect" & iv == "Sports") %>% select(conf.low) %>% my_round("std")`, `r data_tab_comp_std %>% filter(dv == "Positive affect" & iv == "Sports") %>% select(conf.high) %>% my_round("std")`]).
-<!-- Being more satisfied with the current democratic system meaningfully increased life satisfaction ($\beta$ `r data_tab_comp_std %>% filter(dv == "Life satisfaction" & iv == "Satisfaction democracy") %>% select(estimate) %>% my_round("std_txt")` [95% CI `r data_tab_comp_std %>% filter(dv == "Life satisfaction" & iv == "Satisfaction democracy") %>% select(conf.low) %>% my_round("std")`, `r data_tab_comp_std %>% filter(dv == "Life satisfaction" & iv == "Satisfaction democracy") %>% select(conf.high) %>% my_round("std")`]).
--->
 The strongest aspect affecting well-being was internal locus of control.
@@
 Compared to the middle age category Generation X, results showed that if users from Generation Z posted more COVID-19 content than usual this lead to significantly more negative affect ($\beta$ `r fit_aff_neg_lmer_age_tab %>% filter(term == "age_genGen Z:soc_med_post_w") %>% select(estimate) %>% my_round("std_txt")` [95% CI `r fit_aff_neg_lmer_age_tab %>% filter(term == "age_genGen Z:soc_med_post_w") %>% select("2.5 %") %>% my_round("std")`, `r fit_aff_neg_lmer_age_tab %>% filter(term == "age_genGen Z:soc_med_post_w") %>% select("97.5 %") %>% my_round("std")`]).
-<!-- Although this might just be a fluke finding, the low p-value suggests that it might actually be real. -->
 
@@
 
-Based on a panel study with 34 waves largely representative of the Austrian population, this study analyzed the effects of COVID-19 related social media use on well-being.
-Between person correlation analyses showed that more active users of COVID-19 related content on social media also reported decreased well-being. 
-For example, respondents who read more COVID-19 related content than others reported slightly lower levels of life satisfaction, somewhat lower levels of positive affect, and substantially higher levels of negative affect than others.
-To see if these between person correlations would translate to within-person effects, I analyzed if changes in a person's media use led to changes in their well-being.
-The within-person relations showed a different pattern.
-If people consumed more COVID-19 content on social media than usual, this did not meaningfully reduce their well-being.
-Although several statistically significant effects were found, these were very small.
-For example, people who read more COVID-19 related posts than usual reported slightly decreased positive affect.
-People who liked and shared more COVID-19 related posts than usual reported slightly higher levels of negative affect.
-Posting more content about COVID-19 than usual slightly decreased life satisfaction, while increasing both negative affect and positive affect.
-Using Twitter for COVID-19 related content slightly increased negative affect, as did YouTube.
-<!-- was related to Reading COVID-19 related content on social media slightly reduced positive affect. -->
-Again, although all of these within-person effects were statistically significant, they were very small, smaller than the predefined smallest effect size of interest.
-According to the preregistered procedure, they should hence be considered irrelevant. 
-Additional analyses revealed that other factors, for which we would expect to find meaningful effects, such as health or sports, indeed showed substantial and meaningful impacts on well-being.
-In addition, when testing for the longer intervals of one month and four months, again no meaningful effects were found.
-In conclusion, COVID-19 related activity on social media was not a particularly strong influence on peoples' well-being. 
-The results do not support the popular fears that "doomscrolling" or overusing social media during times of crises constitutes a prominent risk for well-being.
+This study, based on a representative panel study spanning 34 waves within the Austrian population, investigated the impact of COVID-19-related social media usage on well-being.
+The between-person correlations revealed that increased engagement with COVID-19 content on social media was associated with decreased well-being.
+For instance, individuals consuming more COVID-19-related content reported slightly lower life satisfaction, somewhat reduced positive affect, and notably elevated negative affect compared to others.
 
-<!-- People who used social media more than usual to learn about COVID-19 didn't show meaningful changes in their well-being. -->
+To explore if these between-person correlations translated into within-person effects, it was examined whether changes in media consumption corresponded with changes in well-being.
+Contrary to expectations, increased consumption of COVID-19 content did not significantly decrease well-being at a meaningful level.
+While several statistically significant effects emerged, their magnitudes were notably small.
+For instance, reading more COVID-19 posts than usual slightly decreased positive affect.
+Liking and sharing more COVID-19 content than usual were associated with slightly higher negative affect.
+Posting more COVID-19 content reduced life satisfaction slightly while elevating both positive and negative affect.
+The effects, although statistically significant, fell within a predefined range considered insignificant.
 
-<!-- The results imply that COVID-19 related social media use doesn't seem to be particularly relevant for well-being. -->
-<!-- Other factors among the third variables that were measured revealed larger effects or relations, suggesting that well-being is rather determined by alternative aspects such as health, satisfaction with democracy, locus of control, or exercising.
--->
+Further analysis demonstrated that factors such as health and physical activity, which would be expected to have substantial impacts on well-being, indeed showed significant influences.
+Additionally, extended assessments covering one and four months did not yield meaningful effects.
+In summary, the influence of COVID-19-related social media activity on well-being was not substantial.
+This counters popular concerns over excessive social media use during crises causing substantial well-being risks.
 
-These specific observations notwithstanding, several general trends can be observed.
-First, overall the results do suggest that effects of COVID-19 related social media use on well-being tend to take place in the negative as opposed to the positive spectrum.
-Although very small, five statistically significant negative results of COVID-19 related social media use on well-being were found.
-Only one positive effect emerged.
-<!-- Also note that in the analyses several control variables were included, ruling out plausible alternative explanations for the negative results. -->
-<!-- For example, it was controlled for as to whether or not participants contracted a COVID-19 infection during a specific wave. -->
-<!-- Hence, we can rule out the alternative explanation that having an infection was the root cause of increased communication and reduced well-being. -->
+There is no consensus among scholars as to when effects become practically relevant and meaningful.
+If we adopt a more liberal and in cautious perspective, some general trends emerged.
+The study indicates a tendency for COVID-19-related social media usage to impact well-being negatively more frequently than positively.
+Notably several small yet statistically significant negative effects were observed, contrasting with a single positive effect.
+Furthermore, several significant associations were found between social media use and affect but only one with life satisfaction, a more stable measure.
 
-Second, six significant outcomes emerged for positive or negative affect, but only one for life satisfaction. 
-Life satisfaction is more stable and not that easily affected by any type or channel of social media communication.
-The more fluctuating positive and negative affect, however, were affected (albeit only slightly).
-Liking, sharing, and posting COVID-19 related content, and spending more time on Twitter and YouTube to browse COVID-19 related content, all slightly negatively influenced affect. 
-This is aligned with prior findings which showed that social media use is associated with increased negative affect but not with life satisfaction [@meierComputermediatedCommunicationSocial2020a].
-<!-- Writing posts implies more substantial engagement, and perhaps also negative exchanges with other users, which could explain the results. -->
-<!-- For example, people who wrote more COVID-19 related posts than usual reported slightly lower levels of life satisfaction than usual. -->
-<!-- Similarly, people who wrote more COVID-19 related posts than usual also reported slightly more negative affect. -->
-<!-- As a potential explanation, when writing posts and comments on social media people explicitly and more deeply engage with COVID-19 related content. -->
-Conversations about COVID-19 on social media are often extreme, negative, or aggressive [@fanStigmatizationSocialMedia2020].
-More deeply engaging with this type of content could negatively affect active authors.
-The hypothesis that tonality could explain the negative effects is especially supported by the observation that spending more time on Twitter and YouTube than usual increased negative affect.
-Communication on both channels is often found to be negative and impolite [e.g., @muellerTwitterMadeMe2022].
-Consuming more negative and misleading information could hence explain the (slightly) increased levels of negative affect.
+These results align with prior research highlighting that social media usage is associated with elevated negative affect but not reduced life satisfaction [meierComputermediatedCommunicationSocial2020a].
+The often extreme, negative, or aggressive tone of COVID-19 discussions on social media [@fanStigmatizationSocialMedia2020] could adversely impact active authors.
+The negative effects may be explained by the consumption of negative and misleading information.
 
-Third, the results show that it makes sense to analyze different communication types and communication channels.
-Reading slightly reduced positive affect, while liking, sharing, and posting slightly increased negative affect.
-Interestingly, posting COVID-19 related comment slightly increased negative affect, while at the same time it also slightly increased positive affect.
-Posting content is often met with strong reaction, both positive by means of likes and negative by means of critical comments. 
-Overall, though, posting led to slightly reduced levels of life satisfaction.
-In conclusion, whereas it was often stated that passive use is bad and active use good [@verduynPassiveFacebookUsage2015], this pattern was only partially found here.
-<!-- Instead, all three types were negative. -->
-The results are aligned with the findings from @valkenburgAssociationsActivePassive2022, who could not confirm that active use is good and that passive use is bad.
-Focusing on communication channels, Twitter and YouTube seem to be more negative, while Instagram, WhatsApp, and Facebook were neutral.
-But, again, all of these effects are very small.
-<!-- Future research might elaborate on these specific relations to probe their stability and relevance. -->
+Moreover, the study suggest that varying communication types and channels warrant separate analyses.
+For instance, reading content slightly reduced positive affect, while liking, sharing, and posting increased negative affect slightly.
+Posting COVID-19-related comments increased both negative and positive affect slightly, while reducing life satisfaction.
+This suggests that posting generates stronger reactions, both positive and negative.
+Together, this challenges the assumption that active use is inherently beneficial and passive use detrimental [@valkenburgAssociationsActivePassive2022].
+Notably, communication channels exhibited differences.
+Twitter and YouTube appeared more negative, whereas Instagram, WhatsApp, and Facebook maintained neutrality.
+Despite these nuanced findings, the effects remained small.
 
-Taken together, the results are hence aligned with the underlying theoretical models and prior empirical results.
-The findings support the differential susceptibility of media effects model [@valkenburgDifferentialSusceptibilityMedia2013], such that effects are generally small and that they depend on the type and channel of communication.
-Additional analyses did not reveal that effects depended on gender.
-Age also large did not play a significant moderation role, but effects of posting COVID-19 related content were found to be more negative for Generation Z.
-Indeed, it has often been argued that effects of social media use are more negative for Gen Z than for prior generations, and this finding can be seen a further tentative support for this hypothesis. 
-From a broader perspective, the results are well-aligned with mood management theory [@zillmannMoodManagementCommunication1988] and the uses and gratifications approach [@katzUsesGratificationsResearch1973], whose premises preclude particularly negative effects of routine and widespread media consumption.
-Both theories posit that if the effects of social media were indeed profoundly negative on average, then people likely would not spend so much time on social media engaging with COVID-19 content.
-Finally, recent empirical studies and meta-analyses reported rather small negative effects, too [@meierComputermediatedCommunicationSocial2020a], 
-<!-- Several studies found that the effects of various types of social media use on well-being are small, often too small to matter [@meierComputermediatedCommunicationSocial2020a],  -->
-echoing the results obtained here.
+In conclusion, the study aligns with theoretical models and past research.
+It supports the notion of small media effects influenced by communication type and channel [@meierComputermediatedCommunicationSocial2020a].
+The differential susceptibility model [@valkenburgDifferentialSusceptibilityMedia2013] was supported, indicating that effects are small and dependent on communication factors.
+Age did not significantly moderate the effects, yet posting COVID-19 content was more negative for Generation Z, potentially reflecting broader negative effects of social media on this generation.
+The study also echoes mood management theory [@zillmannMoodManagementCommunication1988] and the uses and gratifications approach [@katzUsesGratificationsResearch1973], implying that if the effects were highly negative, people would not engage with COVID-19 content as extensively.
 
-<!-- The results allow us to reject a positive effect: Writing more posts on social media will likely not increase well-being. -->
+Despite these insights, limitations exist.
+While the study's focus on within-person effects enhances causal understanding, challenges related to additional confounding variables [@rohrerThinkingClearlyCorrelations2018], best definition of the SESOI [@anvariUsingAnchorbasedMethods2021a], and exact measurement of media use [@scharkowAccuracySelfreportedInternet2016].
+The study's results are applicable primarily to Western societies and may not hold true in different cultural contexts.
 
-<!-- First, in terms of media channels, using Twitter more than usual was related to slightly decreased levels of life satisfaction. -->
-<!-- Twitter is considered to have more negative affordances and tonality as compared to other networks such as Instagram [@waterlooNormsOnlineExpressions2018], which might help explain the results.
--->
-<!-- Second, of all the three COVID-19 related social media activities, people who read about the pandemic more than others showed slightly decreased levels of positive affect, and people who actively posted about the pandemic more than others showed slightly increased levels of negative affect.
--->
-<!-- On the other hand, however, people who posted more about COVID-19 also showed slightly higher levels of positive affects, so taken together the results are ambivalent. -->
-
-<!-- Finally, another interesting observation is that life satisfaction was remarkably stable. -->
-<!-- Hence, even in times of a pandemic, it seems that such broad assessment of life vary only mildly. -->
-<!-- This supports the hypothesis that life satisfaction seems to be determined largely by stable factors such as one's genes [@brownEasyHappinessPie2019]. -->
-
-## Limitations
-
-<!-- The current study analyzed whether changes in media use were related to changes in well-being. -->
-Focusing on within-person effects and controlling for several potential confounders, this study provides an improved perspective on assessing causality.
-However, several challenges remain. 
-<!-- First, it could be that the timing of the interval was wrong; it could be that effects unfold in shorter or longer intervals.  -->
-<!-- Next, it could be that there was not sufficient variance in media use and well-being.  -->
-<!-- Without sufficient variability, statistical models cannot detect existing effects. -->
-<!-- Third,  -->
-In order to correctly establish causality in non-experimental designs, it is necessary to control for _all_ relevant confounding third variables [@rohrerThinkingClearlyCorrelations2018]. 
-Although this study included are large list of confounders, it could still be that crucial variables were missed.
-<!-- [@zillmannMoodManagementCommunication1988]. -->
-<!-- While controlling for potential confounders can support claims of causality, the procedure implemented here cannot prove causality. -->
-<!-- Causality necessitates temporal order, and the cause needs to precede the effect. -->
-<!-- The challenge is that regarding media use, such effects often happen immediately or shortly after use, necessitating intervals in the hours, minutes, or even seconds. -->
-<!-- In many cases only experience sampling studies asking users at the very moment can produce such knowledge. -->
-<!-- However, even then we don't know for certain if we actually measured the right interval. -->
-<!-- Effects depend on the intensity of use or the length of the interval. -->
-<!-- To borrow the words from @rohrerTheseAreNot2021, there is no such thing as "the" effect of social media use on well-being. -->
-<!-- The results reported here are hence contingent on the design I used. -->
-<!-- To document how effects unfold, future research needs to employ different study designs probing different intervals. -->
-More thought needs to be invested in which factors to control for and, equally important, for which factors not to control for.
-<!-- I hope this study provides a first step into this direction. -->
-
-Although I had already reduced the predefined SESOIs to be less conservative, one could argue they were still too large.
-Media use is only one aspect of several factors that simultaneously affect well-being.
-Is it realistic to expect that changing only _one_ of these aspects should already manifest in a detectable change in well-being?
-Or would it make more sense to expect that thoroughly committing to say _two_ activities (e.g.
-regularly exercising _and_ establishing a reading habit) should then cause a detectable improvement in well-being?
-Practically, this would imply a SESOI half the size defined here, namely _b_ = |.15| for life satisfaction and _b_ = |.075| for affect.
-In the case of this study, however, even halving the SESOI would not make a difference.
-All but one effect would still be completely in the null region, and no effect would fall completely outside of the null region.
-I encourage future research to elaborate on what effect sizes are considered meaningful and what not.
-<!-- Again, with this study I hope to provide some first input and guidelines. -->
-
-Both media use and well-being were measured using self-reports.
-Because assessing well-being necessarily requires introspection, using self-reports for affect and life satisfaction is adequate.
-However, for social media use objective measures are preferable, as people often cannot reliably estimate their use [@scharkowAccuracySelfreportedInternet2016].
-<!-- At the same time, most objective measures cannot capture the content or the motivation of use. -->
-<!-- Hence, for the type of research question analyzed here, it still seems necessary to use self-reported measures. -->
-<!-- In many cases they can still be informative [@verbeijSelfreportedMeasuresSocial2021]. -->
-
-Being collected in a single country, the generalizability of the results is limited.
-<!-- The results apply primarily to the more Western sphere. -->
-They might not hold true in other cultures, especially non-Western cultures with a different media landscape or alternative social media channels.
-<!-- That said, being a comparatively large study largely representative of one country, with several waves collected across a large time span, the results should be at least as generalizable as other typical empirical studies collected in the social sciences.
--->
-<!-- Social media use was measured with an ordinal variable, however in the analyses it was treated as a numerical one. -->
-<!-- If treated as an ordinal one, it would have been necessary to analyze four different contrasts for each media measure, which plus the differentiation between between and within factor would have produced eight different measures, we would have made the model exceedingly complex.
--->
-
-## Conclusion
-
-<!-- That said, the results still imply that it does make sense to critically reflect upon COVID-19 related social media use.  -->
-<!-- On average, it seems more beneficial to post less actively about COVID-19 on social media and to spend less time on Twitter and YouTube consuming COVID-19 related content. -->
-
-In this study, COVID-19 related social media use did not meaningfully affect well-being. 
-Very small negative effects were found for writing COVID-19 related posts, sharing COVID-19 related content, and spending more time than usual on Twitter.
-Factors other than social media use, however, were meaningfully related to well-being, including physical health, exercise, satisfaction with democracy, or believing that one is in control of one's life.
-In light of the overall very small effects, engaging in COVID 19-related social media use should not be considered a major concern for one's well-being.
-<!-- Hence, when trying to improve well-being during a pandemic, instead of focusing on social media it seems more fruitful to address other, more pertinent societal problems related to health care, regular exercise, or psychological resilience.
--->
+The study's findings conclude that COVID-19-related social media activity minimally affects well-being, with other factors playing more substantial roles.
+In light of these minimal effects, concerns over COVID-19-related social media engagement on well-being appear to be unwarranted.
 

### 2023-08-24 manuscript.tex
- shortened manuscript for submission

@@
 \keywords{COVID-19, well-being, social media, news use, panel study.}
-\usepackage{lineno}
-
-\linenumbers
 \usepackage{csquotes}
@@
 The rapid spread of news, updates, and opinions can be overwhelming and might exacerbate existing stress or fears about the pandemic (Sharma et al., 2022).
---\textgreater{}
 Social media platforms are known for fostering negativity, with users sometimes engaging in \emph{cyberbullying and harassment}.
@@
 
+\hypertarget{smallest-effect-size-of-interest}{%
+\subsection{Smallest Effect Size of Interest}\label{smallest-effect-size-of-interest}}
+
+Exploring the hypothesis entails establishing the criteria for discerning a `trivial effect size.'
+To achieve this, it is necessary to define the smallest effect size of interest (SESOI) (Anvari \& Lakens, 2021).
+In this context, a trivial effect should fall below the threshold set by the SESOI (detailed below).
+Determining what constitutes a minimally intriguing, nontrivial effect is a matter with normative implications, making it challenging to arrive at a definitive, singular consensus.
+Nonetheless, it remains both essential and beneficial to strive for a reasonable benchmark.
+With that in mind, I propose the following SESOI as a suitable reference point for this study:
+
+\begin{quote}
+SESOI: If a heavy user of COVID-19 related social media news suddenly \emph{stops} using social media altogether, this should have a \emph{noticeable} impact on their overall well-being.
+\end{quote}
+
+What does this mean practically and how can it be operationalized?
+In this study, COVID-19 related social media use was measured on a 5-point scale, ranging from 1 = \emph{never} to 5 = \emph{several times a day}.
+Thus, a change of four units in social media use (e.g., a complete stop) should correspond to a noticeable change in well-being.
+According to Norman et al. (2003), people can reliably notice seven levels of change in satisfaction with health.
+So if satisfaction is measured on a 7-point scale, a four unit change in social media use should result in a one unit change in life satisfaction.
+In this study, life satisfaction was measured on an 11-point scale, and affect on a 5-point scale.
+Transposed to this scaling, the SESOI for life satisfaction is \emph{b} = ±.30, and for positive and negative affect \emph{b} = ±.15 (for more information, see online supplementary material).
+
 \hypertarget{method}{%
@@
 
-\hypertarget{smallest-effect-size-of-interest}{%
-\subsection{Smallest Effect Size of Interest}\label{smallest-effect-size-of-interest}}
-
-Testing the hypothesis necessitates defining what is considered a ``trivial effect size''.
-To this end, we need to define a so-called smallest effect size of interest (SESOI) (Lakens et al., 2018).
-A trivial effect would then need to be smaller than the SESOI (see below).
-What could be a minimally interesting, nontrivial effect?
-Being a normative question, finding a clear, single, or unanimous answer is impossible.
-However, it is still necessary and helpful to work toward a plausible benchmark.
-I suggest the following SESOI for this research question:
-
-\begin{quote}
-SESOI: If a heavy user of COVID-19 related social media news suddenly \emph{stops} using social media altogether, this should have a \emph{noticeable} impact on their overall well-being.
-\end{quote}
-
-What does this mean practically and how can it be operationalized?
-In this study, COVID-19 related social media use was measured on a 5-point scale, ranging from 1 = \emph{never} to 5 = \emph{several times a day}.
-Thus, a change of four units in social media use (e.g., a complete stop) should correspond to a noticeable change in well-being.
-What is a noticeable change in well-being?
-According to Norman et al. (2003), people can reliably distinguish seven levels of satisfaction with health.
-So if satisfaction is measured on a 7-point scale, a four unit change in social media use should result in a one unit change in life satisfaction.
-
-In this study, life satisfaction was measured on an 11-point scale.
-If people can reliably differentiate 7 levels, this corresponds to 11 / 7 = 1.57 unit change on an 11-point scale.
-Hence, a four-point change in media use (e.g., a complete stop) should result in a 1.57-point change in life satisfaction.
-In a statistical regression analysis, \emph{b} estimates the change in the dependent variable if the independent variable increases by one point.
-For life satisfaction, we would therefore define a SESOI of \emph{b} = 1.57 / 4 = 0.39.
-For positive or negative affect, which was measured on a 5-point scale, our SESOI would be \emph{b} = 0.71 / 4 = 0.18.
-Because we are agnostic as to whether the effects are positive or negative, the null region includes both negative and positive effects.
-Finally, in order not to exaggerate precision and to be less conservative, these numbers are reduced to nearby thresholds.\footnote{Note that other researchers also decreased or recommended decreasing thresholds for effect sizes when analyzing within-person or cumulative effects (Beyens et al., 2021; Funder \& Ozer, 2019).}
-Together, this leads to a null region ranging from \emph{b} = -.30 to \emph{b} = .30 for life satisfaction, and \emph{b} = -.15 to \emph{b} = .15 for positive and negative affect.
+\hypertarget{data-analysis}{%
+\subsection{Data Analysis}\label{data-analysis}}
 
 The hypothesis is analyzed using the interval testing approach as proposed by Dienes (2014).
-To illustrate, let us consider the case of life satisfaction {[}SESOI: -.30 : +.30{]}.
+To illustrate, let us consider the case of life satisfaction {[}SESOI: ±.30{]}.
+Here, the null-region lies between -.30 and +.30.
 If the 95\% confidence interval falls completely within the null-region (e.g., \emph{b} = -.05, {[}95\% CI: -.15, .05{]}), the hypothesis that the effect is trivial is supported.
@@
 
-\hypertarget{data-analysis}{%
-\subsection{Data Analysis}\label{data-analysis}}
-
 \hypertarget{causality}{%
@@
 
-When using longitudinal designs to analyze causality, it is important to (a) focus on within-person effects (Hamaker, 2014); to (b) control for confounders (Rohrer \& Murayama, 2023); and to (c) test a plausible interval between measures (Dormann \& Griffin, 2015).
-First, in non-experimental designs it makes much sense to analyze causal effects from an internal, within-person perspective (Hamaker, 2014).
-If a specific person changes their media diet, we need to measure how this behavior affects their well-being.
-Between-person comparisons from longitudinal data cannot provide such insights (Hamaker, 2014).
-To test the hypothesis, I thus consider only the within-person effects.
-
-Second, to identify confounders we should control for variables that affect both media use and well-being, which helps isolate the actual effect (Rohrer, 2018).
-Because we are adopting a within-person perspective, we need to implement \emph{time-varying} confounders (Rohrer \& Murayama, 2023).
-And because we are determining the \emph{overall} causal effect, we need to make sure \emph{not} to control for mediating variables (Rohrer, 2018), for doing so would bias our assessment of the causal effect.
-In this study, I hence preregistered to control for the following variables, which either have already been shown or are likely to affect both social media use and well-being, and which also are not mediators:
-gender, age, education, Austria country of birth, Austria country of birth of parents, residency Vienna, text-based news consumption, video-based news consumption, household size, health, living space, access to garden, access to balcony, employment, work hours per week, being in home-office, household income, outdoor activities, disposition to take risks, and locus of control (Eger \& Maridal, 2015).
-
-Finally, one precondition of causality is temporal order and finding a plausible interval (Dormann \& Griffin, 2015).
-If variables are stable, longer intervals are needed; if they fluctuate, shorter intervals.
-In the case of well-being, we need shorter intervals for the more fluctuating positive and negative affect, and longer ones for the more stable life satisfaction (Dienlin \& Johannes, 2020).
-Whereas using social media can have instant effects on mood (Marciano et al., 2022), effects on life satisfaction often take longer to manifest.
-For example, because media use leads to actual changes in specific behaviors, which then in turn affect life satisfaction (Dienlin et al., 2017).
-
-In this study, I hence analyze how changes in using social media \emph{during the last week} affected changes in positive and negative affect \emph{during the same week}.
-In other words, if people during the last week engaged in more COVID-19 related social media use than usual, did they feel better or worse during that week than usual?
-For life satisfaction, I implemented a longer interval.
-If people \emph{during the last week} used COVID-19 related social media more than they usually do, were they \emph{at the end of the week} more or less satisfied with their lives than they usually are?
-This way it is analyzed if when a person changes their social media diet, are there (a) \emph{simultaneous} changes in their affect and (b) \emph{subsequent} changes in their life satisfaction?
-For the main analyses, the interval is implemented via the wording of the items (see below), not by using lagged measures coming from prior waves waves.
-In additional analyses, I also tested how media use affects well-being one month or four months later.
-All analyses will be controlled for varying confounders (see below), which fosters a causal interpretation.
+When analyzing causality through longitudinal designs, it is important to address several critical aspects.
+An inherent requirement of analyzing causal effects within non-experimental designs is adopting an internal perspective, focusing on \emph{within-person effects} (Hamaker, 2014).
+This entails evaluating how alterations in an individual's media consumption directly impact their well-being.
+Comparisons between individuals, as observed in longitudinal data, lack the granularity needed to understand these causal dynamics (Hamaker, 2014).
+Consequently, this study exclusively investigates within-person effects to shed light on this intricate relationship.
+
+An essential strategy for isolating the genuine impact of variables is to \emph{control for confounding factors} that influence both media use and well-being (Rohrer, 2018).
+In the context of within-person analysis, this necessitates accounting for time-varying confounders (Rohrer \& Murayama, 2023).
+However, a cautious balance must be maintained by not controlling for variables that mediate the relationship (Rohrer, 2018), as their inclusion could distort the assessment of the causal effect.
+This study therefore incorporated several control variables.
+These encompass gender, age, education, place of birth and parental place of birth in Austria, Vienna residency, consumption of text-based and video-based news, household characteristics, health status, living space features, employment-related factors, income, outdoor activities, risk propensity, and locus of control (Eger \& Maridal, 2015).
+These variables have demonstrated connections with both social media use and well-being and are not mediators.
+
+A fundamental prerequisite for establishing causality is determining a \emph{plausible temporal interval} (Dormann \& Griffin, 2015).
+For instance, fluctuations in positive and negative affect call for shorter intervals, while the more enduring nature of life satisfaction implies longer intervals (Dienlin \& Johannes, 2020).
+
+In this study, I examine the linkage between changes in social media use and changes in affect within the same week.
+Specifically, I investigate if heightened COVID-19-related social media use during a week corresponds with enhanced or diminished affect during that same week.
+Additionally, I consider a longer interval for life satisfaction, examining whether increased COVID-19-related social media use over the course of a week influences one's life satisfaction at the week's end.
+This approach aims to uncover whether changes in social media consumption prompt simultaneous changes in affect and subsequent changes in life satisfaction.
+
+Furthermore, the study's main analyses implement this temporal interval through item wording rather than relying on lagged measures from previous waves.
+Supplementary analyses extend this investigation to observe how media use might affect well-being one or four months later.
+In all instances, confounding variables are methodically controlled for, enhancing the study's capacity for causal interpretation.
 
@@
 
-For the variables' means, range, and variance, see Table \ref{tab:tab-descriptives}.
 For a complete list of all items and item characteristics, see \href{https://XMtRA.github.io/Austrian_Corona_Panel_Project}{companion website}.
@@
 For these analyzes I used the predictors from waves 1, 2, 8, 17, 23, and 28, to see if they predicted changes in well-being either one month or four months later.
-
-\begin{table}[tbp]
-
-\begin{center}
-\begin{threeparttable}
-
-\caption{\label{tab:tab-descriptives}Descriptives of the main variables.}
-
-\small{
-
-\begin{tabular}{lllll}
-\toprule
- & \multicolumn{1}{c}{sd} & \multicolumn{1}{c}{min} & \multicolumn{1}{c}{max} & \multicolumn{1}{c}{mean}\\
-\midrule
-Well-being &  &  &  & \\
-\ \ \ Life satisfaction & 2.23 & 6.32 & 6.60 & 6.49\\
-\ \ \ Positive affect & 0.94 & 3.09 & 3.22 & 3.16\\
-\ \ \ Negative affect & 0.77 & 1.75 & 1.86 & 1.81\\
-Social media use &  &  &  & \\
-\ \ \ Read & 1.38 & 1.92 & 2.88 & 2.35\\
-\ \ \ Like \& share & 1.19 & 1.54 & 1.94 & 1.74\\
-\ \ \ Posting & 0.89 & 1.36 & 1.42 & 1.40\\
-Social media channel &  &  &  & \\
-\ \ \ Facebook & 1.58 & 2.02 & 2.72 & 2.37\\
-\ \ \ Twitter & 0.95 & 1.36 & 1.43 & 1.39\\
-\ \ \ Instagram & 1.34 & 2.00 & 2.08 & 2.05\\
-\ \ \ WhatsApp & 1.66 & 2.27 & 2.60 & 2.46\\
-\ \ \ YouTube & 1.28 & 1.91 & 1.98 & 1.95\\
-\bottomrule
-\end{tabular}
-
-}
-
-\end{threeparttable}
-\end{center}
-
-\end{table}
 
@@
 
-\hypertarget{social-media-communication-types}{%
-\subsubsection{Social media communication types}\label{social-media-communication-types}}
-
 The study's main hypothesis was that the causal effects of all types and channels of social media use on all facets of well-being would be trivial.
-Regarding the effects of different communication types (i.e., reading, sharing, of posting about COVID-19 related content), all within-person effects fell completely within the a-priori defined null region (see Figure \ref{fig:fig-within}).
+Regarding the effects of different \emph{communication types} (i.e., reading, sharing, of posting about COVID-19 related content), all within-person effects fell completely within the a-priori defined null region (see Figure \ref{fig:fig-within}).
 For example, respondents who used social media more frequently than usual to like or share COVID-19 related content did not show a simultaneous change in life satisfaction (\emph{b} = -0.02 {[}95\% CI -0.06, 0.01{]}).
@@
 
-\hypertarget{social-media-communication-channels}{%
-\subsubsection{Social media communication channels}\label{social-media-communication-channels}}
-
-Regarding the COVID-19 related use of social media channels (i.e., Facebook, Instagram, WhatsApp, YouTube, and Twitter) the results were comparable (see Figure \ref{fig:fig-within}).
+Regarding the COVID-19 related use of \emph{social media channels} (i.e., Facebook, Instagram, WhatsApp, YouTube, and Twitter) the results were comparable (see Figure \ref{fig:fig-within}).
 Changes in the frequency of using different social media channels to attain information regarding COVID-19 were unrelated to meaningful changes in well-being.
@@
 
-For an overview of all within-person effects, see Table \ref{tab:tab-within} and Figure \ref{fig:fig-within}.
+For an overview of all within-person effects, see Figure \ref{fig:fig-within}.
 
@@
 
-\begin{table}[tbp]
-
-\begin{center}
-\begin{threeparttable}
-
-\caption{\label{tab:tab-within}Overview of all within-person effects.}
-
-\footnotesize{
-
-\begin{tabular}{lrrrrr}
-\toprule
- &  & \multicolumn{2}{c}{Confidence interval}  &  &\\
-\cmidrule(r){3-4}
-Predictor & \multicolumn{1}{c}{b} & \multicolumn{1}{c}{Lower} & \multicolumn{1}{c}{Higher} & \multicolumn{1}{c}{beta} & \multicolumn{1}{c}{p}\\
-\midrule
-Life satisfaction &  &  &  &  & \\
-\ \ \ Reading & 0.01 & -0.03 & 0.05 & 0.01 & .639\\
-\ \ \ Liking \& Sharing & -0.02 & -0.06 & 0.01 & -0.01 & .227\\
-\ \ \ Posting & -0.04 & -0.08 & -0.01 & -0.02 & .025\\
-\ \ \ Facebook & -0.01 & -0.04 & 0.02 & -0.01 & .527\\
-\ \ \ Instagram & 0.03 & -0.01 & 0.07 & 0.02 & .149\\
-\ \ \ WhatsApp & 0.00 & -0.04 & 0.04 & 0.00 & .917\\
-\ \ \ YouTube & 0.01 & -0.03 & 0.04 & 0.00 & .713\\
-\ \ \ Twitter & -0.01 & -0.05 & 0.03 & 0.00 & .503\\
-Positive affect &  &  &  &  & \\
-\ \ \ Reading & -0.03 & -0.05 & -0.02 & -0.04 & < .001\\
-\ \ \ Liking \& Sharing & 0.01 & -0.01 & 0.02 & 0.01 & .508\\
-\ \ \ Posting & 0.02 & 0.01 & 0.04 & 0.02 & .003\\
-\ \ \ Facebook & 0.00 & -0.02 & 0.01 & 0.00 & .671\\
-\ \ \ Instagram & -0.01 & -0.02 & 0.01 & -0.01 & .390\\
-\ \ \ WhatsApp & 0.01 & -0.01 & 0.02 & 0.01 & .374\\
-\ \ \ YouTube & 0.00 & -0.01 & 0.02 & 0.00 & .686\\
-\ \ \ Twitter & 0.01 & 0.00 & 0.03 & 0.01 & .130\\
-Negative affect &  &  &  &  & \\
-\ \ \ Reading & 0.00 & -0.01 & 0.02 & 0.00 & .747\\
-\ \ \ Liking \& Sharing & 0.02 & 0.00 & 0.04 & 0.02 & .022\\
-\ \ \ Posting & 0.05 & 0.04 & 0.07 & 0.05 & < .001\\
-\ \ \ Facebook & 0.00 & -0.01 & 0.01 & 0.00 & .710\\
-\ \ \ Instagram & 0.00 & -0.02 & 0.01 & 0.00 & .654\\
-\ \ \ WhatsApp & 0.00 & -0.01 & 0.01 & 0.01 & .417\\
-\ \ \ YouTube & 0.01 & 0.00 & 0.02 & 0.02 & .011\\
-\ \ \ Twitter & 0.02 & 0.01 & 0.04 & 0.02 & .008\\
-\bottomrule
-\end{tabular}
-
-}
-
-\end{threeparttable}
-\end{center}
-
-\end{table}
-
 \hypertarget{exploratory-analyses}{%
@@
 
-Based on a panel study with 34 waves largely representative of the Austrian population, this study analyzed the effects of COVID-19 related social media use on well-being.
-Between person correlation analyses showed that more active users of COVID-19 related content on social media also reported decreased well-being.
-For example, respondents who read more COVID-19 related content than others reported slightly lower levels of life satisfaction, somewhat lower levels of positive affect, and substantially higher levels of negative affect than others.
-To see if these between person correlations would translate to within-person effects, I analyzed if changes in a person's media use led to changes in their well-being.
-The within-person relations showed a different pattern.
-If people consumed more COVID-19 content on social media than usual, this did not meaningfully reduce their well-being.
-Although several statistically significant effects were found, these were very small.
-For example, people who read more COVID-19 related posts than usual reported slightly decreased positive affect.
-People who liked and shared more COVID-19 related posts than usual reported slightly higher levels of negative affect.
-Posting more content about COVID-19 than usual slightly decreased life satisfaction, while increasing both negative affect and positive affect.
-Using Twitter for COVID-19 related content slightly increased negative affect, as did YouTube.
-Again, although all of these within-person effects were statistically significant, they were very small, smaller than the predefined smallest effect size of interest.
-According to the preregistered procedure, they should hence be considered irrelevant.
-Additional analyses revealed that other factors, for which we would expect to find meaningful effects, such as health or sports, indeed showed substantial and meaningful impacts on well-being.
-In addition, when testing for the longer intervals of one month and four months, again no meaningful effects were found.
-In conclusion, COVID-19 related activity on social media was not a particularly strong influence on peoples' well-being.
-The results do not support the popular fears that ``doomscrolling'' or overusing social media during times of crises constitutes a prominent risk for well-being.
-
-These specific observations notwithstanding, several general trends can be observed.
-First, overall the results do suggest that effects of COVID-19 related social media use on well-being tend to take place in the negative as opposed to the positive spectrum.
-Although very small, five statistically significant negative results of COVID-19 related social media use on well-being were found.
-Only one positive effect emerged.
-
-Second, six significant outcomes emerged for positive or negative affect, but only one for life satisfaction.
-Life satisfaction is more stable and not that easily affected by any type or channel of social media communication.
-The more fluctuating positive and negative affect, however, were affected (albeit only slightly).
-Liking, sharing, and posting COVID-19 related content, and spending more time on Twitter and YouTube to browse COVID-19 related content, all slightly negatively influenced affect.
-This is aligned with prior findings which showed that social media use is associated with increased negative affect but not with life satisfaction (Meier \& Reinecke, 2020).
-Conversations about COVID-19 on social media are often extreme, negative, or aggressive (L. Fan et al., 2020).
-More deeply engaging with this type of content could negatively affect active authors.
-The hypothesis that tonality could explain the negative effects is especially supported by the observation that spending more time on Twitter and YouTube than usual increased negative affect.
-Communication on both channels is often found to be negative and impolite (e.g., Mueller \& Saeltzer, 2022).
-Consuming more negative and misleading information could hence explain the (slightly) increased levels of negative affect.
-
-Third, the results show that it makes sense to analyze different communication types and communication channels.
-Reading slightly reduced positive affect, while liking, sharing, and posting slightly increased negative affect.
-Interestingly, posting COVID-19 related comment slightly increased negative affect, while at the same time it also slightly increased positive affect.
-Posting content is often met with strong reaction, both positive by means of likes and negative by means of critical comments.
-Overall, though, posting led to slightly reduced levels of life satisfaction.
-In conclusion, whereas it was often stated that passive use is bad and active use good (Verduyn et al., 2015), this pattern was only partially found here.
-The results are aligned with the findings from Valkenburg et al. (2022), who could not confirm that active use is good and that passive use is bad.
-Focusing on communication channels, Twitter and YouTube seem to be more negative, while Instagram, WhatsApp, and Facebook were neutral.
-But, again, all of these effects are very small.
-
-Taken together, the results are hence aligned with the underlying theoretical models and prior empirical results.
-The findings support the differential susceptibility of media effects model (Valkenburg \& Peter, 2013), such that effects are generally small and that they depend on the type and channel of communication.
-Additional analyses did not reveal that effects depended on gender.
-Age also large did not play a significant moderation role, but effects of posting COVID-19 related content were found to be more negative for Generation Z.
-Indeed, it has often been argued that effects of social media use are more negative for Gen Z than for prior generations, and this finding can be seen a further tentative support for this hypothesis.
-From a broader perspective, the results are well-aligned with mood management theory (Zillmann, 1988) and the uses and gratifications approach (Katz et al., 1973), whose premises preclude particularly negative effects of routine and widespread media consumption.
-Both theories posit that if the effects of social media were indeed profoundly negative on average, then people likely would not spend so much time on social media engaging with COVID-19 content.
-Finally, recent empirical studies and meta-analyses reported rather small negative effects, too (Meier \& Reinecke, 2020),
-echoing the results obtained here.
-
-\hypertarget{limitations}{%
-\subsection{Limitations}\label{limitations}}
-
-Focusing on within-person effects and controlling for several potential confounders, this study provides an improved perspective on assessing causality.
-However, several challenges remain.
-In order to correctly establish causality in non-experimental designs, it is necessary to control for \emph{all} relevant confounding third variables (Rohrer, 2018).
-Although this study included are large list of confounders, it could still be that crucial variables were missed.
-More thought needs to be invested in which factors to control for and, equally important, for which factors not to control for.
-
-Although I had already reduced the predefined SESOIs to be less conservative, one could argue they were still too large.
-Media use is only one aspect of several factors that simultaneously affect well-being.
-Is it realistic to expect that changing only \emph{one} of these aspects should already manifest in a detectable change in well-being?
-Or would it make more sense to expect that thoroughly committing to say \emph{two} activities (e.g.~regularly exercising \emph{and} establishing a reading habit) should then cause a detectable improvement in well-being?
-Practically, this would imply a SESOI half the size defined here, namely \emph{b} = \textbar.15\textbar{} for life satisfaction and \emph{b} = \textbar.075\textbar{} for affect.
-In the case of this study, however, even halving the SESOI would not make a difference.
-All but one effect would still be completely in the null region, and no effect would fall completely outside of the null region.
-I encourage future research to elaborate on what effect sizes are considered meaningful and what not.
-
-Both media use and well-being were measured using self-reports.
-Because assessing well-being necessarily requires introspection, using self-reports for affect and life satisfaction is adequate.
-However, for social media use objective measures are preferable, as people often cannot reliably estimate their use (Scharkow, 2016).
-
-Being collected in a single country, the generalizability of the results is limited.
-They might not hold true in other cultures, especially non-Western cultures with a different media landscape or alternative social media channels.
-
-\hypertarget{conclusion}{%
-\subsection{Conclusion}\label{conclusion}}
-
-In this study, COVID-19 related social media use did not meaningfully affect well-being.
-Very small negative effects were found for writing COVID-19 related posts, sharing COVID-19 related content, and spending more time than usual on Twitter.
-Factors other than social media use, however, were meaningfully related to well-being, including physical health, exercise, satisfaction with democracy, or believing that one is in control of one's life.
-In light of the overall very small effects, engaging in COVID 19-related social media use should not be considered a major concern for one's well-being.
+This study, based on a representative panel study spanning 34 waves within the Austrian population, investigated the impact of COVID-19-related social media usage on well-being.
+The between-person correlations revealed that increased engagement with COVID-19 content on social media was associated with decreased well-being.
+For instance, individuals consuming more COVID-19-related content reported slightly lower life satisfaction, somewhat reduced positive affect, and notably elevated negative affect compared to others.
+
+To explore if these between-person correlations translated into within-person effects, it was examined whether changes in media consumption corresponded with changes in well-being.
+Contrary to expectations, increased consumption of COVID-19 content did not significantly decrease well-being at a meaningful level.
+While several statistically significant effects emerged, their magnitudes were notably small.
+For instance, reading more COVID-19 posts than usual slightly decreased positive affect.
+Liking and sharing more COVID-19 content than usual were associated with slightly higher negative affect.
+Posting more COVID-19 content reduced life satisfaction slightly while elevating both positive and negative affect.
+The effects, although statistically significant, fell within a predefined range considered insignificant.
+
+Further analysis demonstrated that factors such as health and physical activity, which would be expected to have substantial impacts on well-being, indeed showed significant influences.
+Additionally, extended assessments covering one and four months did not yield meaningful effects.
+In summary, the influence of COVID-19-related social media activity on well-being was not substantial.
+This counters popular concerns over excessive social media use during crises causing substantial well-being risks.
+
+There is no consensus among scholars as to when effects become practically relevant and meaningful.
+If we adopt a more liberal and in cautious perspective, some general trends emerged.
+The study indicates a tendency for COVID-19-related social media usage to impact well-being negatively more frequently than positively.
+Notably several small yet statistically significant negative effects were observed, contrasting with a single positive effect.
+Furthermore, several significant associations were found between social media use and affect but only one with life satisfaction, a more stable measure.
+
+These results align with prior research highlighting that social media usage is associated with elevated negative affect but not reduced life satisfaction {[}meierComputermediatedCommunicationSocial2020a{]}.
+The often extreme, negative, or aggressive tone of COVID-19 discussions on social media (L.
+Fan et al., 2020) could adversely impact active authors.
+The negative effects may be explained by the consumption of negative and misleading information.
+
+Moreover, the study suggest that varying communication types and channels warrant separate analyses.
+For instance, reading content slightly reduced positive affect, while liking, sharing, and posting increased negative affect slightly.
+Posting COVID-19-related comments increased both negative and positive affect slightly, while reducing life satisfaction.
+This suggests that posting generates stronger reactions, both positive and negative.
+Together, this challenges the assumption that active use is inherently beneficial and passive use detrimental (Valkenburg et al., 2022).
+Notably, communication channels exhibited differences.
+Twitter and YouTube appeared more negative, whereas Instagram, WhatsApp, and Facebook maintained neutrality.
+Despite these nuanced findings, the effects remained small.
+
+In conclusion, the study aligns with theoretical models and past research.
+It supports the notion of small media effects influenced by communication type and channel (Meier \& Reinecke, 2020).
+The differential susceptibility model (Valkenburg \& Peter, 2013) was supported, indicating that effects are small and dependent on communication factors.
+Age did not significantly moderate the effects, yet posting COVID-19 content was more negative for Generation Z, potentially reflecting broader negative effects of social media on this generation.
+The study also echoes mood management theory (Zillmann, 1988) and the uses and gratifications approach (Katz et al., 1973), implying that if the effects were highly negative, people would not engage with COVID-19 content as extensively.
+
+Despite these insights, limitations exist.
+While the study's focus on within-person effects enhances causal understanding, challenges related to additional confounding variables (Rohrer, 2018), best definition of the SESOI (Anvari \& Lakens, 2021), and exact measurement of media use (Scharkow, 2016).
+The study's results are applicable primarily to Western societies and may not hold true in different cultural contexts.
+
+The study's findings conclude that COVID-19-related social media activity minimally affects well-being, with other factors playing more substantial roles.
+In light of these minimal effects, concerns over COVID-19-related social media engagement on well-being appear to be unwarranted.
 
@@
 \begin{CSLReferences}{1}{0}
+\leavevmode\vadjust pre{\hypertarget{ref-anvariUsingAnchorbasedMethods2021a}{}}%
+Anvari, F., \& Lakens, D.
+(2021).
+Using anchor-based methods to determine the smallest effect size of interest.
+\emph{Journal of Experimental Social Psychology}, \emph{96}, 104159.
+\url{https://doi.org/10.1016/j.jesp.2021.104159}
+
 \leavevmode\vadjust pre{\hypertarget{ref-atheyDigitalPublicHealth2023}{}}%
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-dienlinDisplacementReinforcementReciprocity2017}{}}%
-Dienlin, T., Masur, P.
-K., \& Trepte, S.
-(2017).
-Displacement or reinforcement?
-{The} reciprocity of {FtF}, {IM}, and {SNS} communication and their effects on loneliness and life satisfaction.
-\emph{Journal of Computer-Mediated Communication}, \emph{22}(2), 71--87.
-\url{https://doi.org/10.1111/jcc4.12183}
-
 \leavevmode\vadjust pre{\hypertarget{ref-dormannOptimalTimeLags2015}{}}%
@@
 \url{https://doi.org/10.1002/pra2.313}
-
-\leavevmode\vadjust pre{\hypertarget{ref-funderEvaluatingEffectSize2019}{}}%
-Funder, D.
-C., \& Ozer, D.
-J.
-(2019).
-Evaluating effect size in psychological research: {Sense} and nonsense.
-\emph{Advances in Methods and Practices in Psychological Science}, \emph{2}(2), 156--168.
-\url{https://doi.org/10.1177/2515245919847202}
 
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-lakensEquivalenceTestingPsychological2018}{}}%
-Lakens, D., Scheel, A.
-M., \& Isager, P.
-M.
-(2018).
-Equivalence testing for psychological research: {A} tutorial.
-\emph{Advances in Methods and Practices in Psychological Science}, \emph{1}(2), 259--269.
-\url{https://doi.org/10.1177/2515245918770963}
-
 \leavevmode\vadjust pre{\hypertarget{ref-latikkaLonelinessPsychologicalDistress2022}{}}%
@@
 \url{https://doi.org/10.1177/0093650220958224}
-
-\leavevmode\vadjust pre{\hypertarget{ref-muellerTwitterMadeMe2022}{}}%
-Mueller, S.
-D., \& Saeltzer, M.
-(2022).
-Twitter made me do it!
-{Twitter}'s tonal platform incentive and its effect on online campaigning.
-\emph{Information, Communication \& Society}, \emph{25}(9), 1247--1272.
-\url{https://doi.org/10.1080/1369118X.2020.1850841}
 
@@
 \leavevmode\vadjust pre{\hypertarget{ref-zillmannMoodManagementCommunication1988}{}}%
-Zillmann, D. (1988). Mood {Management Through Communication Choices}. \emph{American Behavioral Scientist}, \emph{31}(3), 327340. \url{https://doi.org/10.1177/000276488031003005}
+Zillmann, D. (1988). Mood management through communication choices. \emph{American Behavioral Scientist}, \emph{31}(3), 327340. \url{https://doi.org/10.1177/000276488031003005}
 

### 2023-11-03 manuscript.Rmd
- updates before submission

@@
 
-# header-includes:
-#   - |
-#     \makeatletter
-#     \renewcommand{\paragraph}{\@startsection{paragraph}{4}{\parindent}%
-#       {0\baselineskip \@plus 0.2ex \@minus 0.2ex}%
-#       {-1em}%
-#       {\normalfont\normalsize\bfseries\typesectitle}}
-# 
-#     \renewcommand{\subparagraph}[1]{\@startsection{subparagraph}{5}{1em}%
-#       {0\baselineskip \@plus 0.2ex \@minus 0.2ex}%
-#       {-\z@\relax}%
-#       {\normalfont\normalsize\bfseries\itshape\hspace{\parindent}{#1}\textit{\addperi}}{\relax}}
-#     \makeatother
-#     \usepackage{endnotes}
-#     \setlength{\parskip}{0em}
-#     \raggedbottom
-#     \note{\clearpage}
-#     \usepackage{setspace}
-#     \AtBeginEnvironment{tabular}{\singlespacing}
-#     \AtBeginEnvironment{lltable}{\singlespacing}
-#     \AtBeginEnvironment{tablenotes}{\doublespacing}
-#     \captionsetup[table]{font={stretch=1.5}}
-#     \captionsetup[figure]{font={stretch=1.5}}
-# 
-# csl               : "`r system.file('rmd', 'apa7.csl', package = 'papaja')`"
-# documentclass     : "apa7"
-# output            :
-#   papaja::apa6_pdf:
-#     latex_engine  : xelatex
+header-includes:
+  - |
+    \makeatletter
+    \renewcommand{\paragraph}{\@startsection{paragraph}{4}{\parindent}%
+      {0\baselineskip \@plus 0.2ex \@minus 0.2ex}%
+      {-1em}%
+      {\normalfont\normalsize\bfseries\typesectitle}}
+
+    \renewcommand{\subparagraph}[1]{\@startsection{subparagraph}{5}{1em}%
+      {0\baselineskip \@plus 0.2ex \@minus 0.2ex}%
+      {-\z@\relax}%
+      {\normalfont\normalsize\bfseries\itshape\hspace{\parindent}{#1}\textit{\addperi}}{\relax}}
+    \makeatother
+    \usepackage{endnotes}
+    \setlength{\parskip}{0em}
+    \raggedbottom
+    \note{\clearpage}
+    \usepackage{setspace}
+    \AtBeginEnvironment{tabular}{\singlespacing}
+    \AtBeginEnvironment{lltable}{\singlespacing}
+    \AtBeginEnvironment{tablenotes}{\doublespacing}
+    \captionsetup[table]{font={stretch=1.5}}
+    \captionsetup[figure]{font={stretch=1.5}}
+
+csl               : "`r system.file('rmd', 'apa7.csl', package = 'papaja')`"
+documentclass     : "apa7"
+output            :
+  papaja::apa6_pdf:
+    latex_engine  : xelatex
 # output            : papaja::apa6_docx
-output            : html_document
+# output            : html_document
 ---
@@
 Exploring the hypothesis entails establishing the criteria for discerning a 'trivial effect size.' 
-To achieve this, it is necessary to define the smallest effect size of interest (SESOI) [@anvariUsingAnchorbasedMethods2021a]. 
+To achieve this, it is necessary to define the smallest effect size of interest (SESOI) [@anvariUsingAnchorbasedMethods2021]. 
 In this context, a trivial effect should fall below the threshold set by the SESOI (detailed below).
@@
 For instance, individuals consuming more COVID-19-related content reported slightly lower life satisfaction, somewhat reduced positive affect, and notably elevated negative affect compared to others.
+However, to explore if these between-person correlations translated into within-person effects, it was examined whether within-person changes in media consumption corresponded with within-person changes in well-being.
 
-To explore if these between-person correlations translated into within-person effects, it was examined whether changes in media consumption corresponded with changes in well-being.
 Contrary to expectations, increased consumption of COVID-19 content did not significantly decrease well-being at a meaningful level.
@@
 
-There is no consensus among scholars as to when effects become practically relevant and meaningful.
-If we adopt a more liberal and in cautious perspective, some general trends emerged.
-The study indicates a tendency for COVID-19-related social media usage to impact well-being negatively more frequently than positively.
-Notably several small yet statistically significant negative effects were observed, contrasting with a single positive effect.
-Furthermore, several significant associations were found between social media use and affect but only one with life satisfaction, a more stable measure.
+That said, there is no consensus among scholars as to when effects become practically relevant and meaningful.
+If we adopt a more liberal and cautious perspective, the study indicates a tendency for COVID-19-related social media usage to impact well-being negatively more frequently than positively.
+Notably several statistically significant negative effects were observed, contrasting with a single positive effect.
+Hence, although effects were overall very small, the trend was for effects to be negative.
 
-These results align with prior research highlighting that social media usage is associated with elevated negative affect but not reduced life satisfaction [meierComputermediatedCommunicationSocial2020a].
-The often extreme, negative, or aggressive tone of COVID-19 discussions on social media [@fanStigmatizationSocialMedia2020] could adversely impact active authors.
-The negative effects may be explained by the consumption of negative and misleading information.
+These results align with prior research highlighting that social media usage is rather associated with elevated negative affect but not reduced life satisfaction [@meierComputermediatedCommunicationSocial2020a].
+The often extreme, negative, or aggressive tone of COVID-19 discussions on social media [@fanStigmatizationSocialMedia2020] could adversely impact more active authors.
+The negative effects may be explained by the consumption of negative and misleading information [@liYouTubeSourceInformation2020].
 
 Moreover, the study suggest that varying communication types and channels warrant separate analyses.
-For instance, reading content slightly reduced positive affect, while liking, sharing, and posting increased negative affect slightly.
-Posting COVID-19-related comments increased both negative and positive affect slightly, while reducing life satisfaction.
+Reading COVID-19 related content slightly reduced positive affect, while liking, sharing, and posting increased negative affect slightly.
+Posting COVID-19-related comments increased both negative and positive affect slightly, while it reduced life satisfaction.
 This suggests that posting generates stronger reactions, both positive and negative.
@@
 Notably, communication channels exhibited differences.
-Twitter and YouTube appeared more negative, whereas Instagram, WhatsApp, and Facebook maintained neutrality.
-Despite these nuanced findings, the effects remained small.
+Twitter and YouTube appeared more negative, whereas Instagram, WhatsApp, and Facebook were neutral.
 
 In conclusion, the study aligns with theoretical models and past research.
-It supports the notion of small media effects influenced by communication type and channel [@meierComputermediatedCommunicationSocial2020a].
-The differential susceptibility model [@valkenburgDifferentialSusceptibilityMedia2013] was supported, indicating that effects are small and dependent on communication factors.
-Age did not significantly moderate the effects, yet posting COVID-19 content was more negative for Generation Z, potentially reflecting broader negative effects of social media on this generation.
-The study also echoes mood management theory [@zillmannMoodManagementCommunication1988] and the uses and gratifications approach [@katzUsesGratificationsResearch1973], implying that if the effects were highly negative, people would not engage with COVID-19 content as extensively.
+It supports the notion of social media effects being small and contingent on communication type and channel [@meierComputermediatedCommunicationSocial2020a; @valkenburgDifferentialSusceptibilityMedia2013].
+Posting COVID-19 content was more negative for Generation Z, potentially reflecting broader negative effects of social media on this generation.
+Effects not being strongly negative also help explain why people spent so much time engaging with COVID-19 content on social media [@katzUsesGratificationsResearch1973; @zillmannMoodManagementCommunication1988].
 
-Despite these insights, limitations exist.
-While the study's focus on within-person effects enhances causal understanding, challenges related to additional confounding variables [@rohrerThinkingClearlyCorrelations2018], best definition of the SESOI [@anvariUsingAnchorbasedMethods2021a], and exact measurement of media use [@scharkowAccuracySelfreportedInternet2016].
+Several limitations exist.
+While the study's focus on within-person effects enhances causal understanding, challenges exist related to additional relevant confounding variables not included here [@rohrerThinkingClearlyCorrelations2018], correct definition of the SESOI [@anvariUsingAnchorbasedMethods2021], and exact measurement of media use [@scharkowAccuracySelfreportedInternet2016].
 The study's results are applicable primarily to Western societies and may not hold true in different cultural contexts.
@@
 In light of these minimal effects, concerns over COVID-19-related social media engagement on well-being appear to be unwarranted.
+When trying to improve well-being during a pandemic, instead of focusing on social media it seems more fruitful to address other, more pertinent societal problems related to health care, regular exercise, or psychological resilience.
 

### 2023-11-03 manuscript.tex
- updates before submission

@@
 For instance, individuals consuming more COVID-19-related content reported slightly lower life satisfaction, somewhat reduced positive affect, and notably elevated negative affect compared to others.
-
-To explore if these between-person correlations translated into within-person effects, it was examined whether changes in media consumption corresponded with changes in well-being.
+However, to explore if these between-person correlations translated into within-person effects, it was examined whether within-person changes in media consumption corresponded with within-person changes in well-being.
+
 Contrary to expectations, increased consumption of COVID-19 content did not significantly decrease well-being at a meaningful level.
@@
 
-There is no consensus among scholars as to when effects become practically relevant and meaningful.
-If we adopt a more liberal and in cautious perspective, some general trends emerged.
-The study indicates a tendency for COVID-19-related social media usage to impact well-being negatively more frequently than positively.
-Notably several small yet statistically significant negative effects were observed, contrasting with a single positive effect.
-Furthermore, several significant associations were found between social media use and affect but only one with life satisfaction, a more stable measure.
-
-These results align with prior research highlighting that social media usage is associated with elevated negative affect but not reduced life satisfaction {[}meierComputermediatedCommunicationSocial2020a{]}.
+That said, there is no consensus among scholars as to when effects become practically relevant and meaningful.
+If we adopt a more liberal and cautious perspective, the study indicates a tendency for COVID-19-related social media usage to impact well-being negatively more frequently than positively.
+Notably several statistically significant negative effects were observed, contrasting with a single positive effect.
+Hence, although effects were overall very small, the trend was for effects to be negative.
+
+These results align with prior research highlighting that social media usage is rather associated with elevated negative affect but not reduced life satisfaction (Meier \& Reinecke, 2020).
 The often extreme, negative, or aggressive tone of COVID-19 discussions on social media (L.
-Fan et al., 2020) could adversely impact active authors.
-The negative effects may be explained by the consumption of negative and misleading information.
+Fan et al., 2020) could adversely impact more active authors.
+The negative effects may be explained by the consumption of negative and misleading information (Li et al., 2020).
 
 Moreover, the study suggest that varying communication types and channels warrant separate analyses.
-For instance, reading content slightly reduced positive affect, while liking, sharing, and posting increased negative affect slightly.
-Posting COVID-19-related comments increased both negative and positive affect slightly, while reducing life satisfaction.
+Reading COVID-19 related content slightly reduced positive affect, while liking, sharing, and posting increased negative affect slightly.
+Posting COVID-19-related comments increased both negative and positive affect slightly, while it reduced life satisfaction.
 This suggests that posting generates stronger reactions, both positive and negative.
@@
 Notably, communication channels exhibited differences.
-Twitter and YouTube appeared more negative, whereas Instagram, WhatsApp, and Facebook maintained neutrality.
-Despite these nuanced findings, the effects remained small.
+Twitter and YouTube appeared more negative, whereas Instagram, WhatsApp, and Facebook were neutral.
 
 In conclusion, the study aligns with theoretical models and past research.
-It supports the notion of small media effects influenced by communication type and channel (Meier \& Reinecke, 2020).
-The differential susceptibility model (Valkenburg \& Peter, 2013) was supported, indicating that effects are small and dependent on communication factors.
-Age did not significantly moderate the effects, yet posting COVID-19 content was more negative for Generation Z, potentially reflecting broader negative effects of social media on this generation.
-The study also echoes mood management theory (Zillmann, 1988) and the uses and gratifications approach (Katz et al., 1973), implying that if the effects were highly negative, people would not engage with COVID-19 content as extensively.
-
-Despite these insights, limitations exist.
-While the study's focus on within-person effects enhances causal understanding, challenges related to additional confounding variables (Rohrer, 2018), best definition of the SESOI (Anvari \& Lakens, 2021), and exact measurement of media use (Scharkow, 2016).
+It supports the notion of social media effects being small and contingent on communication type and channel (Meier \& Reinecke, 2020; Valkenburg \& Peter, 2013).
+Posting COVID-19 content was more negative for Generation Z, potentially reflecting broader negative effects of social media on this generation.
+Effects not being strongly negative also help explain why people spent so much time engaging with COVID-19 content on social media (Katz et al., 1973; Zillmann, 1988).
+
+Several limitations exist.
+While the study's focus on within-person effects enhances causal understanding, challenges exist related to additional relevant confounding variables not included here (Rohrer, 2018), correct definition of the SESOI (Anvari \& Lakens, 2021), and exact measurement of media use (Scharkow, 2016).
 The study's results are applicable primarily to Western societies and may not hold true in different cultural contexts.
@@
 \begin{CSLReferences}{1}{0}
-\leavevmode\vadjust pre{\hypertarget{ref-anvariUsingAnchorbasedMethods2021a}{}}%
+\leavevmode\vadjust pre{\hypertarget{ref-anvariUsingAnchorbasedMethods2021}{}}%
 Anvari, F., \& Lakens, D.

### 2024-04-04 manuscript.Rmd
Substantial updates:  - Shortened manuscript  - Implemented peer-feedback

@@
 However, because social media are particularly engaging, many find it hard to disconnect, especially during times of crisis.
-Using data from the Austrian Corona Panel Project consisting of 3,485 participants from 34 waves, controlling for several stable and varying confounders, the results showed that COVID-19 related social media use did not meaningfully reduce well-being.
+In this preregistered study, I investigate whether using social media for COVID-19 related reasons affects psychological well-being.
+Using data from the Austrian Corona Panel Project consisting of 3,485 participants from 34 waves, this research question was analyzed using random effects within between models, controlling for several stable and varying confounders.
+The results showed that COVID-19 related social media use did not meaningfully reduce well-being.
 Other factors such as health, income, exercise, or internal locus of control showed larger and meaningful effects."
@@
 documentclass     : "apa7"
-output            :
-  papaja::apa6_pdf:
-    latex_engine  : xelatex
+# output            :
+#   papaja::apa6_pdf:
+#     latex_engine  : xelatex
 # output            : papaja::apa6_docx
-# output            : html_document
+output            : html_document
 ---
@@
 
-During the COVID-19 pandemic it was critical to stay informed regarding the latest developments.
-How dangerous is the virus? 
-In what region is it spreading? 
-How is it transmitted? What are the current safety regulations?
-To obtain relevant information, many people heavily relied on social media, with use being at an all time high [@statistaAverageDailyTime2021].
-Some actually could not stop using social media to learn about COVID-19 related news.
-A new phenomenon termed "doomscrolling" emerged [@sharmaDarkEndTunnel2022].
-Many users were glued to their screens and found it hard to pursue other relevant activities such as working, taking a break, or even looking after their children.
-In the media it was hence increasingly asked whether using social media for COVID-19 related reasons would, next to all other stressors, create an additional burden on mental health [@sandstromDoomscrollingCOVIDNews2021].
-Although research has begun addressing this question
-[e.g., @bendauAssociationsCOVID19Related2021; @edenMediaCopingCOVID192020; @sewallObjectivelyMeasuredDigital2021],
-it still largely unknown if COVID-19 related social media use during the pandemic has had a meaningful impact on well-being.
-This study hence aims to 1) reveal the effect of the different types and channels of social media use on individual well-being, 2) provide generalizable and robust results by analyzing a large-scale longitudinal data-set with 34 waves, and 3) determine the within-person causal effects by analyzing how changes in social media use lead to changes in well-being.
+# Research Transparency Statement
+
+Conflicts of interest: The author declares no conflicts of interest.
+Funding: This research did not receive external funding.
+Artificial intelligence: Chat GPT was used for language polishing.
+Ethics: This research complies with the Declaration of Helsinki (2023), aside from the requirement to preregister human subjects research.
+Ethical review and approval was not required for the study in accordance with the local legislation and institutional requirements.
+The participants provided their written informed consent to participate in this study.
+Computational reproducibility: All analyses and the manuscript itself are completely reproducible (https://XMtRA.github.io/Austrian_Corona_Panel_Project).
+
+Preregistration: The hypotheses, the sample, the measures, the analyses, and the inference criteria (SESOI, _p_-value) were preregistered on the Open Science Framework (https://osf.io/87b24/?view_only=b2289b6fec214fa88ee75a18d45c18f3).
+Because in this study data from an already existing data set was analyzed, the preregistration was done prior to accessing the data.
+In some cases, it was impossible to execute the analyses as originally planned, for example because some properties of the variables only became apparent when seeing the actual data.
+All changes are reported online (https://xmtra.github.io/Austrian_Corona_Panel_Project/preregistration_changes.html).
+Materials: All study materials are publicly available (https://xmtra.github.io/Austrian_Corona_Panel_Project/items_english.html).
+Data: All primary data are publicly available (https://doi.org/10.11587/28KQNS).
+Analysis scripts: All analysis scripts are publicly available (https://XMtRA.github.io/Austrian_Corona_Panel_Project).
+
+\newpage
+
+# Introduction
+
+Amidst the COVID-19 pandemic, staying informed became paramount, prompting heavy reliance on social media for updates, with use being at an all time high [@statistaAverageDailyTime2021]. 
+The phenomenon of "doomscrolling" emerged as individuals struggled to disengage from COVID-19-related news [@buchananBriefExposureSocial2021; @sharmaDarkEndTunnel2022], sparking concerns about its impact on mental health [@sandstromDoomscrollingCOVIDNews2021].
+While initial research began exploring this question [e.g., @bendauAssociationsCOVID19Related2021; @edenMediaCopingCOVID192020; @sewallObjectivelyMeasuredDigital2021], the extent of social media's influence on well-being remains largely unknown.
+This study aims to assess the effects of different social media usage patterns on individual well-being using a comprehensive longitudinal dataset spanning 34 waves, offering insights into within-person causal relationships.
+
+<!-- During the COVID-19 pandemic it was critical to stay informed regarding the latest developments. -->
+<!-- How dangerous is the virus?  -->
+<!-- In what region is it spreading?  -->
+<!-- How is it transmitted?  -->
+<!-- What are the current safety regulations? -->
+<!-- To obtain relevant information, many people heavily relied on social media, with use being at an all time high [@statistaAverageDailyTime2021]. -->
+<!-- Some actually could not stop using social media to learn about COVID-19 related news, and a new phenomenon termed "doomscrolling" emerged [@sharmaDarkEndTunnel2022]. -->
+<!-- Many users were glued to their screens and found it hard to pursue other relevant activities [@buchananBriefExposureSocial2021]. -->
+<!-- In the media it was hence increasingly asked whether using social media for COVID-19 related reasons would, next to all other stressors, create an additional burden on mental health [@sandstromDoomscrollingCOVIDNews2021].
+-->
+<!-- Although research has begun addressing this question -->
+<!-- [e.g., @bendauAssociationsCOVID19Related2021; @edenMediaCopingCOVID192020; @sewallObjectivelyMeasuredDigital2021], -->
+<!-- it still largely unknown if COVID-19 related social media use during the pandemic has had a meaningful impact on well-being. -->
+<!-- This study hence aims to 1) reveal the effect of the different types and channels of social media use on individual well-being, 2) provide generalizable and robust results by analyzing a large-scale longitudinal data-set with 34 waves, and 3) determine the within-person causal effects by analyzing how changes in social media use lead to changes in well-being.
+-->
 
@@
 
-This study investigates how different _facets_ of well-being are affected by different _types_ and different _channels_ of communication [@meierComputermediatedCommunicationSocial2020a].
-Building on the typology of subjective well-being [@dienerAdvancesOpenQuestions2018], three different well-being facets are analyzed: life satisfaction, positive affect, and negative affect.
-Because effects of social media depend on how they are used [@verduynPassiveFacebookUsage2015], I further distinguish three types of use and five popular channels. 
-The types of use include reading, liking and sharing, and posting COVID-19 related content.
-In doing so, this study analyzes social media use focused on COVID-19 related content, which includes posting thoughts about the pandemic, reading posts and comments, or retweeting and liking COVID-19 related news.
-Liking and sharing are combined as they both represent low-threshold, platform-ingrained, easily quantifiable interactions.
+This study investigates how different _facets_ of subjective well-being are affected by different _types_ and different _channels_ of communication [@meierComputermediatedCommunicationSocial2020a].
+Building on the typology of subjective well-being---defined as "referring to the various types of subjective evaluations of one’s life, including both cognitive evaluations and affective
+feelings" [@dienerAdvancesOpenQuestions2018, p. 3]---three different well-being facets are analyzed: life satisfaction, positive affect, and negative affect.
+
+Social media is a broad term, encompassing subdimensions such as social networking sites (e.g., Facebook or Instagram), instant messengers (e.g., WhatsApp or Signal), or interactive video-platforms (e.g., YouTube or TikTok) [@carrSocialMediaDefining2015].
+Effects of social media depend on how they are used: while purposeful, active, social use is generally related to positive outcomes, passive, non-purposeful and asocial use is related to negative outcomes [@frisonExploringRelationshipsDifferent2015; @verduynSocialNetworkingSites2022].
+In this study, I hence distinguish three types of use and five popular channels. 
+The types of use include posting (active use), reading (passive use), and liking and sharing (low-threshold active use) COVID-19 related content.
+<!-- In doing so, this study analyzes social media use focused on COVID-19 related content. -->
+<!-- , which includes posting thoughts about the pandemic, reading posts and comments , or liking and sharing COVID-19 related content. -->
 The five channels to be investigated are Facebook, Twitter, Instagram, WhatsApp, and YouTube, which at the time ranked among the most popular social media services in Austria.
@@
 
-How easily can well-being be affected by external influences?
-In general, according to the set-point theory, well-being is surprisingly stable [@lykkenHappinessWhatStudies1999].
-Although specific events such as marriage or salary increase can have significant impacts on well-being, in most cases effects are short-term, with well-being after some time returning to prior levels [@dienerAdvancesOpenQuestions2018].
-Specific factors such as unemployment, disability, or death, however, can cause long-term changes in well-being [@lucasAdaptationSetpointModel2007].
-So although well-being can be affected by external events and factors, this does not happen easily.
+<!-- How easily can well-being be affected by external influences? -->
+<!-- In general, according to the set-point theory, well-being is considered to be surprisingly stable [@lykkenHappinessWhatStudies1999]. -->
+<!-- Although specific events such as marriage or salary increase can have significant impacts on well-being, effects are often only short-term, with well-being after some time returning to prior levels [@dienerAdvancesOpenQuestions2018].
+-->
+<!-- Specific factors such as unemployment, disability, or death, however, can cause long-term changes in well-being [@lucasAdaptationSetpointModel2007]. -->
+<!-- So although well-being can be affected by external events and factors, it seems this does not happen easily. -->
 
-Can social media use be such a factor?
+<!-- Can social media use be such a factor? -->
 Current literature overviews suggest that social media use on average does seem to decrease well-being [@meierComputermediatedCommunicationSocial2020a].
@@
 These small findings can be explained with the differential susceptibility of media effects model [@valkenburgDifferentialSusceptibilityMedia2013], which states that there is substantial _variation_ of media effects for individual users.
-Whereas for some users social media are more beneficial, for others they are more harmful.
-On average, however, and this is central for this study here, effects tend to be small [@valkenburgDifferentialSusceptibilityMedia2013].
+<!-- Whereas for some users social media are more beneficial, for others they are more harmful. -->
+<!-- On average, however, effects tend to be small [@valkenburgDifferentialSusceptibilityMedia2013]. -->
 For example, in one study it was estimated that roughly one quarter of all users experienced negative effects, another quarter positive effects, while for the rest the effects were neutral [@beyensSocialMediaUse2021].
-Whether or not effects are positive or negative depend on (a) dispositional factors (e.g., personality, temperament, gender), (b) developmental factors (e.g., age, developmental tasks), (c) and social factors (e.g., environment, norms, upbringing).
-Finally, effects depend also on the content that is consumed.
-If the content is aligned with dispositions, developmental capacities, and converging contexts, effects tend to be stronger [@valkenburgDifferentialSusceptibilityMedia2013].
+<!-- Whether or not effects are positive or negative depend on (a) dispositional factors (e.g., personality, temperament, gender), (b) developmental factors (e.g., age, developmental tasks), (c) and social factors (e.g., environment, norms, upbringing) [@valkenburgDifferentialSusceptibilityMedia2013].
+-->
+<!-- Effects depend also on the content that is consumed---if the content is aligned with dispositions, developmental capacities, and converging contexts, effects tend to be stronger [@valkenburgDifferentialSusceptibilityMedia2013].
+-->
 
@@
 First, according to mood management theory [@zillmannMoodManagementCommunication1988], using media can affect people's moods.
-Use can be stimulating or overwhelming, relaxing or boring.
+<!-- Use can be stimulating or overwhelming, relaxing or boring. -->
 After some time, users implicitly learn which media help them balance their mood and affect according to their own situational needs [@zillmannMoodManagementCommunication1988].
@@
 Users select those media that they expect to have a desired effect, for example on mood, knowledge, or entertainment.
-If those beneficial media effects are missing, people will spend their time elsewhere.
-And social media, in general, offer several beneficial effects, explaining why they are used that much.
+<!-- If those beneficial media effects are missing, people will spend their time elsewhere. -->
+And social media, in general, offer several beneficial effects, explaining their ubiquitous use.
 They help find relevant information, maintain and foster relationships, express one's personality, and entertain oneself [@pelletierOneSizeDoesn2020].
-In conclusion, because people spend so much time on social media consuming COVID-19 related content, according to both mood management theory and uses and gratifications theory this indirectly suggest that average effects on well-being are likely not particularly negative.
+
+Third, and closely related, media use affords both both positive and negative mechanisms, which differently impact well-being outcomes.
+These include factors like social comparison, inspiration, social connectedness, social support, displacement of in-person activities, information management, information overload, misinformation, reputation building, romantic connection, coping resources, or emotional contagion [@clarkSocialNetworkSites2018; @hallSocialMediaUse2022; @pelletierOneSizeDoesn2020; @vandenabeeleDigitalWellbeingDynamic2021; @verduynSocialNetworkingSites2022].
+Given these differential mechanisms, strong one-sided effects appear less likely than nuanced and moderate effects.
+
+<!-- In conclusion, because people spend so much time on social media consuming COVID-19 related content, according to mood management theory, uses and gratifications theory, and multiple differential mechanisms, average effects on well-being are likely not particularly negative.
+-->
 
@@
 Despite incorrect information, social media provide a vast platform for disseminating _accurate and timely information_ about COVID-19 [@johnhopkinsuniversityCOVID19Map2023]. 
-Access to reliable information can help people make informed decisions, alleviate uncertainties, and feel empowered during the pandemic.
+<!-- Access to reliable information can help people make informed decisions, alleviate uncertainties, and feel empowered during the pandemic. -->
 Social media platforms enable individuals to _connect with others_ who are experiencing similar challenges during the pandemic [@guazziniSecondWaveAnalysis2022]. 
-Engaging in online communities and support groups can provide emotional support and create a network of like-minded individuals.
-Many mental health organizations and professionals utilize social media to share tips, strategies, and resources for _maintaining mental well-being_ during the pandemic [@twitterTalkingMentalHealthAwarenessTwitter2020].
-Engaging with such content might help individuals prioritize their mental health and develop resilience during challenging times.
-Social media campaigns and initiatives can promote _positive COVID-19 behaviors_, such as mask-wearing, physical distancing, hand hygiene, and vaccination [@atheyDigitalPublicHealth2023; @huntSocialMediaBased2022].
-Public health organizations and influencers leverage the power of social media to spread awareness and encourage responsible actions, contributing to public health efforts and fostering a sense of collective responsibility.
+<!-- Engaging in online communities and support groups can provide emotional support and create a network of like-minded individuals. -->
+Many mental health organizations and professionals utilize social media to share tips, strategies, and resources for _maintaining mental well-being_ during the pandemic. 
+<!-- Engaging with such content might help individuals prioritize their mental health and develop resilience during challenging times. -->
+Social media campaigns and initiatives can promote _positive COVID-19 behaviors_, such as mask-wearing, physical distancing, hand hygiene, and vaccination [@huntSocialMediaBased2022], which ultimately benefit well-being.
+<!-- Public health organizations and influencers leverage the power of social media to spread awareness and encourage responsible actions, contributing to public health efforts and fostering a sense of collective responsibility.
+-->
 
@@
 Social media platforms can easily spread _false or misleading information_ about COVID-19 [@liYouTubeSourceInformation2020]. 
-Due to the ease of sharing and the lack of fact-checking, inaccurate information can go viral and might cause confusion, anxiety, and panic among users.
+<!-- Due to the ease of sharing and the lack of fact-checking, inaccurate information can go viral and might cause confusion, anxiety, and panic among users. -->
 Constant exposure to COVID-19-related content on social media can lead to _information overload_ and contribute to heightened anxiety levels [@fanInformationOverloadWellbeing2021]. 
-The rapid spread of news, updates, and opinions can be overwhelming and might exacerbate existing stress or fears about the pandemic [@sharmaDarkEndTunnel2022].
-Social media platforms are known for fostering negativity, with users sometimes engaging in _cyberbullying and harassment_. 
-Discussions around COVID-19 can become heated and polarized, leading to personal attacks and online conflicts. 
-Such experiences threaten mental well-being and might contribute to feelings of distress and isolation.
+<!-- The rapid spread of news, updates, and opinions can be overwhelming and might exacerbate existing stress or fears about the pandemic [@sharmaDarkEndTunnel2022]. -->
+Discussions around COVID-19 are known for fostering negativity, with users sometimes engaging in _cyberbullying and harassment_. 
+<!-- Discussions around COVID-19 can become heated and polarized, leading to personal attacks and online conflicts.  -->
+<!-- Such experiences threaten mental well-being and might contribute to feelings of distress and isolation. -->
 Social media often showcase the highlights and accomplishments of others, encouraging _social comparison_ [@przybylskiMotivationalEmotionalBehavioral2013]. 
@@
 
-There is still little empirical research on how well-being is affected by social media use that is focused on COVID-19 specifically.
-Echoing the theoretical rationales outlines above, studies have yielded mixed results.
-Some studies found negative effects, indicating that excessive social media use for COVID-19 news led to compulsive behavior and increased stress levels, particularly due to upward social comparison [@stainbackCOVID1924News2020; @yuePassiveSocialMedia2022].
+<!-- There is still limited empirical research on how well-being is affected by social media use that is focused on COVID-19 specifically. -->
+Echoing the theoretical rationales outlines above, empirical studies have yielded mixed results.
+Some studies found negative effects, indicating that excessive social media use for COVID-19 content led to compulsive behavior and increased stress levels, particularly due to upward social comparison [@stainbackCOVID1924News2020].
 Individuals who relied on social media as their primary information source reported higher levels of anxiety and depression symptoms [@bendauAssociationsCOVID19Related2021]. 
-Similarly, increased COVID-19-related media consumption was associated with higher psychological distress.
+Doom-scrolling was associated with negative emotional experiences [@buchananBriefExposureSocial2021].
 On the other hand, some studies reported positive outcomes. 
 Certain individuals experienced increased virtual community and social connectedness during the pandemic through social media, which contributed to their well-being [@guazziniSecondWaveAnalysis2022]. 
-Additionally, engaging more on social media was associated with reduced feelings of loneliness [@latikkaLonelinessPsychologicalDistress2022].
-Several studies reported mostly neutral effects of social media use on well-being indicators [@edenMediaCopingCOVID192020; @sewallObjectivelyMeasuredDigital2021].
-Overall, the literature demonstrates a mixed picture, highlighting both positive and negative effects of social media use focused on or during COVID-19 on well-being.
+Additionally, identification with social media networks was associated with reduced feelings of loneliness [@latikkaLonelinessPsychologicalDistress2022].
+Several studies reported mostly neutral or dual effects of social media use on well-being indicators [@edenMediaCopingCOVID192020; @sewallObjectivelyMeasuredDigital2021].
+<!-- Overall, the literature demonstrates a mixed picture, highlighting both positive and negative effects of social media use focused on or during COVID-19 on well-being. -->
 
@@
 In conclusion, given these mixed empirical results, together with the observation that social media effects on well-being are very small in general, and that several plausible theoretical mechanisms exist for both positive and negative effects, I expect that COVID-19 related communication on social media should not be decidedly positive or negative.
-It seems most likely that both positive and negative coexist, but that on average using social media for COVID-19 related reasons should not have substantial effects on well-being.
+<!-- It seems most likely that both positive and negative coexist, but that on average using social media for COVID-19 related reasons should not have substantial effects on well-being. -->
 
@@
 To achieve this, it is necessary to define the smallest effect size of interest (SESOI) [@anvariUsingAnchorbasedMethods2021].
+<!-- , which helps determine an effect's practical significance.  -->
 In this context, a trivial effect should fall below the threshold set by the SESOI (detailed below).
 Determining what constitutes a minimally intriguing, nontrivial effect is a matter with normative implications, making it challenging to arrive at a definitive, singular consensus. 
-Nonetheless, it remains both essential and beneficial to strive for a reasonable benchmark. 
-With that in mind, I propose the following SESOI as a suitable reference point for this study:
+<!-- Nonetheless, it remains both essential and beneficial to strive for a reasonable benchmark.  -->
+I propose the following SESOI as a suitable reference point for this study:
 
-> SESOI: If a heavy user of COVID-19 related social media news suddenly _stops_ using social media altogether, this should have a _noticeable_ impact on their overall well-being.
+> SESOI: If a heavy user of COVID-19 related social media content suddenly _stops_ using social media altogether, this should have a _noticeable_ impact on their overall well-being.
 
-What does this mean practically and how can it be operationalized?
-In this study, COVID-19 related social media use was measured on a 5-point scale, ranging from 1 = _never_ to 5 = _several times a day_.
+<!-- What does this mean practically and how can it be operationalized? -->
+Put concretely, in this study COVID-19 related social media use was measured on a 5-point scale, ranging from 1 = _never_ to 5 = _several times a day_.
 Thus, a change of four units in social media use (e.g., a complete stop) should correspond to a noticeable change in well-being.
@@
 So if satisfaction is measured on a 7-point scale, a four unit change in social media use should result in a one unit change in life satisfaction. 
-In this study, life satisfaction was measured on an 11-point scale, and affect on a 5-point scale. 
-Transposed to this scaling, the SESOI for life satisfaction is _b_ =  ±.30, and for positive and negative affect _b_ = ±.15 (for more information, see online supplementary material).
+Life satisfaction was measured on an 11-point scale, and affect on a 5-point scale. 
+Transposed to this scaling, the SESOI for life satisfaction is _b_ =  ±.30, and for positive and negative affect _b_ = ±.15 (see online supplementary material).
 
 # Method
-## Preregistration
-
-The hypotheses, the sample, the measures, the analyses, and the inference criteria (SESOI, _p_-value) were preregistered on the Open Science Framework, accessible here: https://osf.io/87b24/?view_only=b2289b6fec214fa88ee75a18d45c18f3.
-Because in this study I analyze data from an already existing large-scale data set, the preregistration was done prior to accessing the data.
-The preregistration was designed on the basis of the panel documentation online [@kittelAustrianCoronaPanel2020].
-In some cases, it was impossible to execute the analyses as I had originally planned, for example because some properties of the variables only became apparent when seeing the actual data.
-The most relevant deviations are reported below, and a complete list of all changes can be found in the online [companion website](https://XMtRA.github.io/Austrian_Corona_Panel_Project) (https://XMtRA.github.io/Austrian_Corona_Panel_Project).
-
 ## Sample
 
-The data come from the Austrian Corona Panel Project [@kittelAustrianCoronaPanel2020], which is a large-scale standalone panel study. 
-The data are hosted on AUSSDA, are publicly available here (https://doi.org/10.11587/28KQNS), and consist of 34 waves.
-Participants were sampled from a pre-existing online access panel provided by the company Marketagent, Austria.
-Panel members were incentivized with 180 credit points for each wave of the study.
+The data come from the Austrian Corona Panel Project [@kittelAustrianCoronaPanel2020], which is a large-scale standalone panel study, consisting of 34 waves.
+<!-- The data are hosted on AUSSDA, are publicly available here (https://doi.org/10.11587/28KQNS). -->
+<!-- Participants were sampled from a pre-existing online access panel provided by the company Marketagent, Austria. -->
+<!-- Panel members were incentivized with 180 credit points for each wave of the study. -->
 The study was conducted between March 2020 and February 2023.
@@
 Each wave consists of at least 1,500 respondents.
-Panel mortality was compensated through a continuous re-acquisition of new participants. 
+Panel mortality (on average ca. 5 percent per wave) was compensated through a continuous re-acquisition of new participants (on average ca. 80 new respondents per wave). 
 The sample size was _N_ = `r d_raw %>% nrow() %>% prettyNum(big.mark = ",")`, with overall `r d_long %>% nrow() %>% prettyNum(big.mark = ",")` observations.
-For an overview of the study set-up, see Figure \@ref(fig:study-desc).
 
-```{r study-desc, warning=F, message=F, fig.cap="Overview of study set-up", out.width = "1\\textwidth", warning=F, message=F, prompt=F}
-knitr::include_graphics("figures/fig_study_description.png")
-```
+<!-- For an overview of the study set-up, see Figure \@ref(fig:study-desc). -->
+
+<!-- ```{r study-desc, warning=F, message=F, fig.cap="Overview of study set-up", out.width = "1\\textwidth", warning=F, message=F, prompt=F} -->
+<!-- knitr::include_graphics("figures/fig_study_description.png") -->
+<!-- ``` -->
 
@@
 All respondents needed to have access to the internet (via computer or mobile devices such as smartphones or tablets). 
-Ethical review and approval was not required for the study in accordance with the local legislation and institutional requirements. 
-The participants provided their written informed consent to participate in this study.
 The average age was `r mean(2021 - d_wide$year_birth, na.rm = T) %>% round(0)` years, `r mean(d_wide$male, na.rm = T) %>% round(., 2) * 100` percent were male, `r (d_wide %>% filter(.$edu_fac %in% c("State college", "Bachelor", "Master", "PhD"))) %>% nrow() %>% divide_by(sum(table(d_wide$edu_fac))) %>% multiply_by(100) %>% round(0)` percent had a University degree, and `r (table(d_wide$employment_fac)["Unemployed"] / sum(table(d_wide$employment_fac))) %>% unname() %>% multiply_by(100) %>% round(., digits = 0)` percent were currently unemployed.
@@
 Here, the null-region lies between -.30 and +.30.
-If the 95% confidence interval falls completely within the null-region (e.g., _b_ = -.05, [95% CI: -.15, .05]), the hypothesis that the effect is trivial is supported.
+<!-- If the 95% confidence interval falls completely within the null-region (e.g., _b_ = -.05, [95% CI: -.15, .05]), the hypothesis that the effect is trivial is supported. -->
 If the confidence interval falls completely outside of the null-region (e.g., _b_ = -.40, [95% CI: -.45, -.35]), the hypothesis is rejected and the existence of a meaningful negative effect is supported.
-If the confidence interval and the null region overlap (e.g., _b_ = -.30, [95% CI: -.35, -.25]), the hypothesis is not supported and the results are considered inconclusive, while a meaningful positive effect is rejected.
-If the confidence interval exceeds both sides of the null region (e.g., _b_ = -.025, [95% CI: -.40, .35]), the hypothesis is not supported and judgement is suspended.
+<!-- If the confidence interval and the null region overlap (e.g., _b_ = -.30, [95% CI: -.35, -.25]), the hypothesis is not supported and the results are considered inconclusive, while a meaningful positive effect is rejected.
+-->
+<!-- If the confidence interval exceeds both sides of the null region (e.g., _b_ = -.025, [95% CI: -.40, .35]), the hypothesis is not supported and judgement is suspended. -->
 For an illustration, see Figure \@ref(fig:sesoi).
@@
 
-When analyzing causality through longitudinal designs, it is important to address several critical aspects.
-An inherent requirement of analyzing causal effects within non-experimental designs is adopting an internal perspective, focusing on _within-person effects_ [@hamakerWhyResearchersShould2014].
-This entails evaluating how alterations in an individual's media consumption directly impact their well-being.
-Comparisons between individuals, as observed in longitudinal data, lack the granularity needed to understand these causal dynamics [@hamakerWhyResearchersShould2014].
-Consequently, this study exclusively investigates within-person effects to shed light on this intricate relationship.
+<!-- When analyzing causality through longitudinal designs, it is important to address several critical aspects.  -->
+Analyzing causal effects within non-experimental designs requires adopting an internal perspective, focusing on _within-person effects_ [@hamakerWhyResearchersShould2014]. 
+This entails evaluating how alterations in an individual's media consumption directly impact changes in their own well-being. 
+<!-- Comparisons between individuals, as observed in longitudinal data, lack the granularity needed to understand these causal dynamics [@hamakerWhyResearchersShould2014].  -->
+Consequently, this study exclusively investigates within-person effects.
+<!-- to shed light on this intricate relationship. -->
 
@@
 However, a cautious balance must be maintained by not controlling for variables that mediate the relationship [@rohrerThinkingClearlyCorrelations2018], as their inclusion could distort the assessment of the causal effect.
-This study therefore incorporated several control variables.
-These encompass gender, age, education, place of birth and parental place of birth in Austria, Vienna residency, consumption of text-based and video-based news, household characteristics, health status, living space features, employment-related factors, income, outdoor activities, risk propensity, and locus of control [@egerStatisticalMetaanalysisWellbeing2015].
-These variables have demonstrated connections with both social media use and well-being and are not mediators.
+This study therefore incorporated several control variables (see below), which have demonstrated connections with both social media use and well-being and are likely not mediators [@egerStatisticalMetaanalysisWellbeing2015].
+<!-- These encompass gender, age, education, place of birth and parental place of birth in Austria, Vienna residency, consumption of text-based and video-based news, household characteristics, health status, living space features, employment-related factors, income, outdoor activities, risk propensity, and locus of control.
+-->
 
-A fundamental prerequisite for establishing causality is determining a _plausible temporal interval_ [@dormannOptimalTimeLags2015].
+Establishing causality necessitates determining a _plausible temporal interval_ [@rohrerTheseAreNot2023]. 
 For instance, fluctuations in positive and negative affect call for shorter intervals, while the more enduring nature of life satisfaction implies longer intervals [@dienlinImpactDigitalTechnology2020].
-
 In this study, I examine the linkage between changes in social media use and changes in affect within the same week. 
 Specifically, I investigate if heightened COVID-19-related social media use during a week corresponds with enhanced or diminished affect during that same week. 
-Additionally, I consider a longer interval for life satisfaction, examining whether increased COVID-19-related social media use over the course of a week influences one's life satisfaction at the week's end.
-This approach aims to uncover whether changes in social media consumption prompt simultaneous changes in affect and subsequent changes in life satisfaction.
-
-Furthermore, the study's main analyses implement this temporal interval through item wording rather than relying on lagged measures from previous waves.
+I consider a longer interval for life satisfaction, examining whether increased COVID-19-related social media use over the course of a week influences one's life satisfaction at the week's end. 
+<!-- This approach aims to uncover whether changes in social media consumption prompt simultaneous changes in affect and subsequent changes in life satisfaction. -->
+<!-- Furthermore, the study's main analyses implement this temporal interval through item wording rather than relying on lagged measures from previous waves.  -->
 Supplementary analyses extend this investigation to observe how media use might affect well-being one or four months later. 
-In all instances, confounding variables are methodically controlled for, enhancing the study's capacity for causal interpretation.
+<!-- In all instances, confounding variables are methodically controlled for, enhancing the study's capacity for causal interpretation. -->
 
@@
 The data were hierarchical, and responses were separately nested in participants and waves (i.e., participants and waves were implemented as random effects).
-Nesting in participants accounts for the longitudinal design. 
+Nesting in participants accounts for the within-person design. 
 Nesting in waves controls for general exogenous developments, such as general decreases in well-being in the population, for example due to lockdown measures.
+<!-- To illustrate, if general levels of well-being were reduced at one wave, for example due to lockdowns, this was accounted for in the model. -->
 Thus, there was no need additionally to control for specific phases or measures of the lockdown.
@@
 They included social media communication types and channels, separated into within and between-person factors, as well as stable and varying covariates.
-Between-person predictors are the predictors centered on the grand mean; within-person predictors are the predictors centered on the person's mean.
-Between-person predictors (which, measuring relations, are not of particular interest in this study) represent how the mean of one respondent differs from the mean of all the other respondents.
+<!-- Between-person predictors are the predictors centered on the grand mean; within-person predictors are the predictors centered on the person's mean. -->
+Between-person predictors (which, measuring relations, are not of particular interest in this study, but are reported online) represent how the mean of one respondent differs from the mean of all the other respondents.
 The within-person predictors represent how much a person at one specific wave differs from their own mean.
-For example, we could find that on Wave 3 a person used social media more than usual, while also experiencing more negative affect than usual.
+<!-- For example, we could find that on Wave 3 a person used social media more than usual, while also experiencing more negative affect than usual. -->
 All predictors were included simultaneously in each of the three models. 
+No collinearity of predictors was observed.
 
@@
 Mean scores were used for positive and negative affect.
+Although imputation has inherent drawbacks, we followed recent recommendations to impute missing data on larger scales [@endersAppliedMissingData2022].
 Missing responses were imputed using multiple imputation with predictive mean matching (five iterations, 30 data-sets), including categorical variables.
@@
 
-To contextualize the results, I conducted additional exploratory analyses. 
+To contextualize the results, I conducted additional exploratory analyses (see online materials)
 I reran the analyses (a) with additional not-preregistered covariates such as trust in media or government, (b) without covariates, (c) with single imputation, and (d) without imputation.
-For more information on the analyses, a complete documentation of the models and results, and all additional analyses, see [companion website](https://XMtRA.github.io/Austrian_Corona_Panel_Project).
+<!-- For more information on the analyses, a complete documentation of the models and results, and all additional analyses, see [companion website](https://XMtRA.github.io/Austrian_Corona_Panel_Project).
+-->
 
@@
 <!-- For the variables' means, range, and variance, see Table \@ref(tab:tab-descriptives). -->
-For a complete list of all items and item characteristics, see [companion website](https://XMtRA.github.io/Austrian_Corona_Panel_Project).
+<!-- For a complete list of all items and item characteristics, see [companion website](https://XMtRA.github.io/Austrian_Corona_Panel_Project). -->
 
@@
 Hence, for the main analyses data from all 34 waves were used.
-In the additional analyses I tested longer intervals, namely if changes in social media use were associated with changes in well-being either one month of four months later.
-For these analyzes I used the predictors from waves 1, 2, 8, 17, 23, and 28, to see if they predicted changes in well-being either one month or four months later.
+<!-- In the additional analyses I tested longer intervals, namely if changes in social media use were associated with changes in well-being either one month of four months later. -->
+<!-- For these analyzes I used the predictors from waves 1, 2, 8, 17, 23, and 28, to see if they predicted changes in well-being either one month or four months later.  -->
 
@@
 
-Looking at the variables from a descriptive perspective, aligned with set-point theory we can see that the level of all well-being measures were surprisingly stable during data collection (see Figure \@ref(fig:fig-descriptives)).
+Looking at the variables from a descriptive perspective, aligned with set-point theory we can see that the level of all well-being measures were comparatively stable during data collection (see Figure \@ref(fig:fig-descriptives)).
 COVID-19 related social media use, however, showed changes. 
@@
 Using the average values across all waves, which provides a stable picture of the general relations, I next looked at the correlations between social media use and well-being (see Figure \@ref(fig:fig-correlations)).
-Several interesting patterns emerged.
+<!-- Several interesting patterns emerged. -->
 In general, people who spend more time engaging with COVID-19 related content on social media reported reduced well-being.
@@
 
-The study's main hypothesis was that the causal effects of all types and channels of social media use on all facets of well-being would be trivial.
+<!-- The study's main hypothesis was that the causal effects of all types and channels of social media use on all facets of well-being would be trivial. -->
 Regarding the effects of different _communication types_ (i.e., reading, sharing, of posting about COVID-19 related content), all within-person effects fell completely within the a-priori defined null region (see Figure \@ref(fig:fig-within)).
@@
 That said, two effects differed statistically from zero.
-Respondents who used Twitter more frequently than usual to attain COVID-19 related news reported slightly higher levels of negative affect than usual (_b_ = `r data_tab_within %>% filter(dv == "Negative affect" & iv == "Twitter" & Variance == "within" & Analysis == "Publication") %>% select(estimate) %>% round(2)` [95% CI `r data_tab_within %>% filter(dv == "Negative affect" & iv == "Twitter" & Variance == "within" & Analysis == "Publication") %>% select(conf.low) %>% my_round("b")`, `r data_tab_within %>% filter(dv == "Negative affect" & iv == "Twitter" & Variance == "within" & Analysis == "Publication") %>% select(conf.high) %>% my_round("b")`]).
+Respondents who used Twitter more frequently than usual to attain COVID-19 related content reported slightly higher levels of negative affect than usual (_b_ = `r data_tab_within %>% filter(dv == "Negative affect" & iv == "Twitter" & Variance == "within" & Analysis == "Publication") %>% select(estimate) %>% round(2)` [95% CI `r data_tab_within %>% filter(dv == "Negative affect" & iv == "Twitter" & Variance == "within" & Analysis == "Publication") %>% select(conf.low) %>% my_round("b")`, `r data_tab_within %>% filter(dv == "Negative affect" & iv == "Twitter" & Variance == "within" & Analysis == "Publication") %>% select(conf.high) %>% my_round("b")`]).
 Likewise, respondents who used YouTube more frequently than usual for COVID-19 related issues reported slightly higher levels of negative affect than usual (_b_ = `r data_tab_within %>% filter(dv == "Negative affect" & iv == "YouTube" & Variance == "within" & Analysis == "Publication") %>% select(estimate) %>% round(2)` [95% CI `r data_tab_within %>% filter(dv == "Negative affect" & iv == "YouTube" & Variance == "within" & Analysis == "Publication") %>% select(conf.low) %>% my_round("b")`, `r data_tab_within %>% filter(dv == "Negative affect" & iv == "YouTube" & Variance == "within" & Analysis == "Publication") %>% select(conf.high) %>% my_round("b")`]).
@@
 The SESOI was b = |0.30| for life satisfaction and b = |0.15| for affect.
-Hence, all of the reported effects are not considered large enough to be meaningful.", cache=F, warning=F, message=F, fig.width=6, fig.height=4}
-fig_results_within
-# knitr::include_graphics("figures/fig_results_within.png")
+Hence, all of the reported effects are not considered large enough to be meaningful.", cache=F, warning=F, message=F, fig.width=6, fig.height=4, out.width="\\textwidth"}
+# fig_results_within
+knitr::include_graphics("figures/fig_results_within.png")
 ```
@@
 To contextualize the results reported above and to see if the study included any meaningful effects at all, I also looked at the effect sizes of the covariates.
-Because each variable featured different response options, which would require defining a SESOI for each variable, I hence report the results of the standardized scales, which allows for a better comparison across the differently scaled variables.
-Here, we can build on Cohen's convention that small effects begin at _r_ = |.10|.
+<!-- Because each variable featured different response options, which would require defining a SESOI for each variable, I hence report the results of the standardized scales, which allows for a better comparison across the differently scaled variables.
+-->
+I report the results of the standardized scales, which allows for a better comparison across the differently scaled variables.
+As a SESOI, we can build on Cohen's convention that small effects begin at _r_ = |.10|.
 
@@
 For instance, individuals consuming more COVID-19-related content reported slightly lower life satisfaction, somewhat reduced positive affect, and notably elevated negative affect compared to others.
-However, to explore if these between-person correlations translated into within-person effects, it was examined whether within-person changes in media consumption corresponded with within-person changes in well-being.
 
-Contrary to expectations, increased consumption of COVID-19 content did not significantly decrease well-being at a meaningful level.
+To explore if these between-person correlations translated into within-person effects, it was examined whether within-person changes in media consumption corresponded with within-person changes in well-being.
+As expected and aligned with the literature, increased consumption of COVID-19 content did not significantly decrease well-being at a meaningful level. 
 While several statistically significant effects emerged, their magnitudes were notably small. 
@@
 
-Further analysis demonstrated that factors such as health and physical activity, which would be expected to have substantial impacts on well-being, indeed showed significant influences.
-Additionally, extended assessments covering one and four months did not yield meaningful effects.
-In summary, the influence of COVID-19-related social media activity on well-being was not substantial.
-This counters popular concerns over excessive social media use during crises causing substantial well-being risks.
+<!-- Further analysis demonstrated that factors such as health and physical activity, which would be expected to have substantial impacts on well-being, indeed showed significant influences.  -->
+<!-- Additionally, extended assessments covering one and four months did not yield meaningful effects.  -->
+<!-- In summary, the influence of COVID-19-related social media activity on well-being was not substantial.  -->
+<!-- This counters popular concerns over social media use during crises causing substantial well-being risks. -->
 
@@
 These results align with prior research highlighting that social media usage is rather associated with elevated negative affect but not reduced life satisfaction [@meierComputermediatedCommunicationSocial2020a].
-The often extreme, negative, or aggressive tone of COVID-19 discussions on social media [@fanStigmatizationSocialMedia2020] could adversely impact more active authors.
-The negative effects may be explained by the consumption of negative and misleading information [@liYouTubeSourceInformation2020].
-
-Moreover, the study suggest that varying communication types and channels warrant separate analyses.
+<!-- The often extreme, negative, or aggressive tone of COVID-19 discussions on social media [@fanStigmatizationSocialMedia2020] could adversely impact more active authors.  -->
+<!-- The negative effects may be explained by the consumption of negative and misleading information [@liYouTubeSourceInformation2020]. -->
+The study suggests that varying communication types and channels warrant separate analyses. 
 Reading COVID-19 related content slightly reduced positive affect, while liking, sharing, and posting increased negative affect slightly. 
 Posting COVID-19-related comments increased both negative and positive affect slightly, while it reduced life satisfaction. 
-This suggests that posting generates stronger reactions, both positive and negative.
-Together, this challenges the assumption that active use is inherently beneficial and passive use detrimental [@valkenburgAssociationsActivePassive2022].
+<!-- This suggests that posting generates stronger reactions, both positive and negative.  -->
+Posting COVID-19 content was more negative for Generation Z, potentially reflecting broader negative effects of social media on this generation. 
+<!-- Together, this finding is aligned with recent work challenging the assumption that active use is inherently beneficial and passive use detrimental [@valkenburgAssociationsActivePassive2022].  -->
 Notably, communication channels exhibited differences. 
@@
 
-In conclusion, the study aligns with theoretical models and past research.
-It supports the notion of social media effects being small and contingent on communication type and channel [@meierComputermediatedCommunicationSocial2020a; @valkenburgDifferentialSusceptibilityMedia2013].
-Posting COVID-19 content was more negative for Generation Z, potentially reflecting broader negative effects of social media on this generation.
-Effects not being strongly negative also help explain why people spent so much time engaging with COVID-19 content on social media [@katzUsesGratificationsResearch1973; @zillmannMoodManagementCommunication1988].
+<!-- In conclusion, the study aligns with theoretical models and past research.  -->
+<!-- It supports the notion of social media effects being small and contingent on communication type and channel [@meierComputermediatedCommunicationSocial2020a; @valkenburgDifferentialSusceptibilityMedia2013].
+-->
+<!-- Effects not being strongly negative also help explain why people spent so much time engaging with COVID-19 content on social media [@katzUsesGratificationsResearch1973; @zillmannMoodManagementCommunication1988].
+-->
 
 Several limitations exist.
-While the study's focus on within-person effects enhances causal understanding, challenges exist related to additional relevant confounding variables not included here [@rohrerThinkingClearlyCorrelations2018], correct definition of the SESOI [@anvariUsingAnchorbasedMethods2021], and exact measurement of media use [@scharkowAccuracySelfreportedInternet2016].
+While the study's focus on within-person effects enhances causal understanding, challenges exist related to additional relevant confounding exogenous variables not included here [@rohrerThinkingClearlyCorrelations2018], correct definition of the SESOI [@anvariUsingAnchorbasedMethods2021], and exact measurement of media use [@scharkowAccuracySelfreportedInternet2016].
+The measures of social media use did not include interpersonal communication about COVID-19 related content, which might result in more positive outcomes [@frisonExploringRelationshipsDifferent2015].
 The study's results are applicable primarily to Western societies and may not hold true in different cultural contexts.
 
-The study's findings conclude that COVID-19-related social media activity minimally affects well-being, with other factors playing more substantial roles.
-In light of these minimal effects, concerns over COVID-19-related social media engagement on well-being appear to be unwarranted.
-When trying to improve well-being during a pandemic, instead of focusing on social media it seems more fruitful to address other, more pertinent societal problems related to health care, regular exercise, or psychological resilience.
+The study's findings conclude that COVID-19-related social media activity minimally affects well-being, with other factors such as health and physical activity playing more substantial roles. 
+In light of these minimal effects, concerns over COVID-19-related social media engagement on well-being could not be substantiated.
+<!-- When trying to improve well-being during a pandemic, instead of focusing on social media it seems more fruitful to address other, more pertinent societal problems related to health care, regular exercise, or psychological resilience.
+-->
 
@@
 
-# Competing Interests
-
-I declare no competing interests.
-
-# Supplementary Material
-
-All the stimuli, presentation materials, analysis scripts, and a reproducible version of the manuscript can be found on the companion website (https://XMtRA.github.io/Austrian_Corona_Panel_Project).
-
-# Data Accessibility Statement
-
-The data are shared on AUSSDA, see https://doi.org/10.11587/28KQNS.
-The data can only be used for scientific purposes.
-
 # Acknowledgements

### 2024-04-04 manuscript.tex
Substantial updates:  - Shortened manuscript  - Implemented peer-feedback

@@
 \fi
+% definitions for citeproc citations
+\NewDocumentCommand\citeproctext{}{}
+\NewDocumentCommand\citeproc{mm}{%
+  \begingroup\def\citeproctext{#2}\cite{#1}\endgroup}
+\makeatletter
+ % allow citations to break across lines
+ \let\@cite@ofmt\@firstofone
+ % avoid brackets around text for \cite:
+ \def\@biblabel#1{}
+ \def\@cite#1#2{{#1\if@tempswa , #2\fi}}
+\makeatother
 \newlength{\cslhangindent}
@@
 \setlength{\csllabelwidth}{3em}
-\newlength{\cslentryspacingunit} % times entry-spacing
-\setlength{\cslentryspacingunit}{\parskip}
-\newenvironment{CSLReferences}[2] % #1 hanging-ident, #2 entry spacing
- {% don't indent paragraphs
-  \setlength{\parindent}{0pt}
+\newenvironment{CSLReferences}[2] % #1 hanging-indent, #2 entry-spacing
+ {\begin{list}{}{%
+  \setlength{\itemindent}{0pt}
+  \setlength{\leftmargin}{0pt}
+  \setlength{\parsep}{0pt}
   % turn on hanging indent if param 1 is 1
   \ifodd #1
-  \let\oldpar\par
-  \def\par{\hangindent=\cslhangindent\oldpar}
+   \setlength{\leftmargin}{\cslhangindent}
+   \setlength{\itemindent}{-1\cslhangindent}
   \fi
   % set entry spacing
-  \setlength{\parskip}{#2\cslentryspacingunit}
- }%
- {}
+  \setlength{\itemsep}{#2\baselineskip}}}
+ {\end{list}}
 \usepackage{calc}
-\newcommand{\CSLBlock}[1]{#1\hfill\break}
-\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\csllabelwidth}{#1}}
-\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth - \csllabelwidth}{#1}\break}
+\newcommand{\CSLBlock}[1]{\hfill\break\parbox[t]{\linewidth}{\strut\ignorespaces#1\strut}}
+\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\csllabelwidth}{\strut#1\strut}}
+\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth - \csllabelwidth}{\strut#1\strut}}
 \newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}
@@
 
-% \usepackage{etoolbox}
 \makeatletter
-\patchcmd{\HyOrg@maketitle}
+\usepackage{etoolbox}
+\patchcmd{\maketitle}
   {\section{\normalfont\normalsize\abstractname}}
@@
   {}{\typeout{Failed to patch abstract.}}
-\patchcmd{\HyOrg@maketitle}
+\patchcmd{\maketitle}
   {\section{\protect\normalfont{\@title}}}
@@
 However, because social media are particularly engaging, many find it hard to disconnect, especially during times of crisis.
-Using data from the Austrian Corona Panel Project consisting of 3,485 participants from 34 waves, controlling for several stable and varying confounders, the results showed that COVID-19 related social media use did not meaningfully reduce well-being.
+In this preregistered study, I investigate whether using social media for COVID-19 related reasons affects psychological well-being.
+Using data from the Austrian Corona Panel Project consisting of 3,485 participants from 34 waves, this research question was analyzed using random effects within between models, controlling for several stable and varying confounders.
+The results showed that COVID-19 related social media use did not meaningfully reduce well-being.
 Other factors such as health, income, exercise, or internal locus of control showed larger and meaningful effects.
@@
 
-During the COVID-19 pandemic it was critical to stay informed regarding the latest developments.
-How dangerous is the virus?
-In what region is it spreading?
-How is it transmitted? What are the current safety regulations?
-To obtain relevant information, many people heavily relied on social media, with use being at an all time high (Statista, 2021).
-Some actually could not stop using social media to learn about COVID-19 related news.
-A new phenomenon termed ``doomscrolling'' emerged (Sharma et al., 2022).
-Many users were glued to their screens and found it hard to pursue other relevant activities such as working, taking a break, or even looking after their children.
-In the media it was hence increasingly asked whether using social media for COVID-19 related reasons would, next to all other stressors, create an additional burden on mental health (Sandstrom et al., 2021).
-Although research has begun addressing this question
-(e.g., Bendau et al., 2021; Eden et al., 2020; Sewall et al., 2021),
-it still largely unknown if COVID-19 related social media use during the pandemic has had a meaningful impact on well-being.
-This study hence aims to 1) reveal the effect of the different types and channels of social media use on individual well-being, 2) provide generalizable and robust results by analyzing a large-scale longitudinal data-set with 34 waves, and 3) determine the within-person causal effects by analyzing how changes in social media use lead to changes in well-being.
+\section{Research Transparency Statement}\label{research-transparency-statement}
 
-\hypertarget{understanding-well-being-and-media-use}{%
-\subsection{Understanding Well-being and Media Use}\label{understanding-well-being-and-media-use}}
+Conflicts of interest: The author declares no conflicts of interest.
+Funding: This research did not receive external funding.
+Artificial intelligence: Chat GPT was used for language polishing.
+Ethics: This research complies with the Declaration of Helsinki (2023), aside from the requirement to preregister human subjects research.
+Ethical review and approval was not required for the study in accordance with the local legislation and institutional requirements.
+The participants provided their written informed consent to participate in this study.
+Computational reproducibility: All analyses and the manuscript itself are completely reproducible (\url{https://XMtRA.github.io/Austrian_Corona_Panel_Project}).
 
-This study investigates how different \emph{facets} of well-being are affected by different \emph{types} and different \emph{channels} of communication (Meier \& Reinecke, 2020).
-Building on the typology of subjective well-being (Diener et al., 2018), three different well-being facets are analyzed: life satisfaction, positive affect, and negative affect.
-Because effects of social media depend on how they are used (Verduyn et al., 2015), I further distinguish three types of use and five popular channels.
-The types of use include reading, liking and sharing, and posting COVID-19 related content.
-In doing so, this study analyzes social media use focused on COVID-19 related content, which includes posting thoughts about the pandemic, reading posts and comments, or retweeting and liking COVID-19 related news.
-Liking and sharing are combined as they both represent low-threshold, platform-ingrained, easily quantifiable interactions.
+Preregistration: The hypotheses, the sample, the measures, the analyses, and the inference criteria (SESOI, \emph{p}-value) were preregistered on the Open Science Framework (\url{https://osf.io/87b24/?view_only=b2289b6fec214fa88ee75a18d45c18f3}).
+Because in this study data from an already existing data set was analyzed, the preregistration was done prior to accessing the data.
+In some cases, it was impossible to execute the analyses as originally planned, for example because some properties of the variables only became apparent when seeing the actual data.
+All changes are reported online (\url{https://xmtra.github.io/Austrian_Corona_Panel_Project/preregistration_changes.html}).
+Materials: All study materials are publicly available (\url{https://xmtra.github.io/Austrian_Corona_Panel_Project/items_english.html}).
+Data: All primary data are publicly available (\url{https://doi.org/10.11587/28KQNS}).
+Analysis scripts: All analysis scripts are publicly available (\url{https://XMtRA.github.io/Austrian_Corona_Panel_Project}).
+
+\newpage
+
+\section{Introduction}\label{introduction}
+
+Amidst the COVID-19 pandemic, staying informed became paramount, prompting heavy reliance on social media for updates, with use being at an all time high (Statista, 2021).
+The phenomenon of ``doomscrolling'' emerged as individuals struggled to disengage from COVID-19-related news (Buchanan et al., 2021; Sharma et al., 2022), sparking concerns about its impact on mental health (Sandstrom et al., 2021).
+While initial research began exploring this question (e.g., Bendau et al., 2021; Eden et al., 2020; Sewall et al., 2021), the extent of social media's influence on well-being remains largely unknown.
+This study aims to assess the effects of different social media usage patterns on individual well-being using a comprehensive longitudinal dataset spanning 34 waves, offering insights into within-person causal relationships.
+
+\subsection{Understanding Well-being and Media Use}\label{understanding-well-being-and-media-use}
+
+This study investigates how different \emph{facets} of subjective well-being are affected by different \emph{types} and different \emph{channels} of communication (Meier \& Reinecke, 2020).
+Building on the typology of subjective well-being---defined as ``referring to the various types of subjective evaluations of one's life, including both cognitive evaluations and affective
+feelings'' (Diener et al., 2018, p. 3)---three different well-being facets are analyzed: life satisfaction, positive affect, and negative affect.
+
+Social media is a broad term, encompassing subdimensions such as social networking sites (e.g., Facebook or Instagram), instant messengers (e.g., WhatsApp or Signal), or interactive video-platforms (e.g., YouTube or TikTok) (Carr \& Hayes, 2015).
+Effects of social media depend on how they are used: while purposeful, active, social use is generally related to positive outcomes, passive, non-purposeful and asocial use is related to negative outcomes (Frison \& Eggermont, 2015; Verduyn et al., 2022).
+In this study, I hence distinguish three types of use and five popular channels.
+The types of use include posting (active use), reading (passive use), and liking and sharing (low-threshold active use) COVID-19 related content.
 The five channels to be investigated are Facebook, Twitter, Instagram, WhatsApp, and YouTube, which at the time ranked among the most popular social media services in Austria.
 
-\hypertarget{social-media-effects-on-well-being}{%
-\subsection{Social Media Effects on Well-Being}\label{social-media-effects-on-well-being}}
+\subsection{Social Media Effects on Well-Being}\label{social-media-effects-on-well-being}
 
-How easily can well-being be affected by external influences?
-In general, according to the set-point theory, well-being is surprisingly stable (Lykken, 1999).
-Although specific events such as marriage or salary increase can have significant impacts on well-being, in most cases effects are short-term, with well-being after some time returning to prior levels (Diener et al., 2018).
-Specific factors such as unemployment, disability, or death, however, can cause long-term changes in well-being (Lucas, 2007).
-So although well-being can be affected by external events and factors, this does not happen easily.
-
-Can social media use be such a factor?
 Current literature overviews suggest that social media use on average does seem to decrease well-being (Meier \& Reinecke, 2020).
@@
 These small findings can be explained with the differential susceptibility of media effects model (Valkenburg \& Peter, 2013), which states that there is substantial \emph{variation} of media effects for individual users.
-Whereas for some users social media are more beneficial, for others they are more harmful.
-On average, however, and this is central for this study here, effects tend to be small (Valkenburg \& Peter, 2013).
 For example, in one study it was estimated that roughly one quarter of all users experienced negative effects, another quarter positive effects, while for the rest the effects were neutral (Beyens et al., 2021).
-Whether or not effects are positive or negative depend on (a) dispositional factors (e.g., personality, temperament, gender), (b) developmental factors (e.g., age, developmental tasks), (c) and social factors (e.g., environment, norms, upbringing).
-Finally, effects depend also on the content that is consumed.
-If the content is aligned with dispositions, developmental capacities, and converging contexts, effects tend to be stronger (Valkenburg \& Peter, 2013).
 
@@
 First, according to mood management theory (Zillmann, 1988), using media can affect people's moods.
-Use can be stimulating or overwhelming, relaxing or boring.
 After some time, users implicitly learn which media help them balance their mood and affect according to their own situational needs (Zillmann, 1988).
@@
 Users select those media that they expect to have a desired effect, for example on mood, knowledge, or entertainment.
-If those beneficial media effects are missing, people will spend their time elsewhere.
-And social media, in general, offer several beneficial effects, explaining why they are used that much.
+And social media, in general, offer several beneficial effects, explaining their ubiquitous use.
 They help find relevant information, maintain and foster relationships, express one's personality, and entertain oneself (Pelletier et al., 2020).
-In conclusion, because people spend so much time on social media consuming COVID-19 related content, according to both mood management theory and uses and gratifications theory this indirectly suggest that average effects on well-being are likely not particularly negative.
 
-\hypertarget{social-media-during-covid-19}{%
-\subsection{Social Media During COVID-19}\label{social-media-during-covid-19}}
+Third, and closely related, media use affords both both positive and negative mechanisms, which differently impact well-being outcomes.
+These include factors like social comparison, inspiration, social connectedness, social support, displacement of in-person activities, information management, information overload, misinformation, reputation building, romantic connection, coping resources, or emotional contagion (Clark et al., 2018; Hall \& Liu, 2022; Pelletier et al., 2020; Vanden Abeele, 2021; Verduyn et al., 2022).
+Given these differential mechanisms, strong one-sided effects appear less likely than nuanced and moderate effects.
+
+\subsection{Social Media During COVID-19}\label{social-media-during-covid-19}
 
@@
 Despite incorrect information, social media provide a vast platform for disseminating \emph{accurate and timely information} about COVID-19 (John Hopkins University, 2023).
-Access to reliable information can help people make informed decisions, alleviate uncertainties, and feel empowered during the pandemic.
 Social media platforms enable individuals to \emph{connect with others} who are experiencing similar challenges during the pandemic (Guazzini et al., 2022).
-Engaging in online communities and support groups can provide emotional support and create a network of like-minded individuals.
-Many mental health organizations and professionals utilize social media to share tips, strategies, and resources for \emph{maintaining mental well-being} during the pandemic (Twitter, 2020).
-Engaging with such content might help individuals prioritize their mental health and develop resilience during challenging times.
-Social media campaigns and initiatives can promote \emph{positive COVID-19 behaviors}, such as mask-wearing, physical distancing, hand hygiene, and vaccination (Athey et al., 2023; Hunt et al., 2022).
-Public health organizations and influencers leverage the power of social media to spread awareness and encourage responsible actions, contributing to public health efforts and fostering a sense of collective responsibility.
+Many mental health organizations and professionals utilize social media to share tips, strategies, and resources for \emph{maintaining mental well-being} during the pandemic.
+Social media campaigns and initiatives can promote \emph{positive COVID-19 behaviors}, such as mask-wearing, physical distancing, hand hygiene, and vaccination (Hunt et al., 2022), which ultimately benefit well-being.
 
@@
 Social media platforms can easily spread \emph{false or misleading information} about COVID-19 (Li et al., 2020).
-Due to the ease of sharing and the lack of fact-checking, inaccurate information can go viral and might cause confusion, anxiety, and panic among users.
-Constant exposure to COVID-19-related content on social media can lead to \emph{information overload} and contribute to heightened anxiety levels (J. Fan \& Smith, 2021).
-The rapid spread of news, updates, and opinions can be overwhelming and might exacerbate existing stress or fears about the pandemic (Sharma et al., 2022).
-Social media platforms are known for fostering negativity, with users sometimes engaging in \emph{cyberbullying and harassment}.
-Discussions around COVID-19 can become heated and polarized, leading to personal attacks and online conflicts.
-Such experiences threaten mental well-being and might contribute to feelings of distress and isolation.
+Constant exposure to COVID-19-related content on social media can lead to \emph{information overload} and contribute to heightened anxiety levels (Fan \& Smith, 2021).
+Discussions around COVID-19 are known for fostering negativity, with users sometimes engaging in \emph{cyberbullying and harassment}.
 Social media often showcase the highlights and accomplishments of others, encouraging \emph{social comparison} (Przybylski et al., 2013).
@@
 
-There is still little empirical research on how well-being is affected by social media use that is focused on COVID-19 specifically.
-Echoing the theoretical rationales outlines above, studies have yielded mixed results.
-Some studies found negative effects, indicating that excessive social media use for COVID-19 news led to compulsive behavior and increased stress levels, particularly due to upward social comparison (Stainback et al., 2020; Yue et al., 2022).
+Echoing the theoretical rationales outlines above, empirical studies have yielded mixed results.
+Some studies found negative effects, indicating that excessive social media use for COVID-19 content led to compulsive behavior and increased stress levels, particularly due to upward social comparison (Stainback et al., 2020).
 Individuals who relied on social media as their primary information source reported higher levels of anxiety and depression symptoms (Bendau et al., 2021).
-Similarly, increased COVID-19-related media consumption was associated with higher psychological distress.
+Doom-scrolling was associated with negative emotional experiences (Buchanan et al., 2021).
 On the other hand, some studies reported positive outcomes.
 Certain individuals experienced increased virtual community and social connectedness during the pandemic through social media, which contributed to their well-being (Guazzini et al., 2022).
-Additionally, engaging more on social media was associated with reduced feelings of loneliness (Latikka et al., 2022).
-Several studies reported mostly neutral effects of social media use on well-being indicators (Eden et al., 2020; Sewall et al., 2021).
-Overall, the literature demonstrates a mixed picture, highlighting both positive and negative effects of social media use focused on or during COVID-19 on well-being.
+Additionally, identification with social media networks was associated with reduced feelings of loneliness (Latikka et al., 2022).
+Several studies reported mostly neutral or dual effects of social media use on well-being indicators (Eden et al., 2020; Sewall et al., 2021).
 
 In conclusion, given these mixed empirical results, together with the observation that social media effects on well-being are very small in general, and that several plausible theoretical mechanisms exist for both positive and negative effects, I expect that COVID-19 related communication on social media should not be decidedly positive or negative.
-It seems most likely that both positive and negative coexist, but that on average using social media for COVID-19 related reasons should not have substantial effects on well-being.
 
@@
 
-\hypertarget{smallest-effect-size-of-interest}{%
-\subsection{Smallest Effect Size of Interest}\label{smallest-effect-size-of-interest}}
+\subsection{Smallest Effect Size of Interest}\label{smallest-effect-size-of-interest}
 
@@
 Determining what constitutes a minimally intriguing, nontrivial effect is a matter with normative implications, making it challenging to arrive at a definitive, singular consensus.
-Nonetheless, it remains both essential and beneficial to strive for a reasonable benchmark.
-With that in mind, I propose the following SESOI as a suitable reference point for this study:
+I propose the following SESOI as a suitable reference point for this study:
 
 \begin{quote}
-SESOI: If a heavy user of COVID-19 related social media news suddenly \emph{stops} using social media altogether, this should have a \emph{noticeable} impact on their overall well-being.
+SESOI: If a heavy user of COVID-19 related social media content suddenly \emph{stops} using social media altogether, this should have a \emph{noticeable} impact on their overall well-being.
 \end{quote}
 
-What does this mean practically and how can it be operationalized?
-In this study, COVID-19 related social media use was measured on a 5-point scale, ranging from 1 = \emph{never} to 5 = \emph{several times a day}.
+Put concretely, in this study COVID-19 related social media use was measured on a 5-point scale, ranging from 1 = \emph{never} to 5 = \emph{several times a day}.
 Thus, a change of four units in social media use (e.g., a complete stop) should correspond to a noticeable change in well-being.
@@
 So if satisfaction is measured on a 7-point scale, a four unit change in social media use should result in a one unit change in life satisfaction.
-In this study, life satisfaction was measured on an 11-point scale, and affect on a 5-point scale.
-Transposed to this scaling, the SESOI for life satisfaction is \emph{b} = ±.30, and for positive and negative affect \emph{b} = ±.15 (for more information, see online supplementary material).
+Life satisfaction was measured on an 11-point scale, and affect on a 5-point scale.
+Transposed to this scaling, the SESOI for life satisfaction is \emph{b} = ±.30, and for positive and negative affect \emph{b} = ±.15 (see online supplementary material).
 
-\hypertarget{method}{%
-\section{Method}\label{method}}
+\section{Method}\label{method}
 
-\hypertarget{preregistration}{%
-\subsection{Preregistration}\label{preregistration}}
+\subsection{Sample}\label{sample}
 
-The hypotheses, the sample, the measures, the analyses, and the inference criteria (SESOI, \emph{p}-value) were preregistered on the Open Science Framework, accessible here: \url{https://osf.io/87b24/?view_only=b2289b6fec214fa88ee75a18d45c18f3}.
-Because in this study I analyze data from an already existing large-scale data set, the preregistration was done prior to accessing the data.
-The preregistration was designed on the basis of the panel documentation online (Kittel et al., 2020).
-In some cases, it was impossible to execute the analyses as I had originally planned, for example because some properties of the variables only became apparent when seeing the actual data.
-The most relevant deviations are reported below, and a complete list of all changes can be found in the online \href{https://XMtRA.github.io/Austrian_Corona_Panel_Project}{companion website} (\url{https://XMtRA.github.io/Austrian_Corona_Panel_Project}).
-
-\hypertarget{sample}{%
-\subsection{Sample}\label{sample}}
-
-The data come from the Austrian Corona Panel Project (Kittel et al., 2020), which is a large-scale standalone panel study.
-The data are hosted on AUSSDA, are publicly available here (\url{https://doi.org/10.11587/28KQNS}), and consist of 34 waves.
-Participants were sampled from a pre-existing online access panel provided by the company Marketagent, Austria.
-Panel members were incentivized with 180 credit points for each wave of the study.
+The data come from the Austrian Corona Panel Project (Kittel et al., 2020), which is a large-scale standalone panel study, consisting of 34 waves.
 The study was conducted between March 2020 and February 2023.
@@
 Each wave consists of at least 1,500 respondents.
-Panel mortality was compensated through a continuous re-acquisition of new participants.
+Panel mortality (on average ca. 5 percent per wave) was compensated through a continuous re-acquisition of new participants (on average ca. 80 new respondents per wave).
 The sample size was \emph{N} = 3,641, with overall 123,794 observations.
-For an overview of the study set-up, see Figure \ref{fig:study-desc}.
-
-\begin{figure}
-\includegraphics[width=1\textwidth]{figures/fig_study_description} \caption{Overview of study set-up}\label{fig:study-desc}
-\end{figure}
 
@@
 All respondents needed to have access to the internet (via computer or mobile devices such as smartphones or tablets).
-Ethical review and approval was not required for the study in accordance with the local legislation and institutional requirements.
-The participants provided their written informed consent to participate in this study.
 The average age was 40 years, 49 percent were male, 14 percent had a University degree, and 5 percent were currently unemployed.
 
-\hypertarget{data-analysis}{%
-\subsection{Data Analysis}\label{data-analysis}}
+\subsection{Data Analysis}\label{data-analysis}
 
@@
 Here, the null-region lies between -.30 and +.30.
-If the 95\% confidence interval falls completely within the null-region (e.g., \emph{b} = -.05, {[}95\% CI: -.15, .05{]}), the hypothesis that the effect is trivial is supported.
 If the confidence interval falls completely outside of the null-region (e.g., \emph{b} = -.40, {[}95\% CI: -.45, -.35{]}), the hypothesis is rejected and the existence of a meaningful negative effect is supported.
-If the confidence interval and the null region overlap (e.g., \emph{b} = -.30, {[}95\% CI: -.35, -.25{]}), the hypothesis is not supported and the results are considered inconclusive, while a meaningful positive effect is rejected.
-If the confidence interval exceeds both sides of the null region (e.g., \emph{b} = -.025, {[}95\% CI: -.40, .35{]}), the hypothesis is not supported and judgement is suspended.
 For an illustration, see Figure \ref{fig:sesoi}.
@@
 
-\hypertarget{causality}{%
-\subsubsection{Causality}\label{causality}}
+\subsubsection{Causality}\label{causality}
 
-When analyzing causality through longitudinal designs, it is important to address several critical aspects.
-An inherent requirement of analyzing causal effects within non-experimental designs is adopting an internal perspective, focusing on \emph{within-person effects} (Hamaker, 2014).
-This entails evaluating how alterations in an individual's media consumption directly impact their well-being.
-Comparisons between individuals, as observed in longitudinal data, lack the granularity needed to understand these causal dynamics (Hamaker, 2014).
-Consequently, this study exclusively investigates within-person effects to shed light on this intricate relationship.
+Analyzing causal effects within non-experimental designs requires adopting an internal perspective, focusing on \emph{within-person effects} (Hamaker, 2014).
+This entails evaluating how alterations in an individual's media consumption directly impact changes in their own well-being.
+Consequently, this study exclusively investigates within-person effects.
 
@@
 However, a cautious balance must be maintained by not controlling for variables that mediate the relationship (Rohrer, 2018), as their inclusion could distort the assessment of the causal effect.
-This study therefore incorporated several control variables.
-These encompass gender, age, education, place of birth and parental place of birth in Austria, Vienna residency, consumption of text-based and video-based news, household characteristics, health status, living space features, employment-related factors, income, outdoor activities, risk propensity, and locus of control (Eger \& Maridal, 2015).
-These variables have demonstrated connections with both social media use and well-being and are not mediators.
+This study therefore incorporated several control variables (see below), which have demonstrated connections with both social media use and well-being and are likely not mediators (Eger \& Maridal, 2015).
 
-A fundamental prerequisite for establishing causality is determining a \emph{plausible temporal interval} (Dormann \& Griffin, 2015).
+Establishing causality necessitates determining a \emph{plausible temporal interval} (Rohrer \& Murayama, 2023).
 For instance, fluctuations in positive and negative affect call for shorter intervals, while the more enduring nature of life satisfaction implies longer intervals (Dienlin \& Johannes, 2020).
-
 In this study, I examine the linkage between changes in social media use and changes in affect within the same week.
 Specifically, I investigate if heightened COVID-19-related social media use during a week corresponds with enhanced or diminished affect during that same week.
-Additionally, I consider a longer interval for life satisfaction, examining whether increased COVID-19-related social media use over the course of a week influences one's life satisfaction at the week's end.
-This approach aims to uncover whether changes in social media consumption prompt simultaneous changes in affect and subsequent changes in life satisfaction.
+I consider a longer interval for life satisfaction, examining whether increased COVID-19-related social media use over the course of a week influences one's life satisfaction at the week's end.
+Supplementary analyses extend this investigation to observe how media use might affect well-being one or four months later.
 
-Furthermore, the study's main analyses implement this temporal interval through item wording rather than relying on lagged measures from previous waves.
-Supplementary analyses extend this investigation to observe how media use might affect well-being one or four months later.
-In all instances, confounding variables are methodically controlled for, enhancing the study's capacity for causal interpretation.
-
-\hypertarget{statistical-model}{%
-\subsubsection{Statistical model}\label{statistical-model}}
+\subsubsection{Statistical model}\label{statistical-model}
 
@@
 The data were hierarchical, and responses were separately nested in participants and waves (i.e., participants and waves were implemented as random effects).
-Nesting in participants accounts for the longitudinal design.
+Nesting in participants accounts for the within-person design.
 Nesting in waves controls for general exogenous developments, such as general decreases in well-being in the population, for example due to lockdown measures.
@@
 They included social media communication types and channels, separated into within and between-person factors, as well as stable and varying covariates.
-Between-person predictors are the predictors centered on the grand mean; within-person predictors are the predictors centered on the person's mean.
-Between-person predictors (which, measuring relations, are not of particular interest in this study) represent how the mean of one respondent differs from the mean of all the other respondents.
+Between-person predictors (which, measuring relations, are not of particular interest in this study, but are reported online) represent how the mean of one respondent differs from the mean of all the other respondents.
 The within-person predictors represent how much a person at one specific wave differs from their own mean.
-For example, we could find that on Wave 3 a person used social media more than usual, while also experiencing more negative affect than usual.
 All predictors were included simultaneously in each of the three models.
+No collinearity of predictors was observed.
 
@@
 Mean scores were used for positive and negative affect.
+Although imputation has inherent drawbacks, we followed recent recommendations to impute missing data on larger scales (Enders, 2022).
 Missing responses were imputed using multiple imputation with predictive mean matching (five iterations, 30 data-sets), including categorical variables.
@@
 
-To contextualize the results, I conducted additional exploratory analyses.
+To contextualize the results, I conducted additional exploratory analyses (see online materials)
 I reran the analyses (a) with additional not-preregistered covariates such as trust in media or government, (b) without covariates, (c) with single imputation, and (d) without imputation.
-For more information on the analyses, a complete documentation of the models and results, and all additional analyses, see \href{https://XMtRA.github.io/Austrian_Corona_Panel_Project}{companion website}.
 
-\hypertarget{measures}{%
-\subsection{Measures}\label{measures}}
+\subsection{Measures}\label{measures}
 
-For a complete list of all items and item characteristics, see \href{https://XMtRA.github.io/Austrian_Corona_Panel_Project}{companion website}.
-
-\hypertarget{well-being}{%
-\subsubsection{Well-being}\label{well-being}}
+\subsubsection{Well-being}\label{well-being}
 
@@
 The response options were 1 (\emph{never}), 2 (\emph{on some days}), 3 (\emph{several times per week}), 4 (\emph{almost every day}), and 5 (\emph{daily}).
-The scale showed good factorial fit, \(\chi^2\)(66) = 69.42, \emph{p} = .363, CFI = 1.00, RMSEA \textless{} .01, 90\% CI {[}\textless{} .01, .02{]}, SRMR = .01.
+The scale showed good factorial fit, \(\chi^2\)(66) = 69.42, \textit{p} = .363, cfi = 1.00, rmsea \textless{} .01, 90\% CI {[}\textless{} .01, .02{]}, srmr = .01.
 Reliability was high, \(\omega\) = .85.
@@
 The response options were 1 (\emph{never}), 2 (\emph{on some days}), 3 (\emph{several times per week}), 4 (\emph{almost every day}), and 5 (\emph{daily}).
-The scale showed good factorial fit, \(\chi^2\)(471) = 4012.14, \emph{p} \textless{} .001, CFI = .98, RMSEA = .07, 90\% CI {[}.07, .08{]}, SRMR = .03.
+The scale showed good factorial fit, \(\chi^2\)(471) = 4012.14, \textit{p} \textless{} .001, cfi = .98, rmsea = .07, 90\% CI {[}.07, .08{]}, srmr = .03.
 Reliability was high, \(\omega\) = .91.
@@
 
-\hypertarget{covid-19-related-social-media-use}{%
-\subsubsection{COVID-19 related social media use}\label{covid-19-related-social-media-use}}
+\subsubsection{COVID-19 related social media use}\label{covid-19-related-social-media-use}
 
@@
 Hence, for the main analyses data from all 34 waves were used.
-In the additional analyses I tested longer intervals, namely if changes in social media use were associated with changes in well-being either one month of four months later.
-For these analyzes I used the predictors from waves 1, 2, 8, 17, 23, and 28, to see if they predicted changes in well-being either one month or four months later.
 
-\hypertarget{control-variables}{%
-\subsubsection{Control variables}\label{control-variables}}
+\subsubsection{Control variables}\label{control-variables}
 
@@
 
-\hypertarget{results}{%
-\section{Results}\label{results}}
+\section{Results}\label{results}
 
-\hypertarget{descriptive-analyses}{%
-\subsection{Descriptive Analyses}\label{descriptive-analyses}}
+\subsection{Descriptive Analyses}\label{descriptive-analyses}
 
-Looking at the variables from a descriptive perspective, aligned with set-point theory we can see that the level of all well-being measures were surprisingly stable during data collection (see Figure \ref{fig:fig-descriptives}).
+Looking at the variables from a descriptive perspective, aligned with set-point theory we can see that the level of all well-being measures were comparatively stable during data collection (see Figure \ref{fig:fig-descriptives}).
 COVID-19 related social media use, however, showed changes.
@@
 Using the average values across all waves, which provides a stable picture of the general relations, I next looked at the correlations between social media use and well-being (see Figure \ref{fig:fig-correlations}).
-Several interesting patterns emerged.
 In general, people who spend more time engaging with COVID-19 related content on social media reported reduced well-being.
@@
 
-\hypertarget{preregistered-analyses}{%
-\subsection{Preregistered Analyses}\label{preregistered-analyses}}
+\subsection{Preregistered Analyses}\label{preregistered-analyses}
 
-The study's main hypothesis was that the causal effects of all types and channels of social media use on all facets of well-being would be trivial.
 Regarding the effects of different \emph{communication types} (i.e., reading, sharing, of posting about COVID-19 related content), all within-person effects fell completely within the a-priori defined null region (see Figure \ref{fig:fig-within}).
@@
 That said, two effects differed statistically from zero.
-Respondents who used Twitter more frequently than usual to attain COVID-19 related news reported slightly higher levels of negative affect than usual (\emph{b} = 0.02 {[}95\% CI 0.01, 0.04{]}).
+Respondents who used Twitter more frequently than usual to attain COVID-19 related content reported slightly higher levels of negative affect than usual (\emph{b} = 0.02 {[}95\% CI 0.01, 0.04{]}).
 Likewise, respondents who used YouTube more frequently than usual for COVID-19 related issues reported slightly higher levels of negative affect than usual (\emph{b} = 0.01 {[}95\% CI \textless{} 0.01, 0.02{]}).
@@
 \begin{figure}
-\centering
-\includegraphics{manuscript_files/figure-latex/fig-within-1.pdf}
-\caption{\label{fig:fig-within}Unstandardized within-person effects of COVID-19 related social media use on well-being.
+\includegraphics[width=\textwidth]{figures/fig_results_within} \caption{Unstandardized within-person effects of COVID-19 related social media use on well-being.
 Note.
-The SESOI was b = \textbar0.30\textbar{} for life satisfaction and b = \textbar0.15\textbar{} for affect.
-Hence, all of the reported effects are not considered large enough to be meaningful.}
+The SESOI was b = |0.30| for life satisfaction and b = |0.15| for affect.
+Hence, all of the reported effects are not considered large enough to be meaningful.}\label{fig:fig-within}
 \end{figure}
 
-\hypertarget{exploratory-analyses}{%
-\subsection{Exploratory Analyses}\label{exploratory-analyses}}
+\subsection{Exploratory Analyses}\label{exploratory-analyses}
 
 To contextualize the results reported above and to see if the study included any meaningful effects at all, I also looked at the effect sizes of the covariates.
-Because each variable featured different response options, which would require defining a SESOI for each variable, I hence report the results of the standardized scales, which allows for a better comparison across the differently scaled variables.
-Here, we can build on Cohen's convention that small effects begin at \emph{r} = \textbar.10\textbar.
+I report the results of the standardized scales, which allows for a better comparison across the differently scaled variables.
+As a SESOI, we can build on Cohen's convention that small effects begin at \emph{r} = \textbar.10\textbar.
 
@@
 
-\hypertarget{discussion}{%
-\section{Discussion}\label{discussion}}
+\section{Discussion}\label{discussion}
 
@@
 For instance, individuals consuming more COVID-19-related content reported slightly lower life satisfaction, somewhat reduced positive affect, and notably elevated negative affect compared to others.
-However, to explore if these between-person correlations translated into within-person effects, it was examined whether within-person changes in media consumption corresponded with within-person changes in well-being.
 
-Contrary to expectations, increased consumption of COVID-19 content did not significantly decrease well-being at a meaningful level.
+To explore if these between-person correlations translated into within-person effects, it was examined whether within-person changes in media consumption corresponded with within-person changes in well-being.
+As expected and aligned with the literature, increased consumption of COVID-19 content did not significantly decrease well-being at a meaningful level.
 While several statistically significant effects emerged, their magnitudes were notably small.
@@
 The effects, although statistically significant, fell within a predefined range considered insignificant.
-
-Further analysis demonstrated that factors such as health and physical activity, which would be expected to have substantial impacts on well-being, indeed showed significant influences.
-Additionally, extended assessments covering one and four months did not yield meaningful effects.
-In summary, the influence of COVID-19-related social media activity on well-being was not substantial.
-This counters popular concerns over excessive social media use during crises causing substantial well-being risks.
 
@@
 These results align with prior research highlighting that social media usage is rather associated with elevated negative affect but not reduced life satisfaction (Meier \& Reinecke, 2020).
-The often extreme, negative, or aggressive tone of COVID-19 discussions on social media (L.
-Fan et al., 2020) could adversely impact more active authors.
-The negative effects may be explained by the consumption of negative and misleading information (Li et al., 2020).
-
-Moreover, the study suggest that varying communication types and channels warrant separate analyses.
+The study suggests that varying communication types and channels warrant separate analyses.
 Reading COVID-19 related content slightly reduced positive affect, while liking, sharing, and posting increased negative affect slightly.
 Posting COVID-19-related comments increased both negative and positive affect slightly, while it reduced life satisfaction.
-This suggests that posting generates stronger reactions, both positive and negative.
-Together, this challenges the assumption that active use is inherently beneficial and passive use detrimental (Valkenburg et al., 2022).
+Posting COVID-19 content was more negative for Generation Z, potentially reflecting broader negative effects of social media on this generation.
 Notably, communication channels exhibited differences.
@@
 
-In conclusion, the study aligns with theoretical models and past research.
-It supports the notion of social media effects being small and contingent on communication type and channel (Meier \& Reinecke, 2020; Valkenburg \& Peter, 2013).
-Posting COVID-19 content was more negative for Generation Z, potentially reflecting broader negative effects of social media on this generation.
-Effects not being strongly negative also help explain why people spent so much time engaging with COVID-19 content on social media (Katz et al., 1973; Zillmann, 1988).
-
 Several limitations exist.
-While the study's focus on within-person effects enhances causal understanding, challenges exist related to additional relevant confounding variables not included here (Rohrer, 2018), correct definition of the SESOI (Anvari \& Lakens, 2021), and exact measurement of media use (Scharkow, 2016).
+While the study's focus on within-person effects enhances causal understanding, challenges exist related to additional relevant confounding exogenous variables not included here (Rohrer, 2018), correct definition of the SESOI (Anvari \& Lakens, 2021), and exact measurement of media use (Scharkow, 2016).
+The measures of social media use did not include interpersonal communication about COVID-19 related content, which might result in more positive outcomes (Frison \& Eggermont, 2015).
 The study's results are applicable primarily to Western societies and may not hold true in different cultural contexts.
 
-The study's findings conclude that COVID-19-related social media activity minimally affects well-being, with other factors playing more substantial roles.
-In light of these minimal effects, concerns over COVID-19-related social media engagement on well-being appear to be unwarranted.
-When trying to improve well-being during a pandemic, instead of focusing on social media it seems more fruitful to address other, more pertinent societal problems related to health care, regular exercise, or psychological resilience.
+The study's findings conclude that COVID-19-related social media activity minimally affects well-being, with other factors such as health and physical activity playing more substantial roles.
+In light of these minimal effects, concerns over COVID-19-related social media engagement on well-being could not be substantiated.
 
@@
 
-\hypertarget{references}{%
-\section{References}\label{references}}
+\section{References}\label{references}
 
-\hypertarget{refs}{}
+\phantomsection\label{refs}
 \begin{CSLReferences}{1}{0}
-\leavevmode\vadjust pre{\hypertarget{ref-anvariUsingAnchorbasedMethods2021}{}}%
+\bibitem[\citeproctext]{ref-anvariUsingAnchorbasedMethods2021}
 Anvari, F., \& Lakens, D.
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-atheyDigitalPublicHealth2023}{}}%
-Athey, S., Grabarz, K., Luca, M., \& Wernerfelt, N.
-(2023).
-Digital public health interventions at scale: {The} impact of social media advertising on beliefs and outcomes related to {COVID} vaccines.
-\emph{Proceedings of the National Academy of Sciences}, \emph{120}(5), e2208110120.
-\url{https://doi.org/10.1073/pnas.2208110120}
-
-\leavevmode\vadjust pre{\hypertarget{ref-bellFixedRandomEffects2019}{}}%
+\bibitem[\citeproctext]{ref-bellFixedRandomEffects2019}
 Bell, A., Fairbrother, M., \& Jones, K.
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-bendauAssociationsCOVID19Related2021}{}}%
+\bibitem[\citeproctext]{ref-bendauAssociationsCOVID19Related2021}
 Bendau, A., Petzold, M.
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-beyensSocialMediaUse2021}{}}%
+\bibitem[\citeproctext]{ref-beyensSocialMediaUse2021}
 Beyens, I., Pouwels, J.
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-dienerAdvancesOpenQuestions2018}{}}%
+\bibitem[\citeproctext]{ref-buchananBriefExposureSocial2021}
+Buchanan, K., Aknin, L.
+B., Lotun, S., \& Sandstrom, G.
+M.
+(2021).
+Brief exposure to social media during the {COVID-19} pandemic: {Doom-scrolling} has negative emotional consequences, but kindness-scrolling does not.
+\emph{PLOS ONE}, \emph{16}(10), e0257728.
+\url{https://doi.org/10.1371/journal.pone.0257728}
+
+\bibitem[\citeproctext]{ref-carrSocialMediaDefining2015}
+Carr, C.
+T., \& Hayes, R.
+A.
+(2015).
+Social {Media}: {Defining}, {Developing}, and {Divining}.
+\emph{Atlantic Journal of Communication}, \emph{23}(1), 46--65.
+\url{https://doi.org/10.1080/15456870.2015.972282}
+
+\bibitem[\citeproctext]{ref-clarkSocialNetworkSites2018}
+Clark, J.
+L., Algoe, S.
+B., \& Green, M.
+C.
+(2018).
+Social {Network Sites} and {Well-Being}: {The Role} of {Social Connection}.
+\emph{Current Directions in Psychological Science}, \emph{27}(1), 32--37.
+\url{https://doi.org/10.1177/0963721417730833}
+
+\bibitem[\citeproctext]{ref-dienerAdvancesOpenQuestions2018}
 Diener, E., Lucas, R.
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-dienesUsingBayesGet2014}{}}%
+\bibitem[\citeproctext]{ref-dienesUsingBayesGet2014}
 Dienes, Z. (2014). Using {Bayes} to get the most out of non-significant results. \emph{Frontiers in Psychology}, \emph{5}. \url{https://doi.org/10.3389/fpsyg.2014.00781}
 
-\leavevmode\vadjust pre{\hypertarget{ref-dienlinImpactDigitalTechnology2020}{}}%
+\bibitem[\citeproctext]{ref-dienlinImpactDigitalTechnology2020}
 Dienlin, T., \& Johannes, N.
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-dormannOptimalTimeLags2015}{}}%
-Dormann, C., \& Griffin, M. A. (2015). Optimal time lags in panel studies. \emph{Psychological Methods}, \emph{20}(4), 489--505. \url{https://doi.org/10.1037/met0000041}
-
-\leavevmode\vadjust pre{\hypertarget{ref-edenMediaCopingCOVID192020}{}}%
+\bibitem[\citeproctext]{ref-edenMediaCopingCOVID192020}
 Eden, A.
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-egerStatisticalMetaanalysisWellbeing2015}{}}%
+\bibitem[\citeproctext]{ref-egerStatisticalMetaanalysisWellbeing2015}
 Eger, R. J., \& Maridal, J. H. (2015). A statistical meta-analysis of the wellbeing literature. \emph{International Journal of Wellbeing}, \emph{5}(2), 45--74. \url{https://doi.org/10.5502/ijw.v5i2.4}
 
-\leavevmode\vadjust pre{\hypertarget{ref-fanInformationOverloadWellbeing2021}{}}%
+\bibitem[\citeproctext]{ref-endersAppliedMissingData2022}
+Enders, C. K. (2022). \emph{Applied missing data analysis} (Second Edition). The Guilford Press.
+
+\bibitem[\citeproctext]{ref-fanInformationOverloadWellbeing2021}
 Fan, J., \& Smith, A. P. (2021). Information overload, wellbeing and {COVID-19}: {A} survey in {China}. \emph{Behavioral Sciences}, \emph{11}(5), 62. \url{https://doi.org/10.3390/bs11050062}
 
-\leavevmode\vadjust pre{\hypertarget{ref-fanStigmatizationSocialMedia2020}{}}%
-Fan, L., Yu, H., \& Yin, Z.
-(2020).
-Stigmatization in social media: {Documenting} and analyzing hate speech for {\textsc{COVID}} -19 on {Twitter}.
-\emph{Proceedings of the Association for Information Science and Technology}, \emph{57}(1).
-\url{https://doi.org/10.1002/pra2.313}
+\bibitem[\citeproctext]{ref-frisonExploringRelationshipsDifferent2015}
+Frison, E., \& Eggermont, S.
+(2015).
+Exploring the relationships between different types of {Facebook} use, perceived online social support, and adolescents' depressed mood.
+\emph{Social Science Computer Review}, 1--19.
+\url{https://doi.org/10.1177/0894439314567449}
 
-\leavevmode\vadjust pre{\hypertarget{ref-guazziniSecondWaveAnalysis2022}{}}%
+\bibitem[\citeproctext]{ref-guazziniSecondWaveAnalysis2022}
 Guazzini, A., Pesce, A., Marotta, L., \& Duradoni, M.
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-hamakerWhyResearchersShould2014}{}}%
+\bibitem[\citeproctext]{ref-hallSocialMediaUse2022}
+Hall, J. A., \& Liu, D. (2022). Social media use, social displacement, and well-being. \emph{Current Opinion in Psychology}, \emph{46}, 101339. \url{https://doi.org/10.1016/j.copsyc.2022.101339}
+
+\bibitem[\citeproctext]{ref-hamakerWhyResearchersShould2014}
 Hamaker, E.
@@
 Csikszentmihalyi (Eds.), \emph{Handbook of research methods for studying daily life} (Paperback ed.).
-{Guilford}.
+Guilford.
 
-\leavevmode\vadjust pre{\hypertarget{ref-huntSocialMediaBased2022}{}}%
+\bibitem[\citeproctext]{ref-huntSocialMediaBased2022}
 Hunt, I.
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-johnhopkinsuniversityCOVID19Map2023}{}}%
+\bibitem[\citeproctext]{ref-johnhopkinsuniversityCOVID19Map2023}
 John Hopkins University. (2023). {COVID-19 Map}. In \emph{Johns Hopkins Coronavirus Resource Center}. https://coronavirus.jhu.edu/map.html.
 
-\leavevmode\vadjust pre{\hypertarget{ref-katzUsesGratificationsResearch1973}{}}%
+\bibitem[\citeproctext]{ref-katzUsesGratificationsResearch1973}
 Katz, E., Blumler, J. G., \& Gurevitch, M. (1973). Uses and {Gratifications Research}. \emph{Public Opinion Quarterly}, \emph{37}(4), 509. \url{https://doi.org/10.1086/268109}
 
-\leavevmode\vadjust pre{\hypertarget{ref-kittelAustrianCoronaPanel2020}{}}%
+\bibitem[\citeproctext]{ref-kittelAustrianCoronaPanel2020}
 Kittel, B., Kritzinger, S., Boomgaarden, H., Prainsack, B., Eberl, J.-M., Kalleitner, F., Lebernegg, N.
@@
 \emph{Austrian {Corona Panel Project} ({SUF} edition)}.
-{AUSSDA}.
+AUSSDA.
 \url{https://doi.org/10.11587/28KQNS}
 
-\leavevmode\vadjust pre{\hypertarget{ref-latikkaLonelinessPsychologicalDistress2022}{}}%
+\bibitem[\citeproctext]{ref-latikkaLonelinessPsychologicalDistress2022}
 Latikka, R., Koivula, A., Oksa, R., Savela, N., \& Oksanen, A.
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-liYouTubeSourceInformation2020}{}}%
+\bibitem[\citeproctext]{ref-liYouTubeSourceInformation2020}
 Li, H.
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-lucasAdaptationSetpointModel2007}{}}%
-Lucas, R.
-E.
-(2007).
-Adaptation and the set-point model of subjective well-being.
-\emph{Current Directions in Psychological Science}, \emph{16}(2), 75--79.
-\url{https://doi.org/10.1111/j.1467-8721.2007.00479.x}
-
-\leavevmode\vadjust pre{\hypertarget{ref-lykkenHappinessWhatStudies1999}{}}%
-Lykken, D. T. (1999). \emph{Happiness: {What} studies on twins show us about nature, nurture, and the happiness set-point}. {Golden Books}.
-
-\leavevmode\vadjust pre{\hypertarget{ref-marcianoDynamicsAdolescentsSmartphone2022}{}}%
+\bibitem[\citeproctext]{ref-marcianoDynamicsAdolescentsSmartphone2022}
 Marciano, L., Driver, C.
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-meierComputermediatedCommunicationSocial2020a}{}}%
+\bibitem[\citeproctext]{ref-meierComputermediatedCommunicationSocial2020a}
 Meier, A., \& Reinecke, L.
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-normanInterpretationChangesHealthrelated2003}{}}%
+\bibitem[\citeproctext]{ref-normanInterpretationChangesHealthrelated2003}
 Norman, G., Sloan, J., \& Wyrwich, K.
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-orbenWindowsDevelopmentalSensitivity2022}{}}%
+\bibitem[\citeproctext]{ref-orbenWindowsDevelopmentalSensitivity2022}
 Orben, A., Przybylski, A.
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-pelletierOneSizeDoesn2020}{}}%
+\bibitem[\citeproctext]{ref-pelletierOneSizeDoesn2020}
 Pelletier, M.
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-przybylskiMotivationalEmotionalBehavioral2013}{}}%
+\bibitem[\citeproctext]{ref-przybylskiMotivationalEmotionalBehavioral2013}
 Przybylski, A.
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-rohrerThinkingClearlyCorrelations2018}{}}%
+\bibitem[\citeproctext]{ref-rohrerThinkingClearlyCorrelations2018}
 Rohrer, J.
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-rohrerTheseAreNot2023}{}}%
+\bibitem[\citeproctext]{ref-rohrerTheseAreNot2023}
 Rohrer, J.
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-sandstromDoomscrollingCOVIDNews2021}{}}%
+\bibitem[\citeproctext]{ref-sandstromDoomscrollingCOVIDNews2021}
 Sandstrom, G., Buchanan, K., Aknin, L., \& Lotun, S.
 (2021).
-Doomscrolling {COVID} news takes an emotional toll \textendash{} here's how to make your social media a happier place.
+Doomscrolling {COVID} news takes an emotional toll -- here's how to make your social media a happier place.
 In \emph{The Conversation}.
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-scharkowAccuracySelfreportedInternet2016}{}}%
+\bibitem[\citeproctext]{ref-scharkowAccuracySelfreportedInternet2016}
 Scharkow, M.
 (2016).
-The accuracy of self-reported {Internet} use\textemdash{{A}} validation study using client log data.
+The accuracy of self-reported {Internet} use---{A} validation study using client log data.
 \emph{Communication Methods and Measures}, \emph{10}(1), 13--27.
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-sewallObjectivelyMeasuredDigital2021}{}}%
+\bibitem[\citeproctext]{ref-sewallObjectivelyMeasuredDigital2021}
 Sewall, C.
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-sharmaDarkEndTunnel2022}{}}%
+\bibitem[\citeproctext]{ref-sharmaDarkEndTunnel2022}
 Sharma, B., Lee, S.
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-stainbackCOVID1924News2020}{}}%
+\bibitem[\citeproctext]{ref-stainbackCOVID1924News2020}
 Stainback, K., Hearne, B.
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-statistaAverageDailyTime2021}{}}%
+\bibitem[\citeproctext]{ref-statistaAverageDailyTime2021}
 Statista.
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-twitterTalkingMentalHealthAwarenessTwitter2020}{}}%
-Twitter. (2020). \emph{Talking \#{MentalHealthAwareness} on {Twitter} with our global partners}. https://blog.twitter.com/en\_us/topics/company/2020/talking-mental-health-awareness-on-twitter.
-
-\leavevmode\vadjust pre{\hypertarget{ref-valkenburgDifferentialSusceptibilityMedia2013}{}}%
+\bibitem[\citeproctext]{ref-valkenburgDifferentialSusceptibilityMedia2013}
 Valkenburg, P. M., \& Peter, J. (2013). The differential susceptibility to media effects model. \emph{Journal of Communication}, \emph{63}(2), 221--243. \url{https://doi.org/10.1111/jcom.12024}
 
-\leavevmode\vadjust pre{\hypertarget{ref-valkenburgAssociationsActivePassive2022}{}}%
-Valkenburg, P.
-M., van Driel, I.
-I., \& Beyens, I.
+\bibitem[\citeproctext]{ref-vandenabeeleDigitalWellbeingDynamic2021}
+Vanden Abeele, M. M. P. (2021). Digital {Wellbeing} as a {Dynamic Construct}. \emph{Communication Theory}, \emph{31}(4), 932--955. \url{https://doi.org/10.1093/ct/qtaa024}
+
+\bibitem[\citeproctext]{ref-verduynSocialNetworkingSites2022}
+Verduyn, P., Gugushvili, N., \& Kross, E.
 (2022).
-The associations of active and passive social media use with well-being: {A} critical scoping review.
-\emph{New Media \& Society}, \emph{24}(2), 530--549.
-\url{https://doi.org/10.1177/14614448211065425}
+Do {Social Networking Sites Influence Well-Being}?
+{The Extended Active-Passive Model}.
+\emph{Current Directions in Psychological Science}, \emph{31}(1), 62--68.
+\url{https://doi.org/10.1177/09637214211053637}
 
-\leavevmode\vadjust pre{\hypertarget{ref-verduynPassiveFacebookUsage2015}{}}%
-Verduyn, P., Lee, D.
-S., Park, J., Shablack, H., Orvell, A., Bayer, J., Ybarra, O., Jonides, J., \& Kross, E.
-(2015).
-Passive {Facebook} usage undermines affective well-being: {Experimental} and longitudinal evidence.
-\emph{Journal of Experimental Psychology: General}, \emph{144}(2), 480--488.
-\url{https://doi.org/10.1037/xge0000057}
-
-\leavevmode\vadjust pre{\hypertarget{ref-wagnerAUTNESOnlinePanel2018}{}}%
+\bibitem[\citeproctext]{ref-wagnerAUTNESOnlinePanel2018}
 Wagner, M., Aichholzer, J., Eberl, J.-M., Meyer, T.
@@
 \emph{{AUTNES Online Panel Study} 2017 ({SUF} edition)}.
-{AUSSDA}.
+AUSSDA.
 \url{https://doi.org/10.11587/I7QIYJ}
 
-\leavevmode\vadjust pre{\hypertarget{ref-worldhealthorganizationWellbeingMeasuresPrimary1998}{}}%
+\bibitem[\citeproctext]{ref-worldhealthorganizationWellbeingMeasuresPrimary1998}
 World Health Organization. (1998). \emph{Wellbeing measures in primary health care/{The Depcare Project}}.
 
-\leavevmode\vadjust pre{\hypertarget{ref-yuePassiveSocialMedia2022}{}}%
-Yue, Z., Zhang, R., \& Xiao, J.
-(2022).
-Passive social media use and psychological well-being during the {COVID-19} pandemic: {The} role of social comparison and emotion regulation.
-\emph{Computers in Human Behavior}, \emph{127}, 107050.
-\url{https://doi.org/10.1016/j.chb.2021.107050}
-
-\leavevmode\vadjust pre{\hypertarget{ref-zillmannMoodManagementCommunication1988}{}}%
+\bibitem[\citeproctext]{ref-zillmannMoodManagementCommunication1988}
 Zillmann, D. (1988). Mood management through communication choices. \emph{American Behavioral Scientist}, \emph{31}(3), 327340. \url{https://doi.org/10.1177/000276488031003005}
@@
 
-\hypertarget{competing-interests}{%
-\section{Competing Interests}\label{competing-interests}}
-
-I declare no competing interests.
-
-\hypertarget{supplementary-material}{%
-\section{Supplementary Material}\label{supplementary-material}}
-
-All the stimuli, presentation materials, analysis scripts, and a reproducible version of the manuscript can be found on the companion website (\url{https://XMtRA.github.io/Austrian_Corona_Panel_Project}).
-
-\hypertarget{data-accessibility-statement}{%
-\section{Data Accessibility Statement}\label{data-accessibility-statement}}
-
-The data are shared on AUSSDA, see \url{https://doi.org/10.11587/28KQNS}.
-The data can only be used for scientific purposes.
-
-\hypertarget{acknowledgements}{%
-\section{Acknowledgements}\label{acknowledgements}}
+\section{Acknowledgements}\label{acknowledgements}
 

### 2024-04-04 preregistration_changes.Rmd
Substantial updates:  - Shortened manuscript  - Implemented peer-feedback

@@
 
-Explanation: The wording of the hypotheses was changed slightly.
-Instead of "trivial", it not states "very small".
-The reason was that trivial contains a normative element (irrelevant).
-Instead, I now state "very small", which seems more objective.
-Some might consider very small effects trivial, others not.
-
-In addition, I link to the control variables explicated below.
+Explanation: The wording of the hypotheses was extended to be more precise. In addition, I link to the control variables explicated below.
 

### 2024-07-17 manuscript.Rmd
- updated before submission

@@
 
-# Research Transparency Statement
-
-Conflicts of interest: The author declares no conflicts of interest.
-Funding: This research did not receive external funding.
-Artificial intelligence: Chat GPT was used for language polishing.
-Ethics: This research complies with the Declaration of Helsinki (2023), aside from the requirement to preregister human subjects research.
-Ethical review and approval was not required for the study in accordance with the local legislation and institutional requirements.
-The participants provided their written informed consent to participate in this study.
-Computational reproducibility: All analyses and the manuscript itself are completely reproducible (https://XMtRA.github.io/Austrian_Corona_Panel_Project).
-
-Preregistration: The hypotheses, the sample, the measures, the analyses, and the inference criteria (SESOI, _p_-value) were preregistered on the Open Science Framework (https://osf.io/87b24/?view_only=b2289b6fec214fa88ee75a18d45c18f3).
-Because in this study data from an already existing data set was analyzed, the preregistration was done prior to accessing the data.
-In some cases, it was impossible to execute the analyses as originally planned, for example because some properties of the variables only became apparent when seeing the actual data.
-All changes are reported online (https://xmtra.github.io/Austrian_Corona_Panel_Project/preregistration_changes.html).
-Materials: All study materials are publicly available (https://xmtra.github.io/Austrian_Corona_Panel_Project/items_english.html).
-Data: All primary data are publicly available (https://doi.org/10.11587/28KQNS).
-Analysis scripts: All analysis scripts are publicly available (https://XMtRA.github.io/Austrian_Corona_Panel_Project).
-
-\newpage
-
 # Introduction
@@
 The phenomenon of "doomscrolling" emerged as individuals struggled to disengage from COVID-19-related news [@buchananBriefExposureSocial2021; @sharmaDarkEndTunnel2022], sparking concerns about its impact on mental health [@sandstromDoomscrollingCOVIDNews2021].
-While initial research began exploring this question [e.g., @bendauAssociationsCOVID19Related2021; @edenMediaCopingCOVID192020; @sewallObjectivelyMeasuredDigital2021], the extent of social media's influence on well-being remains largely unknown.
+While initial research began exploring this question [e.g., @bendauAssociationsCOVID19Related2021; @edenMediaCopingCOVID192020; @sewallObjectivelyMeasuredDigital2021], the impact of COVID-19 related social media use on well-being remains largely unknown.
 This study aims to assess the effects of different social media usage patterns on individual well-being using a comprehensive longitudinal dataset spanning 34 waves, offering insights into within-person causal relationships.
@@
 Effects of social media depend on how they are used: while purposeful, active, social use is generally related to positive outcomes, passive, non-purposeful and asocial use is related to negative outcomes [@frisonExploringRelationshipsDifferent2015; @verduynSocialNetworkingSites2022].
+However, note that these general distinctions have been challenged recently [@meierDoesPassiveSocial2022, @valkenburgAssociationsActivePassive2022].
 In this study, I hence distinguish three types of use and five popular channels. 
@@
 The five channels to be investigated are Facebook, Twitter, Instagram, WhatsApp, and YouTube, which at the time ranked among the most popular social media services in Austria.
+
+This study aims to provide a nuanced analysis of how a specific type of social media use influences various well-being outcomes. 
+A general measure of time spent on social media would be insufficient to address how engagement with COVID-19-related content affects users. 
+Conversely, using overly specific measures might be less broadly relevant or insightful. 
+For example, demonstrating that receiving uplifting messages boosts social support or that encountering hate speech causes distress might yield limited insights. 
+That said, ultimately, the specificity of a measure is not inherently right or wrong but reflects different approaches to answering different questions.
 
@@
 <!-- Can social media use be such a factor? -->
-Current literature overviews suggest that social media use on average does seem to decrease well-being [@meierComputermediatedCommunicationSocial2020a].
+Current literature overviews suggest that more active social media users report on average lower well-being [@meierComputermediatedCommunicationSocial2020a].
 However, for most well-being outcomes, such as life satisfaction, general well-being, or loneliness, the effects are small [@meierComputermediatedCommunicationSocial2020a]. 
-These small findings can be explained with the differential susceptibility of media effects model [@valkenburgDifferentialSusceptibilityMedia2013], which states that there is substantial _variation_ of media effects for individual users.
+In addition, these results are based mostly on correlational research, and a recent meta study based on experimental studies did not find significant causal effects [@fergusonSocialMediaExperiments2024].
+These small or even non-existent findings can be explained with the differential susceptibility of media effects model [@valkenburgDifferentialSusceptibilityMedia2013], which states that there is substantial _variation_ of media effects for individual users.
 <!-- Whereas for some users social media are more beneficial, for others they are more harmful. -->
@@
 
-Why are the effects of social media use on well-being small on average?
+Why are the effects of social media use on well-being small on average, especially given that several scholars theorize strong negative effects [@haidtAnxiousGenerationHow2024, @twengeIGENWhyToday2017]?
 Two prominent media effect theories argue implicitly against strong average negative effects.
@@
 Social media often showcase the highlights and accomplishments of others, encouraging _social comparison_ [@przybylskiMotivationalEmotionalBehavioral2013]. 
-During a pandemic, seeing posts about others' successes or seemingly perfect lives might intensify feelings of inadequacy or FOMO, especially when individuals are unable to participate in similar activities due to restrictions or personal circumstances [@sharmaDarkEndTunnel2022].
+Especially during a pandemic, seeing how others successfully cope with every-day challenges---for example, by baking banana bread, inventing creative games with kids, or exercising at home---might intensify feelings of inadequacy or FOMO, especially when individuals are unable to participate in similar activities due to restrictions or personal circumstances [@sharmaDarkEndTunnel2022].
 
@@
 Posting COVID-19 content was more negative for Generation Z, potentially reflecting broader negative effects of social media on this generation. 
-<!-- Together, this finding is aligned with recent work challenging the assumption that active use is inherently beneficial and passive use detrimental [@valkenburgAssociationsActivePassive2022].  -->
+Together, this finding is aligned with recent work challenging the assumption that active use is inherently beneficial and passive use detrimental [@valkenburgAssociationsActivePassive2022].
 Notably, communication channels exhibited differences. 
@@
 While the study's focus on within-person effects enhances causal understanding, challenges exist related to additional relevant confounding exogenous variables not included here [@rohrerThinkingClearlyCorrelations2018], correct definition of the SESOI [@anvariUsingAnchorbasedMethods2021], and exact measurement of media use [@scharkowAccuracySelfreportedInternet2016].
-The measures of social media use did not include interpersonal communication about COVID-19 related content, which might result in more positive outcomes [@frisonExploringRelationshipsDifferent2015].
+As noted above, different operationalizations of COVID-19 related social media use might lead to different results. 
+For example, the measures of social media use did not include interpersonal communication about COVID-19 related content, which might result in more positive outcomes [@frisonExploringRelationshipsDifferent2015].
+While this study did not find meaningful within-person effects of social media use on well-being, between person analyses revealed strong negative correlations. 
+Alternative causal processes not explored in this paper might explain these correlations.
 The study's results are applicable primarily to Western societies and may not hold true in different cultural contexts.
@@
 -->
+
+\newpage
+
+# Research Transparency Statement
+
+Conflicts of interest: The author declares no conflicts of interest.
+Funding: This research did not receive external funding.
+Artificial intelligence: Chat GPT was used for language polishing.
+Ethics: This research complies with the Declaration of Helsinki (2023), aside from the requirement to preregister human subjects research.
+Ethical review and approval was not required for the study in accordance with the local legislation and institutional requirements.
+The participants provided their written informed consent to participate in this study.
+Computational reproducibility: All analyses and the manuscript itself are completely reproducible (https://XMtRA.github.io/Austrian_Corona_Panel_Project).
+
+Preregistration: The hypotheses, the sample, the measures, the analyses, and the inference criteria (SESOI, _p_-value) were preregistered on the Open Science Framework (https://osf.io/87b24/?view_only=b2289b6fec214fa88ee75a18d45c18f3).
+Because in this study data from an already existing data set was analyzed, the preregistration was done prior to accessing the data.
+In some cases, it was impossible to execute the analyses as originally planned, for example because some properties of the variables only became apparent when seeing the actual data.
+All changes are reported online (https://xmtra.github.io/Austrian_Corona_Panel_Project/preregistration_changes.html).
+Materials: All study materials are publicly available (https://xmtra.github.io/Austrian_Corona_Panel_Project/items_english.html).
+Data: All primary data are publicly available (https://doi.org/10.11587/28KQNS).
+Analysis scripts: All analysis scripts are publicly available (https://XMtRA.github.io/Austrian_Corona_Panel_Project).
 

### 2024-07-17 manuscript.tex
- updated before submission

@@
 \fi
-\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
+\usepackage{bookmark}
 \IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
@@
 
-\section{Research Transparency Statement}\label{research-transparency-statement}
-
-Conflicts of interest: The author declares no conflicts of interest.
-Funding: This research did not receive external funding.
-Artificial intelligence: Chat GPT was used for language polishing.
-Ethics: This research complies with the Declaration of Helsinki (2023), aside from the requirement to preregister human subjects research.
-Ethical review and approval was not required for the study in accordance with the local legislation and institutional requirements.
-The participants provided their written informed consent to participate in this study.
-Computational reproducibility: All analyses and the manuscript itself are completely reproducible (\url{https://XMtRA.github.io/Austrian_Corona_Panel_Project}).
-
-Preregistration: The hypotheses, the sample, the measures, the analyses, and the inference criteria (SESOI, \emph{p}-value) were preregistered on the Open Science Framework (\url{https://osf.io/87b24/?view_only=b2289b6fec214fa88ee75a18d45c18f3}).
-Because in this study data from an already existing data set was analyzed, the preregistration was done prior to accessing the data.
-In some cases, it was impossible to execute the analyses as originally planned, for example because some properties of the variables only became apparent when seeing the actual data.
-All changes are reported online (\url{https://xmtra.github.io/Austrian_Corona_Panel_Project/preregistration_changes.html}).
-Materials: All study materials are publicly available (\url{https://xmtra.github.io/Austrian_Corona_Panel_Project/items_english.html}).
-Data: All primary data are publicly available (\url{https://doi.org/10.11587/28KQNS}).
-Analysis scripts: All analysis scripts are publicly available (\url{https://XMtRA.github.io/Austrian_Corona_Panel_Project}).
-
-\newpage
-
 \section{Introduction}\label{introduction}
@@
 The phenomenon of ``doomscrolling'' emerged as individuals struggled to disengage from COVID-19-related news (Buchanan et al., 2021; Sharma et al., 2022), sparking concerns about its impact on mental health (Sandstrom et al., 2021).
-While initial research began exploring this question (e.g., Bendau et al., 2021; Eden et al., 2020; Sewall et al., 2021), the extent of social media's influence on well-being remains largely unknown.
+While initial research began exploring this question (e.g., Bendau et al., 2021; Eden et al., 2020; Sewall et al., 2021), the impact of COVID-19 related social media use on well-being remains largely unknown.
 This study aims to assess the effects of different social media usage patterns on individual well-being using a comprehensive longitudinal dataset spanning 34 waves, offering insights into within-person causal relationships.
@@
 Effects of social media depend on how they are used: while purposeful, active, social use is generally related to positive outcomes, passive, non-purposeful and asocial use is related to negative outcomes (Frison \& Eggermont, 2015; Verduyn et al., 2022).
+However, note that these general distinctions have been challenged recently Valkenburg et al. (2022).
 In this study, I hence distinguish three types of use and five popular channels.
@@
 
+This study aims to provide a nuanced analysis of how a specific type of social media use influences various well-being outcomes.
+A general measure of time spent on social media would be insufficient to address how engagement with COVID-19-related content affects users.
+Conversely, using overly specific measures might be less broadly relevant or insightful.
+For example, demonstrating that receiving uplifting messages boosts social support or that encountering hate speech causes distress might yield limited insights.
+That said, ultimately, the specificity of a measure is not inherently right or wrong but reflects different approaches to answering different questions.
+
 \subsection{Social Media Effects on Well-Being}\label{social-media-effects-on-well-being}
 
-Current literature overviews suggest that social media use on average does seem to decrease well-being (Meier \& Reinecke, 2020).
+Current literature overviews suggest that more active social media users report on average lower well-being (Meier \& Reinecke, 2020).
 However, for most well-being outcomes, such as life satisfaction, general well-being, or loneliness, the effects are small (Meier \& Reinecke, 2020).
-These small findings can be explained with the differential susceptibility of media effects model (Valkenburg \& Peter, 2013), which states that there is substantial \emph{variation} of media effects for individual users.
+In addition, these results are based mostly on correlational research, and a recent meta study based on experimental studies did not find significant causal effects (Ferguson, 2024).
+These small or even non-existent findings can be explained with the differential susceptibility of media effects model (Valkenburg \& Peter, 2013), which states that there is substantial \emph{variation} of media effects for individual users.
 For example, in one study it was estimated that roughly one quarter of all users experienced negative effects, another quarter positive effects, while for the rest the effects were neutral (Beyens et al., 2021).
 
-Why are the effects of social media use on well-being small on average?
+Why are the effects of social media use on well-being small on average, especially given that several scholars theorize strong negative effects Twenge (2017)?
 Two prominent media effect theories argue implicitly against strong average negative effects.
@@
 Social media often showcase the highlights and accomplishments of others, encouraging \emph{social comparison} (Przybylski et al., 2013).
-During a pandemic, seeing posts about others' successes or seemingly perfect lives might intensify feelings of inadequacy or FOMO, especially when individuals are unable to participate in similar activities due to restrictions or personal circumstances (Sharma et al., 2022).
+Especially during a pandemic, seeing how others successfully cope with every-day challenges---for example, by baking banana bread, inventing creative games with kids, or exercising at home---might intensify feelings of inadequacy or FOMO, especially when individuals are unable to participate in similar activities due to restrictions or personal circumstances (Sharma et al., 2022).
 
@@
 Posting COVID-19 content was more negative for Generation Z, potentially reflecting broader negative effects of social media on this generation.
+Together, this finding is aligned with recent work challenging the assumption that active use is inherently beneficial and passive use detrimental (Valkenburg et al., 2022).
 Notably, communication channels exhibited differences.
@@
 While the study's focus on within-person effects enhances causal understanding, challenges exist related to additional relevant confounding exogenous variables not included here (Rohrer, 2018), correct definition of the SESOI (Anvari \& Lakens, 2021), and exact measurement of media use (Scharkow, 2016).
-The measures of social media use did not include interpersonal communication about COVID-19 related content, which might result in more positive outcomes (Frison \& Eggermont, 2015).
+As noted above, different operationalizations of COVID-19 related social media use might lead to different results.
+For example, the measures of social media use did not include interpersonal communication about COVID-19 related content, which might result in more positive outcomes (Frison \& Eggermont, 2015).
+While this study did not find meaningful within-person effects of social media use on well-being, between person analyses revealed strong negative correlations.
+Alternative causal processes not explored in this paper might explain these correlations.
 The study's results are applicable primarily to Western societies and may not hold true in different cultural contexts.
@@
 In light of these minimal effects, concerns over COVID-19-related social media engagement on well-being could not be substantiated.
+
+\newpage
+
+\section{Research Transparency Statement}\label{research-transparency-statement}
+
+Conflicts of interest: The author declares no conflicts of interest.
+Funding: This research did not receive external funding.
+Artificial intelligence: Chat GPT was used for language polishing.
+Ethics: This research complies with the Declaration of Helsinki (2023), aside from the requirement to preregister human subjects research.
+Ethical review and approval was not required for the study in accordance with the local legislation and institutional requirements.
+The participants provided their written informed consent to participate in this study.
+Computational reproducibility: All analyses and the manuscript itself are completely reproducible (\url{https://XMtRA.github.io/Austrian_Corona_Panel_Project}).
+
+Preregistration: The hypotheses, the sample, the measures, the analyses, and the inference criteria (SESOI, \emph{p}-value) were preregistered on the Open Science Framework (\url{https://osf.io/87b24/?view_only=b2289b6fec214fa88ee75a18d45c18f3}).
+Because in this study data from an already existing data set was analyzed, the preregistration was done prior to accessing the data.
+In some cases, it was impossible to execute the analyses as originally planned, for example because some properties of the variables only became apparent when seeing the actual data.
+All changes are reported online (\url{https://xmtra.github.io/Austrian_Corona_Panel_Project/preregistration_changes.html}).
+Materials: All study materials are publicly available (\url{https://xmtra.github.io/Austrian_Corona_Panel_Project/items_english.html}).
+Data: All primary data are publicly available (\url{https://doi.org/10.11587/28KQNS}).
+Analysis scripts: All analysis scripts are publicly available (\url{https://XMtRA.github.io/Austrian_Corona_Panel_Project}).
 
@@
 
+\bibitem[\citeproctext]{ref-fergusonSocialMediaExperiments2024}
+Ferguson, C.
+J.
+(2024).
+Do social media experiments prove a link with mental health: {A} methodological and meta-analytic review.
+\emph{Psychology of Popular Media}.
+\url{https://doi.org/10.1037/ppm0000541}
+
 \bibitem[\citeproctext]{ref-frisonExploringRelationshipsDifferent2015}
@@
 \url{https://doi.org/10.3390/ijerph19031635}
+
+\bibitem[\citeproctext]{ref-haidtAnxiousGenerationHow2024}
+Haidt, J. (2024). \emph{The anxious generation: How the great rewiring of childhood is causing an epidemic of mental illness}. Penguin Press.
 
@@
 \url{https://doi.org/10.1038/s41598-022-05291-y}
+
+\bibitem[\citeproctext]{ref-meierDoesPassiveSocial2022}
+Meier, A., \& Krause, H.-V. (2022). \emph{Does passive social media use harm well-being? {An} adversarial review} {[}Preprint{]}. PsyArXiv. \url{https://doi.org/10.31234/osf.io/nvbwh}
 
@@
 
+\bibitem[\citeproctext]{ref-twengeIGENWhyToday2017}
+Twenge, J. M. (2017). \emph{{iGEN}: {Why} today's super-connected kids are growing up less rebellious, more tolerant, less happy-and completely unprepared for adulthood}. Atria Books.
+
 \bibitem[\citeproctext]{ref-valkenburgDifferentialSusceptibilityMedia2013}
 Valkenburg, P. M., \& Peter, J. (2013). The differential susceptibility to media effects model. \emph{Journal of Communication}, \emph{63}(2), 221--243. \url{https://doi.org/10.1111/jcom.12024}
+
+\bibitem[\citeproctext]{ref-valkenburgAssociationsActivePassive2022}
+Valkenburg, P.
+M., van Driel, I.
+I., \& Beyens, I.
+(2022).
+The associations of active and passive social media use with well-being: {A} critical scoping review.
+\emph{New Media \& Society}, \emph{24}(2), 530--549.
+\url{https://doi.org/10.1177/14614448211065425}
 

### 2024-07-17 preregistration.Rmd
- updated before submission

---
title: "Preregistration"
output:
  pdf_document:
    latex_engine: xelatex
  # html_document:
  #   toc: TRUE
  #   toc_float:
  #     collapsed: FALSE
  #     toc_depth: 3
---

The actual preregistration can be found [here](https://osf.io/87b24/?view_only=b2289b6fec214fa88ee75a18d45c18f3). I have also pasted it below.

## Hypotheses

The within-person effects of all types of COVID-19 related social media use on all types of well-being indicators will be trivial, while controlling for several stable and varying covariates such as  sociodemographic variables and psychological dispositions.

## Design Plan
#### Study type

Observational Study - Data is collected from study subjects that are not randomly assigned to a treatment. This includes surveys, “natural experiments,” and regression discontinuity designs.

#### Blinding

No blinding is involved in this study.

#### Is there any additional blinding in this study?

No response

## Study design

Secondary data analysis of a questionnaire-based panel study with 24 waves that consist of at least 1,500 participants per wave. A large number of questions was collected each, and a subset will be used for this study.

#### Randomization

No response

#### Sampling Plan

Existing Data

Registration prior to accessing the data

#### Explanation of existing data

I will analyze the Austrian Corona Panel. The data are posted on AUSSDA (https://doi.org/10.11587/28KQNS). I have not yet accessed the data. I have used the study description online to design the preregistration. After finishing the preregistration, I plan to create a login on AUSSDA. I will then download the scientific use file. In case the use file should not contain some of the variables I preselected, or if they are recoded such that the variables cannot be used, I will directly contact the study authors to ask for access to the original data set. In case this won't be possible, I will slightly change the analyses so that I can use the scientific use file.

#### Data collection procedures

I will use the already existing data set. Information on how the data was collected can be found here: https://viecer.univie.ac.at/coronapanel/austrian-corona-panel-data/method-report/

#### Sample size

The sample consists of at least 1,500 respondents per wave. Panel attrition was remedied by collecting new respondents. The overall and final sample size is not reported online, and will know it once I have downloaded the final data set.

#### Sample size rationale

The sample size was predetermined by the original study authors.

#### Stopping rule

Data collection is already finished.

## Variables
#### Manipulated variables

No variables were manipulated.

#### Measured variables

(Note: To improve readability, compared to the actual [preregistration](https://osf.io/87b24/?view_only=b2289b6fec214fa88ee75a18d45c18f3) the following part was reformatted and partially translated. No actual changes were introduced.)

In what follows, I list all the variables that I will use for the analyses (which also means that they cannot be dropped later). Note that the data set contains many more variables that I will not use. I provide the variable name in English, and the response options, answer format, and items in German.

###### Life Satisfaction

"Taken everything together, how satisfied are you currently with your life?"

(0 = extremely unsatisfied, 10 = extremely satisfied)

###### Positive Affect (scale)

"In the last week, how often did you feel"

 - a: ... calm and relaxed
 - e: ... happy
 - i: ... full of energy

(never [1]; on some days [2]; several times per week [3]; almost every day [4]; daily [5])

###### Negative Affect (scale)

"In the last week, how often did you feel"

 - b: ... lonely
 - c: ... aggravated
 - d: ... so depressed, that nothing could lift you up
 - f: ... very nervous
 - g: ... anxious
 - h: ... glum and sad

(never [1]; on some days [2]; several times per week [3]; almost every day [4]; daily [5])

###### Social media use: Reading

"Wie oft haben Sie letzte Woche die folgenden Aktivitäten in Sozialen Netzwerken gesetzt?"

 - "Postings anderer lesen, in denen es um Inhalte zum Coronavirus geht"

(1 = Mehrmals täglich; 2 = Einmal täglich; 3 = Mehrmals in der Woche; 4 = Einmal in der Woche; 5 = Gar nicht/nie)

###### Social media use: Liking & Sharing

"Wie oft haben Sie letzte Woche die folgenden Aktivitäten in Sozialen Netzwerken gesetzt?"

 - "Bei Postings anderer zum Coronavirus auf den „Gefällt mir“ Button drücken, „teilen“ oder „retweeten“"

(nie [1]; an manchen Tagen [2]; mehrmals die Woche [3]; beinahe jeden Tag [4]; täglich [5])

###### Social media use: Active Postings

"Wie oft haben Sie letzte Woche die folgenden Aktivitäten in Sozialen Netzwerken gesetzt?"

 - "Selbst Postings zum Coronavirus in Sozialen Netzwerken verfasst"

(nie [1]; an manchen Tagen [2]; mehrmals die Woche [3]; beinahe jeden Tag [4]; täglich [5])

###### Social media use: Specific platforms

"Wie oft in der letzten Woche haben Sie Informationen zur Corona-Krise in den folgenden Sozialen Medien verfolgt?"

 - Facebook 
 - Twitter
 - Instagram
 - Youtube
 - WhatsApp

(1 = Mehrmals täglich; 2 = Einmal täglich; 3 = Mehrmals in der Woche; 4 = Einmal in der Woche; 5 = Gar nicht/nie)

###### Gender

"Welches Geschlecht haben Sie?"

(a: Männlich; b Weiblich; c: Divers)

###### Age

"In welchem Jahr wurden Sie geboren?"

(Jahr)

###### Education

"Was ist der höchste Schul- oder Bildungsabschluss, den Sie erreicht haben?"

 - a. Volksschule oder weniger
 - b. Hauptschule oder AHS Unterstufe
 - c. Polytechnikum, BMS (Fachschule, z.B. HASCH)
 - d. Lehre, Berufsschule
 - e. AHS mit Matura
 - f. BHS mit Matura (z.B. HTL, HAK, HBLA etc.) 
 - g. Hochschulverwandte Lehranstalt oder Kolleg 
 - h. Bachelor
 - i. Magister / Master / Diplomingenieur / Fachhochschule 
 - j. Doktor / PhD

###### Austria country of birth

"Ist Ihr Geburtsland Österreich?"

  - Ja
  - Nein

###### Austria country of birth of parents

 "Wurde einer oder beide Ihrer Elternteile nicht in Österreich geboren?"

 - a. Ja, einer
 - b. Ja, beide
 - c. Nein

###### News consumption: Text-based

"Nachstehend finden Sie eine Liste mit verschiedenen Tageszeitungen. Wie oft haben Sie sich letzte Woche über das politische Geschehen in der jeweiligen Zeitung (im Internet oder gedruckt) informiert?"

 - Kronen Zeitung oder www.krone.at
 - Der Standard oder derstandard.at
 - Die Presse oder diepresse.com
 - Österreich oder oe24.at
 - Kurier oder kurier.at
 - Salzburger Nachrichten oder salzburg.at
 - Sonstige österreichische Tageszeitungen

###### News consumption: Video-based

"Nachstehend finden Sie eine Liste mit verschiedenen Nachrichtensendungen im Fernsehen. Wie oft haben Sie sich letzte Woche über das politische Geschehen in den folgenden Nachrichtensendungen informiert?"

 - "Nachrichtensendungen im ORF (z.B. Zeit im Bild um 19:30h oder ZiB 2 um 22:00h)"
 - "Nachrichtensendungen im Privatfernsehen (z.B. ATV aktuell, Puls 4/ Pro7/ Sat1 Austria)"

(1 = Mehrmals täglich; 2 = Einmal täglich; 3 = Mehrmals in der Woche; 4 = Einmal in der Woche; 5 = Gar nicht/nie)

###### Residency Vienna

"In welchem Bundesland leben Sie aktuell?"

 - a. Vorarlberg
 - b. Tirol
 - c. Salzburg
 - d. Steiermark
 - e. Kärnten
 - f. Oberösterreich
 - g. Niederösterreich
 - h. Wien
 - i. Burgenland

(Will be transformed to Vienna: yes/no)

###### Household size

"Bitte rechnen Sie hier sich selbst wieder mit ein. Wenn es eine Personengruppe nicht in Ihrem Haushalt gibt, tragen Sie bitte „0“ ein.

 - a. ...Erwachsene über 18 Jahre:
 - b. ...Teenager von 15-17 Jahren:
 - c. ...Kinder von 6-14 Jahren:
 - d. ...Kinder von 3-5 Jahren:
 - e. ...Kinder von 0-2 Jahren:

(Achtung, wenn Anzahl Erwachsene missing, 1 eintragen, da vorher Filterfrage)

###### Health

"Kommen wir jetzt zu ein paar Fragen zu Ihrer Gesundheit. Wie schätzen Sie Ihren allgemeinen Gesundheitszustand ein? Würden Sie sagen, er ist..."

 - a. 1 = Sehr gut
 - b. 2 = Gut
 - c. 3 = Mittelmäßig
 - d. 4 = Schlecht
 - e. 5 = Sehr schlecht

###### Living space

"Wie viele Quadratmeter hat Ihre Wohnung?"

(Einfach-Nennung, Zahleneingabe, min. 0 – max. 600, freiwillig)

###### Access garden

"Haben Sie Zugang zu einer privaten Freifläche (Balkon, Garten)?

 - (Ja, Garten)

###### Access balcony

"Haben Sie Zugang zu einer privaten Freifläche (Balkon, Garten)?"

 - (Ja, Balkon)

###### Employment

Wenn Sie sich selbst zuordnen: Im Februar 2020, welcher der folgenden Gruppen gehören Sie vorwiegend an? (Einfach-Nennung)

 - a. Unselbstständig erwerbstätig (auch Lehrling), Privatwirtschaft
 - b. Unselbstständig erwerbstätig (auch Lehrling), Öffentlicher Dienst
 - c. Selbstständig erwerbstätig
 - d. Arbeitslos
 - e. Pensionist*in
 - f. Ausschließlich haushaltsführend
 - g. Schüler/in, Student/in, Bildungskarenz
 - h. Dauerhaft arbeitsunfähig
 - i. Präsenz-/Zivildiener
 - j. In Elternkarenz

###### Work hours per week

"Wie viele Stunden (inklusive Überstunden) arbeiten Sie in Ihrer Haupttätigkeit jetzt pro Woche? Wenn Sie es nicht genau wissen, genügt Ihre beste Schätzung."

###### Home-Office Yes/No

"Hat sich in Ihrer beruflichen Situation aufgrund der Corona-Krise etwas geändert? Bitte wählen Sie alles Zutreffende aus."

 - a. Ich bin im Home-Office

###### Household income

"Wie viel Geld steht Ihrem Haushalt aktuell zur Verfügung (Netto-Haushaltseinkommen inklusive Sozialleistungen, Rente usw.)? Bitte rechnen Sie Überstunden mit ein, nicht aber den 13./14. Bezug."

 - a. weniger als 1.100 Euro
 - b. 1.100 bis unter 1.500 Euro 
 - c. 1.500 bis unter 1.800 Euro 
 - d. 1.800 bis unter 2.200 Euro 
 - e. 2.200 bis unter 2.700 Euro 
 - f. 2.700 bis unter 3.100 Euro 
 - g. 3.100 bis unter 3.700 Euro 
 - h. 3.700 bis unter 4.300 Euro 
 - i. 4.300 bis unter 5.500 Euro 
 - j. 5.500 Euro und mehr

###### Outdoor activities

"Haben Sie wegen folgenden Gründen in der vergangenen Woche Ihr Zuhause verlassen?"

 - a: Arbeit
 - b: Sport
 - c: Freunde oder Verwandte treffen
 - e: Essen einkaufen
 - f: Haustiere ausführen

(nie [1]; an manchen Tagen [2]; mehrmals die Woche [3]; beinahe jeden Tag [4]; täglich [5])

###### Satisfaction Democracy

"Wie zufrieden oder unzufrieden sind Sie alles in allem mit der Demokratie, so wie sie derzeit in Österreich funktioniert?"

 - a. 1 = Sehr zufrieden
 - b. 2 = Eher zufrieden
 - c. 3 = Teils-teils
 - d. 4 = Eher unzufrieden 
 - e. 5 = Sehr unzufrieden

###### Risk-taking

"Wie sehr sind Sie im Allgemeinen dazu bereit, Risiken einzugehen?"

(0: gar nicht bereit; 10: sehr bereit)

###### Locus of Control

 - "Ich habe mein Leben selbst in der Hand."
 - "Wenn ich mich anstrenge, werde ich auch Erfolg haben."
 - "Egal ob privat oder im Beruf: Mein Leben wird zum großen Teil von anderen bestimmt."
 - "Meine Pläne werden oft vom Schicksal durchkreuzt."

(1 = Trifft voll und ganz zu; 2 = Trifft etwas zu; 3 = Trifft wenig zu; 4 = Trifft nicht zu)
#### Indices

All the measures above that include the label (scale) will be combined into a scale. These variables will be analyzed using CFAs. If, and only if, model fit is below the recommended thresholds, the models will be adapted to achieve sufficient fit (see below). To avoid overfitting, I will use more liberal model fit criteria (CFI > .90, TLI > .90, RMSEA <. .10, SRMR < .10). Of these final models I will export factor scores, which will be used for the subsequent analyses.

## Analysis Plan
#### Statistical models

The data will be analyzed using random effects within between models. The independent variables (all types of social media use) will be analyzed on a between person level (person average) and a within-person level (deviation from person average for each wave).

Because the models are highly complex it is likely that they will not converge immediately. I plan the following analysis pipeline.

Analysis (proceed if doesn't converge):
- REWB Frequentist Model: Random Slope for Social Media Use & Covars
	- REWB Frequentist Model: Random Slope for Social Media Use & Fixed Slope for Covars
		- REWB Frequentist Model: Fixed Slope for Social Media Use  & Fixed Slope for Covars
			- REWB Frequentist Model: Fixed Slope for Social Media Use & Fixed Slope for Covars with standardized effects
				- REWB Bayesian Model
					- REWB Bayesian Model with standardized effects

CFA:
The factorial validity of the measures and the hypotheses will be tested with structural equation modeling (SEM). If Mardia’s test shows that the assumption of multivariate normality is violated, I will use the more robust Satorra-Bentler scaled and mean-adjusted test statistic (MLM) as estimator. I will test each scale in a confirmatory factor analysis. To avoid overfitting, I will use more liberal fit criteria (CFI > .90, TLI > .90, RMSEA <. .10, SRMR < .10) (Kline 2016). If model fit is below the criteria, I will first inspect modification indices, potentially allowing covariance or cross-loadings if theoretically plausible. If these changes do not yield sufficient fit, I will drop malfunctioning items. If fit is still subpar, I will conduct exploratory factor analyses (EFA) to assess the underlying factor structure. EFAs will be run using maximum likelihood estimation and oblimin rotation (Osborne and Costello 2004, 7). If more than one dimension will be revealed, I will implement bifactor model solutions. Bifactor models retain a general measure of the variable, and make it unnecessary to introduce novel (and potentially overfitted) subdimensions. If no adequate bifactor model can be found, I will proceed by deleting items with low loadings on the general factor and/or the specific factors. If also after deletion of individual items no bifactor solution should emerge, I will use a subset of the items to extract a single factor with sufficient factorial validity.

#### Transformations

See list of variables.

#### Inference criteria

Definition of trivial effect size (SESOI)

- b < .3 (life-satisfaction), b < .15 (positive and negative affect)
	- If I need to use standardized effects (see above): beta < .1
- in addition, p < 5%

#### Data exclusion

I will individually check responses for patterns such as straight-lining or missing of inverted items, making sure to remove only clear cases. I will exclude respondents if they answer less than 50% of all questions.

#### Missing data

Missing responses will be imputed using predictive mean matching.

#### Exploratory analysis

I will provide additional analyses online, such as the results without control variables. These aren't alternative results, but rather stem from procedures that I consider -- at this point -- to be inferior.

## Other

It's a deliberate decision not to statistically control the analyses for potential mediators, such as trust in politics, media, etc.
